### 2011-11-09 trunk/book/author.tex
First version

%%%%%%%%%%%%%%%%%%%% author.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% sample root file for your "contribution" to a contributed volume
%
% Use this file as a template for your own input.
%
%%%%%%%%%%%%%%%% Springer %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[graybox]{svmult}

% choose options for [] as required from the list
% in the Reference Guide

\usepackage{mathptmx}       % selects Times Roman as basic font
\usepackage{helvet}         % selects Helvetica as sans-serif font
\usepackage{courier}        % selects Courier as typewriter font
\usepackage{type1cm}        % activate if the above 3 fonts are
                            % not available on your system
%
\usepackage{makeidx}         % allows index generation
\usepackage{graphicx}        % standard LaTeX graphics tool
                             % when including figure files
\usepackage{multicol}        % used for the two-column index
\usepackage[bottom]{footmisc}% places footnotes at page bottom
\usepackage{url}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{amsfonts}		% math fonts.

% see the list of further useful packages
% in the Reference Guide

\makeindex             % used for the subject index
                       % please use the style svind.ist with
                       % your makeindex program

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\Pow}{\mathcal{P}}
\newcommand{\N}{\operatorname{N}}
\newcommand{\bool}{\operatorname*{\mathcal{B}}}
\newcommand{\Pos}{\operatorname{Pos}}
\newcommand{\Nec}{\operatorname{Nec}}
\newcommand{\T}{\mathcal{T}}

\begin{document}

\title*{An overview on temporal databases.\\
 New approaches and proposals.}
% Use \titlerunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\author{Jos\'e Pons \and Christophe Billiet \and Olga Pons}
% Use \authorrunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\institute{Jos\'e Pons \at Department of Computer Science and Artificial Intelligence, Univeristy of Granada, C/ Periodista Daniel Saucedo Aranda s/n \email{jpons@decsai.ugr.es}
\and Christophe Billiet \at Department of Telecommunications and Information Processing, Ghent University, Belgium, St.-Pietersnieuwstraat 41, B-9000 Gent,Belgium. \email{Christophe.Billiet@telin.ugent.be}
\and Olga Pons \at Department of Computer Science and Artificial Intelligence, Univeristy of Granada, C/ Periodista Daniel Saucedo Aranda s/n \email{opc@decsai.ugr.es}
}
%
% Use the package "url.sty" to avoid
% problems with special characters
% used in your e-mail or web address
%
\maketitle

\abstract*{There exist some objects in the nature that have a time-variant (e.g. an employee's working life) or time-related (e.g. the transactions in a bank account) nature. Modelling these objects in a (relational) database is possible, but the time-related attributes have an impact on the consistency of the entire database. Therefore, some temporal database models have been proposed to deal with this. These models deal both with the problem of representation and querying. Time itself is a source of imprecision / vagueness / uncertainty since the measures of time are inherently imprecise. Humans beings manage the time with imprecision and vagueness. Nevertheless, the amount of uncertainty in the temporal indications are defined by socially well-known boundaries. Several proposals for dealing with imprecision in time have been done: some proposals consider the basis of the system the imprecision related to the conversion of diferent time points and intervals. While some other proposals consider the specification of the time as the source of imprecision: the time points are ill-known.}

\abstract{There exist some objects in the nature that have a time-variant (e.g. an employee's working life) or time-related (e.g. the transactions in a bank account) nature. Modelling these objects in a (relational) database is possible, but the time-related attributes have an impact on the consistency of the entire database. Therefore, some temporal database models have been proposed to deal with this. These models deal both with the problem of representation and querying. Time itself is a source of imprecision / vagueness / uncertainty since the measures of time are inherently imprecise. Humans beings manage the time with imprecision and vagueness. Nevertheless, the amount of uncertainty in the temporal indications are defined by socially well-known boundaries. Several proposals for dealing with imprecision in time have been done: some proposals consider the basis of the system the imprecision related to the conversion of diferent time points and intervals. While some other proposals consider the specification of the time as the source of imprecision: the time points are ill-known.}

\section{Introduction}
\label{sec:introduction}
\input{sources/intro.tex}

\section{Time modelling}
\label{sec:time-domain}
\input{sources/timeDomain.tex}

\section{Temporal Databases}
\label{sec:temporal-databases}
\input{sources/temporalDatabases.tex}

\section{Imprecision, Uncertainty and vagueness in temporal databases}
\label{sec:imprecision-uncertainty}
\input{sources/fuzzyTemporalDatabases.tex}

\section{Bipolarity on temporal databases}
\label{sec:bipolarity}
\input{sources/bipolarity.tex}

\section{Further research work}
\label{sec:further-research-work}
\input{sources/frw.tex}

\section*{Acknowledgements}
Part of this research is supported by the grant BES-2009-013805 within the research project TIN2008-02066: \emph{Fuzzy Temporal Information treatment in relational DBMS}.

\input{referenc}
\end{document}

### 2011-11-09 trunk/book/sources/prelim.tex
First version

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Preliminaries
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In this section we will introduce some background concepts about possibility theory and temporal databases. Subsection \ref{subsec:poss} is an introduction of possibility theory. The concepts defined by Dubois and Prade~\cite{Dubois1983} about fuzzy numbers and fuzzy intervals are also explained. Temporal databases and their characteristics are introduced in subsection \ref{subsec:temporal}.

\subsection{\label{subsec:poss}Possibilistic variables}
Informally, a theory of uncertainty considers a set of uncertain events and tries to quantify the confidence about the occurence of those events. Consider therefore a set of outcomes $\Omega$ and let each subset of $\Omega$ be called an \emph{event}. Hereby, $\Omega$ is called the \emph{certain event} and can thus be regarded as a point of reference. Similarly, $\emptyset$ is called the \emph{impossible event}. The quantification of confidence in a theory of uncertainty is achieved by means of a \emph{confidence measure}~\cite{Choquet53}.
\begin{definition}
\label{def:conf-measure}
Consider a universe of outcomes $\Omega$. A confidence measure is defined by a function:
\begin{equation}
g:\Pow(\Omega)\rightarrow[0,1]
\end{equation}
that satisfies:
\begin{eqnarray}
g\left(\emptyset\right)&=&0\\
g\left(\Omega\right)&=&1\\
A\subseteq B&\Rightarrow&g(A)\leq g(B)
\end{eqnarray}
and where $\Pow(\Omega)$ represents the power set of $\Omega$.
\end{definition}
From the constraints in Definition \ref{def:conf-measure}, it can be seen that $\emptyset$ and $\Omega$ are indeed respectively the impossible and the certain event. In addition, the confidence that is assigned to events is monotonic. Due to this monotonicity, we have that:
\begin{eqnarray}
\forall A\in\Omega:\forall B\in\Omega&:&g\left(A\cup B\right)\geq\max\left(g\left(A\right),g\left(B\right)\right)\\
\forall A\in\Omega:\forall B\in\Omega&:&g\left(A\cap B\right)\leq\min\left(g\left(A\right),g\left(B\right)\right).
\end{eqnarray}
We can thus derive two extreme measures of confidence and these measures play a crucial role in the theory of possibility~\cite{Dubois88}. Consider a confidence measure $\Pi$ that satisfies the following constraint:
\begin{equation}
\label{EQ:MAXITIVITY}
\forall A\subseteq\Omega:\forall B\subseteq\Omega:\Pi\left(A\cup B\right) = \max\left(\Pi(A),\Pi(B)\right).
\end{equation}
The function $\Pi$ is then called a \emph{possibility measure} and due to \eqref{EQ:MAXITIVITY}, it is said that $\Pi$ is \emph{maxitive}~\cite{Dubois88}. For each event $A\subseteq\Omega$, $\Pi(A)$ expresses the possibility that $A$ occurs. Due to the maxitivity of $\Pi$ and under the constraint that $\Omega$ is finite, we can easily extract information about the possibility of separate outcomes in the form of a \emph{possibility distribution} $\pi$ defined by:
\begin{equation}
\pi:\Omega\rightarrow[0,1]:\pi(u)=\Pi\left(\left\{u\right\}\right).
\end{equation}
Equivalently, we have that:
\begin{equation}
\forall A\subseteq\Omega:\Pi(A)=\sup_{u\in A}\pi (u).
\end{equation}
A possibility distribution is normalized in the sense that:
\begin{equation}
\sup_{u\in\Omega}\pi(u)=1.
\end{equation}
In order to ensure the existence of a possibility distribution when $\Omega$ is infinite, the maxitivity constraint \eqref{EQ:MAXITIVITY} should be replaced by:
\begin{equation}
\Pi\left(\bigcup_{i}A_i\right)=\sup_{i}\Pi\left(A_i\right)
\end{equation}
for any family $\left\{A_i\right\}$ of indexed subsets of $U$~\cite{Nguyen79}.

Let us now consider a confidence measure $\N$ that satisfies:
\begin{equation}
\label{EQ:MINITIVITY}
\forall A\subseteq\Omega:\forall B\subseteq\Omega:\N\left(A\cap B\right) = \min\left(\Pi(A),\Pi(B)\right).
\end{equation}
The function $\N$ is then called a \emph{necessity measure} and due to \eqref{EQ:MINITIVITY}, it is said that $\N$ is \emph{minitive}. Possibility measures and necessity measures are dual in the sense that:
\begin{equation}
\forall A\subseteq\Omega:\Pi(A)=1-\N(\overline{A}).
\end{equation}
The principles of possibility theory are often used in the scope of a \emph{possibilistic variable}, also known as an \emph{ill-known value}.
\begin{definition}
Consider a universe $U$. A possibilistic variable $X$ over $U$ is a variable taking exactly one value in $U$, but for which this value is (partially) unknown. The available knowledge about the value that $X$ takes, is given by a possibility distribution $\pi_X$. For each $u\in U$, $\pi_X(u)$ represents the possibility that $X$ takes the value $u$.
\end{definition}
A specific application of possibility theory~\cite{Prade82,VanSchooten,DeCooman} is obtained when the universe is the set of Boolean values, i.e. $U=\mathbb{B}=\{T,F\}$. If we consider a Boolean proposition, then this proposition is either true ($T$) or false ($F$). Any Boolean proposition $p$ can thus be regarded as a possibilistic variable for which the possibility and necessity that $p=T$ are denoted in short as $\Pos(p)$ and $\Nec(p)$. The following notations are used:
\begin{eqnarray}
\Pos(p)&=&\pi_p(T)\\
\Nec(p)&=&1-\pi_p(F).
\end{eqnarray}
Keeping in mind the cause of uncertainty that is assumed, it is important to understand that a possibilistic variable $X$ is bounded to take one value only, but this value is not known due to incomplete knowledge. When a possibilistic variable is defined over the universe $\Pow(U)$, it is said to be a \emph{possibilistic multi-valued variable}, also known as an \emph{ill-known set}~\cite{Dubois88b}. Note that an ill-known set conceptually differs from a fuzzy set in the sense that an ill-known set is a crisp set, but for some reason (partially) unknown.
\subsection{\label{subsubsec:fuzzy}Fuzzy numbers and fuzzy intervals}
When we shall discuss the case of fuzzy temporal databases, we shall link the concepts introduced in this paper to the theory of fuzzy intervals. Dubois and Prade~\cite{Dubois1983} proposed the following definition of \emph{fuzzy interval}.
\begin{definition}
A fuzzy interval is a fuzzy set $M$ on the set of real numbers $\mathbb{R}$ such that:
\begin{eqnarray}
\forall (u,v)\in\mathbb{R}^2:\forall w \in [u,v]:&\mu_M(w) \geq\min(\mu_M(u),\mu_M(v)) \ &\mbox{ (convexity) } \\
\exists m \in \mathbb{R}:& \mu_M(m)=1 \ &\mbox{ (normalization) }
\end{eqnarray}
\end{definition}
If $m$ is unique, then $M$ is referred to as a \emph{fuzzy number}, instead of a \emph{fuzzy interval}. In other words, if the core of a fuzzy interval is a singleton, it can be seen as a fuzzy number. In their work, Dubois and Prade propose four different functions (two possibility and two necessity functions) to asses the position of a fuzzy number N relative to that of a fuzzy number M taken as a reference.

The most convenient representation for the membership function of a fuzzy number is a triangular function. The membership function $\mu_M$ for the fuzzy set $M$ has also the properties of convexity and normalization. Three values represent a triangular function (see Figure~\ref{fig:triangular}):
\begin{itemize}
\item
$D$ is the singleton value in the core of $\mu_M(x)$.
\item
$D-a$ is the lower value in the support of $\mu_M(x)$. 
\item
$D+b$ is the upper value in the support of $\mu_M(x)$.
\end{itemize}
Note that the values $a$ and $b$ are values in the underlying ordered domain. E.g. $(a,b) \in \mathbb{R}^2$. The notation for a triangular function adopted here is $[D,a,b]$.
\begin{figure}[h!]
  \centering
  \includegraphics[scale=0.4]{graphs/triangular.eps}
  \caption{Triangular function.}
  \label{fig:triangular}
\end{figure}

The most simple representation for the membership function of a fuzzy interval is a trapezoidal function. This membership function $\mu_T$ for the fuzzy interval $T$ is convex and normalized. Four values represent a trapezoidal membership function (figure  \ref{fig:trapezoidal}): $\left[\alpha,\ \beta,\ \gamma,\ \delta\right]$. The membership function is given by:

\begin{align}
\mu_T(x)
\begin{cases}
1 & \mbox{ if } x \in [\beta,\gamma] \\
0 & \mbox{ if } x > \delta \vee x < \alpha \\
\frac{x-\alpha}{\beta - \alpha} & \mbox{ if } x \in [\alpha,\beta[ \\
\frac{\delta -x}{\delta - \gamma} & \mbox{ if } x \in ]\gamma,\delta] \\
\end{cases}
\end{align}

\begin{figure}[h!]
  \centering
  \includegraphics[scale=0.4]{graphs/trapezoidalDistribution.eps}
  \caption{Trapezoidal distribution.}
  \label{fig:trapezoidal}
\end{figure}

\subsection{\label{subsec:temporal}Temporal databases}

\subsubsection{\label{subsubsec:timeDomain}Time Domain}
The semantics of time and the different approaches to the time model are analyzed in this section; a more detailed explanation is discussed in \ref{sec:timeindatabasedesign}.

The general time model represents time as a set with an imposed partial order. There are two main models of time: linear and cyclic. In the linear model, time advances from the past to the future in a totally ordered fashion (a relationship of total order is imposed to the set). A cyclic model of time is used for recurrent processes. The present proposal focuses on the linear model.
\begin{itemize}
\item
Discrete model:  The discrete model of time is isomorphic in relation to the natural numbers. According to this model, each natural number corresponds to an indivisible time unit. The said time unit is called a chronon or, in other words, a 'time quantum', 'moment' or 'instant'. A chronon is the smallest time unit that can be represented by the model.
\item
Dense model: This model is isomorphic in relation to the rational or the real numbers. It represents one instant of time in the gap between another two instants.
\item
Continuous model: This is an extension of the dense model with no gaps. It is isomorphic in relation to the real numbers.
\end{itemize}
Restrictions on time range may be performed on the base of the following two concepts:
\begin{itemize}
\item
Time boundedness: Time can be bounded orthogonally in the past and in the future.
\item
Time distance: As a metric, time has a distance function which presents four properties:
\begin{enumerate}
\item
Distance is non-negative.
\item
The distance between two different elements is not equal to zero.
\item
The distance from $\alpha$ to $\beta$ is the same as that between $\beta$ and $\alpha$.
\item
Triangle inequality: The distance from $\alpha$ to $\gamma$ is equal or greater than the distance from $\alpha$ to $\beta$ plus the distance from $\beta$ to $\gamma$.
\end{enumerate}
\end{itemize}
On the other hand, time may be relative or absolute (in other words, anchored or unanchored). Sometimes, what one considers 'absolute time' is not as definite as one would hope since our concept of absolute time is based on another time reference (for example, January 1st of year one). Relative time has a direction, differently from distance. It is also possible to use negative references so as to represent relative time (e.g. -3 days = three days ago).

\subsubsection{Time in database design}
\label{sec:timeindatabasedesign}
This section defines the model used to represent time in a database. This model allows working with time instants and time intervals.  Even though there is no general agreement as to whether real-time line is continuous or discrete, a discrete model of time may be the most appropriate for a temporal database.

	A chronon may be defined as the basic unit of time. Mathematically, a chronon is isomorphic to a finite sequence of natural numbers. For instance, if a chronon is \emph{11232010}, it represents a day, namely 11/23/2010. The symbol \emph{c} denotes a chronon. A chronon may represent different time granularity, depending on database or user processing needs (i.e. minutes, seconds, milliseconds, etc.). Real-time instants are smaller than chronons even though they are represented by the said chronons The real-time line is represented by a finite sequence of chronons.

	It is possible to represent time intervals by a sequence of consecutive chronons. A sequence of consecutive chronons could be represented by the pair of starting and terminating chronons. A restriction must be imposed so that the starting instant comes before the ending instant.
\subsubsection{Temporal data models and time domains}
\label{sec:temporaldatamodelsandtimedomains}
A temporal data model is a relational data model which has temporal relations as underlying data structure, and whose operators are all temporal. In \cite{Böhlen_point-versus}, a study between interval-based and point-based temporal data models is shown. We are going to explain the main definitions and the equivalence between the models.

Temporal relations include a temporal attribute. A tuple of a temporal relation may be put under the form $(x_1,...,x_n||t_s)$ where $(x_1,...,x_n)$ are the non-temporal attribute values and \emph{t} is the tuple timestamp. Furthermore, an operator is temporal only if it generates a temporal relation when it is applied to temporal relations.

	The choice of the timestamp representation of the database facts is an important factor. Two options are the most common choices: time points and time intervals. Time intervals may be built from time points. The definition of time point is the following:
Let T be an infinite set.

$T^p=\langle T,< \rangle$ is a time point domain over $T$ iff $<$ defines a total order on $T$. Each element of $T$ corresponds to a time point of $T^p$.

The definition of time interval is:

A time interval $I$ of $T^p$ is any connected subset of $T^p$.

$T^i = \langle I,\subset \rangle $ is the time interval domain over $T^p$ iff $I$ is the set of all the intervals of $T^p$.

Both $T^p$ and $T^i$ are time domains over $T$.

A timestamp over $T^p$ is either a time point or a time interval of $T^p$. Time intervals can be represented as sets of time points which may, in turn, be represented by a time interval with the same values at both of its ends. Sometimes a coalesce operator is defined so as to switch between a set of time points and a time interval. Logically, defining a un-coalescing operator is quite complex. 
\subsubsection{\label{subsubsec:timeDb}Time in databases}
We can classify the time recorded in the database through its semantics. As a result, two kinds of time derive from two orthogonal time domains: valid time and transaction time. Valid-time captures the time-varying nature of the reality model. Transaction time structures the update activity associated with the object. The resulting temporal relations are those emerging from the bitemporal conceptual data model BCDM, on which TSQL2 is based.

The domain of valid time is given as $D_{VT}=\lbrace{c_1^v,c_2^v,...,c_k^v}\rbrace$, whereas the domain of transaction time may be given by $D_{TT}=\lbrace{c_1^t,c_2^t,...,c_k^t}\rbrace$. Thus, a valid-time chronon $c^v$  is a member of $D_{VT}$, a transaction-time chronon $c^t$ is a member of $D_{TT}$ and a bitemporal chronon $c^b=(c^t,c^v)$ is an ordered pair of a transaction-time chronon and valid-time chronon.

We can define the set of names $D_A=\lbrace{A_1,A_2,...,A_n}\rbrace$as attributes and the following set of domains for the said attributes:$D_D=\lbrace{D_1,D_2,...,D_n}\rbrace$. We may also define $\bot_i$ as inapplicable,$\bot_u$ as unknown, and $\bot$ as inapplicable or unknown null values.

The schema for a bitemporal relation $'R'$ consists of an arbitrary number of explicit attributes from $D_A$ with domains in $D_D$ and an implicit timestamp attribute $T$ with domain $2^{\lbrace D_{TT}\bigcup \lbrace UC \rbrace \cdot D_{VT}\rbrace}$, where $UC$ ('until changed') means that the tuple is current in the database. This schema is a special transaction-time marker.

A 'fact' is the information stored in a tuple. We define a tuple as $x=(a_1,a_2,...,a_n|t^b)$ in a bitemporal conceptual relation instance $r(R)$, which consists of a number of attribute values associated with a bitemporal timestamp value. A value $(UC,c^v)$ in a timestamp indicates that the tuple is current in the database.

We say that a fact is true in the reality model during each valid-time chronon when a subset of the valid-time domain is associated with that fact.

We say that a fact is current in the relation during each transaction-time chronon when a subset of the transaction-time chronon domain is associated with the valid-time chronon. Only the subset of transaction times less than UC may be associated with a valid time.

While the definition of a bitemporal chronon is symmetric, the relation between valid time and transaction time is asymmetric. It should also be noted that since no two tuples with identical attribute values are allowed in a bitemporal relation instance, the full history of a fact is contained in a single tuple. In graphical representations of bitemporal space, the x-axis represents the transaction-time dimension and the y-axis denotes the valid-time dimension.

There are two special cases of bitemporal relations:
\begin{itemize}
\item
Valid-time relations, which support only valid time.
\item
Transaction-time relations,which support only transaction time.
\end{itemize}
We use the term 'snapshot relation' for a conventional relation. This relation supports neither transaction nor valid time.
\subsubsection{Implicit versus explicit timestamps}
\label{sec:implicitversusexplicittimestamps}
The association between time and facts may be carried out in two ways: explicitly, where the relation is represented by fully explicit timestamps, and implicitly. This distinction is relevant to three aspects of the data model: the update language, the display of data and the query language.

	If the time model works with transaction time, it means that the update language contains implicit timestamps. The transaction-time chronons of the facts are supplied by the system; in the valid-time model, these chronons are supplied by the user. Thus, in a bitemporal data model, the chronons are also supplied by the user.

	If the time model allows displaying temporal facts, it must represent such data explicitly. It is possible that the model allows the user to choose between several representations of time. In some simple transaction-time models it is common that the temporal facts cannot be displayed.

	The main difficulty about differentiating between implicit and explicit timestamps is the query language. By means of the query language, we may directly access temporal attributes: the temporal attributes are as explicit as the other attributes in the database. If the query language does not allow direct access to temporal attributes, the association schema between timestamps and facts is considered implicit.

\subsubsection{\label{subsubsec:models}Temporal data models}
There are several extensions to the relational model that include the temporal element. In the current section, we briefly describe the model and some of its most interesting features. A more detailed summary can be found in \ref{sec:timeindatabasedesign}. Some models work only with valid or transaction time, whereas a few models can handle both. The current study starts with the models supporting valid time only, continues through the models which support transaction time and finally deals with the bitemporal models.

\paragraph{Valid-time models}
Most models support valid time only.

Brooks:  \cite{brooks56} Proposes a tridimensional view of valid-time database.

Wiederhold: \cite{Blum:1981:DCD:1672611.1672632} Time-oriented database, developed to work with medical applications. This model depicts relations as sets of entity-attribute-time-value quadruples. Time is associated with the visit number of the patient.

Jones LEGOL 2.0: \cite{legol20} A language used in legislative writing. The temporal order of elements and the valid times for objects are important. It was the first time-oriented algebra to be defined, and some of its features are found in subsequent algebras.
Tuples in this model have two time attributes: Start and Stop. Each one indicates the start point and the end point of the interval where the tuple is valid.

Clifford-1: \cite{Clifford82},\cite{Clifford:1983:FST:319983.319986} Each relation schema has an additional time value called 'State'. Also, a boolean attribute is added to indicate if this tuple exists for that state.

Ariav: \cite{Ariav:1986:TOD:7239.7350} Temporally-oriented data model: a valid-time relation is a set of snapshot relation states indexed by valid time. A calculus-based query language is associated with the model: TOSQL.

Navathe: \cite{TSQL} Proposes a temporal extension to SQL called TSQL, the temporal algebra associated to the temporal relational model. This model allows a relationship to handle time-varying attributes, and also relationships without time-varying attributes. The only restriction is that all attributes in the relation must be of the same type. Objects are classified in snapshot relations and valid-time relations. In the latter, each tuple has associated to it an interval of validity with two points: time-start and time-end, like in the LEGOL 2.0 model.

Sadeghi: \cite{913787} Proposes a calculus-based valid-time language, HQL, where all objects represent valid-.time relations. As in Navathe and LEGOL 2.0, two time points are required for the ends of the interval. The model requires coalescing.

Sarda: \cite{Sarda:1990:ESH:627277.627409} This model was designed to support the calculus-based model HSQL. The model associates valid-time with tuples. Objects can either be snapshot or valid-time relations. In this model, valid time is considered to be closed on the left part of the interval and open on the right. Time is represented by an implicit attribute called 'Period' as a single non-atomic value.

The following time data models are not in first normal form (1NF), which means they might have multiple values per attribute. Even though these models are an extension of the 1NF, the representation is not in 1NF, but the operators work on valid-time relations, which are extensions of conventional relations.

Segev: \cite{Segev:1987:LMT:38714.38760} Defines the temporal data model, whose principal time structure is the time sequence. A time sequence is a surrogate value that identifies the object along with a sequence of time-value pairs. There are several types of time sequences depending on the semantics of the data represented.

Clifford-2: \cite{Clifford:1987:HRD:645472.653241},\cite{Clifford:1985:AHR:971699.318922} This model is a refinement of the previous one presented by the same author. The data model allows two types of objects: a set of chronons, termed a lifespan, and a valid-time relation in which each attribute and tuple is assigned a lifespan. A tuple is an ordered value containing the tuple value and its lifespan. Attributes are not atomic since an attribute value in a tuple is a partial function from the chronons' domain to the value domain of the attribute. Relations have key attributes and, at the same chronon, no two tuples are allowed to match the key attributes.

Tansel: \cite{Tansel:1986:ATD:23125.23132},\cite{Clifford:1985:AHR:971699.318922} Designed to support the calculus-based query language (Hquel) and the time by example language. The model supports only one type of object: the valid-time relation. Four types of attributes are supported. If we take the time into account, attributes can be time-varying or non-time-varying. Then, attributes can be atomic-valued or set-valued. There is no need for attributes in the relation to be of the same type, and attribute values in a given tuple do not need to be homogeneous. A triplet containing an element from the attribute's value domain and the boundary points of the time interval represents the value of a time-varying atomic-valued attribute. A set-valued attribute is represented by a set of these triplets.

Gadia-1:  \cite{Gadia:1988:HRM:49346.50065},\cite{Gadia:1985:QLH:325405.325412} This homogeneous model allows two types of objects: valid-time elements and valid-time relations. The model requires that all attribute values in a given tuple be functions on the same valid-time element. Unlike intervals, valid-time elements are closed under union, difference, and complementation.
Bhargava presents an extension of this model to both valid and transaction time.

Gadia-2: \cite{Gadia:1986:TMM:645471.655410},\cite{Chuen},\cite{Gadia:1988:GMR:971701.50233} Is an extension of the homogeneous model known as multihomogeneous model. Temporal elements may be multi-dimensional to model both valid and transaction time. Attribute values are functions from temporal elements onto attribute value domains, but attribute values do not need to be on the same temporal element. The key attributes in the relations must be single-valued in respect to the interval of validity.

Lorentzos: \cite{springerlink:10.1007/3-540-19074-0_71},\cite{Lorentzos:1989:HDT:70777.70787} The temporal relational model allows to specify different granularities and the support of periodic events. This model associates timestamps with individual attribute values rather than with tuples. Timestamps are explicit values updated directly by the user. The said timestamps represent either the chronon during which attribute(s) were valid or a boundary point of the interval of validity. In the model, several timestamps of different granularities can be used together in the specification of a chronon.
This model has the particularity that some columns have a different meaning depending on the context.

\paragraph{Transaction-time models}
The main feature of the transaction-time models is that they are append-only.

Kimball: \cite{kim78} The temporal model is called DATA. In it, the association between facts and times is implicit. Also, the update operations avoid the explicit use of time. However, transaction-time relations cannot be displayed but can only be depicted in the snapshots extracted from the transaction-time relations. At query language level, the association of facts and times is also implicit.

Stonebraker: \cite{Ston87} Proposes the Postgres data model. Like in the previous model, the association between time and facts in the three model features (update language, query language and visualization) is implicit. In this model, the visualization is not restricted to snapshot relation states.
In this model, two timestamps are used when specifying the time for a given tuple that is current in the relation.

Jensen: \cite{Jensen:1991:IIM:627283.627484} In the DM/T model, the association of facts and times is also implicit. DM/T contains a backlog for each user-defined transaction-time relation. This backlog contains the full history of the associated user-defined transaction-time relation. When an insert, delete or update operation is performed, the backlog for this relation is updated.

\paragraph{Bitemporal data models}
The bitemporal data models support both valid and transaction time. This section deals with the following seven models of a bitemporal data model.

Ben-Zvi: \cite{910377} Proposed the time relational model,, which defines two types of objects: snapshot relations and bitemporal relations. Each tuple in a bitemporal relation has five additional implicit attributes:
The two interval ends of the valid-time chronon, called 'Effective-time-start' and 'Effective-time-stop'.
The two interval ends of the transaction-time chronon. 'Registration-time-start' is the transaction-time chronon associated with the 'Effective-time-start', 'Registration-time-stop' is the transaction-time chronon associated with the 'Effective-time-stop'.
Deletion-time. It registers when a tuple which was mistakenly entered is logically deleted.

Ahn: \cite{Snodgrass:1985:TTD:318898.318921},\cite{sno86} This model follows a four-dimensional data model. Each instance is represented as a sequence where each value is composed by a timestamp with a transaction-time chronon and a three dimensional volumes -one of these dimensions being the valid time.

Snodgrass:  \cite{sno86},\cite{Snodgrass:1984:TQL:588011.588041} The data model has four implicit attributes: 
The transaction time when the tuple was inserted.
The transaction time when the tuple was deleted.
The starting point of the valid-time interval.
The end point of the valid time-interval.

	The data query language associated with the model is called TQuel.

McKenzie: \cite{Snodgrass:1993:ATQ:642875.642890},\cite{Mckenzie:1988:ALQ:915060}, \cite{mck81} In this model, the attribute's values are timestamped, but it is required that attributes be single-valued. This makes it possible to perform a Cartesian product. There are two types of objects in the model: the snapshot relations and the valid-time relations. Thus, transaction-time relations are modeled as sequences of snapshot relations. Bitemporal relations are sequences of valid-time relations indexed by transaction time. An attribute's value in a valid-time relation is an ordered pair: a value from the attribute's domain and a set of chronons. Relations are not allowed to have value-equivalent tuples. The model also allows the schema and even the class of the relation (snapshot, transaction-time, valid-time and bitemporal) to change.

Gadia-3: \cite{Bhargava:1993:RDS:642811.642819},\cite{gad92} This data model is associated with the calculus-based query language TempSQL. Attributes are timestamped with finite unions of rectangles in the valid-time x transaction-time space.

Thompson: \cite{171833} This model, based on an accounting system, is called Accounting Data Model. It has four timestamp attributes: two for transaction time, called 'engineering time', and two for valid time, called 'accounting time'. Also, a boolean value called 'timewarp' enables the attribute to change historically.

### 2011-11-09 trunk/book/sources/temporalDatabases.tex
First version

--- 2011-11-09 trunk/book/sources/prelim.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
-% Preliminaries
+% Temporal databases
 %
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-In this section we will introduce some background concepts about possibility theory and temporal databases.
-Subsection \ref{subsec:poss} is an introduction of possibility theory.
-The concepts defined by Dubois and Prade~\cite{Dubois1983} about fuzzy numbers and fuzzy intervals are also explained.
-Temporal databases and their characteristics are introduced in subsection \ref{subsec:temporal}.
-
-\subsection{\label{subsec:poss}Possibilistic variables}
-Informally, a theory of uncertainty considers a set of uncertain events and tries to quantify the confidence about the occurence of those events.
-Consider therefore a set of outcomes $\Omega$ and let each subset of $\Omega$ be called an \emph{event}.
-Hereby, $\Omega$ is called the \emph{certain event} and can thus be regarded as a point of reference.
-Similarly, $\emptyset$ is called the \emph{impossible event}.
-The quantification of confidence in a theory of uncertainty is achieved by means of a \emph{confidence measure}~\cite{Choquet53}.
-\begin{definition}
-\label{def:conf-measure}
-Consider a universe of outcomes $\Omega$. A confidence measure is defined by a function:
-\begin{equation}
-g:\Pow(\Omega)\rightarrow[0,1]
-\end{equation}
-that satisfies:
-\begin{eqnarray}
-g\left(\emptyset\right)&=&0\\
-g\left(\Omega\right)&=&1\\
-A\subseteq B&\Rightarrow&g(A)\leq g(B)
-\end{eqnarray}
-and where $\Pow(\Omega)$ represents the power set of $\Omega$.
-\end{definition}
-From the constraints in Definition \ref{def:conf-measure}, it can be seen that $\emptyset$ and $\Omega$ are indeed respectively the impossible and the certain event.
-In addition, the confidence that is assigned to events is monotonic.
-Due to this monotonicity, we have that:
-\begin{eqnarray}
-\forall A\in\Omega:\forall B\in\Omega&:&g\left(A\cup B\right)\geq\max\left(g\left(A\right),g\left(B\right)\right)\\
-\forall A\in\Omega:\forall B\in\Omega&:&g\left(A\cap B\right)\leq\min\left(g\left(A\right),g\left(B\right)\right).
-\end{eqnarray}
-We can thus derive two extreme measures of confidence and these measures play a crucial role in the theory of possibility~\cite{Dubois88}.
-Consider a confidence measure $\Pi$ that satisfies the following constraint:
-\begin{equation}
-\label{EQ:MAXITIVITY}
-\forall A\subseteq\Omega:\forall B\subseteq\Omega:\Pi\left(A\cup B\right) = \max\left(\Pi(A),\Pi(B)\right).
-\end{equation}
-The function $\Pi$ is then called a \emph{possibility measure} and due to \eqref{EQ:MAXITIVITY}, it is said that $\Pi$ is \emph{maxitive}~\cite{Dubois88}.
-For each event $A\subseteq\Omega$, $\Pi(A)$ expresses the possibility that $A$ occurs.
-Due to the maxitivity of $\Pi$ and under the constraint that $\Omega$ is finite, we can easily extract information about the possibility of separate outcomes in the form of a \emph{possibility distribution} $\pi$ defined by:
-\begin{equation}
-\pi:\Omega\rightarrow[0,1]:\pi(u)=\Pi\left(\left\{u\right\}\right).
-\end{equation}
-Equivalently, we have that:
-\begin{equation}
-\forall A\subseteq\Omega:\Pi(A)=\sup_{u\in A}\pi (u).
-\end{equation}
-A possibility distribution is normalized in the sense that:
-\begin{equation}
-\sup_{u\in\Omega}\pi(u)=1.
-\end{equation}
-In order to ensure the existence of a possibility distribution when $\Omega$ is infinite, the maxitivity constraint \eqref{EQ:MAXITIVITY} should be replaced by:
-\begin{equation}
-\Pi\left(\bigcup_{i}A_i\right)=\sup_{i}\Pi\left(A_i\right)
-\end{equation}
-for any family $\left\{A_i\right\}$ of indexed subsets of $U$~\cite{Nguyen79}.
-
-Let us now consider a confidence measure $\N$ that satisfies:
-\begin{equation}
-\label{EQ:MINITIVITY}
-\forall A\subseteq\Omega:\forall B\subseteq\Omega:\N\left(A\cap B\right) = \min\left(\Pi(A),\Pi(B)\right).
-\end{equation}
-The function $\N$ is then called a \emph{necessity measure} and due to \eqref{EQ:MINITIVITY}, it is said that $\N$ is \emph{minitive}.
-Possibility measures and necessity measures are dual in the sense that:
-\begin{equation}
-\forall A\subseteq\Omega:\Pi(A)=1-\N(\overline{A}).
-\end{equation}
-The principles of possibility theory are often used in the scope of a \emph{possibilistic variable}, also known as an \emph{ill-known value}.
-\begin{definition}
-Consider a universe $U$.
-A possibilistic variable $X$ over $U$ is a variable taking exactly one value in $U$, but for which this value is (partially) unknown.
-The available knowledge about the value that $X$ takes, is given by a possibility distribution $\pi_X$.
-For each $u\in U$, $\pi_X(u)$ represents the possibility that $X$ takes the value $u$.
-\end{definition}
-A specific application of possibility theory~\cite{Prade82,VanSchooten,DeCooman} is obtained when the universe is the set of Boolean values, i.e.
-$U=\mathbb{B}=\{T,F\}$.
-If we consider a Boolean proposition, then this proposition is either true ($T$) or false ($F$).
-Any Boolean proposition $p$ can thus be regarded as a possibilistic variable for which the possibility and necessity that $p=T$ are denoted in short as $\Pos(p)$ and $\Nec(p)$.
-The following notations are used:
-\begin{eqnarray}
-\Pos(p)&=&\pi_p(T)\\
-\Nec(p)&=&1-\pi_p(F).
-\end{eqnarray}
-Keeping in mind the cause of uncertainty that is assumed, it is important to understand that a possibilistic variable $X$ is bounded to take one value only, but this value is not known due to incomplete knowledge.
-When a possibilistic variable is defined over the universe $\Pow(U)$, it is said to be a \emph{possibilistic multi-valued variable}, also known as an \emph{ill-known set}~\cite{Dubois88b}.
-Note that an ill-known set conceptually differs from a fuzzy set in the sense that an ill-known set is a crisp set, but for some reason (partially) unknown.
-\subsection{\label{subsubsec:fuzzy}Fuzzy numbers and fuzzy intervals}
-When we shall discuss the case of fuzzy temporal databases, we shall link the concepts introduced in this paper to the theory of fuzzy intervals.
-Dubois and Prade~\cite{Dubois1983} proposed the following definition of \emph{fuzzy interval}.
-\begin{definition}
-A fuzzy interval is a fuzzy set $M$ on the set of real numbers $\mathbb{R}$ such that:
-\begin{eqnarray}
-\forall (u,v)\in\mathbb{R}^2:\forall w \in [u,v]:&\mu_M(w) \geq\min(\mu_M(u),\mu_M(v)) \ &\mbox{ (convexity) } \\
-\exists m \in \mathbb{R}:& \mu_M(m)=1 \ &\mbox{ (normalization) }
-\end{eqnarray}
-\end{definition}
-If $m$ is unique, then $M$ is referred to as a \emph{fuzzy number}, instead of a \emph{fuzzy interval}.
-In other words, if the core of a fuzzy interval is a singleton, it can be seen as a fuzzy number.
-In their work, Dubois and Prade propose four different functions (two possibility and two necessity functions) to asses the position of a fuzzy number N relative to that of a fuzzy number M taken as a reference.
-
-The most convenient representation for the membership function of a fuzzy number is a triangular function.
-The membership function $\mu_M$ for the fuzzy set $M$ has also the properties of convexity and normalization.
-Three values represent a triangular function (see Figure~\ref{fig:triangular}):
-\begin{itemize}
-\item
-$D$ is the singleton value in the core of $\mu_M(x)$.
-\item
-$D-a$ is the lower value in the support of $\mu_M(x)$. 
-\item
-$D+b$ is the upper value in the support of $\mu_M(x)$.
-\end{itemize}
-Note that the values $a$ and $b$ are values in the underlying ordered domain. E.g. $(a,b) \in \mathbb{R}^2$. The notation for a triangular function adopted here is $[D,a,b]$.
-\begin{figure}[h!]
-  \centering
-  \includegraphics[scale=0.4]{graphs/triangular.eps}
-  \caption{Triangular function.}
-  \label{fig:triangular}
-\end{figure}
-
-The most simple representation for the membership function of a fuzzy interval is a trapezoidal function.
-This membership function $\mu_T$ for the fuzzy interval $T$ is convex and normalized.
-Four values represent a trapezoidal membership function (figure  \ref{fig:trapezoidal}): $\left[\alpha,\ \beta,\ \gamma,\ \delta\right]$.
-The membership function is given by:
-
-\begin{align}
-\mu_T(x)
-\begin{cases}
-1 & \mbox{ if } x \in [\beta,\gamma] \\
-0 & \mbox{ if } x > \delta \vee x < \alpha \\
-\frac{x-\alpha}{\beta - \alpha} & \mbox{ if } x \in [\alpha,\beta[ \\
-\frac{\delta -x}{\delta - \gamma} & \mbox{ if } x \in ]\gamma,\delta] \\
-\end{cases}
-\end{align}
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[scale=0.4]{graphs/trapezoidalDistribution.eps}
-  \caption{Trapezoidal distribution.}
-  \label{fig:trapezoidal}
-\end{figure}
-
-\subsection{\label{subsec:temporal}Temporal databases}
-
-\subsubsection{\label{subsubsec:timeDomain}Time Domain}
-The semantics of time and the different approaches to the time model are analyzed in this section; a more detailed explanation is discussed in \ref{sec:timeindatabasedesign}.
-
-The general time model represents time as a set with an imposed partial order.
-There are two main models of time: linear and cyclic.
-In the linear model, time advances from the past to the future in a totally ordered fashion (a relationship of total order is imposed to the set).
-A cyclic model of time is used for recurrent processes.
-The present proposal focuses on the linear model.
-\begin{itemize}
-\item
-Discrete model:  The discrete model of time is isomorphic in relation to the natural numbers.
-According to this model, each natural number corresponds to an indivisible time unit.
-The said time unit is called a chronon or, in other words, a 'time quantum', 'moment' or 'instant'.
-A chronon is the smallest time unit that can be represented by the model.
-\item
-Dense model: This model is isomorphic in relation to the rational or the real numbers. It represents one instant of time in the gap between another two instants.
-\item
-Continuous model: This is an extension of the dense model with no gaps. It is isomorphic in relation to the real numbers.
-\end{itemize}
-Restrictions on time range may be performed on the base of the following two concepts:
-\begin{itemize}
-\item
-Time boundedness: Time can be bounded orthogonally in the past and in the future.
-\item
-Time distance: As a metric, time has a distance function which presents four properties:
-\begin{enumerate}
-\item
-Distance is non-negative.
-\item
-The distance between two different elements is not equal to zero.
-\item
-The distance from $\alpha$ to $\beta$ is the same as that between $\beta$ and $\alpha$.
-\item
-Triangle inequality: The distance from $\alpha$ to $\gamma$ is equal or greater than the distance from $\alpha$ to $\beta$ plus the distance from $\beta$ to $\gamma$.
-\end{enumerate}
-\end{itemize}
-On the other hand, time may be relative or absolute (in other words, anchored or unanchored).
-Sometimes, what one considers 'absolute time' is not as definite as one would hope since our concept of absolute time is based on another time reference (for example, January 1st of year one).
-Relative time has a direction, differently from distance.
-It is also possible to use negative references so as to represent relative time (e.g.
--3 days = three days ago).
-
-\subsubsection{Time in database design}
-\label{sec:timeindatabasedesign}
 This section defines the model used to represent time in a database.

### 2011-11-09 trunk/book/sources/timeDomain.tex
First version

%
% Time domain
%
In order to define and to work with time it is necessary to study and understand the underlying domain. Moreover, the definition of a temporal domain is the basis for almost any temporal system. In the following subsection \ref{subsec:basic-concepts} we will introduce several concepts that are widely used in the community of temporal databases. Most of the concepts explained here have been introduced in the \emph{'Consensus Glossary of Temporal Database Concepts'}~\cite{Dyreson:1994:CGT:181550.181560}.

\subsection{Basic concepts and properties}
\label{subsec:basic-concepts}

Time is usually represented as a set with an imposed partial order. There are two main models of time: \textbf{linear} and \textbf{cyclic}. In the linear model, time advances from the past to the future in a totally ordered fashion (a relationship of total order is imposed to the set). A cyclic model of time is used for recurrent processes. The majority of the proposals work with a linear model of time.

\begin{svgraybox}
The basic elements in a temporal domain are the following:\\
%\begin{itemize}
%\item
\textbf{Instant}:  Is a time point over an underlying temporal domain.\\
%\item
\textbf{Time interval}: Is the time between two instants.\\
%\item
\textbf{Event}: Is an instantaneous fact. Something that happens at an instant.\\
%\item
\textbf{Chronon}: Is a non-decomposable time interval of fixed duration. It is possible for a temporal model to leave the chronon duration unspecified. The duration will be specified later in the implementation of the model.
%\end{itemize}
\end{svgraybox}

Attending to the density between the chronons, a model can be classified into the following three types:

\begin{itemize}
\item
\textbf{Discrete model}:  The discrete model of time is isomorphic in relation to the natural numbers. According to this model, each natural number corresponds to a \emph{chronon}. 
\item
\textbf{Dense model}: This model is isomorphic in relation to the rational or the real numbers. It represents one instant of time in the gap between another two instants.
\item
\textbf{Continuous model}: This is an extension of the dense model with no gaps. It is isomorphic in relation to the real numbers.
\end{itemize}

Restrictions on time range may be performed on the base of the following two concepts:
\begin{itemize}
\item
\textbf{Time boundedness}: Time can be bounded orthogonally in the past and in the future.
\item
\textbf{Time distance}: As a metric, time has a distance function which presents four properties:
\begin{enumerate}
\item
Distance is non-negative.
\item
The distance between two different elements is not equal to zero.
\item
The distance from $\alpha$ to $\beta$ is the same as that between $\beta$ and $\alpha$.
\item
Triangle inequality: The distance from $\alpha$ to $\gamma$ is equal or greater than the distance from $\alpha$ to $\beta$ plus the distance from $\beta$ to $\gamma$.
\end{enumerate}
\end{itemize}

On the other hand, time may be \textbf{relative} or \textbf{absolute} (in other words, \textbf{anchored} or \textbf{unanchored}). Sometimes, what one considers 'absolute time' is not as definite as one would hope since our concept of absolute time is based on another time reference (for example, January 1st of year one). Relative time has a direction, differently from distance. It is also possible to use negative references so as to represent relative time (e.g. -3 days = three days ago).

%%% time as points and time as intervals:

%%% granularity

\subsection{Temporal Relations}
%an explanation on allen's temporal relations.

\subsection{An example of time domain: Julian Day Number}
% proposal for the underlying Julian Day Number domain.

The Julian Day number \cite{Dir96} is a counter. Its value is incremented in one unit every day from 1 January 4713 B.C. at 12:00 noon. The particularity of starting at noon was useful for astronomers: the observations they took one night belonged to the same Julian Day. Actually it has not any transcendence.\\
Note that the Julian Day represents whole days. There is an extension that allows to represent any precision needed (it is called Julian Date). By default, a Julian Day is expressed in Universal Time. (U.T, also known as Solar Time). However, there are representations in Terrestrial Time (T.T), Epheris Time (J.E.D. or J.D.E.). Any time scale different from Universal Time must be explicited after the Julian Day number.\\

\subsubsection{Representation alternatives}
There are many alternatives to optimize the representation of Julian Day numbers because of its extremely far origin (4713 B.C. year). Table \ref{table:juliandayrepresentations} shows several time domains that can be calculated from the Julian Date, some of them are proposed just for optimization purposes.

\begin{table}
\caption{Julian Day representations}
\label{table:juliandayrepresentations}

\begin{tabular}{p{2cm}p{4cm}p{4cm}p{2cm}}
%header ------------------------------------------------------
\hline\noalign{\smallskip}
Name & From & Formula & Current Value  \\ 
\noalign{\smallskip}\svhline\noalign{\smallskip}
%header ------------------------------------------------------
Julian Date (JD)$^a$ & 12:00 noon Monday 1 January 4713 B.C & & 2455278. 85488  \\ 
Julian Day Number (JDN)$^b$ & 12:00 noon  Monday 1 January 4713 B.C. & JND = floor(JD) & 2455278 \\ 
Reduced Julian Day (RJD)$^c$ & 12:00 noon Tuesday 16 November 1858 A.C. & RJD = JD - 2400000 & 55278.85488  \\ 
Modified Julian Day (MJD)$^d$ & 00:00 Wednesday 17 November 1858 A.C. & MJD = JD - 2400000,5 & 55278.35488 \\ 
Truncated Julian Day (TJD)$^e$ & 00:00 Friday 24 May 1968 A.C. & TJD = JD - 2440000,5 & 15278.35488  \\ 
Truncated Julian Day (TJD)$^f$ & 00:00 Thursday 10 November 1995 A.C. & TJD = (JD- 0,5) mod 10000 & 5278. 35488   \\ 
Dublin Julian Day (DJD))$^g$ & 12:00 noon Sunday 31 December 1899 A.C. & DJD = JD - 2415020 & 40258. 85488 \\ 
Chronological Julian Day (CJD)$^h$  & 00:00 Monday 1 January 4713 B.C. & CJD = JD + 0,5 + timezone adjust. & 2455279. 3548843 (UT)  \\ 
Lilian Day Number$^i$ & Friday 15 October 1582 & floor(JD-2299160,5) & 156118 \\ 
ANSI Date$^j$  & Monday 1 January 1601 & floor(JD-2305812,5) & 149466  \\ 
Rata die$^k$  & Monday 1 January 1 A.C & floor(JD - 1721424.5) & 733854 \\ 
Unix time$^l$  & Thursday 1 January 1970 A.C. & (JD – 2440587.5) × 86400 & 1269333062 \\ 
\noalign{\smallskip}\hline\noalign{\smallskip}
\end{tabular}
$^a$ This is an extension of the Julian Day that allows time representation. \\
$^b$  Each day changes at noon. \\
$^c$  Used by astronomers. \\
$^d$ It starts at midnight. \\
$^e$ Definition from NASA. \cite{Sch}. \\
$^f$ Definition from NIST. \cite{Nis}. \\
$g$ Introduced by the IAU in 1995. \\
$^h$ The timezone must be explicited. Each day changes at midnight. \\
$^i$ The number of days since Gregorian calendar in Universal Time. \\
$^j$ The origin for COBOL integer dates. \\
$^k$  The number of days since actual era. \\
$^l$  It counts the seconds not the day. \\
\end{table}

\subsubsection{Conversion among dates}
\paragraph{Conversion between Gregorian and Julian dates}
The Julian Date consist on the Julian Day number and the time. It is computed with the formula \cite{Usn}:\\
\begin{equation}
\label{eq:julian day}
JD = JDN + \frac{hour-12}{24} + \frac{minutes}{1440} + \frac{seconds}{86400}
\end{equation}

Here the JDN is the Julian Day Number. This equation does not take into account the leap seconds \footnote{A leap second(s) is the amount of second(s) added or deleted every year in the Universal Time.\cite{Wik},\cite{Lea}. }.

\paragraph{Gregorian calendar to Julian day}
Any date from Gregorian calendar \footnote{As Gregorian calendar entered in use in 1582, any use before that date is called \emph{proleptic calendar}.} can be converted into a Julian Day using the following formula:\\
\begin{equation}
\label{eq:JDN}
JDN = \frac{1461x(Y + 4800 + \frac{M-14}{12})}{4}+\frac{367x(M-2-12x\frac{M-14}{12})}{12}-\frac{3x\frac{Y+4900+\frac{M-14}{12}}{100}}{4} 
\end{equation}

This formula was created by Henry Fliegel and Thomas C. van Flandern in 1968. \cite{Fliegel:1968:LEM:364096.364097} It takes into account leap years and that they are applicable only when the century is divisible by 400. The value $Y$ represents the year and the value $M$ represents the month.
\paragraph{Julian calendar to Julian day}
The algorithm to convert between Julian calendar and Julian Day Number is only valid for dates since 4713 B.C (positive values of Julian Day) \cite{Dog}:\\
\begin{equation}
\label{eq:JDN_2_JD}
JDN=367Y-\frac{7 (Y + 5001 + \frac{M-9}{7})}{4}+\frac{275xM}{9}+D+1729777
\end{equation}

Where $Y$ is the year value, $M$ is the month value, and $D$ is the day value.

%\subsubsection{Membership degree computation}
%The main advantage from the Julian day number is the ability to make calculations. Thus, the membership degree computation for a fuzzy number based on a Julian day is explained here. \\
%Let a trapezoidal possibility distribution defined by four JDN named $a$,$b$,$c$ and $d$. The fuzzy validity period \emph{FVP}, can be represented as in figure \ref{fig:fuzzy-validity-period}. The following equation shows the membership degree for a Julian day number \emph{JDN} in a \emph{FVP} represented by means of the possibility distribution of $[a,b,c,d]$. In a valid-time database, the values for the valid starting time \emph{VST} and the valid ending time \emph{VET} can be fuzzyfied as shown in equation \ref{eq:fuzzy-validity-period}.
%
%The membership function for a given $x$ number in \emph{JDN} format is the following:\\
%\begin{equation}
%\label{eq:trapezoidal distribution}
%%\[
%T(x)=
%\begin{cases}
%0, \mbox(if x \leq a, x \geq d)\\
%1, \mbox(if x \in [b,c])\\
%\frac{x-a}{b-a}   \mbox(if x \in (a,b])\\
%\frac{d-x}{d-c} 	 \mbox(if x \in (c,d])\\
%\end{cases}
%%\]
%\end{equation}
%The points outside the trapezium or inside the interval $[a,b]$ are trivial. The points inside the other intervals has the following particularities:\\
%The subtraction operation with Julian day numbers gives as result a scalar value (as explained before). The division operation between two scalars gives as result an scalar in the interval $[0,1]$ which represents the membership degree.

### 2011-11-09 trunk/ipmu2012/preliminaries.tex
First version

\subsection{Possibilistic Variables}
A possibilistic variable $X$ over a universe $U$ is defined as a variable taking exactly one value in $U$, but for which this value is (partially) unknown. The possibility distribution $\pi_X$ gives the available knowledge about the value that $X$ takes. For each $u\in U$, $\pi_X(u)$ represents the possibility that $X$ takes the value $u$.

It is important to understand the difference between the following two concepts:
\begin{itemize}
\item
A \emph{possibilistic variable} $X$ is bounded to take only one value , but this value is not known due to incomplete knowledge. 
\item
An \emph{ill-known set}~\cite{Dubois88b}: a possibilistic variable defined over the universe $\Pow(U)$.
\end{itemize}

Note that while a possibilistic variable refers to one (partially) unknown value, an ill-known set is a crisp set but, for some reason, (partially) unknown.

\subsection{\label{subsec:fuzzy-numbers}Fuzzy numbers and fuzzy intervals}
Dubois and Prade~\cite{Dubois1983} proposed the following definition of \emph{fuzzy interval}.
%\begin{definition}
A fuzzy interval is a fuzzy set $M$ on the set of real numbers $\mathbb{R}$ such that:
\begin{eqnarray}
\forall (u,v)\in\mathbb{R}^2:&\\
\nonumber
\forall w \in [u,v]:&\mu_M(w) \geq\min(\mu_M(u),\mu_M(v))  \\
\exists m \in \mathbb{R}:& \mu_M(m)=1 
\end{eqnarray}
%\end{definition}
If $m$ is unique, then $M$ is referred to as a \emph{fuzzy number}, instead of a \emph{fuzzy interval}. In other words, if the core of a fuzzy interval is a singleton, it can be seen as a fuzzy number. In their work, Dubois and Prade propose four different functions (two possibility and two necessity functions) to asses the position of a fuzzy number N relative to  a fuzzy number M taken as a reference.

The most convenient representation for the membership function of a fuzzy number is a triangular function. The membership function $\mu_M$ for the fuzzy set $M$ has also the properties of convexity and normalization. Three values represent a triangular function (see Figure~\ref{fig:triangular}):
\begin{itemize}
\item
$D$ is the singleton value in the core of $\mu_M(x)$.
\item
$D-a$ is the lower value in the support of $\mu_M(x)$. 
\item
$D+b$ is the upper value in the support of $\mu_M(x)$.
\end{itemize}
Note that the values $a$ and $b$ are values in the underlying ordered domain. E.g. $(a,b) \in \mathbb{R}^2$. The notation for a triangular function adopted here is $[D,a,b]$.
\begin{figure}[h!]
  \centering
  \includegraphics[scale=0.4]{graphs/triangular.eps}
  \caption{Triangular function.}
  \label{fig:triangular}
\end{figure}

The most simple representation for the membership function of a fuzzy interval is a trapezoidal function. This membership function $\mu_T$ for the fuzzy interval $T$ is convex and normalized. Four values represent a trapezoidal membership function (figure  \ref{fig:trapezoidal}): $\left[\alpha,\ \beta,\ \gamma,\ \delta\right]$. The membership function is given by:

\begin{align}
\mu_T(x)
\begin{cases}
1 & \mbox{ if } x \in [\beta,\gamma] \\
0 & \mbox{ if } x > \delta \vee x < \alpha \\
\frac{x-\alpha}{\beta - \alpha} & \mbox{ if } x \in [\alpha,\beta[ \\
\frac{\delta -x}{\delta - \gamma} & \mbox{ if } x \in ]\gamma,\delta] \\
\end{cases}
\end{align}

\begin{figure}[h!]
  \centering
  \includegraphics[scale=0.4]{graphs/trapezoidalDistribution.eps}
  \caption{Trapezoidal distribution.}
  \label{fig:trapezoidal}
\end{figure}

\subsubsection{Set evaluation by ill-known constraints}

\subsubsection{Interval evaluation by ill-known constraints}
The problem of the interval evaluation is explained in \cite{Pon11}: For a crisp interval $I = \left[ a, b \right]$, we want to know if all points in this interval reside between the boundaries of the ill-known interval $\left[ X , Y \right]$, where $X,Y$ are ill-known values on the set of real numbers $\mathbb{R}$, and $\lambda$ is the evaluation function. The following expressions compute the possibility and the necessity:

\begin{eqnarray}
\label{eq:interval-pos}
\Pos\left(\lambda([a,b])\right)=\\
\nonumber
\min\bigg(\sup_{a\geq w}\pi_{X}(w),\sup_{b\leq w}\pi_{Y}(w)\bigg)\\
\label{eq:interval-nec}
\Nec\left(\lambda([a,b])\right)=\\
\nonumber
\min\bigg(\inf_{a<w}1-\pi_{X}(w),\inf_{b>w}1-\pi_{Y}(w)\bigg).
\end{eqnarray}

\paragraph{Example} Consider the ill-known values $X = \left[5, 2, 8\right]$ and $Y = \left[9, 7, 10 \right]$. The knowledge about the evaluation of the interval $\left[a, b \right]$  is given by the expressions \eqref{eq:interval-pos},\eqref{eq:interval-nec}.  Figure~\ref{fig:3d-possibility} shows a 3D plot of the possibility that an interval $[a,b]$ passes the evaluations specified by the ill-known constraints. Note the triangular form for the resulting possibility distribution since the condition $a \leq b$ holds.

\begin{figure}[h!]
\centering
\includegraphics[scale=0.4]{graphs/3D_possibility.eps}
\caption{Possibility of evaluation for the interval $[a,b]$.}
\label{fig:3d-possibility}
\end{figure}
The necessity plot is obtained in a similar way and is shown in Figure~\ref{fig:3d-necessity}. Notice that the necessity measure is not normalized because the supports of $X$ and $Y$ overlap.
\begin{figure}[h!]
\centering
\includegraphics[scale=0.4]{graphs/3D_necessity.eps}
\caption{Necessity of evaluation for the interval $[a,b]$.}
\label{fig:3d-necessity}
\end{figure}

\subsection{Temporal databases}
%In this section, the proposed reasoning is applied to the specific context of intervals on the real line. This setting is of specific interest in the context of fuzzy temporal databases.
A temporal database \cite{Dyreson1994} is a database that manages some aspects of time in its schema. A \textbf{chronon} is the shortest duration of time supported by the database. The time can be represented either as points or intervals \cite{655777}. Fuzzy temporal models \cite{4481150} are proposed when the time points or intervals are not precisely known. There are proposals for the  fuzzyfication of the time point \cite{Dubois89} and the fuzzyfication of the time interval \cite{Garrido2009}.\\
Allen \cite{Allen:1983:MKT:182.358434} studied the comparison between two crisp time intervals. For fuzzy intervals, several proposals \cite{4481150}, \cite{springerlink:10.1007/978-3-540-39964-3_57},\cite{10.1109/TIME.2004.1314418} have been done.

%Time granularity is also associated with the representation of the time. A granularity is the result of partitioning on the set of chronons. The conversion among granularities is a common issue within temporal databases \cite{Lin97efficientconversion}. Granularity is the basis of some systems \cite{Cru97},\cite{624013}.

Despite of user-defined time (an uninterpreted attribute; supported in the standard SQL2 \cite{Mel93}), there are 3 types of time in a temporal database:

\begin{itemize}
	\item
	\textbf{Transaction time}: The time when the fact is stored in the database.
	\item
	\textbf{Valid time}: The time when the fact is true in the modelled reality.
	\item
	\textbf{Decision time} \cite{Nascimento95decisiontime}: The time when an event was decided to happen. 
	\end{itemize}

	 The  database model can be then classified into: transaction time \cite{Ston87},\cite{Jensen:1991:IIM:627283.627484}, Valid time, bi-temporal \cite{Snodgrass:1984:TQL:588011.588041}(both valid and transaction time) or tri-temporal \cite{Nascimento95decisiontime} (valid, transaction and decision time).

Fuzzy temporal models \cite{4481150} deal with time points \cite{Dubois89} or intervals \cite{Garrido2009} that are ill-known. Some fuzzy temporal models assume that the time stored in the database is an interval. The temporal interval is represented by two ill-known time points: $X$  an ill-known starting point and $Y$ an ill-known ending point. The interval $\left[X,Y\right]$ is not a fuzzy interval but an ill-known interval: it is a crisp interval but it is partially unknown which values are in this interval.

### 2011-11-09 book/sources/intro.tex
First version

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Introduction
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In Section \ref{sec:time-domain} an study about the time and its properties is done. Section \ref{sec:temporal-databases} is an overview of the main problems when managing the time in a database, the different proposals for solving them and the temporal database proposals. The section concludes with a comparison among the different proposals. Finally, in Section \ref{sec:fuzzy-temporal-databases} we present a sum up with the proposals for handling the imprecision in temporal databases and the main deficiences we have detected.

### 2011-11-09 book/sources/proposal.tex
First version

% Proposal
%
%
The model deals with uncertainty in the valid time. A valid time period  \emph{VT} is specified with two points: a starting time point \emph{St} and an ending time point \emph{Et}. Thus, a fuzzy validity period, \emph{FVP} is a valid time period in which the starting, the ending or both points are not precisely known. More formally we make the following definitions:

\begin{definition}
Valid Time Period: bien
\end{definition}

\subsection{Representation}

\subsubsection{Undelying domain: Julian Day Number}

\subsubsection{Temporal Operators}

\subsection{Data Manipulation Language}

\subsection{Query specification}

\begin{example}

\end{example}

### 2011-11-09 ipmu2012/introduction.tex
First version

There are several proposals in the literature for inferring uncertainty about sets from uncertainty about values. It is of particular interest for fuzzy temporal databases, the application on temporal intervals. The transformation from two ill-known temporal points to a ill-known set has been discussed in \cite{Garrido2009}. In this work we will analyze the major pitfalls of these approaches and the benefit of using the mechanism proposed in \cite{Pon11}.

The paper is organized as follows. In Section \ref{sec:preliminaries}, some preliminaries and concepts about the possibilistic evaluation framework and the temporal databases are presented. 
Section \ref{sec:proposal} presents the possibilistic model for valid-time databases.
Finally, Section \ref{sec:conclusions} presents the main conclusions and some lines for the future work in this research.

### 2011-11-09 ipmu2012/ipmu.tex
First version

\documentclass{llncs}
%
%

\usepackage[dvips]{graphicx}
\usepackage{makeidx}  % allows for indexgeneration
\usepackage{amsmath}
\usepackage{amsfonts}

\hyphenation{pro-per-ties}
\hyphenation{ge-ne-ral-ly}
\hyphenation{pre-fe-ren-ces}
\hyphenation{u-sing}
\hyphenation{pu-nish-ment}

\newcommand{\Pow}{\mathcal{P}}
\newcommand{\N}{\operatorname{N}}
\newcommand{\bool}{\operatorname*{\mathcal{B}}}
\newcommand{\Pos}{\operatorname{Pos}}
\newcommand{\Nec}{\operatorname{Nec}}
\newcommand{\T}{\mathcal{T}}
\begin{document}

\mainmatter              % start of the contributions
%
\title{A possibilistic valid-time model}
%
%\titlerunning{A Fuzzy Valid-Time Model}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{Jos\'e Enrique Pons \and Christophe Billiet \and Olga Pons Capote \and Guy de Tr\'e}
%Jeffrey Dean \and David Grove \and Craig Chambers \and Kim~B.~Bruce \and
%Elsa Bertino}
%
\authorrunning{Jos\'e Enrique Pons et al.} % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
%\tocauthor{Ivar Ekeland, Roger Temam, Jeffrey Dean, David Grove,
%Craig Chambers, Kim B. Bruce, and Elisa Bertino}
%
\institute{
Department of Computer Science and Artificial Intelligence \\
Universidad de Granada \\
Escuela T\'ecnica Superior de Ingenier\'ia Inform\'atica \\
C/Periodista Daniel Saucedo Aranda s/n \\
E-18071 (Granada-Spain)\\
\email{jpons,opc@decsai.ugr.es}\\ 
%WWW home page:
%\texttt{http://users/\homedir iekeland/web/welcome.html}
%\and
%Universit\'{e} de Paris-Sud,
%Laboratoire d'Analyse Num\'{e}rique, B\^{a}timent 425,\\
%F-91405 Orsay Cedex, France
\and 
Department of Telecommunications and Information Processing
Ghent University
Sint Pietersniewstraat 41 B-9000, Gent, Belgium.
\email{Christophe.Billiet,Guy.De.Tre@ugent.be}
}

\maketitle              % typeset the title of the contribution

\begin{abstract}

\end{abstract}

%
\section{Introduction}
\label{sec:Introduction}
\input{introduction}
%

\section{Preliminaries}
\label{sec:preliminaries}
\input{preliminaries}

\section{Proposal}
\label{sec:proposal}
\input{proposal}

%%%%%%%%%%%%%%%%
%
% FURTHER RESEARCH WORK
%
%%%%%%%%%%%%%%%
%
\section{Conclusions and further research work}
\label{sec:futher-research}
\input{conclusions}
%

%
\section*{Acknowledgements}
%
The researchers are supported by the grant BES-2009-013805 within the research project TIN2008-02066: \emph{Fuzzy Temporal Information treatment in relational DBMS}.

\bibliographystyle{spmpsci}
\bibliography{ipmu}

\end{document}

### 2011-11-09 ipmu2012/proposal.tex
First version

\subsection{Representation of ill-known valid-time intervals}

\subsection{Storage of valid-time intervals}

\subsection{Querying ill-known valid-time intervals}

\subsubsection{Query specification}

\subsubsection{Query evaluation}

\subsubsection{Ranking}

### 2011-11-09 ipmu2012/preliminaries.tex
Update with the introduction and changes in the preliminaries and biblio.

--- 2011-11-09 trunk/ipmu2012/preliminaries.tex
@@
 	\item
-	\textbf{Decision time} \cite{Nascimento95decisiontime}: The time when an event was decided to happen. 
+	\textbf{Decision time} \cite{Nascimento95}: The time when an event was decided to happen. 
 	\end{itemize}
 
-	 The  database model can be then classified into: transaction time \cite{Ston87},\cite{Jensen:1991:IIM:627283.627484}, Valid time, bi-temporal \cite{Snodgrass:1984:TQL:588011.588041}(both valid and transaction time) or tri-temporal \cite{Nascimento95decisiontime} (valid, transaction and decision time).
+	 The  database model can be then classified into: transaction time \cite{Ston87},\cite{Jensen91}, Valid time, bi-temporal \cite{Snodgrass84}(both valid and transaction time) or tri-temporal \cite{Nascimento95} (valid, transaction and decision time).
 

### 2011-11-09 trunk/ipmu2012/introduction.tex
Update with the introduction and changes in the preliminaries and biblio.

Dealing with time in any information system~\cite{Bolour82} is usually a difficult task. First of all, the concept of \emph{time} itself is a complex notion. Klein~\cite{klein94} studied the concept of time in language whereas Devos~\cite{devos94} modelled vague temporal expressions by means of fuzzy sets~\cite{zadeh65}.  Several research works~\cite{cruyssen97,devos98,Cru97} have been done in obtaining theoretical models that allow the representation of time in an information system. Although the main tool for the representation of temporal information are fuzzy sets~\cite{Virant199639,35337,343607,springerlink:10.1007/978-3-540-39964-3_57,Dubois:jucs_9_9:fuzziness_and_uncertainty_in}, some other tools, like rough sets~\cite{pawlak95} have been used~\cite{Qia09}.

In the other hand, the addition of time to a relational database leads to several problems: the modelling of a time-variant object has an impact on the database consistency. Different models have been proposed \cite{Jensen94,TSQL,Sarda90,Jensen91,Snodgrass84,Nascimento95} but all of them consider the time as a crisp value. 
There exist a glossary~\cite{Dyreson1994} with the common terms in the temporal database community.
Further works introduced the concept of vagueness, imprecission in the different models~\cite{Cru97,624013,fuzz2009,gal01}.

There are several proposals in the literature for inferring uncertainty about sets from uncertainty about values. It is of particular interest for fuzzy temporal databases, the application on temporal intervals. The transformation from two ill-known temporal points to a ill-known set has been discussed in \cite{Garrido2009}. In this work we will analyze the major pitfalls of these approaches and the benefit of using the mechanism proposed in \cite{Pon11}.

The paper is organized as follows. In Section \ref{sec:preliminaries}, some preliminaries and concepts about the possibilistic evaluation framework and the temporal databases are presented. 
Section \ref{sec:proposal} presents the possibilistic model for valid-time databases.
Finally, Section \ref{sec:conclusions} presents the main conclusions and some lines for the future work in this research.

### 2011-11-10 trunk/ipmu2012/preliminaries.tex
A more detailed introduction.

@@
 The time can be represented either as points or intervals \cite{655777}.
-Fuzzy temporal models \cite{4481150} are proposed when the time points or intervals are not precisely known.
+Fuzzy temporal models \cite{schockaert08} are proposed when the time points or intervals are not precisely known.
 There are proposals for the  fuzzyfication of the time point \cite{Dubois89} and the fuzzyfication of the time interval \cite{Garrido2009}.\\
-Allen \cite{Allen:1983:MKT:182.358434} studied the comparison between two crisp time intervals.
-For fuzzy intervals, several proposals \cite{4481150}, \cite{springerlink:10.1007/978-3-540-39964-3_57},\cite{10.1109/TIME.2004.1314418} have been done.
+Allen \cite{Allen83} studied the comparison between two crisp time intervals. For fuzzy intervals, several proposals \cite{ohlbach04,nagypal03,schockaert08} have been done.
 
@@
 	\item
-	\textbf{Decision time} \cite{Nascimento95decisiontime}: The time when an event was decided to happen. 
+	\textbf{Decision time} \cite{Nascimento95}: The time when an event was decided to happen. 
 	\end{itemize}
 
-	 The  database model can be then classified into: transaction time \cite{Ston87},\cite{Jensen:1991:IIM:627283.627484}, Valid time, bi-temporal \cite{Snodgrass:1984:TQL:588011.588041}(both valid and transaction time) or tri-temporal \cite{Nascimento95decisiontime} (valid, transaction and decision time).
+	 The  database model can be then classified into: transaction time \cite{Ston87},\cite{Jensen91}, Valid time, bi-temporal \cite{Snodgrass84}(both valid and transaction time) or tri-temporal \cite{Nascimento95} (valid, transaction and decision time).
 
-Fuzzy temporal models \cite{4481150} deal with time points \cite{Dubois89} or intervals \cite{Garrido2009} that are ill-known.
+Fuzzy temporal models \cite{schockaert08} deal with time points \cite{Dubois89} or intervals \cite{Garrido2009} that are ill-known.
 Some fuzzy temporal models assume that the time stored in the database is an interval.

### 2011-11-10 ipmu2012/introduction.tex
A more detailed introduction.

@@
-There are several proposals in the literature for inferring uncertainty about sets from uncertainty about values.
-It is of particular interest for fuzzy temporal databases, the application on temporal intervals.
-The transformation from two ill-known temporal points to a ill-known set has been discussed in \cite{Garrido2009}.
-In this work we will analyze the major pitfalls of these approaches and the benefit of using the mechanism proposed in \cite{Pon11}.
+Dealing with time in any information system~\cite{Bolour82} is usually a difficult task.
+First of all, the concept of \emph{time} itself is a complex notion.
+Klein~\cite{klein94} studied the concept of time in language whereas Devos~\cite{devos94} modelled vague temporal expressions by means of fuzzy sets~\cite{zadeh65}.
+Several research works~\cite{cruyssen97,devos98,Cru97} have been done in obtaining theoretical models that allow the representation of time in an information system.
+Although the main tool for the representation of temporal information are fuzzy sets~\cite{Virant199639,35337,343607,nagypal03,Dubois:jucs_9_9:fuzziness_and_uncertainty_in}, some other tools, like rough sets~\cite{pawlak95} have been used~\cite{Qia09}.
+%granularity
+Granularity is the abstract time levels that the people use.
+The change among granularities~\cite{Lin97} is considered also as a source of imprecision.
+Therefore, some proposals consider the granularity as the base of the temporal model~\cite{Cru97}.
+
+The temporal relations between time points and intervals are studied first by Allen~\cite{Allen83} and recently fuzzified by several authors~\cite{ohlbach04,nagypal03,schockaert08}.
+
+In the other hand, the addition of time to a relational database~\cite{Dyreson1994} leads to several problems: the modelling of a time-variant object has an impact on the database consistency.
+Different models have been proposed \cite{Jensen94,TSQL,Sarda90,Jensen91,Snodgrass84,Nascimento95} but all of them consider the time as a crisp value.
+%There exist a glossary with the common terms in the temporal database community.
+Whereas the concept of time is intrinsically imprecise, the first temporal database models were crisp.
+In recent works, some authors consider the necessity of allowing imprecission and vagueness in a temporal model~\cite{Cru97,624013,fuzz2009,gal01}.
+
+There are several proposals for transforming two fuzzy numbers which represent a fuzzy interval into a unified representation as a single fuzzy number.
+These transformations (the convex hull and the transformation that preserves the imprecision)~\cite{fuzz2009} are misleading.
+
+In this work we will propose a possibilistic valid time model for relational databases.
+The basis of the model is possibilistic framework presented in~\cite{Pon11}.
+The framework deals with the inference of uncertainty about sets from the uncertainty about values.
 
 The paper is organized as follows. In Section \ref{sec:preliminaries}, some preliminaries and concepts about the possibilistic evaluation framework and the temporal databases are presented. 
-Section \ref{sec:proposal} presents the possibilistic model for valid-time databases.
-Finally, Section \ref{sec:conclusions} presents the main conclusions and some lines for the future work in this research.
+Section \ref{sec:proposal} presents the possibilistic model for valid-time databases. In this section is introduced both the representation and the querying.
+Finally, Section \ref{sec:futher-research} presents the main conclusions and some lines for the future work in this research.

### 2011-11-10 ipmu2012/preliminaries.tex
update

@@
+%Introductory text:
+This background section is dedicated to introduce some basic concepts about possibilistic variables, the differences between fuzzy numbers and intervals.
+The interval evaluation by means of ill-known constraints is explained.
+
+The preliminaries section concludes with a brief introduction to temporal databases.
+
 \subsection{Possibilistic Variables}
@@
 
-\paragraph{Example} Consider the ill-known values $X = \left[5, 2, 8\right]$ and $Y = \left[9, 7, 10 \right]$.
+%\paragraph{Example} Consider the ill-known values $X = \left[5, 2, 8\right]$ and $Y = \left[9, 7, 10 \right]$.
 The knowledge about the evaluation of the interval $\left[a, b \right]$  is given by the expressions \eqref{eq:interval-pos},\eqref{eq:interval-nec}.
@@
 Note the triangular form for the resulting possibility distribution since the condition $a \leq b$ holds.
-
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=0.4]{graphs/3D_possibility.eps}
-\caption{Possibility of evaluation for the interval $[a,b]$.}
-\label{fig:3d-possibility}
-\end{figure}
-The necessity plot is obtained in a similar way and is shown in Figure~\ref{fig:3d-necessity}. Notice that the necessity measure is not normalized because the supports of $X$ and $Y$ overlap.
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=0.4]{graphs/3D_necessity.eps}
-\caption{Necessity of evaluation for the interval $[a,b]$.}
-\label{fig:3d-necessity}
-\end{figure}
+%
+%\begin{figure}[h!]
+%\centering
+%\includegraphics[scale=0.4]{graphs/3D_possibility.eps}
+%\caption{Possibility of evaluation for the interval $[a,b]$.}
+%\label{fig:3d-possibility}
+%\end{figure}
+%The necessity plot is obtained in a similar way and is shown in Figure~\ref{fig:3d-necessity}. Notice that the necessity measure is not normalized because the supports of $X$ and $Y$ overlap.
+%\begin{figure}[h!]
+%\centering
+%\includegraphics[scale=0.4]{graphs/3D_necessity.eps}
+%\caption{Necessity of evaluation for the interval $[a,b]$.}
+%\label{fig:3d-necessity}
+%\end{figure}
 
@@
 
-Fuzzy temporal models \cite{4481150} deal with time points \cite{Dubois89} or intervals \cite{Garrido2009} that are ill-known.
+Fuzzy temporal models \cite{schockaert08} deal with time points \cite{Dubois89} or intervals \cite{Garrido2009} that are ill-known.
 Some fuzzy temporal models assume that the time stored in the database is an interval.

### 2011-11-10 ipmu2012/proposal.tex
update

@@
+The proposal consist on a possibilistic valid-time model. The representation and the querying are explained in the following subsections.
+
 \subsection{Representation of ill-known valid-time intervals}
+Valid time is usually represented as an interval. The interval has a starting and an ending points. An ill-known valid-time interval is an interval in witch one or both points are ill-known.
+
+\begin{definition}
+A possibilistic valid-time period \emph{PVT} is a possibilistic interval defined by means of two ill-known points, namely $\left[ X,\ Y \right]$
+\begin{equation}
+PVT = \left[X,\ Y \right] 
+\end{equation}
+X and Y are ill-known values in the set of the real numbers $\mathbb{R}$. The uncertainty about the values taken by $X$ and $Y$ are given by the possibility distributions $\pi_X$ and $\pi_Y$.
+\end{definition}
 

### 2011-11-11 trunk/ipmu2012/preliminaries.tex
Update of the proposal section.

@@
-\subsection{Possibilistic Variables}
+%Introductory text:
+This background section is dedicated to introduce some basic concepts about possibilistic variables, the differences between fuzzy numbers and intervals.
+The interval evaluation by means of ill-known constraints is explained.
+
+The preliminaries section concludes with a brief introduction to temporal databases.
+
+\subsection{\label{subsec:possibilistic-variables}Possibilistic Variables}
 A possibilistic variable $X$ over a universe $U$ is defined as a variable taking exactly one value in $U$, but for which this value is (partially) unknown.
@@
 Dubois and Prade~\cite{Dubois1983} proposed the following definition of \emph{fuzzy interval}.
-%\begin{definition}
+\begin{definition}
 A fuzzy interval is a fuzzy set $M$ on the set of real numbers $\mathbb{R}$ such that:
@@
 \end{eqnarray}
-%\end{definition}
+\end{definition}
 If $m$ is unique, then $M$ is referred to as a \emph{fuzzy number}, instead of a \emph{fuzzy interval}.
@@
 The problem of the interval evaluation is explained in \cite{Pon11}: For a crisp interval $I = \left[ a, b \right]$, we want to know if all points in this interval reside between the boundaries of the ill-known interval $\left[ X , Y \right]$, where $X,Y$ are ill-known values on the set of real numbers $\mathbb{R}$, and $\lambda$ is the evaluation function.
+
+We assume that $X$ specifies the lower bound and $Y$ the upper bound for a given interval, we want to known whether all points in the interval are larger than or equal to $X$ and smaller than or equal to $Y$.
+Therefore, we consider two ill-known constraints:
+
+\begin{eqnarray}
+C_1\stackrel{\triangle}{=}\left(\leq,X\right)\\
+C_2\stackrel{\triangle}{=}\left(\geq,Y\right).
+\end{eqnarray}
+
 The following expressions compute the possibility and the necessity:

### 2011-11-11 ipmu2012/proposal.tex
Update of the proposal section.

@@
 \begin{definition}
-A possibilistic valid-time period \emph{PVT} is a possibilistic interval defined by means of two ill-known points, namely $\left[ X,\ Y \right]$
+A Possibilistic Valid-Time Period \textbf{PVP} is a possibilistic interval defined by means of two ill-known points, namely $\left[ X,\ Y \right]$
 \begin{equation}
-PVT = \left[X,\ Y \right] 
+PVP = \left[X,\ Y \right] 
 \end{equation}
-X and Y are ill-known values in the set of the real numbers $\mathbb{R}$. The uncertainty about the values taken by $X$ and $Y$ are given by the possibility distributions $\pi_X$ and $\pi_Y$.
+$X$ and $Y$ are ill-known values in the set of the real numbers $\mathbb{R}$. The uncertainty about the values taken by $X$ and $Y$ are given by the possibility distributions $\pi_X$ and $\pi_Y$.
 \end{definition}
 
+The possibility distributions $\pi_X$ and $\pi_Y$ are given in the way of a triangular distribution, as explained in subsection \ref{subsec:fuzzy-numbers}.
+This representation allows overlapping (Fig.
+\ref{fig:pvp}).
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[scale=0.2]{graphs/2_triangular.eps}
+  \caption{Two fuzzy numbers $X$ and $Y$ denoting a Possibilistic Valid-Time Period \emph{PVT}.}
+  \label{fig:pvp}
+\end{figure}
+
 \subsection{Storage of valid-time intervals}
+Each database row containing a \emph{PVP} stores it as two triangular possibility distributions.
+In our approach we propose the representation of that as proposed in the  fuzzy interface for relational databases \emph{FIRST}~\cite{Medina94gefred.a,Gal98,Gal06}.
+In this representation it is also possible to represent not only fuzzy numbers but fuzzy constants (see table \ref{table:relational-representation-pvp}):
+
+\begin{itemize}
+\item
+\emph{NULL}: This constant refers to a completely ignorance about the value.
+The possibility distribution for a given fuzzy number $X$ is not defined, therefore, any comparison between a fuzzy number and the \emph{NULL} constant always returns $0$.
+\item
+\emph{UNKNOWN}: The point has a value but it is unknown. The possibility distribution for a given fuzzy number $X$ is $\pi_X=1$
+\item
+\emph{UNDEFINED}: The point does not have a value. The possibility distribution for a given fuzzy number $X$ is $\pi_X=0$
+\end{itemize}
+
+\begin{table}
+\caption{Relational representation for a Possibilistic Valid-Time Period.}
+\centering
+\begin{tabular}{c c c c c c}
+\hline
+Value & FT & F1 & F2 & F3  \\ \hline
+UNKNOWN & 0 & NULL & NULL & NULL  \\ 
+UNDEFINED & 1 & NULL & NULL & NULL  \\ 
+NULL & 2 & NULL & NULL & NULL  \\ 
+$\left[D,\ a,\ b \right]$ & 3 & $D$ & $D-a$ & $D+b$ \\ 
+\hline
+\end{tabular}
+\label{table:relational-representation-pvp}
+\end{table}
 

### 2011-11-14 trunk/ipmu2012/ipmu.tex
I changed the introduction section. Next, I'll change some things about the preliminaries. If you disagree, please send me an e-mail, I am off course open to discussion.

--- 2011-11-09 ipmu2012/ipmu.tex
@@
 %
-\title{A possibilistic valid-time model}
+\title{A Possibilistic Valid-time Model}
 %
@@
 %
-\author{Jos\'e Enrique Pons \and Christophe Billiet \and Olga Pons Capote \and Guy de Tr\'e}
+\author{Jos\'e Enrique Pons \and Christophe Billiet \and Olga Pons Capote \and Guy De Tr\'e}
 %Jeffrey Dean \and David Grove \and Craig Chambers \and Kim~B.~Bruce \and
@@
 Ghent University
-Sint Pietersniewstraat 41 B-9000, Gent, Belgium.
+Sint Pietersnieuwstraat 41 B-9000, Ghent, Belgium.
 \email{Christophe.Billiet,Guy.De.Tre@ugent.be}

### 2011-11-14 trunk/ipmu2012/preliminaries.tex
I changed the introduction section. Next, I'll change some things about the preliminaries. If you disagree, please send me an e-mail, I am off course open to discussion.

@@
 %Introductory text:
-This background section is dedicated to introduce some basic concepts about possibilistic variables, the differences between fuzzy numbers and intervals.
-The interval evaluation by means of ill-known constraints is explained.
+In this section, some basic concepts are introduced, concerning possibility theory, possibilistic variables and fuzzy numbers and fuzzy intervals.
+The framework of interval evaluation by ill-known constraints is explained.
+The section concludes with a brief introduction to temporal databases.
 
-The preliminaries section concludes with a brief introduction to temporal databases.
+\subsection{\label{subsec:possibility-theory}Possibility Theory}
 

### 2011-11-14 ipmu2012/introduction.tex
I changed the introduction section. Next, I'll change some things about the preliminaries. If you disagree, please send me an e-mail, I am off course open to discussion.

--- 2011-11-09 ipmu2012/introduction.tex
@@
-There are several proposals in the literature for inferring uncertainty about sets from uncertainty about values.
-It is of particular interest for fuzzy temporal databases, the application on temporal intervals.
-The transformation from two ill-known temporal points to a ill-known set has been discussed in \cite{Garrido2009}.
-In this work we will analyze the major pitfalls of these approaches and the benefit of using the mechanism proposed in \cite{Pon11}.
+Dealing with time in information systems is usually a difficult task~\cite{Bolour82}.
 
-The paper is organized as follows. In Section \ref{sec:preliminaries}, some preliminaries and concepts about the possibilistic evaluation framework and the temporal databases are presented. 
-Section \ref{sec:proposal} presents the possibilistic model for valid-time databases.
-Finally, Section \ref{sec:conclusions} presents the main conclusions and some lines for the future work in this research.
+First of all, the concept of \emph{time} itself is very complex.
+Klein~\cite{klein94} studied the concept of time in language and Devos~\cite{devos94} modelled vague temporal expressions by means of fuzzy sets~\cite{zadeh65}.
+More towards information systems, several proposals have been concerned with the obtaining of theoretical models that allow the representation of time in information systems~\cite{cruyssen97,devos98,Cru97}.
+Notably, temporal relations between time intervals and to a lesser extent time points, were first studied by Allen~\cite{Allen83}.
+
+Next to the inherent difficulties of time, an issue is \emph{temporal granularity}.
+A temporal granularity is a partitioning of the time line used by a system, usually dependent on the application.
+For example, the age of an adult human being is usually expressed in years: one will use sentences like `Laura is 21 years old' instead of `Laura is 21 years, 3 months and 4 days old'.
+In this example any time period shorter than a year needs no representation and thus the used granularity allows no specification for time periods shorter than a year.
+
+Another issue is the possible inherent imperfection in the description of notions of time.
+The temporal notion in a sentence like `The Alhambra, which I can see from my room, is a few hundred years old' contains imperfection caused by the vagueness of the used expression.
+The temporal notion in a sentence like `The Alhambra was finished somewhere between 1301 A.D.
+and 1400 A.D.' contains imperfection caused by the inaccuracy in the given temporal notion, though precise boundaries are given.
+
+The transition between granularities~\cite{Lin97} is also considered as a source of imprecision.
+In a sentence like `The Alhambra is 612 years old', the temporal notion is very precise if time periods shorter than a year are irrelevant in the used granularity.
+However, if one is concerned with the exact day on which the last stone of the building is laid, a transition to a finer granularity, allowing the specification of time periods shorter than a year, is needed.
+Therefore, some proposals consider granularity as the base of the temporal model~\cite{Cru97}.
+
+To allow information systems to cope with imperfections, many approaches adopt fuzzy sets for the representation of temporal information~\cite{Virant199639,35337,343607,nagypal03,Dubois:jucs_9_9:fuzziness_and_uncertainty_in}, though some other tools, like rough sets~\cite{pawlak95}, have also been used~\cite{Qia09}.
+The temporal relations studied by Allen were recently fuzzified by several authors~\cite{ohlbach04,nagypal03,schockaert08}.
+
+Next to the issues in temporal modelling, the addition of time to a relational database~\cite{Dyreson1994} leads to several more problems, as the modelling of a time-variant object has an impact on the database consistency.
+Different models have been proposed \cite{Jensen94,TSQL,Sarda90,Jensen91,Snodgrass84,Nascimento95}, but all of these consider time as a crisp notion.
+In recent works though, some authors consider the necessity of allowing imprecision in temporal models~\cite{Cru97,624013,fuzz2009,gal01}.
+
+Most of these models use the concepts of fuzzy numbers and fuzzy intervals to represent temporal imprecision and considerable attention has thus been given to the problem of transforming two fuzzy numbers which represent the boundaries of an imprecise time interval into one fuzzy interval representing this interval.
+There are several proposals for this transformation, but some, like those presented in~\cite{fuzz2009} are misleading.
+
+In this work we will propose a possibilistic valid time model for relational databases.
+The base of the model is the possibilistic framework presented in~\cite{Pon11}.
+This framework deals with the inference of uncertainty about the evaluation of sets from the uncertainty about the exact values used in evaluation.
+
+The paper is organized as follows. In section \ref{sec:preliminaries}, some preliminaries and concepts about the possibilistic evaluation framework and temporal databases are presented. 
+Section \ref{sec:proposal} presents the possibilistic model for valid-time databases. In this section, both representation and querying are introduced.
+Finally, section \ref{sec:futher-research} presents the main conclusions and some lines for future work in this research.

### 2011-11-14 ipmu2012/preliminaries.tex
In this revision, I cleaned up the list of references a little. Now it spreads over 1.5 pages approximately. It would be nice to get it on one single page...

@@
 %Introductory text:
-This background section is dedicated to introduce some basic concepts about possibilistic variables, the differences between fuzzy numbers and intervals.
-The interval evaluation by means of ill-known constraints is explained.
+In this section, some basic concepts are introduced, concerning possibility theory, possibilistic variables and fuzzy numbers and intervals.
+The framework of set evaluation by ill-known constraints\cite{Pon11} is explained.
+The section concludes with a brief introduction to temporal databases.
 
-The preliminaries section concludes with a brief introduction to temporal databases.
+\subsection{\label{subsec:possibility-theory}Possibility Theory}
+Possibility theory, like probability theory, deals with uncertainty about the outcome of an experiment.
+In probability theory, this uncertainty is caused by the variability in the outcomes, while in possibility theory, the uncertainty is caused by incomplete knowledge on the experiment.
 
-\subsection{Possibilistic Variables}
+\subsection{\label{subsec:possibilistic-variables}Possibilistic Variables}
 A possibilistic variable $X$ over a universe $U$ is defined as a variable taking exactly one value in $U$, but for which this value is (partially) unknown.
@@
 Fuzzy temporal models \cite{schockaert08} are proposed when the time points or intervals are not precisely known.
-There are proposals for the  fuzzyfication of the time point \cite{Dubois89} and the fuzzyfication of the time interval \cite{Garrido2009}.\\
+There are proposals for the  fuzzyfication of the time point and the fuzzyfication of the time interval.\\
 Allen \cite{Allen83} studied the comparison between two crisp time intervals. For fuzzy intervals, several proposals \cite{ohlbach04,nagypal03,schockaert08} have been done.
@@
 
-Despite of user-defined time (an uninterpreted attribute; supported in the standard SQL2 \cite{Mel93}), there are 3 types of time in a temporal database:
+Despite of user-defined time (an uninterpreted attribute; supported in the standard SQL2), there are 3 types of time in a temporal database:
 
@@
 
-	 The  database model can be then classified into: transaction time \cite{Ston87},\cite{Jensen91}, Valid time, bi-temporal \cite{Snodgrass84}(both valid and transaction time) or tri-temporal \cite{Nascimento95} (valid, transaction and decision time).
+	 The  database model can be then classified into: transaction time \cite{Jensen91}, Valid time, bi-temporal \cite{Snodgrass84}(both valid and transaction time) or tri-temporal \cite{Nascimento95} (valid, transaction and decision time).
 

### 2011-11-14 trunk/ipmu2012/introduction.tex
In this revision, I cleaned up the list of references a little. Now it spreads over 1.5 pages approximately. It would be nice to get it on one single page...

@@
-Dealing with time in any information system~\cite{Bolour82} is usually a difficult task.
-First of all, the concept of \emph{time} itself is a complex notion.
-Klein~\cite{klein94} studied the concept of time in language whereas Devos~\cite{devos94} modelled vague temporal expressions by means of fuzzy sets~\cite{zadeh65}.
-Several research works~\cite{cruyssen97,devos98,Cru97} have been done in obtaining theoretical models that allow the representation of time in an information system.
-Although the main tool for the representation of temporal information are fuzzy sets~\cite{Virant199639,35337,343607,springerlink:10.1007/978-3-540-39964-3_57,Dubois:jucs_9_9:fuzziness_and_uncertainty_in}, some other tools, like rough sets~\cite{pawlak95} have been used~\cite{Qia09}.
+Dealing with time in information systems is usually a difficult task~\cite{Bolour82}.
 
-In the other hand, the addition of time to a relational database leads to several problems: the modelling of a time-variant object has an impact on the database consistency.
-Different models have been proposed \cite{Jensen94,TSQL,Sarda90,Jensen91,Snodgrass84,Nascimento95} but all of them consider the time as a crisp value.
-There exist a glossary~\cite{Dyreson1994} with the common terms in the temporal database community.
-Further works introduced the concept of vagueness, imprecission in the different models~\cite{Cru97,624013,fuzz2009,gal01}.
+First of all, the concept of \emph{time} itself is very complex.
+Klein~\cite{klein94} studied the concept of time in language and Devos~\cite{devos94} modelled vague temporal expressions by means of fuzzy sets~\cite{zadeh65}.
+More towards information systems, several proposals have been concerned with the obtaining of theoretical models that allow the representation of time in information systems~\cite{devos98,Cru97}.
+Notably, temporal relations between time intervals and to a lesser extent time points, were first studied by Allen~\cite{Allen83}.
 
-There are several proposals in the literature for inferring uncertainty about sets from uncertainty about values.
-It is of particular interest for fuzzy temporal databases, the application on temporal intervals.
-The transformation from two ill-known temporal points to a ill-known set has been discussed in \cite{Garrido2009}.
-In this work we will analyze the major pitfalls of these approaches and the benefit of using the mechanism proposed in \cite{Pon11}.
+Next to the inherent difficulties of time, an issue is \emph{temporal granularity}.
+A temporal granularity is a partitioning of the time line used by a system, usually dependent on the application.
+For example, the age of an adult human being is usually expressed in years: one will use sentences like `Laura is 21 years old' instead of `Laura is 21 years, 3 months and 4 days old'.
+In this example any time period shorter than a year needs no representation and thus the used granularity allows no specification for time periods shorter than a year.
 
-The paper is organized as follows. In Section \ref{sec:preliminaries}, some preliminaries and concepts about the possibilistic evaluation framework and the temporal databases are presented. 
-Section \ref{sec:proposal} presents the possibilistic model for valid-time databases.
-Finally, Section \ref{sec:conclusions} presents the main conclusions and some lines for the future work in this research.
+Another issue is the possible inherent imperfection in the description of notions of time.
+The temporal notion in a sentence like `The Alhambra, which I can see from my room, is a few hundred years old' contains imperfection caused by the vagueness of the used expression.
+The temporal notion in a sentence like `The Alhambra was finished somewhere between 1301 A.D.
+and 1400 A.D.' contains imperfection caused by the inaccuracy in the given temporal notion, though precise boundaries are given.
+
+The transition between granularities~\cite{Lin97} is also considered as a source of imprecision.
+In a sentence like `The Alhambra is 612 years old', the temporal notion is very precise if time periods shorter than a year are irrelevant in the used granularity.
+However, if one is concerned with the exact day on which the last stone of the building is laid, a transition to a finer granularity, allowing the specification of time periods shorter than a year, is needed.
+Therefore, some proposals consider granularity as the base of the temporal model~\cite{Cru97}.
+
+To allow information systems to cope with imperfections, many approaches adopt fuzzy sets for the representation of temporal information~\cite{343607,nagypal03,Dubois:jucs_9_9:fuzziness_and_uncertainty_in}, though some other tools, like rough sets, have also been used~\cite{Qia09}.
+The temporal relations studied by Allen were recently fuzzified by several authors~\cite{ohlbach04,nagypal03,schockaert08}.
+
+Next to the issues in temporal modelling, the addition of time to a relational database~\cite{Dyreson1994} leads to several more problems, as the modelling of a time-variant object has an impact on the database consistency.
+Different models have been proposed \cite{Jensen91,Snodgrass84,Nascimento95}, but all of these consider time as a crisp notion.
+In recent works though, some authors consider the necessity of allowing imprecision in temporal models~\cite{Cru97,fuzz2009}.
+
+Most of these models use the concepts of fuzzy numbers and fuzzy intervals to represent temporal imprecision and considerable attention has thus been given to the problem of transforming two fuzzy numbers which represent the boundaries of an imprecise time interval into one fuzzy interval representing this interval.
+There are several proposals for this transformation, but some, like those presented in~\cite{fuzz2009} are misleading.
+
+In this work we will propose a possibilistic valid time model for relational databases.
+The base of the model is the possibilistic framework presented in~\cite{Pon11}.
+This framework deals with the inference of uncertainty about the evaluation of sets from the uncertainty about the exact values used in evaluation.
+
+The paper is organized as follows. In section \ref{sec:preliminaries}, some preliminaries and concepts about the possibilistic evaluation framework and temporal databases are presented. 
+Section \ref{sec:proposal} presents the possibilistic model for valid-time databases. In this section, both representation and querying are introduced.
+Finally, section \ref{sec:futher-research} presents the main conclusions and some lines for future work in this research.

### 2011-11-14 trunk/ipmu2012/proposal.tex
In this revision, I cleaned up the list of references a little. Now it spreads over 1.5 pages approximately. It would be nice to get it on one single page...

--- 2011-11-11 ipmu2012/proposal.tex
@@
 Each database row containing a \emph{PVP} stores it as two triangular possibility distributions.
-In our approach we propose the representation of that as proposed in the  fuzzy interface for relational databases \emph{FIRST}~\cite{Medina94gefred.a,Gal98,Gal06}.
+In our approach we propose the representation of that as proposed in the  fuzzy interface for relational databases \emph{FIRST}~\cite{Medina94gefred.a,Gal98}.
 In this representation it is also possible to represent not only fuzzy numbers but fuzzy constants (see table \ref{table:relational-representation-pvp}):

### 2011-11-15 ipmu2012/proposal.tex

@@
   \centering
-  \includegraphics[scale=0.2]{graphs/2_triangular.eps}
+%  \includegraphics[scale=0.2]{graphs/2_triangular.eps}
   \caption{Two fuzzy numbers $X$ and $Y$ denoting a Possibilistic Valid-Time Period \emph{PVT}.}
@@
 \subsection{Querying ill-known valid-time intervals}
+In order to provide a complete model, we will provide the tools for querying.
+This allows the user to specify both the preferences and an ill-known valid-time interval in the query.
+It is important to notice that the possibilistic / fuzzy data stored in the database has a \emph{disjunctive interpretation} (it is said that we have \emph{uncertainty}: the valid-time interval has only one value but, for some reason the value is ill-known).
+In the query specification, the user is allowed to express \emph{imprecision} or \emph{vagueness} (this is referred as a \emph{conjunctive interpretation}).
+In this subsection we will define the query specification, then the evaluation of the query and finally the ranking for the query.
 
 \subsubsection{Query specification}
+A query in our framework has two different parts: the first one is the query specification for regular attributes. The second part is the temporal specification.
+
+\begin{definition}
+A query $\tilde Q$ is specified by:
+\begin{equation}
+\tilde Q = \left( Q^{time}, Q \right)
+\end{equation}
+\end{definition}
+Where $Q$ are the (possibly fuzzy) preferences of the user  and $Q^{time}$ is the temporal part specified by a possibilistic valid-time period,\emph{PVP}.
+
+%explain the sematics of vt.
 
 \subsubsection{Query evaluation}
+Fuzzy querying of regular (relational) databases, query satisfaction modelling is a matter of degree.
+Usually, criteria satisfaction is modelled by means of a satisfaction degree $s \in \left[ 0, 1\right]$.
+In the model, every record $r$ contains a \emph{PVP} $V_r$ to model the valid-time for that record.
+
+The query evaluation method is the following:
+\begin{itemize}
+\item
+For each record $r$ in the database, the preferences expressed in $Q$ are evaluated in the unit interval.
+\end{itemize}
 

### 2011-11-17 trunk/ipmu2012/introduction.tex
Some changes to the bib and the preliminaries...

@@
 Another issue is the possible inherent imperfection in the description of notions of time.
-The temporal notion in a sentence like `The Alhambra, which I can see from my room, is a few hundred years old' contains imperfection caused by the vagueness of the used expression.
+The temporal notion in a sentence like `The Alhambra, which I can see from my room, is a few hundred years old.' contains imperfection caused by the vagueness of the used expression.
 The temporal notion in a sentence like `The Alhambra was finished somewhere between 1301 A.D.

### 2011-11-17 trunk/ipmu2012/preliminaries.tex
Some changes to the bib and the preliminaries...

@@
 %Introductory text:
-In this section, some basic concepts are introduced, concerning possibility theory, possibilistic variables and fuzzy numbers and fuzzy intervals.
-The framework of interval evaluation by ill-known constraints is explained.
+In this section, some basic concepts are introduced, concerning possibility theory, possibilistic variables and fuzzy numbers and intervals.
+The framework of set evaluation by ill-known constraints\cite{Pon11} is explained.
 The section concludes with a brief introduction to temporal databases.
@@
 \subsection{\label{subsec:possibility-theory}Possibility Theory}
+Possibility theory, like probability theory, deals with uncertainty about the outcome of an experiment.
+In probability theory, this uncertainty is caused by the \emph{variability} in the outcomes, while in possibility theory, the uncertainty is caused by \emph{incomplete knowledge} about the experiment.
+The quantification of confidence in probability theory is achieved by means of a measure of chance, whereas in possibility theory, possibility and necessity measures are used.
+In any case, the quantification of confidence in a theory of uncertainty is achieved using a confidence measure\cite{Shafer:1976:AMathematical}.
+
+\begin{definition}
+Consider a set of outcomes $\Omega$.
+Let $\wp(\Omega)$ denote the powerset of $\Omega$ and let $A$ and $B$ be elements of $\wp(\Omega)$.
+A \emph{confidence measure on $\Omega$} is defined by a function
+	\begin{align}
+	g : \wp(\Omega) & \rightarrow \left[0,1\right]
+	\end{align}
+that satisfies
+	\begin{align}
+	g(\emptyset) &= 0 \\
+	g(\Omega) &= 1 	\label{NormalizationProperty} \\
+	A \subseteq B &\Rightarrow g(A) \leq g(B) \label{MonotonicityProperty}
+	\end{align}
+\end{definition}
+
+Both possibility measures and necessity measures are special cases of confidence measures.
+
+\begin{definition}
+Consider a confidence measure $\Pi$ on a set of outcomes $\Omega$.
+Let $J$ be a countable index set and let $\{ A_{j} | j \in J \wedge A_{j} \subseteq \Omega \}$ be a family of elements of $\wp(\Omega)$.
+$\Pi$ is now a \emph{possibility measure on $\Omega$} if it satisfies:
+	\begin{align}
+	\Pi\left(\bigcup_{j \in J} A_{j} \right) = \sup_{j \in J} \Pi(A_{j})
+	\end{align}
+\end{definition}
+
+In this work, the interpretation is as follows.
+The possibility of an event expresses how plausible the occurrence of the event seems to an observer of the experiment, given the (partial) knowledge of the observer about the experiment.
+
+Information on the possibility of distinct elements of the universe of discourse $\Omega$ can now be given by a \emph{possibility distribution} $\pi$ on $\Omega$, defined by:
+
+\begin{definition}
+Consider a possibility measure $\Pi$ on $\Omega$. A \emph{possibility distribution} $\pi$ underlying the possibility measure $\Pi$ is then a function defined by:
+	\begin{align}
+	\pi : \Omega \rightarrow \left[0, 1\right] : \pi(u) = \Pi(\{u\})
+	\end{align}
+\end{definition}
+
+\begin{definition}
+Consider a confidence measure $N$ on a set of outcomes $\Omega$.
+Let $J$ be a countable index set and let $\{ A_{j} | j \in J \wedge A_{j} \subseteq \Omega \}$ be a family of elements of $\wp(\Omega)$.
+$N$ is now a \emph{necessity measure} on $\Omega$ if it satisfies:
+	\begin{align}
+	N\left(\bigcap_{j \in J} A_{j} \right) = \inf_{j \in J} N(A_{j})
+	\end{align}
+\end{definition}
+
+In this work, the interpretation is as follows.
+The necessity of an event expresses how necessary the occurrence of the event seems to an observer of the experiment, given the (partial) knowledge of the observer about the experiment.
+
+Possibility and necessity measures are dual in the sense that:
+
+\begin{align}
+\forall A \subseteq \Omega : N(A) = 1 - \Pi(\bar{A})
+\end{align}
+
+Regarding interpretation, the above can be seen as: the degree to which an event is necessary is the degree to which every other possible event is not plausible.
 

### 2011-11-17 ipmu2012/introduction.tex
Some small changes. I'll work on the fuzzy numbers and set evaluation parts next.

@@
 Klein~\cite{klein94} studied the concept of time in language and Devos~\cite{devos94} modelled vague temporal expressions by means of fuzzy sets~\cite{zadeh65}.
-More towards information systems, several proposals have been concerned with the obtaining of theoretical models that allow the representation of time in information systems~\cite{cruyssen97,devos98,Cru97}.
+More towards information systems, several proposals have been concerned with the obtaining of theoretical models that allow the representation of time in information systems~\cite{devos98,Cru97}.
 Notably, temporal relations between time intervals and to a lesser extent time points, were first studied by Allen~\cite{Allen83}.
@@
 
-To allow information systems to cope with imperfections, many approaches adopt fuzzy sets for the representation of temporal information~\cite{Virant199639,35337,343607,nagypal03,Dubois:jucs_9_9:fuzziness_and_uncertainty_in}, though some other tools, like rough sets~\cite{pawlak95}, have also been used~\cite{Qia09}.
+To allow information systems to cope with imperfections, many approaches adopt fuzzy sets for the representation of temporal information~\cite{343607,nagypal03,Dubois:jucs_9_9:fuzziness_and_uncertainty_in}, though some other tools, like rough sets, have also been used~\cite{Qia09}.
 The temporal relations studied by Allen were recently fuzzified by several authors~\cite{ohlbach04,nagypal03,schockaert08}.
@@
 Next to the issues in temporal modelling, the addition of time to a relational database~\cite{Dyreson1994} leads to several more problems, as the modelling of a time-variant object has an impact on the database consistency.
-Different models have been proposed \cite{Jensen94,TSQL,Sarda90,Jensen91,Snodgrass84,Nascimento95}, but all of these consider time as a crisp notion.
-In recent works though, some authors consider the necessity of allowing imprecision in temporal models~\cite{Cru97,624013,fuzz2009,gal01}.
+Different models have been proposed \cite{Jensen91,Snodgrass84,Nascimento95}, but all of these consider time as a crisp notion.
+In recent works though, some authors consider the necessity of allowing imprecision in temporal models~\cite{Cru97,Garrido2009}.
 
 Most of these models use the concepts of fuzzy numbers and fuzzy intervals to represent temporal imprecision and considerable attention has thus been given to the problem of transforming two fuzzy numbers which represent the boundaries of an imprecise time interval into one fuzzy interval representing this interval.
-There are several proposals for this transformation, but some, like those presented in~\cite{fuzz2009} are misleading.
+There are several proposals for this transformation, but some, like those presented in~\cite{Garrido2009} are misleading.
 

### 2011-11-17 ipmu2012/preliminaries.tex
Some small changes. I'll work on the fuzzy numbers and set evaluation parts next.

@@
 Possibility theory, like probability theory, deals with uncertainty about the outcome of an experiment.
-In probability theory, this uncertainty is caused by the variability in the outcomes, while in possibility theory, the uncertainty is caused by incomplete knowledge on the experiment.
+In probability theory, this uncertainty is caused by the \emph{variability} in the outcomes, while in possibility theory, the uncertainty is caused by \emph{incomplete knowledge} about the experiment.
+The quantification of confidence in probability theory is achieved by means of a measure of chance, whereas in possibility theory, possibility and necessity measures are used.
+In any case, the quantification of confidence in a theory of uncertainty is achieved using a confidence measure\cite{Shafer:1976:AMathematical}.
+
+\begin{definition}
+Consider a set of outcomes $\Omega$.
+Let $\wp(\Omega)$ denote the powerset of $\Omega$ and let $A$ and $B$ be elements of $\wp(\Omega)$.
+A \emph{confidence measure on $\Omega$} is defined by a function
+	\begin{align}
+	g : \wp(\Omega) & \rightarrow \left[0,1\right]
+	\end{align}
+that satisfies
+	\begin{align}
+	g(\emptyset) &= 0 \\
+	g(\Omega) &= 1 	\label{NormalizationProperty} \\
+	A \subseteq B &\Rightarrow g(A) \leq g(B) \label{MonotonicityProperty}
+	\end{align}
+\end{definition}
+
+Both possibility measures and necessity measures are special cases of confidence measures.
+
+\begin{definition}
+Consider a confidence measure $\Pi$ on a set of outcomes $\Omega$.
+Let $J$ be a countable index set and let $\{ A_{j} | j \in J \wedge A_{j} \subseteq \Omega \}$ be a family of elements of $\wp(\Omega)$.
+$\Pi$ is now a \emph{possibility measure on $\Omega$} if it satisfies:
+	\begin{align}
+	\Pi\left(\bigcup_{j \in J} A_{j} \right) = \sup_{j \in J} \Pi(A_{j})
+	\end{align}
+\end{definition}
+
+In this work, the interpretation is as follows.
+The possibility of an event expresses how plausible the occurrence of the event seems to an observer of the experiment, given the (partial) knowledge of the observer about the experiment.
+
+Information on the possibility of distinct elements of the universe of discourse $\Omega$ can now be given by a \emph{possibility distribution} $\pi$ on $\Omega$, defined by:
+
+\begin{definition}
+Consider a possibility measure $\Pi$ on $\Omega$. A \emph{possibility distribution} $\pi$ on $\Omega$ underlying the possibility measure $\Pi$ is then a function defined by:
+	\begin{align}
+	\pi : \Omega \rightarrow \left[0, 1\right] : \pi(u) = \Pi(\{u\})
+	\end{align}
+\end{definition}
+
+\begin{definition}
+Consider a confidence measure $N$ on a set of outcomes $\Omega$.
+Let $J$ be a countable index set and let $\{ A_{j} | j \in J \wedge A_{j} \subseteq \Omega \}$ be a family of elements of $\wp(\Omega)$.
+$N$ is now a \emph{necessity measure} on $\Omega$ if it satisfies:
+	\begin{align}
+	N\left(\bigcap_{j \in J} A_{j} \right) = \inf_{j \in J} N(A_{j})
+	\end{align}
+\end{definition}
+
+In this work, the interpretation is as follows.
+The necessity of an event expresses how necessary the occurrence of the event seems to an observer of the experiment, given the (partial) knowledge of the observer about the experiment.
+
+Possibility and necessity measures are dual in the sense that:
+
+\begin{align}
+\forall A \subseteq \Omega : N(A) = 1 - \Pi(\bar{A})
+\end{align}
+
+Regarding interpretation, the above can be seen as: the degree to which an event is necessary is the degree to which every other possible event is not plausible.
 
 \subsection{\label{subsec:possibilistic-variables}Possibilistic Variables}
+A \emph{possibilistic variable} can intuitively be seen as a variable for which the possible values are described by its possibility distribution\cite{Pon11}.
+
+\begin{definition}
 A possibilistic variable $X$ over a universe $U$ is defined as a variable taking exactly one value in $U$, but for which this value is (partially) unknown.
-The possibility distribution $\pi_X$ gives the available knowledge about the value that $X$ takes.
-For each $u\in U$, $\pi_X(u)$ represents the possibility that $X$ takes the value $u$.
+The possibility distribution $\pi_X$ on $U$ gives the available knowledge about the value that $X$ takes: for each $u\in U$, $\pi_X(u)$ represents the possibility that $X$ takes the value $u$.
+\end{definition}
 
-It is important to understand the difference between the following two concepts:
-\begin{itemize}
-\item
-A \emph{possibilistic variable} $X$ is bounded to take only one value , but this value is not known due to incomplete knowledge. 
-\item
-An \emph{ill-known set}~\cite{Dubois88b}: a possibilistic variable defined over the universe $\Pow(U)$.
-\end{itemize}
+The exact value a possibilistic variable takes, which is (partially) unknown, is called an \emph{ill-known value} in this work\cite{Dubois88b}.
 
-Note that while a possibilistic variable refers to one (partially) unknown value, an ill-known set is a crisp set but, for some reason, (partially) unknown.
+When a possibilistic variable is defined on the powerset $\Pow(R)$ of some universe $R$, the unique value the variable takes will be a crisp set.
+In this case, the exact set is off course (partially) unknown, but the available knowledge is given in the form of a possibility distribution on the powerset $\Pow(R)$.
+This exact value (a crisp set) the variable takes, is now called an \emph{ill-known set}\cite{Dubois88b}.
+
+In this work, an ill-known set will be defined via its start and end point, which will be ill-known values.
+The correspondences and transitions between these two representations are part of the authors current research.
+
+%It is important to understand the difference between the following two concepts:
+%\begin{itemize}
+%\item
+%A \emph{possibilistic variable} $X$ is bounded to take only one value , but this value is not known due to incomplete knowledge. 
+%\item
+%An \emph{ill-known set}~\cite{Dubois88b}: a possibilistic variable defined over the universe $\Pow(U)$.
+%\end{itemize}
+
+%Note that while a possibilistic variable refers to one (partially) unknown value, an ill-known set is a crisp set but, for some reason, (partially) unknown.
 

### 2011-11-17 ipmu2012/preliminaries.tex
Adaptation of the fuzzy interval and fuzzy number part. Let me know if you don't agree! I will now work on the part on interval evaluation by ill-known constraints.

@@
 When a possibilistic variable is defined on the powerset $\Pow(R)$ of some universe $R$, the unique value the variable takes will be a crisp set.
-In this case, the exact set is off course (partially) unknown, but the available knowledge is given in the form of a possibility distribution on the powerset $\Pow(R)$.
+In this case, the exact set is off course (partially) unknown, but the available knowledge is given in the form of a possibility distribution on the powerset $\Pow(R)$, which describes the possibility of each crisp subset of $R$ to be the value the variable takes.
 This exact value (a crisp set) the variable takes, is now called an \emph{ill-known set}\cite{Dubois88b}.
 
-In this work, an ill-known set will be defined via its start and end point, which will be ill-known values.
+In this work, an ill-known set will be defined and represented via its start and end point, which will be ill-known values.
 The correspondences and transitions between these two representations are part of the authors current research.
@@
 \subsection{\label{subsec:fuzzy-numbers}Fuzzy numbers and fuzzy intervals}
-Dubois and Prade~\cite{Dubois1983} proposed the following definition of \emph{fuzzy interval}.
+Among others, Dubois and Prade~\cite{Dubois1983} use the following definition of a \emph{fuzzy interval}.
 \begin{definition}
-A fuzzy interval is a fuzzy set $M$ on the set of real numbers $\mathbb{R}$ such that:
+A fuzzy interval is a fuzzy set $M$, defined by a membership function $\mu_{M}$, on the set of real numbers $\mathbb{R}$ such that:
 \begin{eqnarray}
-\forall (u,v)\in\mathbb{R}^2:&\\
-\nonumber
-\forall w \in [u,v]:&\mu_M(w) \geq\min(\mu_M(u),\mu_M(v))  \\
-\exists m \in \mathbb{R}:& \mu_M(m)=1 
+\mu_{M} : & \!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\! \mathbb{R} \rightarrow \left[0,1\right] \nonumber \\ 
+\forall (u,v)\in\mathbb{R}^2: \forall w \in [u,v]:&\mu_M(w) \geq\min(\mu_M(u),\mu_M(v))  \\
+\exists m \in \mathbb{R} : & \!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\! \mu_M(m)=1 
 \end{eqnarray}
 \end{definition}
-If $m$ is unique, then $M$ is referred to as a \emph{fuzzy number}, instead of a \emph{fuzzy interval}.
-In other words, if the core of a fuzzy interval is a singleton, it can be seen as a fuzzy number.
-In their work, Dubois and Prade propose four different functions (two possibility and two necessity functions) to asses the position of a fuzzy number N relative to  a fuzzy number M taken as a reference.
+If this modal value $m$ is unique, then $M$ is referred to as a \emph{fuzzy number}, instead of a \emph{fuzzy interval}.
+In other words, if the core of a fuzzy interval is a singleton, it is referred to as a fuzzy number.
 
-The most convenient representation for the membership function of a fuzzy number is a triangular function.
-The membership function $\mu_M$ for the fuzzy set $M$ has also the properties of convexity and normalization.
-Three values represent a triangular function (see Figure~\ref{fig:triangular}):
-\begin{itemize}
-\item
-$D$ is the singleton value in the core of $\mu_M(x)$.
-\item
-$D-a$ is the lower value in the support of $\mu_M(x)$. 
-\item
-$D+b$ is the upper value in the support of $\mu_M(x)$.
-\end{itemize}
-Note that the values $a$ and $b$ are values in the underlying ordered domain. E.g. $(a,b) \in \mathbb{R}^2$. The notation for a triangular function adopted here is $[D,a,b]$.
-\begin{figure}[h!]
-  \centering
-  \includegraphics[scale=0.4]{graphs/triangular.eps}
-  \caption{Triangular function.}
-  \label{fig:triangular}
-\end{figure}
-
-The most simple representation for the membership function of a fuzzy interval is a trapezoidal function.
-This membership function $\mu_T$ for the fuzzy interval $T$ is convex and normalized.
-Four values represent a trapezoidal membership function (figure  \ref{fig:trapezoidal}): $\left[\alpha,\ \beta,\ \gamma,\ \delta\right]$.
-The membership function is given by:
+A simple form of the membership function of a fuzzy interval is a trapezoidal function.
+It can be shown that such a membership function $\mu_T$ for a fuzzy interval $T$ is convex and normalized.
+A visualization is given in figure \ref{fig:trapezoidal}.
+Four reel values, denoted $\alpha$, $\beta$, $\gamma$ and $\delta$ and chosen as in figure \ref{fig:trapezoidal}, suffice to represent a trapezoidal membership function of a fuzzy interval.
+In this work, a fuzzy interval defined as such will be noted as $\left[\alpha, \beta, \gamma, \delta\right]$.
+The corresponding membership function definition for this $\mu_T$ is then given by:
 
 \begin{align}
-\mu_T(x)
+\mu_T : & \quad \mathbb{R} \rightarrow \left[0,1\right] \\
+ : & \quad x \rightarrow
 \begin{cases}
@@
   \includegraphics[scale=0.4]{graphs/trapezoidalDistribution.eps}
-  \caption{Trapezoidal distribution.}
+  \caption{Trapezoidal membership function}
   \label{fig:trapezoidal}
@@
 
-\subsubsection{Set evaluation by ill-known constraints}
+The most convenient form of the membership function of a fuzzy number is a triangular function.
+It can be shown that such a membership function $\mu_M$ for a fuzzy number $M$ is convex and normalized.
+A visualization is given in figure \ref{fig:triangular}.
+Three reel values, denoted $a$, $b$ and $D$, suffice to represent a triangular membership function of a fuzzy number.
+\begin{itemize}
+\item
+$D$ denotes the single value in the core of $M$
+\item
+$D-a$ is then $\inf \{u \in \mathbb{R} : \mu_{M}(u) > 0\}$
+\item
+$D+b$ is then $\sup \{u \in \mathbb{R} : \mu_{M}(u) > 0\}$
+\end{itemize}
+In this work, a fuzzy number defined as such will be noted as $\left[D, a, b \right]$.
+\begin{figure}[h!]
+  \centering
+  \includegraphics[scale=0.4]{graphs/triangular.eps}
+  \caption{Triangular membership function.}
+  \label{fig:triangular}
+\end{figure}
+
+%\subsubsection{Set evaluation by ill-known constraints}
 

### 2011-11-17 ipmu2012/proposal.tex
I modified the part about possibilistic variables because I needed the given information to explain the evaluation of sets...

@@
 
-\subsection{Representation of ill-known valid-time intervals}
+\subsection{Representation of Ill-known Valid-time Intervals}
 Valid time is usually represented as an interval. The interval has a starting and an ending points. An ill-known valid-time interval is an interval in witch one or both points are ill-known.
@@
 
-\subsection{Storage of valid-time intervals}
+\subsection{Storage of Valid-time Intervals}
 Each database row containing a \emph{PVP} stores it as two triangular possibility distributions.
-In our approach we propose the representation of that as proposed in the  fuzzy interface for relational databases \emph{FIRST}~\cite{Medina94gefred.a,Gal98,Gal06}.
+In our approach we propose the representation of that as proposed in the  fuzzy interface for relational databases \emph{FIRST}~\cite{Medina94gefred.a,Gal98}.
 In this representation it is also possible to represent not only fuzzy numbers but fuzzy constants (see table \ref{table:relational-representation-pvp}):
@@
 
-\subsection{Querying ill-known valid-time intervals}
+\subsection{Querying Ill-known Valid-time Intervals}
 In order to provide a complete model, we will provide the tools for querying.

### 2011-11-17 trunk/ipmu2012/preliminaries.tex
I modified the part about possibilistic variables because I needed the given information to explain the evaluation of sets...

@@
 \begin{definition}
-Consider a possibility measure $\Pi$ on $\Omega$. A \emph{possibility distribution} $\pi$ underlying the possibility measure $\Pi$ is then a function defined by:
+Consider a possibility measure $\Pi$ on $\Omega$. A \emph{possibility distribution} $\pi$ on $\Omega$ underlying the possibility measure $\Pi$ is then a function defined by:
 	\begin{align}
@@
 \subsection{\label{subsec:possibilistic-variables}Possibilistic Variables}
+A \emph{possibilistic variable} can intuitively be seen as a variable for which the possible values are described by its possibility distribution\cite{Pon11}.
+
+\begin{definition}
 A possibilistic variable $X$ over a universe $U$ is defined as a variable taking exactly one value in $U$, but for which this value is (partially) unknown.
-The possibility distribution $\pi_X$ gives the available knowledge about the value that $X$ takes.
-For each $u\in U$, $\pi_X(u)$ represents the possibility that $X$ takes the value $u$.
+The possibility distribution $\pi_X$ on $U$ gives the available knowledge about the value that $X$ takes: for each $u\in U$, $\pi_X(u)$ represents the possibility that $X$ takes the value $u$.
+\end{definition}
 
-It is important to understand the difference between the following two concepts:
-\begin{itemize}
-\item
-A \emph{possibilistic variable} $X$ is bounded to take only one value , but this value is not known due to incomplete knowledge. 
-\item
-An \emph{ill-known set}~\cite{Dubois88b}: a possibilistic variable defined over the universe $\Pow(U)$.
-\end{itemize}
+The exact value a possibilistic variable takes, which is (partially) unknown, is called an \emph{ill-known value} in this work\cite{Dubois88b}.
 
-Note that while a possibilistic variable refers to one (partially) unknown value, an ill-known set is a crisp set but, for some reason, (partially) unknown.
+When a possibilistic variable is defined on the powerset $\Pow(R)$ of some universe $R$, the unique value the variable takes will be a crisp set.
+In this case, the exact set is off course (partially) unknown, but the available knowledge is given in the form of a possibility distribution on the powerset $\Pow(R)$, which describes the possibility of each crisp subset of $R$ to be the value the variable takes.
+This exact value (a crisp set) the variable takes, is now called an \emph{ill-known set}\cite{Dubois88b}.
 
-\subsection{\label{subsec:fuzzy-numbers}Fuzzy numbers and fuzzy intervals}
-Dubois and Prade~\cite{Dubois1983} proposed the following definition of \emph{fuzzy interval}.
+In this work, an ill-known set will be defined and represented via its start and end point, which will be ill-known values.
+The correspondences and transitions between these two representations are part of the authors current research.
+
+%It is important to understand the difference between the following two concepts:
+%\begin{itemize}
+%\item
+%A \emph{possibilistic variable} $X$ is bounded to take only one value , but this value is not known due to incomplete knowledge. 
+%\item
+%An \emph{ill-known set}~\cite{Dubois88b}: a possibilistic variable defined over the universe $\Pow(U)$.
+%\end{itemize}
+
+%Note that while a possibilistic variable refers to one (partially) unknown value, an ill-known set is a crisp set but, for some reason, (partially) unknown.
+
+A specific application of possibilistic variables is obtained when the universe under consideration is the set of boolean values $\mathbb{B}$ = $\{T,F\}$.
+Indeed, any boolean propositional variable $p$ takes just one value in $\mathbb{B}$.
+If the knowledge on which value this variable $p$ will take is given by a possibility distribution $\pi_p$, the variable can be seen as a possibilistic variable.
+As the interest lies with the case where the proposition holds, the possibility and necessity that $p$ = $T$ (the proposition holds) demands most attention.
+This possibility and necessity is noted here as:
+\begin{align}
+Pos(p) & = \pi_p(T) \label{propholdsposs} \\
+Nec(p) & = 1-\pi_p(F) \label{propholdsnecc}
+\end{align}
+Here, notation \ref{propholdsposs} denotes the possibility that $p$ = $T$ and the proposition holds, notation \ref{propholdsnecc} denotes the necessity that $p$ = $T$ and the proposition holds.
+
+\subsection{\label{subsec:fuzzy-numbers}Fuzzy Numbers and Fuzzy Intervals}
+Among others, Dubois and Prade~\cite{Dubois1983} use the following definition of a \emph{fuzzy interval}.
 \begin{definition}
-A fuzzy interval is a fuzzy set $M$ on the set of real numbers $\mathbb{R}$ such that:
+A fuzzy interval is a fuzzy set $M$, defined by a membership function $\mu_{M}$, on the set of real numbers $\mathbb{R}$ such that:
 \begin{eqnarray}
-\forall (u,v)\in\mathbb{R}^2:&\\
-\nonumber
-\forall w \in [u,v]:&\mu_M(w) \geq\min(\mu_M(u),\mu_M(v))  \\
-\exists m \in \mathbb{R}:& \mu_M(m)=1 
+\mu_{M} : & \!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\! \mathbb{R} \rightarrow \left[0,1\right] \nonumber \\ 
+\forall (u,v)\in\mathbb{R}^2: \forall w \in [u,v]:&\mu_M(w) \geq\min(\mu_M(u),\mu_M(v))  \\
+\exists m \in \mathbb{R} : & \!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\! \mu_M(m)=1 
 \end{eqnarray}
 \end{definition}
-If $m$ is unique, then $M$ is referred to as a \emph{fuzzy number}, instead of a \emph{fuzzy interval}.
-In other words, if the core of a fuzzy interval is a singleton, it can be seen as a fuzzy number.
-In their work, Dubois and Prade propose four different functions (two possibility and two necessity functions) to asses the position of a fuzzy number N relative to  a fuzzy number M taken as a reference.
+If this modal value $m$ is unique, then $M$ is referred to as a \emph{fuzzy number}, instead of a \emph{fuzzy interval}.
+In other words, if the core of a fuzzy interval is a singleton, it is referred to as a fuzzy number.
 
-The most convenient representation for the membership function of a fuzzy number is a triangular function.
-The membership function $\mu_M$ for the fuzzy set $M$ has also the properties of convexity and normalization.
-Three values represent a triangular function (see Figure~\ref{fig:triangular}):
-\begin{itemize}
-\item
-$D$ is the singleton value in the core of $\mu_M(x)$.
-\item
-$D-a$ is the lower value in the support of $\mu_M(x)$. 
-\item
-$D+b$ is the upper value in the support of $\mu_M(x)$.
-\end{itemize}
-Note that the values $a$ and $b$ are values in the underlying ordered domain. E.g. $(a,b) \in \mathbb{R}^2$. The notation for a triangular function adopted here is $[D,a,b]$.
-\begin{figure}[h!]
-  \centering
-  \includegraphics[scale=0.4]{graphs/triangular.eps}
-  \caption{Triangular function.}
-  \label{fig:triangular}
-\end{figure}
-
-The most simple representation for the membership function of a fuzzy interval is a trapezoidal function.
-This membership function $\mu_T$ for the fuzzy interval $T$ is convex and normalized.
-Four values represent a trapezoidal membership function (figure  \ref{fig:trapezoidal}): $\left[\alpha,\ \beta,\ \gamma,\ \delta\right]$.
-The membership function is given by:
+A simple form of the membership function of a fuzzy interval is a trapezoidal function.
+It can be shown that such a membership function $\mu_T$ for a fuzzy interval $T$ is convex and normalized.
+A visualization is given in figure \ref{fig:trapezoidal}.
+Four reel values, denoted $\alpha$, $\beta$, $\gamma$ and $\delta$ and chosen as in figure \ref{fig:trapezoidal}, suffice to represent a trapezoidal membership function of a fuzzy interval.
+In this work, a fuzzy interval defined as such will be noted as $\left[\alpha, \beta, \gamma, \delta\right]$.
+The corresponding membership function definition for this $\mu_T$ is then given by:
 
 \begin{align}
-\mu_T(x)
+\mu_T : & \quad \mathbb{R} \rightarrow \left[0,1\right] \\
+ : & \quad x \rightarrow
 \begin{cases}
@@
 
-\subsubsection{Set evaluation by ill-known constraints}
+The most convenient form of the membership function of a fuzzy number is a triangular function.
+It can be shown that such a membership function $\mu_M$ for a fuzzy number $M$ is convex and normalized.
+A visualization is given in figure \ref{fig:triangular}.
+Three reel values, denoted $a$, $b$ and $D$, suffice to represent a triangular membership function of a fuzzy number.
+\begin{itemize}
+\item
+$D$ denotes the single value in the core of $M$
+\item
+$D-a$ is then $\inf \{u \in \mathbb{R} : \mu_{M}(u) > 0\}$
+\item
+$D+b$ is then $\sup \{u \in \mathbb{R} : \mu_{M}(u) > 0\}$
+\end{itemize}
+In this work, a fuzzy number defined as such will be noted as $\left[D, a, b \right]$.
+\begin{figure}[h!]
+  \centering
+  \includegraphics[scale=0.4]{graphs/triangular.eps}
+  \caption{Triangular membership function.}
+  \label{fig:triangular}
+\end{figure}
 
-\subsubsection{Interval evaluation by ill-known constraints}
+%\subsubsection{Set evaluation by ill-known constraints}
+
+\subsubsection{Interval Evaluation by Ill-known Constraints}
 The problem of the interval evaluation is explained in \cite{Pon11}: For a crisp interval $I = \left[ a, b \right]$, we want to know if all points in this interval reside between the boundaries of the ill-known interval $\left[ X , Y \right]$, where $X,Y$ are ill-known values on the set of real numbers $\mathbb{R}$, and $\lambda$ is the evaluation function.
@@
 
-\subsection{Temporal databases}
+\subsection{Temporal Databases}
 %In this section, the proposed reasoning is applied to the specific context of intervals on the real line. This setting is of specific interest in the context of fuzzy temporal databases.

### 2011-11-17 trunk/ipmu2012/preliminaries.tex
Last changes for today... I'm getting fuzzy myself...;-)

@@
 \subsubsection{Interval Evaluation by Ill-known Constraints}
-The problem of the interval evaluation is explained in \cite{Pon11}: For a crisp interval $I = \left[ a, b \right]$, we want to know if all points in this interval reside between the boundaries of the ill-known interval $\left[ X , Y \right]$, where $X,Y$ are ill-known values on the set of real numbers $\mathbb{R}$, and $\lambda$ is the evaluation function.
+The problem of the interval evaluation is explained in \cite{Pon11}: for a crisp interval $I = \left[ a, b \right]$, the need exists to know if all points in this interval reside between the boundaries of the ill-known interval $\left[ X , Y \right]$, where $X,Y$ are ill-known values on the set of real numbers $\mathbb{R}$, and $\lambda$ is the evaluation function.
 

### 2011-11-17 ipmu2012/preliminaries.tex
Update for the graphs and the code in the preliminaries.

@@
 
-\subsection{\label{subsec:fuzzy-numbers}Fuzzy numbers and fuzzy intervals}
+A specific application of possibilistic variables is obtained when the universe under consideration is the set of boolean values $\mathbb{B}$ = $\{T,F\}$.
+Indeed, any boolean propositional variable $p$ takes just one value in $\mathbb{B}$.
+If the knowledge on which value this variable $p$ will take is given by a possibility distribution $\pi_p$, the variable can be seen as a possibilistic variable.
+As the interest lies with the case where the proposition holds, the possibility and necessity that $p$ = $T$ (the proposition holds) demands most attention.
+This possibility and necessity is noted here as:
+\begin{align}
+Pos(p) & = \pi_p(T) \label{propholdsposs} \\
+Nec(p) & = 1-\pi_p(F) \label{propholdsnecc}
+\end{align}
+Here, notation \ref{propholdsposs} denotes the possibility that $p$ = $T$ and the proposition holds, notation \ref{propholdsnecc} denotes the necessity that $p$ = $T$ and the proposition holds.
+
+\subsection{\label{subsec:fuzzy-numbers}Fuzzy Numbers and Fuzzy Intervals}
 Among others, Dubois and Prade~\cite{Dubois1983} use the following definition of a \emph{fuzzy interval}.
@@
 
+\def\JPicScale{0.5}
 \begin{figure}[h!]
   \centering
-  \includegraphics[scale=0.4]{graphs/trapezoidalDistribution.eps}
+  \input{./graphs/trapezoidal.tex}
   \caption{Trapezoidal membership function}
@@
   \centering
-  \includegraphics[scale=0.4]{graphs/triangular.eps}
+  \input{./graphs/triangular.tex}
   \caption{Triangular membership function.}
@@
 
-\subsubsection{Interval evaluation by ill-known constraints}
-The problem of the interval evaluation is explained in \cite{Pon11}: For a crisp interval $I = \left[ a, b \right]$, we want to know if all points in this interval reside between the boundaries of the ill-known interval $\left[ X , Y \right]$, where $X,Y$ are ill-known values on the set of real numbers $\mathbb{R}$, and $\lambda$ is the evaluation function.
+\subsubsection{Interval Evaluation by Ill-known Constraints}
+The problem of the interval evaluation is explained in \cite{Pon11}: for a crisp interval $I = \left[ a, b \right]$, the need exists to know if all points in this interval reside between the boundaries of the ill-known interval $\left[ X , Y \right]$, where $X,Y$ are ill-known values on the set of real numbers $\mathbb{R}$, and $\lambda$ is the evaluation function.
 

### 2011-11-21 ipmu2012/proposal.tex
Monday's update!

@@
 \subsection{Representation of Ill-known Valid-time Intervals}
+\label{subsec:representation-ill-known}
 Valid time is usually represented as an interval. The interval has a starting and an ending points. An ill-known valid-time interval is an interval in witch one or both points are ill-known.
@@
 \subsection{Storage of Valid-time Intervals}
+\label{subsec:storage}
 Each database row containing a \emph{PVP} stores it as two triangular possibility distributions.
@@
 It is important to notice that the possibilistic / fuzzy data stored in the database has a \emph{disjunctive interpretation} (it is said that we have \emph{uncertainty}: the valid-time interval has only one value but, for some reason the value is ill-known).
-In the query specification, the user is allowed to express \emph{imprecision} or \emph{vagueness} (this is referred as a \emph{conjunctive interpretation}).
+In the query specification, the user is allowed to express a crisp time interval.
+In this approach, the interpretation is the following: the user w
 In this subsection we will define the query specification, then the evaluation of the query and finally the ranking for the query.
@@
 \begin{equation}
+\label{eq:query-definition}
 \tilde Q = \left( Q^{time}, Q \right)
@@
 \end{definition}
-Where $Q$ are the (possibly fuzzy) preferences of the user  and $Q^{time}$ is the temporal part specified by a possibilistic valid-time period,\emph{PVP}.
+Where $Q$ are the (possibly fuzzy) preferences of the user  and $Q^{time}$ is the temporal part specified by a crisp interval. The temporal part is composed by two arguments:
 
-%explain the sematics of vt.
+\begin{definition}
+ $Q^{time}$ is composed by:
+\begin{equation}
+Q^{time} = \left( \left[a, \ b \right] , Ar \right), a,b \in \mathbb{R}
+\end{equation}
+Where $ \left[a, \ b \right] $ is the specification for the crisp interval and $Ar$ is one of the Allen's relations (table \ref{tab:allen-relations}).
+\end{definition}
+
+\begin{table}[h]
+\centering
+\begin{tabular}{|c|c|c|}
+\hline
+Allen Relation & Constraints & $\bool(p_1,...,p_n)$ \\
+\hline
+I before J & $C_1\stackrel{\triangle}{=} \left(<,X\right)$ & $p_1$ \\
+\hline
+\multirow{4}{*}
+{I equal J} & $C_1\stackrel{\triangle}{=} \left(\geq,X\right)$ & \text{$p_1\wedge\neg p_2\wedge p_3\wedge\neg p_4$}\\
+ & $C_2\stackrel{\triangle}{=} \left(\neq,X\right)$ & \\
+ & $C_3\stackrel{\triangle}{=} \left(\leq,Y\right)$ & \\
+ & $C_4\stackrel{\triangle}{=} \left(\neq,Y\right)$ & \\
+\hline
+\multirow{2}{*}
+{I meets J} & $C_1\stackrel{\triangle}{=} \left(\leq,X\right)$ & $p_1\wedge\neg p_2$\\
+ & $C_2\stackrel{\triangle}{=} \left(\neq,X\right)$ & \\
+\hline
+\multirow{3}{*}
+{I overlaps J} & $C_1\stackrel{\triangle}{=} \left(<,Y\right)$ & $p_1\wedge\neg p_2\wedge\neg p_3$\\
+ & $C_2\stackrel{\triangle}{=} \left(\leq,X\right)$ & \\
+ & $C_3\stackrel{\triangle}{=} \left(\geq,X\right)$ & \\
+\hline
+\multirow{4}{*}
+{I during J} & $C_1\stackrel{\triangle}{=} \left(>,X\right)$ & $(p_1\wedge p_2)\vee(p_3\wedge p_4)$\\
+ & $C_2\stackrel{\triangle}{=} \left(\leq,Y\right)$ & \\
+ & $C_3\stackrel{\triangle}{=} \left(\geq,X\right)$ & \\
+ & $C_4\stackrel{\triangle}{=} \left(<,Y\right)$ & \\
+\hline
+\multirow{2}{*}
+{I starts J} & $C_1\stackrel{\triangle}{=} \left(\geq,X\right)$ & $p_1\wedge\neg p_2$\\
+ &  $C_2\stackrel{\triangle}{=} \left(\neq,X\right)$ & \\
+\hline
+\multirow{2}{*}
+{I finishes J} & $C_1\stackrel{\triangle}{=} \left(\leq,Y\right)$ & $p_1\wedge\neg p_2$\\
+ & $C_2\stackrel{\triangle}{=} \left(\neq,Y\right)$ & \\
+\hline
+\end{tabular}
+\label{tab:allen-relations}
+\caption{Allen's relations represented in the framework.}
+\end{table}
 
@@
 \item
-For each record $r$ in the database, the preferences expressed in $Q$ are evaluated in the unit interval.
+For each record $r$ in the database, the preferences expressed in $Q$ are evaluated in the unit interval. Thus, $e(Q) \in \left[0, 1\right]$ is the evaluation function for the criteria in $Q$.
+\item
+For each record $r$ in the database, the interval $ \left[a, \ b \right]$ is compared to the $PVP$ value by means of the corresponding Allen relation specified by $ar$.
+The result is again in the unit interval.
+Here, $e(Q^{time}) \in \left[0, 1 \right]$ is the evaluation function for the criteria $Q^{time}$.
 \end{itemize}
@@
 \subsubsection{Ranking}
+In order to present the results to the user, a ranking method must be used.
+Both $Q$ and $Q^{time}$ are evaluated independently and also both are evaluated in the unit interval.
+The ranking is computed using a convex combination of both values:
+
+\begin{equation}
+e(\tilde{Q})\ =\ \omega*e(Q)\ +\ (1-\omega)*e(Q^{time})
+\end{equation}
+Where $\omega \in \left[0, 1 \right]$.
+By increasing the value for the parameter $\omega$, the preferences expressed in the query $Q$ can be given more importance.
+By lowering the value for $\omega$, the temporal criteria is emphasized.
+The following example illustrates the querying process.
+
+\begin{example}
+
+\textbf{The database}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Sample table for the car models. 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{table}[ht]
+\caption{Sample database containing the car models.}
+\centering
+\begin{tabular}{c c c c c c c}
+\hline
+ID & IID & Segment & Manufac. & Name & Start & End  \\ [0.5ex]
+\hline
+001 & 1 & B & Peugeot & 205 & [1985,2,3] & [1997,2,1] \\
+002 & 1 & C & Peugeot & 305 & [1977,2,2] & [1989,2,3] \\
+003 & 1 & B & Citroen & C2 & [2003,3,2] & [2006,1,1] \\
+001 & 2 & B & Peugeot & 206 & [1998,1,1] & [2011,2,1] \\
+001 & 3 & B & Peugeot & 207 & [2006,1,1] & [2011,1,1]\\
+\hline
+\end{tabular}
+\label{tb:car-models}
+\end{table}
+
+Consider a database containing car models.
+There are several general attributes (car model's name, car manufacturer, car segment) and one temporal attribute (which is ill-known) containing the approximate date while the car model was sold.
+The temporal data is stored as explained in subsection \ref{subsec:storage}.
+The value for $D$ is stored in \emph{yyyy} format and $a$ and $b$ are represented by an integer, for the convenience of the representation.
+Therefore the chronons in our example will be years.
+The ID field identifies a car model while the field Instance ID (IID) identifies the instance for a car model.
+\\
+
+\textbf{The query}
+Consider the following query:
+
+\emph{The user wants to obtain a list of models in the segment B for the manufacturer Peugeot before the year interval 2001-2005.}
+
+The query is translated to the following expression, using the structure given in equation \eqref{eq:query-definition}:
+
+\begin{center}
+$\tilde{Q} = \left(  c^{time}, c_{segment} \right)$
+\end{center}
+
+Where:
+\begin{itemize}
+\item
+$c^{time}\ = \ ( \left[ 2001,\ 2005 \right],\ $  before $)$.
+\item
+$c_{segment}\ = \ $ Peugeot.
+\end{itemize}
+
+The evaluation for the criteria are presented in the result table \ref{tb:results}.
+
+\begin{table}[ht]
+\caption{Result table and ranking}
+\centering
+\begin{tabular}{c c c c c }
+\hline
+ID & IID &  $c^{time}$ & $c_{segment}$ & rank ($\omega=0.5$) \\ [0.5ex]
+\hline
+001 & 1  & 1 & 1 & 1 \\
+002 & 1  & 1 & 0 & 0.5 \\
+003 & 1  & 0.333  & 0 & 0.166\\
+001 & 2 & 1 & 1 & 1 \\
+001 & 3 & 0 & 1 & 0.5\\
+\hline
+\end{tabular}
+\label{tb:results}
+\end{table}
+\end{example}

### 2011-11-23 estylf2012/estylf2012_en.tex
Estylf paper

\documentclass[twoside,twocolumn,a4paper]{article}

\usepackage[latin1]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{mathptmx}
\usepackage{graphicx}
\usepackage{estylf_en}

% In order to work with mathematical environments (propositions, theorems, etc.), the following packages are suggested:
%\usepackage{amsmath,amsthm}
\usepackage{amsmath, amsthm}
\usepackage{amsfonts}

\hyphenation{pro-per-ties}
\hyphenation{ge-ne-ral-ly}
\hyphenation{pre-fe-ren-ces}
\hyphenation{u-sing}
\hyphenation{pu-nish-ment}

\newcommand{\Pow}{\mathcal{P}}
\newcommand{\N}{\operatorname{N}}
\newcommand{\bool}{\operatorname*{\mathcal{B}}}
\newcommand{\Pos}{\operatorname{Pos}}
\newcommand{\Nec}{\operatorname{Nec}}
\newcommand{\T}{\mathcal{T}}

\begin{document}

\title{Possibilistic evaluation of fuzzy temporal intervals}

%*******************************************************
% REMARKS ABOUT THE AUTHORS
%*******************************************************
% In order to write the name and surname of an author in the the same line, the symbol ~ will be required.
%*******************************************************
% If all the authors share the same affiliation, the following commands are suggested to avoid the superscript:
% \author{Name1~Surname1}
% \address{Address1}{autor1@e-mail1.com}
%*******************************************************

\author[1]{Jos\'e Enrique~Pons}
\author[2]{Antoon~Bronselaer}
\author[1]{Olga~Pons}
\author[2]{Guy~De Tr\'e}
\address[1]{Department of Computer Science and Artificial Intelligence,University of Granada, Spain}{\{jpons,opc\}@decsai.ugr.es}
\address[2]{Department of Telecommunications and Information Processing, Ghent University, Belgium}{\{Antoon.Bronselaer,Guy.De.Tre\}@telin.ugent.be}

\maketitle

\begin{abstract}
Possibility theory tries to quantify the uncertainty caused by incomplete information. A specific case of incomplete information is that of ill-known temporal intervals. The study of these fuzzy intervals is of particular interest in temporal database research. In order to optimize the storage of fuzzy temporal intervals, some transformations have been proposed. In this paper we analyze the possibilistic evaluation of the ill-known temporal intervals and the proposed transformations. It is shown how the reasoning behind the transformations implies a possibilistic information lost.

\textbf{Keywords:} Possibility theory, ill-known values, ill-known sets, intervals, temporal databases.\vspace{9mm}
\end{abstract}

\section{INTRODUCTION}

The connection of possibility theory \cite{Zadeh78},\cite{Shackle61},\cite{Gaines75} and fuzzy sets has been studied among others by Zadeh \cite{Zadeh65}. This connection establishes two different natures for a fuzzy set \cite{Dubois97}:
\begin{itemize}
\item
\textbf{Conjunctive nature}: The fuzzyfication of a regular set. This interpretation corresponds with the following two semantics: \textit{Degree of preference} and \textit{degree of similarity}.
\item
\textbf{Disjunctive nature}: In this case, the disjunctive nature indicates a description of incomplete knowledge. This interpretation corresponds with the semantics for the \textit{degree of uncertainty}.
\end{itemize}

The disjunctive nature can be seen as an additional connection between fuzzy set theory and possibility theory. Dubois and Prade \cite{Dubois88b} have studied the problem of inferring uncertainty about sets from uncertainty about values. They explain how the incomplete knowledge about a set can be represented as a possibility distribution over a universe $\Pow \left(U \right)$, the power set over the universe $U$. The main drawback of the approach is the large overhead that such a distribution introduces.  A mechanism for the propagation of uncertainty about set elements towards uncertainty about sets is introduced in \textit{'Possibilistic evaluation of sets'} \cite{Pon11}.

There are several proposals in the literature for inferring uncertainty about sets from uncertainty about values. It is of particular interest for fuzzy temporal databases, the application on temporal intervals. The transformation from two ill-known temporal points to a ill-known set has been discussed in \cite{Garrido2009}. In this work we will analyze the major pitfalls of these approaches and the benefit of using the mechanism proposed in \cite{Pon11}.

The paper is organized as follows. In Section \ref{sec:preliminaries}, some preliminaries and concepts about the possibilistic evaluation framework and the temporal databases are presented. The proposal for interval evaluation is done in Section \ref{sec:interval-eval}. In Section \ref{sec:analysis}, the analysis among the different transformations is done. Finally, Section \ref{sec:conclusions} presents the main conclusions and some lines for the future work in this research.

\section{\label{sec:preliminaries}PRELIMINARIES}
In this section we will introduce some basic concepts. First of all, an introduction to possibility theory is explained. Then, the concepts of fuzzy numbers and fuzzy intervals introduced by Dubois and Prade \cite{Dubois1983} are discussed. Finally, a brief introduction to temporal databases is done.

\subsection{\label{subsec:possibilistic-variables}Possibilistic variables}
A possibilistic variable $X$ over a universe $U$ is defined as a variable taking exactly one value in $U$, but for which this value is (partially) unknown. The possibility distribution $\pi_X$ gives the available knowledge about the value that $X$ takes. For each $u\in U$, $\pi_X(u)$ represents the possibility that $X$ takes the value $u$.

It is important to understand the difference between the following two concepts:
\begin{itemize}
\item
A \emph{possibilistic variable} $X$ is bounded to take only one value , but this value is not known due to incomplete knowledge. 
\item
An \emph{ill-known set}~\cite{Dubois88b}: a possibilistic variable defined over the universe $\Pow(U)$.
\end{itemize}

Note that while a possibilistic variable refers to one (partially) unknown value, an ill-known set is a crisp set but, for some reason, (partially) unknown.

\subsection{\label{subsec:fuzzy-numbers}Fuzzy numbers and fuzzy intervals}
Dubois and Prade~\cite{Dubois1983} proposed the following definition of \emph{fuzzy interval}.
%\begin{definition}
A fuzzy interval is a fuzzy set $M$ on the set of real numbers $\mathbb{R}$ such that:
\begin{eqnarray}
\forall (u,v)\in\mathbb{R}^2:&\\
\nonumber
\forall w \in [u,v]:&\mu_M(w) \geq\min(\mu_M(u),\mu_M(v))  \\
\exists m \in \mathbb{R}:& \mu_M(m)=1 
\end{eqnarray}
%\end{definition}
If $m$ is unique, then $M$ is referred to as a \emph{fuzzy number}, instead of a \emph{fuzzy interval}. In other words, if the core of a fuzzy interval is a singleton, it can be seen as a fuzzy number. In their work, Dubois and Prade propose four different functions (two possibility and two necessity functions) to asses the position of a fuzzy number N relative to  a fuzzy number M taken as a reference.

The most convenient representation for the membership function of a fuzzy number is a triangular function. The membership function $\mu_M$ for the fuzzy set $M$ has also the properties of convexity and normalization. Three values represent a triangular function (see Figure~\ref{fig:triangular}):
\begin{itemize}
\item
$D$ is the singleton value in the core of $\mu_M(x)$.
\item
$D-a$ is the lower value in the support of $\mu_M(x)$. 
\item
$D+b$ is the upper value in the support of $\mu_M(x)$.
\end{itemize}
Note that the values $a$ and $b$ are values in the underlying ordered domain. E.g. $(a,b) \in \mathbb{R}^2$. The notation for a triangular function adopted here is $[D,a,b]$.
\begin{figure}[h!]
  \centering
  \includegraphics[scale=0.4]{graphs/triangular.pdf}
  \caption{Triangular function.}
  \label{fig:triangular}
\end{figure}

The most simple representation for the membership function of a fuzzy interval is a trapezoidal function. This membership function $\mu_T$ for the fuzzy interval $T$ is convex and normalized. Four values represent a trapezoidal membership function (figure  \ref{fig:trapezoidal}): $\left[\alpha,\ \beta,\ \gamma,\ \delta\right]$. The membership function is given by:

\begin{align}
\mu_T(x)
\begin{cases}
1 & \mbox{ if } x \in [\beta,\gamma] \\
0 & \mbox{ if } x > \delta \vee x < \alpha \\
\frac{x-\alpha}{\beta - \alpha} & \mbox{ if } x \in [\alpha,\beta[ \\
\frac{\delta -x}{\delta - \gamma} & \mbox{ if } x \in ]\gamma,\delta] \\
\end{cases}
\end{align}

\begin{figure}[h!]
  \centering
  \includegraphics[scale=0.4]{graphs/trapezoidalDistribution.pdf}
  \caption{Trapezoidal distribution.}
  \label{fig:trapezoidal}
\end{figure}

\subsection{\label{subsec:temporal}Temporal databases}
%In this section, the proposed reasoning is applied to the specific context of intervals on the real line. This setting is of specific interest in the context of fuzzy temporal databases.
A temporal database \cite{Dyreson1994} is a database that manages some aspects of time in its schema. A \textbf{chronon} is the shortest duration of time supported by the database. The time can be represented either as points or intervals \cite{655777}. Fuzzy temporal models \cite{4481150} are proposed when the time points or intervals are not precisely known. There are proposals for the  fuzzyfication of the time point \cite{Dubois89} and the fuzzyfication of the time interval \cite{Garrido2009}.\\
Allen \cite{Allen:1983:MKT:182.358434} studied the comparison between two crisp time intervals. For fuzzy intervals, several proposals \cite{4481150}, \cite{springerlink:10.1007/978-3-540-39964-3_57},\cite{10.1109/TIME.2004.1314418} have been done.

%Time granularity is also associated with the representation of the time. A granularity is the result of partitioning on the set of chronons. The conversion among granularities is a common issue within temporal databases \cite{Lin97efficientconversion}. Granularity is the basis of some systems \cite{Cru97},\cite{624013}.

Despite of user-defined time (an uninterpreted attribute; supported in the standard SQL2 \cite{Mel93}), there are 3 types of time in a temporal database:

\begin{itemize}
	\item
	\textbf{Transaction time}: The time when the fact is stored in the database.
	\item
	\textbf{Valid time}: The time when the fact is true in the modelled reality.
	\item
	\textbf{Decision time} \cite{Nascimento95decisiontime}: The time when an event was decided to happen. 
	\end{itemize}

	 The  database model can be then classified into: transaction time \cite{Ston87},\cite{Jensen:1991:IIM:627283.627484}, Valid time, bi-temporal \cite{Snodgrass:1984:TQL:588011.588041}(both valid and transaction time) or tri-temporal \cite{Nascimento95decisiontime} (valid, transaction and decision time).

Fuzzy temporal models \cite{4481150} deal with time points \cite{Dubois89} or intervals \cite{Garrido2009} that are ill-known. Some fuzzy temporal models assume that the time stored in the database is an interval. The temporal interval is represented by two ill-known time points: $X$  an ill-known starting point and $Y$ an ill-known ending point. The interval $\left[X,Y\right]$ is not a fuzzy interval but an ill-known interval: it is a crisp interval but it is partially unknown which values are in this interval.

% In such databases, points in time can be ill-known, i.e. the starting, the ending or both points are ill-known. a point in time is modelled as an ill-known value. When two such ill-known points are given, an interesting problem is the inference of uncertainty about the interval that is enclosed by these two ill-known points. In the proposed framework, this problem can be solved by evaluating an interval against two ill-known constraints, where the binary relations are ordering relations on the set of real numbers.

\section{\label{sec:interval-eval}INTERVAL EVALUATION BY ILL-KNOWN CONSTRAINTS}
The problem of the interval evaluation is explained in \cite{Pon11}: For a crisp interval $I = \left[ a, b \right]$, we want to know if all points in this interval reside between the boundaries of the ill-known interval $\left[ X , Y \right]$, where $X,Y$ are ill-known values on the set of real numbers $\mathbb{R}$, and $\lambda$ is the evaluation function. The following expressions compute the possibility and the necessity:

\begin{eqnarray}
\label{eq:interval-pos}
\Pos\left(\lambda([a,b])\right)=\\
\nonumber
\min\bigg(\sup_{a\geq w}\pi_{X}(w),\sup_{b\leq w}\pi_{Y}(w)\bigg)\\
\label{eq:interval-nec}
\Nec\left(\lambda([a,b])\right)=\\
\nonumber
\min\bigg(\inf_{a<w}1-\pi_{X}(w),\inf_{b>w}1-\pi_{Y}(w)\bigg).
\end{eqnarray}

\paragraph{Example} Consider the ill-known values $X = \left[5, 2, 8\right]$ and $Y = \left[9, 7, 10 \right]$. The knowledge about the evaluation of the interval $\left[a, b \right]$  is given by the expressions \eqref{eq:interval-pos},\eqref{eq:interval-nec}.  Figure~\ref{fig:3d-possibility} shows a 3D plot of the possibility that an interval $[a,b]$ passes the evaluations specified by the ill-known constraints. Note the triangular form for the resulting possibility distribution since the condition $a \leq b$ holds.

\begin{figure}[h!]
\centering
\includegraphics[scale=0.4]{graphs/3D_possibility.pdf}
\caption{Possibility of evaluation for the interval $[a,b]$.}
\label{fig:3d-possibility}
\end{figure}
The necessity plot is obtained in a similar way and is shown in Figure~\ref{fig:3d-necessity}. Notice that the necessity measure is not normalized because the supports of $X$ and $Y$ overlap.
\begin{figure}[h!]
\centering
\includegraphics[scale=0.4]{graphs/3D_necessity.pdf}
\caption{Necessity of evaluation for the interval $[a,b]$.}
\label{fig:3d-necessity}
\end{figure}

\section{\label{sec:analysis}ANALYSIS OF PROPOSED TRANSFORMATIONS}

There are several proposals to transform two fuzzy numbers that represent an ill-known interval into one fuzzy interval. This reasoning is however not consistent with possibility theory. We provide several arguments to prove this, which have been previously reasoned by Dubois and Prade \cite{Dubois88b}:

\begin{enumerate}
\item
The fuzzy set given by the two fuzzy numbers is conceptually different from a fuzzy set represented by the fuzzy interval. The fuzzy set given by the two fuzzy numbers belong to $\Pow(\mathbb{R})$ while the fuzzy interval obtained after the transformation is not a possibility distribution on $\Pow(\mathbb{R})$  but on $\mathbb{R}$.
\item
In some cases, (like in the convex hull transformation) because of the underlying ordered domain, the membership function provided for the fuzzy interval preserves the possibility. However, the necessity is not preserved at all.
\end{enumerate}

In the next subsections we will analyze the two major approaches: The transformation that preserves the imprecision and the convex hull.

\subsection{\label{subsubsec:transf-pres-imp}Transformation that preserves the imprecision}
\begin{figure}[h!]
  \centering
  \includegraphics[scale=0.2]{graphs/preservingImprecision.pdf}
  \caption{Transformation of two fuzzy numbers into one fuzzy interval preserving the amount of imprecision.}
  \label{fig:transf-pres-imp}
\end{figure}
This transformation has been proposed in \cite{Garrido2009}. They assume that the imprecision of a fuzzy set is its area. Therefore the transformation consists of preserving the area of both fuzzy numbers that compose the interval. The transformation is performed by means of geometrical computations. The starting point of the fuzzy temporal interval is given by the fuzzy number $S = [d_s,a_s,b_s]$ respectively, the ending point of the fuzzy interval is given by $E = [d_e,a_e,b_e]$ (see Figure \ref{fig:transf-pres-imp}).

The transformation defines a fuzzy interval $T$ by means of a trapezoidal membership function. In order to define the membership function for the fuzzy interval, the following two equalities between the areas from the fuzzy numbers and the fuzzy interval are forced: $S1=S2$ and $S3=S4$. The calculation for the first equality, based on triangle equivalence is:

\begin{equation}
\label{eq:preserving-imprecision}
S1=S2 \Rightarrow \frac{\left(d_s+b_s\right)-\left(d_s-a_s\right)}{2}
\end{equation}        			
The calculation for the right side ($S3=S4$) is analogous.
This transformation is specially widely used when transforming fuzzy time points into fuzzy intervals. The reasoning is the following: the core of the transformed fuzzy interval corresponds with the time when the fact happened. The left and right slopes of the possibility distribution show the imprecision around the starting and the ending times.

The membership function for this fuzzy interval T is represented by the trapezoid $\left[\alpha,\ \beta,\ \gamma,\ \delta\right]$:

\begin{align}
\nonumber
\alpha & =  d_s - a_s \\
\nonumber
\beta & = d_s + b_s \\
\nonumber
\gamma &= d_e - a_e \\
\nonumber
\delta &= d_e + b_e 
\end{align}

Note that when the supports of the two fuzzy numbers overlap, this transformation makes no sense.

%\paragraph{Comparison}
%In this subsection we will study the knowledge we have about the interval $[u,v]$ w.r.t. the fuzzy interval $T$  which is built preserving the imprecision as explained in Section \ref{subsubsec:transf-pres-imp}. Note that this approach is only valid when the supports of the original fuzzy numbers $X$ and $Y$ are disjoint (see Example \ref{ex:preserving-imprecision} below). Now, the knowledge we have about a number $u$ w.r.t. the transformated fuzzy interval $T$ is:
%\begin{eqnarray}
%\Pos\left(u\in T\right) &=& \mu_{T}\left(u\right) \\
%\Nec\left(u\in T\right) &=& 1 - \mu_{T}\left(u\right).
%\end{eqnarray}
%Therefore, the knowledge about an interval $[u,v]$ w.r.t. $T$ is:
%\begin{eqnarray}
%\nonumber
%\Pos\left([u,v]=T\right) &= \min \left(\Pos\left(u\in T\right),\Pos\left(v\in T\right)\right) &= \min \left(\mu_{T}\left(u\right),\mu_{T}\left(v\right)\right) \\
%\nonumber
%\Nec\left([u,v]=T\right) &= \min \left(\Nec\left(u\in T\right),\Nec\left(v\in T\right)\right) &= \min \left(1 - \mu_{T}\left(\overline{u},1 - \mu_{T}\left(\overline{v}\right)\right)\right)
%\end{eqnarray}

%\begin{example}
%\label{ex:preserving-imprecision}
\paragraph{Example}
In this example we are going to transform the two ill-known points $X= \left[5, 3, 3 \right]$ and $Y= \left[9, 2, 1 \right]$ into one fuzzy interval $T$ using the preserving the imprecision approach. Again, the two fuzzy numbers are represented by means of two triangular possibility distributions:

Then, following the equation presented in \eqref{eq:preserving-imprecision} we obtain the following trapezoidal shape:
\begin{equation}
\mu_{T}\left(x\right) = [d_s-a_s,d_s+b_s,d_e-a_e,d_e+b_e] = [2,8,7,10]
\end{equation}
This construction make no sense when the supports of the fuzzy numbers are disjoint. It is easy to build a rational membership function just by ordering the elements in the representation (e.g. $[2,7,8,10]$) but then, the amount of imprecision is not preserved at all. 
%\end{example}

\subsection{\label{subsubsec:trans-convex-hull}Transformation based on the convex hull}
The transformation based on the convex hull is more intuitive than the previous one. The transformation builds a trapezoidal possibility distribution with the left slope equal to the left slope of the starting fuzzy number (in the graphic, from $d_s-a_s$ to $d_s$) , the core equal to the interval $[d_s,d_e]$ and the right slope equal to the right slope of the ending fuzzy number (from $d_e$ to $d_e+b_e$). The interpretation of this convex hull is the ``most possible'' starting point and the ``most possible'' ending point. Figure \ref{fig:convexHull} shows the resulting transformation.
\begin{figure}[h!]
  \centering
  \includegraphics[scale=0.4]{graphs/convexHull.pdf}
  \caption{Transformation of two fuzzy numbers into one fuzzy interval with the convex hull approach.}
  \label{fig:convexHull}
\end{figure}

In this approach, the trapezoidal membership function ($\left[\alpha,\ \beta,\ \gamma,\ \delta\right]$) is built in a different way:

\begin{align}
\nonumber
\alpha & =  d_s - a_s \\
\nonumber
\beta & = d_s \\
\nonumber
\gamma &= d_e  \\
\nonumber
\delta &= d_e + b_e 
\end{align}
%\paragraph{Comparison}
%Now, we will compare the knowledge we have about the interval $[u,v]$ w.r.t. the fuzzy interval $T$, which is the result of the convex hull transformation between the two fuzzy numbers $X$ and $Y$. The membership degree for the convex hull is given by:

%\begin{align}
%\label{eq:ch-possibility}
%\Pos\left([u,v]=T\right) &= \Pos\left(p^{\left(1\right)}\right) \cap \Pos\left(p^{\left(4\right)}\right) \\ 
%&= \min \left( \sup_{w \leq d}\ \mu_X\left(w\right) , \sup_{w \geq d}\ \mu_Y\left(w\right)\right)
%\end{align}

%The necessity measure for the convex hull is not consistent with the necessity measure for the individual fuzzy numbers $X$ and $Y$. This transformation results in an information lost.

%The necessity measures for the following predicates:

%\begin{align}
%\Nec\left(p^{\left(1\right)}\right) &=  \inf_{w \geq d} 1 - \mu_T\left(w\right) \\
%\Nec\left(p^{\left(4\right)}\right) &=  \inf_{w \leq d} 1 - \mu_T\left(w\right)
%\end{align}

%Therefore the necessity for the interval $[u,v]$ to be inside the fuzzy interval $T$ is:

%\begin{align}
%\label{eq:ch-necessity}
%\Nec\left([u,v]=T\right) &= \min \left(\Nec\left(p^{\left(1\right)}\right), p^{\left(4\right)}\right) \\
%& \min \left(\inf_{w \geq d} 1 - \mu_T\left(w\right),inf_{w \leq d} 1 - \mu_T\left(w\right)\right)
%\end{align}

%\begin{example}
\paragraph{Example}
In this example we are going to build the convex hull for the two ill-known starting and ending points $X= \left[5, 3, 3\right]$ and $Y = \left[9, 2, 1\right]$.

\begin{figure}[h!]
  \centering
 % \includegraphics[scale=0.2]{graphs/convexHull/convexHull.pdf}
 \input{scripts/convexhull.tex}
  \caption{Construction of the convex hull $T$ w.r.t the ill-known values $X$ and $Y$.}
  \label{fig:convex-hull-T}
\end{figure}

The convex hull for the fuzzy numbers $X$ and $Y$, represented by means of two triangular possibility distributions results in a trapezoidal possibility distribution with membership function $\mu_T(x)$ (see Figure \ref{fig:convex-hull-T}), represented by $[2,5,9,10]$.

%\begin{figure}[h!]
%  \centering
%  \includegraphics[scale=0.2]{graphs/convexHull/possAfterBefore.eps}
%  \caption{Possibility distributions for  both $u$ possibly greater or equal to $T$ and $u$ possibly smaller or equal to $T$.}
%  \label{fig:convex-hull-pos}
%\end{figure}

%\begin{align}
%\label{eq:ch-possibly-greater}
%\Pos\left(u \geq T\right) & = \sup_{d \leq u} \mu_{T}\left(d\right)  = [2,5,9,+\infty] \\
%\label{eq:ch-possibly-smaller}
%\Pos\left(u \leq T\right) & = \sup_{d \geq u} \mu_{T}\left(d\right)  = [-\infty,5,9,10] 
%\end{align}

%Respectively, the necessity mesaures for a point $u$ w.r.t. the convex hull $T$ are the following (Figure \ref{fig:convex-hull-nec}):

%\begin{align}
%\Nec\left(u \geq T\right) &=  1 - \Pos\left(u \geq T\right) = 1 - \sup_{d \geq u} \mu_{T}\left(d\right) = [9,10,+\infty,+\infty] \\ 
%\Nec\left(u \leq T\right) & = 1 - \Pos\left(u \geq T\right)  = \sup_{d \leq u} \mu_{T}\left(d\right)  = [-\infty,-\infty, 2,5]
%\end{align}

%\begin{figure}[h!]
%  \centering
%  \includegraphics[scale=0.2]{graphs/convexHull/necessity.eps}
%  \caption{Necessity distributions for both $u$ is necessarily greater or equal to $T$ and $u$ is necessarily smaller or equal to $T$.}
%  \label{fig:convex-hull-nec}
%\end{figure}

The membership about an interval $[u,v]$ w.r.t the convex hull $T$ is given by:

\begin{equation}
\label{eq:membership-ch}
\mu_T([u,v]) = \min \left( \mu_T\left(u\right), \mu_T\left(v\right)\right)
\end{equation}

Figure \ref{fig:ch-3d-possibility} shows the 3D plot of the membership function for any interval $[u,v]$ w.r.t. the convex hull $T$ (equation \eqref{eq:membership-ch}). Note that figure Figure \ref{fig:ch-3d-possibility} is equivalent to figure \ref{fig:3d-possibility}. Due to the transformation, it is impossible to obtain a necessity measure.

\begin{figure}[h!]
  \centering
  \includegraphics[scale=0.4]{graphs/convexHull/3dpossibility.pdf}
  \caption{Membership function for interval $[u,v]$ w.r.t. the convex hull $T$.}
  \label{fig:ch-3d-possibility}
\end{figure}

%\begin{figure}[h!]
%  \centering
%  \includegraphics[scale=0.4]{graphs/convexHull/3dnecessity.eps}
%  \caption{Necessity of any interval $[u,v]$ w.r.t. the convex hull $T$.}
%  \label{fig:ch-3d-necessity}
%\end{figure}

%\end{example}

\section{\label{sec:conclusions}CONCLUSIONS AND FUTURE WORK}
The problem of evaluation of sets in the framework of possibility theory has been addressed by Dubois and Prade. They conclude that a correct treatment of sets imply an overly complex possibility distribution. Based on the previous work from the authors, this work is focused on the evaluation of fuzzy temporal interval in the framework of possibility theory. The aim of the paper is to clarify the difference between fuzzy numbers, fuzzy intervals and ill-known intervals. Several proposals can be found in the literature with respect to transforming two fuzzy numbers into a fuzzy interval. We have shown that these techniques are not fully correct in the sense of possibility theory.\\

%Submitted full papers are to be 4 to 6 pages length. Papers must be written on two columns with an overall width of 17.2 cm (8.25 cm each column and 0.7 cm of space between columns). On odd pages, left margin should be of 2.5 cm while the right one should be of 1.3 cm. On even pages, margins are just the contrary (left 1.3 cm, and right 2.5 cm). On all pages, odd and even ones, upper margin should be of 2.5 cm, and lower margin should be of 3.5 cm. Text alignment must be \textit{justify}. The pages must not be numbered.
%
%LaTeX style file (estylf\_en.sty) and the example file
%(estylf2012\_en.tex) are available in the ESTYLF2012 web page.
%
%\section{PRELIMINARIES}
%
%\subsection{SECTIONS AND SUBSECTIONS}
%
%\subsubsection{Subsection}
%All subsections must follow the same format as this one.
%
%\subsection{EQUATIONS}
%All equations must be centered.
%\begin{equation}
%\mu_A = f(a,b,c).
%\end{equation}
%
%\subsection{FIGURES}
%All figures must be centered like Figure~\ref{figure1}. The number and caption of the figure must always appear below the figure.

%\begin{figure}[h]
%\begin{center}
%\includegraphics[width=8cm,height=2.5cm]{estylf2012.jpg}
%\end{center}
%\caption{Example of figure caption.}
%\label{figure1}
%\end{figure}

%\subsection{TABLES}
%All tables must be centered and clear. The number and title always appear above the table (see Table~\ref{table1}).
%
%\begin{table}[ht]
%\begin{center}
%\caption{Example of table title.}
%\label{table1}
%\begin{tabular}{|c|l|} \hline
%\textbf{NAME} & \textbf{DESCRIPTION} \\ \hline
%$A$ & Fuzzy set \\ \hline
%$\mu_{A}$ & Membership function \\
%\hline
%\end{tabular}
%\end{center}
%\end{table}

%\subsection{FOOTNOTES}
%Footnotes must be numbered\footnote{Example of footnote.}, and placed at the bottom of the column where they appear.
%
%\subsection{CITATIONS}
%Citations must include the reference number between square brackets, for example \cite{AlsinaT05}.

\begin{ack}
Part of this research is supported by the grant BES-2009-013805 within the research project TIN2008-02066: \emph{Fuzzy Temporal Information treatment in relational DBMS}.
\end{ack}

\bibliographystyle{acm}
\bibliography{biblio}

%\begin{thebibliography}{99}

%\bibitem{AlsinaT05}
%C. Alsina, E. Trillas: On the symmetric difference of fuzzy sets. \emph{Fuzzy Sets and Systems} 153, pp.~181--194, 2005.
%
%\bibitem{Fishburn73}
%P.C. Fishburn: \emph{Utility Theory for Decision Making}. Wiley, New York, 1970.
%
%\bibitem{Yager03}
%R.R. Yager: E-Z OWA weights. In: \emph{Proceedings of the 10th IFSA World Congress}, Istanbul (Turkey), pp.~39--42, 2003.
%
%\bibitem{Zadeh71}
%L.A. Zadeh: Towards a theory of fuzzy systems. In: R.E. Kalman, N. DeClaris (Eds.), \emph{Aspects of Network and System Theory}, Holt, Rinehart and Winston, New York, pp.~469--490, 1971.

%\end{thebibliography}

%References must be sorted by the author name.

\end{document}

### 2011-11-28 estylf2012/estylf2012_en.tex
update for the estylf

@@
 The membership function $\mu_M$ for the fuzzy set $M$ has also the properties of convexity and normalization.
-Three values represent a triangular function (see Figure~\ref{fig:triangular}):
+Three values represent a triangular function: %(see Figure~\ref{fig:triangular}):
 \begin{itemize}
@@
 Note that the values $a$ and $b$ are values in the underlying ordered domain. E.g. $(a,b) \in \mathbb{R}^2$. The notation for a triangular function adopted here is $[D,a,b]$.
-\begin{figure}[h!]
-  \centering
-  \includegraphics[scale=0.4]{graphs/triangular.pdf}
-  \caption{Triangular function.}
-  \label{fig:triangular}
-\end{figure}
+%\begin{figure}[h!]
+%  \centering
+%  \includegraphics[scale=0.4]{graphs/triangular.pdf}
+%  \caption{Triangular function.}
+%  \label{fig:triangular}
+%\end{figure}
 
@@
 This membership function $\mu_T$ for the fuzzy interval $T$ is convex and normalized.
-Four values represent a trapezoidal membership function (figure  \ref{fig:trapezoidal}): $\left[\alpha,\ \beta,\ \gamma,\ \delta\right]$.
-The membership function is given by:
+Four values represent a trapezoidal membership function %(figure  \ref{fig:trapezoidal}):
+ $\left[\alpha,\ \beta,\ \gamma,\ \delta\right]$. The membership function is given by:
 
-\begin{align}
-\mu_T(x)
-\begin{cases}
-1 & \mbox{ if } x \in [\beta,\gamma] \\
-0 & \mbox{ if } x > \delta \vee x < \alpha \\
-\frac{x-\alpha}{\beta - \alpha} & \mbox{ if } x \in [\alpha,\beta[ \\
-\frac{\delta -x}{\delta - \gamma} & \mbox{ if } x \in ]\gamma,\delta] \\
-\end{cases}
-\end{align}
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[scale=0.4]{graphs/trapezoidalDistribution.pdf}
-  \caption{Trapezoidal distribution.}
-  \label{fig:trapezoidal}
-\end{figure}
+%\begin{align}
+%\mu_T(x)
+%\begin{cases}
+%1 & \mbox{ if } x \in [\beta,\gamma] \\
+%0 & \mbox{ if } x > \delta \vee x < \alpha \\
+%\frac{x-\alpha}{\beta - \alpha} & \mbox{ if } x \in [\alpha,\beta[ \\
+%\frac{\delta -x}{\delta - \gamma} & \mbox{ if } x \in ]\gamma,\delta] \\
+%\end{cases}
+%\end{align}
+%
+%\begin{figure}[h!]
+%  \centering
+%  \includegraphics[scale=0.4]{graphs/trapezoidalDistribution.pdf}
+%  \caption{Trapezoidal distribution.}
+%  \label{fig:trapezoidal}
+%\end{figure}
 
@@
 
-% In such databases, points in time can be ill-known, i.e.
-the starting, the ending or both points are ill-known.
-a point in time is modelled as an ill-known value.
-When two such ill-known points are given, an interesting problem is the inference of uncertainty about the interval that is enclosed by these two ill-known points.
-In the proposed framework, this problem can be solved by evaluating an interval against two ill-known constraints, where the binary relations are ordering relations on the set of real numbers.
-
 \section{\label{sec:interval-eval}INTERVAL EVALUATION BY ILL-KNOWN CONSTRAINTS}
@@
 
-\paragraph{Example} Consider the ill-known values $X = \left[5, 2, 8\right]$ and $Y = \left[9, 7, 10 \right]$.
-The knowledge about the evaluation of the interval $\left[a, b \right]$  is given by the expressions \eqref{eq:interval-pos},\eqref{eq:interval-nec}.
-Figure~\ref{fig:3d-possibility} shows a 3D plot of the possibility that an interval $[a,b]$ passes the evaluations specified by the ill-known constraints.
-Note the triangular form for the resulting possibility distribution since the condition $a \leq b$ holds.
+Note that the expressions in \eqref{eq:interval-pos},\eqref{eq:interval-nec} cannot be simplified in a more efficient expressions. %repensar
+An ill-known point is given by a triangular membership function.
+The ill-known values $X$ and $Y$ are represented by the triangular membership functions $\left[D_X,a_X,b_X \right]$ and $\left[D_Y,a_Y,b_Y \right]$ respectively.
+The equation \eqref{eq:interval-pos} is derived to:
+
+\begin{eqnarray}
+\label{eq:interval-pos-triangular-a}
+\sup_{a\geq w}\pi_{X}(w)=
+\begin{cases}
+0 & \mbox{\ if\ } a \leq \left( D_X - a_X \right) \\
+1 & \mbox{\ if\ } a \geq D_X \\
+ \frac{a-D_X+a_X}{a_X} & \mbox{\ if\ } a \in \left(D_X-a_X,D_X \right)
+\end{cases}\\
+\label{eq:interval-pos-triangular-b}
+\sup_{b\leq w}\pi_{Y}(w)=
+\begin{cases}
+0 & \mbox{\ if\ } b \geq \left( D_Y + b_Y \right) \\
+1 & \mbox{\ if\ } b \leq D_Y \\
+ \frac{D_X+b_X-b}{b_X} & \mbox{\ if\ } b \in \left(D_Y,D_Y+b_Y \right) 
+\end{cases}
+\end{eqnarray}
+
+The necessity expression of \eqref{eq:interval-nec} for triangular membership functions is given by:
+
+\begin{eqnarray}
+\label{eq:interval-nec-triangular-a}
+\inf_{a<w}1-\pi_{X}(w)=
+\begin{cases}
+0 & \mbox{\ if\ } a \leq  D_X   \\
+1 & \mbox{\ if\ } a \geq \left( D_X+b_X \right) \\
+\frac{a}{b_X}- \frac{D_X}{b_X} & \mbox{\ if\ } a \in \left(D_X,D_X+b_X \right)
+\end{cases}\\
+\label{eq:interval-nec-triangular-b}
+\inf_{b>w}1-\pi_{Y}(w)=
+\begin{cases}
+0 & \mbox{\ if\ } b \geq  D_Y \\
+1 & \mbox{\ if\ } b \leq \left( D_Y - a_Y \right) \\
+ \frac{D_Y-b}{a_X} & \mbox{\ if\ } b \in \left(D_Y,D_Y+b_Y \right) 
+\end{cases}
+\end{eqnarray}
+
+\paragraph{Example} Consider the ill-known values $X = \left[3, 2, 1\right]$ and $Y = \left[7, 2, 3 \right]$.
+The knowledge about the evaluation of the interval $\left[a, b \right]$ w.r.t.
+$X$ and $Y$ is given by the expressions \eqref{eq:interval-pos},\eqref{eq:interval-nec}.
+Figure \ref{fig:interval-rep} shows the graphical representation for the ill-known values $X$ and $Y$.
+The representation for equations \eqref{eq:interval-pos-triangular-a},\eqref{eq:interval-pos-triangular-b},\eqref{eq:interval-nec-triangular-a},\eqref{eq:interval-nec-triangular-b} are shown respectively in figures \ref{fig:interval_membership_pos_a},\ref{fig:interval_membership_pos_b},\ref{fig:interval_membership_nec_a},\ref{fig:interval_membership_nec_b}.
+
+Consider now that the value for the interval $\left[a, b \right]$ is $\left[2, 6 \right]$. The evaluation for the possibility and necessity are:
+
+\begin{eqnarray}
+\label{eq:interval-pos}
+\nonumber
+\Pos\left(\lambda([3,6])\right)=\min(1,1)= 1\\
+\label{eq:interval-nec}
+\nonumber
+\Nec\left(\lambda([3,6])\right)=\min(1,0.5 ) = 0.5
+\end{eqnarray}
+
+The main difference between this proposal and the transformations is that this framework always allows to obtain both possibility and necessity measures for each temporal comparison.
+In this case, there are the following possibilities:
+
+\begin{itemize}
+\item
+$\Pos(\lambda([a,b])) = 0 \rightarrow \Nec(\lambda([a,b])) = 0$: The interval $[a,b]$ is not contained the fuzzy interval $[X,Y]$.
+\item
+$\Pos(\lambda([a,b])) > 0 $, $\Nec(\lambda([a,b])) = 0$: The interval $[a,b]$ is contained the fuzzy interval $[X,Y]$ but it is not inside the cores of both $X$ and $Y$.
+\item
+$\Pos(\lambda([a,b])) = 1 $, $\Nec(\lambda([a,b])) = 1$: The interval $[a,b]$ is contained the fuzzy interval $[X,Y]$ and it is inside the cores of both $X$ and $Y$.
+\end{itemize}
 
@@
 \centering
-\includegraphics[scale=0.4]{graphs/3D_possibility.pdf}
-\caption{Possibility of evaluation for the interval $[a,b]$.}
-\label{fig:3d-possibility}
+\input{./graphs/interval.tex}
+\caption{Triangular membership functions for the two ill-known time points $X$ and $Y$.}
+\label{fig:interval-rep}
 \end{figure}
-The necessity plot is obtained in a similar way and is shown in Figure~\ref{fig:3d-necessity}. Notice that the necessity measure is not normalized because the supports of $X$ and $Y$ overlap.
+
 \begin{figure}[h!]
 \centering
-\includegraphics[scale=0.4]{graphs/3D_necessity.pdf}
-\caption{Necessity of evaluation for the interval $[a,b]$.}
-\label{fig:3d-necessity}
+\input{./graphs/interval_membership.tex}
+\caption{Membership function for $\sup_{a\geq w}\pi_{X}(w)$.}
+\label{fig:interval_membership_pos_a}
+\end{figure}
+
+\begin{figure}[h!]
+\centering
+\input{./graphs/interval_membership_b.tex}
+\caption{Membership function for $\sup_{b\leq w}\pi_{Y}(w)$.}
+\label{fig:interval_membership_pos_b}
+\end{figure}
+
+\begin{figure}[h!]
+\centering
+\input{./graphs/interval_membership_nec.tex}
+\caption{Membership function for $\inf_{a<w}1-\pi_{X}(w)$.}
+\label{fig:interval_membership_nec_a}
+\end{figure}
+
+\begin{figure}[h!]
+\centering
+\input{./graphs/interval_membership_nec_b.tex}
+\caption{Membership function for $\inf_{b>w}1-\pi_{Y}(w)$.}
+\label{fig:interval_membership_nec_b}
 \end{figure}
@@
   \centering
-  \includegraphics[scale=0.2]{graphs/preservingImprecision.pdf}
-  \caption{Transformation of two fuzzy numbers into one fuzzy interval preserving the amount of imprecision.}
+  \input{./graphs/preserving_imp.tex}
+  \caption{Transformation of two fuzzy numbers $X$ and $Y$ into one fuzzy interval $T$ preserving the amount of imprecision.}
   \label{fig:transf-pres-imp}
@@
 
-The membership function for this fuzzy interval T is represented by the trapezoid $\left[\alpha,\ \beta,\ \gamma,\ \delta\right]$:
+The membership function for this fuzzy interval T is represented by the trapezoid $\left[\alpha,\ \beta,\ \gamma,\ \delta\right]$ and is equal to $\left[ d_s - a_s, d_s + b_s,d_e - a_e,d_e + b_e\right]$:
 
-\begin{align}
-\nonumber
-\alpha & =  d_s - a_s \\
-\nonumber
-\beta & = d_s + b_s \\
-\nonumber
-\gamma &= d_e - a_e \\
-\nonumber
-\delta &= d_e + b_e 
-\end{align}
+%\begin{align}
+%\nonumber
+%\alpha & =  d_s - a_s \\
+%\nonumber
+%\beta & = d_s + b_s \\
+%\nonumber
+%\gamma &= d_e - a_e \\
+%\nonumber
+%\delta &= d_e + b_e 
+%\end{align}
 
@@
 
-%\paragraph{Comparison}
-%In this subsection we will study the knowledge we have about the interval $[u,v]$ w.r.t.
-the fuzzy interval $T$  which is built preserving the imprecision as explained in Section \ref{subsubsec:transf-pres-imp}.
-Note that this approach is only valid when the supports of the original fuzzy numbers $X$ and $Y$ are disjoint (see Example \ref{ex:preserving-imprecision} below).
-Now, the knowledge we have about a number $u$ w.r.t.
-the transformated fuzzy interval $T$ is:
-%\begin{eqnarray}
-%\Pos\left(u\in T\right) &=& \mu_{T}\left(u\right) \\
-%\Nec\left(u\in T\right) &=& 1 - \mu_{T}\left(u\right).
-%\end{eqnarray}
-%Therefore, the knowledge about an interval $[u,v]$ w.r.t. $T$ is:
-%\begin{eqnarray}
-%\nonumber
-%\Pos\left([u,v]=T\right) &= \min \left(\Pos\left(u\in T\right),\Pos\left(v\in T\right)\right) &= \min \left(\mu_{T}\left(u\right),\mu_{T}\left(v\right)\right) \\
-%\nonumber
-%\Nec\left([u,v]=T\right) &= \min \left(\Nec\left(u\in T\right),\Nec\left(v\in T\right)\right) &= \min \left(1 - \mu_{T}\left(\overline{u},1 - \mu_{T}\left(\overline{v}\right)\right)\right)
-%\end{eqnarray}
+\paragraph{Example}
+Consider the previous two ill-known points representing a time interval:  $X = \left[3, 2, 1\right]$ and $Y = \left[7, 2, 3 \right]$.
+$P_1$ is the value for the transformation $T$, which is $\left[1,4,5,10 \right]$ (see figure \ref{fig:transf-pres-imp}. 
+The value for $\left[a, b \right]$ is $\left[3,6 \right]$ therefore, the value for $P_2$ is $\left[3,3,6,6 \right]$. 
+The knowledge about the interval $\left[a, b \right]$ w.r.t. the transformation $T$ using the preserving the imprecision approach is given by the \emph{CONTAINS} operator \cite{Garrido2009}.
 
-%\begin{example}
-%\label{ex:preserving-imprecision}
-\paragraph{Example}
-In this example we are going to transform the two ill-known points $X= \left[5, 3, 3 \right]$ and $Y= \left[9, 2, 1 \right]$ into one fuzzy interval $T$ using the preserving the imprecision approach.
-Again, the two fuzzy numbers are represented by means of two triangular possibility distributions:
+\begin{eqnarray}
+\nonumber
+\mbox{CONTAINS}\left(P_1,P_2 \right) = \min \sup_{x} \left(1-\mu_{P_2}(x),\mu_{P_1}(x) \right) = \\
+\nonumber
+ \min (\mu_{\sup_x}(3),\mu_{\sup_x}(6) = \\
+\nonumber
+\min (\frac{2}{3},\frac{4}{5}) = \frac{2}{3} = 0.667
+\end{eqnarray}
 
-Then, following the equation presented in \eqref{eq:preserving-imprecision} we obtain the following trapezoidal shape:
-\begin{equation}
-\mu_{T}\left(x\right) = [d_s-a_s,d_s+b_s,d_e-a_e,d_e+b_e] = [2,8,7,10]
-\end{equation}
-This construction make no sense when the supports of the fuzzy numbers are disjoint.
-It is easy to build a rational membership function just by ordering the elements in the representation (e.g.
-$[2,7,8,10]$) but then, the amount of imprecision is not preserved at all.
-%\end{example}
+Figure \ref{fig:example_pi} shows the two membership degrees for both $\left(1-\mu_{P_2}(x)\right)$ and $\left( \mu_{P_1}(x) \right)$.
+
+\begin{figure}[h!]
+  \centering
+  \input{./graphs/example_pi.tex}
+  \caption{Graphical representation for the membership functions $\left(1-\mu_{P_2}(x)\right)$ and $\left( \mu_{P_1}(x) \right) $.}
+  \label{fig:example_pi}
+\end{figure}
+
+In this example it is shown how this transformation is misleading: it is clear that the temporal interval $[3,6]$ is completely contained by the fuzzy temporal interval represented by $[X,Y]$.
+Nevertheless, the \emph{CONTAINS} operator applied to this transformation returns a 0.667 membership degree.
+Therefore, if the user is looking for a temporal interval $[a,b]$ and the database contains a tuple with a fuzzy interval $[X,Y]$, this transformation returns a possibility below 1, thus the user may interprete that the temporal interval $[a,b]$ is not completely inside the fuzzy interval $[X,Y]$.
+Plus, there is not possible to compute a necessity measure for the \emph{CONTAINS} operator in order to obtain more information about $[a,b]$.
 
@@
 The interpretation of this convex hull is the ``most possible'' starting point and the ``most possible'' ending point.
-Figure \ref{fig:convexHull} shows the resulting transformation.
-\begin{figure}[h!]
-  \centering
-  \includegraphics[scale=0.4]{graphs/convexHull.pdf}
-  \caption{Transformation of two fuzzy numbers into one fuzzy interval with the convex hull approach.}
-  \label{fig:convexHull}
-\end{figure}
+Figure \ref{fig:convex-hull-T} shows the resulting transformation.
 
-In this approach, the trapezoidal membership function ($\left[\alpha,\ \beta,\ \gamma,\ \delta\right]$) is built in a different way:
-
-\begin{align}
-\nonumber
-\alpha & =  d_s - a_s \\
-\nonumber
-\beta & = d_s \\
-\nonumber
-\gamma &= d_e  \\
-\nonumber
-\delta &= d_e + b_e 
-\end{align}
-%\paragraph{Comparison}
-%Now, we will compare the knowledge we have about the interval $[u,v]$ w.r.t.
-the fuzzy interval $T$, which is the result of the convex hull transformation between the two fuzzy numbers $X$ and $Y$.
-The membership degree for the convex hull is given by:
+In this approach, the trapezoidal membership function ($\left[\alpha,\ \beta,\ \gamma,\ \delta\right]$) is built in a different way: $\left[d_s - a_s,d_s,d_e,d_e + b_e \right]$
 
 %\begin{align}
-%\label{eq:ch-possibility}
-%\Pos\left([u,v]=T\right) &= \Pos\left(p^{\left(1\right)}\right) \cap \Pos\left(p^{\left(4\right)}\right) \\ 
-%&= \min \left( \sup_{w \leq d}\ \mu_X\left(w\right) , \sup_{w \geq d}\ \mu_Y\left(w\right)\right)
+%\nonumber
+%\alpha & =  d_s - a_s \\
+%\nonumber
+%\beta & = d_s \\
+%\nonumber
+%\gamma &= d_e  \\
+%\nonumber
+%\delta &= d_e + b_e 
 %\end{align}
 
-%The necessity measure for the convex hull is not consistent with the necessity measure for the individual fuzzy numbers $X$ and $Y$. This transformation results in an information lost.
+\paragraph{Example}
+Consider again the previous two ill-known points representing a time interval:  $X = \left[3, 2, 1\right]$ and $Y = \left[7, 2, 3 \right]$.
+$P_1$ is the value for the transformation $T$, which is $\left[1,3,7,10 \right]$ 
+The value for $\left[a, b \right]$ is $\left[3,6 \right]$ therefore, the value for $P_2$ is $\left[3,3,6,6 \right]$. 
+The evaluation for the \emph{CONTAINS} operator is the following:
 
-%The necessity measures for the following predicates:
+\begin{eqnarray}
+\nonumber
+\mbox{CONTAINS}\left(P_1,P_2 \right) = \min \sup_{x} \left(1-\mu_{P_2}(x),\mu_{P_1}(x) \right) = \\
+\nonumber
+\min \sup_{x} \left(1,1 \right) = 1
+\end{eqnarray}
 
-%\begin{align}
-%\Nec\left(p^{\left(1\right)}\right) &=  \inf_{w \geq d} 1 - \mu_T\left(w\right) \\
-%\Nec\left(p^{\left(4\right)}\right) &=  \inf_{w \leq d} 1 - \mu_T\left(w\right)
-%\end{align}
-
-%Therefore the necessity for the interval $[u,v]$ to be inside the fuzzy interval $T$ is:
-
-%\begin{align}
-%\label{eq:ch-necessity}
-%\Nec\left([u,v]=T\right) &= \min \left(\Nec\left(p^{\left(1\right)}\right), p^{\left(4\right)}\right) \\
-%& \min \left(\inf_{w \geq d} 1 - \mu_T\left(w\right),inf_{w \leq d} 1 - \mu_T\left(w\right)\right)
-%\end{align}
-
-%\begin{example}
-\paragraph{Example}
-In this example we are going to build the convex hull for the two ill-known starting and ending points $X= \left[5, 3, 3\right]$ and $Y = \left[9, 2, 1\right]$.
+Figure \ref{fig:example_ch} shows the two membership degrees for both $\left(1-\mu_{P_2}(x)\right)$ and $\left( \mu_{P_1}(x) \right)$.
 
@@
   \centering
- % \includegraphics[scale=0.2]{graphs/convexHull/convexHull.pdf}
- \input{scripts/convexhull.tex}
+  \input{./graphs/example_ch.tex}
+  \caption{Graphical representation for the membership functions $\left(1-\mu_{P_2}(x)\right)$ and $\left( \mu_{P_1}(x) \right) $.}
+  \label{fig:example_ch}
+\end{figure}
+
+This transformation is also misleading.
+Even though the possibility obtained is similar to the possibility obtained in our proposal, there is not possible to compute the necessity for the \emph{CONTAINS} operator.
+Therefore, it is not possible to obtain a complete information for the interval $[a,b]$ w.r.t the fuzzy interval $[X,Y]$.
+
+\begin{figure}[h!]
+  \centering
+ \input{./graphs/convexhull.tex}
   \caption{Construction of the convex hull $T$ w.r.t the ill-known values $X$ and $Y$.}
@@
 \end{figure}
-
-The convex hull for the fuzzy numbers $X$ and $Y$, represented by means of two triangular possibility distributions results in a trapezoidal possibility distribution with membership function $\mu_T(x)$ (see Figure \ref{fig:convex-hull-T}), represented by $[2,5,9,10]$.
-
-%\begin{figure}[h!]
-%  \centering
-%  \includegraphics[scale=0.2]{graphs/convexHull/possAfterBefore.eps}
-%  \caption{Possibility distributions for  both $u$ possibly greater or equal to $T$ and $u$ possibly smaller or equal to $T$.}
-%  \label{fig:convex-hull-pos}
-%\end{figure}
-
-%\begin{align}
-%\label{eq:ch-possibly-greater}
-%\Pos\left(u \geq T\right) & = \sup_{d \leq u} \mu_{T}\left(d\right)  = [2,5,9,+\infty] \\
-%\label{eq:ch-possibly-smaller}
-%\Pos\left(u \leq T\right) & = \sup_{d \geq u} \mu_{T}\left(d\right)  = [-\infty,5,9,10] 
-%\end{align}
-
-%Respectively, the necessity mesaures for a point $u$ w.r.t. the convex hull $T$ are the following (Figure \ref{fig:convex-hull-nec}):
-
-%\begin{align}
-%\Nec\left(u \geq T\right) &=  1 - \Pos\left(u \geq T\right) = 1 - \sup_{d \geq u} \mu_{T}\left(d\right) = [9,10,+\infty,+\infty] \\ 
-%\Nec\left(u \leq T\right) & = 1 - \Pos\left(u \geq T\right)  = \sup_{d \leq u} \mu_{T}\left(d\right)  = [-\infty,-\infty, 2,5]
-%\end{align}
-
-%\begin{figure}[h!]
-%  \centering
-%  \includegraphics[scale=0.2]{graphs/convexHull/necessity.eps}
-%  \caption{Necessity distributions for both $u$ is necessarily greater or equal to $T$ and $u$ is necessarily smaller or equal to $T$.}
-%  \label{fig:convex-hull-nec}
-%\end{figure}
-
-The membership about an interval $[u,v]$ w.r.t the convex hull $T$ is given by:
-
-\begin{equation}
-\label{eq:membership-ch}
-\mu_T([u,v]) = \min \left( \mu_T\left(u\right), \mu_T\left(v\right)\right)
-\end{equation}
-
-Figure \ref{fig:ch-3d-possibility} shows the 3D plot of the membership function for any interval $[u,v]$ w.r.t.
-the convex hull $T$ (equation \eqref{eq:membership-ch}).
-Note that figure Figure \ref{fig:ch-3d-possibility} is equivalent to figure \ref{fig:3d-possibility}.
-Due to the transformation, it is impossible to obtain a necessity measure.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[scale=0.4]{graphs/convexHull/3dpossibility.pdf}
-  \caption{Membership function for interval $[u,v]$ w.r.t. the convex hull $T$.}
-  \label{fig:ch-3d-possibility}
-\end{figure}
-
-%\begin{figure}[h!]
-%  \centering
-%  \includegraphics[scale=0.4]{graphs/convexHull/3dnecessity.eps}
-%  \caption{Necessity of any interval $[u,v]$ w.r.t. the convex hull $T$.}
-%  \label{fig:ch-3d-necessity}
-%\end{figure}
-
-%\end{example}
 
@@
 
-%Submitted full papers are to be 4 to 6 pages length.
-Papers must be written on two columns with an overall width of 17.2 cm (8.25 cm each column and 0.7 cm of space between columns).
-On odd pages, left margin should be of 2.5 cm while the right one should be of 1.3 cm.
-On even pages, margins are just the contrary (left 1.3 cm, and right 2.5 cm).
-On all pages, odd and even ones, upper margin should be of 2.5 cm, and lower margin should be of 3.5 cm.
-Text alignment must be \textit{justify}.
-The pages must not be numbered.
-%
-%LaTeX style file (estylf\_en.sty) and the example file
-%(estylf2012\_en.tex) are available in the ESTYLF2012 web page.
-%
-%\section{PRELIMINARIES}
-%
-%\subsection{SECTIONS AND SUBSECTIONS}
-%
-%\subsubsection{Subsection}
-%All subsections must follow the same format as this one.
-%
-%\subsection{EQUATIONS}
-%All equations must be centered.
-%\begin{equation}
-%\mu_A = f(a,b,c).
-%\end{equation}
-%
-%\subsection{FIGURES}
-%All figures must be centered like Figure~\ref{figure1}. The number and caption of the figure must always appear below the figure.
-
-%\begin{figure}[h]
-%\begin{center}
-%\includegraphics[width=8cm,height=2.5cm]{estylf2012.jpg}
-%\end{center}
-%\caption{Example of figure caption.}
-%\label{figure1}
-%\end{figure}
-
-%\subsection{TABLES}
-%All tables must be centered and clear. The number and title always appear above the table (see Table~\ref{table1}).
-%
-%\begin{table}[ht]
-%\begin{center}
-%\caption{Example of table title.}
-%\label{table1}
-%\begin{tabular}{|c|l|} \hline
-%\textbf{NAME} & \textbf{DESCRIPTION} \\ \hline
-%$A$ & Fuzzy set \\ \hline
-%$\mu_{A}$ & Membership function \\
-%\hline
-%\end{tabular}
-%\end{center}
-%\end{table}
-
-%\subsection{FOOTNOTES}
-%Footnotes must be numbered\footnote{Example of footnote.}, and placed at the bottom of the column where they appear.
-%
-%\subsection{CITATIONS}
-%Citations must include the reference number between square brackets, for example \cite{AlsinaT05}.
-
 \begin{ack}
@@
 
-%\begin{thebibliography}{99}
-
-%\bibitem{AlsinaT05}
-%C. Alsina, E. Trillas: On the symmetric difference of fuzzy sets. \emph{Fuzzy Sets and Systems} 153, pp.~181--194, 2005.
-%
-%\bibitem{Fishburn73}
-%P.C. Fishburn: \emph{Utility Theory for Decision Making}. Wiley, New York, 1970.
-%
-%\bibitem{Yager03}
-%R.R. Yager: E-Z OWA weights. In: \emph{Proceedings of the 10th IFSA World Congress}, Istanbul (Turkey), pp.~39--42, 2003.
-%
-%\bibitem{Zadeh71}
-%L.A. Zadeh: Towards a theory of fuzzy systems. In: R.E. Kalman, N. DeClaris (Eds.), \emph{Aspects of Network and System Theory}, Holt, Rinehart and Winston, New York, pp.~469--490, 1971.
-
-%\end{thebibliography}
-
-%References must be sorted by the author name.
-
 \end{document}

### 2011-11-29 trunk/estylf2012/estylf2012_en.tex
Final version of the paper

--- 2011-11-28 estylf2012/estylf2012_en.tex
@@
 \begin{abstract}
-Possibility theory tries to quantify the uncertainty caused by incomplete information.
-A specific case of incomplete information is that of ill-known temporal intervals.
-The study of these fuzzy intervals is of particular interest in temporal database research.
+%Possibility theory tries to quantify the uncertainty caused by incomplete information. A specific case of incomplete information is that of ill-known temporal intervals. 
+The study of fuzzy intervals is of particular interest in temporal database research.
 In order to optimize the storage of fuzzy temporal intervals, some transformations have been proposed.
-In this paper we analyze the possibilistic evaluation of the ill-known temporal intervals and the proposed transformations.
+In this paper we analyze the possibilistic evaluation of the ill-known temporal intervals.
+We propose a framework to deal with the evaluation of ill-known temporal intervals.
 It is shown how the reasoning behind the transformations implies a possibilistic information lost.
@@
 
-The connection of possibility theory \cite{Zadeh78},\cite{Shackle61},\cite{Gaines75} and fuzzy sets has been studied among others by Zadeh \cite{Zadeh65}.
+The connection between possibility theory \cite{Zadeh78},\cite{Shackle61},\cite{Gaines75} and fuzzy sets has been studied among others by Zadeh \cite{Zadeh65}.
 This connection establishes two different natures for a fuzzy set \cite{Dubois97}:
@@
 
-There are several proposals in the literature for inferring uncertainty about sets from uncertainty about values.
+%There are several proposals in the literature for inferring uncertainty about sets from uncertainty about values.
 It is of particular interest for fuzzy temporal databases, the application on temporal intervals.
@@
 The membership function $\mu_M$ for the fuzzy set $M$ has also the properties of convexity and normalization.
-Three values represent a triangular function: %(see Figure~\ref{fig:triangular}):
+Three values represent a triangular function:
 \begin{itemize}
@@
 Note that the values $a$ and $b$ are values in the underlying ordered domain. E.g. $(a,b) \in \mathbb{R}^2$. The notation for a triangular function adopted here is $[D,a,b]$.
-%\begin{figure}[h!]
-%  \centering
-%  \includegraphics[scale=0.4]{graphs/triangular.pdf}
-%  \caption{Triangular function.}
-%  \label{fig:triangular}
-%\end{figure}
 
@@
 Four values represent a trapezoidal membership function %(figure  \ref{fig:trapezoidal}):
- $\left[\alpha,\ \beta,\ \gamma,\ \delta\right]$. The membership function is given by:
-
-%\begin{align}
-%\mu_T(x)
-%\begin{cases}
-%1 & \mbox{ if } x \in [\beta,\gamma] \\
-%0 & \mbox{ if } x > \delta \vee x < \alpha \\
-%\frac{x-\alpha}{\beta - \alpha} & \mbox{ if } x \in [\alpha,\beta[ \\
-%\frac{\delta -x}{\delta - \gamma} & \mbox{ if } x \in ]\gamma,\delta] \\
-%\end{cases}
-%\end{align}
-%
-%\begin{figure}[h!]
-%  \centering
-%  \includegraphics[scale=0.4]{graphs/trapezoidalDistribution.pdf}
-%  \caption{Trapezoidal distribution.}
-%  \label{fig:trapezoidal}
-%\end{figure}
+ $\left[\alpha,\ \beta,\ \gamma,\ \delta\right]$.
 
@@
 Allen \cite{Allen:1983:MKT:182.358434} studied the comparison between two crisp time intervals.
-For fuzzy intervals, several proposals \cite{4481150}, \cite{springerlink:10.1007/978-3-540-39964-3_57},\cite{10.1109/TIME.2004.1314418} have been done.
-
-%Time granularity is also associated with the representation of the time.
-A granularity is the result of partitioning on the set of chronons.
-The conversion among granularities is a common issue within temporal databases \cite{Lin97efficientconversion}.
-Granularity is the basis of some systems \cite{Cru97},\cite{624013}.
-
+For fuzzy intervals, several proposals \cite{4481150}, \cite{springerlink:10.1007/978-3-540-39964-3_57},\cite{10.1109/TIME.2004.1314418} have been done.\\
 Despite of user-defined time (an uninterpreted attribute; supported in the standard SQL2 \cite{Mel93}), there are 3 types of time in a temporal database:
-
 \begin{itemize}
@@
 \end{itemize}	
-
 The  database model can be then classified into: transaction time \cite{Ston87},\cite{Jensen:1991:IIM:627283.627484}, Valid time, bi-temporal \cite{Snodgrass:1984:TQL:588011.588041}(both valid and transaction time) or tri-temporal \cite{Nascimento95decisiontime} (valid, transaction and decision time).
@@
 
-Note that the expressions in \eqref{eq:interval-pos},\eqref{eq:interval-nec} cannot be simplified in a more efficient expressions. %repensar
+Note that the expressions in \eqref{eq:interval-pos},\eqref{eq:interval-nec} cannot be simplified in a more efficient expressions.
 An ill-known point is given by a triangular membership function.
 The ill-known values $X$ and $Y$ are represented by the triangular membership functions $\left[D_X,a_X,b_X \right]$ and $\left[D_Y,a_Y,b_Y \right]$ respectively.
-The equation \eqref{eq:interval-pos} is derived to:
+The equation \eqref{eq:interval-pos} is then derived to:
 
@@
 
+Note that with this approach, both possibility and necessity measures are obtained.
+When querying a temporal database, these two measures may be used to compute a ranking.
+E.g.
+the results for a query may be ordered by a lexicographical order $(\Pos,\Nec)$.
+Note also that with only the possibility measure, as provided in the transformations, some interesting information is lost: e.g.
+If the user wants to know if the interval $[a,b]$ is inside the ill-known interval $[X,Y]$, and the evaluation returns a possibility of 1 and a necessity of 0.5.
+Then, the user may interpret that the interval $[a,b]$ is inside $[X,Y]$ but it is not between the cores of both $X$ and $Y$.
+
 \paragraph{Example} Consider the ill-known values $X = \left[3, 2, 1\right]$ and $Y = \left[7, 2, 3 \right]$.
@@
 \begin{eqnarray}
-\label{eq:interval-pos}
+%\label{eq:interval-pos}
 \nonumber
 \Pos\left(\lambda([3,6])\right)=\min(1,1)= 1\\
-\label{eq:interval-nec}
+%\label{eq:interval-nec}
 \nonumber
@@
 Nevertheless, the \emph{CONTAINS} operator applied to this transformation returns a 0.667 membership degree.
-Therefore, if the user is looking for a temporal interval $[a,b]$ and the database contains a tuple with a fuzzy interval $[X,Y]$, this transformation returns a possibility below 1, thus the user may interprete that the temporal interval $[a,b]$ is not completely inside the fuzzy interval $[X,Y]$.
+Therefore, if the user is looking for a temporal interval $[a,b]$ and the database contains a tuple with a fuzzy interval $[X,Y]$, this transformation returns a possibility below 1, thus the user may interpret that the temporal interval $[a,b]$ is not completely inside the fuzzy interval $[X,Y]$.
 Plus, there is not possible to compute a necessity measure for the \emph{CONTAINS} operator in order to obtain more information about $[a,b]$.
@@
 In this approach, the trapezoidal membership function ($\left[\alpha,\ \beta,\ \gamma,\ \delta\right]$) is built in a different way: $\left[d_s - a_s,d_s,d_e,d_e + b_e \right]$
-
-%\begin{align}
-%\nonumber
-%\alpha & =  d_s - a_s \\
-%\nonumber
-%\beta & = d_s \\
-%\nonumber
-%\gamma &= d_e  \\
-%\nonumber
-%\delta &= d_e + b_e 
-%\end{align}
 
@@
 \section{\label{sec:conclusions}CONCLUSIONS AND FUTURE WORK}
-The problem of evaluation of sets in the framework of possibility theory has been addressed by Dubois and Prade.
+In this work we presented a framework to deal with the evaluation for the ill-known interval $[X,Y]$ to contains the crisp interval $[a,b]$.
+Our proposal keeps the reasoning in the power set $\Pow(U)$ while the transformations make the reasoning in the set $U$, therefore the transformations result in a possibilistic information lost.
+The proposed framework offer more accurate results, as have been shown in the examples.
+Also the two measures provided: possibility and necessity can be used for the ranking of results.The framework will be extended with the implementation of the Allen's operators and the comparison between two ill-known intervals.
+%
+%
+%The problem of evaluation of sets in the framework of possibility theory has been addressed by Dubois and Prade.
 They conclude that a correct treatment of sets imply an overly complex possibility distribution.
 Based on the previous work from the authors, this work is focused on the evaluation of fuzzy temporal interval in the framework of possibility theory.
-The aim of the paper is to clarify the difference between fuzzy numbers, fuzzy intervals and ill-known intervals.
+%
+%
+%
+%The aim of the paper is to clarify the difference between fuzzy numbers, fuzzy intervals and ill-known intervals.
 Several proposals can be found in the literature with respect to transforming two fuzzy numbers into a fuzzy interval.
 We have shown that these techniques are not fully correct in the sense of possibility theory.\\
+%
 

### 2011-11-30 ipmu2012/conclusions.tex
Update for the proposal and the conclusions section.

We presented a valid-time model to represent and querying ill-known temporal intervals. The main advantage of the framework is that it is always possible to get both a possibility and a necessity measures for every comparison, which is also useful for ranking purposes. The framework models the Allen's relations and it has the flexibility to model specific and more complex relations by means of ill-known constraints.  As future work, the time interval in the query specification would also be ill-known.

### 2011-11-30 ipmu2012/proposal.tex
Update for the proposal and the conclusions section.

@@
 
-The possibility distributions $\pi_X$ and $\pi_Y$ are given in the way of a triangular distribution, as explained in subsection \ref{subsec:fuzzy-numbers}.
+For convenience, the possibility distributions $\pi_X$ and $\pi_Y$ are given in the way of a triangular distribution, as explained in subsection \ref{subsec:fuzzy-numbers}.
 This representation allows overlapping (Fig.
 \ref{fig:pvp}).
+Note that while the two ill-known values $X,Y \in \mathbb{R}$, the fuzzy interval  $[X,Y] \in \mathbb{R}^2$.
 
@@
   \centering
-%  \includegraphics[scale=0.2]{graphs/2_triangular.eps}
-  \caption{Two fuzzy numbers $X$ and $Y$ denoting a Possibilistic Valid-Time Period \emph{PVT}.}
+  \input{./graphs/2_triangular.tex}
+  \caption{Two fuzzy numbers $X$ and $Y$ denoting a Possibilistic Valid-Time Period \emph{PVP}.}
   \label{fig:pvp}
@@
 \item
-\emph{UNKNOWN}: The point has a value but it is unknown. The possibility distribution for a given fuzzy number $X$ is $\pi_X=1$
+\emph{UNKNOWN}: The point has a value but it is unknown.% The possibility distribution for a given fuzzy number $X$ is $\pi_X=1$
 \item
-\emph{UNDEFINED}: The point does not have a value. The possibility distribution for a given fuzzy number $X$ is $\pi_X=0$
+\emph{UNDEFINED}: The point does not have a value. %The possibility distribution for a given fuzzy number $X$ is $\pi_X=0$
 \end{itemize}
@@
 \begin{table}
-\caption{Relational representation for a Possibilistic Valid-Time Period.}
+\caption{Relational representation for an ill-known time point. A \emph{PVP} is represented by two ill-known points.}
 \centering
-\begin{tabular}{c c c c c c}
+\begin{tabular}{c c c c c c c}
 \hline
-Value & FT & F1 & F2 & F3  \\ \hline
-UNKNOWN & 0 & NULL & NULL & NULL  \\ 
-UNDEFINED & 1 & NULL & NULL & NULL  \\ 
-NULL & 2 & NULL & NULL & NULL  \\ 
-$\left[D,\ a,\ b \right]$ & 3 & $D$ & $D-a$ & $D+b$ \\ 
+Value & FT & F1 & F2 & F3 & $\mu(x)$ \\ \hline
+UNKNOWN & 0 & NULL & NULL & NULL  & not defined. \\ 
+UNDEFINED & 1 & NULL & NULL & NULL & $0$ \\ 
+NULL & 2 & NULL & NULL & NULL & $1$  \\ 
+$\left[D,\ a,\ b \right]$ & 3 & $D$ & $D-a$ & $D+b$ & $\mu(\left[D,\ a,\ b \right])$ \\ 
 \hline
@@
 In the query specification, the user is allowed to express a crisp time interval.
-In this approach, the interpretation is the following: the user w
 In this subsection we will define the query specification, then the evaluation of the query and finally the ranking for the query.
@@
 \subsubsection{Query specification}
-A query in our framework has two different parts: the first one is the query specification for regular attributes. The second part is the temporal specification.
+A query in our framework has two different parts: the first one is the temporal specification. The second part is the query specification for regular attributes.
 
@@
 \centering
-\begin{tabular}{|c|c|c|}
+\begin{tabular}{|c|c|c|c|}
 \hline
-Allen Relation & Constraints & $\bool(p_1,...,p_n)$ \\
+Allen Relation & Constraints & $\bool(p_1,...,p_n)$ & $\Pos\left(\lambda \left[a,b \right] \right)$ \\
 \hline
-I before J & $C_1\stackrel{\triangle}{=} \left(<,X\right)$ & $p_1$ \\
+I before J & $C_1\stackrel{\triangle}{=} \left(<,X\right)$ & $p_1$ & $\sup_{a>w}\pi_x(w)$\\
 \hline
 \multirow{4}{*}
-{I equal J} & $C_1\stackrel{\triangle}{=} \left(\geq,X\right)$ & \text{$p_1\wedge\neg p_2\wedge p_3\wedge\neg p_4$}\\
- & $C_2\stackrel{\triangle}{=} \left(\neq,X\right)$ & \\
- & $C_3\stackrel{\triangle}{=} \left(\leq,Y\right)$ & \\
- & $C_4\stackrel{\triangle}{=} \left(\neq,Y\right)$ & \\
+{I equal J} & $C_1\stackrel{\triangle}{=} \left(\geq,X\right)$ & \text{$p_1\wedge\neg p_2\wedge p_3\wedge\neg p_4$} & $\min ( \sup_{a \leq w}\pi_x(w),$\\
+ & $C_2\stackrel{\triangle}{=} \left(\neq,X\right)$ & & $\pi_x(w),$ \\
+ & $C_3\stackrel{\triangle}{=} \left(\leq,Y\right)$ & & $\sup_{b \geq w}\pi_Y(w),$ \\
+ & $C_4\stackrel{\triangle}{=} \left(\neq,Y\right)$ & & $\pi_Y(w))$ \\
 \hline
 \multirow{2}{*}
-{I meets J} & $C_1\stackrel{\triangle}{=} \left(\leq,X\right)$ & $p_1\wedge\neg p_2$\\
- & $C_2\stackrel{\triangle}{=} \left(\neq,X\right)$ & \\
+{I meets J} & $C_1\stackrel{\triangle}{=} \left(\leq,X\right)$ & $p_1\wedge\neg p_2$ & $\min (\sup_{a\geq w} \pi_X(w),$\\
+ & $C_2\stackrel{\triangle}{=} \left(\neq,X\right)$ &  & $\pi_X(w))$ \\
 \hline
 \multirow{3}{*}
-{I overlaps J} & $C_1\stackrel{\triangle}{=} \left(<,Y\right)$ & $p_1\wedge\neg p_2\wedge\neg p_3$\\
- & $C_2\stackrel{\triangle}{=} \left(\leq,X\right)$ & \\
- & $C_3\stackrel{\triangle}{=} \left(\geq,X\right)$ & \\
+{I overlaps J} & $C_1\stackrel{\triangle}{=} \left(<,Y\right)$ & $p_1\wedge\neg p_2\wedge\neg p_3$ & $\min ( \sup_{b>w}\pi_Y(w), $\\
+ & $C_2\stackrel{\triangle}{=} \left(\leq,X\right)$ & &  $\sup_{a \geq w}\pi_X(w),$\\
+ & $C_3\stackrel{\triangle}{=} \left(\geq,X\right)$ & & $\sup_{a \leq w}\pi_X(w))$ \\
 \hline
 \multirow{4}{*}
-{I during J} & $C_1\stackrel{\triangle}{=} \left(>,X\right)$ & $(p_1\wedge p_2)\vee(p_3\wedge p_4)$\\
- & $C_2\stackrel{\triangle}{=} \left(\leq,Y\right)$ & \\
- & $C_3\stackrel{\triangle}{=} \left(\geq,X\right)$ & \\
- & $C_4\stackrel{\triangle}{=} \left(<,Y\right)$ & \\
+{I during J} & $C_1\stackrel{\triangle}{=} \left(>,X\right)$ & $(p_1\wedge p_2)\vee(p_3\wedge p_4)$ & $max ( min ( \sup_{a<w}\pi_X(w),$\\
+ & $C_2\stackrel{\triangle}{=} \left(\leq,Y\right)$ & & $\sup_{b \geq w}\pi_Y(w)),$\\
+ & $C_3\stackrel{\triangle}{=} \left(\geq,X\right)$ & & $\min ( \sup_{a \leq w }\pi_X(w),$\\
+ & $C_4\stackrel{\triangle}{=} \left(<,Y\right)$ & & $\sup_{b>w}\pi_Y(w)$\\
 \hline
 \multirow{2}{*}
-{I starts J} & $C_1\stackrel{\triangle}{=} \left(\geq,X\right)$ & $p_1\wedge\neg p_2$\\
- &  $C_2\stackrel{\triangle}{=} \left(\neq,X\right)$ & \\
+{I starts J} & $C_1\stackrel{\triangle}{=} \left(\geq,X\right)$ & $p_1\wedge\neg p_2$ & $\min( \sup_{a \leq w}\pi_X(w),$\\
+ &  $C_2\stackrel{\triangle}{=} \left(\neq,X\right)$ & & $\pi_X(w))$\\
 \hline
 \multirow{2}{*}
-{I finishes J} & $C_1\stackrel{\triangle}{=} \left(\leq,Y\right)$ & $p_1\wedge\neg p_2$\\
- & $C_2\stackrel{\triangle}{=} \left(\neq,Y\right)$ & \\
+{I finishes J} & $C_1\stackrel{\triangle}{=} \left(\leq,Y\right)$ & $p_1\wedge\neg p_2$ & $\min ( \sup_{b \geq w} \pi_Y(w),$\\
+ & $C_2\stackrel{\triangle}{=} \left(\neq,Y\right)$ & & $\pi_Y(w))$\\
 \hline
@@
 \subsubsection{Query evaluation}
-Fuzzy querying of regular (relational) databases, query satisfaction modelling is a matter of degree.
+In fuzzy querying of regular (relational) databases, the query satisfaction modelling is a matter of degree.
 Usually, criteria satisfaction is modelled by means of a satisfaction degree $s \in \left[ 0, 1\right]$.
-In the model, every record $r$ contains a \emph{PVP} $V_r$ to model the valid-time for that record.
-
-The query evaluation method is the following:
+In the model, every record $r$ contains a \emph{PVP}, $V_r$, to model the valid-time.The query evaluation method is the following:
 \begin{itemize}

### 2011-11-30 trunk/ipmu2012/preliminaries.tex
Some place-saving changes...

--- 2011-11-17 trunk/ipmu2012/preliminaries.tex
@@
 In probability theory, this uncertainty is caused by the \emph{variability} in the outcomes, while in possibility theory, the uncertainty is caused by \emph{incomplete knowledge} about the experiment.
-The quantification of confidence in probability theory is achieved by means of a measure of chance, whereas in possibility theory, possibility and necessity measures are used.
-In any case, the quantification of confidence in a theory of uncertainty is achieved using a confidence measure\cite{Shafer:1976:AMathematical}.
+The quantification of confidence in a theory of uncertainty is achieved using a confidence measure\cite{Shafer:1976:AMathematical}.
+In probability theory this is a measure of chance, in possibility theory, possibility and necessity measures are used.
 
@@
 
-In this work, an ill-known set will be defined and represented via its start and end point, which will be ill-known values.
-The correspondences and transitions between these two representations are part of the authors current research.
+This work will deal with ill-known intervals.
+These are ill-known sets, defined and represented via a start and end point, which will be ill-known values.
+The elements of the set are the values between the start and end point.
+An closed ill-known interval with start point $X$ and end point $Y$ is noted here $\left[X, Y\right]$.
+The correspondences and transitions between the representations of ill-known sets and between the representation of an ill-known set and an ill-known interval are part of the authors current research.
 
@@
 A specific application of possibilistic variables is obtained when the universe under consideration is the set of boolean values $\mathbb{B}$ = $\{T,F\}$.
-Indeed, any boolean propositional variable $p$ takes just one value in $\mathbb{B}$.
-If the knowledge on which value this variable $p$ will take is given by a possibility distribution $\pi_p$, the variable can be seen as a possibilistic variable.
+Indeed, any boolean proposition $p$ takes just one value in $\mathbb{B}$.
+If the knowledge on which value this proposition $p$ will take is given by a possibility distribution $\pi_p$, the proposition can be seen as a possibilistic variable.
 As the interest lies with the case where the proposition holds, the possibility and necessity that $p$ = $T$ (the proposition holds) demands most attention.
@@
 \subsubsection{Interval Evaluation by Ill-known Constraints}
-The problem of the interval evaluation is explained in \cite{Pon11}: For a crisp interval $I = \left[ a, b \right]$, we want to know if all points in this interval reside between the boundaries of the ill-known interval $\left[ X , Y \right]$, where $X,Y$ are ill-known values on the set of real numbers $\mathbb{R}$, and $\lambda$ is the evaluation function.
+The problem of interval evaluation is more generally explained in \cite{Pon11}: for a crisp interval $I = \left[a, b \right]$, the need exists to know if all points in this interval reside between the boundaries of an ill-known interval $\left[ X , Y \right]$.
+
+The notion of an \emph{ill-known constraint} is introduced:
+
+\begin{definition}
+Given a universe $U$, an \emph{ill-known constraint} $C$ on a set $A \subseteq U$ is specified by means of a binary relation $R \subseteq U^{2}$ and a fixed, ill-known value denoted by its variable $X$ over $U$, i.e.:
+\begin{align}
+C \triangleq (R,X)
+\end{align}
+\end{definition}
 

### 2011-11-30 ipmu2012/introduction.tex
Some place-saving changes...

@@
 Klein~\cite{klein94} studied the concept of time in language and Devos~\cite{devos94} modelled vague temporal expressions by means of fuzzy sets~\cite{zadeh65}.
-More towards information systems, several proposals have been concerned with the obtaining of theoretical models that allow the representation of time in information systems~\cite{devos98,Cru97}.
+More towards information systems, several proposals have been concerned with the obtaining of theoretical models that allow the representation of time in information systems~\cite{Cru97}.
 Notably, temporal relations between time intervals and to a lesser extent time points, were first studied by Allen~\cite{Allen83}.
 
-Next to the inherent difficulties of time, an issue is \emph{temporal granularity}.
+%Next to the inherent difficulties of time, an issue is \emph{temporal granularity}.
 A temporal granularity is a partitioning of the time line used by a system, usually dependent on the application.
@@
 
-Another issue is the possible inherent imperfection in the description of notions of time.
+One issue in these attempts at temporal modeling is off course the inherent complexities of the concept of time.
+Another is the possible inherent imperfection in the description of notions of time.
 The temporal notion in a sentence like `The Alhambra, which I can see from my room, is a few hundred years old.' contains imperfection caused by the vagueness of the used expression.
@@
 and 1400 A.D.' contains imperfection caused by the inaccuracy in the given temporal notion, though precise boundaries are given.
+The transition between temporal granularities~\cite{Lin97} is also considered as a source of imprecision in temporal modelling and some proposals consider granularity as the base of the temporal model~\cite{Cru97}.
 
-The transition between granularities~\cite{Lin97} is also considered as a source of imprecision.
+%The transition between granularities~\cite{Lin97} is also considered as a source of imprecision.
 In a sentence like `The Alhambra is 612 years old', the temporal notion is very precise if time periods shorter than a year are irrelevant in the used granularity.
@@
 
-To allow information systems to cope with imperfections, many approaches adopt fuzzy sets for the representation of temporal information~\cite{343607,nagypal03,Dubois:jucs_9_9:fuzziness_and_uncertainty_in}, though some other tools, like rough sets, have also been used~\cite{Qia09}.
-The temporal relations studied by Allen were recently fuzzified by several authors~\cite{ohlbach04,nagypal03,schockaert08}.
+To allow information systems to cope with these imperfections, many approaches adopt fuzzy sets for the representation of temporal information~\cite{343607,Dubois:jucs_9_9:fuzziness_and_uncertainty_in}.
+The temporal relations studied by Allen were recently fuzzified by several authors~\cite{ohlbach04,schockaert08}.
 
-Next to the issues in temporal modelling, the addition of time to a relational database~\cite{Dyreson1994} leads to several more problems, as the modelling of a time-variant object has an impact on the database consistency.
-Different models have been proposed \cite{Jensen91,Snodgrass84,Nascimento95}, but all of these consider time as a crisp notion.
-In recent works though, some authors consider the necessity of allowing imprecision in temporal models~\cite{Cru97,Garrido2009}.
+Next to the issues in temporal modelling, the addition of time to a relational database~\cite{Dyreson1994} leads to several more practical problems, as the modelling of a time-variant object has an impact on the database consistency.
+Different models have been proposed \cite{Jensen91,Snodgrass84} that consider time as a crisp notion.
+In recent works, some authors consider the necessity of allowing imprecision in temporal models~\cite{Cru97,Garrido2009}.
 
@@
 
-In this work we will propose a possibilistic valid time model for relational databases.
-The base of the model is the possibilistic framework presented in~\cite{Pon11}.
-This framework deals with the inference of uncertainty about the evaluation of sets from the uncertainty about the exact values used in evaluation.
+In this work we will propose a possibilistic valid time model for relational databases, based on the framework presented in~\cite{Pon11}, which deals with the inference of uncertainty about the evaluation of sets from the uncertainty about the exact values used in evaluation.
 
-The paper is organized as follows. In section \ref{sec:preliminaries}, some preliminaries and concepts about the possibilistic evaluation framework and temporal databases are presented. 
+The paper is organized as follows. In section \ref{sec:preliminaries}, some preliminaries and concepts are presented. 
 Section \ref{sec:proposal} presents the possibilistic model for valid-time databases. In this section, both representation and querying are introduced.

### 2011-12-01 estylf2012/estylf2012_en.tex
Update for the estylf paper.

@@
 
-There are several proposals in the literature for inferring uncertainty about sets from uncertainty about values.
+%There are several proposals in the literature for inferring uncertainty about sets from uncertainty about values.
 It is of particular interest for fuzzy temporal databases, the application on temporal intervals.
-The transformation from two ill-known temporal points to a ill-known set has been discussed in \cite{Garrido2009}.
+The transformation from two ill-known temporal points to an ill-known set has been discussed in \cite{Garrido2009}.
 In this work we will analyze the major pitfalls of these approaches and the benefit of using the mechanism proposed in \cite{Pon11}.
@@
 In this section we will introduce some basic concepts.
-First of all, an introduction to possibility theory is explained.
+First of all, an introduction to possibility theory is included.
 Then, the concepts of fuzzy numbers and fuzzy intervals introduced by Dubois and Prade \cite{Dubois1983} are discussed.
@@
 Nevertheless, the \emph{CONTAINS} operator applied to this transformation returns a 0.667 membership degree.
-Therefore, if the user is looking for a temporal interval $[a,b]$ and the database contains a tuple with a fuzzy interval $[X,Y]$, this transformation returns a possibility below 1, thus the user may interprete that the temporal interval $[a,b]$ is not completely inside the fuzzy interval $[X,Y]$.
-Plus, there is not possible to compute a necessity measure for the \emph{CONTAINS} operator in order to obtain more information about $[a,b]$.
+Therefore, if the user is looking for a temporal interval $[a,b]$ and the database contains a tuple with a fuzzy interval $[X,Y]$, this transformation returns a possibility below 1, thus the user may interpret that the temporal interval $[a,b]$ is not completely inside the fuzzy interval $[X,Y]$.
+Plus, it is not possible to compute a necessity measure for the \emph{CONTAINS} operator in order to obtain more information about $[a,b]$.
 
@@
 This transformation is also misleading.
-Even though the possibility obtained is similar to the possibility obtained in our proposal, there is not possible to compute the necessity for the \emph{CONTAINS} operator.
+Even though the possibility obtained is similar to the possibility obtained in our proposal, it is not possible to compute the necessity for the \emph{CONTAINS} operator.
 Therefore, it is not possible to obtain a complete information for the interval $[a,b]$ w.r.t the fuzzy interval $[X,Y]$.
@@
 \section{\label{sec:conclusions}CONCLUSIONS AND FUTURE WORK}
-The problem of evaluation of sets in the framework of possibility theory has been addressed by Dubois and Prade.
+In this work we have presented a framework to deal with the evaluation for the ill-known interval $[X,Y]$ to contain the crisp interval $[a,b]$.
+Our proposal keeps the reasoning in the power set $\Pow(U)$ while the transformations make the reasoning in the set $U$, therefore the transformations result in a possibilistic information lost.
+The proposed framework offer more accurate results, as shown in the examples.
+Also the two measures provided: possibility and necessity can be used for the ranking of results.The framework will be extended with the implementation of the Allen's operators and the comparison between two ill-known intervals.
+%
+%
+%The problem of evaluation of sets in the framework of possibility theory has been addressed by Dubois and Prade.
 They conclude that a correct treatment of sets imply an overly complex possibility distribution.
 Based on the previous work from the authors, this work is focused on the evaluation of fuzzy temporal interval in the framework of possibility theory.
-The aim of the paper is to clarify the difference between fuzzy numbers, fuzzy intervals and ill-known intervals.
+%
+%
+%
+%The aim of the paper is to clarify the difference between fuzzy numbers, fuzzy intervals and ill-known intervals.
 Several proposals can be found in the literature with respect to transforming two fuzzy numbers into a fuzzy interval.
 We have shown that these techniques are not fully correct in the sense of possibility theory.\\
+%
 

### 2011-12-01 trunk/ipmu2012/preliminaries.tex
More changes, resulting in one less page...

@@
 \subsection{\label{subsec:possibilistic-variables}Possibilistic Variables}
-A \emph{possibilistic variable} can intuitively be seen as a variable for which the possible values are described by its possibility distribution\cite{Pon11}.
+A \emph{possibilistic variable} is defined as follows\cite{Pon11}.
 
@@
 A possibilistic variable $X$ over a universe $U$ is defined as a variable taking exactly one value in $U$, but for which this value is (partially) unknown.
-The possibility distribution $\pi_X$ on $U$ gives the available knowledge about the value that $X$ takes: for each $u\in U$, $\pi_X(u)$ represents the possibility that $X$ takes the value $u$.
+Its possibility distribution $\pi_X$ on $U$ gives the available knowledge about the value that $X$ takes: for each $u\in U$, $\pi_X(u)$ represents the possibility that $X$ takes the value $u$.
 \end{definition}
@@
 
-When a possibilistic variable is defined on the powerset $\Pow(R)$ of some universe $R$, the unique value the variable takes will be a crisp set.
-In this case, the exact set is off course (partially) unknown, but the available knowledge is given in the form of a possibility distribution on the powerset $\Pow(R)$, which describes the possibility of each crisp subset of $R$ to be the value the variable takes.
+When a possibilistic variable is defined on the powerset $\Pow(R)$ of some universe $R$, the unique value the variable takes will be a crisp set and its possibility distribution on the powerset $\Pow(R)$ will describe the possibility of each crisp subset of $R$ to be the value the variable takes.
 This exact value (a crisp set) the variable takes, is now called an \emph{ill-known set}\cite{Dubois88b}.
-
-This work will deal with ill-known intervals.
-These are ill-known sets, defined and represented via a start and end point, which will be ill-known values.
-The elements of the set are the values between the start and end point.
-An closed ill-known interval with start point $X$ and end point $Y$ is noted here $\left[X, Y\right]$.
-The correspondences and transitions between the representations of ill-known sets and between the representation of an ill-known set and an ill-known interval are part of the authors current research.
 
@@
 If the knowledge on which value this proposition $p$ will take is given by a possibility distribution $\pi_p$, the proposition can be seen as a possibilistic variable.
-As the interest lies with the case where the proposition holds, the possibility and necessity that $p$ = $T$ (the proposition holds) demands most attention.
+As the interest lies with the case where the proposition holds, the possibility and necessity that $p$ = $T$ (the proposition holds) demand most attention.
 This possibility and necessity is noted here as:
 \begin{align}
-Pos(p) & = \pi_p(T) \label{propholdsposs} \\
-Nec(p) & = 1-\pi_p(F) \label{propholdsnecc}
+\text{Possibility that $p$ = $T$ (p holds):} \hspace{50pt} & Pos(p) = \pi_p(T) \label{propholdsposs} \\
+\text{Necessity that $p$ = $T$ (p holds):} \hspace{50pt} & Nec(p) = 1-\pi_p(F) \label{propholdsnecc}
 \end{align}
-Here, notation \ref{propholdsposs} denotes the possibility that $p$ = $T$ and the proposition holds, notation \ref{propholdsnecc} denotes the necessity that $p$ = $T$ and the proposition holds.
+%Here, notation \ref{propholdsposs} denotes the possibility that $p$ = $T$ and the proposition holds, notation \ref{propholdsnecc} denotes the necessity that $p$ = $T$ and the proposition holds.
+
+This work will deal with ill-known intervals.
+These are ill-known sets, defined and represented via a start and end point, which will be ill-known values.
+The elements of the set are the values between the start and end point.
+An closed ill-known interval with start point $X$ and end point $Y$ is noted here $\left[X, Y\right]$.
+The correspondences and transitions between the representations of ill-known sets and between the representation of an ill-known set and an ill-known interval are part of the authors current research.
 
@@
 \end{definition}
-If this modal value $m$ is unique, then $M$ is referred to as a \emph{fuzzy number}, instead of a \emph{fuzzy interval}.
-In other words, if the core of a fuzzy interval is a singleton, it is referred to as a fuzzy number.
+If this modal value $m$ is unique, then $M$ is referred to as a \emph{fuzzy number}. In other words, if the core of a fuzzy interval is a singleton, it is referred to as a fuzzy number.
 
@@
 In this work, a fuzzy interval defined as such will be noted as $\left[\alpha, \beta, \gamma, \delta\right]$.
-The corresponding membership function definition for this $\mu_T$ is then given by:
+%The corresponding membership function definition for this $\mu_T$ is then given by:
 
-\begin{align}
-\mu_T : & \quad \mathbb{R} \rightarrow \left[0,1\right] \\
- : & \quad x \rightarrow
-\begin{cases}
-1 & \mbox{ if } x \in [\beta,\gamma] \\
-0 & \mbox{ if } x > \delta \vee x < \alpha \\
-\frac{x-\alpha}{\beta - \alpha} & \mbox{ if } x \in [\alpha,\beta[ \\
-\frac{\delta -x}{\delta - \gamma} & \mbox{ if } x \in ]\gamma,\delta] \\
-\end{cases}
-\end{align}
+%\begin{align}
+%\mu_T : & \quad \mathbb{R} \rightarrow \left[0,1\right] \\
+% : & \quad x \rightarrow
+%\begin{cases}
+%1 & \mbox{ if } x \in [\beta,\gamma] \\
+%0 & \mbox{ if } x > \delta \vee x < \alpha \\
+%\frac{x-\alpha}{\beta - \alpha} & \mbox{ if } x \in [\alpha,\beta[ \\
+%\frac{\delta -x}{\delta - \gamma} & \mbox{ if } x \in ]\gamma,\delta] \\
+%\end{cases}
+%\end{align}
 
@@
 A visualization is given in figure \ref{fig:triangular}.
-Three reel values, denoted $a$, $b$ and $D$, suffice to represent a triangular membership function of a fuzzy number.
+Three reel values, denoted $a$, $b$ and $D$, suffice to represent a triangular membership function of a fuzzy number and in this work, a fuzzy number defined as such will be noted as $\left[D, a, b \right]$.
+Here:
 \begin{itemize}
@@
 \end{itemize}
-In this work, a fuzzy number defined as such will be noted as $\left[D, a, b \right]$.
 \begin{figure}[h!]
@@
 \subsubsection{Interval Evaluation by Ill-known Constraints}
-The problem of interval evaluation is more generally explained in \cite{Pon11}: for a crisp interval $I = \left[a, b \right]$, the need exists to know if all points in this interval reside between the boundaries of an ill-known interval $\left[ X , Y \right]$.
-
-The notion of an \emph{ill-known constraint} is introduced:
+The problem of interval evaluation is more generally explained in \cite{Pon11}: the need exists to know if all points in a crisp interval $I$ reside between the boundaries of an ill-known interval $\left[ X , Y \right]$.
+In \cite{Pon11}, the notion of an \emph{ill-known constraint} is introduced:
 
 \begin{definition}
-Given a universe $U$, an \emph{ill-known constraint} $C$ on a set $A \subseteq U$ is specified by means of a binary relation $R \subseteq U^{2}$ and a fixed, ill-known value denoted by its variable $X$ over $U$, i.e.:
+Given a universe $U$, an \emph{ill-known constraint} $C$ on a set $A \subseteq U$ is specified by means of a binary relation $R \subseteq U^{2}$ and a fixed, ill-known value denoted by its possibilistic variable $V$ over $U$, i.e.:
 \begin{align}
-C \triangleq (R,X)
+C \triangleq (R,V)
+\end{align}
+Set $A$ now satisfies the constraint if and only if:
+\begin{align}
+\forall a \in A : (a,V) \in R
 \end{align}
@@
 
-We assume that $X$ specifies the lower bound and $Y$ the upper bound for a given interval, we want to known whether all points in the interval are larger than or equal to $X$ and smaller than or equal to $Y$.
+The satisfaction of a constraint $C \triangleq (R,V)$ by a set $A$ is now still a Boolean matter, but due to the uncertainty about the ill-known value $V$, it can be uncertain whether $C$ is satisfied by $A$ or not\cite{Pon11}.
+In fact, this satisfaction now behaves as a proposition.
+Based on the possibility distribution $\pi_{V}$ of $V$, the possibility and necessity that $A$ satisfies $C$ can be found.
+This proposition can thus be seen as a possibilistic variable on $\mathbb{B}$.
+The required possibility and necessity are:
+
+\begin{align}
+\Pos(A\text{ satisfies }C) & = \min_{a \in A}\left(\sup_{(a,w) \in R}\pi_{V}(w))\right) \\
+\Nec(A\text{ satisfies }C) & = \min_{a \in A}\left(\inf_{(a,w) \notin R} 1-\pi_{V}(w)\right)
+\end{align}
+
+Now, to check if crisp interval $I = \left[j, k\right]$ is included in $\left[X, Y\right]$, 2 ill-known constraints are constructed:
+
+%We assume that $X$ specifies the lower bound and $Y$ the upper bound for a given interval, we want to known whether all points in the interval are larger than or equal to $X$ and smaller than or equal to $Y$.
 Therefore, we consider two ill-known constraints:
@@
 \begin{eqnarray}
-C_1\stackrel{\triangle}{=}\left(\leq,X\right)\\
-C_2\stackrel{\triangle}{=}\left(\geq,Y\right).
+C_1\triangleq\left(\geq,X\right)\\
+C_2\triangleq\left(\leq,Y\right).
 \end{eqnarray}
 
-The following expressions compute the possibility and the necessity:
+The possibility and necessity of the set satisfying both constraints is then:
 
-\begin{eqnarray}
+\begin{align}
 \label{eq:interval-pos}
-\Pos\left(\lambda([a,b])\right)=\\
-\nonumber
-\min\bigg(\sup_{a\geq w}\pi_{X}(w),\sup_{b\leq w}\pi_{Y}(w)\bigg)\\
+\Pos(A\text{ satisfies }C_1\ AND\ C_2) & = \min_{a \in A}\left(\sup_{a \geq w}\pi_{X}(w),\sup_{a \leq v}\pi_{Y}(v)\right)\\
 \label{eq:interval-nec}
-\Nec\left(\lambda([a,b])\right)=\\
-\nonumber
-\min\bigg(\inf_{a<w}1-\pi_{X}(w),\inf_{b>w}1-\pi_{Y}(w)\bigg).
-\end{eqnarray}
+\Nec(A\text{ satisfies }C_1\ AND\ C_2) & = \min_{a \in A}\left(\inf_{a < w} 1-\pi_{X}(w),\inf_{a > v} 1-\pi_{Y}(v)\right).
+\end{align}
 
@@
 %In this section, the proposed reasoning is applied to the specific context of intervals on the real line. This setting is of specific interest in the context of fuzzy temporal databases.
-A temporal database \cite{Dyreson1994} is a database that manages some aspects of time in its schema.
-A \textbf{chronon} is the shortest duration of time supported by the database.
-The time can be represented either as points or intervals \cite{655777}.
-Fuzzy temporal models \cite{schockaert08} are proposed when the time points or intervals are not precisely known.
-There are proposals for the  fuzzyfication of the time point and the fuzzyfication of the time interval.\\
-Allen \cite{Allen83} studied the comparison between two crisp time intervals. For fuzzy intervals, several proposals \cite{ohlbach04,nagypal03,schockaert08} have been done.
+A \emph{temporal database} is a database that manages some aspects of time in its schema.
+The reality a temporal database tries to model, contains some temporal notions which have to be handled specifically in order to maintain a consistent modelling behavior.
+A \emph{chronon} is the shortest duration of time supported by the database.
+Time can be represented either as points or intervals \cite{655777}.
+%There are proposals for the fuzzyfication of the time point and the fuzzyfication of the time interval.
 
@@
 
-Despite of user-defined time (an uninterpreted attribute; supported in the standard SQL2), there are 3 types of time in a temporal database:
+The temporal notions in temporal databases can be classified into 4 types based on their interpretation and modelling purpose. User-defined time has no interpretation, but the other types do:
 
@@
 	\item
-	\textbf{Transaction time}: The time when the fact is stored in the database.
+	\emph{Transaction time}: The time when the fact is stored in the database.
 	\item
-	\textbf{Valid time}: The time when the fact is true in the modelled reality.
+	\emph{Valid time}: The time when the fact is true in the modelled reality.
 	\item
-	\textbf{Decision time} \cite{Nascimento95}: The time when an event was decided to happen. 
+	\emph{Decision time}: The time when an event was decided to happen. 
 \end{itemize}
 
-	 The  database model can be then classified into: transaction time \cite{Jensen91}, Valid time, bi-temporal \cite{Snodgrass84}(both valid and transaction time) or tri-temporal \cite{Nascimento95} (valid, transaction and decision time).
+Database models can then be classified into transaction time \cite{Jensen91}, decision time, valid time, bi-temporal \cite{Snodgrass84}(both valid and transaction time) or tri-temporal \cite{Nascimento95} (valid, transaction and decision time) models.
 
-Fuzzy temporal models \cite{schockaert08} deal with time points \cite{Dubois89} or intervals \cite{Garrido2009} that are ill-known.
-Some fuzzy temporal models assume that the time stored in the database is an interval.
+To deal with time points \cite{Dubois89} or intervals \cite{Garrido2009} that are subject to imperfection and can thus not be precisely known, fuzzy temporal models \cite{schockaert08} exist.
+%Some fuzzy temporal models assume that the time stored in the database is an interval.
 The temporal interval is represented by two ill-known time points: $X$  an ill-known starting point and $Y$ an ill-known ending point.

### 2011-12-01 ipmu2012/introduction.tex
More changes, resulting in one less page...

@@
 
-One issue in these attempts at temporal modeling is off course the inherent complexities of the concept of time.
+One issue in these attempts at temporal modelling is off course the inherent complexities of the concept of time.
 Another is the possible inherent imperfection in the description of notions of time.
@@
 
-To allow information systems to cope with these imperfections, many approaches adopt fuzzy sets for the representation of temporal information~\cite{343607,Dubois:jucs_9_9:fuzziness_and_uncertainty_in}.
+To allow information systems to cope with these imperfections, many approaches adopt fuzzy sets for the representation of temporal information~\cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy,Dubois:jucs_9_9:fuzziness_and_uncertainty_in}.
 The temporal relations studied by Allen were recently fuzzified by several authors~\cite{ohlbach04,schockaert08}.
 
-Next to the issues in temporal modelling, the addition of time to a relational database~\cite{Dyreson1994} leads to several more practical problems, as the modelling of a time-variant object has an impact on the database consistency.
-Different models have been proposed \cite{Jensen91,Snodgrass84} that consider time as a crisp notion.
+Next to the issues in temporal modelling, the addition of time to a relational database leads to several more practical problems, as the modelling of a time-variant object has an impact on the database consistency.
+Different models have been proposed \cite{Jensen91,Snodgrass84,Nascimento95} that consider time as a crisp notion.
 In recent works, some authors consider the necessity of allowing imprecision in temporal models~\cite{Cru97,Garrido2009}.

### 2011-12-01 ipmu2012/preliminaries.tex
A very small amend...

@@
 
-Both possibility measures and necessity measures are special cases of confidence measures.
+%Both possibility measures and necessity measures are special cases of confidence measures.
 
@@
 
-When a possibilistic variable is defined on the powerset $\Pow(R)$ of some universe $R$, the unique value the variable takes will be a crisp set.
-In this case, the exact set is off course (partially) unknown, but the available knowledge is given in the form of a possibility distribution on the powerset $\Pow(R)$, which describes the possibility of each crisp subset of $R$ to be the value the variable takes.
+When a possibilistic variable is defined on the powerset $\Pow(R)$ of some universe $R$, the unique value the variable takes will be a crisp set and its possibility distribution on the powerset $\Pow(R)$ will describe the possibility of each crisp subset of $R$ to be the value the variable takes.
 This exact value (a crisp set) the variable takes, is now called an \emph{ill-known set}\cite{Dubois88b}.
-
-In this work, an ill-known set will be defined and represented via its start and end point, which will be ill-known values.
-The correspondences and transitions between these two representations are part of the authors current research.
 
@@
 A specific application of possibilistic variables is obtained when the universe under consideration is the set of boolean values $\mathbb{B}$ = $\{T,F\}$.
-Indeed, any boolean propositional variable $p$ takes just one value in $\mathbb{B}$.
-If the knowledge on which value this variable $p$ will take is given by a possibility distribution $\pi_p$, the variable can be seen as a possibilistic variable.
-As the interest lies with the case where the proposition holds, the possibility and necessity that $p$ = $T$ (the proposition holds) demands most attention.
+Indeed, any boolean proposition $p$ takes just one value in $\mathbb{B}$.
+If the knowledge on which value this proposition $p$ will take is given by a possibility distribution $\pi_p$, the proposition can be seen as a possibilistic variable.
+As the interest lies with the case where the proposition holds, the possibility and necessity that $p$ = $T$ (the proposition holds) demand most attention.
 This possibility and necessity is noted here as:
 \begin{align}
-Pos(p) & = \pi_p(T) \label{propholdsposs} \\
-Nec(p) & = 1-\pi_p(F) \label{propholdsnecc}
+\text{Possibility that $p$ = $T$ (p holds):} \hspace{50pt} & Pos(p) = \pi_p(T) \label{propholdsposs} \\
+\text{Necessity that $p$ = $T$ (p holds):} \hspace{50pt} & Nec(p) = 1-\pi_p(F) \label{propholdsnecc}
 \end{align}
-Here, notation \ref{propholdsposs} denotes the possibility that $p$ = $T$ and the proposition holds, notation \ref{propholdsnecc} denotes the necessity that $p$ = $T$ and the proposition holds.
+%Here, notation \ref{propholdsposs} denotes the possibility that $p$ = $T$ and the proposition holds, notation \ref{propholdsnecc} denotes the necessity that $p$ = $T$ and the proposition holds.
+
+This work will deal with ill-known intervals.
+These are ill-known sets, defined and represented via a start and end point, which will be ill-known values.
+The elements of the set are the values between the start and end point.
+An closed ill-known interval with start point $X$ and end point $Y$ is noted here $\left[X, Y\right]$.
+The correspondences and transitions between the representations of ill-known sets and between the representation of an ill-known set and an ill-known interval are part of the authors current research.
 
@@
 \subsubsection{Interval Evaluation by Ill-known Constraints}
-The problem of the interval evaluation is explained in \cite{Pon11}: for a crisp interval $I = \left[ a, b \right]$, the need exists to know if all points in this interval reside between the boundaries of the ill-known interval $\left[ X , Y \right]$, where $X,Y$ are ill-known values on the set of real numbers $\mathbb{R}$, and $\lambda$ is the evaluation function.
+The problem of interval evaluation is more generally explained in \cite{Pon11}: the need exists to know if all points in a crisp interval $I$ reside between the boundaries of an ill-known interval $\left[ X , Y \right]$.
+In \cite{Pon11}, the notion of an \emph{ill-known constraint} is introduced:
 
-We assume that $X$ specifies the lower bound and $Y$ the upper bound for a given interval, we want to known whether all points in the interval are larger than or equal to $X$ and smaller than or equal to $Y$.
+\begin{definition}
+Given a universe $U$, an \emph{ill-known constraint} $C$ on a set $A \subseteq U$ is specified by means of a binary relation $R \subseteq U^{2}$ and a fixed, ill-known value denoted by its possibilistic variable $V$ over $U$, i.e.:
+\begin{align}
+C \triangleq (R,V)
+\end{align}
+Set $A$ now satisfies the constraint if and only if:
+\begin{align}
+\forall a \in A : (a,V) \in R
+\end{align}
+\end{definition}
+
+The satisfaction of a constraint $C \triangleq (R,V)$ by a set $A$ is now still a Boolean matter, but due to the uncertainty about the ill-known value $V$, it can be uncertain whether $C$ is satisfied by $A$ or not\cite{Pon11}.
+In fact, this satisfaction now behaves as a proposition.
+Based on the possibility distribution $\pi_{V}$ of $V$, the possibility and necessity that $A$ satisfies $C$ can be found.
+This proposition can thus be seen as a possibilistic variable on $\mathbb{B}$.
+The required possibility and necessity are:
+
+\begin{align}
+\Pos(A\text{ satisfies }C) & = \min_{a \in A}\left(\sup_{(a,w) \in R}\pi_{V}(w))\right) \\
+\Nec(A\text{ satisfies }C) & = \min_{a \in A}\left(\inf_{(a,w) \notin R} 1-\pi_{V}(w)\right)
+\end{align}
+
+Now, to check if crisp interval $I = \left[j, k\right]$ is included in $\left[X, Y\right]$, 2 ill-known constraints are constructed:
+
+%We assume that $X$ specifies the lower bound and $Y$ the upper bound for a given interval, we want to known whether all points in the interval are larger than or equal to $X$ and smaller than or equal to $Y$.
 Therefore, we consider two ill-known constraints:

### 2011-12-06 trunk/ipmu2012/preliminaries.tex
Mr. T modifications.

@@
 It can be shown that such a membership function $\mu_T$ for a fuzzy interval $T$ is convex and normalized.
-A visualization is given in figure \ref{fig:trapezoidal}.
-Four reel values, denoted $\alpha$, $\beta$, $\gamma$ and $\delta$ and chosen as in figure \ref{fig:trapezoidal}, suffice to represent a trapezoidal membership function of a fuzzy interval.
+Four reel values, denoted $\alpha$, $\beta$, $\gamma$ and $\delta$ suffice to represent a trapezoidal membership function of a fuzzy interval.
 In this work, a fuzzy interval defined as such will be noted as $\left[\alpha, \beta, \gamma, \delta\right]$.
@@
 
-\def\JPicScale{0.5}
-\begin{figure}[h!]
-  \centering
-  \input{./graphs/trapezoidal.tex}
-  \caption{Trapezoidal membership function}
-  \label{fig:trapezoidal}
-\end{figure}
+%\def\JPicScale{0.5}
+%\begin{figure}[h!]
+%  \centering
+%  \input{./graphs/trapezoidal.tex}
+%  \caption{Trapezoidal membership function}
+%  \label{fig:trapezoidal}
+%\end{figure}
 
@@
 It can be shown that such a membership function $\mu_M$ for a fuzzy number $M$ is convex and normalized.
-A visualization is given in figure \ref{fig:triangular}.
 Three reel values, denoted $a$, $b$ and $D$, suffice to represent a triangular membership function of a fuzzy number and in this work, a fuzzy number defined as such will be noted as $\left[D, a, b \right]$.
@@
 \end{itemize}
-\begin{figure}[h!]
-  \centering
-  \input{./graphs/triangular.tex}
-  \caption{Triangular membership function.}
-  \label{fig:triangular}
-\end{figure}
+%\begin{figure}[h!]
+%  \centering
+%  \input{./graphs/triangular.tex}
+%  \caption{Triangular membership function.}
+%  \label{fig:triangular}
+%\end{figure}
 

### 2011-12-06 trunk/ipmu2012/proposal.tex
Mr. T modifications.

--- 2011-11-14 trunk/ipmu2012/proposal.tex
@@
 
-\subsection{Representation of ill-known valid-time intervals}
+\subsection{Representation of Ill-known Valid-time Intervals}
+\label{subsec:representation-ill-known}
 Valid time is usually represented as an interval. The interval has a starting and an ending points. An ill-known valid-time interval is an interval in witch one or both points are ill-known.
@@
   \centering
-  \includegraphics[scale=0.2]{graphs/2_triangular.eps}
-  \caption{Two fuzzy numbers $X$ and $Y$ denoting a Possibilistic Valid-Time Period \emph{PVT}.}
+  \input{./graphs/2_triangular.tex}
+  \caption{Two fuzzy numbers $X$ and $Y$ denoting a Possibilistic Valid-Time Period \emph{PVP}.}
   \label{fig:pvp}
@@
 
-\subsection{Storage of valid-time intervals}
+\subsection{Storage of Valid-time Intervals}
+\label{subsec:storage}
 Each database row containing a \emph{PVP} stores it as two triangular possibility distributions.
 In our approach we propose the representation of that as proposed in the  fuzzy interface for relational databases \emph{FIRST}~\cite{Medina94gefred.a,Gal98}.
-In this representation it is also possible to represent not only fuzzy numbers but fuzzy constants (see table \ref{table:relational-representation-pvp}):
+In this representation it is also possible to represent not only fuzzy numbers but fuzzy constants (see table \ref{table:relational-representation-pvp}).
+Note that the field \emph{FT} denotes the Fuzzy Type (\emph{NULL}, \emph{UNKNOWN}, \emph{UNDEFINED} or $\left[D,\ a,\ b \right]$) and the fields F1 to F3 store the value.
+Note also that while \emph{NULL} denotes the fuzzy constant, \emph{N} denotes a null value in the database.
 
-\begin{itemize}
-\item
-\emph{NULL}: This constant refers to a completely ignorance about the value.
+%\begin{itemize}
+%\item
+%\emph{NULL}: This constant refers to a completely ignorance about the value.
 The possibility distribution for a given fuzzy number $X$ is not defined, therefore, any comparison between a fuzzy number and the \emph{NULL} constant always returns $0$.
-\item
-\emph{UNKNOWN}: The point has a value but it is unknown. The possibility distribution for a given fuzzy number $X$ is $\pi_X=1$
-\item
-\emph{UNDEFINED}: The point does not have a value. The possibility distribution for a given fuzzy number $X$ is $\pi_X=0$
-\end{itemize}
+%\item
+%\emph{UNKNOWN}: % The possibility distribution for a given fuzzy number $X$ is $\pi_X=1$
+%\item
+%\emph{UNDEFINED}: The point does not have a value. %The possibility distribution for a given fuzzy number $X$ is $\pi_X=0$
+%\end{itemize}
+%
 
 \begin{table}
-\caption{Relational representation for a Possibilistic Valid-Time Period.}
+\caption{Relational representation for an ill-known time point.}%  A \emph{PVP} is represented by two ill-known points.}
 \centering
-\begin{tabular}{c c c c c c}
+\begin{tabular}{c c c c c c c p{2cm}}
 \hline
-Value & FT & F1 & F2 & F3  \\ \hline
-UNKNOWN & 0 & NULL & NULL & NULL  \\ 
-UNDEFINED & 1 & NULL & NULL & NULL  \\ 
-NULL & 2 & NULL & NULL & NULL  \\ 
-$\left[D,\ a,\ b \right]$ & 3 & $D$ & $D-a$ & $D+b$ \\ 
+Value & FT & F1 & F2 & F3 & $\mu(x)$ & Description \\ \hline
+UNKNOWN & 0 & N & N & N  & $1$ & The point has a value but it is unknown. \\ 
+UNDEFINED & 1 & N & N & N & $0$ & The point does not have a value. \\ 
+NULL & 2 & N & N & N &not defined. & Completely ignorance aboute the value. \\ 
+$\left[D,\ a,\ b \right]$ & 3 & $D$ & $a$ & $b$ & $\mu(\left[D,\ a,\ b \right])$ & triangular poss. distr. \\ 
 \hline
@@
 
-\subsection{Querying ill-known valid-time intervals}
+\subsection{Querying Ill-known Valid-time Intervals}
+In order to provide a complete model, we will provide the tools for querying.
+This allows the user to specify both the preferences and an ill-known valid-time interval in the query.
+It is important to notice that the possibilistic / fuzzy data stored in the database has a \emph{disjunctive interpretation} (it is said that we have \emph{uncertainty}: the valid-time interval has only one value but, for some reason the value is ill-known).
+In the query specification, the user is allowed to express a crisp time interval.
+In this subsection we will define the query specification, then the evaluation of the query and finally the ranking for the query.
 
 \subsubsection{Query specification}
+A query in our framework has two different parts: the first one is the temporal specification. The second part is the query specification for regular attributes.
+
+\begin{definition}
+A query $\tilde Q$ is specified by:
+\begin{equation}
+\label{eq:query-definition}
+\tilde Q = \left( Q^{time}, Q \right)
+\end{equation}
+\end{definition}
+Where $Q$ are the (possibly fuzzy) preferences of the user  and $Q^{time}$ is the temporal part specified by a crisp interval.
+\begin{definition}
+ $Q^{time}$ is composed by:
+\begin{equation}
+Q^{time} = \left( \left[a, \ b \right] , Ar \right), a,b \in \mathbb{R}
+\end{equation}
+Where $ \left[a, \ b \right] $ is the specification for the crisp interval and $Ar$ is one of the Allen's relations (table \ref{tab:allen-relations}).
+\end{definition}
+
+\begin{table}[h]
+\label{tab:allen-relations}
+\centering
+\begin{tabular}{|c|c|c|c|}
+\hline
+Allen Relation & Constraints & $\bool(p_1,...,p_n)$ & $\Pos\left(\lambda \left[a,b \right] \right)$ \\
+\hline
+I before J & $C_1\stackrel{\triangle}{=} \left(<,X\right)$ & $p_1$ & $\sup_{a>w}\pi_x(w)$\\
+\hline
+\multirow{2}{*}
+{I equal J} & $C_1\stackrel{\triangle}{=} \left(\geq,X\right)$,$C_2\stackrel{\triangle}{=} \left(\neq,X\right)$ & \text{$p_1\wedge\neg p_2\wedge p_3\wedge\neg p_4$} & $\min ( \sup_{a \leq w}\pi_x(w),\pi_x(w),$\\
+ & $C_3\stackrel{\triangle}{=} \left(\leq,Y\right)$,$C_4\stackrel{\triangle}{=} \left(\neq,Y\right)$ & & $\sup_{b \geq w}\pi_Y(w),\pi_Y(w))$\\
+% &  & &  \\
+% &  & &  \\
+\hline
+%\multirow{2}{*}
+{I meets J} & $C_1\stackrel{\triangle}{=} \left(\leq,X\right)$ ,$C_2\stackrel{\triangle}{=} \left(\neq,X\right)$& $p_1\wedge\neg p_2$ & $\min (\sup_{a\geq w} \pi_X(w),$ $\pi_X(w))$ \\
+% &  &  & \\
+\hline
+\multirow{3}{*}
+{I overlaps J} & $C_1\stackrel{\triangle}{=} \left(<,Y\right)$,$C_2\stackrel{\triangle}{=} \left(\leq,X\right)$ & $p_1\wedge\neg p_2\wedge\neg p_3$ & $\min ( \sup_{b>w}\pi_Y(w), $ \\
+ & $C_3\stackrel{\triangle}{=} \left(\geq,X\right)$ & & $\sup_{a \geq w}\pi_X(w),$ \\
+ &  & & $\sup_{a \leq w}\pi_X(w))$ \\
+\hline
+\multirow{4}{*}
+{I during J} & $C_1\stackrel{\triangle}{=} \left(>,X\right)$, $C_2\stackrel{\triangle}{=} \left(\leq,Y\right)$ & $(p_1\wedge p_2)\vee(p_3\wedge p_4)$ & $max ( min ( \sup_{a<w}\pi_X(w),$ \\
+ & $C_3\stackrel{\triangle}{=} \left(\geq,X\right)$ ,$C_4\stackrel{\triangle}{=} \left(<,Y\right)$ & & $\sup_{b \geq w}\pi_Y(w)),$\\
+ & & & $\min ( \sup_{a \leq w }\pi_X(w),$ \\
+ &  & & $\sup_{b>w}\pi_Y(w)$\\
+\hline
+%\multirow{2}{*}
+{I starts J} & $C_1\stackrel{\triangle}{=} \left(\geq,X\right)$,$C_2\stackrel{\triangle}{=} \left(\neq,X\right)$  & $p_1\wedge\neg p_2$ & $\min( \sup_{a \leq w}\pi_X(w),$ $\pi_X(w))$\\
+% &  & & \\
+\hline
+%\multirow{2}{*}
+{I finishes J} & $C_1\stackrel{\triangle}{=} \left(\leq,Y\right)$, $C_2\stackrel{\triangle}{=} \left(\neq,Y\right)$  & $p_1\wedge\neg p_2$ & $\min ( \sup_{b \geq w} \pi_Y(w),$ $\pi_Y(w))$\\
+% && & \\
+\hline
+\end{tabular}
+
+\caption{Allen's relations represented in the framework.}
+\end{table}
 
 \subsubsection{Query evaluation}
+In fuzzy querying of regular (relational) databases, the query satisfaction modelling is a matter of degree.
+Usually, criteria satisfaction is modelled by means of a satisfaction degree $s \in \left[ 0, 1\right]$.
+In the model, every record $r$ contains a \emph{PVP}, $V_r$, to model the valid-time.The query evaluation method is the following:
+\begin{itemize}
+\item
+For each record $r$ in the database, the preferences expressed in $Q$ are evaluated in the unit interval. Thus, $e(Q) \in \left[0, 1\right]$ is the evaluation function for the criteria in $Q$.
+\item
+For each record $r$ in the database, the interval $ \left[a, \ b \right]$ is compared to the $PVP$ value by means of the corresponding Allen relation specified by $ar$.
+Here, $e(Q^{time}) \in \left[0, 1 \right]$ is the evaluation function for the criteria $Q^{time}$.
+\end{itemize}
 
-\subsubsection{Ranking}
+\subsubsection{Ranking and aggregation}
+In order to present the results to the user, a ranking method must be used. Both $Q$ and $Q^{time}$ are evaluated independently and also both are evaluated in the unit interval.
+First of all, the possibility and the necessity are ranked with the convex combination in \eqref{eq:convex-comb} with $\omega=0.5$.
+Finally this value is aggregated with the rest of the criteria with the same combination.
+%In this case the value for the aggregation is also $\omega = 0.5$
+
+\begin{equation}
+\label{eq:convex-comb}
+e(u,v)\ =\ \omega*v\ +\ (1-\omega)*v, \omega \in \left[0, 1 \right]
+\end{equation}
+
+ By increasing the value for the parameter $\omega$, the preferences expressed in the query $Q$ can be given more importance. By lowering the value for $\omega$, the temporal criteria is emphasized.
+%The following example illustrates the querying process.
+
+\begin{example}
+
+\textbf{The database}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Sample table for the car models. 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{table}[ht]
+\caption{Sample database containing the car models.}
+\centering
+\begin{tabular}{c c c c c c c}
+\hline
+ID & IID & Segment & Manufac. & Name & Start & End  \\ [0.5ex]
+\hline
+001 & 1 & B & Peugeot & 205 & [1985,2,3] & [1997,2,1] \\
+002 & 1 & C & Peugeot & 305 & [1977,2,2] & [1989,2,3] \\
+003 & 1 & B & Citroen & C2 & [2002,2,2] & [2006,1,1] \\
+001 & 2 & B & Peugeot & 206 & [2000,1,2] & [2011,2,1] \\
+001 & 3 & B & Peugeot & 207 & [2006,1,1] & [2011,1,1]\\
+\hline
+\end{tabular}
+\label{tb:car-models}
+\end{table}
+
+Consider a database containing car models.
+There are several general attributes (car model's name, car manufacturer, car segment) and one temporal attribute (which is ill-known) containing the approximate date while the car model was sold.
+The temporal data is stored as explained in subsection \ref{subsec:storage}.
+The value for $D$ is stored in \emph{yyyy} format and $a$ and $b$ are represented by an integer, for the convenience of the representation.
+Therefore the chronons in our example will be years.
+The ID field identifies a car model while the field Instance ID (IID) identifies the instance for a car model.
+\\
+
+\textbf{The query}
+Consider the following query:
+
+\emph{The user wants to obtain a list of models in the segment B for the manufacturer Peugeot before the year interval 2001-2005.}
+
+The query is translated to the following expression, using equation \eqref{eq:query-definition}. The evaluation for the criteria are presented in the result table \ref{tb:results}.
+
+\begin{center}
+$\tilde{Q} = \left(  c^{time}, c_{segment} \right)$
+\end{center}
+
+Where:
+\begin{itemize}
+\item
+$c^{time}\ = \ ( \left[ 2001,\ 2005 \right],\ $  before $)$.
+\item
+$c_{segment}\ = \ $ Peugeot.
+\end{itemize}
+
+\begin{table}[ht]
+\caption{Result table and ranking}
+\centering
+\begin{tabular}{c c c c c c c}
+\hline
+ID & IID &  $\Pos_{c^{time}}$ &$\Nec_{c^{time}}$ & $e(C^{time})$ & $c_{segment}$ & rank ($\omega=0.5$) \\ [0.5ex]
+\hline
+001 & 1 & 1 &  1 & 1 & 1 & 1 \\
+002 & 1 & 1 & 1 & 1 & 0.5 & 0.75 \\
+003 & 1 & 1 & 0.5 & 0.75 &0 & 0.375\\
+001 & 2 & 1 & 0 & 0.5 &1 & 0.75 \\
+001 & 3 & 0 & 0 & 0 &1 & 0.5\\
+\hline
+\end{tabular}
+\label{tb:results}
+\end{table}
+\end{example}

### 2011-12-06 ipmu2012/ipmu.tex
Mr. T modifications.

@@
 \usepackage{amsmath}
+\usepackage{amssymb}
 \usepackage{amsfonts}
-
+\usepackage{multirow}
 \hyphenation{pro-per-ties}
@@
 \maketitle              % typeset the title of the contribution
-
+\def\JPicScale{0.5}
 \begin{abstract}
@@
 %
-\section{Conclusions and further research work}
+\section{Conclusions and Further Research}
 \label{sec:futher-research}

### 2011-12-12 ipmu2012/ipmu.tex
Last temporary changes... this evening final changes.

@@
 \begin{abstract}
-
+Information in databases can be imperfect and this imperfection has several forms and causes.
+In some cases, a single value should be stored, but it is (partially) unknown.
+The uncertainty about which value to store leads to the aforementioned imperfection.
+In temporal databases, uncertainty can arise, concerning which temporal notion needs to be stored.
+Because in temporal databases, temporal notions influence the consistency with which the database models the reality, this uncertainty has a direct impact on the consistency of the model.
+To represent this temporal uncertainty, previous works have adapted fuzzy sets with conjunctive interpretation.
+This approach might prove misleading.
+This work presents a model that represents the uncertainty using possibility and necessity distributions, which are fuzzy sets with disjunctive interpretations.
 \end{abstract}

### 2011-12-12 ipmu2012/preliminaries.tex
Last temporary changes... this evening final changes.

@@
 %Introductory text:
-In this section, some basic concepts are introduced, concerning possibility theory, possibilistic variables and fuzzy numbers and intervals.
-The framework of set evaluation by ill-known constraints\cite{Pon11} is explained.
+In this section, some basic concepts are introduced concerning possibilistic variables and fuzzy numbers and intervals, after which the framework of set evaluation by ill-known constraints\cite{Pon11} is explained.
 The section concludes with a brief introduction to temporal databases.
 
-\subsection{\label{subsec:possibility-theory}Possibility Theory}
-Possibility theory, like probability theory, deals with uncertainty about the outcome of an experiment.
+%\subsection{\label{subsec:possibility-theory}Possibility Theory}
+%Possibility theory, like probability theory, deals with uncertainty about the outcome of an experiment.
 In probability theory, this uncertainty is caused by the \emph{variability} in the outcomes, while in possibility theory, the uncertainty is caused by \emph{incomplete knowledge} about the experiment.
-The quantification of confidence in a theory of uncertainty is achieved using a confidence measure\cite{Shafer:1976:AMathematical}.
+The quantification of confidence in a theory of uncertainty is achieved using a confidence measure\cite{Pon11}.
 In probability theory this is a measure of chance, in possibility theory, possibility and necessity measures are used.
 
-\begin{definition}
-Consider a set of outcomes $\Omega$.
+%\begin{definition}
+%Consider a set of outcomes $\Omega$.
 Let $\wp(\Omega)$ denote the powerset of $\Omega$ and let $A$ and $B$ be elements of $\wp(\Omega)$.
 A \emph{confidence measure on $\Omega$} is defined by a function
-	\begin{align}
-	g : \wp(\Omega) & \rightarrow \left[0,1\right]
-	\end{align}
-that satisfies
-	\begin{align}
-	g(\emptyset) &= 0 \\
-	g(\Omega) &= 1 	\label{NormalizationProperty} \\
-	A \subseteq B &\Rightarrow g(A) \leq g(B) \label{MonotonicityProperty}
-	\end{align}
-\end{definition}
+%	\begin{align}
+%	g : \wp(\Omega) & \rightarrow \left[0,1\right]
+%	\end{align}
+%that satisfies
+%	\begin{align}
+%	g(\emptyset) &= 0 \\
+%	g(\Omega) &= 1 	\label{NormalizationProperty} \\
+%	A \subseteq B &\Rightarrow g(A) \leq g(B) \label{MonotonicityProperty}
+%	\end{align}
+%\end{definition}
 
@@
 
-\begin{definition}
-Consider a confidence measure $\Pi$ on a set of outcomes $\Omega$.
+%\begin{definition}
+%Consider a confidence measure $\Pi$ on a set of outcomes $\Omega$.
 Let $J$ be a countable index set and let $\{ A_{j} | j \in J \wedge A_{j} \subseteq \Omega \}$ be a family of elements of $\wp(\Omega)$.
 $\Pi$ is now a \emph{possibility measure on $\Omega$} if it satisfies:
-	\begin{align}
-	\Pi\left(\bigcup_{j \in J} A_{j} \right) = \sup_{j \in J} \Pi(A_{j})
-	\end{align}
-\end{definition}
+%	\begin{align}
+%	\Pi\left(\bigcup_{j \in J} A_{j} \right) = \sup_{j \in J} \Pi(A_{j})
+%	\end{align}
+%\end{definition}
 
-In this work, the interpretation is as follows.
+%In this work, the interpretation is as follows.
 The possibility of an event expresses how plausible the occurrence of the event seems to an observer of the experiment, given the (partial) knowledge of the observer about the experiment.
 
-Information on the possibility of distinct elements of the universe of discourse $\Omega$ can now be given by a \emph{possibility distribution} $\pi$ on $\Omega$, defined by:
+%Information on the possibility of distinct elements of the universe of discourse $\Omega$ can now be given by a \emph{possibility distribution} $\pi$ on $\Omega$, defined by:
 
-\begin{definition}
-Consider a possibility measure $\Pi$ on $\Omega$. A \emph{possibility distribution} $\pi$ on $\Omega$ underlying the possibility measure $\Pi$ is then a function defined by:
-	\begin{align}
-	\pi : \Omega \rightarrow \left[0, 1\right] : \pi(u) = \Pi(\{u\})
-	\end{align}
-\end{definition}
+%\begin{definition}
+%Consider a possibility measure $\Pi$ on $\Omega$. A \emph{possibility distribution} $\pi$ on $\Omega$ underlying the possibility measure $\Pi$ is then a function defined by:
+%	\begin{align}
+%	\pi : \Omega \rightarrow \left[0, 1\right] : \pi(u) = \Pi(\{u\})
+%	\end{align}
+%\end{definition}
 
-\begin{definition}
-Consider a confidence measure $N$ on a set of outcomes $\Omega$.
+%\begin{definition}
+%Consider a confidence measure $N$ on a set of outcomes $\Omega$.
 Let $J$ be a countable index set and let $\{ A_{j} | j \in J \wedge A_{j} \subseteq \Omega \}$ be a family of elements of $\wp(\Omega)$.
 $N$ is now a \emph{necessity measure} on $\Omega$ if it satisfies:
-	\begin{align}
-	N\left(\bigcap_{j \in J} A_{j} \right) = \inf_{j \in J} N(A_{j})
-	\end{align}
-\end{definition}
+%	\begin{align}
+%	N\left(\bigcap_{j \in J} A_{j} \right) = \inf_{j \in J} N(A_{j})
+%	\end{align}
+%\end{definition}
 
-In this work, the interpretation is as follows.
+%In this work, the interpretation is as follows.
 The necessity of an event expresses how necessary the occurrence of the event seems to an observer of the experiment, given the (partial) knowledge of the observer about the experiment.
 
-Possibility and necessity measures are dual in the sense that:
+%Possibility and necessity measures are dual in the sense that:
 
-\begin{align}
-\forall A \subseteq \Omega : N(A) = 1 - \Pi(\bar{A})
-\end{align}
+%\begin{align}
+%\forall A \subseteq \Omega : N(A) = 1 - \Pi(\bar{A})
+%\end{align}
 
-Regarding interpretation, the above can be seen as: the degree to which an event is necessary is the degree to which every other possible event is not plausible.
+%Regarding interpretation, the above can be seen as: the degree to which an event is necessary is the degree to which every other possible event is not plausible.
 
 \subsection{\label{subsec:possibilistic-variables}Possibilistic Variables}
-A \emph{possibilistic variable} is defined as follows\cite{Pon11}.
+Possibilistic variables rely on possibility theory\cite{Dubois:Prade:1988:PossibilityTheory}. A \emph{possibilistic variable} is defined as follows\cite{Pon11}.
 
@@
 Its possibility distribution $\pi_X$ on $U$ gives the available knowledge about the value that $X$ takes: for each $u\in U$, $\pi_X(u)$ represents the possibility that $X$ takes the value $u$.
+In this work, this possibility is interpreted as a measure of how plausible it is that $X$ takes the value $u$, given (partial) knowledge about the value $X$ takes.
 \end{definition}
@@
 The elements of the set are the values between the start and end point.
-An closed ill-known interval with start point $X$ and end point $Y$ is noted here $\left[X, Y\right]$.
+A closed ill-known interval with start point defined by possibilistic variable $X$ and end point by possibilistic variable $Y$ is noted here $\left[X, Y\right]$.
 The correspondences and transitions between the representations of ill-known sets and between the representation of an ill-known set and an ill-known interval are part of the authors current research.
@@
 \subsection{\label{subsec:fuzzy-numbers}Fuzzy Numbers and Fuzzy Intervals}
-Among others, Dubois and Prade~\cite{Dubois1983} use the following definition of a \emph{fuzzy interval}.
+Among others, Dubois and Prade~\cite{Dubois1983} use fuzzy sets\cite{zadeh65} to define a \emph{fuzzy interval}:
 \begin{definition}
@@
 
-A simple form of the membership function of a fuzzy interval is a trapezoidal function.
+%A simple form of the membership function of a fuzzy interval is a trapezoidal function.
 It can be shown that such a membership function $\mu_T$ for a fuzzy interval $T$ is convex and normalized.
-A visualization is given in figure \ref{fig:trapezoidal}.
-Four reel values, denoted $\alpha$, $\beta$, $\gamma$ and $\delta$ and chosen as in figure \ref{fig:trapezoidal}, suffice to represent a trapezoidal membership function of a fuzzy interval.
+Four reel values, denoted $\alpha$, $\beta$, $\gamma$ and $\delta$ suffice to represent a trapezoidal membership function of a fuzzy interval.
 In this work, a fuzzy interval defined as such will be noted as $\left[\alpha, \beta, \gamma, \delta\right]$.
@@
 It can be shown that such a membership function $\mu_M$ for a fuzzy number $M$ is convex and normalized.
-A visualization is given in figure \ref{fig:triangular}.
-Three reel values, denoted $a$, $b$ and $D$, suffice to represent a triangular membership function of a fuzzy number and in this work, a fuzzy number defined as such will be noted as $\left[D, a, b \right]$.
+Three real values, denoted $a$, $b$ and $D$, suffice to represent a triangular membership function of a fuzzy number and in this work, a fuzzy number defined as such will be noted as $\left[D, a, b \right]$.
 Here:
@@
 \end{itemize}
-\begin{figure}[h!]
-  \centering
-  \input{./graphs/triangular.tex}
-  \caption{Triangular membership function.}
-  \label{fig:triangular}
-\end{figure}
+
+The membership function of such a fuzzy number is then given by:
+
+\begin{align}
+\mu_M : & \quad \mathbb{R} \rightarrow \left[0,1\right] \nonumber\\
+ : & \quad x \rightarrow
+\begin{cases}
+0 & \mbox{ if } \left(x < D-a\right) \vee \left(x > D+b\right) \\
+\frac{1}{a}\left[x-\left(D-a\right)\right] & \mbox{ if } \left(x \geq D-a\right) \wedge (x \leq D)  \\
+-\frac{1}{b}\left[x-\left(D+b\right)\right] & \mbox{ if } \left(x \geq D\right) \wedge (x \leq D+b)
+\end{cases}
+\end{align}
+
+%\begin{figure}[h!]
+%  \centering
+%  \input{./graphs/triangular.tex}
+%  \caption{Triangular membership function.}
+%  \label{fig:triangular}
+%\end{figure}
 
@@
 
-\subsubsection{Interval Evaluation by Ill-known Constraints}
+\subsection{Interval Evaluation by Ill-known Constraints}
 The problem of interval evaluation is more generally explained in \cite{Pon11}: the need exists to know if all points in a crisp interval $I$ reside between the boundaries of an ill-known interval $\left[ X , Y \right]$.
@@
 \begin{align}
-\Pos(A\text{ satisfies }C) & = \min_{a \in A}\left(\sup_{(a,w) \in R}\pi_{V}(w))\right) \\
-\Nec(A\text{ satisfies }C) & = \min_{a \in A}\left(\inf_{(a,w) \notin R} 1-\pi_{V}(w)\right)
+\Pos(A\text{ satisfies }C) & = \min_{a \in A}\left(\sup_{(a,w) \in R}\pi_{V}(w))\right) \label{ill-known-pos}\\
+\Nec(A\text{ satisfies }C) & = \min_{a \in A}\left(\inf_{(a,w) \notin R} 1-\pi_{V}(w)\right) \label{ill-known-nec}
 \end{align}
@@
 
-The possibility and necessity of the set satisfying both constraints is then:
+To calculate possibility and necessity concerning a conjunction of constraints, the $\min$ operator is used. The possibility and necessity of the set satisfying both constraints is then:
 

### 2011-12-12 trunk/ipmu2012/introduction.tex
Last temporary changes... this evening final changes.

@@
 First of all, the concept of \emph{time} itself is very complex.
-Klein~\cite{klein94} studied the concept of time in language and Devos~\cite{devos94} modelled vague temporal expressions by means of fuzzy sets~\cite{zadeh65}.
-More towards information systems, several proposals have been concerned with the obtaining of theoretical models that allow the representation of time in information systems~\cite{devos98,Cru97}.
+More towards information systems, several proposals have been concerned with the obtaining of theoretical models that allow the representation of time in information systems~\cite{Cru97}.
 Notably, temporal relations between time intervals and to a lesser extent time points, were first studied by Allen~\cite{Allen83}.
 
-Next to the inherent difficulties of time, an issue is \emph{temporal granularity}.
+%Next to the inherent difficulties of time, an issue is \emph{temporal granularity}.
 A temporal granularity is a partitioning of the time line used by a system, usually dependent on the application.
@@
 
-Another issue is the possible inherent imperfection in the description of notions of time.
+One issue in these attempts at temporal modelling is off course the inherent complexities of the concept of time.
+Another is the possible inherent imperfection in the description of notions of time.
 The temporal notion in a sentence like `The Alhambra, which I can see from my room, is a few hundred years old.' contains imperfection caused by the vagueness of the used expression.
@@
 
-To allow information systems to cope with imperfections, many approaches adopt fuzzy sets for the representation of temporal information~\cite{343607,nagypal03,Dubois:jucs_9_9:fuzziness_and_uncertainty_in}, though some other tools, like rough sets, have also been used~\cite{Qia09}.
-The temporal relations studied by Allen were recently fuzzified by several authors~\cite{ohlbach04,nagypal03,schockaert08}.
+To allow information systems to cope with these imperfections, many approaches adopt fuzzy sets for the representation of temporal information~\cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy,Dubois:jucs_9_9:fuzziness_and_uncertainty_in}.
+The temporal relations studied by Allen were recently fuzzified by several authors~\cite{ohlbach04,schockaert08}.
 
-Next to the issues in temporal modelling, the addition of time to a relational database~\cite{Dyreson1994} leads to several more problems, as the modelling of a time-variant object has an impact on the database consistency.
-Different models have been proposed \cite{Jensen91,Snodgrass84,Nascimento95}, but all of these consider time as a crisp notion.
-In recent works though, some authors consider the necessity of allowing imprecision in temporal models~\cite{Cru97,fuzz2009}.
+Next to the issues in temporal modelling, the addition of time to a relational database leads to several more practical problems, as the modelling of a time-variant object has an impact on the database consistency.
+Different models have been proposed \cite{Jensen91,Snodgrass84,Nascimento95} that consider time as a crisp notion.
+In recent works, some authors consider the necessity of allowing imprecision in temporal models~\cite{Cru97,Garrido2009}.
 
 Most of these models use the concepts of fuzzy numbers and fuzzy intervals to represent temporal imprecision and considerable attention has thus been given to the problem of transforming two fuzzy numbers which represent the boundaries of an imprecise time interval into one fuzzy interval representing this interval.
-There are several proposals for this transformation, but some, like those presented in~\cite{fuzz2009} are misleading.
+There are several proposals for this transformation, but some, like those presented in~\cite{Garrido2009} are misleading.
 
-In this work we will propose a possibilistic valid time model for relational databases.
-The base of the model is the possibilistic framework presented in~\cite{Pon11}.
-This framework deals with the inference of uncertainty about the evaluation of sets from the uncertainty about the exact values used in evaluation.
+In this work we will propose a possibilistic valid time model for relational databases, based on the framework presented in~\cite{Pon11}, which deals with the inference of uncertainty about the evaluation of sets from the uncertainty about the exact values used in evaluation.
 
-The paper is organized as follows. In section \ref{sec:preliminaries}, some preliminaries and concepts about the possibilistic evaluation framework and temporal databases are presented. 
+The paper is organized as follows. In section \ref{sec:preliminaries}, some preliminaries and concepts are presented. 
 Section \ref{sec:proposal} presents the possibilistic model for valid-time databases. In this section, both representation and querying are introduced.

### 2011-12-12 trunk/ipmu2012/proposal.tex
Last temporary changes... this evening final changes.

--- 2011-11-14 trunk/ipmu2012/proposal.tex
@@
-The proposal consist on a possibilistic valid-time model. The representation and the querying are explained in the following subsections.
+In this section, a Valid-time model will be presented.
+The focus will off course be on the handling of imperfect Valid-time notions.
+First, the representation and storage of Valid-time will be presented.
+Next, a way of querying the model is proposed.
+Finally, an example is given.
+%The proposal consist on a possibilistic valid-time model. The representation and the querying are explained in the following subsections.
 
-\subsection{Representation of ill-known valid-time intervals}
-Valid time is usually represented as an interval. The interval has a starting and an ending points. An ill-known valid-time interval is an interval in witch one or both points are ill-known.
+\subsection{Valid-time Intervals}
+\subsubsection{Representation}
+\label{subsec:representation-ill-known}
+A database models a real object by storing the object's value for each attribute describing a property of the object.
+Thus, a Valid-time database following the presented proposal will model the time during which an object in a certain state is valid, by associating a \emph{Possibilistic Valid-time Period} to the record describing said object state:
 
 \begin{definition}
-A Possibilistic Valid-Time Period \textbf{PVP} is a possibilistic interval defined by means of two ill-known points, namely $\left[ X,\ Y \right]$
-\begin{equation}
-PVP = \left[X,\ Y \right] 
-\end{equation}
-$X$ and $Y$ are ill-known values in the set of the real numbers $\mathbb{R}$. The uncertainty about the values taken by $X$ and $Y$ are given by the possibility distributions $\pi_X$ and $\pi_Y$.
+A \emph{Possibilistic Valid-time Period} (PVP) is an ill-known interval in time, which models a time period during which an object is valid.
 \end{definition}
 
-The possibility distributions $\pi_X$ and $\pi_Y$ are given in the way of a triangular distribution, as explained in subsection \ref{subsec:fuzzy-numbers}.
+Off course, the object of which the PVP models the Valid-time is the object in the particular state modelled by the record that is associated with the PVP.
+Because a PVP is an ill-known interval, it allows modelling the uncertainty about the start or end point of a time interval (and thus about the time interval itself) if such uncertainty exists.
+The interpretation is \emph{disjunctive}: the PVP represents exactly one Valid-time interval, but precisely \emph{which} interval is represented, is (partially) unknown.
+In the presented model, only PVPs are considered in which the possibility distributions of the possibilistic variables defining the start and end point of the ill-known interval have the exact same characteristics of the membership functions of fuzzy numbers.
+A perfectly known start or end point can then be modelled by such an ill-known value defined by a possibilistic variable $P$ for which $\exists !
+x : \mu_{P}(x) > 0$.
+
+As mentioned in \cite{Pon11}, this approach differs from the one where a Valid-time period is represented by one fuzzy set.
+This one fuzzy set is seen as a possibility distribution on $\mathbb{R}$.
+However, in the presented approach, a time period is modelled using an ill-known set, which is defined by a possibility distribution on $\wp(\mathbb{R})$.
+
+%The interval has a starting and an ending points. An ill-known valid-time interval is an interval in witch one or both points are ill-known.
+
+%\begin{definition}
+%A Possibilistic Valid-Time Period \textbf{PVP} is a possibilistic interval defined by means of two ill-known points, namely $\left[ X,\ Y \right]$
+%\begin{equation}
+%PVP = \left[X,\ Y \right] 
+%\end{equation}
+%$X$ and $Y$ are ill-known values in the set of the real numbers $\mathbb{R}$. The uncertainty about the values taken by $X$ and $Y$ are given by the possibility distributions $\pi_X$ and $\pi_Y$.
+%\end{definition}
+
+%For convenience, the possibility distributions $\pi_X$ and $\pi_Y$ are given in the way of a triangular distribution, as explained in subsection \ref{subsec:fuzzy-numbers}.
 This representation allows overlapping (Fig.
 \ref{fig:pvp}).
+Note that while the two ill-known values $X,Y \in \mathbb{R}$, the fuzzy interval  $[X,Y] \in \mathbb{R}^2$.
 
-\begin{figure}[h!]
-  \centering
-  \includegraphics[scale=0.2]{graphs/2_triangular.eps}
-  \caption{Two fuzzy numbers $X$ and $Y$ denoting a Possibilistic Valid-Time Period \emph{PVT}.}
-  \label{fig:pvp}
-\end{figure}
+%\begin{figure}[h!]
+%  \centering
+%  \input{./graphs/2_triangular.tex}
+%  \caption{Two fuzzy numbers $X$ and $Y$ denoting a Possibilistic Valid-Time Period \emph{PVP}.}
+%  \label{fig:pvp}
+%\end{figure}
 
-\subsection{Storage of valid-time intervals}
-Each database row containing a \emph{PVP} stores it as two triangular possibility distributions.
-In our approach we propose the representation of that as proposed in the  fuzzy interface for relational databases \emph{FIRST}~\cite{Medina94gefred.a,Gal98}.
-In this representation it is also possible to represent not only fuzzy numbers but fuzzy constants (see table \ref{table:relational-representation-pvp}):
+\subsubsection{Storage}
+\label{subsec:storage}
+To store a PVP, the possibility distributions defining the ill-known start and end point are stored.
+To store such a possibility distribution, the representation as used in the fuzzy interface for relational databases \emph{FIRST}~\cite{Medina94gefred.a,Gal98} is used.
+Using this representation, it is possible to represent not only fuzzy numbers, but also (fuzzy) constants.
+To store an ill-known point, four values (FT, F1, F2 and F3) are stored.
+They are explained in table \ref{table:relational-representation-pvp}.
+Note that while NULL denotes the fuzzy constant NULL, N denotes a null value in the database.
 
-\begin{itemize}
-\item
-\emph{NULL}: This constant refers to a completely ignorance about the value.
+%\begin{itemize}
+%\item
+%\emph{NULL}: This constant refers to a completely ignorance about the value.
 The possibility distribution for a given fuzzy number $X$ is not defined, therefore, any comparison between a fuzzy number and the \emph{NULL} constant always returns $0$.
-\item
-\emph{UNKNOWN}: The point has a value but it is unknown. The possibility distribution for a given fuzzy number $X$ is $\pi_X=1$
-\item
-\emph{UNDEFINED}: The point does not have a value. The possibility distribution for a given fuzzy number $X$ is $\pi_X=0$
-\end{itemize}
+%\item
+%\emph{UNKNOWN}: % The possibility distribution for a given fuzzy number $X$ is $\pi_X=1$
+%\item
+%\emph{UNDEFINED}: The point does not have a value. %The possibility distribution for a given fuzzy number $X$ is $\pi_X=0$
+%\end{itemize}
+%
 
 \begin{table}
-\caption{Relational representation for a Possibilistic Valid-Time Period.}
 \centering
-\begin{tabular}{c c c c c c}
+\begin{tabular}{c c c c c c l p{2cm}}
 \hline
-Value & FT & F1 & F2 & F3  \\ \hline
-UNKNOWN & 0 & NULL & NULL & NULL  \\ 
-UNDEFINED & 1 & NULL & NULL & NULL  \\ 
-NULL & 2 & NULL & NULL & NULL  \\ 
-$\left[D,\ a,\ b \right]$ & 3 & $D$ & $D-a$ & $D+b$ \\ 
+Value & FT & F1 & F2 & F3 & $\mu(x)$ & Description \\ \hline
+UNKNOWN & 0 & N & N & N  & $1$ & Any value is equally possible\\ 
+UNDEFINED & 1 & N & N & N & $0$ & The value is not defined in the \\
+          &   &   &   &   &     & attribute domain\\ 
+NULL & 2 & N & N & N &not defined & Nothing is known about the value \\ 
+$\left[D,\ a,\ b \right]$ & 3 & $D$ & $a$ & $b$ & $\mu(\left[D,\ a,\ b \right])$ & Ill-known value \\ 
 \hline 
@@
 \label{table:relational-representation-pvp}
+
+\vspace{10pt}
+
+\caption{Relational representation for an ill-known time point.
+FT denotes Fuzzy Type.
+Field FT indicates that the values stored in F1, F2 and F3 denote either NULL, UNKNOWN, UNDEFINED or a triangular possibility distribution $\left[D,\ a,\ b \right]$.
+Fields F1, F2 and F3 contain the actual values.}%  A \emph{PVP} is represented by two ill-known points.}
 \end{table}
 
-\subsection{Querying ill-known valid-time intervals}
+\vspace{-25pt}
 
-\subsubsection{Query specification}
+\subsection{Querying Ill-known Valid-time Intervals}
+In order to present a complete model, the tools for querying will be provided in this subsection.
+The focus will evidently lie on the querying of Valid-time periods.
+First, the query structure will be defined, followed by the evaluation and ranking methods.
+
+%In this subsection we will define the query specification, then the evaluation of the query and finally the ranking for the query.
+
+%It is important to notice that every Valid-time notion in the database is represented using a PVP. Thus, Valid-time intervals in the database can be partially unknown.
+
+%In the query, the user is allowed to specify both non-temporal preferences and a temporal constraint.
+
+%This allows the user to specify both the preferences and an ill-known valid-time interval in the query.
+It is important to notice that the possibilistic / fuzzy data stored in the database has a \emph{disjunctive interpretation} (it is said that we have \emph{uncertainty}: the valid-time interval has only one value but, for some reason the value is ill-known).
+In the query specification, the user is allowed to express a crisp time interval.
+
+\subsubsection{Query Structure}
+It is important to notice that every Valid-time notion in the database is represented using a PVP.
+Thus, Valid-time intervals in the database can be partially unknown.
+In this framework, a query has two separate parts:
+
+% one is a temporal constraint and the other is the collection of query specifications for regular attributes:
+
+%the first one is the temporal specification. The second part is the query specification for regular attributes.
+
+\begin{definition}
+A query $\tilde Q$ is specified by:
+\begin{equation}
+\label{eq:query-definition}
+\tilde Q = \left( Q^{time}, Q \right)
+\end{equation}
+\end{definition}
+Here, $Q$ denotes the collection of (possibly fuzzy) non-temporal preferences of the user and $Q^{time}$ denotes the temporal constraint given by the user:
+\begin{definition}
+ $Q^{time}$ is defined by:
+\begin{equation}
+Q^{time} = \left( I , AR \right)
+\end{equation}
+\end{definition}
+Here, $I$ is a crisp time interval and $AR$ is one of Allen's relations\cite{Allen83}.
+The interpretation is that for a record with Valid-time period given by a PVP $J$, the user requires that $I$ $AR$ $J$ holds.
+
+%In table \ref{tab:allen-relations}, .
+
+%\vspace{-35pt}
 
 \subsubsection{Query evaluation}
+In fuzzy querying of regular (relational) databases, the modelling of query satisfaction is a matter of degree.
+Usually, the evaluation of the query requirements for a record results in a satisfaction degree $s$, where $s$ lies in $\left[0,1\right]$, where 0 denotes total dissatisfaction and 1 denotes complete satisfaction.
+In crisp querying, the evaluation of query requirements for a record results in the accepting or rejecting of the record as a part of the result set.
+This can be modelled using satisfaction degrees, by assigning rejection a degree of $0$ and acceptance a degree of $1$ and not using any other value in $\left[0,1\right]$.
 
-\subsubsection{Ranking}
+The evaluation of a query $\tilde{Q} = \left( Q^{time}, Q \right)$, where $Q^{time} = \left( I, AR \right)$ is now handled as follows.
+For each record $r$ in the database, with the Valid-time notion of $r$ specified by PVP $J$, two things happen independently:
+
+% by a combination of non-temporal query requirements $Q$ is noted $e_{Q}(r) \in \left[0,1\right]$ in this work. The total query evaluation method is now the following.  and Allen relation $AR$
+\begin{itemize}
+\item
+The preferences expressed in $Q$ are evaluated, resulting in a satisfaction degree denoted here $e_{Q}(r)$.
+The model presented here accepts any way of calculating this evaluation, as long as $e_{Q}(r) \in \left[0,1\right]$.
+\item
+Depending on AR, a specific set of ill-known constraints is considered.
+For each AR, this set can be found in table \ref{tab:allen-relations}.
+The possibility $\Pos_{Q^{time}}(r)$ and necessity $\Nec_{Q^{time}}(r)$ that $r$ fulfills all these constraints are calculated using formulas based on equations \ref{ill-known-pos} and \ref{ill-known-nec}.
+The sum of these possibility and necessity scores gives an evaluation score $e_{Q^{time}}(r)$ in interval $\left[0,2\right]$.
+Because necessity cannot exceed $0$ unless possibility is $1$, this gives a natural score.
+\end{itemize}
+
+\begin{table}[h]
+\label{tab:allen-relations}
+\centering
+\begin{tabular}{|c|c|c|}
+\hline
+Allen Relation & Constraints & $\Pos\left(\text{I satisfies all constraints }C_{i}, i = 1,2,...\right)$ \\
+\hline
+I before J & $C_1 \triangleq \left(<,X\right)$ & $\sup_{a>w}\pi_x(w)$\\
+\hline
+
+\multirow{2}{*}
+{I equal J} & $C_1\triangleq \left(\geq,X\right)$,$C_2\triangleq \left(\neq,X\right)$ & $\min ( \sup_{a \leq w}\pi_x(w),\pi_x(w),$ \\
+ & $C_3\triangleq \left(\leq,Y\right)$,$C_4\triangleq \left(\neq,Y\right)$ & $\sup_{b \geq w}\pi_Y(w),\pi_Y(w))$\\
+\hline
+
+I meets J & $C_1\triangleq \left(\leq,X\right)$ ,$C_2\triangleq \left(\neq,X\right)$ & $\min (\sup_{a\geq w} \pi_X(w),$ $\pi_X(w))$ \\
+\hline
+
+\multirow{2}{*}
+{I overlaps J} & $C_1\triangleq \left(<,Y\right)$,$C_2\triangleq \left(\leq,X\right)$ & $\min ( \sup_{b>w}\pi_Y(w), $ \\
+ & $C_3\triangleq \left(\geq,X\right)$ & $\sup_{a \geq w}\pi_X(w),\sup_{a \leq w}\pi_X(w))$ \\
+\hline
+
+\multirow{2}{*}
+{I during J} & $C_1\triangleq \left(>,X\right)$, $C_2\triangleq \left(\leq,Y\right)$ & $\max ( \min ( \sup_{a<w}\pi_X(w),\sup_{b \geq w}\pi_Y(w)),$ \\
+ & $C_3\triangleq \left(\geq,X\right)$ ,$C_4\triangleq \left(<,Y\right)$ & $\min ( \sup_{a \leq w }\pi_X(w),\sup_{b>w}\pi_Y(w)$\\
+\hline
+
+{I starts J} & $C_1\triangleq \left(\geq,X\right)$,$C_2\triangleq \left(\neq,X\right)$  & $\min( \sup_{a \leq w}\pi_X(w),$ $\pi_X(w))$\\
+\hline
+
+{I finishes J} & $C_1\triangleq \left(\leq,Y\right)$, $C_2\triangleq \left(\neq,Y\right)$  & $\min ( \sup_{b \geq w} \pi_Y(w),$ $\pi_Y(w))$ \\
+\hline
+
+\end{tabular}
+
+\vspace{10pt}
+
+\caption{Allen's relations represented in the framework.
+Here, $I = \left[a, b\right]$ denotes the PVP in the query, $J = \left[X, Y\right]$ denotes the PVP of the record, with $\pi_{X}$ and $\pi_{Y}$ the possibility distributions of $X$ and $Y$ respectively.
+For each of Allen's relations $AR$, the corresponding value in column `Constraints' gives the constraints corresponding to $AR$.
+The last column contains the corresponding formula to calculate the possibility that $I$ satisfies all constraints given in column `Constraints' and thus that $I$ $AR$ $J$ holds.
+These formulas are based on equation \ref{ill-known-pos}.}
+
+\vspace{-25pt}
+
+\end{table}
+
+\subsubsection{Ranking and aggregation}
+In order to present the results to the user, a ranking method is used.
+First of all, $e_{Q^{time}}(r)$ is rescaled to the unit interval.
+The final ranking $e_{final}$ is now given by a convex combination:
+
+%the possibility and the necessity are ranked with the convex combination in \eqref{eq:convex-comb} with $\omega=0.5$.
+Finally this value is aggregated with the rest of the criteria with the same combination.
+%In this case the value for the aggregation is also $\omega = 0.5$
+
+\begin{equation}
+\label{eq:convex-comb}
+e_{final}(r)\ =\ \omega*e_{Q}(r)\ +\ (1-\omega)*e_{Q^{time}}(r), \omega \in \left[0, 1 \right]
+\end{equation}
+
+The use of this convex combination allows a record to make up for a low score on the temporal constraint by a good score on the non-temporal constraint (or vice versa).
+Changing $\omega$ also allows granting the temporal constraint more weight with respect to the non-temporal constraint (or vice versa).
+
+%By increasing the value for the parameter $\omega$, the preferences expressed in the query $Q$ can be given more importance. By lowering the value for $\omega$, the temporal criteria is emphasized.
+%The following example illustrates the querying process.
+
+\subsection{Example}
+
+FROM HERE
+
+%\begin{example} 
+\subsubsection{The database}
+Consider a database describing car models, containing general attributes (model name, manufacturer, car segment) and one temporal attribute containing the approximate period of time during which the car model was sold.
+The temporal data is stored as explained in subsection \ref{subsec:storage}.
+The value for $D$ is stored in \emph{yyyy} format and $a$ and $b$ are represented by an integer, for the convenience of the representation.
+Therefore the chronons in our example will be years.
+The ID field identifies a car model while the field Instance ID (IID) identifies the instance for a car model.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Sample table for the car models. 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{table}[ht]
+\centering
+\begin{tabular}{c c c c c c c}
+\hline
+ID & IID & Segment & Manufacturer & Name & Start & End  \\ [0.5ex]
+\hline
+001 & 1 & B & Peugeot & 205 & [1985,2,3] & [1997,2,1] \\
+002 & 1 & C & Peugeot & 305 & [1977,2,2] & [1989,2,3] \\
+003 & 1 & B & Citroen & C2 & [2002,2,2] & [2006,1,1] \\
+001 & 2 & B & Peugeot & 206 & [2000,1,2] & [2011,2,1] \\
+001 & 3 & B & Peugeot & 207 & [2006,1,1] & [2011,1,1]\\
+\hline
+\end{tabular}
+\vspace{10pt}
+\caption{Sample database}
+\label{tb:car-models}
+\end{table}
+
+\textbf{The query}
+Consider the following query:
+
+\emph{The user wants to obtain a list of models in the segment B for the manufacturer Peugeot before the year interval 2001-2005.}
+
+The query is translated to the following expression, using equation \eqref{eq:query-definition}. The evaluation for the criteria are presented in the result table \ref{tb:results}.
+
+\begin{center}
+$\tilde{Q} = \left(  c^{time}, c_{segment} \right)$
+\end{center}
+
+Where:
+\begin{itemize}
+\item
+$c^{time}\ = \ ( \left[ 2001,\ 2005 \right],\ $  before $)$.
+\item
+$c_{segment}\ = \ $ Peugeot.
+\end{itemize}
+
+\begin{table}[ht]
+\caption{Result table and ranking}
+\centering
+\begin{tabular}{c c c c c c c}
+\hline
+ID & IID &  $\Pos_{c^{time}}$ &$\Nec_{c^{time}}$ & $e(C^{time})$ & $c_{segment}$ & rank ($\omega=0.5$) \\ [0.5ex]
+\hline
+001 & 1 & 1 &  1 & 1 & 1 & 1 \\
+002 & 1 & 1 & 1 & 1 & 0.5 & 0.75 \\
+003 & 1 & 1 & 0.5 & 0.75 &0 & 0.375\\
+001 & 2 & 1 & 0 & 0.5 &1 & 0.75 \\
+001 & 3 & 0 & 0 & 0 &1 & 0.5\\
+\hline
+\end{tabular}
+\label{tb:results}
+\end{table}
+%\end{example}

### 2011-12-12 ipmu2012/proposal.tex
Still not the final version!

--- 2011-11-17 ipmu2012/proposal.tex
@@
-The proposal consist on a possibilistic valid-time model. The representation and the querying are explained in the following subsections.
+In this section, a Valid-time model will be presented.
+The focus will off course be on the handling of imperfect Valid-time notions.
+First, the representation and storage of Valid-time will be presented.
+Next, a way of querying the model is proposed.
+Finally, an example is given.
+%The proposal consist on a possibilistic valid-time model. The representation and the querying are explained in the following subsections.
 
-\subsection{Representation of Ill-known Valid-time Intervals}
-Valid time is usually represented as an interval. The interval has a starting and an ending points. An ill-known valid-time interval is an interval in witch one or both points are ill-known.
+\subsection{Valid-time Intervals}
+\subsubsection{Representation}
+\label{subsec:representation-ill-known}
+A database models a real object by storing the object's value for each attribute describing a property of the object.
+Thus, a Valid-time database following the presented proposal will model the time during which an object in a certain state is valid, by associating a \emph{Possibilistic Valid-time Period} to the record describing said object state:
 
 \begin{definition}
-A Possibilistic Valid-Time Period \textbf{PVP} is a possibilistic interval defined by means of two ill-known points, namely $\left[ X,\ Y \right]$
-\begin{equation}
-PVP = \left[X,\ Y \right] 
-\end{equation}
-$X$ and $Y$ are ill-known values in the set of the real numbers $\mathbb{R}$. The uncertainty about the values taken by $X$ and $Y$ are given by the possibility distributions $\pi_X$ and $\pi_Y$.
+A \emph{Possibilistic Valid-time Period} (PVP) is an ill-known interval in time, which models a time period during which an object is valid.
 \end{definition}
 
-The possibility distributions $\pi_X$ and $\pi_Y$ are given in the way of a triangular distribution, as explained in subsection \ref{subsec:fuzzy-numbers}.
+Off course, the object of which the PVP models the Valid-time is the object in the particular state modelled by the record that is associated with the PVP.
+Because a PVP is an ill-known interval, it allows modelling the uncertainty about the start or end point of a time interval (and thus about the time interval itself) if such uncertainty exists.
+The interpretation is \emph{disjunctive}: the PVP represents exactly one Valid-time interval, but precisely \emph{which} interval is represented, is (partially) unknown.
+In the presented model, only PVPs are considered in which the possibility distributions of the possibilistic variables defining the start and end point of the ill-known interval have the exact same characteristics of the membership functions of fuzzy numbers.
+A perfectly known start or end point can then be modelled by such an ill-known value defined by a possibilistic variable $P$ for which $\exists !
+x : \mu_{P}(x) > 0$.
+
+As mentioned in \cite{Pon11}, this approach differs from the one where a Valid-time period is represented by one fuzzy set.
+This one fuzzy set is seen as a possibility distribution on $\mathbb{R}$.
+However, in the presented approach, a time period is modelled using an ill-known set, which is defined by a possibility distribution on $\wp(\mathbb{R})$.
+
+%The interval has a starting and an ending points. An ill-known valid-time interval is an interval in witch one or both points are ill-known.
+
+%\begin{definition}
+%A Possibilistic Valid-Time Period \textbf{PVP} is a possibilistic interval defined by means of two ill-known points, namely $\left[ X,\ Y \right]$
+%\begin{equation}
+%PVP = \left[X,\ Y \right] 
+%\end{equation}
+%$X$ and $Y$ are ill-known values in the set of the real numbers $\mathbb{R}$. The uncertainty about the values taken by $X$ and $Y$ are given by the possibility distributions $\pi_X$ and $\pi_Y$.
+%\end{definition}
+
+%For convenience, the possibility distributions $\pi_X$ and $\pi_Y$ are given in the way of a triangular distribution, as explained in subsection \ref{subsec:fuzzy-numbers}.
 This representation allows overlapping (Fig.
 \ref{fig:pvp}).
+Note that while the two ill-known values $X,Y \in \mathbb{R}$, the fuzzy interval  $[X,Y] \in \mathbb{R}^2$.
 
-\begin{figure}[h!]
-  \centering
-%  \includegraphics[scale=0.2]{graphs/2_triangular.eps}
-  \caption{Two fuzzy numbers $X$ and $Y$ denoting a Possibilistic Valid-Time Period \emph{PVT}.}
-  \label{fig:pvp}
-\end{figure}
+%\begin{figure}[h!]
+%  \centering
+%  \input{./graphs/2_triangular.tex}
+%  \caption{Two fuzzy numbers $X$ and $Y$ denoting a Possibilistic Valid-Time Period \emph{PVP}.}
+%  \label{fig:pvp}
+%\end{figure}
 
-\subsection{Storage of Valid-time Intervals}
-Each database row containing a \emph{PVP} stores it as two triangular possibility distributions.
-In our approach we propose the representation of that as proposed in the  fuzzy interface for relational databases \emph{FIRST}~\cite{Medina94gefred.a,Gal98}.
-In this representation it is also possible to represent not only fuzzy numbers but fuzzy constants (see table \ref{table:relational-representation-pvp}):
+\subsubsection{Storage}
+\label{subsec:storage}
+To store a PVP, the possibility distributions defining the ill-known start and end point are stored.
+To store such a possibility distribution, the representation as used in the fuzzy interface for relational databases \emph{FIRST}~\cite{Medina94gefred.a,Gal98} is used.
+Using this representation, it is possible to represent not only fuzzy numbers, but also (fuzzy) constants.
+To store an ill-known point, four values (FT, F1, F2 and F3) are stored.
+They are explained in table \ref{table:relational-representation-pvp}.
+Note that while NULL denotes the fuzzy constant NULL, N denotes a null value in the database.
 
-\begin{itemize}
-\item
-\emph{NULL}: This constant refers to a completely ignorance about the value.
+%\begin{itemize}
+%\item
+%\emph{NULL}: This constant refers to a completely ignorance about the value.
 The possibility distribution for a given fuzzy number $X$ is not defined, therefore, any comparison between a fuzzy number and the \emph{NULL} constant always returns $0$.
-\item
-\emph{UNKNOWN}: The point has a value but it is unknown. The possibility distribution for a given fuzzy number $X$ is $\pi_X=1$
-\item
-\emph{UNDEFINED}: The point does not have a value. The possibility distribution for a given fuzzy number $X$ is $\pi_X=0$
-\end{itemize}
+%\item
+%\emph{UNKNOWN}: % The possibility distribution for a given fuzzy number $X$ is $\pi_X=1$
+%\item
+%\emph{UNDEFINED}: The point does not have a value. %The possibility distribution for a given fuzzy number $X$ is $\pi_X=0$
+%\end{itemize}
+%
+
+\vspace{-10pt}
 
 \begin{table}
-\caption{Relational representation for a Possibilistic Valid-Time Period.}
 \centering
-\begin{tabular}{c c c c c c}
+\begin{tabular}{c c c c c c l p{2cm}}
 \hline
-Value & FT & F1 & F2 & F3  \\ \hline
-UNKNOWN & 0 & NULL & NULL & NULL  \\ 
-UNDEFINED & 1 & NULL & NULL & NULL  \\ 
-NULL & 2 & NULL & NULL & NULL  \\ 
-$\left[D,\ a,\ b \right]$ & 3 & $D$ & $D-a$ & $D+b$ \\ 
+Value & FT & F1 & F2 & F3 & $\mu(x)$ & Description \\ \hline
+UNKNOWN & 0 & N & N & N  & $1$ & Any value is equally possible\\ 
+UNDEFINED & 1 & N & N & N & $0$ & The value is not defined in the \\
+          &   &   &   &   &     & attribute domain\\ 
+NULL & 2 & N & N & N &not defined & Nothing is known about the value \\ 
+$\left[D,\ a,\ b \right]$ & 3 & $D$ & $a$ & $b$ & $\mu(\left[D,\ a,\ b \right])$ & Ill-known value \\ 
 \hline 
@@
 \label{table:relational-representation-pvp}
+
+\vspace{10pt}
+
+\caption{Relational representation for an ill-known time point.
+FT denotes Fuzzy Type.
+Field FT indicates that the values stored in F1, F2 and F3 denote either NULL, UNKNOWN, UNDEFINED or a triangular possibility distribution $\left[D,\ a,\ b \right]$.
+Fields F1, F2 and F3 contain the actual values.}%  A \emph{PVP} is represented by two ill-known points.}
 \end{table}
 
+\vspace{-25pt}
+
 \subsection{Querying Ill-known Valid-time Intervals}
-In order to provide a complete model, we will provide the tools for querying.
-This allows the user to specify both the preferences and an ill-known valid-time interval in the query.
+In order to present a complete model, the tools for querying will be provided in this subsection.
+The focus will evidently lie on the querying of Valid-time periods.
+First, the query structure will be defined, followed by the evaluation and ranking methods.
+
+%In this subsection we will define the query specification, then the evaluation of the query and finally the ranking for the query.
+
+%It is important to notice that every Valid-time notion in the database is represented using a PVP. Thus, Valid-time intervals in the database can be partially unknown.
+
+%In the query, the user is allowed to specify both non-temporal preferences and a temporal constraint.
+
+%This allows the user to specify both the preferences and an ill-known valid-time interval in the query.
 It is important to notice that the possibilistic / fuzzy data stored in the database has a \emph{disjunctive interpretation} (it is said that we have \emph{uncertainty}: the valid-time interval has only one value but, for some reason the value is ill-known).
-In the query specification, the user is allowed to express \emph{imprecision} or \emph{vagueness} (this is referred as a \emph{conjunctive interpretation}).
-In this subsection we will define the query specification, then the evaluation of the query and finally the ranking for the query.
+In the query specification, the user is allowed to express a crisp time interval.
 
-\subsubsection{Query specification}
-A query in our framework has two different parts: the first one is the query specification for regular attributes. The second part is the temporal specification.
+\subsubsection{Query Structure}
+It is important to notice that every Valid-time notion in the database is represented using a PVP.
+Thus, Valid-time intervals in the database can be partially unknown.
+In this framework, a query has two separate parts:
+
+% one is a temporal constraint and the other is the collection of query specifications for regular attributes:
+
+%the first one is the temporal specification. The second part is the query specification for regular attributes.
 
@@
 \end{definition}
-Where $Q$ are the (possibly fuzzy) preferences of the user  and $Q^{time}$ is the temporal part specified by a possibilistic valid-time period,\emph{PVP}.
+Here, $Q$ denotes the collection of (possibly fuzzy) non-temporal preferences of the user and $Q^{time}$ denotes the temporal constraint given by the user:
+\begin{definition}
+ $Q^{time}$ is defined by:
+\begin{equation}
+Q^{time} = \left( I , AR \right)
+\end{equation}
+\end{definition}
+Here, $I$ is a crisp time interval and $AR$ is one of Allen's relations\cite{Allen83}.
+The interpretation is that for a record with Valid-time period given by a PVP $J$, the user requires that $I$ $AR$ $J$ holds.
 
-%explain the sematics of vt.
+%In table \ref{tab:allen-relations}, .
+
+%\vspace{-35pt}
 
 \subsubsection{Query evaluation}
-Fuzzy querying of regular (relational) databases, query satisfaction modelling is a matter of degree.
-Usually, criteria satisfaction is modelled by means of a satisfaction degree $s \in \left[ 0, 1\right]$.
-In the model, every record $r$ contains a \emph{PVP} $V_r$ to model the valid-time for that record.
+In fuzzy querying of regular (relational) databases, the modelling of query satisfaction is a matter of degree.
+Usually, the evaluation of the query requirements for a record results in a satisfaction degree $s$, where $s$ lies in $\left[0,1\right]$, where 0 denotes total dissatisfaction and 1 denotes complete satisfaction.
+In crisp querying, the evaluation of query requirements for a record results in the accepting or rejecting of the record as a part of the result set.
+This can be modelled using satisfaction degrees, by assigning rejection a degree of $0$ and acceptance a degree of $1$ and not using any other value in $\left[0,1\right]$.
 
-The query evaluation method is the following:
+The evaluation of a query $\tilde{Q} = \left( Q^{time}, Q \right)$, where $Q^{time} = \left( I, AR \right)$ is now handled as follows.
+For each record $r$ in the database, with the Valid-time notion of $r$ specified by PVP $J$, two things happen independently:
+
+% by a combination of non-temporal query requirements $Q$ is noted $e_{Q}(r) \in \left[0,1\right]$ in this work. The total query evaluation method is now the following.  and Allen relation $AR$
 \begin{itemize}
 \item
-For each record $r$ in the database, the preferences expressed in $Q$ are evaluated in the unit interval.
+The preferences expressed in $Q$ are evaluated, resulting in a satisfaction degree denoted here $e_{Q}(r)$.
+The model presented here accepts any way of calculating this evaluation, as long as $e_{Q}(r) \in \left[0,1\right]$.
+\item
+Depending on AR, a specific set of ill-known constraints is considered.
+For each AR, this set can be found in table \ref{tab:allen-relations}.
+The possibility $\Pos_{Q^{time}}(r)$ and necessity $\Nec_{Q^{time}}(r)$ that $r$ fulfills all these constraints are calculated using formulas based on equations \ref{ill-known-pos} and \ref{ill-known-nec}.
+The sum of these possibility and necessity scores gives an evaluation score $e_{Q^{time}}(r)$ in interval $\left[0,2\right]$.
+Because necessity cannot exceed $0$ unless possibility is $1$, this gives a natural score.
 \end{itemize}
 
-\subsubsection{Ranking}
+\begin{table}[h]
+\label{tab:allen-relations}
+\centering
+\begin{tabular}{|c|c|c|}
+\hline
+Allen Relation & Constraints & $\Pos\left(\text{I satisfies all constraints }C_{i}, i = 1,2,...\right)$ \\
+\hline
+I before J & $C_1 \triangleq \left(<,X\right)$ & $\sup_{a>w}\pi_x(w)$\\
+\hline
+
+\multirow{2}{*}
+{I equal J} & $C_1\triangleq \left(\geq,X\right)$,$C_2\triangleq \left(\neq,X\right)$ & $\min ( \sup_{a \leq w}\pi_x(w),\pi_x(w),$ \\
+ & $C_3\triangleq \left(\leq,Y\right)$,$C_4\triangleq \left(\neq,Y\right)$ & $\sup_{b \geq w}\pi_Y(w),\pi_Y(w))$\\
+\hline
+
+I meets J & $C_1\triangleq \left(\leq,X\right)$ ,$C_2\triangleq \left(\neq,X\right)$ & $\min (\sup_{a\geq w} \pi_X(w),$ $\pi_X(w))$ \\
+\hline
+
+\multirow{2}{*}
+{I overlaps J} & $C_1\triangleq \left(<,Y\right)$,$C_2\triangleq \left(\leq,X\right)$ & $\min ( \sup_{b>w}\pi_Y(w), $ \\
+ & $C_3\triangleq \left(\geq,X\right)$ & $\sup_{a \geq w}\pi_X(w),\sup_{a \leq w}\pi_X(w))$ \\
+\hline
+
+\multirow{2}{*}
+{I during J} & $C_1\triangleq \left(>,X\right)$, $C_2\triangleq \left(\leq,Y\right)$ & $\max ( \min ( \sup_{a<w}\pi_X(w),\sup_{b \geq w}\pi_Y(w)),$ \\
+ & $C_3\triangleq \left(\geq,X\right)$ ,$C_4\triangleq \left(<,Y\right)$ & $\min ( \sup_{a \leq w }\pi_X(w),\sup_{b>w}\pi_Y(w)$\\
+\hline
+
+{I starts J} & $C_1\triangleq \left(\geq,X\right)$,$C_2\triangleq \left(\neq,X\right)$  & $\min( \sup_{a \leq w}\pi_X(w),$ $\pi_X(w))$\\
+\hline
+
+{I finishes J} & $C_1\triangleq \left(\leq,Y\right)$, $C_2\triangleq \left(\neq,Y\right)$  & $\min ( \sup_{b \geq w} \pi_Y(w),$ $\pi_Y(w))$ \\
+\hline
+
+\end{tabular}
+
+\vspace{10pt}
+
+\caption{Allen's relations represented in the framework.
+Here, $I = \left[a, b\right]$ denotes the PVP in the query, $J = \left[X, Y\right]$ denotes the PVP of the record, with $\pi_{X}$ and $\pi_{Y}$ the possibility distributions of $X$ and $Y$ respectively.
+For each of Allen's relations $AR$, the corresponding value in column `Constraints' gives the constraints corresponding to $AR$.
+The last column contains the corresponding formula to calculate the possibility that $I$ satisfies all constraints given in column `Constraints' and thus that $I$ $AR$ $J$ holds.
+These formulas are based on equation \ref{ill-known-pos}.}
+
+\vspace{-25pt}
+
+\end{table}
+
+\vspace{-5pt}
+
+\subsubsection{Ranking and aggregation}
+In order to present the results to the user, a ranking method is used.
+First of all, $e_{Q^{time}}(r)$ is rescaled to the unit interval.
+The final ranking $e_{final}$ is now given by a convex combination:
+
+%the possibility and the necessity are ranked with the convex combination in \eqref{eq:convex-comb} with $\omega=0.5$.
+Finally this value is aggregated with the rest of the criteria with the same combination.
+%In this case the value for the aggregation is also $\omega = 0.5$
+
+\begin{equation}
+\label{eq:convex-comb}
+e_{final}(r)\ =\ \omega*e_{Q}(r)\ +\ (1-\omega)*e_{Q^{time}}(r), \omega \in \left[0, 1 \right]
+\end{equation}
+
+The use of this convex combination allows a record to make up for a low score on the temporal constraint by a good score on the non-temporal constraint (or vice versa).
+Changing $\omega$ also allows granting the temporal constraint more weight with respect to the non-temporal constraint (or vice versa).
+
+%By increasing the value for the parameter $\omega$, the preferences expressed in the query $Q$ can be given more importance. By lowering the value for $\omega$, the temporal criteria is emphasized.
+%The following example illustrates the querying process.
+
+\subsection{Example}
+
+%\begin{example} 
+\subsubsection{The database}
+Consider example database \ref{tb:car-models} describing car models, containing general attributes (model name, manufacturer, car segment) and one temporal attribute containing the approximate period of time during which the car model was sold.
+The temporal data is stored as explained in subsection \ref{subsec:storage}.
+Let the chronons in our example be years.
+The value for $D$ is stored in \emph{yyyy} format and $a$ and $b$ are represented by an integer.
+The ID field identifies a car model while the field Instance ID (IID) identifies the instance for a car model, thus a car model in a certain state.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Sample table for the car models. 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{table}[h]
+\centering
+\begin{tabular}{c c c c c c c}
+\hline
+ID & IID & Segment & Manufacturer & Name & Start & End  \\ [0.5ex]
+\hline
+001 & 1 & B & Peugeot & 205 & [1985,2,3] & [1997,2,1] \\
+002 & 1 & C & Peugeot & 305 & [1977,2,2] & [1989,2,3] \\
+003 & 1 & B & Citroen & C2 & [2002,2,2] & [2006,1,1] \\
+001 & 2 & B & Peugeot & 206 & [2000,1,2] & [2011,2,1] \\
+001 & 3 & B & Peugeot & 207 & [2006,1,1] & [2011,1,1]\\
+\hline
+\end{tabular}
+\vspace{10pt}
+\caption{Example database}
+\label{tb:car-models}
+
+\vspace{-30pt}
+
+\end{table}
+
+\subsubsection{The query}
+Consider the following query:
+
+\begin{center}
+\emph{The user wants to obtain a list of models from segment B, made by manufacturer Peugeot before the year interval 2001-2005}
+\end{center}
+
+Using the introduced notations, the query is translated to the following expression:
+
+\vspace{-10pt}
+
+\begin{center}
+$\tilde{Q} = \left(Q^{time}, Q\right)$
+\end{center}
+
+\vspace{-10pt}
+
+In this expression:
+%\begin{itemize}
+%\item
+%$Q^{time}\ = \ ( \left[ 2001,\ 2005 \right],\ $  before $)$.
+%\item
+%$c_{segment}\ = \ $ Peugeot.
+%\end{itemize}
+
+\vspace{-15pt}
+
+\begin{align}
+Q^{time} & = \left(\left[ 2001, 2005 \right], before\right) \\
+Q & = \left(Segment =  B\right) \wedge \left(Manufacturer = Peugeot\right)
+\end{align}
+
+The evaluation for the criteria are presented in the result table \ref{tb:results}.
+
+\begin{table}[ht]
+\caption{Result table and ranking}
+\centering
+\begin{tabular}{c c c c c c c}
+\hline
+ID & IID &  $\Pos_{c^{time}}$ &$\Nec_{c^{time}}$ & $e(C^{time})$ & $c_{segment}$ & rank ($\omega=0.5$) \\ [0.5ex]
+\hline
+001 & 1 & 1 &  1 & 1 & 1 & 1 \\
+002 & 1 & 1 & 1 & 1 & 0.5 & 0.75 \\
+003 & 1 & 1 & 0.5 & 0.75 &0 & 0.375\\
+001 & 2 & 1 & 0 & 0.5 &1 & 0.75 \\
+001 & 3 & 0 & 0 & 0 &1 & 0.5\\
+\hline
+\end{tabular}
+\label{tb:results}
+\end{table}
+%\end{example}

### 2011-12-12 trunk/ipmu2012/preliminaries.tex
Still not the final version!

@@
 
-A simple form of the membership function of a fuzzy interval is a trapezoidal function.
+%A simple form of the membership function of a fuzzy interval is a trapezoidal function.
 It can be shown that such a membership function $\mu_T$ for a fuzzy interval $T$ is convex and normalized.
@@
 It can be shown that such a membership function $\mu_M$ for a fuzzy number $M$ is convex and normalized.
-Three reel values, denoted $a$, $b$ and $D$, suffice to represent a triangular membership function of a fuzzy number and in this work, a fuzzy number defined as such will be noted as $\left[D, a, b \right]$.
+Three real values, denoted $a$, $b$ and $D$, suffice to represent a triangular membership function of a fuzzy number and in this work, a fuzzy number defined as such will be noted as $\left[D, a, b \right]$.
 Here:
@@
 \end{itemize}
+
+The membership function of such a fuzzy number is then given by:
+
+\vspace{-10pt}
+
+\begin{align}
+\mu_M : & \quad \mathbb{R} \rightarrow \left[0,1\right] \nonumber\\
+ : & \quad x \rightarrow
+\begin{cases}
+0 & \mbox{ if } \left(x < D-a\right) \vee \left(x > D+b\right) \\
+\frac{1}{a}\left[x-\left(D-a\right)\right] & \mbox{ if } \left(x \geq D-a\right) \wedge (x \leq D)  \\
+-\frac{1}{b}\left[x-\left(D+b\right)\right] & \mbox{ if } \left(x \geq D\right) \wedge (x \leq D+b)
+\end{cases}
+\end{align}
+
 %\begin{figure}[h!]
@@
 
+\vspace{-10pt}
+
 \begin{align}
-\Pos(A\text{ satisfies }C) & = \min_{a \in A}\left(\sup_{(a,w) \in R}\pi_{V}(w))\right) \\
-\Nec(A\text{ satisfies }C) & = \min_{a \in A}\left(\inf_{(a,w) \notin R} 1-\pi_{V}(w)\right)
+\Pos(A\text{ satisfies }C) & = \min_{a \in A}\left(\sup_{(a,w) \in R}\pi_{V}(w))\right) \label{ill-known-pos}\\
+\Nec(A\text{ satisfies }C) & = \min_{a \in A}\left(\inf_{(a,w) \notin R} 1-\pi_{V}(w)\right) \label{ill-known-nec}
 \end{align}
@@
 
+\vspace{-10pt}
+
 \begin{eqnarray}
@@
 
-The possibility and necessity of the set satisfying both constraints is then:
+To calculate possibility and necessity concerning a conjunction of constraints, the $\min$ operator is used. The possibility and necessity of the set satisfying both constraints is then:
+
+\vspace{-10pt}
 

### 2011-12-13 trunk/ipmu2012/conclusions.tex
Normally, only details left

--- 2011-11-30 ipmu2012/conclusions.tex
@@
-We presented a valid-time model to represent and querying ill-known temporal intervals.
-The main advantage of the framework is that it is always possible to get both a possibility and a necessity measures for every comparison, which is also useful for ranking purposes.
+We presented a Valid-time model to represent and query ill-known temporal intervals.
+The main advantage of this framework is that it is always possible to get both a possibility and a necessity measures for every comparison, which is also useful for ranking purposes.
 The framework models the Allen's relations and it has the flexibility to model specific and more complex relations by means of ill-known constraints.

### 2011-12-13 trunk/ipmu2012/proposal.tex
Normally, only details left

@@
 %
+
+\vspace{-10pt}
 
@@
 
+\vspace{-5pt}
+
 \subsubsection{Ranking and aggregation}
@@
 
-FROM HERE
-
 %\begin{example} 
 \subsubsection{The database}
-Consider a database describing car models, containing general attributes (model name, manufacturer, car segment) and one temporal attribute containing the approximate period of time during which the car model was sold.
+Consider example database \ref{tb:car-models} describing car models, containing general attributes (model name, manufacturer, car segment) and one temporal attribute containing the approximate period of time during which the car model was sold.
 The temporal data is stored as explained in subsection \ref{subsec:storage}.
-The value for $D$ is stored in \emph{yyyy} format and $a$ and $b$ are represented by an integer, for the convenience of the representation.
-Therefore the chronons in our example will be years.
-The ID field identifies a car model while the field Instance ID (IID) identifies the instance for a car model.
+Let the chronons in our example be years.
+The value for $D$ is stored in \emph{yyyy} format and $a$ and $b$ are represented by an integer.
+The ID field identifies a car model while the field Instance ID (IID) identifies the instance for a car model, thus a car model in a certain state.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{table}[ht]
+\begin{table}[h]
 \centering
@@
 \vspace{10pt}
-\caption{Sample database}
+\caption{Example database}
 \label{tb:car-models}
+
+\vspace{-30pt}
+
 \end{table}
 
-\textbf{The query}
+\subsubsection{The query}
 Consider the following query:
 
-\emph{The user wants to obtain a list of models in the segment B for the manufacturer Peugeot before the year interval 2001-2005.}
+\begin{center}
+\emph{The user wants to obtain a list of models from segment B, made by manufacturer Peugeot before the year interval 2001-2005}
+\end{center}
 
-The query is translated to the following expression, using equation \eqref{eq:query-definition}. The evaluation for the criteria are presented in the result table \ref{tb:results}.
+Using the introduced notations, the query is translated to the following expression:
+
+\vspace{-10pt}
 
 \begin{center}
-$\tilde{Q} = \left(  c^{time}, c_{segment} \right)$
+$\tilde{Q} = \left(Q^{time}, Q\right)$
 \end{center}
 
-Where:
-\begin{itemize}
-\item
-$c^{time}\ = \ ( \left[ 2001,\ 2005 \right],\ $  before $)$.
-\item
-$c_{segment}\ = \ $ Peugeot.
-\end{itemize}
+\vspace{-10pt}
+
+In this expression:
+%\begin{itemize}
+%\item
+%$Q^{time}\ = \ ( \left[ 2001,\ 2005 \right],\ $  before $)$.
+%\item
+%$c_{segment}\ = \ $ Peugeot.
+%\end{itemize}
+
+\vspace{-15pt}
+
+\begin{align}
+Q^{time} & = \left(\left[ 2001, 2005 \right], before\right) \\
+Q & = \left(Segment =  B\right) \wedge \left(Manufacturer = Peugeot\right)
+\end{align}
+
+The evaluation for the criteria are presented in the result table \ref{tb:results}.
 
@@
 \hline
-ID & IID &  $\Pos_{c^{time}}$ &$\Nec_{c^{time}}$ & $e(C^{time})$ & $c_{segment}$ & rank ($\omega=0.5$) \\ [0.5ex]
+ID & IID &  $\Pos_{Q^{time}}$ & $\Nec_{Q^{time}}$ & $e_{Q^{time}} (rescaled)$ & $Q$ & $e_{final}$ ($\omega=0.5$) \\ [0.5ex]
 \hline

### 2011-12-13 ipmu2012/conclusions.tex
Also the conclusions...

@@
-We presented a valid-time model to represent and querying ill-known temporal intervals.
-The main advantage of the framework is that it is always possible to get both a possibility and a necessity measures for every comparison, which is also useful for ranking purposes.
-The framework models the Allen's relations and it has the flexibility to model specific and more complex relations by means of ill-known constraints.
-As future work, the time interval in the query specification would also be ill-known.
+We presented a Valid-time model to represent and query ill-known temporal intervals.
+The main advantage of this framework is that it works with ill-known intervals and thus tries to correctly represents and handle querying temporal uncertainty in the database.
+The final ranking step provides a usefull result record set.
+The framework models the Allen relations and it has the flexibility to model specific and more complex relations by means of ill-known constraints.
+As future work, allowing the time interval in the query specification to be ill-known is considered.

### 2011-12-15 trunk/ipmu2012/proposal.tex
Last update

@@
-In this section, a Valid-time model will be presented.
-The focus will off course be on the handling of imperfect Valid-time notions.
-First, the representation and storage of Valid-time will be presented.
+In this section, a valid-time model will be presented.
+The focus will off course be on the handling of imperfect valid-time notions.
+First, the representation and storage of valid-time will be presented.
 Next, a way of querying the model is proposed.
@@
 A database models a real object by storing the object's value for each attribute describing a property of the object.
-Thus, a Valid-time database following the presented proposal will model the time during which an object in a certain state is valid, by associating a \emph{Possibilistic Valid-time Period} to the record describing said object state:
+Thus, a valid-time database following the presented proposal will model the time during which an object in a certain state is valid, by associating a \emph{Possibilistic valid-time Period} to the record describing said object state:
 
 \begin{definition}
-A \emph{Possibilistic Valid-time Period} (PVP) is an ill-known interval in time, which models a time period during which an object is valid.
+A \emph{Possibilistic valid-time Period} (PVP) is an ill-known interval in time, which models a time period during which an object is valid.
 \end{definition}
 
-Off course, the object of which the PVP models the Valid-time is the object in the particular state modelled by the record that is associated with the PVP.
+Off course, the object of which the PVP models the valid-time is the object in the particular state modelled by the record that is associated with the PVP.
 Because a PVP is an ill-known interval, it allows modelling the uncertainty about the start or end point of a time interval (and thus about the time interval itself) if such uncertainty exists.
-The interpretation is \emph{disjunctive}: the PVP represents exactly one Valid-time interval, but precisely \emph{which} interval is represented, is (partially) unknown.
+The interpretation is \emph{disjunctive}: the PVP represents exactly one valid-time interval, but precisely \emph{which} interval is represented, is (partially) unknown.
 In the presented model, only PVPs are considered in which the possibility distributions of the possibilistic variables defining the start and end point of the ill-known interval have the exact same characteristics of the membership functions of fuzzy numbers.
@@
 
-As mentioned in \cite{Pon11}, this approach differs from the one where a Valid-time period is represented by one fuzzy set.
+As mentioned in \cite{Pon11}, this approach differs from the one where a valid-time period is represented by one fuzzy set.
 This one fuzzy set is seen as a possibility distribution on $\mathbb{R}$.
@@
 \centering
+\caption{Relational representation for an ill-known time point.
+FT denotes Fuzzy Type.
+Field FT indicates that the values stored in F1, F2 and F3 denote either NULL, UNKNOWN, UNDEFINED or a triangular possibility distribution $\left[D,\ a,\ b \right]$.
+Fields F1, F2 and F3 contain the actual values.}%  A \emph{PVP} is represented by two ill-known points.}
 \begin{tabular}{c c c c c c l p{2cm}}
@@
 UNKNOWN & 0 & N & N & N  & $1$ & Any value is equally possible\\ 
-UNDEFINED & 1 & N & N & N & $0$ & The value is not defined in the \\
-          &   &   &   &   &     & attribute domain\\ 
+UNDEFINED & 1 & N & N & N & $0$ & The value is not defined \\%in the \\
+    %      &   &   &   &   &     & attribute domain\\ 
 NULL & 2 & N & N & N &not defined & Nothing is known about the value \\ 
@@
 
-\caption{Relational representation for an ill-known time point.
-FT denotes Fuzzy Type.
-Field FT indicates that the values stored in F1, F2 and F3 denote either NULL, UNKNOWN, UNDEFINED or a triangular possibility distribution $\left[D,\ a,\ b \right]$.
-Fields F1, F2 and F3 contain the actual values.}%  A \emph{PVP} is represented by two ill-known points.}
 \end{table}
@@
 \subsection{Querying Ill-known Valid-time Intervals}
-In order to present a complete model, the tools for querying will be provided in this subsection.
-The focus will evidently lie on the querying of Valid-time periods.
+This subsection provides the tools for querying.
+The focus will evidently lie on the querying of valid-time periods.
 First, the query structure will be defined, followed by the evaluation and ranking methods.
@@
 \subsubsection{Query Structure}
-It is important to notice that every Valid-time notion in the database is represented using a PVP.
-Thus, Valid-time intervals in the database can be partially unknown.
+It is important to notice that every valid-time notion in the database is represented using a PVP.
+Thus, valid-time intervals in the database can be partially unknown.
 In this framework, a query has two separate parts:
@@
 Here, $I$ is a crisp time interval and $AR$ is one of Allen's relations\cite{Allen83}.
-The interpretation is that for a record with Valid-time period given by a PVP $J$, the user requires that $I$ $AR$ $J$ holds.
+The interpretation is that for a record with valid-time period given by a PVP $J$, the user requires that $I$ $AR$ $J$ holds.
 
@@
 The evaluation of a query $\tilde{Q} = \left( Q^{time}, Q \right)$, where $Q^{time} = \left( I, AR \right)$ is now handled as follows.
-For each record $r$ in the database, with the Valid-time notion of $r$ specified by PVP $J$, two things happen independently:
+For each record $r$ in the database, with the valid-time notion of $r$ specified by PVP $J$, two things happen independently:
 
@@
 For each AR, this set can be found in table \ref{tab:allen-relations}.
-The possibility $\Pos_{Q^{time}}(r)$ and necessity $\Nec_{Q^{time}}(r)$ that $r$ fulfills all these constraints are calculated using formulas based on equations \ref{ill-known-pos} and \ref{ill-known-nec}.
+The possibility $\Pos_{Q^{time}}(r)$ and necessity $\Nec_{Q^{time}}(r)$ that $r$ fulfills all these constraints are calculated using formulas based on equations \eqref{ill-known-pos} and \eqref{ill-known-nec}.
 The sum of these possibility and necessity scores gives an evaluation score $e_{Q^{time}}(r)$ in interval $\left[0,2\right]$.
@@
 \label{tab:allen-relations}
+\caption{Allen's relations represented in the framework.
+Here, $I = \left[a, b\right]$ denotes the PVP in the query, $J = \left[X, Y\right]$ denotes the PVP of the record, with $\pi_{X}$ and $\pi_{Y}$ the possibility distributions of $X$ and $Y$ respectively.
+For each of Allen's relations $AR$, the corresponding value in column `Constraints' gives the constraints corresponding to $AR$.
+The last column contains the corresponding formula to calculate the possibility that $I$ satisfies all constraints given in column `Constraints' and thus that $I$ $AR$ $J$ holds.
+These formulas are based on equation \eqref{ill-known-pos}.}
+
 \centering
@@
 
-\caption{Allen's relations represented in the framework.
-Here, $I = \left[a, b\right]$ denotes the PVP in the query, $J = \left[X, Y\right]$ denotes the PVP of the record, with $\pi_{X}$ and $\pi_{Y}$ the possibility distributions of $X$ and $Y$ respectively.
-For each of Allen's relations $AR$, the corresponding value in column `Constraints' gives the constraints corresponding to $AR$.
-The last column contains the corresponding formula to calculate the possibility that $I$ satisfies all constraints given in column `Constraints' and thus that $I$ $AR$ $J$ holds.
-These formulas are based on equation \ref{ill-known-pos}.}
-
 \vspace{-25pt}
@@
 \subsubsection{The database}
-Consider example database \ref{tb:car-models} describing car models, containing general attributes (model name, manufacturer, car segment) and one temporal attribute containing the approximate period of time during which the car model was sold.
-The temporal data is stored as explained in subsection \ref{subsec:storage}.
-Let the chronons in our example be years.
+Consider example database \ref{tb:car-models} describing car models, containing general attributes (model name, manufacturer, car segment) and one temporal attribute (stored as explained in subsection \ref{subsec:storage}) containing the approximate period of time  in years during which the car model was sold.
 The value for $D$ is stored in \emph{yyyy} format and $a$ and $b$ are represented by an integer.
@@
 \centering
+\caption{Example database}
 \begin{tabular}{c c c c c c c}
@@
 \vspace{10pt}
-\caption{Example database}
+
 \label{tb:car-models}
@@
 Consider the following query:
-
 \begin{center}
@@
 
-Using the introduced notations, the query is translated to the following expression:
+Using the introduced notations in \eqref{eq:query-definition}, the query is translated to: %the following expression:
 
-\vspace{-10pt}
-
-\begin{center}
-$\tilde{Q} = \left(Q^{time}, Q\right)$
-\end{center}
-
-\vspace{-10pt}
-
-In this expression:
+%\vspace{-10pt}
+%
+%\begin{align}
+%$\tilde{Q} = \left(Q^{time}, Q\right)$
+%\end{center}
+%
+%\vspace{-10pt}
+%
+%In this expression:
 %\begin{itemize}
@@
 
-\vspace{-15pt}
+%\vspace{-15pt}
 
@@
 
-The evaluation for the criteria are presented in the result table \ref{tb:results}.
+%The evaluation for the criteria are presented in the result table \ref{tb:results}.
 
@@
 %\end{example}
+
+Table \ref{tb:results} shows a natural and gradual ranking for the results.
+The last record, (ID 001, IID 3) shows also that with omega=0.5 both temporal and regular criteria have the same importance.% By lowering the value for omega the temporal criteria would have more impact on the final ranking.

### 2011-12-15 ipmu2012/conclusions.tex
Last update

--- 2011-11-30 ipmu2012/conclusions.tex
@@
-We presented a valid-time model to represent and querying ill-known temporal intervals.
-The main advantage of the framework is that it is always possible to get both a possibility and a necessity measures for every comparison, which is also useful for ranking purposes.
-The framework models the Allen's relations and it has the flexibility to model specific and more complex relations by means of ill-known constraints.
-As future work, the time interval in the query specification would also be ill-known.
+We presented a valid-time model to represent and query ill-known temporal intervals.
+The main advantage of this framework is that it works with ill-known intervals and thus tries to correctly represents and handle querying temporal uncertainty in the database by providing both possibility and necessity measures.
+%The final ranking step provides a useful result record set. 
+It is possible to model the Allen's relations and also more complex relations.
+As future work, allowing the time interval in the query specification to be ill-known and new ways for aggregation and ranking are considered.

### 2011-12-15 ipmu2012/introduction.tex
Last update

@@
 Dealing with time in information systems is usually a difficult task~\cite{Bolour82}.
-
 First of all, the concept of \emph{time} itself is very complex.
-Klein~\cite{klein94} studied the concept of time in language and Devos~\cite{devos94} modelled vague temporal expressions by means of fuzzy sets~\cite{zadeh65}.
 More towards information systems, several proposals have been concerned with the obtaining of theoretical models that allow the representation of time in information systems~\cite{Cru97}.
@@
 
-To allow information systems to cope with these imperfections, many approaches adopt fuzzy sets for the representation of temporal information~\cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy,Dubois:jucs_9_9:fuzziness_and_uncertainty_in}.
-The temporal relations studied by Allen were recently fuzzified by several authors~\cite{ohlbach04,schockaert08}.
+To allow information systems to cope with these imperfections, many approaches adopt fuzzy sets for the representation of temporal information~\cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy},\cite{Dubois:jucs_9_9:fuzziness_and_uncertainty_in}.
+The temporal relations studied by Allen were recently fuzzified by several authors~\cite{ohlbach04},\cite{schockaert08}.
 
 Next to the issues in temporal modelling, the addition of time to a relational database leads to several more practical problems, as the modelling of a time-variant object has an impact on the database consistency.
-Different models have been proposed \cite{Jensen91,Snodgrass84,Nascimento95} that consider time as a crisp notion.
-In recent works, some authors consider the necessity of allowing imprecision in temporal models~\cite{Cru97,Garrido2009}.
+Different models have been proposed \cite{Jensen91},\cite{Snodgrass84},\cite{Nascimento95} that consider time as a crisp notion.
+In recent works, some authors consider the necessity of allowing imprecision in temporal models~\cite{Cru97},\cite{Garrido2009}.
 
@@
 
-In this work we will propose a possibilistic valid time model for relational databases, based on the framework presented in~\cite{Pon11}, which deals with the inference of uncertainty about the evaluation of sets from the uncertainty about the exact values used in evaluation.
+In this work we will propose a possibilistic valid-time model for relational databases, based on the framework presented in~\cite{Pon11}, which deals with the inference of uncertainty about the evaluation of sets from the uncertainty about the exact values used in evaluation.
 

### 2011-12-15 ipmu2012/ipmu.tex
Last update

@@
 \usepackage{multirow}
+
 \hyphenation{pro-per-ties}
@@
 %
-The researchers are supported by the grant BES-2009-013805 within the research project TIN2008-02066: \emph{Fuzzy Temporal Information treatment in relational DBMS}.
+The researchers are supported by the grant BES-2009-013805 in the project TIN2008-02066: \emph{Fuzzy Temporal Information treatment in RDBMS}.
 

### 2011-12-15 ipmu2012/preliminaries.tex
Last update

@@
 
+\vspace{-10pt}
+
 \begin{align}
@@
 
-To calculate possibility and necessity concerning a conjunction of constraints, the $\min$ operator is used. The possibility and necessity of the set satisfying both constraints is then:
+To calculate possibility and necessity concerning a conjunction of constraints, the $\min$ operator is used. The possibility and necessity of the set is: %satisfying both constraints is then:
+
+\vspace{-10pt}
 
@@
 %In this section, the proposed reasoning is applied to the specific context of intervals on the real line. This setting is of specific interest in the context of fuzzy temporal databases.
-A \emph{temporal database} is a database that manages some aspects of time in its schema.
+A \emph{temporal database} is a database that manages some aspects of time in its schema \cite{Dyreson1994}.
 The reality a temporal database tries to model, contains some temporal notions which have to be handled specifically in order to maintain a consistent modelling behavior.
 A \emph{chronon} is the shortest duration of time supported by the database.
-Time can be represented either as points or intervals \cite{655777}.
+Time can be represented either as points~\cite{Dubois89} or intervals ~\cite{Garrido2009},\cite{655777} that may be subject to imperfection.
 %There are proposals for the fuzzyfication of the time point and the fuzzyfication of the time interval.
@@
 	\item
-	\emph{Transaction time}: The time when the fact is stored in the database.
+	\emph{Transaction time}~\cite{Jensen91}: The time when the fact is stored in the database.
 	\item
-	\emph{Valid time}: The time when the fact is true in the modelled reality.
+	\emph{Valid time}~\cite{Snodgrass84}: The time when the fact is true in the modelled reality.
 	\item
-	\emph{Decision time}: The time when an event was decided to happen. 
+	\emph{Decision time}~\cite{Nascimento95}: The time when an event was decided to happen. 
 \end{itemize}
 
-Database models can then be classified into transaction time \cite{Jensen91}, decision time, valid time, bi-temporal \cite{Snodgrass84}(both valid and transaction time) or tri-temporal \cite{Nascimento95} (valid, transaction and decision time) models.
+Database models can also be classified into \emph{bi-temporal} (both valid and transaction-time) or \emph{tri-temporal}  (bi-temporal and decision time) models.
 
-To deal with time points \cite{Dubois89} or intervals \cite{Garrido2009} that are subject to imperfection and can thus not be precisely known, fuzzy temporal models \cite{schockaert08} exist.
+%To deal with time points  or intervals  that are subject to imperfection and can thus not be precisely known, fuzzy temporal models \cite{schockaert08} exist.
 %Some fuzzy temporal models assume that the time stored in the database is an interval.

### 2012-01-10 trunk/book/sources/intro.tex
New version after Christmas holidays.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Introduction
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The concept of \emph{time} is easy to understand but very complex to define~\cite{klein94}~\cite{Shackle61}. 
In Information Systems \emph{IS}, several proposals have been concerned with the obtaining of theoretical models that allow the representation of time~\cite{Bolour82}~\cite{Cru97}. 
More towards temporal databases, a lot of work have been done. The first efforts were towards the representation of the historical information related to an object in the database(reference). Some works tried to extend the Entity Relationship Model \emph{ERM} ~\cite{Klopprogge:1983} but without an impact on any database standards like SQL~\cite{Sarda:1990:ESH:627277.627409}.

The ``Consensus Glossary of Temporal Database concepts'' \cite{Dyreson:1994:CGT:181550.181560} is the first publication where the main researchers in temporal databases define the main thesaurus of temporal database concepts. This glossary defines the main types of time in a temporal database and is the basis of furthers researches ~\cite{Sarda:1990:ESH:627277.627409} \cite{Jensen94thetsql2}.

Another main issue in temporal databases is the querying. The user specifies in the query some temporal constraints, e.g. \emph{``before this year''},\emph{``after 1995''}. In order to solve these temporal comparisons, between the time specified in the query and the temporal data stored in the database, some basic relationships have been defined. Allen~\cite{Allen83} first introduced the temporal relations between time intervals and, to a lesser extent, to time points. Some authors~\cite{ohlbach2004},\cite{nagypal2003},\cite{schockaert08} have softened the temporal operators defined by Allen. Therefore it is possible to compute the relationships between two ill-known time points or intervals. In~\cite{garrido2009}, some different temporal operators are defined by a combination of regular fuzzy comparisons. %e.g. the \emph{BEFORE} operator is defined by means of the fuzzy less than (FLT) operator.

To allow information systems to cope with uncertainty, many approaches adopt fuzzy sets~\cite{Zadeh65} for the representation of temporal information~\cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy},\cite{Dubois:jucs_9_9:fuzziness_and_uncertainty_in},~\cite{devos94}. The time points representing the time when the object is valid in the reality might not be precisely known. In order to deal with this, several models have been proposed. Garrido~\cite{garrido2009} proposes a model to deal with uncertainty in the time interval by means of fuzzy intervals. Bronselaer~\cite{Pon11} proposes a consistent framework to deal and represent ill-known time values and their relationships. Rough sets~\cite{Pawlak1995} have been also used to represent time intervals~\cite{Qia09} in databases.

%temporal reasoning and fuzzy temporal reasoning.
Next to the manage of temporal information is the temporal reasoning~\cite{Allen83}. Dubois and Prade ~\cite{Dubois:jucs_9_9:fuzziness_and_uncertainty_in},\cite{Dubois89} deal with fuzzyness and uncertainty in temporal reasoning, but this topic is outside the scope of this work.

%temporal commercial systems.

The work is organized as follows:

In Section \ref{sec:time-domain} an study about the time and its properties is done. Section \ref{sec:temporal-databases} is an overview of the main problems when managing the time in a database, the different proposals for solving them and the temporal database proposals. The section concludes with a comparison among the different proposals. In Section \ref{sec:fuzzy-temporal-databases} we present a sum up with the proposals for handling the imprecision in temporal databases and the main deficiences we have detected. Finally, there are several commercial temporal database management systems  \emph{TDMBS} like \cite{oracle2009},\cite{posgree2009},\cite{teradata2011},\cite{timedb2005}. All of them are analyzed and compared later on this work.

### 2012-01-10 book/author.tex
New version after Christmas holidays.

--- 2011-11-09 trunk/book/author.tex
@@
 % your contribution title if the original one is too long
-\author{Jos\'e Pons \and Christophe Billiet \and Olga Pons}
+\author{Jos\'e Pons \and Christophe Billiet \and Olga Pons \and Guy de Tr\'e}
 % Use \authorrunning{Short Title} for an abbreviated version of
@@
 \and Olga Pons \at Department of Computer Science and Artificial Intelligence, Univeristy of Granada, C/ Periodista Daniel Saucedo Aranda s/n \email{opc@decsai.ugr.es}
+\and Guy De Tr\'e \at Department of Telecommunications and Information Processing, Ghent University, Belgium, St.-Pietersnieuwstraat 41, B-9000 Gent,Belgium. \email{Guy.DeTre@telin.ugent.be}
 }

### 2012-01-10 book/sources/timeDomain.tex
New version after Christmas holidays.

--- 2011-11-09 trunk/book/sources/timeDomain.tex
@@
 \end{svgraybox}
+Any temporal model may represent the time as points as well as intervals. The equivalence between a interval-based and point-based temporal data models is demonstrated in \cite{Böhlen_point-versus}.
 
@@
 On the other hand, time may be \textbf{relative} or \textbf{absolute} (in other words, \textbf{anchored} or \textbf{unanchored}).
-Sometimes, what one considers 'absolute time' is not as definite as one would hope since our concept of absolute time is based on another time reference (for example, January 1st of year one).
+Sometimes, what one considers \emph{absolute time} is not as definite as one would hope since our concept of absolute time is based on another time reference (for example, January 1st of year one).
 Relative time has a direction, differently from distance.
@@
 
-%%% time as points and time as intervals:
+%%% granularity
+%\subsubsection{Granularity}
+Next to the inherent difficulties of time, an issue is \textbf{temporal granularity}.
+A temporal granularity is a partitioning of the time line used by a system, usually dependent on the application.
+For example, the age of an adult human being is usually expressed in years: one will use sentences like \emph{`Laura is 21 years old'} instead of \emph{`Laura is 21 years, 3 months and 4 days old'}.
+In this example any time period shorter than a year needs no representation and thus the used granularity allows no specification for time periods shorter than a year.
 
-%%% granularity
+The transition between granularities~\cite{Lin97} is also considered as a source of imprecision.
+In a sentence like \emph{`The Alhambra is 612 years old'}, the temporal notion is very precise if time periods shorter than a year are irrelevant in the used granularity.
+However, if one is concerned with the exact day on which the last stone of the building is laid, a transition to a finer granularity, allowing the specification of time periods shorter than a year, is needed.
+Therefore, some proposals consider granularity as the base of the temporal model~\cite{Cru97}.
 
@@
 %an explanation on allen's temporal relations.
+%definition of fuzzy temporal interval
 
-\subsection{An example of time domain: Julian Day Number}
+Several operators are defined in order to compare temporal elements.
+Allen~\cite{Allen83} first described the relations between time intervals and to a lesser extent, between time points.
+Figure \ref{fig:allen} shows the temporal relations.
+There are some proposals that can be applied to both crisp and fuzzy temporal interval ~\cite{ohlbach2004},\cite{nagypal2003},\cite{schockaert08}.
+For this fuzzy comparisons~\cite{garrido2009} proposed some different temporal relations that can be easyly implemented by means of regular fuzzy operations.
+
+\begin{figure}
+\centering
+\input{graphs/allen.tex}
+\caption{Allen relations between two time intervals.}
+\label{fig:allen}
+\end{figure}
+
+\subsection{An example of temporal domain: Julian Day Number}
 % proposal for the underlying Julian Day Number domain.
 
-The Julian Day number \cite{Dir96} is a counter.
+The Julian Day number \emph{JDN} \cite{Dir96} is a counter.
 Its value is incremented in one unit every day from 1 January 4713 B.C.
 at 12:00 noon.
-The particularity of starting at noon was useful for astronomers: the observations they took one night belonged to the same Julian Day.
-Actually it has not any transcendence.\\
+The particularity of starting at noon was useful for astronomers: the observations they took one night belonged to the same Julian Day.\\
 Note that the Julian Day represents whole days.
@@
 or J.D.E.).
-Any time scale different from Universal Time must be explicited after the Julian Day number.\\
+Any time scale different from Universal Time must be explicited after the Julian Day number.
+There are several conversion formulas~\cite{Usn}\cite{Wik}\cite{Lea} between a date in Gregorian calendar  format and a date in  JDN format.
+The inverse conversion formula is proposed in~\cite{Fliegel:1968:LEM:364096.364097}\\
 
-\subsubsection{Representation alternatives}
 There are many alternatives to optimize the representation of Julian Day numbers because of its extremely far origin (4713 B.C.
@@
 
-\subsubsection{Conversion among dates}
-\paragraph{Conversion between Gregorian and Julian dates}
-The Julian Date consist on the Julian Day number and the time. It is computed with the formula \cite{Usn}:\\
-\begin{equation}
-\label{eq:julian day}
-JD = JDN + \frac{hour-12}{24} + \frac{minutes}{1440} + \frac{seconds}{86400}
-\end{equation}
-
-Here the JDN is the Julian Day Number.
+%\subsubsection{Conversion among dates}
+%\paragraph{Conversion between Gregorian and Julian dates}
+%The Julian Date consist on the Julian Day number and the time. It is computed with the formula \cite{Usn}:\\
+%\begin{equation}
+%\label{eq:julian day}
+%JD = JDN + \frac{hour-12}{24} + \frac{minutes}{1440} + \frac{seconds}{86400}
+%\end{equation}
+%
+%Here the JDN is the Julian Day Number.
 This equation does not take into account the leap seconds \footnote{A leap second(s) is the amount of second(s) added or deleted every year in the Universal Time.\cite{Wik},\cite{Lea}.
@@
 
-\paragraph{Gregorian calendar to Julian day}
-Any date from Gregorian calendar \footnote{As Gregorian calendar entered in use in 1582, any use before that date is called \emph{proleptic calendar}.} can be converted into a Julian Day using the following formula:\\
-\begin{equation}
-\label{eq:JDN}
-JDN = \frac{1461x(Y + 4800 + \frac{M-14}{12})}{4}+\frac{367x(M-2-12x\frac{M-14}{12})}{12}-\frac{3x\frac{Y+4900+\frac{M-14}{12}}{100}}{4} 
-\end{equation}
-
-This formula was created by Henry Fliegel and Thomas C.
+%\paragraph{Gregorian calendar to Julian day}
+%Any date from Gregorian calendar \footnote{As Gregorian calendar entered in use in 1582, any use before that date is called \emph{proleptic calendar}.} can be converted into a Julian Day using the following formula:\\
+%\begin{equation}
+%\label{eq:JDN}
+%JDN = \frac{1461x(Y + 4800 + \frac{M-14}{12})}{4}+\frac{367x(M-2-12x\frac{M-14}{12})}{12}-\frac{3x\frac{Y+4900+\frac{M-14}{12}}{100}}{4} 
+%\end{equation}
+%
+%This formula was created by Henry Fliegel and Thomas C.
 van Flandern in 1968.
@@
 The value $Y$ represents the year and the value $M$ represents the month.
-\paragraph{Julian calendar to Julian day}
-The algorithm to convert between Julian calendar and Julian Day Number is only valid for dates since 4713 B.C (positive values of Julian Day) \cite{Dog}:\\
-\begin{equation}
-\label{eq:JDN_2_JD}
-JDN=367Y-\frac{7 (Y + 5001 + \frac{M-9}{7})}{4}+\frac{275xM}{9}+D+1729777
-\end{equation}
-
-Where $Y$ is the year value, $M$ is the month value, and $D$ is the day value.
-
-%\subsubsection{Membership degree computation}
+%\paragraph{Julian calendar to Julian day}
+%The algorithm to convert between Julian calendar and Julian Day Number is only valid for dates since 4713 B.C (positive values of Julian Day) \cite{Dog}:\\
+%\begin{equation}
+%\label{eq:JDN_2_JD}
+%JDN=367Y-\frac{7 (Y + 5001 + \frac{M-9}{7})}{4}+\frac{275xM}{9}+D+1729777
+%\end{equation}
+%
+%Where $Y$ is the year value, $M$ is the month value, and $D$ is the day value.
+%
+%%\subsubsection{Membership degree computation}
 %The main advantage from the Julian day number is the ability to make calculations. Thus, the membership degree computation for a fuzzy number based on a Julian day is explained here. \\

### 2012-01-12 trunk/ipmu2012/proposal.tex
The very freaking final but one stupid sentence too long version of the paper.

@@
-In this section, a valid-time model will be presented.
-The focus will off course be on the handling of imperfect valid-time notions.
+In this section, a valid-time model, able to handle imperfect valid-time notions, will be presented.
 First, the representation and storage of valid-time will be presented.
-Next, a way of querying the model is proposed.
+Next, an approach to query the model is proposed.
 Finally, an example is given.
@@
 \label{subsec:representation-ill-known}
-A database models a real object by storing the object's value for each attribute describing a property of the object.
-Thus, a valid-time database following the presented proposal will model the time during which an object in a certain state is valid, by associating a \emph{Possibilistic valid-time Period} to the record describing said object state:
+A database models real objects by storing the values of an object for each attribute describing a property of the object.
+Thus, a valid-time database following the presented proposal will model the time during which an object in a certain state is valid, by associating a \emph{Possibilistic Valid-time Period} (PVP) to the record describing this object state:
 
 \begin{definition}
-A \emph{Possibilistic valid-time Period} (PVP) is an ill-known interval in time, which models a time period during which an object is valid.
+A \emph{Possibilistic Valid-time Period} is an ill-known interval in time, which models a time period during which an object in a certain state is valid.
 \end{definition}
 
-Off course, the object of which the PVP models the valid-time is the object in the particular state modelled by the record that is associated with the PVP.
-Because a PVP is an ill-known interval, it allows modelling the uncertainty about the start or end point of a time interval (and thus about the time interval itself) if such uncertainty exists.
+Because a PVP is an ill-known interval, it allows modelling the uncertainty about the start and/or end point of a time interval (and thus about the time interval itself) if such uncertainty exists.
 The interpretation is \emph{disjunctive}: the PVP represents exactly one valid-time interval, but precisely \emph{which} interval is represented, is (partially) unknown.
-In the presented model, only PVPs are considered in which the possibility distributions of the possibilistic variables defining the start and end point of the ill-known interval have the exact same characteristics of the membership functions of fuzzy numbers.
+In the presented model, only PVPs are considered of which the possibility distributions of the possibilistic variables defining the start and end point of the ill-known interval have exactly the same characteristics as the membership functions of fuzzy numbers.
 A perfectly known start or end point can then be modelled by such an ill-known value defined by a possibilistic variable $P$ for which $\exists !
@@
 As mentioned in \cite{Pon11}, this approach differs from the one where a valid-time period is represented by one fuzzy set.
-This one fuzzy set is seen as a possibility distribution on $\mathbb{R}$.
-However, in the presented approach, a time period is modelled using an ill-known set, which is defined by a possibility distribution on $\wp(\mathbb{R})$.
+Such a fuzzy set is seen as a possibility distribution on $\mathbb{R}$ and thus defines just one ill-known value.
+However, in the presented approach, a time period is modelled using an ill-known set, which is defined by a possibility distribution on $\Pow(\mathbb{R})$.
 
@@
 To store a PVP, the possibility distributions defining the ill-known start and end point are stored.
-To store such a possibility distribution, the representation as used in the fuzzy interface for relational databases \emph{FIRST}~\cite{Medina94gefred.a,Gal98} is used.
+To store such a possibility distribution, the representation as used in the fuzzy interface for relational databases \emph{FIRST}~\cite{Medina94gefred.a}, \cite{Gal98} is used.
 Using this representation, it is possible to represent not only fuzzy numbers, but also (fuzzy) constants.
-To store an ill-known point, four values (FT, F1, F2 and F3) are stored.
+To store an ill-known value, four values (FT, F1, F2 and F3) are stored.
 They are explained in table \ref{table:relational-representation-pvp}.
-Note that while NULL denotes the fuzzy constant NULL, N denotes a null value in the database.
+Note that while NULL denotes the fuzzy constant NULL, N denotes a null value in the database \cite{Medina94gefred.a}.
 
@@
 FT denotes Fuzzy Type.
-Field FT indicates that the values stored in F1, F2 and F3 denote either NULL, UNKNOWN, UNDEFINED or a triangular possibility distribution $\left[D,\ a,\ b \right]$.
+Field FT indicates that the values stored in F1, F2 and F3 denote either NULL, UNKNOWN, UNDEFINED or a triangular possibility distribution $M$ = $\left[D,\ a,\ b \right]$.
 Fields F1, F2 and F3 contain the actual values.}%  A \emph{PVP} is represented by two ill-known points.}
@@
 NULL & 2 & N & N & N &not defined & Nothing is known about the value \\ 
-$\left[D,\ a,\ b \right]$ & 3 & $D$ & $a$ & $b$ & $\mu(\left[D,\ a,\ b \right])$ & Ill-known value \\ 
+$M$ & 3 & $D$ & $a$ & $b$ & $\mu_{M}$ & Ill-known value \\ 
 \hline 
@@
 \subsection{Querying Ill-known Valid-time Intervals}
-This subsection provides the tools for querying.
+This subsection discusses a tool for querying.
 The focus will evidently lie on the querying of valid-time periods.
@@
 \subsubsection{Query Structure}
-It is important to notice that every valid-time notion in the database is represented using a PVP.
-Thus, valid-time intervals in the database can be partially unknown.
-In this framework, a query has two separate parts:
+It is important to notice that every valid-time notion in the database is represented using a PVP. In the presented framework, a query has two separate parts:
 
@@
 \end{definition}
-Here, $I$ is a crisp time interval and $AR$ is one of Allen's relations\cite{Allen83}.
+Here, $I$ is a crisp time interval and $AR$ is one of Allen's relations \cite{Allen83}.
 The interpretation is that for a record with valid-time period given by a PVP $J$, the user requires that $I$ $AR$ $J$ holds.
@@
 The evaluation of a query $\tilde{Q} = \left( Q^{time}, Q \right)$, where $Q^{time} = \left( I, AR \right)$ is now handled as follows.
-For each record $r$ in the database, with the valid-time notion of $r$ specified by PVP $J$, two things happen independently:
+For each record $r$ in the database, with the valid-time notion of $r$ being specified by a PVP $J$, two things happen independently:
 
@@
 The preferences expressed in $Q$ are evaluated, resulting in a satisfaction degree denoted here $e_{Q}(r)$.
-The model presented here accepts any way of calculating this evaluation, as long as $e_{Q}(r) \in \left[0,1\right]$.
+The presented model accepts any sound way of calculating this evaluation, as long as $e_{Q}(r) \in \left[0,1\right]$.
 \item
-Depending on AR, a specific set of ill-known constraints is considered.
-For each AR, this set can be found in table \ref{tab:allen-relations}.
-The possibility $\Pos_{Q^{time}}(r)$ and necessity $\Nec_{Q^{time}}(r)$ that $r$ fulfills all these constraints are calculated using formulas based on equations \eqref{ill-known-pos} and \eqref{ill-known-nec}.
-The sum of these possibility and necessity scores gives an evaluation score $e_{Q^{time}}(r)$ in interval $\left[0,2\right]$.
-Because necessity cannot exceed $0$ unless possibility is $1$, this gives a natural score.
+Depending on AR, a specific set of ill-known constraints is considered, which can be found in table 2.
+The possibility $\Pos_{Q^{time}}(r)$ and necessity $\Nec_{Q^{time}}(r)$ that $r$ fulfills all these constraints are calculated using formulas based on equations \eqref{ill-known-pos} respectively \eqref{ill-known-nec} and aggregated using the $\min$ operator.
+
+%The sum of the possibility score and the necessity score gives an evaluation score $e_{Q^{time}}(r)$ in interval $\left[0,2\right]$.
+Because necessity cannot exceed $0$ unless possibility is $1$, this gives a natural ranking score.
 \end{itemize}
@@
 \label{tab:allen-relations}
-\caption{Allen's relations represented in the framework.
+\caption{Allen's relations used in the framework.
 Here, $I = \left[a, b\right]$ denotes the PVP in the query, $J = \left[X, Y\right]$ denotes the PVP of the record, with $\pi_{X}$ and $\pi_{Y}$ the possibility distributions of $X$ and $Y$ respectively.
 For each of Allen's relations $AR$, the corresponding value in column `Constraints' gives the constraints corresponding to $AR$.
-The last column contains the corresponding formula to calculate the possibility that $I$ satisfies all constraints given in column `Constraints' and thus that $I$ $AR$ $J$ holds.
-These formulas are based on equation \eqref{ill-known-pos}.}
+The last column contains the corresponding formula to calculate the possibility that $I$ satisfies all constraints given in column `Constraints', which are based on equation \eqref{ill-known-pos}.}
 
@@
 \subsubsection{Ranking and aggregation}
-In order to present the results to the user, a ranking method is used.
-First of all, $e_{Q^{time}}(r)$ is rescaled to the unit interval.
-The final ranking $e_{final}$ is now given by a convex combination:
+In order to present the results to the user, a crude ranking method is used: for every record $r$, the sum of $\Pos_{Q^{time}}(r)$ and $\Nec_{Q^{time}}(r)$ gives an evaluation score $e'_{Q^{time}}(r)$ in interval $\left[0,2\right]$.
+Because necessity cannot exceed $0$ unless possibility is $1$, this gives a natural ranking score.
+This $e'_{Q^{time}}(r)$ is then rescaled to the unit interval, resulting in $e_{Q^{time}}(r)$.
+The final ranking $e_{final}(r)$ is now given by a convex combination:
 
@@
 
-The use of this convex combination allows a record to make up for a low score on the temporal constraint by a good score on the non-temporal constraint (or vice versa).
+The use of this convex combination allows a record to make up for a low score for the temporal constraint by a good score for the non-temporal constraint (or vice versa).
 Changing $\omega$ also allows granting the temporal constraint more weight with respect to the non-temporal constraint (or vice versa).
@@
 \subsubsection{The database}
-Consider example database \ref{tb:car-models} describing car models, containing general attributes (model name, manufacturer, car segment) and one temporal attribute (stored as explained in subsection \ref{subsec:storage}) containing the approximate period of time  in years during which the car model was sold.
+Consider the example relation given in table \ref{tb:car-models} describing car models, containing general attributes (model name, manufacturer, car segment) and one temporal attribute (stored as explained in subsection \ref{subsec:storage}) describing the approximate period of time during which the car model was sold.
 The value for $D$ is stored in \emph{yyyy} format and $a$ and $b$ are represented by an integer.
@@
 Table \ref{tb:results} shows a natural and gradual ranking for the results.
-The last record, (ID 001, IID 3) shows also that with omega=0.5 both temporal and regular criteria have the same importance.% By lowering the value for omega the temporal criteria would have more impact on the final ranking.
+The last record, (ID 001, IID 3) shows also that with $\omega$ = 0.5 both temporal and regular criteria have the same importance.% By lowering the value for omega the temporal criteria would have more impact on the final ranking.

### 2012-01-12 ipmu2012/conclusions.tex
The very freaking final but one stupid sentence too long version of the paper.

@@
 We presented a valid-time model to represent and query ill-known temporal intervals.
-The main advantage of this framework is that it works with ill-known intervals and thus tries to correctly represents and handle querying temporal uncertainty in the database by providing both possibility and necessity measures.
+The main advantage of this framework is that it works with ill-known intervals and thus tries to correctly represent and handle querying of temporal uncertainty in a database.
 %The final ranking step provides a useful result record set. 

### 2012-01-12 ipmu2012/introduction.tex
The very freaking final but one stupid sentence too long version of the paper.

@@
 First of all, the concept of \emph{time} itself is very complex.
-More towards information systems, several proposals have been concerned with the obtaining of theoretical models that allow the representation of time in information systems~\cite{Cru97}.
-Notably, temporal relations between time intervals and to a lesser extent time points, were first studied by Allen~\cite{Allen83}.
+More towards information systems, several proposals have been concerned with obtaining theoretical models that allow the representation of time in information systems~\cite{Cru97}.
+Notably, temporal relations between time intervals and to a lesser extent time points, were studied by Allen~\cite{Allen83}.
 
@@
 One issue in these attempts at temporal modelling is off course the inherent complexities of the concept of time.
-Another is the possible inherent imperfection in the description of notions of time.
-The temporal notion in a sentence like `The Alhambra, which I can see from my room, is a few hundred years old.' contains imperfection caused by the vagueness of the used expression.
-The temporal notion in a sentence like `The Alhambra was finished somewhere between 1301 A.D.
-and 1400 A.D.' contains imperfection caused by the inaccuracy in the given temporal notion, though precise boundaries are given.
+Another is the possible inherent imperfections in the description of notions of time.
+The description of the temporal notion in a sentence like `The Alhambra, which I can see from my room, is a few hundred years old.' contains imperfection caused by the vagueness of the used expression.
+The description of the temporal notion in a sentence like `The Alhambra was finished somewhere between 1301 A.D.
+and 1400 A.D.' contains imperfection caused by the uncertainty in the used expression about the exact finishing day of the building.
+Though precise boundaries are given, the exact day on which the building was finished is unknown.
 The transition between temporal granularities~\cite{Lin97} is also considered as a source of imprecision in temporal modelling and some proposals consider granularity as the base of the temporal model~\cite{Cru97}.
@@
 
-Next to the issues in temporal modelling, the addition of time to a relational database leads to several more practical problems, as the modelling of a time-variant object has an impact on the database consistency.
+Next to the issues in temporal modelling, the inclusion of time in a relational database system leads to several more practical problems, as the modelling of the temporal aspects of a time-variant object has an impact on the database consistency.
 Different models have been proposed \cite{Jensen91},\cite{Snodgrass84},\cite{Nascimento95} that consider time as a crisp notion.
-In recent works, some authors consider the necessity of allowing imprecision in temporal models~\cite{Cru97},\cite{Garrido2009}.
+In some specific works, authors consider the necessity of allowing imprecision in temporal models~\cite{Cru97},\cite{Garrido2009}.
 
-Most of these models use the concepts of fuzzy numbers and fuzzy intervals to represent temporal imprecision and considerable attention has thus been given to the problem of transforming two fuzzy numbers which represent the boundaries of an imprecise time interval into one fuzzy interval representing this interval.
+Most of these models use the concepts of fuzzy numbers and fuzzy intervals to represent imprecise temporal notions and considerable attention has thus been given to the problem of transforming two fuzzy numbers which represent the boundaries of an imprecise time interval into one fuzzy interval representing this interval.
 There are several proposals for this transformation, but some, like those presented in~\cite{Garrido2009} are misleading.
@@
 
-The paper is organized as follows. In section \ref{sec:preliminaries}, some preliminaries and concepts are presented. 
+The remainder of the paper is organized as follows. In section \ref{sec:preliminaries}, some preliminaries and concepts are presented. 
 Section \ref{sec:proposal} presents the possibilistic model for valid-time databases. In this section, both representation and querying are introduced.

### 2012-01-12 ipmu2012/ipmu.tex
The very freaking final but one stupid sentence too long version of the paper.

@@
 %
-\author{Jos\'e Enrique Pons \and Christophe Billiet \and Olga Pons Capote \and Guy De Tr\'e}
+\author{Jos\'e Enrique Pons\inst{1} \and Christophe Billiet\inst{2} \and Olga Pons Capote\inst{1} \and Guy De Tr\'e\inst{2}}
 %Jeffrey Dean \and David Grove \and Craig Chambers \and Kim~B.~Bruce \and
@@
 \institute{
-Department of Computer Science and Artificial Intelligence \\
-Universidad de Granada \\
-Escuela T\'ecnica Superior de Ingenier\'ia Inform\'atica \\
-C/Periodista Daniel Saucedo Aranda s/n \\
-E-18071 (Granada-Spain)\\
+Department of Computer Science and Artificial Intelligence, Universidad de Granada, Escuela T\'ecnica Superior de Ingenier\'ia Inform\'atica \\
+C/Periodista Daniel Saucedo Aranda s/n, E-18071, Granada, Spain\\
 \email{jpons,opc@decsai.ugr.es}\\ 
@@
 \and 
-Department of Telecommunications and Information Processing
-Ghent University
-Sint Pietersnieuwstraat 41 B-9000, Ghent, Belgium.
+Department of Telecommunications and Information Processing, Ghent University \\
+Sint-Pietersnieuwstraat 41 B-9000, Ghent, Belgium.
 \email{Christophe.Billiet,Guy.De.Tre@ugent.be}
@@
 Because in temporal databases, temporal notions influence the consistency with which the database models the reality, this uncertainty has a direct impact on the consistency of the model.
-To represent this temporal uncertainty, previous works have adapted fuzzy sets with conjunctive interpretation.
-This approach might prove misleading.
+To represent this temporal uncertainty, previous works have adapted fuzzy sets with conjunctive interpretation, an approach that might prove misleading.
 This work presents a model that represents the uncertainty using possibility and necessity distributions, which are fuzzy sets with disjunctive interpretations.
@@
 %
-The researchers are supported by the grant BES-2009-013805 in the project TIN2008-02066: \emph{Fuzzy Temporal Information treatment in RDBMS}.
+Part of the researchers are supported by the grant BES-2009-013805 in the project TIN2008-02066: \emph{Fuzzy Temporal Information treatment in RDBMS}.
 

### 2012-01-12 ipmu2012/preliminaries.tex
The very freaking final but one stupid sentence too long version of the paper.

@@
 %Introductory text:
-In this section, some basic concepts are introduced concerning possibilistic variables and fuzzy numbers and intervals, after which the framework of set evaluation by ill-known constraints\cite{Pon11} is explained.
+In this section, some basic concepts are introduced concerning possibilistic variables and fuzzy numbers and intervals, after which the framework of set evaluation by ill-known constraints \cite{Pon11} is explained.
 The section concludes with a brief introduction to temporal databases.
@@
 \subsection{\label{subsec:possibilistic-variables}Possibilistic Variables}
-Possibilistic variables rely on possibility theory\cite{Dubois:Prade:1988:PossibilityTheory}. A \emph{possibilistic variable} is defined as follows\cite{Pon11}.
+Possibilistic variables rely on possibility theory \cite{Dubois:Prade:1988:PossibilityTheory}. A \emph{possibilistic variable} is defined as follows \cite{Pon11}.
 
@@
 A possibilistic variable $X$ over a universe $U$ is defined as a variable taking exactly one value in $U$, but for which this value is (partially) unknown.
-Its possibility distribution $\pi_X$ on $U$ gives the available knowledge about the value that $X$ takes: for each $u\in U$, $\pi_X(u)$ represents the possibility that $X$ takes the value $u$.
+Its possibility distribution $\pi_X$ on $U$ models the available knowledge about the value that $X$ takes: for each $u\in U$, $\pi_X(u)$ represents the possibility that $X$ takes the value $u$.
 In this work, this possibility is interpreted as a measure of how plausible it is that $X$ takes the value $u$, given (partial) knowledge about the value $X$ takes.
@@
 
-The exact value a possibilistic variable takes, which is (partially) unknown, is called an \emph{ill-known value} in this work\cite{Dubois88b}.
+The exact value a possibilistic variable takes, which is (partially) unknown, is called an \emph{ill-known value} in this work \cite{Dubois88b}.
 
 When a possibilistic variable is defined on the powerset $\Pow(R)$ of some universe $R$, the unique value the variable takes will be a crisp set and its possibility distribution on the powerset $\Pow(R)$ will describe the possibility of each crisp subset of $R$ to be the value the variable takes.
-This exact value (a crisp set) the variable takes, is now called an \emph{ill-known set}\cite{Dubois88b}.
+This exact value (a crisp set) the variable takes, is now called an \emph{ill-known set} \cite{Dubois88b}.
 
@@
 Indeed, any boolean proposition $p$ takes just one value in $\mathbb{B}$.
-If the knowledge on which value this proposition $p$ will take is given by a possibility distribution $\pi_p$, the proposition can be seen as a possibilistic variable.
+If the knowledge about which value this proposition $p$ will take is given by a possibility distribution $\pi_p$, the proposition can be seen as a possibilistic variable.
 As the interest lies with the case where the proposition holds, the possibility and necessity that $p$ = $T$ (the proposition holds) demand most attention.
@@
 A closed ill-known interval with start point defined by possibilistic variable $X$ and end point by possibilistic variable $Y$ is noted here $\left[X, Y\right]$.
-The correspondences and transitions between the representations of ill-known sets and between the representation of an ill-known set and an ill-known interval are part of the authors current research.
+The correspondences and transitions between the representations of ill-known sets, between the representations of ill-known intervals and between the representations of an ill-known set and an ill-known interval are part of the authors current research.
 
 \subsection{\label{subsec:fuzzy-numbers}Fuzzy Numbers and Fuzzy Intervals}
-Among others, Dubois and Prade~\cite{Dubois1983} use fuzzy sets\cite{zadeh65} to define a \emph{fuzzy interval}:
+Among others, Dubois and Prade~\cite{Dubois1983} use fuzzy sets \cite{zadeh65} to define a \emph{fuzzy interval}:
 \begin{definition}
@@
 
-The most convenient form of the membership function of a fuzzy number is a triangular function.
+The most convenient form of the membership function of a fuzzy number is a triangular form.
 It can be shown that such a membership function $\mu_M$ for a fuzzy number $M$ is convex and normalized.
@@
 \begin{definition}
-Given a universe $U$, an \emph{ill-known constraint} $C$ on a set $A \subseteq U$ is specified by means of a binary relation $R \subseteq U^{2}$ and a fixed, ill-known value denoted by its possibilistic variable $V$ over $U$, i.e.:
+Given a universe $U$, an ill-known constraint $C$ on a set $A \subseteq U$ is specified by means of a binary relation $R \subseteq U^{2}$ and a fixed, ill-known value denoted by its possibilistic variable $V$ over $U$, i.e.:
 \begin{align}
@@
 
-The satisfaction of a constraint $C \triangleq (R,V)$ by a set $A$ is now still a Boolean matter, but due to the uncertainty about the ill-known value $V$, it can be uncertain whether $C$ is satisfied by $A$ or not\cite{Pon11}.
+An example of an ill-known constraint is $C_{ex} \triangleq (<, X)$. Some set $A$ then satisfies $C_{ex}$ if $\forall a \in A : a < X$, given possibilistic variable $X$.
+
+The satisfaction of a constraint $C \triangleq (R,V)$ by a set $A$ is basically still a Boolean matter, but due to the uncertainty about the ill-known value $V$, it can be uncertain whether $C$ is satisfied by $A$ or not \cite{Pon11}.
 In fact, this satisfaction now behaves as a proposition.
@@
 \begin{align}
-\Pos(A\text{ satisfies }C) & = \min_{a \in A}\left(\sup_{(a,w) \in R}\pi_{V}(w))\right) \label{ill-known-pos}\\
+\Pos(A\text{ satisfies }C) & = \min_{a \in A}\left(\sup_{(a,w) \in R}\pi_{V}(w)\right) \label{ill-known-pos}\\
 \Nec(A\text{ satisfies }C) & = \min_{a \in A}\left(\inf_{(a,w) \notin R} 1-\pi_{V}(w)\right) \label{ill-known-nec}
@@
 
-Now, to check if crisp interval $I = \left[j, k\right]$ is included in $\left[X, Y\right]$, 2 ill-known constraints are constructed:
+Now, e.g., to check if crisp interval $I = \left[j, k\right]$ is included in $\left[X, Y\right]$, 2 ill-known constraints are constructed:
 
@@
 \begin{eqnarray}
-C_1\triangleq\left(\geq,X\right)\\
-C_2\triangleq\left(\leq,Y\right).
+C_1 & \triangleq\left(\geq,X\right)\\
+C_2 & \triangleq\left(\leq,Y\right)
 \end{eqnarray}
 
-To calculate possibility and necessity concerning a conjunction of constraints, the $\min$ operator is used. The possibility and necessity of the set is: %satisfying both constraints is then:
+To calculate the possibility and necessity concerning a conjunction of constraints, the $\min$ operator can be used.
+The possibility and necessity of $I$ being included in $\left[X, Y\right]$ are now: %satisfying both constraints is then:
 
@@
 \label{eq:interval-pos}
-\Pos(A\text{ satisfies }C_1\ AND\ C_2) & = \min_{a \in A}\left(\sup_{a \geq w}\pi_{X}(w),\sup_{a \leq v}\pi_{Y}(v)\right)\\
+\Pos(I\text{ satisfies }C_1\ and\ C_2) & = \min_{a \in I}\left(\sup_{a \geq w}\pi_{X}(w),\sup_{a \leq v}\pi_{Y}(v)\right)\\
 \label{eq:interval-nec}
-\Nec(A\text{ satisfies }C_1\ AND\ C_2) & = \min_{a \in A}\left(\inf_{a < w} 1-\pi_{X}(w),\inf_{a > v} 1-\pi_{Y}(v)\right).
+\Nec(I\text{ satisfies }C_1\ and\ C_2) & = \min_{a \in I}\left(\inf_{a < w} 1-\pi_{X}(w),\inf_{a > v} 1-\pi_{Y}(v)\right).
 \end{align}
@@
 
-The temporal notions in temporal databases can be classified into 4 types based on their interpretation and modelling purpose. User-defined time has no interpretation, but the other types do:
+The temporal notions in temporal databases can be classified into four types based on their interpretation and modelling purpose.
+\emph{User-defined time} has no specific impact on the database consistency, but the other types do:
 
@@
 	\item
-	\emph{Transaction time}~\cite{Jensen91}: The time when the fact is stored in the database.
+	\emph{transaction time}~\cite{Jensen91}: time when the fact is stored in the database.
 	\item
-	\emph{Valid time}~\cite{Snodgrass84}: The time when the fact is true in the modelled reality.
+	\emph{valid time}~\cite{Snodgrass84}: time when the fact is true in the modelled reality.
 	\item
-	\emph{Decision time}~\cite{Nascimento95}: The time when an event was decided to happen. 
+	\emph{decision time}~\cite{Nascimento95}: time when an event was decided to happen. 
 \end{itemize}

### 2012-01-12 trunk/book/author.tex
update

@@
 \usepackage{url}
+\usepackage{amsmath}
+\usepackage{amssymb}
+
+\hyphenation{pro-per-ties}
+\hyphenation{ge-ne-ral-ly}
+\hyphenation{pre-fe-ren-ces}
+\hyphenation{u-sing}
+\hyphenation{pu-nish-ment}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\section{Bipolarity on temporal databases}
-\label{sec:bipolarity}
-\input{sources/bipolarity.tex}
+%\section{Bipolarity on temporal databases}
+%\label{sec:bipolarity}
+%\input{sources/bipolarity.tex}
 
@@
 \input{referenc}
+
+\section*{Appendix: Introduction to fuzzy logic}
+\label{app:fuzzy-logic}
+\input{sources/fuzzylogic.tex}
+
 \end{document}

### 2012-01-12 trunk/book/sources/temporalDatabases.tex
update

@@
 %
-This section defines the model used to represent time in a database.
-This model allows working with time instants and time intervals.
-Even though there is no general agreement as to whether real-time line is continuous or discrete, a discrete model of time may be the most appropriate for a temporal database.
+A \emph{temporal database}~\cite{etzion1998} is a database that manages some aspects of time in its schema~\cite{Dyreson1994}.
+The reality a temporal database tries to model, contains some temporal notions which have to be handled specifically in order to maintain a consistent modelling behavior.
+A very extensive list with the most well-known models in temporal databases can be found in ~\cite{Yu1998}.
+Nevertheless, it is neccesary to define some consistent way to query the temporal data.
+There are several languages for querying temporal databases like TSQL~\cite{TSQL}.
+In~\cite{Snodgrass98} a proposal to add temporal support to the standard SQL is given.
 
-	A chronon may be defined as the basic unit of time.
-Mathematically, a chronon is isomorphic to a finite sequence of natural numbers.
-For instance, if a chronon is \emph{11232010}, it represents a day, namely 11/23/2010.
-The symbol \emph{c} denotes a chronon.
-A chronon may represent different time granularity, depending on database or user processing needs (i.e.
-minutes, seconds, milliseconds, etc.).
-Real-time instants are smaller than chronons even though they are represented by the said chronons The real-time line is represented by a finite sequence of chronons.
+\begin{svgraybox}
+The temporal notions in temporal databases can be classified into the following types based on their interpretation and modelling purpose.
+User-defined time has no interpretation, but the other types do:\\
 
-	It is possible to represent time intervals by a sequence of consecutive chronons.
-A sequence of consecutive chronons could be represented by the pair of starting and terminating chronons.
-A restriction must be imposed so that the starting instant comes before the ending instant.
-\subsubsection{Temporal data models and time domains}
-\label{sec:temporaldatamodelsandtimedomains}
-A temporal data model is a relational data model which has temporal relations as underlying data structure, and whose operators are all temporal.
-In \cite{Böhlen_point-versus}, a study between interval-based and point-based temporal data models is shown.
-We are going to explain the main definitions and the equivalence between the models.
+	\textbf{Transaction time} \emph{TT}~\cite{Jensen:1991:IIM:627283.627484}: The time when the fact is stored in the database.\\
 
-Temporal relations include a temporal attribute.
-A tuple of a temporal relation may be put under the form $(x_1,...,x_n||t_s)$ where $(x_1,...,x_n)$ are the non-temporal attribute values and \emph{t} is the tuple timestamp.
-Furthermore, an operator is temporal only if it generates a temporal relation when it is applied to temporal relations.
+	\textbf{Valid time} \emph{VT}~\cite{Snodgrass:1984:TQL:588011.588041}: The time when the fact is true in the modelled reality.\\
 
-	The choice of the timestamp representation of the database facts is an important factor.
-Two options are the most common choices: time points and time intervals.
-Time intervals may be built from time points.
-The definition of time point is the following:
-Let T be an infinite set.
+	\textbf{Decision time} \emph{DT}~\cite{Nascimento95}: The time when an event was decided to happen. \\
+\end{svgraybox}
 
-$T^p=\langle T,< \rangle$ is a time point domain over $T$ iff $<$ defines a total order on $T$. Each element of $T$ corresponds to a time point of $T^p$.
+Database models can also be classified into \emph{bi-temporal} (both valid and transaction-time) or \emph{tri-temporal}  (bi-temporal and decision time) models.
 
-The definition of time interval is:
+\subsection{\label{subsubsec:primary-key}Primary key on temporal databases}
 
-A time interval $I$ of $T^p$ is any connected subset of $T^p$.
+\begin{example}
+\label{ex:pk}
+Consider the following example database in \ref{table:example-database}.
+It contains regular data from employees in a company like the name, the birth date.
+Also two temporal values which represent a valid-time interval.
+Consider now that we want to hire Sarah again.
+That is not possible because of the primary key (ID) does not allow to insert again the row Sarah but with a different start and end time.
+In some models this is resolved by adding to the primary key both values Start and End years.
+The resulting database allows to insert a new row where Sarah the year 2010.
+But this modification also allows to insert spurious values e.g.
+inconsistent time periods (see last row in table \ref{table:example-database-with-new-pk}).
 
-$T^i = \langle I,\subset \rangle $ is the time interval domain over $T^p$ iff $I$ is the set of all the intervals of $T^p$.
+%\vspace{-10pt}
 
-Both $T^p$ and $T^i$ are time domains over $T$.
+\begin{table}
+\centering
+\caption{Example database}
+\begin{tabular}{c c c c c c }
+\hline
+\textbf{ID} & \textbf{Name} & \textbf{Age} & \textbf{Works for} & \textbf{Start year} & \textbf{End year} \\ \hline
+1 & Peter & 24 & John &  2010 & - \\
+2 & Maria & 34 & John & 2001 & - \\
+3 & John & 52 & - &  1999 & - \\
+4 & Sarah & 29 & Maria &  2005 & 2009 \\
+\hline 
+\end{tabular}
+\label{table:example-database}
 
-A timestamp over $T^p$ is either a time point or a time interval of $T^p$.
-Time intervals can be represented as sets of time points which may, in turn, be represented by a time interval with the same values at both of its ends.
-Sometimes a coalesce operator is defined so as to switch between a set of time points and a time interval.
-Logically, defining a un-coalescing operator is quite complex.
-\subsubsection{\label{subsubsec:timeDb}Time in databases}
-We can classify the time recorded in the database through its semantics.
-As a result, two kinds of time derive from two orthogonal time domains: valid time and transaction time.
-Valid-time captures the time-varying nature of the reality model.
-Transaction time structures the update activity associated with the object.
-The resulting temporal relations are those emerging from the bitemporal conceptual data model BCDM, on which TSQL2 is based.
+%\vspace{10pt}
 
-The domain of valid time is given as $D_{VT}=\lbrace{c_1^v,c_2^v,...,c_k^v}\rbrace$, whereas the domain of transaction time may be given by $D_{TT}=\lbrace{c_1^t,c_2^t,...,c_k^t}\rbrace$.
-Thus, a valid-time chronon $c^v$  is a member of $D_{VT}$, a transaction-time chronon $c^t$ is a member of $D_{TT}$ and a bitemporal chronon $c^b=(c^t,c^v)$ is an ordered pair of a transaction-time chronon and valid-time chronon.
+\end{table}
 
-We can define the set of names $D_A=\lbrace{A_1,A_2,...,A_n}\rbrace$as attributes and the following set of domains for the said attributes:$D_D=\lbrace{D_1,D_2,...,D_n}\rbrace$.
-We may also define $\bot_i$ as inapplicable,$\bot_u$ as unknown, and $\bot$ as inapplicable or unknown null values.
+%\vspace{-25pt}
 
-The schema for a bitemporal relation $'R'$ consists of an arbitrary number of explicit attributes from $D_A$ with domains in $D_D$ and an implicit timestamp attribute $T$ with domain $2^{\lbrace D_{TT}\bigcup \lbrace UC \rbrace \cdot D_{VT}\rbrace}$, where $UC$ ('until changed') means that the tuple is current in the database.
-This schema is a special transaction-time marker.
+%\vspace{-10pt}
 
-A 'fact' is the information stored in a tuple.
-We define a tuple as $x=(a_1,a_2,...,a_n|t^b)$ in a bitemporal conceptual relation instance $r(R)$, which consists of a number of attribute values associated with a bitemporal timestamp value.
-A value $(UC,c^v)$ in a timestamp indicates that the tuple is current in the database.
+\begin{table}
+\centering
+\caption{Example database with new primary key and some spurious values}
+\begin{tabular}{c c c c c c }
+\hline
+\textbf{ID} & \textbf{Name} & \textbf{Age} & \textbf{Works for} & \textbf{Start year} & \textbf{End year} \\ \hline
+1 & Peter & 24 & John &  2010 & - \\
+2 & Maria & 34 & John & 2001 & - \\
+3 & John & 52 & - &  1999 & - \\
+4 & Sarah & 29 & Maria &  2005 & 2009 \\
+4 & Sarah & 29 & Maria &  2010 & - \\
+4 & Sarah & 29 & Maria &  2007 & 2008 \\
+\hline 
+\end{tabular}
+\label{table:example-database-with-new-pk}
 
-We say that a fact is true in the reality model during each valid-time chronon when a subset of the valid-time domain is associated with that fact.
+%\vspace{10pt}
 
-We say that a fact is current in the relation during each transaction-time chronon when a subset of the transaction-time chronon domain is associated with the valid-time chronon.
-Only the subset of transaction times less than UC may be associated with a valid time.
+\end{table}
 
-While the definition of a bitemporal chronon is symmetric, the relation between valid time and transaction time is asymmetric.
-It should also be noted that since no two tuples with identical attribute values are allowed in a bitemporal relation instance, the full history of a fact is contained in a single tuple.
-In graphical representations of bitemporal space, the x-axis represents the transaction-time dimension and the y-axis denotes the valid-time dimension.
+%\vspace{-25pt}
 
-There are two special cases of bitemporal relations:
-\begin{itemize}
-\item
-Valid-time relations, which support only valid time.
-\item
-Transaction-time relations,which support only transaction time.
-\end{itemize}
-We use the term 'snapshot relation' for a conventional relation. This relation supports neither transaction nor valid time.
-\subsubsection{Implicit versus explicit timestamps}
-\label{sec:implicitversusexplicittimestamps}
-The association between time and facts may be carried out in two ways: explicitly, where the relation is represented by fully explicit timestamps, and implicitly.
-This distinction is relevant to three aspects of the data model: the update language, the display of data and the query language.
+\end{example}
 
-	If the time model works with transaction time, it means that the update language contains implicit timestamps.
-The transaction-time chronons of the facts are supplied by the system; in the valid-time model, these chronons are supplied by the user.
-Thus, in a bitemporal data model, the chronons are also supplied by the user.
+\subsection{\label{subsubsec:consistency}Consistency on temporal databases}
+As it is shown in example \ref{ex:pk}, it  is necessary to define a consistency mechanism to avoid spurious values in the database.
+The most usual solution is to re-define the DML (\emph{Data Manipulation Language}).
+e.g.
+the \emph{update} sentence is redefined as two sentences an update and a create sentence as illustrated in the following example.
 
-	If the time model allows displaying temporal facts, it must represent such data explicitly.
-It is possible that the model allows the user to choose between several representations of time.
-In some simple transaction-time models it is common that the temporal facts cannot be displayed.
+\begin{example}
+Consider the database in table \ref{table:example-database}. The following \emph{update} sentence:
 
-	The main difficulty about differentiating between implicit and explicit timestamps is the query language.
-By means of the query language, we may directly access temporal attributes: the temporal attributes are as explicit as the other attributes in the database.
-If the query language does not allow direct access to temporal attributes, the association schema between timestamps and facts is considered implicit.
+\begin{center}
+\emph{Update Employees set StartYear = 2010 where ID=3}
+\end{center}
 
-\subsubsection{\label{subsubsec:models}Temporal data models}
-There are several extensions to the relational model that include the temporal element.
-In the current section, we briefly describe the model and some of its most interesting features.
-A more detailed summary can be found in \ref{sec:timeindatabasedesign}.
-Some models work only with valid or transaction time, whereas a few models can handle both.
-The current study starts with the models supporting valid time only, continues through the models which support transaction time and finally deals with the bitemporal models.
+Is translated into the following two sentences: an update for the last version of the row and an insert sentence for the new version. The update sentence showed above is translated into:
 
-\paragraph{Valid-time models}
-Most models support valid time only.
+\begin{center}
+\emph{Update Employees set EndYear = 2010 where ID=3}\\
+\emph{Insert into Employees values (3,John,52,-,2010,-)}
+\end{center}
 
-Brooks:  \cite{brooks56} Proposes a tridimensional view of valid-time database.
+%\vspace{-10pt}
 
-Wiederhold: \cite{Blum:1981:DCD:1672611.1672632} Time-oriented database, developed to work with medical applications.
-This model depicts relations as sets of entity-attribute-time-value quadruples.
-Time is associated with the visit number of the patient.
+\begin{table}
+\centering
+\caption{Example database after update sentence}
+\begin{tabular}{c c c c c c }
+\hline
+\textbf{ID} & \textbf{Name} & \textbf{Age} & \textbf{Works for} & \textbf{Start year} & \textbf{End year} \\ \hline
+1 & Peter & 24 & John &  2010 & - \\
+2 & Maria & 34 & John & 2001 & - \\
+3 & John & 52 & - &  1999 & 2010 \\
+3 & John & 52 & - &  2010 & - \\
+4 & Sarah & 29 & Maria &  2005 & 2009 \\
+\hline 
+\end{tabular}
+\label{table:example-database-update}
 
-Jones LEGOL 2.0: \cite{legol20} A language used in legislative writing.
-The temporal order of elements and the valid times for objects are important.
-It was the first time-oriented algebra to be defined, and some of its features are found in subsequent algebras.
-Tuples in this model have two time attributes: Start and Stop. Each one indicates the start point and the end point of the interval where the tuple is valid.
+%\vspace{10pt}
 
-Clifford-1: \cite{Clifford82},\cite{Clifford:1983:FST:319983.319986} Each relation schema has an additional time value called 'State'.
-Also, a boolean attribute is added to indicate if this tuple exists for that state.
+\end{table}
 
-Ariav: \cite{Ariav:1986:TOD:7239.7350} Temporally-oriented data model: a valid-time relation is a set of snapshot relation states indexed by valid time.
-A calculus-based query language is associated with the model: TOSQL.
+%\vspace{-25pt}
 
-Navathe: \cite{TSQL} Proposes a temporal extension to SQL called TSQL, the temporal algebra associated to the temporal relational model.
-This model allows a relationship to handle time-varying attributes, and also relationships without time-varying attributes.
-The only restriction is that all attributes in the relation must be of the same type.
-Objects are classified in snapshot relations and valid-time relations.
-In the latter, each tuple has associated to it an interval of validity with two points: time-start and time-end, like in the LEGOL 2.0 model.
+Table \ref{table:example-database-update} shows the resulting database after the update sentence.
 
-Sadeghi: \cite{913787} Proposes a calculus-based valid-time language, HQL, where all objects represent valid-.time relations.
-As in Navathe and LEGOL 2.0, two time points are required for the ends of the interval.
-The model requires coalescing.
+\end{example}
 
-Sarda: \cite{Sarda:1990:ESH:627277.627409} This model was designed to support the calculus-based model HSQL.
-The model associates valid-time with tuples.
-Objects can either be snapshot or valid-time relations.
-In this model, valid time is considered to be closed on the left part of the interval and open on the right.
-Time is represented by an implicit attribute called 'Period' as a single non-atomic value.
+\subsection{Comercial Temporal database systems}
 
-The following time data models are not in first normal form (1NF), which means they might have multiple values per attribute.
-Even though these models are an extension of the 1NF, the representation is not in 1NF, but the operators work on valid-time relations, which are extensions of conventional relations.
+\begin{table}
+\centering
+\caption{Commercial implementations for temporal database systems. }
+\begin{tabular}{c c c c c c }
+\hline
+\textbf{Name} & \textbf{Time managed} & \textbf{Comments} & \textbf{References}  \\ \hline
+Oracle Workspace Manager & VT and TT. & Package for Oracle DB. & \cite{oracle2009}\\
+TimeDB & VT and TT. & Interface for Oracle DB. & \cite{timedb2005}\\
+Postgree Temporal & VT. & Package for Postgree SQL. & \cite{posgree2009}\\
+Teradata & VT and TT. & Used for data-mining. & \cite{teradata2011}\\
+\hline 
+\end{tabular}
+\label{table:commercial-temporal-db}
 
-Segev: \cite{Segev:1987:LMT:38714.38760} Defines the temporal data model, whose principal time structure is the time sequence.
-A time sequence is a surrogate value that identifies the object along with a sequence of time-value pairs.
-There are several types of time sequences depending on the semantics of the data represented.
+%\vspace{10pt}
 
-Clifford-2: \cite{Clifford:1987:HRD:645472.653241},\cite{Clifford:1985:AHR:971699.318922} This model is a refinement of the previous one presented by the same author.
-The data model allows two types of objects: a set of chronons, termed a lifespan, and a valid-time relation in which each attribute and tuple is assigned a lifespan.
-A tuple is an ordered value containing the tuple value and its lifespan.
-Attributes are not atomic since an attribute value in a tuple is a partial function from the chronons' domain to the value domain of the attribute.
-Relations have key attributes and, at the same chronon, no two tuples are allowed to match the key attributes.
+\end{table}
 
-Tansel: \cite{Tansel:1986:ATD:23125.23132},\cite{Clifford:1985:AHR:971699.318922} Designed to support the calculus-based query language (Hquel) and the time by example language.
-The model supports only one type of object: the valid-time relation.
-Four types of attributes are supported.
-If we take the time into account, attributes can be time-varying or non-time-varying.
-Then, attributes can be atomic-valued or set-valued.
-There is no need for attributes in the relation to be of the same type, and attribute values in a given tuple do not need to be homogeneous.
-A triplet containing an element from the attribute's value domain and the boundary points of the time interval represents the value of a time-varying atomic-valued attribute.
-A set-valued attribute is represented by a set of these triplets.
+\subsubsection{Oracle Workspace Manager}
+Oracle workspace manager \cite{OraE118602} package allows to get several versions of the data in the same database. It is also possible to version only a table.\\
+When a table is versioned, the system creates a few tables and views as well as auxiliary structures to allow keep several version of the data, while keeping the primary key defined by the user.
+A more detailed explanation of
 
-Gadia-1:  \cite{Gadia:1988:HRM:49346.50065},\cite{Gadia:1985:QLH:325405.325412} This homogeneous model allows two types of objects: valid-time elements and valid-time relations.
-The model requires that all attribute values in a given tuple be functions on the same valid-time element.
-Unlike intervals, valid-time elements are closed under union, difference, and complementation.
-Bhargava presents an extension of this model to both valid and transaction time.
-
-Gadia-2: \cite{Gadia:1986:TMM:645471.655410},\cite{Chuen},\cite{Gadia:1988:GMR:971701.50233} Is an extension of the homogeneous model known as multihomogeneous model.
-Temporal elements may be multi-dimensional to model both valid and transaction time.
-Attribute values are functions from temporal elements onto attribute value domains, but attribute values do not need to be on the same temporal element.
-The key attributes in the relations must be single-valued in respect to the interval of validity.
-
-Lorentzos: \cite{springerlink:10.1007/3-540-19074-0_71},\cite{Lorentzos:1989:HDT:70777.70787} The temporal relational model allows to specify different granularities and the support of periodic events.
-This model associates timestamps with individual attribute values rather than with tuples.
-Timestamps are explicit values updated directly by the user.
-The said timestamps represent either the chronon during which attribute(s) were valid or a boundary point of the interval of validity.
-In the model, several timestamps of different granularities can be used together in the specification of a chronon.
-This model has the particularity that some columns have a different meaning depending on the context.
-
-\paragraph{Transaction-time models}
-The main feature of the transaction-time models is that they are append-only.
-
-Kimball: \cite{kim78} The temporal model is called DATA.
-In it, the association between facts and times is implicit.
-Also, the update operations avoid the explicit use of time.
-However, transaction-time relations cannot be displayed but can only be depicted in the snapshots extracted from the transaction-time relations.
-At query language level, the association of facts and times is also implicit.
-
-Stonebraker: \cite{Ston87} Proposes the Postgres data model.
-Like in the previous model, the association between time and facts in the three model features (update language, query language and visualization) is implicit.
-In this model, the visualization is not restricted to snapshot relation states.
-In this model, two timestamps are used when specifying the time for a given tuple that is current in the relation.
-
-Jensen: \cite{Jensen:1991:IIM:627283.627484} In the DM/T model, the association of facts and times is also implicit.
-DM/T contains a backlog for each user-defined transaction-time relation.
-This backlog contains the full history of the associated user-defined transaction-time relation.
-When an insert, delete or update operation is performed, the backlog for this relation is updated.
-
-\paragraph{Bitemporal data models}
-The bitemporal data models support both valid and transaction time. This section deals with the following seven models of a bitemporal data model.
-
-Ben-Zvi: \cite{910377} Proposed the time relational model,, which defines two types of objects: snapshot relations and bitemporal relations.
-Each tuple in a bitemporal relation has five additional implicit attributes:
-The two interval ends of the valid-time chronon, called 'Effective-time-start' and 'Effective-time-stop'.
-The two interval ends of the transaction-time chronon.
-'Registration-time-start' is the transaction-time chronon associated with the 'Effective-time-start', 'Registration-time-stop' is the transaction-time chronon associated with the 'Effective-time-stop'.
-Deletion-time. It registers when a tuple which was mistakenly entered is logically deleted.
-
-Ahn: \cite{Snodgrass:1985:TTD:318898.318921},\cite{sno86} This model follows a four-dimensional data model.
-Each instance is represented as a sequence where each value is composed by a timestamp with a transaction-time chronon and a three dimensional volumes -one of these dimensions being the valid time.
-
-Snodgrass:  \cite{sno86},\cite{Snodgrass:1984:TQL:588011.588041} The data model has four implicit attributes: 
-The transaction time when the tuple was inserted.
-The transaction time when the tuple was deleted.
-The starting point of the valid-time interval.
-The end point of the valid time-interval.
-
-	The data query language associated with the model is called TQuel.
-
-McKenzie: \cite{Snodgrass:1993:ATQ:642875.642890},\cite{Mckenzie:1988:ALQ:915060}, \cite{mck81} In this model, the attribute's values are timestamped, but it is required that attributes be single-valued.
-This makes it possible to perform a Cartesian product.
-There are two types of objects in the model: the snapshot relations and the valid-time relations.
-Thus, transaction-time relations are modeled as sequences of snapshot relations.
-Bitemporal relations are sequences of valid-time relations indexed by transaction time.
-An attribute's value in a valid-time relation is an ordered pair: a value from the attribute's domain and a set of chronons.
-Relations are not allowed to have value-equivalent tuples.
-The model also allows the schema and even the class of the relation (snapshot, transaction-time, valid-time and bitemporal) to change.
-
-Gadia-3: \cite{Bhargava:1993:RDS:642811.642819},\cite{gad92} This data model is associated with the calculus-based query language TempSQL.
-Attributes are timestamped with finite unions of rectangles in the valid-time x transaction-time space.
-
-Thompson: \cite{171833} This model, based on an accounting system, is called Accounting Data Model.
-It has four timestamp attributes: two for transaction time, called 'engineering time', and two for valid time, called 'accounting time'.
-Also, a boolean value called 'timewarp' enables the attribute to change historically.
+A workspace is a logical group of a set of changes (versions of tuples) and allows consistent access, thus the user always obtain the correct data version.
+Workspaces are ordered in a hierarchy.
+The top level is called LIVE workspace.
+This is used by default and it must be used in production environment.\\
+The system only makes a copy of a tuple if it is changed.
+In order to access different versions of the tuple, the context must be changed.
+Also, it is not necessarily to modify the SQL code to access to versioned tables.\\
+When a table is versioned, it is renamed as table\_LT.
+Here is stored all the data for the information together metadata for the versioning.
+An auxiliary table is created with the workspace metadata with the name table\_AUX.
+A view is created with the name of the original table to allow querying without versioning.

### 2012-01-12 book/sources/fuzzyTemporalDatabases.tex
update

%
% Fuzzy temporal databases
%
Some software applications deal with uncertainty related to time. e.g. a logistics company which transports packages from one destination to another. The time when the package would be at the end point may be stimated but not precisely known. These applications need to manage the uncertainty with respect to the temporal attributes of the objects stored in the database.

There are several approaches to manage uncertainty in the temporal information stored in a database. Valid-time is usually represented as an interval, and therefore it is suitable for the representation of uncertainty in the valid-time interval. Nevertheless, the values for transaction and decision time are usually precisely known.
Some models deal with probability~\cite{Dekhtyar2001} or possibility~\cite{Dubois89} distributions at the endpoints of the time interval. In this section we will explain the most interesting proposals that deal with possibility distributions in the valid-time interval. %Two are the main points developed within this section: the representation of uncertainty in the time interval and the flexible querying of ill-known time periods.

In fuzzy databases~\cite{Galindo2006}, uncertainty is managed at storage level whereas imprecision is managed at querying level. In subsection \ref{subsec:representation-time-intervals} we will explain the main approach for the storage of uncertainty in the valid-time interval. Finally, in subsection \ref{subsec:querying-time-intervals} the proposals for querying in a flexible way valid-time intervals are analyzed.

\subsection{Representation of ill-known temporal intervals}
\label{subsec:representation-time-intervals}
The uncertainty in a valid-time interval may be represented with uncertainty at one or both starting and ending points. The uncertainty is usually represented by means of a possibility distribution in one or both points. It is also possible to represent with rough sets~\cite{Pawlak1995} time intervals~\cite{Qia09}.

Several authors ~\cite{garrido2009} propose some transformations in order to optimize the storage of the valid-time interval. Further research has shown how these transformations drive to a possibilistic information lost~\cite{Pon11}. A comparison between the transformations in ~\cite{garrido2009}  and the framework in~\cite{Pon11} is done in~\cite{pon12}.

\begin{figure}
\centering
\input{graphs/interval.tex}
\caption{Two ill-known values $X$ and $Y$, represented by means of triangular possibility distributions denoting the starting and ending points of an ill-known interval. }
\label{fig:interval}
\end{figure}

An ill-known temporal interval is usually represented by means of two possibility distributions one in each starting and ending points. Figure \ref{fig:interval} shows an ill-known valid-time interval. The starting and ending points are denoted by two fuzzy numbers, $X$ and $Y$, represented by means of two triangular possibility distributions (see the Appendix). For illustration, figure \ref{fig:convexhull} shows one of the transformations based in the convex hull and proposed in~\cite{garrido2009}.

\begin{figure}
\centering
\input{graphs/convexhull.tex}
\caption{Transformation based in the convex hull from the two ill-known points $X$ and $Y$. }
\label{fig:convexhull}
\end{figure}

\subsection{Querying fuzzy temporal databases}
\label{subsec:querying-time-intervals}

\subsubsection{Bipolarity in temporal database querying}

### 2012-01-17 trunk/book/sources/fuzzylogic.tex
A new version with extended definition of granularity and calendar.

Possibility theory, like probability theory, deals with uncertainty about the outcome of an experiment. In probability theory, this uncertainty is caused by the \emph{variability} in the outcomes, while in possibility theory, the uncertainty is caused by \emph{incomplete knowledge} about the experiment. The quantification of confidence in a theory of uncertainty is achieved using a confidence measure\cite{Pon11}. In probability theory this is a measure of chance, in possibility theory, possibility and necessity measures are used.

\begin{definition}
Consider a set of outcomes $\Omega$. Let $\wp(\Omega)$ denote the powerset of $\Omega$ and let $A$ and $B$ be elements of $\wp(\Omega)$. A \emph{confidence measure on $\Omega$} is defined by a function
	\begin{align}
	g : \wp(\Omega) & \rightarrow \left[0,1\right]
	\end{align}
that satisfies
	\begin{align}
	g(\emptyset) &= 0 \\
	g(\Omega) &= 1 	\label{NormalizationProperty} \\
	A \subseteq B &\Rightarrow g(A) \leq g(B) \label{MonotonicityProperty}
	\end{align}
\end{definition}

Both possibility measures and necessity measures are special cases of confidence measures.

\begin{definition}
Consider a confidence measure $\Pi$ on a set of outcomes $\Omega$. Let $J$ be a countable index set and let $\{ A_{j} | j \in J \wedge A_{j} \subseteq \Omega \}$ be a family of elements of $\wp(\Omega)$. $\Pi$ is now a \emph{possibility measure on $\Omega$} if it satisfies:
	\begin{align}
	\Pi\left(\bigcup_{j \in J} A_{j} \right) = \sup_{j \in J} \Pi(A_{j})
	\end{align}
\end{definition}

In this work, the interpretation is as follows. The possibility of an event expresses how plausible the occurrence of the event seems to an observer of the experiment, given the (partial) knowledge of the observer about the experiment.

Information on the possibility of distinct elements of the universe of discourse $\Omega$ can now be given by a \emph{possibility distribution} $\pi$ on $\Omega$, defined by:

\begin{definition}
Consider a possibility measure $\Pi$ on $\Omega$. A \emph{possibility distribution} $\pi$ on $\Omega$ underlying the possibility measure $\Pi$ is then a function defined by:
	\begin{align}
	\pi : \Omega \rightarrow \left[0, 1\right] : \pi(u) = \Pi(\{u\})
	\end{align}
\end{definition}

\begin{definition}
Consider a confidence measure $N$ on a set of outcomes $\Omega$. Let $J$ be a countable index set and let $\{ A_{j} | j \in J \wedge A_{j} \subseteq \Omega \}$ be a family of elements of $\wp(\Omega)$. $N$ is now a \emph{necessity measure} on $\Omega$ if it satisfies:
	\begin{align}
	N\left(\bigcap_{j \in J} A_{j} \right) = \inf_{j \in J} N(A_{j})
	\end{align}
\end{definition}

In this work, the interpretation is as follows. The necessity of an event expresses how necessary the occurrence of the event seems to an observer of the experiment, given the (partial) knowledge of the observer about the experiment.

Possibility and necessity measures are dual in the sense that:

\begin{align}
\forall A \subseteq \Omega : N(A) = 1 - \Pi(\bar{A})
\end{align}

Regarding interpretation, the above can be seen as: the degree to which an event is necessary is the degree to which every other possible event is not plausible.

\subsection*{\label{subsec:fuzzy-numbers}Fuzzy numbers and fuzzy intervals}
Dubois and Prade~\cite{Dubois1983} proposed the following definition of \emph{fuzzy interval}.
%\begin{definition}
A fuzzy interval is a fuzzy set $M$ on the set of real numbers $\mathbb{R}$ such that:
\begin{eqnarray}
\forall (u,v)\in\mathbb{R}^2:&\\
\nonumber
\forall w \in [u,v]:&\mu_M(w) \geq\min(\mu_M(u),\mu_M(v))  \\
\exists m \in \mathbb{R}:& \mu_M(m)=1 
\end{eqnarray}
%\end{definition}
If $m$ is unique, then $M$ is referred to as a \emph{fuzzy number}, instead of a \emph{fuzzy interval}. In other words, if the core of a fuzzy interval is a singleton, it can be seen as a fuzzy number. In their work, Dubois and Prade propose four different functions (two possibility and two necessity functions) to asses the position of a fuzzy number N relative to  a fuzzy number M taken as a reference.

The most convenient representation for the membership function of a fuzzy number is a triangular function (fig. \ref{fig:triangular}). The membership function $\mu_M$ for the fuzzy set $M$ has also the properties of convexity and normalization. Three values represent a triangular function: 
\begin{itemize}
\item
$D$ is the singleton value in the core of $\mu_M(x)$.
\item
$D-a$ is the lower value in the support of $\mu_M(x)$. 
\item
$D+b$ is the upper value in the support of $\mu_M(x)$.
\end{itemize}

\begin{figure}
\centering
\input{graphs/triangular.tex}
\caption{Triangular possibility distribution. }
\label{fig:triangular}
\end{figure}

Note that the values $a$ and $b$ are values in the underlying ordered domain. E.g. $(a,b) \in \mathbb{R}^2$. The notation for a triangular function adopted here is $[D,a,b]$.

The most simple representation for the membership function of a fuzzy interval is a trapezoidal function. This membership function $\mu_T$ for the fuzzy interval $T$ is convex and normalized. Four values represent a trapezoidal membership function (figure  \ref{fig:trapezoidal}):
 $\left[\alpha,\ \beta,\ \gamma,\ \delta\right]$. The membership function is:

\begin{align}
\mu_T(x)
\begin{cases}
1 & \mbox{ if } x \in [\beta,\gamma] \\
0 & \mbox{ if } x > \delta \vee x < \alpha \\
\frac{x-\alpha}{\beta - \alpha} & \mbox{ if } x \in [\alpha,\beta[ \\
\frac{\delta -x}{\delta - \gamma} & \mbox{ if } x \in ]\gamma,\delta] \\
\end{cases}
\end{align}

\begin{figure}
\centering
\input{graphs/trapezoidalDistribution.tex}
\caption{Trapezoidal possibility distribution $\left[\alpha,\ \beta,\ \gamma,\ \delta\right]$. }
\label{fig:trapezoidal}
\end{figure}

### 2012-01-17 book/sources/timeDomain.tex
A new version with extended definition of granularity and calendar.

@@
 \end{svgraybox}
+
+\begin{definition}
+\label{def:set-chronon}
+The \textbf{\emph{set of all the chronons}} in a system is denoted by $\C$. A chronon $c$ is therefore defined as:
+\begin{equation}
+\left \lbrace c \mid c \in \C \right \rbrace
+\end{equation}
+\end{definition}
+
 Any temporal model may represent the time as points as well as intervals. The equivalence between a interval-based and point-based temporal data models is demonstrated in \cite{Böhlen_point-versus}.
 
-Attending to the density between the chronons, a model can be classified into the following three types:
+Attending to the density in the set $\C$ of the chronons, a model can be classified into the following three types:
 
@@
 \item
-\textbf{Discrete model}:  The discrete model of time is isomorphic in relation to the natural numbers. According to this model, each natural number corresponds to a \emph{chronon}. 
+\textbf{Discrete model} ($\C \subset \Nat$):  The discrete model of time is isomorphic in relation to the natural numbers. According to this model, each natural number corresponds to a \emph{chronon}.
 \item
-\textbf{Dense model}: This model is isomorphic in relation to the rational or the real numbers. It represents one instant of time in the gap between another two instants.
+\textbf{Dense model} ($\C \subset \Q$ or $\C \subset \R$): This model is isomorphic in relation to the rational or the real numbers.
+It represents one instant of time in the gap between another two instants.
 \item
-\textbf{Continuous model}: This is an extension of the dense model with no gaps. It is isomorphic in relation to the real numbers.
+\textbf{Continuous model} ($\C \subset \R$): This is an extension of the dense model with no gaps. It is isomorphic in relation to the real numbers. 
 \end{itemize}
@@
 %%% granularity
-%\subsubsection{Granularity}
+\subsection{\label{subsec:granularity}Granularity}
 Next to the inherent difficulties of time, an issue is \textbf{temporal granularity}.
@@
 In this example any time period shorter than a year needs no representation and thus the used granularity allows no specification for time periods shorter than a year.
+The formal definition for granularity is given in ~\cite{Lin97}:
+
+\begin{definition}
+\label{def:granularity}
+A \textbf{\emph{granularity}} $\alpha$ is an ordered set of non-overlapping and continuous time elements called granules. \\
+\end{definition}
+
+\begin{definition}
+\label{def:granule}
+A \textbf{\emph{granule}} is the basic time unit represented by a granularity. E.g. in the hours granularity, an hour is a granule.
+As granularity $\alpha$ it is an ordered set, each \emph{granule} may be indexed by an integer number:\\
+\begin{equation}
+\alpha = \left \lbrace \ldots, 0_\alpha, 1_\alpha, \ldots \right \rbrace
+\end{equation}
+\end{definition}
+
+In a system, the smallest granularity is the \emph{chronons} granularity, denoted with the index $\bot$.
+It is possible to define a function $f$ that maps from  a given granule $i_\alpha$ (in the given granularity $\alpha$) to the set of the corresponding chronons.
+
+\begin{definition}
+\label{def:mapping-function}
+The \textbf{\emph{mapping function}} $f$ in a given granule $i_\alpha$ is denoted by:\\
+%\begin{align}
+$f:i_\alpha \rightarrow \C$\\
+$f(i_\alpha) = \lbrace c_\bot \mid c_\bot$ is in $i_\alpha \rbrace \wedge c_\bot \in \C$\\
+%\end{align}
+\end{definition}
+
+If $\alpha$ is a granularity, then the following properties must hold:\\
+\begin{itemize}
+\item
+$\alpha$ is an ordered set: if $c_\bot \in f(i_\alpha)$ and $c^{'}_\bot \in f(j_\alpha)$ and $c_\bot < c^{'}_\bot$ implies that $i_\alpha < j_\alpha$.
+\begin{figure}
+\centering
+\input{graphs/granularity-prop1.tex}
+\caption{$\alpha$ is an ordered set}
+\label{fig:granularity-prop1}
+\end{figure}
+\item
+$\alpha$ is a set of continuous granules.
+Consider the chronons $c_\bot < c^{'}_\bot < c^{''}_\bot$.
+Then $c_\bot \in f(i_\alpha)$,$c^{''}_\bot \in f(i_\alpha)$ implies that $c^{'}_\bot \in f(i_\alpha)$.
+\begin{figure}
+\centering
+\input{graphs/granularity-prop2.tex}
+\caption{$\alpha$ is a set of continuous granules}
+\label{fig:granularity-prop2}
+\end{figure}
+\item
+The granules in $\alpha$ do not overlap: If $i_\alpha \neq j_\alpha$ then $f(i_\alpha) \cap f(j_\alpha) = \emptyset$.
+\begin{figure}
+\centering
+\input{graphs/granularity-prop3.tex}
+\caption{The granules in $\alpha$ do not overlap}
+\label{fig:granularity-prop3}
+\end{figure}
+\end{itemize}
+
+The mapping function $f$ allows the comparison among granularities:
+
+\begin{definition}
+\label{def:finner-than}
+A granularity $\alpha$ is \emph{``finner than''} granularity $\beta$ when the set of corresponding chronons for a given granule $i_\alpha$ is smaller than the set of the corresponding chronons for a given granule $j_\beta$.
+\begin{equation}
+\label{eq:finner-than}
+\mid f \left( i_\alpha \right) \mid < \mid  f \left( j_\beta \right) \mid
+\end{equation} 
+\end{definition}
+
+\begin{definition}
+\label{def:coarser-than}
+A granularity $\alpha$ is \emph{``coarser than''} granularity $\beta$ when the set of corresponding chronons for a given granule $i_\alpha$ is greater than the set of the corresponding chronons for a given granule $j_\beta$.
+\begin{equation}
+\label{eq:coarser-than}
+\mid f \left( i_\alpha \right) \mid > \mid  f \left( j_\beta \right) \mid
+\end{equation} 
+\end{definition}
+
+A function,\emph{Cast} may be specified to map between two granularities $\alpha$ and $\beta$. The specification for that function is:
+\begin{equation}
+\label{eq:cast-function}
+\mbox{Cast} \left( i_\alpha,\alpha,\beta \right) = j_\beta
+\end{equation}
+This function associates a granule $i_\alpha$ in the granularity $\alpha$ to a corresponding granule $j_\beta$ in the corresponding granularity $\beta$.
+An \emph{upwards mapping} is the mapping from a granularity $\alpha$ to a coarser granularity $\beta$, whereas a downwards mapping is the mapping from a granularity $\gamma$ to a finner granularity $\delta$.
+Figure \ref{fig:granularity-graph-example} shows a granularity graph example with the Gregorian Calendar and the academic days as possible granularities.
+
+\begin{figure}
+\centering
+\includegraphics[scale=0.5]{graphs/granularityGraph.pdf}
+\caption{A granularity graph example}
+\label{fig:granularity-graph-example}
+\end{figure}
+
+The mapping between two granularities may be classified into the following types and are explained within the figure \ref{fig:granularity-graph-example}:
+\begin{itemize}
+\item
+Regular mapping: When the mapping function between two granularities is calculated by means of multiplications or divisions and maybe some anchor adjustment.
+E.g.
+the mapping between hours and minutes.
+This mapping is represented in the figure with a normal arrow.
+\item
+Irregular mapping: When the mapping function can not be calculated by means of multiplications or divisions.
+E.g.
+the mapping between months and days.
+(This mapping is represented in the figure with a bold arrow.
+\item
+Congruent mapping: Two granularities that have the same granules but with different anchos.
+E.g.
+the mapping between the Gregorian Calendar and academic days.
+This mapping is represented in the figure with a straight line without arrow.
+\end{itemize}
 
@@
 
-\subsection{Temporal Relations}
-%an explanation on allen's temporal relations.
-%definition of fuzzy temporal interval
+\subsection{\label{subsec:timedomain-calendar}Time domain: Calendars}
+%maybe introduce previously the concept of time units, linear hierarchy and time point.
+A Calendar is a organization of different granularities (e.g. day, month, year) in a hierarchy. 
+In order to define a calendar, some previous concepts should be introduced~\cite{Kraus1997}:
 
-Several operators are defined in order to compare temporal elements.
-Allen~\cite{Allen83} first described the relations between time intervals and to a lesser extent, between time points.
-Figure \ref{fig:allen} shows the temporal relations.
-There are some proposals that can be applied to both crisp and fuzzy temporal interval ~\cite{ohlbach2004},\cite{nagypal2003},\cite{schockaert08}.
-For this fuzzy comparisons~\cite{garrido2009} proposed some different temporal relations that can be easyly implemented by means of regular fuzzy operations.
+\begin{definition}
+\label{def:granularity-hierarchy}
+\textbf{(Linear hierarchy of granularities)} A linear hierarchy of granularities is a finite collection of granularities with a linear order $\sqsubset$ among these granularities.
+The coarser granularity is called the top of the hierarchy and it must be an infinite set while all the others granularities must be finite sets.
+\end{definition}
 
-\begin{figure}
-\centering
-\input{graphs/allen.tex}
-\caption{Allen relations between two time intervals.}
-\label{fig:allen}
-\end{figure}
+The hierarchy of days $\sqsubset$ months $\sqsubset$ years defines a linear hierarchy.
 
-\subsection{An example of temporal domain: Julian Day Number}
+\begin{definition}
+\label{def:linear-calendar}
+\textbf{(Linear Calendar)} A linear calendar consist of a linear hierarchy $H$ of granularities and a validity predicate which specifies if a granule is valid in  the calendar.
+\end{definition}
+
+The linear hierarchy days $\sqsubset$ months $\sqsubset$ years represents a linear calendar.
+It is necessary to define the validity predicates because 17 January 2012 is a valid element of the calendar whereas 30 February 2012 is not.
+
+\subsubsection{\label{subsubsubsec:julian-day-number}An example of temporal domain: Julian Day Number}
 % proposal for the underlying Julian Day Number domain.
@@
 
-%\subsubsection{Conversion among dates}
-%\paragraph{Conversion between Gregorian and Julian dates}
-%The Julian Date consist on the Julian Day number and the time. It is computed with the formula \cite{Usn}:\\
-%\begin{equation}
-%\label{eq:julian day}
-%JD = JDN + \frac{hour-12}{24} + \frac{minutes}{1440} + \frac{seconds}{86400}
-%\end{equation}
-%
-%Here the JDN is the Julian Day Number.
-This equation does not take into account the leap seconds \footnote{A leap second(s) is the amount of second(s) added or deleted every year in the Universal Time.\cite{Wik},\cite{Lea}.
-}.
+\subsection{Temporal Relations}
+%an explanation on allen's temporal relations.
+%definition of fuzzy temporal interval
 
-%\paragraph{Gregorian calendar to Julian day}
-%Any date from Gregorian calendar \footnote{As Gregorian calendar entered in use in 1582, any use before that date is called \emph{proleptic calendar}.} can be converted into a Julian Day using the following formula:\\
-%\begin{equation}
-%\label{eq:JDN}
-%JDN = \frac{1461x(Y + 4800 + \frac{M-14}{12})}{4}+\frac{367x(M-2-12x\frac{M-14}{12})}{12}-\frac{3x\frac{Y+4900+\frac{M-14}{12}}{100}}{4} 
-%\end{equation}
-%
-%This formula was created by Henry Fliegel and Thomas C.
-van Flandern in 1968.
-\cite{Fliegel:1968:LEM:364096.364097} It takes into account leap years and that they are applicable only when the century is divisible by 400.
-The value $Y$ represents the year and the value $M$ represents the month.
-%\paragraph{Julian calendar to Julian day}
-%The algorithm to convert between Julian calendar and Julian Day Number is only valid for dates since 4713 B.C (positive values of Julian Day) \cite{Dog}:\\
-%\begin{equation}
-%\label{eq:JDN_2_JD}
-%JDN=367Y-\frac{7 (Y + 5001 + \frac{M-9}{7})}{4}+\frac{275xM}{9}+D+1729777
-%\end{equation}
-%
-%Where $Y$ is the year value, $M$ is the month value, and $D$ is the day value.
-%
-%%\subsubsection{Membership degree computation}
-%The main advantage from the Julian day number is the ability to make calculations. Thus, the membership degree computation for a fuzzy number based on a Julian day is explained here. \\
-%Let a trapezoidal possibility distribution defined by four JDN named $a$,$b$,$c$ and $d$.
-The fuzzy validity period \emph{FVP}, can be represented as in figure \ref{fig:fuzzy-validity-period}.
-The following equation shows the membership degree for a Julian day number \emph{JDN} in a \emph{FVP} represented by means of the possibility distribution of $[a,b,c,d]$.
-In a valid-time database, the values for the valid starting time \emph{VST} and the valid ending time \emph{VET} can be fuzzyfied as shown in equation \ref{eq:fuzzy-validity-period}.
-%
-%The membership function for a given $x$ number in \emph{JDN} format is the following:\\
-%\begin{equation}
-%\label{eq:trapezoidal distribution}
-%%\[
-%T(x)=
-%\begin{cases}
-%0, \mbox(if x \leq a, x \geq d)\\
-%1, \mbox(if x \in [b,c])\\
-%\frac{x-a}{b-a}   \mbox(if x \in (a,b])\\
-%\frac{d-x}{d-c} 	 \mbox(if x \in (c,d])\\
-%\end{cases}
-%%\]
-%\end{equation}
-%The points outside the trapezium or inside the interval $[a,b]$ are trivial. The points inside the other intervals has the following particularities:\\
-%The subtraction operation with Julian day numbers gives as result a scalar value (as explained before).
-The division operation between two scalars gives as result an scalar in the interval $[0,1]$ which represents the membership degree.
+Several operators are defined in order to compare temporal elements.
+Allen~\cite{Allen83} first described the relations between time intervals and to a lesser extent, between time points.
+Figure \ref{fig:allen} shows the temporal relations.
+There are some proposals that can be applied to both crisp and fuzzy temporal interval ~\cite{ohlbach2004},\cite{nagypal2003},\cite{schockaert08}.
+For this fuzzy comparisons~\cite{garrido2009} proposed some different temporal relations that can be easyly implemented by means of regular fuzzy operations.
+
+\begin{figure}
+\centering
+\input{graphs/allen.tex}
+\caption{Allen relations between two time intervals.}
+\label{fig:allen}
+\end{figure}

### 2012-01-18 trunk/book/sources/intro.tex
update in the bibliography. fix for the frame  in the allen relations graph. new organization in timeDomain.tex fix for the bibliography in iintro.tex

@@
 
-The ``Consensus Glossary of Temporal Database concepts'' \cite{Dyreson:1994:CGT:181550.181560} is the first publication where the main researchers in temporal databases define the main thesaurus of temporal database concepts.
+The ``Consensus Glossary of Temporal Database concepts'' \cite{Dyreson1994} is the first publication where the main researchers in temporal databases define the main thesaurus of temporal database concepts.
 This glossary defines the main types of time in a temporal database and is the basis of furthers researches ~\cite{Sarda:1990:ESH:627277.627409} \cite{Jensen94thetsql2}.

### 2012-01-18 trunk/book/sources/timeDomain.tex
update in the bibliography. fix for the frame  in the allen relations graph. new organization in timeDomain.tex fix for the bibliography in iintro.tex

@@
 In the following subsection \ref{subsec:basic-concepts} we will introduce several concepts that are widely used in the community of temporal databases.
-Most of the concepts explained here have been introduced in the \emph{'Consensus Glossary of Temporal Database Concepts'}~\cite{Dyreson:1994:CGT:181550.181560}.
+Most of the concepts explained here have been introduced in the \emph{'Consensus Glossary of Temporal Database Concepts'}~\cite{Dyreson1994}.
 
@@
 
-Attending to the density between the chronons, a model can be classified into the following three types:
+\begin{definition}\textbf{(set of chronons)}
+\label{def:set-chronon}
+The \textbf{\emph{set of all the chronons}} in a system is denoted by $\C$. A chronon $c$ is therefore defined as:
+\begin{equation}
+\left \lbrace c \mid c \in \C \right \rbrace
+\end{equation}
+\end{definition}
+
+Any temporal model may represent the time as points as well as intervals. The equivalence between a interval-based and point-based temporal data models is demonstrated in \cite{Böhlen_point-versus}.
+
+Attending to the density in the set $\C$ of the chronons, a model can be classified into the following three types:
 
@@
 
-%%% time as points and time as intervals:
+%%% granularity
+\subsection{\label{subsec:granularity}Granularity}
+Next to the inherent difficulties of time, an issue is \textbf{temporal granularity}.
+A temporal granularity is a partitioning of the time line used by a system, usually dependent on the application.
+For example, the age of an adult human being is usually expressed in years: one will use sentences like \emph{`Laura is 21 years old'} instead of \emph{`Laura is 21 years, 3 months and 4 days old'}.
+In this example any time period shorter than a year needs no representation and thus the used granularity allows no specification for time periods shorter than a year.
+The formal definition for granularity is given in ~\cite{Lin97}:
 
-%%% granularity
+\begin{definition}
+\label{def:granularity}\textbf{(granularity)}
+A \textbf{\emph{granularity}} $\alpha$ is an ordered set of non-overlapping and continuous time elements called granules. \\
+\end{definition}
 
-\subsection{Temporal Relations}
-%an explanation on allen's temporal relations.
+\begin{definition}
+\label{def:granule}\textbf{(granule)}
+A \textbf{\emph{granule}} is the basic time unit represented by a granularity. E.g. in the hours granularity, an hour is a granule.
+As granularity $\alpha$ it is an ordered set, each \emph{granule} may be indexed by an integer number:\\
+\begin{equation}
+\alpha = \left \lbrace \ldots, 0_\alpha, 1_\alpha, \ldots \right \rbrace
+\end{equation}
+\end{definition}
 
-\subsection{An example of time domain: Julian Day Number}
+In a system, the smallest granularity is the \emph{chronons} granularity, denoted with the index $\bot$.
+It is possible to define a function $f$ that maps from  a given granule $i_\alpha$ (in the given granularity $\alpha$) to the set of the corresponding chronons.
+
+\begin{definition}
+\label{def:mapping-function}
+\textbf{(mapping function)}
+The \textbf{\emph{mapping function}} $f$ in a given granule $i_\alpha$ is denoted by:\\
+%\begin{align}
+$f:i_\alpha \rightarrow \C$\\
+$f(i_\alpha) = \lbrace c_\bot \mid c_\bot$ is in $i_\alpha \rbrace \wedge c_\bot \in \C$\\
+%\end{align}
+\end{definition}
+
+If $\alpha$ is a granularity, then the following properties must hold:\\
+\begin{itemize}
+\item
+$\alpha$ is an ordered set: if $c_\bot \in f(i_\alpha)$ and $c^{'}_\bot \in f(j_\alpha)$ and $c_\bot < c^{'}_\bot$ implies that $i_\alpha < j_\alpha$.
+\begin{figure}
+\centering
+\input{graphs/granularity-prop1.tex}
+\caption{$\alpha$ is an ordered set}
+\label{fig:granularity-prop1}
+\end{figure}
+\item
+$\alpha$ is a set of continuous granules.
+Consider the chronons $c_\bot < c^{'}_\bot < c^{''}_\bot$.
+Then $c_\bot \in f(i_\alpha)$,$c^{''}_\bot \in f(i_\alpha)$ implies that $c^{'}_\bot \in f(i_\alpha)$.
+\begin{figure}
+\centering
+\input{graphs/granularity-prop2.tex}
+\caption{$\alpha$ is a set of continuous granules}
+\label{fig:granularity-prop2}
+\end{figure}
+\item
+The granules in $\alpha$ do not overlap: If $i_\alpha \neq j_\alpha$ then $f(i_\alpha) \cap f(j_\alpha) = \emptyset$.
+\begin{figure}
+\centering
+\input{graphs/granularity-prop3.tex}
+\caption{The granules in $\alpha$ do not overlap}
+\label{fig:granularity-prop3}
+\end{figure}
+\end{itemize}
+
+The mapping function $f$ allows the comparison among granularities:
+
+\begin{definition}
+\label{def:finner-than}\textbf{(finner than)}
+A granularity $\alpha$ is \emph{``finner than''} granularity $\beta$ when the set of corresponding chronons for a given granule $i_\alpha$ is smaller than the set of the corresponding chronons for a given granule $j_\beta$.
+\begin{equation}
+\label{eq:finner-than}
+\mid f \left( i_\alpha \right) \mid < \mid  f \left( j_\beta \right) \mid
+\end{equation} 
+\end{definition}
+
+\begin{definition}
+\label{def:coarser-than}\textbf{(coarser than)}
+A granularity $\alpha$ is \emph{``coarser than''} granularity $\beta$ when the set of corresponding chronons for a given granule $i_\alpha$ is greater than the set of the corresponding chronons for a given granule $j_\beta$.
+\begin{equation}
+\label{eq:coarser-than}
+\mid f \left( i_\alpha \right) \mid > \mid  f \left( j_\beta \right) \mid
+\end{equation} 
+\end{definition}
+
+A function,\emph{Cast} may be specified to map between two granularities $\alpha$ and $\beta$. The specification for that function is:
+\begin{equation}
+\label{eq:cast-function}
+\mbox{Cast} \left( i_\alpha,\alpha,\beta \right) = j_\beta
+\end{equation}
+This function associates a granule $i_\alpha$ in the granularity $\alpha$ to a corresponding granule $j_\beta$ in the corresponding granularity $\beta$.
+An \emph{upwards mapping} is the mapping from a granularity $\alpha$ to a coarser granularity $\beta$, whereas a downwards mapping is the mapping from a granularity $\gamma$ to a finner granularity $\delta$.
+Figure \ref{fig:granularity-graph-example} shows a granularity graph example with the Gregorian Calendar and the academic days as possible granularities.
+
+\begin{figure}
+\centering
+\includegraphics[scale=0.5]{graphs/granularityGraph.pdf}
+\caption{A granularity graph example}
+\label{fig:granularity-graph-example}
+\end{figure}
+
+The mapping between two granularities may be classified into the following types and are explained within the figure \ref{fig:granularity-graph-example}:
+\begin{itemize}
+\item
+Regular mapping: When the mapping function between two granularities is calculated by means of multiplications or divisions and maybe some anchor adjustment.
+E.g.
+the mapping between hours and minutes.
+This mapping is represented in the figure with a normal arrow.
+\item
+Irregular mapping: When the mapping function can not be calculated by means of multiplications or divisions.
+E.g.
+the mapping between months and days.
+(This mapping is represented in the figure with a bold arrow.
+\item
+Congruent mapping: Two granularities that have the same granules but with different anchos.
+E.g.
+the mapping between the Gregorian Calendar and academic days.
+This mapping is represented in the figure with a straight line without arrow.
+\end{itemize}
+
+The transition between granularities~\cite{Lin97} is also considered as a source of imprecision.
+In a sentence like \emph{`The Alhambra is 612 years old'}, the temporal notion is very precise if time periods shorter than a year are irrelevant in the used granularity.
+However, if one is concerned with the exact day on which the last stone of the building is laid, a transition to a finer granularity, allowing the specification of time periods shorter than a year, is needed.
+Therefore, some proposals consider granularity as the base of the temporal model~\cite{Cru97}.
+
+\subsection{\label{subsec:timedomain-calendar}Time domain: Calendars}
+%maybe introduce previously the concept of time units, linear hierarchy and time point.
+A Calendar is a organization of different granularities (e.g. day, month, year) in a hierarchy. 
+In order to define a calendar, some previous concepts should be introduced~\cite{Kraus1997}:
+
+\begin{definition}
+\label{def:granularity-hierarchy}
+\textbf{(Linear hierarchy of granularities)} A linear hierarchy of granularities is a finite collection of granularities with a linear order $\sqsubset$ among these granularities.
+The coarser granularity is called the top of the hierarchy and it must be an infinite set while all the others granularities must be finite sets.
+\end{definition}
+
+The hierarchy of days $\sqsubset$ months $\sqsubset$ years defines a linear hierarchy.
+
+\begin{definition}
+\label{def:linear-calendar}
+\textbf{(Linear Calendar)} A linear calendar consist of a linear hierarchy $H$ of granularities and a validity predicate which specifies if a granule is valid in  the calendar.
+\end{definition}
+
+The linear hierarchy days $\sqsubset$ months $\sqsubset$ years represents a linear calendar.
+It is necessary to define the validity predicates because 17 January 2012 is a valid element of the calendar whereas 30 February 2012 is not.
+
+\subsubsection{\label{subsubsubsec:julian-day-number}An example of temporal domain: Julian Day Number}
 % proposal for the underlying Julian Day Number domain.
 
-The Julian Day number \cite{Dir96} is a counter.
+The Julian Day number \emph{JDN} \cite{Dir96} is a counter.
 Its value is incremented in one unit every day from 1 January 4713 B.C.
 at 12:00 noon.
-The particularity of starting at noon was useful for astronomers: the observations they took one night belonged to the same Julian Day.
-Actually it has not any transcendence.\\
+The particularity of starting at noon was useful for astronomers: the observations they took one night belonged to the same Julian Day.\\
 Note that the Julian Day represents whole days.
@@
 
-\subsubsection{Conversion among dates}
-\paragraph{Conversion between Gregorian and Julian dates}
-The Julian Date consist on the Julian Day number and the time. It is computed with the formula \cite{Usn}:\\
-\begin{equation}
-\label{eq:julian day}
-JD = JDN + \frac{hour-12}{24} + \frac{minutes}{1440} + \frac{seconds}{86400}
-\end{equation}
+\subsection{Temporal Relations}
+%an explanation on allen's temporal relations.
+%definition of fuzzy temporal interval
+\def\JPicScale{0.5}
+\begin{figure}[h]
+\centering
+\input{graphs/allen.tex}
+\caption{Allen relations between two time intervals.}
+\label{fig:allen}
+\end{figure}
 
-Here the JDN is the Julian Day Number.
-This equation does not take into account the leap seconds \footnote{A leap second(s) is the amount of second(s) added or deleted every year in the Universal Time.\cite{Wik},\cite{Lea}.
-}.
-
-\paragraph{Gregorian calendar to Julian day}
-Any date from Gregorian calendar \footnote{As Gregorian calendar entered in use in 1582, any use before that date is called \emph{proleptic calendar}.} can be converted into a Julian Day using the following formula:\\
-\begin{equation}
-\label{eq:JDN}
-JDN = \frac{1461x(Y + 4800 + \frac{M-14}{12})}{4}+\frac{367x(M-2-12x\frac{M-14}{12})}{12}-\frac{3x\frac{Y+4900+\frac{M-14}{12}}{100}}{4} 
-\end{equation}
-
-This formula was created by Henry Fliegel and Thomas C.
-van Flandern in 1968.
-\cite{Fliegel:1968:LEM:364096.364097} It takes into account leap years and that they are applicable only when the century is divisible by 400.
-The value $Y$ represents the year and the value $M$ represents the month.
-\paragraph{Julian calendar to Julian day}
-The algorithm to convert between Julian calendar and Julian Day Number is only valid for dates since 4713 B.C (positive values of Julian Day) \cite{Dog}:\\
-\begin{equation}
-\label{eq:JDN_2_JD}
-JDN=367Y-\frac{7 (Y + 5001 + \frac{M-9}{7})}{4}+\frac{275xM}{9}+D+1729777
-\end{equation}
-
-Where $Y$ is the year value, $M$ is the month value, and $D$ is the day value.
-
-%\subsubsection{Membership degree computation}
-%The main advantage from the Julian day number is the ability to make calculations. Thus, the membership degree computation for a fuzzy number based on a Julian day is explained here. \\
-%Let a trapezoidal possibility distribution defined by four JDN named $a$,$b$,$c$ and $d$.
-The fuzzy validity period \emph{FVP}, can be represented as in figure \ref{fig:fuzzy-validity-period}.
-The following equation shows the membership degree for a Julian day number \emph{JDN} in a \emph{FVP} represented by means of the possibility distribution of $[a,b,c,d]$.
-In a valid-time database, the values for the valid starting time \emph{VST} and the valid ending time \emph{VET} can be fuzzyfied as shown in equation \ref{eq:fuzzy-validity-period}.
-%
-%The membership function for a given $x$ number in \emph{JDN} format is the following:\\
-%\begin{equation}
-%\label{eq:trapezoidal distribution}
-%%\[
-%T(x)=
-%\begin{cases}
-%0, \mbox(if x \leq a, x \geq d)\\
-%1, \mbox(if x \in [b,c])\\
-%\frac{x-a}{b-a}   \mbox(if x \in (a,b])\\
-%\frac{d-x}{d-c} 	 \mbox(if x \in (c,d])\\
-%\end{cases}
-%%\]
-%\end{equation}
-%The points outside the trapezium or inside the interval $[a,b]$ are trivial. The points inside the other intervals has the following particularities:\\
-%The subtraction operation with Julian day numbers gives as result a scalar value (as explained before).
-The division operation between two scalars gives as result an scalar in the interval $[0,1]$ which represents the membership degree.
+Several operators are defined in order to compare temporal elements.
+Allen~\cite{Allen83} first described the relations between time intervals and to a lesser extent, between time points.
+Figure \ref{fig:allen} shows the temporal relations.
+There are some proposals that can be applied to both crisp and fuzzy temporal interval ~\cite{ohlbach2004},\cite{nagypal2003},\cite{schockaert08}.
+For this fuzzy comparisons~\cite{garrido2009} proposed some different temporal relations that can be easyly implemented by means of regular fuzzy operations.

### 2012-01-18 trunk/book/sources/fuzzyTemporalDatabases.tex
update of the querying part

--- 2012-01-12 book/sources/fuzzyTemporalDatabases.tex
@@
 \label{subsec:querying-time-intervals}
+One of the main pourposes of a database is to allow the information retrieval.
+The standard query language for databases is SQL ~\cite{Mel93}, however, there are several proposals to extend the SQL language for transaction-time databases~\cite{Sarda90}, valid-time databases~\cite{gad92} and bi-temporal ~\cite{TSQL}.
+Some authors have studied how to support temporal querying in SQL~\cite{Snodgrass98}.
+
+In the querying of a fuzzy temporal database it is possible to distinguish among the following cases:
+
+\begin{itemize}
+\item
+Fuzzy data stored in the database and crisp specification in the query.
+\item
+Crisp data stored in the database and fuzzy data in the query specification.
+\item
+Both data stored in the database and the query specification are fuzzy.
+\end{itemize}
+In this subsection we will explain the query specification for a database that stores ill-known time intervals and crisp values in the query.
+Afterwards, the aggregation and ranking of the temporal results is explained.
+
+\subsubsection{Query structure}
+Consider that in a regular or fuzzy relational database, the query specification for the non-temporal attributes is given by $Q$.
+Therefore, the query specification in a temporal or fuzzy temporal database is given by $\tilde{Q}$:
+
+\begin{definition}
+\textbf{(query specification)}
+A query $\tilde{Q}$ is given by:
+\begin{equation}
+\tilde{Q} = \left( Q^{time}, Q \right)
+\end{equation}
+Here $Q$ are the possibly fuzzy non-temporal contraints and $Q^{time}$ are the temporal constraints specified in the query:
+\end{definition}
+
+\begin{definition}
+\textbf{(temporal query specification)}
+$Q^{time}$ is defined by:
+\begin{equation}
+Q^{time} = \left( I,AR \right)
+\end{equation}
+Where $I$ is a crisp time interval and $AR$ is one of the Allen's relations.
+The interpretation is that given an ill-known time interval $J$, the user request the set of records in which $I$ AR $J$ hold.
+\end{definition}
+
+\subsubsection{Query evaluation}
+Query satisfaction in a (fuzzy) relational database is a matter of degree.
+Tipically, each constraint provided by the user is evaluated and results in a \emph{satisfaction degree} $s \in \left[ 0,1,\right]$.
+The value 0 denotes a total dissatisfaction while the value 1 denotes a total satisfaction.
+
+The query $\tilde{Q}$ is evaluated in two parts that happen simultaneously:
+\begin{itemize}
+\item
+The non-temporal part, $Q$ is evaluated resulting in a satisfaction degree noted as $e_Q(r)$ for each record $r$.
+\item
+The temporal part, $Q^{time}$ is evaluated depending on $AR$.
+A specific set of ill-known constraints is considered depending on the Allen relation in $AR$.
+Then both possibility and necessity measures are computed and aggregated.
+\end{itemize}
+
+\subsubsection{Ranking and aggregation}
+For each temporal attribute $r_{time}$ for the record $r$, both possibility $\Pos \left( Q_{time}\right)$ and necessity $\Nec \left( Q_{time} \right)$ measures are obtained.
 

### 2012-01-23 trunk/book/sources/fuzzyTemporalDatabases.tex
update with the bipolar querying of temporal databases. some typos corrected.

@@
 a logistics company which transports packages from one destination to another.
-The time when the package would be at the end point may be stimated but not precisely known.
+The time when the package would be at the end point may be estimated but not precisely known.
 These applications need to manage the uncertainty with respect to the temporal attributes of the objects stored in the database.
@@
 \label{subsec:querying-time-intervals}
-One of the main pourposes of a database is to allow the information retrieval.
+One of the main purposes of a database is to allow the information retrieval.
 The standard query language for databases is SQL ~\cite{Mel93}, however, there are several proposals to extend the SQL language for transaction-time databases~\cite{Sarda90}, valid-time databases~\cite{gad92} and bi-temporal ~\cite{TSQL}.
@@
 \begin{equation}
+\label{eq:query-specification}
 \tilde{Q} = \left( Q^{time}, Q \right)
 \end{equation}
-Here $Q$ are the possibly fuzzy non-temporal contraints and $Q^{time}$ are the temporal constraints specified in the query:
+Here $Q = \left \lbrace q_1, \cdots, q_n \right \rbrace$ are the possibly fuzzy non-temporal constraints and $Q^{time}$ are the temporal constraints specified in the query:
 \end{definition}
+
+%%specify better this:
+%\begin{svgraybox}
+%\begin{definition}
+%\label{def:query-constraint}
+%\textbf{(query constraint)}
+%$Q$ is the 
+%A query constraint $q_a$ for an attribute $a$ with attribute domain $\mathcal{D}$, is a restriction in the subset of the values in the attribute domain:
+%\begin{equation}
+%\label{eq:query-constraint}
+%q_a = \left \lbrace d \mid d \subseteq \mathcal{D} \right \rbrace
+%\end{equation}
+%\end{definition}
 
@@
 \subsubsection{Query evaluation}
+\label{subsubsec:query-evaluation}
 Query satisfaction in a (fuzzy) relational database is a matter of degree.
-Tipically, each constraint provided by the user is evaluated and results in a \emph{satisfaction degree} $s \in \left[ 0,1,\right]$.
+Typically, each constraint provided by the user is evaluated and results in a \emph{satisfaction degree} $s \in \left[ 0,1,\right]$.
 The value 0 denotes a total dissatisfaction while the value 1 denotes a total satisfaction.
 
-The query $\tilde{Q}$ is evaluated in two parts that happen simultaneously:
+\begin{definition}
+\label{def:evaluation-function}
+\textbf{(evaluation function)}
+An evaluation function $e_{Q} \left( r \right)$ is a mapping from the attribute value from the row $r$ in the database with respect to the constraints $Q = \left \lbrace q_1, \cdots, q_n \right \rbrace$ to the unit interval $\left[ 0,1 \right]$.
+\begin{equation}
+\label{eq:evaluation-function} 
+e_{Q} = t \mid t \in \left[ 0,1 \right]
+\end{equation}
+\end{definition}
+
+Each element in $\tilde{Q}$ is evaluated independently:
 \begin{itemize}
@@
 For each temporal attribute $r_{time}$ for the record $r$, both possibility $\Pos \left( Q_{time}\right)$ and necessity $\Nec \left( Q_{time} \right)$ measures are obtained.
+Then, the measures are combined in order to obtain an evaluation score $e_{Q_{time} \left( r \right)}$.
+The evaluation function is computed by:
+
+\begin{equation}
+\label{eq:temporal-eval}
+e_{Q_{time} \left( r \right)} = \left \lbrace t \mid t = \left( \frac{\Pos \left( Q_{time}\right) + \Nec \left( Q_{time} \right)}{2} \right)   \in \left[ 0,1\right] \right \rbrace
+\end{equation}
+
+This measure provides a natural score because of the following property of the necessity measure: 
+\begin{equation}
+\label{eq:necessity} %iff or iif?
+\Nec \left( Q_{time} \right) > 0  \Longleftrightarrow  \Pos \left( Q_{time}\right) = 1
+\end{equation}
+
+The final ranking is given by a convex combination of both temporal and non-temporal evaluations:
+
+\begin{equation}
+\label{eq:convex-combination}
+e_{final} \left( r \right) = \left \lbrace t \mid t = \omega \ast e_{Q} \left( r \right) + \left( 1- \omega \right) \ast e_{Q^{time}}, t,\omega \in \left[0,1\right] \right \rbrace
+\end{equation}
 
 \subsubsection{Bipolarity in temporal database querying}
+\label{subsubsec:bipolarity}
+Humans express their preferences in both positive and negative statements.
+This property is interesting in querying of databases, because sometimes, the user does not know exactly his / her preferences but what he / she does not want.
+This introduces the necessity of a bipolar way of querying databases.
+Therefore, the bipolar querying of databases is a querying technique which allows to express both positive and negative criteria in the database query.
+
+In this subsection we will address a proposal for the bipolar querying of temporal databases made by Billiet~\cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy}.
+First of all, the query structure is shown.
+Then, both evaluation and ranking proposed by the authors are explained.
+The model deals with a fuzzy valid-time specification as in~\cite{garrido2009}.
+
+\paragraph{Query structure}
+The approach in this work follows the query specification given in \eqref{eq:query-specification}.
+The query has a global time demand, $Q^{time}$ and here, the regular query constraints specified by $Q$ are split into positive and negative preferences:
+
+\begin{equation}
+\label{eq:bipolar-specification}
+\tilde{Q} = \left \lbrace Q^{time}, \left( Q^{pos}, Q^{neg} \right) \right \rbrace
+\end{equation}
+
+Here $Q^{pos}$ and $Q^{neg}$ represent the positive and the negative criteria, respectively.
+The temporal demand in $Q^{time}$ may be specified as in ~\cite{garrido2009} or as in ~ \cite{Pon11}, see section \ref{subsec:representation-time-intervals}.
+.
+\paragraph{Query evaluation}
+As explained previously in the query evaluation section (\ref{subsubsec:query-evaluation}), each element in $\tilde{Q}$ is evaluated independently:
+
+\begin{itemize}
+\item
+The query $Q$ has now two elements: $Q^{pos}$ and $Q^{neg}$.
+Each element is again evaluated independently and the result is a tuple $\left(s,d \right)$ in which $s$ is the satisfaction degree for $Q^{pos}$.
+$d$ is called dissatisfaction degree for $Q^{neg}$.
+The tuple $\left(s,d \right) s,d \in \left[0,1 \right]$ is called \emph{Bipolar Satisfaction Degree}.
+\item
+The temporal specification is evaluated as explained in \ref{subsubsec:query-evaluation} and the result is a value in the unit interval.
+\end{itemize}
+
+\paragraph{Ranking}
+In order to present the results to the user, it is necessary to design a combination function.
+This function allows the classification of the results in the unit interval the preferred way is to made a convex combination as  presented in equation \eqref{eq:convex-combination}.

### 2012-01-23 trunk/book/sources/intro.tex
update with the bipolar querying of temporal databases. some typos corrected.

@@
 Next to the manage of temporal information is the temporal reasoning~\cite{Allen83}.
-Dubois and Prade ~\cite{Dubois:jucs_9_9:fuzziness_and_uncertainty_in},\cite{Dubois89} deal with fuzzyness and uncertainty in temporal reasoning, but this topic is outside the scope of this work.
+Dubois and Prade ~\cite{Dubois:jucs_9_9:fuzziness_and_uncertainty_in},\cite{Dubois89} deal with fuzziness and uncertainty in temporal reasoning, but this topic is outside the scope of this work.
 
@@
 The section concludes with a comparison among the different proposals.
-In Section \ref{sec:fuzzy-temporal-databases} we present a sum up with the proposals for handling the imprecision in temporal databases and the main deficiences we have detected.
+In Section \ref{sec:fuzzy-temporal-databases} we present a sum up with the proposals for handling the imprecision in temporal databases and the main deficiencies we have detected.
 Finally, there are several commercial temporal database management systems  \emph{TDMBS} like \cite{oracle2009},\cite{posgree2009},\cite{teradata2011},\cite{timedb2005}.

### 2012-01-23 trunk/book/sources/prelim.tex
update with the bipolar querying of temporal databases. some typos corrected.

@@
 \subsection{\label{subsec:poss}Possibilistic variables}
-Informally, a theory of uncertainty considers a set of uncertain events and tries to quantify the confidence about the occurence of those events.
+Informally, a theory of uncertainty considers a set of uncertain events and tries to quantify the confidence about the occurrence of those events.
 Consider therefore a set of outcomes $\Omega$ and let each subset of $\Omega$ be called an \emph{event}.

### 2012-01-23 trunk/book/sources/timeDomain.tex
update with the bipolar querying of temporal databases. some typos corrected.

@@
 There are some proposals that can be applied to both crisp and fuzzy temporal interval ~\cite{ohlbach2004},\cite{nagypal2003},\cite{schockaert08}.
-For this fuzzy comparisons~\cite{garrido2009} proposed some different temporal relations that can be easyly implemented by means of regular fuzzy operations.
+For this fuzzy comparisons~\cite{garrido2009} proposed some different temporal relations that can be easily implemented by means of regular fuzzy operations.

### 2012-01-23 book/sources/fuzzyTemporalDatabases.tex
added graph for the workspace manager. Added some text for the commercial temporal databases. Added fuzzy allen relations.

@@
 
+\begin{table}[h]
+
+\caption{Allen's relations used in the framework.
+Here, $I = \left[a, b\right]$ denotes a crisp time interval, $J = \left[X, Y\right]$ denotes an ill-known time interval, with $\pi_{X}$ and $\pi_{Y}$ the possibility distributions of $X$ and $Y$ respectively.
+The second column contains the corresponding formula to calculate the possibility that $I$ satisfies all constraints given by the Allen's relation.}
+
+\centering
+\begin{tabular}{|c|c|}
+\hline
+Allen Relation &  $\Pos\left(\text{I satisfies all constraints }C_{i}, i = 1,2,...\right)$ \\
+\hline
+I before J & $\sup_{a>w}\pi_x(w)$\\
+\hline
+
+\multirow{2}{*}
+{I equal J} &  $\min ( \sup_{a \leq w}\pi_x(w),\pi_x(w),$ \\
+ &  $\sup_{b \geq w}\pi_Y(w),\pi_Y(w))$\\
+\hline
+
+I meets J  & $\min (\sup_{a\geq w} \pi_X(w),$ $\pi_X(w))$ \\
+\hline
+
+\multirow{2}{*}
+{I overlaps J}  & $\min ( \sup_{b>w}\pi_Y(w), $ \\
+ & $\sup_{a \geq w}\pi_X(w),\sup_{a \leq w}\pi_X(w))$ \\
+\hline
+
+\multirow{2}{*}
+{I during J}  & $\max ( \min ( \sup_{a<w}\pi_X(w),\sup_{b \geq w}\pi_Y(w)),$ \\
+ & $\min ( \sup_{a \leq w }\pi_X(w),\sup_{b>w}\pi_Y(w)$\\
+\hline
+
+{I starts J} &  $\min( \sup_{a \leq w}\pi_X(w),$ $\pi_X(w))$\\
+\hline
+
+{I finishes J} &  $\min ( \sup_{b \geq w} \pi_Y(w),$ $\pi_Y(w))$ \\
+\hline
+
+\end{tabular}
+%
+%\vspace{10pt}
+%
+%
+%\vspace{-25pt}
+\label{tab:fuzzy-allen-relations}
+\end{table}
+
+One of the main pourposes of a database is to allow the information retrieval.
+The standard query language for databases is SQL ~\cite{Mel93}, however, there are several proposals to extend the SQL language for transaction-time databases~\cite{Sarda90}, valid-time databases~\cite{gad92} and bi-temporal ~\cite{TSQL}.
+Some authors have studied how to support temporal querying in SQL~\cite{Snodgrass98}.
+
+In the querying of a fuzzy temporal database it is possible to distinguish among the following cases:
+
+\begin{itemize}
+\item
+Fuzzy data stored in the database and crisp specification in the query.
+\item
+Crisp data stored in the database and fuzzy data in the query specification.
+\item
+Both data stored in the database and the query specification are fuzzy.
+\end{itemize}
+In this subsection we will explain the query specification for a database that stores ill-known time intervals and crisp values in the query.
+Afterwards, the aggregation and ranking of the temporal results is explained.
+
+\subsubsection{Query structure}
+Consider that in a regular or fuzzy relational database, the query specification for the non-temporal attributes is given by $Q$.
+Therefore, the query specification in a temporal or fuzzy temporal database is given by $\tilde{Q}$:
+
+\begin{definition}
+\textbf{(query specification)}
+A query $\tilde{Q}$ is given by:
+\begin{equation}
+\label{eq:query-specification}
+\tilde{Q} = \left( Q^{time}, Q \right)
+\end{equation}
+Here $Q = \left \lbrace q_1, \cdots, q_n \right \rbrace$ are the possibly fuzzy non-temporal contraints and $Q^{time}$ are the temporal constraints specified in the query:
+\end{definition}
+
+%%specify better this:
+%\begin{svgraybox}
+%\begin{definition}
+%\label{def:query-constraint}
+%\textbf{(query constraint)}
+%$Q$ is the 
+%A query constraint $q_a$ for an attribute $a$ with attribute domain $\mathcal{D}$, is a restriction in the subset of the values in the attribute domain:
+%\begin{equation}
+%\label{eq:query-constraint}
+%q_a = \left \lbrace d \mid d \subseteq \mathcal{D} \right \rbrace
+%\end{equation}
+%\end{definition}
+
+\begin{definition}
+\textbf{(temporal query specification)}
+$Q^{time}$ is defined by:
+\begin{equation}
+Q^{time} = \left( I,AR \right)
+\end{equation}
+Where $I$ is a crisp time interval and $AR$ is one of the Allen's relations, see table \ref{tab:fuzzy-allen-relations}.
+The interpretation is that given an ill-known time interval $J$, the user request the set of records in which $I$ AR $J$ hold.
+\end{definition}
+
+\subsubsection{Query evaluation}
+\label{subsubsec:query-evaluation}
+Query satisfaction in a (fuzzy) relational database is a matter of degree.
+Tipically, each constraint provided by the user is evaluated and results in a \emph{satisfaction degree} $s \in \left[ 0,1,\right]$.
+The value 0 denotes a total dissatisfaction while the value 1 denotes a total satisfaction.
+
+\begin{definition}
+\label{def:evaluation-function}
+\textbf{(evaluation function)}
+An evaluation function $e_{Q} \left( r \right)$ is a mapping from the attribute value from the row $r$ in the database with respect to the constraints $Q = \left \lbrace q_1, \cdots, q_n \right \rbrace$ to the unit interval $\left[ 0,1 \right]$.
+\begin{equation}
+\label{eq:evaluation-function} 
+e_{Q} = t \mid t \in \left[ 0,1 \right]
+\end{equation}
+\end{definition}
+
+Each element in $\tilde{Q}$ is evaluated independently:
+\begin{itemize}
+\item
+The non-temporal part, $Q$ is evaluated resulting in a satisfaction degree noted as $e_Q(r)$ for each record $r$.
+\item
+The temporal part, $Q^{time}$ is evaluated depending on $AR$.
+A specific set of ill-known constraints is considered depending on the Allen relation in $AR$.
+Then both possibility and necessity measures are computed and aggregated.
+\end{itemize}
+
+\subsubsection{Ranking and aggregation}
+For each temporal attribute $r_{time}$ for the record $r$, both possibility $\Pos \left( Q_{time}\right)$ and necessity $\Nec \left( Q_{time} \right)$ measures are obtained.
+Then, the measures are combined in order to obtain an evaluation score $e_{Q_{time} \left( r \right)}$.
+The evaluation function is computed by:
+
+\begin{equation}
+\label{eq:temporal-eval}
+e_{Q_{time} \left( r \right)} = \left \lbrace t \mid t = \left( \frac{\Pos \left( Q_{time}\right) + \Nec \left( Q_{time} \right)}{2} \right)   \in \left[ 0,1\right] \right \rbrace
+\end{equation}
+
+This measure provides a natural score because of the following property of the necessity measure: 
+\begin{equation}
+\label{eq:necessity} %iff or iif?
+\Nec \left( Q_{time} \right) > 0  \Longleftrightarrow  \Pos \left( Q_{time}\right) = 1
+\end{equation}
+
+The final ranking is given by a convex combination of both temporal and non-temporal evaluations:
+
+\begin{equation}
+\label{eq:convex-combination}
+e_{final} \left( r \right) = \left \lbrace t \mid t = \omega \ast e_{Q} \left( r \right) + \left( 1- \omega \right) \ast e_{Q^{time}}, t,\omega \in \left[0,1\right] \right \rbrace
+\end{equation}
+
 \subsubsection{Bipolarity in temporal database querying}
+\label{subsubsec:bipolarity}
+Humans express their preferences in both positive and negative statements.
+This property is interesting in querying of databases, because sometimes, the user does not know exactly his / her preferences but what he / she does not want.
+This introduces de necessity of a bipolar way of querying databases.
+Therefore, the bipolar querying of databases is a querying technique which allows to express both positive and negative criteria in the database query.
+
+In this subsection we will address a proposal for the bipolar querying of temporal databases made by Billiet~\cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy}.
+First of all, the query structure is shown.
+Then, both evaluation and ranking proposed by the authors are explained.
+The model deals with a fuzzy valid-time specification as in~\cite{garrido2009}.
+
+\paragraph{Query structure}
+The approach in this work follows the query specification given in \eqref{eq:query-specification}.
+The query has a global time demand, $Q^{time}$ and here, the regular query constraints specified by $Q$ are splitted into possitive and negative preferences:
+
+\begin{equation}
+\label{eq:bipolar-specification}
+\tilde{Q} = \left \lbrace Q^{time}, \left( Q^{pos}, Q^{neg} \right) \right \rbrace
+\end{equation}
+
+Here $Q^{pos}$ and $Q^{neg}$ represent the positive and the negative criteria, respectively.
+The temporal demand in $Q^{time}$ may be specified as in ~\cite{garrido2009} or as in ~ \cite{Pon11}, see section \ref{subsec:representation-time-intervals}.
+.
+\paragraph{Query evaluation}
+As explained previously in the query evaluation section (\ref{subsubsec:query-evaluation}), each element in $\tilde{Q}$ is evaluated independently:
+
+\begin{itemize}
+\item
+The query $Q$ has now two elements: $Q^{pos}$ and $Q^{neg}$.
+Each element is again evaluated independently and the result is a tuple $\left(s,d \right)$ in which $s$ is the satisfaction degree for $Q^{pos}$.
+$d$ is called dissatisfaction degree for $Q^{neg}$.
+The tuple $\left(s,d \right) s,d \in \left[0,1 \right]$ is called \emph{Bipolar Satisfaction Degree}.
+\item
+The temporal specification is evaluated as explained in \ref{subsubsec:query-evaluation} and the result is a value in the unit interval.
+\end{itemize}
+
+\paragraph{Ranking}
+In order to present the results to the user, it is necessary to design a combination function.
+This function allows the classification of the results in the unit interval the preffered way is to made a convex combination as  presented in equation \eqref{eq:convex-combination}.

### 2012-01-24 trunk/book/sources/frw.tex
Update of the titles in the file author.tex. First version of the conclusions and further reseach work.

In this work we have presented the main properties of the time in an information system. An overview of the commercial temporal DBMS has also been introduced. Although, to the best of our knowledge, no commercial fuzzy temporal DBMS have been found.  Finally the managing of imprecision in temporal databases is explained. A brief introduction of bipolar querying in temporal databases is done at the end of the section.

Further research work should be done in several ways: first of all, a theoretical model for both dealing with uncertainty in the database as well as in the querying should be defined. Then, another interesting work is to propose an implementation with both DDL (Data Definition Language) and DML (Data Manipulation Language).

### 2012-01-27 book/sources/intro.tex
Cleaned up the title page and abstract, changed the introduction and added some more bibliography...

@@
 
-In Section \ref{sec:time-domain} an study about the time and its properties is done.
+The concept of time itself is very complex to handle and interpret~\cite{klein94},~\cite{Shackle61}, though it is very natural and omnipresent.
+As information systems often attempt the modelling or the representation of natural objects, concepts or processes, they often require modelling or representing temporal information.
+Thus, several proposals have been concerned with the obtaining of theoretical models that allow the modelling or representation of time~\cite{Bolour82},~\cite{Cru97}.
+
+A very specific type of information systems are database systems, which are computer systems designed to manage databases.
+A database contains data representing real objects or concepts.
+Every one of these data is a measurement or description of a property of a real object or concept.
+In reality, some aspects or properties of objects or concepts are time-variant or time-related.
+E.g.
+the moment of a bank transaction is traditionally a moment in time and thus a time-related notion, the function of an employee in a company can change through recorded history and is thus time-variant.
+A temporal database schema is a database schema that models real objects or concepts with time-related or time-variant properties.
+However, the modelling of temporal aspects has a direct impact on the consistency of the temporal database, because the temporal nature of these aspects imposes extra integrity constraints.
+An example.
+Consider a relation in a relational library database, modelling the presence of books in the library.
+Every physical book is represented by a unique identifier.
+Every record in the relation contains such an identifier, a date on which the corresponding book was loaned and a date on which it was subsequently returned (if it was returned).
+Without further precautions, a library employee could add several records with the same book identifier, different `loaned'-dates and no `returned'-dates.
+This would represent the same physical book being loaned several times on different dates and never returned, which is of course impossible.
+A temporal database model will typically constrain record insertion and prevent similar modelling inconsistencies.
+
+A lot of research concerns these temporal database models and their approaches to the modelling of time.
+The first efforts were towards the representation of historical information related to objects represented by records in a database~\cite{Clifford:1985:AHR:971699.318922}.
+Some proposals tried to extend the Entity Relationship Model~\cite{Klopprogge:1983}, without impact on any database standards like SQL~\cite{Sarda:1990:ESH:627277.627409}.
+
+Notably, in 1994, ``A Consensus Glossary of Temporal Database Concepts'' was published~\cite{Dyreson1994}.
+For this publication, 44 temporal database researchers, among which some of the main researchers in this field, cooperated to reach a consensus on the nature and definitions of some of the main temporal database concepts and terminology.
+This glossary was subsequently updated in 1998~\cite{Jensen:Dyreson:1998:TheConsensusGlossary}.
+
+An interesting issue in temporal modelling concerns relationships between temporal notions.
+Notably, Allen~\cite{Allen83} studied temporal relationships between time intervals (and as a special case time points).
+Among others, the querying of temporal databases has greatly profited from these temporal relationships, because they allowed for richer and complexer user-specified temporal query demands, by allowing to express more complex relationships between the temporal notions in the temporal expressions in the query and the temporal indications in the database.
+
+Humans handle temporal information using certain temporal notions like time intervals or time points~\cite{Dyreson1994}.
+Humans are often not hindered by imperfections like imprecisions, vaguenesses, uncertainties or inconsistencies possibly contained in the descriptions of these temporal notions.
+Among many others, these possible imperfections in descriptions of temporal notions determine an important issue in temporal modelling.
+E.g.
+the temporal notion in a sentence like `The Belfry of Bruges was finished on a day somewhere between 1/01/1201 A.D.
+and 31/12/1300 A.D.' contains imperfection because of the uncertainty in the used time-related expression.
+It is known that the building was finished on a single day, but it is not known precisely which day this was.
+
+To allow information systems to cope with these and similar imperfections, many approaches adopt fuzzy sets~\cite{Zadeh65} for the representation of temporal information~\cite{343607},~\cite{nagypal2003},~\cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy},~\cite{Dubois:jucs_9_9:fuzziness_and_uncertainty_in}.
+The temporal relationships studied by Allen were fuzzified by several authors~\cite{ohlbach2004},~\cite{nagypal2003},~\cite{schockaert08}.
+Garrido et al.
+~\cite{garrido2009} present different temporal operators, defined by a combination of regular fuzzy comparisons.
+Both~\cite{garrido2009} and~\cite{Pon11} deal with uncertainty in temporal expressions concerning time intervals.
+Other approaches, like~\cite{Qia09}, use rough sets~\cite{Pawlak1995} to represent time intervals.
+
+Next to temporal modelling, some attention has been spent on temporal reasoning~\cite{Allen83}.
+Although temporal reasoning is not discussed in this chapter, it should be noted that, among others, Dubois and Prade~\cite{Dubois:jucs_9_9:fuzziness_and_uncertainty_in},~\cite{Dubois89} have dealt with fuzziness and uncertainty in temporal reasoning.
+
+The rest of this chapter is structured as follows.
+Section \ref{sec:time-domain} presents some basic concepts and terminology about temporal modelling and discusses some of its important aspects and issues.
+In section \ref{sec:temporal-databases}, some basic concepts and terminology about temporal databases are presented, followed by an overview of some interesting issues concerning temporal databases and a survey of some commercial temporal database systems.
+\textcolor{red}{TODO: REFERENCE SECTION 4!}
+
+%The concept of \emph{time} is easy to understand but very complex to define~\cite{klein94}~\cite{Shackle61}. 
+%In Information Systems \emph{IS}, several proposals have been concerned with the obtaining of theoretical models that allow the representation of time~\cite{Bolour82}~\cite{Cru97}. 
+%More towards temporal databases, a lot of work have been done.
+The first efforts were towards the representation of the historical information related to an object in the database(reference).
+Some works tried to extend the Entity Relationship Model \emph{ERM} ~\cite{Klopprogge:1983} but without an impact on any database standards like SQL~\cite{Sarda:1990:ESH:627277.627409}.
+
+%The ``Consensus Glossary of Temporal Database concepts'' \cite{Dyreson1994} is the first publication where the main researchers in temporal databases define the main thesaurus of temporal database concepts.
+This glossary defines the main types of time in a temporal database and is the basis of furthers researches ~\cite{Sarda:1990:ESH:627277.627409} \cite{Jensen94thetsql2}.
+
+%Another main issue in temporal databases is the querying.
+The user specifies in the query some temporal constraints, e.g.
+\emph{``before this year''},\emph{``after 1995''}.
+In order to solve these temporal comparisons, between the time specified in the query and the temporal data stored in the database, some basic relationships have been defined.
+Allen~\cite{Allen83} first introduced the temporal relations between time intervals and, to a lesser extent, to time points.
+Some authors~\cite{ohlbach2004},\cite{nagypal2003},\cite{schockaert08} have softened the temporal operators defined by Allen.
+Therefore it is possible to compute the relationships between two ill-known time points or intervals.
+In~\cite{garrido2009}, some different temporal operators are defined by a combination of regular fuzzy comparisons.
+%e.g.
+the \emph{BEFORE} operator is defined by means of the fuzzy less than (FLT) operator.
+
+%To allow information systems to cope with uncertainty, many approaches adopt fuzzy sets~\cite{Zadeh65} for the representation of temporal information~\cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy},\cite{Dubois:jucs_9_9:fuzziness_and_uncertainty_in},~\cite{devos94}.
+The time points representing the time when the object is valid in the reality might not be precisely known.
+In order to deal with this, several models have been proposed.
+Garrido~\cite{garrido2009} proposes a model to deal with uncertainty in the time interval by means of fuzzy intervals.
+Bronselaer~\cite{Pon11} proposes a consistent framework to deal and represent ill-known time values and their relationships.
+Rough sets~\cite{Pawlak1995} have been also used to represent time intervals~\cite{Qia09} in databases.
+
+%temporal reasoning and fuzzy temporal reasoning.
+%Next to the manage of temporal information is the temporal reasoning~\cite{Allen83}.
+Dubois and Prade ~\cite{Dubois:jucs_9_9:fuzziness_and_uncertainty_in},\cite{Dubois89} deal with fuzziness and uncertainty in temporal reasoning, but this topic is outside the scope of this work.
+
+%temporal commercial systems.
+
+%The work is organized as follows:
+
+%In Section \ref{sec:time-domain} an study about the time and its properties is done.
 Section \ref{sec:temporal-databases} is an overview of the main problems when managing the time in a database, the different proposals for solving them and the temporal database proposals.
 The section concludes with a comparison among the different proposals.
-Finally, in Section \ref{sec:fuzzy-temporal-databases} we present a sum up with the proposals for handling the imprecision in temporal databases and the main deficiences we have detected.
+In Section \ref{sec:fuzzy-temporal-databases} we present a sum up with the proposals for handling the imprecision in temporal databases and the main deficiencies we have detected.
+Finally, there are several commercial temporal database management systems  \emph{TDMBS} like \cite{oracle2009},\cite{posgree2009},\cite{teradata2011},\cite{timedb2005}.
+All of them are analyzed and compared later on this work.

### 2012-01-27 trunk/book/author.tex
update in the bibliography. fix for the frame  in the allen relations graph. new organization in timeDomain.tex fix for the bibliography in iintro.tex
update of the querying part
Cleaned up the title page and abstract, changed the introduction and added some more bibliography...

@@
 \usepackage{makeidx}         % allows index generation
-\usepackage{graphicx}        % standard LaTeX graphics tool
+\usepackage[dvips]{graphicx}        % standard LaTeX graphics tool
                              % when including figure files
@@
 \usepackage{amssymb}
+\usepackage{multirow}
+\usepackage{color}
 
@@
 \newcommand{\T}{\mathcal{T}}
-
+\newcommand{\C}{\mathcal{C}}
+\newcommand{\Nat}{\mathbb{N}}
+\newcommand{\Q}{\mathbb{Q}}
+\newcommand{\R}{\mathbb{R}}
 \begin{document}
 
-\title*{An overview on temporal databases.\\
- New approaches and proposals.}
+\title*{An Overview on Temporal Databases.\\
+ New Approaches and Proposals.}
 % Use \titlerunning{Short Title} for an abbreviated version of
 % your contribution title if the original one is too long
-\author{Jos\'e Pons \and Christophe Billiet \and Olga Pons \and Guy de Tr\'e}
+\author{Jos\'e Pons \and Christophe Billiet \and Olga Pons \and Guy De Tr\'e}
 % Use \authorrunning{Short Title} for an abbreviated version of
 % your contribution title if the original one is too long
-\institute{Jos\'e Pons \at Department of Computer Science and Artificial Intelligence, Univeristy of Granada, C/ Periodista Daniel Saucedo Aranda s/n \email{jpons@decsai.ugr.es}
-\and Christophe Billiet \at Department of Telecommunications and Information Processing, Ghent University, Belgium, St.-Pietersnieuwstraat 41, B-9000 Gent,Belgium.
-\email{Christophe.Billiet@telin.ugent.be}
-\and Olga Pons \at Department of Computer Science and Artificial Intelligence, Univeristy of Granada, C/ Periodista Daniel Saucedo Aranda s/n \email{opc@decsai.ugr.es}
-\and Guy De Tr\'e \at Department of Telecommunications and Information Processing, Ghent University, Belgium, St.-Pietersnieuwstraat 41, B-9000 Gent,Belgium. \email{Guy.DeTre@telin.ugent.be}
+\institute{
+Jos\'e Pons \at Department of Computer Science and Artificial Intelligence, Escuela T\'ecnica Superior de Ingenier\'ia Inform\'atica, Universidad de Granada,  C/Periodista Daniel Saucedo Aranda s/n, E-18071, Granada, Spain.
+\email{jpons@decsai.ugr.es}
+\and Christophe Billiet \at Department of Telecommunications and Information Processing, Ghent University, St.-Pietersnieuwstraat 41, B-9000 Gent, Belgium. \email{Christophe.Billiet@telin.ugent.be}
+\and Olga Pons \at Department of Computer Science and Artificial Intelligence, Escuela T\'ecnica Superior de Ingenier\'ia Inform\'atica, Universidad de Granada,  C/Periodista Daniel Saucedo Aranda s/n, E-18071, Granada, Spain.
+\email{opc@decsai.ugr.es}
+\and Guy De Tr\'e \at Department of Telecommunications and Information Processing, Ghent University, St.-Pietersnieuwstraat 41, B-9000 Gent, Belgium. \email{Guy.DeTre@telin.ugent.be}
 }
@@
 
-\abstract*{There exist some objects in the nature that have a time-variant (e.g.
+\vspace{-35pt}
+%\abstract*{There exist some objects in the nature that have a time-variant (e.g.
 an employee's working life) or time-related (e.g.
@@
 
-\abstract{There exist some objects in the nature that have a time-variant (e.g.
-an employee's working life) or time-related (e.g.
-the transactions in a bank account) nature.
-Modelling these objects in a (relational) database is possible, but the time-related attributes have an impact on the consistency of the entire database.
+\abstract*{
+In reality, some objects or concepts have properties with a time-variant (e.g.
+an employee's job function) or time-related (e.g.
+financial transactions of a bank account) nature.
+Modelling these kinds of objects or concepts in a (relational) database schema is possible, but time-variant and time-related attributes have an impact on the consistency of the entire database.
 Therefore, some temporal database models have been proposed to deal with this.
-These models deal both with the problem of representation and querying.
-Time itself is a source of imprecision / vagueness / uncertainty since the measures of time are inherently imprecise.
-Humans beings manage the time with imprecision and vagueness.
-Nevertheless, the amount of uncertainty in the temporal indications are defined by socially well-known boundaries.
-Several proposals for dealing with imprecision in time have been done: some proposals consider the basis of the system the imprecision related to the conversion of diferent time points and intervals.
-While some other proposals consider the specification of the time as the source of imprecision: the time points are ill-known.}
+Time itself is a source of imprecision, vagueness and uncertainty, since existing time measuring devices are inherently imperfect.
+Accordingly, human beings manage time using temporal indications and temporal notions, which may contain imprecision, vagueness and uncertainty.
+However, the imperfection in human-used temporal indications is to some degree constrained by socially well-known boundaries.
+Several proposals for dealing with such imperfections exist.
+Some of these proposals consider the basis of the system to be the conversion of the specificity of temporal notions between used temporal expressions.
+Other proposals consider the temporal indications in the used temporal expressions to be the source of imperfection.
+In this chapter, an overview is given, concerning the basic concepts and issues related to the modelling of time as such or in (relational) database models and the imperfections that arise during or as a result of this modelling.
+Next to this, a novel and currently researched technique for handling some of these imperfections is presented.
+}
+
+\abstract{
+In reality, some objects or concepts have properties with a time-variant (e.g.
+an employee's job function) or time-related (e.g.
+financial transactions of a bank account) nature.
+Modelling these kinds of objects or concepts in a (relational) database schema is possible, but time-variant and time-related attributes have an impact on the consistency of the entire database.
+Therefore, some temporal database models have been proposed to deal with this.
+Time itself is a source of imprecision, vagueness and uncertainty, since existing time measuring devices are inherently imperfect.
+Accordingly, human beings manage time using temporal indications and temporal notions, which may contain imprecision, vagueness and uncertainty.
+However, the imperfection in human-used temporal indications is to some degree constrained by socially well-known boundaries.
+Several proposals for dealing with such imperfections exist.
+Some of these proposals consider the basis of the system to be the conversion of the specificity of temporal notions between used temporal expressions.
+Other proposals consider the temporal indications in the used temporal expressions to be the source of imperfection.
+In this chapter, an overview is given, concerning the basic concepts and issues related to the modelling of time as such or in (relational) database models and the imperfections that arise during or as a result of this modelling.
+Next to this, a novel and currently researched technique for handling some of these imperfections is presented.
+}
 
@@
 
-\section{Time modelling}
+\section{Time Modelling}
 \label{sec:time-domain}
 \input{sources/timeDomain.tex}
-
+\def\JPicScale{1}
 \section{Temporal Databases}
@@
 \input{sources/temporalDatabases.tex}
-
-\section{Imprecision, Uncertainty and vagueness in temporal databases}
+\def\JPicScale{1}
+\section{Imprecision, Uncertainty and Vagueness in Temporal Databases}
 \label{sec:imprecision-uncertainty}
@@
 %\input{sources/bipolarity.tex}
-
-\section{Further research work}
+\def\JPicScale{1}
+\section{Conclusions and Further Research Work}
 \label{sec:further-research-work}
 \input{sources/frw.tex}
-
+\def\JPicScale{1}
 \section*{Acknowledgements}
@@
 \input{referenc}
+\def\JPicScale{0.5}
 
+\newpage
 \section*{Appendix: Introduction to fuzzy logic}

### 2012-01-28 trunk/book/sources/fuzzyTemporalDatabases.tex
update

@@
 \label{subsec:querying-time-intervals}
+
+\begin{table}[h]
+
+\caption{Allen's relations used in the framework.
+Here, $I = \left[a, b\right]$ denotes a crisp time interval, $J = \left[X, Y\right]$ denotes an ill-known time interval, with $\pi_{X}$ and $\pi_{Y}$ the possibility distributions of $X$ and $Y$ respectively.
+The second column contains the corresponding formula to calculate the possibility that $I$ satisfies all constraints given by the Allen's relation.}
+
+\centering
+\begin{tabular}{|c|c|}
+\hline
+Allen Relation &  $\Pos\left(\text{I satisfies all constraints }C_{i}, i = 1,2,...\right)$ \\
+\hline
+I before J & $\sup_{a>w}\pi_x(w)$\\
+\hline
+
+\multirow{2}{*}
+{I equal J} &  $\min ( \sup_{a \leq w}\pi_x(w),\pi_x(w),$ \\
+ &  $\sup_{b \geq w}\pi_Y(w),\pi_Y(w))$\\
+\hline
+
+I meets J  & $\min (\sup_{a\geq w} \pi_X(w),$ $\pi_X(w))$ \\
+\hline
+
+\multirow{2}{*}
+{I overlaps J}  & $\min ( \sup_{b>w}\pi_Y(w), $ \\
+ & $\sup_{a \geq w}\pi_X(w),\sup_{a \leq w}\pi_X(w))$ \\
+\hline
+
+\multirow{2}{*}
+{I during J}  & $\max ( \min ( \sup_{a<w}\pi_X(w),\sup_{b \geq w}\pi_Y(w)),$ \\
+ & $\min ( \sup_{a \leq w }\pi_X(w),\sup_{b>w}\pi_Y(w)$\\
+\hline
+
+{I starts J} &  $\min( \sup_{a \leq w}\pi_X(w),$ $\pi_X(w))$\\
+\hline
+
+{I finishes J} &  $\min ( \sup_{b \geq w} \pi_Y(w),$ $\pi_Y(w))$ \\
+\hline
+
+\end{tabular}
+%
+%\vspace{10pt}
+%
+%
+%\vspace{-25pt}
+\label{tab:fuzzy-allen-relations}
+\end{table}
+
 One of the main purposes of a database is to allow the information retrieval.
@@
 \end{equation}
-Where $I$ is a crisp time interval and $AR$ is one of the Allen's relations.
+Where $I$ is a crisp time interval and $AR$ is one of the Allen's relations, see table \ref{tab:fuzzy-allen-relations}.
 The interpretation is that given an ill-known time interval $J$, the user request the set of records in which $I$ AR $J$ hold.
@@
 
-\subsubsection{Bipolarity in temporal database querying}
+\subsection{Bipolarity in temporal databases}
 \label{subsubsec:bipolarity}
@@
 This property is interesting in querying of databases, because sometimes, the user does not know exactly his / her preferences but what he / she does not want.
+\emph{Bipolarity} is therefore a technique to model both positive and negative preferences.
+Sometimes positive and negative information are clearly symmetric and therefore, it is possible to derive one information from another e.g.
+a person may define the concept of tall as 1.80 meters or higher.
+The negative would be the opposite, not tall is therefore, being less than 1.80 meters.
+But in some cases, positive information can not be obtained directly from the negative information and vice-verse.
+E.g.
+a person wants to buy a blue motorbike, this does not mean that the person would reject a black motorbike.
+This heterogeneous semantics are given by the \emph{heterogeneous bipolarity}~\cite{Dubois2006},~\cite{Dubois2008}.\\
+
+Some previous work have been done in ~\cite{Lacroix87}.
+There, instead of positive and negative criteria, desired and mandatory query conditions are explained.
+This can be seen as a way to specify positive and negative information: the inverse of a mandatory condition is what should be rejected (negative information), whereas desired conditions can be seen as positive information.\\
+
 This introduces the necessity of a bipolar way of querying databases.
-Therefore, the bipolar querying of databases is a querying technique which allows to express both positive and negative criteria in the database query.
+Therefore, the bipolar querying of databases~\cite{DeTre2009} is a querying technique which allows the user to express both positive and negative criteria in the database query.\\
 
-In this subsection we will address a proposal for the bipolar querying of temporal databases made by Billiet~\cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy}.
-First of all, the query structure is shown.
-Then, both evaluation and ranking proposed by the authors are explained.
-The model deals with a fuzzy valid-time specification as in~\cite{garrido2009}.
-
-\paragraph{Query structure}
-The approach in this work follows the query specification given in \eqref{eq:query-specification}.
-The query has a global time demand, $Q^{time}$ and here, the regular query constraints specified by $Q$ are split into positive and negative preferences:
-
-\begin{equation}
-\label{eq:bipolar-specification}
-\tilde{Q} = \left \lbrace Q^{time}, \left( Q^{pos}, Q^{neg} \right) \right \rbrace
-\end{equation}
-
-Here $Q^{pos}$ and $Q^{neg}$ represent the positive and the negative criteria, respectively.
-The temporal demand in $Q^{time}$ may be specified as in ~\cite{garrido2009} or as in ~ \cite{Pon11}, see section \ref{subsec:representation-time-intervals}.
-.
-\paragraph{Query evaluation}
-As explained previously in the query evaluation section (\ref{subsubsec:query-evaluation}), each element in $\tilde{Q}$ is evaluated independently:
+Bipolarity is handled by means of different tools:
 
@@
 \item
-The query $Q$ has now two elements: $Q^{pos}$ and $Q^{neg}$.
+The intuitionistic fuzzy sets  \emph{AFS} by Atannasov ~\cite{Atanassov1986}.
+The approach re-define the concept of inverse.
+Instead of considering the inverse of a set just the complementary, a AFS is defined by means of a membership function and a non-membership function.
+\item
+The interval valued fuzzy sets \emph{IVFS} were introduced independently by Zadeh~\cite{Zadeh75a}, Grattan-Guiness~\cite{Grattan76}, Janh~\cite{Jahn75}, Sambuc~\cite{Sambuc75} in the seventies~\cite{Dubois05}.
+%look for the reference!
+\item
+The two fold fuzzy sets \emph{(TFS)} by Dubois and Prade~\cite{Dubois02}. This approach defines desired and mandatory query conditions.
+\end{itemize}
+
+From a theoretical point of view, in a database, bipolarity may be found at:
+\begin{description}
+\item[\textbf{The query}] It is possible to distinguish between:
+	\begin{itemize}
+	\item
+	Inside criterion bipolarity: Each individual criteria may be specified with both positive and negative conditions. E.g. the user wants a black car, but definitively not a red neither a blue one.
+	\item
+	Outside criterion bipolarity: The query is specified with a global positive and negative conditions.
+E.g.
+The user wants a black car with a fuel consumption of 5 liters or less but definitively not a white car with a fuel consumption of 6 liters or more.
+	\end{itemize}
+\item[\textbf{The data storage}] In the database some attributes or even at tuple level it should be possible to specify them with bipolarity.
+Nevertheless, there is not so many research on this topic.
+%	\begin{itemize}
+%	\item
+%	Attribute bipolarity: Each attribute may be specified  
+%	\item
+%	Tuple bipolarity:
+%	\end{itemize}
+\end{description}
+
+A proposal for the bipolar querying of temporal databases has been made by Billiet~\cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy}.
+First of all, the query structure is shown.
+Then, both evaluation and ranking methods are defined.
+The model deals with a fuzzy valid-time specification in~\cite{garrido2009}.
+
+%\paragraph{Query structure}
+%The approach in this work follows the query specification given in \eqref{eq:query-specification}.
+The query has a global time demand, $Q^{time}$ and here, the regular query constraints specified by $Q$ are splitted into possitive and negative preferences:
+%
+%\begin{equation}
+%\label{eq:bipolar-specification}
+%\tilde{Q} = \left \lbrace Q^{time}, \left( Q^{pos}, Q^{neg} \right) \right \rbrace
+%\end{equation}
+%
+%Here $Q^{pos}$ and $Q^{neg}$ represent the positive and the negative criteria, respectively.
+The temporal demand in $Q^{time}$ may be specified as in ~\cite{garrido2009} or as in ~ \cite{Pon11}, see section \ref{subsec:representation-time-intervals}.
+
+%\paragraph{Query evaluation}
+%As explained previously in the query evaluation section (\ref{subsubsec:query-evaluation}), each element in $\tilde{Q}$ is evaluated independently:
+%
+%\begin{itemize}
+%\item
+%The query $Q$ has now two elements: $Q^{pos}$ and $Q^{neg}$.
 Each element is again evaluated independently and the result is a tuple $\left(s,d \right)$ in which $s$ is the satisfaction degree for $Q^{pos}$.
@@
 The tuple $\left(s,d \right) s,d \in \left[0,1 \right]$ is called \emph{Bipolar Satisfaction Degree}.
-\item
-The temporal specification is evaluated as explained in \ref{subsubsec:query-evaluation} and the result is a value in the unit interval.
-\end{itemize}
-
-\paragraph{Ranking}
-In order to present the results to the user, it is necessary to design a combination function.
-This function allows the classification of the results in the unit interval the preferred way is to made a convex combination as  presented in equation \eqref{eq:convex-combination}.
+%\item
+%The temporal specification is evaluated as explained in \ref{subsubsec:query-evaluation} and the result is a value in the unit interval.
+%\end{itemize}
+%
+%\paragraph{Ranking}
+%In order to present the results to the user, it is necessary to design a combination function.
+This function allows the classification of the results in the unit interval the preffered way is to made a convex combination as  presented in equation \eqref{eq:convex-combination}.
+%

### 2012-01-30 trunk/book/sources/temporalDatabases.tex
Ok, some modifications to the second chapter. Let me know what you think...

@@
 A very extensive list with the most well-known models in temporal databases can be found in ~\cite{Yu1998}.
-Nevertheless, it is neccesary to define some consistent way to query the temporal data.
+Nevertheless, it is necessary to define some consistent way to query the temporal data.
 There are several languages for querying temporal databases like TSQL~\cite{TSQL}.
@@
 
-\subsection{Comercial Temporal database systems}
+\subsection{Commercial Temporal database systems}
+There are several commercial temporal DBMS. Table \ref{table:commercial-temporal-db} is a comparative among the most well-known systems. It is also provided a reference for more information.
+
+Oracle workspace manager~\cite{oracle2009} and TimeDB~\cite{timedb2005} are libraries for dealing with time in an OracleDB.
+Nevertheless TimeDB and Postgree Temporal~\cite{posgree2009} are similar.
+Both are simple implementations that implement a subset of the Allen's operators and some operations for the creation and manipulation of a temporal attribute (valid, transaction or both times) in a table.
+Nevertheless, Teradata~\cite{teradata2011} is mainly a business intelligence database designed for data-mining.
+
+The most complete implementation is the Workspace Manager, which is described briefly in the following section.
+Unfortunately, none of these systems work with uncertainty or imprecision neither in the data storage nor in the query.
 
@@
 \subsubsection{Oracle Workspace Manager}
-Oracle workspace manager \cite{OraE118602} package allows to get several versions of the data in the same database. It is also possible to version only a table.\\
+
+%\begin{figure}
+%\centering
+%\includegraphics[scale=0.5]{graphs/workspaceTopology.pdf}
+%\caption{A sample topology in the workspace manager.}
+%\label{fig:workspace-topology}
+%\end{figure}
+
+Oracle workspace manager \cite{OraE118602} package allows to get several versions of the data in the same database. It is also possible to version only a table. The main benefits of this are two:
+
+\begin{itemize}
+\item
+Organization and optimization of data in a hierarchical frames: It is possible to create a frame of time (a workspace) and work only with that partial version of the data.
+It is provided a mechanim to merge data among different workspaces and to solve conflicts.
+The organization of data in these workspaces for very large tables result in smaller time access.
+\item
+Valid time as well as transaction time are managed by the system. Each update sentence makes a change in the versioning of a row.
+\end{itemize}
+
 When a table is versioned, the system creates a few tables and views as well as auxiliary structures to allow keep several version of the data, while keeping the primary key defined by the user.
-A more detailed explanation of
 
@@
 Workspaces are ordered in a hierarchy.
-The top level is called LIVE workspace.
-This is used by default and it must be used in production environment.\\
+The top level is called LIVE workspace (see figure \ref{fig:workspace-topology}).
+\\
+
 The system only makes a copy of a tuple if it is changed.
@@
 Also, it is not necessarily to modify the SQL code to access to versioned tables.\\
-When a table is versioned, it is renamed as table\_LT.
+%When a table is versioned, it is renamed as table\_LT.
 Here is stored all the data for the information together metadata for the versioning.

### 2012-01-30 trunk/book/sources/timeDomain.tex
Ok, some modifications to the second chapter. Let me know what you think...

--- 2011-11-09 trunk/book/sources/timeDomain.tex
@@
 %
-In order to define and to work with time it is necessary to study and understand the underlying domain.
+Before considering the introduction of temporal modelling to information systems, it is necessary to define and explain some main concepts concerning temporal modelling and their corresponding terminology, to situate these and to discuss some properties and issues related to these concepts.
+In this section, several basic concepts and corresponding terminology will be defined, explained and situated.
+Most of these concepts are widely used in the community of temporal databases and their definitions have been agreed upon in the context of~\cite{Dyreson1994}.
+For these concepts, in the entire chapter, the contents of~\cite{Dyreson1994} are followed (and often cited).
+
+In subsection \ref{subsec:basic-concepts}, some basic concepts of time modelling are presented and explained.
+
+%In order to define and to work with time it is necessary to study and understand the underlying domain.
 Moreover, the definition of a temporal domain is the basis for almost any temporal system.
 In the following subsection \ref{subsec:basic-concepts} we will introduce several concepts that are widely used in the community of temporal databases.
-Most of the concepts explained here have been introduced in the \emph{'Consensus Glossary of Temporal Database Concepts'}~\cite{Dyreson:1994:CGT:181550.181560}.
+Most of the concepts explained here have been introduced in the \emph{'Consensus Glossary of Temporal Database Concepts'}~\cite{Dyreson1994}.
 
@@
 \label{subsec:basic-concepts}
+In information systems, time itself is usually percieved as a linear or cyclic concept.
+Therefore, a time domains modelling time is usually represented by a set with an imposed partial order.
+In general, two main types of time models can be discerned: a \emph{linear} model and a \emph{cyclic} model.
+In the linear model, a total order is imposed on the set and the progress of time is seen as a linear matter, while cyclic models are mainly used in the modelling of recurrent processes.
+It should be noted that the majority of proposals use linear time models.
 
-Time is usually represented as a set with an imposed partial order.
-There are two main models of time: \textbf{linear} and \textbf{cyclic}.
-In the linear model, time advances from the past to the future in a totally ordered fashion (a relationship of total order is imposed to the set).
+Data models used by information systems (and in specific, temporal database systems) may represent these underlying time domains using \emph{chronons}.
+However, to explain what chronons are, an explanation of some other temporal concepts is necessary.
+
+\begin{svgraybox}
+\vspace{-10pt}
+\begin{definition}\textbf{(Instant)}\\
+An \emph{\textbf{instant}} is a time point on an underlying time axis.
+\end{definition}
+\vspace{-10pt}
+\end{svgraybox}
+
+Thus, an instant is basically an instantanious time point in a time model.
+
+Orthogonal to the classification of underlying time models as linear or cyclic, they can be classified as \emph{discrete}, \emph{dense} or \emph{continuous} models.
+In a discrete model, the notion exists that every instant has a unique successor and the set of instants is seen as a discrete one.
+Here, intuitively, the set of instants can be seen as isomorphic to the set of natural numbers $\Nat$.
+In a dense model, the notion exists that between any two instants always lies another, while the set of instants can still be seen as a discrete one.
+Here, intuitively, the set of instants can be seen as isomporphic to the set of rational numbers $\Q$ or the set of real numbers $\R$.
+In a continuous model, the notion also exists that between any two instants always lies another one, but unlike dense models, the set of instants is seen as continuous and there are no ``gaps'' between successive instants.
+
+Some other necessary concepts are:
+
+\begin{svgraybox}
+\vspace{-10pt}
+\begin{definition}\textbf{Time Interval}\\
+A \emph{\textbf{time interval}} is the time between two instants.
+\end{definition}
+
+\begin{definition}\textbf{Temporal Element}\\
+A \emph{\textbf{temporal element}} is a finite union of time intervals.
+\end{definition}
+
+\begin{definition}\textbf{Event}\\
+An \emph{\textbf{event}} is an instantaneous fact, i.e. something occurringat an instant.
+\end{definition}
+
+\begin{definition}\textbf{Duration}\\
+A \emph{\textbf{duration}} is an amount of time with known length, but no specific starting or ending instants.
+\end{definition}
+
+\begin{definition}\textbf{`Temporal' as Modifier}\\
+The modifier \emph{\textbf{`temporal'}} is used to indicate that the modified concept concerns some aspect of time.
+\end{definition}
+\vspace{-10pt}
+\end{svgraybox}
+
+%In the linear model, time advances from the past to the future in a totally ordered fashion (a relationship of total order is imposed to the set).
 A cyclic model of time is used for recurrent processes.
@@
 
+Data models might now represent an underlying time domain using chronons:
+
 \begin{svgraybox}
-The basic elements in a temporal domain are the following:\\
+\vspace{-10pt}
+\begin{definition}\textbf{Chronon}\\
+In a data model, a \emph{\textbf{chronon}} is a non-decomposable time interval of some fixed, minimal duration.
+\end{definition}
+\vspace{-10pt}
+\end{svgraybox}
+
+A data model may now represent a time model or time domain by a sequence of consecutive chronons.
+These chronons have identical durations.
+A data model will typically not specify the exact chronon duration, so it can be fixed later by applications implementing the data model.
+
+The fact that chronons are actually time intervals has a particular effect on the representation of instants and time intervals.
+In a data model representing a time domain using chronons, an instant is of course represented as a chronon, but one single chronon may represent multiple instants.
+A time interval may be represented by a set of contiguous chronons, depending on the amount of time the time interval comprises.
+
+%\begin{svgraybox}
+%The basic elements in a temporal domain are the following:\\
 %\begin{itemize}
 %\item
-\textbf{Instant}:  Is a time point over an underlying temporal domain.\\
+%\textbf{Instant}:  Is a time point over an underlying temporal domain.\\
 %\item
-\textbf{Time interval}: Is the time between two instants.\\
+%\textbf{Time interval}: Is the time between two instants.\\
 %\item
-\textbf{Event}: Is an instantaneous fact. Something that happens at an instant.\\
+%\textbf{Event}: Is an instantaneous fact. Something that happens at an instant.\\
 %\item
-\textbf{Chronon}: Is a non-decomposable time interval of fixed duration.
+%\textbf{Chronon}: Is a non-decomposable time interval of fixed duration.
 It is possible for a temporal model to leave the chronon duration unspecified.
@@
 %\end{itemize}
-\end{svgraybox}
+%\end{svgraybox}
 
-Attending to the density between the chronons, a model can be classified into the following three types:
+%\begin{definition}\textbf{(set of chronons)}
+%\label{def:set-chronon}
+%The \textbf{\emph{set of all the chronons}} in a system is denoted by $\C$. A chronon $c$ is therefore defined as:
+%\begin{equation}
+%\left \lbrace c \mid c \in \C \right \rbrace
+%\end{equation}
+%\end{definition}
 
-\begin{itemize}
-\item
-\textbf{Discrete model}:  The discrete model of time is isomorphic in relation to the natural numbers. According to this model, each natural number corresponds to a \emph{chronon}. 
-\item
-\textbf{Dense model}: This model is isomorphic in relation to the rational or the real numbers. It represents one instant of time in the gap between another two instants.
-\item
-\textbf{Continuous model}: This is an extension of the dense model with no gaps. It is isomorphic in relation to the real numbers.
-\end{itemize}
+Another classification of time models concerns the use of points or intervals to model time.
+The equivalence between interval-based and point-based time models is demonstrated in \cite{Böhlen_point-versus}.
 
+%Attending to the density in the set $\C$ of the chronons, a model can be classified into the following three types:
+
+%\begin{itemize}
+%\item
+%\textbf{Discrete model} ($\C \subset \Nat$):  The discrete model of time is isomorphic in relation to the natural numbers.
+%According to this model, each natural number corresponds to a \emph{chronon}.
+%\item
+%\textbf{Dense model} ($\C \subset \Q$ or $\C \subset \R$): This model is isomorphic in relation to the rational or the real %numbers.
+It represents one instant of time in the gap between another two instants.
+%\item
+%\textbf{Continuous model} ($\C \subset \R$): This is an extension of the dense model with no gaps. It is isomorphic in relation to the real numbers. 
+%\end{itemize}
+
+\textcolor{red}{CHECK}
 Restrictions on time range may be performed on the base of the following two concepts:
@@
 \end{itemize}
+\textcolor{red}{END CHECK}
 
+\textcolor{red}{MOVE TO TEMPORAL DATABASE PART}
 On the other hand, time may be \textbf{relative} or \textbf{absolute} (in other words, \textbf{anchored} or \textbf{unanchored}).
-Sometimes, what one considers 'absolute time' is not as definite as one would hope since our concept of absolute time is based on another time reference (for example, January 1st of year one).
+Sometimes, what one considers \emph{absolute time} is not as definite as one would hope since our concept of absolute time is based on another time reference (for example, January 1st of year one).
 Relative time has a direction, differently from distance.
@@
 -3 days = three days ago).
+\textcolor{red}{END MOVE}
 
-%%% time as points and time as intervals:
+Subsection \ref{subsec:granularity} gives a short survey on the concept of temporal granularities.
 
 %%% granularity
+\subsection{Granularities}\label{subsec:granularity}
+An important issue in time modelling concerns the concept of \emph{temporal granularities}. A formal definition for this concept is given in~\cite{Lin97}:
 
-\subsection{Temporal Relations}
-%an explanation on allen's temporal relations.
+\begin{svgraybox}
+\vspace{-10pt}
+\begin{definition}\textbf{(Granularity)}\\
+A \emph{\textbf{granularity}} is an ordered set of non-overlapping and continuous temporal elements called \emph{granules}.
+\end{definition}
 
-\subsection{An example of time domain: Julian Day Number}
+\begin{definition}\textbf{(Granule)}\\
+A \textbf{\emph{granule}} is the basic time unit in a granularity.
+\end{definition}
+\vspace{-10pt}
+\end{svgraybox}
+
+A temporal granularity is in fact a partitioning of the time line (time model) used by a system, usually dependent on the application.
+For example, the age of an adult human being is usually expressed in years: one will use sentences like `Laura is 21 years old' instead of sentences like `Laura is 21 years, 3 months and 4 days old'.
+In this example any duration shorter than a year needs no representation and thus the used granularity allows no specification for durations shorter than a year.
+The granules are years.
+
+As a granularity $G$ is an ordered set, each granule may be represented by an integer.
+In this representation, to keep track of the granularity a granule is an element of, the corresponding granularity is added in subscript:
+
+\begin{equation}
+G = \left\{i_{G}\ |\ i \in \Z \right\}
+\end{equation}
+
+%Next to the inherent difficulties of time, an issue is \textbf{temporal granularity}.
+A temporal granularity is a partitioning of the time line used by a system, usually dependent on the application.
+For example, the age of an adult human being is usually expressed in years: one will use sentences like \emph{`Laura is 21 years old'} instead of \emph{`Laura is 21 years, 3 months and 4 days old'}.
+In this example any time period shorter than a year needs no representation and thus the used granularity allows no specification for time periods shorter than a year.
+%The formal definition for granularity is given in ~\cite{Lin97}:
+
+%\begin{definition}
+%\label{def:granularity}\textbf{(granularity)}
+%A \textbf{\emph{granularity}} $\alpha$ is an ordered set of non-overlapping and continuous time elements called granules. \\
+%\end{definition}
+
+%As granularity $\alpha$ it is an ordered set, each \emph{granule} may be indexed by an integer number:\\
+%\begin{equation}
+%\alpha = \left \lbrace \ldots, 0_\alpha, 1_\alpha, \ldots \right \rbrace
+%\end{equation}
+
+In a system, the granularity with the shortest granules is the \emph{chronons granularity}, which is denoted as $\bot$.
+It is the granularity of which the granules are chronons.
+It is now possible to define a function $f$, called a \emph{mapping function}, that maps a given granule $i_G$, $i \in \Z$, from a given granularity $G$ to a set of corresponding chronons.
+The function definition should look like this:
+\vspace{-10pt}
+
+\begin{align}
+f & : G \rightarrow \Pow(\bot)\nonumber\\
+  & : i_{G} \rightarrow \lbrace c_\bot\ |\ (c_\bot\text{ is contained in }i_{G}) \wedge (c_\bot \in \bot)\rbrace\nonumber
+\end{align}
+
+For a given granularity $G$ and a given mapping function $f$ from $G$ to $\bot$, it is required that the following properties hold:
+
+\begin{itemize}
+\item
+$G$ \emph{is an ordered set}: for a given $i_{G} \in G$ holds: if $c_\bot \in f(i_G)$ and $c^{'}_\bot \in f(j_G)$, then $c_\bot < c^{'}_\bot$ implies that $i_G < j_G$.
+This is illustrated in figure \ref{fig:granularity-prop1}.
+\begin{figure}
+\centering
+\input{graphs/granularity-prop1.tex}
+\caption{$G$ is an ordered set}
+\label{fig:granularity-prop1}
+\end{figure}
+
+\item
+$G$ \emph{is a set of continuous granules}: for a given $i_{G} \in G$ and given chronons $c_\bot$, $c^{'}_\bot$ and $c^{''}_\bot$ holds: if $c_\bot < c^{'}_\bot < c^{''}_\bot$, then $c_\bot \in f(i_G)$ and $c^{''}_\bot \in f(i_G)$ implies that $c^{'}_\bot \in f(i_G)$.
+This is illustrated in figure \ref{fig:granularity-prop2}.
+\begin{figure}
+\centering
+\input{graphs/granularity-prop2.tex}
+\caption{$G$ is a set of continuous granules}
+\label{fig:granularity-prop2}
+\end{figure}
+
+\item
+\emph{The granules in} $G$ \emph{do not overlap}: for given $i_{G}\text{ and } j_{G} \in G$ holds: if $i_G \neq j_G$, then $f(i_G) \cap f(j_G) = \emptyset$.
+This is illustrated in figure \ref{fig:granularity-prop3}.
+\begin{figure}
+\centering
+\input{graphs/granularity-prop3.tex}
+\caption{The granules in $G$ do not overlap}
+\label{fig:granularity-prop3}
+\end{figure}
+\end{itemize}
+
+Mapping functions between granularities and the chronons granularity also allow comparing granularities with respect to the length of their granules.
+In this context, two important concepts can be discerned.
+%A granularity $G$ is said to be \emph{\textbf{finner than}} a granularity $H$ when the set of chronons on which a given granule $i_G$ is mapped by a mapping function $f$ has less elements than the set of chronons on which a given granule $j_H$ is mapped by $f$.
+%A granularity $\alpha$ is \emph{``coarser than''} granularity $\beta$ when the set of corresponding chronons for a given granule $i_\alpha$ is greater than the set of the corresponding chronons for a given granule $j_\beta$.
+
+\begin{svgraybox}
+\vspace{-10pt}
+\begin{definition}\textbf{(Finner Than)}\\
+Consider a mapping function $f$ and let $i_G$ and $j_H$ be elements of granularities $G$ and $H$ respectively. Granularity $G$ is now said to be \emph{\textbf{finner than}} granularity $H$ if:
+\begin{equation}
+\label{eq:finner-than}
+| f \left( i_G \right) |\ <\ |  f \left( j_H \right)| \nonumber
+\end{equation} 
+\end{definition}
+
+\begin{definition}
+\label{def:coarser-than}\textbf{(coarser than)}
+Consider a mapping function $f$ and let $i_G$ and $j_H$ be elements of granularities $G$ and $H$ repectively. Granularity $G$ is now said to be \emph{\textbf{coarser than}} granularity $H$ if:
+\begin{equation}
+\label{eq:coarser-than}
+| f \left( i_G \right)|\ >\ |  f \left( j_H \right)| \nonumber
+\end{equation} 
+\end{definition}
+\vspace{-10pt}
+\end{svgraybox}
+
+It is also possible to describe the relation between different granularities.
+For this, a function, in this chapter denoted as $cast$, may be specified which maps granules from a granularity $G$ to granules from a different granularity $H$.
+The mathematical notation concerning this function is now:
+\begin{equation}
+\label{eq:cast-function}
+\mbox{cast} \left( i_G,G,H \right) = j_H \nonumber
+\end{equation}
+This notation indicates that the function $cast$ associates a granule $i_G$ in $G$ to a corresponding granule $j_H$ in $H$.
+Two kinds of granularity-to-granularity mappings can now be discerned: an \emph{upwards mapping} is a mapping from a granularity $G$ to a coarser granularity $H$, whereas a downwards mapping is a mapping from a granularity $K$ to a finner granularity $L$.
+Different granularity-to-granularity mappings between several granularities can be represented using a granularity graph, which is a directed graph indicating the mapping conversions.
+Figure \ref{fig:granularity-graph-example} shows a granularity graph example.
+Mappings between two granularities may be classified as \emph{regular mappings}, \emph{irregular mappings} or \emph{congruent mappings}.
+\begin{itemize}
+\item
+\emph{Regular mapping}: A regular mapping is a granularity-to-granularity mapping where the mapping function between the two granularities is calculated by means of multiplications and/or divisions and (maybe) an anchor adjustment.
+E.g.
+the mapping between hours and minutes is calculated using a multiplication by $60$.
+Regular mappings are visualised in figure \ref{fig:granularity-graph-example} as normal arrows.
+\item
+\emph{Irregular mapping}: An irregular mapping is a granularity-to-granularity mapping where the mapping function can not be calculated by means of multiplications and/or divisions.
+E.g.
+the mapping between months and days is dependent on the exact month or day.
+Irregular mappings are visualised in figure \ref{fig:granularity-graph-example} as bold arrows.
+\item
+\emph{Congruent mapping}: A congruent mapping is a granularity-to-granularity mapping where the two granularities in the mapping have the same granules but a different anchor.
+E.g.
+the mapping between the days (Gregorian calendar days) and the academic days.
+Both concern the same days, but the academic year starts on octobre 1st, whereas the Gregorian calendar year starts on january 1st.
+Congruent mappings are visualised in figure \ref{fig:granularity-graph-example} as straight lines without arrow heads.
+\end{itemize}
+
+\textcolor{red}{MOVE TO CHAPTER 4}
+The transition between granularities~\cite{Lin97} is also considered as a source of imprecision.
+In a sentence like \emph{`The Alhambra is 612 years old'}, the temporal notion is very precise if time periods shorter than a year are irrelevant in the used granularity.
+However, if one is concerned with the exact day on which the last stone of the building is laid, a transition to a finer granularity, allowing the specification of time periods shorter than a year, is needed.
+Therefore, some proposals consider granularity as the base of the temporal model~\cite{Cru97}.
+
+\textcolor{red}{END MOVE TO CHAPTER 4}
+
+%\begin{figure}
+%\centering
+%\includegraphics[scale=0.5]{graphs/granularityGraph.pdf}
+%\caption{A granularity graph example}
+%\label{fig:granularity-graph-example}
+%\end{figure}
+
+Next, in subsection \ref{subsec:timedomain-calendar}, some basic concepts and issues concerning calendars are explained.
+
+\subsection{Calendars}
+\label{subsec:timedomain-calendar}
+%maybe introduce previously the concept of time units, linear hierarchy and time point.
+
+This subsection will concern \emph{calendars}.
+A calendar provides some sort of human interpretation of time.
+As such, calendars provide meaning or interpretation to temporal values where this meaning or interpretation is relevant or useful to the user.
+More precisely, calendars determine the mapping between human-meaningful time values and an underlying time model~\cite{Dyreson1994}.
+More formally, a calendar is an organization of different granularities (e.g.
+granularities day, month and year) in a hierarchy.
+In order to formally define what a calendar is, some other concepts should first be introduced~\cite{Kraus1997}:
+
+%\begin{svgraybox}
+%\vspace{-10pt}
+%\begin{definition}\textbf{(Calendar)}\\
+%A \emph{\textbf{calendar}} 
+%\end{definition}
+%\vspace{-10pt}
+%\end{svgraybox}
+
+\begin{svgraybox}
+\vspace{-10pt}
+\begin{definition}\textbf{(Linear Hierarchy of Granularities)}\\
+A \emph{\textbf{linear hierarchy of granularities}} is a finite set of granularities with a linear order (denoted $\sqsubset$).
+The most coarse granularity is called \emph{the top of the hierarchy} and must be an infinite set while all other granularities must be finite sets.
+\end{definition}
+\vspace{-10pt}
+\end{svgraybox}
+
+An example of a linear hierarchy is the set $\left\{ \text{days, months, years} \right\}$, where days $\sqsubset$ months and months $\sqsubset$ years.
+With this, a \emph{linear calendar} can be defined.
+
+\begin{svgraybox}
+\vspace{-10pt}
+\begin{definition}\textbf{(Linear Calendar)}\\
+A \emph{\textbf{linear calendar}} consist of a linear hierarchy of granularities and a validity predicate which specifies if a granule is valid in the calendar.
+\end{definition}
+\vspace{-10pt}
+\end{svgraybox}
+
+An example of a linear calendar could be the linear hierarchy $\left\{ \text{days, months, years} \right\}$, where days $\sqsubset$ months and months $\sqsubset$ years, together with a validity predicate that excludes time indications like `32 april 2012' or `30 february 2012'.
+
+%The linear hierarchy days $\sqsubset$ months $\sqsubset$ years represents a linear calendar.
+It is necessary to define the validity predicates because 17 January 2012 is a valid element of the calendar whereas 30 February 2012 is not.
+
+\textcolor{red}{CHECK}
+\subsubsection{\label{subsubsubsec:julian-day-number}An example of temporal domain: Julian Day Number}
 % proposal for the underlying Julian Day Number domain.
 
-The Julian Day number \cite{Dir96} is a counter.
+The Julian Day number \emph{JDN} \cite{Dir96} is a counter.
 Its value is incremented in one unit every day from 1 January 4713 B.C.
 at 12:00 noon.
-The particularity of starting at noon was useful for astronomers: the observations they took one night belonged to the same Julian Day.
-Actually it has not any transcendence.\\
+The particularity of starting at noon was useful for astronomers: the observations they took one night belonged to the same Julian Day.\\
 Note that the Julian Day represents whole days.
@@
 
-\subsubsection{Conversion among dates}
-\paragraph{Conversion between Gregorian and Julian dates}
-The Julian Date consist on the Julian Day number and the time. It is computed with the formula \cite{Usn}:\\
-\begin{equation}
-\label{eq:julian day}
-JD = JDN + \frac{hour-12}{24} + \frac{minutes}{1440} + \frac{seconds}{86400}
-\end{equation}
+\textcolor{red}{END CHECK}
 
-Here the JDN is the Julian Day Number.
-This equation does not take into account the leap seconds \footnote{A leap second(s) is the amount of second(s) added or deleted every year in the Universal Time.\cite{Wik},\cite{Lea}.
-}.
+The next section briefly introduces temporal relationships.
 
-\paragraph{Gregorian calendar to Julian day}
-Any date from Gregorian calendar \footnote{As Gregorian calendar entered in use in 1582, any use before that date is called \emph{proleptic calendar}.} can be converted into a Julian Day using the following formula:\\
-\begin{equation}
-\label{eq:JDN}
-JDN = \frac{1461x(Y + 4800 + \frac{M-14}{12})}{4}+\frac{367x(M-2-12x\frac{M-14}{12})}{12}-\frac{3x\frac{Y+4900+\frac{M-14}{12}}{100}}{4} 
-\end{equation}
+\subsection{Temporal Relationships}
+%an explanation on allen's temporal relations.
+%definition of fuzzy temporal interval
+In this section, a brief introduction can be found, concerning \emph{temporal relationships}, sometimes also called `temporal relations'\cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy}.
+Temporal relationships can be seen as relationships between temporal elements or instants.
 
-This formula was created by Henry Fliegel and Thomas C.
-van Flandern in 1968.
-\cite{Fliegel:1968:LEM:364096.364097} It takes into account leap years and that they are applicable only when the century is divisible by 400.
-The value $Y$ represents the year and the value $M$ represents the month.
-\paragraph{Julian calendar to Julian day}
-The algorithm to convert between Julian calendar and Julian Day Number is only valid for dates since 4713 B.C (positive values of Julian Day) \cite{Dog}:\\
-\begin{equation}
-\label{eq:JDN_2_JD}
-JDN=367Y-\frac{7 (Y + 5001 + \frac{M-9}{7})}{4}+\frac{275xM}{9}+D+1729777
-\end{equation}
+\def\JPicScale{0.5}
+\begin{figure}[h]
+\centering
+\input{graphs/allen.tex}
+\caption{Allen relations between two time intervals.}
+\label{fig:allen}
+\end{figure}
 
-Where $Y$ is the year value, $M$ is the month value, and $D$ is the day value.
+Several operators are defined in order to compare temporal elements or time points.
+Allen~\cite{Allen83} first described the relationships between time intervals and as a special case, between time points.
+Figure \ref{fig:allen} shows the temporal relationships Allen discerned.
+Some proposals can be applied to both crisp and fuzzy temporal intervals~\cite{ohlbach2004},~\cite{nagypal2003},~\cite{schockaert08}.
 
-%\subsubsection{Membership degree computation}
-%The main advantage from the Julian day number is the ability to make calculations. Thus, the membership degree computation for a fuzzy number based on a Julian day is explained here. \\
-%Let a trapezoidal possibility distribution defined by four JDN named $a$,$b$,$c$ and $d$.
-The fuzzy validity period \emph{FVP}, can be represented as in figure \ref{fig:fuzzy-validity-period}.
-The following equation shows the membership degree for a Julian day number \emph{JDN} in a \emph{FVP} represented by means of the possibility distribution of $[a,b,c,d]$.
-In a valid-time database, the values for the valid starting time \emph{VST} and the valid ending time \emph{VET} can be fuzzyfied as shown in equation \ref{eq:fuzzy-validity-period}.
-%
-%The membership function for a given $x$ number in \emph{JDN} format is the following:\\
-%\begin{equation}
-%\label{eq:trapezoidal distribution}
-%%\[
-%T(x)=
-%\begin{cases}
-%0, \mbox(if x \leq a, x \geq d)\\
-%1, \mbox(if x \in [b,c])\\
-%\frac{x-a}{b-a}   \mbox(if x \in (a,b])\\
-%\frac{d-x}{d-c} 	 \mbox(if x \in (c,d])\\
-%\end{cases}
-%%\]
-%\end{equation}
-%The points outside the trapezium or inside the interval $[a,b]$ are trivial. The points inside the other intervals has the following particularities:\\
-%The subtraction operation with Julian day numbers gives as result a scalar value (as explained before).
-The division operation between two scalars gives as result an scalar in the interval $[0,1]$ which represents the membership degree.
+\textcolor{red}{MOVE TO FUZZY PART}
+For this fuzzy comparisons~\cite{garrido2009} proposed some different temporal relations that can be easily implemented by means of regular fuzzy operations.
+\textcolor{red}{END MOVE TO FUZZY PART}
+
+In the next section, some basic concepts and terminology about temporal databases are presented, followed by an overview of some interesting issues concerning temporal databases and a survey of some commercial temporal database systems.

### 2012-02-16 ijcis-pos-model/ap-ijcis_sample_art.tex
style for ijcis

%% International Journal of Computational Intelligence Systems ---
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[11pt,twoside]{article}
\usepackage{ijcis}
%--------------------- ADDITIONAL PACKAGES HERE ---------------------------
% \usepackage{...}
%--------------------------------------------------------------------------
%
\def\labart{yourLabel}      % put a label from your choice here
%\Vol{1}                    % number of the Volume
%\Issue{1}                  % number of the issue
%\Month{January}            % month
%\Year{2008}                % year
%\received{...}
%\revised{...}
%
%---------------------------------------------------------------------------
\thispagestyle{empty}
%%------------------------- YOUR HEADINGS HERE -----------------------------
% Author's initials of first names+last names
\shortauthors{First Author, Second Author}
% Short title
\shorttitle{Instructions for Typesetting Manuscripts Using \LaTeX}
%---------------------------------------------------------------------------
%
%---------------------- YOUR TITLE ----------------------------------------
\title{%
PoVaTiM: A Possibilistic Valid-Time Model
 \\ \vspace*{0.04truein}
%Using \LaTeX\footnote{For the title, try not to use more than 3 lines.
%Typeset the title in 12 pt Times Roman, uppercase and boldface.}
}
%-------------------------- AUTHOR'S NAMES ----------------------------------
\author{%
Jose Enrique Pons\,\up{1}\,\,,
%\author{
Second Author\,\up{2}
}
%----------------------------- ADDRESSES ----------------------------------
\addresses{%
\up{1}
Department of Computer Science and Artificial Intelligence, University of Granada\\
C/ Periodista Daniel Saucedo, S/N,\\
Granada, , Spain%
\footnote{State completely without abbreviations, the affiliation
and mailing address, including country. Typeset in 10~pt Times Italic.}
\\ \vspace*{0.04truein}
E-mail: jpons@decsai.ugr.es
%}\address{%
\\ \vspace*{0.05truein}
\up{2}
Group, Company,\\
Address,\\ City, State ZIP/Zone, Country
\\ \vspace*{0.04truein}
E-mail: address
}
%---------------------------------------------------------------------------
\pagestyle{myheadings}
\begin{document}
\label{\labart-FirstPage}

\maketitle
%-------------------------- ABSTRACT ---------------------------------------
\abstracts{%
The abstract should summarize the context, content and conclusions of the
paper in less than 80 words. It should not contain any references or
displayed equations. Typeset the abstract in 9~pt Times Roman with
baselineskip of 11 pt, making an indentation of 2.5 picas on the left
and right margins.
}
\par\bigskip\par
%-------------------------- KEYWORDS ---------------------------------------
\keywords{List of four to six keywords which characterize the article.}

\vspace*{10pt}\textlineskip
%-------------------------- BEGIN BODY OF TEXT -----------------------------
\begin{multicols}{2}

\section{Introduction}
Contributions are to be in English. Authors are encouraged to
have their contribution checked for grammar. American or British
spelling should be used. Abbreviations are allowed but should be
spelt out in full when first used. Integers ten and below are to
be spelt out. Italicize foreign language phrases (e.g.~Latin,
French).

\section{The Main Text}
The text is to be typeset in 10 pt Times Roman, single spaced with
baselineskip of 13 pt. Text area (excluding running title) is 6.75
inches across and 8.8 inches deep. Final pagination and insertion of
running titles will be done by the publisher.

\section{Major Headings}
Major headings should be typeset in boldface with the first
letter of important words capitalized.

\subsection{Sub-headings}
%\vspace*{-4pt}  %only when needed
Sub-headings should be typeset in bolditalic with the first
letter of first word capitalized and section number in boldface.

%\vspace*{-1pt}  %only when needed
\subsubsection{Sub-subheadings}
%\vspace*{-4pt}  %only when needed

Typeset in italic (Section No. to be in Roman) and capitalize
the first letter of the first word only.

%\vspace*{-1pt}  %only when needed
\subsection{Numbering and spacing}
%\vspace*{-4pt}  %only when needed

Sections, sub-sections and sub-subsections are numbered in Arabic.
Use double spacing after major and subheadings, and single spacing
after\break sub-subheadings.

%\vspace*{-1pt}  %only when needed
\subsection{Lists of items}
%\vspace*{-4pt}  %only when needed

{Lists may be laid out with each item marked by\hfilneg}

\noindent a dot:
\begin{itemize}
\item item one,
\item item two.
\end{itemize}

\setcounter{footnote}{0}
\renewcommand{\thefootnote}{\alph{footnote}}

Items may also be numbered in lowercase Roman numerals:
\begin{enumerate}[(i)]\Nospacing
\item item one
\item item two
    \begin{enumerate}[(a)]\Nospacing
    \item Lists within lists can be numbered with lowercase Roman letters,
    \item second item.
    \end{enumerate}
\end{enumerate}

\subsection{Running Heads}

Please provide a shortened running head (not more than four words,
each starting with a Capital) for the title of your paper. This will
appear with page numbers on the top right-hand side of your paper on
odd pages.

For the running heads for the authors names should appear on your paper
on the even  pages. Please apply the following rules for theses running
heads:
\begin{itemize}
\item for one author: only the initial plus the full last name (e.g., D. Ruan),
\item for two authors: D. Ruan, T. Li,
\item for three authors or more: D. Ruan \emph{et al.}
\end{itemize}

\section{Equations}

Displayed equations should be numbered consecutively in each
section, with the number set flush right and enclosed in
parentheses.
\begin{equation}
\mu(n,t) = \dfrac{\sum^\infty_{i=1} 1(d_i < t, N(d_i) = n)}
{\int^t_{\sigma=0} 1(N(\sigma) = n)\mathrm{d}\sigma}\,\, .
\label{this}
\end{equation}
Equations should be referred to in abbreviated form,
e.g.~``Eq.~(\ref{this})'' or ``(2)''. In multiple-line equations,
the number should be given on the last line.

Displayed equations are to be centered on the page width.
Standard English letters like x are to appear as $x$ (italicized)
in the text if they are used as mathematical symbols. Punctuation
marks are used at the end of equations as if they appeared directly in the text.

\begin{theorem}
Theorems, lemmas, etc. are to be numbered consecutively in the
paper, just by using \\
\verb+\begin{env}My text here...\end{env}+\\
for the environments \emph{theorem}, \emph{lemma},
\emph{proposition}, and \emph{corollary}.\\
\end{theorem}

\vspace*{-5pt}

\begin{proof}
Proofs are produced with the command\\
\verb+\begin{proof}{My proof...}\end{proof}+.\\
It should end with\ $\qed$\kern0.3pt.
\end{proof}

\section{Illustrations and Photographs}

Figures are to be inserted in the text nearest their first reference.
Original india ink drawings or glossy prints are preferred. Please
send one set of originals with copies. If the author requires the
publisher to reduce the figures, ensure that {the figures (including
letterings and numbers) are large enough to be\hfilneg}

%\begin{figure}[htbp] %ORIGINAL SIZE: width=1.4TRUEIN; height=1.5TRUEIN
%\vspace*{13pt}
%\centerline{\psfig{file=ap-ijcis.eps}} %100 percent
\vspace*{10pt}
\fcaption{Labeled tree {\it T}.}
%\end{figure}
\vspace*{13pt}

\noindent
clearly seen after reduction. If photographs are to be used, only
black and white ones are acceptable.

Figures are to be sequentially numbered in Arabic numerals. The
caption must be placed below the figure. For those figures with
multiple parts which appear on different pages, it is best to
place the full caption below the first part, and have
e.g.~``Fig.~1. ({\it Continued})'' below the last part. Typeset
in 9~pt Times Roman with baselineskip of 11~pt. Use double
spacing between a caption and the text that follows immediately.

Previously published material must be accompanied by written
permission from the author and publisher.

\section{Tables}

Tables should be inserted in the text as close to the point of
reference as possible. Some space should be left above and below
the table. Tables should be numbered sequentially in the text in
Arabic numerals. Captions are to be centralized above the
tables. Typeset tables and captions in 9~pt Times Roman with
baselineskip of 11~pt.

If tables need to extend over to a second page, the continuation
of the table should be preceded by a caption, e.g.~``Table~2.
({\it Continued})''

\vglue13pt
%\begin{table}[htbp]
\tcap{Number of tests for WFF triple NA = 5, or\break NA = 8.}
\centerline{\small NP}
\vglue-6pt
\centerline{\small\baselineskip=13pt
\begin{tabular}{l c c c c c}\\
\hline
{} &{} &3 &4 &8 &10\\
\hline
{} &\phantom03 &1200 &2000 &\phantom02500 &\phantom03000\\
NC &\phantom05 &2000 &2200 &\phantom02700 &\phantom03400\\
{} &\phantom08 &2500 &2700 &16000 &22000\\
{} &10 &3000 &3400 &22000 &28000\\
\hline\\
\end{tabular}}
%\end{table}

\section{References}

References in the text are to be numbered consecutively in Arabic
numerals, in the order of first appearance. They are to be cited as
superscripts without parentheses or brackets after punctuation marks
like commas and periods but before punctuation marks like colons,
semi-colons and question marks. Where superscripts might cause
ambiguity, cite references in parentheses in abbreviated form,
e.g.~(Ref.~12).

\section{Footnotes}

Footnotes should be numbered sequentially in superscript
lowercase Roman letters.\fnm{a}\fnt{a}{Footnotes should be
typeset in 8~pt Times Roman at the bottom of the page.}

\section*{Note Added}

Additional note can be added before Acknowledgment.

\section*{Acknowledgments}

This section should come before the References. Funding
information may also be included here.

\appendix{}

Appendices should be used only when absolutely necessary. They should
come immediately\break before the References. If there is more than
one appendix, number them alphabetically. Number%\break
displayed equations occurring in the Appendix in this way, e.g.~(\ref{that}),
(A.2), etc.
\begin{equation}
\mu(n,t) = {\displaystyle\sum^\infty_{i=1} 1(d_i < t, N(d_i) = n) \over
\displaystyle\int^t_{\sigma=0} 1(N(\sigma) = n)d\sigma}\,\, .\label{that}
\end{equation}

\section*{References}
References are to be listed in the order cited in the text. Use the style shown
in the following examples. For journal names, use the standard abbreviations.
Typeset references in 9~pt Times Roman.

%----------------------------- END BODY OF TEXT -----------------------------

\begin{thebibliography}{000}
\bibitem{1}
J. J. Hopfield, ``Neurons with graded response have collective
computational properties like two-state neurons,'' {\it Proc. Natl. Acad. Sci.},
{\bf 81}, 3088--3092 (1984).

\bibitem{2}
D. W. Tank and J. J. Hopfield, ``Simple `neural' optimization networks:
An A/D converter, signal decision circuit, and a linear programming circuit,''
{\it IEEE Trans. on Circuits and Systems}, {\bf 33}, 533--541 (1986).

\bibitem{3}
Y. S. Foo and Y. Takefuji, ``Integer linear programming neural networks for
job-shop scheduling,'' {\it Proc. IEEE Intl. Conf. on Neural Networks},
{\bf II}, 341--348 (1988).
%\phantom{00}
\label{\labart-LastPage}
\end{thebibliography}
\end{multicols}
\end{document}

### 2012-02-17 trunk/book/sources/fuzzyTemporalDatabases.tex
fix: all the definitions are in a svgraybox and start with a capital letter.

@@
 
+\begin{svgraybox}
 \begin{definition}
-\textbf{(query specification)}
+\textbf{Query specification}
 A query $\tilde{Q}$ is given by:
@@
 \end{definition}
-
+\end{svgraybox}
 %%specify better this:
@@
 
+\begin{svgraybox}
 \begin{definition}
-\textbf{(temporal query specification)}
+\textbf{Temporal query specification}
 $Q^{time}$ is defined by:
@@
 \end{definition}
+\end{svgraybox}
 
@@
 
+\begin{svgraybox}
 \begin{definition}
 \label{def:evaluation-function}
-\textbf{(evaluation function)}
+\textbf{Evaluation function}
 An evaluation function $e_{Q} \left( r \right)$ is a mapping from the attribute value from the row $r$ in the database with respect to the constraints $Q = \left \lbrace q_1, \cdots, q_n \right \rbrace$ to the unit interval $\left[ 0,1 \right]$.

### 2012-02-17 trunk/book/sources/timeDomain.tex
fix: all the definitions are in a svgraybox and start with a capital letter.

@@
 \vspace{-10pt}
-\begin{definition}\textbf{(Instant)}\\
+\begin{definition}\textbf{Instant}\\
 An \emph{\textbf{instant}} is a time point on an underlying time axis.
@@
 \vspace{-10pt}
-\begin{definition}\textbf{(Granularity)}\\
+\begin{definition}\textbf{Granularity}\\
 A \emph{\textbf{granularity}} is an ordered set of non-overlapping and continuous temporal elements called \emph{granules}.
@@
 
-\begin{definition}\textbf{(Granule)}\\
+\begin{definition}\textbf{Granule}\\
 A \textbf{\emph{granule}} is the basic time unit in a granularity.
@@
 \vspace{-10pt}
-\begin{definition}\textbf{(Finner Than)}\\
+\begin{definition}\textbf{Finner Than}\\
 Consider a mapping function $f$ and let $i_G$ and $j_H$ be elements of granularities $G$ and $H$ respectively. Granularity $G$ is now said to be \emph{\textbf{finner than}} granularity $H$ if:
@@
 \begin{definition}
-\label{def:coarser-than}\textbf{(coarser than)}
+\label{def:coarser-than}\textbf{Coarser Than}
 Consider a mapping function $f$ and let $i_G$ and $j_H$ be elements of granularities $G$ and $H$ repectively. Granularity $G$ is now said to be \emph{\textbf{coarser than}} granularity $H$ if:
@@
 \vspace{-10pt}
-\begin{definition}\textbf{(Linear Hierarchy of Granularities)}\\
+\begin{definition}\textbf{Linear Hierarchy of Granularities}\\
 A \emph{\textbf{linear hierarchy of granularities}} is a finite set of granularities with a linear order (denoted $\sqsubset$).
@@
 \vspace{-10pt}
-\begin{definition}\textbf{(Linear Calendar)}\\
+\begin{definition}\textbf{Linear Calendar}\\
 A \emph{\textbf{linear calendar}} consist of a linear hierarchy of granularities and a validity predicate which specifies if a granule is valid in the calendar.

### 2012-02-21 trunk/book/sources/timeDomain.tex
bibliography update and cleanup.

@@
 \label{subsec:basic-concepts}
-In information systems, time itself is usually percieved as a linear or cyclic concept.
+In information systems, time itself is usually percieved as a linear or cyclic concept~\cite{Jensen94thetsql2}.
 Therefore, a time domains modelling time is usually represented by a set with an imposed partial order.
-In general, two main types of time models can be discerned: a \emph{linear} model and a \emph{cyclic} model.
+In general, two main types of time models can be discerned: a \emph{linear} model~\cite{benthem82} and a \emph{cyclic} model~\cite{lorentzos88}.
 In the linear model, a total order is imposed on the set and the progress of time is seen as a linear matter, while cyclic models are mainly used in the modelling of recurrent processes.
@@
 
-Data models used by information systems (and in specific, temporal database systems) may represent these underlying time domains using \emph{chronons}.
+Data models used by information systems (and in specific, temporal database systems) may represent these underlying time domains using \emph{chronons}~\cite{Dyreson1994}.
 However, to explain what chronons are, an explanation of some other temporal concepts is necessary.
@@
 \vspace{-10pt}
-\begin{definition}\textbf{Instant}\\
+\begin{definition}\textbf{Instant}~\cite{Dyreson1994}\\
 An \emph{\textbf{instant}} is a time point on an underlying time axis.
@@
 Orthogonal to the classification of underlying time models as linear or cyclic, they can be classified as \emph{discrete}, \emph{dense} or \emph{continuous} models.
-In a discrete model, the notion exists that every instant has a unique successor and the set of instants is seen as a discrete one.
+In a discrete model~\cite{Clifford:1985:AHR:971699.318922}, the notion exists that every instant has a unique successor and the set of instants is seen as a discrete one.
 Here, intuitively, the set of instants can be seen as isomorphic to the set of natural numbers $\Nat$.
@@
 
-Some other necessary concepts are:
+Some other necessary concepts are ~\cite{Dyreson1994}:
 
@@
 \textcolor{red}{CHECK}
-Restrictions on time range may be performed on the base of the following two concepts:
+Restrictions on time range may be performed on the base of the following two concepts~\cite{Jensen94thetsql2}:
 \begin{itemize}
@@
 \textcolor{red}{MOVE TO TEMPORAL DATABASE PART}
-On the other hand, time may be \textbf{relative} or \textbf{absolute} (in other words, \textbf{anchored} or \textbf{unanchored}).
+On the other hand, time may be~\cite{Jensen94thetsql2} \textbf{relative} or \textbf{absolute} (in other words, \textbf{anchored} or \textbf{unanchored}).
 Sometimes, what one considers \emph{absolute time} is not as definite as one would hope since our concept of absolute time is based on another time reference (for example, January 1st of year one).
@@
 \vspace{-10pt}
-\begin{definition}\textbf{Granularity}\\
+\begin{definition}\textbf{Granularity}~\cite{wang93}\\
 A \emph{\textbf{granularity}} is an ordered set of non-overlapping and continuous temporal elements called \emph{granules}.
@@
 
-For a given granularity $G$ and a given mapping function $f$ from $G$ to $\bot$, it is required that the following properties hold:
+For a given granularity $G$ and a given mapping function $f$ from $G$ to $\bot$, it is required that the following properties hold~\cite{Lin97}:
 

### 2012-02-22 trunk/ijcis-pos-model/sources/temporalModel.tex
first update.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Temporal model.tex
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{\label{subsec:time-interval-constraint}The time interval}
The representation of a temporal interval is given by two points: S the starting and E the ending points respectively.

\vglue13pt
%\begin{table}[htbp]
\tcap{Data types for a time point}
\centerline{\small DATA TYPES}
\vglue-6pt
\centerline{\small\baselineskip=13pt
\begin{tabular}{c p{2cm} p{2cm}}\\
\hline
1 & A single time point &  $1/x, x \in \T$\\
2 & A possibility distribution in the numeric domain & A fuzzy number or a fuzzy interval.\\
3 & An unknown value & \textbf{UNKNOWN}$= \lbrace1/t, t \in \T \rbrace$\\
4 & An undefined value & \textbf{UNDEFINED}$= \lbrace 0/t, t \in \T \rbrace$\\
5 & A null value & \textbf{NULL} $=\lbrace 1/$Unknown, 1/Undefined $\rbrace$\\
\hline\\
\end{tabular}
\label{tbl:time-point-types}
}
%\end{table}

\textcolor{red}{MAYBE COMPATIBILITY WITH GEFRED TYPES??}

A valid temporal interval $\left[S,E\right]$ in the system is a combination for the values of $S$,$E$ as shown in table \ref{tbl:valid-time-interval}.

\textcolor{red}{HAS IT ANY SENSE TO ALLOW THE COMBINATION [UNKNOWN, UNDEFINED] AND SO ON??}

\vglue13pt
%\begin{table}[htbp]
\tcap{Valid-Time interval data types}
\centerline{\small VALID TIME DATA TYPES}
\vglue-6pt
\centerline{\small\baselineskip=13pt
\begin{tabular}{c p{2cm} p{2cm}}\\
\hline
Type for S & Type for E & Semantics \\
\hline
Unknown & Unkown & Unknown interval. \\
Undefined & Undefined & Undefined interval. \\
Null & Null & Null interval. \\
\hline
1,2 & 1,2 & A possibilistic time interval. \\
\hline\\
\end{tabular}
\label{tbl:valid-time-interval}}
%\end{table}

The following is a study of the constraints in the range of values for both starting and ending times to represent a valid time interval.

The time period inside a valid time interval is denoted by two constraints:
\begin{eqnarray}
C_1\stackrel{\triangle}{=}\left(\geq,X\right)\\
C_2\stackrel{\triangle}{=}\left(\leq,Y\right).
\end{eqnarray}
Applying the inference of uncertainty as proposed in our general reasoning, we find for the first constraint that:
\begin{eqnarray}
\Pos\left(C_1([a,b])\right) &=& \min_{r\in[a,b]}\Big(\sup_{r\leq w}\pi_{X}(w)\Big)\\
\Nec\left(C_1([a,b])\right) &=& \min_{r\in[a,b]}\Big(\inf_{r>w}1-\pi_{X}(w)\Big)
\end{eqnarray}
which can be simplified to:
\begin{eqnarray}
\Pos\left(C_1([a,b])\right) &=& \sup_{a\leq w}\pi_{X}(w)\\
\Nec\left(C_1([a,b])\right) &=& \inf_{a>w}1-\pi_{X}(w).
\end{eqnarray}
For the second constraint, we find that:
\begin{eqnarray}
\Pos\left(C_2([a,b])\right) &=& \min_{r\in[a,b]}\Big(\sup_{r\geq w}\pi_{Y}(w)\Big)\\
\Nec\left(C_2([a,b])\right) &=& \min_{r\in[a,b]}\Big(\inf_{r<w}1-\pi_{Y}(w)\Big)
\end{eqnarray}
which can be simplified to:
\begin{eqnarray}
\Pos\left(C_2([a,b])\right) &=& \sup_{b\geq w}\pi_{Y}(w)\\
\Nec\left(C_2([a,b])\right) &=& \inf_{b<w}1-\pi_{Y}(w).
\end{eqnarray}
The uncertainty about the inclusion of an interval $I=[a,b]$ in the interval with ill-known boundaries can now be found by evaluating $[a,b]$ against the evaluation function $\lambda$ with $\bool=\wedge$. Application of \eqref{eq:conjunctive1} and \eqref{eq:conjunctive2} leads to:
\begin{eqnarray}
\Pos\left(\lambda([a,b])\right)&=&\min\bigg(\Pos(C_1([a,b])),\Pos(C_2([a,b]))\bigg)\\
\Nec\left(\lambda([a,b])\right)&=&\min\bigg(\Nec(C_1([a,b])),\Nec(C_2([a,b]))\bigg).
\end{eqnarray}
These last expressions can also be expanded as:
\begin{eqnarray}
\label{eq:interval-pos}
\Pos\left(\lambda([a,b])\right)&=&\min\bigg(\sup_{a\leq w}\pi_{X}(w),\sup_{b\geq w}\pi_{Y}(w)\bigg)\\
\label{eq:interval-nec}
\Nec\left(\lambda([a,b])\right)&=&\min\bigg(\inf_{a>w}1-\pi_{X}(w),\inf_{b<w}1-\pi_{Y}(w)\bigg).
\end{eqnarray}
Note that the interval $[X,Y]$ used here, is certainly not a fuzzy interval. Instead, we are dealing with an ill-known interval, i.e. it is a crisp interval, but it is partially unknown which values are in this interval. The uncertainty stems from the fact that the interval boundaries are ill-known.

Within this framework for the reasoning of the knowledge about the interval, we are going to study the semantics from the point of view of the valid time as well as from the point of view of the temporal database.

\subsubsection{S = Unknown, E = Unknown}
Consider that the possibility distribution for S is Unknown and for E is also Unknown. The possibility distributions are $\pi_S$ and $\pi_E$ respectively:

\begin{eqnarray}
\pi_{S}(t \in \T) = 1\\
\pi_{E}(t \in \T) = 1
\end{eqnarray}

Therefore the constraints $C_1$ and $C_2$ are:

\begin{eqnarray}
\Pos\left(C_1([a,b])\right) &=& \sup_{a\leq w}\pi_{S}(w) = 1\\
\Nec\left(C_1([a,b])\right) &=& \inf_{a>w}1-\pi_{S}(w) = 0\\
\Pos\left(C_2([a,b])\right) &=& \sup_{b\geq w}\pi_{E}(w)=1\\
\Nec\left(C_2([a,b])\right) &=& \inf_{b<w}1-\pi_{E}(w)=0
\end{eqnarray}

\begin{eqnarray}
\Pos\left(\lambda([a,b])\right)&=&\min\bigg(\sup_{a\leq w}\pi_{S}(w),\sup_{b\geq w}\pi_{E}(w)\bigg) \\
&=& \min(1,1)  = 1\\
\Nec\left(\lambda([a,b])\right)&=&\min\bigg(\inf_{a>w}1-\pi_{S}(w),\inf_{b<w}1-\pi_{E}(w)\bigg)\\
&=& \min(0,0) = 0
\end{eqnarray}

\begin{example}
A constant object in the database. It is an object that is always valid in the database.
The main issue in a valid time database is that the update sentence have to modify the starting or ending point distribution. The most natural way to convert a snapshot database to a temporal database is to consider the objects belonging to the interval [Unknown, Unknown].
\end{example}

\subsubsection{[Unknown,value] or [value,Unknown]}
These two cases are managed in this section. Here we will note $\pi_{S}(t)$ or $\pi_{E}(t)$ as the possibility distribution for 'value' in the starting or ending points.

First of all the case [Unknown,value] in \eqref{eq:interval-pos} and in \eqref{eq:interval-nec}:

\begin{eqnarray}
\Pos\left(\lambda([a,b])\right)&=&\min (1,\sup_{b\geq w}\pi_{E}(w))\\
 &=& \sup_{b\geq w}\pi_{E}(w)\\
\Nec\left(\lambda([a,b])\right)&=&\min (0,\inf_{b<w}1-\pi_{E}(w))\\
&=& 0
\end{eqnarray}

Similarily for [value,Unknown] the possibility and neccesity are:
\begin{eqnarray}
\Pos\left(\lambda([a,b])\right)&=&\min (\sup_{a\leq w}\pi_{S}(w),1)\\
 &=& \sup_{a\leq w}\pi_{S}(w)\\
\Nec\left(\lambda([a,b])\right)&=&\min (\inf_{a>w}1-\pi_{S}(w),0)\\
&=& 0
\end{eqnarray}

The semantics for an object with valid time given by [Unknown,value] is an object that has been valid from the beggining to an ending time. Analogously, the semantics for an object with a valid time given by the interval [value,Unknown] is an object that it is still valid in the database.

\subsubsection{[Undefined,Undefined]}
Maybe this combination has no sense for a valid time object. The possibility distributions for both starting and ending points are:

\begin{eqnarray}
\pi_{S}(t \in \T) = 0\\
\pi_{E}(t \in \T) = 0
\end{eqnarray}

Therefore the constraints $C_1$ and $C_2$ are:

\begin{eqnarray}
\Pos\left(C_1([a,b])\right) &=& \sup_{a\leq w}\pi_{S}(w) = 0\\
\Nec\left(C_1([a,b])\right) &=& \inf_{a>w}1-\pi_{S}(w) = 1\\
\Pos\left(C_2([a,b])\right) &=& \sup_{b\geq w}\pi_{E}(w)=0\\
\Nec\left(C_2([a,b])\right) &=& \inf_{b<w}1-\pi_{E}(w)=1
\end{eqnarray}

Taking these values into  \eqref{eq:interval-pos} and in \eqref{eq:interval-nec}:
\begin{eqnarray}
\Pos\left(\lambda([a,b])\right)&=&\min (0,0)\\
 &=& 0\\
\Nec\left(\lambda([a,b])\right)&=&\min (1,1)\\
&=& 1
\end{eqnarray}

\textcolor{red}{TAKE A LOOK INTO THIS}

Which is an inconsistency.

\subsection{[Undefined,value] or [value,Undefined]}

The possibility and neccesity measures for [Undefined,value] are :
\begin{eqnarray}
\Pos\left(\lambda([a,b])\right)&=&\min (0,\sup_{b\geq w}\pi_{E}(w))\\
 &=& 0\\
\Nec\left(\lambda([a,b])\right)&=&\min (1,\inf_{b<w}1-\pi_{E}(w))\\
&=& \inf_{b<w}1-\pi_{E}(w)
\end{eqnarray}

The other way around, [value,Undefined]:
\begin{eqnarray}
\Pos\left(\lambda([a,b])\right)&=&\min (\sup_{a\leq w}\pi_{S}(w),0)\\
 &=& 0\\
\Nec\left(\lambda([a,b])\right)&=&\min (\inf_{a>w}1-\pi_{S}(w),1)\\
&=& \inf_{a>w}1-\pi_{S}(w)
\end{eqnarray}

\textcolor{red}{TAKE A LOOK INTO THIS}
This is also an inconsistency.

\subsubsection{[NULL,NULL] and maybe any combination with NULL}
I think the framework we have can not deal with NULL computations. Maybe just for non-specified time or something like that.

\subsubsection{[Undefined,Unknown] or [Unknown,Undefined]}
The semantic for this could be also unconsistent.

\begin{example}
Consider a snapshot database (table \ref{tbl:snapshot-car-models}) that stores information about car models. The primary key is given by the set $\lbrace$ Brand, Model $\rbrace$.

\vglue13pt
%\begin{table}[htbp]
\tcap{Car model snapshot database}
\centerline{\small CAR MODELS}
\vglue-6pt
\centerline{\small\baselineskip=13pt
\begin{tabular}{c c c}\\
\hline
\textbf{Brand} & \textbf{Model} & Segment \\
\hline
Peugeot & 205 & B \\
Peugeot & 206 & B \\
Peugeot & 207 & B \\
Peugeto & 208 & B \\
Citroen & C2 & B \\
\hline\\
\end{tabular}
\label{tbl:snapshot-car-models}
}

Consider now that the user wants to make that table a valid-time table, then, from the point of view of the storage, the following changes should be done:

\begin{itemize}
\item
A new definition for the primary key, including a version field.
\item
The addition of two new columns to store the valid-time interval. By default the initial values are Unknown for both starting and ending points of the interval.
\end{itemize}
Therefore, the table \ref{tbl:snapshot-car-models} is transformed into table \ref{tbl:initial-valid-time-car-models}:

 \vglue13pt
%\begin{table}[htbp]
\tcap{Car model valid-time database. Note that U is the abbreviation for the Unknown constant.}
\centerline{\small VALID TIME CAR MODELS}
\vglue-6pt
\centerline{\small\baselineskip=13pt
\begin{tabular}{c c c c c c}\\
\hline
\textbf{Brand} & \textbf{Model} & \textbf{VersionID} & Segment & S & E \\
\hline
Peugeot & 205 & 001 & B & U & U \\
Peugeot & 206 & 001 & B & U & U \\
Peugeot & 207 & 001 & B & U & U \\
Peugeto & 208 & 001 & B & U & U \\
Citroen & C2 & 001 & B & U & U \\
\hline\\
\end{tabular}
\label{tbl:initial-valid-time-car-models}
}

\end{example}

\subsection{\label{subsec:temporal-model}The generalized temporal model}

\begin{definition}
Generalized fuzzy temporal domain.\\
If $\T$ is the temporal domain, $\tilde \Pow\left( \T\right)$ is the set of all possibility distributions defined on $\T$.
The \textbf{Generalized Fuzzy Temporal Domain}, $\T_G$ is
\begin{equation}
\T_G \subseteq \left \lbrace \tilde \Pow\left( \T\right) \cup NULL \right \rbrace
\end{equation}
\end{definition}

### 2012-02-24 trunk/book/sources/timeDomain.tex
update for the time domain and the temporal databases

@@
 \label{subsec:basic-concepts}
-In information systems, time itself is usually percieved as a linear or cyclic concept~\cite{Jensen94thetsql2}.
+In information systems, time itself is usually perceived as a linear or cyclic concept~\cite{Jensen94thetsql2}.
 Therefore, a time domains modelling time is usually represented by a set with an imposed partial order.
@@
 
-Thus, an instant is basically an instantanious time point in a time model.
+Thus, an instant is basically an instantaneous time point in a time model.
 
@@
 \begin{definition}\textbf{Event}\\
-An \emph{\textbf{event}} is an instantaneous fact, i.e. something occurringat an instant.
+An \emph{\textbf{event}} is an instantaneous fact, i.e. something occurring at an instant.
 \end{definition}
@@
 
-In a system, the granularity with the shortest granules is the \emph{chronons granularity}, which is denoted as $\bot$.
-It is the granularity of which the granules are chronons.
-It is now possible to define a function $f$, called a \emph{mapping function}, that maps a given granule $i_G$, $i \in \Z$, from a given granularity $G$ to a set of corresponding chronons.
+In a system, the granularity with the shortest granules is the \emph{chronons granularity}, which is denoted as $\bot$. It is the granularity of which the granules are chronons. 
+%It is now possible to define a function $f$, called a \emph{mapping function}, that maps a given granule $i_G$, $i \in \Z$, from a given granularity $G$ to a set of corresponding chronons.
 The function definition should look like this:
+
+\begin{svgraybox}
+\begin{definition}
+\label{def:mapping-function}
+\textbf{Mapping function}~\cite{Lin97}\\ 
+A mapping function $f$ is a function that translates from a given granule $i_G$, $i \in \Z$, in a given granularity $G$, to a set of corresponding chronons.\\
+
 \vspace{-10pt}
@@
 \end{align}
+\end{definition}
+\end{svgraybox}
 
-For a given granularity $G$ and a given mapping function $f$ from $G$ to $\bot$, it is required that the following properties hold~\cite{Lin97}:
+Note that a mapping function $f$ always maps from a given granularity $G$ to the set of chronons $\bot$. Therefore, the output for a mapping function is a subset of $\Pow(\bot)$. \\
+ A mapping function $f$ requires that the following properties hold~\cite{Lin97}:
 
-\begin{itemize}
-\item
+%
+%
+%For a given granularity $G$ and a given mapping function $f$ from $G$ to $\bot$, it is required that 
+\begin{property}
+%\item
 $G$ \emph{is an ordered set}: for a given $i_{G} \in G$ holds: if $c_\bot \in f(i_G)$ and $c^{'}_\bot \in f(j_G)$, then $c_\bot < c^{'}_\bot$ implies that $i_G < j_G$.
@@
 \end{figure}
+\end{property}
 
-\item
+%\item
+\begin{property}
 $G$ \emph{is a set of continuous granules}: for a given $i_{G} \in G$ and given chronons $c_\bot$, $c^{'}_\bot$ and $c^{''}_\bot$ holds: if $c_\bot < c^{'}_\bot < c^{''}_\bot$, then $c_\bot \in f(i_G)$ and $c^{''}_\bot \in f(i_G)$ implies that $c^{'}_\bot \in f(i_G)$.
@@
 \end{figure}
+\end{property}
 
-\item
+%\item
+\begin{property}
 \emph{The granules in} $G$ \emph{do not overlap}: for given $i_{G}\text{ and } j_{G} \in G$ holds: if $i_G \neq j_G$, then $f(i_G) \cap f(j_G) = \emptyset$.
@@
 \end{figure}
-\end{itemize}
+\end{property}
+%\end{itemize}
 
@@
 \label{def:coarser-than}\textbf{Coarser Than}
-Consider a mapping function $f$ and let $i_G$ and $j_H$ be elements of granularities $G$ and $H$ repectively. Granularity $G$ is now said to be \emph{\textbf{coarser than}} granularity $H$ if:
+Consider a mapping function $f$ and let $i_G$ and $j_H$ be elements of granularities $G$ and $H$ respectively. Granularity $G$ is now said to be \emph{\textbf{coarser than}} granularity $H$ if:
 \begin{equation}
@@
 
-It is also possible to describe the relation between different granularities.
-For this, a function, in this chapter denoted as $cast$, may be specified which maps granules from a granularity $G$ to granules from a different granularity $H$.
-The mathematical notation concerning this function is now:
+It is also possible to describe the relation between different granularities. This is called a casting function:
+\begin{svgraybox}
+\begin{definition}
+\textbf{Casting function}~\cite{Lin97}\\
+\label{def:casting-fucntion}
+A casting function $cast$ maps granules from a granularity $G$ to granules from a different granularity $H$.
 \begin{equation}
@@
 \end{equation}
+\end{definition}
+\end{svgraybox}
+
 This notation indicates that the function $cast$ associates a granule $i_G$ in $G$ to a corresponding granule $j_H$ in $H$.
@@
 Different granularity-to-granularity mappings between several granularities can be represented using a granularity graph, which is a directed graph indicating the mapping conversions.
-Figure \ref{fig:granularity-graph-example} shows a granularity graph example.
+
  Mappings between two granularities may be classified as \emph{regular mappings}, \emph{irregular mappings} or \emph{congruent mappings}.
@@
 the mapping between hours and minutes is calculated using a multiplication by $60$.
-Regular mappings are visualised in figure \ref{fig:granularity-graph-example} as normal arrows.
 \item
@@
 the mapping between months and days is dependent on the exact month or day.
-Irregular mappings are visualised in figure \ref{fig:granularity-graph-example} as bold arrows.
 \item
@@
 the mapping between the days (Gregorian calendar days) and the academic days.
-Both concern the same days, but the academic year starts on octobre 1st, whereas the Gregorian calendar year starts on january 1st.
-Congruent mappings are visualised in figure \ref{fig:granularity-graph-example} as straight lines without arrow heads.
 \end{itemize}
+
+\begin{example}
+Consider a systems that models both Gregorian Calendar dates as well as Academic dates.
+The chronons granularity in this system are milliseconds.
+Figure \ref{fig:granularity-graph-example} shows the complete granularity graph example.
+The transitions between Chronons and Seconds and between Seconds and Minutes and so on is called a \emph{regular mapping}.
+The transition between Days and Months is not defined by a multiplication or division, this is called a \emph{irregular mapping}.
+In the graph, these \emph{irregular mappings} are noted with a bold arrow.
+Finally, the transition between a Gregorian Calendar Day and an Academic Day is called a \emph{Congruent mapping}: two granularities with the same granules but with different anchors.
+Both concern the same days, but the academic year starts on October 1st, whereas the Gregorian calendar year starts on January 1st.
+Congruent mappings are visualised as straight lines without arrow heads.
+
+\end{example}
 
@@
 
-%\begin{figure}
-%\centering
-%\includegraphics[scale=0.5]{graphs/granularityGraph.pdf}
-%\caption{A granularity graph example}
-%\label{fig:granularity-graph-example}
-%\end{figure}
+\begin{figure}
+\centering
+\includegraphics[scale=0.5]{graphs/granularityGraph.eps}
+\caption{A granularity graph example}
+\label{fig:granularity-graph-example}
+\end{figure}
 
@@
 
-An example of a linear calendar could be the linear hierarchy $\left\{ \text{days, months, years} \right\}$, where days $\sqsubset$ months and months $\sqsubset$ years, together with a validity predicate that excludes time indications like `32 april 2012' or `30 february 2012'.
+An example of a linear calendar could be the linear hierarchy $\left\{ \text{days, months, years} \right\}$, where days $\sqsubset$ months and months $\sqsubset$ years, together with a validity predicate that excludes time indications like `32 April 2012' or `30 February 2012'.
 
@@
 or J.D.E.).
-Any time scale different from Universal Time must be explicited after the Julian Day number.
+Any time scale different from Universal Time must be explicit after the Julian Day number.
 There are several conversion formulas~\cite{Usn}\cite{Wik}\cite{Lea} between a date in Gregorian calendar  format and a date in  JDN format.

### 2012-02-25 book/sources/timeDomain.tex
update for the graphs/allen.eps

--- 2012-01-10 book/sources/timeDomain.tex
@@
 \label{subsec:basic-concepts}
+In information systems, time itself is usually perceived as a linear or cyclic concept~\cite{Jensen94thetsql2}.
+Therefore, a time domains modelling time is usually represented by a set with an imposed partial order.
+In general, two main types of time models can be discerned: a \emph{linear} model~\cite{benthem82} and a \emph{cyclic} model~\cite{lorentzos88}.
+In the linear model, a total order is imposed on the set and the progress of time is seen as a linear matter, while cyclic models are mainly used in the modelling of recurrent processes.
+It should be noted that the majority of proposals use linear time models.
 
-Time is usually represented as a set with an imposed partial order.
-There are two main models of time: \textbf{linear} and \textbf{cyclic}.
-In the linear model, time advances from the past to the future in a totally ordered fashion (a relationship of total order is imposed to the set).
+Data models used by information systems (and in specific, temporal database systems) may represent these underlying time domains using \emph{chronons}~\cite{Dyreson1994}.
+However, to explain what chronons are, an explanation of some other temporal concepts is necessary.
+
+\begin{svgraybox}
+\vspace{-10pt}
+\begin{definition}\textbf{Instant}~\cite{Dyreson1994}\\
+An \emph{\textbf{instant}} is a time point on an underlying time axis.
+\end{definition}
+\vspace{-10pt}
+\end{svgraybox}
+
+Thus, an instant is basically an instantaneous time point in a time model.
+
+Orthogonal to the classification of underlying time models as linear or cyclic, they can be classified as \emph{discrete}, \emph{dense} or \emph{continuous} models.
+In a discrete model~\cite{Clifford:1985:AHR:971699.318922}, the notion exists that every instant has a unique successor and the set of instants is seen as a discrete one.
+Here, intuitively, the set of instants can be seen as isomorphic to the set of natural numbers $\Nat$.
+In a dense model, the notion exists that between any two instants always lies another, while the set of instants can still be seen as a discrete one.
+Here, intuitively, the set of instants can be seen as isomporphic to the set of rational numbers $\Q$ or the set of real numbers $\R$.
+In a continuous model, the notion also exists that between any two instants always lies another one, but unlike dense models, the set of instants is seen as continuous and there are no ``gaps'' between successive instants.
+
+Some other necessary concepts are ~\cite{Dyreson1994}:
+
+\begin{svgraybox}
+\vspace{-10pt}
+\begin{definition}\textbf{Time Interval}\\
+A \emph{\textbf{time interval}} is the time between two instants.
+\end{definition}
+
+\begin{definition}\textbf{Temporal Element}\\
+A \emph{\textbf{temporal element}} is a finite union of time intervals.
+\end{definition}
+
+\begin{definition}\textbf{Event}\\
+An \emph{\textbf{event}} is an instantaneous fact, i.e. something occurring at an instant.
+\end{definition}
+
+\begin{definition}\textbf{Duration}\\
+A \emph{\textbf{duration}} is an amount of time with known length, but no specific starting or ending instants.
+\end{definition}
+
+\begin{definition}\textbf{`Temporal' as Modifier}\\
+The modifier \emph{\textbf{`temporal'}} is used to indicate that the modified concept concerns some aspect of time.
+\end{definition}
+\vspace{-10pt}
+\end{svgraybox}
+
+%In the linear model, time advances from the past to the future in a totally ordered fashion (a relationship of total order is imposed to the set).
 A cyclic model of time is used for recurrent processes.
@@
 %\end{itemize}
-\end{svgraybox}
-Any temporal model may represent the time as points as well as intervals. The equivalence between a interval-based and point-based temporal data models is demonstrated in \cite{Böhlen_point-versus}.
+%\end{svgraybox}
 
-Attending to the density between the chronons, a model can be classified into the following three types:
+%\begin{definition}\textbf{(set of chronons)}
+%\label{def:set-chronon}
+%The \textbf{\emph{set of all the chronons}} in a system is denoted by $\C$. A chronon $c$ is therefore defined as:
+%\begin{equation}
+%\left \lbrace c \mid c \in \C \right \rbrace
+%\end{equation}
+%\end{definition}
 
-\begin{itemize}
-\item
-\textbf{Discrete model}:  The discrete model of time is isomorphic in relation to the natural numbers. According to this model, each natural number corresponds to a \emph{chronon}. 
-\item
-\textbf{Dense model}: This model is isomorphic in relation to the rational or the real numbers. It represents one instant of time in the gap between another two instants.
-\item
-\textbf{Continuous model}: This is an extension of the dense model with no gaps. It is isomorphic in relation to the real numbers.
-\end{itemize}
+Another classification of time models concerns the use of points or intervals to model time.
+The equivalence between interval-based and point-based time models is demonstrated in \cite{Böhlen_point-versus}.
 
-Restrictions on time range may be performed on the base of the following two concepts:
+%Attending to the density in the set $\C$ of the chronons, a model can be classified into the following three types:
+
+%\begin{itemize}
+%\item
+%\textbf{Discrete model} ($\C \subset \Nat$):  The discrete model of time is isomorphic in relation to the natural numbers.
+%According to this model, each natural number corresponds to a \emph{chronon}.
+%\item
+%\textbf{Dense model} ($\C \subset \Q$ or $\C \subset \R$): This model is isomorphic in relation to the rational or the real %numbers.
+It represents one instant of time in the gap between another two instants.
+%\item
+%\textbf{Continuous model} ($\C \subset \R$): This is an extension of the dense model with no gaps. It is isomorphic in relation to the real numbers. 
+%\end{itemize}
+
+\textcolor{red}{CHECK}
+Restrictions on time range may be performed on the base of the following two concepts~\cite{Jensen94thetsql2}:
 \begin{itemize}
@@
 \end{itemize}
+\textcolor{red}{END CHECK}
 
-On the other hand, time may be \textbf{relative} or \textbf{absolute} (in other words, \textbf{anchored} or \textbf{unanchored}).
+\textcolor{red}{MOVE TO TEMPORAL DATABASE PART}
+On the other hand, time may be~\cite{Jensen94thetsql2} \textbf{relative} or \textbf{absolute} (in other words, \textbf{anchored} or \textbf{unanchored}).
 Sometimes, what one considers \emph{absolute time} is not as definite as one would hope since our concept of absolute time is based on another time reference (for example, January 1st of year one).
@@
 -3 days = three days ago).
+\textcolor{red}{END MOVE}
+
+Subsection \ref{subsec:granularity} gives a short survey on the concept of temporal granularities.
 
 %%% granularity
-%\subsubsection{Granularity}
-Next to the inherent difficulties of time, an issue is \textbf{temporal granularity}.
+\subsection{Granularities}\label{subsec:granularity}
+An important issue in time modelling concerns the concept of \emph{temporal granularities}. A formal definition for this concept is given in~\cite{Lin97}:
+
+\begin{svgraybox}
+\vspace{-10pt}
+\begin{definition}\textbf{Granularity}~\cite{wang93}\\
+A \emph{\textbf{granularity}} is an ordered set of non-overlapping and continuous temporal elements called \emph{granules}.
+\end{definition}
+
+\begin{definition}\textbf{Granule}\\
+A \textbf{\emph{granule}} is the basic time unit in a granularity.
+\end{definition}
+\vspace{-10pt}
+\end{svgraybox}
+
+A temporal granularity is in fact a partitioning of the time line (time model) used by a system, usually dependent on the application.
+For example, the age of an adult human being is usually expressed in years: one will use sentences like `Laura is 21 years old' instead of sentences like `Laura is 21 years, 3 months and 4 days old'.
+In this example any duration shorter than a year needs no representation and thus the used granularity allows no specification for durations shorter than a year.
+The granules are years.
+
+As a granularity $G$ is an ordered set, each granule may be represented by an integer.
+In this representation, to keep track of the granularity a granule is an element of, the corresponding granularity is added in subscript:
+
+\begin{equation}
+G = \left\{i_{G}\ |\ i \in \Z \right\}
+\end{equation}
+
+%Next to the inherent difficulties of time, an issue is \textbf{temporal granularity}.
 A temporal granularity is a partitioning of the time line used by a system, usually dependent on the application.
@@
 In this example any time period shorter than a year needs no representation and thus the used granularity allows no specification for time periods shorter than a year.
+%The formal definition for granularity is given in ~\cite{Lin97}:
 
+%\begin{definition}
+%\label{def:granularity}\textbf{(granularity)}
+%A \textbf{\emph{granularity}} $\alpha$ is an ordered set of non-overlapping and continuous time elements called granules. \\
+%\end{definition}
+
+%As granularity $\alpha$ it is an ordered set, each \emph{granule} may be indexed by an integer number:\\
+%\begin{equation}
+%\alpha = \left \lbrace \ldots, 0_\alpha, 1_\alpha, \ldots \right \rbrace
+%\end{equation}
+
+In a system, the granularity with the shortest granules is the \emph{chronons granularity}, which is denoted as $\bot$. It is the granularity of which the granules are chronons. 
+%It is now possible to define a function $f$, called a \emph{mapping function}, that maps a given granule $i_G$, $i \in \Z$, from a given granularity $G$ to a set of corresponding chronons.
+The function definition should look like this:
+
+\begin{svgraybox}
+\begin{definition}
+\label{def:mapping-function}
+\textbf{Mapping function}~\cite{Lin97}\\ 
+A mapping function $f$ is a function that translates from a given granule $i_G$, $i \in \Z$, in a given granularity $G$, to a set of corresponding chronons.\\
+
+\vspace{-10pt}
+
+\begin{align}
+f & : G \rightarrow \Pow(\bot)\nonumber\\
+  & : i_{G} \rightarrow \lbrace c_\bot\ |\ (c_\bot\text{ is contained in }i_{G}) \wedge (c_\bot \in \bot)\rbrace\nonumber
+\end{align}
+\end{definition}
+\end{svgraybox}
+
+Note that a mapping function $f$ always maps from a given granularity $G$ to the set of chronons $\bot$. Therefore, the output for a mapping function is a subset of $\Pow(\bot)$. \\
+ A mapping function $f$ requires that the following properties hold~\cite{Lin97}:
+
+%
+%
+%For a given granularity $G$ and a given mapping function $f$ from $G$ to $\bot$, it is required that 
+\begin{property}
+%\item
+$G$ \emph{is an ordered set}: for a given $i_{G} \in G$ holds: if $c_\bot \in f(i_G)$ and $c^{'}_\bot \in f(j_G)$, then $c_\bot < c^{'}_\bot$ implies that $i_G < j_G$.
+This is illustrated in figure \ref{fig:granularity-prop1}.
+\begin{figure}
+\centering
+\input{graphs/granularity-prop1.tex}
+\caption{$G$ is an ordered set}
+\label{fig:granularity-prop1}
+\end{figure}
+\end{property}
+
+%\item
+\begin{property}
+$G$ \emph{is a set of continuous granules}: for a given $i_{G} \in G$ and given chronons $c_\bot$, $c^{'}_\bot$ and $c^{''}_\bot$ holds: if $c_\bot < c^{'}_\bot < c^{''}_\bot$, then $c_\bot \in f(i_G)$ and $c^{''}_\bot \in f(i_G)$ implies that $c^{'}_\bot \in f(i_G)$.
+This is illustrated in figure \ref{fig:granularity-prop2}.
+\begin{figure}
+\centering
+\input{graphs/granularity-prop2.tex}
+\caption{$G$ is a set of continuous granules}
+\label{fig:granularity-prop2}
+\end{figure}
+\end{property}
+
+%\item
+\begin{property}
+\emph{The granules in} $G$ \emph{do not overlap}: for given $i_{G}\text{ and } j_{G} \in G$ holds: if $i_G \neq j_G$, then $f(i_G) \cap f(j_G) = \emptyset$.
+This is illustrated in figure \ref{fig:granularity-prop3}.
+\begin{figure}
+\centering
+\input{graphs/granularity-prop3.tex}
+\caption{The granules in $G$ do not overlap}
+\label{fig:granularity-prop3}
+\end{figure}
+\end{property}
+%\end{itemize}
+
+Mapping functions between granularities and the chronons granularity also allow comparing granularities with respect to the length of their granules.
+In this context, two important concepts can be discerned.
+%A granularity $G$ is said to be \emph{\textbf{finner than}} a granularity $H$ when the set of chronons on which a given granule $i_G$ is mapped by a mapping function $f$ has less elements than the set of chronons on which a given granule $j_H$ is mapped by $f$.
+%A granularity $\alpha$ is \emph{``coarser than''} granularity $\beta$ when the set of corresponding chronons for a given granule $i_\alpha$ is greater than the set of the corresponding chronons for a given granule $j_\beta$.
+
+\begin{svgraybox}
+\vspace{-10pt}
+\begin{definition}\textbf{Finner Than}\\
+Consider a mapping function $f$ and let $i_G$ and $j_H$ be elements of granularities $G$ and $H$ respectively. Granularity $G$ is now said to be \emph{\textbf{finner than}} granularity $H$ if:
+\begin{equation}
+\label{eq:finner-than}
+| f \left( i_G \right) |\ <\ |  f \left( j_H \right)| \nonumber
+\end{equation} 
+\end{definition}
+
+\begin{definition}
+\label{def:coarser-than}\textbf{Coarser Than}
+Consider a mapping function $f$ and let $i_G$ and $j_H$ be elements of granularities $G$ and $H$ respectively. Granularity $G$ is now said to be \emph{\textbf{coarser than}} granularity $H$ if:
+\begin{equation}
+\label{eq:coarser-than}
+| f \left( i_G \right)|\ >\ |  f \left( j_H \right)| \nonumber
+\end{equation} 
+\end{definition}
+\vspace{-10pt}
+\end{svgraybox}
+
+It is also possible to describe the relation between different granularities. This is called a casting function:
+\begin{svgraybox}
+\begin{definition}
+\textbf{Casting function}~\cite{Lin97}\\
+\label{def:casting-fucntion}
+A casting function $cast$ maps granules from a granularity $G$ to granules from a different granularity $H$.
+\begin{equation}
+\label{eq:cast-function}
+\mbox{cast} \left( i_G,G,H \right) = j_H \nonumber
+\end{equation}
+\end{definition}
+\end{svgraybox}
+
+This notation indicates that the function $cast$ associates a granule $i_G$ in $G$ to a corresponding granule $j_H$ in $H$.
+Two kinds of granularity-to-granularity mappings can now be discerned: an \emph{upwards mapping} is a mapping from a granularity $G$ to a coarser granularity $H$, whereas a downwards mapping is a mapping from a granularity $K$ to a finner granularity $L$.
+Different granularity-to-granularity mappings between several granularities can be represented using a granularity graph, which is a directed graph indicating the mapping conversions.
+
+ Mappings between two granularities may be classified as \emph{regular mappings}, \emph{irregular mappings} or \emph{congruent mappings}.
+\begin{itemize}
+\item
+\emph{Regular mapping}: A regular mapping is a granularity-to-granularity mapping where the mapping function between the two granularities is calculated by means of multiplications and/or divisions and (maybe) an anchor adjustment.
+E.g.
+the mapping between hours and minutes is calculated using a multiplication by $60$.
+\item
+\emph{Irregular mapping}: An irregular mapping is a granularity-to-granularity mapping where the mapping function can not be calculated by means of multiplications and/or divisions.
+E.g.
+the mapping between months and days is dependent on the exact month or day.
+\item
+\emph{Congruent mapping}: A congruent mapping is a granularity-to-granularity mapping where the two granularities in the mapping have the same granules but a different anchor.
+E.g.
+the mapping between the days (Gregorian calendar days) and the academic days.
+\end{itemize}
+
+\begin{example}
+Consider a systems that models both Gregorian Calendar dates as well as Academic dates.
+The chronons granularity in this system are milliseconds.
+Figure \ref{fig:granularity-graph-example} shows the complete granularity graph example.
+The transitions between Chronons and Seconds and between Seconds and Minutes and so on is called a \emph{regular mapping}.
+The transition between Days and Months is not defined by a multiplication or division, this is called a \emph{irregular mapping}.
+In the graph, these \emph{irregular mappings} are noted with a bold arrow.
+Finally, the transition between a Gregorian Calendar Day and an Academic Day is called a \emph{Congruent mapping}: two granularities with the same granules but with different anchors.
+Both concern the same days, but the academic year starts on October 1st, whereas the Gregorian calendar year starts on January 1st.
+Congruent mappings are visualised as straight lines without arrow heads.
+
+\end{example}
+
+\textcolor{red}{MOVE TO CHAPTER 4}
 The transition between granularities~\cite{Lin97} is also considered as a source of imprecision.
@@
 
-\subsection{Temporal Relations}
-%an explanation on allen's temporal relations.
-%definition of fuzzy temporal interval
-
-Several operators are defined in order to compare temporal elements.
-Allen~\cite{Allen83} first described the relations between time intervals and to a lesser extent, between time points.
-Figure \ref{fig:allen} shows the temporal relations.
-There are some proposals that can be applied to both crisp and fuzzy temporal interval ~\cite{ohlbach2004},\cite{nagypal2003},\cite{schockaert08}.
-For this fuzzy comparisons~\cite{garrido2009} proposed some different temporal relations that can be easyly implemented by means of regular fuzzy operations.
+\textcolor{red}{END MOVE TO CHAPTER 4}
 
@@
 \centering
-\input{graphs/allen.tex}
-\caption{Allen relations between two time intervals.}
-\label{fig:allen}
+\includegraphics[scale=0.5]{graphs/granularityGraph.eps}
+\caption{A granularity graph example}
+\label{fig:granularity-graph-example}
 \end{figure}
 
-\subsection{An example of temporal domain: Julian Day Number}
+Next, in subsection \ref{subsec:timedomain-calendar}, some basic concepts and issues concerning calendars are explained.
+
+\subsection{Calendars}
+\label{subsec:timedomain-calendar}
+%maybe introduce previously the concept of time units, linear hierarchy and time point.
+
+This subsection will concern \emph{calendars}.
+A calendar provides some sort of human interpretation of time.
+As such, calendars provide meaning or interpretation to temporal values where this meaning or interpretation is relevant or useful to the user.
+More precisely, calendars determine the mapping between human-meaningful time values and an underlying time model~\cite{Dyreson1994}.
+More formally, a calendar is an organization of different granularities (e.g.
+granularities day, month and year) in a hierarchy.
+In order to formally define what a calendar is, some other concepts should first be introduced~\cite{Kraus1997}:
+
+%\begin{svgraybox}
+%\vspace{-10pt}
+%\begin{definition}\textbf{(Calendar)}\\
+%A \emph{\textbf{calendar}} 
+%\end{definition}
+%\vspace{-10pt}
+%\end{svgraybox}
+
+\begin{svgraybox}
+\vspace{-10pt}
+\begin{definition}\textbf{Linear Hierarchy of Granularities}\\
+A \emph{\textbf{linear hierarchy of granularities}} is a finite set of granularities with a linear order (denoted $\sqsubset$).
+The most coarse granularity is called \emph{the top of the hierarchy} and must be an infinite set while all other granularities must be finite sets.
+\end{definition}
+\vspace{-10pt}
+\end{svgraybox}
+
+An example of a linear hierarchy is the set $\left\{ \text{days, months, years} \right\}$, where days $\sqsubset$ months and months $\sqsubset$ years.
+With this, a \emph{linear calendar} can be defined.
+
+\begin{svgraybox}
+\vspace{-10pt}
+\begin{definition}\textbf{Linear Calendar}\\
+A \emph{\textbf{linear calendar}} consist of a linear hierarchy of granularities and a validity predicate which specifies if a granule is valid in the calendar.
+\end{definition}
+\vspace{-10pt}
+\end{svgraybox}
+
+An example of a linear calendar could be the linear hierarchy $\left\{ \text{days, months, years} \right\}$, where days $\sqsubset$ months and months $\sqsubset$ years, together with a validity predicate that excludes time indications like `32 April 2012' or `30 February 2012'.
+
+%The linear hierarchy days $\sqsubset$ months $\sqsubset$ years represents a linear calendar.
+It is necessary to define the validity predicates because 17 January 2012 is a valid element of the calendar whereas 30 February 2012 is not.
+
+\textcolor{red}{CHECK}
+\subsubsection{\label{subsubsubsec:julian-day-number}An example of temporal domain: Julian Day Number}
 % proposal for the underlying Julian Day Number domain.
@@
 
-%\subsubsection{Conversion among dates}
-%\paragraph{Conversion between Gregorian and Julian dates}
-%The Julian Date consist on the Julian Day number and the time. It is computed with the formula \cite{Usn}:\\
-%\begin{equation}
-%\label{eq:julian day}
-%JD = JDN + \frac{hour-12}{24} + \frac{minutes}{1440} + \frac{seconds}{86400}
-%\end{equation}
-%
-%Here the JDN is the Julian Day Number.
-This equation does not take into account the leap seconds \footnote{A leap second(s) is the amount of second(s) added or deleted every year in the Universal Time.\cite{Wik},\cite{Lea}.
-}.
+\textcolor{red}{END CHECK}
 
-%\paragraph{Gregorian calendar to Julian day}
-%Any date from Gregorian calendar \footnote{As Gregorian calendar entered in use in 1582, any use before that date is called \emph{proleptic calendar}.} can be converted into a Julian Day using the following formula:\\
-%\begin{equation}
-%\label{eq:JDN}
-%JDN = \frac{1461x(Y + 4800 + \frac{M-14}{12})}{4}+\frac{367x(M-2-12x\frac{M-14}{12})}{12}-\frac{3x\frac{Y+4900+\frac{M-14}{12}}{100}}{4} 
-%\end{equation}
-%
-%This formula was created by Henry Fliegel and Thomas C.
-van Flandern in 1968.
-\cite{Fliegel:1968:LEM:364096.364097} It takes into account leap years and that they are applicable only when the century is divisible by 400.
-The value $Y$ represents the year and the value $M$ represents the month.
-%\paragraph{Julian calendar to Julian day}
-%The algorithm to convert between Julian calendar and Julian Day Number is only valid for dates since 4713 B.C (positive values of Julian Day) \cite{Dog}:\\
-%\begin{equation}
-%\label{eq:JDN_2_JD}
-%JDN=367Y-\frac{7 (Y + 5001 + \frac{M-9}{7})}{4}+\frac{275xM}{9}+D+1729777
-%\end{equation}
-%
-%Where $Y$ is the year value, $M$ is the month value, and $D$ is the day value.
-%
-%%\subsubsection{Membership degree computation}
-%The main advantage from the Julian day number is the ability to make calculations. Thus, the membership degree computation for a fuzzy number based on a Julian day is explained here. \\
-%Let a trapezoidal possibility distribution defined by four JDN named $a$,$b$,$c$ and $d$.
-The fuzzy validity period \emph{FVP}, can be represented as in figure \ref{fig:fuzzy-validity-period}.
-The following equation shows the membership degree for a Julian day number \emph{JDN} in a \emph{FVP} represented by means of the possibility distribution of $[a,b,c,d]$.
-In a valid-time database, the values for the valid starting time \emph{VST} and the valid ending time \emph{VET} can be fuzzyfied as shown in equation \ref{eq:fuzzy-validity-period}.
-%
-%The membership function for a given $x$ number in \emph{JDN} format is the following:\\
-%\begin{equation}
-%\label{eq:trapezoidal distribution}
-%%\[
-%T(x)=
-%\begin{cases}
-%0, \mbox(if x \leq a, x \geq d)\\
-%1, \mbox(if x \in [b,c])\\
-%\frac{x-a}{b-a}   \mbox(if x \in (a,b])\\
-%\frac{d-x}{d-c} 	 \mbox(if x \in (c,d])\\
-%\end{cases}
-%%\]
-%\end{equation}
-%The points outside the trapezium or inside the interval $[a,b]$ are trivial. The points inside the other intervals has the following particularities:\\
-%The subtraction operation with Julian day numbers gives as result a scalar value (as explained before).
-The division operation between two scalars gives as result an scalar in the interval $[0,1]$ which represents the membership degree.
+The next section briefly introduces temporal relationships.
+
+\subsection{Temporal Relationships}
+%an explanation on allen's temporal relations.
+%definition of fuzzy temporal interval
+In this section, a brief introduction can be found, concerning \emph{temporal relationships}, sometimes also called `temporal relations'\cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy}.
+Temporal relationships can be seen as relationships between temporal elements or instants.
+
+\def\JPicScale{0.5}
+\begin{figure}[h]
+\centering
+\includegraphics[scale=0.5]{graphs/allen.eps}
+\caption{Allen relations between two time intervals.}
+\label{fig:allen}
+\end{figure}
+
+Several operators are defined in order to compare temporal elements or time points.
+Allen~\cite{Allen83} first described the relationships between time intervals and as a special case, between time points.
+Figure \ref{fig:allen} shows the temporal relationships Allen discerned.
+Some proposals can be applied to both crisp and fuzzy temporal intervals~\cite{ohlbach2004},~\cite{nagypal2003},~\cite{schockaert08}.
+
+\textcolor{red}{MOVE TO FUZZY PART}
+For this fuzzy comparisons~\cite{garrido2009} proposed some different temporal relations that can be easily implemented by means of regular fuzzy operations.
+\textcolor{red}{END MOVE TO FUZZY PART}
+
+In the next section, some basic concepts and terminology about temporal databases are presented, followed by an overview of some interesting issues concerning temporal databases and a survey of some commercial temporal database systems.

### 2012-02-25 trunk/book/sources/temporalDatabases.tex
update with the simplification of the text.

@@
 %
-A \emph{temporal database}~\cite{etzion1998} is a database that manages some aspects of time in its schema~\cite{Dyreson1994}.
+A temporal database can generally be seen as a database that manages some temporal aspects in its schema \cite{etzion1998}.
+In subsection \ref{subsec:td-general-concepts}, some main concepts and properties concerning temporal databases and their definitions are presented and explained.
+In subsections \ref{subsubsec:primary-key} and \ref{subsubsec:consistency}, some main issues of relational temporal databases are presented and discussed.
+Finally, subsection \ref{Comm-temp} presents an overview of some commercial temporal database systems.
+
+\subsection{\label{subsec:td-general-concepts}Basic Concepts and Properties}
+A database models some part of reality.
+As mentioned in the introduction, the part of reality a temporal database tries to model, contains some temporal aspects.
+For example, in this part of reality, some concepts or objects could have time-related or time-variant properties.
+The modelling of these temporal aspects has to be handled specifically in order for the database to maintain a consistent model of reality.
+
+Thus, a temporal database will contain temporal values, i.e.
+values representing (indications of) time.
+Temporal values in a temporal database can be classified into the following types based on their interpretation and modelling purpose.
+The definitions and explanations of these types can be found in \cite{Dyreson1994} and \cite{Nascimento95decisiontime} and more information can be found in \cite{Jensen:1991:IIM:627283.627484}, \cite{Snodgrass:1984:TQL:588011.588041} and \cite{Nascimento95decisiontime}.
+
+\begin{svgraybox}
+\vspace{-10pt}
+\begin{definition}\textbf{Valid Time}\\
+The \emph{\textbf{valid time}} (VT) of a fact is the time when the fact is true in the modeled reality.
+\end{definition}
+
+\begin{definition}\textbf{Transaction Time}\\
+A database fact is stored in a database at some point in time, and after it is stored, it is current until logically deleted.
+The \emph{\textbf{transaction time}} (TT) of a database fact is the time when the fact is current in the database and may be retrieved.
+\end{definition}
+
+\begin{definition}\textbf{Decision Time}\\
+\emph{\textbf{Decision time}} (DT) denotes the time when an event was decided to happen.
+\end{definition}
+
+\begin{definition}\textbf{User-defined Time}\\
+\emph{\textbf{User-defined time}} (UDT) is an uninterpreted attribute domain of date and time.
+\end{definition}
+\vspace{-10pt}
+\end{svgraybox}
+
+\emph{Valid times} are usually supplied by the user, whereas \emph{transaction times} are usually system-generated and -supplied \cite{Dyreson1994}.
+Temporal values of type UDT are not given any special interpretation and have thus no special query language support \cite{Dyreson1994}.
+
+A \emph{temporal database} can now formally be defined as follows \cite{Dyreson1994}.
+
+\begin{svgraybox}
+\vspace{-10pt}
+\begin{definition}\textbf{Temporal Database}\\
+A \emph{\textbf{temporal database}} supports some aspect of time, not counting user-defined time.
+\end{definition}
+\vspace{-10pt}
+\end{svgraybox}
+
+In a relational temporal database, temporal values will of course be in the tuples of the extentions of temporal relations:
+
+\begin{svgraybox}
+\vspace{-10pt}
+\begin{definition}\textbf{Valid-time Relation}\\
+A \emph{\textbf{valid-time relation}} is a relation with exactly one system supported valid time \cite{Dyreson1994}.
+\end{definition}
+
+\begin{definition}\textbf{Transaction-time Relation}\\
+A \emph{\textbf{transaction-time relation}} is a relation with exactly one system supported transaction time \cite{Dyreson1994}.
+\end{definition}
+\vspace{-10pt}
+\end{svgraybox}
+
+A \emph{valid-time}, respectively \emph{transaction-time} \emph{relational database} is now defined as containing one or more valid-time, respectively transaction-time relations \cite{Dyreson1994}.
+Next to this, \emph{bitemporal} relational databases contain both valid time and transaction time and tritemporal databases contain valid time, transaction time and decision time.
+
+%Database models can also be classified into \emph{bi-temporal} (both valid and transaction-time) or \emph{tri-temporal}  (bi-temporal and decision time) models.
+
+A very extensive list of the most well-known temporal database models can be found in ~\cite{Yu1998}.
+As it is of course necessary to define a consistent way to query the temporal data, there are several proposals concerned with query languages and query language adaptations for temporal databases like ~\cite{TSQL} and ~\cite{Snodgrass98}.
+
+%A \emph{temporal database}~\cite{etzion1998} is a database that manages some aspects of time in its schema~\cite{Dyreson1994}.
 The reality a temporal database tries to model, contains some temporal notions which have to be handled specifically in order to maintain a consistent modelling behavior.
@@
 
-\begin{svgraybox}
-The temporal notions in temporal databases can be classified into the following types based on their interpretation and modelling purpose.
+%\begin{svgraybox}
+%The temporal notions in temporal databases can be classified into the following types based on their interpretation and modelling purpose.
 User-defined time has no interpretation, but the other types do:\\
 
-	\textbf{Transaction time} \emph{TT}~\cite{Jensen:1991:IIM:627283.627484}: The time when the fact is stored in the database.\\
+%	\textbf{Transaction time} \emph{TT}~\cite{Jensen:1991:IIM:627283.627484}: The time when the fact is stored in the database.\\
 
-	\textbf{Valid time} \emph{VT}~\cite{Snodgrass:1984:TQL:588011.588041}: The time when the fact is true in the modelled reality.\\
+%	\textbf{Valid time} \emph{VT}~\cite{Snodgrass:1984:TQL:588011.588041}: The time when the fact is true in the modelled reality.\\
 
-	\textbf{Decision time} \emph{DT}~\cite{Nascimento95}: The time when an event was decided to happen. \\
-\end{svgraybox}
+%	\textbf{Decision time} \emph{DT}~\cite{Nascimento95}: The time when an event was decided to happen. \\
+%\end{svgraybox}
 
-Database models can also be classified into \emph{bi-temporal} (both valid and transaction-time) or \emph{tri-temporal}  (bi-temporal and decision time) models.
+In the following sections, one of the main points of attention concerning relational temporal databases is explained and discussed.
 
-\subsection{\label{subsubsec:primary-key}Primary key on temporal databases}
+\subsection{\label{subsubsec:primary-key}Primary Keys in Temporal Relations}
+When designing a relation, a subset of the relation's attribute set is ususally chosen as primary key.
+The values of a record for these attributes will then uniquely determine this record and no two records may have the same values for every attribute in this primary key.
+In temporal relations, different records can exist with the same values for every attribute except the attributes describing the temporal aspects of the objects or concepts represented by the records (the temporal attributes).
+These different records represent different temporal versions of the same real objects or concepts.
+To allow the existence of such records, the most common solution is to extend the original primary key with the temporal attributes.
+%One of the first tasks in the design of a database is the choice for the primary key.
+In a temporal database, the most common solution for the choice of the primary key is to extend the primary key with the temporal value.
+Nevertheless, when dealing with uncertainty in the temporal values, the solution is usually to create a version identifier otherwise the primary key have uncertainty which should be avoided.
+
+The following example illustrates this primary key issue in a valid-time relation.
 
@@
 \label{ex:pk}
-Consider the following example database in \ref{table:example-database}.
-It contains regular data from employees in a company like the name, the birth date.
-Also two temporal values which represent a valid-time interval.
+Consider the example database visualized in table \ref{table:example-database}.
+It contains data representing employees in a company, and two temporal values which represent a valid-time interval.
 Consider now that we want to hire Sarah again.
@@
 \subsection{\label{subsubsec:consistency}Consistency on temporal databases}
-As it is shown in example \ref{ex:pk}, it  is necessary to define a consistency mechanism to avoid spurious values in the database.
+The consistence mechanism in a temporal database is usually re-defined.
+The main problem is the insertion of spurious values in the database, as shown in example \ref{ex:pk}, in table \ref{table:example-database-with-new-pk}.
 The most usual solution is to re-define the DML (\emph{Data Manipulation Language}).
@@
 
-\begin{center}
-\emph{Update Employees set StartYear = 2010 where ID=3}
-\end{center}
+\begin{verbatim}
+Update Employees set StartYear = 2010 where ID=3;
+\end{verbatim}
 
@@
 
-\begin{center}
-\emph{Update Employees set EndYear = 2010 where ID=3}\\
-\emph{Insert into Employees values (3,John,52,-,2010,-)}
-\end{center}
+\begin{verbatim}
+Update Employees set EndYear = 2010 where ID=3;
+Insert into Employees values (3,John,52,-,2010,-);
+\end{verbatim}
 
@@
 
-\subsection{Commercial Temporal database systems}
+\subsection{\label{Comm-temp}Commercial Temporal database systems}
 There are several commercial temporal DBMS. Table \ref{table:commercial-temporal-db} is a comparative among the most well-known systems. It is also provided a reference for more information.
@@
 
-The most complete implementation is the Workspace Manager, which is described briefly in the following section.
-Unfortunately, none of these systems work with uncertainty or imprecision neither in the data storage nor in the query.
+The most complete implementation is the Workspace Manager. Unfortunately, none of these systems work with uncertainty or imprecision neither in the data storage nor in the query.
 
@@
 
-\subsubsection{Oracle Workspace Manager}
-
+%\subsubsection{Oracle Workspace Manager}
+%
 %\begin{figure}
 %\centering
-%\includegraphics[scale=0.5]{graphs/workspaceTopology.pdf}
+%\includegraphics[scale=0.5]{graphs/workspaceTopology.eps}
 %\caption{A sample topology in the workspace manager.}
@@
 %\end{figure}
-
-Oracle workspace manager \cite{OraE118602} package allows to get several versions of the data in the same database. It is also possible to version only a table. The main benefits of this are two:
-
-\begin{itemize}
-\item
-Organization and optimization of data in a hierarchical frames: It is possible to create a frame of time (a workspace) and work only with that partial version of the data.
-It is provided a mechanim to merge data among different workspaces and to solve conflicts.
+%
+%Oracle workspace manager \cite{OraE118602} package allows to get several versions of the data in the same database. It is also possible to version only a table. The main benefits of this are two:
+%
+%\begin{itemize}
+%\item
+%Organization and optimization of data in a hierarchical frames: It is possible to create a frame of time (a workspace) and work only with that partial version of the data.
+It is provided a mechanism to merge data among different workspaces and to solve conflicts.
 The organization of data in these workspaces for very large tables result in smaller time access.
-\item
-Valid time as well as transaction time are managed by the system. Each update sentence makes a change in the versioning of a row.
-\end{itemize}
-
-When a table is versioned, the system creates a few tables and views as well as auxiliary structures to allow keep several version of the data, while keeping the primary key defined by the user.
-
-A workspace is a logical group of a set of changes (versions of tuples) and allows consistent access, thus the user always obtain the correct data version.
+%\item
+%Valid time as well as transaction time are managed by the system. Each update sentence makes a change in the versioning of a row.
+%\end{itemize}
+%
+%When a table is versioned, the system creates a few tables and views as well as auxiliary structures to allow keep several version of the data, while keeping the primary key defined by the user.
+%
+%A workspace is a logical group of a set of changes (versions of tuples) and allows consistent access, thus the user always obtain the correct data version.
 Workspaces are ordered in a hierarchy.
@@
 \\
-
-The system only makes a copy of a tuple if it is changed.
+%
+%The system only makes a copy of a tuple if it is changed.
 In order to access different versions of the tuple, the context must be changed.
 Also, it is not necessarily to modify the SQL code to access to versioned tables.\\
-%When a table is versioned, it is renamed as table\_LT.
+%%When a table is versioned, it is renamed as table\_LT.
 Here is stored all the data for the information together metadata for the versioning.

### 2012-02-25 book/sources/timeDomain.tex
update with the simplification of the text.

@@
 
+ \begin{itemize}
+ \item
+ $G$ \emph{is an ordered set}.
+ \item
+ $G$ \emph{is a set of continuous granules}.
+ \item
+ \emph{The granules in} $G$ \emph{do not overlap}.
+ \end{itemize}
+
+%%
+%%
+%%For a given granularity $G$ and a given mapping function $f$ from $G$ to $\bot$, it is required that 
+%\begin{property}
+%%\item
+%$G$ \emph{is an ordered set}: for a given $i_{G} \in G$ holds: if $c_\bot \in f(i_G)$ and $c^{'}_\bot \in f(j_G)$, then $c_\bot < c^{'}_\bot$ implies that $i_G < j_G$.
+This is illustrated in figure \ref{fig:granularity-prop1}.
+%\begin{figure}
+%\centering
+%\input{graphs/granularity-prop1.tex}
+%\caption{$G$ is an ordered set}
+%\label{fig:granularity-prop1}
+%\end{figure}
+%\end{property}
 %
+%%\item
+%\begin{property}
+%$G$ \emph{is a set of continuous granules}: for a given $i_{G} \in G$ and given chronons $c_\bot$, $c^{'}_\bot$ and $c^{''}_\bot$ holds: if $c_\bot < c^{'}_\bot < c^{''}_\bot$, then $c_\bot \in f(i_G)$ and $c^{''}_\bot \in f(i_G)$ implies that $c^{'}_\bot \in f(i_G)$.
+This is illustrated in figure \ref{fig:granularity-prop2}.
+%\begin{figure}
+%\centering
+%\input{graphs/granularity-prop2.tex}
+%\caption{$G$ is a set of continuous granules}
+%\label{fig:granularity-prop2}
+%\end{figure}
+%\end{property}
 %
-%For a given granularity $G$ and a given mapping function $f$ from $G$ to $\bot$, it is required that 
-\begin{property}
-%\item
-$G$ \emph{is an ordered set}: for a given $i_{G} \in G$ holds: if $c_\bot \in f(i_G)$ and $c^{'}_\bot \in f(j_G)$, then $c_\bot < c^{'}_\bot$ implies that $i_G < j_G$.
-This is illustrated in figure \ref{fig:granularity-prop1}.
-\begin{figure}
-\centering
-\input{graphs/granularity-prop1.tex}
-\caption{$G$ is an ordered set}
-\label{fig:granularity-prop1}
-\end{figure}
-\end{property}
-
-%\item
-\begin{property}
-$G$ \emph{is a set of continuous granules}: for a given $i_{G} \in G$ and given chronons $c_\bot$, $c^{'}_\bot$ and $c^{''}_\bot$ holds: if $c_\bot < c^{'}_\bot < c^{''}_\bot$, then $c_\bot \in f(i_G)$ and $c^{''}_\bot \in f(i_G)$ implies that $c^{'}_\bot \in f(i_G)$.
-This is illustrated in figure \ref{fig:granularity-prop2}.
-\begin{figure}
-\centering
-\input{graphs/granularity-prop2.tex}
-\caption{$G$ is a set of continuous granules}
-\label{fig:granularity-prop2}
-\end{figure}
-\end{property}
-
-%\item
-\begin{property}
-\emph{The granules in} $G$ \emph{do not overlap}: for given $i_{G}\text{ and } j_{G} \in G$ holds: if $i_G \neq j_G$, then $f(i_G) \cap f(j_G) = \emptyset$.
+%%\item
+%\begin{property}
+%\emph{The granules in} $G$ \emph{do not overlap}: for given $i_{G}\text{ and } j_{G} \in G$ holds: if $i_G \neq j_G$, then $f(i_G) \cap f(j_G) = \emptyset$.
 This is illustrated in figure \ref{fig:granularity-prop3}.
-\begin{figure}
-\centering
-\input{graphs/granularity-prop3.tex}
-\caption{The granules in $G$ do not overlap}
-\label{fig:granularity-prop3}
-\end{figure}
-\end{property}
-%\end{itemize}
+%\begin{figure}
+%\centering
+%\input{graphs/granularity-prop3.tex}
+%\caption{The granules in $G$ do not overlap}
+%\label{fig:granularity-prop3}
+%\end{figure}
+%\end{property}
+%%\end{itemize}
 
@@
 
-\textcolor{red}{CHECK}
-\subsubsection{\label{subsubsubsec:julian-day-number}An example of temporal domain: Julian Day Number}
-% proposal for the underlying Julian Day Number domain.
-
-The Julian Day number \emph{JDN} \cite{Dir96} is a counter.
+%\textcolor{red}{CHECK}
+%\subsubsection{\label{subsubsubsec:julian-day-number}An example of temporal domain: Julian Day Number}
+%% proposal for the underlying Julian Day Number domain.
+%
+%The Julian Day number \emph{JDN} \cite{Dir96} is a counter.
 Its value is incremented in one unit every day from 1 January 4713 B.C.
@@
 The particularity of starting at noon was useful for astronomers: the observations they took one night belonged to the same Julian Day.\\
-Note that the Julian Day represents whole days.
+%Note that the Julian Day represents whole days.
 There is an extension that allows to represent any precision needed (it is called Julian Date).
@@
 The inverse conversion formula is proposed in~\cite{Fliegel:1968:LEM:364096.364097}\\
-
-There are many alternatives to optimize the representation of Julian Day numbers because of its extremely far origin (4713 B.C.
+%
+%There are many alternatives to optimize the representation of Julian Day numbers because of its extremely far origin (4713 B.C.
 year).
 Table \ref{table:juliandayrepresentations} shows several time domains that can be calculated from the Julian Date, some of them are proposed just for optimization purposes.
-
-\begin{table}
-\caption{Julian Day representations}
-\label{table:juliandayrepresentations}
-
-\begin{tabular}{p{2cm}p{4cm}p{4cm}p{2cm}}
-%header ------------------------------------------------------
-\hline\noalign{\smallskip}
-Name & From & Formula & Current Value  \\ 
-\noalign{\smallskip}\svhline\noalign{\smallskip}
-%header ------------------------------------------------------
-Julian Date (JD)$^a$ & 12:00 noon Monday 1 January 4713 B.C & & 2455278. 85488  \\ 
-Julian Day Number (JDN)$^b$ & 12:00 noon  Monday 1 January 4713 B.C. & JND = floor(JD) & 2455278 \\ 
-Reduced Julian Day (RJD)$^c$ & 12:00 noon Tuesday 16 November 1858 A.C. & RJD = JD - 2400000 & 55278.85488  \\ 
-Modified Julian Day (MJD)$^d$ & 00:00 Wednesday 17 November 1858 A.C. & MJD = JD - 2400000,5 & 55278.35488 \\ 
-Truncated Julian Day (TJD)$^e$ & 00:00 Friday 24 May 1968 A.C. & TJD = JD - 2440000,5 & 15278.35488  \\ 
-Truncated Julian Day (TJD)$^f$ & 00:00 Thursday 10 November 1995 A.C. & TJD = (JD- 0,5) mod 10000 & 5278. 35488   \\ 
-Dublin Julian Day (DJD))$^g$ & 12:00 noon Sunday 31 December 1899 A.C. & DJD = JD - 2415020 & 40258. 85488 \\ 
-Chronological Julian Day (CJD)$^h$  & 00:00 Monday 1 January 4713 B.C. & CJD = JD + 0,5 + timezone adjust. & 2455279. 3548843 (UT)  \\ 
-Lilian Day Number$^i$ & Friday 15 October 1582 & floor(JD-2299160,5) & 156118 \\ 
-ANSI Date$^j$  & Monday 1 January 1601 & floor(JD-2305812,5) & 149466  \\ 
-Rata die$^k$  & Monday 1 January 1 A.C & floor(JD - 1721424.5) & 733854 \\ 
-Unix time$^l$  & Thursday 1 January 1970 A.C. & (JD – 2440587.5) × 86400 & 1269333062 \\ 
-\noalign{\smallskip}\hline\noalign{\smallskip}
-\end{tabular}
-$^a$ This is an extension of the Julian Day that allows time representation. \\
-$^b$  Each day changes at noon. \\
-$^c$  Used by astronomers. \\
-$^d$ It starts at midnight. \\
-$^e$ Definition from NASA. \cite{Sch}. \\
-$^f$ Definition from NIST. \cite{Nis}. \\
-$g$ Introduced by the IAU in 1995. \\
-$^h$ The timezone must be explicited. Each day changes at midnight. \\
-$^i$ The number of days since Gregorian calendar in Universal Time. \\
-$^j$ The origin for COBOL integer dates. \\
-$^k$  The number of days since actual era. \\
-$^l$  It counts the seconds not the day. \\
-\end{table}
-
-\textcolor{red}{END CHECK}
+%
+%
+%\begin{table}
+%\caption{Julian Day representations}
+%\label{table:juliandayrepresentations}
+%
+%\begin{tabular}{p{2cm}p{4cm}p{4cm}p{2cm}}
+%%header ------------------------------------------------------
+%\hline\noalign{\smallskip}
+%Name & From & Formula & Current Value  \\ 
+%\noalign{\smallskip}\svhline\noalign{\smallskip}
+%%header ------------------------------------------------------
+%Julian Date (JD)$^a$ & 12:00 noon Monday 1 January 4713 B.C & & 2455278. 85488  \\ 
+%Julian Day Number (JDN)$^b$ & 12:00 noon  Monday 1 January 4713 B.C. & JND = floor(JD) & 2455278 \\ 
+%Reduced Julian Day (RJD)$^c$ & 12:00 noon Tuesday 16 November 1858 A.C. & RJD = JD - 2400000 & 55278.85488  \\ 
+%Modified Julian Day (MJD)$^d$ & 00:00 Wednesday 17 November 1858 A.C. & MJD = JD - 2400000,5 & 55278.35488 \\ 
+%Truncated Julian Day (TJD)$^e$ & 00:00 Friday 24 May 1968 A.C. & TJD = JD - 2440000,5 & 15278.35488  \\ 
+%Truncated Julian Day (TJD)$^f$ & 00:00 Thursday 10 November 1995 A.C. & TJD = (JD- 0,5) mod 10000 & 5278. 35488   \\ 
+%Dublin Julian Day (DJD))$^g$ & 12:00 noon Sunday 31 December 1899 A.C. & DJD = JD - 2415020 & 40258. 85488 \\ 
+%Chronological Julian Day (CJD)$^h$  & 00:00 Monday 1 January 4713 B.C. & CJD = JD + 0,5 + timezone adjust. & 2455279. 3548843 (UT)  \\ 
+%Lilian Day Number$^i$ & Friday 15 October 1582 & floor(JD-2299160,5) & 156118 \\ 
+%ANSI Date$^j$  & Monday 1 January 1601 & floor(JD-2305812,5) & 149466  \\ 
+%Rata die$^k$  & Monday 1 January 1 A.C & floor(JD - 1721424.5) & 733854 \\ 
+%Unix time$^l$  & Thursday 1 January 1970 A.C. & (JD – 2440587.5) × 86400 & 1269333062 \\ 
+%\noalign{\smallskip}\hline\noalign{\smallskip}
+%\end{tabular}
+%$^a$ This is an extension of the Julian Day that allows time representation. \\
+%$^b$  Each day changes at noon. \\
+%$^c$  Used by astronomers. \\
+%$^d$ It starts at midnight. \\
+%$^e$ Definition from NASA. \cite{Sch}. \\
+%$^f$ Definition from NIST. \cite{Nis}. \\
+%$g$ Introduced by the IAU in 1995. \\
+%$^h$ The timezone must be explicited. Each day changes at midnight. \\
+%$^i$ The number of days since Gregorian calendar in Universal Time. \\
+%$^j$ The origin for COBOL integer dates. \\
+%$^k$  The number of days since actual era. \\
+%$^l$  It counts the seconds not the day. \\
+%\end{table}
+%
+%\textcolor{red}{END CHECK}
 

### 2012-02-27 trunk/book/sources/fuzzyTemporalDatabases.tex
Almost done!

@@
 %
-Some software applications deal with uncertainty related to time.
+Consider a logistics company which transports packages.
+The time when a package would arrive at its destination may be estimated and can eventually be researched, but it is typically not always precisely known.
+For such companies and in many other situations, information systems able to handle imperfection corresponding to certain temporal aspects of the objects modelled by the system are necessary.
+The transition between granularities~\cite{Lin97} is also sometimes considered as a source of imprecision.
+Therefore, some proposals consider granularity as the base of the temporal model~\cite{Cru97}.
+%Some software applications deal with uncertainty related to time.
 e.g.
@@
 
-There are several approaches to manage uncertainty in the temporal information stored in a database.
-Valid-time is usually represented as an interval, and therefore it is suitable for the representation of uncertainty in the valid-time interval.
-Nevertheless, the values for transaction and decision time are usually precisely known.
-Some models deal with probability~\cite{Dekhtyar2001} or possibility~\cite{Dubois89} distributions at the endpoints of the time interval.
-In this section we will explain the most interesting proposals that deal with possibility distributions in the valid-time interval.
+Data imperfection is usually the focus of research in fuzzy databases.
+In fuzzy databases~\cite{Galindo2006}, uncertainty is usually expected to appear at storage level whereas imprecision is usually expected to appear in querying.
+Concerning temporal databases, there are several approaches to handle uncertainty in temporal data stored in a database.
+
+Valid-time notions usually take the form of an interval.
+Some models consider probability~\cite{Dekhtyar2001} or possibility~\cite{Dubois89} distributions to describe the endpoints of a time interval.
+In this section, we will present a novel and currently researched technique to describe valid-time interval start- and endpoints using possibility distributions.
+In subsection \ref{subsec:representation-time-intervals} we will present an approach to modelling uncertainty in valid-time intervals.
+Finally, in subsection \ref{subsec:querying-time-intervals} the proposals for querying valid-time intervals in a flexible way are presented and discussed.%In this section we will explain the most interesting proposals that deal with possibility distributions in the valid-time interval.
 %Two are the main points developed within this section: the representation of uncertainty in the time interval and the flexible querying of ill-known time periods.
 
-In fuzzy databases~\cite{Galindo2006}, uncertainty is managed at storage level whereas imprecision is managed at querying level.
+%In fuzzy databases~\cite{Galindo2006}, uncertainty is managed at storage level whereas imprecision is managed at querying level.
 In subsection \ref{subsec:representation-time-intervals} we will explain the main approach for the storage of uncertainty in the valid-time interval.
@@
 
-\subsection{Representation of ill-known temporal intervals}
+\subsection{Ill-known Time Intervals}
 \label{subsec:representation-time-intervals}
-The uncertainty in a valid-time interval may be represented with uncertainty at one or both starting and ending points.
+Uncertainty in a valid-time interval may be modelled by modelling the uncertainty concerning the exact start, respectively end point of the valid-time interval.
+The uncertainty concerning these points can by modelled by means of possibility distributions.
+It is of course also possible to use rough sets~\cite{Pawlak1995} for this modelling~\cite{Qia09}.
+The work presented in this section relies on possibilistic variables, which rely on possibility theory \cite{Dubois:Prade:1988:PossibilityTheory}.
+
+A \emph{possibilistic variable} is defined as follows \cite{Pon11}.
+
+\begin{svgraybox}
+\vspace{-10pt}
+\begin{definition}
+A possibilistic variable $X$ over a universe $U$ is defined as a variable taking exactly one value in $U$, but for which this value is (partially) unknown.
+Its possibility distribution $\pi_X$ on $U$ models the available knowledge about the value that $X$ takes: for each $u\in U$, $\pi_X(u)$ represents the possibility that $X$ takes the value $u$.
+In this work, this possibility is interpreted as a measure of how plausible it is that $X$ takes the value $u$, given (partial) knowledge about the value $X$ takes.
+\end{definition}
+\vspace{-10pt}
+\end{svgraybox}
+
+The exact value a possibilistic variable takes, which is (partially) unknown, is called an \emph{ill-known value} in this work \cite{Dubois88b}.
+
+When a possibilistic variable is defined on the powerset $\Pow(R)$ of some universe $R$, the unique value the variable takes will be a crisp set and its possibility distribution on the powerset $\Pow(R)$ will describe the possibility of each crisp subset of $R$ to be the value the variable takes.
+This exact value (a crisp set) the variable takes, is now called an \emph{ill-known set} \cite{Dubois88b}.
+
+This work will deal with ill-known time intervals.
+These are ill-known intervals in time.
+Ill-known intervals are ill-known sets, defined and represented via a start and end point, which will be ill-known values.
+The elements of the set are the values between the start and end point.
+A closed ill-known interval with start point defined by possibilistic variable $X$ and end point by possibilistic variable $Y$ is noted here $\left[X, Y\right]$.
+The correspondences and transitions between the representations of ill-known sets, between the representations of ill-known intervals and between the representations of an ill-known set and an ill-known interval are part of the authors current research.
+Figure \ref{fig:interval} illustrates this.
+
+%Uncertainty in a valid-time interval may be modelled with uncertainty at one or both starting and ending points.
 The uncertainty is usually represented by means of a possibility distribution in one or both points.
@@
 
-Several authors ~\cite{garrido2009} propose some transformations in order to optimize the storage of the valid-time interval.
-Further research has shown how these transformations drive to a possibilistic information lost~\cite{Pon11}.
-A comparison between the transformations in ~\cite{garrido2009}  and the framework in~\cite{Pon11} is done in~\cite{pon12}.
+Several authors ~\cite{garrido2009} propose transformations in order to optimize the storage of such ill-known valid-time intervals, though recent research might seem to indicate some minor issues with respect to some of these transformations ~\cite{Pon11}.
+A comparison between the transformations in ~\cite{garrido2009} and the framework in~\cite{Pon11} is presented in~\cite{pon12}.
+Figure \ref{fig:convexhull} illustrates a transformation based on the `convex hull' approach REF.
+
+%An ill-known time interval is usually represented by means of two possibility distributions one in each starting and ending points.
+Figure \ref{fig:interval} shows an ill-known valid-time interval.
+The starting and ending points are denoted by two fuzzy numbers, $X$ and $Y$, represented by means of two triangular possibility distributions (see the Appendix).
+For illustration, figure \ref{fig:convexhull} shows one of the transformations based in the convex hull and proposed in~\cite{garrido2009}.
 
@@
 \input{graphs/interval.tex}
-\caption{Two ill-known values $X$ and $Y$, represented by means of triangular possibility distributions denoting the starting and ending points of an ill-known interval. }
+\caption{A closed ill-known time interval $\left[X, Y\right]$, where triangular possibility distributions describe the ill-known values defining the start and end points.}
 \label{fig:interval}
 \end{figure}
-
-An ill-known temporal interval is usually represented by means of two possibility distributions one in each starting and ending points.
-Figure \ref{fig:interval} shows an ill-known valid-time interval.
-The starting and ending points are denoted by two fuzzy numbers, $X$ and $Y$, represented by means of two triangular possibility distributions (see the Appendix).
-For illustration, figure \ref{fig:convexhull} shows one of the transformations based in the convex hull and proposed in~\cite{garrido2009}.
 
@@
 
-\subsection{Querying fuzzy temporal databases}
+In section \ref{subsec:querying-time-intervals}, a technique allowing the querying of fuzzy valid-time databases which contain ill-known valid-time intervals is presented and briefly explained.
+The technique relies on the framework for set evaluation presented in \cite{Pon11}.
+
+\subsection{Querying Fuzzy Valid-Time Databases with Ill-Known Valid-Time Intervals}
 \label{subsec:querying-time-intervals}
+
+One of the main purposes of the existence of databases is to allow information retrieval.
+The standard query language for databases is SQL ~\cite{Mel93}.
+Several proposals to extend the SQL language for transaction-time databases~\cite{Sarda90}, valid-time databases~\cite{gad92} and bitemporal databases ~\cite{TSQL} exist and some authors have studied how to support temporal querying in standard SQL~\cite{Snodgrass98}.
+
+In the following subsections, the query structure, the query evaluation method and the technique to meaningfully rank and aggregate the query results, proposed in this work, are presented.
+
+%In the querying of a fuzzy temporal database it is possible to distinguish among the following cases:
+
+%\begin{itemize}
+%\item
+%Fuzzy data stored in the database and crisp specification in the query.
+%\item
+%Crisp data stored in the database and fuzzy data in the query specification.
+%\item
+%Both data stored in the database and the query specification are fuzzy.
+%\end{itemize}
+%In this subsection we will explain the query specification for a database that stores ill-known time intervals and crisp values in the query.
+Afterwards, the aggregation and ranking of the temporal results is explained.
+
+\subsubsection{Query Structure}
+In this work, a query $\tilde{Q}$ is made up of two separate constructs of user query demands.
+
+%Consider that in a regular or fuzzy relational database, the query specification for the non-temporal attributes is given by $Q$.
+Therefore, the query specification in a temporal or fuzzy temporal database is given by $\tilde{Q}$:
+
+\begin{svgraybox}
+\vspace{-10pt}
+\begin{definition}
+\textbf{Query}
+A query $\tilde{Q}$ in this work is given by:
+\begin{equation}
+\label{eq:query-specification}
+\tilde{Q} = \left( Q^{time}, Q \right)
+\end{equation}
+Here $Q = \left \lbrace q_1,\ \cdots,\ q_n \right \rbrace$ contains all (possibly fuzzy) non-temporal user-defined query demands.
+These comprise all constraints unrelated to the valid-time indications in the queried relations.
+$Q^{time}$ denotes the temporal constraint specified by the user.
+\end{definition}
+\vspace{-10pt}
+\end{svgraybox}
+%%specify better this:
+%\begin{svgraybox}
+%\begin{definition}
+%\label{def:query-constraint}
+%\textbf{(query constraint)}
+%$Q$ is the 
+%A query constraint $q_a$ for an attribute $a$ with attribute domain $\mathcal{D}$, is a restriction in the subset of the values in the attribute domain:
+%\begin{equation}
+%\label{eq:query-constraint}
+%q_a = \left \lbrace d \mid d \subseteq \mathcal{D} \right \rbrace
+%\end{equation}
+%\end{definition}
+
+In this work, the query structure allows the user to specify a single temporal constraint.
+
+\begin{svgraybox}
+\vspace{-10pt}
+\begin{definition}
+\textbf{Temporal Constraint $Q^{time}$}
+In this work, $Q^{time}$ is defined by:
+\begin{equation}
+Q^{time} = \left( I,AR \right)
+\end{equation}
+where $I$ denotes a crisp time interval and $AR$ is one of the Allen relations.
+\end{definition}
+\vspace{-10pt}
+\end{svgraybox}
+
+The interpretation is that, for a record with an ill-known valid-time interval $J$, the user demands that $I$ AR $J$ hold.
+
+\subsubsection{Query evaluation}
+\label{subsubsec:query-evaluation}
+Query satisfaction in a fuzzy relational database is a matter of degree.
+Typically, the query evaluation results in a \emph{satisfaction degree} $s$, which is typically in the unit interval.
+$s \in \left[ 0,1,\right]$.
+Here, a satisfaction degree of $0$ denotes a total dissatisfaction while a degree of $1$ denotes a total satisfaction.
+
+%\begin{svgraybox}
+%\begin{definition}
+%\label{def:evaluation-function}
+%\textbf{Evaluation function}
+%An evaluation function $e_{Q} \left( r \right)$ is a mapping from the attribute value from the row $r$ in the database with respect to the constraints $Q = \left \lbrace q_1, \cdots, q_n \right \rbrace$ to the unit interval $\left[ 0,1 \right]$.
+%\begin{equation}
+%\label{eq:evaluation-function} 
+%e_{Q} = t \mid t \in \left[ 0,1 \right]
+%\end{equation}
+%\end{definition}
+%\end{svgraybox}
+
+In the presented approach, for every record $r$, each part of $\tilde{Q}$ is evaluated independently:
+\begin{itemize}
+\item
+The user preferences expressed in the non-temporal part, $Q$, are evaluated, resulting in a satisfaction degree denoted as $e_Q(r)$.
+The presented approach accepts any valid, sound method of calculating this evaluation, as long as the method is well-founded and $e_Q(r) \in \left[0,1\right]$.
+\item
+The evaluation of the temporal demand expressed in the temporal part, $Q^{time}$, depends on $AR$.
+A specific set of ill-known constraints is considered depending on the Allen relation denoted by $AR$.
+The exact set chosen can be found in table \ref{tab:fuzzy-allen-relations}, for every possible value of $AR$.
+Then both possibility and necessity measures are computed and aggregated.
+\end{itemize}
 
@@
 \end{table}
-
-One of the main purposes of a database is to allow the information retrieval.
-The standard query language for databases is SQL ~\cite{Mel93}, however, there are several proposals to extend the SQL language for transaction-time databases~\cite{Sarda90}, valid-time databases~\cite{gad92} and bi-temporal ~\cite{TSQL}.
-Some authors have studied how to support temporal querying in SQL~\cite{Snodgrass98}.
-
-In the querying of a fuzzy temporal database it is possible to distinguish among the following cases:
-
-\begin{itemize}
-\item
-Fuzzy data stored in the database and crisp specification in the query.
-\item
-Crisp data stored in the database and fuzzy data in the query specification.
-\item
-Both data stored in the database and the query specification are fuzzy.
-\end{itemize}
-In this subsection we will explain the query specification for a database that stores ill-known time intervals and crisp values in the query.
-Afterwards, the aggregation and ranking of the temporal results is explained.
-
-\subsubsection{Query structure}
-Consider that in a regular or fuzzy relational database, the query specification for the non-temporal attributes is given by $Q$.
-Therefore, the query specification in a temporal or fuzzy temporal database is given by $\tilde{Q}$:
-
-\begin{svgraybox}
-\begin{definition}
-\textbf{Query specification}
-A query $\tilde{Q}$ is given by:
-\begin{equation}
-\label{eq:query-specification}
-\tilde{Q} = \left( Q^{time}, Q \right)
-\end{equation}
-Here $Q = \left \lbrace q_1, \cdots, q_n \right \rbrace$ are the possibly fuzzy non-temporal constraints and $Q^{time}$ are the temporal constraints specified in the query:
-\end{definition}
-\end{svgraybox}
-%%specify better this:
-%\begin{svgraybox}
-%\begin{definition}
-%\label{def:query-constraint}
-%\textbf{(query constraint)}
-%$Q$ is the 
-%A query constraint $q_a$ for an attribute $a$ with attribute domain $\mathcal{D}$, is a restriction in the subset of the values in the attribute domain:
-%\begin{equation}
-%\label{eq:query-constraint}
-%q_a = \left \lbrace d \mid d \subseteq \mathcal{D} \right \rbrace
-%\end{equation}
-%\end{definition}
-
-\begin{svgraybox}
-\begin{definition}
-\textbf{Temporal query specification}
-$Q^{time}$ is defined by:
-\begin{equation}
-Q^{time} = \left( I,AR \right)
-\end{equation}
-Where $I$ is a crisp time interval and $AR$ is one of the Allen's relations, see table \ref{tab:fuzzy-allen-relations}.
-The interpretation is that given an ill-known time interval $J$, the user request the set of records in which $I$ AR $J$ hold.
-\end{definition}
-\end{svgraybox}
-
-\subsubsection{Query evaluation}
-\label{subsubsec:query-evaluation}
-Query satisfaction in a (fuzzy) relational database is a matter of degree.
-Typically, each constraint provided by the user is evaluated and results in a \emph{satisfaction degree} $s \in \left[ 0,1,\right]$.
-The value 0 denotes a total dissatisfaction while the value 1 denotes a total satisfaction.
-
-\begin{svgraybox}
-\begin{definition}
-\label{def:evaluation-function}
-\textbf{Evaluation function}
-An evaluation function $e_{Q} \left( r \right)$ is a mapping from the attribute value from the row $r$ in the database with respect to the constraints $Q = \left \lbrace q_1, \cdots, q_n \right \rbrace$ to the unit interval $\left[ 0,1 \right]$.
-\begin{equation}
-\label{eq:evaluation-function} 
-e_{Q} = t \mid t \in \left[ 0,1 \right]
-\end{equation}
-\end{definition}
-\end{svgraybox}
-
-Each element in $\tilde{Q}$ is evaluated independently:
-\begin{itemize}
-\item
-The non-temporal part, $Q$ is evaluated resulting in a satisfaction degree noted as $e_Q(r)$ for each record $r$.
-\item
-The temporal part, $Q^{time}$ is evaluated depending on $AR$.
-A specific set of ill-known constraints is considered depending on the Allen relation in $AR$.
-Then both possibility and necessity measures are computed and aggregated.
-\end{itemize}
 

### 2012-02-27 trunk/book/sources/fuzzylogic.tex
Almost done!

@@
-Possibility theory, like probability theory, deals with uncertainty about the outcome of an experiment.
+%Possibility theory, like probability theory, deals with uncertainty about the outcome of an experiment.
 In probability theory, this uncertainty is caused by the \emph{variability} in the outcomes, while in possibility theory, the uncertainty is caused by \emph{incomplete knowledge} about the experiment.
@@
 
-\begin{definition}
-Consider a set of outcomes $\Omega$.
+%\begin{definition}
+%Consider a set of outcomes $\Omega$.
 Let $\wp(\Omega)$ denote the powerset of $\Omega$ and let $A$ and $B$ be elements of $\wp(\Omega)$.
 A \emph{confidence measure on $\Omega$} is defined by a function
-	\begin{align}
-	g : \wp(\Omega) & \rightarrow \left[0,1\right]
-	\end{align}
-that satisfies
-	\begin{align}
-	g(\emptyset) &= 0 \\
-	g(\Omega) &= 1 	\label{NormalizationProperty} \\
-	A \subseteq B &\Rightarrow g(A) \leq g(B) \label{MonotonicityProperty}
-	\end{align}
-\end{definition}
+%	\begin{align}
+%	g : \wp(\Omega) & \rightarrow \left[0,1\right]
+%	\end{align}
+%that satisfies
+%	\begin{align}
+%	g(\emptyset) &= 0 \\
+%	g(\Omega) &= 1 	\label{NormalizationProperty} \\
+%	A \subseteq B &\Rightarrow g(A) \leq g(B) \label{MonotonicityProperty}
+%	\end{align}
+%\end{definition}
 
-Both possibility measures and necessity measures are special cases of confidence measures.
+%Both possibility measures and necessity measures are special cases of confidence measures.
 
-\begin{definition}
-Consider a confidence measure $\Pi$ on a set of outcomes $\Omega$.
+%\begin{definition}
+%Consider a confidence measure $\Pi$ on a set of outcomes $\Omega$.
 Let $J$ be a countable index set and let $\{ A_{j} | j \in J \wedge A_{j} \subseteq \Omega \}$ be a family of elements of $\wp(\Omega)$.
 $\Pi$ is now a \emph{possibility measure on $\Omega$} if it satisfies:
-	\begin{align}
-	\Pi\left(\bigcup_{j \in J} A_{j} \right) = \sup_{j \in J} \Pi(A_{j})
-	\end{align}
-\end{definition}
+%	\begin{align}
+%	\Pi\left(\bigcup_{j \in J} A_{j} \right) = \sup_{j \in J} \Pi(A_{j})
+%	\end{align}
+%\end{definition}
 
-In this work, the interpretation is as follows.
+%In this work, the interpretation is as follows.
 The possibility of an event expresses how plausible the occurrence of the event seems to an observer of the experiment, given the (partial) knowledge of the observer about the experiment.
 
-Information on the possibility of distinct elements of the universe of discourse $\Omega$ can now be given by a \emph{possibility distribution} $\pi$ on $\Omega$, defined by:
+%Information on the possibility of distinct elements of the universe of discourse $\Omega$ can now be given by a \emph{possibility distribution} $\pi$ on $\Omega$, defined by:
 
-\begin{definition}
-Consider a possibility measure $\Pi$ on $\Omega$. A \emph{possibility distribution} $\pi$ on $\Omega$ underlying the possibility measure $\Pi$ is then a function defined by:
-	\begin{align}
-	\pi : \Omega \rightarrow \left[0, 1\right] : \pi(u) = \Pi(\{u\})
-	\end{align}
-\end{definition}
+%\begin{definition}
+%Consider a possibility measure $\Pi$ on $\Omega$. A \emph{possibility distribution} $\pi$ on $\Omega$ underlying the possibility measure $\Pi$ is then a function defined by:
+%	\begin{align}
+%	\pi : \Omega \rightarrow \left[0, 1\right] : \pi(u) = \Pi(\{u\})
+%	\end{align}
+%\end{definition}
 
-\begin{definition}
-Consider a confidence measure $N$ on a set of outcomes $\Omega$.
+%\begin{definition}
+%Consider a confidence measure $N$ on a set of outcomes $\Omega$.
 Let $J$ be a countable index set and let $\{ A_{j} | j \in J \wedge A_{j} \subseteq \Omega \}$ be a family of elements of $\wp(\Omega)$.
 $N$ is now a \emph{necessity measure} on $\Omega$ if it satisfies:
-	\begin{align}
-	N\left(\bigcap_{j \in J} A_{j} \right) = \inf_{j \in J} N(A_{j})
-	\end{align}
-\end{definition}
+%	\begin{align}
+%	N\left(\bigcap_{j \in J} A_{j} \right) = \inf_{j \in J} N(A_{j})
+%	\end{align}
+%\end{definition}
 
-In this work, the interpretation is as follows.
+%In this work, the interpretation is as follows.
 The necessity of an event expresses how necessary the occurrence of the event seems to an observer of the experiment, given the (partial) knowledge of the observer about the experiment.
 
-Possibility and necessity measures are dual in the sense that:
+%Possibility and necessity measures are dual in the sense that:
 
-\begin{align}
-\forall A \subseteq \Omega : N(A) = 1 - \Pi(\bar{A})
-\end{align}
+%\begin{align}
+%\forall A \subseteq \Omega : N(A) = 1 - \Pi(\bar{A})
+%\end{align}
 
-Regarding interpretation, the above can be seen as: the degree to which an event is necessary is the degree to which every other possible event is not plausible.
+%Regarding interpretation, the above can be seen as: the degree to which an event is necessary is the degree to which every other possible event is not plausible.
 

### 2012-02-27 trunk/book/sources/temporalDatabases.tex
Almost done!

@@
 
-In the following sections, one of the main points of attention concerning relational temporal databases is explained and discussed.
+In the following sections, some of the main points of attention concerning relational temporal databases are explained and discussed.
 
@@
 In temporal relations, different records can exist with the same values for every attribute except the attributes describing the temporal aspects of the objects or concepts represented by the records (the temporal attributes).
-These different records represent different temporal versions of the same real objects or concepts.
+These different records represent different temporal versions of the same real object or concept.
 To allow the existence of such records, the most common solution is to extend the original primary key with the temporal attributes.
@@
 \label{ex:pk}
-Consider the example database visualized in table \ref{table:example-database}.
+Consider the example valid-time relation visualized in table \ref{table:example-database}, which contains data representing employees in a company.
+The attributes `Name', `Age' and `Supervisor' describe the name, age and supervisor of an employee, respectively.
+The attributes `Start' and `End' describe the year when an employee started, respectively finished working for the company.
+For example, employee Sarah started working for the company in 2005 and finished in 2009.
+The attribute `ID' uniquely identifies an employee.
+Consider `ID' being the primary key and consider the company wanting to hire Sarah again.
+Another tuple should thus be inserted into the relation, containing value 4 as `ID'.
+This will of course not be allowed by the primary key.
+This problem is solved by defining a new primary key: `ID', `Start', `End'.
+Now, the resulting relation allows inserting a new tuple representing that Sarah was rehired in 2010, as shown in table \ref{table:example-database-with-new-pk}.
+%Consider the example database visualized in table \ref{table:example-database}.
 It contains data representing employees in a company, and two temporal values which represent a valid-time interval.
@@
 
+\end{example}
+
+Extending a primary key to include temporal attributes allows for the insertion of spurious values, for example representing temporal inconsistencies.
+This is shown in table \ref{table:example-database-with-new-pk}, where it seems Sarah was rehired in 2007 and fired in 2008.
+
 %\vspace{-10pt}
@@
 \centering
-\caption{Example database}
+\caption{Example relation modelling the employees of a company.}
 \begin{tabular}{c c c c c c }
 \hline
-\textbf{ID} & \textbf{Name} & \textbf{Age} & \textbf{Works for} & \textbf{Start year} & \textbf{End year} \\ \hline
+\textbf{ID} & \textbf{Name} & \textbf{Age} & \textbf{Supervisor} & \textbf{Start} & \textbf{End} \\ \hline
 1 & Peter & 24 & John &  2010 & - \\
@@
 \centering
-\caption{Example database with new primary key and some spurious values}
+\caption{Example relation from example \ref{ex:pk} with new primary key and some spurious values.}
 \begin{tabular}{c c c c c c }
 \hline
-\textbf{ID} & \textbf{Name} & \textbf{Age} & \textbf{Works for} & \textbf{Start year} & \textbf{End year} \\ \hline
+\textbf{ID} & \textbf{Name} & \textbf{Age} & \textbf{Supervisor} & \textbf{Start year} & \textbf{End year} \\ \hline
 1 & Peter & 24 & John &  2010 & - \\
@@
 %\vspace{-25pt}
+This latter issue is the concern of the next subsection.
 
-\end{example}
+\subsection{\label{subsubsec:consistency}Consistency in Temporal Relations}
+As stated in subsection \ref{subsubsec:primary-key}, one of the main issues concerns the insertion of spurious values in a relation.
+The most usual approach to deal with this is to change the database DML to enforce the consistency with which the temporal relation models the temporal properties of the part of reality the relation models.
+An example of such DML adaptation is given below.
 
-\subsection{\label{subsubsec:consistency}Consistency on temporal databases}
-The consistence mechanism in a temporal database is usually re-defined.
+%The consistence mechanism in a temporal database is usually re-defined.
 The main problem is the insertion of spurious values in the database, as shown in example \ref{ex:pk}, in table \ref{table:example-database-with-new-pk}.
@@
 \begin{example}
-Consider the database in table \ref{table:example-database}. The following \emph{update} sentence:
+\textcolor{red}{PROBLEM}
+Consider again the example relation visualized in table \ref{table:example-database}. Imagine John left the company, but nobody updated the relation, so it is not known when
+
+The following \emph{update} sentence:
 
@@
 \centering
-\caption{Example database after update sentence}
+\caption{Example relation from example \ref{ex:pk} updated maintaining consistency.}
 \begin{tabular}{c c c c c c }
@@
 
-\subsection{\label{Comm-temp}Commercial Temporal database systems}
-There are several commercial temporal DBMS. Table \ref{table:commercial-temporal-db} is a comparative among the most well-known systems. It is also provided a reference for more information.
+\subsection{\label{Comm-temp}Commercial Temporal Database Systems}
+There are several commercial temporal DBMS. Table \ref{table:commercial-temporal-db} describes the most well-knownand provides references for more information.
 
-Oracle workspace manager~\cite{oracle2009} and TimeDB~\cite{timedb2005} are libraries for dealing with time in an OracleDB.
-Nevertheless TimeDB and Postgree Temporal~\cite{posgree2009} are similar.
-Both are simple implementations that implement a subset of the Allen's operators and some operations for the creation and manipulation of a temporal attribute (valid, transaction or both times) in a table.
-Nevertheless, Teradata~\cite{teradata2011} is mainly a business intelligence database designed for data-mining.
+Oracle workspace manager~\cite{oracle2009} and TimeDB~\cite{timedb2005} are libraries for dealing with time in OracleDB.
+Nevertheless, TimeDB and Postgree Temporal~\cite{posgree2009} are similar: both are simple implementations that implement a subset of the Allen operators and some operations for the creation and manipulation of temporal attributes (valid time, transaction time or both times are supported).
+Teradata~\cite{teradata2011} is mainly a business intelligence system designed for data mining.
+The most complete implementation is Workspace Manager.
 
-The most complete implementation is the Workspace Manager. Unfortunately, none of these systems work with uncertainty or imprecision neither in the data storage nor in the query.
+Unfortunately, none of these systems take imperfect information into account, neither in data storage nor in querying.
 
@@
 \centering
-\caption{Commercial implementations for temporal database systems. }
+\caption{Commercial Temporal Database Systems. }
 \begin{tabular}{c c c c c c }
 \hline
-\textbf{Name} & \textbf{Time managed} & \textbf{Comments} & \textbf{References}  \\ \hline
+\textbf{Name} & \textbf{Time Supported} & \textbf{Comments} & \textbf{Reference}  \\ \hline
 Oracle Workspace Manager & VT and TT. & Package for Oracle DB. & \cite{oracle2009}\\

### 2012-02-27 trunk/book/sources/timeDomain.tex
Almost done!

@@
 
-\textcolor{red}{MOVE TO TEMPORAL DATABASE PART}
-On the other hand, time may be~\cite{Jensen94thetsql2} \textbf{relative} or \textbf{absolute} (in other words, \textbf{anchored} or \textbf{unanchored}).
+%\textcolor{red}{MOVE TO TEMPORAL DATABASE PART}
+%On the other hand, time may be~\cite{Jensen94thetsql2} \textbf{relative} or \textbf{absolute} (in other words, \textbf{anchored} or \textbf{unanchored}).
 Sometimes, what one considers \emph{absolute time} is not as definite as one would hope since our concept of absolute time is based on another time reference (for example, January 1st of year one).
@@
 -3 days = three days ago).
-\textcolor{red}{END MOVE}
+%\textcolor{red}{END MOVE}
 
@@
 \centering
-\input{graphs/allen.tex}
+\includegraphics[scale=0.5]{graphs/allen.eps}
 \caption{Allen relations between two time intervals.}

### 2012-02-27 book/author.tex
A new version with extended definition of granularity and calendar.
added graph for the workspace manager. Added some text for the commercial temporal databases. Added fuzzy allen relations.
Update of the titles in the file author.tex. First version of the conclusions and further reseach work.
Almost done!

--- 2011-11-09 trunk/book/author.tex
@@
 \usepackage{url}
+\usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage{multirow}
+\usepackage{color}
+
+\hyphenation{pro-per-ties}
+\hyphenation{ge-ne-ral-ly}
+\hyphenation{pre-fe-ren-ces}
+\hyphenation{u-sing}
+\hyphenation{pu-nish-ment}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \newcommand{\T}{\mathcal{T}}
-
+\newcommand{\C}{\mathcal{C}}
+\newcommand{\Nat}{\mathbb{N}}
+\newcommand{\Q}{\mathbb{Q}}
+\newcommand{\R}{\mathbb{R}}
+\newcommand{\Z}{\mathbb{Z}}
 \begin{document}
 
-\title*{An overview on temporal databases.\\
- New approaches and proposals.}
+\title*{An Overview on Temporal Databases.\\
+ New Approaches and Proposals.}
 % Use \titlerunning{Short Title} for an abbreviated version of
 % your contribution title if the original one is too long
-\author{Jos\'e Pons \and Christophe Billiet \and Olga Pons}
+\author{Jos\'e Pons \and Christophe Billiet \and Olga Pons \and Guy De Tr\'e}
 % Use \authorrunning{Short Title} for an abbreviated version of
 % your contribution title if the original one is too long
-\institute{Jos\'e Pons \at Department of Computer Science and Artificial Intelligence, Univeristy of Granada, C/ Periodista Daniel Saucedo Aranda s/n \email{jpons@decsai.ugr.es}
-\and Christophe Billiet \at Department of Telecommunications and Information Processing, Ghent University, Belgium, St.-Pietersnieuwstraat 41, B-9000 Gent,Belgium.
-\email{Christophe.Billiet@telin.ugent.be}
-\and Olga Pons \at Department of Computer Science and Artificial Intelligence, Univeristy of Granada, C/ Periodista Daniel Saucedo Aranda s/n \email{opc@decsai.ugr.es}
+\institute{
+Jos\'e Pons \at Department of Computer Science and Artificial Intelligence, Escuela T\'ecnica Superior de Ingenier\'ia Inform\'atica, Universidad de Granada,  C/Periodista Daniel Saucedo Aranda s/n, E-18071, Granada, Spain.
+\email{jpons@decsai.ugr.es}
+\and Christophe Billiet \at Department of Telecommunications and Information Processing, Ghent University, St.-Pietersnieuwstraat 41, B-9000 Gent, Belgium. \email{Christophe.Billiet@telin.ugent.be}
+\and Olga Pons \at Department of Computer Science and Artificial Intelligence, Escuela T\'ecnica Superior de Ingenier\'ia Inform\'atica, Universidad de Granada,  C/Periodista Daniel Saucedo Aranda s/n, E-18071, Granada, Spain.
+\email{opc@decsai.ugr.es}
+\and Guy De Tr\'e \at Department of Telecommunications and Information Processing, Ghent University, St.-Pietersnieuwstraat 41, B-9000 Gent, Belgium. \email{Guy.DeTre@telin.ugent.be}
 }
@@
 \input{sources/temporalDatabases.tex}
-
-\section{Imprecision, Uncertainty and vagueness in temporal databases}
+\def\JPicScale{1}
+\section{Imperfection in Temporal Databases}
 \label{sec:imprecision-uncertainty}
@@
 
-\section{Bipolarity on temporal databases}
-\label{sec:bipolarity}
-\input{sources/bipolarity.tex}
-
-\section{Further research work}
+%\section{Bipolarity on temporal databases}
+%\label{sec:bipolarity}
+%\input{sources/bipolarity.tex}
+\def\JPicScale{1}
+\section{Conclusions and Further Research Work}
 \label{sec:further-research-work}
 \input{sources/frw.tex}
-
+\def\JPicScale{1}
 \section*{Acknowledgements}
@@
 \input{referenc}
+\def\JPicScale{0.5}
+
+\newpage
+\section*{Appendix: Introduction to fuzzy logic}
+\label{app:fuzzy-logic}
+\input{sources/fuzzylogic.tex}
+
 \end{document}

### 2012-02-27 trunk/book/sources/frw.tex
Last draft of the paper. This is the final draft.

@@
-In this work we have presented the main properties of the time in an information system.
+In this chapter, some of the main concepts concerning temporal modelling and temporal modelling in information systems and the terminology corresponding with these concepts are introduced and explained and some of the main properties of and issues with these concepts are presented and discussed.
+An overview of some commercial temporal DBMS is briefly introduced.
+Finally, a novel technique for querying valid-time relations using imperfect query specifications is presented.
+
+ %we have presented the main properties of the time in an information system.
 An overview of the commercial temporal DBMS has also been introduced.
@@
 
-Further research work should be done in several ways: first of all, a theoretical model for both dealing with uncertainty in the database as well as in the querying should be defined.
+Further research work could follow several general directions.
+First of all, a theoretical model for dealing with uncertainty in both the database and the query at the same time could be researched and defined.
+Next, implementations including both DDL and DML could be proposed and constructed.
+
+%Further research work should be done in several ways: first of all, a theoretical model for both dealing with uncertainty in the database as well as in the querying should be defined.
 Then, another interesting work is to propose an implementation with both DDL (Data Definition Language) and DML (Data Manipulation Language).

### 2012-02-27 trunk/book/sources/fuzzyTemporalDatabases.tex
update with necessity in allen relations
Last draft of the paper. This is the final draft.

@@
 The evaluation of the temporal demand expressed in the temporal part, $Q^{time}$, depends on $AR$.
-A specific set of ill-known constraints is considered depending on the Allen relation denoted by $AR$.
+A specific set of ill-known constraints\cite{Pon11} is considered depending on the Allen relation denoted by $AR$.
 The exact set chosen can be found in table \ref{tab:fuzzy-allen-relations}, for every possible value of $AR$.
-Then both possibility and necessity measures are computed and aggregated.
+Then, using these formulas given in table \ref{tab:fuzzy-allen-relations}, both the possibility degree $\Pos_{Q^{time}}(r)$ and the necessity degree $\Nec_{Q^{time}}(r)$ are computed.
 \end{itemize}
+
+\textcolor{red}{TODO: maybe part on what these and mean? Maybe part on where the formulas in table \ref{tab:fuzzy-allen-relations} come from?}
 
@@
 \centering
-\begin{tabular}{|c|c|}
+\begin{tabular}{|c|c|c|}
 \hline
-Allen Relation &  $\Pos\left(\text{I satisfies all constraints }C_{i}, i = 1,2,...\right)$ \\
+Allen Relation &  $\Pos\left(\text{I satisfies all constraints }C_{i}, i = 1,2,...\right)$ & $\Nec\left(C_{i}, i = 1,2,... \right)$ \\
 \hline
-I before J & $\sup_{a>w}\pi_x(w)$\\
+I before J & $\sup_{a>w}\pi_X(w)$ & $\inf_{a \leq w} 1- \pi_X(w)$\\
+\hline
+
+\multirow{4}{*}
+{I equal J} &  $\min ( \sup_{a \leq w}\pi_X(w),$  & $\min ( \inf_{a>w} 1-\pi_X(w),$\\
+& $\pi_X(w),$ & $\inf_{w \in \left[a,b\right]} \pi_X(w), $ \\
+ &  $\sup_{b \geq w}\pi_Y(w),$ & $\inf_{b>w} 1-\pi_Y(w),$\\
+ & $\pi_Y(w))$ & $\inf_{w \in \left[a,b\right]} \pi_Y(w) )$ \\
 \hline
@@
 \multirow{2}{*}
-{I equal J} &  $\min ( \sup_{a \leq w}\pi_x(w),\pi_x(w),$ \\
- &  $\sup_{b \geq w}\pi_Y(w),\pi_Y(w))$\\
+I meets J  & $\min (\sup_{a\geq w} \pi_X(w),$ & $\min (\inf_{a<w} 1 - \pi_X(w),$\\
+& $\pi_X(w))$  & $ \inf_{w \in \left[a,b\right]} \pi_X(w) )$ \\
 \hline
 
-I meets J  & $\min (\sup_{a\geq w} \pi_X(w),$ $\pi_X(w))$ \\
+\multirow{3}{*}
+{I overlaps J}  & $\min ( \sup_{b>w}\pi_Y(w), $ & $\min ( \inf_{b \geq w} 1 - \pi_Y(w),$\\
+ & $\sup_{a \geq w}\pi_X(w),$& $1- \inf_{a>w}1-\pi_X(w),$\\
+ & $\sup_{a \leq w}\pi_X(w))$ & $1- \inf_{a<w}1-\pi_X(w) )$ \\
 \hline
 
+\multirow{4}{*}
+{I during J}  & $\max ( \min ( \sup_{a<w}\pi_X(w),,$ & $\max (\min (\inf_{a \leq w}1-\pi_X(w), $\\
+ & $\sup_{b \geq w}\pi_Y(w)) $ & $ \inf_{b>w} 1- \pi_Y(w)),$ \\
+ & $\min ( \sup_{a \leq w }\pi_X(w),$ & $\min ( \inf_{a<w}1-\pi_X(w),$\\
+ & $\sup_{b>w}\pi_Y(w))$ & $\inf_{b\geq w}1-\pi_Y(w)))$ \\
+\hline
 \multirow{2}{*}
-{I overlaps J}  & $\min ( \sup_{b>w}\pi_Y(w), $ \\
- & $\sup_{a \geq w}\pi_X(w),\sup_{a \leq w}\pi_X(w))$ \\
+{I starts J} &  $\min( \sup_{a \leq w}\pi_X(w),$  & $\min (\inf_{a<w}1-\pi_X(w),$\\
+& $\pi_X(w))$ & $inf_{w \in \left[a,b\right]} 1- \pi_X(w))$ \\
 \hline
-
 \multirow{2}{*}
-{I during J}  & $\max ( \min ( \sup_{a<w}\pi_X(w),\sup_{b \geq w}\pi_Y(w)),$ \\
- & $\min ( \sup_{a \leq w }\pi_X(w),\sup_{b>w}\pi_Y(w)$\\
-\hline
-
-{I starts J} &  $\min( \sup_{a \leq w}\pi_X(w),$ $\pi_X(w))$\\
-\hline
-
-{I finishes J} &  $\min ( \sup_{b \geq w} \pi_Y(w),$ $\pi_Y(w))$ \\
+{I finishes J} &  $\min ( \sup_{b \geq w} \pi_Y(w),$  & $\min(\inf_{b>w}1-\pi_Y(w),$\\
+& $\pi_Y(w))$ & $\inf_{w \in [a,b]} 1- \pi_Y(w))$ \\
 \hline
@@
 
-\subsubsection{Ranking and aggregation}
-For each temporal attribute $r_{time}$ for the record $r$, both possibility $\Pos \left( Q_{time}\right)$ and necessity $\Nec \left( Q_{time} \right)$ measures are obtained.
+\subsubsection{Aggregation and Ranking}
+To be able to present the most appropriate results to the user, for every record $r$, an aggregation method is used to aggregate $\Pos_{Q^{time}}(r)$ and $\Nec_{Q^{time}}(r)$ into a temporal record rank $e_{Q^{time}}(r)$ and after this, a convex combination combining $e_{Q^{time}}(r)$ and $e_{Q}(r)$ will provide the final record rank $e_{final}(r)$.
+
+%For each temporal attribute $r_{time}$ for the record $r$, both possibility $\Pos \left( Q_{time}\right)$ and necessity $\Nec \left( Q_{time} \right)$ measures are obtained.
 Then, the measures are combined in order to obtain an evaluation score $e_{Q_{time} \left( r \right)}$.
@@
 
+To calculate $e_{Q^{time}}(r)$, an a simple and crude method is used:
+
 \begin{equation}
 \label{eq:temporal-eval}
-e_{Q_{time} \left( r \right)} = \left \lbrace t \mid t = \left( \frac{\Pos \left( Q_{time}\right) + \Nec \left( Q_{time} \right)}{2} \right)   \in \left[ 0,1\right] \right \rbrace
+e_{Q^{time}}(r) = \left(\frac{\Pos_{Q^{time}}(r)+\Nec_{Q^{time}}(r)}{2}\right)
+%e_{Q_{time} \left( r \right)} = \left \lbrace t \mid t = \left( \frac{\Pos \left( Q_{time}\right) + \Nec \left( Q_{time} \right)}{2} \right)   \in \left[ 0,1\right] \right \rbrace
 \end{equation}
 
-This measure provides a natural score because of the following property of the necessity measure: 
-\begin{equation}
-\label{eq:necessity} %iff or iif?
-\Nec \left( Q_{time} \right) > 0  \Longleftrightarrow  \Pos \left( Q_{time}\right) = 1
-\end{equation}
+This method aims to provide the result records with a natural ranking based on the users temporal constraint.
+$e_{Q^{time}}(r)$ will of course be a value in $\left[0,1\right]$, as both $\Pos_{Q^{time}}(r) \in \left[0,1\right]$ and $\Nec_{Q^{time}}(r) \in \left[0,1\right]$.
+The purpose is that records which fit the users temporal demand better get a higher score than records fitting the temporal demand worse.
+Here, this aim is reached because the necessity degree $\Nec_{Q^{time}}(r)$ cannot exceed $0$ unless the possibility degree $\Pos_{Q^{time}}(r)$ equals $1$.
 
-The final ranking is given by a convex combination of both temporal and non-temporal evaluations:
+%This measure provides a natural score because of the following property of the necessity measure: 
+%\begin{equation}
+%\label{eq:necessity} %iff or iif?
+%\Nec \left( Q_{time} \right) > 0  \Longleftrightarrow  \Pos \left( Q_{time}\right) = 1
+%\end{equation}
+
+The final ranking $e_{final}(r)$ for a record $r$ is now given by a convex combination of both temporal and non-temporal evaluation scores.
 
@@
 \label{eq:convex-combination}
-e_{final} \left( r \right) = \left \lbrace t \mid t = \omega \ast e_{Q} \left( r \right) + \left( 1- \omega \right) \ast e_{Q^{time}}, t,\omega \in \left[0,1\right] \right \rbrace
+e_{final} \left( r \right) = \omega \ast e_{Q} \left( r \right) + \left( 1- \omega \right) \ast e_{Q^{time}}
 \end{equation}
+
+A convex combination is used mainly for 2 reasons:
+\begin{itemize}
+	\item The use of this convex combination allows a record to make up for a low temporal evaluation score with a high non-temporal evaluation score and vice versa.
+	\item The exact value of $\omega$ can now be modified to ascribe more value to either the fulfillment of the user's temporal demands or the fulfillment of the user's non-temporal constraints.
+\end{itemize}
 
@@
 \label{subsubsec:bipolarity}
-Humans express their preferences in both positive and negative statements.
-This property is interesting in querying of databases, because sometimes, the user does not know exactly his / her preferences but what he / she does not want.
-\emph{Bipolarity} is therefore a technique to model both positive and negative preferences.
-Sometimes positive and negative information are clearly symmetric and therefore, it is possible to derive one information from another e.g.
-a person may define the concept of tall as 1.80 meters or higher.
-The negative would be the opposite, not tall is therefore, being less than 1.80 meters.
-But in some cases, positive information can not be obtained directly from the negative information and vice-verse.
+Humans express their preferences using both positive and negative statements, where positive statements express what is desired or acceptable and negative statements express what is undesired or unacceptable \cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy}.
+This realization is interesting with regard to database querying, because sometimes a user does not exactly know his or her preferences or can't express them in only positive statements, but prefers to use negative statements to express what he or she dislikes or doesn't need.
+This introduces the need for \emph{bipolar querying}, a technique to model both positive and negative user preferences in a database query.
+Sometimes positive and negative preferences are clearly symmetric, making it possible to derive one from the other.
+For example, a person may define the concept of `tall' as `1.80 meters or higher'.
+The negative would then be the opposite: not tall would be `anyhing less than 1.80 meters'.
+However, in some cases, positive preferences can not be directly obtained from negative preferences or vice versa.
 E.g.
-a person wants to buy a blue motorbike, this does not mean that the person would reject a black motorbike.
-This heterogeneous semantics are given by the \emph{heterogeneous bipolarity}~\cite{Dubois2006},~\cite{Dubois2008}.\\
+when a person prefers to buy a black motorbike, this does not necessarily mean the person would totally reject a very dark blue motorbike.
+This phenomenon is called \emph{heterogeneous bipolarity}~\cite{Dubois2006},~\cite{Dubois2008}.
 
-Some previous work have been done in ~\cite{Lacroix87}.
+The use of imprecise query preference formulation in bipolar querying is well discussed in existing literature \cite{DeTre2009}, \cite{Dubois2008}, \cite{Lacroix87}.
+In \cite{Lacroix87}, desired and mandatory query conditions are used, instead of positive and negative preferences.
+However, the inverse of a mandatory preference expresses what should be rejected and this could be seen as negative information, whereas desired query conditions can be seen as positive preferences.
+However, the combination of bipolar querying and the use of imprecise query preferences in the context of temporal databases is not so well discussed in existing literature.
+A proposal for the bipolar querying of valid-time databases has been made by Billiet et al.~\cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy}.
+The model presented there deals with a fuzzy valid-time specification based on~\cite{garrido2009}.
+
+%Some previous work have been done in ~\cite{Lacroix87}.
 There, instead of positive and negative criteria, desired and mandatory query conditions are explained.
@@
 
-This introduces the necessity of a bipolar way of querying databases.
+%This introduces the necessity of a bipolar way of querying databases.
 Therefore, the bipolar querying of databases~\cite{DeTre2009} is a querying technique which allows the user to express both positive and negative criteria in the database query.\\
 
-Bipolarity is handled by means of different tools:
+Bipolarity can be handled using different concepts, such as intuitionistic fuzzy sets~\cite{Atanassov1986}, interval valued fuzzy sets~\cite{Zadeh75a} Grattan-Guiness~\cite{Grattan76}, Janh~\cite{Jahn75}, Sambuc~\cite{Sambuc75},~\cite{Dubois05} or two fold fuzzy sets \cite{Dubois02}.
 
+%\begin{itemize}
+%\item
+%Intuitionistic Fuzzy Sets~\cite{Atanassov1986} are sometimes used.
+The approach re-define the concept of inverse.
+Instead of considering the inverse of a set just the complementary, a AFS is defined by means of a membership function and a non-membership function.
+%\item
+%The interval valued fuzzy sets \emph{IVFS} were introduced independently by Zadeh~\cite{Zadeh75a}, Grattan-Guiness~\cite{Grattan76}, Janh~\cite{Jahn75}, Sambuc~\cite{Sambuc75} in the seventies~\cite{Dubois05}.
+%look for the reference!
+%\item
+%The two fold fuzzy sets \emph{(TFS)} by Dubois and Prade~\cite{Dubois02}. This approach defines desired and mandatory query conditions.
+%\end{itemize}
+
+From a theoretical point of view, bipolarity might be found either in the queries presented to a database system or in a database managed by a database system.
+
+When bipolarity is found in queries, it is possible to distinguish between:
 \begin{itemize}
 	\item
-The intuitionistic fuzzy sets  \emph{AFS} by Atannasov ~\cite{Atanassov1986}.
-The approach re-define the concept of inverse.
-Instead of considering the inverse of a set just the complementary, a AFS is defined by means of a membership function and a non-membership function.
+	Bipolarity inside query criteria: each individual query criterion may be specified using both positive and negative preferences.
+For example when querying a car database, the user can express that he or she wants a black car, but definitely not a red neither a blue one.
+Bipolarity resides here within the car color criterion.
 	\item
-The interval valued fuzzy sets \emph{IVFS} were introduced independently by Zadeh~\cite{Zadeh75a}, Grattan-Guiness~\cite{Grattan76}, Janh~\cite{Jahn75}, Sambuc~\cite{Sambuc75} in the seventies~\cite{Dubois05}.
-%look for the reference!
-\item
-The two fold fuzzy sets \emph{(TFS)} by Dubois and Prade~\cite{Dubois02}. This approach defines desired and mandatory query conditions.
+	Bipolarity outside query criteria: the query is specified using a global positive and a global negative preference part.
+For example when querying a car database, the user can express that he or she wants a black car, but definitely not a car with a fuel consumption of 6 liters or more.
 	\end{itemize}
 
-From a theoretical point of view, in a database, bipolarity may be found at:
-\begin{description}
-\item[\textbf{The query}] It is possible to distinguish between:
-	\begin{itemize}
-	\item
-	Inside criterion bipolarity: Each individual criteria may be specified with both positive and negative conditions. E.g. the user wants a black car, but definitively not a red neither a blue one.
-	\item
-	Outside criterion bipolarity: The query is specified with a global positive and negative conditions.
-E.g.
-The user wants a black car with a fuel consumption of 5 liters or less but definitively not a white car with a fuel consumption of 6 liters or more.
-	\end{itemize}
-\item[\textbf{The data storage}] In the database some attributes or even at tuple level it should be possible to specify them with bipolarity.
-Nevertheless, there is not so many research on this topic.
+Concerning bipolarity inside a database, it should be possible to specify both positive and negative real world object or concept aspects, even at record level.
+Nevertheless, not so much research exists concerning bipolarity in databases.
+
+%some attributes or even at tuple level it should be possible to specify them with bipolarity. Nevertheless, there is not so many research on this topic.
 %	\begin{itemize}
@@
 %	\end{itemize}
-\end{description}
-
-A proposal for the bipolar querying of temporal databases has been made by Billiet~\cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy}.
-First of all, the query structure is shown.
-Then, both evaluation and ranking methods are defined.
-The model deals with a fuzzy valid-time specification in~\cite{garrido2009}.
 

### 2012-02-27 trunk/book/sources/temporalDatabases.tex
Last draft of the paper. This is the final draft.

@@
 
+\begin{note}
+In the following tables (\ref{table:example-database},\ref{table:example-database-with-new-pk},\ref{table:example-database-update}) the relation `Supervisor' stores the ID of the employee.
+To improve the readability of the example, the tables show the name of the employee.
+Note also that the field for the age is expressed as the age in years whereas the value stored in the database is the birthday.
+\end{note}
+
 Extending a primary key to include temporal attributes allows for the insertion of spurious values, for example representing temporal inconsistencies.
@@
 %\vspace{-25pt}
-This latter issue is the concern of the next subsection.
+The next subsection is concerned with another interresting issue.
 
@@
 As stated in subsection \ref{subsubsec:primary-key}, one of the main issues concerns the insertion of spurious values in a relation.
+Another issue concerns the possible breach of consistency when updating records.
 The most usual approach to deal with this is to change the database DML to enforce the consistency with which the temporal relation models the temporal properties of the part of reality the relation models.
@@
 \begin{example}
-\textcolor{red}{PROBLEM}
-Consider again the example relation visualized in table \ref{table:example-database}. Imagine John left the company, but nobody updated the relation, so it is not known when
+Consider the database in table \ref{table:example-database}. The employee with ID=3 (John) works now for the employee with ID= 4 (Sarah). The update is made in the following sentence:
 
@@
 \begin{verbatim}
-Update Employees set StartYear = 2010 where ID=3;
+Update Employees set 'Works for' = 4 where ID=3;
 \end{verbatim}
 
-Is translated into the following two sentences: an update for the last version of the row and an insert sentence for the new version. The update sentence showed above is translated into:
+Is translated into the following two sentences: an update for the last version of the row and an insert sentence for the new version.
+Consider that the time in the  system is the year 2010.
+Then, the update sentence showed above is translated into:
 
@@
 Update Employees set EndYear = 2010 where ID=3;
-Insert into Employees values (3,John,52,-,2010,-);
+Insert into Employees values (3,John,52,Sarah,2010,-);
 \end{verbatim}
@@
 3 & John & 52 & - &  1999 & 2010 \\
-3 & John & 52 & - &  2010 & - \\
+3 & John & 52 & Sarah &  2010 & - \\
 4 & Sarah & 29 & Maria &  2005 & 2009 \\

### 2012-02-27 trunk/book/sources/timeDomain.tex
Last draft of the paper. This is the final draft.

@@
 
-\textcolor{red}{MOVE TO CHAPTER 4}
-The transition between granularities~\cite{Lin97} is also considered as a source of imprecision.
+%\textcolor{red}{MOVE TO CHAPTER 4}
+%The transition between granularities~\cite{Lin97} is also considered as a source of imprecision.
 In a sentence like \emph{`The Alhambra is 612 years old'}, the temporal notion is very precise if time periods shorter than a year are irrelevant in the used granularity.
@@
 
-\textcolor{red}{END MOVE TO CHAPTER 4}
+%\textcolor{red}{END MOVE TO CHAPTER 4}
 
@@
 Figure \ref{fig:allen} shows the temporal relationships Allen discerned.
-Some proposals can be applied to both crisp and fuzzy temporal intervals~\cite{ohlbach2004},~\cite{nagypal2003},~\cite{schockaert08}.
-
-\textcolor{red}{MOVE TO FUZZY PART}
-For this fuzzy comparisons~\cite{garrido2009} proposed some different temporal relations that can be easily implemented by means of regular fuzzy operations.
-\textcolor{red}{END MOVE TO FUZZY PART}
+Some proposals can be applied to both crisp and fuzzy temporal intervals~\cite{ohlbach2004},~\cite{nagypal2003},~\cite{schockaert08},~\cite{garrido2009}.
 

### 2012-03-01 book/sources/intro.tex
Almost done!
The new form of the chapter...

@@
 The concept of time itself is very complex to handle and interpret~\cite{klein94},~\cite{Shackle61}, though it is very natural and omnipresent.
-As information systems often attempt the modelling or the representation of natural objects, concepts or processes, they often require modelling or representing temporal information.
+As information systems often attempt the modelling of natural objects, concepts or processes, they often require modelling temporal aspects or concepts.
 Thus, several proposals have been concerned with the obtaining of theoretical models that allow the modelling or representation of time~\cite{Bolour82},~\cite{Cru97}.
@@
 A database contains data representing real objects or concepts.
-Every one of these data is a measurement or description of a property of a real object or concept.
+Every one of these data is a result value of a measurement of a property or a description of a property of a real object or concept.
 In reality, some aspects or properties of objects or concepts are time-variant or time-related.
@@
 Without further precautions, a library employee could add several records with the same book identifier, different `loaned'-dates and no `returned'-dates.
-This would represent the same physical book being loaned several times on different dates and never returned, which is of course impossible.
+This group of records would represent the same physical book being loaned several times on different dates and never returned, which is of course impossible.
 A temporal database model will typically constrain record insertion and prevent similar modelling inconsistencies.
 
-A lot of research concerns these temporal database models and their approaches to the modelling of time.
+A lot of research concerns temporal database models and their approaches to the modelling of time.
 The first efforts were towards the representation of historical information related to objects represented by records in a database~\cite{Clifford:1985:AHR:971699.318922}.
@@
 
-Humans handle temporal information using certain temporal notions like time intervals or time points~\cite{Dyreson1994}.
-Humans are often not hindered by imperfections like imprecisions, vaguenesses, uncertainties or inconsistencies possibly contained in the descriptions of these temporal notions.
+Humans handle temporal information using certain temporal notions like time intervals or time points~\cite{Dyreson1994}, and they are often not hindered by imperfections like imprecisions, vaguenesses, uncertainties or inconsistencies possibly contained in the descriptions of these temporal notions.
 Among many others, these possible imperfections in descriptions of temporal notions determine an important issue in temporal modelling.
 E.g.
-the temporal notion in a sentence like `The Belfry of Bruges was finished on a day somewhere between 1/01/1201 A.D.
+the description of the temporal notion in a sentence like `The Belfry of Bruges was finished on a day somewhere between 1/01/1201 A.D.
 and 31/12/1300 A.D.' contains imperfection because of the uncertainty in the used time-related expression.
@@
 In section \ref{sec:temporal-databases}, some basic concepts and terminology about temporal databases are presented, followed by an overview of some interesting issues concerning temporal databases and a survey of some commercial temporal database systems.
-\textcolor{red}{TODO: REFERENCE SECTION 4!}
+Finally, in section \ref{sec:imprecision-uncertainty}, an approach to querying temporal databases containing imperfect temporal information is presented, followed by some conclusions and some suggestions for future work in section \ref{sec:further-research-work}.
 

### 2012-03-01 trunk/book/author.tex
Ok, some modifications to the second chapter. Let me know what you think...
Last draft of the paper. This is the final draft.
The new form of the chapter...

@@
 \newcommand{\R}{\mathbb{R}}
+\newcommand{\Z}{\mathbb{Z}}
 \begin{document}
@@
 % your contribution title if the original one is too long
+%\institute{
+%Jos\'e Pons \at Department of Computer Science and Artificial Intelligence, Escuela T\'ecnica Superior de Ingenier\'ia Inform\'atica, Universidad de Granada,  C/Periodista Daniel Saucedo Aranda s/n, E-18071, Granada, Spain.
+%\email{jpons@decsai.ugr.es}
+%\and Christophe Billiet \at Department of Telecommunications and Information Processing, Ghent University, St.-Pietersnieuwstraat 41, B-9000 Gent, Belgium. \email{Christophe.Billiet@telin.ugent.be}
+%\and Olga Pons \at Department of Computer Science and Artificial Intelligence, Escuela T\'ecnica Superior de Ingenier\'ia Inform\'atica, Universidad de Granada,  C/Periodista Daniel Saucedo Aranda s/n, E-18071, Granada, Spain.
+%\email{opc@decsai.ugr.es}
+%\and Guy De Tr\'e \at Department of Telecommunications and Information Processing, Ghent University, St.-Pietersnieuwstraat 41, B-9000 Gent, Belgium. \email{Guy.DeTre@telin.ugent.be}
+%}
+
 \institute{
-Jos\'e Pons \at Department of Computer Science and Artificial Intelligence, Escuela T\'ecnica Superior de Ingenier\'ia Inform\'atica, Universidad de Granada,  C/Periodista Daniel Saucedo Aranda s/n, E-18071, Granada, Spain.
-\email{jpons@decsai.ugr.es}
-\and Christophe Billiet \at Department of Telecommunications and Information Processing, Ghent University, St.-Pietersnieuwstraat 41, B-9000 Gent, Belgium. \email{Christophe.Billiet@telin.ugent.be}
-\and Olga Pons \at Department of Computer Science and Artificial Intelligence, Escuela T\'ecnica Superior de Ingenier\'ia Inform\'atica, Universidad de Granada,  C/Periodista Daniel Saucedo Aranda s/n, E-18071, Granada, Spain.
-\email{opc@decsai.ugr.es}
-\and Guy De Tr\'e \at Department of Telecommunications and Information Processing, Ghent University, St.-Pietersnieuwstraat 41, B-9000 Gent, Belgium. \email{Guy.DeTre@telin.ugent.be}
+Jos\'e Pons, Olga Pons \at Department of Computer Science and Artificial Intelligence, Escuela T\'ecnica Superior de Ingenier\'ia Inform\'atica, Universidad de Granada,  C/Periodista Daniel Saucedo Aranda s/n, E-18071, Granada, Spain.
+\email{jpons, opc@decsai.ugr.es}
+\and
+Christophe Billiet, Guy De Tr\'e \at Department of Telecommunications and Information Processing, Ghent University, St.-Pietersnieuwstraat 41, B-9000 Gent, Belgium.
+\email{Christophe.Billiet, Guy.DeTre@telin.ugent.be}
 }
+
 %
@@
 
-\vspace{-35pt}
+\vspace{-70pt}
 %\abstract*{There exist some objects in the nature that have a time-variant (e.g.
@@
 \abstract*{
-In reality, some objects or concepts have properties with a time-variant (e.g.
-an employee's job function) or time-related (e.g.
-financial transactions of a bank account) nature.
+In reality, some objects or concepts have properties with a time-variant or time-related nature.
 Modelling these kinds of objects or concepts in a (relational) database schema is possible, but time-variant and time-related attributes have an impact on the consistency of the entire database.
-Therefore, some temporal database models have been proposed to deal with this.
-Time itself is a source of imprecision, vagueness and uncertainty, since existing time measuring devices are inherently imperfect.
+Therefore, temporal database models have been proposed to deal with this.
+Time itself can be at the source of imprecision, vagueness and uncertainty, since existing time measuring devices are inherently imperfect.
 Accordingly, human beings manage time using temporal indications and temporal notions, which may contain imprecision, vagueness and uncertainty.
 However, the imperfection in human-used temporal indications is to some degree constrained by socially well-known boundaries.
-Several proposals for dealing with such imperfections exist.
+Several proposals for dealing with such imperfections when modelling temporal aspects exist.
 Some of these proposals consider the basis of the system to be the conversion of the specificity of temporal notions between used temporal expressions.
 Other proposals consider the temporal indications in the used temporal expressions to be the source of imperfection.
-In this chapter, an overview is given, concerning the basic concepts and issues related to the modelling of time as such or in (relational) database models and the imperfections that arise during or as a result of this modelling.
+In this chapter, an overview is given, concerning the basic concepts and issues related to the modelling of time as such or in (relational) database models and the imperfections that may arise during or as a result of this modelling.
 Next to this, a novel and currently researched technique for handling some of these imperfections is presented.
@@
 \abstract{
-In reality, some objects or concepts have properties with a time-variant (e.g.
-an employee's job function) or time-related (e.g.
-financial transactions of a bank account) nature.
+In reality, some objects or concepts have properties with a time-variant or time-related nature.
 Modelling these kinds of objects or concepts in a (relational) database schema is possible, but time-variant and time-related attributes have an impact on the consistency of the entire database.
-Therefore, some temporal database models have been proposed to deal with this.
-Time itself is a source of imprecision, vagueness and uncertainty, since existing time measuring devices are inherently imperfect.
+Therefore, temporal database models have been proposed to deal with this.
+Time itself can be at the source of imprecision, vagueness and uncertainty, since existing time measuring devices are inherently imperfect.
 Accordingly, human beings manage time using temporal indications and temporal notions, which may contain imprecision, vagueness and uncertainty.
 However, the imperfection in human-used temporal indications is to some degree constrained by socially well-known boundaries.
-Several proposals for dealing with such imperfections exist.
+Several proposals for dealing with such imperfections when modelling temporal aspects exist.
 Some of these proposals consider the basis of the system to be the conversion of the specificity of temporal notions between used temporal expressions.
 Other proposals consider the temporal indications in the used temporal expressions to be the source of imperfection.
-In this chapter, an overview is given, concerning the basic concepts and issues related to the modelling of time as such or in (relational) database models and the imperfections that arise during or as a result of this modelling.
+In this chapter, an overview is given, concerning the basic concepts and issues related to the modelling of time as such or in (relational) database models and the imperfections that may arise during or as a result of this modelling.
 Next to this, a novel and currently researched technique for handling some of these imperfections is presented.
@@
 
-\section{Time Modelling}
+\section{Basic Concepts and Issues in Time Modelling}
 \label{sec:time-domain}
@@
 \def\JPicScale{1}
-\section{Temporal Databases}
+
+\section{Data Imperfections in Time Modelling}
+\label{sec:fuzzy-time-mod}
+\input{sources/fuzzyTimeDomain.tex}
+
+\section{Basic Concepts and Issues in Temporal Databases}
 \label{sec:temporal-databases}
@@
 \def\JPicScale{1}
-\section{Imprecision, Uncertainty and Vagueness in Temporal Databases}
+
+\section{Data Imperfections in Temporal Databases}
 \label{sec:imprecision-uncertainty}
@@
 
-\newpage
-\section*{Appendix: Introduction to fuzzy logic}
-\label{app:fuzzy-logic}
-\input{sources/fuzzylogic.tex}
+%\newpage
+%\section*{Appendix: Introduction to fuzzy logic}
+%\label{app:fuzzy-logic}
+%\input{sources/fuzzylogic.tex}
 

### 2012-03-01 trunk/book/sources/timeDomain.tex
The new form of the chapter...

@@
 Before considering the introduction of temporal modelling to information systems, it is necessary to define and explain some main concepts concerning temporal modelling and their corresponding terminology, to situate these and to discuss some properties and issues related to these concepts.
-In this section, several basic concepts and corresponding terminology will be defined, explained and situated.
+In this section, several basic concepts and their corresponding terminology will be defined, explained and situated.
 Most of these concepts are widely used in the community of temporal databases and their definitions have been agreed upon in the context of~\cite{Dyreson1994}.
@@
 
-In subsection \ref{subsec:basic-concepts}, some basic concepts of time modelling are presented and explained.
+%In subsection \ref{subsec:basic-concepts}, some basic concepts of time modelling are presented and explained.
 
@@
 
-\subsection{Basic concepts and properties}
+\subsection{Basic Concepts and Properties}
 \label{subsec:basic-concepts}
 In information systems, time itself is usually perceived as a linear or cyclic concept~\cite{Jensen94thetsql2}.
-Therefore, a time domains modelling time is usually represented by a set with an imposed partial order.
+Therefore, a time domain modelling time is usually represented by a set with an imposed partial order.
 In general, two main types of time models can be discerned: a \emph{linear} model~\cite{benthem82} and a \emph{cyclic} model~\cite{lorentzos88}.
@@
 
-Orthogonal to the classification of underlying time models as linear or cyclic, they can be classified as \emph{discrete}, \emph{dense} or \emph{continuous} models.
+Orthogonal to the classification of underlying time models as linear or cyclic, they can be classified as \emph{discrete}, \emph{dense} or \emph{continuous} models\cite{Dyreson1994}.
 In a discrete model~\cite{Clifford:1985:AHR:971699.318922}, the notion exists that every instant has a unique successor and the set of instants is seen as a discrete one.
 Here, intuitively, the set of instants can be seen as isomorphic to the set of natural numbers $\Nat$.
-In a dense model, the notion exists that between any two instants always lies another, while the set of instants can still be seen as a discrete one.
-Here, intuitively, the set of instants can be seen as isomporphic to the set of rational numbers $\Q$ or the set of real numbers $\R$.
-In a continuous model, the notion also exists that between any two instants always lies another one, but unlike dense models, the set of instants is seen as continuous and there are no ``gaps'' between successive instants.
+In a dense model, the notion exists that between any two instants always lies another.
+Here, intuitively, the set of instants can be seen as isomporphic to the set of rational numbers $\Q$ (when the set of instants is a discrete one) or the set of real numbers $\R$ (when the set of instants is a continuous one).
+In a continuous model, the notion also exists that between any two instants always lies another one, but the set of instants is always seen as continuous and there are no ``gaps'' between successive instants.
 
@@
 The fact that chronons are actually time intervals has a particular effect on the representation of instants and time intervals.
-In a data model representing a time domain using chronons, an instant is of course represented as a chronon, but one single chronon may represent multiple instants.
+In a data model representing a time domain using chronons, an instant is of course represented by a chronon.
 A time interval may be represented by a set of contiguous chronons, depending on the amount of time the time interval comprises.
@@
 
-\textcolor{red}{CHECK}
-Restrictions on time range may be performed on the base of the following two concepts~\cite{Jensen94thetsql2}:
-\begin{itemize}
-\item
-\textbf{Time boundedness}: Time can be bounded orthogonally in the past and in the future.
-\item
-\textbf{Time distance}: As a metric, time has a distance function which presents four properties:
-\begin{enumerate}
-\item
-Distance is non-negative.
-\item
-The distance between two different elements is not equal to zero.
-\item
-The distance from $\alpha$ to $\beta$ is the same as that between $\beta$ and $\alpha$.
-\item
-Triangle inequality: The distance from $\alpha$ to $\gamma$ is equal or greater than the distance from $\alpha$ to $\beta$ plus the distance from $\beta$ to $\gamma$.
-\end{enumerate}
-\end{itemize}
-\textcolor{red}{END CHECK}
+%\textcolor{red}{CHECK}
+Restrictions on time range may exist, as time may be bounded orthogonally in the past and in the future~\cite{Jensen94thetsql2}.
+%\begin{itemize}
+%\item
+%\textbf{Time boundedness}: Time can be bounded orthogonally in the past and in the future.
+%\item
+%\textbf{Time distance}: As a metric, time has a distance function which presents four properties:
+%\begin{enumerate}
+%\item
+%Distance is non-negative.
+%\item
+%The distance between two different elements is not equal to zero.
+%\item
+%The distance from $\alpha$ to $\beta$ is the same as that between $\beta$ and $\alpha$.
+%\item
+%Triangle inequality: The distance from $\alpha$ to $\gamma$ is equal or greater than the distance from $\alpha$ to $\beta$ plus the distance from $\beta$ to $\gamma$.
+%\end{enumerate}
+%\end{itemize}
+%\textcolor{red}{END CHECK}
 
@@
 
-Subsection \ref{subsec:granularity} gives a short survey on the concept of temporal granularities.
+%Subsection \ref{subsec:granularity} gives a short survey on the concept of temporal granularities.
 
@@
 As a granularity $G$ is an ordered set, each granule may be represented by an integer.
-In this representation, to keep track of the granularity a granule is an element of, the corresponding granularity is added in subscript:
+In this representation, to keep track of the granularity a granule is an element of, the corresponding granularity name is added in subscript:
 
@@
 
-In a system, the granularity with the shortest granules is the \emph{chronons granularity}, which is denoted as $\bot$. It is the granularity of which the granules are chronons. 
+In a system, the granularity with the shortest granules is the \emph{chronons granularity}, which is denoted as `$\bot$'. It is the granularity of which the granules are chronons. 
 %It is now possible to define a function $f$, called a \emph{mapping function}, that maps a given granule $i_G$, $i \in \Z$, from a given granularity $G$ to a set of corresponding chronons.
@@
 \begin{svgraybox}
+\vspace{-10pt}
 \begin{definition}
@@
 \textbf{Mapping function}~\cite{Lin97}\\ 
-A mapping function $f$ is a function that translates from a given granule $i_G$, $i \in \Z$, in a given granularity $G$, to a set of corresponding chronons.\\
+A mapping function $f$ is a function that maps a given granule $i_G$, $i \in \Z$, in a given granularity $G$, to a set of corresponding chronons:\\
 
@@
 \end{definition}
+\vspace{-10pt}
 \end{svgraybox}
 
-Note that a mapping function $f$ always maps from a given granularity $G$ to the set of chronons $\bot$. Therefore, the output for a mapping function is a subset of $\Pow(\bot)$. \\
+Note that a mapping function $f$ always maps from a given granularity $G$ to the powerset of the set of chronons $\bot$.
+Therefore, the output for a mapping function is an element of $\Pow(\bot)$ and thus a subset of $\bot$.
+\\
  A mapping function $f$ requires that the following properties hold~\cite{Lin97}:
@@
 
-Mapping functions between granularities and the chronons granularity also allow comparing granularities with respect to the length of their granules.
+The existence of mapping functions between granularities and the chronons granularity also allows comparing granularities with respect to the length of their granules.
 In this context, two important concepts can be discerned.
@@
 \label{def:casting-fucntion}
-A casting function $cast$ maps granules from a granularity $G$ to granules from a different granularity $H$.
-\begin{equation}
+Consider two different granularities $G$ and $H$. A granularity-to-granularity \emph{\textbf{casting function}} $cast$ is then a function mapping granules from $G$ to granules from $H$: 
+\begin{align}
 \label{eq:cast-function}
-\mbox{cast} \left( i_G,G,H \right) = j_H \nonumber
-\end{equation}
+%\mbox{cast} \left( i_G,G,H \right) = j_H \nonumber
+cast & : G \times \mathbb{G} \times \mathbb{G} \rightarrow H \nonumber\\
+		 & : (i_{G}, G, H) \rightarrow j_{H} \nonumber
+\end{align}
+where $i_{G} \in G$ and $j_{H} \in H$ and where $\mathbb{G}$ denotes the set of all granularities.
 \end{definition}
@@
 
-This notation indicates that the function $cast$ associates a granule $i_G$ in $G$ to a corresponding granule $j_H$ in $H$.
+Thus, the function $cast$ associates a granule $i_G$ in $G$ to a corresponding granule $j_H$ in $H$.
 Two kinds of granularity-to-granularity mappings can now be discerned: an \emph{upwards mapping} is a mapping from a granularity $G$ to a coarser granularity $H$, whereas a downwards mapping is a mapping from a granularity $K$ to a finner granularity $L$.
@@
 
- Mappings between two granularities may be classified as \emph{regular mappings}, \emph{irregular mappings} or \emph{congruent mappings}.
+Mappings between two granularities may be classified as \emph{regular mappings}, \emph{irregular mappings} or \emph{congruent mappings}~\cite{Lin97},~\cite{DyresonSnodgrass1994}.
 \begin{itemize}
 \item
-\emph{Regular mapping}: A regular mapping is a granularity-to-granularity mapping where the mapping function between the two granularities is calculated by means of multiplications and/or divisions and (maybe) an anchor adjustment.
+\emph{Regular mapping}: A regular mapping is a granularity-to-granularity mapping where the mapping function value is calculated by means of multiplications and/or divisions and (maybe) an anchor adjustment.
 E.g.
-the mapping between hours and minutes is calculated using a multiplication by $60$.
+the mapping value of the mapping between hours and minutes is calculated using a multiplication by $60$.
 \item
-\emph{Irregular mapping}: An irregular mapping is a granularity-to-granularity mapping where the mapping function can not be calculated by means of multiplications and/or divisions.
+\emph{Irregular mapping}: An irregular mapping is a granularity-to-granularity mapping where the mapping function value can not be calculated by means of multiplications and/or divisions.
 E.g.
-the mapping between months and days is dependent on the exact month or day.
+the mapping value of the mapping between months and days is dependent on the exact month or day.
 \item
-\emph{Congruent mapping}: A congruent mapping is a granularity-to-granularity mapping where the two granularities in the mapping have the same granules but a different anchor.
+\emph{Congruent mapping}: A congruent mapping is a granularity-to-granularity mapping where the two granularities involved in the mapping have the same granules but a different anchor.
 E.g.
-the mapping between the days (Gregorian calendar days) and the academic days.
+the mapping between the days (Gregorian calendar days) and the academic days is a congruent mapping.
 \end{itemize}
 
+The above is illustrated in the following example.
+
 \begin{example}
-Consider a systems that models both Gregorian Calendar dates as well as Academic dates.
-The chronons granularity in this system are milliseconds.
-Figure \ref{fig:granularity-graph-example} shows the complete granularity graph example.
+Consider a system that models both Gregorian Calendar dates as well as Academic dates.
+The chronons granularity in this system contains milliseconds.
+Figure \ref{fig:granularity-graph-example} shows the complete granularity graph corresponding with this example.
 The transitions between Chronons and Seconds and between Seconds and Minutes and so on is called a \emph{regular mapping}.

### 2012-03-05 trunk/book/sources/fuzzyTimeDomain.tex
update with the title, some new bibliography and the time domain.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Fuzzy time domain
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Humans usually deals with time in a imprecise way. e.g. `A few days ago we received this package'. This issue has been studied in the field of Artificial Intelligence ~\cite{Tre97} and language understanding~\cite{DeCaluwe:1997:FTI:285506.285516}. Some proposals ~\cite{knight1993},~\cite{Cru97},~\cite{nagypal2003} conclude that the best representation for incomplete temporal knowledge are time intervals instead of time points. This means that, as Allen proprosed in \cite{Allen83} the primitive units (the chronons) in the temporal system are intervals.

When modelling temporal knowledge, the following types of information might be found~\cite{nagypal2003}:

\begin{description}
\item[\emph{Uncertainty.}] The temporal specification of the event is uncertain. This is usually found in historical documents when some documents state contradictory facts about some event.

\item[\emph{Vagueness.}] Some of the temporal specifications are defined in a vague way. e.g. `during summer', `late night'. Note also that temporal specifications become vague when the granularity is changed ~\cite{Cru97},~\cite{nagypal2003} . e.g. `The book was written by July 2012' is usefull if we are interested in the month and the year but it turns into a incomplete expression when we try to obtain the exact day. It is possible to set the following classification~\cite{Dev98}:

\begin{itemize}
\item
Non-numerical indications. e.g. \emph{often, usually}.
\item
Approximative time indications. e.g. \emph{around 10 a.m.}
\item
Indications on temporal relations. e.g. \emph{after 12, during summer}.
\end{itemize}

\item[\emph{Subjectivity} or \emph{Ambiguity}.] The specification of a temporal fact may be affected by subjectivity.  It is possible to distinguish between:
\begin{itemize}
\item Historical periods~\cite{nagypal2003} like `late romanticism' or `early middle ages'. 
\item The interpretation of the temporal event~\cite{Dev98} depends on the side of the hearer. e.g. A speaker says `We will meet each other at six', the hearer has to choose between 6 p.m. or 6 a.m. The speaker, however knows exactly which one is meant. 
\end{itemize}
\end{description}

Another important task in the modelling of incomplete knowledge is when dealing with transitions among the different granularities.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% End fuzzy time domain
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2012-03-09 book/sources/timeDomain.tex
A temporary version of the proces of implementing Guy's remarks

@@
 
-\textcolor{red}{CHECK}
-Restrictions on time range may be performed on the base of the following two concepts~\cite{Jensen94thetsql2}:
-\begin{itemize}
-\item
-\textbf{Time boundedness}: Time can be bounded orthogonally in the past and in the future.
-\item
-\textbf{Time distance}: As a metric, time has a distance function which presents four properties:
-\begin{enumerate}
-\item
-Distance is non-negative.
-\item
-The distance between two different elements is not equal to zero.
-\item
-The distance from $\alpha$ to $\beta$ is the same as that between $\beta$ and $\alpha$.
-\item
-Triangle inequality: The distance from $\alpha$ to $\gamma$ is equal or greater than the distance from $\alpha$ to $\beta$ plus the distance from $\beta$ to $\gamma$.
-\end{enumerate}
-\end{itemize}
-\textcolor{red}{END CHECK}
+%\textcolor{red}{CHECK}
+Restrictions on time range may exist, as time may be bounded orthogonally in the past and in the future~\cite{Jensen94thetsql2}.
+%\begin{itemize}
+%\item
+%\textbf{Time boundedness}: Time can be bounded orthogonally in the past and in the future.
+%\item
+%\textbf{Time distance}: As a metric, time has a distance function which presents four properties:
+%\begin{enumerate}
+%\item
+%Distance is non-negative.
+%\item
+%The distance between two different elements is not equal to zero.
+%\item
+%The distance from $\alpha$ to $\beta$ is the same as that between $\beta$ and $\alpha$.
+%\item
+%Triangle inequality: The distance from $\alpha$ to $\gamma$ is equal or greater than the distance from $\alpha$ to $\beta$ plus the distance from $\beta$ to $\gamma$.
+%\end{enumerate}
+%\end{itemize}
+%\textcolor{red}{END CHECK}
 
-\textcolor{red}{MOVE TO TEMPORAL DATABASE PART}
-On the other hand, time may be~\cite{Jensen94thetsql2} \textbf{relative} or \textbf{absolute} (in other words, \textbf{anchored} or \textbf{unanchored}).
+%\textcolor{red}{MOVE TO TEMPORAL DATABASE PART}
+%On the other hand, time may be~\cite{Jensen94thetsql2} \textbf{relative} or \textbf{absolute} (in other words, \textbf{anchored} or \textbf{unanchored}).
 Sometimes, what one considers \emph{absolute time} is not as definite as one would hope since our concept of absolute time is based on another time reference (for example, January 1st of year one).
@@
 -3 days = three days ago).
-\textcolor{red}{END MOVE}
+%\textcolor{red}{END MOVE}
 
-Subsection \ref{subsec:granularity} gives a short survey on the concept of temporal granularities.
+%Subsection \ref{subsec:granularity} gives a short survey on the concept of temporal granularities.
 
@@
 \label{def:casting-fucntion}
-A casting function $cast$ maps granules from a granularity $G$ to granules from a different granularity $H$.
-\begin{equation}
+Consider two different granularities $G$ and $H$. A granularity-to-granularity \emph{\textbf{casting function}} $cast$ is then a function mapping granules from $G$ to granules from $H$: 
+\begin{align}
 \label{eq:cast-function}
-\mbox{cast} \left( i_G,G,H \right) = j_H \nonumber
-\end{equation}
+%\mbox{cast} \left( i_G,G,H \right) = j_H \nonumber
+cast & : G \times \mathbb{G} \times \mathbb{G} \rightarrow H \nonumber\\
+		 & : (i_{G}, G, H) \rightarrow j_{H} \nonumber
+\end{align}
+where $i_{G} \in G$ and $j_{H} \in H$ and where $\mathbb{G}$ denotes the set of all granularities.
 \end{definition}
+\vspace{-10pt}
 \end{svgraybox}
 
-This notation indicates that the function $cast$ associates a granule $i_G$ in $G$ to a corresponding granule $j_H$ in $H$.
+Thus, the function $cast$ associates a granule $i_G$ in $G$ to a corresponding granule $j_H$ in $H$.
 Two kinds of granularity-to-granularity mappings can now be discerned: an \emph{upwards mapping} is a mapping from a granularity $G$ to a coarser granularity $H$, whereas a downwards mapping is a mapping from a granularity $K$ to a finner granularity $L$.
-Different granularity-to-granularity mappings between several granularities can be represented using a granularity graph, which is a directed graph indicating the mapping conversions.
-
- Mappings between two granularities may be classified as \emph{regular mappings}, \emph{irregular mappings} or \emph{congruent mappings}.
+Orthogonal to this classification, mappings between two granularities may be classified as \emph{regular mappings}, \emph{irregular mappings} or \emph{congruent mappings}~\cite{Lin97},~\cite{DyresonSnodgrass1994}.
 \begin{itemize}
 \item
-\emph{Regular mapping}: A regular mapping is a granularity-to-granularity mapping where the mapping function between the two granularities is calculated by means of multiplications and/or divisions and (maybe) an anchor adjustment.
+\emph{Regular mapping}: A regular mapping is a granularity-to-granularity mapping where the mapping function value is calculated by means of multiplications and/or divisions and (maybe) an anchor adjustment.
 E.g.
-the mapping between hours and minutes is calculated using a multiplication by $60$.
+the mapping value of the mapping between hours and minutes is calculated using a multiplication by $60$.
 \item
-\emph{Irregular mapping}: An irregular mapping is a granularity-to-granularity mapping where the mapping function can not be calculated by means of multiplications and/or divisions.
+\emph{Irregular mapping}: An irregular mapping is a granularity-to-granularity mapping where the mapping function value can not be calculated by means of multiplications and/or divisions.
 E.g.
-the mapping between months and days is dependent on the exact month or day.
+the mapping value of the mapping between months and days is dependent on the exact month or day.
 \item
-\emph{Congruent mapping}: A congruent mapping is a granularity-to-granularity mapping where the two granularities in the mapping have the same granules but a different anchor.
+\emph{Congruent mapping}: A congruent mapping is a granularity-to-granularity mapping where the two granularities involved in the mapping have the same granules but a different anchor.
 E.g.
-the mapping between the days (Gregorian calendar days) and the academic days.
+the mapping between the days (Gregorian calendar days) and the academic days is a congruent mapping.
 \end{itemize}
 
+Different granularity-to-granularity mappings between several granularities can be represented using a granularity graph, which is a directed graph indicating the mapping conversions.
+The above is illustrated in the following example.
+
 \begin{example}
-Consider a systems that models both Gregorian Calendar dates as well as Academic dates.
-The chronons granularity in this system are milliseconds.
-Figure \ref{fig:granularity-graph-example} shows the complete granularity graph example.
-The transitions between Chronons and Seconds and between Seconds and Minutes and so on is called a \emph{regular mapping}.
-The transition between Days and Months is not defined by a multiplication or division, this is called a \emph{irregular mapping}.
-In the graph, these \emph{irregular mappings} are noted with a bold arrow.
-Finally, the transition between a Gregorian Calendar Day and an Academic Day is called a \emph{Congruent mapping}: two granularities with the same granules but with different anchors.
+Consider a system that models both Gregorian calendar dates as well as academic calendar dates.
+In this system, the chronons granularity is a set of milliseconds.
+Figure \ref{fig:granularity-graph-example} shows the complete granularity graph corresponding to this example.
+The transitions between the chronons granularity and the seconds granularity and between the seconds granularity and the minutes granularity are examples of regular mappings.
+They are represented by normal arrows in the visualisation of the graph.
+The transition between the days granularity and the months granularity is an example of an irregular mapping.
+In the graph visualisation, irregular mappings are represented by a bold arrow.
+Finally, the transition between the Gregorian calendar day granularity and the academic day granularity is an example of a congruent mapping.
 Both concern the same days, but the academic year starts on October 1st, whereas the Gregorian calendar year starts on January 1st.
-Congruent mappings are visualised as straight lines without arrow heads.
+In the graph visualisation, congruent mappings are visualised as straight lines without arrow heads.
 
@@
 
-Next, in subsection \ref{subsec:timedomain-calendar}, some basic concepts and issues concerning calendars are explained.
+%Next, in subsection \ref{subsec:timedomain-calendar}, some basic concepts and issues concerning calendars are explained.
 
-\subsection{Calendars}
-\label{subsec:timedomain-calendar}
+%\subsection{Calendars}
+%\label{subsec:timedomain-calendar}
 %maybe introduce previously the concept of time units, linear hierarchy and time point.
 
-This subsection will concern \emph{calendars}.
+%This subsection will concern \emph{calendars}.
 A calendar provides some sort of human interpretation of time.
@@
 
-\begin{svgraybox}
-\vspace{-10pt}
-\begin{definition}\textbf{Linear Hierarchy of Granularities}\\
-A \emph{\textbf{linear hierarchy of granularities}} is a finite set of granularities with a linear order (denoted $\sqsubset$).
+%\begin{svgraybox}
+%\vspace{-10pt}
+%\begin{definition}\textbf{Linear Hierarchy of Granularities}\\
+%A \emph{\textbf{linear hierarchy of granularities}} is a finite set of granularities with a linear order (denoted $\sqsubset$).
 The most coarse granularity is called \emph{the top of the hierarchy} and must be an infinite set while all other granularities must be finite sets.
-\end{definition}
-\vspace{-10pt}
-\end{svgraybox}
+%\end{definition}
+%\vspace{-10pt}
+%\end{svgraybox}
 
-An example of a linear hierarchy is the set $\left\{ \text{days, months, years} \right\}$, where days $\sqsubset$ months and months $\sqsubset$ years.
+%An example of a linear hierarchy is the set $\left\{ \text{days, months, years} \right\}$, where days $\sqsubset$ months and months $\sqsubset$ years.
 With this, a \emph{linear calendar} can be defined.
 
-\begin{svgraybox}
-\vspace{-10pt}
-\begin{definition}\textbf{Linear Calendar}\\
-A \emph{\textbf{linear calendar}} consist of a linear hierarchy of granularities and a validity predicate which specifies if a granule is valid in the calendar.
-\end{definition}
-\vspace{-10pt}
-\end{svgraybox}
+%\begin{svgraybox}
+%\vspace{-10pt}
+%\begin{definition}\textbf{Linear Calendar}\\
+%A \emph{\textbf{linear calendar}} consist of a linear hierarchy of granularities and a validity predicate which specifies if a granule is valid in the calendar.
+%\end{definition}
+%\vspace{-10pt}
+%\end{svgraybox}
 
-An example of a linear calendar could be the linear hierarchy $\left\{ \text{days, months, years} \right\}$, where days $\sqsubset$ months and months $\sqsubset$ years, together with a validity predicate that excludes time indications like `32 April 2012' or `30 February 2012'.
+%An example of a linear calendar could be the linear hierarchy $\left\{ \text{days, months, years} \right\}$, where days $\sqsubset$ months and months $\sqsubset$ years, together with a validity predicate that excludes time indications like `32 April 2012' or `30 February 2012'.
 
@@
 
-The next section briefly introduces temporal relationships.
+%The next section briefly introduces temporal relationships.
 

### 2012-03-09 trunk/book/sources/temporalDatabases.tex
A temporary version of the proces of implementing Guy's remarks

--- 2012-02-27 trunk/book/sources/temporalDatabases.tex
@@
 %
-A temporal database can generally be seen as a database that manages some temporal aspects in its schema \cite{etzion1998}.
+A temporal database can generally be seen as a database that manages some temporal aspects in its schema \cite{etzion1998}~\cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy}.
 In subsection \ref{subsec:td-general-concepts}, some main concepts and properties concerning temporal databases and their definitions are presented and explained.
@@
 \subsection{\label{subsec:td-general-concepts}Basic Concepts and Properties}
-A database models some part of reality.
-As mentioned in the introduction, the part of reality a temporal database tries to model, contains some temporal aspects.
+A database schema models some part of reality.
+As mentioned in the introduction, the part of reality a temporal database schema tries to model, contains some temporal aspects.
 For example, in this part of reality, some concepts or objects could have time-related or time-variant properties.
@@
 
-Thus, a temporal database will contain temporal values, i.e.
+Thus, a temporal database will contain \emph{temporal values}, i.e.
 values representing (indications of) time.
@@
 \emph{Valid times} are usually supplied by the user, whereas \emph{transaction times} are usually system-generated and -supplied \cite{Dyreson1994}.
-Temporal values of type UDT are not given any special interpretation and have thus no special query language support \cite{Dyreson1994}.
+Temporal values of type UDT are not given any extraordinary interpretation and have thus no extraordinary query language support \cite{Dyreson1994}.
 
@@
 A \emph{valid-time}, respectively \emph{transaction-time} \emph{relational database} is now defined as containing one or more valid-time, respectively transaction-time relations \cite{Dyreson1994}.
-Next to this, \emph{bitemporal} relational databases contain both valid time and transaction time and tritemporal databases contain valid time, transaction time and decision time.
+Next to this, \emph{bitemporal} relational databases contain both valid time and transaction time~\cite{Dyreson1994} and tritemporal databases contain valid time, transaction time and decision time~\cite{Nascimento95decisiontime}.
 
@@
 
-In the following sections, some of the main points of attention concerning relational temporal databases are explained and discussed.
+%In the following sections, some of the main points of attention concerning relational temporal databases are explained and discussed.
 
-\subsection{\label{subsubsec:primary-key}Primary Keys in Temporal Relations}
-When designing a relation, a subset of the relation's attribute set is ususally chosen as primary key.
-The values of a record for these attributes will then uniquely determine this record and no two records may have the same values for every attribute in this primary key.
-In temporal relations, different records can exist with the same values for every attribute except the attributes describing the temporal aspects of the objects or concepts represented by the records (the temporal attributes).
-These different records represent different temporal versions of the same real object or concept.
-To allow the existence of such records, the most common solution is to extend the original primary key with the temporal attributes.
+In the rest of the chapter, the focus will be on concepts and issues concerning valid-time relations and aspects of valid-time relations.
+For this reason, the next two sections will present and discuss some main issues concerning temporal databases, specifically applied to or presented in the context of valid-time relations.
+
+\subsection{\label{subsubsec:primary-key}Primary Keys in Valid-time Relation Design}
+Generally, when designing a relation based on a relational database model, a subset of the relation's attribute set is ususally chosen as primary key.
+The values of a tuple for these attributes will then uniquely determine this tuple and no two distinct tuples may have the exact same values for every attribute in this primary key.
+Next to attributes unrelated to time, a valid-time relation schema will typically contain one or more attributes which model the valid-time aspects and behavior of the real objects and concepts modelled by the relation schema.
+In this work, these attributes are called \emph{valid-time attributes}.
+In valid-time relation extentions, distinct tuples can exist containing the exact same values for every attribute except the valid-time attributes.
+These distinct tuples represent distinct versions of the same real object or concept, valid during different time periods.
+To allow the existence of such tuples when designing a valid-time relation using a relational database model, the most common solution is to include the valid-time attributes in the primary key.
 %One of the first tasks in the design of a database is the choice for the primary key.
@@
 
-The following example illustrates this primary key issue in a valid-time relation.
+The following example illustrates this primary key issue.
 
@@
 \label{ex:pk}
-Consider the example valid-time relation visualized in table \ref{table:example-database}, which contains data representing employees in a company.
-The attributes `Name', `Age' and `Supervisor' describe the name, age and supervisor of an employee, respectively.
-The attributes `Start' and `End' describe the year when an employee started, respectively finished working for the company.
-For example, employee Sarah started working for the company in 2005 and finished in 2009.
-The attribute `ID' uniquely identifies an employee.
-Consider `ID' being the primary key and consider the company wanting to hire Sarah again.
-Another tuple should thus be inserted into the relation, containing value 4 as `ID'.
-This will of course not be allowed by the primary key.
-This problem is solved by defining a new primary key: `ID', `Start', `End'.
-Now, the resulting relation allows inserting a new tuple representing that Sarah was rehired in 2010, as shown in table \ref{table:example-database-with-new-pk}.
+Consider the example valid-time relation visualized in table \ref{table:example-database}, which models when certain people worked as employees in a certain company and under whose supervision they worked during that time.
+The valid-time attributes `Start' and `End' describe the year when an employee started, respectively finished working for the company.
+For example, the last tuple visualized in table \ref{table:example-database} represents that the employee represented by this tuple started working for the company in 2005 and finished in 2009.
+The attributes `Name', `Birthday' and `Supervisor' describe respectively the name, birthday date and unique identifier of the supervisor of an employee in the time during which he or she worked for the company.
+When correct, the date of an employee's birthday never changes and as such, the modelling of birthday dates has no effect on the database consistency.
+The `Birthday' attribute thus describes UDT values.
+The attribute `ID' describes employee identifiers.
+For each tuple, this identifier (a number) uniquely identifies the employee represented by the tuple.
+
+Now consider $\{$ID$\}$ being the primary key and consider the company wanting to hire Sarah again in 2010.
+This would be represented by another tuple in the relation, containing value 4 for attribute `ID'.
+The existence of such a tuple is of course not allowed by the primary key, because it would mean the existence of two distinct tuples containing value 4 for attribute `ID'.
+This problem can now be solved by defining a new primary key: $\{$ID, Start, End$\}$, which allows for the existence of distinct tuples with value 4 for attribute `ID', as long as they have different values for attributes `Start' or `End'.
+The resulting relation is shown in table~\ref{table:example-database-with-new-pk}.
 %Consider the example database visualized in table \ref{table:example-database}.
@@
 \end{example}
+%\begin{note}
+%The attribute `Age' in the presented relations visualized in tables~\ref{table:example-database},~\ref{table:example-database-with-new-pk} and~\ref{table:example-database-update} describes the age of an employee and is included as an example of .
+In reality, the birthday of the employee should be stored in the database, rather than the age, because the age
+%In the presented relations visualized in tables~\ref{table:example-database},~\ref{table:example-database-with-new-pk} and~\ref{table:example-database-update}, the value of a tuple for attribute `Supervisor' should actually contain the unique identifier of the supervisor of the employee represented by the tuple.
+the relation `Supervisor' stores the ID of the employee.
+To improve the readability of the example, the tables show the name of the employee.
+Note also that the field for the age is expressed as the age in years whereas the value stored in the database is the birthday.
+%\end{note}
 
-Extending a primary key to include temporal attributes allows for the insertion of spurious values, for example representing temporal inconsistencies.
+%Extending a primary key to include temporal attributes allows for the insertion of spurious values, for example representing temporal inconsistencies.
 This is shown in table \ref{table:example-database-with-new-pk}, where it seems Sarah was rehired in 2007 and fired in 2008.
@@
 \centering
-\caption{Example relation modelling the employees of a company.}
+\caption{Example relation modelling the employees of a company. Values for the `Birthday' attribute are visualized here in `dd/mm/yyyy' format.}
 \begin{tabular}{c c c c c c }
 \hline
-\textbf{ID} & \textbf{Name} & \textbf{Age} & \textbf{Supervisor} & \textbf{Start} & \textbf{End} \\ \hline
-1 & Peter & 24 & John &  2010 & - \\
-2 & Maria & 34 & John & 2001 & - \\
-3 & John & 52 & - &  1999 & - \\
-4 & Sarah & 29 & Maria &  2005 & 2009 \\
+\textbf{ID} & \textbf{Name} & \textbf{Birthday} & \textbf{Supervisor} & \textbf{Start} & \textbf{End} \\ \hline
+1 & Peter & 24/10/1985 & 3 &  2010 & - \\
+2 & Maria & 03/04/1984 & 3 & 2001 & - \\
+3 & John & 21/02/1964 & - &  1999 & - \\
+4 & Sarah & 29/11/1985 & 2 &  2005 & 2009 \\
 \hline 
@@
 \centering
-\caption{Example relation from example \ref{ex:pk} with new primary key and some spurious values.}
+\caption{Example relation after including the valid-time attributes in the primary key and adding a tuple.}
 \begin{tabular}{c c c c c c }
 \hline
-\textbf{ID} & \textbf{Name} & \textbf{Age} & \textbf{Supervisor} & \textbf{Start year} & \textbf{End year} \\ \hline
-1 & Peter & 24 & John &  2010 & - \\
-2 & Maria & 34 & John & 2001 & - \\
-3 & John & 52 & - &  1999 & - \\
-4 & Sarah & 29 & Maria &  2005 & 2009 \\
-4 & Sarah & 29 & Maria &  2010 & - \\
-4 & Sarah & 29 & Maria &  2007 & 2008 \\
+\textbf{ID} & \textbf{Name} & \textbf{Birthday} & \textbf{Supervisor} & \textbf{Start} & \textbf{End} \\ \hline
+1 & Peter & 24/10/1985 & 3 &  2010 & - \\
+2 & Maria & 03/04/1984 & 3 & 2001 & - \\
+3 & John & 21/02/1964 & - &  1999 & - \\
+4 & Sarah & 29/11/1985 & 2 &  2005 & 2009 \\
+4 & Sarah & 29/11/1985 & 2 &  2010 & - \\
+%4 & Sarah & 29/11/1985 & 2 &  2007 & 2008 \\
 \hline 
@@
 \label{table:example-database-with-new-pk}
-
-%\vspace{10pt}
-
 \end{table}
 
-%\vspace{-25pt}
-This latter issue is the concern of the next subsection.
+%The next subsection is concerned with another interresting issue.
 
-\subsection{\label{subsubsec:consistency}Consistency in Temporal Relations}
-As stated in subsection \ref{subsubsec:primary-key}, one of the main issues concerns the insertion of spurious values in a relation.
+\subsection{\label{subsubsec:consistency}Consistency in Valid-time Relation Querying}
+TODO\\
+TODO\\
+TODO\\
+
+The solution presented in subsection~\ref{subsubsec:primary-key} concerns relation design and consists of including the valid-time attributes in the primary key.
+Unfortunately, implementing this solution as such allows for the existence of records whose values imply inconsistencies with respect to the modelling of reality.
+
+Consider a valid-time relation of which the primary key can be partitioned into two sets of attributes.
+One set contains attributes totally unrelated to time, for which the values of a record allow to uniquely identify the object or concept represented by the record.
+The other set contains the valid-time attribute(s).
+Because the valid-time attribute(s) is(are) included in the primary key, the existence of distinct records with exactly the same values for all time-unrelated attributes and distinct values for at least one valid-time attribute is not prohibited.
+Thus, inserting such records into the relation is not prohibited either, even if the information represented by the values for the valid-time attributes shows clear inconsistencies.
+An example.
+
+\begin{example}
+\label{ex:prob}
+Consider the example valid-time relation visualized in table \ref{table:erconsistency}, which is based on the relation visualized in table~\ref{table:example-database}.
+The primary key is again $\{$ID, Start, End$\}$.
+The last record in the relation represents a person named `Sarah' started working for the company in 2007 and finished in 2008, with supervisor `John'.
+However, the fourth record represents the same person (the value for attribute `ID' is the same) started working for the company in 2005 and finished in 2009, with supervisor `Maria'.
+The intention is clear: Sarah worked in the company from 2005 to 2009, first for Maria, then for John, then again for Maria.
+It is of course possible for an employee to change supervisors, but it is of course impossible for a person to start working in the same company twice at different times, for different supervisors, without stopping to work for one in between, as it is impossible to stop working for a supervisor twice at different times, without working for another one in between.
+The valid-time information represented by the last record is clearly not consistent with the valid-time information represented by the fourth record, or vice versa.
+\end{example}
+
+\begin{table}
+\centering
+\caption{Example relation with records whose values for the valid-time attributes violate consistency.}
+\begin{tabular}{c c c c c c }
+\hline
+\textbf{ID} & \textbf{Name} & \textbf{Birthday} & \textbf{Supervisor} & \textbf{Start} & \textbf{End} \\ \hline
+1 & Peter & 24/10/1985 & 3 &  2010 & - \\
+2 & Maria & 03/04/1984 & 3 & 2001 & - \\
+3 & John & 21/02/1964 & - &  1999 & - \\
+4 & Sarah & 29/11/1985 & 2 &  2005 & 2009 \\
+%4 & Sarah & 29/11/1985 & 2 &  2010 & - \\
+4 & Sarah & 29/11/1985 & 3 &  2007 & 2008 \\
+\hline 
+\end{tabular}
+\label{table:erconsistency}
+\end{table}
+
+The most usual approach to deal with this inconsistency problem is to adapt the DML used by the DBMS, as to enforce consistency towards time with respect to the modelled reality.
+
+\begin{example}
+Consider the problem presented in example~\ref{ex:prob}.
+The inconsistency arises when the last record in table~\ref{table:erconsistency} is inserted.
+Because the record's values for the valid-time attributes differ from those of the fourth record, the last record is accepted.
+The DML statement used was (the table is called `Employees'):
+
+\begin{verbatim}
+INSERT INTO Employees VALUES
+ (4, `Sarah', `29/11/1985', 3, 2007, 2008);
+\end{verbatim}
+
+The inconsistency problem can now be solved by replacing this statement with the next:
+
+\begin{verbatim}
+UPDATE Employees SET `End' = `2007' WHERE
+ (ID = 4) AND (Start = 2005) AND (End = 2009);
+INSERT INTO Employees VALUES
+ (4, `Sarah', `29/11/1985', 3, 2007, 2008);
+INSERT INTO Employees VALUES
+ (4, `Sarah', `29/11/1985', 2, 2008, 2009);
+\end{verbatim}
+
+The resulting relation is visualized in table~\ref{table:example-database-update}.
+
+\end{example}
+
+%As stated in subsection \ref{subsubsec:primary-key}, one of the main issues concerns the insertion of spurious values in a relation.
+Another issue concerns the possible breach of consistency when updating records.
 The most usual approach to deal with this is to change the database DML to enforce the consistency with which the temporal relation models the temporal properties of the part of reality the relation models.
@@
 
-\begin{example}
-\textcolor{red}{PROBLEM}
-Consider again the example relation visualized in table \ref{table:example-database}. Imagine John left the company, but nobody updated the relation, so it is not known when
+%\begin{example}
+%Consider the database in table \ref{table:example-database}. The employee with ID=3 (John) works now for the employee with ID= 4 (Sarah). The update is made in the following sentence:
 
-The following \emph{update} sentence:
+%The following \emph{update} sentence:
 
-\begin{verbatim}
-Update Employees set StartYear = 2010 where ID=3;
-\end{verbatim}
+%\begin{verbatim}
+%Update Employees set 'Works for' = 4 where ID=3;
+%\end{verbatim}
 
-Is translated into the following two sentences: an update for the last version of the row and an insert sentence for the new version. The update sentence showed above is translated into:
+%Is translated into the following two sentences: an update for the last version of the row and an insert sentence for the new version.
+Consider that the time in the  system is the year 2010.
+Then, the update sentence showed above is translated into:
 
-\begin{verbatim}
-Update Employees set EndYear = 2010 where ID=3;
-Insert into Employees values (3,John,52,-,2010,-);
-\end{verbatim}
-
-%\vspace{-10pt}
+%\begin{verbatim}
+%Update Employees set EndYear = 2010 where ID=3;
+%Insert into Employees values (3,John,52,Sarah,2010,-);
+%\end{verbatim}
 
@@
 \centering
-\caption{Example relation from example \ref{ex:pk} updated maintaining consistency.}
+\caption{Example relation updated maintaining consistency.}
 \begin{tabular}{c c c c c c }
 \hline
-\textbf{ID} & \textbf{Name} & \textbf{Age} & \textbf{Works for} & \textbf{Start year} & \textbf{End year} \\ \hline
-1 & Peter & 24 & John &  2010 & - \\
-2 & Maria & 34 & John & 2001 & - \\
-3 & John & 52 & - &  1999 & 2010 \\
-3 & John & 52 & - &  2010 & - \\
-4 & Sarah & 29 & Maria &  2005 & 2009 \\
+\textbf{ID} & \textbf{Name} & \textbf{Birthday} & \textbf{Supervisor} & \textbf{Start} & \textbf{End} \\ \hline
+1 & Peter & 24/10/1985 & 3 &  2010 & - \\
+2 & Maria & 03/04/1984 & - & 2001 & - \\
+3 & John & 21/02/1964 & - &  1999 & 2010 \\
+3 & John & 21/02/1964 & - &  2010 & - \\
+4 & Sarah & 29/11/1985 & 2 &  2005 & 2007 \\
+4 & Sarah & 29/11/1985 & 3 &  2007 & 2008 \\
+4 & Sarah & 29/11/1985 & 2 &  2008 & 2009 \\
 \hline 
@@
 
-Table \ref{table:example-database-update} shows the resulting database after the update sentence.
+%Table \ref{table:example-database-update} shows the resulting database after the update sentence.
 
-\end{example}
+%\end{example}
 
 \subsection{\label{Comm-temp}Commercial Temporal Database Systems}
-There are several commercial temporal DBMS. Table \ref{table:commercial-temporal-db} describes the most well-knownand provides references for more information.
+Several commercial temporal DBMS exist. Table \ref{table:commercial-temporal-db} presents some information on some of the more well-known temporal DBMS and provides references for more information.
 
 Oracle workspace manager~\cite{oracle2009} and TimeDB~\cite{timedb2005} are libraries for dealing with time in OracleDB.
-Nevertheless, TimeDB and Postgree Temporal~\cite{posgree2009} are similar: both are simple implementations that implement a subset of the Allen operators and some operations for the creation and manipulation of temporal attributes (valid time, transaction time or both times are supported).
+On another note, TimeDB and Postgree Temporal~\cite{posgree2009} are similar: both are simple implementations that implement a subset of the Allen operators and some operations for the creation and manipulation of temporal attributes (valid time, transaction time or both times are supported).
 Teradata~\cite{teradata2011} is mainly a business intelligence system designed for data mining.
@@
 
-Unfortunately, none of these systems take imperfect information into account, neither in data storage nor in querying.
+Unfortunately, none of these systems take data imperfections into account, neither in data storage nor in querying.
 

### 2012-03-13 trunk/book/sources/fuzzyTemporalDatabases.tex
Almost done! In this one, the references are a little improved and the piece about fuzzy temporal reasoning (without databases) is written out.

@@
 Consider a logistics company which transports packages.
-The time when a package would arrive at its destination may be estimated and can eventually be researched, but it is typically not always precisely known.
+At the moment a package leaves, the time when it will arrive at its destination may be estimated, but will typically not be known precisely.
+For such companies and in many other situations, information systems able to handle imperfection with respect to certain temporal aspects of the objects modelled by the system are necessary.
+
+\subsection{Handling Data Imperfections in Temporal Databases}
+
+%The time when a package would arrive at its destination may be estimated and can eventually be researched, but it is typically not always precisely known.
 For such companies and in many other situations, information systems able to handle imperfection corresponding to certain temporal aspects of the objects modelled by the system are necessary.
@@
 
-Data imperfection is usually the focus of research in fuzzy databases.
+Data imperfections and techniques to represent them correctly in databases and queries are usually the focus of research in fuzzy databases.
 In fuzzy databases~\cite{Galindo2006}, uncertainty is usually expected to appear at storage level whereas imprecision is usually expected to appear in querying.
@@
 Finally, in subsection \ref{subsec:querying-time-intervals} the proposals for querying in a flexible way valid-time intervals are analyzed.
+\subsection{Handling Uncertainty in a Valid-time Relation}
 
-\subsection{Ill-known Time Intervals}
+\subsubsection{Ill-known Time Intervals}
 \label{subsec:representation-time-intervals}
@@
 
-\subsection{Querying Fuzzy Valid-Time Databases with Ill-Known Valid-Time Intervals}
+\subsubsection{Querying Fuzzy Valid-Time Databases with Ill-Known Valid-Time Intervals}
 \label{subsec:querying-time-intervals}
@@
 
-\subsubsection{Query Structure}
+\subsubsection{Query Structure} TO BE ANOTHER PART
 In this work, a query $\tilde{Q}$ is made up of two separate constructs of user query demands.
@@
 
-\subsubsection{Query evaluation}
+\subsubsection{Query evaluation} TO BE ANOTHER PART
 \label{subsubsec:query-evaluation}
@@
 
-\subsubsection{Aggregation and Ranking}
+\subsubsection{Aggregation and Ranking} TO BE ANOTHER PART
 To be able to present the most appropriate results to the user, for every record $r$, an aggregation method is used to aggregate $\Pos_{Q^{time}}(r)$ and $\Nec_{Q^{time}}(r)$ into a temporal record rank $e_{Q^{time}}(r)$ and after this, a convex combination combining $e_{Q^{time}}(r)$ and $e_{Q}(r)$ will provide the final record rank $e_{final}(r)$.
@@
 
-\subsection{Bipolarity in temporal databases}
+\subsection{Bipolarity in Temporal Databases}
 \label{subsubsec:bipolarity}

### 2012-03-13 book/author.tex
Almost done! In this one, the references are a little improved and the piece about fuzzy temporal reasoning (without databases) is written out.

@@
 
-\title*{An Overview on Temporal Databases.\\
- New Approaches and Proposals.}
+\title*{Aspects of Dealing with Imperfect Data in Temporal Databases}
 % Use \titlerunning{Short Title} for an abbreviated version of
@@
 \abstract*{
-In reality, some objects or concepts have properties with a time-variant (e.g.
-an employee's job function) or time-related (e.g.
-financial transactions of a bank account) nature.
+In reality, some objects or concepts have properties with a time-variant or time-related nature.
 Modelling these kinds of objects or concepts in a (relational) database schema is possible, but time-variant and time-related attributes have an impact on the consistency of the entire database.
-Therefore, some temporal database models have been proposed to deal with this.
-Time itself is a source of imprecision, vagueness and uncertainty, since existing time measuring devices are inherently imperfect.
+Therefore, temporal database models have been proposed to deal with this.
+Time itself can be at the source of imprecision, vagueness and uncertainty, since existing time measuring devices are inherently imperfect.
 Accordingly, human beings manage time using temporal indications and temporal notions, which may contain imprecision, vagueness and uncertainty.
-However, the imperfection in human-used temporal indications is to some degree constrained by socially well-known boundaries.
-Several proposals for dealing with such imperfections exist.
+However, the imperfection in human-used temporal indications is supported by human interpretation, whereas information systems need extraordinary support for this.
+Several proposals for dealing with such imperfections when modelling temporal aspects exist.
 Some of these proposals consider the basis of the system to be the conversion of the specificity of temporal notions between used temporal expressions.
 Other proposals consider the temporal indications in the used temporal expressions to be the source of imperfection.
-In this chapter, an overview is given, concerning the basic concepts and issues related to the modelling of time as such or in (relational) database models and the imperfections that arise during or as a result of this modelling.
+In this chapter, an overview is given, concerning the basic concepts and issues related to the modelling of time as such or in (relational) database models and the imperfections that may arise during or as a result of this modelling.
 Next to this, a novel and currently researched technique for handling some of these imperfections is presented.
@@
 \abstract{
-In reality, some objects or concepts have properties with a time-variant (e.g.
-an employee's job function) or time-related (e.g.
-financial transactions of a bank account) nature.
+In reality, some objects or concepts have properties with a time-variant or time-related nature.
 Modelling these kinds of objects or concepts in a (relational) database schema is possible, but time-variant and time-related attributes have an impact on the consistency of the entire database.
-Therefore, some temporal database models have been proposed to deal with this.
-Time itself is a source of imprecision, vagueness and uncertainty, since existing time measuring devices are inherently imperfect.
+Therefore, temporal database models have been proposed to deal with this.
+Time itself can be at the source of imprecision, vagueness and uncertainty, since existing time measuring devices are inherently imperfect.
 Accordingly, human beings manage time using temporal indications and temporal notions, which may contain imprecision, vagueness and uncertainty.
-However, the imperfection in human-used temporal indications is to some degree constrained by socially well-known boundaries.
-Several proposals for dealing with such imperfections exist.
+However, the imperfection in human-used temporal indications is supported by human interpretation, whereas information systems need extraordinary support for this.
+Several proposals for dealing with such imperfections when modelling temporal aspects exist.
 Some of these proposals consider the basis of the system to be the conversion of the specificity of temporal notions between used temporal expressions.
 Other proposals consider the temporal indications in the used temporal expressions to be the source of imperfection.
-In this chapter, an overview is given, concerning the basic concepts and issues related to the modelling of time as such or in (relational) database models and the imperfections that arise during or as a result of this modelling.
+In this chapter, an overview is given, concerning the basic concepts and issues related to the modelling of time as such or in (relational) database models and the imperfections that may arise during or as a result of this modelling.
 Next to this, a novel and currently researched technique for handling some of these imperfections is presented.
@@
 \def\JPicScale{1}
-\section{Imperfection in Temporal Databases}
+
+\section{Data Imperfections in Temporal Databases}
 \label{sec:imprecision-uncertainty}
@@
 \def\JPicScale{1}
-\section{Conclusions and Further Research Work}
+\section{Conclusions and Further Research}
 \label{sec:further-research-work}

### 2012-03-13 book/sources/fuzzyTimeDomain.tex
Almost done! In this one, the references are a little improved and the piece about fuzzy temporal reasoning (without databases) is written out.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Fuzzy time domain
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%INTRO
As explained in the introduction, humans handle temporal information using temporal notions like time intervals or time points \cite{Dyreson1994}. While the used temporal notions may contain imperfections~\cite{Dev98},~\cite{Dubois:jucs_9_9:fuzziness_and_uncertainty_in},~\cite{nagypal2003},~\cite{Dubois89}, humans often gracefully deal with these, as their inherent interpretation capability accounts for a lot of them. This phenomenon has been studied a.o. in the field of artificial intelligence~\cite{Tre97},~\cite{5151} and language understanding~\cite{DeCaluwe:1997:FTI:285506.285516},~\cite{nagypal2003},~\cite{Dev98}. An information system, however, cannot appeal to a similar interpretation functionality. Thus, many proposals have been concerned with the combination of time and imperfections in the context of information systems~\cite{nagypal2003}. In this section, some main concepts and issues concerning this combination are presented.

\subsection{Types of Imperfections in Temporal Modelling}
Generally, in temporal modelling, a distinction is made between the following types of imperfections~\cite{nagypal2003}.

\begin{itemize}
	\item \emph{Uncertainty}: Temporal information or data may contain uncertainty. This means that the exact temporal value is (partially) unknown, however, generally some knowledge is present anyhow, possibly describing the value~\cite{Dubois:jucs_9_9:fuzziness_and_uncertainty_in},~\cite{nagypal2003},~\cite{Dubois89},~\cite{343607}. E.g., the temporal notion described in a sentence like `The Belfry of Bruges was finished on a day somewhere between 01/01/1201 A.D. and 31/12/1300 A.D.' contains uncertainty: it is known that the belfry of Bruges was finished on a single day and that this day lays somewhere between 01/01/1201 A.D. and 31/12/1300 A.D., but it is not known exactly which day it is.
	\item \emph{Vagueness}: Temporal information or data might contain inherent vagueness, as a precise instant or time interval may be intended, but the description of it is certainly vague~\cite{schockaert08},~\cite{nagypal2003},~\cite{Dev98}. E.g., the temporal notion described in a sentence like `It happened during summer.' contains vagueness, as even the boundaries of the mentioned temporal notion are not clearly expressed.
	\item \emph{Subjectivity or ambiguity}: Temporal notions might be subject to subjectivity or ambiguity. In certain cases, the temporal notion concerns a historical period like `late romanticism' or `the early middle ages' and thus contains subjectivity~\cite{nagypal2003}. In other cases, the interpretation of the temporal notion depends on extra factors. E.g., consider a person saying to another person `Let's meet each other at six.' The person hearing these words doesn't now if 6 a.m. or 6 p.m. is intended, though the person saying the words does.
\end{itemize}

As to the sources of the imperfections in temporal information, most proposals consider no specific source~\cite{5151},~\cite{schockaert08},~\cite{nagypal2003},~\cite{Dubois89},~\cite{343607},~\cite{ohlbach2004},~\cite{Virant199639}. Some proposals, however, deal with the imperfections specifically resulting from aspects of language~\cite{Dev98} and other proposals consider transitions between different granularities to be the source of imperfection in temporal information~\cite{Lin97}. Therefore, some proposals consider granularity as the base of the temporal model~\cite{Cru97}.

\subsection{Imperfections in Temporal Relationships}
As the existence of temporal relationships allows to compare temporal notions, many approaches have been concerned with finding similar temporal relationships, able to support imperfections in the temporal information which is described by temporal notions or even by the temporal relationships themselves~\cite{ohlbach2004},~\cite{nagypal2003},~\cite{schockaert08},~\cite{Dubois:jucs_9_9:fuzziness_and_uncertainty_in}. These approaches are often based on Allen's operators~\cite{Allen83}.

%%%TODO

%Humans usually deals with time in a imprecise way. e.g. `A few days ago we received this package'. This issue has been studied in the field of Artificial Intelligence ~\cite{Tre97} and language understanding~\cite{DeCaluwe:1997:FTI:285506.285516}. Some proposals ~\cite{knight1993},~\cite{Cru97},~\cite{nagypal2003} conclude that the best representation for incomplete temporal knowledge are time intervals instead of time points. This means that, as Allen proprosed in \cite{Allen83} the primitive units (the chronons) in the temporal system are intervals.

%When modelling temporal knowledge, the following types of information might be found~\cite{nagypal2003}:

%\begin{description}
%\item[\emph{Uncertainty.}] The temporal specification of the event is uncertain. This is usually found in historical documents when some documents state contradictory facts about some event.

%\item[\emph{Vagueness.}] Some of the temporal specifications are defined in a vague way. e.g. `during summer', `late night'. Note also that temporal specifications become vague when the granularity is changed ~\cite{Cru97},~\cite{nagypal2003} . e.g. `The book was written by July 2012' is usefull if we are interested in the month and the year but it turns into a incomplete expression when we try to obtain the exact day. It is possible to set the following classification~\cite{Dev98}:

%\begin{itemize}
%\item
%Non-numerical indications. e.g. \emph{often, usually}.
%\item
%Approximative time indications. e.g. \emph{around 10 a.m.}
%\item
%Indications on temporal relations. e.g. \emph{after 12, during summer}.
%\end{itemize}

%\item[\emph{Subjectivity} or \emph{Ambiguity}.] The specification of a temporal fact may be affected by subjectivity.  It is possible to distinguish between:
%\begin{itemize}
%\item Historical periods~\cite{nagypal2003} like `late romanticism' or `early middle ages'. 
%\item The interpretation of the temporal event~\cite{Dev98} depends on the side of the hearer. e.g. A speaker says `We will meet each other at six', the hearer has to choose between 6 p.m. or 6 a.m. The speaker, however knows exactly which one is meant. 
%\end{itemize}
%\end{description}

%Another important task in the modelling of incomplete knowledge is when dealing with transitions among the different granularities.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% End fuzzy time domain
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2012-03-13 book/sources/intro.tex
Almost done! In this one, the references are a little improved and the piece about fuzzy temporal reasoning (without databases) is written out.

@@
 A database contains data representing real objects or concepts.
-Every one of these data is a result value of a measurement of a property or a description of a property of a real object or concept.
+Each (atomic) part of these data is a result value of a measurement of a property or a description of a property of a real object or concept.
 In reality, some aspects or properties of objects or concepts are time-variant or time-related.
-E.g.
-the moment of a bank transaction is traditionally a moment in time and thus a time-related notion, the function of an employee in a company can change through recorded history and is thus time-variant.
+E.g., the moment of a bank transaction is traditionally a moment in time and thus a time-related notion, the function of an employee in a company can change through recorded history and is thus time-variant.
 A temporal database schema is a database schema that models real objects or concepts with time-related or time-variant properties.
@@
 Notably, Allen~\cite{Allen83} studied temporal relationships between time intervals (and as a special case time points).
-Among others, the querying of temporal databases has greatly profited from these temporal relationships, because they allowed for richer and complexer user-specified temporal query demands, by allowing to express more complex relationships between the temporal notions in the temporal expressions in the query and the temporal indications in the database.
+Among others, the querying of temporal databases has greatly profited from these temporal relationships, because they allowed for richer and more complex user-specified temporal query demands, by allowing to express more complex relationships between the temporal notions in the temporal expressions in the query and the temporal indications in the database.
+E.g., in a relation modelling who was department head of an institution during which periods of time, a query like `who were the department heads when Thomas worked for the institution' can be evaluated using similar relationships.
 
-Humans handle temporal information using certain temporal notions like time intervals or time points~\cite{Dyreson1994}, and they are often not hindered by imperfections like imprecisions, vaguenesses, uncertainties or inconsistencies possibly contained in the descriptions of these temporal notions.
+Humans handle temporal information using certain temporal notions like time intervals or time points~\cite{Dyreson1994}, and they often have to deal with imperfections like imprecisions, vaguenesses, uncertainties or inconsistencies possibly contained in the descriptions of these temporal notions.
 Among many others, these possible imperfections in descriptions of temporal notions determine an important issue in temporal modelling.
-E.g.
-the description of the temporal notion in a sentence like `The Belfry of Bruges was finished on a day somewhere between 1/01/1201 A.D.
+E.g., the description of the temporal notion in a sentence like `The Belfry of Bruges was finished on a day somewhere between 01/01/1201 A.D.
 and 31/12/1300 A.D.' contains imperfection because of the uncertainty in the used time-related expression.
@@
 
-To allow information systems to cope with these and similar imperfections, many approaches adopt fuzzy sets~\cite{Zadeh65} for the representation of temporal information~\cite{343607},~\cite{nagypal2003},~\cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy},~\cite{Dubois:jucs_9_9:fuzziness_and_uncertainty_in}.
+To allow information systems to cope with these and similar data imperfections, many approaches adopt fuzzy sets~\cite{Zadeh65} for the representation of temporal information~\cite{343607},~\cite{nagypal2003},~\cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy},~\cite{Dubois:jucs_9_9:fuzziness_and_uncertainty_in}.
 The temporal relationships studied by Allen were fuzzified by several authors~\cite{ohlbach2004},~\cite{nagypal2003},~\cite{schockaert08}.
@@
 
+The aim of this chapter is to present and explain some main concepts regarding time in information systems and to present and discuss some issues and techniques concerned with handling data imperfections related to time.
 The rest of this chapter is structured as follows.
-Section \ref{sec:time-domain} presents some basic concepts and terminology about temporal modelling and discusses some of its important aspects and issues.
+Section \ref{sec:time-domain} presents some basic concepts and terminology about temporal modelling and discusses some of its important aspects and issues, while section~\ref{sec:fuzzy-time-mod} presents some important issues concerning the combination of data imperfections and temporal modelling.
 In section \ref{sec:temporal-databases}, some basic concepts and terminology about temporal databases are presented, followed by an overview of some interesting issues concerning temporal databases and a survey of some commercial temporal database systems.

### 2012-03-13 book/sources/timeDomain.tex
Almost done! In this one, the references are a little improved and the piece about fuzzy temporal reasoning (without databases) is written out.

@@
 
-Data models used by information systems (and in specific, temporal database systems) may represent these underlying time domains using \emph{chronons}~\cite{Dyreson1994}.
+Data models used by information systems (and in specific, temporal database systems) may represent these underlying time domains using \emph{chronons}~\cite{Dyreson1994}, which can informally be discribed as the smallest distinguishable time units that can be used in the system.
 However, to explain what chronons are, an explanation of some other temporal concepts is necessary.
@@
 
+\begin{definition}\textbf{Duration}\\
+A \emph{\textbf{duration}} is an amount of time with known length, but no specific starting or ending instants.
+\end{definition}
+\vspace{-10pt}
+\end{svgraybox}
+
+A time interval as such is bounded by two instants, whereas a duration is not. Also, it should be noted that an instant is in fact a singular case of a time interval.
+
+\begin{svgraybox}
+\vspace{-10pt}
 \begin{definition}\textbf{Temporal Element}\\
@@
 An \emph{\textbf{event}} is an instantaneous fact, i.e. something occurring at an instant.
-\end{definition}
-
-\begin{definition}\textbf{Duration}\\
-A \emph{\textbf{duration}} is an amount of time with known length, but no specific starting or ending instants.
 \end{definition}
@@
 
-Data models might now represent an underlying time domain using chronons:
+Data models for time modelling might now represent an underlying time domain using chronons:
 
@@
 
-In a system, the granularity with the shortest granules is the \emph{chronons granularity}, which is denoted as `$\bot$'. It is the granularity of which the granules are chronons. 
+In a system, the granularity with the shortest granules is the \emph{chronons granularity}, which is denoted by `$\bot$'. It is the granularity of which the granules are chronons. 
 %It is now possible to define a function $f$, called a \emph{mapping function}, that maps a given granule $i_G$, $i \in \Z$, from a given granularity $G$ to a set of corresponding chronons.
@@
 f & : G \rightarrow \Pow(\bot)\nonumber\\
-  & : i_{G} \rightarrow \lbrace c_\bot\ |\ (c_\bot\text{ is contained in }i_{G}) \wedge (c_\bot \in \bot)\rbrace\nonumber
+  &  i_{G} \mapsto \lbrace c_\bot\ |\ (c_\bot\text{ is contained in }i_{G}) \wedge (c_\bot \in \bot)\rbrace\nonumber
 \end{align}
@@
 \emph{Regular mapping}: A regular mapping is a granularity-to-granularity mapping where the mapping function value is calculated by means of multiplications and/or divisions and (maybe) an anchor adjustment.
-E.g.
-the mapping value of the mapping between hours and minutes is calculated using a multiplication by $60$.
+E.g., the mapping value of the mapping between hours and minutes is calculated using a multiplication by $60$.
 \item
 \emph{Irregular mapping}: An irregular mapping is a granularity-to-granularity mapping where the mapping function value can not be calculated by means of multiplications and/or divisions.
-E.g.
-the mapping value of the mapping between months and days is dependent on the exact month or day.
+E.g., the mapping value of the mapping between months and days is dependent on the exact month or day.
 \item
 \emph{Congruent mapping}: A congruent mapping is a granularity-to-granularity mapping where the two granularities involved in the mapping have the same granules but a different anchor.
-E.g.
-the mapping between the days (Gregorian calendar days) and the academic days is a congruent mapping.
+E.g., the mapping between the days (Gregorian calendar days) and the academic days is a congruent mapping.
 \end{itemize}
@@
 \begin{example}
+\label{examplegrangraph}
 Consider a system that models both Gregorian calendar dates as well as academic calendar dates.
@@
 Figure \ref{fig:granularity-graph-example} shows the complete granularity graph corresponding to this example.
-The transitions between the chronons granularity and the seconds granularity and between the seconds granularity and the minutes granularity are examples of regular mappings.
-They are represented by normal arrows in the visualisation of the graph.
+The transition between the chronons granularity and the seconds granularity is an example of a regular mapping.
+Regular mappings are represented by thin arrows in the visualisation of the graph.
 The transition between the days granularity and the months granularity is an example of an irregular mapping.
@@
 \includegraphics[scale=0.5]{graphs/granularityGraph.eps}
-\caption{A granularity graph example}
+\caption{The granularity graph corresponding to example \ref{examplegrangraph}}
 \label{fig:granularity-graph-example}
 \end{figure}
+
+\vspace{-25pt}
 
@@
 In this section, a brief introduction can be found, concerning \emph{temporal relationships}, sometimes also called `temporal relations'\cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy}.
-Temporal relationships can be seen as relationships between temporal elements or instants.
+Temporal relationships can be seen as relationships between temporal elements belonging to the same time domain.
+These relationships express how the temporal elements are related to one another, with respect to temporal precedence and overlap.
 
@@
 \includegraphics[scale=0.5]{graphs/allen.eps}
-\caption{Allen relations between two time intervals.}
+\caption{Allen relations between two time intervals I and J.}
 \label{fig:allen}
@@
 
-Several operators are defined in order to compare temporal elements or time points.
-Allen~\cite{Allen83} first described the relationships between time intervals and as a special case, between time points.
+Several (groups of) operators have been proposed in order to compare temporal elements and model the temporal relationships between them.
+Allen~\cite{Allen83} most notably described such relationships between time intervals and as a special case, between instants.
 Figure \ref{fig:allen} shows the temporal relationships Allen discerned.
-Some proposals can be applied to both crisp and fuzzy temporal intervals~\cite{ohlbach2004},~\cite{nagypal2003},~\cite{schockaert08},~\cite{garrido2009}.
+Some proposals can be applied to both crisp and other time intervals~\cite{ohlbach2004},~\cite{nagypal2003},~\cite{schockaert08},~\cite{garrido2009}.
 
-In the next section, some basic concepts and terminology about temporal databases are presented, followed by an overview of some interesting issues concerning temporal databases and a survey of some commercial temporal database systems.
+%In the next section, some basic concepts and terminology about temporal databases are presented, followed by an overview of some interesting issues concerning temporal databases and a survey of some commercial temporal database systems.

### 2012-03-14 book/sources/fuzzyTimeDomain.tex
update in the text and in the bibliography

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 %INTRO
@@
 In this section, some main concepts and issues concerning this combination are presented.
+
+\subsection{\label{subsec:representation}Representation}
+
+As mentioned before, humans deals with time with vagueness.
+Even for some specific events or facts, the temporal specification becomes imprecise.
+Therefore, as a result of the imprecise managment of the time that humans do, a time point is specified or managed by means of a time interval which boundaries may not be precissely known.
+
+\begin{example}
+Consider a speaker and a hearer. The speaker wants to make an appointment with the hearer. Consider now that the speaker says:\\
+\begin{center}
+\emph{`We will meet each other tomorow around 10'}\\
+\end{center}
+The hearer usually agree that the appointment will be in the time interval 9.55 to 10.05.
+\end{example}
+
+The study of the semantics of `around' in temporal~\cite{Dev98} indications has shown that also the size of the temporal interval associated with the imprecise specification of the time depends on the distance with respect to the current time.
+E.g.
+Consider now that the speaker is talking about something that happened \emph{`during last week'} then the hearer would consider a time interval of more or less 10 days.
+
+Some proposals ~\cite{knight1993},~\cite{Cru97},~\cite{nagypal2003},~\cite{Chountas2000} conclude that the best representation for incomplete temporal knowledge are therefore time intervals even if they refer to a fact that happen in a time point.
+This means that, as Allen proprosed in \cite{Allen83} the primitive units (the chronons) in the temporal system are intervals.
+
+In order to manage uncertain temporal information properly, several theoretical frameworks have been proposed:
+
+\begin{description}
+\item[\emph{Probability theory}]
+The probability theory~\cite{Dey1996},~\cite{Lakshmanan1997},~\cite{Haddawy1996} is employed when a probability is associated to the time interval.
+It is very usual in logistics information systems.
+e.g.
+\emph{`The package will arrive at its destination on monday morning with a probability of 0.8'}.
+
+\item[\emph{Possibility theory}]
+The possibility theory~\cite{Zadeh65} associates a possibility degree to the temporal fact or event.
+This theoretical framework is widely used to model uncertainty and vaguenes in time~\cite{Dubois:jucs_9_9:fuzziness_and_uncertainty_in},\cite{Dubois89},\cite{devos94},\cite{nagypal2003}.
+Several works~\cite{schockaert08},~\cite{ohlbach2004} presents a fuzzy version of the temporal relations given by Allen~\cite{Allen83}.
+The aim of these works is to obtain a flexible way to compare uncertain, ill-known temporal intervals by means of the temporal relationships.
+
+\item[\emph{Rough sets}]
+Rough set theory~\cite{Pawlak1995} has been used to represent uncertainty in time intervals.
+The two dimensional representation of time intervals and the temporal relationships between them has been studied in~\cite{Qia09}.
+\end{description}
+
+%\subsection{\label{subsec:classification}Classification of imperfect temporal information}
 
@@
 
+Temporal information is usually related to facts or events~\cite{Chountas2000}. Therefore, the following types of temporal information may be found:
+
+\begin{description}
+\item[\emph{Definite temporal information}]
+It is found when all times associated with the fact are absolute times. The temporal information is precisely known.
+
+\item[\emph{Indefinite temporal information}]~\cite{Dey1996}
+The time associated with the fact has not been fully defined. Consider an event that in fact occur but it is not known exactly when.
+
+\item[\emph{Infinite temporal uncertain information}]~\cite{Kabanza1990}
+An infinite number of times are associated with a fact.
+This is usually found in recurrent events like meetings.
+(e.g.
+\emph{The meetings are every Wednesday at noon.}).
+Some systems (usually with different information providers) may dispute the occurence of a fact and / or the duration of a fact.
+\end{description}
+
 \subsection{Imperfections in Temporal Relationships}
@@
 
-%%%TODO
-
-%Humans usually deals with time in a imprecise way.
-e.g.
-`A few days ago we received this package'.
-This issue has been studied in the field of Artificial Intelligence ~\cite{Tre97} and language understanding~\cite{DeCaluwe:1997:FTI:285506.285516}.
-Some proposals ~\cite{knight1993},~\cite{Cru97},~\cite{nagypal2003} conclude that the best representation for incomplete temporal knowledge are time intervals instead of time points.
-This means that, as Allen proprosed in \cite{Allen83} the primitive units (the chronons) in the temporal system are intervals.
-
-%When modelling temporal knowledge, the following types of information might be found~\cite{nagypal2003}:
-
-%\begin{description}
-%\item[\emph{Uncertainty.}] The temporal specification of the event is uncertain. This is usually found in historical documents when some documents state contradictory facts about some event.
-
-%\item[\emph{Vagueness.}] Some of the temporal specifications are defined in a vague way.
-e.g.
-`during summer', `late night'.
-Note also that temporal specifications become vague when the granularity is changed ~\cite{Cru97},~\cite{nagypal2003} .
-e.g.
-`The book was written by July 2012' is usefull if we are interested in the month and the year but it turns into a incomplete expression when we try to obtain the exact day.
-It is possible to set the following classification~\cite{Dev98}:
-
-%\begin{itemize}
-%\item
-%Non-numerical indications. e.g. \emph{often, usually}.
-%\item
-%Approximative time indications. e.g. \emph{around 10 a.m.}
-%\item
-%Indications on temporal relations. e.g. \emph{after 12, during summer}.
-%\end{itemize}
-
-%\item[\emph{Subjectivity} or \emph{Ambiguity}.] The specification of a temporal fact may be affected by subjectivity.  It is possible to distinguish between:
-%\begin{itemize}
-%\item Historical periods~\cite{nagypal2003} like `late romanticism' or `early middle ages'. 
-%\item The interpretation of the temporal event~\cite{Dev98} depends on the side of the hearer.
-e.g.
-A speaker says `We will meet each other at six', the hearer has to choose between 6 p.m.
-or 6 a.m.
-The speaker, however knows exactly which one is meant.
-%\end{itemize}
-%\end{description}
-
-%Another important task in the modelling of incomplete knowledge is when dealing with transitions among the different granularities.
+\subsection{\label{subsec:temporal-granularities}Imperfections on temporal granularities}
 

### 2012-03-14 book/sources/fuzzyTimeDomain.tex
update in the bibliography, and in the graph for the example in the fuzzy time part. Update also in the commercial database systems, the addition of secondo as a temporal database system.

@@
 These approaches are often based on Allen's operators~\cite{Allen83}.
+Example \ref{ex:fuzzy-allen-relation} is a case of use for one of the Allen's relations.
 
-\subsection{\label{subsec:temporal-granularities}Imperfections on temporal granularities}
+\begin{example}
+\label{ex:fuzzy-allen-relation}
+Consider an event which is bounded at $[A,B]$ both the starting and ending points respectively (see (1) in figure \ref{fig:fuzzy-allen-relationship} ).
+The classical Allen's after relation returns an interval from $]B,\infty[$ as shown in (2) in figure \ref{fig:fuzzy-allen-relationship}.
+A fuzzified version of the Allen's operators is illustrated in (3), figure \ref{fig:fuzzy-allen-relationship}.
+The comparison between two time intervals results in a possibility degree in the unit interval.
+The shape for the membership function is shown in (3), figure \ref{fig:fuzzy-allen-relationship}.
+Note that all the points strictly after the point $B$ results in a possibility degree of 1 whereas there is an area near the point $B$ in which the possibility degree runs smoothly between 0 and 1.
+
+Consider now the interval given by $[C,D]$, illustrated in (4), figure \ref{fig:fuzzy-allen-relationship}.
+The user wants to know if $[C,D]$ is after $[A,B]$.
+The crisp version of the \emph{`after'} operator would return \emph{`no'} as an answer.
+The fuzzy version for the same operator would return \emph{`yes, with a possibility of 0.5'}.
+
+\begin{figure}
+\centering
+\includegraphics[scale=0.5]{graphs/fuzzyAllen.eps}
+\caption{Example for the Allen's relation, \emph{`after'}.
+(1) The event bounded within the time points $A$ and $B$.
+(2) The crisp version for the \emph{`after'} operator.
+(3) A fuzzy version for the after operator.
+(4) Another event bounded within the time points $[C,D]$.}
+\label{fig:fuzzy-allen-relationship}
+\end{figure}
+\end{example}
 

### 2012-03-14 trunk/book/sources/temporalDatabases.tex
update in the bibliography, and in the graph for the example in the fuzzy time part. Update also in the commercial database systems, the addition of secondo as a temporal database system.

@@
 %
-A temporal database can generally be seen as a database that manages some temporal aspects in its schema \cite{etzion1998}~\cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy}.
+A temporal database can generally be seen as a database that manages some temporal aspects in its schema \cite{etzion1998},~\cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy}.
 In subsection \ref{subsec:td-general-concepts}, some main concepts and properties concerning temporal databases and their definitions are presented and explained.
@@
 
-\emph{Valid times} are usually supplied by the user, whereas \emph{transaction times} are usually system-generated and -supplied \cite{Dyreson1994}.
+\emph{Valid times} are usually provided by the user, whereas \emph{transaction times} are usually system-generated and -supplied \cite{Dyreson1994}.
 Temporal values of type UDT are not given any extraordinary interpretation and have thus no extraordinary query language support \cite{Dyreson1994}.
@@
 Generally, when designing a relation based on a relational database model, a subset of the relation's attribute set is ususally chosen as primary key.
-The values of a tuple for these attributes will then uniquely determine this tuple and no two distinct tuples may have the exact same values for every attribute in this primary key.
+The values of a tuple for these attributes will then uniquely determine this tuple, hence no two distinct tuples may have the exact same values for every attribute in this primary key.
 Next to attributes unrelated to time, a valid-time relation schema will typically contain one or more attributes which model the valid-time aspects and behavior of the real objects and concepts modelled by the relation schema.
@@
 
-\subsection{\label{subsubsec:consistency}Consistency in Valid-time Relation Querying}
-TODO\\
-TODO\\
-TODO\\
-
+\subsection{\label{subsubsec:consistency}Consistency in Valid-time Relation Content Modification}
 The solution presented in subsection~\ref{subsubsec:primary-key} concerns relation design and consists of including the valid-time attributes in the primary key.
@@
 
-The inconsistency problem can now be solved by replacing this statement with the next:
+\noindent
+The inconsistency problem can now be solved by replacing this statement with:
 
@@
 
+\noindent
 The resulting relation is visualized in table~\ref{table:example-database-update}.
@@
 \subsection{\label{Comm-temp}Commercial Temporal Database Systems}
-Several commercial temporal DBMS exist. Table \ref{table:commercial-temporal-db} presents some information on some of the more well-known temporal DBMS and provides references for more information.
+Several commercial temporal DBMS exist. Table \ref{table:commercial-temporal-db} gives an overview of some of the more well-known temporal DBMS and provides references for more information.
 
@@
 Teradata~\cite{teradata2011} is mainly a business intelligence system designed for data mining.
+Secondo~\cite{Dieker2000} is an extensible database system in which the core of the database may be replaced by a customized algebras.
+It is designed for non-standard applications and it support both valid and transaction times.
+
  The most complete implementation is Workspace Manager.
@@
 Teradata & VT and TT. & Used for data-mining. & \cite{teradata2011}\\
+Secondo & VT and TT. & Spatio-temporal database. & \cite{Guting} \\
 \hline 

### 2012-03-15 book/sources/fuzzyTemporalDatabases.tex
A little commit inbetween (don't mind the mess)

--- 2012-01-12 book/sources/fuzzyTemporalDatabases.tex
@@
 %
-Some software applications deal with uncertainty related to time.
+Consider a logistics company which transports packages.
+At the moment a package leaves, the time when it will arrive at its destination may be estimated, but will typically not be known precisely.
+For such companies and in many other situations, information systems able to handle imperfection with respect to certain temporal aspects of the objects modelled by the system are necessary.
+
+\subsection{Data Imperfections in Temporal Databases}
+
+%The time when a package would arrive at its destination may be estimated and can eventually be researched, but it is typically not always precisely known.
+For such companies and in many other situations, information systems able to handle imperfection corresponding to certain temporal aspects of the objects modelled by the system are necessary.
+The transition between granularities~\cite{Lin97} is also sometimes considered as a source of imprecision.
+Therefore, some proposals consider granularity as the base of the temporal model~\cite{Cru97}.
+%Some software applications deal with uncertainty related to time.
 e.g.
 a logistics company which transports packages from one destination to another.
-The time when the package would be at the end point may be stimated but not precisely known.
+The time when the package would be at the end point may be estimated but not precisely known.
 These applications need to manage the uncertainty with respect to the temporal attributes of the objects stored in the database.
 
-There are several approaches to manage uncertainty in the temporal information stored in a database.
-Valid-time is usually represented as an interval, and therefore it is suitable for the representation of uncertainty in the valid-time interval.
-Nevertheless, the values for transaction and decision time are usually precisely known.
-Some models deal with probability~\cite{Dekhtyar2001} or possibility~\cite{Dubois89} distributions at the endpoints of the time interval.
-In this section we will explain the most interesting proposals that deal with possibility distributions in the valid-time interval.
+Data and information imperfections and techniques to represent them correctly in databases and queries are usually the focus of research in fuzzy databases.
+Proposals from this field may present an approach based somehow on fuzzy set theory~\cite{Zadeh65} or possibility theory~\cite{Dubois:Prade:1988:PossibilityTheory}, although other theories support information imperfections too.
+Comparably, many proposals concerning the introduction of data imperfections or information imperfections in temporal databases present approaches based somehow on fuzzy set theory~\cite{garrido2009},~\cite{gal01},~\cite{nagypal2003},~\cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy} or possibility theory~\cite{Dubois89},~\cite{Pon11},~\cite{129895}, although proposals based on other theories exist~\cite{Dyreson1998},~\cite{Dekhtyar2001},~\cite{Qia09}.
+As possibility theory is usually seen as a theory of confidence, aimed at dealing with uncertainty, in some proposals, possibility theory is used specifically to handle uncertainty in temporal information.
+In fuzzy databases~\cite{Galindo2006}, uncertainty is usually expected to appear at storage level (in the database), whereas other types of imperfection, notably imprecision, are usually expected to appear in querying.
+
+Concerning temporal databases, there are several approaches to handle uncertainty in temporal data stored in a database.
+Many of these approaches concern several different types of time notions (VT, TT or DT), but most of these approaches focus somehow on valid time~\cite{garrido2009},~\cite{gal01}.
+
+In the following subsection, a novel approach to representing uncertainty concerning valid time notions and a corresponding technique to query similar valid time indications in a valid-time relation are proposed.
+The presented proposal is based on concepts introduced in~\cite{Dubois88b} and on the framework proposed in~\cite{Pon11}.
+
+\subsection{Handling Uncertainty in a Valid-time Relation}
+A valid time indication usually takes the form of a time interval.
+Such a valid time interval can be described (and stored in a valid time relation record) using its boundaries (endpoints) or using one endpoint and the interval length.
+Usually, a valid-time interval is represented using its endpoints, which is also the approach adopted by the presented proposal.
+
+Generally, the uncertainty concerning a set of values might be described by a possibility distribution on the powerset of which one of the elements can be the intended set~\cite{Dubois88b}.
+This representation, however, introduces some issues in practice or in practical applications.
+Therefore, in the presented proposal, possibility theory is used to model uncertainty, but only uncertainty concerning the exact values of the start and end point of a valid-time interval is considered and the uncertainty in both the start point and the end point are modelled using possibility distributions.
+
+In fact, to model the uncertainty related to a valid-time interval using possibility theory, the presented proposal relies on the concept of ill-known sets~\cite{Dubois88b}.
+
+%Some models consider probability~\cite{Dekhtyar2001} or possibility~\cite{Dubois89} distributions to describe the endpoints of a time interval.
+In this section, we will present a novel and currently researched technique to describe valid-time interval start- and endpoints using possibility distributions.
+In subsection \ref{subsec:representation-time-intervals} we will present an approach to modelling uncertainty in valid-time intervals.
+Finally, in subsection \ref{subsec:querying-time-intervals} the proposals for querying valid-time intervals in a flexible way are presented and discussed.%In this section we will explain the most interesting proposals that deal with possibility distributions in the valid-time interval.
 %Two are the main points developed within this section: the representation of uncertainty in the time interval and the flexible querying of ill-known time periods.
 
-In fuzzy databases~\cite{Galindo2006}, uncertainty is managed at storage level whereas imprecision is managed at querying level.
+%In fuzzy databases~\cite{Galindo2006}, uncertainty is managed at storage level whereas imprecision is managed at querying level.
 In subsection \ref{subsec:representation-time-intervals} we will explain the main approach for the storage of uncertainty in the valid-time interval.
@@
 
-\subsection{Representation of ill-known temporal intervals}
+\subsubsection{Ill-known Time Intervals}
 \label{subsec:representation-time-intervals}
-The uncertainty in a valid-time interval may be represented with uncertainty at one or both starting and ending points.
+%Uncertainty in a valid-time interval may be modelled by modelling the uncertainty concerning the exact start, respectively end point of the valid-time interval.
+The uncertainty concerning these points can by modelled by means of possibility distributions.
+It is of course also possible to use rough sets~\cite{Pawlak1995} for this modelling~\cite{Qia09}.
+The work presented in this section relies on possibilistic variables, which rely on possibility theory \cite{Dubois:Prade:1988:PossibilityTheory}.
+
+The presented proposal relies on the concept of ill-known sets~\cite{Dubois88b}.
+To correctly explain the concept of ill-known sets, the concept of possibilistic variables should be introduced first.
+In the presented proposal, the definition of possibilistic variables of~\cite{Pon11} is followed.
+A \emph{possibilistic variable} is defined as follows~\cite{Pon11}.
+
+\begin{svgraybox}
+\vspace{-10pt}
+\begin{definition}
+A possibilistic variable $X$ over a universe $U$ is defined as a variable taking exactly one value in $U$, but for which this value is (partially) unknown.
+Its possibility distribution $\pi_X$ on $U$ models the available knowledge about the value that $X$ takes: for each $u\in U$, $\pi_X(u)$ represents the possibility that $X$ takes the value $u$.
+In the presented work, this possibility is interpreted as a measure of how plausible it is that $X$ takes the value $u$, given (partial) knowledge about the value $X$ takes.
+\end{definition}
+\vspace{-10pt}
+\end{svgraybox}
+
+BASE ON IPMU 2
+
+Consider a set $R$, which contains single values (and not collections of values).
+When a possibilistic variable $X_{v}$ is defined on such a set $R$, the unique value $X_{v}$ takes, which is (partially) unknown, is called an \emph{ill-known value} in this work \cite{Dubois88b}.
+
+%The exact value a possibilistic variable takes, which is (partially) unknown, is called an \emph{ill-known value} in this work \cite{Dubois88b}.
+
+When a possibilistic variable is defined on the powerset $\Pow(R)$ of some universe $R$, the unique value the variable takes will be a crisp set and its possibility distribution on the powerset $\Pow(R)$ will describe the possibility of each crisp subset of $R$ to be the value the variable takes.
+This exact value (a crisp set) the variable takes, is now called an \emph{ill-known set} \cite{Dubois88b}.
+
+This work will deal with ill-known time intervals.
+These are ill-known intervals in time.
+Ill-known intervals are ill-known sets, defined and represented via a start and end point, which will be ill-known values.
+The elements of the set are the values between the start and end point.
+A closed ill-known interval with start point defined by possibilistic variable $X$ and end point by possibilistic variable $Y$ is noted here $\left[X, Y\right]$.
+The correspondences and transitions between the representations of ill-known sets, between the representations of ill-known intervals and between the representations of an ill-known set and an ill-known interval are part of the authors current research.
+Figure \ref{fig:interval} illustrates this.
+
+%Uncertainty in a valid-time interval may be modelled with uncertainty at one or both starting and ending points.
 The uncertainty is usually represented by means of a possibility distribution in one or both points.
@@
 
-\subsection{Querying fuzzy temporal databases}
+In section \ref{subsec:querying-time-intervals}, a technique allowing the querying of fuzzy valid-time databases which contain ill-known valid-time intervals is presented and briefly explained.
+The technique relies on the framework for set evaluation presented in \cite{Pon11}.
+
+\subsubsection{Ill-known Time Constraints}
+
+\subsubsection{Querying Fuzzy Valid-Time Databases with Ill-Known Valid-Time Intervals}
 \label{subsec:querying-time-intervals}
 
-\subsubsection{Bipolarity in temporal database querying}
+One of the main purposes of the existence of databases is to allow information retrieval.
+The standard query language for databases is SQL ~\cite{Mel93}.
+Several proposals to extend the SQL language for transaction-time databases~\cite{Sarda90}, valid-time databases~\cite{gad92} and bitemporal databases ~\cite{TSQL} exist and some authors have studied how to support temporal querying in standard SQL~\cite{Snodgrass98}.
+
+In the following subsections, the query structure, the query evaluation method and the technique to meaningfully rank and aggregate the query results, proposed in this work, are presented.
+
+%In the querying of a fuzzy temporal database it is possible to distinguish among the following cases:
+
+%\begin{itemize}
+%\item
+%Fuzzy data stored in the database and crisp specification in the query.
+%\item
+%Crisp data stored in the database and fuzzy data in the query specification.
+%\item
+%Both data stored in the database and the query specification are fuzzy.
+%\end{itemize}
+%In this subsection we will explain the query specification for a database that stores ill-known time intervals and crisp values in the query.
+Afterwards, the aggregation and ranking of the temporal results is explained.
+
+\paragraph{\textbf{Query Structure}}
+In this work, a query $\tilde{Q}$ is made up of two separate constructs of user query demands.
+
+%Consider that in a regular or fuzzy relational database, the query specification for the non-temporal attributes is given by $Q$.
+Therefore, the query specification in a temporal or fuzzy temporal database is given by $\tilde{Q}$:
+
+\begin{svgraybox}
+\vspace{-10pt}
+\begin{definition}
+\textbf{Query}
+A query $\tilde{Q}$ in this work is given by:
+\begin{equation}
+\label{eq:query-specification}
+\tilde{Q} = \left( Q^{time}, Q \right)
+\end{equation}
+Here $Q = \left \lbrace q_1,\ \cdots,\ q_n \right \rbrace$ contains all (possibly fuzzy) non-temporal user-defined query demands.
+These comprise all constraints unrelated to the valid-time indications in the queried relations.
+$Q^{time}$ denotes the temporal constraint specified by the user.
+\end{definition}
+\vspace{-10pt}
+\end{svgraybox}
+%%specify better this:
+%\begin{svgraybox}
+%\begin{definition}
+%\label{def:query-constraint}
+%\textbf{(query constraint)}
+%$Q$ is the 
+%A query constraint $q_a$ for an attribute $a$ with attribute domain $\mathcal{D}$, is a restriction in the subset of the values in the attribute domain:
+%\begin{equation}
+%\label{eq:query-constraint}
+%q_a = \left \lbrace d \mid d \subseteq \mathcal{D} \right \rbrace
+%\end{equation}
+%\end{definition}
+
+In this work, the query structure allows the user to specify a single temporal constraint.
+
+\begin{svgraybox}
+\vspace{-10pt}
+\begin{definition}
+\textbf{Temporal Constraint $Q^{time}$}
+In this work, $Q^{time}$ is defined by:
+\begin{equation}
+Q^{time} = \left( I,AR \right)
+\end{equation}
+where $I$ denotes a crisp time interval and $AR$ is one of the Allen relations.
+\end{definition}
+\vspace{-10pt}
+\end{svgraybox}
+
+The interpretation is that, for a record with an ill-known valid-time interval $J$, the user demands that $I$ AR $J$ hold.
+
+\paragraph{\textbf{Query evaluation}}
+\label{subsubsec:query-evaluation}
+Query satisfaction in a fuzzy relational database is a matter of degree.
+Typically, the query evaluation results in a \emph{satisfaction degree} $s$, which is typically in the unit interval.
+$s \in \left[ 0,1,\right]$.
+Here, a satisfaction degree of $0$ denotes a total dissatisfaction while a degree of $1$ denotes a total satisfaction.
+
+%\begin{svgraybox}
+%\begin{definition}
+%\label{def:evaluation-function}
+%\textbf{Evaluation function}
+%An evaluation function $e_{Q} \left( r \right)$ is a mapping from the attribute value from the row $r$ in the database with respect to the constraints $Q = \left \lbrace q_1, \cdots, q_n \right \rbrace$ to the unit interval $\left[ 0,1 \right]$.
+%\begin{equation}
+%\label{eq:evaluation-function} 
+%e_{Q} = t \mid t \in \left[ 0,1 \right]
+%\end{equation}
+%\end{definition}
+%\end{svgraybox}
+
+In the presented approach, for every record $r$, each part of $\tilde{Q}$ is evaluated independently:
+\begin{itemize}
+\item
+The user preferences expressed in the non-temporal part, $Q$, are evaluated, resulting in a satisfaction degree denoted as $e_Q(r)$.
+The presented approach accepts any valid, sound method of calculating this evaluation, as long as the method is well-founded and $e_Q(r) \in \left[0,1\right]$.
+\item
+The evaluation of the temporal demand expressed in the temporal part, $Q^{time}$, depends on $AR$.
+A specific set of ill-known constraints\cite{Pon11} is considered depending on the Allen relation denoted by $AR$.
+The exact set chosen can be found in table \ref{tab:fuzzy-allen-relations}, for every possible value of $AR$.
+Then, using these formulas given in table \ref{tab:fuzzy-allen-relations}, both the possibility degree $\Pos_{Q^{time}}(r)$ and the necessity degree $\Nec_{Q^{time}}(r)$ are computed.
+\end{itemize}
+
+\textcolor{red}{TODO: maybe part on what these and mean? Maybe part on where the formulas in table \ref{tab:fuzzy-allen-relations} come from?}
+
+\begin{table}[h]
+
+\caption{Allen's relations used in the framework.
+Here, $I = \left[a, b\right]$ denotes a crisp time interval, $J = \left[X, Y\right]$ denotes an ill-known time interval, with $\pi_{X}$ and $\pi_{Y}$ the possibility distributions of $X$ and $Y$ respectively.
+The second column contains the corresponding formula to calculate the possibility that $I$ satisfies all constraints given by the Allen's relation.}
+
+\centering
+\begin{tabular}{|c|c|c|}
+\hline
+Allen Relation &  $\Pos\left(\text{I satisfies all constraints }C_{i}, i = 1,2,...\right)$ & $\Nec\left(C_{i}, i = 1,2,... \right)$ \\
+\hline
+I before J & $\sup_{a>w}\pi_X(w)$ & $\inf_{a \leq w} 1- \pi_X(w)$\\
+\hline
+
+\multirow{4}{*}
+{I equal J} &  $\min ( \sup_{a \leq w}\pi_X(w),$  & $\min ( \inf_{a>w} 1-\pi_X(w),$\\
+& $\pi_X(w),$ & $\inf_{w \in \left[a,b\right]} \pi_X(w), $ \\
+ &  $\sup_{b \geq w}\pi_Y(w),$ & $\inf_{b>w} 1-\pi_Y(w),$\\
+ & $\pi_Y(w))$ & $\inf_{w \in \left[a,b\right]} \pi_Y(w) )$ \\
+\hline
+
+\multirow{2}{*}
+I meets J  & $\min (\sup_{a\geq w} \pi_X(w),$ & $\min (\inf_{a<w} 1 - \pi_X(w),$\\
+& $\pi_X(w))$  & $ \inf_{w \in \left[a,b\right]} \pi_X(w) )$ \\
+\hline
+
+\multirow{3}{*}
+{I overlaps J}  & $\min ( \sup_{b>w}\pi_Y(w), $ & $\min ( \inf_{b \geq w} 1 - \pi_Y(w),$\\
+ & $\sup_{a \geq w}\pi_X(w),$& $1- \inf_{a>w}1-\pi_X(w),$\\
+ & $\sup_{a \leq w}\pi_X(w))$ & $1- \inf_{a<w}1-\pi_X(w) )$ \\
+\hline
+
+\multirow{4}{*}
+{I during J}  & $\max ( \min ( \sup_{a<w}\pi_X(w),,$ & $\max (\min (\inf_{a \leq w}1-\pi_X(w), $\\
+ & $\sup_{b \geq w}\pi_Y(w)) $ & $ \inf_{b>w} 1- \pi_Y(w)),$ \\
+ & $\min ( \sup_{a \leq w }\pi_X(w),$ & $\min ( \inf_{a<w}1-\pi_X(w),$\\
+ & $\sup_{b>w}\pi_Y(w))$ & $\inf_{b\geq w}1-\pi_Y(w)))$ \\
+\hline
+\multirow{2}{*}
+{I starts J} &  $\min( \sup_{a \leq w}\pi_X(w),$  & $\min (\inf_{a<w}1-\pi_X(w),$\\
+& $\pi_X(w))$ & $inf_{w \in \left[a,b\right]} 1- \pi_X(w))$ \\
+\hline
+\multirow{2}{*}
+{I finishes J} &  $\min ( \sup_{b \geq w} \pi_Y(w),$  & $\min(\inf_{b>w}1-\pi_Y(w),$\\
+& $\pi_Y(w))$ & $\inf_{w \in [a,b]} 1- \pi_Y(w))$ \\
+\hline
+
+\end{tabular}
+%
+%\vspace{10pt}
+%
+%
+%\vspace{-25pt}
+\label{tab:fuzzy-allen-relations}
+\end{table}
+
+\subsubsection{Aggregation and Ranking} TO BE ANOTHER PART
+To be able to present the most appropriate results to the user, for every record $r$, an aggregation method is used to aggregate $\Pos_{Q^{time}}(r)$ and $\Nec_{Q^{time}}(r)$ into a temporal record rank $e_{Q^{time}}(r)$ and after this, a convex combination combining $e_{Q^{time}}(r)$ and $e_{Q}(r)$ will provide the final record rank $e_{final}(r)$.
+
+%For each temporal attribute $r_{time}$ for the record $r$, both possibility $\Pos \left( Q_{time}\right)$ and necessity $\Nec \left( Q_{time} \right)$ measures are obtained.
+Then, the measures are combined in order to obtain an evaluation score $e_{Q_{time} \left( r \right)}$.
+The evaluation function is computed by:
+
+To calculate $e_{Q^{time}}(r)$, an a simple and crude method is used:
+
+\begin{equation}
+\label{eq:temporal-eval}
+e_{Q^{time}}(r) = \left(\frac{\Pos_{Q^{time}}(r)+\Nec_{Q^{time}}(r)}{2}\right)
+%e_{Q_{time} \left( r \right)} = \left \lbrace t \mid t = \left( \frac{\Pos \left( Q_{time}\right) + \Nec \left( Q_{time} \right)}{2} \right)   \in \left[ 0,1\right] \right \rbrace
+\end{equation}
+
+This method aims to provide the result records with a natural ranking based on the users temporal constraint.
+$e_{Q^{time}}(r)$ will of course be a value in $\left[0,1\right]$, as both $\Pos_{Q^{time}}(r) \in \left[0,1\right]$ and $\Nec_{Q^{time}}(r) \in \left[0,1\right]$.
+The purpose is that records which fit the users temporal demand better get a higher score than records fitting the temporal demand worse.
+Here, this aim is reached because the necessity degree $\Nec_{Q^{time}}(r)$ cannot exceed $0$ unless the possibility degree $\Pos_{Q^{time}}(r)$ equals $1$.
+
+%This measure provides a natural score because of the following property of the necessity measure: 
+%\begin{equation}
+%\label{eq:necessity} %iff or iif?
+%\Nec \left( Q_{time} \right) > 0  \Longleftrightarrow  \Pos \left( Q_{time}\right) = 1
+%\end{equation}
+
+The final ranking $e_{final}(r)$ for a record $r$ is now given by a convex combination of both temporal and non-temporal evaluation scores.
+
+\begin{equation}
+\label{eq:convex-combination}
+e_{final} \left( r \right) = \omega \ast e_{Q} \left( r \right) + \left( 1- \omega \right) \ast e_{Q^{time}}
+\end{equation}
+
+A convex combination is used mainly for 2 reasons:
+\begin{itemize}
+	\item The use of this convex combination allows a record to make up for a low temporal evaluation score with a high non-temporal evaluation score and vice versa.
+	\item The exact value of $\omega$ can now be modified to ascribe more value to either the fulfillment of the user's temporal demands or the fulfillment of the user's non-temporal constraints.
+\end{itemize}
+
+\subsection{Bipolarity in Temporal Databases}
+\label{subsubsec:bipolarity}
+Humans express their preferences using both positive and negative statements, where positive statements express what is desired or acceptable and negative statements express what is undesired or unacceptable \cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy}.
+This realization is interesting with regard to database querying, because sometimes a user does not exactly know his or her preferences or can't express them in only positive statements, but prefers to use negative statements to express what he or she dislikes or doesn't need.
+This introduces the need for \emph{bipolar querying}, a technique to model both positive and negative user preferences in a database query.
+Sometimes positive and negative preferences are clearly symmetric, making it possible to derive one from the other.
+For example, a person may define the concept of `tall' as `1.80 meters or higher'.
+The negative would then be the opposite: not tall would be `anyhing less than 1.80 meters'.
+However, in some cases, positive preferences can not be directly obtained from negative preferences or vice versa.
+E.g.
+when a person prefers to buy a black motorbike, this does not necessarily mean the person would totally reject a very dark blue motorbike.
+This phenomenon is called \emph{heterogeneous bipolarity}~\cite{Dubois2006},~\cite{Dubois2008}.
+
+The use of imprecise query preference formulation in bipolar querying is well discussed in existing literature \cite{DeTre2009}, \cite{Dubois2008}, \cite{Lacroix87}.
+In \cite{Lacroix87}, desired and mandatory query conditions are used, instead of positive and negative preferences.
+However, the inverse of a mandatory preference expresses what should be rejected and this could be seen as negative information, whereas desired query conditions can be seen as positive preferences.
+However, the combination of bipolar querying and the use of imprecise query preferences in the context of temporal databases is not so well discussed in existing literature.
+A proposal for the bipolar querying of valid-time databases has been made by Billiet et al.~\cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy}.
+The model presented there deals with a fuzzy valid-time specification based on~\cite{garrido2009}.
+
+%Some previous work have been done in ~\cite{Lacroix87}.
+There, instead of positive and negative criteria, desired and mandatory query conditions are explained.
+This can be seen as a way to specify positive and negative information: the inverse of a mandatory condition is what should be rejected (negative information), whereas desired conditions can be seen as positive information.\\
+
+%This introduces the necessity of a bipolar way of querying databases.
+Therefore, the bipolar querying of databases~\cite{DeTre2009} is a querying technique which allows the user to express both positive and negative criteria in the database query.\\
+
+Bipolarity can be handled using different concepts, such as intuitionistic fuzzy sets~\cite{Atanassov1986}, interval valued fuzzy sets~\cite{Zadeh75a} Grattan-Guiness~\cite{Grattan76}, Janh~\cite{Jahn75}, Sambuc~\cite{Sambuc75},~\cite{Dubois05} or two fold fuzzy sets \cite{Dubois02}.
+
+%\begin{itemize}
+%\item
+%Intuitionistic Fuzzy Sets~\cite{Atanassov1986} are sometimes used.
+The approach re-define the concept of inverse.
+Instead of considering the inverse of a set just the complementary, a AFS is defined by means of a membership function and a non-membership function.
+%\item
+%The interval valued fuzzy sets \emph{IVFS} were introduced independently by Zadeh~\cite{Zadeh75a}, Grattan-Guiness~\cite{Grattan76}, Janh~\cite{Jahn75}, Sambuc~\cite{Sambuc75} in the seventies~\cite{Dubois05}.
+%look for the reference!
+%\item
+%The two fold fuzzy sets \emph{(TFS)} by Dubois and Prade~\cite{Dubois02}. This approach defines desired and mandatory query conditions.
+%\end{itemize}
+
+From a theoretical point of view, bipolarity might be found either in the queries presented to a database system or in a database managed by a database system.
+
+When bipolarity is found in queries, it is possible to distinguish between:
+\begin{itemize}
+	\item
+	Bipolarity inside query criteria: each individual query criterion may be specified using both positive and negative preferences.
+For example when querying a car database, the user can express that he or she wants a black car, but definitely not a red neither a blue one.
+Bipolarity resides here within the car color criterion.
+	\item
+	Bipolarity outside query criteria: the query is specified using a global positive and a global negative preference part.
+For example when querying a car database, the user can express that he or she wants a black car, but definitely not a car with a fuel consumption of 6 liters or more.
+	\end{itemize}
+
+Concerning bipolarity inside a database, it should be possible to specify both positive and negative real world object or concept aspects, even at record level.
+Nevertheless, not so much research exists concerning bipolarity in databases.
+
+%some attributes or even at tuple level it should be possible to specify them with bipolarity. Nevertheless, there is not so many research on this topic.
+%	\begin{itemize}
+%	\item
+%	Attribute bipolarity: Each attribute may be specified  
+%	\item
+%	Tuple bipolarity:
+%	\end{itemize}
+
+%\paragraph{Query structure}
+%The approach in this work follows the query specification given in \eqref{eq:query-specification}.
+The query has a global time demand, $Q^{time}$ and here, the regular query constraints specified by $Q$ are splitted into possitive and negative preferences:
+%
+%\begin{equation}
+%\label{eq:bipolar-specification}
+%\tilde{Q} = \left \lbrace Q^{time}, \left( Q^{pos}, Q^{neg} \right) \right \rbrace
+%\end{equation}
+%
+%Here $Q^{pos}$ and $Q^{neg}$ represent the positive and the negative criteria, respectively.
+The temporal demand in $Q^{time}$ may be specified as in ~\cite{garrido2009} or as in ~ \cite{Pon11}, see section \ref{subsec:representation-time-intervals}.
+
+%\paragraph{Query evaluation}
+%As explained previously in the query evaluation section (\ref{subsubsec:query-evaluation}), each element in $\tilde{Q}$ is evaluated independently:
+%
+%\begin{itemize}
+%\item
+%The query $Q$ has now two elements: $Q^{pos}$ and $Q^{neg}$.
+Each element is again evaluated independently and the result is a tuple $\left(s,d \right)$ in which $s$ is the satisfaction degree for $Q^{pos}$.
+$d$ is called dissatisfaction degree for $Q^{neg}$.
+The tuple $\left(s,d \right) s,d \in \left[0,1 \right]$ is called \emph{Bipolar Satisfaction Degree}.
+%\item
+%The temporal specification is evaluated as explained in \ref{subsubsec:query-evaluation} and the result is a value in the unit interval.
+%\end{itemize}
+%
+%\paragraph{Ranking}
+%In order to present the results to the user, it is necessary to design a combination function.
+This function allows the classification of the results in the unit interval the preffered way is to made a convex combination as  presented in equation \eqref{eq:convex-combination}.
+%

### 2012-03-16 trunk/book/sources/frw.tex
To my opinion, the content is now perfect.

@@
-In this chapter, some of the main concepts concerning temporal modelling and temporal modelling in information systems and the terminology corresponding with these concepts are introduced and explained and some of the main properties of and issues with these concepts are presented and discussed.
+In this chapter, some of the main concepts concerning information imperfections in temporal modelling and information imperfections in temporal modelling in information systems and the terminology corresponding with these concepts are introduced and explained and some of the main properties of and issues with these concepts are presented and discussed.
 An overview of some commercial temporal DBMS is briefly introduced.

### 2012-03-16 trunk/book/sources/fuzzyTimeDomain.tex
To my opinion, the content is now perfect.

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%INTRO
+As explained in the introduction, humans handle temporal information using temporal notions like time intervals or time points \cite{Dyreson1994}.
+While the used temporal notions may contain imperfections~\cite{Dev98},~\cite{Dubois:jucs_9_9:fuzziness_and_uncertainty_in},~\cite{nagypal2003},~\cite{Dubois89}, humans often gracefully deal with these, as their inherent interpretation capability accounts for a lot of them.
+This phenomenon has been studied a.o.
+in the field of artificial intelligence~\cite{Tre97},~\cite{5151} and language understanding~\cite{DeCaluwe:1997:FTI:285506.285516},~\cite{nagypal2003},~\cite{Dev98}.
+An information system, however, cannot appeal to a similar interpretation functionality.
+Thus, many proposals have been concerned with the combination of time and imperfections in the context of information systems~\cite{nagypal2003}.
+In this section, some main concepts and issues concerning this combination are presented.
 
-Humans usually deals with time in a imprecise way.
-e.g.
-`A few days ago we received this package'.
-This issue has been studied in the field of Artificial Intelligence ~\cite{Tre97} and language understanding~\cite{DeCaluwe:1997:FTI:285506.285516}.
-Some proposals ~\cite{knight1993},~\cite{Cru97},~\cite{nagypal2003} conclude that the best representation for incomplete temporal knowledge are time intervals instead of time points.
-This means that, as Allen proprosed in \cite{Allen83} the primitive units (the chronons) in the temporal system are intervals.
+%\subsection{\label{subsec:classification}Classification of imperfect temporal information}
 
-When modelling temporal knowledge, the following types of information might be found~\cite{nagypal2003}:
-
-\begin{description}
-\item[\emph{Uncertainty.}] The temporal specification of the event is uncertain. This is usually found in historical documents when some documents state contradictory facts about some event.
-
-\item[\emph{Vagueness.}] Some of the temporal specifications are defined in a vague way.
-e.g.
-`during summer', `late night'.
-Note also that temporal specifications become vague when the granularity is changed ~\cite{Cru97},~\cite{nagypal2003} .
-e.g.
-`The book was written by July 2012' is usefull if we are interested in the month and the year but it turns into a incomplete expression when we try to obtain the exact day.
-It is possible to set the following classification~\cite{Dev98}:
+\subsection{Types of Imperfections in Temporal Modelling}
+Generally, in temporal modelling, a distinction is made between the following types of imperfections~\cite{nagypal2003}.
 
 \begin{itemize}
-\item
-Non-numerical indications. e.g. \emph{often, usually}.
-\item
-Approximative time indications. e.g. \emph{around 10 a.m.}
-\item
-Indications on temporal relations. e.g. \emph{after 12, during summer}.
+	\item \emph{Uncertainty}: Temporal information or data may contain uncertainty.
+This means that the exact temporal value is (partially) unknown, however, generally some knowledge is present anyhow, possibly describing the value~\cite{Dubois:jucs_9_9:fuzziness_and_uncertainty_in},~\cite{nagypal2003},~\cite{Dubois89},~\cite{343607}.
+E.g., the temporal notion described in a sentence like `The Belfry of Bruges was finished on a day somewhere between 01/01/1201 A.D.
+and 31/12/1300 A.D.' contains uncertainty: it is known that the belfry of Bruges was finished on a single day and that this day lays somewhere between 01/01/1201 A.D.
+and 31/12/1300 A.D., but it is not known exactly which day it is.
+	\item \emph{Vagueness}: Temporal information or data might contain inherent vagueness, as a precise instant or time interval may be intended, but the description of it is certainly vague~\cite{schockaert08},~\cite{nagypal2003},~\cite{Dev98}.
+E.g., the temporal notion described in a sentence like `It happened during summer.' contains vagueness, as even the boundaries of the mentioned temporal notion are not clearly expressed.
+	\item \emph{Subjectivity or ambiguity}: Temporal notions might be subject to subjectivity or ambiguity.
+In certain cases, the temporal notion concerns a historical period like `late romanticism' or `the early middle ages' and thus contains subjectivity~\cite{nagypal2003}.
+In other cases, the interpretation of the temporal notion depends on extra factors.
+E.g., consider a person saying to another person `Let's meet each other at six.' The person hearing these words doesn't now if 6 a.m.
+or 6 p.m.
+is intended, though the person saying the words does.
 \end{itemize}
 
-\item[\emph{Subjectivity} or \emph{Ambiguity}.] The specification of a temporal fact may be affected by subjectivity.  It is possible to distinguish between:
+As to the sources of the imperfections in temporal information, most proposals consider no specific source~\cite{5151},~\cite{schockaert08},~\cite{nagypal2003},~\cite{Dubois89},~\cite{343607},~\cite{ohlbach2004},~\cite{Virant199639}.
+Some proposals, however, deal with the imperfections specifically resulting from aspects of language~\cite{Dev98} and other proposals consider transitions between different granularities to be the source of imperfection in temporal information~\cite{Lin97}.
+Therefore, some proposals consider granularity as the base of the temporal model~\cite{Cru97}.
+
+In an information system, temporal information is usually related to facts or events~\cite{Chountas2000}.
+In light of this, a classification of temporal information can be considered, in which the following types of temporal information may be found:
+
 \begin{itemize}
-\item Historical periods~\cite{nagypal2003} like `late romanticism' or `early middle ages'. 
-\item The interpretation of the temporal event~\cite{Dev98} depends on the side of the hearer.
-e.g.
-A speaker says `We will meet each other at six', the hearer has to choose between 6 p.m.
-or 6 a.m.
-The speaker, however knows exactly which one is meant.
+\item \emph{Definite temporal information}:
+Definite temporal information contains information describing a situation in which all time indications associated with some fact are absolute time indications.
+The temporal information is precisely known.
+
+\item \emph{Indefinite temporal information}~\cite{Dey1996}:
+Indefinite temporal information contains information describing a situation in which the time indication associated with some fact has not been fully defined.
+E.g., consider an event that in fact occured but it is not known exactly when.
+
+\item \emph{Infinite temporal uncertain information}~\cite{Kabanza1990}:
+Infinite temporal uncertain information contains information describing a situation in which an infinite number of time indications are associated with some fact.
+This is usually found in recurrent events like meetings.
+E.g., consider meetings that take place every Wednesday at noon.
+Some systems (usually with different information providers) may dispute the occurence and/or the duration of a fact.
 \end{itemize}
-\end{description}
 
-Another important task in the modelling of incomplete knowledge is when dealing with transitions among the different granularities.
+\subsection{\label{subsec:representation}Representation of Imperfect Information}
+As mentioned before, information systems may have to deal with time indications which contain vagueness.
+Even for some specific events or facts, the temporal indications may become imprecise.
+Therefore, a time point might be specified by means of a time interval of which the boundaries may not be precisely known.
+An example.
+
+\begin{example}
+Consider a speaker and a hearer. The speaker wants to make an appointment with the hearer. Now, consider the speaker saysing: \\
+\begin{center}
+\emph{`We will meet each other tomorrow around 10'}\\
+\end{center}
+The hearer will now usually instinctively agree that the appointment will be in e.g. the time interval between 9.55h and 10.05h.
+\end{example}
+
+The study of the semantics of `around' in temporal~\cite{Dev98} indications has shown that the size of the time interval associated with the imprecise specification of a time point depends on the distance with respect to the current time.
+E.g., consider now that the speaker is talking about something that happened \emph{`during last week'}, then the hearer would consider a time interval of more or less 10 days.
+
+Some proposals~\cite{knight1993},~\cite{Cru97},~\cite{nagypal2003},~\cite{Chountas2000} conclude that the best representation for incomplete temporal knowledge is therefore based on time intervals, even if they refer to a fact that happen at a time point.
+This means that, as Allen proposed in~\cite{Allen83}, the primitive units (the chronons) in a time domain, used in an information system should be intervals.
+
+In order to represent and manage uncertain temporal information properly, several theoretical frameworks have been proposed:
+
+\begin{itemize}
+\item \emph{Probability theory}:
+Probability theory~\cite{Dey1996},~\cite{Lakshmanan1997},~\cite{Haddawy1996} is usually employed when uncertainty concerning a time interval allows a probability to be associated to the time interval.
+The use of probability theory is very usual in logistics information systems.
+E.g., \emph{`The package will arrive at its destination on monday morning with a probability of 0.8'}.
+
+\item \emph{Possibility theory}:
+Using possibility theory~\cite{Dubois:Prade:1988:PossibilityTheory}, a possibility degree is associated to the temporal fact or event.
+Possibility theory is widely used to model uncertainty and vaguenes in time~\cite{Dubois:jucs_9_9:fuzziness_and_uncertainty_in},\cite{Dubois89},\cite{devos94},\cite{nagypal2003}.
+Several works~\cite{schockaert08},~\cite{ohlbach2004} present fuzzy versions of the temporal relations proposed by Allen~\cite{Allen83}.
+The aim of these works is generally to obtain a flexible way to compare uncertain, ill-known temporal intervals by means of temporal relationships.
+
+\item \emph{Rough sets}:
+Rough set theory~\cite{Pawlak1995} has been used to represent uncertainty in time intervals.
+The two dimensional representation of time intervals and the temporal relationships between them has been studied in~\cite{Qia09}.
+\end{itemize}
+
+\subsection{Imperfections in Temporal Relationships}
+As the existence of temporal relationships allows to compare temporal notions, many approaches have been concerned with finding similar temporal relationships, able to support imperfections in the temporal information which is described by temporal notions or even by the temporal relationships themselves~\cite{ohlbach2004},~\cite{nagypal2003},~\cite{schockaert08},~\cite{Dubois:jucs_9_9:fuzziness_and_uncertainty_in}.
+These approaches are often based on Allen's operators~\cite{Allen83}.
+Example \ref{ex:fuzzy-allen-relation} presents a short example concerning one of Allen's relationships.
+
+\begin{example}
+\label{ex:fuzzy-allen-relation}
+Consider an event which takes place between time points $A$ and $B$.
+Thus, the event comprises time interval $[A,B]$ (this is visualized in part (1) of figure \ref{fig:fuzzy-allen-relationship}).
+The classical Allen relationship `after' returns an interval $]B,\infty[$ as shown in part (2) of figure \ref{fig:fuzzy-allen-relationship}.
+A fuzzified version of Allen's `after' operator is illustrated in part (3) of figure \ref{fig:fuzzy-allen-relationship}.
+The comparison between two time intervals results in a possibility degree in the unit interval.
+The shape of the possibility distribution is shown in part (3) of figure \ref{fig:fuzzy-allen-relationship}.
+Note that all the points strictly after the point $B$ results in a possibility degree of 1 whereas there is an area near the point $B$ in which the possibility degree runs smoothly between 0 and 1.
+
+Consider now the interval given by $[C,D]$, illustrated in part (4) of figure \ref{fig:fuzzy-allen-relationship}.
+The user wants to know if $[C,D]$ is after $[A,B]$.
+The crisp version of the `after' operator would return \emph{`no'} as an answer.
+The fuzzy version for the same operator would return \emph{`yes, with a possibility of 0.5'}.
+
+\begin{figure}
+\centering
+\includegraphics[scale=0.5]{graphs/fuzzyAllen.eps}
+\caption{Example for the Allen relationship \emph{`after'}.
+(1) The event bounded within time points $A$ and $B$.
+(2) The crisp version of the \emph{`after'} operator.
+(3) A fuzzy version of the after operator.
+(4) Another event, bounded within time points $[C,D]$.}
+\label{fig:fuzzy-allen-relationship}
+\end{figure}
+\end{example}
 

### 2012-03-16 trunk/book/sources/temporalDatabases.tex
To my opinion, the content is now perfect.

@@
 Teradata~\cite{teradata2011} is mainly a business intelligence system designed for data mining.
-Secondo~\cite{Dieker2000} is an extensible database system in which the core of the database may be replaced by a customized algebras.
-It is designed for non-standard applications and it support both valid and transaction times.
+Secondo~\cite{Dieker2000} is an extensible database system in which the core of the database may be replaced by a customized algebra.
+It is designed for non-standard applications and it supports both valid and transaction times.
 

### 2012-03-16 book/sources/fuzzyTemporalDatabases.tex
To my opinion, the content is now perfect.

@@
 As possibility theory is usually seen as a theory of confidence, aimed at dealing with uncertainty, in some proposals, possibility theory is used specifically to handle uncertainty in temporal information.
-In fuzzy databases~\cite{Galindo2006}, uncertainty is usually expected to appear at storage level (in the database), whereas other types of imperfection, notably imprecision, are usually expected to appear in querying.
+In fuzzy databases~\cite{Galindo2006}, uncertainty is usually expected to appear in the database content, whereas other types of imperfection, notably imprecision, are usually expected to appear in querying.
 
@@
 
-In fact, to model the uncertainty related to a valid-time interval using possibility theory, the presented proposal relies on the concept of ill-known sets~\cite{Dubois88b}.
+In fact, to model the uncertainty related to a valid-time interval using possibility theory, the presented proposal introduces so-called ill-known time intervals, relying on the concept of ill-known sets~\cite{Dubois88b}.
 
@@
 
-The presented proposal relies on the concept of ill-known sets~\cite{Dubois88b}.
+To represent valid time indications which might contain uncertainty, the presented proposal introduces the concept of ill-known valid time intervals, which relies on the concept of ill-known sets~\cite{Dubois88b}.
 To correctly explain the concept of ill-known sets, the concept of possibilistic variables should be introduced first.
 In the presented proposal, the definition of possibilistic variables of~\cite{Pon11} is followed.
-A \emph{possibilistic variable} is defined as follows~\cite{Pon11}.
+In~\cite{Pon11}, a \emph{possibilistic variable} is defined as follows.
 
@@
 
-BASE ON IPMU 2
-
-Consider a set $R$, which contains single values (and not collections of values).
+Now, consider a set $R$, which contains single values (and not collections of values).
 When a possibilistic variable $X_{v}$ is defined on such a set $R$, the unique value $X_{v}$ takes, which is (partially) unknown, is called an \emph{ill-known value} in this work \cite{Dubois88b}.
 
-%The exact value a possibilistic variable takes, which is (partially) unknown, is called an \emph{ill-known value} in this work \cite{Dubois88b}.
+When a possibilistic variable is defined on the powerset $\Pow(R)$ of some universe $R$, the unique value the variable takes will be a crisp set and its possibility distribution on the powerset $\Pow(R)$ will describe the possibility of each crisp subset of $R$ to be the value the variable takes.
+This value (a crisp set) the variable takes, which is (partially) unknown, is now called an \emph{ill-known set} \cite{Dubois88b}.
 
-When a possibilistic variable is defined on the powerset $\Pow(R)$ of some universe $R$, the unique value the variable takes will be a crisp set and its possibility distribution on the powerset $\Pow(R)$ will describe the possibility of each crisp subset of $R$ to be the value the variable takes.
-This exact value (a crisp set) the variable takes, is now called an \emph{ill-known set} \cite{Dubois88b}.
+Finally, consider a set $R$, which contains single values (and not collections of values) and its powerset $\Pow(R)$.
+Now consider a subset $\Pow_{I}(R)$ of $\Pow(R)$ and let this subset contain every element of $\Pow(R)$ that is an interval, but no other elements.
+When a possibilistic variable $X_{i}$ is defined on the subset $\Pow_{I}(R)$ of the powerset $\Pow(R)$ of some set $R$, the unique value $X_{i}$ takes will be a crisp interval and the possibility distribution $\pi_{X_{i}}$ of $X_{i}$ will be a possibility distribution on $\Pow_{I}(R)$.
+This $\pi_{X_{i}}$ will define the possibility of each value of $\Pow_{I}(R)$ (a value of $\Pow_{I}(R)$ is a crisp interval which is a subset of $R$) being the value $X_{i}$ takes.
+This exact value (a crisp interval) the variable takes, which is (partially) unknown, is called an \emph{ill-known interval} here.
 
-This work will deal with ill-known time intervals.
-These are ill-known intervals in time.
-Ill-known intervals are ill-known sets, defined and represented via a start and end point, which will be ill-known values.
-The elements of the set are the values between the start and end point.
-A closed ill-known interval with start point defined by possibilistic variable $X$ and end point by possibilistic variable $Y$ is noted here $\left[X, Y\right]$.
-The correspondences and transitions between the representations of ill-known sets, between the representations of ill-known intervals and between the representations of an ill-known set and an ill-known interval are part of the authors current research.
+The presented proposal will deal with ill-known time intervals.
+Ill-known time intervals are ill-known intervals in a time domain.
+In the presented proposal, an ill-known time interval will be defined and represented via its start and end point, which will be ill-known values.
+The elements of the ill-known time interval are the values between its start and end point, including the start and end points themselves\footnote{The presented proposal only deals with closed ill-known time intervals.
+Dealing with halfopen or open ill-known intervals is part of the current research of the authors.}.
+It should be clear that this approach to representing ill-known time intervals differs from the approach based on a single possibility distribution on a set $\Pow_{I}(R)$ of a set $R$.
+These approaches have a different behavior and can be used to describe different ill-known time intervals.
+The correspondences, interactions and transitions between these different representations of ill-known intervals and their interpretations are part of the authors current research.
+
+In the presented proposal, a closed ill-known time interval with start point defined by possibilistic variable $X$ and end point by possibilistic variable $Y$ is noted $\left[X, Y\right]$.
+Figure \ref{fig:interval} shows a closed ill-known time interval.
+%The correspondences and transitions between the representations of ill-known sets, between the representations of ill-known intervals and between the representations of an ill-known set and an ill-known interval are part of the authors current research.
 Figure \ref{fig:interval} illustrates this.
-
-%Uncertainty in a valid-time interval may be modelled with uncertainty at one or both starting and ending points.
-The uncertainty is usually represented by means of a possibility distribution in one or both points.
-It is also possible to represent with rough sets~\cite{Pawlak1995} time intervals~\cite{Qia09}.
-
-Several authors ~\cite{garrido2009} propose transformations in order to optimize the storage of such ill-known valid-time intervals, though recent research might seem to indicate some minor issues with respect to some of these transformations ~\cite{Pon11}.
-A comparison between the transformations in ~\cite{garrido2009} and the framework in~\cite{Pon11} is presented in~\cite{pon12}.
-Figure \ref{fig:convexhull} illustrates a transformation based on the `convex hull' approach REF.
-
-%An ill-known time interval is usually represented by means of two possibility distributions one in each starting and ending points.
-Figure \ref{fig:interval} shows an ill-known valid-time interval.
-The starting and ending points are denoted by two fuzzy numbers, $X$ and $Y$, represented by means of two triangular possibility distributions (see the Appendix).
-For illustration, figure \ref{fig:convexhull} shows one of the transformations based in the convex hull and proposed in~\cite{garrido2009}.
 
@@
 
+%Uncertainty in a valid-time interval may be modelled with uncertainty at one or both starting and ending points.
+The uncertainty is usually represented by means of a possibility distribution in one or both points.
+It is also possible to represent with rough sets~\cite{Pawlak1995} time intervals~\cite{Qia09}.
+
+Several authors work with concepts very similar to these ill-known time intervals and some of them~\cite{garrido2009} propose transformations of the describing functions in order to optimize the storage of such ill-known valid time intervals, though recent research might seem to indicate some minor issues with respect to some of these transformations~\cite{Pon11}.
+A comparison between the transformations in~\cite{garrido2009} and the framework in~\cite{Pon11} is presented in~\cite{pon12}.
+Figure \ref{fig:convexhull} illustrates a transformation based on the `convex hull' approach~\cite{garrido2009}.
+
+%An ill-known time interval is usually represented by means of two possibility distributions one in each starting and ending points.
+Figure \ref{fig:interval} shows an ill-known valid-time interval.
+The starting and ending points are denoted by two fuzzy numbers, $X$ and $Y$, represented by means of two triangular possibility distributions (see the Appendix).
+For illustration, figure \ref{fig:convexhull} shows one of the transformations based in the convex hull and proposed in~\cite{garrido2009}.
 \begin{figure}
@@
 
-In section \ref{subsec:querying-time-intervals}, a technique allowing the querying of fuzzy valid-time databases which contain ill-known valid-time intervals is presented and briefly explained.
+%In section \ref{subsec:querying-time-intervals}, a technique allowing the querying of fuzzy valid-time databases which contain ill-known valid-time intervals is presented and briefly explained.
 The technique relies on the framework for set evaluation presented in \cite{Pon11}.
 
-\subsubsection{Ill-known Time Constraints}
+In the presented proposal, ill-known valid time intervals will be used to represent valid time indications in a valid-time relation and to model the uncertainty these may contain.
 
-\subsubsection{Querying Fuzzy Valid-Time Databases with Ill-Known Valid-Time Intervals}
+To evaluate the temporal demands in queries issued by users to query a valid-time relation containing ill-known valid time intervals, the presented proposal introduces a technique based on the concept of ill-known time constraints, which is based on the concept of ill-known constraints as presented in~\cite{Pon11}.
+Both concepts are treated in subsection \ref{iktc}.
+
+Before ill-known time constraints can be introduced, another notion related to possibilistic variables shoud be paid attention to.
+In fact, a specific application of possibilistic variables is obtained when the set under consideration is the set of boolean values, denoted $\mathbb{B}$ = $\{T,F\}$, where $T$ denotes `true' and $F$ denotes `false' \cite{Pon11}.
+Indeed, any boolean proposition $p$ takes exactly one value in $\mathbb{B}$.
+If the knowledge about which value this proposition $p$ takes, is given by a possibility distribution $\pi_p$, then proposition $p$ can be seen as a possibilistic variable.
+In the presented proposal, the interest lies with the case where the proposition holds (denoted $p$ = $T$) and the possibility and necessity that $p$ = $T$ demand most attention.
+In the following sections, the following notations are used, based on previous notations:
+
+\begin{align}
+\text{Possibility that $p$ = $T$:} \hspace{50pt} & Pos(p) = \pi_p(T) \\
+\text{Necessity that $p$ = $T$:} \hspace{50pt} & Nec(p) = 1-\pi_p(F) 
+\end{align}
+
+\subsubsection{\label{iktc}Ill-known Time Constraints}
+The presented proposal contains a technique for evaluating user queries used to query a valid time relation in which the valid time indications are represented by ill-known time intervals.
+Part of this query evaluation technique relies on the concept of ill-known time constraints, which is based on the concept of ill-known constraints as presented in~\cite{Pon11}.
+These concepts are presented below.
+Following~\cite{Pon11}, an \emph{ill-known constraint} is defined as follows.
+
+\begin{svgraybox}
+\vspace{-10pt}
+\begin{definition}
+Given a universe $U$, an \emph{ill-known constraint} $C$ is specified by means of a binary relation $R \subseteq U^{2}$ and a fixed, ill-known value defined by its possibilistic variable $V$ on $U$, i.e.:
+\begin{align}
+C \triangleq (V,R) \nonumber
+\end{align}
+A set $A \subseteq U$ now satisfies this constraint $C$ if and only if:
+\begin{align}
+\forall a \in A : (V,a) \in R \nonumber
+\end{align}
+\end{definition}
+\vspace{-10pt}
+\end{svgraybox}
+
+An example of an ill-known constraint is given by:
+\begin{align}
+C_{<} \triangleq (X, <) \nonumber
+\end{align}
+Some set $A$ then satisfies $C_{<}$ if 
+\begin{align}
+\forall a \in A : X <a \nonumber
+\end{align}
+
+The satisfaction of a constraint $C \triangleq (V,R)$ by a set $A$ is basically a Boolean matter (either the set satisfies the constraint or not) and can thus be seen as a boolean proposition, but due to the uncertainty inherent to the ill-known value $V$, it can be uncertain whether $C$ is satisfied by $A$ or not \cite{Pon11}.
+Based on the possibility distribution $\pi_{V}$ of $V$, the possibility and necessity that $A$ satisfies $C$ can be found.
+This proposition can thus be seen as a possibilistic variable on $\mathbb{B}$.
+The required possibility and necessity are calculated using the following formulas~\cite{Pon11}.
+
+\begin{align}
+Pos(A\text{ satisfies }C) & = \min_{a \in A}\left(\sup_{(w,a) \in R}\pi_{V}(w))\right) \label{eq:pos}\\
+Nec(A\text{ satisfies }C) & = \min_{a \in A}\left(\inf_{(w,a) \notin R} 1-\pi_{V}(w)\right) \label{eq:nec}
+\end{align}
+
+Now, to calculate the possibility or necessity of a set $A$ satisfying multiple constraints, the $\min$ t-norm operator is used to express a conjunction of constraints. For example:
+
+\begin{align}
+Pos((A\text{ satisfies }C_{1}) \text{ and } (A\text{ satisfies }C_{2})) & = \nonumber\\
+ \min_{a \in A}(Pos(A\text{ satisfies }C_{1}), Pos(A\text{ satisfies }C_{2})) \nonumber \\
+Nec((A\text{ satisfies }C_{1}) \text{ and } (A\text{ satisfies }C_{2})) & = \nonumber\\
+ \min_{a \in A}(Nec(A\text{ satisfies }C_{1}), Nec(A\text{ satisfies }C_{2})) \nonumber
+\end{align}
+
+Accordingly, the $\max$ t-conorm operator is used to express a disjunction of constraints. For example:
+
+\begin{align}
+Pos((A\text{ satisfies }C_{1}) \text{ or } (A\text{ satisfies }C_{2})) & = \nonumber\\
+ \max_{a \in A}(Pos(A\text{ satisfies }C_{1}), Pos(A\text{ satisfies }C_{2})) \nonumber \\
+Nec((A\text{ satisfies }C_{1}) \text{ or } (A\text{ satisfies }C_{2})) & = \nonumber\\
+ \max_{a \in A}(Nec(A\text{ satisfies }C_{1}), Nec(A\text{ satisfies }C_{2})) \nonumber
+\end{align}
+
+%Negation?
+
+In the presented proposal, \emph{ill-known time constraints} are considered, which are ill-known constraints of which the considered universe is a time domain.
+
+In the next subsection, the core of the presented proposal is described.
+
+\subsubsection{Querying Valid-time Relations containing Ill-known Valid-time Intervals}
 \label{subsec:querying-time-intervals}
 
-One of the main purposes of the existence of databases is to allow information retrieval.
-The standard query language for databases is SQL ~\cite{Mel93}.
-Several proposals to extend the SQL language for transaction-time databases~\cite{Sarda90}, valid-time databases~\cite{gad92} and bitemporal databases ~\cite{TSQL} exist and some authors have studied how to support temporal querying in standard SQL~\cite{Snodgrass98}.
+One of the main purposes of the existence of (relational) databases is to allow information retrieval.
+The standard query language for (relational) databases is SQL~\cite{Mel93}, but several proposals to extend the SQL language for transaction-time databases~\cite{Sarda90}, valid-time databases~\cite{gad92} and bitemporal databases~\cite{TSQL} exist and some authors have studied how to support temporal querying in standard SQL~\cite{Snodgrass98}.
 
-In the following subsections, the query structure, the query evaluation method and the technique to meaningfully rank and aggregate the query results, proposed in this work, are presented.
+As mentioned before, the presented proposal deals with querying a valid time relation.
+In this subsection, the core of the presented proposal is described.
+First the particular structure of the relation is described, along with the nature and structure of its supposed contents.
+Next, the particular query structure is presented.
+Finally, the particular method for evaluating queries and for ranking the result records are presented.
+
+%In the following subsections, the nature and structure of the relation content, the query structure, the query evaluation method and the technique to meaningfully rank and aggregate the query results, proposed in this work, are presented.
 
@@
 
+\paragraph{\textbf{Relation Structure}}
+In the presented proposal, a valid-time relation is considered, in which every record contains just one valid-time indication.
+This valid-time indication is represented by a closed ill-known time interval, to allow uncertainty in the valid-time information.
+As explained above, the ill-known time intervals used here will be defined and represented via their start and end points, which are ill-known values in the valid-time domain.
+
 \paragraph{\textbf{Query Structure}}
-In this work, a query $\tilde{Q}$ is made up of two separate constructs of user query demands.
+In the presented proposal, a query consists of two separate constructs of user demands.
 
@@
 \begin{definition}
-\textbf{Query}
-A query $\tilde{Q}$ in this work is given by:
+\textbf{Query}\\
+A query $\tilde{Q}$ is given by:
 \begin{equation}
@@
 \end{equation}
-Here $Q = \left \lbrace q_1,\ \cdots,\ q_n \right \rbrace$ contains all (possibly fuzzy) non-temporal user-defined query demands.
+
+Here, $Q$ denotes the construct of all (possibly fuzzy) non-temporal user query demands.
+These comprise all constraints and demands unrelated to valid-time and thus unrelated to the valid-time indications in the queried relation.
+$Q^{time}$ denotes the temporal demand specified by the user.
+
+%Here $Q = \left \lbrace q_1,\ \cdots,\ q_n \right \rbrace$ contains all (possibly fuzzy) non-temporal user-defined query demands.
 These comprise all constraints unrelated to the valid-time indications in the queried relations.
@@
 
-In this work, the query structure allows the user to specify a single temporal constraint.
+The presented query structure allows the user to specify a single temporal demand, denoted by $Q^{time}$.
 
@@
 \begin{definition}
-\textbf{Temporal Constraint $Q^{time}$}
-In this work, $Q^{time}$ is defined by:
+\textbf{Temporal demand}\\
+A temporal demand $Q^{time}$ is defined by:
 \begin{equation}
@@
 \end{equation}
-where $I$ denotes a crisp time interval and $AR$ is one of the Allen relations.
+Here, $I$ denotes a crisp time interval (which can be specified in any way required) and $AR$ denotes one of the Allen relations (cf. section~\ref{allen-crisp}).
 \end{definition}
@@
 
-The interpretation is that, for a record with an ill-known valid-time interval $J$, the user demands that $I$ AR $J$ hold.
+The interpretation of such a temporal demand $Q^{time} = \left( I,AR \right)$ is that, for a record with an ill-known valid-time interval $J$, the user demands that $I$ AR $J$ holds.
 
@@
 \label{subsubsec:query-evaluation}
-Query satisfaction in a fuzzy relational database is a matter of degree.
-Typically, the query evaluation results in a \emph{satisfaction degree} $s$, which is typically in the unit interval.
-$s \in \left[ 0,1,\right]$.
-Here, a satisfaction degree of $0$ denotes a total dissatisfaction while a degree of $1$ denotes a total satisfaction.
+Query satisfaction in a fuzzy relational database is usually a matter of degree.
+Typically, the evaluation of the query demands for a record results in a \emph{satisfaction degree} $s$, which is typically in the unit interval, i.e.
+$s \in \left[ 0,1\right]$.
+This satisfaction degree then models the extent to which the record satisfies the query demands.
+As such, a satisfaction degree of $0$ indicates total dissatisfaction while a degree of $1$ indicates total satisfaction and every level of satisfaction between total satisfaction and total dissatisfaction is indicated by a satisfaction degre between $0$ and $1$.
 
@@
 
-In the presented approach, for every record $r$, each part of $\tilde{Q}$ is evaluated independently:
+In the presented approach, for every record $r$, each part of a query $\tilde{Q} = \left( Q^{time}, Q \right)$ is evaluated independently:
 \begin{itemize}
 \item
-The user preferences expressed in the non-temporal part, $Q$, are evaluated, resulting in a satisfaction degree denoted as $e_Q(r)$.
+The user preferences expressed in the non-temporal part $Q$, are evaluated, resulting in a satisfaction degree denoted $e_Q(r)$.
 The presented approach accepts any valid, sound method of calculating this evaluation, as long as the method is well-founded and $e_Q(r) \in \left[0,1\right]$.
 \item
-The evaluation of the temporal demand expressed in the temporal part, $Q^{time}$, depends on $AR$.
-A specific set of ill-known constraints\cite{Pon11} is considered depending on the Allen relation denoted by $AR$.
-The exact set chosen can be found in table \ref{tab:fuzzy-allen-relations}, for every possible value of $AR$.
+The evaluation of the temporal demand expressed in the temporal part, $Q^{time} = \left( I,AR \right)$, depends on $AR$.
+A specific construct of ill-known constraints (cf.
+section~\ref{iktc}) is considered, depending on the Allen relation denoted by $AR$.
+The exact construct of constraints is an instantiation based on the formulas which can be found in table \ref{table:const}, for every possible value of $AR$.
+The form and capacity of these constraints are based on~\cite{Pon11}.
+Then, using equations \eqref{eq:pos} and \eqref{eq:nec}, the exact formulas to calculate the possibility $\Pos_{Q^{time}}(r)$ and the necessity $\Nec_{Q^{time}}(r)$ that record $r$ satisfies this construct of ill-known time constraints are derived from this construct of constraints.
+As mentioned, $\Pos_{Q^{time}}(r)$ and $\Nec_{Q^{time}}(r)$ denote the possibility, respectively the necessity that the considered record $r$ satisfies the construct of constraints corresponding to the temporal demand $Q^{time}$ and thus the possibility, respectively the necessity that $r$ satisfies $Q^{time}$.
+
+%The exact set of constraints is an instantiation based on the formulas which can be found in table \ref{tab:fuzzy-allen-relations}, for every possible value of $AR$.
 Then, using these formulas given in table \ref{tab:fuzzy-allen-relations}, both the possibility degree $\Pos_{Q^{time}}(r)$ and the necessity degree $\Nec_{Q^{time}}(r)$ are computed.
@@
 
-\textcolor{red}{TODO: maybe part on what these and mean? Maybe part on where the formulas in table \ref{tab:fuzzy-allen-relations} come from?}
+%\textcolor{red}{TODO: maybe part on what these pos and nec mean? Maybe part on where the formulas in table \ref{tab:fuzzy-allen-relations} come from?}
 
 \begin{table}[h]
+\caption{Constructs of constraints related to their respective Allen relations, as used in the presented work.
+In this table, the ill-known time interval $J = \left[X, Y\right]$ in a record $r$ has a start point described by possibilistic variable $X$ and an end point described by possibilistic variable $Y$.
+The crisp time interval in the user's temporal demand is denoted $I$.}
 
-\caption{Allen's relations used in the framework.
+\centering
+\begin{tabular}{|c|c|}
+\hline
+Allen Relation & Construct of constraints\\
+\hline
+I before J & $C_1 \triangleq \left(<,X\right)$ \\
+\hline
+I equal J & $\left(C_1 \triangleq \left(\geq,X\right)\right) \wedge \neg \left(C_2 \triangleq \left(\neq,X\right) \right) \wedge \left(C_3 \triangleq \left(\leq,Y\right) \right) \wedge \neg \left(C_4 \triangleq \left(\neq,Y\right)\right)$ \\
+\hline
+I meets J & $\left(C_1 \triangleq \left(\leq,X\right)\right) \wedge \neg \left(C_2 \triangleq \left(\neq,X\right) \right)$ \\
+\hline
+I overlaps J & $\left(C_1 \triangleq \left(<,Y\right)\right) \wedge \neg \left(C_2 \triangleq \left(\leq,X\right) \right) \wedge \neg \left(C_3 \triangleq \left(\geq,X\right) \right)$ \\
+\hline
+I during J & $\left( \left (C_1 \triangleq \left(>,X\right) \right) \wedge \left(C_2 \triangleq \left(\leq,Y\right) \right) \right) \vee \left( \left(C_3 \triangleq \left(\geq,X\right) \right) \wedge  \left(C_4 \triangleq \left(<,Y\right) \right) \right)$ \\
+\hline
+I starts J & $\left(C_1 \triangleq \left(\geq,X\right) \right) \wedge \neg \left(C_2\triangleq \left(\neq,X\right)\right)$ \\
+\hline
+I finishes J & $\left(C_1 \triangleq \left(\leq,Y\right) \right) \wedge \neg \left(C_2\triangleq \left(\neq,Y\right)\right)$ \\
+\hline
+\end{tabular}
+\label{table:const}
+\end{table}
+
+%\begin{table}[h]
+
+%\caption{Allen's relations used in the framework.
 Here, $I = \left[a, b\right]$ denotes a crisp time interval, $J = \left[X, Y\right]$ denotes an ill-known time interval, with $\pi_{X}$ and $\pi_{Y}$ the possibility distributions of $X$ and $Y$ respectively.
@@
 
-\centering
-\begin{tabular}{|c|c|c|}
-\hline
-Allen Relation &  $\Pos\left(\text{I satisfies all constraints }C_{i}, i = 1,2,...\right)$ & $\Nec\left(C_{i}, i = 1,2,... \right)$ \\
-\hline
-I before J & $\sup_{a>w}\pi_X(w)$ & $\inf_{a \leq w} 1- \pi_X(w)$\\
-\hline
+%\centering
+%\begin{tabular}{|c|c|c|}
+%\hline
+%Allen Relation &  $\Pos\left(\text{I satisfies all constraints }C_{i}, i = 1,2,...\right)$ & $\Nec\left(C_{i}, i = 1,2,... \right)$ \\
+%\hline
+%I before J & $\sup_{a>w}\pi_X(w)$ & $\inf_{a \leq w} 1- \pi_X(w)$\\
+%\hline
 
-\multirow{4}{*}
-{I equal J} &  $\min ( \sup_{a \leq w}\pi_X(w),$  & $\min ( \inf_{a>w} 1-\pi_X(w),$\\
-& $\pi_X(w),$ & $\inf_{w \in \left[a,b\right]} \pi_X(w), $ \\
- &  $\sup_{b \geq w}\pi_Y(w),$ & $\inf_{b>w} 1-\pi_Y(w),$\\
- & $\pi_Y(w))$ & $\inf_{w \in \left[a,b\right]} \pi_Y(w) )$ \\
-\hline
+%\multirow{4}{*}
+%{I equal J} &  $\min ( \sup_{a \leq w}\pi_X(w),$  & $\min ( \inf_{a>w} 1-\pi_X(w),$\\
+%& $\pi_X(w),$ & $\inf_{w \in \left[a,b\right]} \pi_X(w), $ \\
+% &  $\sup_{b \geq w}\pi_Y(w),$ & $\inf_{b>w} 1-\pi_Y(w),$\\
+% & $\pi_Y(w))$ & $\inf_{w \in \left[a,b\right]} \pi_Y(w) )$ \\
+%\hline
 
-\multirow{2}{*}
-I meets J  & $\min (\sup_{a\geq w} \pi_X(w),$ & $\min (\inf_{a<w} 1 - \pi_X(w),$\\
-& $\pi_X(w))$  & $ \inf_{w \in \left[a,b\right]} \pi_X(w) )$ \\
-\hline
+%\multirow{2}{*}
+%I meets J  & $\min (\sup_{a\geq w} \pi_X(w),$ & $\min (\inf_{a<w} 1 - \pi_X(w),$\\
+%& $\pi_X(w))$  & $ \inf_{w \in \left[a,b\right]} \pi_X(w) )$ \\
+%\hline
 
-\multirow{3}{*}
-{I overlaps J}  & $\min ( \sup_{b>w}\pi_Y(w), $ & $\min ( \inf_{b \geq w} 1 - \pi_Y(w),$\\
- & $\sup_{a \geq w}\pi_X(w),$& $1- \inf_{a>w}1-\pi_X(w),$\\
- & $\sup_{a \leq w}\pi_X(w))$ & $1- \inf_{a<w}1-\pi_X(w) )$ \\
-\hline
+%\multirow{3}{*}
+%{I overlaps J}  & $\min ( \sup_{b>w}\pi_Y(w), $ & $\min ( \inf_{b \geq w} 1 - \pi_Y(w),$\\
+% & $\sup_{a \geq w}\pi_X(w),$& $1- \inf_{a>w}1-\pi_X(w),$\\
+% & $\sup_{a \leq w}\pi_X(w))$ & $1- \inf_{a<w}1-\pi_X(w) )$ \\
+%\hline
 
-\multirow{4}{*}
-{I during J}  & $\max ( \min ( \sup_{a<w}\pi_X(w),,$ & $\max (\min (\inf_{a \leq w}1-\pi_X(w), $\\
- & $\sup_{b \geq w}\pi_Y(w)) $ & $ \inf_{b>w} 1- \pi_Y(w)),$ \\
- & $\min ( \sup_{a \leq w }\pi_X(w),$ & $\min ( \inf_{a<w}1-\pi_X(w),$\\
- & $\sup_{b>w}\pi_Y(w))$ & $\inf_{b\geq w}1-\pi_Y(w)))$ \\
-\hline
-\multirow{2}{*}
-{I starts J} &  $\min( \sup_{a \leq w}\pi_X(w),$  & $\min (\inf_{a<w}1-\pi_X(w),$\\
-& $\pi_X(w))$ & $inf_{w \in \left[a,b\right]} 1- \pi_X(w))$ \\
-\hline
-\multirow{2}{*}
-{I finishes J} &  $\min ( \sup_{b \geq w} \pi_Y(w),$  & $\min(\inf_{b>w}1-\pi_Y(w),$\\
-& $\pi_Y(w))$ & $\inf_{w \in [a,b]} 1- \pi_Y(w))$ \\
-\hline
+%\multirow{4}{*}
+%{I during J}  & $\max ( \min ( \sup_{a<w}\pi_X(w),,$ & $\max (\min (\inf_{a \leq w}1-\pi_X(w), $\\
+% & $\sup_{b \geq w}\pi_Y(w)) $ & $ \inf_{b>w} 1- \pi_Y(w)),$ \\
+% & $\min ( \sup_{a \leq w }\pi_X(w),$ & $\min ( \inf_{a<w}1-\pi_X(w),$\\
+% & $\sup_{b>w}\pi_Y(w))$ & $\inf_{b\geq w}1-\pi_Y(w)))$ \\
+%\hline
+%\multirow{2}{*}
+%{I starts J} &  $\min( \sup_{a \leq w}\pi_X(w),$  & $\min (\inf_{a<w}1-\pi_X(w),$\\
+%& $\pi_X(w))$ & $inf_{w \in \left[a,b\right]} 1- \pi_X(w))$ \\
+%\hline
+%\multirow{2}{*}
+%{I finishes J} &  $\min ( \sup_{b \geq w} \pi_Y(w),$  & $\min(\inf_{b>w}1-\pi_Y(w),$\\
+%& $\pi_Y(w))$ & $\inf_{w \in [a,b]} 1- \pi_Y(w))$ \\
+%\hline
 
-\end{tabular}
+%\end{tabular}
 %
@@
 %\vspace{-25pt}
-\label{tab:fuzzy-allen-relations}
-\end{table}
+%\label{tab:fuzzy-allen-relations}
+%\end{table}
 
-\subsubsection{Aggregation and Ranking} TO BE ANOTHER PART
-To be able to present the most appropriate results to the user, for every record $r$, an aggregation method is used to aggregate $\Pos_{Q^{time}}(r)$ and $\Nec_{Q^{time}}(r)$ into a temporal record rank $e_{Q^{time}}(r)$ and after this, a convex combination combining $e_{Q^{time}}(r)$ and $e_{Q}(r)$ will provide the final record rank $e_{final}(r)$.
+%Thus, the evaluation of a query $\tilde{Q}$ for a record $r$ thus results in a satisfaction degree related to the non-temporal query demands and a couple of possibility and necessity related to the temporal demand.
+To be able to present the most appropriate results to the user most prominently, these evaluation results are aggregated and combined to result in a record rank.
+
+\paragraph{\textbf{Aggregation and Ranking}} 
+In this subsection, the notations used in the previous subsection are followed.
+To be able to present the most appropriate results to the user most prominently, for every record $r$, an aggregation method is used to aggregate $\Pos_{Q^{time}}(r)$ and $\Nec_{Q^{time}}(r)$ into a temporal record rank $e_{Q^{time}}(r)$ and after this, a convex combination combining $e_{Q^{time}}(r)$ and $e_{Q}(r)$ will provide the final record rank $e_{final}(r)$.
 
@@
 \end{itemize}
+
+In the next subsection, some main concepts and issues concerning bipolarity in the context of temporal databases are presented and discussed.
 

### 2012-03-16 trunk/book/sources/intro.tex
update

--- 2012-01-18 trunk/book/sources/intro.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-The concept of \emph{time} is easy to understand but very complex to define~\cite{klein94}~\cite{Shackle61}. 
-In Information Systems \emph{IS}, several proposals have been concerned with the obtaining of theoretical models that allow the representation of time~\cite{Bolour82}~\cite{Cru97}. 
-More towards temporal databases, a lot of work have been done.
+
+The concept of time itself is very complex to handle and interpret~\cite{klein94},~\cite{Shackle61}, though it is very natural and omnipresent.
+As information systems often attempt the modelling of natural objects, concepts or processes, they often require modelling temporal aspects or concepts.
+Thus, several proposals have been concerned with the obtaining of theoretical models that allow the modelling or representation of time~\cite{Bolour82},~\cite{Cru97}.
+
+A very specific type of information systems are database systems, which are computer systems designed to manage databases.
+A database contains data representing real objects or concepts.
+Each (atomic) part of these data is a result value of a measurement of a property or a description of a property of a real object or concept.
+In reality, some aspects or properties of objects or concepts are time-variant or time-related.
+e.g., the moment of a bank transaction is traditionally a moment in time and thus a time-related notion, the function of an employee in a company can change through recorded history and is thus time-variant.
+A temporal database schema is a database schema that models real objects or concepts with time-related or time-variant properties.
+However, the modelling of temporal aspects has a direct impact on the consistency of the temporal database, because the temporal nature of these aspects imposes extra integrity constraints.
+An example.
+Consider a relation in a relational library database, modelling the presence of books in the library.
+Every physical book is represented by a unique identifier.
+Every record in the relation contains such an identifier, a date on which the corresponding book was loaned and a date on which it was subsequently returned (if it was returned).
+Without further precautions, a library employee could add several records with the same book identifier, different `loaned'-dates and no `returned'-dates.
+This group of records would represent the same physical book being loaned several times on different dates and never returned, which is of course impossible.
+A temporal database model will typically constrain record insertion and prevent similar modelling inconsistencies.
+
+A lot of research concerns temporal database models and their approaches to the modelling of time.
+The first efforts were towards the representation of historical information related to objects represented by records in a database~\cite{Clifford:1985:AHR:971699.318922}.
+Some proposals tried to extend the Entity Relationship Model~\cite{Klopprogge:1983}, without impact on any database standards like SQL~\cite{Sarda:1990:ESH:627277.627409}.
+
+Notably, in 1994, ``A Consensus Glossary of Temporal Database Concepts'' was published~\cite{Dyreson1994}.
+For this publication, 44 temporal database researchers, among which some of the main researchers in this field, cooperated to reach a consensus on the nature and definitions of some of the main temporal database concepts and terminology.
+This glossary was subsequently updated in 1998~\cite{Jensen:Dyreson:1998:TheConsensusGlossary}.
+
+An interesting issue in temporal modelling concerns relationships between temporal notions.
+Notably, Allen~\cite{Allen83} studied temporal relationships between time intervals (and as a special case time points).
+Among others, the querying of temporal databases has greatly profited from these temporal relationships, because they allowed for richer and more complex user-specified temporal query demands, by allowing to express more complex relationships between the temporal notions in the temporal expressions in the query and the temporal indications in the database.
+e.g., in a relation modelling who was department head of an institution during which periods of time, a query like `who were the department heads when Thomas worked for the institution' can be evaluated using similar relationships.
+
+Humans handle temporal information using certain temporal notions like time intervals or time points~\cite{Dyreson1994}, and they often have to deal with imperfections like imprecisions, vaguenesses, uncertainties or inconsistencies possibly contained in the descriptions of these temporal notions.
+Among many others, these possible imperfections in descriptions of temporal notions determine an important issue in temporal modelling.
+e.g., the description of the temporal notion in a sentence like `The Belfry of Bruges was finished on a day somewhere between 01/01/1201 A.D.
+and 31/12/1300 A.D.' contains imperfection because of the uncertainty in the used time-related expression.
+It is known that the building was finished on a single day, but it is not known precisely which day this was.
+
+To allow information systems to cope with these and similar data imperfections, many approaches adopt fuzzy sets~\cite{Zadeh65} for the representation of temporal information~\cite{343607},~\cite{nagypal2003},~\cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy},~\cite{Dubois:jucs_9_9:fuzziness_and_uncertainty_in}.
+The temporal relationships studied by Allen were fuzzified by several authors~\cite{ohlbach2004},~\cite{nagypal2003},~\cite{schockaert08}.
+Garrido et al.
+~\cite{garrido2009} present different temporal operators, defined by a combination of regular fuzzy comparisons.
+Both~\cite{garrido2009} and~\cite{Pon11} deal with uncertainty in temporal expressions concerning time intervals.
+Other approaches, like~\cite{Qia09}, use rough sets~\cite{Pawlak1995} to represent time intervals.
+
+Next to temporal modelling, some attention has been spent on temporal reasoning~\cite{Allen83}.
+Although temporal reasoning is not discussed in this chapter, it should be noted that, among others, Dubois and Prade et al.~\cite{Dubois:jucs_9_9:fuzziness_and_uncertainty_in},~\cite{Dubois89} have dealt with fuzziness and uncertainty in temporal reasoning.
+
+The aim of this chapter is to present and explain some main concepts regarding time in information systems and to present and discuss some issues and techniques concerned with handling data imperfections related to time.
+The rest of this chapter is structured as follows.
+Section \ref{sec:time-domain} presents some basic concepts and terminology about temporal modelling and discusses some of its important aspects and issues, while section~\ref{sec:fuzzy-time-mod} presents some important issues concerning the combination of data imperfections and temporal modelling.
+In section \ref{sec:temporal-databases}, some basic concepts and terminology about temporal databases are presented, followed by an overview of some interesting issues concerning temporal databases and a survey of some commercial temporal database systems.
+Finally, in section \ref{sec:imprecision-uncertainty}, an approach to querying temporal databases containing imperfect temporal information is presented, followed by some conclusions and some suggestions for future work in section \ref{sec:further-research-work}.
+
+%The concept of \emph{time} is easy to understand but very complex to define~\cite{klein94}~\cite{Shackle61}. 
+%In Information Systems \emph{IS}, several proposals have been concerned with the obtaining of theoretical models that allow the representation of time~\cite{Bolour82}~\cite{Cru97}. 
+%More towards temporal databases, a lot of work have been done.
 The first efforts were towards the representation of the historical information related to an object in the database(reference).
@@
 
-The ``Consensus Glossary of Temporal Database concepts'' \cite{Dyreson1994} is the first publication where the main researchers in temporal databases define the main thesaurus of temporal database concepts.
+%The ``Consensus Glossary of Temporal Database concepts'' \cite{Dyreson1994} is the first publication where the main researchers in temporal databases define the main thesaurus of temporal database concepts.
 This glossary defines the main types of time in a temporal database and is the basis of furthers researches ~\cite{Sarda:1990:ESH:627277.627409} \cite{Jensen94thetsql2}.
 
-Another main issue in temporal databases is the querying.
-The user specifies in the query some temporal constraints, e.g.
-\emph{``before this year''},\emph{``after 1995''}.
+%Another main issue in temporal databases is the querying.
+The user specifies in the query some temporal constraints, e.g., \emph{``before this year''},\emph{``after 1995''}.
 In order to solve these temporal comparisons, between the time specified in the query and the temporal data stored in the database, some basic relationships have been defined.
@@
 In~\cite{garrido2009}, some different temporal operators are defined by a combination of regular fuzzy comparisons.
-%e.g.
-the \emph{BEFORE} operator is defined by means of the fuzzy less than (FLT) operator.
+%e.g., the \emph{BEFORE} operator is defined by means of the fuzzy less than (FLT) operator.
 
-To allow information systems to cope with uncertainty, many approaches adopt fuzzy sets~\cite{Zadeh65} for the representation of temporal information~\cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy},\cite{Dubois:jucs_9_9:fuzziness_and_uncertainty_in},~\cite{devos94}.
+%To allow information systems to cope with uncertainty, many approaches adopt fuzzy sets~\cite{Zadeh65} for the representation of temporal information~\cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy},\cite{Dubois:jucs_9_9:fuzziness_and_uncertainty_in},~\cite{devos94}.
 The time points representing the time when the object is valid in the reality might not be precisely known.
@@
 %temporal reasoning and fuzzy temporal reasoning.
-Next to the manage of temporal information is the temporal reasoning~\cite{Allen83}.
-Dubois and Prade ~\cite{Dubois:jucs_9_9:fuzziness_and_uncertainty_in},\cite{Dubois89} deal with fuzzyness and uncertainty in temporal reasoning, but this topic is outside the scope of this work.
+%Next to the manage of temporal information is the temporal reasoning~\cite{Allen83}.
+Dubois and Prade ~\cite{Dubois:jucs_9_9:fuzziness_and_uncertainty_in},\cite{Dubois89} deal with fuzziness and uncertainty in temporal reasoning, but this topic is outside the scope of this work.
 
@@
 
-The work is organized as follows:
+%The work is organized as follows:
 
-In Section \ref{sec:time-domain} an study about the time and its properties is done.
+%In Section \ref{sec:time-domain} an study about the time and its properties is done.
 Section \ref{sec:temporal-databases} is an overview of the main problems when managing the time in a database, the different proposals for solving them and the temporal database proposals.
 The section concludes with a comparison among the different proposals.
-In Section \ref{sec:fuzzy-temporal-databases} we present a sum up with the proposals for handling the imprecision in temporal databases and the main deficiences we have detected.
+In Section \ref{sec:fuzzy-temporal-databases} we present a sum up with the proposals for handling the imprecision in temporal databases and the main deficiencies we have detected.
 Finally, there are several commercial temporal database management systems  \emph{TDMBS} like \cite{oracle2009},\cite{posgree2009},\cite{teradata2011},\cite{timedb2005}.

### 2012-03-16 book/sources/fuzzyTemporalDatabases.tex
update

@@
 %Some software applications deal with uncertainty related to time.
-e.g.
-a logistics company which transports packages from one destination to another.
+e.g., a logistics company which transports packages from one destination to another.
 The time when the package would be at the end point may be estimated but not precisely known.
@@
 \begin{definition}
+\textbf{Possibilistic variable}~\cite{Pon11}\\
 A possibilistic variable $X$ over a universe $U$ is defined as a variable taking exactly one value in $U$, but for which this value is (partially) unknown.
@@
 For example, a person may define the concept of `tall' as `1.80 meters or higher'.
-The negative would then be the opposite: not tall would be `anyhing less than 1.80 meters'.
+The negative would then be the opposite: not tall would be `anything less than 1.80 meters'.
 However, in some cases, positive preferences can not be directly obtained from negative preferences or vice versa.
-E.g.
-when a person prefers to buy a black motorbike, this does not necessarily mean the person would totally reject a very dark blue motorbike.
+e.g., when a person prefers to buy a black motorbike, this does not necessarily mean the person would totally reject a very dark blue motorbike.
 This phenomenon is called \emph{heterogeneous bipolarity}~\cite{Dubois2006},~\cite{Dubois2008}.

### 2012-03-16 book/sources/timeDomain.tex
To my opinion, the content is now perfect.
update

@@
 
-Data models used by information systems (and in specific, temporal database systems) may represent these underlying time domains using \emph{chronons}~\cite{Dyreson1994}, which can informally be discribed as the smallest distinguishable time units that can be used in the system.
+Data models used by information systems (and in specific, temporal database systems) may represent these underlying time domains using \emph{chronons}~\cite{Dyreson1994}, which can informally be described as the smallest distinguishable time units that can be used in the system.
 However, to explain what chronons are, an explanation of some other temporal concepts is necessary.
@@
 In a dense model, the notion exists that between any two instants always lies another.
-Here, intuitively, the set of instants can be seen as isomporphic to the set of rational numbers $\Q$ (when the set of instants is a discrete one) or the set of real numbers $\R$ (when the set of instants is a continuous one).
+Here, intuitively, the set of instants can be seen as isomorphic to the set of rational numbers $\Q$ (when the set of instants is a discrete one) or the set of real numbers $\R$ (when the set of instants is a continuous one).
 In a continuous model, the notion also exists that between any two instants always lies another one, but the set of instants is always seen as continuous and there are no ``gaps'' between successive instants.
 
-Some other necessary concepts are ~\cite{Dyreson1994}:
+Some other necessary concepts are:
 
@@
 \vspace{-10pt}
-\begin{definition}\textbf{Time Interval}\\
+\begin{definition}\textbf{Time Interval}~\cite{Dyreson1994}\\
 A \emph{\textbf{time interval}} is the time between two instants.
@@
 
-\begin{definition}\textbf{Duration}\\
+\begin{definition}\textbf{Duration}~\cite{Dyreson1994}\\
 A \emph{\textbf{duration}} is an amount of time with known length, but no specific starting or ending instants.
@@
 \vspace{-10pt}
-\begin{definition}\textbf{Temporal Element}\\
+\begin{definition}\textbf{Temporal Element}~\cite{Dyreson1994}\\
 A \emph{\textbf{temporal element}} is a finite union of time intervals.
@@
 
-\begin{definition}\textbf{Event}\\
+\begin{definition}\textbf{Event}~\cite{Dyreson1994}\\
 An \emph{\textbf{event}} is an instantaneous fact, i.e. something occurring at an instant.
@@
 
-\begin{definition}\textbf{`Temporal' as Modifier}\\
+\begin{definition}\textbf{`Temporal' as Modifier}~\cite{Dyreson1994}\\
 The modifier \emph{\textbf{`temporal'}} is used to indicate that the modified concept concerns some aspect of time.
@@
 \vspace{-10pt}
-\begin{definition}\textbf{Chronon}\\
+\begin{definition}\textbf{Chronon}~\cite{Dyreson1994}\\
 In a data model, a \emph{\textbf{chronon}} is a non-decomposable time interval of some fixed, minimal duration.
@@
 
-Another classification of time models concerns the use of points or intervals to model time.
-The equivalence between interval-based and point-based time models is demonstrated in \cite{Böhlen_point-versus}.
+Another classification of time models concerns the use of points or intervals to model time. The equivalence between interval-based and point-based time models is demonstrated in \cite{Bohlen1998}.
 
@@
 Relative time has a direction, differently from distance.
-It is also possible to use negative references so as to represent relative time (e.g.
--3 days = three days ago).
+It is also possible to use negative references so as to represent relative time (e.g., -3 days = three days ago).
 %\textcolor{red}{END MOVE}
@@
 
-\begin{definition}\textbf{Granule}\\
+\begin{definition}\textbf{Granule}~\cite{wang93}\\
 A \textbf{\emph{granule}} is the basic time unit in a granularity.
@@
 \vspace{-10pt}
-\begin{definition}\textbf{Finner Than}\\
+\begin{definition}\textbf{Finner Than}~\cite{Lin97}\\
 Consider a mapping function $f$ and let $i_G$ and $j_H$ be elements of granularities $G$ and $H$ respectively. Granularity $G$ is now said to be \emph{\textbf{finner than}} granularity $H$ if:
@@
 \begin{definition}
-\label{def:coarser-than}\textbf{Coarser Than}
+\label{def:coarser-than}\textbf{Coarser Than}~\cite{Lin97}\\
 Consider a mapping function $f$ and let $i_G$ and $j_H$ be elements of granularities $G$ and $H$ respectively. Granularity $G$ is now said to be \emph{\textbf{coarser than}} granularity $H$ if:
@@
 \emph{Regular mapping}: A regular mapping is a granularity-to-granularity mapping where the mapping function value is calculated by means of multiplications and/or divisions and (maybe) an anchor adjustment.
-E.g., the mapping value of the mapping between hours and minutes is calculated using a multiplication by $60$.
+e.g., the mapping value of the mapping between hours and minutes is calculated using a multiplication by $60$.
 \item
 \emph{Irregular mapping}: An irregular mapping is a granularity-to-granularity mapping where the mapping function value can not be calculated by means of multiplications and/or divisions.
-E.g., the mapping value of the mapping between months and days is dependent on the exact month or day.
+e.g., the mapping value of the mapping between months and days is dependent on the exact month or day.
 \item
 \emph{Congruent mapping}: A congruent mapping is a granularity-to-granularity mapping where the two granularities involved in the mapping have the same granules but a different anchor.
-E.g., the mapping between the days (Gregorian calendar days) and the academic days is a congruent mapping.
+e.g., the mapping between the days (Gregorian calendar days) and the academic days is a congruent mapping.
 \end{itemize}
@@
 More precisely, calendars determine the mapping between human-meaningful time values and an underlying time model~\cite{Dyreson1994}.
-More formally, a calendar is an organization of different granularities (e.g.
-granularities day, month and year) in a hierarchy.
+More formally, a calendar is an organization of different granularities (e.g., granularities day, month and year) in a hierarchy.
 In order to formally define what a calendar is, some other concepts should first be introduced~\cite{Kraus1997}:
@@
 
-\subsection{Temporal Relationships}
+\subsection{\label{allen-crisp}Temporal Relationships}
 %an explanation on allen's temporal relations.

### 2012-03-16 book/sources/prelim.tex
update
update from the task list.

--- 2012-01-23 trunk/book/sources/prelim.tex
@@
 \end{itemize}
-Note that the values $a$ and $b$ are values in the underlying ordered domain. E.g. $(a,b) \in \mathbb{R}^2$. The notation for a triangular function adopted here is $[D,a,b]$.
+Note that the values $a$ and $b$ are values in the underlying ordered domain. e.g., $(a,b) \in \mathbb{R}^2$. The notation for a triangular function adopted here is $[D,a,b]$.
 \begin{figure}[h!]
@@
 Relative time has a direction, differently from distance.
-It is also possible to use negative references so as to represent relative time (e.g.
--3 days = three days ago).
+It is also possible to use negative references so as to represent relative time (e.g., -3 days = three days ago).
 
@@
 We can classify the time recorded in the database through its semantics.
-As a result, two kinds of time derive from two orthogonal time domains: valid time and transaction time.
+As a result, two kinds of time derive from two orthogonal time domains: valid-time and transaction-time.
 Valid-time captures the time-varying nature of the reality model.
@@
 
-The domain of valid time is given as $D_{VT}=\lbrace{c_1^v,c_2^v,...,c_k^v}\rbrace$, whereas the domain of transaction time may be given by $D_{TT}=\lbrace{c_1^t,c_2^t,...,c_k^t}\rbrace$.
+The domain of valid-time is given as $D_{VT}=\lbrace{c_1^v,c_2^v,...,c_k^v}\rbrace$, whereas the domain of transaction-time may be given by $D_{TT}=\lbrace{c_1^t,c_2^t,...,c_k^t}\rbrace$.
 Thus, a valid-time chronon $c^v$  is a member of $D_{VT}$, a transaction-time chronon $c^t$ is a member of $D_{TT}$ and a bitemporal chronon $c^b=(c^t,c^v)$ is an ordered pair of a transaction-time chronon and valid-time chronon.
@@
 We say that a fact is current in the relation during each transaction-time chronon when a subset of the transaction-time chronon domain is associated with the valid-time chronon.
-Only the subset of transaction times less than UC may be associated with a valid time.
+Only the subset of transaction-times less than UC may be associated with a valid-time.
 
-While the definition of a bitemporal chronon is symmetric, the relation between valid time and transaction time is asymmetric.
+While the definition of a bitemporal chronon is symmetric, the relation between valid-time and transaction-time is asymmetric.
 It should also be noted that since no two tuples with identical attribute values are allowed in a bitemporal relation instance, the full history of a fact is contained in a single tuple.
@@
 \item
-Valid-time relations, which support only valid time.
+Valid-time relations, which support only valid-time.
 \item
-Transaction-time relations,which support only transaction time.
+Transaction-time relations,which support only transaction-time.
 \end{itemize}
-We use the term 'snapshot relation' for a conventional relation. This relation supports neither transaction nor valid time.
+We use the term 'snapshot relation' for a conventional relation. This relation supports neither transaction nor valid-time.
 \subsubsection{Implicit versus explicit timestamps}
@@
 
-	If the time model works with transaction time, it means that the update language contains implicit timestamps.
+	If the time model works with transaction-time, it means that the update language contains implicit timestamps.
 The transaction-time chronons of the facts are supplied by the system; in the valid-time model, these chronons are supplied by the user.
@@
 A more detailed summary can be found in \ref{sec:timeindatabasedesign}.
-Some models work only with valid or transaction time, whereas a few models can handle both.
-The current study starts with the models supporting valid time only, continues through the models which support transaction time and finally deals with the bitemporal models.
+Some models work only with valid or transaction-time, whereas a few models can handle both.
+The current study starts with the models supporting valid-time only, continues through the models which support transaction-time and finally deals with the bitemporal models.
 
 \paragraph{Valid-time models}
-Most models support valid time only.
+Most models support valid-time only.
 
@@
 Jones LEGOL 2.0: \cite{legol20} A language used in legislative writing.
-The temporal order of elements and the valid times for objects are important.
+The temporal order of elements and the valid-times for objects are important.
 It was the first time-oriented algebra to be defined, and some of its features are found in subsequent algebras.
@@
 
-Ariav: \cite{Ariav:1986:TOD:7239.7350} Temporally-oriented data model: a valid-time relation is a set of snapshot relation states indexed by valid time.
+Ariav: \cite{Ariav:1986:TOD:7239.7350} Temporally-oriented data model: a valid-time relation is a set of snapshot relation states indexed by valid-time.
 A calculus-based query language is associated with the model: TOSQL.
@@
 Objects can either be snapshot or valid-time relations.
-In this model, valid time is considered to be closed on the left part of the interval and open on the right.
+In this model, valid-time is considered to be closed on the left part of the interval and open on the right.
 Time is represented by an implicit attribute called 'Period' as a single non-atomic value.
@@
 Unlike intervals, valid-time elements are closed under union, difference, and complementation.
-Bhargava presents an extension of this model to both valid and transaction time.
+Bhargava presents an extension of this model to both valid and transaction-time.
 
 Gadia-2: \cite{Gadia:1986:TMM:645471.655410},\cite{Chuen},\cite{Gadia:1988:GMR:971701.50233} Is an extension of the homogeneous model known as multihomogeneous model.
-Temporal elements may be multi-dimensional to model both valid and transaction time.
+Temporal elements may be multi-dimensional to model both valid and transaction-time.
 Attribute values are functions from temporal elements onto attribute value domains, but attribute values do not need to be on the same temporal element.
@@
 \paragraph{Bitemporal data models}
-The bitemporal data models support both valid and transaction time. This section deals with the following seven models of a bitemporal data model.
+The bitemporal data models support both valid and transaction-time. This section deals with the following seven models of a bitemporal data model.
 
@@
 Ahn: \cite{Snodgrass:1985:TTD:318898.318921},\cite{sno86} This model follows a four-dimensional data model.
-Each instance is represented as a sequence where each value is composed by a timestamp with a transaction-time chronon and a three dimensional volumes -one of these dimensions being the valid time.
+Each instance is represented as a sequence where each value is composed by a timestamp with a transaction-time chronon and a three dimensional volumes -one of these dimensions being the valid-time.
 
 Snodgrass:  \cite{sno86},\cite{Snodgrass:1984:TQL:588011.588041} The data model has four implicit attributes: 
-The transaction time when the tuple was inserted.
-The transaction time when the tuple was deleted.
+The transaction-time when the tuple was inserted.
+The transaction-time when the tuple was deleted.
 The starting point of the valid-time interval.
-The end point of the valid time-interval.
+The end point of the valid-time-interval.
 
@@
 Thus, transaction-time relations are modeled as sequences of snapshot relations.
-Bitemporal relations are sequences of valid-time relations indexed by transaction time.
+Bitemporal relations are sequences of valid-time relations indexed by transaction-time.
 An attribute's value in a valid-time relation is an ordered pair: a value from the attribute's domain and a set of chronons.
@@
 Thompson: \cite{171833} This model, based on an accounting system, is called Accounting Data Model.
-It has four timestamp attributes: two for transaction time, called 'engineering time', and two for valid time, called 'accounting time'.
+It has four timestamp attributes: two for transaction-time, called 'engineering time', and two for valid-time, called 'accounting time'.
 Also, a boolean value called 'timewarp' enables the attribute to change historically.

### 2012-03-16 book/sources/proposal.tex
update from the task list.

@@
 %
-The model deals with uncertainty in the valid time.
-A valid time period  \emph{VT} is specified with two points: a starting time point \emph{St} and an ending time point \emph{Et}.
-Thus, a fuzzy validity period, \emph{FVP} is a valid time period in which the starting, the ending or both points are not precisely known.
+The model deals with uncertainty in the valid-time.
+A valid-time period  \emph{VT} is specified with two points: a starting time point \emph{St} and an ending time point \emph{Et}.
+Thus, a fuzzy validity period, \emph{FVP} is a valid-time period in which the starting, the ending or both points are not precisely known.
 More formally we make the following definitions:

### 2012-03-16 trunk/book/sources/fuzzyTemporalDatabases.tex
update from the task list.

@@
 
-\subsection{Handling Data Imperfections in Temporal Databases}
+\subsection{Data Imperfections in Temporal Databases}
 
@@
 
-Data imperfections and techniques to represent them correctly in databases and queries are usually the focus of research in fuzzy databases.
-In fuzzy databases~\cite{Galindo2006}, uncertainty is usually expected to appear at storage level whereas imprecision is usually expected to appear in querying.
+Data and information imperfections and techniques to represent them correctly in databases and queries are usually the focus of research in fuzzy databases.
+Proposals from this field may present an approach based somehow on fuzzy set theory~\cite{Zadeh65} or possibility theory~\cite{Dubois:Prade:1988:PossibilityTheory}, although other theories support information imperfections too.
+Comparably, many proposals concerning the introduction of data imperfections or information imperfections in temporal databases present approaches based somehow on fuzzy set theory~\cite{garrido2009},~\cite{gal01},~\cite{nagypal2003},~\cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy} or possibility theory~\cite{Dubois89},~\cite{Pon11},~\cite{129895}, although proposals based on other theories exist~\cite{Dyreson1998},~\cite{Dekhtyar2001},~\cite{Qia09}.
+As possibility theory is usually seen as a theory of confidence, aimed at dealing with uncertainty, in some proposals, possibility theory is used specifically to handle uncertainty in temporal information.
+In fuzzy databases~\cite{Galindo2006}, uncertainty is usually expected to appear in the database content, whereas other types of imperfection, notably imprecision, are usually expected to appear in querying.
+
 Concerning temporal databases, there are several approaches to handle uncertainty in temporal data stored in a database.
+Many of these approaches concern several different types of time notions (VT, TT or DT), but most of these approaches focus somehow on valid-time~\cite{garrido2009},~\cite{gal01}.
 
-Valid-time notions usually take the form of an interval.
-Some models consider probability~\cite{Dekhtyar2001} or possibility~\cite{Dubois89} distributions to describe the endpoints of a time interval.
+In the following subsection, a novel approach to representing uncertainty concerning valid-time notions and a corresponding technique to query similar valid-time indications in a valid-time relation are proposed.
+The presented proposal is based on concepts introduced in~\cite{Dubois88b} and on the framework proposed in~\cite{Pon11}.
+
+\subsection{Handling Uncertainty in a Valid-time Relation}
+A valid-time indication usually takes the form of a time interval.
+Such a valid-time interval can be described (and stored in a valid-time relation record) using its boundaries (endpoints) or using one endpoint and the interval length.
+Usually, a valid-time interval is represented using its endpoints, which is also the approach adopted by the presented proposal.
+
+Generally, the uncertainty concerning a set of values might be described by a possibility distribution on the powerset of which one of the elements can be the intended set~\cite{Dubois88b}.
+This representation, however, introduces some issues in practice or in practical applications.
+Therefore, in the presented proposal, possibility theory is used to model uncertainty, but only uncertainty concerning the exact values of the start and end point of a valid-time interval is considered and the uncertainty in both the start point and the end point are modelled using possibility distributions.
+
+In fact, to model the uncertainty related to a valid-time interval using possibility theory, the presented proposal introduces so-called ill-known time intervals, relying on the concept of ill-known sets~\cite{Dubois88b}.
+
+%Some models consider probability~\cite{Dekhtyar2001} or possibility~\cite{Dubois89} distributions to describe the endpoints of a time interval.
 In this section, we will present a novel and currently researched technique to describe valid-time interval start- and endpoints using possibility distributions.
@@
 Finally, in subsection \ref{subsec:querying-time-intervals} the proposals for querying in a flexible way valid-time intervals are analyzed.
-\subsection{Handling Uncertainty in a Valid-time Relation}
 
@@
 \label{subsec:representation-time-intervals}
-Uncertainty in a valid-time interval may be modelled by modelling the uncertainty concerning the exact start, respectively end point of the valid-time interval.
+%Uncertainty in a valid-time interval may be modelled by modelling the uncertainty concerning the exact start, respectively end point of the valid-time interval.
 The uncertainty concerning these points can by modelled by means of possibility distributions.
@@
 
-A \emph{possibilistic variable} is defined as follows \cite{Pon11}.
+To represent valid-time indications which might contain uncertainty, the presented proposal introduces the concept of ill-known valid-time intervals, which relies on the concept of ill-known sets~\cite{Dubois88b}.
+To correctly explain the concept of ill-known sets, the concept of possibilistic variables should be introduced first.
+In the presented proposal, the definition of possibilistic variables of~\cite{Pon11} is followed.
+In~\cite{Pon11}, a \emph{possibilistic variable} is defined as follows.
 
@@
 \begin{definition}
+\textbf{Possibilistic variable}~\cite{Pon11}\\
 A possibilistic variable $X$ over a universe $U$ is defined as a variable taking exactly one value in $U$, but for which this value is (partially) unknown.
 Its possibility distribution $\pi_X$ on $U$ models the available knowledge about the value that $X$ takes: for each $u\in U$, $\pi_X(u)$ represents the possibility that $X$ takes the value $u$.
-In this work, this possibility is interpreted as a measure of how plausible it is that $X$ takes the value $u$, given (partial) knowledge about the value $X$ takes.
+In the presented work, this possibility is interpreted as a measure of how plausible it is that $X$ takes the value $u$, given (partial) knowledge about the value $X$ takes.
 \end{definition}
@@
 
-The exact value a possibilistic variable takes, which is (partially) unknown, is called an \emph{ill-known value} in this work \cite{Dubois88b}.
+Now, consider a set $R$, which contains single values (and not collections of values).
+When a possibilistic variable $X_{v}$ is defined on such a set $R$, the unique value $X_{v}$ takes, which is (partially) unknown, is called an \emph{ill-known value} in this work \cite{Dubois88b}.
 
 When a possibilistic variable is defined on the powerset $\Pow(R)$ of some universe $R$, the unique value the variable takes will be a crisp set and its possibility distribution on the powerset $\Pow(R)$ will describe the possibility of each crisp subset of $R$ to be the value the variable takes.
-This exact value (a crisp set) the variable takes, is now called an \emph{ill-known set} \cite{Dubois88b}.
+This value (a crisp set) the variable takes, which is (partially) unknown, is now called an \emph{ill-known set} \cite{Dubois88b}.
 
-This work will deal with ill-known time intervals.
-These are ill-known intervals in time.
-Ill-known intervals are ill-known sets, defined and represented via a start and end point, which will be ill-known values.
-The elements of the set are the values between the start and end point.
-A closed ill-known interval with start point defined by possibilistic variable $X$ and end point by possibilistic variable $Y$ is noted here $\left[X, Y\right]$.
-The correspondences and transitions between the representations of ill-known sets, between the representations of ill-known intervals and between the representations of an ill-known set and an ill-known interval are part of the authors current research.
+Finally, consider a set $R$, which contains single values (and not collections of values) and its powerset $\Pow(R)$.
+Now consider a subset $\Pow_{I}(R)$ of $\Pow(R)$ and let this subset contain every element of $\Pow(R)$ that is an interval, but no other elements.
+When a possibilistic variable $X_{i}$ is defined on the subset $\Pow_{I}(R)$ of the powerset $\Pow(R)$ of some set $R$, the unique value $X_{i}$ takes will be a crisp interval and the possibility distribution $\pi_{X_{i}}$ of $X_{i}$ will be a possibility distribution on $\Pow_{I}(R)$.
+This $\pi_{X_{i}}$ will define the possibility of each value of $\Pow_{I}(R)$ (a value of $\Pow_{I}(R)$ is a crisp interval which is a subset of $R$) being the value $X_{i}$ takes.
+This exact value (a crisp interval) the variable takes, which is (partially) unknown, is called an \emph{ill-known interval} here.
+
+The presented proposal will deal with ill-known time intervals.
+Ill-known time intervals are ill-known intervals in a time domain.
+In the presented proposal, an ill-known time interval will be defined and represented via its start and end point, which will be ill-known values.
+The elements of the ill-known time interval are the values between its start and end point, including the start and end points themselves\footnote{The presented proposal only deals with closed ill-known time intervals.
+Dealing with halfopen or open ill-known intervals is part of the current research of the authors.}.
+It should be clear that this approach to representing ill-known time intervals differs from the approach based on a single possibility distribution on a set $\Pow_{I}(R)$ of a set $R$.
+These approaches have a different behavior and can be used to describe different ill-known time intervals.
+The correspondences, interactions and transitions between these different representations of ill-known intervals and their interpretations are part of the authors current research.
+
+In the presented proposal, a closed ill-known time interval with start point defined by possibilistic variable $X$ and end point by possibilistic variable $Y$ is noted $\left[X, Y\right]$.
+Figure \ref{fig:interval} shows a closed ill-known time interval.
+%The correspondences and transitions between the representations of ill-known sets, between the representations of ill-known intervals and between the representations of an ill-known set and an ill-known interval are part of the authors current research.
 Figure \ref{fig:interval} illustrates this.
-
-%Uncertainty in a valid-time interval may be modelled with uncertainty at one or both starting and ending points.
-The uncertainty is usually represented by means of a possibility distribution in one or both points.
-It is also possible to represent with rough sets~\cite{Pawlak1995} time intervals~\cite{Qia09}.
-
-Several authors ~\cite{garrido2009} propose transformations in order to optimize the storage of such ill-known valid-time intervals, though recent research might seem to indicate some minor issues with respect to some of these transformations ~\cite{Pon11}.
-A comparison between the transformations in ~\cite{garrido2009} and the framework in~\cite{Pon11} is presented in~\cite{pon12}.
-Figure \ref{fig:convexhull} illustrates a transformation based on the `convex hull' approach REF.
-
-%An ill-known time interval is usually represented by means of two possibility distributions one in each starting and ending points.
-Figure \ref{fig:interval} shows an ill-known valid-time interval.
-The starting and ending points are denoted by two fuzzy numbers, $X$ and $Y$, represented by means of two triangular possibility distributions (see the Appendix).
-For illustration, figure \ref{fig:convexhull} shows one of the transformations based in the convex hull and proposed in~\cite{garrido2009}.
 
@@
 
+%Uncertainty in a valid-time interval may be modelled with uncertainty at one or both starting and ending points.
+The uncertainty is usually represented by means of a possibility distribution in one or both points.
+It is also possible to represent with rough sets~\cite{Pawlak1995} time intervals~\cite{Qia09}.
+
+Several authors work with concepts very similar to these ill-known time intervals and some of them~\cite{garrido2009} propose transformations of the describing functions in order to optimize the storage of such ill-known valid-time intervals, though recent research might seem to indicate some minor issues with respect to some of these transformations~\cite{Pon11}.
+A comparison between the transformations in~\cite{garrido2009} and the framework in~\cite{Pon11} is presented in~\cite{pon12}.
+Figure \ref{fig:convexhull} illustrates a transformation based on the `convex hull' approach~\cite{garrido2009}.
+
+%An ill-known time interval is usually represented by means of two possibility distributions one in each starting and ending points.
+Figure \ref{fig:interval} shows an ill-known valid-time interval.
+The starting and ending points are denoted by two fuzzy numbers, $X$ and $Y$, represented by means of two triangular possibility distributions (see the Appendix).
+For illustration, figure \ref{fig:convexhull} shows one of the transformations based in the convex hull and proposed in~\cite{garrido2009}.
 \begin{figure}
@@
 
-In section \ref{subsec:querying-time-intervals}, a technique allowing the querying of fuzzy valid-time databases which contain ill-known valid-time intervals is presented and briefly explained.
+%In section \ref{subsec:querying-time-intervals}, a technique allowing the querying of fuzzy valid-time databases which contain ill-known valid-time intervals is presented and briefly explained.
 The technique relies on the framework for set evaluation presented in \cite{Pon11}.
 
-\subsubsection{Querying Fuzzy Valid-Time Databases with Ill-Known Valid-Time Intervals}
+In the presented proposal, ill-known valid-time intervals will be used to represent valid-time indications in a valid-time relation and to model the uncertainty these may contain.
+
+To evaluate the temporal demands in queries issued by users to query a valid-time relation containing ill-known valid-time intervals, the presented proposal introduces a technique based on the concept of ill-known time constraints, which is based on the concept of ill-known constraints as presented in~\cite{Pon11}.
+Both concepts are treated in subsection \ref{iktc}.
+
+Before ill-known time constraints can be introduced, another notion related to possibilistic variables shoud be paid attention to.
+In fact, a specific application of possibilistic variables is obtained when the set under consideration is the set of boolean values, denoted $\mathbb{B}$ = $\{T,F\}$, where $T$ denotes `true' and $F$ denotes `false' \cite{Pon11}.
+Indeed, any boolean proposition $p$ takes exactly one value in $\mathbb{B}$.
+If the knowledge about which value this proposition $p$ takes, is given by a possibility distribution $\pi_p$, then proposition $p$ can be seen as a possibilistic variable.
+In the presented proposal, the interest lies with the case where the proposition holds (denoted $p$ = $T$) and the possibility and necessity that $p$ = $T$ demand most attention.
+In the following sections, the following notations are used, based on previous notations:
+
+\begin{align}
+\text{Possibility that $p$ = $T$:} \hspace{50pt} & Pos(p) = \pi_p(T) \\
+\text{Necessity that $p$ = $T$:} \hspace{50pt} & Nec(p) = 1-\pi_p(F) 
+\end{align}
+
+\subsubsection{\label{iktc}Ill-known Time Constraints}
+The presented proposal contains a technique for evaluating user queries used to query a valid-time relation in which the valid-time indications are represented by ill-known time intervals.
+Part of this query evaluation technique relies on the concept of ill-known time constraints, which is based on the concept of ill-known constraints as presented in~\cite{Pon11}.
+These concepts are presented below.
+Following~\cite{Pon11}, an \emph{ill-known constraint} is defined as follows.
+
+\begin{svgraybox}
+\vspace{-10pt}
+\begin{definition}
+Given a universe $U$, an \emph{ill-known constraint} $C$ is specified by means of a binary relation $R \subseteq U^{2}$ and a fixed, ill-known value defined by its possibilistic variable $V$ on $U$, i.e.:
+\begin{align}
+C \triangleq (V,R) \nonumber
+\end{align}
+A set $A \subseteq U$ now satisfies this constraint $C$ if and only if:
+\begin{align}
+\forall a \in A : (V,a) \in R \nonumber
+\end{align}
+\end{definition}
+\vspace{-10pt}
+\end{svgraybox}
+
+An example of an ill-known constraint is given by:
+\begin{align}
+C_{<} \triangleq (X, <) \nonumber
+\end{align}
+Some set $A$ then satisfies $C_{<}$ if 
+\begin{align}
+\forall a \in A : X <a \nonumber
+\end{align}
+
+The satisfaction of a constraint $C \triangleq (V,R)$ by a set $A$ is basically a Boolean matter (either the set satisfies the constraint or not) and can thus be seen as a boolean proposition, but due to the uncertainty inherent to the ill-known value $V$, it can be uncertain whether $C$ is satisfied by $A$ or not \cite{Pon11}.
+Based on the possibility distribution $\pi_{V}$ of $V$, the possibility and necessity that $A$ satisfies $C$ can be found.
+This proposition can thus be seen as a possibilistic variable on $\mathbb{B}$.
+The required possibility and necessity are calculated using the following formulas~\cite{Pon11}.
+
+\begin{align}
+Pos(A\text{ satisfies }C) & = \min_{a \in A}\left(\sup_{(w,a) \in R}\pi_{V}(w))\right) \label{eq:pos}\\
+Nec(A\text{ satisfies }C) & = \min_{a \in A}\left(\inf_{(w,a) \notin R} 1-\pi_{V}(w)\right) \label{eq:nec}
+\end{align}
+
+Now, to calculate the possibility or necessity of a set $A$ satisfying multiple constraints, the $\min$ t-norm operator is used to express a conjunction of constraints. For example:
+
+\begin{align}
+Pos((A\text{ satisfies }C_{1}) \text{ and } (A\text{ satisfies }C_{2})) & = \nonumber\\
+ \min_{a \in A}(Pos(A\text{ satisfies }C_{1}), Pos(A\text{ satisfies }C_{2})) \nonumber \\
+Nec((A\text{ satisfies }C_{1}) \text{ and } (A\text{ satisfies }C_{2})) & = \nonumber\\
+ \min_{a \in A}(Nec(A\text{ satisfies }C_{1}), Nec(A\text{ satisfies }C_{2})) \nonumber
+\end{align}
+
+Accordingly, the $\max$ t-conorm operator is used to express a disjunction of constraints. For example:
+
+\begin{align}
+Pos((A\text{ satisfies }C_{1}) \text{ or } (A\text{ satisfies }C_{2})) & = \nonumber\\
+ \max_{a \in A}(Pos(A\text{ satisfies }C_{1}), Pos(A\text{ satisfies }C_{2})) \nonumber \\
+Nec((A\text{ satisfies }C_{1}) \text{ or } (A\text{ satisfies }C_{2})) & = \nonumber\\
+ \max_{a \in A}(Nec(A\text{ satisfies }C_{1}), Nec(A\text{ satisfies }C_{2})) \nonumber
+\end{align}
+
+%Negation?
+
+In the presented proposal, \emph{ill-known time constraints} are considered, which are ill-known constraints of which the considered universe is a time domain.
+
+In the next subsection, the core of the presented proposal is described.
+
+\subsubsection{Querying Valid-time Relations containing Ill-known Valid-time Intervals}
 \label{subsec:querying-time-intervals}
 
-One of the main purposes of the existence of databases is to allow information retrieval.
-The standard query language for databases is SQL ~\cite{Mel93}.
-Several proposals to extend the SQL language for transaction-time databases~\cite{Sarda90}, valid-time databases~\cite{gad92} and bitemporal databases ~\cite{TSQL} exist and some authors have studied how to support temporal querying in standard SQL~\cite{Snodgrass98}.
+One of the main purposes of the existence of (relational) databases is to allow information retrieval.
+The standard query language for (relational) databases is SQL~\cite{Mel93}, but several proposals to extend the SQL language for transaction-time databases~\cite{Sarda90}, valid-time databases~\cite{gad92} and bitemporal databases~\cite{TSQL} exist and some authors have studied how to support temporal querying in standard SQL~\cite{Snodgrass98}.
 
-In the following subsections, the query structure, the query evaluation method and the technique to meaningfully rank and aggregate the query results, proposed in this work, are presented.
+As mentioned before, the presented proposal deals with querying a valid-time relation.
+In this subsection, the core of the presented proposal is described.
+First the particular structure of the relation is described, along with the nature and structure of its supposed contents.
+Next, the particular query structure is presented.
+Finally, the particular method for evaluating queries and for ranking the result records are presented.
+
+%In the following subsections, the nature and structure of the relation content, the query structure, the query evaluation method and the technique to meaningfully rank and aggregate the query results, proposed in this work, are presented.
 
@@
 
-\subsubsection{Query Structure} TO BE ANOTHER PART
-In this work, a query $\tilde{Q}$ is made up of two separate constructs of user query demands.
+\paragraph{\textbf{Relation Structure}}
+In the presented proposal, a valid-time relation is considered, in which every record contains just one valid-time indication.
+This valid-time indication is represented by a closed ill-known time interval, to allow uncertainty in the valid-time information.
+As explained above, the ill-known time intervals used here will be defined and represented via their start and end points, which are ill-known values in the valid-time domain.
+
+\paragraph{\textbf{Query Structure}}
+In the presented proposal, a query consists of two separate constructs of user demands.
 
@@
 
-The interpretation is that, for a record with an ill-known valid-time interval $J$, the user demands that $I$ AR $J$ hold.
+The interpretation of such a temporal demand $Q^{time} = \left( I,AR \right)$ is that, for a record with an ill-known valid-time interval $J$, the user demands that $I$ AR $J$ holds.
 
-\subsubsection{Query evaluation} TO BE ANOTHER PART
+\paragraph{\textbf{Query evaluation}}
 \label{subsubsec:query-evaluation}
-Query satisfaction in a fuzzy relational database is a matter of degree.
-Typically, the query evaluation results in a \emph{satisfaction degree} $s$, which is typically in the unit interval.
-$s \in \left[ 0,1,\right]$.
-Here, a satisfaction degree of $0$ denotes a total dissatisfaction while a degree of $1$ denotes a total satisfaction.
+Query satisfaction in a fuzzy relational database is usually a matter of degree.
+Typically, the evaluation of the query demands for a record results in a \emph{satisfaction degree} $s$, which is typically in the unit interval, i.e.
+$s \in \left[ 0,1\right]$.
+This satisfaction degree then models the extent to which the record satisfies the query demands.
+As such, a satisfaction degree of $0$ indicates total dissatisfaction while a degree of $1$ indicates total satisfaction and every level of satisfaction between total satisfaction and total dissatisfaction is indicated by a satisfaction degre between $0$ and $1$.
 
@@
 
+In the next subsection, some main concepts and issues concerning bipolarity in the context of temporal databases are presented and discussed.
+
 \subsection{Bipolarity in Temporal Databases}

### 2012-03-16 trunk/book/sources/temporalDatabases.tex
Still cool...

@@
 %
-A temporal database can generally be seen as a database that manages some temporal aspects in its schema \cite{etzion1998},~\cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy}.
-In subsection \ref{subsec:td-general-concepts}, some main concepts and properties concerning temporal databases and their definitions are presented and explained.
-In subsections \ref{subsubsec:primary-key} and \ref{subsubsec:consistency}, some main issues of relational temporal databases are presented and discussed.
-Finally, subsection \ref{Comm-temp} presents an overview of some commercial temporal database systems.
+A temporal database can generally be seen as a database that manages some temporal aspects in its schema~\cite{etzion1998},~\cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy}.
+In subsection~\ref{subsec:td-general-concepts}, some main concepts and properties concerning temporal databases and their definitions are presented and explained.
+In subsections~\ref{subsubsec:primary-key} and~\ref{subsubsec:consistency}, some main issues of relational temporal databases are presented and discussed.
+Finally, subsection~\ref{Comm-temp} presents an overview of some commercial temporal database systems.
 
@@
 Temporal values in a temporal database can be classified into the following types based on their interpretation and modelling purpose.
-The definitions and explanations of these types can be found in \cite{Dyreson1994} and \cite{Nascimento95decisiontime} and more information can be found in \cite{Jensen:1991:IIM:627283.627484}, \cite{Snodgrass:1984:TQL:588011.588041} and \cite{Nascimento95decisiontime}.
+The definitions and explanations of these types can be found in~\cite{Dyreson1994} and~\cite{Nascimento95decisiontime} and more information can be found in~\cite{Jensen:1991:IIM:627283.627484},~\cite{Snodgrass:1984:TQL:588011.588041} and~\cite{Nascimento95decisiontime}.
 
@@
 \vspace{-10pt}
-\begin{definition}\textbf{Valid Time}\\
-The \emph{\textbf{valid time}} (VT) of a fact is the time when the fact is true in the modeled reality.
+\begin{definition}\textbf{Valid Time}~\cite{Dyreson1994}\\
+The \emph{\textbf{valid-time}} (VT) of a fact is the time when the fact is true in the modeled reality.
 \end{definition}
 
-\begin{definition}\textbf{Transaction Time}\\
+\begin{definition}\textbf{Transaction Time}~\cite{Dyreson1994}\\
 A database fact is stored in a database at some point in time, and after it is stored, it is current until logically deleted.
-The \emph{\textbf{transaction time}} (TT) of a database fact is the time when the fact is current in the database and may be retrieved.
+The \emph{\textbf{transaction-time}} (TT) of a database fact is the time when the fact is current in the database and may be retrieved.
 \end{definition}
 
-\begin{definition}\textbf{Decision Time}\\
+\begin{definition}\textbf{Decision Time}~\cite{Nascimento95decisiontime}\\
 \emph{\textbf{Decision time}} (DT) denotes the time when an event was decided to happen.
@@
 
-\begin{definition}\textbf{User-defined Time}\\
+\begin{definition}\textbf{User-defined Time}~\cite{Dyreson1994}\\
 \emph{\textbf{User-defined time}} (UDT) is an uninterpreted attribute domain of date and time.
@@
 
-\emph{Valid times} are usually provided by the user, whereas \emph{transaction times} are usually system-generated and -supplied \cite{Dyreson1994}.
-Temporal values of type UDT are not given any extraordinary interpretation and have thus no extraordinary query language support \cite{Dyreson1994}.
+\emph{Valid times} are usually provided by the user, whereas \emph{transaction-times} are usually system-generated and -supplied~\cite{Dyreson1994}.
+Temporal values of type UDT are not given any extraordinary interpretation and have thus no extraordinary query language support~\cite{Dyreson1994}.
 
-A \emph{temporal database} can now formally be defined as follows \cite{Dyreson1994}.
+A \emph{temporal database} can now formally be defined as follows:
 
@@
 \vspace{-10pt}
-\begin{definition}\textbf{Temporal Database}\\
+\begin{definition}\textbf{Temporal Database}~\cite{Dyreson1994}\\
 A \emph{\textbf{temporal database}} supports some aspect of time, not counting user-defined time.
@@
 
-In a relational temporal database, temporal values will of course be in the tuples of the extentions of temporal relations:
+In a relational temporal database, temporal values will of course be in the tuples of the extensions of temporal relations:
 
@@
 \vspace{-10pt}
-\begin{definition}\textbf{Valid-time Relation}\\
-A \emph{\textbf{valid-time relation}} is a relation with exactly one system supported valid time \cite{Dyreson1994}.
+\begin{definition}\textbf{Valid-time Relation}~\cite{Dyreson1994}\\
+A \emph{\textbf{valid-time relation}} is a relation with exactly one system supported valid-time.
 \end{definition}
 
-\begin{definition}\textbf{Transaction-time Relation}\\
-A \emph{\textbf{transaction-time relation}} is a relation with exactly one system supported transaction time \cite{Dyreson1994}.
+\begin{definition}\textbf{Transaction-time Relation}~\cite{Dyreson1994}\\
+A \emph{\textbf{transaction-time relation}} is a relation with exactly one system supported transaction-time.
 \end{definition}
@@
 
-A \emph{valid-time}, respectively \emph{transaction-time} \emph{relational database} is now defined as containing one or more valid-time, respectively transaction-time relations \cite{Dyreson1994}.
-Next to this, \emph{bitemporal} relational databases contain both valid time and transaction time~\cite{Dyreson1994} and tritemporal databases contain valid time, transaction time and decision time~\cite{Nascimento95decisiontime}.
+A \emph{valid-time}, respectively \emph{transaction-time} \emph{relational database} is now defined as containing one or more valid-time, respectively transaction-time relations~\cite{Dyreson1994}.
+Next to this, \emph{bitemporal} relational databases contain both valid-time and transaction-time~\cite{Dyreson1994} and tritemporal databases contain valid-time, transaction-time and decision-time~\cite{Nascimento95decisiontime}.
 
-%Database models can also be classified into \emph{bi-temporal} (both valid and transaction-time) or \emph{tri-temporal}  (bi-temporal and decision time) models.
+%Database models can also be classified into \emph{bi-temporal} (both valid and transaction-time) or \emph{tri-temporal}  (bi-temporal and decision-time) models.
 
@@
 \subsection{\label{subsubsec:primary-key}Primary Keys in Valid-time Relation Design}
-Generally, when designing a relation based on a relational database model, a subset of the relation's attribute set is ususally chosen as primary key.
+Generally, when designing a relation based on a relational database model, a subset of the relation's attribute set is usually chosen as primary key.
 The values of a tuple for these attributes will then uniquely determine this tuple, hence no two distinct tuples may have the exact same values for every attribute in this primary key.
@@
 In this work, these attributes are called \emph{valid-time attributes}.
-In valid-time relation extentions, distinct tuples can exist containing the exact same values for every attribute except the valid-time attributes.
+In valid-time relation extensions, distinct tuples can exist containing the exact same values for every attribute except the valid-time attributes.
 These distinct tuples represent distinct versions of the same real object or concept, valid during different time periods.
@@
 \label{ex:pk}
-Consider the example valid-time relation visualized in table \ref{table:example-database}, which models when certain people worked as employees in a certain company and under whose supervision they worked during that time.
+Consider the example valid-time relation visualized in table~\ref{table:example-database}, which models when certain people worked as employees in a certain company and under whose supervision they worked during that time.
 The valid-time attributes `Start' and `End' describe the year when an employee started, respectively finished working for the company.
-For example, the last tuple visualized in table \ref{table:example-database} represents that the employee represented by this tuple started working for the company in 2005 and finished in 2009.
+For example, the last tuple visualized in table~\ref{table:example-database} represents that the employee represented by this tuple started working for the company in 2005 and finished in 2009.
 The attributes `Name', `Birthday' and `Supervisor' describe respectively the name, birthday date and unique identifier of the supervisor of an employee in the time during which he or she worked for the company.
@@
 The resulting database allows to insert a new row where Sarah the year 2010.
-But this modification also allows to insert spurious values e.g.
-inconsistent time periods (see last row in table \ref{table:example-database-with-new-pk}).
+But this modification also allows to insert spurious values e.g., inconsistent time periods (see last row in table \ref{table:example-database-with-new-pk}).
 
@@
 \label{ex:prob}
-Consider the example valid-time relation visualized in table \ref{table:erconsistency}, which is based on the relation visualized in table~\ref{table:example-database}.
+Consider the example valid-time relation visualized in table~\ref{table:erconsistency}, which is based on the relation visualized in table~\ref{table:example-database}.
 The primary key is again $\{$ID, Start, End$\}$.
@@
 The most usual solution is to re-define the DML (\emph{Data Manipulation Language}).
-e.g.
-the \emph{update} sentence is redefined as two sentences an update and a create sentence as illustrated in the following example.
+e.g., the \emph{update} sentence is redefined as two sentences an update and a create sentence as illustrated in the following example.
 
@@
 \subsection{\label{Comm-temp}Commercial Temporal Database Systems}
-Several commercial temporal DBMS exist. Table \ref{table:commercial-temporal-db} gives an overview of some of the more well-known temporal DBMS and provides references for more information.
+Several commercial temporal DBMS exist. Table~\ref{table:commercial-temporal-db} gives an overview of some of the more well-known temporal DBMS and provides references for more information.
 
 Oracle workspace manager~\cite{oracle2009} and TimeDB~\cite{timedb2005} are libraries for dealing with time in OracleDB.
-On another note, TimeDB and Postgree Temporal~\cite{posgree2009} are similar: both are simple implementations that implement a subset of the Allen operators and some operations for the creation and manipulation of temporal attributes (valid time, transaction time or both times are supported).
+On another note, TimeDB and Postgree Temporal~\cite{posgree2009} are similar: both are simple implementations that implement a subset of the Allen operators and some operations for the creation and manipulation of temporal attributes (valid-time, transaction-time or both times are supported).
 Teradata~\cite{teradata2011} is mainly a business intelligence system designed for data mining.
 Secondo~\cite{Dieker2000} is an extensible database system in which the core of the database may be replaced by a customized algebra.
-It is designed for non-standard applications and it supports both valid and transaction times.
+It is designed for non-standard applications and it supports both valid and transaction-times.
 
@@
 %\item
-%Valid time as well as transaction time are managed by the system. Each update sentence makes a change in the versioning of a row.
+%Valid time as well as transaction-time are managed by the system. Each update sentence makes a change in the versioning of a row.
 %\end{itemize}

### 2012-03-16 book/sources/fuzzyTemporalDatabases.tex
Still cool...

@@
 Concerning temporal databases, there are several approaches to handle uncertainty in temporal data stored in a database.
-Many of these approaches concern several different types of time notions (VT, TT or DT), but most of these approaches focus somehow on valid time~\cite{garrido2009},~\cite{gal01}.
+Many of these approaches concern several different types of time notions (VT, TT or DT), but most of these approaches focus somehow on valid-time~\cite{garrido2009},~\cite{gal01}.
 
-In the following subsection, a novel approach to representing uncertainty concerning valid time notions and a corresponding technique to query similar valid time indications in a valid-time relation are proposed.
+In the following subsection, a novel approach to representing uncertainty concerning valid-time notions and a corresponding technique to query similar valid-time indications in a valid-time relation are proposed.
 The presented proposal is based on concepts introduced in~\cite{Dubois88b} and on the framework proposed in~\cite{Pon11}.
@@
 \subsection{Handling Uncertainty in a Valid-time Relation}
-A valid time indication usually takes the form of a time interval.
-Such a valid time interval can be described (and stored in a valid time relation record) using its boundaries (endpoints) or using one endpoint and the interval length.
+A valid-time indication usually takes the form of a time interval.
+Such a valid-time interval can be described (and stored in a valid-time relation record) using its boundaries (endpoints) or using one endpoint and the interval length.
 Usually, a valid-time interval is represented using its endpoints, which is also the approach adopted by the presented proposal.
@@
 
-To represent valid time indications which might contain uncertainty, the presented proposal introduces the concept of ill-known valid time intervals, which relies on the concept of ill-known sets~\cite{Dubois88b}.
+To represent valid-time indications which might contain uncertainty, the presented proposal introduces the concept of ill-known valid-time intervals, which relies on the concept of ill-known sets~\cite{Dubois88b}.
 To correctly explain the concept of ill-known sets, the concept of possibilistic variables should be introduced first.
@@
 Now, consider a set $R$, which contains single values (and not collections of values).
-When a possibilistic variable $X_{v}$ is defined on such a set $R$, the unique value $X_{v}$ takes, which is (partially) unknown, is called an \emph{ill-known value} in this work \cite{Dubois88b}.
+When a possibilistic variable $X_{v}$ is defined on such a set $R$, the unique value $X_{v}$ takes, which is (partially) unknown, is called an \emph{ill-known value} in this work~\cite{Dubois88b}.
 
 When a possibilistic variable is defined on the powerset $\Pow(R)$ of some universe $R$, the unique value the variable takes will be a crisp set and its possibility distribution on the powerset $\Pow(R)$ will describe the possibility of each crisp subset of $R$ to be the value the variable takes.
-This value (a crisp set) the variable takes, which is (partially) unknown, is now called an \emph{ill-known set} \cite{Dubois88b}.
+This value (a crisp set) the variable takes, which is (partially) unknown, is now called an \emph{ill-known set}~\cite{Dubois88b}.
 
@@
 In the presented proposal, a closed ill-known time interval with start point defined by possibilistic variable $X$ and end point by possibilistic variable $Y$ is noted $\left[X, Y\right]$.
-Figure \ref{fig:interval} shows a closed ill-known time interval.
+Figure~\ref{fig:interval} shows a closed ill-known time interval.
 %The correspondences and transitions between the representations of ill-known sets, between the representations of ill-known intervals and between the representations of an ill-known set and an ill-known interval are part of the authors current research.
@@
 
-Several authors work with concepts very similar to these ill-known time intervals and some of them~\cite{garrido2009} propose transformations of the describing functions in order to optimize the storage of such ill-known valid time intervals, though recent research might seem to indicate some minor issues with respect to some of these transformations~\cite{Pon11}.
+Several authors work with concepts very similar to these ill-known time intervals and some of them~\cite{garrido2009} propose transformations of the describing functions in order to optimize the storage of such ill-known valid-time intervals, though recent research might seem to indicate some minor issues with respect to some of these transformations~\cite{Pon11}.
 A comparison between the transformations in~\cite{garrido2009} and the framework in~\cite{Pon11} is presented in~\cite{pon12}.
-Figure \ref{fig:convexhull} illustrates a transformation based on the `convex hull' approach~\cite{garrido2009}.
+Figure~\ref{fig:convexhull} illustrates a transformation based on the `convex hull' approach~\cite{garrido2009}.
 
@@
 
-In the presented proposal, ill-known valid time intervals will be used to represent valid time indications in a valid-time relation and to model the uncertainty these may contain.
+In the presented proposal, ill-known valid-time intervals will be used to represent valid-time indications in a valid-time relation and to model the uncertainty these may contain.
 
-To evaluate the temporal demands in queries issued by users to query a valid-time relation containing ill-known valid time intervals, the presented proposal introduces a technique based on the concept of ill-known time constraints, which is based on the concept of ill-known constraints as presented in~\cite{Pon11}.
-Both concepts are treated in subsection \ref{iktc}.
+To evaluate the temporal demands in queries issued by users to query a valid-time relation containing ill-known valid-time intervals, the presented proposal introduces a technique based on the concept of ill-known time constraints, which is based on the concept of ill-known constraints as presented in~\cite{Pon11}.
+Both concepts are treated in subsection~\ref{iktc}.
 
 Before ill-known time constraints can be introduced, another notion related to possibilistic variables shoud be paid attention to.
-In fact, a specific application of possibilistic variables is obtained when the set under consideration is the set of boolean values, denoted $\mathbb{B}$ = $\{T,F\}$, where $T$ denotes `true' and $F$ denotes `false' \cite{Pon11}.
+In fact, a specific application of possibilistic variables is obtained when the set under consideration is the set of boolean values, denoted $\mathbb{B}$ = $\{T,F\}$, where $T$ denotes `true' and $F$ denotes `false'~\cite{Pon11}.
 Indeed, any boolean proposition $p$ takes exactly one value in $\mathbb{B}$.
@@
 \subsubsection{\label{iktc}Ill-known Time Constraints}
-The presented proposal contains a technique for evaluating user queries used to query a valid time relation in which the valid time indications are represented by ill-known time intervals.
+The presented proposal contains a technique for evaluating user queries used to query a valid-time relation in which the valid-time indications are represented by ill-known time intervals.
 Part of this query evaluation technique relies on the concept of ill-known time constraints, which is based on the concept of ill-known constraints as presented in~\cite{Pon11}.
@@
 
-The satisfaction of a constraint $C \triangleq (V,R)$ by a set $A$ is basically a Boolean matter (either the set satisfies the constraint or not) and can thus be seen as a boolean proposition, but due to the uncertainty inherent to the ill-known value $V$, it can be uncertain whether $C$ is satisfied by $A$ or not \cite{Pon11}.
+The satisfaction of a constraint $C \triangleq (V,R)$ by a set $A$ is basically a Boolean matter (either the set satisfies the constraint or not) and can thus be seen as a boolean proposition, but due to the uncertainty inherent to the ill-known value $V$, it can be uncertain whether $C$ is satisfied by $A$ or not~\cite{Pon11}.
 Based on the possibility distribution $\pi_{V}$ of $V$, the possibility and necessity that $A$ satisfies $C$ can be found.
@@
 
-As mentioned before, the presented proposal deals with querying a valid time relation.
+As mentioned before, the presented proposal deals with querying a valid-time relation.
 In this subsection, the core of the presented proposal is described.
@@
 section~\ref{iktc}) is considered, depending on the Allen relation denoted by $AR$.
-The exact construct of constraints is an instantiation based on the formulas which can be found in table \ref{table:const}, for every possible value of $AR$.
+The exact construct of constraints is an instantiation based on the formulas which can be found in table~\ref{table:const}, for every possible value of $AR$.
 The form and capacity of these constraints are based on~\cite{Pon11}.
@@
 \label{subsubsec:bipolarity}
-Humans express their preferences using both positive and negative statements, where positive statements express what is desired or acceptable and negative statements express what is undesired or unacceptable \cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy}.
+Humans express their preferences using both positive and negative statements, where positive statements express what is desired or acceptable and negative statements express what is undesired or unacceptable~\cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy}.
 This realization is interesting with regard to database querying, because sometimes a user does not exactly know his or her preferences or can't express them in only positive statements, but prefers to use negative statements to express what he or she dislikes or doesn't need.
@@
 However, in some cases, positive preferences can not be directly obtained from negative preferences or vice versa.
-e.g., when a person prefers to buy a black motorbike, this does not necessarily mean the person would totally reject a very dark blue motorbike.
+E.g., when a person prefers to buy a black motorbike, this does not necessarily mean the person would totally reject a very dark blue motorbike.
 This phenomenon is called \emph{heterogeneous bipolarity}~\cite{Dubois2006},~\cite{Dubois2008}.
 
-The use of imprecise query preference formulation in bipolar querying is well discussed in existing literature \cite{DeTre2009}, \cite{Dubois2008}, \cite{Lacroix87}.
-In \cite{Lacroix87}, desired and mandatory query conditions are used, instead of positive and negative preferences.
+The use of imprecise query preference formulation in bipolar querying is well discussed in existing literature~\cite{DeTre2009},~\cite{Dubois2008},~\cite{Lacroix87}.
+In~\cite{Lacroix87}, desired and mandatory query conditions are used, instead of positive and negative preferences.
 However, the inverse of a mandatory preference expresses what should be rejected and this could be seen as negative information, whereas desired query conditions can be seen as positive preferences.
@@
 
-Bipolarity can be handled using different concepts, such as intuitionistic fuzzy sets~\cite{Atanassov1986}, interval valued fuzzy sets~\cite{Zadeh75a} Grattan-Guiness~\cite{Grattan76}, Janh~\cite{Jahn75}, Sambuc~\cite{Sambuc75},~\cite{Dubois05} or two fold fuzzy sets \cite{Dubois02}.
+Bipolarity can be handled using different concepts, such as intuitionistic fuzzy sets~\cite{Atanassov1986}, interval valued fuzzy sets~\cite{Zadeh75a} Grattan-Guiness~\cite{Grattan76}, Janh~\cite{Jahn75}, Sambuc~\cite{Sambuc75},~\cite{Dubois05} or two fold fuzzy sets~\cite{Dubois02}.
 

### 2012-03-16 book/sources/fuzzyTimeDomain.tex
Still cool...

--- 2012-03-13 book/sources/fuzzyTimeDomain.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 %INTRO
-As explained in the introduction, humans handle temporal information using temporal notions like time intervals or time points \cite{Dyreson1994}.
+As explained in the introduction, humans handle temporal information using temporal notions like time intervals or time points~\cite{Dyreson1994}.
 While the used temporal notions may contain imperfections~\cite{Dev98},~\cite{Dubois:jucs_9_9:fuzziness_and_uncertainty_in},~\cite{nagypal2003},~\cite{Dubois89}, humans often gracefully deal with these, as their inherent interpretation capability accounts for a lot of them.
@@
 
+In an information system, temporal information is usually related to facts or events~\cite{Chountas2000}.
+In light of this, a classification of temporal information can be considered, in which the following types of temporal information may be found:
+
+\begin{itemize}
+\item \emph{Definite temporal information}:
+Definite temporal information contains information describing a situation in which all time indications associated with some fact are absolute time indications.
+The temporal information is precisely known.
+
+\item \emph{Indefinite temporal information}~\cite{Dey1996}:
+Indefinite temporal information contains information describing a situation in which the time indication associated with some fact has not been fully defined.
+E.g., consider an event that in fact occured but it is not known exactly when.
+
+\item \emph{Infinite temporal uncertain information}~\cite{Kabanza1990}:
+Infinite temporal uncertain information contains information describing a situation in which an infinite number of time indications are associated with some fact.
+This is usually found in recurrent events like meetings.
+E.g., consider meetings that take place every Wednesday at noon.
+Some systems (usually with different information providers) may dispute the occurence and/or the duration of a fact.
+\end{itemize}
+
+\subsection{\label{subsec:representation}Representation of Imperfect Information}
+As mentioned before, information systems may have to deal with time indications which contain vagueness.
+Even for some specific events or facts, the temporal indications may become imprecise.
+Therefore, a time point might be specified by means of a time interval of which the boundaries may not be precisely known.
+An example.
+
+\begin{example}
+Consider a speaker and a hearer. The speaker wants to make an appointment with the hearer. Now, consider the speaker saysing: \\
+\begin{center}
+\emph{`We will meet each other tomorrow around 10'}\\
+\end{center}
+The hearer will now usually instinctively agree that the appointment will be in e.g. the time interval between 9.55h and 10.05h.
+\end{example}
+
+The study of the semantics of `around' in temporal~\cite{Dev98} indications has shown that the size of the time interval associated with the imprecise specification of a time point depends on the distance with respect to the current time.
+E.g., consider now that the speaker is talking about something that happened \emph{`during last week'}, then the hearer would consider a time interval of more or less 10 days.
+
+Some proposals~\cite{knight1993},~\cite{Cru97},~\cite{nagypal2003},~\cite{Chountas2000} conclude that the best representation for incomplete temporal knowledge is therefore based on time intervals, even if they refer to a fact that happen at a time point.
+This means that, as Allen proposed in~\cite{Allen83}, the primitive units (the chronons) in a time domain, used in an information system should be intervals.
+
+In order to represent and manage uncertain temporal information properly, several theoretical frameworks have been proposed:
+
+\begin{itemize}
+\item \emph{Probability theory}:
+Probability theory~\cite{Dey1996},~\cite{Lakshmanan1997},~\cite{Haddawy1996} is usually employed when uncertainty concerning a time interval allows a probability to be associated to the time interval.
+The use of probability theory is very usual in logistics information systems.
+E.g., \emph{`The package will arrive at its destination on monday morning with a probability of 0.8'}.
+
+\item \emph{Possibility theory}:
+Using possibility theory~\cite{Dubois:Prade:1988:PossibilityTheory}, a possibility degree is associated to the temporal fact or event.
+Possibility theory is widely used to model uncertainty and vaguenes in time~\cite{Dubois:jucs_9_9:fuzziness_and_uncertainty_in},\cite{Dubois89},\cite{devos94},\cite{nagypal2003}.
+Several works~\cite{schockaert08},~\cite{ohlbach2004} present fuzzy versions of the temporal relations proposed by Allen~\cite{Allen83}.
+The aim of these works is generally to obtain a flexible way to compare uncertain, ill-known temporal intervals by means of temporal relationships.
+
+\item \emph{Rough sets}:
+Rough set theory~\cite{Pawlak1995} has been used to represent uncertainty in time intervals.
+The two dimensional representation of time intervals and the temporal relationships between them has been studied in~\cite{Qia09}.
+\end{itemize}
+
 \subsection{Imperfections in Temporal Relationships}
@@
 These approaches are often based on Allen's operators~\cite{Allen83}.
+Example~\ref{ex:fuzzy-allen-relation} presents a short example concerning one of Allen's relationships.
 
-%%%TODO
+\begin{example}
+\label{ex:fuzzy-allen-relation}
+Consider an event which takes place between time points $A$ and $B$.
+Thus, the event comprises time interval $[A,B]$ (this is visualized in part (1) of figure~\ref{fig:fuzzy-allen-relationship}).
+The classical Allen relationship `after' returns an interval $]B,\infty[$ as shown in part (2) of figure~\ref{fig:fuzzy-allen-relationship}.
+A fuzzified version of Allen's `after' operator is illustrated in part (3) of figure~\ref{fig:fuzzy-allen-relationship}.
+The comparison between two time intervals results in a possibility degree in the unit interval.
+The shape of the possibility distribution is shown in part (3) of figure~\ref{fig:fuzzy-allen-relationship}.
+Note that all the points strictly after the point $B$ results in a possibility degree of 1 whereas there is an area near the point $B$ in which the possibility degree runs smoothly between 0 and 1.
 
-%Humans usually deals with time in a imprecise way.
-e.g.
-`A few days ago we received this package'.
-This issue has been studied in the field of Artificial Intelligence ~\cite{Tre97} and language understanding~\cite{DeCaluwe:1997:FTI:285506.285516}.
-Some proposals ~\cite{knight1993},~\cite{Cru97},~\cite{nagypal2003} conclude that the best representation for incomplete temporal knowledge are time intervals instead of time points.
-This means that, as Allen proprosed in \cite{Allen83} the primitive units (the chronons) in the temporal system are intervals.
+Consider now the interval given by $[C,D]$, illustrated in part (4) of figure~\ref{fig:fuzzy-allen-relationship}.
+The user wants to know if $[C,D]$ is after $[A,B]$.
+The crisp version of the `after' operator would return \emph{`no'} as an answer.
+The fuzzy version for the same operator would return \emph{`yes, with a possibility of 0.5'}.
 
-%When modelling temporal knowledge, the following types of information might be found~\cite{nagypal2003}:
-
-%\begin{description}
-%\item[\emph{Uncertainty.}] The temporal specification of the event is uncertain. This is usually found in historical documents when some documents state contradictory facts about some event.
-
-%\item[\emph{Vagueness.}] Some of the temporal specifications are defined in a vague way.
-e.g.
-`during summer', `late night'.
-Note also that temporal specifications become vague when the granularity is changed ~\cite{Cru97},~\cite{nagypal2003} .
-e.g.
-`The book was written by July 2012' is usefull if we are interested in the month and the year but it turns into a incomplete expression when we try to obtain the exact day.
-It is possible to set the following classification~\cite{Dev98}:
-
-%\begin{itemize}
-%\item
-%Non-numerical indications. e.g. \emph{often, usually}.
-%\item
-%Approximative time indications. e.g. \emph{around 10 a.m.}
-%\item
-%Indications on temporal relations. e.g. \emph{after 12, during summer}.
-%\end{itemize}
-
-%\item[\emph{Subjectivity} or \emph{Ambiguity}.] The specification of a temporal fact may be affected by subjectivity.  It is possible to distinguish between:
-%\begin{itemize}
-%\item Historical periods~\cite{nagypal2003} like `late romanticism' or `early middle ages'. 
-%\item The interpretation of the temporal event~\cite{Dev98} depends on the side of the hearer.
-e.g.
-A speaker says `We will meet each other at six', the hearer has to choose between 6 p.m.
-or 6 a.m.
-The speaker, however knows exactly which one is meant.
-%\end{itemize}
-%\end{description}
-
-%Another important task in the modelling of incomplete knowledge is when dealing with transitions among the different granularities.
+\begin{figure}
+\centering
+\includegraphics[scale=0.5]{graphs/fuzzyAllen.eps}
+\caption{Example for the Allen relationship \emph{`after'}.
+(1) The event bounded within time points $A$ and $B$.
+(2) The crisp version of the \emph{`after'} operator.
+(3) A fuzzy version of the after operator.
+(4) Another event, bounded within time points $[C,D]$.}
+\label{fig:fuzzy-allen-relationship}
+\end{figure}
+\end{example}
 

### 2012-03-16 book/sources/intro.tex
Still cool...

@@
 In reality, some aspects or properties of objects or concepts are time-variant or time-related.
-E.g., the moment of a bank transaction is traditionally a moment in time and thus a time-related notion, the function of an employee in a company can change through recorded history and is thus time-variant.
+e.g., the moment of a bank transaction is traditionally a moment in time and thus a time-related notion, the function of an employee in a company can change through recorded history and is thus time-variant.
 A temporal database schema is a database schema that models real objects or concepts with time-related or time-variant properties.
@@
 Among others, the querying of temporal databases has greatly profited from these temporal relationships, because they allowed for richer and more complex user-specified temporal query demands, by allowing to express more complex relationships between the temporal notions in the temporal expressions in the query and the temporal indications in the database.
-E.g., in a relation modelling who was department head of an institution during which periods of time, a query like `who were the department heads when Thomas worked for the institution' can be evaluated using similar relationships.
+e.g., in a relation modelling who was department head of an institution during which periods of time, a query like `who were the department heads when Thomas worked for the institution' can be evaluated using similar relationships.
 
@@
 Among many others, these possible imperfections in descriptions of temporal notions determine an important issue in temporal modelling.
-E.g., the description of the temporal notion in a sentence like `The Belfry of Bruges was finished on a day somewhere between 01/01/1201 A.D.
+e.g., the description of the temporal notion in a sentence like `The Belfry of Bruges was finished on a day somewhere between 01/01/1201 A.D.
 and 31/12/1300 A.D.' contains imperfection because of the uncertainty in the used time-related expression.
@@
 Next to temporal modelling, some attention has been spent on temporal reasoning~\cite{Allen83}.
-Although temporal reasoning is not discussed in this chapter, it should be noted that, among others, Dubois and Prade~\cite{Dubois:jucs_9_9:fuzziness_and_uncertainty_in},~\cite{Dubois89} have dealt with fuzziness and uncertainty in temporal reasoning.
+Although temporal reasoning is not discussed in this chapter, it should be noted that, among others, Dubois and Prade et al.~\cite{Dubois:jucs_9_9:fuzziness_and_uncertainty_in},~\cite{Dubois89} have dealt with fuzziness and uncertainty in temporal reasoning.
 
@@
 Section \ref{sec:time-domain} presents some basic concepts and terminology about temporal modelling and discusses some of its important aspects and issues, while section~\ref{sec:fuzzy-time-mod} presents some important issues concerning the combination of data imperfections and temporal modelling.
-In section \ref{sec:temporal-databases}, some basic concepts and terminology about temporal databases are presented, followed by an overview of some interesting issues concerning temporal databases and a survey of some commercial temporal database systems.
-Finally, in section \ref{sec:imprecision-uncertainty}, an approach to querying temporal databases containing imperfect temporal information is presented, followed by some conclusions and some suggestions for future work in section \ref{sec:further-research-work}.
+In section~\ref{sec:temporal-databases}, some basic concepts and terminology about temporal databases are presented, followed by an overview of some interesting issues concerning temporal databases and a survey of some commercial temporal database systems.
+Finally, in section~\ref{sec:imprecision-uncertainty}, an approach to querying temporal databases containing imperfect temporal information is presented, followed by some conclusions and some suggestions for future work in section~\ref{sec:further-research-work}.
 
@@
 %Another main issue in temporal databases is the querying.
-The user specifies in the query some temporal constraints, e.g.
-\emph{``before this year''},\emph{``after 1995''}.
+The user specifies in the query some temporal constraints, e.g., \emph{``before this year''},\emph{``after 1995''}.
 In order to solve these temporal comparisons, between the time specified in the query and the temporal data stored in the database, some basic relationships have been defined.
@@
 In~\cite{garrido2009}, some different temporal operators are defined by a combination of regular fuzzy comparisons.
-%e.g.
-the \emph{BEFORE} operator is defined by means of the fuzzy less than (FLT) operator.
+%e.g., the \emph{BEFORE} operator is defined by means of the fuzzy less than (FLT) operator.
 

### 2012-03-16 book/sources/timeDomain.tex
Still cool...

@@
 
-Another classification of time models concerns the use of points or intervals to model time. The equivalence between interval-based and point-based time models is demonstrated in \cite{Bohlen1998}.
+Another classification of time models concerns the use of points or intervals to model time. The equivalence between interval-based and point-based time models is demonstrated in~\cite{Bohlen1998}.
 
@@
 In this system, the chronons granularity is a set of milliseconds.
-Figure \ref{fig:granularity-graph-example} shows the complete granularity graph corresponding to this example.
+Figure~\ref{fig:granularity-graph-example} shows the complete granularity graph corresponding to this example.
 The transition between the chronons granularity and the seconds granularity is an example of a regular mapping.
@@
 \includegraphics[scale=0.5]{graphs/granularityGraph.eps}
-\caption{The granularity graph corresponding to example \ref{examplegrangraph}}
+\caption{The granularity graph corresponding to example~\ref{examplegrangraph}}
 \label{fig:granularity-graph-example}
@@
 Allen~\cite{Allen83} most notably described such relationships between time intervals and as a special case, between instants.
-Figure \ref{fig:allen} shows the temporal relationships Allen discerned.
+Figure~\ref{fig:allen} shows the temporal relationships Allen discerned.
 Some proposals can be applied to both crisp and other time intervals~\cite{ohlbach2004},~\cite{nagypal2003},~\cite{schockaert08},~\cite{garrido2009}.

### 2012-03-16 trunk/book/sources/timeDomain.tex
And (hopefully) the last commit. We're as good as done.

@@
 
-Data models used by information systems (and in specific, temporal database systems) may represent these underlying time domains using \emph{chronons}~\cite{Dyreson1994}.
+Data models used by information systems (and in specific, temporal database systems) may represent these underlying time domains using \emph{chronons}~\cite{Dyreson1994}, which can informally be described as the smallest distinguishable time units that can be used in the system.
 However, to explain what chronons are, an explanation of some other temporal concepts is necessary.
@@
 
-\begin{definition}\textbf{Temporal Element}\\
+\begin{definition}\textbf{Duration}~\cite{Dyreson1994}\\
+A \emph{\textbf{duration}} is an amount of time with known length, but no specific starting or ending instants.
+\end{definition}
+\vspace{-10pt}
+\end{svgraybox}
+
+A time interval as such is bounded by two instants, whereas a duration is not. Also, it should be noted that an instant is in fact a singular case of a time interval.
+
+\begin{svgraybox}
+\vspace{-10pt}
+\begin{definition}\textbf{Temporal Element}~\cite{Dyreson1994}\\
 A \emph{\textbf{temporal element}} is a finite union of time intervals.
@@
 
-\begin{definition}\textbf{Duration}\\
-A \emph{\textbf{duration}} is an amount of time with known length, but no specific starting or ending instants.
-\end{definition}
-
-\begin{definition}\textbf{`Temporal' as Modifier}\\
+\begin{definition}\textbf{`Temporal' as Modifier}~\cite{Dyreson1994}\\
 The modifier \emph{\textbf{`temporal'}} is used to indicate that the modified concept concerns some aspect of time.
@@
 
-Another classification of time models concerns the use of points or intervals to model time.
-The equivalence between interval-based and point-based time models is demonstrated in \cite{Böhlen_point-versus}.
+Another classification of time models concerns the use of points or intervals to model time. The equivalence between interval-based and point-based time models is demonstrated in~\cite{Bohlen1998}.
 
@@
 \end{definition}
+\vspace{-10pt}
 \end{svgraybox}
@@
 Two kinds of granularity-to-granularity mappings can now be discerned: an \emph{upwards mapping} is a mapping from a granularity $G$ to a coarser granularity $H$, whereas a downwards mapping is a mapping from a granularity $K$ to a finner granularity $L$.
-Different granularity-to-granularity mappings between several granularities can be represented using a granularity graph, which is a directed graph indicating the mapping conversions.
-
-Mappings between two granularities may be classified as \emph{regular mappings}, \emph{irregular mappings} or \emph{congruent mappings}~\cite{Lin97},~\cite{DyresonSnodgrass1994}.
+Orthogonal to this classification, mappings between two granularities may be classified as \emph{regular mappings}, \emph{irregular mappings} or \emph{congruent mappings}~\cite{Lin97},~\cite{DyresonSnodgrass1994}.
 \begin{itemize}
@@
 \emph{Regular mapping}: A regular mapping is a granularity-to-granularity mapping where the mapping function value is calculated by means of multiplications and/or divisions and (maybe) an anchor adjustment.
-E.g.
-the mapping value of the mapping between hours and minutes is calculated using a multiplication by $60$.
+e.g., the mapping value of the mapping between hours and minutes is calculated using a multiplication by $60$.
 \item
 \emph{Irregular mapping}: An irregular mapping is a granularity-to-granularity mapping where the mapping function value can not be calculated by means of multiplications and/or divisions.
-E.g.
-the mapping value of the mapping between months and days is dependent on the exact month or day.
+e.g., the mapping value of the mapping between months and days is dependent on the exact month or day.
 \item
 \emph{Congruent mapping}: A congruent mapping is a granularity-to-granularity mapping where the two granularities involved in the mapping have the same granules but a different anchor.
-E.g.
-the mapping between the days (Gregorian calendar days) and the academic days is a congruent mapping.
+e.g., the mapping between the days (Gregorian calendar days) and the academic days is a congruent mapping.
 \end{itemize}
 
+Different granularity-to-granularity mappings between several granularities can be represented using a granularity graph, which is a directed graph indicating the mapping conversions.
 The above is illustrated in the following example.
@@
 \begin{example}
-Consider a system that models both Gregorian Calendar dates as well as Academic dates.
-The chronons granularity in this system contains milliseconds.
-Figure \ref{fig:granularity-graph-example} shows the complete granularity graph corresponding with this example.
-The transitions between Chronons and Seconds and between Seconds and Minutes and so on is called a \emph{regular mapping}.
-The transition between Days and Months is not defined by a multiplication or division, this is called a \emph{irregular mapping}.
-In the graph, these \emph{irregular mappings} are noted with a bold arrow.
-Finally, the transition between a Gregorian Calendar Day and an Academic Day is called a \emph{Congruent mapping}: two granularities with the same granules but with different anchors.
+\label{examplegrangraph}
+Consider a system that models both Gregorian calendar dates as well as academic calendar dates.
+In this system, the chronons granularity is a set of milliseconds.
+Figure~\ref{fig:granularity-graph-example} shows the complete granularity graph corresponding to this example.
+The transition between the chronons granularity and the seconds granularity is an example of a regular mapping.
+Regular mappings are represented by thin arrows in the visualisation of the graph.
+The transition between the days granularity and the months granularity is an example of an irregular mapping.
+In the graph visualisation, irregular mappings are represented by a bold arrow.
+Finally, the transition between the Gregorian calendar day granularity and the academic day granularity is an example of a congruent mapping.
 Both concern the same days, but the academic year starts on October 1st, whereas the Gregorian calendar year starts on January 1st.
-Congruent mappings are visualised as straight lines without arrow heads.
+In the graph visualisation, congruent mappings are visualised as straight lines without arrow heads.
 
@@
 \includegraphics[scale=0.5]{graphs/granularityGraph.eps}
-\caption{A granularity graph example}
+\caption{The granularity graph corresponding to example~\ref{examplegrangraph}.}
 \label{fig:granularity-graph-example}
@@
 
-Next, in subsection \ref{subsec:timedomain-calendar}, some basic concepts and issues concerning calendars are explained.
+\vspace{-25pt}
 
-\subsection{Calendars}
-\label{subsec:timedomain-calendar}
+%Next, in subsection \ref{subsec:timedomain-calendar}, some basic concepts and issues concerning calendars are explained.
+
+%\subsection{Calendars}
+%\label{subsec:timedomain-calendar}
 %maybe introduce previously the concept of time units, linear hierarchy and time point.
 
-This subsection will concern \emph{calendars}.
+%This subsection will concern \emph{calendars}.
 A calendar provides some sort of human interpretation of time.
@@
 
-The next section briefly introduces temporal relationships.
+%The next section briefly introduces temporal relationships.
 
-\subsection{Temporal Relationships}
+\subsection{\label{allen-crisp}Temporal Relationships}
 %an explanation on allen's temporal relations.
@@
 
-Several operators are defined in order to compare temporal elements or time points.
-Allen~\cite{Allen83} first described the relationships between time intervals and as a special case, between time points.
-Figure \ref{fig:allen} shows the temporal relationships Allen discerned.
-Some proposals can be applied to both crisp and fuzzy temporal intervals~\cite{ohlbach2004},~\cite{nagypal2003},~\cite{schockaert08},~\cite{garrido2009}.
+Several (collections of) operators have been proposed in order to compare temporal elements and model the temporal relationships between them.
+Allen~\cite{Allen83} most notably described such relationships between time intervals and as a special case, between instants.
+Figure~\ref{fig:allen} shows the temporal relationships Allen discerned.
+Some proposals can be applied to both crisp and other time intervals~\cite{ohlbach2004},~\cite{nagypal2003},~\cite{schockaert08},~\cite{garrido2009}.
 
-In the next section, some basic concepts and terminology about temporal databases are presented, followed by an overview of some interesting issues concerning temporal databases and a survey of some commercial temporal database systems.
+%In the next section, some basic concepts and terminology about temporal databases are presented, followed by an overview of some interesting issues concerning temporal databases and a survey of some commercial temporal database systems.

### 2012-03-29 ipmu2012/introduction.tex
Modifications sugested by reviewers + new bibliography.

@@
 
-One issue in these attempts at temporal modelling is off course the inherent complexities of the concept of time.
+One issue in these attempts at temporal modelling is of course the inherent complexities of the concept of time.
 Another is the possible inherent imperfections in the description of notions of time.
@@
 The description of the temporal notion in a sentence like `The Alhambra was finished somewhere between 1301 A.D.
-and 1400 A.D.' contains imperfection caused by the uncertainty in the used expression about the exact finishing day of the building.
+and 1400 A.D.' contains imperfection caused by the imprecision in the used expression about the exact finishing day of the building.
 Though precise boundaries are given, the exact day on which the building was finished is unknown.
+An example of uncertainty is given by the following sentence: `I believe that the Alhambra was finished in 1300 A.D.'
+
 The transition between temporal granularities~\cite{Lin97} is also considered as a source of imprecision in temporal modelling and some proposals consider granularity as the base of the temporal model~\cite{Cru97}.
@@
 
-To allow information systems to cope with these imperfections, many approaches adopt fuzzy sets for the representation of temporal information~\cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy},\cite{Dubois:jucs_9_9:fuzziness_and_uncertainty_in}.
+To allow information systems to cope with these imperfections, there are some approaches that work with probability~\cite{Dekhtyar2001},~\cite{Parisi2010} whereas some other approaches adopt fuzzy sets for the representation of temporal information~\cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy},\cite{Dubois:jucs_9_9:fuzziness_and_uncertainty_in}.
 The temporal relations studied by Allen were recently fuzzified by several authors~\cite{ohlbach04},\cite{schockaert08}.
@@
 Different models have been proposed \cite{Jensen91},\cite{Snodgrass84},\cite{Nascimento95} that consider time as a crisp notion.
+Some of these models uses TSQL~\cite{Jensen1995} (a temporal extension to SQL) as query language.
 In some specific works, authors consider the necessity of allowing imprecision in temporal models~\cite{Cru97},\cite{Garrido2009}.

### 2012-03-29 trunk/ipmu2012/ipmu.tex
I modified the part about possibilistic variables because I needed the given information to explain the evaluation of sets...
Monday's update!
Some place-saving changes...
Modifications sugested by reviewers + new bibliography.

@@
 \usepackage{amsmath}
+\usepackage{amssymb}
 \usepackage{amsfonts}
+\usepackage{multirow}
 
@@
 \maketitle              % typeset the title of the contribution
-
+\def\JPicScale{0.5}
 \begin{abstract}
-
+Information in databases can be imperfect and this imperfection has several forms and causes.
+In some cases, a single value should be stored, but it is (partially) unknown.
+The uncertainty about which value to store leads to the aforementioned imperfection.
+In temporal databases, uncertainty can arise, concerning which temporal notion needs to be stored.
+Because in temporal databases, temporal notions influence the consistency with which the database models the reality, this uncertainty has a direct impact on the consistency of the model.
+To represent this temporal uncertainty, previous works have adapted fuzzy sets with conjunctive interpretation, an approach that might prove misleading.
+This work presents a model that represents the uncertainty using possibility and necessity measures, which are fuzzy sets with disjunctive interpretations.
 \end{abstract}
@@
 %
-The researchers are supported by the grant BES-2009-013805 within the research project TIN2008-02066: \emph{Fuzzy Temporal Information treatment in relational DBMS}.
+Part of the researchers are supported by the grant BES-2009-013805 in the project TIN2008-02066: \emph{Fuzzy Temporal Information treatment in RDBMS}.
 

### 2012-03-29 trunk/ipmu2012/preliminaries.tex
Modifications sugested by reviewers + new bibliography.

@@
 
-A specific application of possibilistic variables is obtained when the universe under consideration is the set of boolean values $\mathbb{B}$ = $\{T,F\}$.
-Indeed, any boolean proposition $p$ takes just one value in $\mathbb{B}$.
-If the knowledge on which value this proposition $p$ will take is given by a possibility distribution $\pi_p$, the proposition can be seen as a possibilistic variable.
+A specific application of possibilistic variables is obtained when the universe under consideration is the set of Boolean values $\mathbb{B}$ = $\{T,F\}$.
+Indeed, any Boolean proposition $p$ takes just one value in $\mathbb{B}$.
+If the knowledge about which value this proposition $p$ will take is given by a possibility distribution $\pi_p$, the proposition can be seen as a possibilistic variable.
 As the interest lies with the case where the proposition holds, the possibility and necessity that $p$ = $T$ (the proposition holds) demand most attention.
@@
 
-The most convenient form of the membership function of a fuzzy number is a triangular function.
+The most convenient form of the membership function of a fuzzy number is a triangular form.
 It can be shown that such a membership function $\mu_M$ for a fuzzy number $M$ is convex and normalized.
-Three real values, denoted $a$, $b$ and $D$, suffice to represent a triangular membership function of a fuzzy number and in this work, a fuzzy number defined as such will be noted as $\left[D, a, b \right]$.
+Three real values, denoted by $a$, $b$ and $D$, suffice to represent a triangular membership function of a fuzzy number and in this work, a fuzzy number defined as such will be noted as $\left[D, a, b \right]$.
 Here:
@@
 
-The membership function of such a fuzzy number is then given by:
-
-\vspace{-10pt}
-
-\begin{align}
-\mu_M : & \quad \mathbb{R} \rightarrow \left[0,1\right] \nonumber\\
- : & \quad x \rightarrow
-\begin{cases}
-0 & \mbox{ if } \left(x < D-a\right) \vee \left(x > D+b\right) \\
-\frac{1}{a}\left[x-\left(D-a\right)\right] & \mbox{ if } \left(x \geq D-a\right) \wedge (x \leq D)  \\
--\frac{1}{b}\left[x-\left(D+b\right)\right] & \mbox{ if } \left(x \geq D\right) \wedge (x \leq D+b)
-\end{cases}
-\end{align}
+%The membership function of such a fuzzy number is then given by:
+%
+%\vspace{-10pt}
+%
+%\begin{align}
+%\mu_M : & \quad \mathbb{R} \rightarrow \left[0,1\right] \nonumber\\
+% : & \quad x \rightarrow
+%\begin{cases}
+%0 & \mbox{ if } \left(x < D-a\right) \vee \left(x > D+b\right) \\
+%\frac{1}{a}\left[x-\left(D-a\right)\right] & \mbox{ if } \left(x \geq D-a\right) \wedge (x \leq D)  \\
+%-\frac{1}{b}\left[x-\left(D+b\right)\right] & \mbox{ if } \left(x \geq D\right) \wedge (x \leq D+b)
+%\end{cases}
+%\end{align}
 
@@
 \begin{eqnarray}
-C_1\triangleq\left(\geq,X\right)\\
-C_2\triangleq\left(\leq,Y\right).
+C_1 & \triangleq\left(\geq,X\right)\\
+C_2 & \triangleq\left(\leq,Y\right)
 \end{eqnarray}
 
-To calculate possibility and necessity concerning a conjunction of constraints, the $\min$ operator is used. The possibility and necessity of the set satisfying both constraints is then:
+To calculate the possibility and necessity concerning a conjunction of constraints, the $\min$ operator can be used.
+The possibility and necessity of $I$ being included in $\left[X, Y\right]$ are now: %satisfying both constraints is then:
 
@@
 	\item
-	\emph{Transaction time}: The time when the fact is stored in the database.
+	\emph{transaction time}~\cite{Jensen91}: time when the fact is stored in the database.
 	\item
-	\emph{Valid time}: The time when the fact is true in the modelled reality.
+	\emph{valid time}~\cite{Snodgrass84}: time when the fact is true in the modelled reality.
 	\item
-	\emph{Decision time}: The time when an event was decided to happen. 
+	\emph{decision time}~\cite{Nascimento95}: time when an event was decided to happen. 
 \end{itemize}
 
-Database models can then be classified into transaction time \cite{Jensen91}, decision time, valid time, bi-temporal \cite{Snodgrass84}(both valid and transaction time) or tri-temporal \cite{Nascimento95} (valid, transaction and decision time) models.
+Database models can also be classified into \emph{bi-temporal} (both valid and transaction-time) or \emph{tri-temporal}  (bi-temporal and decision time) models.
 
-To deal with time points \cite{Dubois89} or intervals \cite{Garrido2009} that are subject to imperfection and can thus not be precisely known, fuzzy temporal models \cite{schockaert08} exist.
+%To deal with time points  or intervals  that are subject to imperfection and can thus not be precisely known, fuzzy temporal models \cite{schockaert08} exist.
 %Some fuzzy temporal models assume that the time stored in the database is an interval.

### 2012-03-29 trunk/ipmu2012/proposal.tex
Modifications sugested by reviewers + new bibliography.

@@
 To store an ill-known value, four values (FT, F1, F2 and F3) are stored.
-They are explained in table \ref{table:relational-representation-pvp}.
+They are explained in Table \ref{table:relational-representation-pvp}.
 Note that while NULL denotes the fuzzy constant NULL, N denotes a null value in the database \cite{Medina94gefred.a}.
@@
 This subsection discusses a tool for querying.
+Although TSQL~\cite{Jensen1995} is the temporal extension to SQL, the query language proposed is an extension to FSQL~\cite{Gal98} with temporal operators~\cite{Garrido2009}.
 The focus will evidently lie on the querying of valid-time periods.
@@
 \item
-Depending on AR, a specific set of ill-known constraints is considered, which can be found in table 2.
+Depending on AR, a specific set of ill-known constraints is considered, which can be found in Table 2.
 The possibility $\Pos_{Q^{time}}(r)$ and necessity $\Nec_{Q^{time}}(r)$ that $r$ fulfills all these constraints are calculated using formulas based on equations \eqref{ill-known-pos} respectively \eqref{ill-known-nec} and aggregated using the $\min$ operator.
@@
 Because necessity cannot exceed $0$ unless possibility is $1$, this gives a natural ranking score.
+Some authors ~\cite{Bosc2010a} mentioned before that the possibility and necessity measures result in a total order in the set of events.
 This $e'_{Q^{time}}(r)$ is then rescaled to the unit interval, resulting in $e_{Q^{time}}(r)$.
@@
 
-\subsection{Example}
+%\subsection{Example}
 
-%\begin{example} 
-\subsubsection{The database}
-Consider the example relation given in table \ref{tb:car-models} describing car models, containing general attributes (model name, manufacturer, car segment) and one temporal attribute (stored as explained in subsection \ref{subsec:storage}) describing the approximate period of time during which the car model was sold.
+\begin{example}
+
+%\subsubsection{The database}
+\paragraph{The database}
+Consider the example relation given in Table \ref{tb:car-models} describing car models, containing general attributes (model name, manufacturer, car segment) and one temporal attribute (stored as explained in subsection \ref{subsec:storage}) describing the approximate period of time during which the car model was sold.
 The value for $D$ is stored in \emph{yyyy} format and $a$ and $b$ are represented by an integer.
@@
 
-\subsubsection{The query}
+%\subsubsection{The query}
+%\paragraph{The database}
 Consider the following query:
@@
 \end{table}
-%\end{example}
 
@@
 The last record, (ID 001, IID 3) shows also that with $\omega$ = 0.5 both temporal and regular criteria have the same importance.% By lowering the value for omega the temporal criteria would have more impact on the final ranking.
+
+\end{example}

### 2012-04-05 ijcis-pos-model/sources/temporalModel.tex
Preliminary version for the ijcis paper

--- 2012-02-22 trunk/ijcis-pos-model/sources/temporalModel.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{\label{subsec:time-interval-constraint}The time interval}
-The representation of a temporal interval is given by two points: S the starting and E the ending points respectively.
-
-\vglue13pt
-%\begin{table}[htbp]
-\tcap{Data types for a time point}
-\centerline{\small DATA TYPES}
-\vglue-6pt
-\centerline{\small\baselineskip=13pt
-\begin{tabular}{c p{2cm} p{2cm}}\\
-\hline
-1 & A single time point &  $1/x, x \in \T$\\
-2 & A possibility distribution in the numeric domain & A fuzzy number or a fuzzy interval.\\
-3 & An unknown value & \textbf{UNKNOWN}$= \lbrace1/t, t \in \T \rbrace$\\
-4 & An undefined value & \textbf{UNDEFINED}$= \lbrace 0/t, t \in \T \rbrace$\\
-5 & A null value & \textbf{NULL} $=\lbrace 1/$Unknown, 1/Undefined $\rbrace$\\
-\hline\\
-\end{tabular}
-\label{tbl:time-point-types}
-}
-%\end{table}
-
-\textcolor{red}{MAYBE COMPATIBILITY WITH GEFRED TYPES??}
-
-A valid temporal interval $\left[S,E\right]$ in the system is a combination for the values of $S$,$E$ as shown in table \ref{tbl:valid-time-interval}.
-
-\textcolor{red}{HAS IT ANY SENSE TO ALLOW THE COMBINATION [UNKNOWN, UNDEFINED] AND SO ON??}
 
@@
 Unknown & Unkown & Unknown interval. \\
-Undefined & Undefined & Undefined interval. \\
-Null & Null & Null interval. \\
 \hline

### 2012-04-05 trunk/ipmu2012/proposal.tex
modifications to fit 10 pages

@@
 %\subsubsection{The database}
-\paragraph{The database}
+%\paragraph{The database}
 Consider the example relation given in Table \ref{tb:car-models} describing car models, containing general attributes (model name, manufacturer, car segment) and one temporal attribute (stored as explained in subsection \ref{subsec:storage}) describing the approximate period of time during which the car model was sold.
@@
 The ID field identifies a car model while the field Instance ID (IID) identifies the instance for a car model, thus a car model in a certain state.
+
+\end{example}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \vspace{10pt}
-
 \label{tb:car-models}
-
 \vspace{-30pt}
-
 \end{table}
@@
 %\subsubsection{The query}
-%\paragraph{The database}
-Consider the following query:
+%\paragraph{The query}
+Consider the following query:\\
 \begin{center}
-\emph{The user wants to obtain a list of models from segment B, made by manufacturer Peugeot before the year interval 2001-2005}
+\emph{The user wants to obtain a list of models from segment B, made by manufacturer Peugeot before the year interval 2001-2005}\\
 \end{center}
@@
 The last record, (ID 001, IID 3) shows also that with $\omega$ = 0.5 both temporal and regular criteria have the same importance.% By lowering the value for omega the temporal criteria would have more impact on the final ranking.
-
-\end{example}

### 2012-04-07 ipmu2012/proposal.tex
removed latex warnings

@@
-In this section, a Valid-time model will be presented.
-The focus will off course be on the handling of imperfect Valid-time notions.
-First, the representation and storage of Valid-time will be presented.
-Next, a way of querying the model is proposed.
+In this section, a valid-time model, able to handle imperfect valid-time notions, will be presented.
+First, the representation and storage of valid-time will be presented.
+Next, an approach to query the model is proposed.
 Finally, an example is given.
@@
 \label{subsec:representation-ill-known}
-A database models a real object by storing the object's value for each attribute describing a property of the object.
-Thus, a Valid-time database following the presented proposal will model the time during which an object in a certain state is valid, by associating a \emph{Possibilistic Valid-time Period} to the record describing said object state:
+A database models real objects by storing the values of an object for each attribute describing a property of the object.
+Thus, a valid-time database following the presented proposal will model the time during which an object in a certain state is valid, by associating a \emph{Possibilistic Valid-time Period} (PVP) to the record describing this object state:
 
 \begin{definition}
-A \emph{Possibilistic Valid-time Period} (PVP) is an ill-known interval in time, which models a time period during which an object is valid.
+A \emph{Possibilistic Valid-time Period} is an ill-known interval in time, which models a time period during which an object in a certain state is valid.
 \end{definition}
 
-Off course, the object of which the PVP models the Valid-time is the object in the particular state modelled by the record that is associated with the PVP.
-Because a PVP is an ill-known interval, it allows modelling the uncertainty about the start or end point of a time interval (and thus about the time interval itself) if such uncertainty exists.
-The interpretation is \emph{disjunctive}: the PVP represents exactly one Valid-time interval, but precisely \emph{which} interval is represented, is (partially) unknown.
-In the presented model, only PVPs are considered in which the possibility distributions of the possibilistic variables defining the start and end point of the ill-known interval have the exact same characteristics of the membership functions of fuzzy numbers.
+Because a PVP is an ill-known interval, it allows modelling the uncertainty about the start and/or end point of a time interval (and thus about the time interval itself) if such uncertainty exists.
+The interpretation is \emph{disjunctive}: the PVP represents exactly one valid-time interval, but precisely \emph{which} interval is represented, is (partially) unknown.
+In the presented model, only PVPs are considered of which the possibility distributions of the possibilistic variables defining the start and end point of the ill-known interval have exactly the same characteristics as the membership functions of fuzzy numbers.
 A perfectly known start or end point can then be modelled by such an ill-known value defined by a possibilistic variable $P$ for which $\exists !
@@
 
-As mentioned in \cite{Pon11}, this approach differs from the one where a Valid-time period is represented by one fuzzy set.
-This one fuzzy set is seen as a possibility distribution on $\mathbb{R}$.
-However, in the presented approach, a time period is modelled using an ill-known set, which is defined by a possibility distribution on $\wp(\mathbb{R})$.
+As mentioned in \cite{Pon11}, this approach differs from the one where a valid-time period is represented by one fuzzy set.
+Such a fuzzy set is seen as a possibility distribution on $\mathbb{R}$ and thus defines just one ill-known value.
+However, in the presented approach, a time period is modelled using an ill-known set, which is defined by a possibility distribution on $\Pow(\mathbb{R})$.
 
@@
 To store a PVP, the possibility distributions defining the ill-known start and end point are stored.
-To store such a possibility distribution, the representation as used in the fuzzy interface for relational databases \emph{FIRST}~\cite{Medina94gefred.a,Gal98} is used.
+To store such a possibility distribution, the representation as used in the fuzzy interface for relational databases \emph{FIRST}~\cite{Medina94gefred.a}, \cite{Gal98} is used.
 Using this representation, it is possible to represent not only fuzzy numbers, but also (fuzzy) constants.
-To store an ill-known point, four values (FT, F1, F2 and F3) are stored.
-They are explained in table \ref{table:relational-representation-pvp}.
-Note that while NULL denotes the fuzzy constant NULL, N denotes a null value in the database.
+To store an ill-known value, four values (FT, F1, F2 and F3) are stored.
+They are explained in Table \ref{table:relational-representation-pvp}.
+Note that while NULL denotes the fuzzy constant NULL, N denotes a null value in the database \cite{Medina94gefred.a}.
 
@@
 \centering
+\caption{Relational representation for an ill-known time point.
+FT denotes Fuzzy Type.
+Field FT indicates that the values stored in F1, F2 and F3 denote either NULL, UNKNOWN, UNDEFINED or a triangular possibility distribution $M$ = $\left[D,\ a,\ b \right]$.
+Fields F1, F2 and F3 contain the actual values.}%  A \emph{PVP} is represented by two ill-known points.}
 \begin{tabular}{c c c c c c l p{2cm}}
@@
 UNKNOWN & 0 & N & N & N  & $1$ & Any value is equally possible\\ 
-UNDEFINED & 1 & N & N & N & $0$ & The value is not defined in the \\
-          &   &   &   &   &     & attribute domain\\ 
+UNDEFINED & 1 & N & N & N & $0$ & The value is not defined \\%in the \\
+    %      &   &   &   &   &     & attribute domain\\ 
 NULL & 2 & N & N & N &not defined & Nothing is known about the value \\ 
-$\left[D,\ a,\ b \right]$ & 3 & $D$ & $a$ & $b$ & $\mu(\left[D,\ a,\ b \right])$ & Ill-known value \\ 
+$M$ & 3 & $D$ & $a$ & $b$ & $\mu_{M}$ & Ill-known value \\ 
 \hline 
@@
 \subsection{Querying Ill-known Valid-time Intervals}
-In order to present a complete model, the tools for querying will be provided in this subsection.
-The focus will evidently lie on the querying of Valid-time periods.
+This subsection discusses a tool for querying.
+Although TSQL~\cite{Jensen1995} is the temporal extension to SQL, the query language proposed is an extension to FSQL~\cite{Gal98} with temporal operators~\cite{Garrido2009}.
+The focus will evidently lie on the querying of valid-time periods.
 First, the query structure will be defined, followed by the evaluation and ranking methods.
@@
 \subsubsection{Query Structure}
-It is important to notice that every Valid-time notion in the database is represented using a PVP.
-Thus, Valid-time intervals in the database can be partially unknown.
-In this framework, a query has two separate parts:
+It is important to notice that every valid-time notion in the database is represented using a PVP. In the presented framework, a query has two separate parts:
 
@@
 \end{definition}
-Here, $I$ is a crisp time interval and $AR$ is one of Allen's relations\cite{Allen83}.
-The interpretation is that for a record with Valid-time period given by a PVP $J$, the user requires that $I$ $AR$ $J$ holds.
+Here, $I$ is a crisp time interval and $AR$ is one of Allen's relations \cite{Allen83}.
+The interpretation is that for a record with valid-time period given by a PVP $J$, the user requires that $I$ $AR$ $J$ holds.
 
@@
 The evaluation of a query $\tilde{Q} = \left( Q^{time}, Q \right)$, where $Q^{time} = \left( I, AR \right)$ is now handled as follows.
-For each record $r$ in the database, with the Valid-time notion of $r$ specified by PVP $J$, two things happen independently:
+For each record $r$ in the database, with the valid-time notion of $r$ being specified by a PVP $J$, two things happen independently:
 
@@
 The preferences expressed in $Q$ are evaluated, resulting in a satisfaction degree denoted here $e_{Q}(r)$.
-The model presented here accepts any way of calculating this evaluation, as long as $e_{Q}(r) \in \left[0,1\right]$.
+The presented model accepts any sound way of calculating this evaluation, as long as $e_{Q}(r) \in \left[0,1\right]$.
 \item
-Depending on AR, a specific set of ill-known constraints is considered.
-For each AR, this set can be found in table \ref{tab:allen-relations}.
-The possibility $\Pos_{Q^{time}}(r)$ and necessity $\Nec_{Q^{time}}(r)$ that $r$ fulfills all these constraints are calculated using formulas based on equations \ref{ill-known-pos} and \ref{ill-known-nec}.
-The sum of these possibility and necessity scores gives an evaluation score $e_{Q^{time}}(r)$ in interval $\left[0,2\right]$.
-Because necessity cannot exceed $0$ unless possibility is $1$, this gives a natural score.
+Depending on AR, a specific set of ill-known constraints is considered, which can be found in Table 2.
+The possibility and necessity that $r$ fulfills all these constraints are calculated using formulas based on equations \eqref{ill-known-pos} respectively \eqref{ill-known-nec} and aggregated using the $\min$ operator.
+
+%The sum of the possibility score and the necessity score gives an evaluation score $e_{Q^{time}}(r)$ in interval $\left[0,2\right]$.
+Because necessity cannot exceed $0$ unless possibility is $1$, this gives a natural ranking score.
 \end{itemize}
@@
 \label{tab:allen-relations}
+\caption{Allen's relations used in the framework.
+Here, $I = \left[a, b\right]$ denotes the PVP in the query, $J = \left[X, Y\right]$ denotes the PVP of the record, with $\pi_{X}$ and $\pi_{Y}$ the possibility distributions of $X$ and $Y$ respectively.
+For each of Allen's relations $AR$, the corresponding value in column `Constraints' gives the constraints corresponding to $AR$.
+The last column contains the corresponding formula to calculate the possibility that $I$ satisfies all constraints given in column `Constraints', which are based on equation \eqref{ill-known-pos}.}
+
 \centering
-\begin{tabular}{|c|c|c|}
+\begin{tabular}{|c|l|l|}
 \hline
-Allen Relation & Constraints & $\Pos\left(\text{I satisfies all constraints }C_{i}, i = 1,2,...\right)$ \\
+\multirow{2}{*}
+{Allen Relation}  & {Constraints} & $\Pos($ I satisfies all constraints  \\
+ & & $C_{i}, i = 1,2,\ldots)$ \\
 \hline
@@
 \subsubsection{Ranking and aggregation}
-In order to present the results to the user, a ranking method is used.
-First of all, $e_{Q^{time}}(r)$ is rescaled to the unit interval.
-The final ranking $e_{final}$ is now given by a convex combination:
+In order to present the results to the user, a crude ranking method is used: for every record $r$, the sum of $\Pos_{Q^{time}}(r)$ and $\Nec_{Q^{time}}(r)$ gives an evaluation score $e'_{Q^{time}}(r)$ in interval $\left[0,2\right]$.
+Because necessity cannot exceed $0$ unless possibility is $1$, this gives a natural ranking score.
+Some authors ~\cite{Bosc2010a} mentioned before that the possibility and necessity measures result in a total order in the set of events.
+This $e'_{Q^{time}}(r)$ is then rescaled to the unit interval, resulting in $e_{Q^{time}}(r)$.
+The final ranking $e_{final}(r)$ is now given by a convex combination:
 
@@
 
-\subsection{Example}
+%\subsection{Example}
 
-%\begin{example} 
-\subsubsection{The database}
-Consider example database \ref{tb:car-models} describing car models, containing general attributes (model name, manufacturer, car segment) and one temporal attribute containing the approximate period of time during which the car model was sold.
-The temporal data is stored as explained in subsection \ref{subsec:storage}.
-Let the chronons in our example be years.
+\begin{example}
+
+%\subsubsection{The database}
+%\paragraph{The database}
+Consider the example relation given in Table \ref{tb:car-models} describing car models, containing general attributes (model name, manufacturer, car segment) and one temporal attribute (stored as explained in subsection \ref{subsec:storage}) describing the approximate period of time during which the car model was sold.
 The value for $D$ is stored in \emph{yyyy} format and $a$ and $b$ are represented by an integer.
 The ID field identifies a car model while the field Instance ID (IID) identifies the instance for a car model, thus a car model in a certain state.
+
+\end{example}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \vspace{10pt}
-\caption{Example database}
 \label{tb:car-models}
-
 \vspace{-30pt}
-
 \end{table}
 
-\subsubsection{The query}
+%\subsubsection{The query}
+%\paragraph{The query}
 Consider the following query:
-
 \begin{center}
-\emph{The user wants to obtain a list of models from segment B, made by manufacturer Peugeot before the year interval 2001-2005}
+\emph{The user wants to obtain a list of models from segment B, made by manufacturer Peugeot before the year interval 2001-2005}\\
 \end{center}
 
-Using the introduced notations, the query is translated to the following expression:
+Using the introduced notations in \eqref{eq:query-definition}, the query is translated to: %the following expression:
 
-\vspace{-10pt}
-
-\begin{center}
-$\tilde{Q} = \left(Q^{time}, Q\right)$
-\end{center}
-
-\vspace{-10pt}
-
-In this expression:
+%\vspace{-10pt}
+%
+%\begin{align}
+%$\tilde{Q} = \left(Q^{time}, Q\right)$
+%\end{center}
+%
+%\vspace{-10pt}
+%
+%In this expression:
 %\begin{itemize}
@@
 \end{table}
-%\end{example}
+
+Table \ref{tb:results} shows a natural and gradual ranking for the results.
+The last record, (ID 001, IID 3) shows also that with $\omega$ = 0.5 both temporal and regular criteria have the same importance.% By lowering the value for omega the temporal criteria would have more impact on the final ranking.

### 2012-04-09 ipmu2012/ipmu.tex
I think this should be the final version... but how knows..

@@
 To represent this temporal uncertainty, previous works have adapted fuzzy sets with conjunctive interpretation, an approach that might prove misleading.
-This work presents a model that represents the uncertainty using possibility and necessity distributions, which are fuzzy sets with disjunctive interpretations.
+This work presents a model that represents the uncertainty using possibility and necessity measures, which are fuzzy sets with disjunctive interpretations.
 \end{abstract}
+
+\begin{keywords}
+temporal databases, fuzzy databases, information systems, incompleteness
+\end{keywords}
 

### 2012-04-09 ipmu2012/preliminaries.tex
I think this should be the final version... but how knows..

@@
-%Introductory text:
-In this section, some basic concepts are introduced concerning possibilistic variables and fuzzy numbers and intervals, after which the framework of set evaluation by ill-known constraints \cite{Pon11} is explained.
+In this section, some basic concepts are introduced concerning possibilistic variables and fuzzy numbers and intervals, after which the framework of set evaluation by ill-known constraints \cite{Pons2011} is explained.
 The section concludes with a brief introduction to temporal databases.
 
-%\subsection{\label{subsec:possibility-theory}Possibility Theory}
-%Possibility theory, like probability theory, deals with uncertainty about the outcome of an experiment.
-In probability theory, this uncertainty is caused by the \emph{variability} in the outcomes, while in possibility theory, the uncertainty is caused by \emph{incomplete knowledge} about the experiment.
-The quantification of confidence in a theory of uncertainty is achieved using a confidence measure\cite{Pon11}.
-In probability theory this is a measure of chance, in possibility theory, possibility and necessity measures are used.
-
-%\begin{definition}
-%Consider a set of outcomes $\Omega$.
-Let $\wp(\Omega)$ denote the powerset of $\Omega$ and let $A$ and $B$ be elements of $\wp(\Omega)$.
-A \emph{confidence measure on $\Omega$} is defined by a function
-%	\begin{align}
-%	g : \wp(\Omega) & \rightarrow \left[0,1\right]
-%	\end{align}
-%that satisfies
-%	\begin{align}
-%	g(\emptyset) &= 0 \\
-%	g(\Omega) &= 1 	\label{NormalizationProperty} \\
-%	A \subseteq B &\Rightarrow g(A) \leq g(B) \label{MonotonicityProperty}
-%	\end{align}
-%\end{definition}
-
-%Both possibility measures and necessity measures are special cases of confidence measures.
-
-%\begin{definition}
-%Consider a confidence measure $\Pi$ on a set of outcomes $\Omega$.
-Let $J$ be a countable index set and let $\{ A_{j} | j \in J \wedge A_{j} \subseteq \Omega \}$ be a family of elements of $\wp(\Omega)$.
-$\Pi$ is now a \emph{possibility measure on $\Omega$} if it satisfies:
-%	\begin{align}
-%	\Pi\left(\bigcup_{j \in J} A_{j} \right) = \sup_{j \in J} \Pi(A_{j})
-%	\end{align}
-%\end{definition}
-
-%In this work, the interpretation is as follows.
-The possibility of an event expresses how plausible the occurrence of the event seems to an observer of the experiment, given the (partial) knowledge of the observer about the experiment.
-
-%Information on the possibility of distinct elements of the universe of discourse $\Omega$ can now be given by a \emph{possibility distribution} $\pi$ on $\Omega$, defined by:
-
-%\begin{definition}
-%Consider a possibility measure $\Pi$ on $\Omega$. A \emph{possibility distribution} $\pi$ on $\Omega$ underlying the possibility measure $\Pi$ is then a function defined by:
-%	\begin{align}
-%	\pi : \Omega \rightarrow \left[0, 1\right] : \pi(u) = \Pi(\{u\})
-%	\end{align}
-%\end{definition}
-
-%\begin{definition}
-%Consider a confidence measure $N$ on a set of outcomes $\Omega$.
-Let $J$ be a countable index set and let $\{ A_{j} | j \in J \wedge A_{j} \subseteq \Omega \}$ be a family of elements of $\wp(\Omega)$.
-$N$ is now a \emph{necessity measure} on $\Omega$ if it satisfies:
-%	\begin{align}
-%	N\left(\bigcap_{j \in J} A_{j} \right) = \inf_{j \in J} N(A_{j})
-%	\end{align}
-%\end{definition}
-
-%In this work, the interpretation is as follows.
-The necessity of an event expresses how necessary the occurrence of the event seems to an observer of the experiment, given the (partial) knowledge of the observer about the experiment.
-
-%Possibility and necessity measures are dual in the sense that:
-
-%\begin{align}
-%\forall A \subseteq \Omega : N(A) = 1 - \Pi(\bar{A})
-%\end{align}
-
-%Regarding interpretation, the above can be seen as: the degree to which an event is necessary is the degree to which every other possible event is not plausible.
-
 \subsection{\label{subsec:possibilistic-variables}Possibilistic Variables}
-Possibilistic variables rely on possibility theory \cite{Dubois:Prade:1988:PossibilityTheory}. A \emph{possibilistic variable} is defined as follows \cite{Pon11}.
+Possibilistic variables rely on possibility theory \cite{Dubois1988a}. A \emph{possibilistic variable} is defined as follows \cite{Pons2011}.
 
@@
 
-The exact value a possibilistic variable takes, which is (partially) unknown, is called an \emph{ill-known value} in this work \cite{Dubois88b}.
+The exact value a possibilistic variable takes, which is (partially) unknown, is called an \emph{ill-known value} in this work \cite{Dubois1988a}.
 
 When a possibilistic variable is defined on the powerset $\Pow(R)$ of some universe $R$, the unique value the variable takes will be a crisp set and its possibility distribution on the powerset $\Pow(R)$ will describe the possibility of each crisp subset of $R$ to be the value the variable takes.
-This exact value (a crisp set) the variable takes, is now called an \emph{ill-known set} \cite{Dubois88b}.
+This exact value (a crisp set) the variable takes, is now called an \emph{ill-known set} \cite{Dubois1988a}.
 
-%It is important to understand the difference between the following two concepts:
-%\begin{itemize}
-%\item
-%A \emph{possibilistic variable} $X$ is bounded to take only one value , but this value is not known due to incomplete knowledge. 
-%\item
-%An \emph{ill-known set}~\cite{Dubois88b}: a possibilistic variable defined over the universe $\Pow(U)$.
-%\end{itemize}
-
-%Note that while a possibilistic variable refers to one (partially) unknown value, an ill-known set is a crisp set but, for some reason, (partially) unknown.
-
-A specific application of possibilistic variables is obtained when the universe under consideration is the set of boolean values $\mathbb{B}$ = $\{T,F\}$.
-Indeed, any boolean proposition $p$ takes just one value in $\mathbb{B}$.
+A specific application of possibilistic variables is obtained when the universe under consideration is the set of Boolean values $\mathbb{B}$ = $\{T,F\}$.
+Indeed, any Boolean proposition $p$ takes just one value in $\mathbb{B}$.
 If the knowledge about which value this proposition $p$ will take is given by a possibility distribution $\pi_p$, the proposition can be seen as a possibilistic variable.
@@
 \end{align}
-%Here, notation \ref{propholdsposs} denotes the possibility that $p$ = $T$ and the proposition holds, notation \ref{propholdsnecc} denotes the necessity that $p$ = $T$ and the proposition holds.
 
@@
 \subsection{\label{subsec:fuzzy-numbers}Fuzzy Numbers and Fuzzy Intervals}
-Among others, Dubois and Prade~\cite{Dubois1983} use fuzzy sets \cite{zadeh65} to define a \emph{fuzzy interval}:
+Among others, Dubois and Prade~\cite{Dubois1983} use fuzzy sets \cite{Zadeh1965} to define a \emph{fuzzy interval}:
 \begin{definition}
@@
 
-%A simple form of the membership function of a fuzzy interval is a trapezoidal function.
-It can be shown that such a membership function $\mu_T$ for a fuzzy interval $T$ is convex and normalized.
-Four reel values, denoted $\alpha$, $\beta$, $\gamma$ and $\delta$ suffice to represent a trapezoidal membership function of a fuzzy interval.
-In this work, a fuzzy interval defined as such will be noted as $\left[\alpha, \beta, \gamma, \delta\right]$.
-%The corresponding membership function definition for this $\mu_T$ is then given by:
-
-%\begin{align}
-%\mu_T : & \quad \mathbb{R} \rightarrow \left[0,1\right] \\
-% : & \quad x \rightarrow
-%\begin{cases}
-%1 & \mbox{ if } x \in [\beta,\gamma] \\
-%0 & \mbox{ if } x > \delta \vee x < \alpha \\
-%\frac{x-\alpha}{\beta - \alpha} & \mbox{ if } x \in [\alpha,\beta[ \\
-%\frac{\delta -x}{\delta - \gamma} & \mbox{ if } x \in ]\gamma,\delta] \\
-%\end{cases}
-%\end{align}
-
-%\def\JPicScale{0.5}
-%\begin{figure}[h!]
-%  \centering
-%  \input{./graphs/trapezoidal.tex}
-%  \caption{Trapezoidal membership function}
-%  \label{fig:trapezoidal}
-%\end{figure}
-
 The most convenient form of the membership function of a fuzzy number is a triangular form.
 It can be shown that such a membership function $\mu_M$ for a fuzzy number $M$ is convex and normalized.
-Three real values, denoted $a$, $b$ and $D$, suffice to represent a triangular membership function of a fuzzy number and in this work, a fuzzy number defined as such will be noted as $\left[D, a, b \right]$.
+Three real values, denoted by $a$, $b$ and $D$, suffice to represent a triangular membership function of a fuzzy number and in this work, a fuzzy number defined as such will be noted as $\left[D, a, b \right]$.
 Here:
@@
 
-The membership function of such a fuzzy number is then given by:
-
-\vspace{-10pt}
-
-\begin{align}
-\mu_M : & \quad \mathbb{R} \rightarrow \left[0,1\right] \nonumber\\
- : & \quad x \rightarrow
-\begin{cases}
-0 & \mbox{ if } \left(x < D-a\right) \vee \left(x > D+b\right) \\
-\frac{1}{a}\left[x-\left(D-a\right)\right] & \mbox{ if } \left(x \geq D-a\right) \wedge (x \leq D)  \\
--\frac{1}{b}\left[x-\left(D+b\right)\right] & \mbox{ if } \left(x \geq D\right) \wedge (x \leq D+b)
-\end{cases}
-\end{align}
-
-%\begin{figure}[h!]
-%  \centering
-%  \input{./graphs/triangular.tex}
-%  \caption{Triangular membership function.}
-%  \label{fig:triangular}
-%\end{figure}
-
-%\subsubsection{Set evaluation by ill-known constraints}
-
 \subsection{Interval Evaluation by Ill-known Constraints}
-The problem of interval evaluation is more generally explained in \cite{Pon11}: the need exists to know if all points in a crisp interval $I$ reside between the boundaries of an ill-known interval $\left[ X , Y \right]$.
-In \cite{Pon11}, the notion of an \emph{ill-known constraint} is introduced:
+The problem of interval evaluation is more generally explained in \cite{Pons2011}: the need exists to know if all points in a crisp interval $I$ reside between the boundaries of an ill-known interval $\left[ X , Y \right]$.
+In \cite{Pons2011}, the notion of an \emph{ill-known constraint} is introduced:
 
@@
 
-The satisfaction of a constraint $C \triangleq (R,V)$ by a set $A$ is basically still a Boolean matter, but due to the uncertainty about the ill-known value $V$, it can be uncertain whether $C$ is satisfied by $A$ or not \cite{Pon11}.
+The satisfaction of a constraint $C \triangleq (R,V)$ by a set $A$ is basically still a Boolean matter, but due to the uncertainty about the ill-known value $V$, it can be uncertain whether $C$ is satisfied by $A$ or not \cite{Pons2011}.
 In fact, this satisfaction now behaves as a proposition.
@@
 
-%We assume that $X$ specifies the lower bound and $Y$ the upper bound for a given interval, we want to known whether all points in the interval are larger than or equal to $X$ and smaller than or equal to $Y$.
-Therefore, we consider two ill-known constraints:
-
 \vspace{-10pt}
@@
 
-To calculate the possibility and necessity concerning a conjunction of constraints, the $\min$ operator can be used.
-The possibility and necessity of $I$ being included in $\left[X, Y\right]$ are now: %satisfying both constraints is then:
+To calculate the possibility and necessity concerning a conjunction of constraints, the $\min$ operator can be used. The possibility and necessity of $I$ being included in $\left[X, Y\right]$ are now:
 
@@
 
-%\paragraph{Example} Consider the ill-known values $X = \left[5, 2, 8\right]$ and $Y = \left[9, 7, 10 \right]$.
-The knowledge about the evaluation of the interval $\left[a, b \right]$  is given by the expressions \eqref{eq:interval-pos},\eqref{eq:interval-nec}.
-Figure~\ref{fig:3d-possibility} shows a 3D plot of the possibility that an interval $[a,b]$ passes the evaluations specified by the ill-known constraints.
-Note the triangular form for the resulting possibility distribution since the condition $a \leq b$ holds.
-%
-%\begin{figure}[h!]
-%\centering
-%\includegraphics[scale=0.4]{graphs/3D_possibility.eps}
-%\caption{Possibility of evaluation for the interval $[a,b]$.}
-%\label{fig:3d-possibility}
-%\end{figure}
-%The necessity plot is obtained in a similar way and is shown in Figure~\ref{fig:3d-necessity}. Notice that the necessity measure is not normalized because the supports of $X$ and $Y$ overlap.
-%\begin{figure}[h!]
-%\centering
-%\includegraphics[scale=0.4]{graphs/3D_necessity.eps}
-%\caption{Necessity of evaluation for the interval $[a,b]$.}
-%\label{fig:3d-necessity}
-%\end{figure}
-
 \subsection{Temporal Databases}
-%In this section, the proposed reasoning is applied to the specific context of intervals on the real line. This setting is of specific interest in the context of fuzzy temporal databases.
 A \emph{temporal database} is a database that manages some aspects of time in its schema \cite{Dyreson1994}.
@@
 A \emph{chronon} is the shortest duration of time supported by the database.
-Time can be represented either as points~\cite{Dubois89} or intervals ~\cite{Garrido2009},\cite{655777} that may be subject to imperfection.
-%There are proposals for the fuzzyfication of the time point and the fuzzyfication of the time interval.
-
-%Time granularity is also associated with the representation of the time.
-A granularity is the result of partitioning on the set of chronons.
-The conversion among granularities is a common issue within temporal databases \cite{Lin97efficientconversion}.
-Granularity is the basis of some systems \cite{Cru97},\cite{624013}.
+Time can be represented either as points~\cite{DuBois1989} or intervals ~\cite{Garrido2009},\cite{Bohlen1998} that may be subject to imperfection.
 
@@
 	\item
-	\emph{transaction time}~\cite{Jensen91}: time when the fact is stored in the database.
+	\emph{transaction time}~\cite{Jensen1991}: time when the fact is stored in the database.
 	\item
-	\emph{valid time}~\cite{Snodgrass84}: time when the fact is true in the modelled reality.
+	\emph{valid time}~\cite{Snodgrass1984}: time when the fact is true in the modelled reality.
 	\item
-	\emph{decision time}~\cite{Nascimento95}: time when an event was decided to happen. 
+	\emph{decision time}~\cite{Nascimento1995}: time when an event was decided to happen. 
 \end{itemize}
@@
 Database models can also be classified into \emph{bi-temporal} (both valid and transaction-time) or \emph{tri-temporal}  (bi-temporal and decision time) models.
-
-%To deal with time points  or intervals  that are subject to imperfection and can thus not be precisely known, fuzzy temporal models \cite{schockaert08} exist.
-%Some fuzzy temporal models assume that the time stored in the database is an interval.
-The temporal interval is represented by two ill-known time points: $X$  an ill-known starting point and $Y$ an ill-known ending point.
-The interval $\left[X,Y\right]$ is not a fuzzy interval but an ill-known interval: it is a crisp interval but it is partially unknown which values are in this interval.

### 2012-04-09 ipmu2012/proposal.tex
last version
I think this should be the final version... but how knows..

@@
 Finally, an example is given.
-%The proposal consist on a possibilistic valid-time model. The representation and the querying are explained in the following subsections.
 
@@
 
-As mentioned in \cite{Pon11}, this approach differs from the one where a valid-time period is represented by one fuzzy set.
+As mentioned in \cite{Pons2011}, this approach differs from the one where a valid-time period is represented by one fuzzy set.
 Such a fuzzy set is seen as a possibility distribution on $\mathbb{R}$ and thus defines just one ill-known value.
 However, in the presented approach, a time period is modelled using an ill-known set, which is defined by a possibility distribution on $\Pow(\mathbb{R})$.
-
-%The interval has a starting and an ending points. An ill-known valid-time interval is an interval in witch one or both points are ill-known.
-
-%\begin{definition}
-%A Possibilistic Valid-Time Period \textbf{PVP} is a possibilistic interval defined by means of two ill-known points, namely $\left[ X,\ Y \right]$
-%\begin{equation}
-%PVP = \left[X,\ Y \right] 
-%\end{equation}
-%$X$ and $Y$ are ill-known values in the set of the real numbers $\mathbb{R}$. The uncertainty about the values taken by $X$ and $Y$ are given by the possibility distributions $\pi_X$ and $\pi_Y$.
-%\end{definition}
-
-%For convenience, the possibility distributions $\pi_X$ and $\pi_Y$ are given in the way of a triangular distribution, as explained in subsection \ref{subsec:fuzzy-numbers}.
-This representation allows overlapping (Fig.
-\ref{fig:pvp}).
-Note that while the two ill-known values $X,Y \in \mathbb{R}$, the fuzzy interval  $[X,Y] \in \mathbb{R}^2$.
-
-%\begin{figure}[h!]
-%  \centering
-%  \input{./graphs/2_triangular.tex}
-%  \caption{Two fuzzy numbers $X$ and $Y$ denoting a Possibilistic Valid-Time Period \emph{PVP}.}
-%  \label{fig:pvp}
-%\end{figure}
 
@@
 To store a PVP, the possibility distributions defining the ill-known start and end point are stored.
-To store such a possibility distribution, the representation as used in the fuzzy interface for relational databases \emph{FIRST}~\cite{Medina94gefred.a}, \cite{Gal98} is used.
+To store such a possibility distribution, the representation as used in the fuzzy interface for relational databases \emph{FIRST}~\cite{Medina1994}, \cite{Galindo1998} is used.
 Using this representation, it is possible to represent not only fuzzy numbers, but also (fuzzy) constants.
@@
 They are explained in Table \ref{table:relational-representation-pvp}.
-Note that while NULL denotes the fuzzy constant NULL, N denotes a null value in the database \cite{Medina94gefred.a}.
-
-%\begin{itemize}
-%\item
-%\emph{NULL}: This constant refers to a completely ignorance about the value.
-The possibility distribution for a given fuzzy number $X$ is not defined, therefore, any comparison between a fuzzy number and the \emph{NULL} constant always returns $0$.
-%\item
-%\emph{UNKNOWN}: % The possibility distribution for a given fuzzy number $X$ is $\pi_X=1$
-%\item
-%\emph{UNDEFINED}: The point does not have a value. %The possibility distribution for a given fuzzy number $X$ is $\pi_X=0$
-%\end{itemize}
-%
+Note that while NULL denotes the fuzzy constant NULL, N denotes a null value in the database \cite{Medina1994}.
 
@@
 Field FT indicates that the values stored in F1, F2 and F3 denote either NULL, UNKNOWN, UNDEFINED or a triangular possibility distribution $M$ = $\left[D,\ a,\ b \right]$.
-Fields F1, F2 and F3 contain the actual values.}%  A \emph{PVP} is represented by two ill-known points.}
+Fields F1, F2 and F3 contain the actual values.}
+
 \begin{tabular}{c c c c c c l p{2cm}}
@@
 UNKNOWN & 0 & N & N & N  & $1$ & Any value is equally possible\\ 
-UNDEFINED & 1 & N & N & N & $0$ & The value is not defined \\%in the \\
-    %      &   &   &   &   &     & attribute domain\\ 
+UNDEFINED & 1 & N & N & N & $0$ & The value is not defined \\
 NULL & 2 & N & N & N &not defined & Nothing is known about the value \\ 
@@
 This subsection discusses a tool for querying.
-Although TSQL~\cite{Jensen1995} is the temporal extension to SQL, the query language proposed is an extension to FSQL~\cite{Gal98} with temporal operators~\cite{Garrido2009}.
 The focus will evidently lie on the querying of valid-time periods.
@@
 
-%In this subsection we will define the query specification, then the evaluation of the query and finally the ranking for the query.
-
-%It is important to notice that every Valid-time notion in the database is represented using a PVP. Thus, Valid-time intervals in the database can be partially unknown.
-
-%In the query, the user is allowed to specify both non-temporal preferences and a temporal constraint.
-
-%This allows the user to specify both the preferences and an ill-known valid-time interval in the query.
-It is important to notice that the possibilistic / fuzzy data stored in the database has a \emph{disjunctive interpretation} (it is said that we have \emph{uncertainty}: the valid-time interval has only one value but, for some reason the value is ill-known).
-In the query specification, the user is allowed to express a crisp time interval.
-
 \subsubsection{Query Structure}
 It is important to notice that every valid-time notion in the database is represented using a PVP. In the presented framework, a query has two separate parts:
-
-% one is a temporal constraint and the other is the collection of query specifications for regular attributes:
-
-%the first one is the temporal specification. The second part is the query specification for regular attributes.
 
@@
 \end{definition}
-Here, $I$ is a crisp time interval and $AR$ is one of Allen's relations \cite{Allen83}.
+Here, $I$ is a crisp time interval and $AR$ is one of Allen's relations \cite{Allen1983}.
 The interpretation is that for a record with valid-time period given by a PVP $J$, the user requires that $I$ $AR$ $J$ holds.
 
-%In table \ref{tab:allen-relations}, .
-
-%\vspace{-35pt}
-
-\subsubsection{Query evaluation}
+\subsubsection{Query Evaluation}
 In fuzzy querying of regular (relational) databases, the modelling of query satisfaction is a matter of degree.
@@
 
-% by a combination of non-temporal query requirements $Q$ is noted $e_{Q}(r) \in \left[0,1\right]$ in this work. The total query evaluation method is now the following.  and Allen relation $AR$
 \begin{itemize}
@@
 The possibility and necessity that $r$ fulfills all these constraints are calculated using formulas based on equations \eqref{ill-known-pos} respectively \eqref{ill-known-nec} and aggregated using the $\min$ operator.
-
-%The sum of the possibility score and the necessity score gives an evaluation score $e_{Q^{time}}(r)$ in interval $\left[0,2\right]$.
-Because necessity cannot exceed $0$ unless possibility is $1$, this gives a natural ranking score.
 \end{itemize}
@@
 
-\subsubsection{Ranking and aggregation}
+\subsubsection{Ranking and Aggregation}
 In order to present the results to the user, a crude ranking method is used: for every record $r$, the sum of $\Pos_{Q^{time}}(r)$ and $\Nec_{Q^{time}}(r)$ gives an evaluation score $e'_{Q^{time}}(r)$ in interval $\left[0,2\right]$.
@@
 The final ranking $e_{final}(r)$ is now given by a convex combination:
-
-%the possibility and the necessity are ranked with the convex combination in \eqref{eq:convex-comb} with $\omega=0.5$.
-Finally this value is aggregated with the rest of the criteria with the same combination.
-%In this case the value for the aggregation is also $\omega = 0.5$
 
@@
 
-%By increasing the value for the parameter $\omega$, the preferences expressed in the query $Q$ can be given more importance. By lowering the value for $\omega$, the temporal criteria is emphasized.
-%The following example illustrates the querying process.
-
-%\subsection{Example}
-
 \begin{example} 
-
-%\subsubsection{The database}
-%\paragraph{The database}
 Consider the example relation given in Table \ref{tb:car-models} describing car models, containing general attributes (model name, manufacturer, car segment) and one temporal attribute (stored as explained in subsection \ref{subsec:storage}) describing the approximate period of time during which the car model was sold.
@@
 The ID field identifies a car model while the field Instance ID (IID) identifies the instance for a car model, thus a car model in a certain state.
+\end{example}
 
-\end{example}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Sample table for the car models. 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{table}[h]
@@
 
-%\subsubsection{The query}
-%\paragraph{The query}
 Consider the following query:
@@
 
-Using the introduced notations in \eqref{eq:query-definition}, the query is translated to: %the following expression:
-
-%\vspace{-10pt}
-%
-%\begin{align}
-%$\tilde{Q} = \left(Q^{time}, Q\right)$
-%\end{center}
-%
-%\vspace{-10pt}
-%
-%In this expression:
-%\begin{itemize}
-%\item
-%$Q^{time}\ = \ ( \left[ 2001,\ 2005 \right],\ $  before $)$.
-%\item
-%$c_{segment}\ = \ $ Peugeot.
-%\end{itemize}
-
-%\vspace{-15pt}
+Using the introduced notations in \eqref{eq:query-definition}, the query is translated to:
 
@@
 \end{align}
-
-%The evaluation for the criteria are presented in the result table \ref{tb:results}.
 
@@
 Table \ref{tb:results} shows a natural and gradual ranking for the results.
-The last record, (ID 001, IID 3) shows also that with $\omega$ = 0.5 both temporal and regular criteria have the same importance.% By lowering the value for omega the temporal criteria would have more impact on the final ranking.
+The last record, (ID 001, IID 3) shows also that with $\omega$ = 0.5 both temporal and regular criteria have the same importance.

### 2012-04-09 trunk/ipmu2012/conclusions.tex
I think this should be the final version... but how knows..

@@
-We presented a Valid-time model to represent and query ill-known temporal intervals.
-The main advantage of this framework is that it is always possible to get both a possibility and a necessity measures for every comparison, which is also useful for ranking purposes.
-The framework models the Allen's relations and it has the flexibility to model specific and more complex relations by means of ill-known constraints.
-As future work, the time interval in the query specification would also be ill-known.
+We presented a valid-time model to represent and query ill-known temporal intervals.
+The framework tries to correctly represent and handle querying of temporal uncertainty in a database.
+It is possible to model the Allen's relations and also more complex relations.
+As future work, new ways for aggregation and ranking are considered.

### 2012-04-09 trunk/ipmu2012/introduction.tex
I think this should be the final version... but how knows..

@@
-Dealing with time in information systems is usually a difficult task~\cite{Bolour82}.
+Dealing with time in information systems is usually a difficult task~\cite{Bolour1982}.
+First of all, the concept of \emph{time} itself is very complex.
+More towards information systems, several proposals have been concerned with obtaining theoretical models that allow the representation of time in information systems~\cite{VanderCruyssen1997a}.
+Notably, temporal relations between time intervals and to a lesser extent time points, were studied by Allen~\cite{Allen1983}.
 
-First of all, the concept of \emph{time} itself is very complex.
-More towards information systems, several proposals have been concerned with the obtaining of theoretical models that allow the representation of time in information systems~\cite{Cru97}.
-Notably, temporal relations between time intervals and to a lesser extent time points, were first studied by Allen~\cite{Allen83}.
+One issue in these attempts at temporal modelling is of course the inherent complexities of the concept of time.
+Another is the possible inherent imperfections in the description of notions of time.
+The description of the temporal notion in a sentence like `The Alhambra, which I can see from my room, is a few hundred years old.' contains imperfection caused by the vagueness of the used expression.
+The description of the temporal notion in a sentence like `The Alhambra was finished somewhere between 1301 A.D.
+and 1400 A.D.' contains imperfection caused by the imprecision in the used expression about the exact finishing day of the building.
+Though precise boundaries are given, the exact day on which the building was finished is unknown.
+An example of uncertainty is given by the following sentence: `I believe that the Alhambra was finished in 1300 A.D.'
 
-%Next to the inherent difficulties of time, an issue is \emph{temporal granularity}.
-A temporal granularity is a partitioning of the time line used by a system, usually dependent on the application.
-For example, the age of an adult human being is usually expressed in years: one will use sentences like `Laura is 21 years old' instead of `Laura is 21 years, 3 months and 4 days old'.
-In this example any time period shorter than a year needs no representation and thus the used granularity allows no specification for time periods shorter than a year.
+The transition between temporal granularities~\cite{Lin1997a} is also considered as a source of imprecision in temporal modelling and some proposals consider granularity as the base of the temporal model~\cite{VanderCruyssen1997a}.
 
-One issue in these attempts at temporal modelling is off course the inherent complexities of the concept of time.
-Another is the possible inherent imperfection in the description of notions of time.
-The temporal notion in a sentence like `The Alhambra, which I can see from my room, is a few hundred years old.' contains imperfection caused by the vagueness of the used expression.
-The temporal notion in a sentence like `The Alhambra was finished somewhere between 1301 A.D.
-and 1400 A.D.' contains imperfection caused by the inaccuracy in the given temporal notion, though precise boundaries are given.
-The transition between temporal granularities~\cite{Lin97} is also considered as a source of imprecision in temporal modelling and some proposals consider granularity as the base of the temporal model~\cite{Cru97}.
+To allow information systems to cope with these imperfections, there are some approaches that work with probability~\cite{Dekhtyar2001},~\cite{Parisi2010} whereas some other approaches adopt fuzzy sets for the representation of temporal information~\cite{Billiet2011},\cite{Dubois2003}.
+The temporal relations studied by Allen were recently fuzzified by several authors~\cite{Ohlbach2004},\cite{Schockaert2008}.
 
-%The transition between granularities~\cite{Lin97} is also considered as a source of imprecision.
-In a sentence like `The Alhambra is 612 years old', the temporal notion is very precise if time periods shorter than a year are irrelevant in the used granularity.
-However, if one is concerned with the exact day on which the last stone of the building is laid, a transition to a finer granularity, allowing the specification of time periods shorter than a year, is needed.
-Therefore, some proposals consider granularity as the base of the temporal model~\cite{Cru97}.
+Next to the issues in temporal modelling, the inclusion of time in a relational database system leads to several more practical problems, as the modelling of the temporal aspects of a time-variant object has an impact on the database consistency.
+Different models have been proposed \cite{Jensen1991},\cite{Snodgrass1984},\cite{Nascimento1995} that consider time as a crisp notion.
+Some of these models uses TSQL~\cite{Jensen1995} (a temporal extension to SQL) as query language.
+In some specific works, authors consider the necessity of allowing imprecision in temporal models~\cite{VanderCruyssen1997a},\cite{Garrido2009}.
 
-To allow information systems to cope with these imperfections, many approaches adopt fuzzy sets for the representation of temporal information~\cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy,Dubois:jucs_9_9:fuzziness_and_uncertainty_in}.
-The temporal relations studied by Allen were recently fuzzified by several authors~\cite{ohlbach04,schockaert08}.
-
-Next to the issues in temporal modelling, the addition of time to a relational database leads to several more practical problems, as the modelling of a time-variant object has an impact on the database consistency.
-Different models have been proposed \cite{Jensen91,Snodgrass84,Nascimento95} that consider time as a crisp notion.
-In recent works, some authors consider the necessity of allowing imprecision in temporal models~\cite{Cru97,Garrido2009}.
-
-Most of these models use the concepts of fuzzy numbers and fuzzy intervals to represent temporal imprecision and considerable attention has thus been given to the problem of transforming two fuzzy numbers which represent the boundaries of an imprecise time interval into one fuzzy interval representing this interval.
+Most of these models use the concepts of fuzzy numbers and fuzzy intervals to represent imprecise temporal notions and considerable attention has thus been given to the problem of transforming two fuzzy numbers which represent the boundaries of an imprecise time interval into one fuzzy interval representing this interval.
 There are several proposals for this transformation, but some, like those presented in~\cite{Garrido2009} are misleading.
 
-In this work we will propose a possibilistic valid time model for relational databases, based on the framework presented in~\cite{Pon11}, which deals with the inference of uncertainty about the evaluation of sets from the uncertainty about the exact values used in evaluation.
+In this work we will propose a possibilistic valid-time model for relational databases, based on the framework presented in~\cite{Pons2011}, which deals with the inference of uncertainty about the evaluation of sets from the uncertainty about the exact values used in evaluation.
 
-The paper is organized as follows. In section \ref{sec:preliminaries}, some preliminaries and concepts are presented. 
+The remainder of the paper is organized as follows. In section \ref{sec:preliminaries}, some preliminaries and concepts are presented. 
 Section \ref{sec:proposal} presents the possibilistic model for valid-time databases. In this section, both representation and querying are introduced.

### 2012-04-10 ipmu2012_2/ipmu.tex
The second paper for IPMU 2012

\documentclass{llncs}
%
%

\usepackage[dvips]{graphicx}
\usepackage{makeidx}  % allows for indexgeneration
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{multirow}
\usepackage{color}
\hyphenation{pro-per-ties}
\hyphenation{ge-ne-ral-ly}
\hyphenation{pre-fe-ren-ces}
\hyphenation{u-sing}
\hyphenation{pu-nish-ment}

%\newcommand{\Pow}{\mathcal{P}}
%\newcommand{\N}{\operatorname{N}}
%\newcommand{\bool}{\operatorname*{\mathcal{B}}}
%\newcommand{\Pos}{\operatorname{Pos}}
%\newcommand{\Nec}{\operatorname{Nec}}
%\newcommand{\T}{\mathcal{T}}

\newenvironment{keywords}{
       \list{}{\advance\topsep by0.35cm\relax\small
       \leftmargin=1cm
       \labelwidth=0.35cm
       \listparindent=0.35cm
       \itemindent\listparindent
       \rightmargin\leftmargin}\item[\hskip\labelsep
                                     \bfseries Keywords:]}
{\endlist}

\begin{document}

\mainmatter              % start of the contributions
%
\title{Evaluating Possibilistic Valid-Time Queries}
%
\author{Christophe Billiet\inst{1} \and Jos\'e Enrique Pons\inst{2} \and Olga Pons Capote\inst{2} \and Guy De Tr\'e\inst{1}}

%
\authorrunning{Christophe Billiet et al.} % abbreviated author list (for running head)
%
\institute{Department of Telecommunications and Information Processing, Ghent University \\
Sint Pietersnieuwstraat 41, B-9000 Ghent, Belgium \\
\email{Christophe.Billiet,Guy.DeTre@ugent.be}
\and
Department of Computer Science and Artificial Intelligence, 
Escuela T\'ecnica Superior de Ingenier\'ia Inform\'atica, Universidad de Granada \\
C/Periodista Daniel Saucedo Aranda s/n, E-18071 Granada, Spain \\
\email{jpons,opc@decsai.ugr.es}  
}

\maketitle              % typeset the title of the contribution

\begin{abstract}
A temporal database schema models objects or concepts with time-related or time-variant properties and a derived database contains measurements or descriptions concerning these temporal properties. The modelling of temporal information in a temporal database has an impact on the consistency of the database. Of course, this temporal information can be queried. In some cases, information in the query's temporal demand is (partially) unknown. This paper presents a technique for querying a valid-time relation using, among others, a temporal constraint in which the time indication in the temporal expression contains uncertainty and a method for evaluating such queries. This should allow querying using a (partially) unknown temporal demand.
\end{abstract}

\begin{keywords}
valid-time database, uncertainty, ill-known intervals, uncertain temporal querying
\end{keywords}

\section{Introduction}
\label{sec:Introduction}
\input{introduction}
%
\vspace{-10pt}
\section{Preliminaries}
\label{sec:preliminaries}

%\vspace{-5pt}
\input{preliminaries}

\section{Possibilistic Valid-Time Query Evaluation}
\label{sec:proposal}
\input{proposal}

\section{Conclusions}
\label{sec:futher-research}
\input{conclusions}
%

%
\subsubsection*{Acknowledgements.}
%
Part of the research is supported by grant BES-2009-013805 within research project TIN2008-02066: \emph{Fuzzy Temporal Information Treatment in Relational DBMS}.

\bibliographystyle{spmpsci}
\bibliography{ipmu}

\end{document}

### 2012-04-10 ipmu2012_2/conclusions.tex
Part of the paper

In this paper, a technique for querying valid-time relations is presented. This technique allows defining a temporal constraint based on a (partially) unknown time interval. The main advantage of this technique is that it tries to correctly represent and handle uncertainty in the temporal constraint, by using ill-known intervals. The technique allows the usage of both the Allen relations and more complex constructions. For future work, allowing the temporal specifications in the database to be (partially) unknown and new methods for aggregation and ranking are considered.

%We presented a valid-time model to represent and querying ill-known temporal intervals. The main advantage of the framework is that it is always possible to get both a possibility and a necessity measures for every comparison, which is also useful for ranking purposes. The framework models the Allen's relations and it has the flexibility to model specific and more complex relations by means of ill-known constraints.  As future work, the time interval in the query specification would also be ill-known.

### 2012-04-10 trunk/ipmu2012_2/introduction.tex
Part of the paper

%A database schema models a part of reality by modelling real objects or concepts. For this, a database contains data. 
A database contains data representing real objects or concepts. Every one of these data is a measurement or description of a property of a real object or concept. In reality, some aspects or properties of objects or concepts are time-variant or time-related. E.g. the moment of a bank transaction is traditionally a moment in time and thus a time-related notion, the function of an employee in a company can change through recorded history and is thus time-variant. A \emph{temporal database schema} \cite{Dyreson1994} is a database schema that models real objects or concepts with time-related or -variant properties. However, the modelling of temporal aspects has a direct impact on the consistency of the temporal database, because the temporal nature of these aspects imposes extra integrity constraints. An example. Consider a relation in a relational library database, modelling the presence of books in the library. Every physical book is represented by a unique identifier. Every record in the relation contains such an identifier, a date on which the corresponding book was loaned and a date on which it was subsequently returned (if it is returned). Without further precautions, a library employee could add several records with the same book identifier, different `loaned'-dates and no `returned'-dates. This would represent a situation in which the same physical book was loaned several times on different dates and never returned, which is of course impossible. A temporal database schema will typically constrain record insertion and prevent similar modelling inconsistencies.

Humans handle temporal information using certain temporal notions like time intervals or time points \cite{Dyreson1994}. The modelling and handling of these temporal notions in information systems has always been seen as a difficult task \cite{Bolour82}. Klein \cite{klein94} studied the concept of time in language and Devos \cite{devos94} modelled vague temporal expressions by means of fuzzy sets \cite{zadeh65}. Notably, temporal relationships between time intervals (and as a special case instants~\cite{Dyreson1994}) were studied by Allen \cite{Allen83}. Among many others, an issue in temporal modelling is the possible imperfection in (descriptions of) temporal notions. E.g. the temporal notion in a sentence like `The Belfry of Bruges was finished on one single day somewhere between 1/01/1201 A.D. and 31/12/1300 A.D.' contains imperfection because of the uncertainty in the used time-related expression. It is known that the building was finished on a single day, but it is not known precisely which day this was.

To allow information systems to cope with these and similar imperfections, many approaches adopt fuzzy sets for the representation of temporal information \cite{343607}, \cite{nagypal03}, \cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy}, \cite{Dubois:jucs_9_9:fuzziness_and_uncertainty_in}. The temporal relationships studied by Allen were fuzzified by several authors \cite{ohlbach04}, \cite{nagypal03}, \cite{schockaert08}.

Next to this, the inclusion of temporal indications in a relational database leads to several practical problems. In some specific works, authors consider the necessity of allowing imprecision in the representation of temporal indications \cite{Cru97}, \cite{Garrido2009}. Most of these approaches use the concepts of fuzzy numbers and fuzzy intervals to represent imprecise temporal indications and considerable attention has thus been given to the problem of transforming two fuzzy numbers which represent the boundaries of an imprecise time interval into one fuzzy interval representing this interval. There are several proposals for this transformation, but some of them might show signs of minor issues, as pointed out in \cite{Pon11}.

In this work, an approach is proposed that allows users to query a valid-time relation using temporal demands containing uncertainty. The novelty of this approach is that it allows uncertainty in the query's temporal demand. The presented approach is based on the possibilistic framework for set evaluation found in \cite{Pon11}. In section \ref{sec:preliminaries}, some main concepts are explained and the part of the framework on which this work is based, is briefly presented. In section \ref{sec:proposal}, this paper's approach is presented. Finally, in section \ref{sec:futher-research}, conclusions are drawn and some directions for future research are presented.

%In the attempts at temporal modelling, one issue is the inherent complexities of the concept of time. Klein~\cite{klein94} studied the concept of time in language and Devos~\cite{devos94} modelled vague temporal expressions by means of fuzzy sets~\cite{zadeh65}. 
%\\
%\\
%Transition between temporal granularities~\cite{Lin97} is also considered as a source of imprecision in temporal modelling and some proposals consider granularity as the base of the temporal model~\cite{Cru97}.
%Different approaches have been proposed \cite{Jensen91,Snodgrass84,Nascimento95} that consider time as a crisp notion.

%\\
%Specifically towards information systems, several proposals have been concerned with the obtaining of theoretical models that allow the representation of time in information systems~\cite{Cru97}. Notably, temporal relations between time intervals and to a lesser extent time points, were first studied by Allen~\cite{Allen83}.
%\\
%To allow information systems to cope with these imperfections, many approaches adopt fuzzy sets for the representation of temporal information~\cite{343607,nagypal03,Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy,Dubois:jucs_9_9:fuzziness_and_uncertainty_in}. The temporal relations studied by Allen were recently fuzzified by several authors~\cite{ohlbach04,nagypal03,schockaert08}.

%Next to the issues in temporal modelling, the addition of temporal notions to a relational database leads to several practical problems, as the attempt to model or introduce a time-variant object in a relational databse has an impact on the database consistency. Different approaches have been proposed \cite{Jensen91,Snodgrass84,Nascimento95} that consider time as a crisp notion. In recent works, some authors consider the necessity of allowing imprecision in temporal representation~\cite{Cru97,Garrido2009}.
%\\
%Most of these approaches use the concepts of fuzzy numbers and fuzzy intervals to represent temporal imprecision and considerable attention has thus been given to the problem of transforming two fuzzy numbers which represent the boundaries of an imprecise time interval into one fuzzy interval representing this interval. There are several proposals for this transformation, but some, like those presented in~\cite{Garrido2009}, are misleading.

%In this work, an approach is proposed that allows users to query a Valid-time database using imperfectly defined temporal demands. This approach is based on the possibilistic framework found in \cite{Pon11}. In section \ref{sec:preliminaries}, some main concepts are explained and the part of the framework on which this work is based, is briefly presented.

### 2012-04-10 trunk/ipmu2012_2/preliminaries.tex
Part of the paper

%Introductory text:
This section starts with a brief introduction to temporal databases. Next, some basic concepts are introduced, concerning possibilistic variables and ill-known values, sets and intervals. Last, the framework of set evaluation using ill-known constraints \cite{Pon11} is explained.

\subsection{Temporal Databases}
%In this section, the proposed reasoning is applied to the specific context of intervals on the real line. This setting is of specific interest in the context of fuzzy temporal databases.
As explained in the introduction, a \emph{temporal database schema} models objects or concepts with time-related or -variant properties and a \emph{temporal database} contains measurements or descriptions of temporal properties of these objects or concepts. Also, the modelling of temporal aspects has a direct impact on the consistency of the temporal database.

In \cite{Dyreson1994}, a consensus glossary on temporal terminology has been presented. Next, some of the important concepts from this glossary are described.

A \emph{chronon} is a non-decomposable time interval with minimal duration and the database is thus unable to distinguish time intervals shorter than a chronon.

Based on their interpretation and modelling purpose, temporal data in a temporal database can be classified into four types. Of these types, \emph{user-defined time} contains temporal data without impact on the consistency of the temporal database. These data are not handled specifically. The other types are:

%A \emph{chronon} is the shortest duration of time supported by the TDB and thus the database is unable to distinguish time periods shorter than a chronon.

%Time in a TDB can be represented either as points or intervals \cite{655777}.
%There are proposals for the fuzzyfication of the time point and the fuzzyfication of the time interval.

%Time granularity is also associated with the representation of the time. A granularity is the result of partitioning on the set of chronons. The conversion among granularities is a common issue within temporal databases \cite{Lin97efficientconversion}. Granularity is the basis of some systems \cite{Cru97},\cite{624013}.

\vspace{-5pt}
\begin{itemize}
	\item \emph{transaction time} \cite{Dyreson1994} contains temporal data describing when a fact is stored in the database and not yet logically deleted
	\item	\emph{valid time} \cite{Dyreson1994} contains temporal data describing when a fact is true in the modelled reality
	\item	\emph{decision time} \cite{Nascimento95} contains temporal data describing when an event was decided to happen
\end{itemize}
\vspace{-5pt}

Other possible types of database models are \emph{bi-temporal} (both valid and transaction time) \cite{Dyreson1994} or \emph{tri-temporal} (valid, transaction and decision time) \cite{Nascimento95} models. To deal with descriptions of time points \cite{Dubois89} or intervals \cite{Garrido2009} that are subject to imperfection, fuzzy temporal models \cite{schockaert08} exist. %Some fuzzy temporal models assume that the time stored in the database is an interval. The temporal interval is represented by two ill-known time points: $X$  an ill-known starting point and $Y$ an ill-known ending point. The interval $\left[X,Y\right]$ is not a fuzzy interval but an ill-known interval: it is a crisp interval but it is partially unknown which values are in this interval.

%Database models can then be classified into transaction time \cite{Jensen91}, decision time, valid time, bi-temporal (both valid and transaction time) \cite{Snodgrass84} or tri-temporal (valid, transaction and decision time) \cite{Nascimento95} models.

%\subsection{\label{subsec:possibility-theory}Possibility Theory}
%Possibility theory, like probability theory, deals with uncertainty about the outcome of an experiment. In probability theory, this uncertainty is caused by the \emph{variability} in the outcomes, while in possibility theory, the uncertainty is caused by \emph{incomplete knowledge} about the experiment. The quantification of confidence in a theory of uncertainty is achieved using a confidence measure\cite{Shafer:1976:AMathematical}. In probability theory this is a measure of chance, in possibility theory, possibility and necessity measures are used.

%\begin{definition}
%Consider a set of outcomes $\Omega$. Let $\wp(\Omega)$ denote the powerset of $\Omega$ and let $A$ and $B$ be elements of $\wp(\Omega)$. A \emph{confidence measure on $\Omega$} is defined by a function
%	\begin{align}
%	g : \wp(\Omega) & \rightarrow \left[0,1\right]
%	\end{align}
%that satisfies
%	\begin{align}
%	g(\emptyset) &= 0 \\
%	g(\Omega) &= 1 	\label{NormalizationProperty} \\
%	A \subseteq B &\Rightarrow g(A) \leq g(B) \label{MonotonicityProperty}
%	\end{align}
%\end{definition}

%Both possibility measures and necessity measures are special cases of confidence measures.

%\begin{definition}
%Consider a confidence measure $\Pi$ on a set of outcomes $\Omega$. Let $J$ be a countable index set and let $\{ A_{j} | j \in J \wedge A_{j} \subseteq \Omega \}$ be a family of elements of $\wp(\Omega)$. $\Pi$ is now a \emph{possibility measure on $\Omega$} if it satisfies:
%	\begin{align}
%	\Pi\left(\bigcup_{j \in J} A_{j} \right) = \sup_{j \in J} \Pi(A_{j})
%	\end{align}
%\end{definition}

%In this work, the interpretation is as follows. The possibility of an event expresses how plausible the occurrence of the event seems to an observer of the experiment, given the (partial) knowledge of the observer about the experiment.

%Information on the possibility of distinct elements of the universe of discourse $\Omega$ can now be given by a \emph{possibility distribution} $\pi$ on $\Omega$, defined by:

%\begin{definition}
%Consider a possibility measure $\Pi$ on $\Omega$. A \emph{possibility distribution} $\pi$ on $\Omega$ underlying the possibility measure $\Pi$ is then a function defined by:
%	\begin{align}
%	\pi : \Omega \rightarrow \left[0, 1\right] : \pi(u) = \Pi(\{u\})
%	\end{align}
%\end{definition}

%\begin{definition}
%Consider a confidence measure $N$ on a set of outcomes $\Omega$. Let $J$ be a countable index set and let $\{ A_{j} | j \in J \wedge A_{j} \subseteq \Omega \}$ be a family of elements of $\wp(\Omega)$. $N$ is now a \emph{necessity measure} on $\Omega$ if it satisfies:
%	\begin{align}
%	N\left(\bigcap_{j \in J} A_{j} \right) = \inf_{j \in J} N(A_{j})
%	\end{align}
%\end{definition}

%In this work, the interpretation is as follows. The necessity of an event expresses how necessary the occurrence of the event seems to an observer of the experiment, given the (partial) knowledge of the observer about the experiment.

%Possibility and necessity measures are dual in the sense that:

%\begin{align}
%\forall A \subseteq \Omega : N(A) = 1 - \Pi(\bar{A})
%\end{align}

%Regarding interpretation, the above can be seen as: the degree to which an event is necessary is the degree to which every other possible event is not plausible.

\subsection{\label{subsec:possibilistic-variables}Possibilistic Variables and Ill-known Values, Sets and Intervals}
In this section, possibilistic variables and the concepts of ill-known values, ill-known sets and ill-known intervals are introduced, based on \cite{Pon11}. These concepts rely heavily on possibility theory \cite{Dubois:Prade:1988:PossibilityTheory} and its concepts like `possibility' and `possibility distribution'. In this work, `possibility' is always interpreted as a measure of plausibility. A \emph{possibilistic variable} is defined as follows \cite{Pon11}.

%It is assumed that the reader is familiar with possibility theory. For more information on possibility theory, the reader is referred to \cite{Dubois:Prade:1988:PossibilityTheory}.

\begin{definition}
A \emph{possibilistic variable} $X$ on a universe $U$ is defined as a variable taking exactly one value in $U$, but for which this value is (partially) unknown. The variable's possibility distribution $\pi_X$ on $U$ models the available knowledge about the value that $X$ takes: for each $u \in U$, $\pi_X(u)$ represents the possibility that $X$ takes the value $u$. This possibility is interpreted as a measure of plausibility and thus as a measure of how plausible it is that $X$ takes the value $u$, given (partial) knowledge about the value $X$ takes.
\end{definition}
Consider a set $R$ containing single values (and not collections of values). When a possibilistic variable $X_{v}$ is defined on such a set $R$, the unique value $X_{v}$ takes, which is (partially) unknown, is called an \emph{ill-known value} in this work \cite{Dubois88b}.

Now, consider a set $R$ containing single values and its powerset $\wp(R)$. When a possibilistic variable $X_{s}$ is defined on the powerset $\wp(R)$ of such a set $R$, the unique value $X_{s}$ takes will be a crisp set and the possibility distribution $\pi_{X_{s}}$ of $X_{s}$ will be a possibility distribution on $\wp(R)$. This $\pi_{X_{s}}$ will define the possibility of each value of $\wp(R)$ (a value of $\wp(R)$ is a crisp subset of $R$) being the value $X_{s}$ takes. This exact value $X_{s}$ takes, is called an \emph{ill-known set} \cite{Dubois88b}.

Consider a set $R$ containing single values and its powerset $\wp(R)$. Now consider a subset $\wp_{I}(R)$ of $\wp(R)$ and let this subset contain every element of $\wp(R)$ that is an interval, but no other elements. When a possibilistic variable $X_{i}$ is defined on the subset $\wp_{I}(R)$ of the powerset $\wp(R)$ of some set $R$, the unique value $X_{i}$ takes will be a crisp interval and the possibility distribution $\pi_{X_{i}}$ of $X_{i}$ will be a possibility distribution on $\wp_{I}(R)$. This $\pi_{X_{i}}$ will define the possibility of each value of $\wp_{I}(R)$ (a value of $\wp_{I}(R)$ is a crisp interval in $R$) being the value $X_{i}$ takes. This exact value the variable takes, is called an \emph{ill-known interval} here.

In this work, another approach to defining and describing an ill-known interval is used. Here, an ill-known interval $I$ is defined and described by its start and end point, which are ill-known values. Thus, an ill-known interval is seen as an interval of which the exact start and end point are (partially) unknown, which implies that the interval itself is (partially) unknown. Thus, the start and end point of an ill-known interval are mutually independent ill-known values, which are defined by mutually independent possibilistic variables.

In intentions, both approaches are the same: they attempt to model a single interval for which some uncertainty exists about which values are in the interval and which are not. The approach mentioned first does this by defining the possibility that an interval is the meant interval, for every interval in the considered interval set, whereas the second approach describes the possibility of a single point being the start point of the meant interval and the possibility of a single point being the end point of the meant interval, for every point imaginable. The actual ill-known interval is then inferred from these start and end points.

Whatsoever, the interactions and behaviors of these representations and the correspondences, interactions and transformations between them are part of the current research of the authors.

%describes a set of values that is certainly in the interval and uncertainty about a set of others (near the start and end point). Thus, this last approach contains more information: it gives a set of values that are certainly in the interval...}

%It is important to understand the difference between the following two concepts:
%\begin{itemize}
%\item
%A \emph{possibilistic variable} $X$ is bounded to take only one value , but this value is not known due to incomplete knowledge. 
%\item
%An \emph{ill-known set}~\cite{Dubois88b}: a possibilistic variable defined over the universe $\Pow(U)$.
%\end{itemize}

%Note that while a possibilistic variable refers to one (partially) unknown value, an ill-known set is a crisp set but, for some reason, (partially) unknown.
A specific application of possibilistic variables is obtained when the universe under consideration is the set of boolean values, denoted $\mathbb{B}$ = $\{T,F\}$, where $T$ denotes `true' and $F$ denotes `false' \cite{Pon11}. Indeed, any boolean proposition $p$ takes exactly one value in $\mathbb{B}$. If the knowledge about which value this proposition $p$ takes is given by a possibility distribution $\pi_p$, proposition $p$ can be seen as a possibilistic variable. As the interest lies with the case where the proposition holds (denoted $p$ = $T$), the possibility and necessity that $p$ = $T$ demand most attention. In this work, the following notations are used:

\vspace{-15pt}
\begin{align}
\text{Possibility that $p$ = $T$:} \hspace{50pt} & Pos(p) = \pi_p(T) \label{propholdsposs} \\
\text{Necessity that $p$ = $T$:} \hspace{50pt} & Nec(p) = 1-\pi_p(F) \label{propholdsnecc}
\end{align}

%Here, notation \ref{propholdsposs} denotes the possibility that $p$ = $T$ and the proposition holds, notation \ref{propholdsnecc} denotes the necessity that $p$ = $T$ and the proposition holds.

%This work will deal with ill-known intervals. These are ill-known sets, defined and represented via a start and end point, which will be ill-known values. The elements of the set are the values between the start and end point. An closed ill-known interval with start point $X$ and end point $Y$ is noted here $\left[X, Y\right]$. The correspondences and transitions between the representations of ill-known sets and between the representation of an ill-known set and an ill-known interval are part of the authors current research.

%\subsection{\label{subsec:fuzzy-numbers}Fuzzy Numbers and Fuzzy Intervals}
%Among others, Dubois and Prade~\cite{Dubois1983} use the following definition of a \emph{fuzzy interval}.
%\begin{definition}
%A fuzzy interval is a fuzzy set $M$, defined by a membership function $\mu_{M}$, on the set of real numbers $\mathbb{R}$ such that:
%\begin{eqnarray}
%\mu_{M} : & \!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\! \mathbb{R} \rightarrow \left[0,1\right] \nonumber \\ 
%\forall (u,v)\in\mathbb{R}^2: \forall w \in [u,v]:&\mu_M(w) \geq\min(\mu_M(u),\mu_M(v))  \\
%\exists m \in \mathbb{R} : & \!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\! \mu_M(m)=1 
%\end{eqnarray}
%\end{definition}
%If this modal value $m$ is unique, then $M$ is referred to as a \emph{fuzzy number}. In other words, if the core of a fuzzy interval is a singleton, it is referred to as a fuzzy number.

%A simple form of the membership function of a fuzzy interval is a trapezoidal function. It can be shown that such a membership function $\mu_T$ for a fuzzy interval $T$ is convex and normalized. A visualization is given in figure \ref{fig:trapezoidal}. Four reel values, denoted $\alpha$, $\beta$, $\gamma$ and $\delta$ and chosen as in figure \ref{fig:trapezoidal}, suffice to represent a trapezoidal membership function of a fuzzy interval. In this work, a fuzzy interval defined as such will be noted as $\left[\alpha, \beta, \gamma, \delta\right]$. %The corresponding membership function definition for this $\mu_T$ is then given by:

%\begin{align}
%\mu_T : & \quad \mathbb{R} \rightarrow \left[0,1\right] \\
% : & \quad x \rightarrow
%\begin{cases}
%1 & \mbox{ if } x \in [\beta,\gamma] \\
%0 & \mbox{ if } x > \delta \vee x < \alpha \\
%\frac{x-\alpha}{\beta - \alpha} & \mbox{ if } x \in [\alpha,\beta[ \\
%\frac{\delta -x}{\delta - \gamma} & \mbox{ if } x \in ]\gamma,\delta] \\
%\end{cases}
%\end{align}

%\def\JPicScale{0.5}
%\begin{figure}[h!]
%  \centering
%  \input{./graphs/trapezoidal.tex}
%  \caption{Trapezoidal membership function}
%  \label{fig:trapezoidal}
%\end{figure}

%The most convenient form of the membership function of a fuzzy number is a triangular function. It can be shown that such a membership function $\mu_M$ for a fuzzy number $M$ is convex and normalized. A visualization is given in figure \ref{fig:triangular}. Three reel values, denoted $a$, $b$ and $D$, suffice to represent a triangular membership function of a fuzzy number and in this work, a fuzzy number defined as such will be noted as $\left[D, a, b \right]$. Here:
%\begin{itemize}
%\item
%$D$ denotes the single value in the core of $M$
%\item
%$D-a$ is then $\inf \{u \in \mathbb{R} : \mu_{M}(u) > 0\}$
%\item
%$D+b$ is then $\sup \{u \in \mathbb{R} : \mu_{M}(u) > 0\}$
%\end{itemize}
%\begin{figure}[h!]
%  \centering
%  \input{./graphs/triangular.tex}
%  \caption{Triangular membership function.}
%  \label{fig:triangular}
%\end{figure}

%\subsubsection{Set evaluation by ill-known constraints}

\subsection{Interval Evaluation by Ill-known Constraints}
The problem of interval evaluation is more generally explained in \cite{Pon11}: basically, the need exists to check how all points in a crisp set are positioned with respect to one or more ill-known values.

In \cite{Pon11}, the notion of an ill-known constraint is introduced:

\begin{definition}
Given a universe $U$, an \emph{ill-known constraint} $C$ is specified by means of a binary relation $R \subseteq U^{2}$ and a fixed, ill-known value defined by its possibilistic variable $V$ on $U$, i.e.:
\begin{align}
C \triangleq (V,R)
\end{align}
Some set $A \subseteq U$ now satisfies this constraint $C$ if and only if:
\begin{align}
\forall a \in A : (V,a) \in R
\end{align}
\end{definition}

An example of an ill-known constraint is $C_{<} \triangleq (X, <)$. Some set $A$ then satisfies $C_{<}$ if $\forall a \in A : X <a$.

The satisfaction of a constraint $C \triangleq (V,R)$ by a set $A$ is basically a Boolean matter and can thus be seen as a boolean proposition, but due to the uncertainty inherent to the ill-known value $V$, it can be uncertain whether $C$ is satisfied by $A$ or not \cite{Pon11}. Based on the possibility distribution $\pi_{V}$ of $V$, the possibility and necessity that $A$ satisfies $C$ can be found. This proposition can thus be seen as a possibilistic variable on $\mathbb{B}$. The required possibility and necessity are:

\vspace{-10pt}

\begin{align}
Pos(A\text{ satisfies }C) & = \min_{a \in A}\left(\sup_{(w,a) \in R}\pi_{V}(w))\right) \label{eq:pos}\\
Nec(A\text{ satisfies }C) & = \min_{a \in A}\left(\inf_{(w,a) \notin R} 1-\pi_{V}(w)\right) \label{eq:nec}
\end{align}

Now, to calculate the possibility or necessity of a set $A$ satisfying multiple constraints, the $\min$ t-norm operator is used. For example:

\vspace{-10pt}

\begin{align}
Pos((A\text{ satisfies }C_{1}) \text{ and } (A\text{ satisfies }C_{2})) & = \nonumber\\
 \min_{a \in A}(Pos(A\text{ satisfies }C_{1}), Pos(A\text{ satisfies }C_{2})) \nonumber \\
Nec((A\text{ satisfies }C_{1}) \text{ and } (A\text{ satisfies }C_{2})) & = \nonumber\\
 \min_{a \in A}(Nec(A\text{ satisfies }C_{1}), Nec(A\text{ satisfies }C_{2})) \nonumber
\end{align}

%\vspace{-10pt}

%Now, to check if crisp interval $I = \left[j, k\right]$ is included in $\left[X, Y\right]$, 2 ill-known constraints are constructed:

%We assume that $X$ specifies the lower bound and $Y$ the upper bound for a given interval, we want to known whether all points in the interval are larger than or equal to $X$ and smaller than or equal to $Y$. Therefore, we consider two ill-known constraints:

%\begin{eqnarray}
%C_1\triangleq\left(\geq,X\right)\\
%C_2\triangleq\left(\leq,Y\right).
%\end{eqnarray}

%The possibility and necessity of the set satisfying both constraints is then:

%\begin{align}
%\label{eq:interval-pos}
%Pos(A\text{ satisfies }C_1\ AND\ C_2) & = \min_{a \in A}\left(\sup_{a \geq w}\pi_{X}(w),\sup_{a \leq v}\pi_{Y}(v)\right)\\
%\label{eq:interval-nec}
%Nec(A\text{ satisfies }C_1\ AND\ C_2) & = \min_{a \in A}\left(\inf_{a < w} 1-\pi_{X}(w),\inf_{a > v} 1-\pi_{Y}(v)\right).
%\end{align}

%\paragraph{Example} Consider the ill-known values $X = \left[5, 2, 8\right]$ and $Y = \left[9, 7, 10 \right]$. The knowledge about the evaluation of the interval $\left[a, b \right]$  is given by the expressions \eqref{eq:interval-pos},\eqref{eq:interval-nec}.  Figure~\ref{fig:3d-possibility} shows a 3D plot of the possibility that an interval $[a,b]$ passes the evaluations specified by the ill-known constraints. Note the triangular form for the resulting possibility distribution since the condition $a \leq b$ holds.
%
%\begin{figure}[h!]
%\centering
%\includegraphics[scale=0.4]{graphs/3D_possibility.eps}
%\caption{Possibility of evaluation for the interval $[a,b]$.}
%\label{fig:3d-possibility}
%\end{figure}
%The necessity plot is obtained in a similar way and is shown in Figure~\ref{fig:3d-necessity}. Notice that the necessity measure is not normalized because the supports of $X$ and $Y$ overlap.
%\begin{figure}[h!]
%\centering
%\includegraphics[scale=0.4]{graphs/3D_necessity.eps}
%\caption{Necessity of evaluation for the interval $[a,b]$.}
%\label{fig:3d-necessity}
%\end{figure}

### 2012-04-10 trunk/ipmu2012_2/proposal.tex
Part of the paper

In this section, the proposal of this paper is presented. The proposal consists of an approach for querying a valid-time relation \cite{Dyreson1994}. The novelty of this approach is that it allows uncertainty in the temporal demand in the query. To be able to present the query results to the user, this section also presents an approach for ranking the records in the queried relation, integrating the consequences of the query uncertainty into the ranking.

To illustrate the proposal of this paper, an example will be presented and examined in the course of the proposal presentation.

%This section presents the proposal of this paper. The proposal consists of an approach for querying a relational valid-time database, in which this approach allows uncertainty in the query, and an approach for ranking the results of such queries, respecting the consequences of the uncertainty in the query

\subsection{The Valid-Time Relation}
An approach for querying a valid-time relation \cite{Dyreson1994} is presented. Here, such a relation should contain one or more attributes, called valid-time attributes \cite{Dyreson1994}, specifying exactly one valid-time interval \cite{Dyreson1994} for each record. This valid-time interval should be a crisp interval in the temporal domain, representing the uninterrupted period of time during which the object or concept represented by the corresponding record is real or true in the modelled reality.

An example relation is visualized in table \ref{table:example}. Every record represents a certain rental car owned by a car rental service, in a certain state. The attributes `Startdate' and `Enddate' describe the dates on which a time interval starts respectively ends (this interval includes the start- and enddates themselves) during which a car represented by a record, in the state represented by this record, is available for rent. The chronons are days. The attribute `Mileage' describes the amount of kilometers the car has already driven. The attribute `Color' describes the color of the car. Every unique value for the `ID' attribute corresponds to a unique physical car. For the same car, every different value for attribute `IID' corresponds to a different state of the car. %For example, the first and last records describe the same physical car (because their value for the attribute `ID' is the same), though in the first record, the car resides in Catania and in the last record, it resides in Bruges.
\vspace{-15pt}
\begin{table}
\caption{An example relation containing car rental information. Mileage is in kilometers, temporal data is in dd/mm/yyyy format.}
\centering
\begin{tabular}{c c c c c c}
\hline
\textbf{ID} & \textbf{IID} & \textbf{Color} & \textbf{Mileage} & \textbf{Startdate} & \textbf{Enddate} \\
\hline
001 & 1 & red & 20345 & 15/06/2012 & 14/08/2012 \\
002 & 1 & blue & 23404 & 10/06/2012 & 10/08/2012 \\
003 & 1 & blue & 25340 & 30/06/2012 & 30/08/2012 \\
004 & 1 & blue & 33367 & 15/06/2012 & 31/07/2012 \\
001 & 2 & red & 42420 & 15/08/2012 & 14/09/2012 \\
\hline
\end{tabular}
\label{table:example}
\end{table}
\vspace{-25pt}

%The database is a standard relational valid-time database. To focus attention, we describe: it's just a normal relational database, but every relation can have a valid-time attribute. If a relation has a valid-time attribute, the entries in it are crisp time intervals.

\subsection{The Query}
In this subsection, the proposed querying approach is presented. First, the query structure is presented, then the specific approach to query evaluation, followed by a ranking method.

\subsubsection{Query Structure}
In the presented approach, the user's query demands may consist of several non-temporal demands and a single temporal demand. Non-temporal demands are demands concerning any attribute that is not a valid-time attribute, whereas the temporal demand concerns valid-time. The interpretation is that the user queries the relation for records satisfying both the non-temporal demands and the temporal demand.

Thus, in this framework, every query $Q$ consists of two parts:

\vspace{-10pt}
\begin{align}
Q & = (Q_{T}, Q_{N})
\end{align}

Here, $Q_{N}$ denotes the collection of (possibly fuzzy) non-temporal user preferences. $Q_{T}$ denotes the temporal demand. A temporal demand $Q_{T}$ consists of two parts:

\vspace{-20pt}
\begin{align}
Q_{T} & = (AR, IKI)
\end{align}

Here, $AR$ denotes an Allen relation \cite{Allen83} and $IKI$ denotes an ill-known interval in the valid-time domain used by the valid-time attributes of the relation. The interpretation of this temporal demand is that the user prefers a record which represents an object that is valid in the modelled reality during a time interval related to $IKI$. The nature of this relation is given by $AR$. The usage of an ill-known interval in the temporal demand allows for some uncertainty in the user's temporal preference: the user can query the relation searching for records representing objects valid during time intervals related by an Allen relation to a time interval that is partially unknown to the user.

As mentioned in \cite{Pon11}, this approach differs from the one where a valid-time interval is represented by one fuzzy set. Such a fuzzy set is seen as a possibility distribution on some time domain $\mathbb{N}^{*}$ and thus defines just one ill-known value. However, in the presented approach, a time interval is represented by an ill-known interval, which is defined by two possibility distributions on this $\mathbb{N}^{*}$, describing start and end point of the interval. As explained in section \ref{sec:preliminaries}, this ill-known interval can be seen as defined by a possibility distribution on $\wp(\mathbb{N}^{*})$.

An example is in order: consider a user who wants to rent a car during approximately the whole of july, preferably a blue one with a mileage below 30000 km. Due to schedual issues and possible unforseen conditions, the user is not certain when exactly the rental car will be needed and thus on which exact day the rental car should start and end being available. These preferences can be translated into query $Q_{ex}$:

\vspace{-10pt}
\begin{align}
Q_{ex} & = (Q_{T,ex}, Q_{N,ex}) \nonumber\\
Q_{T,ex} & = (AR_{ex}, IKI_{ex}) \nonumber
\end{align}

Here, $Q_{N,ex}$ will denote that the car must be blue and have a mileage below 30000 km. $AR_{ex}$ here denotes the Allen relation `during'. Now to define $IKI_{ex}$, a definition of its start and end points are necessary. These start and end points are ill-known values, defined by possibilistic variables $S$ respectively $E$, which are defined by their possibility distributions $\pi_{S}$ respectively $\pi_{E}$. To define these possibility distributions, a translation of the notion `approximately July' is necessary. This translation can be given by the user or suggested or constructed by the system. Imagine the user expressing that he will certainly not need the car on or before the 29th of June or on or after the 2nd of August. Also, the possibility that the user will need the car from the 1st of July to the 31th of July, boundaries included, is the highest. Lastly, the possibility that the user will need the car between the described periods varies linearly. The resulting possibility distributions are given by the equations below. In these equations, $\mathbb{T}$ is the set of all days in time, always denoted in dd/mm/yyyy format.

\vspace{-10pt}
\begin{table}
\centering
\begin{tabular}{c l c c l}
$\pi_{S}$ & : $\mathbb{T} \rightarrow \left[0, 1\right]$ & \hspace{10pt} & $\pi_{E}$ & : $\mathbb{T} \rightarrow \left[0, 1\right]$ \\
 & : $x \rightarrow 0, \hspace{27pt} x \leq 29/06/12$ & & \hspace{10pt} & : $x \rightarrow 1, \hspace{27pt} x \leq 31/07/12$ \\
 & : $x \rightarrow 0.5, \hspace{20pt} x = 30/06/12$ & & \hspace{10pt} & : $x \rightarrow 0.5, \hspace{20pt} x = 1/08/12$ \\
 & : $x \rightarrow 1, \hspace{27pt} x \geq 1/07/12$ & & \hspace{10pt} & : $x \rightarrow 0, \hspace{27pt} x \geq 2/08/12$ \\

\end{tabular}
\label{tab:eqns}
\end{table}

%\begin{align}
%\pi_{S} & : \mathbb{T} \rightarrow \left[0, 1\right] \nonumber \\
% & : x \rightarrow 0, \hspace{28pt} x \leq 29/06/12 \nonumber \\
% & : x \rightarrow 0.5, \hspace{20pt} x = 30/06/12 \nonumber \\
% & : x \rightarrow 1, \hspace{28pt} x \geq 1/07/12 \nonumber \\
%\pi_{E} & : \mathbb{T} \rightarrow \left[0, 1\right] \nonumber \\
% & : x \rightarrow 1, \hspace{28pt} x \leq 31/07/12 \nonumber \\
% & : x \rightarrow 0.5, \hspace{20pt} x = 1/08/12 \nonumber \\
% & : x \rightarrow 0, \hspace{28pt} x \geq 2/08/12 \nonumber
%\end{align}

%DUS: iki met 2 grenspunten: schrijf functies grenspunten neer. Leg uit dat IKI daarmee gedefinieer is.

% (IN TIME? MAYBE WORK WITH PVPS AGAIN?)

%ELABORATE ON TEMPORAL CONSTRAINT: EXPLAIN WHAT IT MEANS, IN WHICH SITUATIONS IT CAN BE USEFUL, ...

%EXPLAIN THAT ONE CAN SET A TEMPORAL CONSTRAINT, BUT JUST ONE...  The temporal demand is a single constraint concerning valid-time.

%In this approach, the user's query demands may consist of several non-temporal demands and a single temporal demand. Non-temporal demands are demands concerning any attribute that is not a valid-time attribute. The temporal demand is a single constraint

%Query with two parts: normal things and temporal part. Temporal part is made of Allen relation and ill-known interval. Thus specified by 2 poss dist. How this is done is whatever (if enough place: launch the triangular stuff). 
%Interpretation of query.

%Running example.
\vspace{-20pt}
\subsubsection{Query Evaluation}
The evaluation of a crisp query for a record in a regular (relational) database results in the accepting or rejecting of the record as a part of the result set presented to the user. In fuzzy querying, query satisfaction modelling is a matter of degree, as the evaluation of a fuzzy query for a record usually results in some satisfaction degree $s$, where $s \in \left[0,1\right]$, where 0 denotes total dissatisfaction and 1 denotes complete satisfaction. Now, the evaluation of a crisp query can also be modelled using similar satisfaction degrees, by assigning rejection a degree of $0$ and acceptance a degree of $1$ and not using any other value in $\left[0,1\right]$.

In this work, the evaluation of a query $Q$ = $(Q_{T}, Q_{N})$, with $Q_{T}$ = $(AR, IKI)$ is handled as follows. For each record $r$ in the database, two things happen independently:

\vspace{-5pt}
\begin{itemize}
	\item The non-temporal preferences in $Q_{N}$ are evaluated. This results in a satisfaction degree $e_{Q_{N}}(r)$. The presented model accepts any valid and consistent way of calculating this evaluation, as long as $e_{Q_{N}}(r) \in \left[0,1\right]$.
	\item The temporal demand in $Q_{T}$ is evaluated. Depending on $AR$ and $IKI$, a specific set of ill-known constraints is constructed, which can be found in table \ref{table:ikc}. Based on equations \eqref{eq:pos} and \eqref{eq:nec} and using the $\min$ operator for aggregation, formulas are calculated to determine the possibility $Pos_{Q^{T}}(r) \in \left[0,1\right]$ and necessity $Nec_{Q^{T}}(r) \in \left[0,1\right]$ that $r$ fulfills all these constraints. Because these formulas only depend on $AR$ and $IKI$, they only need to be constructed once and not for every considered record.
\end{itemize}

\vspace{-15pt}
\begin{table}
\caption{The Allen relations with corresponding expressions for ill-known constraints. $A$ denotes a crisp interval. In the specification of the ill-known constraints, it is assumed that the corresponding ill-known interval $IKI$ is defined by a start point and an end point, respectively defined by possibilistic variables S and E.}
\centering
\begin{tabular}{c|c}
\hline
\textbf{Allen Relation }& \textbf{Constraints }\\
\hline

IKI before A & $C_1 \triangleq \left(E,<\right)$ \\
\hline

IKI equal A & $C_1\triangleq \left(S,\leq\right)$ $\wedge$ $C_2\triangleq \left(S, \neq \right)$ $\wedge$ $C_3\triangleq \left(E,\geq\right)$ $\wedge$ $C_4\triangleq \left(E,\neq\right)$ \\
\hline

IKI meets A & $C_1\triangleq \left(E,\leq\right)$ $\wedge$ $\neg C_2\triangleq \left(E, \neq\right)$ \\
\hline

IKI overlaps A & $C_1\triangleq \left(S,<\right)$ $\wedge$ $\neg C_2\triangleq \left(E, \leq\right)$ $\wedge$ $\neg C_3\triangleq \left(E, \geq\right)$ \\
\hline

IKI during A & $\neg C_1\triangleq \left(S, \leq \right)$ $\wedge$ $\neg C_2\triangleq \left(E, \geq \right)$ \\
\hline

IKI starts A & $C_1\triangleq \left(S, \leq \right)$ $\wedge$ $\neg C_2\triangleq \left(S, \neq\right)$ $\wedge$ $\neg C_3\triangleq \left(E, \geq \right)$ \\
\hline

IKI finishes A & $C_1\triangleq \left(E, \geq \right)$ $\wedge$ $\neg C_2\triangleq \left(E, \neq\right)$ $\wedge$ $\neg C_3\triangleq \left(S, \leq\right)$ \\

\hline
\end{tabular}
\label{table:ikc}
\end{table}

%In fuzzy querying of regular (relational) databases, the modelling of query satisfaction is a matter of degree. Usually, the evaluation of the query requirements for a record results in a satisfaction degree $s$, where $s$ lies in $\left[0,1\right]$, where 0 denotes total dissatisfaction and 1 denotes complete satisfaction. In crisp querying, the evaluation of query requirements for a record results in the accepting or rejecting of the record as a part of the result set. This can be modelled using satisfaction degrees, by assigning rejection a degree of $0$ and acceptance a degree of $1$ and not using any other value in $\left[0,1\right]$.

%First of all, non-temporal part is evaluated. Satisfaction degree in unit interval is obtained. Explain interpretation if necessary. Give example
%Next, temporal part is evaluated, resulting in possibility and necessity (explain!) degree. Explain interpretation. Give example. Explain that calculation does not need to be repeated => speed!

\subsubsection{Ranking}
As a final step, every record $r$ is given a final rank $e_{final}(r)$ depending on $e_{Q_{N}}(r)$ and a value $e_{Q_{T}}(r)$ based on both $Pos_{Q^{T}}(r)$ and $Nec_{Q^{T}}(r)$. For every record $r$, this is done as follows.

First, $e_{Q_{T}}(r)$ is calculated using the expression in equation \eqref{eq:finalrank}.

\vspace{-10pt}
\begin{align}
e_{Q_{T}}(r) & = \frac{Pos_{Q^{T}}(r) + Nec_{Q^{T}}(r)}{2} \label{eq:finalrank}
\end{align}

%First, $e_{Q_{T}}(r)$ is $0$. If $Pos_{Q^{T}}(r)$ is exactly $1$, then $1$ is added to $e_{Q_{T}}(r)$ and the value of $Nec_{Q^{T}}(r)$ is added to $e_{Q_{T}}(r)$. Else, only the value of $Pos_{Q^{T}}(r)$ is added to $e_{Q_{T}}(r)$. Because both necessity and possibility are in the unit interval, this gives a value between $0$ and $2$. This value is now normalized by dividing it by $2$.

Because necessity cannot exceed $0$ unless possibility is $1$ and $Pos_{Q^{T}}(r) \in \left[0,1\right]$ and $Nec_{Q^{T}}(r) \in \left[0,1\right]$, the sum in the numerator gives a natural ranking score in $\left[0,2\right]$. The function of the denominator is to normalize this score to a value in $\left[0,1\right]$. The final ranking $e_{final}(r)$ is now given by a convex combination:

\vspace{-10pt}
\begin{equation}
\label{eq:convex-comb}
e_{final}(r)\ =\ \omega*e_{Q_{N}}(r)\ +\ (1-\omega)*e_{Q_{T}}(r),\ \omega \in \left[0, 1 \right]
\end{equation}

The use of this convex combination allows a record to make up for a low score for the temporal constraint by a good score for the non-temporal constraint (or vice versa). Changing $\omega$ also allows granting the temporal constraint more weight with respect to the non-temporal constraint (or vice versa). The result of the evaluation and ranking steps on the example are shown in table \ref{table:final}. With $\omega$ = $0.5$, it is clear that both non-temporal and temporal criteria have the same importance and the final ranking is natural.

\vspace{-15pt}
\begin{table}
\caption{Scores and final ranking for the example records, using $\omega = 0.5$.}
\centering
\begin{tabular}{c c c c c c c}
\hline

\textbf{ID} & \textbf{IID} & \textbf{$e_{Q_{N}}(r)$} & \textbf{$Pos_{Q^{T}}(r)$} & \textbf{$Nec_{Q^{T}}(r)$} & \textbf{$e_{Q_{T}}(r)$} & \textbf{$e_{final}(r)$} \\
\hline
001 & 1 & 0 & 1 & 1 & 1 & 0.5 \\
002 & 1 & 1 & 1 & 1 & 1 & 1 \\
003 & 1 & 1 & 0.5 & 0 & 0.25 & 0.625 \\
004 & 1 & 0 & 1 & 1 & 1 & 0.5 \\
001 & 2 & 0 & 0 & 0 & 0 & 0 \\

\hline
\end{tabular}
\label{table:final}
\end{table}
\vspace{-25pt}
%The proposal consist on a possibilistic valid-time model. The representation and the querying are explained in the following subsections.

%\subsection{Representation of Ill-known Valid-time Intervals}
%\label{subsec:representation-ill-known}
%Valid time is usually represented as an interval. The interval has a starting and an ending points. An ill-known valid-time interval is an interval in witch one or both points are ill-known.

%\begin{definition}
%A Possibilistic Valid-Time Period \textbf{PVP} is a possibilistic interval defined by means of two ill-known points, namely $\left[ X,\ Y \right]$
%\begin{equation}
%PVP = \left[X,\ Y \right] 
%\end{equation}
%$X$ and $Y$ are ill-known values in the set of the real numbers $\mathbb{R}$. The uncertainty about the values taken by $X$ and $Y$ are given by the possibility distributions $\pi_X$ and $\pi_Y$.
%\end{definition}

%For convenience, the possibility distributions $\pi_X$ and $\pi_Y$ are given in the way of a triangular distribution, as explained in subsection \ref{subsec:fuzzy-numbers}. This representation allows overlapping (Fig. \ref{fig:pvp}). Note that while the two ill-known values $X,Y \in \mathbb{R}$, the fuzzy interval  $[X,Y] \in \mathbb{R}^2$.

%\begin{figure}[h!]
%  \centering
%  \input{./graphs/2_triangular.tex}
%  \caption{Two fuzzy numbers $X$ and $Y$ denoting a Possibilistic Valid-Time Period \emph{PVP}.}
%  \label{fig:pvp}
%\end{figure}

%\subsection{Storage of Valid-time Intervals}
%\label{subsec:storage}
%Each database row containing a \emph{PVP} stores it as two triangular possibility distributions. In our approach we propose the representation of that as proposed in the  fuzzy interface for relational databases \emph{FIRST}~\cite{Medina94gefred.a,Gal98}. In this representation it is also possible to represent not only fuzzy numbers but fuzzy constants (see table \ref{table:relational-representation-pvp}):

%\begin{itemize}
%\item
%\emph{NULL}: This constant refers to a completely ignorance about the value. The possibility distribution for a given fuzzy number $X$ is not defined, therefore, any comparison between a fuzzy number and the \emph{NULL} constant always returns $0$.
%\item
%\emph{UNKNOWN}: The point has a value but it is unknown.% The possibility distribution for a given fuzzy number $X$ is $\pi_X=1$
%\item
%\emph{UNDEFINED}: The point does not have a value. %The possibility distribution for a given fuzzy number $X$ is $\pi_X=0$
%\end{itemize}

%\begin{table}
%\caption{Relational representation for an ill-known time point. A \emph{PVP} is represented by two ill-known points.}
%\centering
%\begin{tabular}{c c c c c c c}
%\hline
%Value & FT & F1 & F2 & F3 & $\mu(x)$ \\ \hline
%UNKNOWN & 0 & NULL & NULL & NULL  & not defined. \\ 
%UNDEFINED & 1 & NULL & NULL & NULL & $0$ \\ 
%NULL & 2 & NULL & NULL & NULL & $1$  \\ 
%$\left[D,\ a,\ b \right]$ & 3 & $D$ & $D-a$ & $D+b$ & $\mu(\left[D,\ a,\ b \right])$ \\ 
%\hline
%\end{tabular}
%\label{table:relational-representation-pvp}
%\end{table}

%\subsection{Querying Ill-known Valid-time Intervals}
%In order to provide a complete model, we will provide the tools for querying. This allows the user to specify both the preferences and an ill-known valid-time interval in the query. It is important to notice that the possibilistic / fuzzy data stored in the database has a \emph{disjunctive interpretation} (it is said that we have \emph{uncertainty}: the valid-time interval has only one value but, for some reason the value is ill-known). In the query specification, the user is allowed to express a crisp time interval. 
%In this subsection we will define the query specification, then the evaluation of the query and finally the ranking for the query.

%\subsubsection{Query specification}
%A query in our framework has two different parts: the first one is the temporal specification. The second part is the query specification for regular attributes.

%\begin{definition}
%A query $\tilde Q$ is specified by:
%\begin{equation}
%\label{eq:query-definition}
%\tilde Q = \left( Q^{time}, Q \right)
%\end{equation}
%\end{definition}
%Where $Q$ are the (possibly fuzzy) preferences of the user  and $Q^{time}$ is the temporal part specified by a crisp interval. The temporal part is composed by two arguments:

%\begin{definition}
% $Q^{time}$ is composed by:
%\begin{equation}
%Q^{time} = \left( \left[a, \ b \right] , Ar \right), a,b \in \mathbb{R}
%\end{equation}
%Where $ \left[a, \ b \right] $ is the specification for the crisp interval and $Ar$ is one of the Allen's relations (table \ref{tab:allen-relations}).
%\end{definition}

%\begin{table}[h]
%\centering
%\begin{tabular}{|c|c|c|c|}
%\hline
%Allen Relation & Constraints & $\bool(p_1,...,p_n)$ & $\Pos\left(\lambda \left[a,b \right] \right)$ \\
%\hline
%I before J & $C_1\stackrel{\triangle}{=} \left(<,X\right)$ & $p_1$ & $\sup_{a>w}\pi_x(w)$\\
%\hline
%\multirow{4}{*}
%{I equal J} & $C_1\stackrel{\triangle}{=} \left(\geq,X\right)$ & \text{$p_1\wedge\neg p_2\wedge p_3\wedge\neg p_4$} & $\min ( \sup_{a \leq w}\pi_x(w),$\\
% & $C_2\stackrel{\triangle}{=} \left(\neq,X\right)$ & & $\pi_x(w),$ \\
% & $C_3\stackrel{\triangle}{=} \left(\leq,Y\right)$ & & $\sup_{b \geq w}\pi_Y(w),$ \\
% & $C_4\stackrel{\triangle}{=} \left(\neq,Y\right)$ & & $\pi_Y(w))$ \\
%\hline
%\multirow{2}{*}
%{I meets J} & $C_1\stackrel{\triangle}{=} \left(\leq,X\right)$ & $p_1\wedge\neg p_2$ & $\min (\sup_{a\geq w} \pi_X(w),$\\
% & $C_2\stackrel{\triangle}{=} \left(\neq,X\right)$ &  & $\pi_X(w))$ \\
%\hline
%\multirow{3}{*}
%{I overlaps J} & $C_1\stackrel{\triangle}{=} \left(<,Y\right)$ & $p_1\wedge\neg p_2\wedge\neg p_3$ & $\min ( \sup_{b>w}\pi_Y(w), $\\
% & $C_2\stackrel{\triangle}{=} \left(\leq,X\right)$ & &  $\sup_{a \geq w}\pi_X(w),$\\
% & $C_3\stackrel{\triangle}{=} \left(\geq,X\right)$ & & $\sup_{a \leq w}\pi_X(w))$ \\
%\hline
%\multirow{4}{*}
%{I during J} & $C_1\stackrel{\triangle}{=} \left(>,X\right)$ & $(p_1\wedge p_2)\vee(p_3\wedge p_4)$ & $max ( min ( \sup_{a<w}\pi_X(w),$\\
% & $C_2\stackrel{\triangle}{=} \left(\leq,Y\right)$ & & $\sup_{b \geq w}\pi_Y(w)),$\\
% & $C_3\stackrel{\triangle}{=} \left(\geq,X\right)$ & & $\min ( \sup_{a \leq w }\pi_X(w),$\\
% & $C_4\stackrel{\triangle}{=} \left(<,Y\right)$ & & $\sup_{b>w}\pi_Y(w)$\\
%\hline
%\multirow{2}{*}
%{I starts J} & $C_1\stackrel{\triangle}{=} \left(\geq,X\right)$ & $p_1\wedge\neg p_2$ & $\min( \sup_{a \leq w}\pi_X(w),$\\
% &  $C_2\stackrel{\triangle}{=} \left(\neq,X\right)$ & & $\pi_X(w))$\\
%\hline
%\multirow{2}{*}
%{I finishes J} & $C_1\stackrel{\triangle}{=} \left(\leq,Y\right)$ & $p_1\wedge\neg p_2$ & $\min ( \sup_{b \geq w} \pi_Y(w),$\\
% & $C_2\stackrel{\triangle}{=} \left(\neq,Y\right)$ & & $\pi_Y(w))$\\
%\hline
%\end{tabular}
%\label{tab:allen-relations}
%\caption{Allen's relations represented in the framework.}
%\end{table}

%\subsubsection{Query evaluation}
%In fuzzy querying of regular (relational) databases, the query satisfaction modelling is a matter of degree. Usually, criteria satisfaction is modelled by means of a satisfaction degree $s \in \left[ 0, 1\right]$. In the model, every record $r$ contains a \emph{PVP}, $V_r$, to model the valid-time.The query evaluation method is the following:
%\begin{itemize}
%\item
%For each record $r$ in the database, the preferences expressed in $Q$ are evaluated in the unit interval. Thus, $e(Q) \in \left[0, 1\right]$ is the evaluation function for the criteria in $Q$.
%\item
%For each record $r$ in the database, the interval $ \left[a, \ b \right]$ is compared to the $PVP$ value by means of the corresponding Allen relation specified by $ar$. The result is again in the unit interval. Here, $e(Q^{time}) \in \left[0, 1 \right]$ is the evaluation function for the criteria $Q^{time}$.
%\end{itemize}

%\subsubsection{Ranking}
%In order to present the results to the user, a ranking method must be used. Both $Q$ and $Q^{time}$ are evaluated independently and also both are evaluated in the unit interval. The ranking is computed using a convex combination of both values:

%\begin{equation}
%e(\tilde{Q})\ =\ \omega*e(Q)\ +\ (1-\omega)*e(Q^{time})
%\end{equation}
%Where $\omega \in \left[0, 1 \right]$. By increasing the value for the parameter $\omega$, the preferences expressed in the query $Q$ can be given more importance. By lowering the value for $\omega$, the temporal criteria is emphasized.
%The following example illustrates the querying process.

%\begin{example}

%\textbf{The database}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Sample table for the car models. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{table}[ht]
%\caption{Sample database containing the car models.}
%\centering
%\begin{tabular}{c c c c c c c}
%\hline
%ID & IID & Segment & Manufac. & Name & Start & End  \\ [0.5ex]
%\hline
%001 & 1 & B & Peugeot & 205 & [1985,2,3] & [1997,2,1] \\
%002 & 1 & C & Peugeot & 305 & [1977,2,2] & [1989,2,3] \\
%003 & 1 & B & Citroen & C2 & [2003,3,2] & [2006,1,1] \\
%001 & 2 & B & Peugeot & 206 & [1998,1,1] & [2011,2,1] \\
%001 & 3 & B & Peugeot & 207 & [2006,1,1] & [2011,1,1]\\
%\hline
%\end{tabular}
%\label{tb:car-models}
%\end{table}

%Consider a database containing car models. There are several general attributes (car model's name, car manufacturer, car segment) and one temporal attribute (which is ill-known) containing the approximate date while the car model was sold. The temporal data is stored as explained in subsection \ref{subsec:storage}. The value for $D$ is stored in \emph{yyyy} format and $a$ and $b$ are represented by an integer, for the convenience of the representation. Therefore the chronons in our example will be years. The ID field identifies a car model while the field Instance ID (IID) identifies the instance for a car model. \\

%\textbf{The query}
%Consider the following query:

%\emph{The user wants to obtain a list of models in the segment B for the manufacturer Peugeot before the year interval 2001-2005.}

%The query is translated to the following expression, using the structure given in equation \eqref{eq:query-definition}:

%\begin{center}
%$\tilde{Q} = \left(  c^{time}, c_{segment} \right)$
%\end{center}

%Where:
%\begin{itemize}
%\item
%$c^{time}\ = \ ( \left[ 2001,\ 2005 \right],\ $  before $)$.
%\item
%$c_{segment}\ = \ $ Peugeot.
%\end{itemize}

%The evaluation for the criteria are presented in the result table \ref{tb:results}.

%\begin{table}[ht]
%\caption{Result table and ranking}
%\centering
%\begin{tabular}{c c c c c }
%\hline
%ID & IID &  $c^{time}$ & $c_{segment}$ & rank ($\omega=0.5$) \\ [0.5ex]
%\hline
%001 & 1  & 1 & 1 & 1 \\
%002 & 1  & 1 & 0 & 0.5 \\
%003 & 1  & 0.333  & 0 & 0.166\\
%001 & 2 & 1 & 1 & 1 \\
%001 & 3 & 0 & 1 & 0.5\\
%\hline
%\end{tabular}
%\label{tb:results}
%\end{table}
%\end{example}

### 2012-04-10 ijcis-pos-model/sources/temporalModel.tex
Update of the draft.

@@
 
-\vglue13pt
-%\begin{table}[htbp]
-\tcap{Valid-Time interval data types}
-\centerline{\small VALID TIME DATA TYPES}
-\vglue-6pt
-\centerline{\small\baselineskip=13pt
-\begin{tabular}{c p{2cm} p{2cm}}\\
-\hline
-Type for S & Type for E & Semantics \\
-\hline
-Unknown & Unkown & Unknown interval. \\
-\hline
-1,2 & 1,2 & A possibilistic time interval. \\
-\hline\\
-\end{tabular}
-\label{tbl:valid-time-interval}}
-%\end{table}
-
-The following is a study of the constraints in the range of values for both starting and ending times to represent a valid time interval.
-
-The time period inside a valid time interval is denoted by two constraints:
-\begin{eqnarray}
-C_1\stackrel{\triangle}{=}\left(\geq,X\right)\\
-C_2\stackrel{\triangle}{=}\left(\leq,Y\right).
-\end{eqnarray}
-Applying the inference of uncertainty as proposed in our general reasoning, we find for the first constraint that:
-\begin{eqnarray}
-\Pos\left(C_1([a,b])\right) &=& \min_{r\in[a,b]}\Big(\sup_{r\leq w}\pi_{X}(w)\Big)\\
-\Nec\left(C_1([a,b])\right) &=& \min_{r\in[a,b]}\Big(\inf_{r>w}1-\pi_{X}(w)\Big)
-\end{eqnarray}
-which can be simplified to:
-\begin{eqnarray}
-\Pos\left(C_1([a,b])\right) &=& \sup_{a\leq w}\pi_{X}(w)\\
-\Nec\left(C_1([a,b])\right) &=& \inf_{a>w}1-\pi_{X}(w).
-\end{eqnarray}
-For the second constraint, we find that:
-\begin{eqnarray}
-\Pos\left(C_2([a,b])\right) &=& \min_{r\in[a,b]}\Big(\sup_{r\geq w}\pi_{Y}(w)\Big)\\
-\Nec\left(C_2([a,b])\right) &=& \min_{r\in[a,b]}\Big(\inf_{r<w}1-\pi_{Y}(w)\Big)
-\end{eqnarray}
-which can be simplified to:
-\begin{eqnarray}
-\Pos\left(C_2([a,b])\right) &=& \sup_{b\geq w}\pi_{Y}(w)\\
-\Nec\left(C_2([a,b])\right) &=& \inf_{b<w}1-\pi_{Y}(w).
-\end{eqnarray}
-The uncertainty about the inclusion of an interval $I=[a,b]$ in the interval with ill-known boundaries can now be found by evaluating $[a,b]$ against the evaluation function $\lambda$ with $\bool=\wedge$.
+%
+%\vglue13pt
+%%\begin{table}[htbp]
+%\tcap{Valid-Time interval data types}
+%\centerline{\small VALID TIME DATA TYPES}
+%\vglue-6pt
+%\centerline{\small\baselineskip=13pt
+%\begin{tabular}{c p{2cm} p{2cm}}\\
+%\hline
+%Type for S & Type for E & Semantics \\
+%\hline
+%Unknown & Unkown & Unknown interval. \\
+%\hline
+%1,2 & 1,2 & A possibilistic time interval. \\
+%\hline\\
+%\end{tabular}
+%\label{tbl:valid-time-interval}}
+%%\end{table}
+%
+%The following is a study of the constraints in the range of values for both starting and ending times to represent a valid time interval.
+%
+%The time period inside a valid time interval is denoted by two constraints:
+%\begin{eqnarray}
+%C_1\stackrel{\triangle}{=}\left(\geq,X\right)\\
+%C_2\stackrel{\triangle}{=}\left(\leq,Y\right).
+%\end{eqnarray}
+%Applying the inference of uncertainty as proposed in our general reasoning, we find for the first constraint that:
+%\begin{eqnarray}
+%\Pos\left(C_1([a,b])\right) &=& \min_{r\in[a,b]}\Big(\sup_{r\leq w}\pi_{X}(w)\Big)\\
+%\Nec\left(C_1([a,b])\right) &=& \min_{r\in[a,b]}\Big(\inf_{r>w}1-\pi_{X}(w)\Big)
+%\end{eqnarray}
+%which can be simplified to:
+%\begin{eqnarray}
+%\Pos\left(C_1([a,b])\right) &=& \sup_{a\leq w}\pi_{X}(w)\\
+%\Nec\left(C_1([a,b])\right) &=& \inf_{a>w}1-\pi_{X}(w).
+%\end{eqnarray}
+%For the second constraint, we find that:
+%\begin{eqnarray}
+%\Pos\left(C_2([a,b])\right) &=& \min_{r\in[a,b]}\Big(\sup_{r\geq w}\pi_{Y}(w)\Big)\\
+%\Nec\left(C_2([a,b])\right) &=& \min_{r\in[a,b]}\Big(\inf_{r<w}1-\pi_{Y}(w)\Big)
+%\end{eqnarray}
+%which can be simplified to:
+%\begin{eqnarray}
+%\Pos\left(C_2([a,b])\right) &=& \sup_{b\geq w}\pi_{Y}(w)\\
+%\Nec\left(C_2([a,b])\right) &=& \inf_{b<w}1-\pi_{Y}(w).
+%\end{eqnarray}
+%The uncertainty about the inclusion of an interval $I=[a,b]$ in the interval with ill-known boundaries can now be found by evaluating $[a,b]$ against the evaluation function $\lambda$ with $\bool=\wedge$.
 Application of \eqref{eq:conjunctive1} and \eqref{eq:conjunctive2} leads to:
-\begin{eqnarray}
-\Pos\left(\lambda([a,b])\right)&=&\min\bigg(\Pos(C_1([a,b])),\Pos(C_2([a,b]))\bigg)\\
-\Nec\left(\lambda([a,b])\right)&=&\min\bigg(\Nec(C_1([a,b])),\Nec(C_2([a,b]))\bigg).
-\end{eqnarray}
-These last expressions can also be expanded as:
-\begin{eqnarray}
-\label{eq:interval-pos}
-\Pos\left(\lambda([a,b])\right)&=&\min\bigg(\sup_{a\leq w}\pi_{X}(w),\sup_{b\geq w}\pi_{Y}(w)\bigg)\\
-\label{eq:interval-nec}
-\Nec\left(\lambda([a,b])\right)&=&\min\bigg(\inf_{a>w}1-\pi_{X}(w),\inf_{b<w}1-\pi_{Y}(w)\bigg).
-\end{eqnarray}
-Note that the interval $[X,Y]$ used here, is certainly not a fuzzy interval.
+%\begin{eqnarray}
+%\Pos\left(\lambda([a,b])\right)&=&\min\bigg(\Pos(C_1([a,b])),\Pos(C_2([a,b]))\bigg)\\
+%\Nec\left(\lambda([a,b])\right)&=&\min\bigg(\Nec(C_1([a,b])),\Nec(C_2([a,b]))\bigg).
+%\end{eqnarray}
+%These last expressions can also be expanded as:
+%\begin{eqnarray}
+%\label{eq:interval-pos}
+%\Pos\left(\lambda([a,b])\right)&=&\min\bigg(\sup_{a\leq w}\pi_{X}(w),\sup_{b\geq w}\pi_{Y}(w)\bigg)\\
+%\label{eq:interval-nec}
+%\Nec\left(\lambda([a,b])\right)&=&\min\bigg(\inf_{a>w}1-\pi_{X}(w),\inf_{b<w}1-\pi_{Y}(w)\bigg).
+%\end{eqnarray}
+%Note that the interval $[X,Y]$ used here, is certainly not a fuzzy interval.
 Instead, we are dealing with an ill-known interval, i.e.
@@
 The uncertainty stems from the fact that the interval boundaries are ill-known.
-
-Within this framework for the reasoning of the knowledge about the interval, we are going to study the semantics from the point of view of the valid time as well as from the point of view of the temporal database.
-
-\subsubsection{S = Unknown, E = Unknown}
-Consider that the possibility distribution for S is Unknown and for E is also Unknown. The possibility distributions are $\pi_S$ and $\pi_E$ respectively:
-
-\begin{eqnarray}
-\pi_{S}(t \in \T) = 1\\
-\pi_{E}(t \in \T) = 1
-\end{eqnarray}
-
-Therefore the constraints $C_1$ and $C_2$ are:
-
-\begin{eqnarray}
-\Pos\left(C_1([a,b])\right) &=& \sup_{a\leq w}\pi_{S}(w) = 1\\
-\Nec\left(C_1([a,b])\right) &=& \inf_{a>w}1-\pi_{S}(w) = 0\\
-\Pos\left(C_2([a,b])\right) &=& \sup_{b\geq w}\pi_{E}(w)=1\\
-\Nec\left(C_2([a,b])\right) &=& \inf_{b<w}1-\pi_{E}(w)=0
-\end{eqnarray}
-
-\begin{eqnarray}
-\Pos\left(\lambda([a,b])\right)&=&\min\bigg(\sup_{a\leq w}\pi_{S}(w),\sup_{b\geq w}\pi_{E}(w)\bigg) \\
-&=& \min(1,1)  = 1\\
-\Nec\left(\lambda([a,b])\right)&=&\min\bigg(\inf_{a>w}1-\pi_{S}(w),\inf_{b<w}1-\pi_{E}(w)\bigg)\\
-&=& \min(0,0) = 0
-\end{eqnarray}
-
-\begin{example}
-A constant object in the database. It is an object that is always valid in the database.
-The main issue in a valid time database is that the update sentence have to modify the starting or ending point distribution.
+%
+%Within this framework for the reasoning of the knowledge about the interval, we are going to study the semantics from the point of view of the valid time as well as from the point of view of the temporal database.
+%
+%\subsubsection{S = Unknown, E = Unknown}
+%Consider that the possibility distribution for S is Unknown and for E is also Unknown. The possibility distributions are $\pi_S$ and $\pi_E$ respectively:
+%
+%\begin{eqnarray}
+%\pi_{S}(t \in \T) = 1\\
+%\pi_{E}(t \in \T) = 1
+%\end{eqnarray}
+%
+%Therefore the constraints $C_1$ and $C_2$ are:
+%
+%\begin{eqnarray}
+%\Pos\left(C_1([a,b])\right) &=& \sup_{a\leq w}\pi_{S}(w) = 1\\
+%\Nec\left(C_1([a,b])\right) &=& \inf_{a>w}1-\pi_{S}(w) = 0\\
+%\Pos\left(C_2([a,b])\right) &=& \sup_{b\geq w}\pi_{E}(w)=1\\
+%\Nec\left(C_2([a,b])\right) &=& \inf_{b<w}1-\pi_{E}(w)=0
+%\end{eqnarray}
+%
+%
+%\begin{eqnarray}
+%\Pos\left(\lambda([a,b])\right)&=&\min\bigg(\sup_{a\leq w}\pi_{S}(w),\sup_{b\geq w}\pi_{E}(w)\bigg) \\
+%&=& \min(1,1)  = 1\\
+%\Nec\left(\lambda([a,b])\right)&=&\min\bigg(\inf_{a>w}1-\pi_{S}(w),\inf_{b<w}1-\pi_{E}(w)\bigg)\\
+%&=& \min(0,0) = 0
+%\end{eqnarray}
+%
+%\begin{example}
+%A constant object in the database. It is an object that is always valid in the database.
+%The main issue in a valid time database is that the update sentence have to modify the starting or ending point distribution.
 The most natural way to convert a snapshot database to a temporal database is to consider the objects belonging to the interval [Unknown, Unknown].
-\end{example}
-
-\subsubsection{[Unknown,value] or [value,Unknown]}
-These two cases are managed in this section. Here we will note $\pi_{S}(t)$ or $\pi_{E}(t)$ as the possibility distribution for 'value' in the starting or ending points.
-
-First of all the case [Unknown,value] in \eqref{eq:interval-pos} and in \eqref{eq:interval-nec}:
-
-\begin{eqnarray}
-\Pos\left(\lambda([a,b])\right)&=&\min (1,\sup_{b\geq w}\pi_{E}(w))\\
- &=& \sup_{b\geq w}\pi_{E}(w)\\
-\Nec\left(\lambda([a,b])\right)&=&\min (0,\inf_{b<w}1-\pi_{E}(w))\\
-&=& 0
-\end{eqnarray}
-
-Similarily for [value,Unknown] the possibility and neccesity are:
-\begin{eqnarray}
-\Pos\left(\lambda([a,b])\right)&=&\min (\sup_{a\leq w}\pi_{S}(w),1)\\
- &=& \sup_{a\leq w}\pi_{S}(w)\\
-\Nec\left(\lambda([a,b])\right)&=&\min (\inf_{a>w}1-\pi_{S}(w),0)\\
-&=& 0
-\end{eqnarray}
-
-The semantics for an object with valid time given by [Unknown,value] is an object that has been valid from the beggining to an ending time.
+%\end{example}
+%
+%\subsubsection{[Unknown,value] or [value,Unknown]}
+%These two cases are managed in this section. Here we will note $\pi_{S}(t)$ or $\pi_{E}(t)$ as the possibility distribution for 'value' in the starting or ending points.
+%
+%First of all the case [Unknown,value] in \eqref{eq:interval-pos} and in \eqref{eq:interval-nec}:
+%
+%\begin{eqnarray}
+%\Pos\left(\lambda([a,b])\right)&=&\min (1,\sup_{b\geq w}\pi_{E}(w))\\
+% &=& \sup_{b\geq w}\pi_{E}(w)\\
+%\Nec\left(\lambda([a,b])\right)&=&\min (0,\inf_{b<w}1-\pi_{E}(w))\\
+%&=& 0
+%\end{eqnarray}
+%
+%Similarily for [value,Unknown] the possibility and neccesity are:
+%\begin{eqnarray}
+%\Pos\left(\lambda([a,b])\right)&=&\min (\sup_{a\leq w}\pi_{S}(w),1)\\
+% &=& \sup_{a\leq w}\pi_{S}(w)\\
+%\Nec\left(\lambda([a,b])\right)&=&\min (\inf_{a>w}1-\pi_{S}(w),0)\\
+%&=& 0
+%\end{eqnarray}
+%
+%The semantics for an object with valid time given by [Unknown,value] is an object that has been valid from the beggining to an ending time.
 Analogously, the semantics for an object with a valid time given by the interval [value,Unknown] is an object that it is still valid in the database.
-
-\subsubsection{[Undefined,Undefined]}
-Maybe this combination has no sense for a valid time object. The possibility distributions for both starting and ending points are:
-
-\begin{eqnarray}
-\pi_{S}(t \in \T) = 0\\
-\pi_{E}(t \in \T) = 0
-\end{eqnarray}
-
-Therefore the constraints $C_1$ and $C_2$ are:
-
-\begin{eqnarray}
-\Pos\left(C_1([a,b])\right) &=& \sup_{a\leq w}\pi_{S}(w) = 0\\
-\Nec\left(C_1([a,b])\right) &=& \inf_{a>w}1-\pi_{S}(w) = 1\\
-\Pos\left(C_2([a,b])\right) &=& \sup_{b\geq w}\pi_{E}(w)=0\\
-\Nec\left(C_2([a,b])\right) &=& \inf_{b<w}1-\pi_{E}(w)=1
-\end{eqnarray}
-
-Taking these values into  \eqref{eq:interval-pos} and in \eqref{eq:interval-nec}:
-\begin{eqnarray}
-\Pos\left(\lambda([a,b])\right)&=&\min (0,0)\\
- &=& 0\\
-\Nec\left(\lambda([a,b])\right)&=&\min (1,1)\\
-&=& 1
-\end{eqnarray}
-
-\textcolor{red}{TAKE A LOOK INTO THIS}
-
-Which is an inconsistency.
-
-\subsection{[Undefined,value] or [value,Undefined]}
-
-The possibility and neccesity measures for [Undefined,value] are :
-\begin{eqnarray}
-\Pos\left(\lambda([a,b])\right)&=&\min (0,\sup_{b\geq w}\pi_{E}(w))\\
- &=& 0\\
-\Nec\left(\lambda([a,b])\right)&=&\min (1,\inf_{b<w}1-\pi_{E}(w))\\
-&=& \inf_{b<w}1-\pi_{E}(w)
-\end{eqnarray}
-
-The other way around, [value,Undefined]:
-\begin{eqnarray}
-\Pos\left(\lambda([a,b])\right)&=&\min (\sup_{a\leq w}\pi_{S}(w),0)\\
- &=& 0\\
-\Nec\left(\lambda([a,b])\right)&=&\min (\inf_{a>w}1-\pi_{S}(w),1)\\
-&=& \inf_{a>w}1-\pi_{S}(w)
-\end{eqnarray}
-
-\textcolor{red}{TAKE A LOOK INTO THIS}
-This is also an inconsistency.
-
-\subsubsection{[NULL,NULL] and maybe any combination with NULL}
-I think the framework we have can not deal with NULL computations. Maybe just for non-specified time or something like that.
-
-\subsubsection{[Undefined,Unknown] or [Unknown,Undefined]}
-The semantic for this could be also unconsistent.
-
-\begin{example}
-Consider a snapshot database (table \ref{tbl:snapshot-car-models}) that stores information about car models. The primary key is given by the set $\lbrace$ Brand, Model $\rbrace$.
-
-\vglue13pt
-%\begin{table}[htbp]
-\tcap{Car model snapshot database}
-\centerline{\small CAR MODELS}
-\vglue-6pt
-\centerline{\small\baselineskip=13pt
-\begin{tabular}{c c c}\\
-\hline
-\textbf{Brand} & \textbf{Model} & Segment \\
-\hline
-Peugeot & 205 & B \\
-Peugeot & 206 & B \\
-Peugeot & 207 & B \\
-Peugeto & 208 & B \\
-Citroen & C2 & B \\
-\hline\\
-\end{tabular}
-\label{tbl:snapshot-car-models}
-}
-
-Consider now that the user wants to make that table a valid-time table, then, from the point of view of the storage, the following changes should be done:
-
-\begin{itemize}
-\item
-A new definition for the primary key, including a version field.
-\item
-The addition of two new columns to store the valid-time interval. By default the initial values are Unknown for both starting and ending points of the interval.
-\end{itemize}
-Therefore, the table \ref{tbl:snapshot-car-models} is transformed into table \ref{tbl:initial-valid-time-car-models}:
-
- \vglue13pt
-%\begin{table}[htbp]
-\tcap{Car model valid-time database. Note that U is the abbreviation for the Unknown constant.}
-\centerline{\small VALID TIME CAR MODELS}
-\vglue-6pt
-\centerline{\small\baselineskip=13pt
-\begin{tabular}{c c c c c c}\\
-\hline
-\textbf{Brand} & \textbf{Model} & \textbf{VersionID} & Segment & S & E \\
-\hline
-Peugeot & 205 & 001 & B & U & U \\
-Peugeot & 206 & 001 & B & U & U \\
-Peugeot & 207 & 001 & B & U & U \\
-Peugeto & 208 & 001 & B & U & U \\
-Citroen & C2 & 001 & B & U & U \\
-\hline\\
-\end{tabular}
-\label{tbl:initial-valid-time-car-models}
-}
-
-\end{example}
+%
+%
+%\subsubsection{[Undefined,Undefined]}
+%Maybe this combination has no sense for a valid time object. The possibility distributions for both starting and ending points are:
+%
+%\begin{eqnarray}
+%\pi_{S}(t \in \T) = 0\\
+%\pi_{E}(t \in \T) = 0
+%\end{eqnarray}
+%
+%Therefore the constraints $C_1$ and $C_2$ are:
+%
+%\begin{eqnarray}
+%\Pos\left(C_1([a,b])\right) &=& \sup_{a\leq w}\pi_{S}(w) = 0\\
+%\Nec\left(C_1([a,b])\right) &=& \inf_{a>w}1-\pi_{S}(w) = 1\\
+%\Pos\left(C_2([a,b])\right) &=& \sup_{b\geq w}\pi_{E}(w)=0\\
+%\Nec\left(C_2([a,b])\right) &=& \inf_{b<w}1-\pi_{E}(w)=1
+%\end{eqnarray}
+%
+%Taking these values into  \eqref{eq:interval-pos} and in \eqref{eq:interval-nec}:
+%\begin{eqnarray}
+%\Pos\left(\lambda([a,b])\right)&=&\min (0,0)\\
+% &=& 0\\
+%\Nec\left(\lambda([a,b])\right)&=&\min (1,1)\\
+%&=& 1
+%\end{eqnarray}
+%
+%\textcolor{red}{TAKE A LOOK INTO THIS}
+%
+%Which is an inconsistency.
+%
+%\subsection{[Undefined,value] or [value,Undefined]}
+%
+%The possibility and neccesity measures for [Undefined,value] are :
+%\begin{eqnarray}
+%\Pos\left(\lambda([a,b])\right)&=&\min (0,\sup_{b\geq w}\pi_{E}(w))\\
+% &=& 0\\
+%\Nec\left(\lambda([a,b])\right)&=&\min (1,\inf_{b<w}1-\pi_{E}(w))\\
+%&=& \inf_{b<w}1-\pi_{E}(w)
+%\end{eqnarray}
+%
+%The other way around, [value,Undefined]:
+%\begin{eqnarray}
+%\Pos\left(\lambda([a,b])\right)&=&\min (\sup_{a\leq w}\pi_{S}(w),0)\\
+% &=& 0\\
+%\Nec\left(\lambda([a,b])\right)&=&\min (\inf_{a>w}1-\pi_{S}(w),1)\\
+%&=& \inf_{a>w}1-\pi_{S}(w)
+%\end{eqnarray}
+%
+%\textcolor{red}{TAKE A LOOK INTO THIS}
+%This is also an inconsistency.
+%
+%\subsubsection{[NULL,NULL] and maybe any combination with NULL}
+%I think the framework we have can not deal with NULL computations. Maybe just for non-specified time or something like that.
+%
+%\subsubsection{[Undefined,Unknown] or [Unknown,Undefined]}
+%The semantic for this could be also unconsistent.
+%
+%
+%\begin{example}
+%Consider a snapshot database (table \ref{tbl:snapshot-car-models}) that stores information about car models. The primary key is given by the set $\lbrace$ Brand, Model $\rbrace$.
+%
+%\vglue13pt
+%%\begin{table}[htbp]
+%\tcap{Car model snapshot database}
+%\centerline{\small CAR MODELS}
+%\vglue-6pt
+%\centerline{\small\baselineskip=13pt
+%\begin{tabular}{c c c}\\
+%\hline
+%\textbf{Brand} & \textbf{Model} & Segment \\
+%\hline
+%Peugeot & 205 & B \\
+%Peugeot & 206 & B \\
+%Peugeot & 207 & B \\
+%Peugeto & 208 & B \\
+%Citroen & C2 & B \\
+%\hline\\
+%\end{tabular}
+%\label{tbl:snapshot-car-models}
+%}
+%
+%Consider now that the user wants to make that table a valid-time table, then, from the point of view of the storage, the following changes should be done:
+%
+%\begin{itemize}
+%\item
+%A new definition for the primary key, including a version field.
+%\item
+%The addition of two new columns to store the valid-time interval. By default the initial values are Unknown for both starting and ending points of the interval.
+%\end{itemize}
+%Therefore, the table \ref{tbl:snapshot-car-models} is transformed into table \ref{tbl:initial-valid-time-car-models}:
+%
+% \vglue13pt
+%%\begin{table}[htbp]
+%\tcap{Car model valid-time database. Note that U is the abbreviation for the Unknown constant.}
+%\centerline{\small VALID TIME CAR MODELS}
+%\vglue-6pt
+%\centerline{\small\baselineskip=13pt
+%\begin{tabular}{c c c c c c}\\
+%\hline
+%\textbf{Brand} & \textbf{Model} & \textbf{VersionID} & Segment & S & E \\
+%\hline
+%Peugeot & 205 & 001 & B & U & U \\
+%Peugeot & 206 & 001 & B & U & U \\
+%Peugeot & 207 & 001 & B & U & U \\
+%Peugeto & 208 & 001 & B & U & U \\
+%Citroen & C2 & 001 & B & U & U \\
+%\hline\\
+%\end{tabular}
+%\label{tbl:initial-valid-time-car-models}
+%}
+%
+%
+%
+%\end{example}
 

### 2012-04-10 ijcis-pos-model/sources/temporaldatabases.tex
Update of the draft.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Time
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\label{subsec:time-in-databases}Time in Databases}
The concept of time has been studied in databases for a long time. A true standard for adding temporal aspects to relational databases does not exist, but there is a consensus in the literature \cite{Dyreson1994} on what is called a \emph{temporal database}: a temporal database is a database dealing with some aspects of time in its schema.
In a temporal DBMS, a \textbf{chronon} is the shortest duration of time supported by the system. In temporal databases, some temporal attributes can be managed without treating the attribute differently from non-temporal attributes. The time described by such an attribute is called \textbf{user defined time} (\emph{UDT}). In addition to UDT, the following types of time can be discerned in a temporal database, all of which are handled exceptionally by the DBMS:

\begin{itemize}
	\item
	\textbf{Transaction time} (\emph{TT}) \cite{Rowe1987},\cite{Jensen1991} denotes the time when the fact (object) is stored in the database. It is usually append-only: as the past can not be changed, TT can not be changed neither. Furthermore, at the moment of insertion, a TT can be neither in the past nor in the future.
	\item
	\textbf{Valid time} (\emph{VT}) \cite{Jensen1994},\cite{Sarda1990},\cite{McKenzie1981} denotes the time when the fact (object) is true in the modelled reality. A fuzzy extension has been proposed by \cite{Garrido2009}. 
%	\item
%	User defined time: It is an uninterpreted attribute. The domain is date/time. The query language has no special support for it.
	\item
	\textbf{Decision time} (\emph{DT}, proposed in \cite{Nascimento1995}) denotes the time when an event was decided to happen. 
	\end{itemize}

	E.g., consider a database containing employee contract descriptions. The time when the employee's contract is valid, represented by an interval, is VT. The time when the employee's contract is stored in the database is the TT. The time when the decision for hiring this employee was made is the DT.

% is a non-decomposable unit of time.  it  There are two ways to represent a chronon: as a point or as an interval \cite{655777}.

When working with these time concepts, the Data Manipulation Language (\emph{DML}, which is part of the standard database querying language SQL) is extended to deal with possible temporal inconsistencies within the data and to handle more complex (temporal) queries. 
%Next to these concepts, also \textbf{user defined time} (\emph{UDT}) is discerned. UDT is an uninterpreted attribute in the date/time domain. This means that the attribute uses the date/time domain, but the database model does not treat the attribute differently from non-temporal attributes.
Depending on the time managed, a database is classified as either a \textbf{Valid Time Database} (\emph{VTDB}), a \textbf{Transaction Time Database} (\emph{TTDB}), a \textbf{bi-temporal database} (both valid and transaction time are managed) or a \textbf{tri-temporal database} (valid time, transaction time and decision time are managed).

%\subsection{Temporal Elements}
%%	
%	In order to define a temporal database model, some basic elements should be defined \cite{Dyreson:1994:CGT:181550.181560}:
%	\begin{itemize}
%	\item A	\textbf{chronon} is a non-decomposable unit of time. In a temporal DBMS, it is the shortest duration of time supported by the system. There are two ways to represent a chronon: as a point or as an interval \cite{655777}.
%	\item
%	\emph{Event}: An instant of time. Usually, an event is said to be occur during time $t$ if it occurs during the chronon represented by $t$.
%	\item
%	\emph{Interval}: The time between two events. The representation is very often a set of contiguous chronons.
	%\item
	%\emph{Span}: A directed duration of time. A duration is an amount of time with known  lenght but no specific starting or ending chronons. The span can be either positive, denoting a forward motion of time or negative, denoting a backward motion of time.
%	\item A	\textbf{timestamp} is a time value associated with some object in the database.

%	\item The	\textbf{lifespan} (of a database object)is the time over which it is defined. The lifespan of a valid time object denotes the time when the corresponding object exists. The lifespan of a transaction time object is the value of the timestamp.

%	Depending on the type of time the meaning is different:
%		\begin{itemize}
%		\item
%		Lifespan of a valid time object: Refers the time when the corresponding object exists.	
%		\item
%		Lifespan of a transaction time object: The transaction time of a database object refers when the object is stored in the database. The lifespan is the value of the timestamp.		
%		\end{itemize}
%	\end{itemize}

%	In the temporal database thesaurus, \emph{'Snapshot'} is the word for non-temporal matters. As a temporal database is a generalization for relational databases, an \emph{snapshot database} is a relational database. Furthermore, a \emph{snapshot relation} is a relation incorporating neither valid nor transaction time.

%\subsection{Main issues when dealing with time}
%Among others, the following problems are present when dealing with time in a database:

%	\begin{itemize}	
%	\item \textbf{Granularity} denotes a partition on the set of chronons. The conversion between several granularities is studied in \cite{Lin97efficientconversion}. Granularity is the base of the temporal model in \cite{Cru97}. An object-oriented implementation is in \cite{624013}.
%	\item	To ensure \textbf{consistency}, temporal databases usually redefine the primary key of a relation. The new primary key takes into account the presence of the time. In order to keep the consistency, the DML is re-defined. For example, in a VTDB, the temporal update sentence is usually composed by an update sentence (closing the old row) plus an insert sentence (creating the new row).
%	\end{itemize}
%Guy's suggestion: instead of imprecision, use imperfection which is more general.
\subsubsection{Imperfection and time}
Representing imprecision and its semantics when dealing with time has been studied for a long time. Several proposals for representing and computing imprecise time indications can be found in \cite{DeCaluwe1997} and \cite{DeTre1997}. Also, the changes between several granularities can be seen as a source of imprecision \cite{Devos1998}.

In the proposal section we will consider two kinds of imprecision:
\begin{itemize}
\item \textbf{Uncertainty in the database} denotes the uncertainty that arises when the knowledge about the temporal data in the database is uncertain. E.g., a database record shows that \emph{`The car is in the garage around April.'}
 \item \textbf{Imprecision in the query specification} denotes the imprecision in the specification of temporal criteria by the user, when querying. E.g., \emph{`The user wants to obtain a car which is red and which is in the garage around April.'}
\end{itemize}

\subsubsection{Representation}
Several proposals for managing uncertain time in a database exist. Some proposals work with rough sets \cite{Qiang2009}, other proposals rely on possibility distributions for representing uncertainty in time \cite{Dyreson1998},\cite{Garrido2009}, \cite{Galindo2001}. In order to compare temporal possibility distributions, extensions of the classical Allen's operators \cite{Allen1983} are defined in \cite{Ohlbach2004}, \cite{Nagypal2003},\cite{Dubois2003a} and \cite{Schockaert2008}
%In the proposal section, we will follow the representation by means of possibility distributions, in order to work with both satisfaction and dissatisfaction degrees. Also, in order to work properly with fuzzy operators, the underlying domain should be numeric. 
% Take a look into this:
%In this paper, the representation for the dates will follow the Julian Day Number (JDN) representation \cite{Dir96}.

%If the starting points and/or the end points of the interval representing the time are not known precisely, it is easy to fuzzify them, using, e.g., two triangular possibility distributions.

In order to deal with uncertainty in time intervals, several proposals are made. Here, two approaches are introduced: the first one, based on \emph{Fuzzy Validity Period}~\cite{Garrido2009} and the second one based on \emph{Possibilistic Valid-time Period}~\cite{JoseEnriquePons2012}.

\begin{definition}
A \textbf{Fuzzy Validity Period} (\emph{FVP}) is defined as a fuzzy time interval specifying when an object is valid. A fuzzy time interval is then the fuzzification of a crisp time interval.
\end{definition}
Several options to transform possibility distributions corresponding to the fuzzy starting point and the fuzzy end point into one consistent FVP exist \cite{Garrido2009}, e.g (Fig. \ref{fig:fuzzy-validity-period}):
\begin{itemize}
\item The \textbf{convex hull} approach is the most intuitive approach. The resulting FVP is the convex hull of the union of both fuzzy sets.
\item The \textbf{uncertainty preserving} approach is less intuitive but more realistic. The amount of uncertainty is maintained at the edges of the possibility distribution representing the FVP \cite{Garrido2009}.
\end{itemize}

%%%%%%
% FUZZY VALIDITY PERIOD
%%%%%%
\vspace*{13pt}
\begin{center}
{
\includegraphics[scale=0.25]{./graphs/trapezoidalDistribution.pdf}

}
\end{center}
%\centerline{ \psfig{file=./graphs/Y-time-point.eps}}
\vspace*{10pt}
\fcaption{\label{fig:fuzzy-validity-period}Transformation to obtain the FVP. The top graph shows the two triangular possibility distributions. The middle graph shows the convex hull validity period, the bottom one shows the result of the second transformation, which maintains the imprecision.}
%  \label{fig:fuzzy-validity-period}
\vspace*{13pt}

\begin{definition}
A \textbf{Possibilistic Valid-time Period} (\emph{PVP}) is an ill-known interval in time specifying when an object is valid.
\end{definition}
A PVP is an ill-known interval in the sense of definition \ref{def;possibilistic-variable}, section \ref{subsec:possibilistic-variables}. Note that this representation is \emph{disjunctive}: the PVP represents only one crisp time interval, but that for some reason it is (partially) unknown.

The ill-known interval approach has many advantages from the representation by the FVP as demonstrated in \cite{Pons2011},\cite{Pons2012}. Table \ref{tbl:comparative-pvp-fvp} is a comparative between PVP and FVP. The following list defines the items in the comparative:

\begin{enumerate}
\item Domain: The domain of the possibility distribution modelled by the approach.
\item Implementation of relationships: How to implement a relationship.
\item Allen's relations: Are the Allen's relations defined?
\item Storage: The way the data is stored in the database.
\item Possibility measures: Does the framework provides always a possibility measure for any relation between the temporal elements?
\item Necessity measures:  Does the framework provides always a necessity measure for any relation between the temporal elements?
\end{enumerate}

%%
%% comparison table between pvp and fvp.
%%
\vglue13pt
%\begin{table}[htbp]
\tcap{Comparative PVP vs FVP}
%\centerline{\small DATA TYPES}
\vglue-6pt
\centerline{\small\baselineskip=13pt
\begin{tabular}{c p{2cm} p{2cm}}\\
Item & PVP & FVP\\
\hline
(1) & $\Pow(\R)$ & $\R$ \\
(2) & Ill-known constraints. & Ad-hoc operators. \\
(3) & $\checkmark$ & - \\
(4) & Two distributions one for each endpoint. & Only one distribution. \\ 
(5) & $\checkmark$ & $\checkmark$ \\
(6) & $\checkmark$ & - \\
\hline\\
\end{tabular}
\label{tbl:comparative-pvp-fvp}
}

In the rest of the paper we will work only with PVP to represent valid-time intervals.

\subsection{Understanding Valid-time Databases}
This subsection is devoted to describe the behaviour of a crisp valid-time database. For the shake of simplicity, only the three main operations in the Data Manipulation Language \emph{DML} a.k.a \emph{CRUD} (CReate Update, Delete) operations are shown. Usually the DML operations in a temporal database are re-defined e.g. a typical update sentence in SQL could be re-defined by means of several insert and update sentences. Therefore, in order to be clear, these high level primitives in the DMl for a valid-time database are usually noted as Insert, Modify and Delete. In the following subsections each primitive is defined and explained. Finally a illustrative example is given. For a more complete information on the behaviour of a bitemporal database, please refer to ~\cite{Jensen1994}.

\begin{definition}
Consider a relation $R$ in a temporal database, an object $A$ given by the attributes $\left(a_1, \ldots, a_n \right)$ and the crisp time interval $I$ given by $\left[a,\ b\right] : a,b \in \T, a \leq b$. A valid-time object $A$ in the relation $R$ is specified by:
\begin{equation}
\label{eq:rel-def}
\left( R, A , I \right)
\end{equation}
\end{definition}

\subsubsection{Insert}
The interpretation of the insert operation is the following: The user wants to store in the database that an object $A$  is true for some period(s) of time $I$ in the given relation $R$. To indicate that the object $A$ is current in the relation, the special value Until Current, $\UC$  is used.

There are three main cases when performing a create operation:
\begin{enumerate}
\item The object $A$ was never in the relation $R$: The object is added with the valid-time indicated by the crisp interval $I$.

\item The object $A$ was part of some previously recorded state: A new version of the object is inserted in the relation $R$ with the valid-time indicated by $I$.

\item The object $A$ is current in the relation $R$: The insert operation is rejected, a modify operation should be used.
\end{enumerate}

The algorithm for the implementation of the insert operation is the following:

\begin{align}
\label{eq:insert}
\INS \left(R ,A, I\right) =\\
\begin{cases}
\nonumber
R \cup \left(A, I \right), & \mbox{ if }  A \not \in R\\
R \cup  & \mbox{ if }  A \in R \wedge \\
R, & \mbox{ otherwise }  
\end{cases} 	
\end{align}

\subsubsection{Delete}
The delete operation logically removes a current object $A$ which is valid in the relation $R$:
\begin{align}
\label{eq:delete}
\DEL \left(R ,A \right) =\\
\begin{cases}
\nonumber
R - \left(A, I \right), & \mbox{ if }  A \not \in R\\
R, & \mbox{ otherwise }  
\end{cases} 	
\end{align}

\subsubsection{Modify}
The modification of an object $A$ is usually defined by a delete operation followed by an insert operation. The modify operation is defined then by:

\begin{align}
\label{eq:modify}
\MOD \left( R, A, I \right) &=&\\
\nonumber
\INS \left( \DEL \left(R, A \right), A, I \right)
\end{align}

### 2012-04-10 trunk/ijcis-pos-model/sources/preliminaries.tex
Update of the draft.

%Introductory text:
In this section, some basic concepts are introduced concerning possibilistic variables and fuzzy numbers and intervals, after which the framework of set evaluation by ill-known constraints \cite{Pons2011} is explained. The section concludes with a brief introduction to temporal databases.

\subsection{\label{subsec:possibility-theory}Possibility Theory}
Possibility theory, like probability theory, deals with uncertainty about the outcome of an experiment. In probability theory, this uncertainty is caused by the \emph{variability} in the outcomes, while in possibility theory, the uncertainty is caused by \emph{incomplete knowledge} about the experiment. The quantification of confidence in a theory of uncertainty is achieved using a confidence measure\cite{Pons2011}. In probability theory this is a measure of chance, in possibility theory, possibility and necessity measures are used.

\begin{definition}
Consider a set of outcomes $\Omega$. Let $\wp(\Omega)$ denote the powerset of $\Omega$ and let $A$ and $B$ be elements of $\wp(\Omega)$. A \emph{confidence measure on $\Omega$} is defined by a function
	\begin{align}
	g : \wp(\Omega) & \rightarrow \left[0,1\right]
	\end{align}
that satisfies
	\begin{align}
	g(\emptyset) &= 0 \\
	g(\Omega) &= 1 	\label{NormalizationProperty} \\
	A \subseteq B &\Rightarrow g(A) \leq g(B) \label{MonotonicityProperty}
	\end{align}
\end{definition}

Both possibility measures and necessity measures are special cases of confidence measures.

\begin{definition}
Consider a confidence measure $\Pi$ on a set of outcomes $\Omega$. Let $J$ be a countable index set and let $\{ A_{j} | j \in J \wedge A_{j} \subseteq \Omega \}$ be a family of elements of $\wp(\Omega)$. $\Pi$ is now a \emph{possibility measure on $\Omega$} if it satisfies:
	\begin{align}
	\Pi\left(\bigcup_{j \in J} A_{j} \right) = \sup_{j \in J} \Pi(A_{j})
	\end{align}
\end{definition}

In this work, the interpretation is as follows. The possibility of an event expresses how plausible the occurrence of the event seems to an observer of the experiment, given the (partial) knowledge of the observer about the experiment.

Information on the possibility of distinct elements of the universe of discourse $\Omega$ can now be given by a \emph{possibility distribution} $\pi$ on $\Omega$, defined by:

\begin{definition}
Consider a possibility measure $\Pi$ on $\Omega$. A \emph{possibility distribution} $\pi$ on $\Omega$ underlying the possibility measure $\Pi$ is then a function defined by:
	\begin{align}
	\pi : \Omega \rightarrow \left[0, 1\right] : \pi(u) = \Pi(\{u\})
	\end{align}
\end{definition}

\begin{definition}
Consider a confidence measure $N$ on a set of outcomes $\Omega$. Let $J$ be a countable index set and let $\{ A_{j} | j \in J \wedge A_{j} \subseteq \Omega \}$ be a family of elements of $\wp(\Omega)$. $N$ is now a \emph{necessity measure} on $\Omega$ if it satisfies:
	\begin{align}
	N\left(\bigcap_{j \in J} A_{j} \right) = \inf_{j \in J} N(A_{j})
	\end{align}
\end{definition}

In this work, the interpretation is as follows. The necessity of an event expresses how necessary the occurrence of the event seems to an observer of the experiment, given the (partial) knowledge of the observer about the experiment.

Possibility and necessity measures are dual in the sense that:

\begin{align}
\forall A \subseteq \Omega : N(A) = 1 - \Pi(\bar{A})
\end{align}

Regarding interpretation, the above can be seen as: the degree to which an event is necessary is the degree to which every other possible event is not plausible.

\subsection{\label{subsec:possibilistic-variables}Possibilistic Variables}
Possibilistic variables rely on possibility theory \cite{Dubois1988a}. A \emph{possibilistic variable} is defined as follows \cite{Pons2011}.

\begin{definition}
\label{def;possibilistic-variable}
A possibilistic variable $X$ over a universe $U$ is defined as a variable taking exactly one value in $U$, but for which this value is (partially) unknown. Its possibility distribution $\pi_X$ on $U$ models the available knowledge about the value that $X$ takes: for each $u\in U$, $\pi_X(u)$ represents the possibility that $X$ takes the value $u$. In this work, this possibility is interpreted as a measure of how plausible it is that $X$ takes the value $u$, given (partial) knowledge about the value $X$ takes.
\end{definition}

The exact value a possibilistic variable takes, which is (partially) unknown, is called an \emph{ill-known value} in this work \cite{Dubois1988a}.

When a possibilistic variable is defined on the powerset $\Pow(R)$ of some universe $R$, the unique value the variable takes will be a crisp set and its possibility distribution on the powerset $\Pow(R)$ will describe the possibility of each crisp subset of $R$ to be the value the variable takes. This exact value (a crisp set) the variable takes, is now called an \emph{ill-known set} \cite{Dubois1988a}.

It is important to understand the difference between the following two concepts:
\begin{itemize}
\item
A \emph{possibilistic variable} $X$ is bounded to take only one value , but this value is not known due to incomplete knowledge. 
\item
An \emph{ill-known set}~\cite{Dubois1988a}: a possibilistic variable defined over the universe $\Pow(U)$.
\end{itemize}

Note that while a possibilistic variable refers to one (partially) unknown value, an ill-known set is a crisp set but, for some reason, (partially) unknown.

A specific application of possibilistic variables is obtained when the universe under consideration is the set of Boolean values $\mathbb{B}$ = $\{T,F\}$. Indeed, any Boolean proposition $p$ takes just one value in $\mathbb{B}$. If the knowledge about which value this proposition $p$ will take is given by a possibility distribution $\pi_p$, the proposition can be seen as a possibilistic variable. As the interest lies with the case where the proposition holds, the possibility and necessity that $p$ = $T$ (the proposition holds) demand most attention. This possibility and necessity is noted here as:
\begin{align}
\text{Possibility that $p$ = $T$ (p holds):} &\\
\nonumber
Pos(p) = \pi_p(T) \label{propholdsposs} \\
\text{Necessity that $p$ = $T$ (p holds):} & \\
\nonumber
Nec(p) = 1-\pi_p(F) \label{propholdsnecc}
\end{align}

Here, notation \ref{propholdsposs} denotes the possibility that $p$ = $T$ and the proposition holds, notation \ref{propholdsnecc} denotes the necessity that $p$ = $T$ and the proposition holds.

This work will deal with ill-known intervals. These are ill-known sets, defined and represented via a start and end point, which will be ill-known values. The elements of the set are the values between the start and end point. A closed ill-known interval with start point defined by possibilistic variable $X$ and end point by possibilistic variable $Y$ is noted here $\left[X, Y\right]$. The correspondences and transitions between the representations of ill-known sets, between the representations of ill-known intervals and between the representations of an ill-known set and an ill-known interval are part of the authors current research.

\subsection{\label{subsec:fuzzy-numbers}Fuzzy Numbers and Fuzzy Intervals}
Among others, Dubois and Prade~\cite{Dubois1983} use fuzzy sets \cite{Zadeh1965} to define a \emph{fuzzy interval}:
\begin{definition}
A fuzzy interval is a fuzzy set $M$, defined by a membership function $\mu_{M}$, on the set of real numbers $\mathbb{R}$ such that:
\begin{eqnarray}
\mu_{M} :  \mathbb{R} \rightarrow \left[0,1\right] \nonumber \\ 
\forall (u,v)\in\mathbb{R}^2: \forall w \in [u,v]:\\
\mu_M(w) \geq\min(\mu_M(u),\mu_M(v))  \\
\exists m \in \mathbb{R} :  \mu_M(m)=1 
\end{eqnarray}
\end{definition}
If this modal value $m$ is unique, then $M$ is referred to as a \emph{fuzzy number}. In other words, if the core of a fuzzy interval is a singleton, it is referred to as a fuzzy number.

A simple form of the membership function of a fuzzy interval is a trapezoidal function. It can be shown that such a membership function $\mu_T$ for a fuzzy interval $T$ is convex and normalized. Four reel values, denoted $\alpha$, $\beta$, $\gamma$ and $\delta$ suffice to represent a trapezoidal membership function of a fuzzy interval. In this work, a fuzzy interval defined as such will be noted as $\left[\alpha, \beta, \gamma, \delta\right]$. The corresponding membership function definition for this $\mu_T$ is then given by:

\begin{align}
\mu_T : & \quad \mathbb{R} \rightarrow \left[0,1\right] \\
 : & \quad x \rightarrow
\begin{cases}
1 & \mbox{ if } x \in [\beta,\gamma] \\
0 & \mbox{ if } x > \delta \vee x < \alpha \\
\frac{x-\alpha}{\beta - \alpha} & \mbox{ if } x \in [\alpha,\beta[ \\
\frac{\delta -x}{\delta - \gamma} & \mbox{ if } x \in ]\gamma,\delta] \\
\end{cases}
\end{align}

%\def\JPicScale{0.5}
%\begin{figure}[h!]
%  \centering
%  \input{./graphs/trapezoidal.tex}
%  \caption{Trapezoidal membership function}
%  \label{fig:trapezoidal}
%\end{figure}

The most convenient form of the membership function of a fuzzy number is a triangular form. It can be shown that such a membership function $\mu_M$ for a fuzzy number $M$ is convex and normalized. Three real values, denoted by $a$, $b$ and $D$, suffice to represent a triangular membership function of a fuzzy number and in this work, a fuzzy number defined as such will be noted as $\left[D, a, b \right]$. Here:
\begin{itemize}
\item
$D$ denotes the single value in the core of $M$
\item
$D-a$ is then $\inf \{u \in \mathbb{R} : \mu_{M}(u) > 0\}$
\item
$D+b$ is then $\sup \{u \in \mathbb{R} : \mu_{M}(u) > 0\}$
\end{itemize}

The membership function of such a fuzzy number is then given by:

\vspace{-10pt}

\begin{align}
\mu_M :   \mathbb{R} \rightarrow \left[0,1\right] :  x \rightarrow \nonumber\\
\begin{cases}
\nonumber
0 & \mbox{ if } \left(x < D-a\right) \vee \left(x > D+b\right) \\
\frac{\left[x-\left(D-a\right)\right]}{a} & \mbox{ if } \left(x \geq D-a\right) \wedge (x \leq D)  \\
-\frac{\left[x-\left(D+b\right)\right]}{b} & \mbox{ if } \left(x \geq D\right) \wedge (x \leq D+b)
\end{cases}
\end{align}

%\begin{figure}[h!]
%  \centering
%  \input{./graphs/triangular.tex}
%  \caption{Triangular membership function.}
%  \label{fig:triangular}
%\end{figure}

\subsection{Interval Evaluation by Ill-known Constraints}
The problem of interval evaluation is more generally explained in \cite{Pons2011}: the need exists to know if all points in a crisp interval $I$ reside between the boundaries of an ill-known interval $\left[ X , Y \right]$. In \cite{Pons2011}, the notion of an \emph{ill-known constraint} is introduced:

\begin{definition}
Given a universe $U$, an ill-known constraint $C$ on a set $A \subseteq U$ is specified by means of a binary relation $R \subseteq U^{2}$ and a fixed, ill-known value denoted by its possibilistic variable $V$ over $U$, i.e.:
\begin{align}
C \triangleq (R,V)
\end{align}
Set $A$ now satisfies the constraint if and only if:
\begin{align}
\forall a \in A : (a,V) \in R
\end{align}
\end{definition}

An example of an ill-known constraint is $C_{ex} \triangleq (<, X)$. Some set $A$ then satisfies $C_{ex}$ if $\forall a \in A : a < X$, given possibilistic variable $X$.

The satisfaction of a constraint $C \triangleq (R,V)$ by a set $A$ is basically still a Boolean matter, but due to the uncertainty about the ill-known value $V$, it can be uncertain whether $C$ is satisfied by $A$ or not \cite{Pons2011}. In fact, this satisfaction now behaves as a proposition. Based on the possibility distribution $\pi_{V}$ of $V$, the possibility and necessity that $A$ satisfies $C$ can be found. This proposition can thus be seen as a possibilistic variable on $\mathbb{B}$. The required possibility and necessity are:

\vspace{-10pt}

\begin{align}
\Pos(A\text{ satisfies }C) & =\\
\nonumber
\min_{a \in A}\left(\sup_{(a,w) \in R}\pi_{V}(w)\right) \label{ill-known-pos}\\
\Nec(A\text{ satisfies }C) & =\\
\nonumber
\min_{a \in A}\left(\inf_{(a,w) \notin R} 1-\pi_{V}(w)\right) \label{ill-known-nec}
\end{align}

So far, we have shown how it can be verified whether or not a set satisfies an ill-known constraint. However, from the examples at the beginning of this section, it is observed that Boolean combinations of constraints are required. For example, the problem of interval evaluation as explained earlier requires that all elements of an interval $[a,b]$ are larger than a value $X$ and at the same time smaller than a value $Y$, which implies that a conjunctive Boolean combination of both constraints must be satisfied. To allow Boolean combinations of constraints, the following definitions are introduced.
\begin{definition}
\label{def:set-evaluation}
Consider a universe $U$, an $n$-ary vector $\mathbf{C}$ of constraints and a Boolean function $\bool:\mathbb{B}^{n}\rightarrow\mathbb{B}$. An evaluation function is defined by:
\begin{equation}
\lambda:\Pow(U)\rightarrow\mathbb{B}:A\mapsto\bool\Big(C_{1}(A),...,C_{n}(A)\Big).
\end{equation}
\end{definition}
Definition \ref{def:set-evaluation} presents the definition of an evaluation function that evaluates a Boolean combination of some basic constraints. Informally, it states that a set $A$ passes the evaluation made by $\lambda$ if the Boolean combination of some propositions equals $T$. This crisp definition can be generalized to the case of ill-known constraints.
\begin{definition}
\label{def:ill-known-sets}
Consider a universe $U$, an $n$-ary vector $\mathbf{C}$ of ill-known constraints and a Boolean function $\bool:\mathbb{B}^{n}\rightarrow\mathbb{B}$. The uncertainty about the evaluation of a set $A$ by an evaluation function $\lambda$ is then given by:
\begin{equation}
\forall A\in\Pow(U):\pi_{\lambda(A)}=\widetilde{\bool}\Big(\pi_{C_1(A)},...,\pi_{C_n(A)}\Big)\\
\end{equation}
Hereby, $\widetilde{\bool}$ is the possibilistic extension of $\bool$.
\end{definition}
It is well known that any Boolean function $\bool$ can be cast to a canonical form~\cite{McCluskey1965}, requiring only the logical conjunction $\wedge$, logical disjunction $\vee$ and logical negation. Therefore, only the case of Boolean conjunction, Boolean disjunction and Boolean negation will be treated within the scope of this paper. By applying the possibilistic extensions of $\wedge$, $\vee$ and $\neg$, concrete equations are obtained for the calculations of uncertainty about the evaluation of a set by means of an evaluation function $\lambda$. In the case of conjunction (i.e., $\bool=\wedge$), the inference of uncertainty about the evaluation of a set reduces to:
\begin{eqnarray}
\label{eq:conjunctive1}
\forall A\in\Pow(U):\Pos(\lambda(A))&=\\
\nonumber
\min_{i=1}^n\Pos\left(C_i(A)\right)\\
\label{eq:conjunctive2}
\forall A\in\Pow(U):\Nec(\lambda(A))&=\\
\nonumber
\min_{i=1}^n\Nec\left(C_i(A)\right).
\end{eqnarray}
In the case of disjunction (i.e. $\bool=\vee$), the inference of uncertainty about the evaluation of a set reduces to:
\begin{eqnarray}
\label{eq:disjunctive}
\forall A\in\Pow(U):\Pos(\lambda(A))&=\\
\nonumber
\max_{i=1}^n\Pos\left(C_i(A)\right)\\
\forall A\in\Pow(U):\Nec(\lambda(A))&=\\
\nonumber
\max_{i=1}^n\Nec\left(C_i(A)\right).
\end{eqnarray}
Note that by using the functions $\min$ and $\max$ here, there is an implicit assumption that the possibilistic variables $\pi_{C_i}$ are mutual $\min$-dependent in the sense of De Cooman (i.e. non-interactive). For an extensive reading on (in)dependency of possibilistic variables, the reader is referred to~\cite{GertDeCooman1997b},\cite{GertDeCooman1997a},\cite{GertDeCooman1997}. In case of $\neg$, we get:
\begin{eqnarray}
\label{eq:negation}
\forall A\in\Pow(U):\Pos(\neg\lambda(A))&=\\
\nonumber
1-\Nec(\lambda(A))\\
\forall A\in\Pow(U):\Nec(\neg\lambda(A))&=\\
\nonumber
1-\Pos(\lambda(A)).
\end{eqnarray}

Now, e.g., to check if crisp interval $I = \left[j, k\right]$ is included in $\left[X, Y\right]$, 2 ill-known constraints are constructed:

%We assume that $X$ specifies the lower bound and $Y$ the upper bound for a given interval, we want to known whether all points in the interval are larger than or equal to $X$ and smaller than or equal to $Y$. Therefore, we consider two ill-known constraints:

\vspace{-10pt}

\begin{eqnarray}
C_1 & \triangleq\left(\geq,X\right)\\
C_2 & \triangleq\left(\leq,Y\right)
\end{eqnarray}

To calculate the possibility and necessity concerning a conjunction of constraints, the $\min$ operator can be used. The possibility and necessity of $I$ being included in $\left[X, Y\right]$ are now: %satisfying both constraints is then:

\vspace{-10pt}

\begin{align}
\label{eq:interval-pos}
\Pos(I\text{ satisfies }C_1\ and\ C_2) & =\\
\nonumber
\min_{a \in I}\left(\sup_{a \geq w}\pi_{X}(w),\sup_{a \leq v}\pi_{Y}(v)\right)\\
\label{eq:interval-nec}
\Nec(I\text{ satisfies }C_1\ and\ C_2) & =\\
\nonumber
\min_{a \in I}\left(\inf_{a < w} 1-\pi_{X}(w),\inf_{a > v} 1-\pi_{Y}(v)\right).
\end{align}

Sometimes, an ill-known value might be specified by some boolean combination of ill-known constraints. This allow to define ill-known values by means of relationships with respect to other ill-known points.

%\begin{definition}
%\emph{Ill-known point defined by ill-known constraints}
%Consider a universe $U$, an n-ary vector $\mathbf{C}$ of ill-known constraints and a Boolean function $\bool:\mathbb{B}^{n}\rightarrow\mathbb{B}$. An ill-known value $X$ is defined by:
%\begin{align}
%\label{eq:ill-known-value-def-by-const}
%X \in \Pow(U):\pi_{X}=\widetilde{\bool}\Big(\pi_{C_1(Z_1)},...,\pi_{C_n(Z_n)}\Big)\\
%\end{align} 
%\end{definition}

\begin{definition}
Consider a universe $U$, an n-ary vector $\mathbf{C}$ of ill-known constraints over the variables $Z_1, \ldots, Z_n$ and a Boolean function $\bool:\mathbb{B}^{n}\rightarrow\mathbb{B}$. The uncertainty about the evaluation of an ill-known value $X$ is given by:
\begin{align}
\label{eq:ill-known-value-def-by-const-unc}
X \in \Pow(U):\pi_{X}=\widetilde{\bool}\Big(\pi_{C_1(Z_1)},\ldots,\pi_{C_n(Z_n)}\Big)
\end{align} 
\end{definition}

\begin{example}
Consider the ill-known values $X = \left[12, 2, 2\right]$ and $Y = \left[18, 2, 1 \right]$. The ill-known value $Z$ is defined using ill-known constraints w.r.t. the ill-known values $X$ and $Y$:
\begin{align}
\nonumber
Z \triangleq \left \lbrace C_1\left(>,X\right) \wedge C_2(\leq,Y) \right \rbrace
\end{align}
$Z$ is a fuzzy interval defined by a trapezoidal shape given by $\left[12,\ 14,\ 18,\ 19 \right]$
Figure \ref{fig:example-ill-known-by-const} illustrates the relations among the variables $X$, $Y$ and $Z$.
\end{example}

\vspace*{13pt}
\begin{center}
{
\includegraphics[scale=1]{./graphs/ill-known-by-constraints.pdf}
\label{fig:example-ill-known-by-const}
}
\end{center}
%\centerline{ \psfig{file=./graphs/Y-time-point.eps}}
\vspace*{10pt}
\fcaption{Ill-known values $X$ and $Y$. The grey area represents the ill-known value $Z$ defined by two ill-known constraints.}
\vspace*{13pt}

%\paragraph{Example} Consider the ill-known values $X = \left[5, 2, 8\right]$ and $Y = \left[9, 7, 10 \right]$. The knowledge about the evaluation of the interval $\left[a, b \right]$  is given by the expressions \eqref{eq:interval-pos},\eqref{eq:interval-nec}.  Figure~\ref{fig:3d-possibility} shows a 3D plot of the possibility that an interval $[a,b]$ passes the evaluations specified by the ill-known constraints. Note the triangular form for the resulting possibility distribution since the condition $a \leq b$ holds.
%
%\begin{figure}[h!]
%\centering
%\includegraphics[scale=0.4]{graphs/3D_possibility.eps}
%\caption{Possibility of evaluation for the interval $[a,b]$.}
%\label{fig:3d-possibility}
%\end{figure}
%The necessity plot is obtained in a similar way and is shown in Figure~\ref{fig:3d-necessity}. Notice that the necessity measure is not normalized because the supports of $X$ and $Y$ overlap.
%\begin{figure}[h!]
%\centering
%\includegraphics[scale=0.4]{graphs/3D_necessity.eps}
%\caption{Necessity of evaluation for the interval $[a,b]$.}
%\label{fig:3d-necessity}
%\end{figure}

### 2012-04-10 trunk/ijcis-pos-model/sources/timeRepresentation.tex
Update of the draft.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Time representation
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

This section is devoted to specify the representation of time within the framework of the possibility theory. First of all, the specification for a single ill-known temporal point shall be explained. Then, the formal specification and the related constraints are given for an ill-known valid-time interval.

\subsection{\label{subsec:ill-known-point-rep}Ill-known time point}
An ill-known time point $X$ is a precise time point that for some reason, it is not fully specified. Note that $X$ has only one possible value but that value is unspecified.

\begin{definition}
\label{def:ill-known-time-point}
\emph{Ill-known time point}\\
Consider a time domain $\T$, uncertainty about the values of the ill-known time point $X$ is given by the possibility distribution $\pi_X$:
\begin{equation}
\label{eq:ill-known-time-point}
\Pi(X) = \pi_X(t) \in \left[0,1\right] , t \in \T
\end{equation}
\end{definition}

\begin{definition}
\label{def:ill-known-domain}
\emph{Domain for an ill-known time point}\\
Consider $\Pow(\T)$ the set of all the possibility distributions over $\T$, and the three fuzzy constants \emph{UNKNOWN} $=\left\lbrace 1/t, t \in \T \right\rbrace$, \emph{UNDEFINED} $=\left\lbrace 0/t, t \in \T \right\rbrace$ and \emph{NULL} $=\lbrace 1/$ \emph{UNKNOWN}, $1/$ \emph{UNDEFINED} $\rbrace$. The domain for an ill-known time point $X$ is given by: 
\begin{align}
\label{eq:ill-known-domain}
\Dom (X) =  \lbrace \Pow(\T) & \cup \mbox{\emph{UNKNOWN}} \\
\nonumber
&\cup \mbox{\emph{UNDEFINED}} \\
\nonumber
&\cup \mbox{\emph{NULL}} \rbrace
%
\end{align}
\end{definition}

It is also possible to specify an ill-known time point by just a set of ill-known constraints plus

\subsubsection{\label{subsubsec:ill-known-time-datatypes}Datatypes}
The datatypes for an ill-known time point $X$ are shown in Table \ref{tbl:time-point-types}.
\vglue13pt
%\begin{table}[htbp]
\tcap{Data types for a time point}
\centerline{\small DATA TYPES}
\vglue-6pt
\centerline{\small\baselineskip=13pt
\begin{tabular}{c p{2cm} p{2cm}}\\
\hline
1 & A single time point &  $1/x, x \in \T$\\
2 & A possibility distribution in the numeric domain & A fuzzy number or a fuzzy interval.\\
3 & An unknown value & \textbf{UNKNOWN}$= \lbrace1/t, t \in \T \rbrace$\\
4 & An undefined value & \textbf{UNDEFINED}$= \lbrace 0/t, t \in \T \rbrace$\\
5 & A null value & \textbf{NULL} $=\lbrace 1/$Unknown, 1/Undefined $\rbrace$\\
\hline\\
\end{tabular}
\label{tbl:time-point-types}
}

%Example of different uses of the ill-known time value.
\begin{example}
Consider a historical database with data from diplomatic documents. The following fields are stored: the digital identifier \emph{ID} which is the primary key  and the stimated time when the document was sent (field \emph{Date}). Table \ref{tbl:sample-time-point} contains some example data from this database. 
\end{example}

\vglue13pt
%\begin{table}[htbp]
\tcap{Sample of the historical database}
%\centerline{\small DATA TYPES}
\vglue-6pt
\centerline{\small\baselineskip=13pt
\begin{tabular}{c c}\\
\textbf{ID} & Date\\
\hline
23454 & Unknown \\ 
34563 & 11/12/1204 \\
12211 & $\left[7/2/1204,30,30\right]$ \\
23455 & $\left[10,10/6/1204,20/6/1204,15 \right]$ \\
\hline\\
\end{tabular}
\label{tbl:sample-time-point}
}

In that database, for the document with ID=23454, all the dates in the domain are equally possible. Nevertheless, the document 34563 was sent in the crisp (exact) date 11/12/1204. The time for documents 12211 and 23455 are specified with several possibility distributions. The first one is also known as a fuzzy number whereas the second one is also known as a fuzzy interval.
\textcolor{red}{References to the preliminaries}.

%\end{example}

\subsection{\label{subsec:ill-known-interval}Ill-known time interval}

\begin{definition}
\emph{Ill-known time interval}\\
An ill-known time interval denoted by $\left[X, Y\right]$ is a precise time interval that for some reason, its boudaries are not precissely known. The ill-known interval is defined by two ill-known points, namely $X$,$Y$.
In order to know if all the points in the crisp time interval $I=\left[a, b\right]$ are within the boudaries of the ill-known time interval $\left[X, Y\right]$ we define the following two ill-known constraints $C_{1}, C_{2}$ 
\begin{eqnarray}
C_1\stackrel{\triangle}{=}\left(\geq,X\right) \wedge \\
C_2\stackrel{\triangle}{=}\left(\leq,Y\right)
\end{eqnarray}
Note that the boolean operator applied to the constraints is $\bool=\wedge$.
This means that we want to know if all the points in the interval $I$ are greater or equal to $X$ and, at the same time, smaller or equal to $Y$.
\end{definition}

%% equations for the poss and nec!!
The simplified equations for the possibility and the necessity measures are~\cite{Pons2011}:
\begin{eqnarray}
\label{eq:interval-pos}
\Pos\left(\lambda([a,b])\right)&=&\\
\nonumber
\min\bigg(\sup_{a\leq w}\pi_{X}(w),\sup_{b\geq w}\pi_{Y}(w)\bigg)\\
\label{eq:interval-nec}
\Nec\left(\lambda([a,b])\right)&=&\\
\nonumber
\min\bigg(\inf_{a>w}1-\pi_{X}(w),\inf_{b<w}1-\pi_{Y}(w)\bigg).
\end{eqnarray}

\subsubsection{\label{subsubsec:open-interval}Open ill-known time intervals}
Quite often, the user may want to specify time intervals with open boudaries in one or both endpoints. Consider an ill-known time interval $\left[X, Y\right]$. Then it is possible to distinguish between the following two types of open intervals:
\begin{enumerate}
\item
\emph{Completely unknown time interval}: Both starting and ending points are unknown, therefore the hole interval is specified by the \emph{UNKNOWN} constant. Due to the representation, both endpoints, say $X$ and $Y$ are stored. Thus, the interval is stored as $[$ \emph{UNKNOWN}, \emph{UNKNOWN} $]$.
\item
\emph{Semi-open time interval}: Only one of the two ill-known endpoints for the time interval $\left[X, Y\right]$ is unknown. Note that due to the ill-known constraints $C_{1}, C_{2}$ it is not possible a representation like $[$ \emph{UNKNOWN} $, Y]$ or $[X, $ \emph{UNKNOWN} $]$. The solution adopted for this problem is explained in the following paragraph.
\end{enumerate}

\paragraph{Representation of semi-open time intervals}
As mentioned before, the problem resides in the representation of these kind of interval. In order to get a short representation for these intervals, two constants are defined: $FB$ \emph{From the Beggining}: it is only valid for the left value of the interval and $UC$ \emph{Until Changed} (this name is usual in the temporal database community, see ~\cite{Jensen1994}) which is only valid for the right value of the interval. Note that these two constants are aliases for a function that obtains both possibility and necessity measures.

Because of the ill-known constraints $C_{1}, C_{2}$, a function called \emph{Open} should be defined in order to deal with a proper representation of these intervals:

\begin{definition}
\label{def:open-func}
$Open(C)$\\
The function $\Open(C)$ provides both possibility and necessity measures for all the points in the open part of a semi-open ill-known time interval.
The parameter $C$ is an ill-known constraint $C\stackrel{\triangle}{=}\left(R,T\right)$ as defined in equation \eqref{rand} and in~\cite{Pons2011}.
The possibility and necessity measures are defined by:
\begin{eqnarray}
\label{eq:open-pos-nec}
\Pos\left(\Open(C)\right) &=& \Big(\sup_{r\in\T \Rp\  w}\pi_{T}(w)\Big)\\
\Nec\left(\Open(C)\right) &=& \Big(\inf_{r\in\T \Rn\  w}1-\pi_{T}(w)\Big)
\end{eqnarray}
Where the values for the binary relations $\Rp$ and $\Rn$ are shown in Table \ref{tbl:open-pos-nec-rels}.
\end{definition}

\vglue13pt
%\begin{table}[htbp]
\tcap{Relations for the $\Open(C)$ function.}
\centerline{\small Relations}
\vglue-6pt
\centerline{\small\baselineskip=13pt
\begin{tabular}{c c c}\\
\hline
$R$ & $\Rp$ & $\Rn$\\ \hline
$<$ & $>$ & $\leq$ \\
$\leq$ & $\geq$ & $<$ \\
$>$ & $<$ & $\geq$ \\
$\geq$ & $\leq$ & $>$ \\
\hline\\
\end{tabular}
\label{tbl:open-pos-nec-rels}
}

As explained before, the constants $\FB$ and $\UC$ are aliases for the function $\Open$ with the following parameters:
\begin{eqnarray}
\FB &=& \Open(C_2)\\
\UC &=& \Open(C_1)
\end{eqnarray}

\begin{example}
Consider an ill-known time interval given by $\left[\FB, Y \right]$. Consider also that in this case, $Y= \left[15/10/2012,3,4 \right] $. Figure \ref{fig:example-open-time-interval} shows the representation for $Y$. The user wants to obtain the possibility and the necessity measures for the $\FB$ part of the interval are:
\end{example}

\begin{eqnarray}
\nonumber
\FB = \Open(C_2) \mbox{ with } C_2\stackrel{\triangle}{=}\left(\leq,Y\right)\\
\nonumber
\Pos\left(\Open(C)\right) = \Big(\sup_{r\in\T \geq\  w}\pi_{Y}(w)\Big)\\
\nonumber
\Nec\left(\Open(C)\right) = \Big(\inf_{r\in\T <\  w}1-\pi_{Y}(w)\Big)
\end{eqnarray}

\vspace*{13pt}
\begin{center}
{
\includegraphics[scale=1]{./graphs/open-time-interval.pdf}
\label{fig:example-open-time-interval}
}
\end{center}
%\centerline{ \psfig{file=./graphs/Y-time-point.eps}}
\vspace*{10pt}
\fcaption{Possibility distribution for $Y$, and possibility and necessity measures for the open ill-known point, $X$}
\vspace*{13pt}

\subsubsection{\label{subsubsec:ill-known-time-interval-datatypes}Datatypes}
In order to represent properly an ill-known time interval in a database, some datatypes are needed. Because of the ill-known constraints, not all the combinations of datatypes for each ill-known time point (see Table \ref{tbl:time-point-types}) are allowed. Table \ref{tbl:time-interval-types} shows all the possible combinations for the datatypes for an ill-known time interval denoted by $\left[X, Y\right]$.

\vglue13pt
%\begin{table}[htbp]
\tcap{Data types for an ill-known time interval $\left[X, Y\right]$.}
\centerline{\small DATA TYPES}
\vglue-6pt
\centerline{\small\baselineskip=13pt
\begin{tabular}{c c p{2cm}}\\
\hline
$X$  & $Y$  & Description\\
\hline
$\lbrace1,2 \rbrace$ & $\lbrace1,2 \rbrace$ &  an ill-known time interval.\\
3 & 3 & An unknown time interval. \\
$\FB$ & $\lbrace1,2 \rbrace$ & a left-open time interval.\\
$\lbrace1,2 \rbrace$ & $\UC$ & a right-open time interval.\\
\hline\\
\end{tabular}
\label{tbl:time-interval-types}
}

%\subsection{\label{subsec:time-interval-constraint}Ill-known valid time-interval}
%The representation of a possibilistic valid-time interval is given by two ill-known points: $\left[S,E \right]$ the starting and ending points respectively.  
%
%
%A valid temporal interval $\left[S,E\right]$ in the system is a combination for the values of $S$,$E$. Note that the only allowed combination of types is shown in table \ref{tbl:valid-time-interval}.

### 2012-04-12 ijcis-pos-model/sources/timeRepresentation.tex
new modifications

--- 2012-04-10 trunk/ijcis-pos-model/sources/timeRepresentation.tex
@@
 
+\vglue13pt
+%\begin{table}[htbp]
+\tcap{Relations for the $\Open(C)$ function.}
+\centerline{\small Relations}
+\vglue-6pt
+\centerline{\small\baselineskip=13pt
+\begin{tabular}{c c c}\\
+\hline
+$R$ & $\Rp$ & $\Rn$\\ \hline
+$<$ & $>$ & $\leq$ \\
+$\leq$ & $\geq$ & $<$ \\
+$>$ & $<$ & $\geq$ \\
+$\geq$ & $\leq$ & $>$ \\
+\hline\\
+\end{tabular}
+\label{tbl:open-pos-nec-rels}
+}
+
 \subsubsection{\label{subsubsec:ill-known-time-interval-datatypes}Datatypes}
@@
 
+\subsection{\label{subsec:fuzzy-allen-relations}Allen's relations}
+
+\vglue13pt
+%\begin{table}[htbp]
+\tcap{\label{tbl:fuzzy-allen-relations}Constructs of constraints related to their respective Allen relations, as used in the presented work.
+In this table, the ill-known time interval $J = \left[X, Y\right]$ in a record $r$ has a start point described by possibilistic variable $X$ and an end point described by possibilistic variable $Y$.
+The crisp time interval in the user's temporal demand is denoted $I$.}
+\centerline{\small Relations}
+\vglue-6pt
+\centerline{\small\baselineskip=13pt
+\begin{tabular}{c p{4cm} }\\
+\hline
+Allen Relation & Construct of constraints\\
+\hline
+I before J & $C_1 \triangleq \left(<,X\right)$ \\
+\hline
+I equal J & $\left(C_1 \triangleq \left(\geq,X\right)\right) \wedge \neg \left(C_2 \triangleq \left(\neq,X\right) \right) \wedge \left(C_3 \triangleq \left(\leq,Y\right) \right) \wedge \neg \left(C_4 \triangleq \left(\neq,Y\right)\right)$ \\
+\hline
+I meets J & $\left(C_1 \triangleq \left(\leq,X\right)\right) \wedge \neg \left(C_2 \triangleq \left(\neq,X\right) \right)$ \\
+\hline
+I overlaps J & $\left(C_1 \triangleq \left(<,Y\right)\right) \wedge \neg \left(C_2 \triangleq \left(\leq,X\right) \right) \wedge \neg \left(C_3 \triangleq \left(\geq,X\right) \right)$ \\
+\hline
+I during J & $\left( \left (C_1 \triangleq \left(>,X\right) \right) \wedge \left(C_2 \triangleq \left(\leq,Y\right) \right) \right) \vee \left( \left(C_3 \triangleq \left(\geq,X\right) \right) \wedge  \left(C_4 \triangleq \left(<,Y\right) \right) \right)$ \\
+\hline
+I starts J & $\left(C_1 \triangleq \left(\geq,X\right) \right) \wedge \neg \left(C_2\triangleq \left(\neq,X\right)\right)$ \\
+\hline
+I finishes J & $\left(C_1 \triangleq \left(\leq,Y\right) \right) \wedge \neg \left(C_2\triangleq \left(\neq,Y\right)\right)$ \\
+\hline\\
+\end{tabular}
+}
+
 %\subsection{\label{subsec:time-interval-constraint}Ill-known valid time-interval}

### 2012-04-12 trunk/ijcis-pos-model/sources/preliminaries.tex
new modifications

@@
 \end{definition}
-If this modal value $m$ is unique, then $M$ is referred to as a \emph{fuzzy number}. In other words, if the core of a fuzzy interval is a singleton, it is referred to as a fuzzy number.
 
-A simple form of the membership function of a fuzzy interval is a trapezoidal function.
+\vspace*{13pt}
+\begin{center}
+{
+\includegraphics[scale=0.25]{./graphs/triangular.pdf}
+
+}
+\end{center}
+%\centerline{ \psfig{file=./graphs/Y-time-point.eps}}
+\vspace*{10pt}
+\fcaption{\label{fig:triangular-dist}Trapezoidal membership function.}
+%  \label{fig:fuzzy-validity-period}
+\vspace*{13pt}
+
+If this modal value $m$ is unique, then $M$ is referred to as a \emph{fuzzy number}.
+In other words, if the core of a fuzzy interval is a singleton, it is referred to as a fuzzy number (Figure \ref{fig:triangular-dist}).
+
+A simple form of the membership function of a fuzzy interval is a trapezoidal function (Figure \ref{fig:trapezoidal-dist}.
 It can be shown that such a membership function $\mu_T$ for a fuzzy interval $T$ is convex and normalized.
@@
 
+\vspace*{13pt}
+\begin{center}
+{
+\includegraphics[scale=0.25]{./graphs/trapezoidalDistribution.pdf}
+
+}
+\end{center}
+%\centerline{ \psfig{file=./graphs/Y-time-point.eps}}
+\vspace*{10pt}
+\fcaption{\label{fig:trapezoidal-dist}Trapezoidal membership function.}
+%  \label{fig:fuzzy-validity-period}
+\vspace*{13pt}
 %\def\JPicScale{0.5}
@@
 
-Now, e.g., to check if crisp interval $I = \left[j, k\right]$ is included in $\left[X, Y\right]$, 2 ill-known constraints are constructed:
+\begin{example}
+Consider that we want to check if crisp interval $I = \left[j, k\right]$ is included in $\left[X, Y\right]$, 2 ill-known constraints are constructed:
 
@@
 \end{align}
+\end{example}
 
-Sometimes, an ill-known value might be specified by some boolean combination of ill-known constraints.
+There is a special boolean combination of constraints that is of particular interest:
+\begin{definition}
+\emph{Convex combination of ill-known constraints}. Consider the $C_1$ and $C_2$ two ill-known constraints. A convex combination of the constraints is given by:
+\begin{align}
+\label{eq:convex-combination}
+CC \triangleq \left \lbrace C_1 \wedge C_2 \right \rbrace
+\end{align}
+In a more general way, it is possible to define the convex combination of an n-ary vector  $\mathbf{C}$ of constraints:
+\begin{align}
+\label{eq:nary-convex-combination}
+CC \triangleq \left \lbrace C_1 \wedge \ldots \wedge C_n \right \rbrace
+\end{align}
+\end{definition}
+
+\begin{theorem}
+\label{th:convex-combination-ill-known-constraints}
+Consider the convex combination $CC$ of any n-ary vector $\mathbf{C} = \left \lbrace C_{1_{Z_1}}, \ldots, C_{n_{Z_n}} \right \rbrace$ of constraints over the ill-known variables $Z_1, \ldots, Z_n$.
+Then, if $\pi_{C_1(Z_1)}, \ldots, \pi_{C_n(Z_n)}$ are convex, then  $\pi_{CC}$ is also convex.
+\end{theorem}
+
+\begin{proof}
+Let $CC \triangleq \left \lbrace C_{1_{Z_1}}, \ldots, C_{n_{Z_n}}  \right \rbrace$. Then:
+\begin{align}
+\label{eq:proof-convex1}
+\pi_{CC} \left( \lambda x_1 + \left( 1 - \lambda \right)x_2 \right) = \\
+\min  \big( \pi_{C_1(Z_1)}\left( \lambda x_1 + \left( 1 - \lambda \right)x_2 \right), \ldots,\\
+   \pi_{C_n(Z_n)}\left( \lambda x_1 + \left( 1 - \lambda \right)x_2 \right)  \big)
+\end{align}
+Since $\pi_{C_1(Z_1)}, \ldots, \pi_{C_n(Z_n)}$ are convex:
+\begin{align}
+\label{eq:proof-convex2}
+\pi_{C_1(Z_1)} \left(\lambda x_1 + \left( 1 - \lambda \right)x_2  \right) \geqq \\
+\nonumber
+\min \big(\pi_{C_1(Z_1)} \left( x_1 \right),\pi_{C_1(Z_1)} \left( x_2 \right)  \big)\\
+\nonumber
+\vdots \\
+\nonumber
+\pi_{C_n(Z_n)} \left(\lambda x_1 + \left( 1 - \lambda \right)x_2  \right) \geqq \\
+\nonumber
+\min \big(\pi_{C_n(Z_n)} \left( x_1 \right),\pi_{C_n(Z_n)} \left( x_2 \right) \big) 
+\end{align}
+Then, by using equation \eqref{eq:proof-convex1}:
+\begin{align}
+\label{eq:proof-convex3}
+\pi_{CC} \left( \lambda x_1 + \left( 1 - \lambda \right)x_2 \right) \geqq  \\
+\nonumber
+\min \big( \min \left(\pi_{C_1(Z_1)} \left( x_1 \right),\pi_{C_1(Z_1)} \left( x_2 \right) \right),\\
+\nonumber
+\ldots, \min \left(\pi_{C_n(Z_n)} \left( x_1 \right),\pi_{C_n(Z_n)} \left( x_2 \right) \right) \big)
+\end{align}
+Which is equivalent to the following:
+\begin{align}
+\label{eq:proof-convex4}
+\pi_{CC} \left( \lambda x_1 + \left( 1 - \lambda \right)x_2 \right) \geqq  \\
+\nonumber
+\min \big( \min \left(\pi_{C_1(Z_1)} \left( x_1 \right), \ldots, \pi_{C_n(Z_n)} \left( x_1 \right) \right),\\
+\nonumber
+\ldots, \min \left(\pi_{C_1(Z_1)} \left( x_2 \right), \ldots, \pi_{C_n(Z_n)} \left( x_2 \right) \right) \big)
+\end{align}
+Finally we obtain:
+\begin{align}
+\label{eq:proof-convex5}
+\pi_{CC} \left( \lambda x_1 + \left( 1 - \lambda \right)x_2 \right) \geqq  \\
+\nonumber
+\min \big(\pi_{CC} \left( x_1 \right), \pi_{CC} \left( x_2 \right) \big)
+\end{align}
+\end{proof}
+
+Sometimes, an ill-known value might be specified by a convex combination of ill-known constraints.
 This allow to define ill-known values by means of relationships with respect to other ill-known points.
@@
 \begin{definition}
-Consider a universe $U$, an n-ary vector $\mathbf{C}$ of ill-known constraints over the variables $Z_1, \ldots, Z_n$ and a Boolean function $\bool:\mathbb{B}^{n}\rightarrow\mathbb{B}$.
+Consider a universe $U$, $X$ an ill-known value and $CC \triangleq \left \lbrace C_{1_{Z_1}}, \ldots, C_{n_{Z_n}}  \right \rbrace$ a convex combination of ill-known constraints over the variables $Z_1, \ldots, Z_n$.
+The value $X$ is defined by means of the convex combination:
+\begin{align}
+\label{eq:ill-known-value-by-convex-constraints}
+X \triangleq CC 
+\end{align}
+
 The uncertainty about the evaluation of an ill-known value $X$ is given by:
@@
 \label{eq:ill-known-value-def-by-const-unc}
-X \in \Pow(U):\pi_{X}=\widetilde{\bool}\Big(\pi_{C_1(Z_1)},\ldots,\pi_{C_n(Z_n)}\Big)
+X \in \Pow(U):\pi_{X}=\pi_{CC}
 \end{align} 
+Note that $\pi_{X}$ is convex since $\pi_{CC}$ is convex as demostrated in theorem \ref{th:convex-combination-ill-known-constraints}.
 \end{definition}
@@
 \begin{example}
-Consider the ill-known values $X = \left[12, 2, 2\right]$ and $Y = \left[18, 2, 1 \right]$. The ill-known value $Z$ is defined using ill-known constraints w.r.t. the ill-known values $X$ and $Y$:
+Consider the ill-known values $X = \left[12, 2, 2\right]$ and $Y = \left[18, 2, 1 \right]$. The ill-known value $Z$ is defined using the convex combination $CC$ of constraints $C_1$ and $C_2$:
 \begin{align}
 \nonumber
-Z \triangleq \left \lbrace C_1\left(>,X\right) \wedge C_2(\leq,Y) \right \rbrace
+CC \triangleq \left \lbrace C_1\left(>,X\right) \wedge C_2(\leq,Y) \right \rbrace \\
+\nonumber
+Z \triangleq CC
 \end{align}
@@
 \vspace*{10pt}
-\fcaption{Ill-known values $X$ and $Y$. The grey area represents the ill-known value $Z$ defined by two ill-known constraints.}
+\fcaption{Ill-known values $X$ and $Y$. The grey area represents the ill-known value $Z$ defined by the convex combination of the two ill-known constraints $C_1$ and $C_2$.}
 \vspace*{13pt}

### 2012-04-13 trunk/ijcis-pos-model/sources/temporaldatabases.tex
new modifications
updated version.

--- 2012-04-10 ijcis-pos-model/sources/temporaldatabases.tex
@@
 This means that the attribute uses the date/time domain, but the database model does not treat the attribute differently from non-temporal attributes.
-Depending on the time managed, a database is classified as either a \textbf{Valid Time Database} (\emph{VTDB}), a \textbf{Transaction Time Database} (\emph{TTDB}), a \textbf{bi-temporal database} (both valid and transaction time are managed) or a \textbf{tri-temporal database} (valid time, transaction time and decision time are managed).
+Depending on the time managed, a database is classified as either a \textbf{Valid Time Database} (\emph{VTDB}), a \textbf{Transaction Time Database} (\emph{TTDB}), a \textbf{bi-temporal database} (both valid and transaction time are managed) or a \textbf{tri-temporal} or \emph{multitemporal} database (valid time, transaction time and decision time are managed).
 
@@
 {
-\includegraphics[scale=0.25]{./graphs/trapezoidalDistribution.pdf}
+\includegraphics[scale=0.25]{./graphs/comparisoncv.pdf}
 
@@
 a typical update sentence in SQL could be re-defined by means of several insert and update sentences.
-Therefore, in order to be clear, these high level primitives in the DMl for a valid-time database are usually noted as Insert, Modify and Delete.
+Therefore, in order to be clear, these high level primitives in the DML for a valid-time database are usually noted as Insert, Modify and Delete.
 In the following subsections each primitive is defined and explained.
@@
 
+In order to simplify the algorithms for the manipulation of data, some auxiliary functions are defined:
+
+\begin{definition}
+\label{def:close-a-crisp-interval}
+Consider two crisp intervals $I= \left[a ,b \right]$ and $J= \left[c ,d \right]$, $c > a$.
+The CloseL$\left(I, J\right)$ function allows to close the left-open interval $I$ with respect to the first value $d$ in $J$:
+\begin{align}
+\mbox{CloseL} \left( I, J \right) &=& \\ 
+\begin{cases}
+\nonumber
+I & \mbox{ if } b \neq +\infty \\
+I=\left[a, c \right[ & \mbox{ if } b = +\infty \wedge J > I
+\end{cases}
+\end{align}
+\end{definition}
+
+Analogously, it is defined the function to close an right open interval: 
+\begin{definition}
+\label{def:close-a-crisp-interval-r}
+CloseR$\left(I, J\right)$ 
+\begin{align}
+\mbox{CloseR} \left( I, J \right) &=& \\ 
+\begin{cases}
+\nonumber
+I & \mbox{ if } a \neq -\infty \\
+I=\left]d, b \right[ & \mbox{ if } a = -\infty \wedge J < I
+\end{cases}
+\end{align}
+\end{definition}
+
+\begin{definition}
+\label{def:current-in-relation}
+Consider an object $A$ and a relation $R$ in a temporal database.
+The fucntion Current$\left(R, A \right)$ obtains the tuple $\left(A, I \right)$ that is current in the relation $R$, that is it, $I = \left[a, +\infty \right]$.
+
+\begin{align}
+\label{eq:current-in-relation}
+\mbox{Current} \left(R, A \right) &=& \\ 
+\begin{cases}
+\nonumber
+I & \mbox{ if } \exists b \in I : (A,I) \in R \wedge b = +\infty \\
+\emptyset & \mbox{ in any other case }
+\end{cases}
+\end{align}
+\end{definition}
+
+Now it is possible to close the current version of an object by combining the two previously defined functions.
+
+\begin{definition}
+\label{def:close-current-version}
+Consider an object $A$, the relation $R$ and a time interval $J$.
+Then, the function close-current$\left(R,A,J \right)$ closes any current version of the object $A$ if it exists and add the new version $\left(A, J \right)$.
+
+\begin{eqnarray}
+%\label{eq:close-current}
+\text{Close-current} \left(R, A \right) =\\
+\begin{cases}
+\nonumber
+\big \lbrace R - \left(A, I \right) \cup \left \lbrace \left(A, \mbox{ Close } \left(I, J\right) \right) \cup \left(A, J\right)\right \rbrace  \big \rbrace \\
+\nonumber
+\mbox{ if } I = \mbox{ Current } \left(R, A \right) \wedge J > \mbox{ Close } \left(I, J \right)   \\
+\nonumber R , \text{ in any other case}
+\end{cases}
+\end{eqnarray}
+\end{definition}
+
 \subsubsection{Insert}
-The interpretation of the insert operation is the following: The user wants to store in the database that an object $A$  is true for some period(s) of time $I$ in the given relation $R$.
+The user wants to store an object $A$ which is valid in the relation $R$ during the time interval specified by $I = \left(a, b \right)$.
+%
+%
+%The interpretation of the insert operation is the following: The user wants to store in the database that an object $A$  is true for some period(s) of time $I$ in the given relation $R$.
 To indicate that the object $A$ is current in the relation, the special value Until Current, $\UC$  is used.
-
 There are three main cases when performing a create operation:
@@
 
-\item The object $A$ was part of some previously recorded state: A new version of the object is inserted in the relation $R$ with the valid-time indicated by $I$.
+\item The object $A$ is in the relation $R$, but it is not current
 
@@
 R \cup \left(A, I \right), & \mbox{ if }  A \not \in R\\
-R \cup  & \mbox{ if }  A \in R \wedge \\
+R \cup  \left(A, I \right) & \mbox{ if }  A \in R \wedge \\
 R, & \mbox{ otherwise }  

### 2012-04-15 trunk/ijcis-pos-model/sources/conclusions.tex
update witha new version

We presented a valid-time model to represent and handle ill-known temporal intervals. The semantics and the implementation of the DML operations are described within this work. As future work , the definition of the Data Definiton Language as well as the querying will be considered.

### 2012-04-15 ijcis-pos-model/sources/intro.tex
update witha new version

--- 2012-03-16 trunk/book/sources/intro.tex
@@
 
-The concept of time itself is very complex to handle and interpret~\cite{klein94},~\cite{Shackle61}, though it is very natural and omnipresent.
+The concept of time itself is very complex to handle and interpret~\cite{Klein1994},~\cite{Shackle1961}, though it is very natural and omnipresent.
 As information systems often attempt the modelling of natural objects, concepts or processes, they often require modelling temporal aspects or concepts.
-Thus, several proposals have been concerned with the obtaining of theoretical models that allow the modelling or representation of time~\cite{Bolour82},~\cite{Cru97}.
+Thus, several proposals have been concerned with the obtaining of theoretical models that allow the modelling or representation of time~\cite{Bolour1982},~\cite{VanderCruyssen1997}.
 
@@
 A lot of research concerns temporal database models and their approaches to the modelling of time.
-The first efforts were towards the representation of historical information related to objects represented by records in a database~\cite{Clifford:1985:AHR:971699.318922}.
-Some proposals tried to extend the Entity Relationship Model~\cite{Klopprogge:1983}, without impact on any database standards like SQL~\cite{Sarda:1990:ESH:627277.627409}.
+The first efforts were towards the representation of historical information related to objects represented by records in a database~\cite{Clifford1985}.
+Some proposals tried to extend the Entity Relationship Model~\cite{Klopprogge1983}, without impact on any database standards like SQL~\cite{Sarda1990}.
 
@@
 For this publication, 44 temporal database researchers, among which some of the main researchers in this field, cooperated to reach a consensus on the nature and definitions of some of the main temporal database concepts and terminology.
-This glossary was subsequently updated in 1998~\cite{Jensen:Dyreson:1998:TheConsensusGlossary}.
+This glossary was subsequently updated in 1998~\cite{Dyreson1998}.
 
 An interesting issue in temporal modelling concerns relationships between temporal notions.
-Notably, Allen~\cite{Allen83} studied temporal relationships between time intervals (and as a special case time points).
+Notably, Allen~\cite{Allen1983} studied temporal relationships between time intervals (and as a special case time points).
 Among others, the querying of temporal databases has greatly profited from these temporal relationships, because they allowed for richer and more complex user-specified temporal query demands, by allowing to express more complex relationships between the temporal notions in the temporal expressions in the query and the temporal indications in the database.
@@
 
-To allow information systems to cope with these and similar data imperfections, many approaches adopt fuzzy sets~\cite{Zadeh65} for the representation of temporal information~\cite{343607},~\cite{nagypal2003},~\cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy},~\cite{Dubois:jucs_9_9:fuzziness_and_uncertainty_in}.
-The temporal relationships studied by Allen were fuzzified by several authors~\cite{ohlbach2004},~\cite{nagypal2003},~\cite{schockaert08}.
+To allow information systems to cope with these and similar data imperfections, many approaches adopt fuzzy sets~\cite{Zadeh1965} for the representation of temporal information~\cite{Mitra1994},~\cite{Nagypal2003},~\cite{Billiet2011},~\cite{Dubois2003}.
+The temporal relationships studied by Allen were fuzzified by several authors~\cite{Ohlbach2004},~\cite{Nagypal2003},~\cite{Schockaert2008}.
 Garrido et al.
-~\cite{garrido2009} present different temporal operators, defined by a combination of regular fuzzy comparisons.
-Both~\cite{garrido2009} and~\cite{Pon11} deal with uncertainty in temporal expressions concerning time intervals.
-Other approaches, like~\cite{Qia09}, use rough sets~\cite{Pawlak1995} to represent time intervals.
+~\cite{Garrido2009} present different temporal operators, defined by a combination of regular fuzzy comparisons.
+Both~\cite{Garrido2009} and~\cite{Pons2011} deal with uncertainty in temporal expressions concerning time intervals.
+Other approaches, like~\cite{Qiang2009}, use rough sets~\cite{Pawlak1995} to represent time intervals.
 
-Next to temporal modelling, some attention has been spent on temporal reasoning~\cite{Allen83}.
-Although temporal reasoning is not discussed in this chapter, it should be noted that, among others, Dubois and Prade et al.~\cite{Dubois:jucs_9_9:fuzziness_and_uncertainty_in},~\cite{Dubois89} have dealt with fuzziness and uncertainty in temporal reasoning.
+Next to temporal modelling, some attention has been spent on temporal reasoning~\cite{Allen1983}.
+Although temporal reasoning is not discussed in this chapter, it should be noted that, among others, Dubois and Prade et al.~\cite{Dubois2003},~\cite{DuBois1989} have dealt with fuzziness and uncertainty in temporal reasoning.
 
-The aim of this chapter is to present and explain some main concepts regarding time in information systems and to present and discuss some issues and techniques concerned with handling data imperfections related to time.
-The rest of this chapter is structured as follows.
-Section \ref{sec:time-domain} presents some basic concepts and terminology about temporal modelling and discusses some of its important aspects and issues, while section~\ref{sec:fuzzy-time-mod} presents some important issues concerning the combination of data imperfections and temporal modelling.
-In section \ref{sec:temporal-databases}, some basic concepts and terminology about temporal databases are presented, followed by an overview of some interesting issues concerning temporal databases and a survey of some commercial temporal database systems.
-Finally, in section \ref{sec:imprecision-uncertainty}, an approach to querying temporal databases containing imperfect temporal information is presented, followed by some conclusions and some suggestions for future work in section \ref{sec:further-research-work}.
-
-%The concept of \emph{time} is easy to understand but very complex to define~\cite{klein94}~\cite{Shackle61}. 
-%In Information Systems \emph{IS}, several proposals have been concerned with the obtaining of theoretical models that allow the representation of time~\cite{Bolour82}~\cite{Cru97}. 
-%More towards temporal databases, a lot of work have been done.
-The first efforts were towards the representation of the historical information related to an object in the database(reference).
-Some works tried to extend the Entity Relationship Model \emph{ERM} ~\cite{Klopprogge:1983} but without an impact on any database standards like SQL~\cite{Sarda:1990:ESH:627277.627409}.
-
-%The ``Consensus Glossary of Temporal Database concepts'' \cite{Dyreson1994} is the first publication where the main researchers in temporal databases define the main thesaurus of temporal database concepts.
-This glossary defines the main types of time in a temporal database and is the basis of furthers researches ~\cite{Sarda:1990:ESH:627277.627409} \cite{Jensen94thetsql2}.
-
-%Another main issue in temporal databases is the querying.
-The user specifies in the query some temporal constraints, e.g., \emph{``before this year''},\emph{``after 1995''}.
-In order to solve these temporal comparisons, between the time specified in the query and the temporal data stored in the database, some basic relationships have been defined.
-Allen~\cite{Allen83} first introduced the temporal relations between time intervals and, to a lesser extent, to time points.
-Some authors~\cite{ohlbach2004},\cite{nagypal2003},\cite{schockaert08} have softened the temporal operators defined by Allen.
-Therefore it is possible to compute the relationships between two ill-known time points or intervals.
-In~\cite{garrido2009}, some different temporal operators are defined by a combination of regular fuzzy comparisons.
-%e.g., the \emph{BEFORE} operator is defined by means of the fuzzy less than (FLT) operator.
-
-%To allow information systems to cope with uncertainty, many approaches adopt fuzzy sets~\cite{Zadeh65} for the representation of temporal information~\cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy},\cite{Dubois:jucs_9_9:fuzziness_and_uncertainty_in},~\cite{devos94}.
-The time points representing the time when the object is valid in the reality might not be precisely known.
-In order to deal with this, several models have been proposed.
-Garrido~\cite{garrido2009} proposes a model to deal with uncertainty in the time interval by means of fuzzy intervals.
-Bronselaer~\cite{Pon11} proposes a consistent framework to deal and represent ill-known time values and their relationships.
-Rough sets~\cite{Pawlak1995} have been also used to represent time intervals~\cite{Qia09} in databases.
-
-%temporal reasoning and fuzzy temporal reasoning.
-%Next to the manage of temporal information is the temporal reasoning~\cite{Allen83}.
-Dubois and Prade ~\cite{Dubois:jucs_9_9:fuzziness_and_uncertainty_in},\cite{Dubois89} deal with fuzziness and uncertainty in temporal reasoning, but this topic is outside the scope of this work.
-
-%temporal commercial systems.
-
-%The work is organized as follows:
-
-%In Section \ref{sec:time-domain} an study about the time and its properties is done.
-Section \ref{sec:temporal-databases} is an overview of the main problems when managing the time in a database, the different proposals for solving them and the temporal database proposals.
-The section concludes with a comparison among the different proposals.
-In Section \ref{sec:fuzzy-temporal-databases} we present a sum up with the proposals for handling the imprecision in temporal databases and the main deficiencies we have detected.
-Finally, there are several commercial temporal database management systems  \emph{TDMBS} like \cite{oracle2009},\cite{posgree2009},\cite{teradata2011},\cite{timedb2005}.
-All of them are analyzed and compared later on this work.
+The aim of this work is to present and explain the main operations of the manipulation language for a temporal database which stores the valid-time periods of the objects with uncertainty.
+The rest of the work is organized as follows.
+Section \ref{sec:prelim} presents some background concepts about both possibility theory and temporal databases.
+In section \ref{sec:time-rep} the representation of the valid-time intervals in the database is explained.
+Section \ref{sec:temporal-model} explains the main concepts and behaviour of the temporal model.
+The Data Manipulation Language (DML) is described and implemented.
+Finally in Section \ref{sec:conclusions} presents the conclusions and some possibilities for future research work.

### 2012-04-15 ijcis-pos-model/sources/temporaldatabases.tex
update witha new version

@@
 	\item
-	\textbf{Decision time} (\emph{DT}, proposed in \cite{Nascimento1995}) denotes the time when an event was decided to happen. 
+	\textbf{Decision time} (\emph{DT}) \cite{Nascimento1995},\cite{Chakravarthy1994},\cite{Etzion1992},\cite{Ozsoyoglu1995} denotes the time when an event was decided to happen. 
 	\end{itemize}
@@
 
+In order to simplify the algorithms for the manipulation of data, some auxiliary functions are defined:
+
+\begin{definition}
+\label{def:close-a-crisp-interval-r}
+Consider two crisp intervals $I= \left[a ,b \right]$ and $J= \left[c ,d \right]$, $c > a$.
+The CloseR$\left(I, J\right)$ function allows to close the right-open interval $I$ with respect to the first value $d$ in $J$:
+\begin{align}
+\mbox{CloseR} \left( I, J \right) &=& \\ 
+\begin{cases}
+\nonumber
+I & \mbox{ if } b \neq +\infty \\
+I=\left[a, c \right[ & \mbox{ if } b = +\infty \wedge J > I
+\end{cases}
+\end{align}
+\end{definition}
+
+Analogously, it is defined the function to close a left-open interval: 
+\begin{definition}
+\label{def:close-a-crisp-interval-l}
+CloseL$\left(I, J\right)$ 
+\begin{align}
+\mbox{CloseR} \left( I, J \right) &=& \\ 
+\begin{cases}
+\nonumber
+I & \mbox{ if } a \neq -\infty \\
+I=\left]d, b \right] & \mbox{ if } a = -\infty \wedge J < I
+\end{cases}
+\end{align}
+\end{definition}
+
+\begin{definition}
+\label{def:current-in-relation}
+Consider an object $A$ and a relation $R$ in a temporal database.
+The function Current$\left(R, A \right)$ obtains the tuple $\left(A, I \right)$ that is current in the relation $R$, that is it, $I = \left[a, +\infty \right]$.
+
+\begin{align}
+\label{eq:current-in-relation}
+\mbox{Current} \left(R, A \right) &=& \\ 
+\begin{cases}
+\nonumber
+I & \mbox{ if } \exists b \in I : (A,I) \in R \wedge b = +\infty \\
+\emptyset & \mbox{ in any other case }
+\end{cases}
+\end{align}
+\end{definition}
+
+\begin{definition}
+\label{def:first-in-relation}
+Analogously, the function First$\left(R, A \right)$ obtains the tuple $\left(A, I \right)$ where $I$ is a right-open interval and which is the first version in the relation $R$, that is it, $I = \left[-\infty, b \right]$.
+\begin{align}
+\label{eq:first-in-relation}
+\mbox{First} \left(R, A \right) &=& \\ 
+\begin{cases}
+\nonumber
+I & \mbox{ if } \exists b \in I : (A,I) \in R \wedge a = -\infty \\
+\emptyset & \mbox{ in any other case }
+\end{cases}
+\end{align}
+
+\end{definition}
+
+Now it is possible to close the current version of an object by using \eqref{def:close-a-crisp-interval-r} and \eqref{eq:current-in-relation}
+
+\begin{definition}
+\label{def:close-current-version}
+Consider an object $A$, the relation $R$ and a time interval $J$.
+Then, the function close-current$\left(R,A,J \right)$ closes any current version of the object $A$ if it exists and add the new version $\left(A, J \right)$.
+
+\begin{eqnarray}
+%\label{eq:close-current}
+\text{Close-current} \left(R, A, J \right) =\\
+\begin{cases}
+\nonumber
+\big \lbrace R - \left(A, I \right) \cup \left \lbrace \left(A, \mbox{ CloseR } \left(I, J\right) \right) \cup \left(A, J\right)\right \rbrace  \big \rbrace \\
+\nonumber
+\mbox{ if } I = \mbox{ Current } \left(R, A \right) \wedge J > \mbox{ CloseR } \left(I, J \right)   \\
+\nonumber R , \text{ in any other case}
+\end{cases}
+\end{eqnarray}
+\end{definition}
+
+It is also possible to append at the beggining by using \eqref{def:close-a-crisp-interval-l} and \eqref{def:first-in-relation}.
+\begin{definition}
+\label{def:append-first-version}
+Consider an object $A$, the relation $R$ and a time interval $J$.
+Then, the function append-first$\left(R,A,J \right)$ closes any right open version of the object $A$ if it exists and add the new version $\left(A, J \right)$.
+
+\begin{eqnarray}
+%\label{eq:close-current}
+\text{Append-first} \left(R, A, J \right) =\\
+\begin{cases}
+\nonumber
+\big \lbrace R - \left(A, I \right) \cup \left \lbrace \left(A, \mbox{ CloseL } \left(I, J\right) \right) \cup \left(A, J\right)\right \rbrace  \big \rbrace \\
+\nonumber
+\mbox{ if } I = \mbox{ First } \left(R, A \right) \wedge J < \mbox{ CloseL } \left(I, J \right)   \\
+\nonumber R , \text{ in any other case}
+\end{cases}
+\end{eqnarray}
+\end{definition}
+
 \subsubsection{Insert}
-The interpretation of the insert operation is the following: The user wants to store in the database that an object $A$  is true for some period(s) of time $I$ in the given relation $R$.
+The user wants to store an object $A$ which is valid in the relation $R$ during the time interval specified by $I = \left(a, b \right)$.
+%
+%
+%The interpretation of the insert operation is the following: The user wants to store in the database that an object $A$  is true for some period(s) of time $I$ in the given relation $R$.
 To indicate that the object $A$ is current in the relation, the special value Until Current, $\UC$  is used.
-
-There are three main cases when performing a create operation:
+There are two main cases when performing a create operation:
 \begin{enumerate}
@@
 
-\item The object $A$ was part of some previously recorded state: A new version of the object is inserted in the relation $R$ with the valid-time indicated by $I$.
+\item The object $A$ is in the relation $R$. Depending on the value of the time interval $I$, there are two possibilities:
+	\begin{enumerate}
+	\item Insert $\left(A, I\right)$ in the relation $R$. If the time interval $I$ does not overlap any other valid-time interval in the relation $R$.
+	\item Reject the insertion, if the time interval $I$ do overlap any existing valid-time interval for the object $A$ in the relation $R$.
+	\end{enumerate}
 
-\item The object $A$ is current in the relation $R$: The insert operation is rejected, a modify operation should be used.
 \end{enumerate}
@@
 R \cup \left(A, I \right), & \mbox{ if }  A \not \in R\\
-R \cup  & \mbox{ if }  A \in R \wedge \\
+R \cup   \left(A, I \right) & \mbox{ if }  A \in R \wedge \\
+& \big \lbrace \left( \exists J: \left(A, J \right) \in R \wedge J<I  \right) \vee \\
+& \left( \exists K: \left(A, K \right) \in R \wedge K>I \vee \right) \wedge\\
+&  J<I<K  \big \rbrace\\
 R, & \mbox{ otherwise }  

### 2012-04-16 trunk/ijcis-pos-model/sources/temporalModel.tex
version sent to olga and nicolas.

@@
 
-\subsection{\label{subsec:time-interval-constraint}The time interval}
-The representation of a temporal interval is given by two points: S the starting and E the ending points respectively.
-
-\vglue13pt
-%\begin{table}[htbp]
-\tcap{Data types for a time point}
-\centerline{\small DATA TYPES}
-\vglue-6pt
-\centerline{\small\baselineskip=13pt
-\begin{tabular}{c p{2cm} p{2cm}}\\
-\hline
-1 & A single time point &  $1/x, x \in \T$\\
-2 & A possibility distribution in the numeric domain & A fuzzy number or a fuzzy interval.\\
-3 & An unknown value & \textbf{UNKNOWN}$= \lbrace1/t, t \in \T \rbrace$\\
-4 & An undefined value & \textbf{UNDEFINED}$= \lbrace 0/t, t \in \T \rbrace$\\
-5 & A null value & \textbf{NULL} $=\lbrace 1/$Unknown, 1/Undefined $\rbrace$\\
-\hline\\
-\end{tabular}
-\label{tbl:time-point-types}
-}
-%\end{table}
-
-\textcolor{red}{MAYBE COMPATIBILITY WITH GEFRED TYPES??}
-
-A valid temporal interval $\left[S,E\right]$ in the system is a combination for the values of $S$,$E$ as shown in table \ref{tbl:valid-time-interval}.
-
-\textcolor{red}{HAS IT ANY SENSE TO ALLOW THE COMBINATION [UNKNOWN, UNDEFINED] AND SO ON??}
-
-\vglue13pt
-%\begin{table}[htbp]
-\tcap{Valid-Time interval data types}
-\centerline{\small VALID TIME DATA TYPES}
-\vglue-6pt
-\centerline{\small\baselineskip=13pt
-\begin{tabular}{c p{2cm} p{2cm}}\\
-\hline
-Type for S & Type for E & Semantics \\
-\hline
-Unknown & Unkown & Unknown interval. \\
-Undefined & Undefined & Undefined interval. \\
-Null & Null & Null interval. \\
-\hline
-1,2 & 1,2 & A possibilistic time interval. \\
-\hline\\
-\end{tabular}
-\label{tbl:valid-time-interval}}
-%\end{table}
-
-The following is a study of the constraints in the range of values for both starting and ending times to represent a valid time interval.
-
-The time period inside a valid time interval is denoted by two constraints:
-\begin{eqnarray}
-C_1\stackrel{\triangle}{=}\left(\geq,X\right)\\
-C_2\stackrel{\triangle}{=}\left(\leq,Y\right).
-\end{eqnarray}
-Applying the inference of uncertainty as proposed in our general reasoning, we find for the first constraint that:
-\begin{eqnarray}
-\Pos\left(C_1([a,b])\right) &=& \min_{r\in[a,b]}\Big(\sup_{r\leq w}\pi_{X}(w)\Big)\\
-\Nec\left(C_1([a,b])\right) &=& \min_{r\in[a,b]}\Big(\inf_{r>w}1-\pi_{X}(w)\Big)
-\end{eqnarray}
-which can be simplified to:
-\begin{eqnarray}
-\Pos\left(C_1([a,b])\right) &=& \sup_{a\leq w}\pi_{X}(w)\\
-\Nec\left(C_1([a,b])\right) &=& \inf_{a>w}1-\pi_{X}(w).
-\end{eqnarray}
-For the second constraint, we find that:
-\begin{eqnarray}
-\Pos\left(C_2([a,b])\right) &=& \min_{r\in[a,b]}\Big(\sup_{r\geq w}\pi_{Y}(w)\Big)\\
-\Nec\left(C_2([a,b])\right) &=& \min_{r\in[a,b]}\Big(\inf_{r<w}1-\pi_{Y}(w)\Big)
-\end{eqnarray}
-which can be simplified to:
-\begin{eqnarray}
-\Pos\left(C_2([a,b])\right) &=& \sup_{b\geq w}\pi_{Y}(w)\\
-\Nec\left(C_2([a,b])\right) &=& \inf_{b<w}1-\pi_{Y}(w).
-\end{eqnarray}
-The uncertainty about the inclusion of an interval $I=[a,b]$ in the interval with ill-known boundaries can now be found by evaluating $[a,b]$ against the evaluation function $\lambda$ with $\bool=\wedge$.
+%
+%\vglue13pt
+%%\begin{table}[htbp]
+%\tcap{Valid-Time interval data types}
+%\centerline{\small VALID TIME DATA TYPES}
+%\vglue-6pt
+%\centerline{\small\baselineskip=13pt
+%\begin{tabular}{c p{2cm} p{2cm}}\\
+%\hline
+%Type for S & Type for E & Semantics \\
+%\hline
+%Unknown & Unkown & Unknown interval. \\
+%\hline
+%1,2 & 1,2 & A possibilistic time interval. \\
+%\hline\\
+%\end{tabular}
+%\label{tbl:valid-time-interval}}
+%%\end{table}
+%
+%The following is a study of the constraints in the range of values for both starting and ending times to represent a valid time interval.
+%
+%The time period inside a valid time interval is denoted by two constraints:
+%\begin{eqnarray}
+%C_1\stackrel{\triangle}{=}\left(\geq,X\right)\\
+%C_2\stackrel{\triangle}{=}\left(\leq,Y\right).
+%\end{eqnarray}
+%Applying the inference of uncertainty as proposed in our general reasoning, we find for the first constraint that:
+%\begin{eqnarray}
+%\Pos\left(C_1([a,b])\right) &=& \min_{r\in[a,b]}\Big(\sup_{r\leq w}\pi_{X}(w)\Big)\\
+%\Nec\left(C_1([a,b])\right) &=& \min_{r\in[a,b]}\Big(\inf_{r>w}1-\pi_{X}(w)\Big)
+%\end{eqnarray}
+%which can be simplified to:
+%\begin{eqnarray}
+%\Pos\left(C_1([a,b])\right) &=& \sup_{a\leq w}\pi_{X}(w)\\
+%\Nec\left(C_1([a,b])\right) &=& \inf_{a>w}1-\pi_{X}(w).
+%\end{eqnarray}
+%For the second constraint, we find that:
+%\begin{eqnarray}
+%\Pos\left(C_2([a,b])\right) &=& \min_{r\in[a,b]}\Big(\sup_{r\geq w}\pi_{Y}(w)\Big)\\
+%\Nec\left(C_2([a,b])\right) &=& \min_{r\in[a,b]}\Big(\inf_{r<w}1-\pi_{Y}(w)\Big)
+%\end{eqnarray}
+%which can be simplified to:
+%\begin{eqnarray}
+%\Pos\left(C_2([a,b])\right) &=& \sup_{b\geq w}\pi_{Y}(w)\\
+%\Nec\left(C_2([a,b])\right) &=& \inf_{b<w}1-\pi_{Y}(w).
+%\end{eqnarray}
+%The uncertainty about the inclusion of an interval $I=[a,b]$ in the interval with ill-known boundaries can now be found by evaluating $[a,b]$ against the evaluation function $\lambda$ with $\bool=\wedge$.
 Application of \eqref{eq:conjunctive1} and \eqref{eq:conjunctive2} leads to:
-\begin{eqnarray}
-\Pos\left(\lambda([a,b])\right)&=&\min\bigg(\Pos(C_1([a,b])),\Pos(C_2([a,b]))\bigg)\\
-\Nec\left(\lambda([a,b])\right)&=&\min\bigg(\Nec(C_1([a,b])),\Nec(C_2([a,b]))\bigg).
-\end{eqnarray}
-These last expressions can also be expanded as:
-\begin{eqnarray}
-\label{eq:interval-pos}
-\Pos\left(\lambda([a,b])\right)&=&\min\bigg(\sup_{a\leq w}\pi_{X}(w),\sup_{b\geq w}\pi_{Y}(w)\bigg)\\
-\label{eq:interval-nec}
-\Nec\left(\lambda([a,b])\right)&=&\min\bigg(\inf_{a>w}1-\pi_{X}(w),\inf_{b<w}1-\pi_{Y}(w)\bigg).
-\end{eqnarray}
-Note that the interval $[X,Y]$ used here, is certainly not a fuzzy interval.
+%\begin{eqnarray}
+%\Pos\left(\lambda([a,b])\right)&=&\min\bigg(\Pos(C_1([a,b])),\Pos(C_2([a,b]))\bigg)\\
+%\Nec\left(\lambda([a,b])\right)&=&\min\bigg(\Nec(C_1([a,b])),\Nec(C_2([a,b]))\bigg).
+%\end{eqnarray}
+%These last expressions can also be expanded as:
+%\begin{eqnarray}
+%\label{eq:interval-pos}
+%\Pos\left(\lambda([a,b])\right)&=&\min\bigg(\sup_{a\leq w}\pi_{X}(w),\sup_{b\geq w}\pi_{Y}(w)\bigg)\\
+%\label{eq:interval-nec}
+%\Nec\left(\lambda([a,b])\right)&=&\min\bigg(\inf_{a>w}1-\pi_{X}(w),\inf_{b<w}1-\pi_{Y}(w)\bigg).
+%\end{eqnarray}
+%Note that the interval $[X,Y]$ used here, is certainly not a fuzzy interval.
 Instead, we are dealing with an ill-known interval, i.e.
@@
 The uncertainty stems from the fact that the interval boundaries are ill-known.
-
-Within this framework for the reasoning of the knowledge about the interval, we are going to study the semantics from the point of view of the valid time as well as from the point of view of the temporal database.
-
-\subsubsection{S = Unknown, E = Unknown}
-Consider that the possibility distribution for S is Unknown and for E is also Unknown. The possibility distributions are $\pi_S$ and $\pi_E$ respectively:
-
-\begin{eqnarray}
-\pi_{S}(t \in \T) = 1\\
-\pi_{E}(t \in \T) = 1
-\end{eqnarray}
-
-Therefore the constraints $C_1$ and $C_2$ are:
-
-\begin{eqnarray}
-\Pos\left(C_1([a,b])\right) &=& \sup_{a\leq w}\pi_{S}(w) = 1\\
-\Nec\left(C_1([a,b])\right) &=& \inf_{a>w}1-\pi_{S}(w) = 0\\
-\Pos\left(C_2([a,b])\right) &=& \sup_{b\geq w}\pi_{E}(w)=1\\
-\Nec\left(C_2([a,b])\right) &=& \inf_{b<w}1-\pi_{E}(w)=0
-\end{eqnarray}
-
-\begin{eqnarray}
-\Pos\left(\lambda([a,b])\right)&=&\min\bigg(\sup_{a\leq w}\pi_{S}(w),\sup_{b\geq w}\pi_{E}(w)\bigg) \\
-&=& \min(1,1)  = 1\\
-\Nec\left(\lambda([a,b])\right)&=&\min\bigg(\inf_{a>w}1-\pi_{S}(w),\inf_{b<w}1-\pi_{E}(w)\bigg)\\
-&=& \min(0,0) = 0
-\end{eqnarray}
-
-\begin{example}
-A constant object in the database. It is an object that is always valid in the database.
-The main issue in a valid time database is that the update sentence have to modify the starting or ending point distribution.
+%
+%Within this framework for the reasoning of the knowledge about the interval, we are going to study the semantics from the point of view of the valid time as well as from the point of view of the temporal database.
+%
+%\subsubsection{S = Unknown, E = Unknown}
+%Consider that the possibility distribution for S is Unknown and for E is also Unknown. The possibility distributions are $\pi_S$ and $\pi_E$ respectively:
+%
+%\begin{eqnarray}
+%\pi_{S}(t \in \T) = 1\\
+%\pi_{E}(t \in \T) = 1
+%\end{eqnarray}
+%
+%Therefore the constraints $C_1$ and $C_2$ are:
+%
+%\begin{eqnarray}
+%\Pos\left(C_1([a,b])\right) &=& \sup_{a\leq w}\pi_{S}(w) = 1\\
+%\Nec\left(C_1([a,b])\right) &=& \inf_{a>w}1-\pi_{S}(w) = 0\\
+%\Pos\left(C_2([a,b])\right) &=& \sup_{b\geq w}\pi_{E}(w)=1\\
+%\Nec\left(C_2([a,b])\right) &=& \inf_{b<w}1-\pi_{E}(w)=0
+%\end{eqnarray}
+%
+%
+%\begin{eqnarray}
+%\Pos\left(\lambda([a,b])\right)&=&\min\bigg(\sup_{a\leq w}\pi_{S}(w),\sup_{b\geq w}\pi_{E}(w)\bigg) \\
+%&=& \min(1,1)  = 1\\
+%\Nec\left(\lambda([a,b])\right)&=&\min\bigg(\inf_{a>w}1-\pi_{S}(w),\inf_{b<w}1-\pi_{E}(w)\bigg)\\
+%&=& \min(0,0) = 0
+%\end{eqnarray}
+%
+%\begin{example}
+%A constant object in the database. It is an object that is always valid in the database.
+%The main issue in a valid time database is that the update sentence have to modify the starting or ending point distribution.
 The most natural way to convert a snapshot database to a temporal database is to consider the objects belonging to the interval [Unknown, Unknown].
-\end{example}
-
-\subsubsection{[Unknown,value] or [value,Unknown]}
-These two cases are managed in this section. Here we will note $\pi_{S}(t)$ or $\pi_{E}(t)$ as the possibility distribution for 'value' in the starting or ending points.
-
-First of all the case [Unknown,value] in \eqref{eq:interval-pos} and in \eqref{eq:interval-nec}:
-
-\begin{eqnarray}
-\Pos\left(\lambda([a,b])\right)&=&\min (1,\sup_{b\geq w}\pi_{E}(w))\\
- &=& \sup_{b\geq w}\pi_{E}(w)\\
-\Nec\left(\lambda([a,b])\right)&=&\min (0,\inf_{b<w}1-\pi_{E}(w))\\
-&=& 0
-\end{eqnarray}
-
-Similarily for [value,Unknown] the possibility and neccesity are:
-\begin{eqnarray}
-\Pos\left(\lambda([a,b])\right)&=&\min (\sup_{a\leq w}\pi_{S}(w),1)\\
- &=& \sup_{a\leq w}\pi_{S}(w)\\
-\Nec\left(\lambda([a,b])\right)&=&\min (\inf_{a>w}1-\pi_{S}(w),0)\\
-&=& 0
-\end{eqnarray}
-
-The semantics for an object with valid time given by [Unknown,value] is an object that has been valid from the beggining to an ending time.
+%\end{example}
+%
+%\subsubsection{[Unknown,value] or [value,Unknown]}
+%These two cases are managed in this section. Here we will note $\pi_{S}(t)$ or $\pi_{E}(t)$ as the possibility distribution for 'value' in the starting or ending points.
+%
+%First of all the case [Unknown,value] in \eqref{eq:interval-pos} and in \eqref{eq:interval-nec}:
+%
+%\begin{eqnarray}
+%\Pos\left(\lambda([a,b])\right)&=&\min (1,\sup_{b\geq w}\pi_{E}(w))\\
+% &=& \sup_{b\geq w}\pi_{E}(w)\\
+%\Nec\left(\lambda([a,b])\right)&=&\min (0,\inf_{b<w}1-\pi_{E}(w))\\
+%&=& 0
+%\end{eqnarray}
+%
+%Similarily for [value,Unknown] the possibility and neccesity are:
+%\begin{eqnarray}
+%\Pos\left(\lambda([a,b])\right)&=&\min (\sup_{a\leq w}\pi_{S}(w),1)\\
+% &=& \sup_{a\leq w}\pi_{S}(w)\\
+%\Nec\left(\lambda([a,b])\right)&=&\min (\inf_{a>w}1-\pi_{S}(w),0)\\
+%&=& 0
+%\end{eqnarray}
+%
+%The semantics for an object with valid time given by [Unknown,value] is an object that has been valid from the beggining to an ending time.
 Analogously, the semantics for an object with a valid time given by the interval [value,Unknown] is an object that it is still valid in the database.
-
-\subsubsection{[Undefined,Undefined]}
-Maybe this combination has no sense for a valid time object. The possibility distributions for both starting and ending points are:
-
-\begin{eqnarray}
-\pi_{S}(t \in \T) = 0\\
-\pi_{E}(t \in \T) = 0
-\end{eqnarray}
-
-Therefore the constraints $C_1$ and $C_2$ are:
-
-\begin{eqnarray}
-\Pos\left(C_1([a,b])\right) &=& \sup_{a\leq w}\pi_{S}(w) = 0\\
-\Nec\left(C_1([a,b])\right) &=& \inf_{a>w}1-\pi_{S}(w) = 1\\
-\Pos\left(C_2([a,b])\right) &=& \sup_{b\geq w}\pi_{E}(w)=0\\
-\Nec\left(C_2([a,b])\right) &=& \inf_{b<w}1-\pi_{E}(w)=1
-\end{eqnarray}
-
-Taking these values into  \eqref{eq:interval-pos} and in \eqref{eq:interval-nec}:
-\begin{eqnarray}
-\Pos\left(\lambda([a,b])\right)&=&\min (0,0)\\
- &=& 0\\
-\Nec\left(\lambda([a,b])\right)&=&\min (1,1)\\
-&=& 1
-\end{eqnarray}
-
-\textcolor{red}{TAKE A LOOK INTO THIS}
-
-Which is an inconsistency.
-
-\subsection{[Undefined,value] or [value,Undefined]}
-
-The possibility and neccesity measures for [Undefined,value] are :
-\begin{eqnarray}
-\Pos\left(\lambda([a,b])\right)&=&\min (0,\sup_{b\geq w}\pi_{E}(w))\\
- &=& 0\\
-\Nec\left(\lambda([a,b])\right)&=&\min (1,\inf_{b<w}1-\pi_{E}(w))\\
-&=& \inf_{b<w}1-\pi_{E}(w)
-\end{eqnarray}
-
-The other way around, [value,Undefined]:
-\begin{eqnarray}
-\Pos\left(\lambda([a,b])\right)&=&\min (\sup_{a\leq w}\pi_{S}(w),0)\\
- &=& 0\\
-\Nec\left(\lambda([a,b])\right)&=&\min (\inf_{a>w}1-\pi_{S}(w),1)\\
-&=& \inf_{a>w}1-\pi_{S}(w)
-\end{eqnarray}
-
-\textcolor{red}{TAKE A LOOK INTO THIS}
-This is also an inconsistency.
-
-\subsubsection{[NULL,NULL] and maybe any combination with NULL}
-I think the framework we have can not deal with NULL computations. Maybe just for non-specified time or something like that.
-
-\subsubsection{[Undefined,Unknown] or [Unknown,Undefined]}
-The semantic for this could be also unconsistent.
-
-\begin{example}
-Consider a snapshot database (table \ref{tbl:snapshot-car-models}) that stores information about car models. The primary key is given by the set $\lbrace$ Brand, Model $\rbrace$.
-
-\vglue13pt
-%\begin{table}[htbp]
-\tcap{Car model snapshot database}
-\centerline{\small CAR MODELS}
-\vglue-6pt
-\centerline{\small\baselineskip=13pt
-\begin{tabular}{c c c}\\
-\hline
-\textbf{Brand} & \textbf{Model} & Segment \\
-\hline
-Peugeot & 205 & B \\
-Peugeot & 206 & B \\
-Peugeot & 207 & B \\
-Peugeto & 208 & B \\
-Citroen & C2 & B \\
-\hline\\
-\end{tabular}
-\label{tbl:snapshot-car-models}
-}
-
-Consider now that the user wants to make that table a valid-time table, then, from the point of view of the storage, the following changes should be done:
-
-\begin{itemize}
-\item
-A new definition for the primary key, including a version field.
-\item
-The addition of two new columns to store the valid-time interval. By default the initial values are Unknown for both starting and ending points of the interval.
-\end{itemize}
-Therefore, the table \ref{tbl:snapshot-car-models} is transformed into table \ref{tbl:initial-valid-time-car-models}:
-
- \vglue13pt
-%\begin{table}[htbp]
-\tcap{Car model valid-time database. Note that U is the abbreviation for the Unknown constant.}
-\centerline{\small VALID TIME CAR MODELS}
-\vglue-6pt
-\centerline{\small\baselineskip=13pt
-\begin{tabular}{c c c c c c}\\
-\hline
-\textbf{Brand} & \textbf{Model} & \textbf{VersionID} & Segment & S & E \\
-\hline
-Peugeot & 205 & 001 & B & U & U \\
-Peugeot & 206 & 001 & B & U & U \\
-Peugeot & 207 & 001 & B & U & U \\
-Peugeto & 208 & 001 & B & U & U \\
-Citroen & C2 & 001 & B & U & U \\
-\hline\\
-\end{tabular}
-\label{tbl:initial-valid-time-car-models}
-}
-
-\end{example}
+%
+%
+%\subsubsection{[Undefined,Undefined]}
+%Maybe this combination has no sense for a valid time object. The possibility distributions for both starting and ending points are:
+%
+%\begin{eqnarray}
+%\pi_{S}(t \in \T) = 0\\
+%\pi_{E}(t \in \T) = 0
+%\end{eqnarray}
+%
+%Therefore the constraints $C_1$ and $C_2$ are:
+%
+%\begin{eqnarray}
+%\Pos\left(C_1([a,b])\right) &=& \sup_{a\leq w}\pi_{S}(w) = 0\\
+%\Nec\left(C_1([a,b])\right) &=& \inf_{a>w}1-\pi_{S}(w) = 1\\
+%\Pos\left(C_2([a,b])\right) &=& \sup_{b\geq w}\pi_{E}(w)=0\\
+%\Nec\left(C_2([a,b])\right) &=& \inf_{b<w}1-\pi_{E}(w)=1
+%\end{eqnarray}
+%
+%Taking these values into  \eqref{eq:interval-pos} and in \eqref{eq:interval-nec}:
+%\begin{eqnarray}
+%\Pos\left(\lambda([a,b])\right)&=&\min (0,0)\\
+% &=& 0\\
+%\Nec\left(\lambda([a,b])\right)&=&\min (1,1)\\
+%&=& 1
+%\end{eqnarray}
+%
+%\textcolor{red}{TAKE A LOOK INTO THIS}
+%
+%Which is an inconsistency.
+%
+%\subsection{[Undefined,value] or [value,Undefined]}
+%
+%The possibility and neccesity measures for [Undefined,value] are :
+%\begin{eqnarray}
+%\Pos\left(\lambda([a,b])\right)&=&\min (0,\sup_{b\geq w}\pi_{E}(w))\\
+% &=& 0\\
+%\Nec\left(\lambda([a,b])\right)&=&\min (1,\inf_{b<w}1-\pi_{E}(w))\\
+%&=& \inf_{b<w}1-\pi_{E}(w)
+%\end{eqnarray}
+%
+%The other way around, [value,Undefined]:
+%\begin{eqnarray}
+%\Pos\left(\lambda([a,b])\right)&=&\min (\sup_{a\leq w}\pi_{S}(w),0)\\
+% &=& 0\\
+%\Nec\left(\lambda([a,b])\right)&=&\min (\inf_{a>w}1-\pi_{S}(w),1)\\
+%&=& \inf_{a>w}1-\pi_{S}(w)
+%\end{eqnarray}
+%
+%\textcolor{red}{TAKE A LOOK INTO THIS}
+%This is also an inconsistency.
+%
+%\subsubsection{[NULL,NULL] and maybe any combination with NULL}
+%I think the framework we have can not deal with NULL computations. Maybe just for non-specified time or something like that.
+%
+%\subsubsection{[Undefined,Unknown] or [Unknown,Undefined]}
+%The semantic for this could be also unconsistent.
+%
+%
+%\begin{example}
+%Consider a snapshot database (table \ref{tbl:snapshot-car-models}) that stores information about car models. The primary key is given by the set $\lbrace$ Brand, Model $\rbrace$.
+%
+%\vglue13pt
+%%\begin{table}[htbp]
+%\tcap{Car model snapshot database}
+%\centerline{\small CAR MODELS}
+%\vglue-6pt
+%\centerline{\small\baselineskip=13pt
+%\begin{tabular}{c c c}\\
+%\hline
+%\textbf{Brand} & \textbf{Model} & Segment \\
+%\hline
+%Peugeot & 205 & B \\
+%Peugeot & 206 & B \\
+%Peugeot & 207 & B \\
+%Peugeto & 208 & B \\
+%Citroen & C2 & B \\
+%\hline\\
+%\end{tabular}
+%\label{tbl:snapshot-car-models}
+%}
+%
+%Consider now that the user wants to make that table a valid-time table, then, from the point of view of the storage, the following changes should be done:
+%
+%\begin{itemize}
+%\item
+%A new definition for the primary key, including a version field.
+%\item
+%The addition of two new columns to store the valid-time interval. By default the initial values are Unknown for both starting and ending points of the interval.
+%\end{itemize}
+%Therefore, the table \ref{tbl:snapshot-car-models} is transformed into table \ref{tbl:initial-valid-time-car-models}:
+%
+% \vglue13pt
+%%\begin{table}[htbp]
+%\tcap{Car model valid-time database. Note that U is the abbreviation for the Unknown constant.}
+%\centerline{\small VALID TIME CAR MODELS}
+%\vglue-6pt
+%\centerline{\small\baselineskip=13pt
+%\begin{tabular}{c c c c c c}\\
+%\hline
+%\textbf{Brand} & \textbf{Model} & \textbf{VersionID} & Segment & S & E \\
+%\hline
+%Peugeot & 205 & 001 & B & U & U \\
+%Peugeot & 206 & 001 & B & U & U \\
+%Peugeot & 207 & 001 & B & U & U \\
+%Peugeto & 208 & 001 & B & U & U \\
+%Citroen & C2 & 001 & B & U & U \\
+%\hline\\
+%\end{tabular}
+%\label{tbl:initial-valid-time-car-models}
+%}
+%
+%
+%
+%\end{example}
 
 \subsection{\label{subsec:temporal-model}The generalized temporal model}
+The model is based on the GEFRED~\cite{Medina1994} (Generalized Model of Fuzzy Relational DB) model.
+This model is extended by adding valid-time support.
+The information in the system is defined by the following elements:
+
+\begin{definiton}
+\label{def:generalized-fuzzy-domain}
+$D$ is the discourse domain, $\tilde \Pow\left(D \right)$ is the set of all possibility distributions defined on $\D$, plus the NULL constant. The generalized fuzzy domain $D_G$ is defined as:
+\begin{equation}
+\label{eq:generalized-fuzzy-domain}
+D_G \subseteq \tilde \Pow\left(D \right)\cup \text{NULL}
+\end{equation}
+\end{definition}
+
+It is possible to define a more specific generalized temporal domain, $\T_G$:
 
 \begin{definition}
+\label{def:generalized-fuzzy-temporal-domain}
 Generalized fuzzy temporal domain.\\
-If $\T$ is the temporal domain, $\tilde \Pow\left( \T\right)$ is the set of all possibility distributions defined on $\T$.
+If $\T$ is the temporal domain, $\tilde \Pow\left( \T\right)$ is the set of all \emph{normalized} possibility distributions (see Section \ref{subsec:possibility-theory}, equation \eqref{NormalizationProperty}) defined on $\T$.
 The \textbf{Generalized Fuzzy Temporal Domain}, $\T_G$ is
 \begin{equation}
-\T_G \subseteq \left \lbrace \tilde \Pow\left( \T\right) \cup NULL \right \rbrace
+\T_G \subseteq \left \lbrace \tilde \Pow\left( \T\right) \cup \text{NULL} \right \rbrace
 \end{equation}
 \end{definition}
+
+Note that $\T_G \subseteq D_G$.

### 2012-04-16 trunk/ijcis-pos-model/sources/temporaldatabases.tex
version sent to olga and nicolas.

@@
 \begin{definition}
-\label{def:close-a-crisp-interval}
+\label{def:close-a-crisp-interval-r}
 Consider two crisp intervals $I= \left[a ,b \right]$ and $J= \left[c ,d \right]$, $c > a$.
-The CloseL$\left(I, J\right)$ function allows to close the left-open interval $I$ with respect to the first value $d$ in $J$:
+The CloseR$\left(I, J\right)$ function allows to close the right-open interval $I$ with respect to the first value $d$ in $J$:
 \begin{align}
-\mbox{CloseL} \left( I, J \right) &=& \\ 
+\mbox{CloseR} \left( I, J \right) &=& \\ 
 \begin{cases}
@@
 
-Analogously, it is defined the function to close an right open interval: 
-\begin{definition}
-\label{def:close-a-crisp-interval-r}
-CloseR$\left(I, J\right)$ 
-\begin{align}
-\mbox{CloseR} \left( I, J \right) &=& \\ 
-\begin{cases}
-\nonumber
-I & \mbox{ if } a \neq -\infty \\
-I=\left]d, b \right[ & \mbox{ if } a = -\infty \wedge J < I
-\end{cases}
-\end{align}
-\end{definition}
+%Analogously, it is defined the function to close a left-open interval: 
+%\begin{definition}
+%\label{def:close-a-crisp-interval-l}
+%CloseL$\left(I, J\right)$ 
+%\begin{align}
+%\mbox{CloseL} \left( I, J \right) &=& \\ 
+%\begin{cases}
+%\nonumber
+%I & \mbox{ if } a \neq -\infty \\
+%I=\left]d, b \right] & \mbox{ if } a = -\infty \wedge J < I
+%\end{cases}
+%\end{align}
+%\end{definition}
 
@@
 Consider an object $A$ and a relation $R$ in a temporal database.
-The fucntion Current$\left(R, A \right)$ obtains the tuple $\left(A, I \right)$ that is current in the relation $R$, that is it, $I = \left[a, +\infty \right]$.
+The function Current$\left(R, A \right)$ obtains the tuple $\left(A, I \right)$ that is current in the relation $R$, that is it, $I = \left[a, +\infty \right]$.
 
@@
 
-Now it is possible to close the current version of an object by combining the two previously defined functions.
+%\begin{definition}
+%\label{def:first-in-relation}
+%Analogously, the function First$\left(R, A \right)$ obtains the tuple $\left(A, I \right)$ where $I$ is a right-open interval and which is the first version in the relation $R$, that is it, $I = \left[-\infty, b \right]$.
+%\begin{align}
+%\label{eq:first-in-relation}
+%\mbox{First} \left(R, A \right) &=& \\ 
+%\begin{cases}
+%\nonumber
+%I & \mbox{ if } \exists b \in I : (A,I) \in R \wedge a = -\infty \\
+%\emptyset & \mbox{ in any other case }
+%\end{cases}
+%\end{align}
+%
+%\end{definition}
+
+Now it is possible to close the current version of an object by using \eqref{def:close-a-crisp-interval-r} and \eqref{eq:current-in-relation}
 
@@
 %\label{eq:close-current}
-\text{Close-current} \left(R, A \right) =\\
+\text{Close-current} \left(R, A, J \right) =\\
 \begin{cases}
 \nonumber
-\big \lbrace R - \left(A, I \right) \cup \left \lbrace \left(A, \mbox{ Close } \left(I, J\right) \right) \cup \left(A, J\right)\right \rbrace  \big \rbrace \\
+\big \lbrace R - \left(A, I \right) \cup \left \lbrace \left(A, \mbox{ CloseR } \left(I, J\right) \right) \cup \left(A, J\right)\right \rbrace  \big \rbrace \\
 \nonumber
-\mbox{ if } I = \mbox{ Current } \left(R, A \right) \wedge J > \mbox{ Close } \left(I, J \right)   \\
+\mbox{ if } I = \mbox{ Current } \left(R, A \right) \wedge J > \mbox{ CloseR } \left(I, J \right)   \\
 \nonumber R , \text{ in any other case}
@@
 
-\subsubsection{Insert}
+%It is also possible to append at the beggining by using \eqref{def:close-a-crisp-interval-l} and \eqref{def:first-in-relation}.
+%\begin{definition}
+%\label{def:append-first-version}
+%Consider an object $A$, the relation $R$ and a time interval $J$.
+Then, the function append-first$\left(R,A,J \right)$ closes any right open version of the object $A$ if it exists and add the new version $\left(A, J \right)$.
+%
+%\begin{eqnarray}
+%%\label{eq:close-current}
+%\text{Append-first} \left(R, A, J \right) =\\
+%\begin{cases}
+%\nonumber
+%\big \lbrace R - \left(A, I \right) \cup \left \lbrace \left(A, \mbox{ CloseL } \left(I, J\right) \right) \cup \left(A, J\right)\right \rbrace  \big \rbrace \\
+%\nonumber
+%\mbox{ if } I = \mbox{ First } \left(R, A \right) \wedge J < \mbox{ CloseL } \left(I, J \right)   \\
+%\nonumber R , \text{ in any other case}
+%\end{cases}
+%\end{eqnarray}
+%\end{definition}
+
+\subsubsection{\label{subsubsec:insert}Insert}
 The user wants to store an object $A$ which is valid in the relation $R$ during the time interval specified by $I = \left(a, b \right)$.
@@
 To indicate that the object $A$ is current in the relation, the special value Until Current, $\UC$  is used.
-There are three main cases when performing a create operation:
+There are two main cases when performing a create operation:
 \begin{enumerate}
@@
 
-\item The object $A$ is in the relation $R$, but it is not current
+\item The object $A$ is in the relation $R$. Depending on the value of the time interval $I$, there are two possibilities:
+	\begin{enumerate}
+	\item Insert $\left(A, I\right)$ in the relation $R$. If the time interval $I$ does not overlap any other valid-time interval in the relation $R$.
+	\item Reject the insertion, if the time interval $I$ do overlap any existing valid-time interval for the object $A$ in the relation $R$.
+	\end{enumerate}
 
-\item The object $A$ is current in the relation $R$: The insert operation is rejected, a modify operation should be used.
 \end{enumerate}
@@
 R \cup   \left(A, I \right) & \mbox{ if }  A \in R \wedge \\
+& \big \lbrace \left( \exists J: \left(A, J \right) \in R \wedge J<I  \right) \vee \\
+& \left( \exists K: \left(A, K \right) \in R \wedge K>I \vee \right) \wedge\\
+&  J<I<K  \big \rbrace\\
 R, & \mbox{ otherwise }  
@@
 
-\subsubsection{Delete}
+\subsubsection{\label{subsubsec:del}Delete}
 The delete operation logically removes a current object $A$ which is valid in the relation $R$:
@@
 \nonumber
-R - \left(A, I \right), & \mbox{ if }  A \not \in R\\
+R - \left(A, I \right), & \mbox{ if } \left( A, I \right) \in R\\
 R, & \mbox{ otherwise }  
@@
 
-\subsubsection{Modify}
-The modification of an object $A$ is usually defined by a delete operation followed by an insert operation. The modify operation is defined then by:
+\subsubsection{\label{subsubsec:modify}Modify}
+This operation adds new information about an existing object $A$ in the relation $R$. The modify operation does not remove any previous value of the object $A$.
+Note that the modify operation is only applicable when the object $A$ is current in the relation $R$; $ \left(A, I \right) \in R, I = \left[a, +\infty \right]$.
 
@@
 \label{eq:modify}
-\MOD \left( R, A, I \right) &=&\\
+\MOD \left( R, A, I \right) =\\
+\begin{cases}
 \nonumber
-\INS \left( \DEL \left(R, A \right), A, I \right)
+\mbox{ Close-current }\left(R, A, \mbox{Current}\left(R, A \right) \right) \\
+ \mbox{ if Current}\left(R ,A \right) \neq \emptyset \\
+ R,  \mbox{ otherwise }
+\end{cases}
 \end{align}
+
+\subsubsection{\label{subsubsec:revise}Revise}
+This operation changes the value of the object $A$ in the relation $R$, at a given time interval $I$. This operation is defined when a correction on the values for the object $A$ should be done.
+\begin{align}
+\label{eq:revise}
+\mbox{ Revise}\left(R, A, I \right)=\\
+\begin{cases}
+\nonumber
+ R - \left(A', I \right) \cup \left(A, I \right)  \\
+  \mbox{ if } \left(A', I \right) \in R: A = \left(a_1, \ldots, a_n \right),\\
+A'= \left(a_1', \ldots, a_n' \right)   \\   
+R,  \mbox{ otherwise }
+\end{cases}
+\end{align}

### 2012-04-16 ijcis-pos-model/sources/preliminaries.tex
version sent to olga and nicolas.

--- 2012-04-12 trunk/ijcis-pos-model/sources/preliminaries.tex
@@
 
-\subsection{Interval Evaluation by Ill-known Constraints}
+\subsection{\label{subsec:interval-evaluation-by-ill-known-constraints}Interval Evaluation by Ill-known Constraints}
 The problem of interval evaluation is more generally explained in \cite{Pons2011}: the need exists to know if all points in a crisp interval $I$ reside between the boundaries of an ill-known interval $\left[ X , Y \right]$.
@@
 \begin{align}
+\label{eq:ill-known-constraint}
 C \triangleq (R,V)
@@
 \end{align} 
-Note that $\pi_{X}$ is convex since $\pi_{CC}$ is convex as demostrated in theorem \ref{th:convex-combination-ill-known-constraints}.
+Note that $\pi_{X}$ is convex since $\pi_{CC}$ is convex as demonstrated in theorem \ref{th:convex-combination-ill-known-constraints}.
 \end{definition}

### 2012-04-16 ijcis-pos-model/sources/timeRepresentation.tex
version sent to olga and nicolas.

@@
 Consider a historical database with data from diplomatic documents.
-The following fields are stored: the digital identifier \emph{ID} which is the primary key  and the stimated time when the document was sent (field \emph{Date}).
+The following fields are stored: the digital identifier \emph{ID} which is the primary key  and the estimated time when the document was sent (field \emph{Date}).
 Table \ref{tbl:sample-time-point} contains some example data from this database.
@@
 The time for documents 12211 and 23455 are specified with several possibility distributions.
-The first one is also known as a fuzzy number whereas the second one is also known as a fuzzy interval.
-\textcolor{red}{References to the preliminaries}.
+The first one is also known as a fuzzy number whereas the second one is also known as a fuzzy interval, as explained in Section \ref{sec:prelim}.
 
@@
 \emph{Ill-known time interval}\\
-An ill-known time interval denoted by $\left[X, Y\right]$ is a precise time interval that for some reason, its boudaries are not precissely known.
+An ill-known time interval denoted by $\left[X, Y\right]$ is a precise time interval that for some reason, its boundaries are not precisely known.
 The ill-known interval is defined by two ill-known points, namely $X$,$Y$.
-In order to know if all the points in the crisp time interval $I=\left[a, b\right]$ are within the boudaries of the ill-known time interval $\left[X, Y\right]$ we define the following two ill-known constraints $C_{1}, C_{2}$
+In order to know if all the points in the crisp time interval $I=\left[a, b\right]$ are within the boundaries of the ill-known time interval $\left[X, Y\right]$ we define the following two ill-known constraints $C_{1}, C_{2}$
 \begin{eqnarray}
@@
 \subsubsection{\label{subsubsec:open-interval}Open ill-known time intervals}
-Quite often, the user may want to specify time intervals with open boudaries in one or both endpoints.
+Quite often, the user may want to specify time intervals with open boundaries in one or both endpoints.
 Consider an ill-known time interval $\left[X, Y\right]$.
@@
 As mentioned before, the problem resides in the representation of these kind of interval.
-In order to get a short representation for these intervals, two constants are defined: $FB$ \emph{From the Beggining}: it is only valid for the left value of the interval and $UC$ \emph{Until Changed} (this name is usual in the temporal database community, see ~\cite{Jensen1994}) which is only valid for the right value of the interval.
+In order to get a short representation for these intervals, two constants are defined: $FB$ \emph{From the Beginning}: it is only valid for the left value of the interval and $UC$ \emph{Until Changed} (this name is usual in the temporal database community, see ~\cite{Jensen1994}) which is only valid for the right value of the interval.
 Note that these two constants are aliases for a function that obtains both possibility and necessity measures.
@@
 The function $\Open(C)$ provides both possibility and necessity measures for all the points in the open part of a semi-open ill-known time interval.
-The parameter $C$ is an ill-known constraint $C\stackrel{\triangle}{=}\left(R,T\right)$ as defined in equation \eqref{rand} and in~\cite{Pons2011}.
+The parameter $C$ is an ill-known constraint $C\stackrel{\triangle}{=}\left(R,T\right)$ as defined in \ref{subsec:interval-evaluation-by-ill-known-constraints}, in the equation \eqref{eq:ill-known-constraint} and in~\cite{Pons2011}.
 The possibility and necessity measures are defined by:

### 2012-04-17 trunk/ijcis-pos-model/sources/temporalModel.tex
updated version

@@
 
-\begin{definiton}
+\begin{definition}
 \label{def:generalized-fuzzy-domain}
-$D$ is the discourse domain, $\tilde \Pow\left(D \right)$ is the set of all possibility distributions defined on $\D$, plus the NULL constant. The generalized fuzzy domain $D_G$ is defined as:
+$D$ is the discourse domain, $\tilde \Pow\left(D \right)$ is the set of all possibility distributions defined on $D$, plus the NULL constant. The generalized fuzzy domain $D_G$ is defined as:
 \begin{equation}
@@
 \end{definition}
+The datatypes for $D_G$ are shown in table \ref{tablaaaaaaaaa}
+\textcolor{red}{poner tabla}
 
@@
 Note that $\T_G \subseteq D_G$.
+The datatypes for this domain have been studied previously in section \ref{sec:time-rep} and are shown in tables \ref{tbl:time-point-types},\ref{tbl:time-interval-types}.
+
+\begin{definition}
+A generalized fuzzy temporal relation $R_{FTG}$ is given by:
+\label{def:generalized-fuzzy-temporal-relation}
+\begin{equation}
+\label{eq:generalized-fuzzy-temporal-relation}
+R_{FTG} = \left(\Head, \Body \right)
+\end{equation}
+Where $\Head$ is the Head of the relation and consist on a fixed set of triplets attribute- domain - compatibility with an optional the valid-time attribute:
+
+\begin{align}
+\label{eq:head-valid-time}
+\Head = \big \lbrace \left(A_{G1}:D_{G1}\left[,C_{A_{G1}} \right] \right),\\
+\nonumber
+ \ldots,\\
+ \nonumber
+  \left(A_{Gn}:D_{Gn}\left[,C_{A_{Gn}} \right] \right),\\
+  \nonumber
+  \Big[  \left( \text{PVP}, D_{\text{PVP}}\left[,C_{A_{\text{PVP}}} \right] \right) \Big] \big \rbrace
+\end{align}
+Note that $D_{Gj}$ ($j = 1, \ldots, n$) is the domain for the attribute $A_{Gj}$. $C_{A_{Gj}}$ is the compatibility attribute in the unit interval $\left[0, 1 \right]$.
+
+$\Body$ is the body of the relation and it consists on a set of tuples. Each tuple is a triplet attribute- value- degree with an optional valid-time attribute:
+
+\begin{align}
+\label{eq:body-valid-time}
+\Body = \big \lbrace \left(A_{G1}:\tilde{d}_{i1}\left[,c_{i1} \right] \right),\\
+\nonumber
+ \ldots,\\
+ \nonumber
+  \left(A_{Gn}:\tilde{d}_{in}\left[,c_{in} \right] \right),\\
+  \nonumber
+   \Big[  \left( \text{PVP}, \tilde{d}_{\text{PVP}} \left[,C_{A_{\text{PVP}}} \right] \right)  \Big] \big \rbrace
+\end{align}
+
+\end{definition}
+
+The definition in \cite{Medina1994} for $R_{FTG}$ shows that classical relations are a particular case of this model.
+
+\begin{definition}
+\label{def:value-component}
+The value component $R^{v}_{FTG}$ of a fuzzy temporal relation $R_{FTG}$ is a set with the value components for both the head and the body of the relation:
+\begin{align}
+\label{eq:value-component}
+R^{v}_{FTG} = \left \lbrace \Head^{v},\Body^{v} \right \rbrace \\
+\nonumber
+\text{Where: } \\
+\nonumber
+\Head^{v} = \left \lbrace A_{G1}:D_{G1}, \ldots,  A_{Gn}:D_{Gn} \right \rbrace \\
+\nonumber
+\Body^ {v} = \left \lbrace A_{G1}:\tilde{d}_{i1}, \ldots,  A_{Gn}:\tilde{d}_{in}\ \right \rbrace \\
+\end{align}
+\end{definition}
+
+\begin{definition}
+\label{def:compatibility-component}
+The compatibility component $R^{c}_{FTG}$ of a fuzzy temporal relation $R_{FTG}$ is a set with the compatibility components for both the head and the body of the relation:
+\begin{align}
+\label{eq:compatibility-component}
+R^{c}_{FTG} = \left \lbrace \Head^{c},\Body^{c} \right \rbrace \\
+\nonumber
+\text{Where: } \\
+\nonumber
+\Head^{c} = \left \lbrace \left[C_{A_{G1}} \right], \ldots,  \left[C_{A_{Gn}} \right] \right \rbrace \\
+\nonumber
+\Body^ {c} = \left \lbrace \left[ c_{i1} \right], \ldots, \left[ c_{in} \right] \right \rbrace \\
+\end{align}
+\end{definition}
+
+\begin{definition}
+\label{def:temporal-component}
+The temporal component $R^{t}_{FTG}$ of a fuzzy temporal relation $R_{FTG}$ is a set with the temporal components for both the head and the body of the relation:
+\begin{align}
+\label{eq:temporal-component}
+R^{t}_{FTG} = \left \lbrace \Head^{t},\Body^{t} \right \rbrace \\
+\nonumber
+\text{Where: } \\
+\nonumber
+\Head^{t} = \left \lbrace \left( \text{PVP}, D_{\text{PVP}}\left[,C_{A_{\text{PVP}}} \right] \right) \right \rbrace \\
+\nonumber
+\Body^ {t} = \left \lbrace  \left[ \text{PVP}, \tilde{d}_{\text{PVP}} \left[,C_{A_{\text{PVP}}} \right] \right]  \right \rbrace \\
+\end{align}
+\end{definition}
+
+Analogously, it is possible to define the the value component for the temporal part and the compatibility component for the temporal part.
+
+\begin{definition}
+\label{def:generalized-primary-key}
+A generalized primary key, $K_G$ is a subset of the head:
+\begin{align}
+\label{eq:generalized-primary-key}
+K_G \subseteq \Head, K_G = \left \lbrace  \left(A_{Gs}:D_{Gs} \right) \right \rbrace \\
+\nonumber
+s\in S \subseteq \left(1, \ldots, n \right) \\
+\nonumber
+\forall s \in S, \text{Typeof } \left(D_{Gs} \right) \in \left \lbrace 1, 2 \right \rbrace \\
+\nonumber
+\forall i, i' \in \left \lbrace 1, \ldots, m\right \rbrace , \exists s \in S: \left(A_{Gs}:d_{is} \right) \neq \left(A_{Gs}:d_{i's} \right)
+\end{align}
+\end{definition}
+
+\texcolor{red}{primary temporal key}
+
+\begin{definiton}
+\label{def:generalized-fuzzy-temporal-key}
+A generalized fuzzy temporal key, $K_{GT}$ is a subset od the head:
+
+\end{definition}

### 2012-04-18 ijcis-pos-model/sources/preliminaries.tex
update

@@
 \begin{definition}
+\label{def:convex-combination-ill-known-constraints}
 Consider a universe $U$, $X$ an ill-known value and $CC \triangleq \left \lbrace C_{1_{Z_1}}, \ldots, C_{n_{Z_n}}  \right \rbrace$ a convex combination of ill-known constraints over the variables $Z_1, \ldots, Z_n$.

### 2012-04-18 ijcis-pos-model/sources/temporalModel.tex
update

--- 2012-04-05 ijcis-pos-model/sources/temporalModel.tex
@@
 
-\vglue13pt
-%\begin{table}[htbp]
-\tcap{Valid-Time interval data types}
-\centerline{\small VALID TIME DATA TYPES}
-\vglue-6pt
-\centerline{\small\baselineskip=13pt
-\begin{tabular}{c p{2cm} p{2cm}}\\
-\hline
-Type for S & Type for E & Semantics \\
-\hline
-Unknown & Unkown & Unknown interval. \\
-\hline
-1,2 & 1,2 & A possibilistic time interval. \\
-\hline\\
-\end{tabular}
-\label{tbl:valid-time-interval}}
-%\end{table}
+\subsection{\label{subsec:temporal-model}The generalized temporal model}
+The model is based on the GEFRED~\cite{Medina1994} (Generalized Model of Fuzzy Relational DB) model.
+This model is extended by adding valid-time support.
+The information in the system is defined by the following elements:
 
-The following is a study of the constraints in the range of values for both starting and ending times to represent a valid time interval.
+\begin{definition}
+\label{def:generalized-fuzzy-domain}
+$D$ is the discourse domain, $\tilde \Pow\left(D \right)$ is the set of all possibility distributions defined on $D$, plus the NULL constant. The generalized fuzzy domain $D_G$ is defined as:
+\begin{equation}
+\label{eq:generalized-fuzzy-domain}
+D_G \subseteq \tilde \Pow\left(D \right)\cup \text{NULL}
+\end{equation}
+\end{definition}
+The datatypes for $D_G$ are shown in table \ref{tbl:gefred-data-types}.
 
-The time period inside a valid time interval is denoted by two constraints:
-\begin{eqnarray}
-C_1\stackrel{\triangle}{=}\left(\geq,X\right)\\
-C_2\stackrel{\triangle}{=}\left(\leq,Y\right).
-\end{eqnarray}
-Applying the inference of uncertainty as proposed in our general reasoning, we find for the first constraint that:
-\begin{eqnarray}
-\Pos\left(C_1([a,b])\right) &=& \min_{r\in[a,b]}\Big(\sup_{r\leq w}\pi_{X}(w)\Big)\\
-\Nec\left(C_1([a,b])\right) &=& \min_{r\in[a,b]}\Big(\inf_{r>w}1-\pi_{X}(w)\Big)
-\end{eqnarray}
-which can be simplified to:
-\begin{eqnarray}
-\Pos\left(C_1([a,b])\right) &=& \sup_{a\leq w}\pi_{X}(w)\\
-\Nec\left(C_1([a,b])\right) &=& \inf_{a>w}1-\pi_{X}(w).
-\end{eqnarray}
-For the second constraint, we find that:
-\begin{eqnarray}
-\Pos\left(C_2([a,b])\right) &=& \min_{r\in[a,b]}\Big(\sup_{r\geq w}\pi_{Y}(w)\Big)\\
-\Nec\left(C_2([a,b])\right) &=& \min_{r\in[a,b]}\Big(\inf_{r<w}1-\pi_{Y}(w)\Big)
-\end{eqnarray}
-which can be simplified to:
-\begin{eqnarray}
-\Pos\left(C_2([a,b])\right) &=& \sup_{b\geq w}\pi_{Y}(w)\\
-\Nec\left(C_2([a,b])\right) &=& \inf_{b<w}1-\pi_{Y}(w).
-\end{eqnarray}
-The uncertainty about the inclusion of an interval $I=[a,b]$ in the interval with ill-known boundaries can now be found by evaluating $[a,b]$ against the evaluation function $\lambda$ with $\bool=\wedge$.
-Application of \eqref{eq:conjunctive1} and \eqref{eq:conjunctive2} leads to:
-\begin{eqnarray}
-\Pos\left(\lambda([a,b])\right)&=&\min\bigg(\Pos(C_1([a,b])),\Pos(C_2([a,b]))\bigg)\\
-\Nec\left(\lambda([a,b])\right)&=&\min\bigg(\Nec(C_1([a,b])),\Nec(C_2([a,b]))\bigg).
-\end{eqnarray}
-These last expressions can also be expanded as:
-\begin{eqnarray}
-\label{eq:interval-pos}
-\Pos\left(\lambda([a,b])\right)&=&\min\bigg(\sup_{a\leq w}\pi_{X}(w),\sup_{b\geq w}\pi_{Y}(w)\bigg)\\
-\label{eq:interval-nec}
-\Nec\left(\lambda([a,b])\right)&=&\min\bigg(\inf_{a>w}1-\pi_{X}(w),\inf_{b<w}1-\pi_{Y}(w)\bigg).
-\end{eqnarray}
-Note that the interval $[X,Y]$ used here, is certainly not a fuzzy interval.
-Instead, we are dealing with an ill-known interval, i.e.
-it is a crisp interval, but it is partially unknown which values are in this interval.
-The uncertainty stems from the fact that the interval boundaries are ill-known.
+\begin{definition}
+\label{def:typeof-domain}
+The function typeof$\left(D \right)$ returns the datatype associated with the domain $D$.
 
-Within this framework for the reasoning of the knowledge about the interval, we are going to study the semantics from the point of view of the valid time as well as from the point of view of the temporal database.
-
-\subsubsection{S = Unknown, E = Unknown}
-Consider that the possibility distribution for S is Unknown and for E is also Unknown. The possibility distributions are $\pi_S$ and $\pi_E$ respectively:
-
-\begin{eqnarray}
-\pi_{S}(t \in \T) = 1\\
-\pi_{E}(t \in \T) = 1
-\end{eqnarray}
-
-Therefore the constraints $C_1$ and $C_2$ are:
-
-\begin{eqnarray}
-\Pos\left(C_1([a,b])\right) &=& \sup_{a\leq w}\pi_{S}(w) = 1\\
-\Nec\left(C_1([a,b])\right) &=& \inf_{a>w}1-\pi_{S}(w) = 0\\
-\Pos\left(C_2([a,b])\right) &=& \sup_{b\geq w}\pi_{E}(w)=1\\
-\Nec\left(C_2([a,b])\right) &=& \inf_{b<w}1-\pi_{E}(w)=0
-\end{eqnarray}
-
-\begin{eqnarray}
-\Pos\left(\lambda([a,b])\right)&=&\min\bigg(\sup_{a\leq w}\pi_{S}(w),\sup_{b\geq w}\pi_{E}(w)\bigg) \\
-&=& \min(1,1)  = 1\\
-\Nec\left(\lambda([a,b])\right)&=&\min\bigg(\inf_{a>w}1-\pi_{S}(w),\inf_{b<w}1-\pi_{E}(w)\bigg)\\
-&=& \min(0,0) = 0
-\end{eqnarray}
-
-\begin{example}
-A constant object in the database. It is an object that is always valid in the database.
-The main issue in a valid time database is that the update sentence have to modify the starting or ending point distribution.
-The most natural way to convert a snapshot database to a temporal database is to consider the objects belonging to the interval [Unknown, Unknown].
-\end{example}
-
-\subsubsection{[Unknown,value] or [value,Unknown]}
-These two cases are managed in this section. Here we will note $\pi_{S}(t)$ or $\pi_{E}(t)$ as the possibility distribution for 'value' in the starting or ending points.
-
-First of all the case [Unknown,value] in \eqref{eq:interval-pos} and in \eqref{eq:interval-nec}:
-
-\begin{eqnarray}
-\Pos\left(\lambda([a,b])\right)&=&\min (1,\sup_{b\geq w}\pi_{E}(w))\\
- &=& \sup_{b\geq w}\pi_{E}(w)\\
-\Nec\left(\lambda([a,b])\right)&=&\min (0,\inf_{b<w}1-\pi_{E}(w))\\
-&=& 0
-\end{eqnarray}
-
-Similarily for [value,Unknown] the possibility and neccesity are:
-\begin{eqnarray}
-\Pos\left(\lambda([a,b])\right)&=&\min (\sup_{a\leq w}\pi_{S}(w),1)\\
- &=& \sup_{a\leq w}\pi_{S}(w)\\
-\Nec\left(\lambda([a,b])\right)&=&\min (\inf_{a>w}1-\pi_{S}(w),0)\\
-&=& 0
-\end{eqnarray}
-
-The semantics for an object with valid time given by [Unknown,value] is an object that has been valid from the beggining to an ending time.
-Analogously, the semantics for an object with a valid time given by the interval [value,Unknown] is an object that it is still valid in the database.
-
-\subsubsection{[Undefined,Undefined]}
-Maybe this combination has no sense for a valid time object. The possibility distributions for both starting and ending points are:
-
-\begin{eqnarray}
-\pi_{S}(t \in \T) = 0\\
-\pi_{E}(t \in \T) = 0
-\end{eqnarray}
-
-Therefore the constraints $C_1$ and $C_2$ are:
-
-\begin{eqnarray}
-\Pos\left(C_1([a,b])\right) &=& \sup_{a\leq w}\pi_{S}(w) = 0\\
-\Nec\left(C_1([a,b])\right) &=& \inf_{a>w}1-\pi_{S}(w) = 1\\
-\Pos\left(C_2([a,b])\right) &=& \sup_{b\geq w}\pi_{E}(w)=0\\
-\Nec\left(C_2([a,b])\right) &=& \inf_{b<w}1-\pi_{E}(w)=1
-\end{eqnarray}
-
-Taking these values into  \eqref{eq:interval-pos} and in \eqref{eq:interval-nec}:
-\begin{eqnarray}
-\Pos\left(\lambda([a,b])\right)&=&\min (0,0)\\
- &=& 0\\
-\Nec\left(\lambda([a,b])\right)&=&\min (1,1)\\
-&=& 1
-\end{eqnarray}
-
-\textcolor{red}{TAKE A LOOK INTO THIS}
-
-Which is an inconsistency.
-
-\subsection{[Undefined,value] or [value,Undefined]}
-
-The possibility and neccesity measures for [Undefined,value] are :
-\begin{eqnarray}
-\Pos\left(\lambda([a,b])\right)&=&\min (0,\sup_{b\geq w}\pi_{E}(w))\\
- &=& 0\\
-\Nec\left(\lambda([a,b])\right)&=&\min (1,\inf_{b<w}1-\pi_{E}(w))\\
-&=& \inf_{b<w}1-\pi_{E}(w)
-\end{eqnarray}
-
-The other way around, [value,Undefined]:
-\begin{eqnarray}
-\Pos\left(\lambda([a,b])\right)&=&\min (\sup_{a\leq w}\pi_{S}(w),0)\\
- &=& 0\\
-\Nec\left(\lambda([a,b])\right)&=&\min (\inf_{a>w}1-\pi_{S}(w),1)\\
-&=& \inf_{a>w}1-\pi_{S}(w)
-\end{eqnarray}
-
-\textcolor{red}{TAKE A LOOK INTO THIS}
-This is also an inconsistency.
-
-\subsubsection{[NULL,NULL] and maybe any combination with NULL}
-I think the framework we have can not deal with NULL computations. Maybe just for non-specified time or something like that.
-
-\subsubsection{[Undefined,Unknown] or [Unknown,Undefined]}
-The semantic for this could be also unconsistent.
-
-\begin{example}
-Consider a snapshot database (table \ref{tbl:snapshot-car-models}) that stores information about car models. The primary key is given by the set $\lbrace$ Brand, Model $\rbrace$.
+\end{definition}
 
@@
 %\begin{table}[htbp]
-\tcap{Car model snapshot database}
-\centerline{\small CAR MODELS}
+\tcap{\label{tbl:gefred-data-types}Data types}
+%\centerline{\small TITLE}
 \vglue-6pt
 \centerline{\small\baselineskip=13pt
-\begin{tabular}{c c c}\\
-\hline
-\textbf{Brand} & \textbf{Model} & Segment \\
-\hline
-Peugeot & 205 & B \\
-Peugeot & 206 & B \\
-Peugeot & 207 & B \\
-Peugeto & 208 & B \\
-Citroen & C2 & B \\
+\begin{tabular}{c p{4cm} }\\
+No. & Datatype \\ \hline
+1 & A single scalar. \\
+2 & A single number. \\
+3 & A set of mutually exclusive possible scalar assignations. \\
+4 & A set of mutually exclusive possible numeric assignations. \\
+5 & A possibility distribution in a scalar domain. \\
+6 & A possibility distribution in a numeric domain. \\
+7 & A real number in $\left[0, 1 \right]$ referring to degree of matching. \\
+8 & An \emph{UNKNOWN} value. \\
+9 &  An \emph{UNDEFINED} value. \\
+10 & A \emph{NULL} value. \\
 \hline\\
 \end{tabular}
-\label{tbl:snapshot-car-models}
 }
 
-Consider now that the user wants to make that table a valid-time table, then, from the point of view of the storage, the following changes should be done:
+It is possible to define a more specific generalized temporal domain, $\T_G$:
 
-\begin{itemize}
-\item
-A new definition for the primary key, including a version field.
-\item
-The addition of two new columns to store the valid-time interval. By default the initial values are Unknown for both starting and ending points of the interval.
-\end{itemize}
-Therefore, the table \ref{tbl:snapshot-car-models} is transformed into table \ref{tbl:initial-valid-time-car-models}:
+\begin{definition}
+\label{def:generalized-fuzzy-temporal-domain}
+Generalized fuzzy temporal domain.\\
+If $\T$ is the temporal domain, $\tilde \Pow\left( \T\right)$ is the set of all \emph{normalized} possibility distributions (see Section \ref{subsec:possibility-theory}, equation \eqref{NormalizationProperty}) defined on $\T$.
+The \textbf{Generalized Fuzzy Temporal Domain}, $\T_G$ is
+\begin{equation}
+\T_G \subseteq \left \lbrace \tilde \Pow\left( \T\right) \cup \text{NULL} \right \rbrace
+\end{equation}
+\end{definition}
 
- \vglue13pt
-%\begin{table}[htbp]
-\tcap{Car model valid-time database. Note that U is the abbreviation for the Unknown constant.}
-\centerline{\small VALID TIME CAR MODELS}
-\vglue-6pt
-\centerline{\small\baselineskip=13pt
-\begin{tabular}{c c c c c c}\\
-\hline
-\textbf{Brand} & \textbf{Model} & \textbf{VersionID} & Segment & S & E \\
-\hline
-Peugeot & 205 & 001 & B & U & U \\
-Peugeot & 206 & 001 & B & U & U \\
-Peugeot & 207 & 001 & B & U & U \\
-Peugeto & 208 & 001 & B & U & U \\
-Citroen & C2 & 001 & B & U & U \\
-\hline\\
-\end{tabular}
-\label{tbl:initial-valid-time-car-models}
-}
+Note that $\T_G \subseteq D_G$.
+The datatypes for this domain have been studied previously in section \ref{sec:time-rep} and are shown in tables \ref{tbl:time-point-types},\ref{tbl:time-interval-types}.
+
+\begin{definition}
+A generalized fuzzy temporal relation $R_{FTG}$ is given by:
+\label{def:generalized-fuzzy-temporal-relation}
+\begin{equation}
+\label{eq:generalized-fuzzy-temporal-relation}
+R_{FTG} = \left(\Head, \Body \right)
+\end{equation}
+Where $\Head$ is the Head of the relation and consist on a fixed set of triplets attribute- domain - compatibility with an optional the valid-time attribute:
+
+\begin{align}
+\label{eq:head-valid-time}
+\Head = \big \lbrace \left(A_{G1}:D_{G1}\left[,C_{A_{G1}} \right] \right),\\
+\nonumber
+ \ldots,\\
+ \nonumber
+  \left(A_{Gn}:D_{Gn}\left[,C_{A_{Gn}} \right] \right),\\
+  \nonumber
+  \Big[  \left( \text{PVP}, D_{\text{PVP}}\left[,C_{A_{\text{PVP}}} \right] \right) \Big] \big \rbrace
+\end{align}
+Note that $D_{Gj}$ ($j = 1, \ldots, n$) is the domain for the attribute $A_{Gj}$. $C_{A_{Gj}}$ is the compatibility attribute in the unit interval $\left[0, 1 \right]$.
+
+$\Body$ is the body of the relation and it consists on a set of tuples. Each tuple is a triplet attribute- value- degree with an optional valid-time attribute:
+
+\begin{align}
+\label{eq:body-valid-time}
+\Body = \big \lbrace \left(A_{G1}:\tilde{d}_{i1}\left[,c_{i1} \right] \right),\\
+\nonumber
+ \ldots,\\
+ \nonumber
+  \left(A_{Gn}:\tilde{d}_{in}\left[,c_{in} \right] \right),\\
+  \nonumber
+   \Big[  \left( \text{PVP}, \tilde{d}_{\text{PVP}} \left[,C_{A_{\text{PVP}}} \right] \right)  \Big] \big \rbrace
+\end{align}
+
+\end{definition}
+
+The definition in \cite{Medina1994} for $R_{FTG}$ shows that classical relations are a particular case of this model.
+
+\begin{definition}
+\label{def:value-component}
+The value component $R^{v}_{FTG}$ of a fuzzy temporal relation $R_{FTG}$ is a set with the value components for both the head and the body of the relation:
+\begin{align}
+\label{eq:value-component}
+R^{v}_{FTG} = \left \lbrace \Head^{v},\Body^{v} \right \rbrace \\
+\nonumber
+\text{Where: } \\
+\nonumber
+\Head^{v} = \left \lbrace A_{G1}:D_{G1}, \ldots,  A_{Gn}:D_{Gn} \right \rbrace \\
+\nonumber
+\Body^ {v} = \left \lbrace A_{G1}:\tilde{d}_{i1}, \ldots,  A_{Gn}:\tilde{d}_{in}\ \right \rbrace \\
+\end{align}
+\end{definition}
+
+\begin{definition}
+\label{def:compatibility-component}
+The compatibility component $R^{c}_{FTG}$ of a fuzzy temporal relation $R_{FTG}$ is a set with the compatibility components for both the head and the body of the relation:
+\begin{align}
+\label{eq:compatibility-component}
+R^{c}_{FTG} = \left \lbrace \Head^{c},\Body^{c} \right \rbrace \\
+\nonumber
+\text{Where: } \\
+\nonumber
+\Head^{c} = \left \lbrace \left[C_{A_{G1}} \right], \ldots,  \left[C_{A_{Gn}} \right] \right \rbrace \\
+\nonumber
+\Body^ {c} = \left \lbrace \left[ c_{i1} \right], \ldots, \left[ c_{in} \right] \right \rbrace \\
+\end{align}
+\end{definition}
+
+\begin{definition}
+\label{def:temporal-component}
+The temporal component $R^{t}_{FTG}$ of a fuzzy temporal relation $R_{FTG}$ is a set with the temporal components for both the head and the body of the relation:
+\begin{align}
+\label{eq:temporal-component}
+R^{t}_{FTG} = \left \lbrace \Head^{t},\Body^{t} \right \rbrace \\
+\nonumber
+\text{Where: } \\
+\nonumber
+\Head^{t} = \left \lbrace \left( \text{PVP}, D_{\text{PVP}}\left[,C_{A_{\text{PVP}}} \right] \right) \right \rbrace \\
+\nonumber
+\Body^ {t} = \left \lbrace  \left[ \text{PVP}, \tilde{d}_{\text{PVP}} \left[,C_{A_{\text{PVP}}} \right] \right]  \right \rbrace \\
+\end{align}
+\end{definition}
+
+Analogously, it is possible to define the the value component for the temporal part and the compatibility component for the temporal part.
+
+\begin{definition}
+\label{def:generalized-primary-key}
+A generalized primary key, $K_G$ is a subset of the head:
+\begin{align}
+\label{eq:generalized-primary-key}
+K_G \subseteq \Head, K_G = \left \lbrace  \left(A_{Gs}:D_{Gs} \right) \right \rbrace \\
+\nonumber
+s\in S \subseteq \left(1, \ldots, n \right) \\
+\nonumber
+\forall s \in S, \text{Typeof } \left(D_{Gs} \right) \in \left \lbrace 1, 2 \right \rbrace \\
+\nonumber
+\forall i, i' \in \left \lbrace 1, \ldots, m\right \rbrace , \exists s \in S: \\
+\nonumber
+\left(A_{Gs}:d_{is} \right) \neq \left(A_{Gs}:d_{i's} \right)
+\end{align}
+\end{definition}
+
+\begin{definition}
+\label{def:generalized-fuzzy-temporal-key}
+A generalized fuzzy temporal key, $K_{GT}$ is a subset of the head. An attribute, namely \emph{version} $V$ is added when dealing with valid-time.
+\begin{align}
+\label{eq:generalized-fuzzy-temporal-key}
+K_{GT} \subseteq \Head, K_{GT} = \left \lbrace  \left(A_{Gs}:D_{Gs} \right) \right.  \\
+\nonumber
+ \left. \cup  \left(V_{ID}:D_{ID} \right)	\right \rbrace \\
+\nonumber
+\text{Typeof }\left(D_{ID} \right) = \\
+\nonumber
+s\in S \subseteq \left(1, \ldots, n \right) \\
+\nonumber
+\forall s \in S, \text{Typeof } \left(D_{Gs} \right) \in \left \lbrace 1, 2 \right \rbrace \\
+\nonumber
+\forall i, i' \in \left \lbrace 1, \ldots, m\right \rbrace , \exists s \in S: \\
+\nonumber
+\left(A_{Gs}:d_{is} \right) \neq \left(A_{Gs}:d_{i's} \right)
+\end{align}
+
+\end{definition}
+
+\begin{example}
 
@@
 
-\subsection{\label{subsec:temporal-model}The generalized temporal model}
+\subsection{\label{subsec:data-manipulation}Data manipulation language}
+
+The algebra for the data manipulation is defined in \cite{Medina1994}.
+The Generalized Fuzzy Relational Algebra manipulates relations like $R_{FG}$.
+The operations defined are: \emph{Union, Intersection, Difference, Cartesian Product, Projection, Join} and \emph{Selection}.
+Thus, in this section we will describe and implement the following operations for temporal databases, as described in \ref{subsubsec:Understanding-valid-time-databases}.
+The operations implemented are: \emph{Insert, Modify, Delete, } and \emph{Revise}.
+The semantics of the operations will be the same that those defined for a crisp temporal database, whereas the temporal representation is made by the possibilistic valid-time period and the ill-known constraints (see sections \ref{sec:prelim} and \ref{sec:time-rep}.
+
+It is important to nottice that whereas the result of the evaluation of any comparison between crisp time intervals is boolean, the evaluation of any comparison between \emph{PVPs} is in the unit interval.
+Therefore, for a valid-time object, say $A \in R$, in a crisp temporal database it is not possible an overlaping among any valid-time interval ($\forall I, \not \exists J: \left(A, I \right) \in R \wedge \left(A, J \right) \in R \wedge \not \Overlap\left(I, J \right)$).
+Thus, from the point of view of a fuzzy temporal database, there are the following options:
+
+\begin{itemize}
+\item [\textbf{Strictly consistent}] It is not possible for two PVPs of the same object to be valid at the same time.
+In other words, if $A$ is an object, $R$ is a relation and $\pi_{I}\left(t \right)$ and $\pi_{J}\left(t \right)$ are the possibilities for a time point $t \in \T$ to be in the PVPs $I$ and $J$ respectively,
+\begin{align}
+\label{eq:stricly-consistent}
+\forall I,J, \forall t \in \T :\\
+%\nonumber
+%\left(A, I \right) \in R \wedge \\
+%\nonumber
+%\left(A, J \right) \in R : \\
+\nonumber
+\pi_{I}(t) + \pi_{J}(t) \leq 1
+\end{align}
+
+\item [\textbf{Ill-consistent}] The overlapping of two PVPs of the same object do overlap. It can be distinghised among the following sub-types:
+	\begin{itemize}
+	\item [\emph{Co-existence}] Two versions of the same object $A$ may exist at the same time.
+	\begin{align}
+	\label{eq:co-existence}
+	\forall I, \exists ! J, \exists t \in \T : \\
+	\nonumber
+	\pi_{I}(t) + \pi_{J}(t) \geq 1
+	\end{align}
+	\item [\emph{Weak-consistence}] Several versions of the same object $A$ may exist at the same time.
+	\begin{align}
+	\label{eq:weak-consistence}
+	\forall I, \exists J, \exists t \in \T : \\
+	\nonumber
+	\pi_{I}(t) + \pi_{J}(t) \geq 1
+	\end{align}
+	\end{itemize}
+\end{itemize}
+
+In the implementation of the DML operations we will consider a strictly consistent temporal database.
+Since the time intervals are now possibilistic valid-time periods, PVPs, the auxiliary functions defined in equations \eqref{eq:close-a-crisp-interval} to \eqref{eq:close-current} should be re-defined.
 
 \begin{definition}
-Generalized fuzzy temporal domain.\\
-If $\T$ is the temporal domain, $\tilde \Pow\left( \T\right)$ is the set of all possibility distributions defined on $\T$.
-The \textbf{Generalized Fuzzy Temporal Domain}, $\T_G$ is
-\begin{equation}
-\T_G \subseteq \left \lbrace \tilde \Pow\left( \T\right) \cup NULL \right \rbrace
-\end{equation}
+\label{def:close-r-a-pvp}
+Consider two PVPs given by $I = \left(X, Y \right)$ and $J = \left(X', Y' \right)$.
+The CloseR function closes the PVP given by $I$ with a convex combination of ill-known constraints (see section \ref{subsec:interval-evaluation-by-ill-known-constraints}):
+\begin{align}
+\label{eq:close-r-a-pvp}
+\text{CloseR}\left(I, J\right) = \\
+\begin{cases}
+\nonumber
+I   & \mbox{ if } Y \not \in \text{UNKNOWN}  \\
+I = \left[X, Z \right] & \mbox{ if } Y \in \text{UNKNOWN}, \\
+& Z \triangleq \left \lbrace C_1\left(>, X \right), C_2\left(<, X' \right) \right \rbrace
+\end{cases}
+\end{align}
 \end{definition}
+
+\begin{definition}
+\label{def:pvp-current-in-relation}
+Consider an object $A$ and a relation $R$ in a fuzzy temporal database.
+The function Current$\left(R, A \right)$ obtains the tuple $\left(A, I \right), I = \left(X, Y \right)$ that is current in the relation $R$, that is it, $I = \left[X , \text{UNKNOWN} \right]$.
+
+\begin{align}
+\label{eq:pvp-current-in-relation}
+\mbox{Current} \left(R, A \right) &=& \\ 
+\begin{cases}
+\nonumber
+I & \mbox{ if } \exists  \in I : \\
+&  (A,I) \in R \wedge Y \in \text{UNKNOWN}\\
+\emptyset & \mbox{ in any other case }
+\end{cases}
+\end{align}
+\end{definition}
+
+Now it is possible to close the current version of an object by using \eqref{eq:close-r-a-pvp} and \eqref{eq:pvp-current-in-relation}.
+
+\begin{definition}
+\label{def:pvp-close-current-version}
+Consider an object $A$, the relation $R$ and a pvp, $J$.
+Then, the function close - current$\left(R, A, J \right)$ closes any current version of the object $A$ if it exists and add the new version $\left(A, J \right)$.
+
+\begin{eqnarray}
+\label{eq:pvp-close-current}
+\text{Close-current} \left(R, A, J \right) =\\
+\begin{cases}
+\nonumber
+\big \lbrace R - \left(A, I \right) \cup \left \lbrace \left(A, \mbox{ CloseR } \left(I, J\right) \right) \cup \left(A, J\right)\right \rbrace  \big \rbrace \\
+\nonumber
+\mbox{ if } I = \mbox{ Current } \left(R, A \right) \wedge J > \mbox{ CloseR } \left(I, J \right)   \\
+\nonumber R , \text{ in any other case}
+\end{cases}
+\end{eqnarray}
+\end{definition}
+
+\subsubsection{\label{subsubsec:insert-fuzzy-temporal}Insert}
+
+\subsubsection{\label{subsubsec:modify-fuzzy-temporal}Modify}
+
+\subsubsection{\label{subsubsec:delete-fuzzy-temporal}Delete}
+
+\subsubsection{\label{subsubsec:revise-fuzzy-temporal}Revise}

### 2012-04-18 trunk/ijcis-pos-model/sources/temporaldatabases.tex
update

@@
 
-\subsection{Understanding Valid-time Databases}
+\subsection{\label{subsubsec:Understanding-valid-time-databases}Understanding Valid-time Databases}
 This subsection is devoted to describe the behaviour of a crisp valid-time database.
@@
 \begin{align}
+\label{eq:close-a-crisp-interval}
 \mbox{CloseR} \left( I, J \right) &=& \\ 
@@
 \begin{eqnarray}
-%\label{eq:close-current}
+\label{eq:close-current}
 \text{Close-current} \left(R, A, J \right) =\\

### 2012-04-19 ijcis-pos-model/sources/temporalModel.tex
draft.

@@
 \subsubsection{\label{subsubsec:insert-fuzzy-temporal}Insert}
+The user wants to store an object $A$ which is valid in the relation $R$ during the time interval specified by the PVP, $I = \left(X, Y \right)$.
+%
+%
+%The interpretation of the insert operation is the following: The user wants to store in the database that an object $A$  is true for some period(s) of time $I$ in the given relation $R$.
+To indicate that the object $A$ is current in the relation, the special value Until Current, $\UC$  is used.
+There are two main cases when performing an insert operation:
+\begin{enumerate}
+\item The object $A$ was never in the relation $R$: The object is added with the valid-time indicated by the PVP $I$.
 
+\item The object $A$ is in the relation $R$. Depending on the value of the time interval $I$, there are two possibilities:
+	\begin{enumerate}
+	\item Insert $\left(A, I\right)$ in the relation $R$. If the time interval $I$ does not overlap any other valid-time interval in the relation $R$.
+	\item Reject the insertion, if the time interval $I$ do overlap any existing valid-time interval for the object $A$ in the relation $R$.
+	\end{enumerate}
+
+\end{enumerate}
+
+The algorithm for the implementation of the insert operation is the following:
+
+\begin{align}
+\label{eq:insert-fuzzy-temporal}
+\INS \left(R ,A, I\right) =\\
+\begin{cases}
+\nonumber
+R \cup \left(A, I \right), & \mbox{ if }  A \not \in R\\
+R \cup   \left(A, I \right) & \mbox{ if }  A \in R \wedge \\
+\big \lbrace \left( \exists J= \left(X', Y' \right): \left(A, J \right) \in R \wedge \right. \\
+\left. Y' \in \left \lbrace \left(<, X \right)\right \rbrace  \right) \vee \\
+\left( \exists K= \left(X'', Y'' \right): \left(A, K \right) \in R \wedge \right. \\
+\left. X'' \in \left \lbrace \left(>, Y \right)\right \rbrace \vee \right) \big \rbrace\\
+%&  J<I<K  \big \rbrace\\
+R, & \mbox{ otherwise }  
+\end{cases} 	
+\end{align}
 \subsubsection{\label{subsubsec:modify-fuzzy-temporal}Modify}
+The delete operation logically removes a current object $A$ which is valid in the relation $R$:
+\begin{align}
+\label{eq:delete-fuzzy-temporal}
+\DEL \left(R ,A \right) =\\
+\begin{cases}
+\nonumber
+R - \left(A, I \right), & \mbox{ if } \left( A, I \right) \in R\\
+R, & \mbox{ otherwise }  
+\end{cases} 	
+\end{align}
 
 \subsubsection{\label{subsubsec:delete-fuzzy-temporal}Delete}
+This operation adds new information about an existing object $A$ in the relation $R$. The modify operation does not remove any previous value of the object $A$.
+Note that the modify operation is only applicable when the object $A$ is current in the relation $R$; $ \left(A, I \right) \in R, I = \left[X, \UC \right]$.
+
+\begin{align}
+\label{eq:modify-fuzzy-temporal}
+\MOD \left( R, A, I \right) =\\
+\begin{cases}
+\nonumber
+\mbox{ Close-current }\left(R, A, \mbox{Current}\left(R, A \right) \right) \\
+ \mbox{ if Current}\left(R ,A \right) \neq \emptyset \\
+ R,  \mbox{ otherwise }
+\end{cases}
+\end{align}
 
 \subsubsection{\label{subsubsec:revise-fuzzy-temporal}Revise}
+This operation changes the value of the object $A$ in the relation $R$, at a given time interval $I$. This operation is defined when a correction on the values for the object $A$ should be done.
+\begin{align}
+\label{eq:revise-fuzzy-temporal}
+\mbox{ Revise}\left(R, A, I \right)=\\
+\begin{cases}
+\nonumber
+ R - \left(A', I \right) \cup \left(A, I \right)  \\
+  \mbox{ if } \left(A', I \right) \in R: A = \left(a_1, \ldots, a_n \right),\\
+A'= \left(a_1', \ldots, a_n' \right)   \\   
+R,  \mbox{ otherwise }
+\end{cases}
+\end{align}

### 2012-05-23 ijcis-pos-model/sources/abstract.tex
First minor changes suggested by Nicolas and Olga.
Only introduction and abstract.

--- 2012-01-10 book/author.tex
@@
-%%%%%%%%%%%%%%%%%%%% author.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%
 %
-% sample root file for your "contribution" to a contributed volume
+% ABSTRACT
 %
-% Use this file as a template for your own input.
-%
-%%%%%%%%%%%%%%%% Springer %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\documentclass[graybox]{svmult}
-
-% choose options for [] as required from the list
-% in the Reference Guide
-
-\usepackage{mathptmx}       % selects Times Roman as basic font
-\usepackage{helvet}         % selects Helvetica as sans-serif font
-\usepackage{courier}        % selects Courier as typewriter font
-\usepackage{type1cm}        % activate if the above 3 fonts are
-                            % not available on your system
-%
-\usepackage{makeidx}         % allows index generation
-\usepackage{graphicx}        % standard LaTeX graphics tool
-                             % when including figure files
-\usepackage{multicol}        % used for the two-column index
-\usepackage[bottom]{footmisc}% places footnotes at page bottom
-\usepackage{url}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\usepackage{amsfonts}		% math fonts.
-
-% see the list of further useful packages
-% in the Reference Guide
-
-\makeindex             % used for the subject index
-                       % please use the style svind.ist with
-                       % your makeindex program
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\newcommand{\Pow}{\mathcal{P}}
-\newcommand{\N}{\operatorname{N}}
-\newcommand{\bool}{\operatorname*{\mathcal{B}}}
-\newcommand{\Pos}{\operatorname{Pos}}
-\newcommand{\Nec}{\operatorname{Nec}}
-\newcommand{\T}{\mathcal{T}}
-
-\begin{document}
-
-\title*{An overview on temporal databases.\\
- New approaches and proposals.}
-% Use \titlerunning{Short Title} for an abbreviated version of
-% your contribution title if the original one is too long
-\author{Jos\'e Pons \and Christophe Billiet \and Olga Pons \and Guy de Tr\'e}
-% Use \authorrunning{Short Title} for an abbreviated version of
-% your contribution title if the original one is too long
-\institute{Jos\'e Pons \at Department of Computer Science and Artificial Intelligence, Univeristy of Granada, C/ Periodista Daniel Saucedo Aranda s/n \email{jpons@decsai.ugr.es}
-\and Christophe Billiet \at Department of Telecommunications and Information Processing, Ghent University, Belgium, St.-Pietersnieuwstraat 41, B-9000 Gent,Belgium.
-\email{Christophe.Billiet@telin.ugent.be}
-\and Olga Pons \at Department of Computer Science and Artificial Intelligence, Univeristy of Granada, C/ Periodista Daniel Saucedo Aranda s/n \email{opc@decsai.ugr.es}
-\and Guy De Tr\'e \at Department of Telecommunications and Information Processing, Ghent University, Belgium, St.-Pietersnieuwstraat 41, B-9000 Gent,Belgium. \email{Guy.DeTre@telin.ugent.be}
-}
-%
-% Use the package "url.sty" to avoid
-% problems with special characters
-% used in your e-mail or web address
-%
-\maketitle
-
-\abstract*{There exist some objects in the nature that have a time-variant (e.g.
-an employee's working life) or time-related (e.g.
-the transactions in a bank account) nature.
-Modelling these objects in a (relational) database is possible, but the time-related attributes have an impact on the consistency of the entire database.
-Therefore, some temporal database models have been proposed to deal with this.
-These models deal both with the problem of representation and querying.
-Time itself is a source of imprecision / vagueness / uncertainty since the measures of time are inherently imprecise.
-Humans beings manage the time with imprecision and vagueness.
-Nevertheless, the amount of uncertainty in the temporal indications are defined by socially well-known boundaries.
-Several proposals for dealing with imprecision in time have been done: some proposals consider the basis of the system the imprecision related to the conversion of diferent time points and intervals.
-While some other proposals consider the specification of the time as the source of imprecision: the time points are ill-known.}
-
-\abstract{There exist some objects in the nature that have a time-variant (e.g.
-an employee's working life) or time-related (e.g.
-the transactions in a bank account) nature.
-Modelling these objects in a (relational) database is possible, but the time-related attributes have an impact on the consistency of the entire database.
-Therefore, some temporal database models have been proposed to deal with this.
-These models deal both with the problem of representation and querying.
-Time itself is a source of imprecision / vagueness / uncertainty since the measures of time are inherently imprecise.
-Humans beings manage the time with imprecision and vagueness.
-Nevertheless, the amount of uncertainty in the temporal indications are defined by socially well-known boundaries.
-Several proposals for dealing with imprecision in time have been done: some proposals consider the basis of the system the imprecision related to the conversion of diferent time points and intervals.
-While some other proposals consider the specification of the time as the source of imprecision: the time points are ill-known.}
-
-\section{Introduction}
-\label{sec:introduction}
-\input{sources/intro.tex}
-
-\section{Time modelling}
-\label{sec:time-domain}
-\input{sources/timeDomain.tex}
-
-\section{Temporal Databases}
-\label{sec:temporal-databases}
-\input{sources/temporalDatabases.tex}
-
-\section{Imprecision, Uncertainty and vagueness in temporal databases}
-\label{sec:imprecision-uncertainty}
-\input{sources/fuzzyTemporalDatabases.tex}
-
-\section{Bipolarity on temporal databases}
-\label{sec:bipolarity}
-\input{sources/bipolarity.tex}
-
-\section{Further research work}
-\label{sec:further-research-work}
-\input{sources/frw.tex}
-
-\section*{Acknowledgements}
-Part of this research is supported by the grant BES-2009-013805 within the research project TIN2008-02066: \emph{Fuzzy Temporal Information treatment in relational DBMS}.
-
-\input{referenc}
-\end{document}
+%%%%%%%%%%%%%%%%%%%%%%%%
+In reality, some objects or concepts have properties with a time-variant or time-related nature.
+Modelling these kinds of objects or concepts in a (relational) database schema is possible, but time-variant and time-related attributes have an impact on the consistency of the entire database.
+Therefore, temporal database models have been proposed to deal with this.
+Time itself can be at the source of imprecision, vagueness and uncertainty, since existing time measuring devices are inherently imperfect.
+Accordingly, human beings manage time using temporal indications and temporal notions, which may contain imprecision, vagueness and uncertainty.
+However, the imperfection in human-used temporal indications is supported by human interpretation, whereas information systems need support for this.
+Several proposals for dealing with such imperfections when modelling temporal aspects exist.
+Some of these proposals transform the temporal expression into a compact representation.
+Other proposals consider the temporal indications in the used temporal expressions to be the source of imperfection.
+In this work we present a novel model to deal with imperfections in valid-time databases. Next to that, the data manipulation language, \emph{DML} is defined and implemented.

### 2012-05-23 trunk/ijcis-pos-model/sources/intro.tex
First minor changes suggested by Nicolas and Olga.
Only introduction and abstract.

--- 2012-04-15 ijcis-pos-model/sources/intro.tex
@@
 
-The concept of time itself is very complex to handle and interpret~\cite{Klein1994},~\cite{Shackle1961}, though it is very natural and omnipresent.
+The concept of time itself is very complex to handle and interpret~\cite{Klein1994,Shackle1961}, though it is very natural and omnipresent.
 As information systems often attempt the modelling of natural objects, concepts or processes, they often require modelling temporal aspects or concepts.
-Thus, several proposals have been concerned with the obtaining of theoretical models that allow the modelling or representation of time~\cite{Bolour1982},~\cite{VanderCruyssen1997}.
+Thus, several proposals have been concerned with the obtaining of theoretical models that allow the modelling or representation of time~\cite{Bolour1982,VanderCruyssen1997}.
 
@@
 An example.
-Consider a relation in a relational library database, modelling the presence of books in the library.
-Every physical book is represented by a unique identifier.
-Every record in the relation contains such an identifier, a date on which the corresponding book was loaned and a date on which it was subsequently returned (if it was returned).
-Without further precautions, a library employee could add several records with the same book identifier, different `loaned'-dates and no `returned'-dates.
-This group of records would represent the same physical book being loaned several times on different dates and never returned, which is of course impossible.
+Consider a library database, modelling the presence of books in the library.
+Two dates are stored: the loan and the return date.
+It is clear that a book cannot be loaned again until it is returned.
+Without further precautions, a library employee could loan the same book several times even if it is not returned.
 A temporal database model will typically constrain record insertion and prevent similar modelling inconsistencies.
@@
 Among others, the querying of temporal databases has greatly profited from these temporal relationships, because they allowed for richer and more complex user-specified temporal query demands, by allowing to express more complex relationships between the temporal notions in the temporal expressions in the query and the temporal indications in the database.
-e.g., in a relation modelling who was department head of an institution during which periods of time, a query like `who were the department heads when Thomas worked for the institution' can be evaluated using similar relationships.
+For example, in a relation modelling who was department head of an institution during which periods of time, a query like `who were the department heads when Thomas worked for the institution' can be evaluated using similar relationships.
 
@@
 Among many others, these possible imperfections in descriptions of temporal notions determine an important issue in temporal modelling.
-e.g., the description of the temporal notion in a sentence like `The Belfry of Bruges was finished on a day somewhere between 01/01/1201 A.D.
+Consider as an example the description of the temporal notion in a sentence like `The Belfry of Bruges was finished on a day somewhere between 01/01/1201 A.D.
 and 31/12/1300 A.D.' contains imperfection because of the uncertainty in the used time-related expression.
@@
 
-To allow information systems to cope with these and similar data imperfections, many approaches adopt fuzzy sets~\cite{Zadeh1965} for the representation of temporal information~\cite{Mitra1994},~\cite{Nagypal2003},~\cite{Billiet2011},~\cite{Dubois2003}.
-The temporal relationships studied by Allen were fuzzified by several authors~\cite{Ohlbach2004},~\cite{Nagypal2003},~\cite{Schockaert2008}.
+To allow information systems to cope with these and similar data imperfections, many approaches adopt fuzzy sets~\cite{Zadeh1965} for the representation of temporal information~\cite{Mitra1994,Nagypal2003,Billiet2011,Dubois2003}.
+The temporal relationships studied by Allen were fuzzified by several authors~\cite{Ohlbach2004,Nagypal2003,Schockaert2008}.
 Garrido et al.
 ~\cite{Garrido2009} present different temporal operators, defined by a combination of regular fuzzy comparisons.
-Both~\cite{Garrido2009} and~\cite{Pons2011} deal with uncertainty in temporal expressions concerning time intervals.
+Also,~\cite{Garrido2009,Pons2011} studied uncertainty in temporal expressions concerning time intervals.
 Other approaches, like~\cite{Qiang2009}, use rough sets~\cite{Pawlak1995} to represent time intervals.
@@
 Next to temporal modelling, some attention has been spent on temporal reasoning~\cite{Allen1983}.
-Although temporal reasoning is not discussed in this chapter, it should be noted that, among others, Dubois and Prade et al.~\cite{Dubois2003},~\cite{DuBois1989} have dealt with fuzziness and uncertainty in temporal reasoning.
+Although temporal reasoning is not discussed in this paper, it should be noted that, among others, Dubois and Prade et al.~\cite{Dubois2003,DuBois1989} have dealt with fuzziness and uncertainty in temporal reasoning.
 

### 2012-05-24 ijcis-pos-model/sources/preliminaries.tex
last update

@@
 \begin{align}
+\label{propholdsposs}
 \text{Possibility that $p$ = $T$ (p holds):} &\\
 \nonumber
-Pos(p) = \pi_p(T) \label{propholdsposs} \\
+Pos(p) = \pi_p(T)  \\
+\label{propholdsnecc}
 \text{Necessity that $p$ = $T$ (p holds):} & \\
 \nonumber
-Nec(p) = 1-\pi_p(F) \label{propholdsnecc}
+Nec(p) = 1-\pi_p(F) 
 \end{align}
 
-Here, notation \ref{propholdsposs} denotes the possibility that $p$ = $T$ and the proposition holds, notation \ref{propholdsnecc} denotes the necessity that $p$ = $T$ and the proposition holds.
+Here, equation \eqref{propholdsposs} denotes the possibility that $p$ = $T$ and the proposition holds, equation \eqref{propholdsnecc} denotes the necessity that $p$ = $T$ and the proposition holds.
 
@@
 These are ill-known sets, defined and represented via a start and end point, which will be ill-known values.
-The elements of the set are the values between the start and end point.
-A closed ill-known interval with start point defined by possibilistic variable $X$ and end point by possibilistic variable $Y$ is noted here $\left[X, Y\right]$.
+The elements of the set are the values between the starting and ending points.
+A closed ill-known interval with starting point defined by possibilistic variable $X$ and ending point by possibilistic variable $Y$ is noted here $\left[X, Y\right]$.
 The correspondences and transitions between the representations of ill-known sets, between the representations of ill-known intervals and between the representations of an ill-known set and an ill-known interval are part of the authors current research.
@@
 \vspace*{10pt}
-\fcaption{\label{fig:triangular-dist}Trapezoidal membership function.}
+\fcaption{\label{fig:triangular-dist}Example of fuzzy number.}
 %  \label{fig:fuzzy-validity-period}
@@
 If this modal value $m$ is unique, then $M$ is referred to as a \emph{fuzzy number}. 
-In other words, if the core of a fuzzy interval is a singleton, it is referred to as a fuzzy number (Figure \ref{fig:triangular-dist}).
+%In other words, if the core of a fuzzy interval is a singleton, it is referred to as a fuzzy number (Figure \ref{fig:triangular-dist}).
 
-A simple form of the membership function of a fuzzy interval is a trapezoidal function (Figure \ref{fig:trapezoidal-dist}.
+A simple form of the membership function of a fuzzy interval is a trapezoidal function (Figure \ref{fig:trapezoidal-dist}).
 It can be shown that such a membership function $\mu_T$ for a fuzzy interval $T$ is convex and normalized.
-Four reel values, denoted $\alpha$, $\beta$, $\gamma$ and $\delta$ suffice to represent a trapezoidal membership function of a fuzzy interval.
+Four real values, denoted $\alpha$, $\beta$, $\gamma$ and $\delta$ suffice to represent a trapezoidal membership function of a fuzzy interval.
 In this work, a fuzzy interval defined as such will be noted as $\left[\alpha, \beta, \gamma, \delta\right]$.
@@
 \vspace*{10pt}
-\fcaption{\label{fig:trapezoidal-dist}Trapezoidal membership function.}
+\fcaption{\label{fig:trapezoidal-dist}Example of fuzzy interval.}
 %  \label{fig:fuzzy-validity-period}
@@
 
-So far, we have shown how it can be verified whether or not a set satisfies an ill-known constraint.
+So far, we have shown how it can be verified if a set satisfies or not an ill-known constraint.
+
+\begin{example}
+
+\vspace*{13pt}
+\begin{center}
+{
+\includegraphics[scale=1]{./graphs/example-ill-known.pdf}
+}
+\end{center}
+%\centerline{ \psfig{file=./graphs/Y-time-point.eps}}
+\vspace*{10pt}
+\fcaption{\label{fig:example-ill-known}Example of the evaluation of the ill-known constraint $C \triangleq (\leq, X)$.
+Possibility and necessity measures are shown in grey the upper and lower graphics respectively.
+}
+%  \label{fig:fuzzy-validity-period}
+\vspace*{13pt}
+
+ Consider an ill-known value given by $X = \left[5, 3, 2 \right]$.
+$C \triangleq (\leq, X)$ is the ill-known constraint.
+The set $A$ is $A \subseteq \R^+$.
+Then, the evaluation of the possibility and the neccesity are obtained from \eqref{ill-known-pos} and \eqref{ill-known-nec} respectively.
+(See Figure \ref{fig:example-ill-known}).
+
+\begin{align}
+\Pos(A\text{ satisfies }C) & =\\
+\nonumber
+\min_{a \in A}\left(\sup_{a \leq R}\pi_{X}(w)\right)\\
+\Nec(A\text{ satisfies }C) & =\\
+\nonumber
+\min_{a \in A}\left(\inf_{a > w} 1-\pi_{X}(w)\right) 
+\end{align}
+
+\end{example}
+
  However, from the examples at the beginning of this section, it is observed that Boolean combinations of constraints are required.
@@
 \end{align}
-$Z$ is a fuzzy interval defined by a trapezoidal shape given by $\left[12,\ 14,\ 18,\ 19 \right]$
+$Z$ is a fuzzy interval defined by a trapezoidal shape given by $\left[12,\ 14,\ 18,\ 19 \right]$.
 Figure \ref{fig:example-ill-known-by-const} illustrates the relations among the variables $X$, $Y$ and $Z$.
@@
 \includegraphics[scale=1]{./graphs/ill-known-by-constraints.pdf}
-\label{fig:example-ill-known-by-const}
+
 }
@@
 \vspace*{10pt}
-\fcaption{Ill-known values $X$ and $Y$. The grey area represents the ill-known value $Z$ defined by the convex combination of the two ill-known constraints $C_1$ and $C_2$.}
+\fcaption{\label{fig:example-ill-known-by-const}Ill-known values $X$ and $Y$.
+The grey area represents the ill-known value $Z$ defined by the convex combination of the two ill-known constraints $C_1$ and $C_2$.}
 \vspace*{13pt}

### 2012-05-24 ijcis-pos-model/sources/temporalModel.tex
last update

@@
 \begin{itemize}
-\item [\textbf{Strictly consistent}] It is not possible for two PVPs of the same object to be valid at the same time.
+\item 
+\begin{definition}
+\textbf{Strictly consistent} It is not possible for two PVPs of the same object to be valid at the same time.
 In other words, if $A$ is an object, $R$ is a relation and $\pi_{I}\left(t \right)$ and $\pi_{J}\left(t \right)$ are the possibilities for a time point $t \in \T$ to be in the PVPs $I$ and $J$ respectively,
@@
 \end{align}
+\end{definition}
 
-\item [\textbf{Ill-consistent}] The overlapping of two PVPs of the same object do overlap. It can be distinghised among the following sub-types:
+\item 
+\begin{definition}
+ \textbf{Ill-consistent} The overlapping of two PVPs of the same object do overlap. It can be distinghised among the following sub-types:
 	\begin{itemize}
-	\item [\emph{Co-existence}] Two versions of the same object $A$ may exist at the same time.
+	\item 
+	      \begin{definition}
+	      \emph{Co-existence} Two versions of the same object $A$ may exist at the same time.
 		      \begin{align}
@@
 		      \end{align}
-	\item [\emph{Weak-consistence}] Several versions of the same object $A$ may exist at the same time.
+	      \end{definition}
+
+	\item 
+	      \begin{definition}
+	      \emph{Weak-consistence} Several versions of the same object $A$ may exist at the same time.
 		      \begin{align}
@@
 		      \end{align}
+	      \end{definition}
+
 	\end{itemize}
+\end{definition}
+
 \end{itemize}

### 2012-05-24 ijcis-pos-model/sources/temporaldatabases.tex
last update

@@
 	\item
-	\textbf{Transaction time} (\emph{TT}) \cite{Rowe1987},\cite{Jensen1991} denotes the time when the fact (object) is stored in the database.
-It is usually append-only: as the past can not be changed, TT can not be changed neither.
+	\textbf{Transaction time} (\emph{TT}) \cite{Rowe1987,Jensen1991} denotes the time when the fact (object) is stored in the database.
+It is usually append-only: as the past cannot be changed, TT cannot be changed either.
 Furthermore, at the moment of insertion, a TT can be neither in the past nor in the future.
 	\item
-	\textbf{Valid time} (\emph{VT}) \cite{Jensen1994},\cite{Sarda1990},\cite{McKenzie1981} denotes the time when the fact (object) is true in the modelled reality.
+	\textbf{Valid time} (\emph{VT}) \cite{Jensen1994,Sarda1990,McKenzie1981} denotes the time when the fact (object) is true in the modelled reality.
 A fuzzy extension has been proposed by \cite{Garrido2009}.
@@
 	\item
-	\textbf{Decision time} (\emph{DT}) \cite{Nascimento1995},\cite{Chakravarthy1994},\cite{Etzion1992},\cite{Ozsoyoglu1995} denotes the time when an event was decided to happen. 
+	\textbf{Decision time} (\emph{DT}) \cite{Nascimento1995,Chakravarthy1994,Etzion1992,Ozsoyoglu1995} denotes the time when an event was decided to happen. 
 	\end{itemize}
@@
 Representing imprecision and its semantics when dealing with time has been studied for a long time.
-Several proposals for representing and computing imprecise time indications can be found in \cite{DeCaluwe1997} and \cite{DeTre1997}.
+Several proposals for representing and computing imprecise time indications can be found in \cite{DeCaluwe1997,DeTre1997}.
 Also, the changes between several granularities can be seen as a source of imprecision \cite{Devos1998}.
@@
 \begin{itemize}
-\item \textbf{Uncertainty in the database} denotes the uncertainty that arises when the knowledge about the temporal data in the database is uncertain.
+\item \textbf{Imperfection in the database} denotes the uncertainty that arises when the knowledge about the temporal data in the database is uncertain.
 E.g., a database record shows that \emph{`The car is in the garage around April.'}
@@
 Several proposals for managing uncertain time in a database exist.
-Some proposals work with rough sets \cite{Qiang2009}, other proposals rely on possibility distributions for representing uncertainty in time \cite{Dyreson1998},\cite{Garrido2009}, \cite{Galindo2001}.
-In order to compare temporal possibility distributions, extensions of the classical Allen's operators \cite{Allen1983} are defined in \cite{Ohlbach2004}, \cite{Nagypal2003},\cite{Dubois2003a} and \cite{Schockaert2008}
+Some proposals work with rough sets \cite{Qiang2009}, other proposals rely on possibility distributions for representing uncertainty in time \cite{Dyreson1998,Garrido2009,Galindo2001}.
+In order to compare temporal possibility distributions, extensions of the classical Allen's operators \cite{Allen1983} are defined in \cite{Ohlbach2004,Nagypal2003,Dubois2003a,Schockaert2008}.
 %In the proposal section, we will follow the representation by means of possibility distributions, in order to work with both satisfaction and dissatisfaction degrees.
@@
 \begin{definition}
-A \textbf{Fuzzy Validity Period} (\emph{FVP}) is defined as a fuzzy time interval specifying when an object is valid. A fuzzy time interval is then the fuzzification of a crisp time interval.
+A \textbf{Fuzzy Validity Period} (\emph{FVP}) is defined as a fuzzy time interval specifying when the data regarding an object is valid.
+A fuzzy time interval is then the fuzzification of a crisp time interval.
 \end{definition}
@@
 \begin{definition}
-A \textbf{Possibilistic Valid-time Period} (\emph{PVP}) is an ill-known interval in time specifying when an object is valid.
+A \textbf{Possibilistic Valid-time Period} (\emph{PVP}) is an ill-known interval in time specifying when the data regarding an object is valid.
 \end{definition}
@@
 
-Analogously, it is defined the function to close a left-open interval: 
-\begin{definition}
-\label{def:close-a-crisp-interval-l}
-CloseL$\left(I, J\right)$ 
-\begin{align}
-\mbox{CloseR} \left( I, J \right) &=& \\ 
-\begin{cases}
-\nonumber
-I & \mbox{ if } a \neq -\infty \\
-I=\left]d, b \right] & \mbox{ if } a = -\infty \wedge J < I
-\end{cases}
-\end{align}
-\end{definition}
+%Analogously, it is defined the function to close a left-open interval: 
+%\begin{definition}
+%\label{def:close-a-crisp-interval-l}
+%CloseL$\left(I, J\right)$ 
+%\begin{align}
+%\mbox{CloseL} \left( I, J \right) &=& \\ 
+%\begin{cases}
+%\nonumber
+%I & \mbox{ if } a \neq -\infty \\
+%I=\left]d, b \right] & \mbox{ if } a = -\infty \wedge J < I
+%\end{cases}
+%\end{align}
+%\end{definition}
 
@@
 
-\begin{definition}
-\label{def:first-in-relation}
-Analogously, the function First$\left(R, A \right)$ obtains the tuple $\left(A, I \right)$ where $I$ is a right-open interval and which is the first version in the relation $R$, that is it, $I = \left[-\infty, b \right]$.
-\begin{align}
-\label{eq:first-in-relation}
-\mbox{First} \left(R, A \right) &=& \\ 
-\begin{cases}
-\nonumber
-I & \mbox{ if } \exists b \in I : (A,I) \in R \wedge a = -\infty \\
-\emptyset & \mbox{ in any other case }
-\end{cases}
-\end{align}
-
-\end{definition}
+%\begin{definition}
+%\label{def:first-in-relation}
+%Analogously, the function First$\left(R, A \right)$ obtains the tuple $\left(A, I \right)$ where $I$ is a right-open interval and which is the first version in the relation $R$, that is it, $I = \left[-\infty, b \right]$.
+%\begin{align}
+%\label{eq:first-in-relation}
+%\mbox{First} \left(R, A \right) &=& \\ 
+%\begin{cases}
+%\nonumber
+%I & \mbox{ if } \exists b \in I : (A,I) \in R \wedge a = -\infty \\
+%\emptyset & \mbox{ in any other case }
+%\end{cases}
+%\end{align}
+%
+%\end{definition}
 
@@
 
-It is also possible to append at the beggining by using \eqref{def:close-a-crisp-interval-l} and \eqref{def:first-in-relation}.
-\begin{definition}
-\label{def:append-first-version}
-Consider an object $A$, the relation $R$ and a time interval $J$.
+%It is also possible to append at the beggining by using \eqref{def:close-a-crisp-interval-l} and \eqref{def:first-in-relation}.
+%\begin{definition}
+%\label{def:append-first-version}
+%Consider an object $A$, the relation $R$ and a time interval $J$.
 Then, the function append-first$\left(R,A,J \right)$ closes any right open version of the object $A$ if it exists and add the new version $\left(A, J \right)$.
+%
+%\begin{eqnarray}
+%%\label{eq:close-current}
+%\text{Append-first} \left(R, A, J \right) =\\
+%\begin{cases}
+%\nonumber
+%\big \lbrace R - \left(A, I \right) \cup \left \lbrace \left(A, \mbox{ CloseL } \left(I, J\right) \right) \cup \left(A, J\right)\right \rbrace  \big \rbrace \\
+%\nonumber
+%\mbox{ if } I = \mbox{ First } \left(R, A \right) \wedge J < \mbox{ CloseL } \left(I, J \right)   \\
+%\nonumber R , \text{ in any other case}
+%\end{cases}
+%\end{eqnarray}
+%\end{definition}
 
-\begin{eqnarray}
-%\label{eq:close-current}
-\text{Append-first} \left(R, A, J \right) =\\
-\begin{cases}
-\nonumber
-\big \lbrace R - \left(A, I \right) \cup \left \lbrace \left(A, \mbox{ CloseL } \left(I, J\right) \right) \cup \left(A, J\right)\right \rbrace  \big \rbrace \\
-\nonumber
-\mbox{ if } I = \mbox{ First } \left(R, A \right) \wedge J < \mbox{ CloseL } \left(I, J \right)   \\
-\nonumber R , \text{ in any other case}
-\end{cases}
-\end{eqnarray}
-\end{definition}
-
-\subsubsection{Insert}
+\subsubsection{\label{subsubsec:insert}Insert}
 The user wants to store an object $A$ which is valid in the relation $R$ during the time interval specified by $I = \left(a, b \right)$.

### 2012-05-24 trunk/ijcis-pos-model/sources/temporalModel.tex
some comments from olga and nico

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%
-%\vglue13pt
-%%\begin{table}[htbp]
-%\tcap{Valid-Time interval data types}
-%\centerline{\small VALID TIME DATA TYPES}
-%\vglue-6pt
-%\centerline{\small\baselineskip=13pt
-%\begin{tabular}{c p{2cm} p{2cm}}\\
-%\hline
-%Type for S & Type for E & Semantics \\
-%\hline
-%Unknown & Unkown & Unknown interval. \\
-%\hline
-%1,2 & 1,2 & A possibilistic time interval. \\
-%\hline\\
-%\end{tabular}
-%\label{tbl:valid-time-interval}}
-%%\end{table}
-%
-%The following is a study of the constraints in the range of values for both starting and ending times to represent a valid time interval.
-%
-%The time period inside a valid time interval is denoted by two constraints:
-%\begin{eqnarray}
-%C_1\stackrel{\triangle}{=}\left(\geq,X\right)\\
-%C_2\stackrel{\triangle}{=}\left(\leq,Y\right).
-%\end{eqnarray}
-%Applying the inference of uncertainty as proposed in our general reasoning, we find for the first constraint that:
-%\begin{eqnarray}
-%\Pos\left(C_1([a,b])\right) &=& \min_{r\in[a,b]}\Big(\sup_{r\leq w}\pi_{X}(w)\Big)\\
-%\Nec\left(C_1([a,b])\right) &=& \min_{r\in[a,b]}\Big(\inf_{r>w}1-\pi_{X}(w)\Big)
-%\end{eqnarray}
-%which can be simplified to:
-%\begin{eqnarray}
-%\Pos\left(C_1([a,b])\right) &=& \sup_{a\leq w}\pi_{X}(w)\\
-%\Nec\left(C_1([a,b])\right) &=& \inf_{a>w}1-\pi_{X}(w).
-%\end{eqnarray}
-%For the second constraint, we find that:
-%\begin{eqnarray}
-%\Pos\left(C_2([a,b])\right) &=& \min_{r\in[a,b]}\Big(\sup_{r\geq w}\pi_{Y}(w)\Big)\\
-%\Nec\left(C_2([a,b])\right) &=& \min_{r\in[a,b]}\Big(\inf_{r<w}1-\pi_{Y}(w)\Big)
-%\end{eqnarray}
-%which can be simplified to:
-%\begin{eqnarray}
-%\Pos\left(C_2([a,b])\right) &=& \sup_{b\geq w}\pi_{Y}(w)\\
-%\Nec\left(C_2([a,b])\right) &=& \inf_{b<w}1-\pi_{Y}(w).
-%\end{eqnarray}
-%The uncertainty about the inclusion of an interval $I=[a,b]$ in the interval with ill-known boundaries can now be found by evaluating $[a,b]$ against the evaluation function $\lambda$ with $\bool=\wedge$.
-Application of \eqref{eq:conjunctive1} and \eqref{eq:conjunctive2} leads to:
-%\begin{eqnarray}
-%\Pos\left(\lambda([a,b])\right)&=&\min\bigg(\Pos(C_1([a,b])),\Pos(C_2([a,b]))\bigg)\\
-%\Nec\left(\lambda([a,b])\right)&=&\min\bigg(\Nec(C_1([a,b])),\Nec(C_2([a,b]))\bigg).
-%\end{eqnarray}
-%These last expressions can also be expanded as:
-%\begin{eqnarray}
-%\label{eq:interval-pos}
-%\Pos\left(\lambda([a,b])\right)&=&\min\bigg(\sup_{a\leq w}\pi_{X}(w),\sup_{b\geq w}\pi_{Y}(w)\bigg)\\
-%\label{eq:interval-nec}
-%\Nec\left(\lambda([a,b])\right)&=&\min\bigg(\inf_{a>w}1-\pi_{X}(w),\inf_{b<w}1-\pi_{Y}(w)\bigg).
-%\end{eqnarray}
-%Note that the interval $[X,Y]$ used here, is certainly not a fuzzy interval.
-Instead, we are dealing with an ill-known interval, i.e.
-it is a crisp interval, but it is partially unknown which values are in this interval.
-The uncertainty stems from the fact that the interval boundaries are ill-known.
-%
-%Within this framework for the reasoning of the knowledge about the interval, we are going to study the semantics from the point of view of the valid time as well as from the point of view of the temporal database.
-%
-%\subsubsection{S = Unknown, E = Unknown}
-%Consider that the possibility distribution for S is Unknown and for E is also Unknown. The possibility distributions are $\pi_S$ and $\pi_E$ respectively:
-%
-%\begin{eqnarray}
-%\pi_{S}(t \in \T) = 1\\
-%\pi_{E}(t \in \T) = 1
-%\end{eqnarray}
-%
-%Therefore the constraints $C_1$ and $C_2$ are:
-%
-%\begin{eqnarray}
-%\Pos\left(C_1([a,b])\right) &=& \sup_{a\leq w}\pi_{S}(w) = 1\\
-%\Nec\left(C_1([a,b])\right) &=& \inf_{a>w}1-\pi_{S}(w) = 0\\
-%\Pos\left(C_2([a,b])\right) &=& \sup_{b\geq w}\pi_{E}(w)=1\\
-%\Nec\left(C_2([a,b])\right) &=& \inf_{b<w}1-\pi_{E}(w)=0
-%\end{eqnarray}
-%
-%
-%\begin{eqnarray}
-%\Pos\left(\lambda([a,b])\right)&=&\min\bigg(\sup_{a\leq w}\pi_{S}(w),\sup_{b\geq w}\pi_{E}(w)\bigg) \\
-%&=& \min(1,1)  = 1\\
-%\Nec\left(\lambda([a,b])\right)&=&\min\bigg(\inf_{a>w}1-\pi_{S}(w),\inf_{b<w}1-\pi_{E}(w)\bigg)\\
-%&=& \min(0,0) = 0
-%\end{eqnarray}
-%
-%\begin{example}
-%A constant object in the database. It is an object that is always valid in the database.
-%The main issue in a valid time database is that the update sentence have to modify the starting or ending point distribution.
-The most natural way to convert a snapshot database to a temporal database is to consider the objects belonging to the interval [Unknown, Unknown].
-%\end{example}
-%
-%\subsubsection{[Unknown,value] or [value,Unknown]}
-%These two cases are managed in this section. Here we will note $\pi_{S}(t)$ or $\pi_{E}(t)$ as the possibility distribution for 'value' in the starting or ending points.
-%
-%First of all the case [Unknown,value] in \eqref{eq:interval-pos} and in \eqref{eq:interval-nec}:
-%
-%\begin{eqnarray}
-%\Pos\left(\lambda([a,b])\right)&=&\min (1,\sup_{b\geq w}\pi_{E}(w))\\
-% &=& \sup_{b\geq w}\pi_{E}(w)\\
-%\Nec\left(\lambda([a,b])\right)&=&\min (0,\inf_{b<w}1-\pi_{E}(w))\\
-%&=& 0
-%\end{eqnarray}
-%
-%Similarily for [value,Unknown] the possibility and neccesity are:
-%\begin{eqnarray}
-%\Pos\left(\lambda([a,b])\right)&=&\min (\sup_{a\leq w}\pi_{S}(w),1)\\
-% &=& \sup_{a\leq w}\pi_{S}(w)\\
-%\Nec\left(\lambda([a,b])\right)&=&\min (\inf_{a>w}1-\pi_{S}(w),0)\\
-%&=& 0
-%\end{eqnarray}
-%
-%The semantics for an object with valid time given by [Unknown,value] is an object that has been valid from the beggining to an ending time.
-Analogously, the semantics for an object with a valid time given by the interval [value,Unknown] is an object that it is still valid in the database.
-%
-%
-%\subsubsection{[Undefined,Undefined]}
-%Maybe this combination has no sense for a valid time object. The possibility distributions for both starting and ending points are:
-%
-%\begin{eqnarray}
-%\pi_{S}(t \in \T) = 0\\
-%\pi_{E}(t \in \T) = 0
-%\end{eqnarray}
-%
-%Therefore the constraints $C_1$ and $C_2$ are:
-%
-%\begin{eqnarray}
-%\Pos\left(C_1([a,b])\right) &=& \sup_{a\leq w}\pi_{S}(w) = 0\\
-%\Nec\left(C_1([a,b])\right) &=& \inf_{a>w}1-\pi_{S}(w) = 1\\
-%\Pos\left(C_2([a,b])\right) &=& \sup_{b\geq w}\pi_{E}(w)=0\\
-%\Nec\left(C_2([a,b])\right) &=& \inf_{b<w}1-\pi_{E}(w)=1
-%\end{eqnarray}
-%
-%Taking these values into  \eqref{eq:interval-pos} and in \eqref{eq:interval-nec}:
-%\begin{eqnarray}
-%\Pos\left(\lambda([a,b])\right)&=&\min (0,0)\\
-% &=& 0\\
-%\Nec\left(\lambda([a,b])\right)&=&\min (1,1)\\
-%&=& 1
-%\end{eqnarray}
-%
-%\textcolor{red}{TAKE A LOOK INTO THIS}
-%
-%Which is an inconsistency.
-%
-%\subsection{[Undefined,value] or [value,Undefined]}
-%
-%The possibility and neccesity measures for [Undefined,value] are :
-%\begin{eqnarray}
-%\Pos\left(\lambda([a,b])\right)&=&\min (0,\sup_{b\geq w}\pi_{E}(w))\\
-% &=& 0\\
-%\Nec\left(\lambda([a,b])\right)&=&\min (1,\inf_{b<w}1-\pi_{E}(w))\\
-%&=& \inf_{b<w}1-\pi_{E}(w)
-%\end{eqnarray}
-%
-%The other way around, [value,Undefined]:
-%\begin{eqnarray}
-%\Pos\left(\lambda([a,b])\right)&=&\min (\sup_{a\leq w}\pi_{S}(w),0)\\
-% &=& 0\\
-%\Nec\left(\lambda([a,b])\right)&=&\min (\inf_{a>w}1-\pi_{S}(w),1)\\
-%&=& \inf_{a>w}1-\pi_{S}(w)
-%\end{eqnarray}
-%
-%\textcolor{red}{TAKE A LOOK INTO THIS}
-%This is also an inconsistency.
-%
-%\subsubsection{[NULL,NULL] and maybe any combination with NULL}
-%I think the framework we have can not deal with NULL computations. Maybe just for non-specified time or something like that.
-%
-%\subsubsection{[Undefined,Unknown] or [Unknown,Undefined]}
-%The semantic for this could be also unconsistent.
-%
-%
-%\begin{example}
-%Consider a snapshot database (table \ref{tbl:snapshot-car-models}) that stores information about car models. The primary key is given by the set $\lbrace$ Brand, Model $\rbrace$.
-%
-%\vglue13pt
-%%\begin{table}[htbp]
-%\tcap{Car model snapshot database}
-%\centerline{\small CAR MODELS}
-%\vglue-6pt
-%\centerline{\small\baselineskip=13pt
-%\begin{tabular}{c c c}\\
-%\hline
-%\textbf{Brand} & \textbf{Model} & Segment \\
-%\hline
-%Peugeot & 205 & B \\
-%Peugeot & 206 & B \\
-%Peugeot & 207 & B \\
-%Peugeto & 208 & B \\
-%Citroen & C2 & B \\
-%\hline\\
-%\end{tabular}
-%\label{tbl:snapshot-car-models}
-%}
-%
-%Consider now that the user wants to make that table a valid-time table, then, from the point of view of the storage, the following changes should be done:
-%
-%\begin{itemize}
-%\item
-%A new definition for the primary key, including a version field.
-%\item
-%The addition of two new columns to store the valid-time interval. By default the initial values are Unknown for both starting and ending points of the interval.
-%\end{itemize}
-%Therefore, the table \ref{tbl:snapshot-car-models} is transformed into table \ref{tbl:initial-valid-time-car-models}:
-%
-% \vglue13pt
-%%\begin{table}[htbp]
-%\tcap{Car model valid-time database. Note that U is the abbreviation for the Unknown constant.}
-%\centerline{\small VALID TIME CAR MODELS}
-%\vglue-6pt
-%\centerline{\small\baselineskip=13pt
-%\begin{tabular}{c c c c c c}\\
-%\hline
-%\textbf{Brand} & \textbf{Model} & \textbf{VersionID} & Segment & S & E \\
-%\hline
-%Peugeot & 205 & 001 & B & U & U \\
-%Peugeot & 206 & 001 & B & U & U \\
-%Peugeot & 207 & 001 & B & U & U \\
-%Peugeto & 208 & 001 & B & U & U \\
-%Citroen & C2 & 001 & B & U & U \\
-%\hline\\
-%\end{tabular}
-%\label{tbl:initial-valid-time-car-models}
-%}
-%
-%
-%
-%\end{example}
 
@@
 \end{definition}
-The datatypes for $D_G$ are shown in table \ref{tablaaaaaaaaa}
-\textcolor{red}{poner tabla}
+The datatypes for $D_G$ are shown in table \ref{tbl:gefred-data-types}.
+
+\begin{definition}
+\label{def:typeof-domain}
+The function typeof$\left(D \right)$ returns the datatype associated with the domain $D$.
+
+\end{definition}
+
+\vglue13pt
+%\begin{table}[htbp]
+\tcap{\label{tbl:gefred-data-types}Data types}
+%\centerline{\small TITLE}
+\vglue-6pt
+\centerline{\small\baselineskip=13pt
+\begin{tabular}{c p{4cm} }\\
+No. & Datatype \\ \hline
+1 & A single scalar. \\
+2 & A single number. \\
+3 & A set of mutually exclusive possible scalar assignations. \\
+4 & A set of mutually exclusive possible numeric assignations. \\
+5 & A possibility distribution in a scalar domain. \\
+6 & A possibility distribution in a numeric domain. \\
+7 & A real number in $\left[0, 1 \right]$ referring to degree of matching. \\
+8 & An \emph{UNKNOWN} value. \\
+9 &  An \emph{UNDEFINED} value. \\
+10 & A \emph{NULL} value. \\
+\hline\\
+\end{tabular}
+}
 
@@
 \nonumber
-\forall i, i' \in \left \lbrace 1, \ldots, m\right \rbrace , \exists s \in S: \left(A_{Gs}:d_{is} \right) \neq \left(A_{Gs}:d_{i's} \right)
+\forall i, i' \in \left \lbrace 1, \ldots, m\right \rbrace , \exists s \in S: \\
+\nonumber
+\left(A_{Gs}:d_{is} \right) \neq \left(A_{Gs}:d_{i's} \right)
 \end{align}
@@
 
-\texcolor{red}{primary temporal key}
-
-\begin{definiton}
+\begin{definition}
 \label{def:generalized-fuzzy-temporal-key}
-A generalized fuzzy temporal key, $K_{GT}$ is a subset od the head:
+A generalized fuzzy temporal key, $K_{GT}$ is a subset of the head. An attribute, namely \emph{version} $V$ is added when dealing with valid-time.
+\begin{align}
+\label{eq:generalized-fuzzy-temporal-key}
+K_{GT} \subseteq \Head, K_{GT} = \left \lbrace  \left(A_{Gs}:D_{Gs} \right) \right.  \\
+\nonumber
+ \left. \cup  \left(V_{ID}:D_{ID} \right)	\right \rbrace \\
+\nonumber
+\text{Typeof }\left(D_{ID} \right) = \\
+\nonumber
+s\in S \subseteq \left(1, \ldots, n \right) \\
+\nonumber
+\forall s \in S, \text{Typeof } \left(D_{Gs} \right) \in \left \lbrace 1, 2 \right \rbrace \\
+\nonumber
+\forall i, i' \in \left \lbrace 1, \ldots, m\right \rbrace , \exists s \in S: \\
+\nonumber
+\left(A_{Gs}:d_{is} \right) \neq \left(A_{Gs}:d_{i's} \right)
+\end{align}
 
 \end{definition}
+
+\begin{example}
+
+\end{example}
+
+\subsection{\label{subsec:data-manipulation}Data manipulation language}
+
+The algebra for the data manipulation is defined in \cite{Medina1994}.
+The Generalized Fuzzy Relational Algebra manipulates relations like $R_{FG}$.
+The operations defined are: \emph{Union, Intersection, Difference, Cartesian Product, Projection, Join} and \emph{Selection}.
+Thus, in this section we will describe and implement the following operations for temporal databases, as described in \ref{subsubsec:Understanding-valid-time-databases}.
+The operations implemented are: \emph{Insert, Modify, Delete, } and \emph{Revise}.
+The semantics of the operations will be the same that those defined for a crisp temporal database, whereas the temporal representation is made by the possibilistic valid-time period and the ill-known constraints (see sections \ref{sec:prelim} and \ref{sec:time-rep}.
+
+It is important to notice that whereas the result of the evaluation of any comparison between crisp time intervals is boolean, the evaluation of any comparison between \emph{PVPs} is in the unit interval.
+Therefore, for a valid-time object, say $A \in R$, in a crisp temporal database it is not possible an overlaping among any valid-time interval ($\forall I, \not \exists J: \left(A, I \right) \in R \wedge \left(A, J \right) \in R \wedge \not \Overlap\left(I, J \right)$).
+Thus, from the point of view of a fuzzy temporal database, there are the following options:
+
+\begin{itemize}
+\item 
+\begin{definition}
+\textbf{Strictly consistent} It is not possible for two PVPs of the same object to be valid at the same time.
+In other words, if $A$ is an object, $R$ is a relation and $\pi_{I}\left(t \right)$ and $\pi_{J}\left(t \right)$ are the possibilities for a time point $t \in \T$ to be in the PVPs $I$ and $J$ respectively,
+\begin{align}
+\label{eq:stricly-consistent}
+\forall I,J, \forall t \in \T :\\
+%\nonumber
+%\left(A, I \right) \in R \wedge \\
+%\nonumber
+%\left(A, J \right) \in R : \\
+\nonumber
+\pi_{I}(t) + \pi_{J}(t) \leq 1
+\end{align}
+\end{definition}
+
+\item 
+\begin{definition}
+ \textbf{Ill-consistent} The overlapping of two PVPs of the same object do overlap. It can be distinghised among the following sub-types:
+	\begin{itemize}
+	\item 
+	      \begin{definition}
+	      \emph{Co-existence} Two versions of the same object $A$ may exist at the same time.
+		      \begin{align}
+		      \label{eq:co-existence}
+		      \forall I, \exists ! J, \exists t \in \T : \\
+		      \nonumber
+		      \pi_{I}(t) + \pi_{J}(t) \geq 1
+		      \end{align}
+	      \end{definition}
+
+	\item 
+	      \begin{definition}
+	      \emph{Weak-consistence} Several versions of the same object $A$ may exist at the same time.
+		      \begin{align}
+		      \label{eq:weak-consistence}
+		      \forall I, \exists J, \exists t \in \T : \\
+		      \nonumber
+		      \pi_{I}(t) + \pi_{J}(t) \geq 1
+		      \end{align}
+	      \end{definition}
+
+	\end{itemize}
+\end{definition}
+
+\end{itemize}
+
+In the implementation of the DML operations we will consider a strictly consistent temporal database.
+Since the time intervals are now possibilistic valid-time periods, PVPs, the auxiliary functions defined in equations \eqref{eq:close-a-crisp-interval} to \eqref{eq:close-current} should be re-defined.
+
+\begin{definition}
+\label{def:close-r-a-pvp}
+Consider two PVPs given by $I = \left(X, Y \right)$ and $J = \left(X', Y' \right)$.
+The CloseR function closes the PVP given by $I$ with a convex combination of ill-known constraints (see section \ref{subsec:interval-evaluation-by-ill-known-constraints}):
+\begin{align}
+\label{eq:close-r-a-pvp}
+\text{CloseR}\left(I, J\right) = \\
+\begin{cases}
+\nonumber
+I   & \mbox{ if } Y \not \in \text{UNKNOWN}  \\
+I = \left[X, Z \right] & \mbox{ if } Y \in \text{UNKNOWN}, \\
+& Z \triangleq \left \lbrace C_1\left(>, X \right), C_2\left(<, X' \right) \right \rbrace
+\end{cases}
+\end{align}
+\end{definition}
+
+\begin{definition}
+\label{def:pvp-current-in-relation}
+Consider an object $A$ and a relation $R$ in a fuzzy temporal database.
+The function Current$\left(R, A \right)$ obtains the tuple $\left(A, I \right), I = \left(X, Y \right)$ that is current in the relation $R$, that is it, $I = \left[X , \text{UNKNOWN} \right]$.
+
+\begin{align}
+\label{eq:pvp-current-in-relation}
+\mbox{Current} \left(R, A \right) &=& \\ 
+\begin{cases}
+\nonumber
+I & \mbox{ if } \exists  \in I : \\
+&  (A,I) \in R \wedge Y \in \text{UNKNOWN}\\
+\emptyset & \mbox{ in any other case }
+\end{cases}
+\end{align}
+\end{definition}
+
+Now it is possible to close the current version of an object by using \eqref{eq:close-r-a-pvp} and \eqref{eq:pvp-current-in-relation}.
+
+\begin{definition}
+\label{def:pvp-close-current-version}
+Consider an object $A$, the relation $R$ and a pvp, $J$.
+Then, the function close - current$\left(R, A, J \right)$ closes any current version of the object $A$ if it exists and add the new version $\left(A, J \right)$.
+
+\begin{eqnarray}
+\label{eq:pvp-close-current}
+\text{Close-current} \left(R, A, J \right) =\\
+\begin{cases}
+\nonumber
+\big \lbrace R - \left(A, I \right) \cup \left \lbrace \left(A, \mbox{ CloseR } \left(I, J\right) \right) \cup \left(A, J\right)\right \rbrace  \big \rbrace \\
+\nonumber
+\mbox{ if } I = \mbox{ Current } \left(R, A \right) \wedge J > \mbox{ CloseR } \left(I, J \right)   \\
+\nonumber R , \text{ in any other case}
+\end{cases}
+\end{eqnarray}
+\end{definition}
+
+\subsubsection{\label{subsubsec:insert-fuzzy-temporal}Insert}
+The user wants to store an object $A$ which is valid in the relation $R$ during the time interval specified by the PVP, $I = \left(X, Y \right)$.
+%
+%
+%The interpretation of the insert operation is the following: The user wants to store in the database that an object $A$  is true for some period(s) of time $I$ in the given relation $R$.
+To indicate that the object $A$ is current in the relation, the special value Until Current, $\UC$  is used.
+There are two main cases when performing an insert operation:
+\begin{enumerate}
+\item The object $A$ was never in the relation $R$: The object is added with the valid-time indicated by the PVP $I$.
+
+\item The object $A$ is in the relation $R$. Depending on the value of the time interval $I$, there are two possibilities:
+	\begin{enumerate}
+	\item Insert $\left(A, I\right)$ in the relation $R$. If the time interval $I$ does not overlap any other valid-time interval in the relation $R$.
+	\item Reject the insertion, if the time interval $I$ do overlap any existing valid-time interval for the object $A$ in the relation $R$.
+	\end{enumerate}
+
+\end{enumerate}
+
+The algorithm for the implementation of the insert operation is the following:
+
+\begin{align}
+\label{eq:insert-fuzzy-temporal}
+\INS \left(R ,A, I\right) =\\
+\begin{cases}
+\nonumber
+R \cup \left(A, I \right), & \mbox{ if }  A \not \in R\\
+R \cup   \left(A, I \right) & \mbox{ if }  A \in R \wedge \\
+\big \lbrace \left( \exists J= \left(X', Y' \right): \left(A, J \right) \in R \wedge \right. \\
+\left. Y' \in \left \lbrace \left(<, X \right)\right \rbrace  \right) \vee \\
+\left( \exists K= \left(X'', Y'' \right): \left(A, K \right) \in R \wedge \right. \\
+\left. X'' \in \left \lbrace \left(>, Y \right)\right \rbrace \vee \right) \big \rbrace\\
+%&  J<I<K  \big \rbrace\\
+R, & \mbox{ otherwise }  
+\end{cases} 	
+\end{align}
+\subsubsection{\label{subsubsec:modify-fuzzy-temporal}Modify}
+The delete operation logically removes a current object $A$ which is valid in the relation $R$:
+\begin{align}
+\label{eq:delete-fuzzy-temporal}
+\DEL \left(R ,A \right) =\\
+\begin{cases}
+\nonumber
+R - \left(A, I \right), & \mbox{ if } \left( A, I \right) \in R\\
+R, & \mbox{ otherwise }  
+\end{cases} 	
+\end{align}
+
+\subsubsection{\label{subsubsec:delete-fuzzy-temporal}Delete}
+This operation adds new information about an existing object $A$ in the relation $R$. The modify operation does not remove any previous value of the object $A$.
+Note that the modify operation is only applicable when the object $A$ is current in the relation $R$; $ \left(A, I \right) \in R, I = \left[X, \UC \right]$.
+
+\begin{align}
+\label{eq:modify-fuzzy-temporal}
+\MOD \left( R, A, I \right) =\\
+\begin{cases}
+\nonumber
+\mbox{ Close-current }\left(R, A, \mbox{Current}\left(R, A \right) \right) \\
+ \mbox{ if Current}\left(R ,A \right) \neq \emptyset \\
+ R,  \mbox{ otherwise }
+\end{cases}
+\end{align}
+
+\subsubsection{\label{subsubsec:revise-fuzzy-temporal}Revise}
+This operation changes the value of the object $A$ in the relation $R$, at a given time interval $I$. This operation is defined when a correction on the values for the object $A$ should be done.
+\begin{align}
+\label{eq:revise-fuzzy-temporal}
+\mbox{ Revise}\left(R, A, I \right)=\\
+\begin{cases}
+\nonumber
+ R - \left(A', I \right) \cup \left(A, I \right)  \\
+  \mbox{ if } \left(A', I \right) \in R: A = \left(a_1, \ldots, a_n \right),\\
+A'= \left(a_1', \ldots, a_n' \right)   \\   
+R,  \mbox{ otherwise }
+\end{cases}
+\end{align}

### 2012-05-24 ijcis-pos-model/sources/timeRepresentation.tex
some comments from olga and nico

@@
 \subsection{\label{subsec:ill-known-point-rep}Ill-known time point}
-An ill-known time point $X$ is a precise time point that for some reason, it is not fully specified. Note that $X$ has only one possible value but that value is unspecified.
+An ill-known time point $X$ is a precise time point that, for some reason, it is not fully specified. Note that $X$ has only one possible value but that value is unspecified.
 
@@
 
+It is also possible to specify an ill-known time point by a convex combination of ill-known constraints, (Definition \ref{def:convex-combination-ill-known-constraints}) as shown by equation \eqref{eq:ill-known-value-def-by-const-unc}.
+
 \begin{definition}
@@
 \emph{Domain for an ill-known time point}\\
-Consider $\Pow(\T)$ the set of all the possibility distributions over $\T$, and the three fuzzy constants \emph{UNKNOWN} $=\left\lbrace 1/t, t \in \T \right\rbrace$, \emph{UNDEFINED} $=\left\lbrace 0/t, t \in \T \right\rbrace$ and \emph{NULL} $=\lbrace 1/$ \emph{UNKNOWN}, $1/$ \emph{UNDEFINED} $\rbrace$.
+Consider $\Pow(\T)$ the set of all the possibility distributions over $\T$, and the three fuzzy constants \emph{UNKNOWN} $=\left\lbrace 1/t, \forall t \in \T \right\rbrace$, \emph{UNDEFINED} $=\left\lbrace 0/t,\forall t \in \T \right\rbrace$ and \emph{NULL} $=\lbrace 1/$ \emph{UNKNOWN}, $1/$ \emph{UNDEFINED} $\rbrace$.
 The domain for an ill-known time point $X$ is given by:
@@
 \end{definition}
-
-It is also possible to specify an ill-known time point by just a set of ill-known constraints plus
 
@@
 2 & A possibility distribution in the numeric domain & A fuzzy number or a fuzzy interval.\\
-3 & An unknown value & \textbf{UNKNOWN}$= \lbrace1/t, t \in \T \rbrace$\\
-4 & An undefined value & \textbf{UNDEFINED}$= \lbrace 0/t, t \in \T \rbrace$\\
+3 & An unknown value & \textbf{UNKNOWN}$= \lbrace1/t,\forall t \in \T \rbrace$\\
+4 & An undefined value & \textbf{UNDEFINED}$= \lbrace 0/t,\forall t \in \T \rbrace$\\
 5 & A null value & \textbf{NULL} $=\lbrace 1/$Unknown, 1/Undefined $\rbrace$\\
@@
 \emph{Ill-known time interval}\\
-An ill-known time interval denoted by $\left[X, Y\right]$ is a precise time interval that for some reason, its boundaries are not precisely known.
+An ill-known time interval denoted by $\left[X, Y\right]$ is a precise time interval whose boundaries are not precisely known.
 The ill-known interval is defined by two ill-known points, namely $X$,$Y$.
@@
 \begin{eqnarray}
-C_1\stackrel{\triangle}{=}\left(\geq,X\right) \wedge \\
+C_1\stackrel{\triangle}{=}\left(\geq,X\right)  \\
 C_2\stackrel{\triangle}{=}\left(\leq,Y\right)
 \end{eqnarray}
-Note that the boolean operator applied to the constraints is $\bool=\wedge$.
+Note that the boolean operator applied to the constraints is $\bool=\wedge$. Hence, the construct of both constraints is $C_1 \wedge C_2$
 This means that we want to know if all the points in the interval $I$ are greater or equal to $X$ and, at the same time, smaller or equal to $Y$.
@@
 \item
-\emph{Completely unknown time interval}: Both starting and ending points are unknown, therefore the hole interval is specified by the \emph{UNKNOWN} constant.
-Due to the representation, both endpoints, say $X$ and $Y$ are stored.
+\begin{definition}
+\emph{Completely unknown time interval}: Both starting and ending points are unknown, therefore the whole interval is specified by the \emph{UNKNOWN} constant.
+Due to the representation, both endpoints, say $X$ and $Y$, are stored.
 Thus, the interval is stored as $[$ \emph{UNKNOWN}, \emph{UNKNOWN} $]$.
+\end{definition}
+
 \item
+\begin{definition}
 \emph{Semi-open time interval}: Only one of the two ill-known endpoints for the time interval $\left[X, Y\right]$ is unknown.
@@
 The solution adopted for this problem is explained in the following paragraph.
+\end{definition}
 \end{enumerate}
@@
 
-\vglue13pt
-%\begin{table}[htbp]
-\tcap{Relations for the $\Open(C)$ function.}
-\centerline{\small Relations}
-\vglue-6pt
-\centerline{\small\baselineskip=13pt
-\begin{tabular}{c c c}\\
-\hline
-$R$ & $\Rp$ & $\Rn$\\ \hline
-$<$ & $>$ & $\leq$ \\
-$\leq$ & $\geq$ & $<$ \\
-$>$ & $<$ & $\geq$ \\
-$\geq$ & $\leq$ & $>$ \\
-\hline\\
-\end{tabular}
-\label{tbl:open-pos-nec-rels}
-}
+% \vglue13pt
+% %\begin{table}[htbp]
+% \tcap{Relations for the $\Open(C)$ function.}
+% \centerline{\small Relations}
+% \vglue-6pt
+% \centerline{\small\baselineskip=13pt
+% \begin{tabular}{c c c}\\
+% \hline
+% $R$ & $\Rp$ & $\Rn$\\ \hline
+% $<$ & $>$ & $\leq$ \\
+% $\leq$ & $\geq$ & $<$ \\
+% $>$ & $<$ & $\geq$ \\
+% $\geq$ & $\leq$ & $>$ \\
+% \hline\\
+% \end{tabular}
+% \label{tbl:open-pos-nec-rels}
+% }
 

### 2012-05-27 trunk/ijcis-pos-model/sources/temporaldatabases.tex
modifications on the operations

@@
 \begin{definition}
-Consider a relation $R$ in a temporal database, an object $A$ given by the attributes $\left(a_1, \ldots, a_n \right)$ and the crisp time interval $I$ given by $\left[a,\ b\right] : a,b \in \T, a \leq b$.
-A valid-time object $A$ in the relation $R$ is specified by:
+Consider a relation $R$ in a temporal database, an entity $A$ given by the attributes $\left(a_1, \ldots, a_n \right)$ and the crisp time interval $I$ given by $\left[a,\ b\right] : a,b \in \T, a \leq b$.
+The pair $\left(A, I\right)$ expresses that the data regarding the entity A is valid during the crisp time interval I.
+Note that the relation $R = \left(A, I \right) x \left(A, I \right)$.
+The following equation indicates that the pair $\left(A, I\right)$ is in the relation R:
 \begin{equation}
 \label{eq:rel-def}
-\left( R, A , I \right)
+\left( R, \left(A , I\right) \right)
+\end{equation}
+\end{definition}
+\begin{example}
+ Consider a library database.
+Each physical book is stored with an identifier (ID), a title and the start and end loan dates respectively.
+Table \ref{tbl:library-sample} illustrates the first version of the database, after three insertions.
+In this example, $A = ($ ID, Title $)$ and $I = ($ Start Loan, End Loan $)$.
+\end{example}
+
+\vglue13pt
+%\begin{table}[htbp]
+\tcap{\label{tbl:library-sample}Sample library database}
+\centerline{\small DATA TYPES}
+\vglue-6pt
+\centerline{\small\baselineskip=13pt
+\begin{tabular}{c c c c}\\
+\hline
+\textbf{ID}  & Title  & Start Loan & End Loan \\
+\hline
+3 & Dracula & 15/3/2012 & 30/3/2012 \\
+4 & Frankenstein & 25/3/2012 & 4/4/2012 \\
+5 & Harry Potter & 18/3/2012 & 2/4/2012\\
+3 & Dracula & 4/4/2012 & UC \\
+\hline\\
+\end{tabular}
+}
+
+In order to simplify the algorithms for the manipulation of data, some auxiliary functions and constants are defined:
+
+\begin{definition}
+ \label{def:until-changed}
+\textbf{Until changed (UC):} To show that the data related to the entity $\left(A, I \right)$ is up-to-date, a special constant called \emph{Until Changed} is defined for the endpoint of the time interval $I$.
+The value for this constant is considered to be $+\infty$.
+\end{definition}
+
+%quizas poner un ejemplo relacionado con lo de la liberia.
+
+If the pair $\left(A, I \right)$ with $I = \left(c, UC \right), c \in \T$ is in the relation $R$, then, it is said that the pair $\left(A, I \right)$ is \emph{current} in the relation.
+
+For example, last row in Table \ref{tbl:library-sample}.
+The value for the time interval is $I = \left(4/4/2012 , UC \right)$.
+The meaning is that the book with ID = 3 was loaned the  4/4/2012 and it is still loaned.
+The book with ID = 3 is current in the relation.
+
+\begin{definition}
+\textbf{Current $\left(R, A \right)$:}
+\label{def:current-in-relation}
+This function returns a crisp time interval $I$ if the entity $A$ is current in the relation $R$ or $\emptyset$ if the entity $A$ is not current in the relation.
+
+%Consider an object $A$ and a relation $R$ in a temporal database.
+The function Current$\left(R, A \right)$ obtains the tuple $\left(A, I \right)$ that is current in the relation $R$, that is it, $I = \left[a, +\infty \right]$.
+
+\begin{align}
+\label{eq:current-in-relation}
+\mbox{Current} \left(R, A \right) &=& \\ 
+\begin{cases}
+\nonumber
+I & \mbox{ if } (A,I) \in R \wedge I = \left(c, UC \right), c \in \T\\
+\emptyset & \mbox{ in any other case }
+\end{cases}
+\end{align}
+\end{definition}
+
+For example, the function Current(R, (3,`Dracula')) returns the time interval $I = \left(4/4/2012 , UC  \right)$. On the other hand, Current(R, (4,`Frankenstein')) returns the empty set.
+
+The following two Allen's relations (see Figure \ref{fig:allen-rel}) are defined like in~\cite{Nagypal2003} to implement the auxiliary functions.
+
+\vspace*{13pt}
+\begin{center}
+{
+\includegraphics[scale=0.5]{./graphs/allen.pdf}
+
+}
+\end{center}
+%\centerline{ \psfig{file=./graphs/Y-time-point.eps}}
+\vspace*{10pt}
+\fcaption{\label{fig:allen-rel}Allen's relations for two crisp time intervals $I$ and $J$.}
+\vspace*{13pt}
+
+\begin{definition}
+\textbf{Overlaps:}
+ \label{def:overlaps}
+Given two time intervals $I = \left(c, d \right)$ and $J = \left(e, f\right)$, it is said that I overlaps J if:
+\begin{equation}
+ \label{eqn:overlaps}
+\text{I Overlaps J}  = \left(c > e  \right) \wedge \left(d < f  \right) 
 \end{equation}
@@
 
-In order to simplify the algorithms for the manipulation of data, some auxiliary functions are defined:
+\begin{definition}
+\textbf{During:}
+ \label{def:during}
+Given two time intervals $I = \left(c, d \right)$ and $J = \left(e, f\right)$, it is said that I during J if:
+\begin{equation}
+ \label{eqn:during}
+\text{I During J}  = \left(c < e  \right) \wedge \left(d > e  \right) \wedge \left(d < f \right)
+\end{equation}
+\end{definition}
 
 \begin{definition}
+\textbf{CloseR$\left( I, J \right)$:}
 \label{def:close-a-crisp-interval-r}
-Consider two crisp intervals $I= \left[a ,b \right]$ and $J= \left[c ,d \right]$, $c > a$.
-The CloseR$\left(I, J\right)$ function allows to close the right-open interval $I$ with respect to the first value $d$ in $J$:
+Consider two crisp intervals $I= \left[c ,d \right]$ and $J= \left[e ,f \right]$.
+The CloseR$\left(I, J\right)$ function allows to close the right-open interval $I$ with respect to the first value $e$ in $J$:
 \begin{align}
@@
 \nonumber
-I & \mbox{ if } b \neq +\infty \\
-I=\left[a, c \right[ & \mbox{ if } b = +\infty \wedge J > I
+%I & \mbox{ if } b \neq +\infty \\
+%I=\left[a, c \right[ & \mbox{ if } b = +\infty \wedge J > I
+\left[c, e \right[ & \mbox{ if } d = UC \wedge J \mbox{ During } I \\
+I & \mbox{ in any other case.}
 \end{cases}
@@
 \end{definition}
+
+For example, consider two time intervals, $I = \left(4/4/2012, UC \right)$ and $J = \left(24/4/2012, UC \right)$.
+The result of applying the CloseR$\left(I, J \right)$ is $I = \left(4/4/2012, 23/4/2012 \right)$.
 
@@
 %\end{definition}
-
-\begin{definition}
-\label{def:current-in-relation}
-Consider an object $A$ and a relation $R$ in a temporal database.
-The function Current$\left(R, A \right)$ obtains the tuple $\left(A, I \right)$ that is current in the relation $R$, that is it, $I = \left[a, +\infty \right]$.
-
-\begin{align}
-\label{eq:current-in-relation}
-\mbox{Current} \left(R, A \right) &=& \\ 
-\begin{cases}
-\nonumber
-I & \mbox{ if } \exists b \in I : (A,I) \in R \wedge b = +\infty \\
-\emptyset & \mbox{ in any other case }
-\end{cases}
-\end{align}
-\end{definition}
 
@@
 
-Now it is possible to close the current version of an object by using \eqref{def:close-a-crisp-interval-r} and \eqref{eq:current-in-relation}
+Now it is possible to close the current version of an entity by using \eqref{eq:close-a-crisp-interval} and \eqref{eq:current-in-relation}.
+This functionality is required to append new information about an existing entity in the relation.
 
@@
 \label{def:close-current-version}
-Consider an object $A$, the relation $R$ and a time interval $J$.
-Then, the function close-current$\left(R,A,J \right)$ closes any current version of the object $A$ if it exists and add the new version $\left(A, J \right)$.
+Consider an entity $A$, the relation $R$ and a time interval $J$.
+Then, the function close-current$\left(R,\left(A,J\right) \right)$ closes any current version of the object $A$ if it exists and add the new version $\left(A, J \right)$.
 
@@
 \label{eq:close-current}
-\text{Close-current} \left(R, A, J \right) =\\
+\text{Close-current} \left(R, \left(A, J\right) \right) =\\
 \begin{cases}
 \nonumber
-\big \lbrace R - \left(A, I \right) \cup \left \lbrace \left(A, \mbox{ CloseR } \left(I, J\right) \right) \cup \left(A, J\right)\right \rbrace  \big \rbrace \\
+R - \left(A, I \right) \cup \left \lbrace \left(A, \mbox{ CloseR } \left(I, J\right) \right)\right \rbrace \cup \left \lbrace\left(A, J\right) \right \rbrace  \\
 \nonumber
-\mbox{ if } I = \mbox{ Current } \left(R, A \right) \wedge J > \mbox{ CloseR } \left(I, J \right)   \\
+\mbox{ if } I = \mbox{ Current } \left(R, A \right)\\ %\wedge J > \mbox{ CloseR } \left(I, J \right)   \\
 \nonumber R , \text{ in any other case}
@@
 \end{definition}
+
+For example, consider the book with ID = 3 in Table \ref{tbl:library-sample}.
+The current version was loaned on 4/4/2012.
+The book was returned to the library but, for some reason, the end loan date was not registered.
+The book is loaned again on 24/4/2012.
+Then, the Close-current function allows to close the current version of the book and insert a new version.
+First, the function CloseR is applied with $I = \left(4/4/2012, UC \right)$ and $J = \left(24/4/2012, UC \right)$.
+Hence, the value for the time interval $I$ is $ \left(4/4/2012, 23/4/2012 \right)$.
+Then, the modifications on the value for the time interval $I$ are stored.
+Finally, a new row with the current values of the book and the time interval $J$ is stored.
+The result of this operation is illustrated on Table \ref{tbl:library-sample-2}.
 
@@
 %\end{definition}
+\subsubsection{\label{subsubsec:modify}Modify}
+This operation adds new information about an existing entity $A$ in the relation $R$.
+The modify operation does not remove any previous value of the object $A$.
+It closes the current version for the entity $A$ and adds a new version.
+
+%Note that the modify operation is only applicable when the object $A$ is current in the relation $R$; $ \left(A, I \right) \in R, I = \left[a, +\infty \right]$.
+
+\begin{align}
+\label{eq:modify}
+\MOD \left( R, \left(A, I \right)\right) =\\
+%\begin{cases}
+\nonumber
+\mbox{ Close-current }\left(R, A, \mbox{Current}\left(R, A \right) \right) %\\
+% \mbox{ if Current}\left(R ,A \right) \neq \emptyset \\
+% R,  \mbox{ otherwise }
+%\end{cases}
+\end{align}
 
@@
 \begin{enumerate}
-\item The object $A$ was never in the relation $R$: The object is added with the valid-time indicated by the crisp interval $I$.
+\item The object $A$ was never in the relation $R$: The object is added with the valid-time indicated by the crisp interval $I$. In the example, this correspond with the following insertion sentences:
+\begin{verbatim}
+Insert(3,'Dracula',
+(15/3/2012, 30/3/2012));
+Insert(4,'Frankenstein',
+(25/3/2012, 4/4/2012));
+Insert(5,'Harry Potter',
+(18/3/2012, 2/4/2012));
+\end{verbatim}
 
-\item The object $A$ is in the relation $R$. Depending on the value of the time interval $I$, there are two possibilities:
+\item The object $A$ is in the relation $R$. Depending on the value of the time interval $I$, there are three possibilities:
 	\begin{enumerate}
-	\item Insert $\left(A, I\right)$ in the relation $R$. If the time interval $I$ does not overlap any other valid-time interval in the relation $R$.
-	\item Reject the insertion, if the time interval $I$ do overlap any existing valid-time interval for the object $A$ in the relation $R$.
+	\item Insert $\left(A, I\right)$ in the relation $R$.
+If the time interval $I$ does not overlap any other valid-time interval in the relation $R$.
+For example, consider that the book with ID 3 is loaned on 4/4/2012 but the book has not been returned yet.
+This is illustrated in Table \ref{tbl:library-sample}.
+The insert sentence is:
+	      \begin{verbatim}
+Insert(3,'Dracula',
+(4/4/2012, UC));
+	      \end{verbatim}
+
+	\item Modify and close the current version of $A$ and insert $\left(A, I\right)$.
+For example, consider now that the book with ID 3 is loaned  the 24/4/2012.
+Here the problem is that the book with ID 3 was loaned on 4/4/2012, but not returned.
+If the book is already in the library for a new loan, then it is necessary to set the return date and add a new row with the new loan date.
+This is illustrated in Table \ref{tbl:library-sample-2}, the insert sentence is:
+	    \begin{verbatim}
+Insert(3,'Dracula',
+(24/4/2012, UC));
+	      \end{verbatim}
+	\item Reject the insertion, if the time interval $I$ does overlap any existing valid-time interval for the object $A$ in the relation $R$.
+For example, consider now that the librarian wants to introduce a past loan for the book with ID 3.
+The book was loaned on 6/4/12/2012 and returned on 25/4/2012.
+As this interval does overlaps other time intervals, it is not possible that the book was loaned in that time interval.Therefore, the insertion is rejected.
+The insert sentence is:
+
+	      \begin{verbatim}
+Insert(3,'Dracula',
+(6/4/12/2012,25/4/2012));
+	    \end{verbatim}
+
 	\end{enumerate}
@@
 \end{enumerate}
+
+\vglue13pt
+%\begin{table}[htbp]
+\tcap{\label{tbl:library-sample-2}Sample library database}
+%\centerline{\small DATA TYPES}
+\vglue-6pt
+\centerline{\small\baselineskip=13pt
+\begin{tabular}{c c c c}\\
+\hline
+\textbf{ID}  & Title  & Start Loan & End Loan \\
+\hline
+3 & Dracula & 15/3/2012 & 30/3/2012 \\
+4 & Frankenstein & 25/3/2012 & 4/4/2012 \\
+5 & Harry Potter & 18/3/2012 & 2/4/2012\\
+3 & Dracula & 4/4/2012 & 23/4/2012 \\
+3 & Dracula & 24/4/2012 & UC \\
+\hline\\
+\end{tabular}
+}
 
@@
 \label{eq:insert}
-\INS \left(R ,A, I\right) =\\
+\INS \left(R ,\left(A, I\right)\right) =\\
 \begin{cases}
 \nonumber
-R \cup \left(A, I \right), & \mbox{ if }  A \not \in R\\
-R \cup   \left(A, I \right) & \mbox{ if }  A \in R \wedge \\
-& \big \lbrace \left( \exists J: \left(A, J \right) \in R \wedge J<I  \right) \vee \\
-& \left( \exists K: \left(A, K \right) \in R \wedge K>I \vee \right) \wedge\\
-&  J<I<K  \big \rbrace\\
-R, & \mbox{ otherwise }  
+R \cup \left(A, I \right) & \mbox{ if }  (A, I ) \not \in R  \vee \\
+ &   \left(A, J\right) \in R \wedge \lnot \left(I \mbox{ Overlaps } J\right) \\
+\mbox{modify(R,(A,I))} & \mbox{ otherwise }    %\mbox{ if } \exists J \left(A,J\right) \in R \wedge \mbox{Current(A,J)} \\ 
+%R, &   
 \end{cases} 	
@@
 \subsubsection{\label{subsubsec:del}Delete}
-The delete operation logically removes a current object $A$ which is valid in the relation $R$:
+The delete operation logically removes a current entity $A$ which is valid in the relation $R$:
 \begin{align}
@@
 \nonumber
-R - \left(A, I \right), & \mbox{ if } \left( A, I \right) \in R\\
+R - \left(A, I \right), & \forall \left( A, I \right) \in R\\
 R, & \mbox{ otherwise }  
@@
 
-\subsubsection{\label{subsubsec:modify}Modify}
-This operation adds new information about an existing object $A$ in the relation $R$. The modify operation does not remove any previous value of the object $A$.
-Note that the modify operation is only applicable when the object $A$ is current in the relation $R$; $ \left(A, I \right) \in R, I = \left[a, +\infty \right]$.
+For example, consider that the librarian wants to delete the history for the book with ID = 3. The result of this operation is shown in Table \ref{tbl:library-sample-3}. The delete sentence is:
 
+\begin{verbatim}
+ Delete(3,'Dracula');
+\end{verbatim}
+
+\vglue13pt
+%\begin{table}[htbp]
+\tcap{\label{tbl:library-sample-3}Sample library database}
+%\centerline{\small DATA TYPES}
+\vglue-6pt
+\centerline{\small\baselineskip=13pt
+\begin{tabular}{c c c c}\\
+\hline
+\textbf{ID}  & Title  & Start Loan & End Loan \\
+\hline
+4 & Frankenstein & 25/3/2012 & 4/4/2012 \\
+5 & Harry Potter IV & 18/3/2012 & 2/4/2012\\
+\hline\\
+\end{tabular}
+}
+
+\subsubsection{\label{subsubsec:revise}Revise}
+This operation replaces the old values for the entity $A'= \left(a_1', \ldots, a_n' \right) $ with the new values specified by $A = \left(a_1, \ldots, a_n \right)$ in the relation $R$, at a given time interval $I$.
+This operation is defined when a correction on the values for the entity $A$ should be done.
 \begin{align}
-\label{eq:modify}
-\MOD \left( R, A, I \right) =\\
+\REV\left(R, \left(A, I \right) \right)=\\
 \begin{cases}
 \nonumber
-\mbox{ Close-current }\left(R, A, \mbox{Current}\left(R, A \right) \right) \\
- \mbox{ if Current}\left(R ,A \right) \neq \emptyset \\
- R,  \mbox{ otherwise }
+ R - \left(A', I \right) \cup \left(A, I \right) & \mbox{ if } \left(A', I \right) \in R  \\
+   %: A = \left(a_1, \ldots, a_n \right),\\
+%A'= \left(a_1', \ldots, a_n' \right)   \\   
+R & \mbox{ otherwise }
 \end{cases}
@@
 
-\subsubsection{\label{subsubsec:revise}Revise}
-This operation changes the value of the object $A$ in the relation $R$, at a given time interval $I$. This operation is defined when a correction on the values for the object $A$ should be done.
-\begin{align}
-\label{eq:revise}
-\mbox{ Revise}\left(R, A, I \right)=\\
-\begin{cases}
-\nonumber
- R - \left(A', I \right) \cup \left(A, I \right)  \\
-  \mbox{ if } \left(A', I \right) \in R: A = \left(a_1, \ldots, a_n \right),\\
-A'= \left(a_1', \ldots, a_n' \right)   \\   
-R,  \mbox{ otherwise }
-\end{cases}
-\end{align}
+For example, consider now that the librarian wants to make a correction in the title for the book with ID = 5.
+The new title is `Harry Potter IV'.
+The result of this operation is illustrated in Table \ref{tbl:library-sample-3}.
+ Then, the revise sentence is:
+
+\begin{verbatim}
+Revise(5,'Harry Potter IV', 
+(18/3/2012, 2/4/2012));
+\end{verbatim}

### 2012-06-05 ijcis-pos-model/sources/temporaldatabases.tex
last modifications

@@
 
-In the proposal section we will consider two kinds of imprecision:
+In the proposal section we will consider two kinds of imperfection:
 \begin{itemize}
-\item \textbf{Imperfection in the database} denotes the uncertainty that arises when the knowledge about the temporal data in the database is uncertain.
-E.g., a database record shows that \emph{`The car is in the garage around April.'}
+\item \textbf{Imperfection in the database} the knowledge about the temporal data contains some imperfection. E.g., a database record shows that \emph{`The car is in the garage around April.'}
  \item \textbf{Imprecision in the query specification} denotes the imprecision in the specification of temporal criteria by the user, when querying.
@@
 \begin{definition}
-Consider a relation $R$ in a temporal database, an object $A$ given by the attributes $\left(a_1, \ldots, a_n \right)$ and the crisp time interval $I$ given by $\left[a,\ b\right] : a,b \in \T, a \leq b$.
-A valid-time object $A$ in the relation $R$ is specified by:
+Consider a relation $R$ in a temporal database, an entity $A$ given by the attributes $\left(a_1, \ldots, a_n \right)$ and the crisp time interval $I$ given by $\left[a,\ b\right] : a,b \in \T, a \leq b$.
+The pair $\left(A, I\right)$ expresses that the data regarding the entity A is valid during the crisp time interval I.
+Note that the relation $R = \left(A, I \right) x \left(A, I \right)$.
+The following equation indicates that the pair $\left(A, I\right)$ is in the relation R:
 \begin{equation}
 \label{eq:rel-def}
-\left( R, A , I \right)
+\left( R, \left(A , I\right) \right)
+\end{equation}
+\end{definition}
+\begin{example}
+\label{ex:library-database}
+ Consider a library database.
+Each physical book is stored with an identifier (ID), a title and the start and end loan dates respectively.
+Table \ref{tbl:library-sample} illustrates the first version of the database, after three insertions.
+In this example, $A = ($ ID, Title $)$ and $I = ($ Start Loan, End Loan $)$.
+\end{example}
+
+\vglue13pt
+%\begin{table}[htbp]
+\tcap{\label{tbl:library-sample}Sample library database}
+\centerline{\small DATA TYPES}
+\vglue-6pt
+\centerline{\small\baselineskip=13pt
+\begin{tabular}{c c c c}\\
+\hline
+\textbf{ID}  & Title  & Start Loan & End Loan \\
+\hline
+3 & Dracula & 15/3/2012 & 30/3/2012 \\
+4 & Frankenstein & 25/3/2012 & 4/4/2012 \\
+5 & Harry Potter & 18/3/2012 & 2/4/2012\\
+3 & Dracula & 4/4/2012 & UC \\
+\hline\\
+\end{tabular}
+}
+
+In order to simplify the algorithms for the manipulation of data, some auxiliary functions and constants are defined:
+
+\begin{definition}
+ \label{def:until-changed}
+\textbf{Until changed (UC):} To show that the data related to the entity $\left(A, I \right)$ is up-to-date, a special constant called \emph{Until Changed} is defined for the endpoint, $d$, of the time interval $I = \left(c, d\right)$.
+The value for this constant is considered to be $+\infty$.
+\end{definition}
+
+If the pair $\left(A, I \right)$ with $I = \left(c, UC \right), c \in \T$ is in the relation $R$, then, it is said that the pair $\left(A, I \right)$ is \emph{current} in the relation.
+
+For example, last row in Table \ref{tbl:library-sample}.
+The value for the time interval is $I = \left(4/4/2012 , UC \right)$.
+The meaning is that the book with ID = 3 was loaned the  4/4/2012 and it is still loaned.
+The book with ID = 3 is current in the relation.
+
+\begin{definition}
+\textbf{Current $\left(R, A \right)$:}
+\label{def:current-in-relation}
+This function returns a crisp time interval $I$ if the entity $A$ is current in the relation $R$ or the empty set, if the entity $A$ is not current in the relation.
+
+%Consider an object $A$ and a relation $R$ in a temporal database.
+The function Current$\left(R, A \right)$ obtains the tuple $\left(A, I \right)$ that is current in the relation $R$, that is it, $I = \left[a, +\infty \right]$.
+
+\begin{align}
+\label{eq:current-in-relation}
+\mbox{Current} \left(R, A \right) &=& \\ 
+\begin{cases}
+\nonumber
+I & \mbox{ if } (A,I) \in R \wedge I = \left(c, UC \right), c \in \T\\
+\emptyset & \mbox{ in any other case }
+\end{cases}
+\end{align}
+\end{definition}
+
+For example, the function Current(R, (3,`Dracula')) returns the time interval $I = \left(4/4/2012 , UC  \right)$. On the other hand, Current(R, (4,`Frankenstein')) returns the empty set.
+
+The following two Allen's relations (see Figure \ref{fig:allen-rel}) are defined~\cite{Nagypal2003} to implement the auxiliary functions.
+
+\vspace*{13pt}
+\begin{center}
+{
+\includegraphics[scale=0.5]{./graphs/allen.pdf}
+
+}
+\end{center}
+%\centerline{ \psfig{file=./graphs/Y-time-point.eps}}
+\vspace*{10pt}
+\fcaption{\label{fig:allen-rel}Allen's relations for two crisp time intervals $I$ and $J$.}
+\vspace*{13pt}
+
+\begin{definition}
+\textbf{Overlaps:}
+ \label{def:overlaps}
+Given two time intervals $I = \left(c, d \right)$ and $J = \left(e, f\right)$, it is said that I overlaps J if:
+\begin{equation}
+ \label{eqn:overlaps}
+\text{I Overlaps J}  = \left(c > e  \right) \wedge \left(d < f  \right) 
 \end{equation}
@@
 %\end{definition}
+\subsubsection{\label{subsubsec:modify}Modify}
+This operation adds new information about an existing entity $A$ in the relation $R$.
+The modify operation does not remove any previous value of the object $A$.
+It closes the current version for the entity $A$ and adds a new version.
+
+%Note that the modify operation is only applicable when the object $A$ is current in the relation $R$; $ \left(A, I \right) \in R, I = \left[a, +\infty \right]$.
+
+\begin{align}
+\label{eq:modify}
+\MOD \left( R, \left(A, I \right)\right) =\\
+%\begin{cases}
+\nonumber
+\mbox{ Close-current }\left(R, \left(A, I\right) \right) %\\
+% \mbox{ if Current}\left(R ,A \right) \neq \emptyset \\
+% R,  \mbox{ otherwise }
+%\end{cases}
+\end{align}
 
@@
 
-\subsubsection{\label{subsubsec:modify}Modify}
-This operation adds new information about an existing object $A$ in the relation $R$. The modify operation does not remove any previous value of the object $A$.
-Note that the modify operation is only applicable when the object $A$ is current in the relation $R$; $ \left(A, I \right) \in R, I = \left[a, +\infty \right]$.
+For example, consider that the librarian wants to delete the history for the book with ID = 3. The result of this operation is shown in Table \ref{tbl:library-sample-3}. The delete sentence is:
 
+\begin{verbatim}
+ Delete(3,'Dracula');
+\end{verbatim}
+
+\vglue13pt
+%\begin{table}[htbp]
+\tcap{\label{tbl:library-sample-3}Sample library database}
+%\centerline{\small DATA TYPES}
+\vglue-6pt
+\centerline{\small\baselineskip=13pt
+\begin{tabular}{c c c c}\\
+\hline
+\textbf{ID}  & Title  & Start Loan & End Loan \\
+\hline
+4 & Frankenstein & 25/3/2012 & 4/4/2012 \\
+5 & Harry Potter IV & 18/3/2012 & 2/4/2012\\
+\hline\\
+\end{tabular}
+}
+
+\subsubsection{\label{subsubsec:revise}Revise}
+This operation replaces the old values for the entity $A'= \left(a_1', \ldots, a_n' \right) $ with the new values specified by $A = \left(a_1, \ldots, a_n \right)$ in the relation $R$, at a given time interval $I$.
+This operation is used when a correction on the values for the entity $A$ should be done.
 \begin{align}
-\label{eq:modify}
-\MOD \left( R, A, I \right) =\\
+\REV\left(R, \left(A, I \right) \right)=\\
 \begin{cases}
 \nonumber
-\mbox{ Close-current }\left(R, A, \mbox{Current}\left(R, A \right) \right) \\
- \mbox{ if Current}\left(R ,A \right) \neq \emptyset \\
- R,  \mbox{ otherwise }
+ R - \left(A', I \right) \cup \left \lbrace \left(A, I \right) \right \rbrace & \mbox{ if } \left(A', I \right) \in R  \\
+   %: A = \left(a_1, \ldots, a_n \right),\\
+%A'= \left(a_1', \ldots, a_n' \right)   \\   
+R & \mbox{ otherwise }
 \end{cases}

### 2012-06-05 ijcis-pos-model/sources/timeRepresentation.tex
last modifications

@@
 \label{def:ill-known-time-point}
-\emph{Ill-known time point}\\
+\textbf{Ill-known time point.}\\
 Consider a time domain $\T$, uncertainty about the values of the ill-known time point $X$ is given by the possibility distribution $\pi_X$:
@@
 \label{def:ill-known-domain}
-\emph{Domain for an ill-known time point}\\
+\textbf{Domain for an ill-known time point.}\\
 Consider $\Pow(\T)$ the set of all the possibility distributions over $\T$, and the three fuzzy constants \emph{UNKNOWN} $=\left\lbrace 1/t, \forall t \in \T \right\rbrace$, \emph{UNDEFINED} $=\left\lbrace 0/t,\forall t \in \T \right\rbrace$ and \emph{NULL} $=\lbrace 1/$ \emph{UNKNOWN}, $1/$ \emph{UNDEFINED} $\rbrace$.
@@
 \subsubsection{\label{subsubsec:ill-known-time-datatypes}Datatypes}
-The datatypes for an ill-known time point $X$ are shown in Table \ref{tbl:time-point-types}.
+The data type for the representation of an ill-known time point allows the representation of the values shown in Table \ref{tbl:time-point-types}
+%The datatypes for an  $X$ are shown in Table \ref{}.
 \vglue13pt
 %\begin{table}[htbp]
-\tcap{Data types for a time point}
-\centerline{\small DATA TYPES}
+\tcap{\label{tbl:time-point-types}Values for the time point data type.}
+\centerline{\small TIME POINT DATA TYPE}
 \vglue-6pt
 \centerline{\small\baselineskip=13pt
-\begin{tabular}{c p{2cm} p{2cm}}\\
+\begin{tabular}{c p{0.4\columnwidth} p{0.3\columnwidth}}\\
+Subtype & Value & Representation \\
 \hline
@@
 \end{tabular}
-\label{tbl:time-point-types}
 }
@@
 \begin{definition}
-\emph{Ill-known time interval}\\
+\textbf{Ill-known time interval}\\
 An ill-known time interval denoted by $\left[X, Y\right]$ is a precise time interval whose boundaries are not precisely known.
@@
 \begin{eqnarray}
+\label{eq:constraint-c1}
 C_1\stackrel{\triangle}{=}\left(\geq,X\right)  \\
+\label{eq:constraint-c2}
 C_2\stackrel{\triangle}{=}\left(\leq,Y\right)
@@
 Then it is possible to distinguish between the following two types of open intervals:
-\begin{enumerate}
-\item
+%\begin{enumerate}
+%\item
 \begin{definition}
-\emph{Completely unknown time interval}: Both starting and ending points are unknown, therefore the whole interval is specified by the \emph{UNKNOWN} constant.
-Due to the representation, both endpoints, say $X$ and $Y$, are stored.
-Thus, the interval is stored as $[$ \emph{UNKNOWN}, \emph{UNKNOWN} $]$.
+\textbf{Completely unknown time interval}: Both starting and ending points are unknown, therefore the whole interval is unknown.
+%specified by the \emph{UNKNOWN} constant. Due to the representation, both endpoints, say $X$ and $Y$, are stored. Thus, the interval is stored as $[$ \emph{UNKNOWN}, \emph{UNKNOWN} $]$. 
 \end{definition}
 
-\item
+%\item
 \begin{definition}
-\emph{Semi-open time interval}: Only one of the two ill-known endpoints for the time interval $\left[X, Y\right]$ is unknown.
-Note that due to the ill-known constraints $C_{1}, C_{2}$ it is not possible a representation like $[$ \emph{UNKNOWN} $, Y]$ or $[X, $ \emph{UNKNOWN} $]$.
+\textbf{Semi-open time interval}: Only one of the two ill-known endpoints for the time interval $\left[X, Y\right]$ is unknown.
+Example \ref{ex:open-time-interval} and Figure \ref{fig:example-open-time-interval} illustrates a left open time interval.
+% Note that due to the ill-known constraints $C_{1}, C_{2}$ it is not possible a representation like $[$ \emph{UNKNOWN} $, Y]$ or $[X, $ \emph{UNKNOWN} $]$.
 The solution adopted for this problem is explained in the following paragraph.
 \end{definition}
-\end{enumerate}
+%\end{enumerate}
 
-\paragraph{Representation of semi-open time intervals}
+\subsubsection{\label{subsubsec:representation-semi-open}Representation of semi-open time intervals}
 As mentioned before, the problem resides in the representation of these kind of interval.
@@
 %\begin{table}[htbp]
-\tcap{Relations for the $\Open(C)$ function.}
+\tcap{Relations for the $\Open(C)$ function. Depending on the relation $R = \left(\leq, <, >, \geq\right)$ in the constraint $C$, the values for $\Rp$ and $\Rn$ are shown.}
 \centerline{\small Relations}
@@
 \hline
-$R$ & $\Rp$ & $\Rn$\\ \hline
-$<$ & $>$ & $\leq$ \\
-$\leq$ & $\geq$ & $<$ \\
-$>$ & $<$ & $\geq$ \\
-$\geq$ & $\leq$ & $>$ \\
+Constraint & $\Rp$ & $\Rn$\\ \hline
+$C\stackrel{\triangle}{=}\left(<,T\right)$ & $>$ & $\leq$ \\
+$C\stackrel{\triangle}{=}\left(\leq,T\right)$ & $\geq$ & $<$ \\
+$C\stackrel{\triangle}{=}\left(>,T\right)$ & $<$ & $\geq$ \\
+$C\stackrel{\triangle}{=}\left(\geq,T\right)$ & $\leq$ & $>$ \\
 \hline\\
@@
 
+Where the constraints $C_1$ and $C_2$ are given in equations \eqref{eq:constraint-c1} and \eqref{eq:constraint-c2}.
+
 \begin{example}
+\label{ex:open-time-interval}
 Consider an ill-known time interval given by $\left[\FB, Y \right]$.
@@
 %\begin{table}[htbp]
-\tcap{Data types for an ill-known time interval $\left[X, Y\right]$.}
-\centerline{\small DATA TYPES}
+\tcap{All the possible combination of values for the time interval $\left[X, Y\right]$. The subtypes refer to Table \ref{tbl:time-point-types}.}
+\centerline{\small TIME INTERVAL DATA TYPE}
 \vglue-6pt
 \centerline{\small\baselineskip=13pt
-\begin{tabular}{c c p{2cm}}\\
+\begin{tabular}{p{0.15\columnwidth} p{0.15\columnwidth} p{0.5\columnwidth}}\\
 \hline
-$X$  & $Y$  & Description\\
+Subtype for $X$  & Subtype for $Y$  & Description\\
 \hline
-$\lbrace1,2 \rbrace$ & $\lbrace1,2 \rbrace$ &  an ill-known time interval.\\
+1 or 2 & 1 or 2 &  An ill-known time interval.\\
 3 & 3 & An unknown time interval. \\
-$\FB$ & $\lbrace1,2 \rbrace$ & a left-open time interval.\\
-$\lbrace1,2 \rbrace$ & $\UC$ & a right-open time interval.\\
+$\FB$ & 1 or 2 & A left-open time interval.\\
+1 or 2 & $\UC$ & A right-open time interval.\\
 \hline\\
@@
 
+% Maybe an example with a database.
+% \begin{example}
+% 
+% \end{example}
+
 \subsection{\label{subsec:fuzzy-allen-relations}Allen's relations}
+The Allen's relations can be represented by a Boolean combination of ill-known constraints.
+The constructs for the constraints are shown in Table \ref{tbl:fuzzy-allen-relations}.
+The possibility and necessity measures are obtained for each Boolean function as explained in Section \ref{subsec:interval-evaluation-by-ill-known-constraints} in equations \eqref{eq:conjunctive1}-\eqref{eq:negation}.
 
@@
 \centerline{\small\baselineskip=13pt
-\begin{tabular}{c p{4cm} }\\
+\begin{tabular}{p{0.2\columnwidth} p{0.8\columnwidth} }\\
 \hline

### 2012-06-05 trunk/ijcis-pos-model/sources/temporalModel.tex
last modifications

@@
 The model is based on the GEFRED~\cite{Medina1994} (Generalized Model of Fuzzy Relational DB) model.
-This model is extended by adding valid-time support.
+This model is extended by adding valid-time support which will be illustrated with an example.
 The information in the system is defined by the following elements:
@@
 \label{def:generalized-fuzzy-domain}
+\textbf{Generalized fuzzy domain.}
 $D$ is the discourse domain, $\tilde \Pow\left(D \right)$ is the set of all possibility distributions defined on $D$, plus the NULL constant. The generalized fuzzy domain $D_G$ is defined as:
@@
 \label{def:typeof-domain}
-The function typeof$\left(D \right)$ returns the datatype associated with the domain $D$.
-
+\textbf{Typeof(D).}
+The function typeof$\left(D \right)$ returns the datatype associated with the domain $D$. For datatype, the function returns the associated number as shown in Table \ref{tbl:gefred-data-types}.
 \end{definition}
@@
 \label{def:generalized-fuzzy-temporal-domain}
-Generalized fuzzy temporal domain.\\
+\textbf{Generalized fuzzy temporal domain.}
 If $\T$ is the temporal domain, $\tilde \Pow\left( \T\right)$ is the set of all \emph{normalized} possibility distributions (see Section \ref{subsec:possibility-theory}, equation \eqref{NormalizationProperty}) defined on $\T$.
-The \textbf{Generalized Fuzzy Temporal Domain}, $\T_G$ is
+The Generalized Fuzzy Temporal Domain, $\T_G$ is
 \begin{equation}
@@
 Note that $\T_G \subseteq D_G$.
-The datatypes for this domain have been studied previously in section \ref{sec:time-rep} and are shown in tables \ref{tbl:time-point-types},\ref{tbl:time-interval-types}.
+The datatypes for this domain have been studied previously in section \ref{sec:time-rep} and are shown in tables \ref{tbl:time-point-types} and \ref{tbl:time-interval-types}.
 
 \begin{definition}
+\textbf{Generalized fuzzy temporal relation.}
 A generalized fuzzy temporal relation $R_{FTG}$ is given by:
@@
 
+\begin{example}
+\label{ex:sample-ft-database}
+ Consider a library database as explained in example \ref{ex:library-database}.
+But now, the time is stored with imprecision.
+For simplicity, the ill-known time points are represented as triangular fuzzy numbers.
+An ill-known time point is given by $[dd/mm/yyyy,a,b]$.
+The values for $a$ and $b$ are integers.
+Thus, the date given by $[15/3/2012,5,2]$ is a triangular fuzzy number with the left bound the 10/3/2012, the core on the 15/3/2012 and the right bound on 17/3/2012.
+Table \ref{tbl:fuzzy-library-sample} shows the elements Head, $\Head$ and Body, $\Body$. 
+\begin{align}
+ \label{eq:head-ex}
+\Head = & \lbrace \left(\text{ID}:D_{\text{ID}}\right),\\
+\nonumber
+&  \left(\text{Title}:D_{\text{title}}\right),\\
+\nonumber
+& \left(\text{Loan}:D_{\text{Loan}}\right) \\
+\nonumber
+& \left(\text{PVP}:D_{\text{PVP}}\right) \rbrace
+\end{align}
+
+When the compatibility degree is 1, the component is omitted. The body, $\Body$ consists on all the tuples shown in Table \ref{tbl:fuzzy-library-sample}.
+
+\end{example}
+
+\vglue13pt
+%\begin{table}[htbp]
+\tcap{\label{tbl:fuzzy-library-sample}Sample library database}
+%\centerline{\small DATA TYPES}
+\vglue-6pt
+\centerline{\small\baselineskip=13pt
+\begin{tabular}{c c c |c c}\\
+\hline
+& & & \multicolumn{2}{c}{PVP} \\
+$\Head$ &  \textbf{ID}  & Title  &  Start Loan & End Loan\\
+\hline
+\multirow{3}{*}{$\Body$ \Big\lbrace} &  3 & Dracula & [15/3/2012, 5, 2] & [30/3/2012, 1, 1] \\
+ &4 & Frankenstein & [25/3/2012, 3, 2] & [4/4/2012, 1, 7] \\
+&5 & Harry Potter & [18/3/2012, 4, 1] & [2/4/2012, 2, 2]\\
+% \ 3 & Dracula & [4/4/2012, 3, 3] & UC \\
+\hline\\
+\end{tabular}
+}
+
 \begin{definition}
 \label{def:value-component}
+\textbf{Value component.}
 The value component $R^{v}_{FTG}$ of a fuzzy temporal relation $R_{FTG}$ is a set with the value components for both the head and the body of the relation:
@@
 \nonumber
-\Body^ {v} = \left \lbrace A_{G1}:\tilde{d}_{i1}, \ldots,  A_{Gn}:\tilde{d}_{in}\ \right \rbrace \\
+\Body^ {v} = \left \lbrace A_{G1}:\tilde{d}_{i1}, \ldots,  A_{Gn}:\tilde{d}_{in}\ \right \rbrace 
 \end{align}
@@
 
+For example, in the case of the record with ID = 3:
+\begin{align}
+%\label{eq:value-component-s}
+\nonumber
+\Head^{v} &=  \lbrace \left(\text{ID}:D_{\text{ID}}\right), \left(\text{Title}:D_{\text{title}}\right),\\
+\nonumber
+& \left(\text{Loan}:D_{\text{Loan}}\right), \left(\text{PVP}:D_{\text{PVP}}\right) \rbrace \\
+\nonumber
+\Body^ {v} &= \left \lbrace 3:, \text{``Dracula``}: ,  \left[15/3/2012, 30/3/2012\right] \right \rbrace 
+\end{align}
+
 \begin{definition}
 \label{def:compatibility-component}
+\textbf{Compatibility component.}
 The compatibility component $R^{c}_{FTG}$ of a fuzzy temporal relation $R_{FTG}$ is a set with the compatibility components for both the head and the body of the relation:
@@
 \nonumber
-\Body^ {c} = \left \lbrace \left[ c_{i1} \right], \ldots, \left[ c_{in} \right] \right \rbrace \\
+\Body^ {c} = \left \lbrace \left[ c_{i1} \right], \ldots, \left[ c_{in} \right] \right \rbrace 
 \end{align}
@@
 
+For example, in the case of the record with ID = 3:
+\begin{align}
+%\label{eq:sample-compatibility-component}
+\nonumber
+\Head^{c} = & \lbrace \left(C_\text{ID}\right), \left(C_{\text{title}}\right), \left(C_{\text{Loan}}\right), \left(C_{\text{PVP}}\right) \rbrace \\
+% \nonumber
+% &  \\
+% \nonumber
+% & \left(C_{\text{Loan}}\right) \\
+% \nonumber
+% & \left(C_{\text{PVP}}\right) \rbrace \\
+\nonumber
+\Body^ {c} = &\left \lbrace 1, 1, 1, 1 \right \rbrace 
+\end{align}
+
 \begin{definition}
 \label{def:temporal-component}
+\textbf{Temporal component.}
 The temporal component $R^{t}_{FTG}$ of a fuzzy temporal relation $R_{FTG}$ is a set with the temporal components for both the head and the body of the relation:
@@
 \nonumber
-\Body^ {t} = \left \lbrace  \left[ \text{PVP}, \tilde{d}_{\text{PVP}} \left[,C_{A_{\text{PVP}}} \right] \right]  \right \rbrace \\
+\Body^ {t} = \left \lbrace  \left[ \text{PVP}, \tilde{d}_{\text{PVP}} \left[,C_{A_{\text{PVP}}} \right] \right]  \right \rbrace 
 \end{align}
 \end{definition}
+
+For example, in the case of the record with ID = 3:
+\begin{align}
+%\label{eq:sample-temporal-component}
+\nonumber
+\Head^{t} = & \lbrace \left( \text{PVP}, D_{\text{PVP}}\right),\\
+\nonumber
+\Body^ {t} =& \left \lbrace 4/4/2012 , UC \right \rbrace 
+\end{align}
 
@@
 \label{def:generalized-primary-key}
+\textbf{Generalized primary key.}
 A generalized primary key, $K_G$ is a subset of the head:
@@
 
+For example, consider the database in Table \ref{tbl:fuzzy-library-sample}. Without any temporal constraint, the primary key $K_G$ is:
+
+\begin{align}
+\nonumber
+K_G \subseteq \Head, K_G = \left \lbrace  \left(\text{ID}:D_{\text{ID}} \right) \right \rbrace
+\end{align}
+In this case, the function \emph{Typeof(ID) = 2} (see Definition \ref{def:typeof-domain} and Table \ref{tbl:gefred-data-types}).
+The primary key for the table is the attribute \emph{ID}, a unique number.
+Two different books have two different values for the \emph{ID} attribute.
+
+In order to add valid-time support, the primary key should be re-defined.
+E.g., consider the library database.
+If the primary key is the \emph{ID} attribute, a book could not be loaned more than once.
+To resolve this problem, we extend the given primary key with a version identifier.
+%Thus, when a book is loaned, a new entry in the table is added. The \emph{ID} attribute identifies the physical book whereas
+
 \begin{definition}
 \label{def:generalized-fuzzy-temporal-key}
+\textbf{Generalized fuzzy temporal key.}
 A generalized fuzzy temporal key, $K_{GT}$ is a subset of the head. An attribute, namely \emph{version} $V$ is added when dealing with valid-time.
@@
 \end{align}
-
 \end{definition}
 
-\begin{example}
+Consider now the database in Table \ref{tbl:fuzzy-library-sample-valid-time}. The primary key is now:
 
-\end{example}
+\begin{align}
+ \nonumber
+K_{GT} \subseteq \Head, K_{GT} = \left \lbrace  \left(\text{ID}:D_{\text{ID}} \right) \cup  \left(V_{\text{ID}}:D_{\text{ID}} \right) \right \rbrace	
+%\nonumber
+ %\left. \cup  \left(V_{ID}:D_{ID} \right)	\right \rbrace \\
+\end{align}
+
+\vglue13pt
+%\begin{table}[htbp]
+\tcap{\label{tbl:fuzzy-library-sample-valid-time}Sample library database}
+%\centerline{\small DATA TYPES}
+\vglue-6pt
+\centerline{\small\baselineskip=13pt
+\begin{tabular}{c c c c | c c}\\
+\hline
+& & & &\multicolumn{2}{c}{PVP} \\
+$\Head$  & \textbf{ID}  & \textbf{V} & Title  & Start Loan & End Loan \\
+\hline
+\multirow{4}{*}{$\Body  \Big\lbrace$}  & 3 & 001 & Dracula & [15/3/2012, 5, 2] & [30/3/2012, 1, 1] \\
+& 4 & 001 & Frankenstein & [25/3/2012, 3, 2] & [4/4/2012, 1, 7]\\
+& 5 & 001 & Harry Potter & [18/3/2012, 4, 1] & [2/4/2012, 2, 2]\\
+& 3 & 002 & Dracula & [4/4/2012, 3, 3] & UC \\
+\hline\\
+\end{tabular}
+}
 
@@
 It is important to notice that whereas the result of the evaluation of any comparison between crisp time intervals is boolean, the evaluation of any comparison between \emph{PVPs} is in the unit interval.
-Therefore, for a valid-time object, say $A \in R$, in a crisp temporal database it is not possible an overlaping among any valid-time interval ($\forall I, \not \exists J: \left(A, I \right) \in R \wedge \left(A, J \right) \in R \wedge \not \Overlap\left(I, J \right)$).
-Thus, from the point of view of a fuzzy temporal database, there are the following options:
+Therefore, for a valid-time object, say $A \in R$, in a crisp temporal database it is not possible an overlapping among any valid-time interval ($\forall I, \not \exists J: \left(A, I \right) \in R \wedge \left(A, J \right) \in R \wedge \lnot \Overlap\left(I, J \right)$).
+Thus, from the point of view of a fuzzy temporal database, there are the following options (see Figure \ref{fig:tdb-class}):
 
-\begin{itemize}
-\item 
+% \begin{itemize}
+% \item 
 \begin{definition}
@@
 
-\item 
+%\item 
 \begin{definition}
- \textbf{Ill-consistent} The overlapping of two PVPs of the same object do overlap. It can be distinghised among the following sub-types:
-	\begin{itemize}
-	\item 
+ \textbf{Ill-consistent} The overlapping of two PVPs of the same object do overlap. It can be distinguished among the following sub-types:
+\end{definition}
+% 	\begin{itemize}
+% 	\item 
 	      \begin{definition}
-	      \emph{Co-existence} Two versions of the same object $A$ may exist at the same time.
+	      \textbf{Co-existence} Two versions of the same object $A$ may exist at the same time.
 		      \begin{align}
@@
 
-	\item 
+	%\item 
 	      \begin{definition}
-	      \emph{Weak-consistence} Several versions of the same object $A$ may exist at the same time.
+	      \textbf{Weak-consistence} Several versions of the same object $A$ may exist at the same time.
 		      \begin{align}
@@
 
-	\end{itemize}
-\end{definition}
+	%\end{itemize}
 
-\end{itemize}
+\vspace*{13pt}
+\begin{center}
+{
+\includegraphics[scale=1.25]{./graphs/tdb-clas.pdf}
+
+}
+\end{center}
+%\centerline{ \psfig{file=./graphs/Y-time-point.eps}}
+\vspace*{10pt}
+\fcaption{\label{fig:tdb-class}Classification for a fuzzy temporal database.
+The validity for an entity is represented by a trapezoid.
+Graphic (a) shows an example of strictly consistent database.
+Between two versions of the entity it is possible some amount of overlapping, but always holds equation \eqref{eq:stricly-consistent}.
+Graphic (b) shows an example of a co-existence.
+Two versions of the same entity may overlap in any degree, but not more than two versions.
+Graphic (c) shows an example of weak-consistence.
+The different versions of the same entity are shown in an unfolded way, to clarify the multiple overlapping.
+}
+\vspace*{13pt}
+
+%\end{itemize}
 
 In the implementation of the DML operations we will consider a strictly consistent temporal database.
+This is a natural extension of the crisp temporal databases.
+This approach allows a easy implementation for the DML operations defined in Section \ref{subsubsec:Understanding-valid-time-databases}.
+
 Since the time intervals are now possibilistic valid-time periods, PVPs, the auxiliary functions defined in equations \eqref{eq:close-a-crisp-interval} to \eqref{eq:close-current} should be re-defined.
@@
 \label{def:close-r-a-pvp}
+\textbf{CloseR$\left(I, J\right)$.}
 Consider two PVPs given by $I = \left(X, Y \right)$ and $J = \left(X', Y' \right)$.
@@
 \label{eq:close-r-a-pvp}
-\text{CloseR}\left(I, J\right) = \\
+\text{CloseR}\left(I, J\right) &= \\
 \begin{cases}
 \nonumber
-I   & \mbox{ if } Y \not \in \text{UNKNOWN}  \\
-I = \left[X, Z \right] & \mbox{ if } Y \in \text{UNKNOWN}, \\
+I   & \mbox{ if } I \neq \left[X, UC \right] \\
+I = \left[X, Z \right] & \mbox{ if } I = \left[X, UC \right] \\
 & Z \triangleq \left \lbrace C_1\left(>, X \right), C_2\left(<, X' \right) \right \rbrace
@@
 
+For example, consider $I = \left[\left[4/4/2012, 3, 3\right] , UC\right]$ and $J = \left[ \left[15/4/2012,2,1\right], UC\right]$.
+The result for CloseR$\left(I, J\right)$  is $I = \left[ \left[4/4/2012, 3, 3\right], \left[4,7,13,15 \right] \right]$.
+The value that closes $I$ is a trapezoid in the form $\left[ \alpha, \beta, \gamma, \delta \right]$ as defined in Section \ref{subsec:fuzzy-numbers}.
+
 \begin{definition}
 \label{def:pvp-current-in-relation}
-Consider an object $A$ and a relation $R$ in a fuzzy temporal database.
+\textbf{Current$\left(R, A \right)$.}
+Consider an entity $A$ and a relation $R$ in a fuzzy temporal database.
 The function Current$\left(R, A \right)$ obtains the tuple $\left(A, I \right), I = \left(X, Y \right)$ that is current in the relation $R$, that is it, $I = \left[X , \text{UNKNOWN} \right]$.
@@
 \nonumber
-I & \mbox{ if } \exists  \in I : \\
-&  (A,I) \in R \wedge Y \in \text{UNKNOWN}\\
+I & \mbox{ if } I = \left[X, UC \right] \\
+%&  (A,I) \in R \wedge Y \in \text{UNKNOWN}\\
 \emptyset & \mbox{ in any other case }
@@
 
+For example, consider the database in Table \ref{tbl:fuzzy-library-sample-valid-time}.
+The function Current$\left(R, A \right)$ with $A =\left(ID=3\right)$ returns time interval for the current version of $A$, $I = \left[ \left[4/4/2012, 3, 3 \right], UC \right]$.
+Nevertheless, the time interval for the entity with ID = 4 is the empty set.
 Now it is possible to close the current version of an object by using \eqref{eq:close-r-a-pvp} and \eqref{eq:pvp-current-in-relation}.
@@
 \label{def:pvp-close-current-version}
+\textbf{Close-current$\left(R, \left(A, J \right) \right)$.}
 Consider an object $A$, the relation $R$ and a pvp, $J$.
@@
 \label{eq:pvp-close-current}
-\text{Close-current} \left(R, A, J \right) =\\
+\text{Close-current} \left(R, \left(A, J \right) \right) =\\
 \begin{cases}
 \nonumber
-\big \lbrace R - \left(A, I \right) \cup \left \lbrace \left(A, \mbox{ CloseR } \left(I, J\right) \right) \cup \left(A, J\right)\right \rbrace  \big \rbrace \\
+\big \lbrace R - \left(A, I \right) \cup \left \lbrace \left(A, \mbox{ CloseR } \left(I, J\right) \right) \right \rbrace \cup \left \lbrace \left(A, J\right)\right \rbrace  \big \rbrace \\
 \nonumber
-\mbox{ if } I = \mbox{ Current } \left(R, A \right) \wedge J > \mbox{ CloseR } \left(I, J \right)   \\
+\mbox{ if } I = \mbox{ Current } \left(R, A \right) \\ %\wedge J > \mbox{ CloseR } \left(I, J \right)   \\
 \nonumber R , \text{ in any other case}
@@
 
+For example, consider the database in Table  \ref{tbl:fuzzy-library-sample-valid-time}.
+The function Close-current1$\left(R, \left(ID=3\right) \right.,$ $\left.
+\left[ \left[15/4/2012,2,1\right], UC\right]\right)$ closes the current version of the book with ID=3 and creates a new version with the specified time interval.
+
+\subsubsection{\label{subsubsec:modify-fuzzy-temporal}Modify}
+This operation adds new information about an existing entity $A$ in the relation $R$. The modify operation does not remove any previous value of the entity $A$.
+Note that the modify operation is only applicable when the object $A$ is current in the relation $R$; $ \left(A, I \right) \in R, I = \left[X, \UC \right]$.
+
+\begin{align}
+\label{eq:modify-fuzzy-temporal}
+\MOD \left( R, \left( A, I \right) \right) =\\
+%\begin{cases}
+\nonumber
+\mbox{ Close-current }\left(R, \left(A, I\right) \right) 
+% \mbox{ if Current}\left(R ,A \right) \neq \emptyset \\
+% R,  \mbox{ otherwise }
+%\end{cases}
+\end{align}
+
 \subsubsection{\label{subsubsec:insert-fuzzy-temporal}Insert}
-The user wants to store an object $A$ which is valid in the relation $R$ during the time interval specified by the PVP, $I = \left(X, Y \right)$.
+The user wants to store an entity $A$ which is valid in the relation $R$ during the time interval specified by the PVP, $I = \left(X, Y \right)$.
 %
@@
 To indicate that the object $A$ is current in the relation, the special value Until Current, $\UC$  is used.
-There are two main cases when performing an insert operation:
+There are the following cases when performing an insert operation:
 \begin{enumerate}
-\item The object $A$ was never in the relation $R$: The object is added with the valid-time indicated by the PVP $I$.
+\item The entity $A$ was never in the relation $R$: The entity is added with the valid-time indicated by the PVP, $I$.
+For example, consider the database given by Table \ref{tbl:fuzzy-library-sample-valid-time}. The following sentences correspond with the insertion of the first version of each book.
 
-\item The object $A$ is in the relation $R$. Depending on the value of the time interval $I$, there are two possibilities:
+\begin{verbatim}
+Insert(3,'Dracula',
+[[15/3/2012, 5, 2],
+        [30/3/2012, 1, 1]]);
+Insert(4,'Frankenstein',
+[[25/3/2012, 3, 2], 
+        [4/4/2012, 1, 7]]);
+Insert(5,'Harry Potter',
+[[18/3/2012, 4, 1], 
+        [2/4/2012, 2, 2]]);
+\end{verbatim}
+
+\item The entity $A$ is in the relation $R$. Depending on the value of the time interval $I$, there are three possibilities:
 	\begin{enumerate}
-	\item Insert $\left(A, I\right)$ in the relation $R$. If the time interval $I$ does not overlap any other valid-time interval in the relation $R$.
-	\item Reject the insertion, if the time interval $I$ do overlap any existing valid-time interval for the object $A$ in the relation $R$.
+	\item Insert $\left(A, I\right)$ in the relation $R$.
+If the time interval $I$ does not overlap any other valid-time interval in the relation $R$.
+Note that here, the result of the Overlaps operator is in the unit interval.
+For example, the librarian loans the book with ID=3 but the book is not returned yet. The insert sentence is the following.
+ \begin{verbatim}
+Insert(3,'Dracula',
+[[4/4/2012, 3, 3], UC]);
+	      \end{verbatim}
+
+	\item Modify and close the current version of the entity $A$ and insert $\left(A, I \right)$.
+For example, consider now that the book with ID 3 is loaned  around the 24/4/2012 (this is model by a triangular fuzzy number: $\left[24/4/2012, 1, 1 \right]$).
+Here the problem is that the book with ID 3 was loaned around 4/4/2012, but not returned.
+If the book is already in the library for a new loan, then it is necessary to set the return date and add a new row with the new loan date.
+%This is illustrated in Table \ref{tbl:library-sample-2}, the insert sentence is:
+	    \begin{verbatim}
+Insert(3,'Dracula',
+[[24/4/2012, 1, 1], UC]]);
+	      \end{verbatim}
+
+	\item Reject the insertion, if the time interval $I$ do overlap with a degree of 1 any existing valid-time interval for the entity $A$ in the relation $R$.
+For example, consider now that the librarian wants to introduce a past loan for the book with ID 3.
+The book was loaned around 6/4/12/2012 and returned around 25/4/2012.
+As this interval does overlaps other time intervals with a degree of 1, it is not possible that the book was loaned in that time interval.Therefore, the insertion is rejected.
+The insert sentence is:
+
+	      \begin{verbatim}
+Insert(3,'Dracula',
+[[6/4/12/2012, 1, 1],
+         [25/4/2012, 1, 1]]);
+	    \end{verbatim}
 	\end{enumerate}
@@
 \label{eq:insert-fuzzy-temporal}
-\INS \left(R ,A, I\right) =\\
+\INS \left(R ,\left(A, I \right) \right) &=\\
 \begin{cases}
 \nonumber
-R \cup \left(A, I \right), & \mbox{ if }  A \not \in R\\
-R \cup   \left(A, I \right) & \mbox{ if }  A \in R \wedge \\
-\big \lbrace \left( \exists J= \left(X', Y' \right): \left(A, J \right) \in R \wedge \right. \\
-\left. Y' \in \left \lbrace \left(<, X \right)\right \rbrace  \right) \vee \\
-\left( \exists K= \left(X'', Y'' \right): \left(A, K \right) \in R \wedge \right. \\
-\left. X'' \in \left \lbrace \left(>, Y \right)\right \rbrace \vee \right) \big \rbrace\\
+R \cup \left(A, I \right), & \mbox{ if }  \left(A, I \right) \not \in R \vee \\
+&  \left(A, J \right) \in R \wedge I \text{Overlaps} J < 1\\
+%\big \lbrace \left( \exists J= \left(X', Y' \right): \left(A, J \right) \in R \wedge \right. \\
+%\left. Y' \in \left \lbrace \left(<, X \right)\right \rbrace  \right) \vee \\
+%\left( \exists K= \left(X'', Y'' \right): \left(A, K \right) \in R \wedge \right. \\
+%\left. X'' \in \left \lbrace \left(>, Y \right)\right \rbrace \vee \right) \big \rbrace\\
 %&  J<I<K  \big \rbrace\\
-R, & \mbox{ otherwise }  
+\MOD \left(R, \left(A, I \right)\right), & \mbox{ otherwise }  
 \end{cases} 	
 \end{align}
-\subsubsection{\label{subsubsec:modify-fuzzy-temporal}Modify}
-The delete operation logically removes a current object $A$ which is valid in the relation $R$:
+\subsubsection{\label{subsubsec:delete-fuzzy-temporal}Delete}
+The delete operation logically removes an entity $A$ which is valid in the relation $R$:
 \begin{align}
@@
 \nonumber
-R - \left(A, I \right), & \mbox{ if } \left( A, I \right) \in R\\
+R - \left(A, I \right), &  \forall \left( A, I \right) \in R\\
 R, & \mbox{ otherwise }  
@@
 
-\subsubsection{\label{subsubsec:delete-fuzzy-temporal}Delete}
-This operation adds new information about an existing object $A$ in the relation $R$. The modify operation does not remove any previous value of the object $A$.
-Note that the modify operation is only applicable when the object $A$ is current in the relation $R$; $ \left(A, I \right) \in R, I = \left[X, \UC \right]$.
+For example, consider that the librarian wants to delete the history for the book with ID = 3. The following sentence deletes all the versions for the book with ID = 3.
 
+\begin{verbatim}
+ Delete(3,'Dracula');
+\end{verbatim}
+
+\subsubsection{\label{subsubsec:revise-fuzzy-temporal}Revise}
+This operation replaces the old values for the entity $A'= \left(a_1', \ldots, a_n' \right) $ with the new values specified by $A = \left(a_1, \ldots, a_n \right)$ in the relation $R$, at a given time interval $I$.
+This operation is used when a correction on the values for the entity $A$ should be done.
 \begin{align}
-\label{eq:modify-fuzzy-temporal}
-\MOD \left( R, A, I \right) =\\
+\label{eq:revise-fuzzy-temporal}
+\REV\left(R, \left(A, I \right)\right)=\\
 \begin{cases}
 \nonumber
-\mbox{ Close-current }\left(R, A, \mbox{Current}\left(R, A \right) \right) \\
- \mbox{ if Current}\left(R ,A \right) \neq \emptyset \\
- R,  \mbox{ otherwise }
+ R - \left(A', I \right) \cup \left \lbrace\left(A, I \right)\right \rbrace  & \mbox{ if } \left(A', I \right) \in R \\%: A = \left(a_1, \ldots, a_n \right),\\
+%A'= \left(a_1', \ldots, a_n' \right)   \\   
+R, & \mbox{ otherwise }
 \end{cases}
@@
 
-\subsubsection{\label{subsubsec:revise-fuzzy-temporal}Revise}
-This operation changes the value of the object $A$ in the relation $R$, at a given time interval $I$. This operation is defined when a correction on the values for the object $A$ should be done.
-\begin{align}
-\label{eq:revise-fuzzy-temporal}
-\mbox{ Revise}\left(R, A, I \right)=\\
-\begin{cases}
-\nonumber
- R - \left(A', I \right) \cup \left(A, I \right)  \\
-  \mbox{ if } \left(A', I \right) \in R: A = \left(a_1, \ldots, a_n \right),\\
-A'= \left(a_1', \ldots, a_n' \right)   \\   
-R,  \mbox{ otherwise }
-\end{cases}
-\end{align}
+For example, consider now that the librarian wants to make a correction in the title for the book with ID = 5. The new title is `Harry Potter IV'. The revise sentence is:
+
+\begin{verbatim}
+Revise(5,'Harry Potter IV', 
+[[18/3/2012, 4, 1], 
+          [2/4/2012, 2, 2]]);
+\end{verbatim}

### 2012-06-12 trunk/ijcis-pos-model/sources/intro.tex
My contributions so far...

--- 2012-04-15 ijcis-pos-model/sources/intro.tex
@@
 
+%From information systems in general to database systems in particular:
+Generally, \emph{information systems} model the structure and behavior of real objects, concepts or processes.
+A specific type of information systems are \emph{database systems}, which are computer systems designed to manage \emph{databases}.
+A database is basically a collection of (persistent) data.
+These (typically atomic) data represent real objects or concepts.
+In the context of database design, such objects or concepts are typically called \emph{entities} and a collection of similar entities is typically modelled by an \emph{entitytype}, which is basically a combination of a name and a list of \emph{attributes}, which describe properties of the entities.
+Typically, in an early stage of the database development process, the structures of and interactions or relationships between used entities are modelled in a \emph{design schema}, following some design model.
+A database thus contains (typically atomic) data.
+Each (atomic) part of these data is a result value of a measurement of a property of an entity or a description of a property of an entity and will correspond to the attribute of the entity's entitytype, which describes the property.
+
+%From database models in general to the relational model in particular:
+The structure and behavior of a database, along with some integrity and security restrictions, is dictated by a \emph{database model}.
+A database model is basically a collection of instructions and regulations, used to (logically) describe the structure and behavior of a database, along with some integrity and security restrictions.
+Applying a database model to a concrete design schema results in a \emph{database schema}, which models the logical structure and behavior of a database.
+Several different database models exist, but the most popular is the \emph{relational database model}~\cite{Codd:1970:RMD:362384.362685}.
+Following the relational database model, entitytypes are modelled as \emph{relations}, which comprise a name and a list of attributes, modelling the entitytype's attributes.
+
+%From time in general to time in information systems
+In reality, some aspects or properties of objects, concepts or processes are time-variant or time-related.
+For example, the moment of a bank transaction is traditionally a moment in time and thus a time-related notion, the function of an employee in a company can change through recorded history and is thus time-variant.
 The concept of time itself is very complex to handle and interpret~\cite{Klein1994},~\cite{Shackle1961}, though it is very natural and omnipresent.
 As information systems often attempt the modelling of natural objects, concepts or processes, they often require modelling temporal aspects or concepts.
-Thus, several proposals have been concerned with the obtaining of theoretical models that allow the modelling or representation of time~\cite{Bolour1982},~\cite{VanderCruyssen1997}.
+Thus, several proposals have been concerned with obtaining theoretical models that allow the modelling or representation of time~\cite{Bolour1982},~\cite{VanderCruyssen1997}.
 
-A very specific type of information systems are database systems, which are computer systems designed to manage databases.
-A database contains data representing real objects or concepts.
-Each (atomic) part of these data is a result value of a measurement of a property or a description of a property of a real object or concept.
-In reality, some aspects or properties of objects or concepts are time-variant or time-related.
-e.g., the moment of a bank transaction is traditionally a moment in time and thus a time-related notion, the function of an employee in a company can change through recorded history and is thus time-variant.
-A temporal database schema is a database schema that models real objects or concepts with time-related or time-variant properties.
-However, the modelling of temporal aspects has a direct impact on the consistency of the temporal database, because the temporal nature of these aspects imposes extra integrity constraints.
+%From time in information systems to time in temporal databases
+Time is a complex concept, which makes modelling time in database systems complex too.
+A \emph{temporal database} is a database that deals with certain time-related aspects in its schema: a \emph{temporal database schema}~\cite{Dyreson1994} is a database schema that models entities (and interactions or relationships between entities) with time-related or time-variant properties.
+However, the modelling of temporal properties or aspects has a direct impact on the consistency of the resulting temporal database, because the temporal nature of these aspects or properties imposes extra integrity constraints.
 An example.
-Consider a relation in a relational library database, modelling the presence of books in the library.
+Consider a relation in a relational library database, representing the physical absence (or presence) of books in the library.
 Every physical book is represented by a unique identifier.
 Every record in the relation contains such an identifier, a date on which the corresponding book was loaned and a date on which it was subsequently returned (if it was returned).
+As such, every record represents a period of time during which the physical book corresponding with the identifier is loand to a library customer.
 Without further precautions, a library employee could add several records with the same book identifier, different `loaned'-dates and no `returned'-dates.
@@
 
+%Research about temporal databases
 A lot of research concerns temporal database models and their approaches to the modelling of time.
-The first efforts were towards the representation of historical information related to objects represented by records in a database~\cite{Clifford1985}.
+Some of the first proposals concerned the representation of historical information related to entities~\cite{Clifford1985}.
 Some proposals tried to extend the Entity Relationship Model~\cite{Klopprogge1983}, without impact on any database standards like SQL~\cite{Sarda1990}.
+Notably, in 1994, `A Consensus Glossary of Temporal Database Concepts' was published~\cite{Dyreson1994}.
+For this publication, 44 temporal database researchers, among which some of the main researchers in this field, cooperated to reach a consensus on the nature and definitions of some of the main temporal database concepts and their terminology.
+This glossary was subsequently updated in 1998~\cite{Dyreson1998}.
+In the presented work, the concepts and terminology from this glossary are used and followed.
 
-Notably, in 1994, ``A Consensus Glossary of Temporal Database Concepts'' was published~\cite{Dyreson1994}.
-For this publication, 44 temporal database researchers, among which some of the main researchers in this field, cooperated to reach a consensus on the nature and definitions of some of the main temporal database concepts and terminology.
-This glossary was subsequently updated in 1998~\cite{Dyreson1998}.
+%Research about temporal relationships
+An interesting issue in temporal modelling concerns the relationships between temporal notions.
+Notably, Allen~\cite{Allen1983} studied temporal relationships between time intervals~\cite{Dyreson1994} (and, as a special case, instants~\cite{Dyreson1994}).
+Among others, the querying of temporal databases has greatly profited from these temporal relationships, because they allow for richer and more complex user-specified temporal query demands, by allowing to express more complex relationships between the temporal notions in the temporal expressions in the query and the temporal indications in the database.
+For example, given a relation representing who was department head of an institution during which time intervals, a query like `Who were the department heads during the time intervals when Thomas worked for the institution?' can be evaluated using similar relationships.
+%TODO: CHECK THIS LAST SENTENCE
 
-An interesting issue in temporal modelling concerns relationships between temporal notions.
-Notably, Allen~\cite{Allen1983} studied temporal relationships between time intervals (and as a special case time points).
-Among others, the querying of temporal databases has greatly profited from these temporal relationships, because they allowed for richer and more complex user-specified temporal query demands, by allowing to express more complex relationships between the temporal notions in the temporal expressions in the query and the temporal indications in the database.
-e.g., in a relation modelling who was department head of an institution during which periods of time, a query like `who were the department heads when Thomas worked for the institution' can be evaluated using similar relationships.
+%Now introducing imperfections in representations of time...
+Humans handle temporal information using certain temporal notions like specific time intervals or instants~\cite{Dyreson1994} and they often have to deal with imperfections like imprecisions, vaguenesses, uncertainties or inconsistencies possibly contained in (the descriptions of) these temporal notions.
+Among many others, these possible imperfections in (descriptions of) temporal notions determine an important issue in temporal modelling.
+As an example, consider the temporal notion described in a sentence like `The Belfry of Bruges was finished on a day somewhere between 01/01/1201 A.D.
+and 31/12/1300 A.D.'.
+The description of the temporal notion reads `a day somewhere between 01/01/1201 A.D.
+and 31/12/1300 A.D.'.
+This temporal notion thus contains imperfection in the form of the uncertainty about the exact day the last stone of the building was laid: it is known that the building was finished on a single day, but it is not known which day this was.
+To allow information systems to cope with these and similar temporal data imperfections, many approaches adopt fuzzy sets~\cite{Zadeh1965} and/or fuzzy logic to model temporal information~\cite{Mitra1994},~\cite{Nagypal2003},~\cite{Billiet2011},~\cite{Dubois2003}.
 
-Humans handle temporal information using certain temporal notions like time intervals or time points~\cite{Dyreson1994}, and they often have to deal with imperfections like imprecisions, vaguenesses, uncertainties or inconsistencies possibly contained in the descriptions of these temporal notions.
-Among many others, these possible imperfections in descriptions of temporal notions determine an important issue in temporal modelling.
-e.g., the description of the temporal notion in a sentence like `The Belfry of Bruges was finished on a day somewhere between 01/01/1201 A.D.
-and 31/12/1300 A.D.' contains imperfection because of the uncertainty in the used time-related expression.
-It is known that the building was finished on a single day, but it is not known precisely which day this was.
-
-To allow information systems to cope with these and similar data imperfections, many approaches adopt fuzzy sets~\cite{Zadeh1965} for the representation of temporal information~\cite{Mitra1994},~\cite{Nagypal2003},~\cite{Billiet2011},~\cite{Dubois2003}.
+%Now introducing imperfections in temporal relationships
 The temporal relationships studied by Allen were fuzzified by several authors~\cite{Ohlbach2004},~\cite{Nagypal2003},~\cite{Schockaert2008}.
 Garrido et al.
-~\cite{Garrido2009} present different temporal operators, defined by a combination of regular fuzzy comparisons.
-Both~\cite{Garrido2009} and~\cite{Pons2011} deal with uncertainty in temporal expressions concerning time intervals.
+~\cite{Garrido2009} present different temporal operators, defined by a combination of fuzzy comparisons.
+Also,~\cite{Garrido2009},~\cite{Pons2011} studied uncertainty in the context of time intervals.
 Other approaches, like~\cite{Qiang2009}, use rough sets~\cite{Pawlak1995} to represent time intervals.
 
-Next to temporal modelling, some attention has been spent on temporal reasoning~\cite{Allen1983}.
-Although temporal reasoning is not discussed in this chapter, it should be noted that, among others, Dubois and Prade et al.~\cite{Dubois2003},~\cite{DuBois1989} have dealt with fuzziness and uncertainty in temporal reasoning.
+%Imperfections in temporal reasoning
+Next to temporal modelling, some attention has gone to temporal reasoning~\cite{Allen1983}.
+Although the focus of this paper is temporal modelling, it should be noted that, among others, Dubois and Prade et al.~\cite{Dubois2003},~\cite{DuBois1989} have dealt with fuzziness and uncertainty in temporal reasoning.
+
+%Overview of this work
+%Include: what is the paper about? (situation and explanation of the problem we will solve and situation of our solution (we use constraints...))
+%Include: how does the solution work? (short summary of how the solution will solve the problem/attend to the problem)
+%Include: What exactly will be presented
+%In all of these: stress every single innovative/novel point of the presented work!
+%TODO: build the overview when the paper is finished
 

### 2012-06-12 trunk/ijcis-pos-model/sources/temporaldatabases.tex
My contributions so far...

--- 2012-04-16 trunk/ijcis-pos-model/sources/temporaldatabases.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%Time in databases: temporal databases: the crisp stuff. General things.
 \subsection{\label{subsec:time-in-databases}Time in Databases}
-The concept of time has been studied in databases for a long time.
-A true standard for adding temporal aspects to relational databases does not exist, but there is a consensus in the literature \cite{Dyreson1994} on what is called a \emph{temporal database}: a temporal database is a database dealing with some aspects of time in its schema.
-In a temporal DBMS, a \textbf{chronon} is the shortest duration of time supported by the system.
-In temporal databases, some temporal attributes can be managed without treating the attribute differently from non-temporal attributes.
-The time described by such an attribute is called \textbf{user defined time} (\emph{UDT}).
-In addition to UDT, the following types of time can be discerned in a temporal database, all of which are handled exceptionally by the DBMS:
+The concept of adding time to databases has been profoundly studied~\cite{Darwen1998},~\cite{Etzion1998},~\cite{Etzion1998b},~\cite{Gal1998},~\cite{Jensen1999},~\cite{Lomet2006},~\cite{Nascimento1995},~\cite{Sarda1990},~\cite{Snodgrass1998},~\cite{Toman1998},~\cite{Wu1998}.
+A true standard for adding temporal aspects to relational databases does not exist, but there is a consensus on what is called a \emph{temporal database}: a temporal database is a database that supports some aspects of time~\cite{Dyreson1994}.
+Thus, a temporal database will contain data representing \emph{temporal values}.
+Notions of time, such as instants~\cite{Dyreson1994}, time intervals~\cite{Dyreson1994} or temporal elements~\cite{Dyreson1994} are generally called `temporal values' in the context of temporal databases.
+Based on their interpretation and modelling purpose, time notions and the corresponding times can be classified into four types.
+Of these types, \emph{user-defined time} comprises time indications of which the modelling has no impact on the consistency of the temporal database.
+The data representing these temporal values are not handled differently from data of non-temporal attributes (by the temporal DBMS)~\cite{Dyreson1994}.
+The other types contain time indications typically represented by data which are handled exceptionally by the temporal DBMS.
+The definitions and explanations of these types can be found in~\cite{Dyreson1994} and~\cite{Nascimento1995} and more information can be found in~\cite{Jensen1991},~\cite{Snodgrass1984} and~\cite{Nascimento1995}.
 
 \begin{itemize}
-	\item
-	\textbf{Transaction time} (\emph{TT}) \cite{Rowe1987},\cite{Jensen1991} denotes the time when the fact (object) is stored in the database.
-It is usually append-only: as the past can not be changed, TT can not be changed neither.
-Furthermore, at the moment of insertion, a TT can be neither in the past nor in the future.
-	\item
-	\textbf{Valid time} (\emph{VT}) \cite{Jensen1994},\cite{Sarda1990},\cite{McKenzie1981} denotes the time when the fact (object) is true in the modelled reality.
-A fuzzy extension has been proposed by \cite{Garrido2009}.
-%	\item
-%	User defined time: It is an uninterpreted attribute. The domain is date/time. The query language has no special support for it.
-	\item
-	\textbf{Decision time} (\emph{DT}) \cite{Nascimento1995},\cite{Chakravarthy1994},\cite{Etzion1992},\cite{Ozsoyoglu1995} denotes the time when an event was decided to happen. 
+	\item \emph{Valid time}: The \emph{valid-time} (VT) of a fact is the time when the fact is true in the modeled reality~\cite{Dyreson1994}.
+	\item \emph{Transaction time}: A database fact is stored in a database at some point in time, and after it is stored, it is current until logically deleted.
+The \emph{transaction-time} (TT) of a database fact is the time when the fact is current in the database and may be retrieved~\cite{Dyreson1994}.
+	\item \emph{Decision time}: The \emph{Decision time} (DT) of an event is the time when an event was decided to happen~\cite{Nascimento1995}.
 \end{itemize}
 
-	E.g., consider a database containing employee contract descriptions.
-The time when the employee's contract is valid, represented by an interval, is VT.
-The time when the employee's contract is stored in the database is the TT.
-The time when the decision for hiring this employee was made is the DT.
+An example: consider a database containing data representing the capacities of various aspects of employee contracts.
+The time during which an employee's contract is valid and the employee is hired, will then be VT.
+The moment on which an employee's contract is stored in the database, will then be TT.
+The time when the decision for hiring this employee was made, will then be DT.
 
-% is a non-decomposable unit of time.  it  There are two ways to represent a chronon: as a point or as an interval \cite{655777}.
+Depending on the time represented in the database, a database is classified as either a \emph{valid-time database} (VTDB), a \emph{transaction-time database} (TTDB), a \emph{decision-time database} (DTDB), a \emph{bitemporal database} (both valid-time indications and transaction-time indications are represented), or a \emph{tritemporal} or \emph{multitemporal database} (valid-time indications, transaction-time indications and decision-time indications are represented).
 
-When working with these time concepts, the Data Manipulation Language (\emph{DML}, which is part of the standard database querying language SQL) is extended to deal with possible temporal inconsistencies within the data and to handle more complex (temporal) queries.
-%Next to these concepts, also \textbf{user defined time} (\emph{UDT}) is discerned.
-UDT is an uninterpreted attribute in the date/time domain.
-This means that the attribute uses the date/time domain, but the database model does not treat the attribute differently from non-temporal attributes.
-Depending on the time managed, a database is classified as either a \textbf{Valid Time Database} (\emph{VTDB}), a \textbf{Transaction Time Database} (\emph{TTDB}), a \textbf{bi-temporal database} (both valid and transaction time are managed) or a \textbf{tri-temporal} or \emph{multitemporal} database (valid time, transaction time and decision time are managed).
+%Now in the context of a relational database: what is a VT/TT/... relation? We will use VT relations, thus we should explain what they are.
+In the context of relational databases, more specific definitions exist.
+For example, a \emph{valid-time relational database} is a relational database containing one or more \emph{valid-time relations}, a \emph{transaction-time relational database} is a relational database containing one or more \emph{transaction-time relations} and a \emph{bitemporal relational database} is a relational database containing one or more \emph{bitemporal relations}~\cite{Dyreson1994}.
+A valid-time relation is then a relation with exactly one system-supported VT, a transaction-time relation is a relation with exactly one system-supported TT and a bitemporal relation is a relation with exactly one system-supported VT and exactly one system-supported TT~\cite{Dyreson1994}.
+The attributes used in a valid-time relation to describe VT are called \emph{valid-time attributes}.
+The attributes used in a transaction-time relation to describe TT are called \emph{transaction-time attributes}.
+Any attribute unrelated to VT, TT or DT is usually called a \emph{non-temporal attribute}.
 
-%\subsection{Temporal Elements}
-%%	
-%	In order to define a temporal database model, some basic elements should be defined \cite{Dyreson:1994:CGT:181550.181560}:
-%	\begin{itemize}
-%	\item A	\textbf{chronon} is a non-decomposable unit of time.
-In a temporal DBMS, it is the shortest duration of time supported by the system.
-There are two ways to represent a chronon: as a point or as an interval \cite{655777}.
-%	\item
-%	\emph{Event}: An instant of time. Usually, an event is said to be occur during time $t$ if it occurs during the chronon represented by $t$.
-%	\item
-%	\emph{Interval}: The time between two events. The representation is very often a set of contiguous chronons.
-	%\item
-	%\emph{Span}: A directed duration of time.
-A duration is an amount of time with known  lenght but no specific starting or ending chronons.
-The span can be either positive, denoting a forward motion of time or negative, denoting a backward motion of time.
-%	\item A	\textbf{timestamp} is a time value associated with some object in the database.
+%In a temporal database system, a \emph{chronon} is the shortest duration of time supported by the system.
+% is a non-decomposable unit of time.
+it  There are two ways to represent a chronon: as a point or as an interval \cite{655777}.
+%WE NEVER USE CHRONONS IN THIS PAPER!
+Typically, when time modelling is supported by a database system, some behavioral aspects of the database model are extended or enhanced and some extra behavioral aspects are introduced, to deal with possible temporal inconsistencies and to allow more complex (temporal) queries.
+The presented work only deals with valid-time relational databases.
+In the presented work, a general extended/enhanced version of the relational database model is used.
+This extended version is based on several works, mainly the work presented in~\cite{Jensen1994}.
+In subsection~\ref{subsubsec:Understanding-valid-time-databases}, the extensions to the relational database model used in this work are presented on a logical level, but first, in subsection~\ref{example-situation}, an example situation is presented, which will be used throughout this paper to explain the various concepts treated in light of the presented work.
 
-%	\item The	\textbf{lifespan} (of a database object)is the time over which it is defined.
-The lifespan of a valid time object denotes the time when the corresponding object exists.
-The lifespan of a transaction time object is the value of the timestamp.
+%Thus, in the presented work, an extended version of the Data Manipulation Language DML, which is part of the standard database querying language SQL, is used.
+This extended version is based on several works, mainly the work presented in~\cite{Jensen1994}.
+In subsection~\ref{subsubsec:Understanding-valid-time-databases}, the extensions to the DML used in this work are presented, but first, in subsection~\ref{example-situation}, an example situation is presented, which will be used throughout this paper to explain the various concepts treated in light of the presented work.
 
-%	Depending on the type of time the meaning is different:
-%		\begin{itemize}
-%		\item
-%		Lifespan of a valid time object: Refers the time when the corresponding object exists.	
-%		\item
-%		Lifespan of a transaction time object: The transaction time of a database object refers when the object is stored in the database. The lifespan is the value of the timestamp.		
-%		\end{itemize}
-%	\end{itemize}
+\subsection{\label{example-situation}Example Situation}
+Consider a valid-time relation modelling the absence through loaning of books in a library.
+The relation contains 4 attributes: `ID', `Title', `Start Loan' and `End Loan'.
+Every physical book is given a unique identifier.
+This identifier is represented by a number and described by attribute `ID'.
+Attribute `Title' describes the title of a book.
+Attributes `Start Loan' and `End Loan' describe the starting date, respectively ending date on which a valid-time interval starts, respectively ends.
+This time interval is the time interval during which the book, represented by the value of the tuple for the `ID' attribute, was loaned to a customer.
+Table \ref{tbl:library-sample} visualises this relation with a few tuples in the relation extention.
 
-%	In the temporal database thesaurus, \emph{'Snapshot'} is the word for non-temporal matters.
-As a temporal database is a generalization for relational databases, an \emph{snapshot database} is a relational database.
-Furthermore, a \emph{snapshot relation} is a relation incorporating neither valid nor transaction time.
-
-%\subsection{Main issues when dealing with time}
-%Among others, the following problems are present when dealing with time in a database:
-
-%	\begin{itemize}	
-%	\item \textbf{Granularity} denotes a partition on the set of chronons.
-The conversion between several granularities is studied in \cite{Lin97efficientconversion}.
-Granularity is the base of the temporal model in \cite{Cru97}.
-An object-oriented implementation is in \cite{624013}.
-%	\item	To ensure \textbf{consistency}, temporal databases usually redefine the primary key of a relation.
-The new primary key takes into account the presence of the time.
-In order to keep the consistency, the DML is re-defined.
-For example, in a VTDB, the temporal update sentence is usually composed by an update sentence (closing the old row) plus an insert sentence (creating the new row).
-%	\end{itemize}
-%Guy's suggestion: instead of imprecision, use imperfection which is more general.
-\subsubsection{Imperfection and time}
-Representing imprecision and its semantics when dealing with time has been studied for a long time.
-Several proposals for representing and computing imprecise time indications can be found in \cite{DeCaluwe1997} and \cite{DeTre1997}.
-Also, the changes between several granularities can be seen as a source of imprecision \cite{Devos1998}.
-
-In the proposal section we will consider two kinds of imprecision:
-\begin{itemize}
-\item \textbf{Uncertainty in the database} denotes the uncertainty that arises when the knowledge about the temporal data in the database is uncertain.
-E.g., a database record shows that \emph{`The car is in the garage around April.'}
- \item \textbf{Imprecision in the query specification} denotes the imprecision in the specification of temporal criteria by the user, when querying.
-E.g., \emph{`The user wants to obtain a car which is red and which is in the garage around April.'}
-\end{itemize}
-
-\subsubsection{Representation}
-Several proposals for managing uncertain time in a database exist.
-Some proposals work with rough sets \cite{Qiang2009}, other proposals rely on possibility distributions for representing uncertainty in time \cite{Dyreson1998},\cite{Garrido2009}, \cite{Galindo2001}.
-In order to compare temporal possibility distributions, extensions of the classical Allen's operators \cite{Allen1983} are defined in \cite{Ohlbach2004}, \cite{Nagypal2003},\cite{Dubois2003a} and \cite{Schockaert2008}
-%In the proposal section, we will follow the representation by means of possibility distributions, in order to work with both satisfaction and dissatisfaction degrees.
-Also, in order to work properly with fuzzy operators, the underlying domain should be numeric.
-% Take a look into this:
-%In this paper, the representation for the dates will follow the Julian Day Number (JDN) representation \cite{Dir96}.
-
-%If the starting points and/or the end points of the interval representing the time are not known precisely, it is easy to fuzzify them, using, e.g., two triangular possibility distributions.
-
-In order to deal with uncertainty in time intervals, several proposals are made.
-Here, two approaches are introduced: the first one, based on \emph{Fuzzy Validity Period}~\cite{Garrido2009} and the second one based on \emph{Possibilistic Valid-time Period}~\cite{JoseEnriquePons2012}.
-
-\begin{definition}
-A \textbf{Fuzzy Validity Period} (\emph{FVP}) is defined as a fuzzy time interval specifying when an object is valid. A fuzzy time interval is then the fuzzification of a crisp time interval.
-\end{definition}
-Several options to transform possibility distributions corresponding to the fuzzy starting point and the fuzzy end point into one consistent FVP exist \cite{Garrido2009}, e.g (Fig.
-\ref{fig:fuzzy-validity-period}):
-\begin{itemize}
-\item The \textbf{convex hull} approach is the most intuitive approach. The resulting FVP is the convex hull of the union of both fuzzy sets.
-\item The \textbf{uncertainty preserving} approach is less intuitive but more realistic.
-The amount of uncertainty is maintained at the edges of the possibility distribution representing the FVP \cite{Garrido2009}.
-\end{itemize}
-
-%%%%%%
-% FUZZY VALIDITY PERIOD
-%%%%%%
-\vspace*{13pt}
-\begin{center}
-{
-\includegraphics[scale=0.25]{./graphs/comparisoncv.pdf}
-
-}
-\end{center}
-%\centerline{ \psfig{file=./graphs/Y-time-point.eps}}
-\vspace*{10pt}
-\fcaption{\label{fig:fuzzy-validity-period}Transformation to obtain the FVP.
-The top graph shows the two triangular possibility distributions.
-The middle graph shows the convex hull validity period, the bottom one shows the result of the second transformation, which maintains the imprecision.}
-%  \label{fig:fuzzy-validity-period}
-\vspace*{13pt}
-
-\begin{definition}
-A \textbf{Possibilistic Valid-time Period} (\emph{PVP}) is an ill-known interval in time specifying when an object is valid.
-\end{definition}
-A PVP is an ill-known interval in the sense of definition \ref{def;possibilistic-variable}, section \ref{subsec:possibilistic-variables}.
-Note that this representation is \emph{disjunctive}: the PVP represents only one crisp time interval, but that for some reason it is (partially) unknown.
-
-The ill-known interval approach has many advantages from the representation by the FVP as demonstrated in \cite{Pons2011},\cite{Pons2012}.
-Table \ref{tbl:comparative-pvp-fvp} is a comparative between PVP and FVP.
-The following list defines the items in the comparative:
-
-\begin{enumerate}
-\item Domain: The domain of the possibility distribution modelled by the approach.
-\item Implementation of relationships: How to implement a relationship.
-\item Allen's relations: Are the Allen's relations defined?
-\item Storage: The way the data is stored in the database.
-\item Possibility measures: Does the framework provides always a possibility measure for any relation between the temporal elements?
-\item Necessity measures:  Does the framework provides always a necessity measure for any relation between the temporal elements?
-\end{enumerate}
-
-%%
-%% comparison table between pvp and fvp.
-%%
-\vglue13pt
-%\begin{table}[htbp]
-\tcap{Comparative PVP vs FVP}
+\tcap{\label{tbl:library-sample}Example library relation.}
 %\centerline{\small DATA TYPES}
@@
 \centerline{\small\baselineskip=13pt
-\begin{tabular}{c p{2cm} p{2cm}}\\
-Item & PVP & FVP\\
+\begin{tabular}{c c c c}\\
 \hline
-(1) & $\Pow(\R)$ & $\R$ \\
-(2) & Ill-known constraints. & Ad-hoc operators. \\
-(3) & $\checkmark$ & - \\
-(4) & Two distributions one for each endpoint. & Only one distribution. \\ 
-(5) & $\checkmark$ & $\checkmark$ \\
-(6) & $\checkmark$ & - \\
+\textbf{ID}  & Title  & Start Loan & End Loan \\
+\hline
+3 & Dracula & 15/3/2012 & 30/3/2012 \\
+4 & Frankenstein & 25/3/2012 & 4/4/2012 \\
+5 & Harry Potter & 18/3/2012 & 2/4/2012\\
+3 & Dracula & 4/4/2012 & UC \\
 \hline\\
 \end{tabular}
-\label{tbl:comparative-pvp-fvp}
 }
 
-In the rest of the paper we will work only with PVP to represent valid-time intervals.
+%Valid-time databases and relations: general things AND an explanation about the exact things we will use in the rest of the paper.
+(+ vocabulary/concept explanation).
+Also explain the working of crisp valid-time databases/relations.
+\subsection{\label{subsubsec:Understanding-valid-time-databases}General Valid-time Databases}
+This subsection is devoted to describing the general behavior of general valid-time relational databases.
+This behavior will be described on a logical level, by (logically) defining and describing general forms of the four basic functions typically present in any relational database model.
+These functions are `create', `read', `update' and `delete' and will be denoted `the CRUD functions'.
+Usually, the CRUD functions in a valid-time relational database model are defined using the CRUD functions in a non-temporal relational database model.
+For example, the update function in a valid-time relational database model will typically be defined using several non-temporal create and update functions.
+In the presented work, the CRUD functions in the non-temporal relational database model will be given a subscript `NT', while the CRUD functions in the valid-time relational database model will be given a subscript `VT'.
 
-\subsection{Understanding Valid-time Databases}
-This subsection is devoted to describe the behaviour of a crisp valid-time database.
-For the shake of simplicity, only the three main operations in the Data Manipulation Language \emph{DML} a.k.a \emph{CRUD} (CReate Update, Delete) operations are shown.
+Before defining and describing the CRUD functions used in the presented work, some concepts should be described and defined and the corresponding terminology and denotation style of these concepts should be determined.
+
+%For the sake of simplicity, only the three main operations in the Data Manipulation Language \emph{DML} a.k.a \emph{CRUD} (CReate Update, Delete) operations are shown.
 Usually the DML operations in a temporal database are re-defined e.g.
@@
 
+Every valid-time relation has one VT.
+In the presented work, indications of VT which are represented in a valid-time relation are always time intervals.
+In the presented work, such a valid-time interval is represented in the valid-time relation using two instants~\cite{Dyreson1994}: a starting instant and an ending instant.
+
+Consider a valid-time relation $R$ with attribute tuple $\left(A_{0},\ A_{1},\ A_{2},\ \ldots,\ A_{n},\ Start,\ End\right)$, where every $A_{i}$, $i \in \Nat$ and $i \geq 1$ and $i \leq n$ and $n \in \Nat_{0}$, denotes a non-temporal attribute and where $Start$ and $End$ denote temporal attributes, respectively describing starting and ending instants.
+Now consider an entity $E$ of which the properties are described by the aforementioned attributes $A_{i}$ and a time interval $I$ defined by instants $I_{s} \text{ and } I_{e}$ where $I_{s}$ and $I_{e}$ denote instants.
+Let the values representing these instants be denoted $s \in \T$ and $e \in \T$ respectively, $\T$ being the domain of the temporal attributes and let the representation of the time interval I be denoted $\left[s,\ e\right]$.
+Let the value representing the state of $E$ for attribute $A_{i}$ during time interval $I$ be denoted $a_{i}$, for every $i \in \Nat$ for which $i \geq 1$, $i \leq n$, and let the n+1-tuple $\left(a_{0},\ a_{1},\ a_{2},\ \ldots,\ a_{n}\right)$ be denoted $a$.
+The denotation
+
+\begin{equation}
+\label{eq:def-state}
+\left(a,\ \left[s,\ e\right]\right)
+\end{equation}
+
+now expresses that the state of $E$ represented by $a$ is valid in the modelled reality during time interval $I$.
+This denotation is also used to indicate the n+3-tuple $\left(a_{0},\ a_{1},\ a_{2},\ \ldots,\ a_{n},\ s,\ e\right)$.
+The denotation
+
+\begin{equation}
+\label{eq:def-element}
+\left(R,\ \left(a,\ \left[s,\ e\right]\right)\right)
+\end{equation}
+
+now expresses that the n+3-tuple $\left(a,\ \left[s,\ e\right]\right)$ = $\left(a_{0},\ a_{1},\ a_{2},\ \ldots,\ a_{n},\ s,\ e\right)$ is an element of the extention of $R$.
+
+%\begin{definition}
+%Consider a relation $R$ in a temporal database, an entity $A$ given by the attributes $\left(a_1, \ldots, a_n \right)$ and the crisp time interval $I$ given by $\left[a,\ b\right] : a,b \in \T, a \leq b$.
+The pair $\left(A, I\right)$ expresses that the data regarding the entity A is valid during the crisp time interval I.
+Note that the relation $R = \left(A, I \right) x \left(A, I \right)$.
+The following equation indicates that the pair $\left(A, I\right)$ is in the relation R:
+%\begin{equation}
+%\label{eq:rel-def}
+%\left( R, \left(A , I\right) \right)
+%\end{equation}
+%\end{definition}
+
+%\begin{example} \textcolor{red}{TODO: INTEGRATE THIS EXAMPLE IN THE `GENERAL EXAMPLE'}
+%\label{ex:library-database}
+%Consider a library database.
+Each physical book is stored with an identifier (ID), a title and the start and end loan dates respectively.
+Table \ref{tbl:library-sample} illustrates the first version of the database, after three insertions.
+In this example, $A = ($ ID, Title $)$ and $I = ($ Start Loan, End Loan $)$.
+%\end{example}
+
+%\vglue13pt
+%\begin{table}[htbp]
+
+%In order to simplify the algorithms for the manipulation of data, some auxiliary functions and constants are defined:
+In order to simplify the definition of the CRUD functions, some auxiliary functions and constants are first defined.
+
+For the valid-time relation $R$ to be able to represent that the state of entity $E$, represented by $a$, is currently valid, the time domain used by the temporal attributes ($Start$ and $End$) should contain a special constant called \emph{Until Changed} (\emph{UC}).
+
 \begin{definition}
-Consider a relation $R$ in a temporal database, an object $A$ given by the attributes $\left(a_1, \ldots, a_n \right)$ and the crisp time interval $I$ given by $\left[a,\ b\right] : a,b \in \T, a \leq b$.
-A valid-time object $A$ in the relation $R$ is specified by:
+ \label{def:until-changed}
+\textbf{Until Changed (\emph{UC}).}\\
+Given a time domain $\T$, \emph{Until Changed} is a single value in $\T$ representing an instant infinitely far in the future.
+\end{definition}
+\vspace{10pt}
+
+This constant is special in that sense that if the value of a tuple of $R$ for attribute $End$ is $UC$, the state, represented by the values of this tuple for the non-temporal attributes, is considered to be valid currently and during an unlimited amount of time in the future.
+
+%To show that the data related to the entity $\left(A, I \right)$ is up-to-date, a special constant called \emph{Until Changed} is defined for the endpoint, $d$, of the time interval $I = \left(c, d\right)$.
+The value for this constant is considered to be $+\infty$.
+
+\begin{definition}
+\label{def:current-tuple}
+\textbf{Current Tuple.}\\
+The n+1-tuple $a$ is \emph{current} in relation $R$ if and only if:
+
 \begin{equation}
-\label{eq:rel-def}
-\left( R, A , I \right)
+\exists\ c\ \in\ \T|\ \left(R,\ \left(a,\ \left[c,\ UC\right]\right)\right) \nonumber
 \end{equation}
+
 \end{definition}
 
-In order to simplify the algorithms for the manipulation of data, some auxiliary functions are defined:
+%If the pair $\left(A, I \right)$ with $I = \left(c, UC \right), c \in \T$ is in the relation $R$, then, it is said that the pair $\left(A, I \right)$ is \emph{current} in the relation.
 
-\begin{definition}
-\label{def:close-a-crisp-interval-r}
-Consider two crisp intervals $I= \left[a ,b \right]$ and $J= \left[c ,d \right]$, $c > a$.
-The CloseR$\left(I, J\right)$ function allows to close the right-open interval $I$ with respect to the first value $d$ in $J$:
+For example, the last row in Table \ref{tbl:library-sample} visualizes the values representing a time interval $\left[4/4/2012 , UC \right]$.
+The interpretation is that the book with ID = 3 was loaned on 4/4/2012 and it is still loaned.
+The tuple representing the book with ID = 3 is current in the relation.
+
+The concepts above now allow the introduction of the function \emph{Current}:
+
 \begin{align}
-\mbox{CloseR} \left( I, J \right) &=& \\ 
+\text{Current} & :\ \Rel \times R_{NT} & \rightarrow & Ex_{R} \nonumber \\
+               & :\ \left(R,\ a\right) & \rightarrow &
  \begin{cases}
+  \left\{\left(a,\ \left[s,\ e\right]\right)\right\} & \text{if } \exists \left(a,\ \left[s,\ e\right]\right)| \\
+  									 & (\left(R,\ \left(a,\ \left[s,\ e\right]\right)\right)\\
+  									 & \wedge \left(\left[s,\ e\right]\ =\ \left[c,\ UC\right]\right)), \\
+  									 & c \in \T \\
+  \emptyset & \text{else}
+ \end{cases}
  \nonumber
-I & \mbox{ if } b \neq +\infty \\
-I=\left[a, c \right[ & \mbox{ if } b = +\infty \wedge J > I
+\end{align}
+
+Here, $Ex_{R}$ denotes the extention of $R$, $\Rel$ denotes the set of all relations in the database schema and $A_{NT}$ denotes the set of all n+1-tuples consisting of the non-temporal values of the tuples of relation $R$.
+
+%\begin{definition}
+%\label{def:current-in-relation}
+%\textbf{Function `Current'.}\\
+
+%This function returns a crisp time interval $I$ if the entity $A$ is current in the relation $R$ or the empty set, if the entity $A$ is not current in the relation.
+
+%Consider an object $A$ and a relation $R$ in a temporal database.
+The function Current$\left(R, A \right)$ obtains the tuple $\left(A, I \right)$ that is current in the relation $R$, that is it, $I = \left[a, +\infty \right]$.
+
+%\begin{align}
+%\label{eq:current-in-relation}
+%\mbox{Current} \left(R, A \right) &=& \\ 
+%\begin{cases}
+%\nonumber
+%I & \mbox{ if } (A,I) \in R \wedge I = \left(c, UC \right), c \in \T\\
+%\emptyset & \mbox{ in any other case }
+%\end{cases}
+%\end{align}
+%\end{definition}
+
+For example, the expression Current($R_{ex}$, $\left\{3,`Dracula'\right\}$), where $R_{ex}$ denotes the example relation visualized in table \ref{tbl:library-sample}, returns the tuple $\left\{3, `Dracula', 4/4/2012 , UC\right\}$.
+On the other hand, Current($R_{ex}$, $\left\{4,`Frankenstein'\right\}$) returns the empty set.
+
+%The following two Allen's relations (see Figure \ref{fig:allen-rel}) are defined~\cite{Nagypal2003} to implement the auxiliary functions.
+
+Given two time intervals $I$, represented by $\left[s_{I},\ e_{I}\right]$, and $J$, represented by $\left[s_{J},\ e_{J}\right]$, the following two functions \emph{During} and \emph{Overlaps} are functions used to compare the time intervals $I$ and $J$, based on sequential comparisons of the values representing their starting and ending instants.
+
+\begin{align}
+\text{During} & :\ \Pow(\T)\ \times\ \Pow(\T) & \rightarrow & \bool \nonumber \\
+              & :\ \left(\left[s_{I},\ e_{I}\right],\ \left[s_{J},\ e_{J}\right]\right) & \rightarrow & \left(s_{I}\ >\ s_{J}\right)\ \wedge\ \left(e_{I}\ <\ e_{J}\right)
+              \nonumber
+\end{align}
+
+\begin{align}
+\text{Overlaps} & :\ \Pow(\T)\ \times\ \Pow(\T) & \rightarrow & \bool \nonumber \\
+							  & :\ \left(\left[s_{I},\ e_{I}\right],\ \left[s_{J},\ e_{J}\right]\right) & \rightarrow & \left(s_{I}\ <\ s_{J}\right)\ \wedge\ \left(e_{I}\ <\ e_{J}\right)\nonumber \\
+							  & & & \wedge\ \left(e_{I}\ >\ s_{J}\right)
+							  \nonumber
+\end{align}
+
+Here, $\Pow(\T)$ is the powerset of $\T$ and $\bool$ is the Boolean domain $\bool\ =\ \left\{T,\ F\right\}$, where $T$ is a value representing the logical constant `True' and $F$ is a value representing the logical constant `False'.
+Now, the function \emph{CloseR} can be introduced.
+CloseR is used to `close' the representation of a time interval with respect to the representation of another time interval.
+
+\begin{align}
+\text{CloseR} & :\ \Pow(\T)\ \times\ \Pow(\T) & \rightarrow & \Pow(\T) \nonumber \\
+							& :\ \left(\left[s_{I},\ e_{I}\right],\ \left[s_{J},\ e_{J}\right]\right) & \rightarrow & 
+							\begin{cases}
+							\left[s_{I},\ s_{J}\right[ & \text{if } \left(s_{J}\ =\ UC\right)\ \wedge\\
+							 & \text{During}(\left[s_{J},\ e_{J}\right], \\
+							 & \left[s_{I},\ e_{I}\right]) \\
+							\left[s_{I},\ e_{I}\right] & \text{else }
 							\end{cases}
+							\nonumber
 \end{align}
-\end{definition}
+
+For example, consider two time intervals, represented by $I = \left[4/4/2012, UC \right]$ and $J = \left[24/4/2012, UC \right]$ respectively.
+Then, CloseR$\left(I, J \right) = \left[4/4/2012, 23/4/2012 \right]$.
 
@@
 
-Now it is possible to close the current version of an object by using \eqref{def:close-a-crisp-interval-r} and \eqref{eq:current-in-relation}
+The functionality provided by functions Current and CloseR now allows to introduce the necessary CRUD-functions.
 
-\begin{definition}
-\label{def:close-current-version}
-Consider an object $A$, the relation $R$ and a time interval $J$.
-Then, the function close-current$\left(R,A,J \right)$ closes any current version of the object $A$ if it exists and add the new version $\left(A, J \right)$.
+%Now it is possible to `close' the current version of an entity by using \eqref{eq:close-a-crisp-interval} and \eqref{eq:current-in-relation}.
+This functionality is required to append new information about an existing entity in the relation.
 
-\begin{eqnarray}
+%\begin{definition}
+%\label{def:close-current-version}
+%Consider an entity $A$, the relation $R$ and a time interval $J$.
+Then, the function close-current$\left(R,\left(A,J\right) \right)$ closes any current version of the object $A$ if it exists and add the new version $\left(A, J \right)$.
+
+%\begin{eqnarray}
 %\label{eq:close-current}
-\text{Close-current} \left(R, A, J \right) =\\
-\begin{cases}
-\nonumber
-\big \lbrace R - \left(A, I \right) \cup \left \lbrace \left(A, \mbox{ CloseR } \left(I, J\right) \right) \cup \left(A, J\right)\right \rbrace  \big \rbrace \\
-\nonumber
-\mbox{ if } I = \mbox{ Current } \left(R, A \right) \wedge J > \mbox{ CloseR } \left(I, J \right)   \\
-\nonumber R , \text{ in any other case}
-\end{cases}
-\end{eqnarray}
-\end{definition}
+%\text{Close-current} \left(R, \left(A, J\right) \right) =\\
+%\begin{cases}
+%\nonumber
+%R - \left(A, I \right) \cup \left \lbrace \left(A, \mbox{ CloseR } \left(I, J\right) \right)\right \rbrace \cup \left \lbrace\left(A, J\right) \right \rbrace  \\
+%\nonumber
+%\mbox{ if } I = \mbox{ Current } \left(R, A \right)\\ %\wedge J > \mbox{ CloseR } \left(I, J \right)   \\
+%\nonumber R , \text{ in any other case}
+%\end{cases}
+%\end{eqnarray}
+%\end{definition}
+
+%For example, consider the book with ID = 3 in Table \ref{tbl:library-sample}.
+The current version was loaned on 4/4/2012.
+The book was returned to the library but, for some reason, the end loan date was not registered.
+The book is loaned again on 24/4/2012.
+Then, the Close-current function allows to close the current version of the book and insert a new version.
+First, the function CloseR is applied with $I = \left(4/4/2012, UC \right)$ and $J = \left(24/4/2012, UC \right)$.
+Hence, the value for the time interval $I$ is $ \left(4/4/2012, 23/4/2012 \right)$.
+Then, the modifications on the value for the time interval $I$ are stored.
+Finally, a new row with the current values of the book and the time interval $J$ is stored.
+%The result of this operation is illustrated on Table \ref{tbl:library-sample-2}.
 
@@
 %\end{definition}
+\subsubsection{\label{subsubsec:modify}CRUD function `update'}
+This operation adds new information about an existing entity $E$ in the relation $R$.
+The update function does not remove any previous information about $E$, but it `closes' the current version for the entity $E$ and adds a new version.
 
-\subsubsection{\label{subsubsec:insert}Insert}
-The user wants to store an object $A$ which is valid in the relation $R$ during the time interval specified by $I = \left(a, b \right)$.
-%
-%
-%The interpretation of the insert operation is the following: The user wants to store in the database that an object $A$  is true for some period(s) of time $I$ in the given relation $R$.
-To indicate that the object $A$ is current in the relation, the special value Until Current, $\UC$  is used.
-There are two main cases when performing a create operation:
+%Note that the modify operation is only applicable when the object $A$ is current in the relation $R$; $ \left(A, I \right) \in R, I = \left[a, +\infty \right]$.
+
+\begin{align}
+\text{update}_{VT} & :\ \Rel\ \times\ Ex_{R} & \rightarrow & \Rel \nonumber \\
+							& :\ \left(R,\ \left(a,\ \left[s_{J},\ e_{J}\right]\right)\right) & \rightarrow & 
+							\begin{cases}
+							COND & \text{if } \\
+							& Current(R,\ a)\ \neq\ \emptyset \\							
+							R & \text{else }
+							\end{cases}
+							\nonumber
+\end{align}
+
+Here, COND stands for the expression:
+
+\begin{align}
+& \left(Ex_{R}\ \backslash\ Current\left(R,\ a\right)\right) \nonumber \\
+& \cup\ \left\{\left(a,\ CloseR\left(\left[c,\ UC\right],\ \left[s_{J},\ e_{J}\right]\right)\right)\right\} \nonumber \\
+& \cup\ \left\{\left(a,\ \left[s_{J},\ e_{J}\right]\right)\right\} \nonumber
+\end{align}
+
+In all of this, the `previous' version of the entity was valid during the time interval represented by $\left[c,\ UC\right]$ and the `new' version of the entity is valid during the time interval represented by $\left[s_{J},\ e_{J}\right]$.
+
+%\begin{align}
+%\label{eq:modify}
+%\MOD \left( R, \left(A, I \right)\right) =\\
+%\begin{cases}
+%\nonumber
+%\mbox{ Close-current }\left(R, \left(A, I\right) \right) %\\
+% \mbox{ if Current}\left(R ,A \right) \neq \emptyset \\
+% R,  \mbox{ otherwise }
+%\end{cases}
+%\end{align}
+
+\subsubsection{\label{subsubsec:insert}CRUD function `create'}
+The purpose of this function is to allow storing data describing an entity $E$ which is valid during the time interval represented by $\left[s_{I},\ e_{I}\right]$.
+The exact behavior of this function depends on the situation of the relation $R$.
+Two main possibilities exist.
 \begin{enumerate}
-\item The object $A$ was never in the relation $R$: The object is added with the valid-time indicated by the crisp interval $I$.
+\item The entity $E$ was never represented by any tuple in $R$.
+In this case, a tuple is simply added to the extention of $R$, representing the state of $E$ during $\left[s_{I},\ e_{I}\right]$.
+%: The object is added with the valid-time indicated by the crisp interval $I$.
+In the example, this correspond with the following insertion sentences:
 
-\item The object $A$ is in the relation $R$. Depending on the value of the time interval $I$, there are two possibilities:
+\begin{align}
+\text{create}_{VT} & :\ \Rel\ \times\ Ex_{R} & \rightarrow & \Rel \nonumber \\
+							& :\ \left(R,\ \left(a,\ \left[s_{I},\ e_{I}\right]\right)\right) & \rightarrow \nonumber \\ 
+							& create_{NT}\left(R,\ \left(a,\ \left[s_{I},\ e_{I}\right]\right)\right)
+							\nonumber
+\end{align}
+
+For example, the first 3 tuples in the relation visualized in table \ref{tbl:library-sample} were created using the create function:
+
+\begin{align}
+create_{VT}\left(R_{ex}, \left\{3, `Dracula', 15/3/2012, 30/3/2012\right\}\right) \nonumber \\
+create_{VT}\left(R_{ex}, \left\{4, `Frankenstein', 25/3/2012, 4/4/2012\right\}\right) \nonumber \\
+create_{VT}\left(R_{ex}, \left\{5, `Harry Potter', 18/3/2012, 2/4/2012\right\}\right) \nonumber
+\end{align}
+
+%\begin{verbatim}
+%Insert(3,'Dracula',
+%(15/3/2012, 30/3/2012));
+%Insert(4,'Frankenstein',
+%(25/3/2012, 4/4/2012));
+%Insert(5,'Harry Potter',
+%(18/3/2012, 2/4/2012));
+%\end{verbatim}
+
+\item $R$ contains at least one tuple representing a certain state of $E$.
+The exact behavior of the create function now depends on the exact time interval represented by $\left[s_{I},\ e_{I}\right]$.
+Three possibilities exist.
 	\begin{enumerate}
-	\item Insert $\left(A, I\right)$ in the relation $R$. If the time interval $I$ does not overlap any other valid-time interval in the relation $R$.
-	\item Reject the insertion, if the time interval $I$ do overlap any existing valid-time interval for the object $A$ in the relation $R$.
+	\item If the time interval represented by $\left[s_{I},\ e_{I}\right]$ doesn't overlap with any other valid-time interval in $R$ associated to $E$, a tuple is simply added to the extention of $R$, representing the state of $E$ during $\left[s_{I},\ e_{I}\right]$.
+
+\begin{align}
+\text{create}_{VT} & :\ \Rel\ \times\ Ex_{R} & \rightarrow & \Rel \nonumber \\
+							& :\ \left(R,\ \left(a,\ \left[s_{I},\ e_{I}\right]\right)\right) & \rightarrow \nonumber \\ 
+							& create_{NT}\left(R,\ \left(a,\ \left[s_{I},\ e_{I}\right]\right)\right)
+							\nonumber
+\end{align}
+%	Insert $\left(A, I\right)$ in the relation $R$. If the time interval $I$ does not overlap any other valid-time interval in the relation $R$.
+
+For example, consider that the book with ID 3 is loaned on 4/4/2012 but the book has not been returned yet.
+This is illustrated in table \ref{tbl:library-sample}.
+The last row of this table was created by:
+
+\begin{align}
+create_{VT}\left(R_{ex}, \left\{3, `Dracula', 4/4/2012, UC\right\}\right) \nonumber
+\end{align}
+
+	\item If $R$ contains a current tuple describing the state of $E$, this current version should be closed and a the new version should be inserted.
+
+\begin{align}
+\text{create}_{VT} & :\ \Rel\ \times\ Ex_{R} & \rightarrow & \Rel \nonumber \\
+							& :\ \left(R,\ \left(a,\ \left[s_{I},\ e_{I}\right]\right)\right) & \rightarrow \nonumber \\ 
+							& update_{VT}\left(R,\ \left(a,\ \left[s_{I},\ e_{I}\right]\right)\right)
+							\nonumber
+\end{align}
+
+For example, consider now that the book with ID = 3 is loaned on 24/4/2012.
+Here, the problem is that the book with ID = 3 was loaned on 4/4/2012, but not returned.
+If the book is already in the library, a new loan being possible, then it is necessary to set the return date and add a new tuple with the new loan date.
+The result of this is illustrated in table \ref{tbl:library-sample-2}.
+
+	\item If $R$ contains a tuple describing the state of $E$, which is not current, a mistake has obviously been made and the insertion of the new tuple should be rejected.
+%	Reject the insertion, if the time interval $I$ does overlap any existing valid-time interval for the object $A$ in the relation $R$. 
+For example, consider now that the librarian wants to introduce a past loan for the book with ID 3.
+The book was loaned on 6/4/12/2012 and returned on 25/4/2012.
+As this interval does overlaps other time intervals, it is not possible that the book was loaned in that time interval.
+Therefore, the insertion is rejected.
+
 	\end{enumerate}
@@
 
-The algorithm for the implementation of the insert operation is the following:
+\vglue13pt
+%\begin{table}[htbp]
+\tcap{\label{tbl:library-sample-2}Sample library database}
+%\centerline{\small DATA TYPES}
+\vglue-6pt
+\centerline{\small\baselineskip=13pt
+\begin{tabular}{c c c c}\\
+\hline
+\textbf{ID}  & Title  & Start Loan & End Loan \\
+\hline
+3 & Dracula & 15/3/2012 & 30/3/2012 \\
+4 & Frankenstein & 25/3/2012 & 4/4/2012 \\
+5 & Harry Potter & 18/3/2012 & 2/4/2012\\
+3 & Dracula & 4/4/2012 & 23/4/2012 \\
+3 & Dracula & 24/4/2012 & UC \\
+\hline\\
+\end{tabular}
+}
+
+%The algorithm for the implementation of the insert operation is the following:
+
+%\begin{align}
+%\label{eq:insert}
+%\INS \left(R ,\left(A, I\right)\right) =\\
+%\begin{cases}
+%\nonumber
+%R \cup \left(A, I \right) & \mbox{ if }  (A, I ) \not \in R  \vee \\
+% &   \left(A, J\right) \in R \wedge \lnot \left(I \mbox{ Overlaps } J\right) \\
+%\mbox{modify(R,(A,I))} & \mbox{ otherwise }    %\mbox{ if } \exists J \left(A,J\right) \in R \wedge \mbox{Current(A,J)} \\ 
+%R, &   
+%\end{cases} 	
+%\end{align}
+
+\subsubsection{\label{subsubsec:del}CRUD function `delete'}
+The delete function logically removes all information about an entity $E$ from valid-time relation $R$.
 
 \begin{align}
-\label{eq:insert}
-\INS \left(R ,A, I\right) =\\
-\begin{cases}
+\text{delete}_{VT} & :\ \Rel\ \times\ Ex_{R} & \rightarrow & \Rel \nonumber \\
+							& :\ \left(R,\ \left(a,\ \left[s_{I},\ e_{I}\right]\right)\right) & \rightarrow \nonumber \\ 
+							& delete_{NT}\left(R,\ \left(a,\ \left[s,\ e\right]\right)\right), \nonumber \\
+							& \forall\ \left(a,\ \left[s,\ e\right]\right)\ \in\ R
 							\nonumber
-R \cup \left(A, I \right), & \mbox{ if }  A \not \in R\\
-R \cup   \left(A, I \right) & \mbox{ if }  A \in R \wedge \\
-& \big \lbrace \left( \exists J: \left(A, J \right) \in R \wedge J<I  \right) \vee \\
-& \left( \exists K: \left(A, K \right) \in R \wedge K>I \vee \right) \wedge\\
-&  J<I<K  \big \rbrace\\
-R, & \mbox{ otherwise }  
-\end{cases} 	
 \end{align}
 
-\subsubsection{\label{subsubsec:del}Delete}
-The delete operation logically removes a current object $A$ which is valid in the relation $R$:
+Here, $\left(a,\ \left[s,\ e\right]\right)$ indicates any tuple representing information about $E$.
+
+%\begin{align}
+%\label{eq:delete}
+%\DEL \left(R ,A \right) =\\
+%\begin{cases}
+%\nonumber
+%R - \left(A, I \right), & \forall \left( A, I \right) \in R\\
+%R, & \mbox{ otherwise }  
+%\end{cases} 	
+%\end{align}
+
+For example, consider that the librarian wants to delete the information about the book with ID = 3. The result of this operation is shown in table \ref{tbl:library-sample-3}. The expression used is:
+
 \begin{align}
-\label{eq:delete}
-\DEL \left(R ,A \right) =\\
-\begin{cases}
-\nonumber
-R - \left(A, I \right), & \mbox{ if } \left( A, I \right) \in R\\
-R, & \mbox{ otherwise }  
-\end{cases} 	
+delete_{VT}\left(R_{ex}, \left\{3, `Dracula', 24/4/2012, UC\right\}\right) \nonumber
 \end{align}
 
-\subsubsection{\label{subsubsec:modify}Modify}
-This operation adds new information about an existing object $A$ in the relation $R$. The modify operation does not remove any previous value of the object $A$.
-Note that the modify operation is only applicable when the object $A$ is current in the relation $R$; $ \left(A, I \right) \in R, I = \left[a, +\infty \right]$.
+\vglue13pt
+%\begin{table}[htbp]
+\tcap{\label{tbl:library-sample-3}Sample library database}
+%\centerline{\small DATA TYPES}
+\vglue-6pt
+\centerline{\small\baselineskip=13pt
+\begin{tabular}{c c c c}\\
+\hline
+\textbf{ID}  & Title  & Start Loan & End Loan \\
+\hline
+4 & Frankenstein & 25/3/2012 & 4/4/2012 \\
+5 & Harry Potter IV & 18/3/2012 & 2/4/2012\\
+\hline\\
+\end{tabular}
+}
+
+\subsubsection{\label{subsubsec:revise}Function `revise'}
+The function \emph{revise} is a necessary function to be able to change faulty data in the valid-time database, without the database system interpreting the correction as an addition of data.
 
 \begin{align}
-\label{eq:modify}
-\MOD \left( R, A, I \right) =\\
-\begin{cases}
-\nonumber
-\mbox{ Close-current }\left(R, A, \mbox{Current}\left(R, A \right) \right) \\
- \mbox{ if Current}\left(R ,A \right) \neq \emptyset \\
- R,  \mbox{ otherwise }
-\end{cases}
+\text{revise}_{VT} & :\ \Rel\ \times\ Ex_{R}\ \times\ Ex_{R} & \rightarrow & \Rel \nonumber \\
+							& :\ \left(R,\ \left(a,\ \left[s_{I},\ e_{I}\right]\right),\ \left(a',\ \left[s_{I}',\ e_{I}'\right]\right)\right) & \rightarrow \nonumber
 \end{align}
 
-\subsubsection{\label{subsubsec:revise}Revise}
-This operation changes the value of the object $A$ in the relation $R$, at a given time interval $I$. This operation is defined when a correction on the values for the object $A$ should be done.
 \begin{align}
-\label{eq:revise}
-\mbox{ Revise}\left(R, A, I \right)=\\
 							\begin{cases}
+							COND & \text{if } \nonumber \\
+							& \left(a,\ \left[s_{I},\ e_{I}\right]\right)\ \in\ Ex_{R}\\							
+							R & \text{else }
+							\end{cases}
 							\nonumber
- R - \left(A', I \right) \cup \left(A, I \right)  \\
-  \mbox{ if } \left(A', I \right) \in R: A = \left(a_1, \ldots, a_n \right),\\
-A'= \left(a_1', \ldots, a_n' \right)   \\   
-R,  \mbox{ otherwise }
-\end{cases}
 \end{align}
+
+Here, $\left(a,\ \left[s_{I},\ e_{I}\right]\right)$ denotes the original tuple, $\left(a',\ \left[s_{I}',\ e_{I}'\right]\right)$ denotes the intended tuple and COND is:
+
+\begin{align}
+R\ \backslash\ \left\{\left(a,\ \left[s_{I},\ e_{I}\right]\right)\right\}\ \cup\ \left\{\left(a',\ \left[s_{I}',\ e_{I}'\right]\right)\right\} \nonumber
+\end{align}
+
+%This operation replaces the old values for the entity $A'= \left(a_1', \ldots, a_n' \right) $ with the new values specified by $A = \left(a_1, \ldots, a_n \right)$ in the relation $R$, at a given time interval $I$.
+This operation is used when a correction on the values for the entity $A$ should be done.
+%\begin{align}
+%\REV\left(R, \left(A, I \right) \right)=\\
+%\begin{cases}
+%\nonumber
+% R - \left(A', I \right) \cup \left \lbrace \left(A, I \right) \right \rbrace & \mbox{ if } \left(A', I \right) \in R  \\
+   %: A = \left(a_1, \ldots, a_n \right),\\
+%A'= \left(a_1', \ldots, a_n' \right)   \\   
+%R & \mbox{ otherwise }
+%\end{cases}
+%\end{align}
+
+For example, consider now that the librarian wants to make a correction in the title for the book with ID = 5.
+The new title is `Harry Potter IV'.
+The result of this operation is illustrated in Table \ref{tbl:library-sample-3}.
+
+%\vspace*{13pt}
+%\begin{center}
+%{
+%\includegraphics[scale=0.5]{./graphs/allen.pdf}
+
+%}
+%\end{center}
+%\centerline{ \psfig{file=./graphs/Y-time-point.eps}}
+%\vspace*{10pt}
+%\fcaption{\label{fig:allen-rel}Allen's relations for two crisp time intervals $I$ and $J$.}
+%\vspace*{13pt}
+
+%\begin{definition}
+%\textbf{Overlaps:}
+% \label{def:overlaps}
+%Given two time intervals $I = \left(c, d \right)$ and $J = \left(e, f\right)$, it is said that I overlaps J if:
+%\begin{equation}
+% \label{eqn:overlaps}
+%\text{I Overlaps J}  = \left(c > e  \right) \wedge \left(d < f  \right) 
+%\end{equation}
+%\end{definition}
+
+%\begin{definition}
+%\textbf{During:}
+% \label{def:during}
+%Given two time intervals $I = \left(c, d \right)$ and $J = \left(e, f\right)$, it is said that I during J if:
+%\begin{equation}
+% \label{eqn:during}
+%\text{I During J}  = \left(c < e  \right) \wedge \left(d > e  \right) \wedge \left(d < f \right)
+%\end{equation}
+%\end{definition}
+
+%\begin{definition}
+%\textbf{CloseR$\left( I, J \right)$:}
+%\label{def:close-a-crisp-interval-r}
+%Consider two crisp intervals $I= \left[c ,d \right]$ and $J= \left[e ,f \right]$.
+The CloseR$\left(I, J\right)$ function allows to close the right-open interval $I$ with respect to the first value $e$ in $J$:
+%\begin{align}
+%\label{eq:close-a-crisp-interval}
+%\mbox{CloseR} \left( I, J \right) &=& \\ 
+%\begin{cases}
+%\nonumber
+%I & \mbox{ if } b \neq +\infty \\
+%I=\left[a, c \right[ & \mbox{ if } b = +\infty \wedge J > I
+%\left[c, e \right[ & \mbox{ if } d = UC \wedge J \mbox{ During } I \\
+%I & \mbox{ in any other case.}
+%\end{cases}
+%\end{align}
+%\end{definition}

### 2012-06-12 ijcis-pos-model/sources/preliminaries.tex
My contributions so far...

--- 2012-04-16 ijcis-pos-model/sources/preliminaries.tex
@@
-%Introductory text:
-In this section, some basic concepts are introduced concerning possibilistic variables and fuzzy numbers and intervals, after which the framework of set evaluation by ill-known constraints \cite{Pons2011} is explained.
-The section concludes with a brief introduction to temporal databases.
+%Introductory text: present the entire structure (and purpose?) of the preliminaries section
+In this section, first some basic concepts and vocabulary concerning relational databases are reviewed.
+After this, some attention is given to temporal databases in general and to valid-time relations in particular.
+An example is presented, which will be used throughout the entire paper.
+Next, some concepts concerning possibility theory and fuzzy numbers and intervals are explained and defined, along with their terminology.
+The framework of set evaluation by ill-known constraints~\cite{Pons2011} is presented.
+Finally, some attention is given to the modelling and handling of imperfections in temporal modelling.
 
-\subsection{\label{subsec:possibility-theory}Possibility Theory}
-Possibility theory, like probability theory, deals with uncertainty about the outcome of an experiment.
-In probability theory, this uncertainty is caused by the \emph{variability} in the outcomes, while in possibility theory, the uncertainty is caused by \emph{incomplete knowledge} about the experiment.
-The quantification of confidence in a theory of uncertainty is achieved using a confidence measure\cite{Pons2011}.
-In probability theory this is a measure of chance, in possibility theory, possibility and necessity measures are used.
+%Something about relational databases in general and relations in specific (to make the vocabulary clear)
+\subsection{Relational Databases}
+In this subsection, a few structural and behavioral aspects of the relational database model will be presented.
 
-\begin{definition}
-Consider a set of outcomes $\Omega$.
-Let $\wp(\Omega)$ denote the powerset of $\Omega$ and let $A$ and $B$ be elements of $\wp(\Omega)$.
-A \emph{confidence measure on $\Omega$} is defined by a function
-	\begin{align}
-	g : \wp(\Omega) & \rightarrow \left[0,1\right]
-	\end{align}
-that satisfies
-	\begin{align}
-	g(\emptyset) &= 0 \\
-	g(\Omega) &= 1 	\label{NormalizationProperty} \\
-	A \subseteq B &\Rightarrow g(A) \leq g(B) \label{MonotonicityProperty}
-	\end{align}
-\end{definition}
+In a relational database, information is structured in \emph{relations}.
+As explained in the introduction, a collection of similar entities is modelled by an entitytype, which, using the relational database model, will be modelled by a relation.
+A relation consists of a \emph{relation schema} and an \emph{extention}.
+The relation schema basically determines the relation structure, by dictating which data (describing entities) will be kept and how these data will be represented.
+Each (atomic) part of these data is a result value of a measurement of a property of an entity or a description of a property of an entity.
+The extention contains these data.
+A relation schema consists of a \emph{relation name} and a finite set of \emph{attributes}.
+The relation name is of course the name of the entire relation, while the attributes describe common properties of the entities modelled by the relation.
+Every attribute consists of an attribute name and an associated \emph{data type}.
+Each data type has its \emph{domain} and its operators.
+The domain of a data type is the set of values which data of the data type in question can assume.
+Operators of a data type can be applied to data of a data type.
+The extention of a relation is a set of \emph{tuples}.
+Each tuple represents an enitity, by containing values: for every attribute in the attribute set of the relation schema, the tuple contains the value describing the exact state of the property of the entity.
+%Remember: a tuple is just an ordered tuple of values.
 
-Both possibility measures and necessity measures are special cases of confidence measures.
-
-\begin{definition}
-Consider a confidence measure $\Pi$ on a set of outcomes $\Omega$.
-Let $J$ be a countable index set and let $\{ A_{j} | j \in J \wedge A_{j} \subseteq \Omega \}$ be a family of elements of $\wp(\Omega)$.
-$\Pi$ is now a \emph{possibility measure on $\Omega$} if it satisfies:
-	\begin{align}
-	\Pi\left(\bigcup_{j \in J} A_{j} \right) = \sup_{j \in J} \Pi(A_{j})
-	\end{align}
-\end{definition}
-
-In this work, the interpretation is as follows.
-The possibility of an event expresses how plausible the occurrence of the event seems to an observer of the experiment, given the (partial) knowledge of the observer about the experiment.
-
-Information on the possibility of distinct elements of the universe of discourse $\Omega$ can now be given by a \emph{possibility distribution} $\pi$ on $\Omega$, defined by:
-
-\begin{definition}
-Consider a possibility measure $\Pi$ on $\Omega$. A \emph{possibility distribution} $\pi$ on $\Omega$ underlying the possibility measure $\Pi$ is then a function defined by:
-	\begin{align}
-	\pi : \Omega \rightarrow \left[0, 1\right] : \pi(u) = \Pi(\{u\})
-	\end{align}
-\end{definition}
-
-\begin{definition}
-Consider a confidence measure $N$ on a set of outcomes $\Omega$.
-Let $J$ be a countable index set and let $\{ A_{j} | j \in J \wedge A_{j} \subseteq \Omega \}$ be a family of elements of $\wp(\Omega)$.
-$N$ is now a \emph{necessity measure} on $\Omega$ if it satisfies:
-	\begin{align}
-	N\left(\bigcap_{j \in J} A_{j} \right) = \inf_{j \in J} N(A_{j})
-	\end{align}
-\end{definition}
-
-In this work, the interpretation is as follows.
-The necessity of an event expresses how necessary the occurrence of the event seems to an observer of the experiment, given the (partial) knowledge of the observer about the experiment.
-
-Possibility and necessity measures are dual in the sense that:
-
-\begin{align}
-\forall A \subseteq \Omega : N(A) = 1 - \Pi(\bar{A})
-\end{align}
-
-Regarding interpretation, the above can be seen as: the degree to which an event is necessary is the degree to which every other possible event is not plausible.
-
-\subsection{\label{subsec:possibilistic-variables}Possibilistic Variables}
-Possibilistic variables rely on possibility theory \cite{Dubois1988a}. A \emph{possibilistic variable} is defined as follows \cite{Pons2011}.
-
-\begin{definition}
-\label{def;possibilistic-variable}
-A possibilistic variable $X$ over a universe $U$ is defined as a variable taking exactly one value in $U$, but for which this value is (partially) unknown.
-Its possibility distribution $\pi_X$ on $U$ models the available knowledge about the value that $X$ takes: for each $u\in U$, $\pi_X(u)$ represents the possibility that $X$ takes the value $u$.
-In this work, this possibility is interpreted as a measure of how plausible it is that $X$ takes the value $u$, given (partial) knowledge about the value $X$ takes.
-\end{definition}
-
-The exact value a possibilistic variable takes, which is (partially) unknown, is called an \emph{ill-known value} in this work \cite{Dubois1988a}.
-
-When a possibilistic variable is defined on the powerset $\Pow(R)$ of some universe $R$, the unique value the variable takes will be a crisp set and its possibility distribution on the powerset $\Pow(R)$ will describe the possibility of each crisp subset of $R$ to be the value the variable takes.
-This exact value (a crisp set) the variable takes, is now called an \emph{ill-known set} \cite{Dubois1988a}.
-
-It is important to understand the difference between the following two concepts:
-\begin{itemize}
-\item
-A \emph{possibilistic variable} $X$ is bounded to take only one value , but this value is not known due to incomplete knowledge. 
-\item
-An \emph{ill-known set}~\cite{Dubois1988a}: a possibilistic variable defined over the universe $\Pow(U)$.
-\end{itemize}
-
-Note that while a possibilistic variable refers to one (partially) unknown value, an ill-known set is a crisp set but, for some reason, (partially) unknown.
-
-A specific application of possibilistic variables is obtained when the universe under consideration is the set of Boolean values $\mathbb{B}$ = $\{T,F\}$.
-Indeed, any Boolean proposition $p$ takes just one value in $\mathbb{B}$.
-If the knowledge about which value this proposition $p$ will take is given by a possibility distribution $\pi_p$, the proposition can be seen as a possibilistic variable.
-As the interest lies with the case where the proposition holds, the possibility and necessity that $p$ = $T$ (the proposition holds) demand most attention.
-This possibility and necessity is noted here as:
-\begin{align}
-\text{Possibility that $p$ = $T$ (p holds):} &\\
-\nonumber
-Pos(p) = \pi_p(T) \label{propholdsposs} \\
-\text{Necessity that $p$ = $T$ (p holds):} & \\
-\nonumber
-Nec(p) = 1-\pi_p(F) \label{propholdsnecc}
-\end{align}
-
-Here, notation \ref{propholdsposs} denotes the possibility that $p$ = $T$ and the proposition holds, notation \ref{propholdsnecc} denotes the necessity that $p$ = $T$ and the proposition holds.
-
-This work will deal with ill-known intervals.
-These are ill-known sets, defined and represented via a start and end point, which will be ill-known values.
-The elements of the set are the values between the start and end point.
-A closed ill-known interval with start point defined by possibilistic variable $X$ and end point by possibilistic variable $Y$ is noted here $\left[X, Y\right]$.
-The correspondences and transitions between the representations of ill-known sets, between the representations of ill-known intervals and between the representations of an ill-known set and an ill-known interval are part of the authors current research.
-
-\subsection{\label{subsec:fuzzy-numbers}Fuzzy Numbers and Fuzzy Intervals}
-Among others, Dubois and Prade~\cite{Dubois1983} use fuzzy sets \cite{Zadeh1965} to define a \emph{fuzzy interval}:
-\begin{definition}
-A fuzzy interval is a fuzzy set $M$, defined by a membership function $\mu_{M}$, on the set of real numbers $\mathbb{R}$ such that:
-\begin{eqnarray}
-\mu_{M} :  \mathbb{R} \rightarrow \left[0,1\right] \nonumber \\ 
-\forall (u,v)\in\mathbb{R}^2: \forall w \in [u,v]:\\
-\mu_M(w) \geq\min(\mu_M(u),\mu_M(v))  \\
-\exists m \in \mathbb{R} :  \mu_M(m)=1 
-\end{eqnarray}
-\end{definition}
-
-\vspace*{13pt}
-\begin{center}
-{
-\includegraphics[scale=0.25]{./graphs/triangular.pdf}
-
-}
-\end{center}
-%\centerline{ \psfig{file=./graphs/Y-time-point.eps}}
-\vspace*{10pt}
-\fcaption{\label{fig:triangular-dist}Trapezoidal membership function.}
-%  \label{fig:fuzzy-validity-period}
-\vspace*{13pt}
-
-If this modal value $m$ is unique, then $M$ is referred to as a \emph{fuzzy number}.
-In other words, if the core of a fuzzy interval is a singleton, it is referred to as a fuzzy number (Figure \ref{fig:triangular-dist}).
-
-A simple form of the membership function of a fuzzy interval is a trapezoidal function (Figure \ref{fig:trapezoidal-dist}.
-It can be shown that such a membership function $\mu_T$ for a fuzzy interval $T$ is convex and normalized.
-Four reel values, denoted $\alpha$, $\beta$, $\gamma$ and $\delta$ suffice to represent a trapezoidal membership function of a fuzzy interval.
-In this work, a fuzzy interval defined as such will be noted as $\left[\alpha, \beta, \gamma, \delta\right]$.
-The corresponding membership function definition for this $\mu_T$ is then given by:
-
-\begin{align}
-\mu_T : & \quad \mathbb{R} \rightarrow \left[0,1\right] \\
- : & \quad x \rightarrow
-\begin{cases}
-1 & \mbox{ if } x \in [\beta,\gamma] \\
-0 & \mbox{ if } x > \delta \vee x < \alpha \\
-\frac{x-\alpha}{\beta - \alpha} & \mbox{ if } x \in [\alpha,\beta[ \\
-\frac{\delta -x}{\delta - \gamma} & \mbox{ if } x \in ]\gamma,\delta] \\
-\end{cases}
-\end{align}
-
-\vspace*{13pt}
-\begin{center}
-{
-\includegraphics[scale=0.25]{./graphs/trapezoidalDistribution.pdf}
-
-}
-\end{center}
-%\centerline{ \psfig{file=./graphs/Y-time-point.eps}}
-\vspace*{10pt}
-\fcaption{\label{fig:trapezoidal-dist}Trapezoidal membership function.}
-%  \label{fig:fuzzy-validity-period}
-\vspace*{13pt}
-%\def\JPicScale{0.5}
-%\begin{figure}[h!]
-%  \centering
-%  \input{./graphs/trapezoidal.tex}
-%  \caption{Trapezoidal membership function}
-%  \label{fig:trapezoidal}
-%\end{figure}
-
-The most convenient form of the membership function of a fuzzy number is a triangular form.
-It can be shown that such a membership function $\mu_M$ for a fuzzy number $M$ is convex and normalized.
-Three real values, denoted by $a$, $b$ and $D$, suffice to represent a triangular membership function of a fuzzy number and in this work, a fuzzy number defined as such will be noted as $\left[D, a, b \right]$.
-Here:
-\begin{itemize}
-\item
-$D$ denotes the single value in the core of $M$
-\item
-$D-a$ is then $\inf \{u \in \mathbb{R} : \mu_{M}(u) > 0\}$
-\item
-$D+b$ is then $\sup \{u \in \mathbb{R} : \mu_{M}(u) > 0\}$
-\end{itemize}
-
-The membership function of such a fuzzy number is then given by:
-
-\vspace{-10pt}
-
-\begin{align}
-\mu_M :   \mathbb{R} \rightarrow \left[0,1\right] :  x \rightarrow \nonumber\\
-\begin{cases}
-\nonumber
-0 & \mbox{ if } \left(x < D-a\right) \vee \left(x > D+b\right) \\
-\frac{\left[x-\left(D-a\right)\right]}{a} & \mbox{ if } \left(x \geq D-a\right) \wedge (x \leq D)  \\
--\frac{\left[x-\left(D+b\right)\right]}{b} & \mbox{ if } \left(x \geq D\right) \wedge (x \leq D+b)
-\end{cases}
-\end{align}
-
-%\begin{figure}[h!]
-%  \centering
-%  \input{./graphs/triangular.tex}
-%  \caption{Triangular membership function.}
-%  \label{fig:triangular}
-%\end{figure}
-
-\subsection{\label{subsec:interval-evaluation-by-ill-known-constraints}Interval Evaluation by Ill-known Constraints}
-The problem of interval evaluation is more generally explained in \cite{Pons2011}: the need exists to know if all points in a crisp interval $I$ reside between the boundaries of an ill-known interval $\left[ X , Y \right]$.
-In \cite{Pons2011}, the notion of an \emph{ill-known constraint} is introduced:
-
-\begin{definition}
-Given a universe $U$, an ill-known constraint $C$ on a set $A \subseteq U$ is specified by means of a binary relation $R \subseteq U^{2}$ and a fixed, ill-known value denoted by its possibilistic variable $V$ over $U$, i.e.:
-\begin{align}
-\label{eq:ill-known-constraint}
-C \triangleq (R,V)
-\end{align}
-Set $A$ now satisfies the constraint if and only if:
-\begin{align}
-\forall a \in A : (a,V) \in R
-\end{align}
-\end{definition}
-
-An example of an ill-known constraint is $C_{ex} \triangleq (<, X)$. Some set $A$ then satisfies $C_{ex}$ if $\forall a \in A : a < X$, given possibilistic variable $X$.
-
-The satisfaction of a constraint $C \triangleq (R,V)$ by a set $A$ is basically still a Boolean matter, but due to the uncertainty about the ill-known value $V$, it can be uncertain whether $C$ is satisfied by $A$ or not \cite{Pons2011}.
-In fact, this satisfaction now behaves as a proposition.
-Based on the possibility distribution $\pi_{V}$ of $V$, the possibility and necessity that $A$ satisfies $C$ can be found.
-This proposition can thus be seen as a possibilistic variable on $\mathbb{B}$.
-The required possibility and necessity are:
-
-\vspace{-10pt}
-
-\begin{align}
-\Pos(A\text{ satisfies }C) & =\\
-\nonumber
-\min_{a \in A}\left(\sup_{(a,w) \in R}\pi_{V}(w)\right) \label{ill-known-pos}\\
-\Nec(A\text{ satisfies }C) & =\\
-\nonumber
-\min_{a \in A}\left(\inf_{(a,w) \notin R} 1-\pi_{V}(w)\right) \label{ill-known-nec}
-\end{align}
-
-So far, we have shown how it can be verified whether or not a set satisfies an ill-known constraint.
-However, from the examples at the beginning of this section, it is observed that Boolean combinations of constraints are required.
-For example, the problem of interval evaluation as explained earlier requires that all elements of an interval $[a,b]$ are larger than a value $X$ and at the same time smaller than a value $Y$, which implies that a conjunctive Boolean combination of both constraints must be satisfied.
-To allow Boolean combinations of constraints, the following definitions are introduced.
-\begin{definition}
-\label{def:set-evaluation}
-Consider a universe $U$, an $n$-ary vector $\mathbf{C}$ of constraints and a Boolean function $\bool:\mathbb{B}^{n}\rightarrow\mathbb{B}$. An evaluation function is defined by:
-\begin{equation}
-\lambda:\Pow(U)\rightarrow\mathbb{B}:A\mapsto\bool\Big(C_{1}(A),...,C_{n}(A)\Big).
-\end{equation}
-\end{definition}
-Definition \ref{def:set-evaluation} presents the definition of an evaluation function that evaluates a Boolean combination of some basic constraints.
-Informally, it states that a set $A$ passes the evaluation made by $\lambda$ if the Boolean combination of some propositions equals $T$.
-This crisp definition can be generalized to the case of ill-known constraints.
-\begin{definition}
-\label{def:ill-known-sets}
-Consider a universe $U$, an $n$-ary vector $\mathbf{C}$ of ill-known constraints and a Boolean function $\bool:\mathbb{B}^{n}\rightarrow\mathbb{B}$.
-The uncertainty about the evaluation of a set $A$ by an evaluation function $\lambda$ is then given by:
-\begin{equation}
-\forall A\in\Pow(U):\pi_{\lambda(A)}=\widetilde{\bool}\Big(\pi_{C_1(A)},...,\pi_{C_n(A)}\Big)\\
-\end{equation}
-Hereby, $\widetilde{\bool}$ is the possibilistic extension of $\bool$.
-\end{definition}
-It is well known that any Boolean function $\bool$ can be cast to a canonical form~\cite{McCluskey1965}, requiring only the logical conjunction $\wedge$, logical disjunction $\vee$ and logical negation.
-Therefore, only the case of Boolean conjunction, Boolean disjunction and Boolean negation will be treated within the scope of this paper.
-By applying the possibilistic extensions of $\wedge$, $\vee$ and $\neg$, concrete equations are obtained for the calculations of uncertainty about the evaluation of a set by means of an evaluation function $\lambda$.
-In the case of conjunction (i.e., $\bool=\wedge$), the inference of uncertainty about the evaluation of a set reduces to:
-\begin{eqnarray}
-\label{eq:conjunctive1}
-\forall A\in\Pow(U):\Pos(\lambda(A))&=\\
-\nonumber
-\min_{i=1}^n\Pos\left(C_i(A)\right)\\
-\label{eq:conjunctive2}
-\forall A\in\Pow(U):\Nec(\lambda(A))&=\\
-\nonumber
-\min_{i=1}^n\Nec\left(C_i(A)\right).
-\end{eqnarray}
-In the case of disjunction (i.e. $\bool=\vee$), the inference of uncertainty about the evaluation of a set reduces to:
-\begin{eqnarray}
-\label{eq:disjunctive}
-\forall A\in\Pow(U):\Pos(\lambda(A))&=\\
-\nonumber
-\max_{i=1}^n\Pos\left(C_i(A)\right)\\
-\forall A\in\Pow(U):\Nec(\lambda(A))&=\\
-\nonumber
-\max_{i=1}^n\Nec\left(C_i(A)\right).
-\end{eqnarray}
-Note that by using the functions $\min$ and $\max$ here, there is an implicit assumption that the possibilistic variables $\pi_{C_i}$ are mutual $\min$-dependent in the sense of De Cooman (i.e.
-non-interactive).
-For an extensive reading on (in)dependency of possibilistic variables, the reader is referred to~\cite{GertDeCooman1997b},\cite{GertDeCooman1997a},\cite{GertDeCooman1997}.
-In case of $\neg$, we get:
-\begin{eqnarray}
-\label{eq:negation}
-\forall A\in\Pow(U):\Pos(\neg\lambda(A))&=\\
-\nonumber
-1-\Nec(\lambda(A))\\
-\forall A\in\Pow(U):\Nec(\neg\lambda(A))&=\\
-\nonumber
-1-\Pos(\lambda(A)).
-\end{eqnarray}
-
-\begin{example}
-Consider that we want to check if crisp interval $I = \left[j, k\right]$ is included in $\left[X, Y\right]$, 2 ill-known constraints are constructed:
-
-%We assume that $X$ specifies the lower bound and $Y$ the upper bound for a given interval, we want to known whether all points in the interval are larger than or equal to $X$ and smaller than or equal to $Y$.
-Therefore, we consider two ill-known constraints:
-
-\vspace{-10pt}
-
-\begin{eqnarray}
-C_1 & \triangleq\left(\geq,X\right)\\
-C_2 & \triangleq\left(\leq,Y\right)
-\end{eqnarray}
-
-To calculate the possibility and necessity concerning a conjunction of constraints, the $\min$ operator can be used.
-The possibility and necessity of $I$ being included in $\left[X, Y\right]$ are now: %satisfying both constraints is then:
-
-\vspace{-10pt}
-
-\begin{align}
-\label{eq:interval-pos}
-\Pos(I\text{ satisfies }C_1\ and\ C_2) & =\\
-\nonumber
-\min_{a \in I}\left(\sup_{a \geq w}\pi_{X}(w),\sup_{a \leq v}\pi_{Y}(v)\right)\\
-\label{eq:interval-nec}
-\Nec(I\text{ satisfies }C_1\ and\ C_2) & =\\
-\nonumber
-\min_{a \in I}\left(\inf_{a < w} 1-\pi_{X}(w),\inf_{a > v} 1-\pi_{Y}(v)\right).
-\end{align}
-\end{example}
-
-There is a special boolean combination of constraints that is of particular interest:
-\begin{definition}
-\emph{Convex combination of ill-known constraints}. Consider the $C_1$ and $C_2$ two ill-known constraints. A convex combination of the constraints is given by:
-\begin{align}
-\label{eq:convex-combination}
-CC \triangleq \left \lbrace C_1 \wedge C_2 \right \rbrace
-\end{align}
-In a more general way, it is possible to define the convex combination of an n-ary vector  $\mathbf{C}$ of constraints:
-\begin{align}
-\label{eq:nary-convex-combination}
-CC \triangleq \left \lbrace C_1 \wedge \ldots \wedge C_n \right \rbrace
-\end{align}
-\end{definition}
-
-\begin{theorem}
-\label{th:convex-combination-ill-known-constraints}
-Consider the convex combination $CC$ of any n-ary vector $\mathbf{C} = \left \lbrace C_{1_{Z_1}}, \ldots, C_{n_{Z_n}} \right \rbrace$ of constraints over the ill-known variables $Z_1, \ldots, Z_n$.
-Then, if $\pi_{C_1(Z_1)}, \ldots, \pi_{C_n(Z_n)}$ are convex, then  $\pi_{CC}$ is also convex.
-\end{theorem}
-
-\begin{proof}
-Let $CC \triangleq \left \lbrace C_{1_{Z_1}}, \ldots, C_{n_{Z_n}}  \right \rbrace$. Then:
-\begin{align}
-\label{eq:proof-convex1}
-\pi_{CC} \left( \lambda x_1 + \left( 1 - \lambda \right)x_2 \right) = \\
-\min  \big( \pi_{C_1(Z_1)}\left( \lambda x_1 + \left( 1 - \lambda \right)x_2 \right), \ldots,\\
-   \pi_{C_n(Z_n)}\left( \lambda x_1 + \left( 1 - \lambda \right)x_2 \right)  \big)
-\end{align}
-Since $\pi_{C_1(Z_1)}, \ldots, \pi_{C_n(Z_n)}$ are convex:
-\begin{align}
-\label{eq:proof-convex2}
-\pi_{C_1(Z_1)} \left(\lambda x_1 + \left( 1 - \lambda \right)x_2  \right) \geqq \\
-\nonumber
-\min \big(\pi_{C_1(Z_1)} \left( x_1 \right),\pi_{C_1(Z_1)} \left( x_2 \right)  \big)\\
-\nonumber
-\vdots \\
-\nonumber
-\pi_{C_n(Z_n)} \left(\lambda x_1 + \left( 1 - \lambda \right)x_2  \right) \geqq \\
-\nonumber
-\min \big(\pi_{C_n(Z_n)} \left( x_1 \right),\pi_{C_n(Z_n)} \left( x_2 \right) \big) 
-\end{align}
-Then, by using equation \eqref{eq:proof-convex1}:
-\begin{align}
-\label{eq:proof-convex3}
-\pi_{CC} \left( \lambda x_1 + \left( 1 - \lambda \right)x_2 \right) \geqq  \\
-\nonumber
-\min \big( \min \left(\pi_{C_1(Z_1)} \left( x_1 \right),\pi_{C_1(Z_1)} \left( x_2 \right) \right),\\
-\nonumber
-\ldots, \min \left(\pi_{C_n(Z_n)} \left( x_1 \right),\pi_{C_n(Z_n)} \left( x_2 \right) \right) \big)
-\end{align}
-Which is equivalent to the following:
-\begin{align}
-\label{eq:proof-convex4}
-\pi_{CC} \left( \lambda x_1 + \left( 1 - \lambda \right)x_2 \right) \geqq  \\
-\nonumber
-\min \big( \min \left(\pi_{C_1(Z_1)} \left( x_1 \right), \ldots, \pi_{C_n(Z_n)} \left( x_1 \right) \right),\\
-\nonumber
-\ldots, \min \left(\pi_{C_1(Z_1)} \left( x_2 \right), \ldots, \pi_{C_n(Z_n)} \left( x_2 \right) \right) \big)
-\end{align}
-Finally we obtain:
-\begin{align}
-\label{eq:proof-convex5}
-\pi_{CC} \left( \lambda x_1 + \left( 1 - \lambda \right)x_2 \right) \geqq  \\
-\nonumber
-\min \big(\pi_{CC} \left( x_1 \right), \pi_{CC} \left( x_2 \right) \big)
-\end{align}
-\end{proof}
-
-Sometimes, an ill-known value might be specified by a convex combination of ill-known constraints.
-This allow to define ill-known values by means of relationships with respect to other ill-known points.
-
-%\begin{definition}
-%\emph{Ill-known point defined by ill-known constraints}
-%Consider a universe $U$, an n-ary vector $\mathbf{C}$ of ill-known constraints and a Boolean function $\bool:\mathbb{B}^{n}\rightarrow\mathbb{B}$. An ill-known value $X$ is defined by:
-%\begin{align}
-%\label{eq:ill-known-value-def-by-const}
-%X \in \Pow(U):\pi_{X}=\widetilde{\bool}\Big(\pi_{C_1(Z_1)},...,\pi_{C_n(Z_n)}\Big)\\
-%\end{align} 
-%\end{definition}
-
-\begin{definition}
-Consider a universe $U$, $X$ an ill-known value and $CC \triangleq \left \lbrace C_{1_{Z_1}}, \ldots, C_{n_{Z_n}}  \right \rbrace$ a convex combination of ill-known constraints over the variables $Z_1, \ldots, Z_n$.
-The value $X$ is defined by means of the convex combination:
-\begin{align}
-\label{eq:ill-known-value-by-convex-constraints}
-X \triangleq CC 
-\end{align}
-
-The uncertainty about the evaluation of an ill-known value $X$ is given by:
-\begin{align}
-\label{eq:ill-known-value-def-by-const-unc}
-X \in \Pow(U):\pi_{X}=\pi_{CC}
-\end{align} 
-Note that $\pi_{X}$ is convex since $\pi_{CC}$ is convex as demonstrated in theorem \ref{th:convex-combination-ill-known-constraints}.
-\end{definition}
-
-\begin{example}
-Consider the ill-known values $X = \left[12, 2, 2\right]$ and $Y = \left[18, 2, 1 \right]$. The ill-known value $Z$ is defined using the convex combination $CC$ of constraints $C_1$ and $C_2$:
-\begin{align}
-\nonumber
-CC \triangleq \left \lbrace C_1\left(>,X\right) \wedge C_2(\leq,Y) \right \rbrace \\
-\nonumber
-Z \triangleq CC
-\end{align}
-$Z$ is a fuzzy interval defined by a trapezoidal shape given by $\left[12,\ 14,\ 18,\ 19 \right]$
-Figure \ref{fig:example-ill-known-by-const} illustrates the relations among the variables $X$, $Y$ and $Z$.
-\end{example}
-
-\vspace*{13pt}
-\begin{center}
-{
-\includegraphics[scale=1]{./graphs/ill-known-by-constraints.pdf}
-\label{fig:example-ill-known-by-const}
-}
-\end{center}
-%\centerline{ \psfig{file=./graphs/Y-time-point.eps}}
-\vspace*{10pt}
-\fcaption{Ill-known values $X$ and $Y$. The grey area represents the ill-known value $Z$ defined by the convex combination of the two ill-known constraints $C_1$ and $C_2$.}
-\vspace*{13pt}
-
-%\paragraph{Example} Consider the ill-known values $X = \left[5, 2, 8\right]$ and $Y = \left[9, 7, 10 \right]$.
-The knowledge about the evaluation of the interval $\left[a, b \right]$  is given by the expressions \eqref{eq:interval-pos},\eqref{eq:interval-nec}.
-Figure~\ref{fig:3d-possibility} shows a 3D plot of the possibility that an interval $[a,b]$ passes the evaluations specified by the ill-known constraints.
-Note the triangular form for the resulting possibility distribution since the condition $a \leq b$ holds.
-%
-%\begin{figure}[h!]
-%\centering
-%\includegraphics[scale=0.4]{graphs/3D_possibility.eps}
-%\caption{Possibility of evaluation for the interval $[a,b]$.}
-%\label{fig:3d-possibility}
-%\end{figure}
-%The necessity plot is obtained in a similar way and is shown in Figure~\ref{fig:3d-necessity}. Notice that the necessity measure is not normalized because the supports of $X$ and $Y$ overlap.
-%\begin{figure}[h!]
-%\centering
-%\includegraphics[scale=0.4]{graphs/3D_necessity.eps}
-%\caption{Necessity of evaluation for the interval $[a,b]$.}
-%\label{fig:3d-necessity}
-%\end{figure}
+In this work, a specific variant of the relational database model will be presented.

### 2012-06-12 trunk/ijcis-pos-model/sources/imperfectionInDB.tex
Forgot something...

--- 2012-04-12 trunk/ijcis-pos-model/sources/preliminaries.tex
@@
-%Introductory text:
-In this section, some basic concepts are introduced concerning possibilistic variables and fuzzy numbers and intervals, after which the framework of set evaluation by ill-known constraints \cite{Pons2011} is explained.
-The section concludes with a brief introduction to temporal databases.
+%Introductory text: explain the problem around uncertainty. General things. Announce what will come.
+%Remember: clearly state that everything before this point concerned the crisp case.
+
+%Possibility theory: general things. DO NOT FORGET THE INTERPRETATION PART.
+
+%Possibilistic variables
+
+%Ill-known constraints. Make certain it is clear that entire sets can be evaluated, but we only use intervals...
+
+%Fuzzy numbers and intervals. Give an introduction to make the transition smooth
 

### 2012-06-12 ijcis-pos-model/sources/imperfectionInValidtimeDB.tex
Forgot something...

--- 2012-05-24 ijcis-pos-model/sources/temporaldatabases.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% Time
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{\label{subsec:time-in-databases}Time in Databases}
-The concept of time has been studied in databases for a long time.
-A true standard for adding temporal aspects to relational databases does not exist, but there is a consensus in the literature \cite{Dyreson1994} on what is called a \emph{temporal database}: a temporal database is a database dealing with some aspects of time in its schema.
-In a temporal DBMS, a \textbf{chronon} is the shortest duration of time supported by the system.
-In temporal databases, some temporal attributes can be managed without treating the attribute differently from non-temporal attributes.
-The time described by such an attribute is called \textbf{user defined time} (\emph{UDT}).
-In addition to UDT, the following types of time can be discerned in a temporal database, all of which are handled exceptionally by the DBMS:
+%Imperfection and time: modelling imperfect time (generally)
 
-\begin{itemize}
-	\item
-	\textbf{Transaction time} (\emph{TT}) \cite{Rowe1987,Jensen1991} denotes the time when the fact (object) is stored in the database.
-It is usually append-only: as the past cannot be changed, TT cannot be changed either.
-Furthermore, at the moment of insertion, a TT can be neither in the past nor in the future.
-	\item
-	\textbf{Valid time} (\emph{VT}) \cite{Jensen1994,Sarda1990,McKenzie1981} denotes the time when the fact (object) is true in the modelled reality.
-A fuzzy extension has been proposed by \cite{Garrido2009}.
-%	\item
-%	User defined time: It is an uninterpreted attribute. The domain is date/time. The query language has no special support for it.
-	\item
-	\textbf{Decision time} (\emph{DT}) \cite{Nascimento1995,Chakravarthy1994,Etzion1992,Ozsoyoglu1995} denotes the time when an event was decided to happen. 
-	\end{itemize}
+%Modelling imperfect time in temporal databases
 
-	E.g., consider a database containing employee contract descriptions.
-The time when the employee's contract is valid, represented by an interval, is VT.
-The time when the employee's contract is stored in the database is the TT.
-The time when the decision for hiring this employee was made is the DT.
-
-% is a non-decomposable unit of time.  it  There are two ways to represent a chronon: as a point or as an interval \cite{655777}.
-
-When working with these time concepts, the Data Manipulation Language (\emph{DML}, which is part of the standard database querying language SQL) is extended to deal with possible temporal inconsistencies within the data and to handle more complex (temporal) queries.
-%Next to these concepts, also \textbf{user defined time} (\emph{UDT}) is discerned.
-UDT is an uninterpreted attribute in the date/time domain.
-This means that the attribute uses the date/time domain, but the database model does not treat the attribute differently from non-temporal attributes.
-Depending on the time managed, a database is classified as either a \textbf{Valid Time Database} (\emph{VTDB}), a \textbf{Transaction Time Database} (\emph{TTDB}), a \textbf{bi-temporal database} (both valid and transaction time are managed) or a \textbf{tri-temporal} or \emph{multitemporal} database (valid time, transaction time and decision time are managed).
-
-%\subsection{Temporal Elements}
-%%	
-%	In order to define a temporal database model, some basic elements should be defined \cite{Dyreson:1994:CGT:181550.181560}:
-%	\begin{itemize}
-%	\item A	\textbf{chronon} is a non-decomposable unit of time.
-In a temporal DBMS, it is the shortest duration of time supported by the system.
-There are two ways to represent a chronon: as a point or as an interval \cite{655777}.
-%	\item
-%	\emph{Event}: An instant of time. Usually, an event is said to be occur during time $t$ if it occurs during the chronon represented by $t$.
-%	\item
-%	\emph{Interval}: The time between two events. The representation is very often a set of contiguous chronons.
-	%\item
-	%\emph{Span}: A directed duration of time.
-A duration is an amount of time with known  lenght but no specific starting or ending chronons.
-The span can be either positive, denoting a forward motion of time or negative, denoting a backward motion of time.
-%	\item A	\textbf{timestamp} is a time value associated with some object in the database.
-
-%	\item The	\textbf{lifespan} (of a database object)is the time over which it is defined.
-The lifespan of a valid time object denotes the time when the corresponding object exists.
-The lifespan of a transaction time object is the value of the timestamp.
-
-%	Depending on the type of time the meaning is different:
-%		\begin{itemize}
-%		\item
-%		Lifespan of a valid time object: Refers the time when the corresponding object exists.	
-%		\item
-%		Lifespan of a transaction time object: The transaction time of a database object refers when the object is stored in the database. The lifespan is the value of the timestamp.		
-%		\end{itemize}
-%	\end{itemize}
-
-%	In the temporal database thesaurus, \emph{'Snapshot'} is the word for non-temporal matters.
-As a temporal database is a generalization for relational databases, an \emph{snapshot database} is a relational database.
-Furthermore, a \emph{snapshot relation} is a relation incorporating neither valid nor transaction time.
-
-%\subsection{Main issues when dealing with time}
-%Among others, the following problems are present when dealing with time in a database:
-
-%	\begin{itemize}	
-%	\item \textbf{Granularity} denotes a partition on the set of chronons.
-The conversion between several granularities is studied in \cite{Lin97efficientconversion}.
-Granularity is the base of the temporal model in \cite{Cru97}.
-An object-oriented implementation is in \cite{624013}.
-%	\item	To ensure \textbf{consistency}, temporal databases usually redefine the primary key of a relation.
-The new primary key takes into account the presence of the time.
-In order to keep the consistency, the DML is re-defined.
-For example, in a VTDB, the temporal update sentence is usually composed by an update sentence (closing the old row) plus an insert sentence (creating the new row).
-%	\end{itemize}
 %Guy's suggestion: instead of imprecision, use imperfection which is more general.
@@
 In the rest of the paper we will work only with PVP to represent valid-time intervals.
-
-\subsection{\label{subsubsec:Understanding-valid-time-databases}Understanding Valid-time Databases}
-This subsection is devoted to describe the behaviour of a crisp valid-time database.
-For the shake of simplicity, only the three main operations in the Data Manipulation Language \emph{DML} a.k.a \emph{CRUD} (CReate Update, Delete) operations are shown.
-Usually the DML operations in a temporal database are re-defined e.g.
-a typical update sentence in SQL could be re-defined by means of several insert and update sentences.
-Therefore, in order to be clear, these high level primitives in the DML for a valid-time database are usually noted as Insert, Modify and Delete.
-In the following subsections each primitive is defined and explained.
-Finally a illustrative example is given.
-For a more complete information on the behaviour of a bitemporal database, please refer to ~\cite{Jensen1994}.
-
-\begin{definition}
-Consider a relation $R$ in a temporal database, an object $A$ given by the attributes $\left(a_1, \ldots, a_n \right)$ and the crisp time interval $I$ given by $\left[a,\ b\right] : a,b \in \T, a \leq b$.
-A valid-time object $A$ in the relation $R$ is specified by:
-\begin{equation}
-\label{eq:rel-def}
-\left( R, A , I \right)
-\end{equation}
-\end{definition}
-
-In order to simplify the algorithms for the manipulation of data, some auxiliary functions are defined:
-
-\begin{definition}
-\label{def:close-a-crisp-interval-r}
-Consider two crisp intervals $I= \left[a ,b \right]$ and $J= \left[c ,d \right]$, $c > a$.
-The CloseR$\left(I, J\right)$ function allows to close the right-open interval $I$ with respect to the first value $d$ in $J$:
-\begin{align}
-\label{eq:close-a-crisp-interval}
-\mbox{CloseR} \left( I, J \right) &=& \\ 
-\begin{cases}
-\nonumber
-I & \mbox{ if } b \neq +\infty \\
-I=\left[a, c \right[ & \mbox{ if } b = +\infty \wedge J > I
-\end{cases}
-\end{align}
-\end{definition}
-
-%Analogously, it is defined the function to close a left-open interval: 
-%\begin{definition}
-%\label{def:close-a-crisp-interval-l}
-%CloseL$\left(I, J\right)$ 
-%\begin{align}
-%\mbox{CloseL} \left( I, J \right) &=& \\ 
-%\begin{cases}
-%\nonumber
-%I & \mbox{ if } a \neq -\infty \\
-%I=\left]d, b \right] & \mbox{ if } a = -\infty \wedge J < I
-%\end{cases}
-%\end{align}
-%\end{definition}
-
-\begin{definition}
-\label{def:current-in-relation}
-Consider an object $A$ and a relation $R$ in a temporal database.
-The function Current$\left(R, A \right)$ obtains the tuple $\left(A, I \right)$ that is current in the relation $R$, that is it, $I = \left[a, +\infty \right]$.
-
-\begin{align}
-\label{eq:current-in-relation}
-\mbox{Current} \left(R, A \right) &=& \\ 
-\begin{cases}
-\nonumber
-I & \mbox{ if } \exists b \in I : (A,I) \in R \wedge b = +\infty \\
-\emptyset & \mbox{ in any other case }
-\end{cases}
-\end{align}
-\end{definition}
-
-%\begin{definition}
-%\label{def:first-in-relation}
-%Analogously, the function First$\left(R, A \right)$ obtains the tuple $\left(A, I \right)$ where $I$ is a right-open interval and which is the first version in the relation $R$, that is it, $I = \left[-\infty, b \right]$.
-%\begin{align}
-%\label{eq:first-in-relation}
-%\mbox{First} \left(R, A \right) &=& \\ 
-%\begin{cases}
-%\nonumber
-%I & \mbox{ if } \exists b \in I : (A,I) \in R \wedge a = -\infty \\
-%\emptyset & \mbox{ in any other case }
-%\end{cases}
-%\end{align}
-%
-%\end{definition}
-
-Now it is possible to close the current version of an object by using \eqref{def:close-a-crisp-interval-r} and \eqref{eq:current-in-relation}
-
-\begin{definition}
-\label{def:close-current-version}
-Consider an object $A$, the relation $R$ and a time interval $J$.
-Then, the function close-current$\left(R,A,J \right)$ closes any current version of the object $A$ if it exists and add the new version $\left(A, J \right)$.
-
-\begin{eqnarray}
-\label{eq:close-current}
-\text{Close-current} \left(R, A, J \right) =\\
-\begin{cases}
-\nonumber
-\big \lbrace R - \left(A, I \right) \cup \left \lbrace \left(A, \mbox{ CloseR } \left(I, J\right) \right) \cup \left(A, J\right)\right \rbrace  \big \rbrace \\
-\nonumber
-\mbox{ if } I = \mbox{ Current } \left(R, A \right) \wedge J > \mbox{ CloseR } \left(I, J \right)   \\
-\nonumber R , \text{ in any other case}
-\end{cases}
-\end{eqnarray}
-\end{definition}
-
-%It is also possible to append at the beggining by using \eqref{def:close-a-crisp-interval-l} and \eqref{def:first-in-relation}.
-%\begin{definition}
-%\label{def:append-first-version}
-%Consider an object $A$, the relation $R$ and a time interval $J$.
-Then, the function append-first$\left(R,A,J \right)$ closes any right open version of the object $A$ if it exists and add the new version $\left(A, J \right)$.
-%
-%\begin{eqnarray}
-%%\label{eq:close-current}
-%\text{Append-first} \left(R, A, J \right) =\\
-%\begin{cases}
-%\nonumber
-%\big \lbrace R - \left(A, I \right) \cup \left \lbrace \left(A, \mbox{ CloseL } \left(I, J\right) \right) \cup \left(A, J\right)\right \rbrace  \big \rbrace \\
-%\nonumber
-%\mbox{ if } I = \mbox{ First } \left(R, A \right) \wedge J < \mbox{ CloseL } \left(I, J \right)   \\
-%\nonumber R , \text{ in any other case}
-%\end{cases}
-%\end{eqnarray}
-%\end{definition}
-
-\subsubsection{\label{subsubsec:insert}Insert}
-The user wants to store an object $A$ which is valid in the relation $R$ during the time interval specified by $I = \left(a, b \right)$.
-%
-%
-%The interpretation of the insert operation is the following: The user wants to store in the database that an object $A$  is true for some period(s) of time $I$ in the given relation $R$.
-To indicate that the object $A$ is current in the relation, the special value Until Current, $\UC$  is used.
-There are two main cases when performing a create operation:
-\begin{enumerate}
-\item The object $A$ was never in the relation $R$: The object is added with the valid-time indicated by the crisp interval $I$.
-
-\item The object $A$ is in the relation $R$. Depending on the value of the time interval $I$, there are two possibilities:
-	\begin{enumerate}
-	\item Insert $\left(A, I\right)$ in the relation $R$. If the time interval $I$ does not overlap any other valid-time interval in the relation $R$.
-	\item Reject the insertion, if the time interval $I$ do overlap any existing valid-time interval for the object $A$ in the relation $R$.
-	\end{enumerate}
-
-\end{enumerate}
-
-The algorithm for the implementation of the insert operation is the following:
-
-\begin{align}
-\label{eq:insert}
-\INS \left(R ,A, I\right) =\\
-\begin{cases}
-\nonumber
-R \cup \left(A, I \right), & \mbox{ if }  A \not \in R\\
-R \cup   \left(A, I \right) & \mbox{ if }  A \in R \wedge \\
-& \big \lbrace \left( \exists J: \left(A, J \right) \in R \wedge J<I  \right) \vee \\
-& \left( \exists K: \left(A, K \right) \in R \wedge K>I \vee \right) \wedge\\
-&  J<I<K  \big \rbrace\\
-R, & \mbox{ otherwise }  
-\end{cases} 	
-\end{align}
-
-\subsubsection{\label{subsubsec:del}Delete}
-The delete operation logically removes a current object $A$ which is valid in the relation $R$:
-\begin{align}
-\label{eq:delete}
-\DEL \left(R ,A \right) =\\
-\begin{cases}
-\nonumber
-R - \left(A, I \right), & \mbox{ if } \left( A, I \right) \in R\\
-R, & \mbox{ otherwise }  
-\end{cases} 	
-\end{align}
-
-\subsubsection{\label{subsubsec:modify}Modify}
-This operation adds new information about an existing object $A$ in the relation $R$. The modify operation does not remove any previous value of the object $A$.
-Note that the modify operation is only applicable when the object $A$ is current in the relation $R$; $ \left(A, I \right) \in R, I = \left[a, +\infty \right]$.
-
-\begin{align}
-\label{eq:modify}
-\MOD \left( R, A, I \right) =\\
-\begin{cases}
-\nonumber
-\mbox{ Close-current }\left(R, A, \mbox{Current}\left(R, A \right) \right) \\
- \mbox{ if Current}\left(R ,A \right) \neq \emptyset \\
- R,  \mbox{ otherwise }
-\end{cases}
-\end{align}
-
-\subsubsection{\label{subsubsec:revise}Revise}
-This operation changes the value of the object $A$ in the relation $R$, at a given time interval $I$. This operation is defined when a correction on the values for the object $A$ should be done.
-\begin{align}
-\label{eq:revise}
-\mbox{ Revise}\left(R, A, I \right)=\\
-\begin{cases}
-\nonumber
- R - \left(A', I \right) \cup \left(A, I \right)  \\
-  \mbox{ if } \left(A', I \right) \in R: A = \left(a_1, \ldots, a_n \right),\\
-A'= \left(a_1', \ldots, a_n' \right)   \\   
-R,  \mbox{ otherwise }
-\end{cases}
-\end{align}

### 2012-06-12 trunk/ijcis-pos-model/sources/intro.tex
preliminary final modifications

--- 2012-05-23 trunk/ijcis-pos-model/sources/intro.tex
@@
 
-The concept of time itself is very complex to handle and interpret~\cite{Klein1994,Shackle1961}, though it is very natural and omnipresent.
-As information systems often attempt the modelling of natural objects, concepts or processes, they often require modelling temporal aspects or concepts.
-Thus, several proposals have been concerned with the obtaining of theoretical models that allow the modelling or representation of time~\cite{Bolour1982,VanderCruyssen1997}.
+The concept of time is very complex to handle and interpret~\cite{Klein1994,Shackle1961}, though it is very natural and omnipresent in real world data.
+As information systems attempt the modelling of natural objects, concepts or processes, they often require modelling temporal aspects or concepts.
+Thus, several proposals have arose to obtain theoretical models that allow the modelling or representation of time~\cite{Bolour1982,VanderCruyssen1997}.
 
-A very specific type of information systems are database systems, which are computer systems designed to manage databases.
+A very specific type of information systems are database systems.
 A database contains data representing real objects or concepts.
-Each (atomic) part of these data is a result value of a measurement of a property or a description of a property of a real object or concept.
 In reality, some aspects or properties of objects or concepts are time-variant or time-related.
-e.g., the moment of a bank transaction is traditionally a moment in time and thus a time-related notion, the function of an employee in a company can change through recorded history and is thus time-variant.
-A temporal database schema is a database schema that models real objects or concepts with time-related or time-variant properties.
+e.g., the moment of a bank transaction is traditionally a point in time and thus a time-related notion, the function of an employee in a company can change through the recorded history and is thus time-variant.
+A temporal database schema is a database schema that models objects or concepts with time-related or time-variant properties.
 However, the modelling of temporal aspects has a direct impact on the consistency of the temporal database, because the temporal nature of these aspects imposes extra integrity constraints.
 An example.
-Consider a library database, modelling the presence of books in the library.
-Two dates are stored: the loan and the return date.
-It is clear that a book cannot be loaned again until it is returned.
-Without further precautions, a library employee could loan the same book several times even if it is not returned.
+
+Consider a hospital database containing data about the state of the patients in the hospital.
+Two dates are stored: the date when the patient arrives to and leaves from the hospital.
+It is clear that a patient cannot left the hospital if he or she has not arrived.
+Without further precautions, a hospital employee could insert a patient who is already in the hospital.
 A temporal database model will typically constrain record insertion and prevent similar modelling inconsistencies.
@@
 Notably, in 1994, ``A Consensus Glossary of Temporal Database Concepts'' was published~\cite{Dyreson1994}.
-For this publication, 44 temporal database researchers, among which some of the main researchers in this field, cooperated to reach a consensus on the nature and definitions of some of the main temporal database concepts and terminology.
-This glossary was subsequently updated in 1998~\cite{Dyreson1998}.
+For this publication, 44 temporal database researchers, among these some of the main researchers in this field, cooperated to reach a consensus on the nature and definitions of some of the main temporal database concepts and terminology.
+This glossary was updated in 1998~\cite{Dyreson1998}.
+
+% %From time in general to time in information systems
+% In real world, some aspects or properties of objects, concepts or processes are time-variant or time-related.
+For example, the moment of a bank transaction is traditionally a point in time and thus a time-related notion, the function of an employee in a company can change through recorded history and is thus time-variant.
+As information systems often attempt the modelling of natural objects, concepts or processes, they often require modelling temporal aspects or concepts.
+Thus, several proposals have been concerned with obtaining theoretical models that allow the modelling or representation of time~\cite{Bolour1982},~\cite{VanderCruyssen1997}.
+
+% %From time in information systems to time in temporal databases
+% Time is a complex concept, which makes modelling time in database systems complex too.
+A \emph{temporal database} is a database that deals with certain time-related aspects in its schema: a \emph{temporal database schema}~\cite{Dyreson1994} is a database schema that models entities (and interactions or relationships between entities) with time-related or time-variant properties.
+However, the modelling of temporal properties or aspects has a direct impact on the consistency of the resulting temporal database, because the temporal nature of these aspects or properties imposes extra integrity constraints.
+An example.
+Consider a relation in a relational library database, representing the physical absence (or presence) of books in the library.
+Every physical book is represented by a unique identifier.
+Every record in the relation contains such an identifier, a date on which the corresponding book was loaned and a date on which it was subsequently returned (if it was returned).
+As such, every record represents a period of time during which the physical book corresponding with the identifier is loand to a library customer.
+Without further precautions, a library employee could add several records with the same book identifier, different `loaned'-dates and no `returned'-dates.
+This group of records would represent the same physical book being loaned several times on different dates and never returned, which is of course impossible.
+A temporal database model will typically constrain record insertion and prevent similar modelling inconsistencies.
 
 An interesting issue in temporal modelling concerns relationships between temporal notions.
-Notably, Allen~\cite{Allen1983} studied temporal relationships between time intervals (and as a special case time points).
-Among others, the querying of temporal databases has greatly profited from these temporal relationships, because they allowed for richer and more complex user-specified temporal query demands, by allowing to express more complex relationships between the temporal notions in the temporal expressions in the query and the temporal indications in the database.
-For example, in a relation modelling who was department head of an institution during which periods of time, a query like `who were the department heads when Thomas worked for the institution' can be evaluated using similar relationships.
+In this sense, Allen~\cite{Allen1983} studied temporal relationships between time intervals (and as a special case time points).
+Among others, the querying of temporal databases has greatly profited from these temporal relationships, because they allowed richer and more complex user-specified temporal query demands, by allowing to express more complex relationships between the temporal notions in the temporal expressions in the query.
+For example, a query like `who were the department heads when Thomas worked for the institution' can be evaluated using operators similar to Allen's ones.
 
-Humans handle temporal information using certain temporal notions like time intervals or time points~\cite{Dyreson1994}, and they often have to deal with imperfections like imprecisions, vaguenesses, uncertainties or inconsistencies possibly contained in the descriptions of these temporal notions.
+Humans handle temporal information using certain temporal notions like time intervals or time points~\cite{Dyreson1994}, and they often have to deal with imperfections like imprecision, vagueness, uncertainty or inconsistencies possibly contained in the descriptions of these temporal notions.
 Among many others, these possible imperfections in descriptions of temporal notions determine an important issue in temporal modelling.
-Consider as an example the description of the temporal notion in a sentence like `The Belfry of Bruges was finished on a day somewhere between 01/01/1201 A.D.
+Consider as an example the description of the temporal notion in a sentence like `The Belfry of Bruges was finished between 01/01/1201 A.D.
 and 31/12/1300 A.D.' contains imperfection because of the uncertainty in the used time-related expression.
-It is known that the building was finished on a single day, but it is not known precisely which day this was.
+It is known that the building was finished on a single day, but it is not precisely known  which day this was.
 
@@
 The temporal relationships studied by Allen were fuzzified by several authors~\cite{Ohlbach2004,Nagypal2003,Schockaert2008}.
-Garrido et al.
-~\cite{Garrido2009} present different temporal operators, defined by a combination of regular fuzzy comparisons.
-Also,~\cite{Garrido2009,Pons2011} studied uncertainty in temporal expressions concerning time intervals.
+Although,  Garrido et al.
+~\cite{Garrido2009} present a compact representation for the time, they defined different relationships among time intervals by using a combination of regular fuzzy comparisons.
+Also ~\cite{Garrido2009,Pons2011} studied uncertainty in temporal expressions concerning time intervals.
 Other approaches, like~\cite{Qiang2009}, use rough sets~\cite{Pawlak1995} to represent time intervals.
 
-Next to temporal modelling, some attention has been spent on temporal reasoning~\cite{Allen1983}.
+%Now introducing imperfections in representations of time...
+% Humans handle temporal information using certain temporal notions like specific time intervals or instants~\cite{Dyreson1994} and they often have to deal with imperfections like imprecisions, vaguenesses, uncertainties or inconsistencies possibly contained in (the descriptions of) these temporal notions.
+Among many others, these possible imperfections in (descriptions of) temporal notions determine an important issue in temporal modelling.
+As an example, consider the temporal notion described in a sentence like `The Belfry of Bruges was finished on a day somewhere between 01/01/1201 A.D.
+and 31/12/1300 A.D.'.
+The description of the temporal notion reads `a day somewhere between 01/01/1201 A.D.
+and 31/12/1300 A.D.'.
+This temporal notion thus contains imperfection in the form of the uncertainty about the exact day the last stone of the building was laid: it is known that the building was finished on a single day, but it is not known which day this was.
+To allow information systems to cope with these and similar temporal data imperfections, many approaches adopt fuzzy sets~\cite{Zadeh1965} and/or fuzzy logic to model temporal information~\cite{Mitra1994},~\cite{Nagypal2003},~\cite{Billiet2011},~\cite{Dubois2003}.
+
+In addition to temporal modelling, some attention has been spent on temporal reasoning~\cite{Allen1983}.
 Although temporal reasoning is not discussed in this paper, it should be noted that, among others, Dubois and Prade et al.~\cite{Dubois2003,DuBois1989} have dealt with fuzziness and uncertainty in temporal reasoning.
 
-The aim of this work is to present and explain the main operations of the manipulation language for a temporal database which stores the valid-time periods of the objects with uncertainty.
+The present work defines and implement a model for properly represent and manage uncertainty in valid-time specification in a relational database.
+None of the previous research offer a database model for acomplish this task.
+
+The aim of this work is to present and explain the main operations of the manipulation language for a temporal database which stores the valid-time periods of the objects affected by imprecision.
+The rest of the work is organized as follows.
+Section \ref{sec:prelim} presents some background concepts about both possibility theory and temporal databases.
+In section \ref{sec:time-rep} the representation of the valid-time intervals in the database is explained.
+Section \ref{sec:temporal-model} explains the main concepts and behaviour of the temporal model and the Data Manipulation Language (DML) is described and implemented.
+Finally, Section \ref{sec:conclusions} presents the conclusions and some possibilities for future research work.
+
+% 
+% 
+% %From information systems in general to database systems in particular:
+% Generally, \emph{information systems} model the structure and behavior of real objects, concepts or processes.
+A specific type of information systems are \emph{database systems}, which are computer systems designed to manage \emph{databases}.
+A database is basically a collection of (persistent) data.
+These (typically atomic) data represent real objects or concepts.
+In the context of database design, such objects or concepts are typically called \emph{entities} and a collection of similar entities is typically modelled by an \emph{entitytype}, which is basically a combination of a name and a list of \emph{attributes}, which describe properties of the entities.
+Typically, in an early stage of the database development process, the structures of and interactions or relationships between used entities are modelled in a \emph{design schema}, following some design model.
+A database thus contains (typically atomic) data.
+Each (atomic) part of these data is a result value of a measurement of a property of an entity or a description of a property of an entity and will correspond to the attribute of the entity's entitytype, which describes the property.
+% 
+% %From database models in general to the relational model in particular:
+% The structure and behavior of a database, along with some integrity and security restrictions, is dictated by a \emph{database model}.
+A database model is basically a collection of instructions and regulations, used to (logically) describe the structure and behavior of a database, along with some integrity and security restrictions.
+Applying a database model to a concrete design schema results in a \emph{database schema}, which models the logical structure and behavior of a database.
+Several different database models exist, but the most popular is the \emph{relational database model}~\cite{Codd:1970:RMD:362384.362685}.
+Following the relational database model, entitytypes are modelled as \emph{relations}, which comprise a name and a list of attributes, modelling the entitytype's attributes.
+% 
+% %From time in general to time in information systems
+% In reality, some aspects or properties of objects, concepts or processes are time-variant or time-related.
+For example, the moment of a bank transaction is traditionally a moment in time and thus a time-related notion, the function of an employee in a company can change through recorded history and is thus time-variant.
+The concept of time itself is very complex to handle and interpret~\cite{Klein1994},~\cite{Shackle1961}, though it is very natural and omnipresent.
+As information systems often attempt the modelling of natural objects, concepts or processes, they often require modelling temporal aspects or concepts.
+Thus, several proposals have been concerned with obtaining theoretical models that allow the modelling or representation of time~\cite{Bolour1982},~\cite{VanderCruyssen1997}.
+% 
+% %From time in information systems to time in temporal databases
+% Time is a complex concept, which makes modelling time in database systems complex too.
+A \emph{temporal database} is a database that deals with certain time-related aspects in its schema: a \emph{temporal database schema}~\cite{Dyreson1994} is a database schema that models entities (and interactions or relationships between entities) with time-related or time-variant properties.
+However, the modelling of temporal properties or aspects has a direct impact on the consistency of the resulting temporal database, because the temporal nature of these aspects or properties imposes extra integrity constraints.
+An example.
+Consider a relation in a relational library database, representing the physical absence (or presence) of books in the library.
+Every physical book is represented by a unique identifier.
+Every record in the relation contains such an identifier, a date on which the corresponding book was loaned and a date on which it was subsequently returned (if it was returned).
+As such, every record represents a period of time during which the physical book corresponding with the identifier is loand to a library customer.
+Without further precautions, a library employee could add several records with the same book identifier, different `loaned'-dates and no `returned'-dates.
+This group of records would represent the same physical book being loaned several times on different dates and never returned, which is of course impossible.
+A temporal database model will typically constrain record insertion and prevent similar modelling inconsistencies.
+% 
+% %Research about temporal databases
+% A lot of research concerns temporal database models and their approaches to the modelling of time.
+Some of the first proposals concerned the representation of historical information related to entities~\cite{Clifford1985}.
+Some proposals tried to extend the Entity Relationship Model~\cite{Klopprogge1983}, without impact on any database standards like SQL~\cite{Sarda1990}.
+Notably, in 1994, `A Consensus Glossary of Temporal Database Concepts' was published~\cite{Dyreson1994}.
+For this publication, 44 temporal database researchers, among which some of the main researchers in this field, cooperated to reach a consensus on the nature and definitions of some of the main temporal database concepts and their terminology.
+This glossary was subsequently updated in 1998~\cite{Dyreson1998}.
+In the presented work, the concepts and terminology from this glossary are used and followed.
+% 
+% %Research about temporal relationships
+% An interesting issue in temporal modelling concerns the relationships between temporal notions.
+Notably, Allen~\cite{Allen1983} studied temporal relationships between time intervals~\cite{Dyreson1994} (and, as a special case, instants~\cite{Dyreson1994}).
+Among others, the querying of temporal databases has greatly profited from these temporal relationships, because they allow for richer and more complex user-specified temporal query demands, by allowing to express more complex relationships between the temporal notions in the temporal expressions in the query and the temporal indications in the database.
+For example, given a relation representing who was department head of an institution during which time intervals, a query like `Who were the department heads during the time intervals when Thomas worked for the institution?' can be evaluated using similar relationships.
+%TODO: CHECK THIS LAST SENTENCE
+% 
+% %Now introducing imperfections in representations of time...
+% Humans handle temporal information using certain temporal notions like specific time intervals or instants~\cite{Dyreson1994} and they often have to deal with imperfections like imprecisions, vaguenesses, uncertainties or inconsistencies possibly contained in (the descriptions of) these temporal notions.
+Among many others, these possible imperfections in (descriptions of) temporal notions determine an important issue in temporal modelling.
+As an example, consider the temporal notion described in a sentence like `The Belfry of Bruges was finished on a day somewhere between 01/01/1201 A.D.
+and 31/12/1300 A.D.'.
+The description of the temporal notion reads `a day somewhere between 01/01/1201 A.D.
+and 31/12/1300 A.D.'.
+This temporal notion thus contains imperfection in the form of the uncertainty about the exact day the last stone of the building was laid: it is known that the building was finished on a single day, but it is not known which day this was.
+To allow information systems to cope with these and similar temporal data imperfections, many approaches adopt fuzzy sets~\cite{Zadeh1965} and/or fuzzy logic to model temporal information~\cite{Mitra1994},~\cite{Nagypal2003},~\cite{Billiet2011},~\cite{Dubois2003}.
+% 
+% %Now introducing imperfections in temporal relationships
+% The temporal relationships studied by Allen were fuzzified by several authors~\cite{Ohlbach2004},~\cite{Nagypal2003},~\cite{Schockaert2008}.
+Garrido et al.
+~\cite{Garrido2009} present different temporal operators, defined by a combination of fuzzy comparisons.
+Also,~\cite{Garrido2009},~\cite{Pons2011} studied uncertainty in the context of time intervals.
+Other approaches, like~\cite{Qiang2009}, use rough sets~\cite{Pawlak1995} to represent time intervals.
+% 
+% %Imperfections in temporal reasoning
+% Next to temporal modelling, some attention has gone to temporal reasoning~\cite{Allen1983}.
+Although the focus of this paper is temporal modelling, it should be noted that, among others, Dubois and Prade et al.~\cite{Dubois2003},~\cite{DuBois1989} have dealt with fuzziness and uncertainty in temporal reasoning.
+% 
+% %Overview of this work
+% %Include: what is the paper about? (situation and explanation of the problem we will solve and situation of our solution (we use constraints...))
+% %Include: how does the solution work? (short summary of how the solution will solve the problem/attend to the problem)
+% %Include: What exactly will be presented
+% %In all of these: stress every single innovative/novel point of the presented work!
+% %TODO: build the overview when the paper is finished
+% 
+% The aim of this work is to present and explain the main operations of the manipulation language for a temporal database which stores the valid-time periods of the objects with uncertainty.
 The rest of the work is organized as follows.
@@
 Finally in Section \ref{sec:conclusions} presents the conclusions and some possibilities for future research work.
+%

### 2012-06-12 trunk/ijcis-pos-model/sources/temporaldatabases.tex
preliminary final modifications

--- 2012-05-27 trunk/ijcis-pos-model/sources/temporaldatabases.tex
@@
 \subsection{\label{subsec:time-in-databases}Time in Databases}
-The concept of time has been studied in databases for a long time.
+The concept of time has been studied in the database framework for a long time.
 A true standard for adding temporal aspects to relational databases does not exist, but there is a consensus in the literature \cite{Dyreson1994} on what is called a \emph{temporal database}: a temporal database is a database dealing with some aspects of time in its schema.
@@
 
-	E.g., consider a database containing employee contract descriptions.
-The time when the employee's contract is valid, represented by an interval, is VT.
+	For example, consider a database containing employee contract descriptions.
+The time when the employee's contract is valid, represented by an interval, is the VT.
 The time when the employee's contract is stored in the database is the TT.
@@
 \subsubsection{Imperfection and time}
-Representing imprecision and its semantics when dealing with time has been studied for a long time.
-Several proposals for representing and computing imprecise time indications can be found in \cite{DeCaluwe1997,DeTre1997}.
+The representation of imprecision and its semantics when dealing with time has been studied for a long time.
+Several proposals for representing and handling imprecise time indications can be found in \cite{DeCaluwe1997,DeTre1997}.
 Also, the changes between several granularities can be seen as a source of imprecision \cite{Devos1998}.
 
-In the proposal section we will consider two kinds of imprecision:
+In this paper we will consider two kinds of imperfection:
 \begin{itemize}
-\item \textbf{Imperfection in the database} denotes the uncertainty that arises when the knowledge about the temporal data in the database is uncertain.
-E.g., a database record shows that \emph{`The car is in the garage around April.'}
+\item \textbf{Imperfection in the database} the knowledge about the temporal data contains some imperfection. E.g., a database record shows that \emph{`The car was in the garage around April.'}
  \item \textbf{Imprecision in the query specification} denotes the imprecision in the specification of temporal criteria by the user, when querying.
-E.g., \emph{`The user wants to obtain a car which is red and which is in the garage around April.'}
+E.g., \emph{`The user wants a car which was in the garage around April.'}
 \end{itemize}
@@
 Several proposals for managing uncertain time in a database exist.
-Some proposals work with rough sets \cite{Qiang2009}, other proposals rely on possibility distributions for representing uncertainty in time \cite{Dyreson1998,Garrido2009,Galindo2001}.
-In order to compare temporal possibility distributions, extensions of the classical Allen's operators \cite{Allen1983} are defined in \cite{Ohlbach2004,Nagypal2003,Dubois2003a,Schockaert2008}.
+Some proposals work with rough sets \cite{Qiang2009}, and some others rely on possibility distributions for representing uncertainty in time \cite{Dyreson1998,Garrido2009,Galindo2001}.
+In order to compare temporal possibility distributions, the extensions of the classical Allen's operators \cite{Allen1983} are defined in \cite{Ohlbach2004,Nagypal2003,Dubois2003a,Schockaert2008}.
 %In the proposal section, we will follow the representation by means of possibility distributions, in order to work with both satisfaction and dissatisfaction degrees.
@@
 
-In order to deal with uncertainty in time intervals, several proposals are made.
-Here, two approaches are introduced: the first one, based on \emph{Fuzzy Validity Period}~\cite{Garrido2009} and the second one based on \emph{Possibilistic Valid-time Period}~\cite{JoseEnriquePons2012}.
+In order to deal with uncertainty in time intervals, several proposals have been made.
+Here, two approaches are introduced: the first one, based on \emph{Fuzzy Validity Periods}~\cite{Garrido2009} and the second one, based on \emph{Possibilistic Valid-time Periods}~\cite{JoseEnriquePons2012}.
 
 \begin{definition}
-A \textbf{Fuzzy Validity Period} (\emph{FVP}) is defined as a fuzzy time interval specifying when the data regarding an object is valid.
-A fuzzy time interval is then the fuzzification of a crisp time interval.
+A \textbf{Fuzzy Validity Period} (\emph{FVP}) is defined as a fuzzy time interval specifying when the data regarding an object are valid.
+A fuzzy time interval is then the fuzzification of a crisp time interval.\\
 \end{definition}
+
 Several options to transform possibility distributions corresponding to the fuzzy starting point and the fuzzy end point into one consistent FVP exist \cite{Garrido2009}, e.g (Fig.
@@
 \begin{itemize}
-\item The \textbf{convex hull} approach is the most intuitive approach. The resulting FVP is the convex hull of the union of both fuzzy sets.
-\item The \textbf{uncertainty preserving} approach is less intuitive but more realistic.
-The amount of uncertainty is maintained at the edges of the possibility distribution representing the FVP \cite{Garrido2009}.
+\item The \textbf{convex hull} approach is the most intuitive approach. The resulting FVP is the convex hull of the union of both possibility distributions.
+\item The \textbf{uncertainty preserving} approach. The amount of uncertainty is maintained at the edges of the possibility distribution representing the FVP \cite{Garrido2009}.
 \end{itemize}
@@
 
+The main feature for the FVP is the optimization for the storage.
+The compact representation is the result of a conjunctive semantic.
+The object is valid within all the time points inside the starting and ending points.
+
 \begin{definition}
-A \textbf{Possibilistic Valid-time Period} (\emph{PVP}) is an ill-known interval in time specifying when the data regarding an object is valid.
+A \textbf{Possibilistic Valid-time Period} (\emph{PVP}) is an ill-known interval of time specifying when the data regarding an object are valid.
 \end{definition}
-A PVP is an ill-known interval in the sense of definition \ref{def;possibilistic-variable}, section \ref{subsec:possibilistic-variables}.
-Note that this representation is \emph{disjunctive}: the PVP represents only one crisp time interval, but that for some reason it is (partially) unknown.
+A PVP is an ill-known interval in the sense of definition \ref{def;possibilistic-variable}.
+Note that this representation is \emph{disjunctive}: the PVP represents only one crisp time interval, but for some reason, it is (partially) unknown.
 
-The ill-known interval approach has many advantages from the representation by the FVP as demonstrated in \cite{Pons2011},\cite{Pons2012}.
+The main feature for the PVP is that it preserves all the information for both starting and ending points\cite{Pons2011,Pons2012}.
 Table \ref{tbl:comparative-pvp-fvp} is a comparative between PVP and FVP.
@@
 \item Allen's relations: Are the Allen's relations defined?
-\item Storage: The way the data is stored in the database.
-\item Possibility measures: Does the framework provides always a possibility measure for any relation between the temporal elements?
-\item Necessity measures:  Does the framework provides always a necessity measure for any relation between the temporal elements?
+\item Storage: The way the data are stored in the database.
+\item Possibility measures: Does the framework provide always a possibility measure for any relation between the temporal elements?
+\item Necessity measures:  Does the framework provide always a necessity measure for any relation between the temporal elements?
 \end{enumerate}
@@
 This subsection is devoted to describe the behaviour of a crisp valid-time database.
-For the shake of simplicity, only the three main operations in the Data Manipulation Language \emph{DML} a.k.a \emph{CRUD} (CReate Update, Delete) operations are shown.
-Usually the DML operations in a temporal database are re-defined e.g.
-a typical update sentence in SQL could be re-defined by means of several insert and update sentences.
-Therefore, in order to be clear, these high level primitives in the DML for a valid-time database are usually noted as Insert, Modify and Delete.
+For the sake of simplicity, only the three main operations in the Data Manipulation Language (CReate Update, Delete) are shown.
+Usually the DML operations in a temporal database are re-defined (a typical update sentence in SQL could be re-defined by means of several insert and update sentences).
+Therefore, for the sake of clarity, these high level primitives in the DML for a valid-time database are usually noted as Insert, Modify and Delete.
 In the following subsections each primitive is defined and explained.
-Finally a illustrative example is given.
+Finally, an illustrative example is given.
 For a more complete information on the behaviour of a bitemporal database, please refer to ~\cite{Jensen1994}.
@@
 \begin{definition}
-Consider a relation $R$ in a temporal database, an entity $A$ given by the attributes $\left(a_1, \ldots, a_n \right)$ and the crisp time interval $I$ given by $\left[a,\ b\right] : a,b \in \T, a \leq b$.
-The pair $\left(A, I\right)$ expresses that the data regarding the entity A is valid during the crisp time interval I.
-Note that the relation $R = \left(A, I \right) x \left(A, I \right)$.
-The following equation indicates that the pair $\left(A, I\right)$ is in the relation R:
+\emph{Valid-time relation.}
+Consider $A=\left(A_1, \ldots, A_n \right)$ be the set of attributes that define an entity and $I = \left(s_i,e_i \right)$ the set of attribute defining the time interval.
+The pair $\left(A, I\right)$ expresses that the data regarding the entities A are valid during the crisp time interval I.
+Let $R \subseteq A\  x\  I$ be a valid-time relation.
+Then, the following equation indicates that the pair $\left(A, I\right)$ is in the relation R:
+
 \begin{equation}
@@
 \end{definition}
-\begin{example}
- Consider a library database.
-Each physical book is stored with an identifier (ID), a title and the start and end loan dates respectively.
-Table \ref{tbl:library-sample} illustrates the first version of the database, after three insertions.
-In this example, $A = ($ ID, Title $)$ and $I = ($ Start Loan, End Loan $)$.
-\end{example}
-
-\vglue13pt
-%\begin{table}[htbp]
-\tcap{\label{tbl:library-sample}Sample library database}
-\centerline{\small DATA TYPES}
-\vglue-6pt
-\centerline{\small\baselineskip=13pt
-\begin{tabular}{c c c c}\\
-\hline
-\textbf{ID}  & Title  & Start Loan & End Loan \\
-\hline
-3 & Dracula & 15/3/2012 & 30/3/2012 \\
-4 & Frankenstein & 25/3/2012 & 4/4/2012 \\
-5 & Harry Potter & 18/3/2012 & 2/4/2012\\
-3 & Dracula & 4/4/2012 & UC \\
-\hline\\
-\end{tabular}
-}
 
@@
 \begin{definition}
- \label{def:until-changed}
-\textbf{Until changed (UC):} To show that the data related to the entity $\left(A, I \right)$ is up-to-date, a special constant called \emph{Until Changed} is defined for the endpoint of the time interval $I$.
-The value for this constant is considered to be $+\infty$.
+ \label{def:from-beginning}
+\emph{From the beginning (FB).} 
+Consider $A=\left(A_1, \ldots, A_n \right)$ be the set of attributes that define an entity and $I = \left(s_i,e_i \right)$ the set of attribute defining the time interval.
+Let $R \subseteq A\  x\  I$ be a valid-time relation.
+The from the beginning constant is used when the value for $s_i$ is equal to $-\infty$.
 \end{definition}
 
-%quizas poner un ejemplo relacionado con lo de la liberia.
+\begin{definition}
+ \label{def:until-changed}
+\emph{Until changed (UC).} 
+Consider $A=\left(A_1, \ldots, A_n \right)$ be the set of attributes that define an entity and $I = \left(s_i,e_i \right)$ the set of attribute defining the time interval.
+Let $R \subseteq A\  x\  I$ be a valid-time relation.
+The until changed constant is used when the value for $e_i$ is equal to $+\infty$.
+\end{definition}
 
@@
 The value for the time interval is $I = \left(4/4/2012 , UC \right)$.
-The meaning is that the book with ID = 3 was loaned the  4/4/2012 and it is still loaned.
-The book with ID = 3 is current in the relation.
+The meaning is that the patient with ID = 3 was in the hospital the  4/4/2012 and it is still there.
+The patient with ID = 3 is current in the relation.
+
+\begin{example}
+\label{ex:library-database}
+ Consider a hospital database.
+Each patient is stored in the database with an identifier (ID), the name and the dates when the patient entered and left the hospital.
+Table \ref{tbl:library-sample} illustrates the first version of the database, after three insertions.
+In this example, $A = ($ ID, Name $)$ and $I = ($ Arrival, Leave $)$.
+\end{example}
+
+\vglue13pt
+%\begin{table}[htbp]
+\tcap{\label{tbl:library-sample}Sample hospital database}
+%\centerline{\small DATA TYPES}
+\vglue-6pt
+\centerline{\small\baselineskip=13pt
+\begin{tabular}{c c c c}\\
+\hline
+\textbf{ID}  & Name  & Arrival & Leave \\
+\hline
+3 & John & 15/3/2012 & 30/3/2012 \\
+4 & Antoon & 25/3/2012 & 4/4/2012 \\
+5 & Daan & 18/3/2012 & 2/4/2012\\
+3 & John & 4/4/2012 & UC \\
+\hline\\
+\end{tabular}
+}
 
 \begin{definition}
-\textbf{Current $\left(R, A \right)$:}
 \label{def:current-in-relation}
-This function returns a crisp time interval $I$ if the entity $A$ is current in the relation $R$ or $\emptyset$ if the entity $A$ is not current in the relation.
+\emph{Current $\left(R, A \right)$.}
+Consider $A=\left(A_1, \ldots, A_n \right)$ be the set of attributes that define an entity and $I = \left(s_i,UC \right)$ the set of attribute defining the time interval.
+Let $R \subseteq A\  x\  I$ be a valid-time relation.
+The function Current $\left(R, A \right)$ returns the crisp time interval $I$ if the entity $A$ is currently in the relation $R$ or the empty set, if the entity $A$ is not currently in the relation.
 
@@
 \nonumber
-I & \mbox{ if } (A,I) \in R \wedge I = \left(c, UC \right), c \in \T\\
+I & \mbox{ if } (A,I) \in R \mbox{ and } I = \left(c, UC \right), c \in \T\\
 \emptyset & \mbox{ in any other case }
@@
 
-For example, the function Current(R, (3,`Dracula')) returns the time interval $I = \left(4/4/2012 , UC  \right)$. On the other hand, Current(R, (4,`Frankenstein')) returns the empty set.
+For example, the function Current(R, (3,`John')) returns the time interval $I = \left(4/4/2012 , UC  \right)$. On the other hand, Current(R, (4,`Antoon')) returns the empty set.
 
-The following two Allen's relations (see Figure \ref{fig:allen-rel}) are defined like in~\cite{Nagypal2003} to implement the auxiliary functions.
+The Allen's relations between two time intervals are shown in Figure \ref{fig:allen-rel}.
+The complete implementation of the relations using only the starting and the ending points is defined in~\cite{Nagypal2003}.
 
@@
 \vspace*{10pt}
-\fcaption{\label{fig:allen-rel}Allen's relations for two crisp time intervals $I$ and $J$.}
+\fcaption{\label{fig:allen-rel}Allen's relations for two crisp time intervals $I$ and $J$.
+Note that the crisp time interval $I$ is fixed and the different positions of the interval $J$ illustrate the Allen's relations.}
 \vspace*{13pt}
@@
 \begin{definition}
-\textbf{Overlaps:}
+\emph{Overlaps:}
  \label{def:overlaps}
@@
 \begin{definition}
-\textbf{During:}
+\emph{During:}
  \label{def:during}
@@
 
+In the rest of the paper, and without losing generality, we will consider that the time granularity are days. Also, the dates will be given in the format \emph{dd/mm/yyyy}.
+
 \begin{definition}
-\textbf{CloseR$\left( I, J \right)$:}
+\emph{CloseR$\left( I, J \right)$:}
 \label{def:close-a-crisp-interval-r}
@@
 %I=\left[a, c \right[ & \mbox{ if } b = +\infty \wedge J > I
-\left[c, e \right[ & \mbox{ if } d = UC \wedge J \mbox{ During } I \\
+\left[c, e \right[ & \mbox{ if } d = UC \mbox{ and } J \mbox{ During } I \\
 I & \mbox{ in any other case.}
@@
 Now it is possible to close the current version of an entity by using \eqref{eq:close-a-crisp-interval} and \eqref{eq:current-in-relation}.
-This functionality is required to append new information about an existing entity in the relation.
+This functionality is required to add or update new information about an existing entity in the relation.
 
@@
 \label{def:close-current-version}
-Consider an entity $A$, the relation $R$ and a time interval $J$.
-Then, the function close-current$\left(R,\left(A,J\right) \right)$ closes any current version of the object $A$ if it exists and add the new version $\left(A, J \right)$.
+\emph{Close-current$\left(R, \left(A, J  \right)\right)$}.
+Consider $A=\left(A_1, \ldots, A_n \right)$ be the set of attributes that define an entity and $J = \left(s_j,e_j \right)$ the set of attribute defining the time interval.
+Let $R \subseteq A\  x\  I$ be a valid-time relation.
+Then, the function close-current$\left(R,\left(A,J\right) \right)$ closes any current version of the object $A$ if it exists and add the new version $\left(A, J\right)$.
 
@@
 \nonumber
-\mbox{ if } I = \mbox{ Current } \left(R, A \right)\\ %\wedge J > \mbox{ CloseR } \left(I, J \right)   \\
+\mbox{ if  Current } \left(R, A \right) \neq \emptyset\\ %\wedge J > \mbox{ CloseR } \left(I, J \right)   \\
 \nonumber R , \text{ in any other case}
@@
 
-For example, consider the book with ID = 3 in Table \ref{tbl:library-sample}.
-The current version was loaned on 4/4/2012.
-The book was returned to the library but, for some reason, the end loan date was not registered.
-The book is loaned again on 24/4/2012.
-Then, the Close-current function allows to close the current version of the book and insert a new version.
+For example, consider the patient with ID = 3 in Table \ref{tbl:library-sample}.
+The current version arrived on 4/4/2012.
+The patient left the hospital but, for some reason, the date  was not registered.
+The patient arrived again at the hospital on 24/4/2012.
+Then, the Close-current function allows to close the current version of the patient and insert a new version.
 First, the function CloseR is applied with $I = \left(4/4/2012, UC \right)$ and $J = \left(24/4/2012, UC \right)$.
@@
 Then, the modifications on the value for the time interval $I$ are stored.
-Finally, a new row with the current values of the book and the time interval $J$ is stored.
+Finally, a new row with the current values of the patient and the time interval $J$ is stored.
 The result of this operation is illustrated on Table \ref{tbl:library-sample-2}.
@@
 \nonumber
-\mbox{ Close-current }\left(R, A, \mbox{Current}\left(R, A \right) \right) %\\
+\mbox{ Close-current }\left(R, \left(A, I\right) \right) %\\
 % \mbox{ if Current}\left(R ,A \right) \neq \emptyset \\
@@
 \begin{enumerate}
-\item The object $A$ was never in the relation $R$: The object is added with the valid-time indicated by the crisp interval $I$. In the example, this correspond with the following insertion sentences:
-\begin{verbatim}
-Insert(3,'Dracula',
-(15/3/2012, 30/3/2012));
-Insert(4,'Frankenstein',
-(25/3/2012, 4/4/2012));
-Insert(5,'Harry Potter',
-(18/3/2012, 2/4/2012));
-\end{verbatim}
+\item The object $A$ was never in the relation $R$: The object is added with the valid-time indicated by the crisp interval $I$. 
+% In the example, this correspond with the following insertion sentences:
+% \begin{verbatim}
+% Insert(3,'Dracula',
+% (15/3/2012, 30/3/2012));
+% Insert(4,'Frankenstein',
+% (25/3/2012, 4/4/2012));
+% Insert(5,'Harry Potter',
+% (18/3/2012, 2/4/2012));
+% \end{verbatim}
 
@@
 This is illustrated in Table \ref{tbl:library-sample}.
-The insert sentence is:
-	      \begin{verbatim}
-Insert(3,'Dracula',
-(4/4/2012, UC));
-	      \end{verbatim}
+%  The insert sentence is:
+% 	      \begin{verbatim}
+% Insert(3,'Dracula',
+% (4/4/2012, UC));
+% 	      \end{verbatim}
 
@@
 If the book is already in the library for a new loan, then it is necessary to set the return date and add a new row with the new loan date.
-This is illustrated in Table \ref{tbl:library-sample-2}, the insert sentence is:
-	    \begin{verbatim}
-Insert(3,'Dracula',
-(24/4/2012, UC));
-	      \end{verbatim}
+This is illustrated in Table \ref{tbl:library-sample-2}.
+%  the insert sentence is:
+% 	    \begin{verbatim}
+% Insert(3,'Dracula',
+% (24/4/2012, UC));
+% 	      \end{verbatim}
 	\item Reject the insertion, if the time interval $I$ does overlap any existing valid-time interval for the object $A$ in the relation $R$.
@@
 As this interval does overlaps other time intervals, it is not possible that the book was loaned in that time interval.Therefore, the insertion is rejected.
-The insert sentence is:
-
-	      \begin{verbatim}
-Insert(3,'Dracula',
-(6/4/12/2012,25/4/2012));
-	    \end{verbatim}
+%  The insert sentence is:
+% 
+% 	      \begin{verbatim}
+% Insert(3,'Dracula',
+% (6/4/12/2012,25/4/2012));
+% 	    \end{verbatim}
 
@@
 \hline
-\textbf{ID}  & Title  & Start Loan & End Loan \\
+\textbf{ID}  & Name  & Arrival & Leave \\
 \hline
-3 & Dracula & 15/3/2012 & 30/3/2012 \\
-4 & Frankenstein & 25/3/2012 & 4/4/2012 \\
-5 & Harry Potter & 18/3/2012 & 2/4/2012\\
-3 & Dracula & 4/4/2012 & 23/4/2012 \\
-3 & Dracula & 24/4/2012 & UC \\
+3 & John & 15/3/2012 & 30/3/2012 \\
+4 & Antoon & 25/3/2012 & 4/4/2012 \\
+5 & Daan Potter & 18/3/2012 & 2/4/2012\\
+3 & John & 4/4/2012 & 23/4/2012 \\
+3 & John & 24/4/2012 & UC \\
 \hline\\
@@
 \nonumber
-R \cup \left(A, I \right) & \mbox{ if }  (A, I ) \not \in R  \vee \\
- &   \left(A, J\right) \in R \wedge \lnot \left(I \mbox{ Overlaps } J\right) \\
+R \cup \left(A, I \right) & \mbox{ if }  (A, I ) \not \in R  \mbox{ or } \\
+ &   \left(A, J\right) \in R \mbox{ and } \lnot \left(I \mbox{ Overlaps } J\right) \\
 \mbox{modify(R,(A,I))} & \mbox{ otherwise }    %\mbox{ if } \exists J \left(A,J\right) \in R \wedge \mbox{Current(A,J)} \\ 
@@
 \nonumber
-R - \left(A, I \right), & \forall \left( A, I \right) \in R\\
+R - \left \lbrace \left(A, I_1 \right),  \ldots, \left(A, I_1 \right)   \right \rbrace, & \forall I_i : \left( A, I_i \right) \in R\\
 R, & \mbox{ otherwise }  
@@
 
-For example, consider that the librarian wants to delete the history for the book with ID = 3. The result of this operation is shown in Table \ref{tbl:library-sample-3}. The delete sentence is:
+For example, consider that the hospital wants to delete the history for the patient with ID = 3. The result of this operation is shown in Table \ref{tbl:library-sample-3}.
 
-\begin{verbatim}
- Delete(3,'Dracula');
-\end{verbatim}
+% The delete sentence is:
+% 
+% \begin{verbatim}
+%  Delete(3,'Dracula');
+% \end{verbatim}
 
@@
 %\begin{table}[htbp]
-\tcap{\label{tbl:library-sample-3}Sample library database}
+\tcap{\label{tbl:library-sample-3}Sample hospital database}
 %\centerline{\small DATA TYPES}
@@
 \hline
-\textbf{ID}  & Title  & Start Loan & End Loan \\
+\textbf{ID}  & Name  & Arrival & Leave\\
 \hline
-4 & Frankenstein & 25/3/2012 & 4/4/2012 \\
-5 & Harry Potter IV & 18/3/2012 & 2/4/2012\\
+4 & Antoon & 25/3/2012 & 4/4/2012 \\
+5 & Daan & 18/3/2012 & 2/4/2012\\
 \hline\\
@@
 
-\subsubsection{\label{subsubsec:revise}Revise}
-This operation replaces the old values for the entity $A'= \left(a_1', \ldots, a_n' \right) $ with the new values specified by $A = \left(a_1, \ldots, a_n \right)$ in the relation $R$, at a given time interval $I$.
-This operation is defined when a correction on the values for the entity $A$ should be done.
-\begin{align}
-\REV\left(R, \left(A, I \right) \right)=\\
-\begin{cases}
-\nonumber
- R - \left(A', I \right) \cup \left(A, I \right) & \mbox{ if } \left(A', I \right) \in R  \\
-   %: A = \left(a_1, \ldots, a_n \right),\\
-%A'= \left(a_1', \ldots, a_n' \right)   \\   
-R & \mbox{ otherwise }
-\end{cases}
-\end{align}
-
-For example, consider now that the librarian wants to make a correction in the title for the book with ID = 5.
+% \subsubsection{\label{subsubsec:revise}Revise}
+% This operation replaces the old values for the entity $A'= \left(a_1', \ldots, a_n' \right) $ with the new values specified by $A = \left(a_1, \ldots, a_n \right)$ in the relation $R$, at a given time interval $I$.
+This operation is used when a correction on the values for the entity $A$ should be done.
+% \begin{align}
+% \REV\left(R, \left(A, I \right) \right)=\\
+% \begin{cases}
+% \nonumber
+%  R - \left(A', I \right) \cup \left \lbrace \left(A, I \right) \right \rbrace & \mbox{ if } \left(A', I \right) \in R  \\
+%    %: A = \left(a_1, \ldots, a_n \right),\\
+% %A'= \left(a_1', \ldots, a_n' \right)   \\   
+% R & \mbox{ otherwise }
+% \end{cases}
+% \end{align} 
+% 
+% For example, consider now that the librarian wants to make a correction in the title for the book with ID = 5.
 The new title is `Harry Potter IV'.
 The result of this operation is illustrated in Table \ref{tbl:library-sample-3}.
- Then, the revise sentence is:
-
-\begin{verbatim}
-Revise(5,'Harry Potter IV', 
-(18/3/2012, 2/4/2012));
-\end{verbatim}
+%  Then, the revise sentence is:
+% 
+% \begin{verbatim}
+% Revise(5,'Harry Potter IV', 
+% (18/3/2012, 2/4/2012));
+% \end{verbatim}

### 2012-06-12 trunk/ijcis-pos-model/sources/timeRepresentation.tex
preliminary final modifications

@@
 This section is devoted to specify the representation of time within the framework of the possibility theory.
-First of all, the specification for a single ill-known temporal point shall be explained.
+First of all, the specification for a single ill-known temporal point will be explained.
 Then, the formal specification and the related constraints are given for an ill-known valid-time interval.
@@
 \subsection{\label{subsec:ill-known-point-rep}Ill-known time point}
-An ill-known time point $X$ is a precise time point that for some reason, it is not fully specified. Note that $X$ has only one possible value but that value is unspecified.
+An ill-known time point $X$ is a precise time point that, for some reason, is not fully specified. Note that $X$ has only one possible value but that value is unspecified.
 
@@
 \label{def:ill-known-time-point}
-\emph{Ill-known time point}\\
-Consider a time domain $\T$, uncertainty about the values of the ill-known time point $X$ is given by the possibility distribution $\pi_X$:
+\textbf{Ill-known time point.}\\
+Consider a time domain $\T$; the uncertainty about the values of the ill-known time point $X$ is given by the possibility distribution $\pi_X$:
 \begin{equation}
@@
 
+It is also possible to specify an ill-known time point by a convex combination of ill-known constraints, as shown by equation \eqref{eq:ill-known-value-def-by-const-unc}.
+
 \begin{definition}
 \label{def:ill-known-domain}
-\emph{Domain for an ill-known time point}\\
-Consider $\Pow(\T)$ the set of all the possibility distributions over $\T$, and the three fuzzy constants \emph{UNKNOWN} $=\left\lbrace 1/t, t \in \T \right\rbrace$, \emph{UNDEFINED} $=\left\lbrace 0/t, t \in \T \right\rbrace$ and \emph{NULL} $=\lbrace 1/$ \emph{UNKNOWN}, $1/$ \emph{UNDEFINED} $\rbrace$.
+\textbf{Domain for an ill-known time point.}\\
+Consider $\Pow(\T)$ the set of all the possibility distributions over $\T$, and the three fuzzy constants \emph{UNKNOWN} $=\left\lbrace 1/t, \forall t \in \T \right\rbrace$, \emph{UNDEFINED} $=\left\lbrace 0/t,\forall t \in \T \right\rbrace$ and \emph{NULL} $=\lbrace 1/$ \emph{UNKNOWN}, $1/$ \emph{UNDEFINED} $\rbrace$.
 The domain for an ill-known time point $X$ is given by:
@@
 
-It is also possible to specify an ill-known time point by just a set of ill-known constraints plus
-
 \subsubsection{\label{subsubsec:ill-known-time-datatypes}Datatypes}
-The datatypes for an ill-known time point $X$ are shown in Table \ref{tbl:time-point-types}.
+The data type for the representation of an ill-known time point allows the representation of the values shown in Table \ref{tbl:time-point-types}
+%The datatypes for an  $X$ are shown in Table \ref{}.
 \vglue13pt
 %\begin{table}[htbp]
-\tcap{Data types for a time point}
-\centerline{\small DATA TYPES}
+\tcap{\label{tbl:time-point-types}Values for the time point data type.}
+\centerline{\small TIME POINT DATA TYPES}
 \vglue-6pt
 \centerline{\small\baselineskip=13pt
-\begin{tabular}{c p{2cm} p{2cm}}\\
+\begin{tabular}{c p{0.4\columnwidth} p{0.3\columnwidth}}\\
+Subtype & Value & Representation \\
 \hline
@@
 \begin{example}
-Consider a historical database with data from diplomatic documents.
-The following fields are stored: the digital identifier \emph{ID} which is the primary key  and the stimated time when the document was sent (field \emph{Date}).
+Consider a historical database with data from medieval diplomatic documents.
+The following fields are stored: the digital identifier \emph{ID} which is the primary key  and the estimated time when the document was sent (field \emph{Date}).
 Table \ref{tbl:sample-time-point} contains some example data from this database.
@@
 
+%\begin{definition}
+\textbf{Ill-known time interval}\\
+An ill-known time interval denoted by $\left[X, Y\right]$ is a precise time interval whose boundaries are not precisely known.
+The ill-known time interval is defined by two ill-known  time points, namely $X$,$Y$.
+The evaluation of the ill-known time interval is given iby equations \eqref{ill-known-pos},\eqref{ill-known-nec}.
+
+% In order to know whether all the points in the crisp time interval $I=\left[a, b\right]$ are within the boundaries of the ill-known time interval $\left[X, Y\right]$ we define the following two ill-known constraints $C_{1}, C_{2}$
+% \begin{eqnarray}
+% \label{eq:constraint-c1}
+% C_1\stackrel{\triangle}{=}\left(\geq,X\right)  \\
+% \label{eq:constraint-c2}
+% C_2\stackrel{\triangle}{=}\left(\leq,Y\right)
+% \end{eqnarray}
+% Note that the boolean operator applied to the constraints is $\bool=\wedge$. Hence, the construct of both constraints is $C_1 \wedge C_2$
+% This means that we want to know if all the points in the interval $I$ are greater than or equal to $X$ and, at the same time, smaller than or equal to $Y$.
+% %\end{definition}
+% 
+% %% equations for the poss and nec!!
+% The simplified equations for the possibility and the necessity measures are~\cite{Pons2011}:
+% \begin{eqnarray}
+% \label{eq:time-interval-pos}
+% \Pos\left(\lambda([a,b])\right)&=&\\
+% \nonumber
+% \min\bigg(\sup_{a\leq w}\pi_{X}(w),\sup_{b\geq w}\pi_{Y}(w)\bigg)\\
+% \label{eq:time-interval-nec}
+% \Nec\left(\lambda([a,b])\right)&=&\\
+% \nonumber
+% \min\bigg(\inf_{a>w}1-\pi_{X}(w),\inf_{b<w}1-\pi_{Y}(w)\bigg).
+% \end{eqnarray}
+
+\subsubsection{\label{subsubsec:open-interval}Open ill-known time intervals}
+Quite often, the user may want to specify time intervals with open boundaries in one or both endpoints.
+Consider an ill-known time interval $\left[X, Y\right]$.
+Then it is possible to distinguish between the following two types of open intervals:
+%\begin{enumerate}
+%\item
 \begin{definition}
-\emph{Ill-known time interval}\\
-An ill-known time interval denoted by $\left[X, Y\right]$ is a precise time interval that for some reason, its boudaries are not precissely known.
-The ill-known interval is defined by two ill-known points, namely $X$,$Y$.
-In order to know if all the points in the crisp time interval $I=\left[a, b\right]$ are within the boudaries of the ill-known time interval $\left[X, Y\right]$ we define the following two ill-known constraints $C_{1}, C_{2}$
-\begin{eqnarray}
-C_1\stackrel{\triangle}{=}\left(\geq,X\right) \wedge \\
-C_2\stackrel{\triangle}{=}\left(\leq,Y\right)
-\end{eqnarray}
-Note that the boolean operator applied to the constraints is $\bool=\wedge$.
-This means that we want to know if all the points in the interval $I$ are greater or equal to $X$ and, at the same time, smaller or equal to $Y$.
+\textbf{Completely unknown time interval}: Both starting and ending points are unknown, therefore the whole interval is unknown.
+%specified by the \emph{UNKNOWN} constant. Due to the representation, both endpoints, say $X$ and $Y$, are stored. Thus, the interval is stored as $[$ \emph{UNKNOWN}, \emph{UNKNOWN} $]$. 
 \end{definition}
 
-%% equations for the poss and nec!!
-The simplified equations for the possibility and the necessity measures are~\cite{Pons2011}:
-\begin{eqnarray}
-\label{eq:interval-pos}
-\Pos\left(\lambda([a,b])\right)&=&\\
-\nonumber
-\min\bigg(\sup_{a\leq w}\pi_{X}(w),\sup_{b\geq w}\pi_{Y}(w)\bigg)\\
-\label{eq:interval-nec}
-\Nec\left(\lambda([a,b])\right)&=&\\
-\nonumber
-\min\bigg(\inf_{a>w}1-\pi_{X}(w),\inf_{b<w}1-\pi_{Y}(w)\bigg).
-\end{eqnarray}
+%\item
+\begin{definition}
+\textbf{Semi-open time interval}: Only one of the two ill-known endpoints for the time interval $\left[X, Y\right]$ is unknown.
+Example \ref{ex:open-time-interval} and Figure \ref{fig:example-open-time-interval} illustrates a left open time interval.
+% Note that due to the ill-known constraints $C_{1}, C_{2}$ it is not possible a representation like $[$ \emph{UNKNOWN} $, Y]$ or $[X, $ \emph{UNKNOWN} $]$.
+The solution adopted for this problem is explained in the following paragraph.
+\end{definition}
+%\end{enumerate}
 
-\subsubsection{\label{subsubsec:open-interval}Open ill-known time intervals}
-Quite often, the user may want to specify time intervals with open boudaries in one or both endpoints.
-Consider an ill-known time interval $\left[X, Y\right]$.
-Then it is possible to distinguish between the following two types of open intervals:
-\begin{enumerate}
-\item
-\emph{Completely unknown time interval}: Both starting and ending points are unknown, therefore the hole interval is specified by the \emph{UNKNOWN} constant.
-Due to the representation, both endpoints, say $X$ and $Y$ are stored.
-Thus, the interval is stored as $[$ \emph{UNKNOWN}, \emph{UNKNOWN} $]$.
-\item
-\emph{Semi-open time interval}: Only one of the two ill-known endpoints for the time interval $\left[X, Y\right]$ is unknown.
-Note that due to the ill-known constraints $C_{1}, C_{2}$ it is not possible a representation like $[$ \emph{UNKNOWN} $, Y]$ or $[X, $ \emph{UNKNOWN} $]$.
-The solution adopted for this problem is explained in the following paragraph.
-\end{enumerate}
+\subsubsection{\label{subsubsec:representation-semi-open}Representation of semi-open time intervals}
+As mentioned before, the problem resides in the representation of this kind of interval.
 
-\paragraph{Representation of semi-open time intervals}
-As mentioned before, the problem resides in the representation of these kind of interval.
-In order to get a short representation for these intervals, two constants are defined: $FB$ \emph{From the Beggining}: it is only valid for the left value of the interval and $UC$ \emph{Until Changed} (this name is usual in the temporal database community, see ~\cite{Jensen1994}) which is only valid for the right value of the interval.
+% In order to get a short representation for these intervals, two constants are defined: $FB$ -\emph{From the Beginning}- it is only valid for the left value of the interval and $UC$ -\emph{Until Changed}- (this name is usual in the temporal database community, see ~\cite{Jensen1994}) which is only valid for the right value of the interval.
 Note that these two constants are aliases for a function that obtains both possibility and necessity measures.
 
-Because of the ill-known constraints $C_{1}, C_{2}$, a function called \emph{Open} should be defined in order to deal with a proper representation of these intervals:
+Because of the ill-known constraints $C_{1}, C_{2}$, a function called \emph{Open} should be defined in order to deal with a proper representation of these intervals.
 
@@
 $Open(C)$\\
-The function $\Open(C)$ provides both possibility and necessity measures for all the points in the open part of a semi-open ill-known time interval.
-The parameter $C$ is an ill-known constraint $C\stackrel{\triangle}{=}\left(R,T\right)$ as defined in equation \eqref{rand} and in~\cite{Pons2011}.
+Consider an ill-known value $T$, a binary relationship $B_r$ one of $\left(\leq, <, >, \geq\right)$ and the constraint $C\stackrel{\triangle}{=}\left(R,T\right)$.
+The function $\Open(C) = \left(I_p(C), I_n(C) \right)$ provides both possibility and necessity measures for all the points in the open part of a semi-open ill-known time interval.
+
 The possibility and necessity measures are defined by:
@@
 \label{eq:open-pos-nec}
-\Pos\left(\Open(C)\right) &=& \Big(\sup_{r\in\T \Rp\  w}\pi_{T}(w)\Big)\\
-\Nec\left(\Open(C)\right) &=& \Big(\inf_{r\in\T \Rn\  w}1-\pi_{T}(w)\Big)
+I_p \left(C \right) &=& \Big(\sup_{r\in\T \Rp\  w}\pi_{T}(w)\Big)\\
+I_n \left(C \right) &=& \Big(\inf_{r\in\T \Rn\  w}1-\pi_{T}(w)\Big)
 \end{eqnarray}
@@
 %\begin{table}[htbp]
-\tcap{Relations for the $\Open(C)$ function.}
+\tcap{Relations for the $\Open(C)$ function. Depending on the relation $B_r = \left(\leq, <, >, \geq\right)$ in the constraint $C$, the values for $\Rp$ and $\Rn$ are shown.}
 \centerline{\small Relations}
@@
 Figure \ref{fig:example-open-time-interval} shows the representation for $Y$.
-The user wants to obtain the possibility and the necessity measures for the $\FB$ part of the interval are:
+The user wants to obtain the possibility and the necessity measures for the $\FB$ part of the interval.
 \end{example}
@@
 \nonumber
-\Pos\left(\Open(C)\right) = \Big(\sup_{r\in\T \geq\  w}\pi_{Y}(w)\Big)\\
+I_p(C_2)= \Big(\sup_{r\in\T \geq\  w}\pi_{Y}(w)\Big)\\
 \nonumber
-\Nec\left(\Open(C)\right) = \Big(\inf_{r\in\T <\  w}1-\pi_{Y}(w)\Big)
+I_n(C_2) = \Big(\inf_{r\in\T <\  w}1-\pi_{Y}(w)\Big)
 \end{eqnarray}
@@
 
+% \vglue13pt
+% %\begin{table}[htbp]
+% \tcap{Relations for the $\Open(C)$ function.}
+% \centerline{\small Relations}
+% \vglue-6pt
+% \centerline{\small\baselineskip=13pt
+% \begin{tabular}{c c c}\\
+% \hline
+% $R$ & $\Rp$ & $\Rn$\\ \hline
+% $<$ & $>$ & $\leq$ \\
+% $\leq$ & $\geq$ & $<$ \\
+% $>$ & $<$ & $\geq$ \\
+% $\geq$ & $\leq$ & $>$ \\
+% \hline\\
+% \end{tabular}
+% \label{tbl:open-pos-nec-rels}
+% }
+
 \subsubsection{\label{subsubsec:ill-known-time-interval-datatypes}Datatypes}
-In order to represent properly an ill-known time interval in a database, some datatypes are needed.
+In order to properly represent  an ill-known time interval in a database, some datatypes are needed.
 Because of the ill-known constraints, not all the combinations of datatypes for each ill-known time point (see Table \ref{tbl:time-point-types}) are allowed.
-Table \ref{tbl:time-interval-types} shows all the possible combinations for the datatypes for an ill-known time interval denoted by $\left[X, Y\right]$.
+Table \ref{tbl:time-interval-types} shows all the possible datatypes combinations for the  for an ill-known time interval denoted by $\left[X, Y\right]$.
 
@@
 }
+
+% Maybe an example with a database.
+% \begin{example}
+% 
+% \end{example}
+
+% \subsection{\label{subsec:fuzzy-allen-relations}Allen's relations}
+% The Allen's relations can be represented by a Boolean combination of ill-known constraints.
+The constructs for the constraints are shown in Table \ref{tbl:fuzzy-allen-relations}.
+The possibility and necessity measures are obtained for each Boolean function as explained in Section \ref{subsec:interval-evaluation-by-ill-known-constraints} in equations \eqref{eq:conjunctive1}-\eqref{eq:negation}.
+% 
+% 
+% \vglue13pt
+% %\begin{table}[htbp]
+% \tcap{\label{tbl:fuzzy-allen-relations}Constructs of constraints related to their respective Allen relations, as used in the presented work.
+In this table, the ill-known time interval $J = \left[X, Y\right]$ in a record $r$ has a start point described by possibilistic variable $X$ and an end point described by possibilistic variable $Y$.
+The crisp time interval in the user's temporal demand is denoted $I$.}
+% \centerline{\small Relations}
+% \vglue-6pt
+% \centerline{\small\baselineskip=13pt
+% \begin{tabular}{p{0.2\columnwidth} p{0.8\columnwidth} }\\
+% \hline
+% Allen Relation & Construct of constraints\\
+% \hline
+% I before J & $C_1 \triangleq \left(<,X\right)$ \\
+% \hline
+% I equal J & $\left(C_1 \triangleq \left(\geq,X\right)\right) \wedge \neg \left(C_2 \triangleq \left(\neq,X\right) \right) \wedge \left(C_3 \triangleq \left(\leq,Y\right) \right) \wedge \neg \left(C_4 \triangleq \left(\neq,Y\right)\right)$ \\
+% \hline
+% I meets J & $\left(C_1 \triangleq \left(\leq,X\right)\right) \wedge \neg \left(C_2 \triangleq \left(\neq,X\right) \right)$ \\
+% \hline
+% I overlaps J & $\left(C_1 \triangleq \left(<,Y\right)\right) \wedge \neg \left(C_2 \triangleq \left(\leq,X\right) \right) \wedge \neg \left(C_3 \triangleq \left(\geq,X\right) \right)$ \\
+% \hline
+% I during J & $\left( \left (C_1 \triangleq \left(>,X\right) \right) \wedge \left(C_2 \triangleq \left(\leq,Y\right) \right) \right) \vee \left( \left(C_3 \triangleq \left(\geq,X\right) \right) \wedge  \left(C_4 \triangleq \left(<,Y\right) \right) \right)$ \\
+% \hline
+% I starts J & $\left(C_1 \triangleq \left(\geq,X\right) \right) \wedge \neg \left(C_2\triangleq \left(\neq,X\right)\right)$ \\
+% \hline
+% I finishes J & $\left(C_1 \triangleq \left(\leq,Y\right) \right) \wedge \neg \left(C_2\triangleq \left(\neq,Y\right)\right)$ \\
+% \hline\\
+% \end{tabular}
+% }
 

### 2012-06-12 ijcis-pos-model/sources/abstract.tex
preliminary final modifications

@@
 %%%%%%%%%%%%%%%%%%%%%%%%
-In reality, some objects or concepts have properties with a time-variant or time-related nature.
-Modelling these kinds of objects or concepts in a (relational) database schema is possible, but time-variant and time-related attributes have an impact on the consistency of the entire database.
-Therefore, temporal database models have been proposed to deal with this.
-Time itself can be at the source of imprecision, vagueness and uncertainty, since existing time measuring devices are inherently imperfect.
-Accordingly, human beings manage time using temporal indications and temporal notions, which may contain imprecision, vagueness and uncertainty.
-However, the imperfection in human-used temporal indications is supported by human interpretation, whereas information systems need support for this.
-Several proposals for dealing with such imperfections when modelling temporal aspects exist.
-Some of these proposals transform the temporal expression into a compact representation.
-Other proposals consider the temporal indications in the used temporal expressions to be the source of imperfection.
-In this work we present a novel model to deal with imperfections in valid-time databases. Next to that, the data manipulation language, \emph{DML} is defined and implemented.
+In real world, it is very common that some objects or concepts have properties with a time-variant or time-related nature.
+Modelling this kind of objects or concepts in a (relational) database schema is possible, but time-variant and time-related attributes have an impact on the consistency of the entire database.
+Therefore, temporal database models have been proposed to deal with this problem in the literature.
+Time can be affected by imprecision, vagueness and / or  uncertainty, since existing time measuring devices are inherently imperfect.
+Additionally, human beings manage time using temporal indications and temporal notions, which may also be imprecise.
+However, the imperfection in human-used temporal indications is supported by human interpretation, whereas information systems need appropiate support in order to accomplish this task.
+Several proposals for dealing with such imperfections when modelling temporal data exist.
+Some of these proposals transform the temporal data into a compact representation.
+There is not a model for managing and handling uncertainty regarding temporal information.
+In this work we present a novel model to deal with imprecision in valid-time databases toghether with the data manipulation language, \emph{DML} is defined and implemented.

### 2012-06-12 ijcis-pos-model/sources/preliminaries.tex
preliminary final modifications

--- 2012-05-24 ijcis-pos-model/sources/preliminaries.tex
@@
 %Introductory text:
-In this section, some basic concepts are introduced concerning possibilistic variables and fuzzy numbers and intervals, after which the framework of set evaluation by ill-known constraints \cite{Pons2011} is explained.
+In this section, we introduce some basic concepts concerning possibilistic variables and fuzzy numbers and intervals.
+Then, the framework of set evaluation by ill-known constraints \cite{Pons2011} is explained.
 The section concludes with a brief introduction to temporal databases.
 
+% %Something about relational databases in general and relations in specific (to make the vocabulary clear)
+% \subsection{Relational Databases}
+% In this subsection, a few structural and behavioral aspects of the relational database model will be presented.
+% 
+% In a relational database, information is structured in \emph{relations}.
+As explained in the introduction, a collection of similar entities is modelled by an entitytype, which, using the relational database model, will be modelled by a relation.
+A relation consists of a \emph{relation schema} and an \emph{extention}.
+The relation schema basically determines the relation structure, by dictating which data (describing entities) will be kept and how these data will be represented.
+Each (atomic) part of these data is a result value of a measurement of a property of an entity or a description of a property of an entity.
+The extention contains these data.
+A relation schema consists of a \emph{relation name} and a finite set of \emph{attributes}.
+The relation name is of course the name of the entire relation, while the attributes describe common properties of the entities modelled by the relation.
+Every attribute consists of an attribute name and an associated \emph{data type}.
+Each data type has its \emph{domain} and its operators.
+The domain of a data type is the set of values which data of the data type in question can assume.
+Operators of a data type can be applied to data of a data type.
+The extention of a relation is a set of \emph{tuples}.
+Each tuple represents an enitity, by containing values: for every attribute in the attribute set of the relation schema, the tuple contains the value describing the exact state of the property of the entity.
+%Remember: a tuple is just an ordered tuple of values.
+
 \subsection{\label{subsec:possibility-theory}Possibility Theory}
-Possibility theory, like probability theory, deals with uncertainty about the outcome of an experiment.
-In probability theory, this uncertainty is caused by the \emph{variability} in the outcomes, while in possibility theory, the uncertainty is caused by \emph{incomplete knowledge} about the experiment.
-The quantification of confidence in a theory of uncertainty is achieved using a confidence measure\cite{Pons2011}.
-In probability theory this is a measure of chance, in possibility theory, possibility and necessity measures are used.
+Possibility theory is devoted to the handling of incomplete information. In order to capture partial ignorance, two measures are provided: possibility and necessity.
 
@@
 
-Both possibility measures and necessity measures are special cases of confidence measures.
+Both possibility and necessity measures are special cases of confidence measures.
 
@@
 
-Regarding interpretation, the above can be seen as: the degree to which an event is necessary is the degree to which every other possible event is not plausible.
+That is, the degree to which an event is necessary is the extent to every other possible event is not plausible.
 
 \subsection{\label{subsec:possibilistic-variables}Possibilistic Variables}
-Possibilistic variables rely on possibility theory \cite{Dubois1988a}. A \emph{possibilistic variable} is defined as follows \cite{Pons2011}.
+Possibilistic variables rely on possibility theory \cite{Dubois1988a} and are defined as follows \cite{Pons2011}.
 
@@
 A possibilistic variable $X$ over a universe $U$ is defined as a variable taking exactly one value in $U$, but for which this value is (partially) unknown.
-Its possibility distribution $\pi_X$ on $U$ models the available knowledge about the value that $X$ takes: for each $u\in U$, $\pi_X(u)$ represents the possibility that $X$ takes the value $u$.
+Possibility distribution $\pi_X$ on $U$ models the available knowledge about the value $X$ takes; for each $u\in U$, $\pi_X(u)$ represents the possibility that $X$ takes the value $u$.
 In this work, this possibility is interpreted as a measure of how plausible it is that $X$ takes the value $u$, given (partial) knowledge about the value $X$ takes.
@@
 
-When a possibilistic variable is defined on the powerset $\Pow(R)$ of some universe $R$, the unique value the variable takes will be a crisp set and its possibility distribution on the powerset $\Pow(R)$ will describe the possibility of each crisp subset of $R$ to be the value the variable takes.
+When a possibilistic variable is defined on the powerset $\Pow(U)$ of some universe $U$, the unique value the variable takes will be a crisp set and its possibility distribution on the powerset $\Pow(U)$ will describe the possibility of each crisp subset of $U$ to be the value the variable takes.
 This exact value (a crisp set) the variable takes, is now called an \emph{ill-known set} \cite{Dubois1988a}.
@@
 Indeed, any Boolean proposition $p$ takes just one value in $\mathbb{B}$.
-If the knowledge about which value this proposition $p$ will take is given by a possibility distribution $\pi_p$, the proposition can be seen as a possibilistic variable.
-As the interest lies with the case where the proposition holds, the possibility and necessity that $p$ = $T$ (the proposition holds) demand most attention.
+If the knowledge about which value this proposition $p$ will take is given by a possibility distribution $\pi_p$; the proposition can be seen as a possibilistic variable.
+The possibility and necessity that $p$ = $T$ (the proposition holds) demand more attention.
 This possibility and necessity is noted here as:
@@
 
-Here, equation \eqref{propholdsposs} denotes the possibility that $p$ = $T$ and the proposition holds, equation \eqref{propholdsnecc} denotes the necessity that $p$ = $T$ and the proposition holds.
+Here, equation \eqref{propholdsposs} denotes the possibility that $p$ = $T$ and the proposition holds, while equation \eqref{propholdsnecc} denotes the necessity that $p$ = $T$ and the proposition holds.
 
-This work will deal with ill-known intervals.
-These are ill-known sets, defined and represented via a start and end point, which will be ill-known values.
+This work deals with ill-known intervals.
+These are ill-known sets, defined and represented via a starting and ending points, which are ill-known values.
 The elements of the set are the values between the starting and ending points.
 A closed ill-known interval with starting point defined by possibilistic variable $X$ and ending point by possibilistic variable $Y$ is noted here $\left[X, Y\right]$.
-The correspondences and transitions between the representations of ill-known sets, between the representations of ill-known intervals and between the representations of an ill-known set and an ill-known interval are part of the authors current research.
+%The correspondences and transitions between the representations of ill-known sets, between the representations of ill-known intervals and between the representations of an ill-known set and an ill-known interval are part of the authors current research.
 
@@
 \begin{definition}
-A fuzzy interval is a fuzzy set $M$, defined by a membership function $\mu_{M}$, on the set of real numbers $\mathbb{R}$ such that:
+A fuzzy interval is a fuzzy set $M$, defined by a membership function $\mu_{M}$ on the set of real numbers $\mathbb{R}$ such that:
 \begin{eqnarray}
-\mu_{M} :  \mathbb{R} \rightarrow \left[0,1\right] \nonumber \\ 
+\mu_{M} :  \mathbb{R} \rightarrow \left[0,1\right]  \\ 
+\nonumber
 \forall (u,v)\in\mathbb{R}^2: \forall w \in [u,v]:\\
+\nonumber
 \mu_M(w) \geq\min(\mu_M(u),\mu_M(v))  \\
+\nonumber
 \exists m \in \mathbb{R} :  \mu_M(m)=1 
@@
 It can be shown that such a membership function $\mu_T$ for a fuzzy interval $T$ is convex and normalized.
-Four real values, denoted $\alpha$, $\beta$, $\gamma$ and $\delta$ suffice to represent a trapezoidal membership function of a fuzzy interval.
-In this work, a fuzzy interval defined as such will be noted as $\left[\alpha, \beta, \gamma, \delta\right]$.
-The corresponding membership function definition for this $\mu_T$ is then given by:
+The values $\left[\alpha, \beta, \gamma, \delta\right]$ represent a trapezoidal possibility distribution defined by $\mu_T$:
+
+% Four real values, denoted $\alpha$, $\beta$, $\gamma$ and $\delta$ suffice to represent a trapezoidal membership function of a fuzzy interval.
+In this work, a fuzzy interval defined as such will be noted as .
+The corresponding membership function definition for this  is then given by:
 
@@
 \mu_T : & \quad \mathbb{R} \rightarrow \left[0,1\right] \\
+\nonumber
  : & \quad x \rightarrow
@@
 
-The most convenient form of the membership function of a fuzzy number is a triangular form.
+The representation for a triangular possibility distribution is given by three values $\left[D,a,b \right]$.
+The distribution is obtained from a trapezoidal possibility distribution by replacing the values $\left[D-a, D, D, D+b \right]$.
+
+% The most convenient form of the membership function of a fuzzy number is a triangular form.
 It can be shown that such a membership function $\mu_M$ for a fuzzy number $M$ is convex and normalized.
@@
 Here:
-\begin{itemize}
-\item
-$D$ denotes the single value in the core of $M$
-\item
-$D-a$ is then $\inf \{u \in \mathbb{R} : \mu_{M}(u) > 0\}$
-\item
-$D+b$ is then $\sup \{u \in \mathbb{R} : \mu_{M}(u) > 0\}$
-\end{itemize}
-
-The membership function of such a fuzzy number is then given by:
-
-\vspace{-10pt}
-
-\begin{align}
-\mu_M :   \mathbb{R} \rightarrow \left[0,1\right] :  x \rightarrow \nonumber\\
-\begin{cases}
-\nonumber
-0 & \mbox{ if } \left(x < D-a\right) \vee \left(x > D+b\right) \\
-\frac{\left[x-\left(D-a\right)\right]}{a} & \mbox{ if } \left(x \geq D-a\right) \wedge (x \leq D)  \\
--\frac{\left[x-\left(D+b\right)\right]}{b} & \mbox{ if } \left(x \geq D\right) \wedge (x \leq D+b)
-\end{cases}
-\end{align}
+% \begin{itemize}
+% \item
+% $D$ denotes the single value in the core of $M$
+% \item
+% $D-a$ is then $\inf \{u \in \mathbb{R} : \mu_{M}(u) > 0\}$
+% \item
+% $D+b$ is then $\sup \{u \in \mathbb{R} : \mu_{M}(u) > 0\}$
+% \end{itemize}
+% 
+% The membership function of such a fuzzy number is then given by:
+% 
+% \vspace{-10pt}
+% 
+% \begin{align}
+% \mu_M :   \mathbb{R} \rightarrow \left[0,1\right] :  x \rightarrow \nonumber\\
+% \begin{cases}
+% \nonumber
+% 0 & \mbox{ if } \left(x < D-a\right) \vee \left(x > D+b\right) \\
+% \frac{\left[x-\left(D-a\right)\right]}{a} & \mbox{ if } \left(x \geq D-a\right) \wedge (x \leq D)  \\
+% -\frac{\left[x-\left(D+b\right)\right]}{b} & \mbox{ if } \left(x \geq D\right) \wedge (x \leq D+b)
+% \end{cases}
+% \end{align}
 
@@
 \subsection{\label{subsec:interval-evaluation-by-ill-known-constraints}Interval Evaluation by Ill-known Constraints}
-The problem of interval evaluation is more generally explained in \cite{Pons2011}: the need exists to know if all points in a crisp interval $I$ reside between the boundaries of an ill-known interval $\left[ X , Y \right]$.
-In \cite{Pons2011}, the notion of an \emph{ill-known constraint} is introduced:
+In our case, we need to know if all points in a crisp interval $I$ reside between the boundaries of an ill-known interval $\left[ X , Y \right]$.
+To compute that, we introduce the concept of \emph{ill-known constraint}\cite{Pons2011}.
 
@@
 \end{align}
-Set $A$ now satisfies the constraint if and only if:
+Set $A$ satisfies the constraint if and only if:
 \begin{align}
@@
 
-An example of an ill-known constraint is $C_{ex} \triangleq (<, X)$. Some set $A$ then satisfies $C_{ex}$ if $\forall a \in A : a < X$, given possibilistic variable $X$.
+An example of an ill-known constraint is $C_{ex} = (<, X)$. Some set $A$ satisfies $C_{ex}$ if $\forall a \in A : a < X$, given the possibilistic variable $X$.
 
-The satisfaction of a constraint $C \triangleq (R,V)$ by a set $A$ is basically still a Boolean matter, but due to the uncertainty about the ill-known value $V$, it can be uncertain whether $C$ is satisfied by $A$ or not \cite{Pons2011}.
+The satisfaction of a constraint $C \triangleq (R,V)$ by a set $A$ is still a Boolean matter, but due to the uncertainty about the ill-known value $V$, it can be uncertain whether $C$ is satisfied by $A$ or not \cite{Pons2011}.
 In fact, this satisfaction now behaves as a proposition.
-Based on the possibility distribution $\pi_{V}$ of $V$, the possibility and necessity that $A$ satisfies $C$ can be found.
+Based on the possibility distribution $\pi_{V}$ of $V$, the possibility and necessity that $A$ satisfies $C$ can be computed.
 This proposition can thus be seen as a possibilistic variable on $\mathbb{B}$.
@@
 \begin{align}
+\label{ill-known-pos}\
 \Pos(A\text{ satisfies }C) & =\\
 \nonumber
-\min_{a \in A}\left(\sup_{(a,w) \in R}\pi_{V}(w)\right) \label{ill-known-pos}\\
+\min_{a \in A}\left(\sup_{(a,w) \in R}\pi_{V}(w)\right) \\
+\label{ill-known-nec}
 \Nec(A\text{ satisfies }C) & =\\
 \nonumber
-\min_{a \in A}\left(\inf_{(a,w) \notin R} 1-\pi_{V}(w)\right) \label{ill-known-nec}
+\min_{a \in A}\left(\inf_{(a,w) \notin R} 1-\pi_{V}(w)\right) 
 \end{align}
 
-So far, we have shown how it can be verified if a set satisfies or not an ill-known constraint.
+So far, we have shown how it can be verified whether a set satisfies or not an ill-known constraint. The interval evaluation problem is explained in a more general context in \cite{Pons2011}.
 
 \begin{example}
+As example, consider  $X = \left[5, 3, 2 \right]$ and let $C =(\leq, X)$ the ill-known constraint.
+Then, the evaluation of the possibility and the necessity are obtained from \eqref{ill-known-pos} and \eqref{ill-known-nec} respectively.
+(See Figure \ref{fig:example-ill-known}).
 
@@
 \vspace*{10pt}
-\fcaption{\label{fig:example-ill-known}Example of the evaluation of the ill-known constraint $C \triangleq (\leq, X)$.
-Possibility and necessity measures are shown in grey the upper and lower graphics respectively.
-}
+\fcaption{\label{fig:example-ill-known}Example of the evaluation of the ill-known constraint $C \triangleq (\leq, X)$. Possibility and necessity measures are shown in grey. }
 %  \label{fig:fuzzy-validity-period}
 \vspace*{13pt}
-
- Consider an ill-known value given by $X = \left[5, 3, 2 \right]$.
-$C \triangleq (\leq, X)$ is the ill-known constraint.
-The set $A$ is $A \subseteq \R^+$.
-Then, the evaluation of the possibility and the neccesity are obtained from \eqref{ill-known-pos} and \eqref{ill-known-nec} respectively.
-(See Figure \ref{fig:example-ill-known}).
 
@@
 
- However, from the examples at the beginning of this section, it is observed that Boolean combinations of constraints are required.
-For example, the problem of interval evaluation as explained earlier requires that all elements of an interval $[a,b]$ are larger than a value $X$ and at the same time smaller than a value $Y$, which implies that a conjunctive Boolean combination of both constraints must be satisfied.
+It is observed that Boolean combinations of constraints are required.
+For example, the problem of interval evaluation (explained earlier) requires that all elements of an interval $[a,b]$ are larger than a value $X$ and, at the same time, smaller than a value $Y$, which implies that a conjunctive Boolean combination of both constraints must be satisfied.
 To allow Boolean combinations of constraints, the following definitions are introduced.
@@
 \begin{equation}
-\lambda:\Pow(U)\rightarrow\mathbb{B}:A\mapsto\bool\Big(C_{1}(A),...,C_{n}(A)\Big).
+\lambda:\Pow(U)\rightarrow\mathbb{B}:\lambda(A)\mapsto\bool\Big(C_{1}(A),...,C_{n}(A)\Big).
 \end{equation}
@@
 \end{definition}
-It is well known that any Boolean function $\bool$ can be cast to a canonical form~\cite{McCluskey1965}, requiring only the logical conjunction $\wedge$, logical disjunction $\vee$ and logical negation.
-Therefore, only the case of Boolean conjunction, Boolean disjunction and Boolean negation will be treated within the scope of this paper.
+It is well known that any Boolean function $\bool$ can be cast to a canonical form~\cite{McCluskey1965}, requiring only logical conjunction $\wedge$, logical disjunction $\vee$ and logical negation.
+Therefore, only those connectives will be treated within the scope of this paper.
 By applying the possibilistic extensions of $\wedge$, $\vee$ and $\neg$, concrete equations are obtained for the calculations of uncertainty about the evaluation of a set by means of an evaluation function $\lambda$.
@@
 \nonumber
-\min_{i=1}^n\Pos\left(C_i(A)\right)\\
+\min_{i=1 \ldots n}\Pos\left(C_i(A)\right)\\
 \label{eq:conjunctive2}
@@
 \nonumber
-\min_{i=1}^n\Nec\left(C_i(A)\right).
+\min_{i=1 \ldots n}\Nec\left(C_i(A)\right).
 \end{eqnarray}
@@
 \nonumber
-\max_{i=1}^n\Pos\left(C_i(A)\right)\\
+\max_{i=1 \ldots n}\Pos\left(C_i(A)\right)\\
 \forall A\in\Pow(U):\Nec(\lambda(A))&=\\
 \nonumber
-\max_{i=1}^n\Nec\left(C_i(A)\right).
+\max_{i=1 \ldots n}\Nec\left(C_i(A)\right).
 \end{eqnarray}
@@
 \begin{example}
-Consider that we want to check if crisp interval $I = \left[j, k\right]$ is included in $\left[X, Y\right]$, 2 ill-known constraints are constructed:
+Consider that we want to check if crisp interval $I = \left[j, k\right]$ is included in $\left[X, Y\right]$. In this situation, two ill-known constraints are constructed.
 
@@
 
-There is a special boolean combination of constraints that is of particular interest:
+There is a special boolean combination of constraints that is of particular interest; let us see it.
 \begin{definition}
-\emph{Convex combination of ill-known constraints}. Consider the $C_1$ and $C_2$ two ill-known constraints. A convex combination of the constraints is given by:
+\emph{Conjunctive combination of ill-known constraints}.
+Consider a universe $U$.
+Let $R_1, R_2 \subseteq U$ be two binary relations in U.
+Let $X_1, X_2$ be ywo fixed ill-known values in $U$.
+Let  $C_1 = (R_1, X_1)$ and $C_2 = (R_2, X_2)$ be two ill-known constraints.
+A conjunctive combination of both constraints is given by:
 \begin{align}
@@
 \end{align}
-In a more general way, it is possible to define the convex combination of an n-ary vector  $\mathbf{C}$ of constraints:
+In a more general way, it is possible to define the conjunctive combination of an n-ary vector of constraints:
 \begin{align}
@@
 \label{th:convex-combination-ill-known-constraints}
-Consider the convex combination $CC$ of any n-ary vector $\mathbf{C} = \left \lbrace C_{1_{Z_1}}, \ldots, C_{n_{Z_n}} \right \rbrace$ of constraints over the ill-known variables $Z_1, \ldots, Z_n$.
+Consider the conjunctive combination $CC$ of any n-ary vector $\left \lbrace C_{1_{Z_1}}, \ldots, C_{n_{Z_n}} \right \rbrace$ of constraints over the ill-known variables $Z_1, \ldots, Z_n$.
 Then, if $\pi_{C_1(Z_1)}, \ldots, \pi_{C_n(Z_n)}$ are convex, then  $\pi_{CC}$ is also convex.
@@
 \begin{proof}
-Let $CC \triangleq \left \lbrace C_{1_{Z_1}}, \ldots, C_{n_{Z_n}}  \right \rbrace$. Then:
+Let $CC \triangleq \left \lbrace C_{1_{Z_1}} \wedge \ldots \wedge C_{n_{Z_n}}  \right \rbrace$. Then:
 \begin{align}
@@
 \pi_{CC} \left( \lambda x_1 + \left( 1 - \lambda \right)x_2 \right) = \\
+\nonumber
 \min  \big( \pi_{C_1(Z_1)}\left( \lambda x_1 + \left( 1 - \lambda \right)x_2 \right), \ldots,\\
+\nonumber
    \pi_{C_n(Z_n)}\left( \lambda x_1 + \left( 1 - \lambda \right)x_2 \right)  \big)
@@
 \label{def:convex-combination-ill-known-constraints}
-Consider a universe $U$, $X$ an ill-known value and $CC \triangleq \left \lbrace C_{1_{Z_1}}, \ldots, C_{n_{Z_n}}  \right \rbrace$ a convex combination of ill-known constraints over the variables $Z_1, \ldots, Z_n$.
-The value $X$ is defined by means of the convex combination:
+\emph{Ill-known value defined by conjunctive combination of constraints.}
+Consider a universe $U$, $X$ an ill-known value and $CC \triangleq \left \lbrace C_{1_{Z_1}} \wedge \ldots \wedge C_{n_{Z_n}}  \right \rbrace$ a convex combination of ill-known constraints over the variables $Z_1, \ldots, Z_n$.
+Then, the value $X$ is defined by means of the convex combination:
 \begin{align}
@@
 \end{align} 
-Note that $\pi_{X}$ is convex since $\pi_{CC}$ is convex as demonstrated in theorem \ref{th:convex-combination-ill-known-constraints}.
+Note that $\pi_{X}$ is convex since $\pi_{CC}$ is convex as demonstrated in Theorem \ref{th:convex-combination-ill-known-constraints}.
 \end{definition}
@@
 \begin{example}
-Consider the ill-known values $X = \left[12, 2, 2\right]$ and $Y = \left[18, 2, 1 \right]$. The ill-known value $Z$ is defined using the convex combination $CC$ of constraints $C_1$ and $C_2$:
+As an example, consider the hospital database.
+When a patient enters o leaves a medical service in hospital, a record tracking the patient is stored.
+The exact time for this is not precissely known.
+Consider now that $X = \left[12, 2, 2\right]$ is the time in hours when a patient left the emergency service and $Y = \left[18, 2, 1 \right]$ is the time in hours when the patient left the hospital.
+It is possible to obtain $Z$ (the period of time between the emergency service and the left of the patient) by using a conjunctive combination of constraints.
+
 \begin{align}
 \nonumber
-CC \triangleq \left \lbrace C_1\left(>,X\right) \wedge C_2(\leq,Y) \right \rbrace \\
+CC = \left \lbrace C_1\left(>,X\right) \wedge C_2(\leq,Y) \right \rbrace \\
 \nonumber
-Z \triangleq CC
+Z = CC
 \end{align}
@@
 %\end{figure}
+
+%enlace:
+We have seen the main theoretical concepts about ill-known values.
+Now, we are going to explain the main concepts about the treatment of time and the imperfection related to the time in databases.
+In the following subsections we will show how to represent imperfect time intervals by means of ill-known values.

### 2012-06-12 ijcis-pos-model/sources/temporalModel.tex
preliminary final modifications

--- 2012-04-18 ijcis-pos-model/sources/temporalModel.tex
@@
 \subsection{\label{subsec:temporal-model}The generalized temporal model}
-The model is based on the GEFRED~\cite{Medina1994} (Generalized Model of Fuzzy Relational DB) model.
-This model is extended by adding valid-time support.
+The model is based on GEFRED~\cite{Medina1994} (Generalized Model of Fuzzy Relational DB) model.
+This model is extended by adding valid-time support which will be illustrated with an example.
 The information in the system is defined by the following elements:
@@
 \label{def:typeof-domain}
-The function typeof$\left(D \right)$ returns the datatype associated with the domain $D$.
-
+\textbf{Typeof(D).}
+The function typeof$\left(D \right)$ returns the datatype associated with the domain $D$. For every datatype, the function returns the associated number as shown in Table \ref{tbl:gefred-data-types}.
 \end{definition}
@@
 
-It is possible to define a more specific generalized temporal domain, $\T_G$:
+It is possible to define a more specific generalized temporal domain, $\T_G$.
 
@@
 
+\begin{example}
+\label{ex:sample-ft-database}
+ Consider a library database as explained in example \ref{ex:library-database}.
+But now, the time is stored with imprecision.
+For simplicity, the ill-known time points are represented as triangular fuzzy numbers.
+An ill-known time point is given by $[dd/mm/yyyy,a,b]$.
+The values for $a$ and $b$ are integers.
+Thus, the date given by $[15/3/2012,5,2]$ is a triangular fuzzy number with the left bound the 10/3/2012, the core on the 15/3/2012 and the right bound on 17/3/2012.
+Table \ref{tbl:fuzzy-library-sample} shows the elements Head, $\Head$ and Body, $\Body$. 
+\begin{align}
+ \label{eq:head-ex}
+\Head = & \lbrace \left(\text{ID}:D_{\text{ID}}\right),\\
+\nonumber
+&  \left(\text{Title}:D_{\text{title}}\right),\\
+\nonumber
+& \left(\text{Loan}:D_{\text{Loan}}\right) \\
+\nonumber
+& \left(\text{PVP}:D_{\text{PVP}}\right) \rbrace
+\end{align}
+
+When the compatibility degree is 1, the component is omitted. The body, $\Body$ consists on all the tuples shown in Table \ref{tbl:fuzzy-library-sample}.
+
+\end{example}
+
+\vglue13pt
+%\begin{table}[htbp]
+\tcap{\label{tbl:fuzzy-library-sample}Sample library database}
+%\centerline{\small DATA TYPES}
+\vglue-6pt
+\centerline{\small\baselineskip=13pt
+\begin{tabular}{c c c |p{0.35\columnwidth}}\\
+\hline
+& & & \centerline{PVP} \\
+$\Head$ &  \textbf{ID}  & Title  &  Start Loan , End Loan\\
+\hline
+\multirow{3}{*}{$\Body$ } &  3 & Dracula & [15/3/2012, 5, 2] , [30/3/2012, 1, 1] \\
+ &4 & Frankenstein & [25/3/2012, 3, 2] , [4/4/2012, 1, 7] \\
+&5 & Harry Potter & [18/3/2012, 4, 1] , [2/4/2012, 2, 2]\\
+% \ 3 & Dracula & [4/4/2012, 3, 3] & UC \\
+\hline\\
+\end{tabular}
+}
+
 \begin{definition}
 \label{def:value-component}
+\textbf{Value component.}
 The value component $R^{v}_{FTG}$ of a fuzzy temporal relation $R_{FTG}$ is a set with the value components for both the head and the body of the relation:
@@
 \end{align}
-
 \end{definition}
 
-\begin{example}
+Consider now the database in Table \ref{tbl:fuzzy-library-sample-valid-time}. The primary key is now:
 
-\end{example}
+\begin{align}
+ \nonumber
+K_{GT} \subseteq \Head, K_{GT} = \left \lbrace  \left(\text{ID}:D_{\text{ID}} \right) \cup  \left(V_{\text{ID}}:D_{\text{ID}} \right) \right \rbrace	
+%\nonumber
+ %\left. \cup  \left(V_{ID}:D_{ID} \right)	\right \rbrace \\
+\end{align}
+
+\vglue13pt
+%\begin{table}[htbp]
+\tcap{\label{tbl:fuzzy-library-sample-valid-time}Sample library database}
+%\centerline{\small DATA TYPES}
+\vglue-6pt
+\centerline{\small\baselineskip=13pt
+\begin{tabular}{c c c c | p{0.35\columnwidth} }\\
+\hline
+% & & & \multicolumn{2}{c}{PVP} \\
+%& & & & \centerline{PVP}\\
+{$\Head$}  &  &  & &(Start Loan\\
+& \textbf{ID}  & \textbf{V} & Title& , End Loan)	\\
+ \hline
+\multirow{4}{*}{$\Body$}  & 3 & 001 & Dracula & [15/3/2012, 5, 2], [30/3/2012, 1, 1] \\
+& 4 & 001 & Frankenstein & [25/3/2012, 3, 2] , [4/4/2012, 1, 7]\\
+& 5 & 001 & Harry Potter & [18/3/2012, 4, 1] , [2/4/2012, 2, 2]\\
+ & 3 & 002 & Dracula & [4/4/2012, 3, 3] , UC \\
+\hline\\
+\end{tabular}
+}
 
@@
 
-It is important to nottice that whereas the result of the evaluation of any comparison between crisp time intervals is boolean, the evaluation of any comparison between \emph{PVPs} is in the unit interval.
-Therefore, for a valid-time object, say $A \in R$, in a crisp temporal database it is not possible an overlaping among any valid-time interval ($\forall I, \not \exists J: \left(A, I \right) \in R \wedge \left(A, J \right) \in R \wedge \not \Overlap\left(I, J \right)$).
-Thus, from the point of view of a fuzzy temporal database, there are the following options:
+It is important to notice that whereas the result of the evaluation of any comparison between crisp time intervals is boolean, the evaluation of any comparison between \emph{PVPs} is in the unit interval.
+Therefore, for a valid-time object, say $A \in R$, in a crisp temporal database it is not possible an overlapping among any valid-time interval ($\forall I, \not \exists J: \left(A, I \right) \in R \wedge \left(A, J \right) \in R \wedge \lnot \Overlap\left(I, J \right)$).
+Thus, from the point of view of a fuzzy temporal database, there are the following options (see Figure \ref{fig:tdb-class}):
 
-\begin{itemize}
-\item [\textbf{Strictly consistent}] It is not possible for two PVPs of the same object to be valid at the same time.
+% \begin{itemize}
+% \item 
+\begin{definition}
+\textbf{Strictly consistent} It is not possible for two PVPs of the same object to be valid at the same time.
 In other words, if $A$ is an object, $R$ is a relation and $\pi_{I}\left(t \right)$ and $\pi_{J}\left(t \right)$ are the possibilities for a time point $t \in \T$ to be in the PVPs $I$ and $J$ respectively,
@@
 \end{align}
+\end{definition}
 
-\item [\textbf{Ill-consistent}] The overlapping of two PVPs of the same object do overlap. It can be distinghised among the following sub-types:
-	\begin{itemize}
-	\item [\emph{Co-existence}] Two versions of the same object $A$ may exist at the same time.
+%\item 
+\begin{definition}
+ \textbf{Ill-consistent} The overlapping of two PVPs of the same object do overlap. It can be distinguished among the following sub-types:
+\end{definition}
+% 	\begin{itemize}
+% 	\item 
+	      \begin{definition}
+	      \textbf{Co-existence} Two versions of the same object $A$ may exist at the same time.
 		      \begin{align}
@@
 		      \end{align}
-	\item [\emph{Weak-consistence}] Several versions of the same object $A$ may exist at the same time.
+	      \end{definition}
+
+	%\item 
+	      \begin{definition}
+	      \textbf{Weak-consistence} Several versions of the same object $A$ may exist at the same time.
 		      \begin{align}
@@
 		      \end{align}
-	\end{itemize}
-\end{itemize}
+	      \end{definition}
+
+	%\end{itemize}
+
+\vspace*{13pt}
+\begin{center}
+{
+\includegraphics[scale=1.25]{./graphs/tdb-clas.pdf}
+
+}
+\end{center}
+%\centerline{ \psfig{file=./graphs/Y-time-point.eps}}
+\vspace*{10pt}
+\fcaption{\label{fig:tdb-class}Classification for a fuzzy temporal database.
+The validity for an entity is represented by a trapezoid.
+Graphic (a) shows an example of strictly consistent database.
+Between two versions of the entity it is possible some amount of overlapping, but always holds equation \eqref{eq:stricly-consistent}.
+Graphic (b) shows an example of a co-existence.
+Two versions of the same entity may overlap in any degree, but not more than two versions.
+Graphic (c) shows an example of weak-consistence.
+The different versions of the same entity are shown in an unfolded way, to clarify the multiple overlapping.
+}
+\vspace*{13pt}
+
+%\end{itemize}
 
 In the implementation of the DML operations we will consider a strictly consistent temporal database.
+This is a natural extension of the crisp temporal databases.
+This approach allows a easy implementation for the DML operations defined in Section \ref{subsubsec:Understanding-valid-time-databases}.
+
 Since the time intervals are now possibilistic valid-time periods, PVPs, the auxiliary functions defined in equations \eqref{eq:close-a-crisp-interval} to \eqref{eq:close-current} should be re-defined.
@@
 
-\subsubsection{\label{subsubsec:insert-fuzzy-temporal}Insert}
+For example, consider the database in Table  \ref{tbl:fuzzy-library-sample-valid-time}.
+The function Close-current1$\left(R, \left(ID=3\right) \right.,$ $\left.
+\left[ \left[15/4/2012,2,1\right], UC\right]\right)$ closes the current version of the book with ID=3 and creates a new version with the specified time interval.
 
 \subsubsection{\label{subsubsec:modify-fuzzy-temporal}Modify}
+This operation adds new information about an existing entity $A$ in the relation $R$. The modify operation does not remove any previous value of the entity $A$.
+Note that the modify operation is only applicable when the object $A$ is current in the relation $R$; $ \left(A, I \right) \in R, I = \left[X, \UC \right]$.
 
+\begin{align}
+\label{eq:modify-fuzzy-temporal}
+\MOD \left( R, \left( A, I \right) \right) =\\
+%\begin{cases}
+\nonumber
+\mbox{ Close-current }\left(R, \left(A, I\right) \right) 
+% \mbox{ if Current}\left(R ,A \right) \neq \emptyset \\
+% R,  \mbox{ otherwise }
+%\end{cases}
+\end{align}
+
+\subsubsection{\label{subsubsec:insert-fuzzy-temporal}Insert}
+The user wants to store an entity $A$ which is valid in the relation $R$ during the time interval specified by the PVP, $I = \left(X, Y \right)$.
+%
+%
+%The interpretation of the insert operation is the following: The user wants to store in the database that an object $A$  is true for some period(s) of time $I$ in the given relation $R$.
+To indicate that the object $A$ is current in the relation, the special value Until Current, $\UC$  is used.
+There are the following cases when performing an insert operation:
+\begin{enumerate}
+\item The entity $A$ was never in the relation $R$: The entity is added with the valid-time indicated by the PVP, $I$.
+For example, consider the database given by Table \ref{tbl:fuzzy-library-sample-valid-time}. The following sentences correspond with the insertion of the first version of each book.
+
+\begin{verbatim}
+Insert(3,'Dracula',
+[[15/3/2012, 5, 2],
+        [30/3/2012, 1, 1]]);
+Insert(4,'Frankenstein',
+[[25/3/2012, 3, 2], 
+        [4/4/2012, 1, 7]]);
+Insert(5,'Harry Potter',
+[[18/3/2012, 4, 1], 
+        [2/4/2012, 2, 2]]);
+\end{verbatim}
+
+\item The entity $A$ is in the relation $R$. Depending on the value of the time interval $I$, there are three possibilities:
+	\begin{enumerate}
+	\item Insert $\left(A, I\right)$ in the relation $R$.
+If the time interval $I$ does not overlap any other valid-time interval in the relation $R$.
+Note that here, the result of the Overlaps operator is in the unit interval.
+For example, the librarian loans the book with ID=3 but the book is not returned yet. The insert sentence is the following.
+ \begin{verbatim}
+Insert(3,'Dracula',
+[[4/4/2012, 3, 3], UC]);
+	      \end{verbatim}
+
+	\item Modify and close the current version of the entity $A$ and insert $\left(A, I \right)$.
+For example, consider now that the book with ID 3 is loaned  around the 24/4/2012 (this is model by a triangular fuzzy number: $\left[24/4/2012, 1, 1 \right]$).
+Here the problem is that the book with ID 3 was loaned around 4/4/2012, but not returned.
+If the book is already in the library for a new loan, then it is necessary to set the return date and add a new row with the new loan date.
+%This is illustrated in Table \ref{tbl:library-sample-2}, the insert sentence is:
+	    \begin{verbatim}
+Insert(3,'Dracula',
+[[24/4/2012, 1, 1], UC]]);
+	      \end{verbatim}
+
+	\item Reject the insertion, if the time interval $I$ do overlap with a degree of 1 any existing valid-time interval for the entity $A$ in the relation $R$.
+For example, consider now that the librarian wants to introduce a past loan for the book with ID 3.
+The book was loaned around 6/4/12/2012 and returned around 25/4/2012.
+As this interval does overlaps other time intervals with a degree of 1, it is not possible that the book was loaned in that time interval.Therefore, the insertion is rejected.
+The insert sentence is:
+
+	      \begin{verbatim}
+Insert(3,'Dracula',
+[[6/4/12/2012, 1, 1],
+         [25/4/2012, 1, 1]]);
+	    \end{verbatim}
+	\end{enumerate}
+
+\end{enumerate}
+
+The algorithm for the implementation of the insert operation is the following:
+
+\begin{align}
+\label{eq:insert-fuzzy-temporal}
+\INS \left(R ,\left(A, I \right) \right) &=\\
+\begin{cases}
+\nonumber
+R \cup \left(A, I \right), & \mbox{ if }  \left(A, I \right) \not \in R \vee \\
+&  \left(A, J \right) \in R \wedge I \text{Overlaps} J < 1\\
+%\big \lbrace \left( \exists J= \left(X', Y' \right): \left(A, J \right) \in R \wedge \right. \\
+%\left. Y' \in \left \lbrace \left(<, X \right)\right \rbrace  \right) \vee \\
+%\left( \exists K= \left(X'', Y'' \right): \left(A, K \right) \in R \wedge \right. \\
+%\left. X'' \in \left \lbrace \left(>, Y \right)\right \rbrace \vee \right) \big \rbrace\\
+%&  J<I<K  \big \rbrace\\
+\MOD \left(R, \left(A, I \right)\right), & \mbox{ otherwise }  
+\end{cases} 	
+\end{align}
 \subsubsection{\label{subsubsec:delete-fuzzy-temporal}Delete}
+The delete operation logically removes an entity $A$ which is valid in the relation $R$:
+\begin{align}
+\label{eq:delete-fuzzy-temporal}
+\DEL \left(R ,A \right) =\\
+\begin{cases}
+\nonumber
+R - \left(A, I \right), &  \forall \left( A, I \right) \in R\\
+R, & \mbox{ otherwise }  
+\end{cases} 	
+\end{align}
+
+For example, consider that the librarian wants to delete the history for the book with ID = 3. The following sentence deletes all the versions for the book with ID = 3.
+
+\begin{verbatim}
+ Delete(3,'Dracula');
+\end{verbatim}
 
 \subsubsection{\label{subsubsec:revise-fuzzy-temporal}Revise}
+This operation replaces the old values for the entity $A'= \left(a_1', \ldots, a_n' \right) $ with the new values specified by $A = \left(a_1, \ldots, a_n \right)$ in the relation $R$, at a given time interval $I$.
+This operation is used when a correction on the values for the entity $A$ should be done.
+\begin{align}
+\label{eq:revise-fuzzy-temporal}
+\REV\left(R, \left(A, I \right)\right)=\\
+\begin{cases}
+\nonumber
+ R - \left(A', I \right) \cup \left \lbrace\left(A, I \right)\right \rbrace  & \mbox{ if } \left(A', I \right) \in R \\%: A = \left(a_1, \ldots, a_n \right),\\
+%A'= \left(a_1', \ldots, a_n' \right)   \\   
+R, & \mbox{ otherwise }
+\end{cases}
+\end{align}
+
+For example, consider now that the librarian wants to make a correction in the title for the book with ID = 5. The new title is `Harry Potter IV'. The revise sentence is:
+
+\begin{verbatim}
+Revise(5,'Harry Potter IV', 
+[[18/3/2012, 4, 1], 
+          [2/4/2012, 2, 2]]);
+\end{verbatim}

### 2012-06-13 ijcis-pos-model/sources/timeRepresentation.tex
version sent to Olga and Nicolas

@@
 \subsection{\label{subsec:ill-known-point-rep}Ill-known time point}
-An ill-known time point $X$ is a precise time point that, for some reason, it is not fully specified. Note that $X$ has only one possible value but that value is unspecified.
+An ill-known time point $X$ is a precise time point that, for some reason, is not fully specified. Note that $X$ has only one possible value but that value is unspecified.
 
@@
 \textbf{Ill-known time point.}\\
-Consider a time domain $\T$, uncertainty about the values of the ill-known time point $X$ is given by the possibility distribution $\pi_X$:
+Consider a time domain $\T$; the uncertainty about the values of the ill-known time point $X$ is given by the possibility distribution $\pi_X$:
 \begin{equation}
@@
 
-It is also possible to specify an ill-known time point by a convex combination of ill-known constraints, (Definition \ref{def:convex-combination-ill-known-constraints}) as shown by equation \eqref{eq:ill-known-value-def-by-const-unc}.
+It is also possible to specify an ill-known time point by a convex combination of ill-known constraints, as shown by equation \eqref{eq:ill-known-value-def-by-const-unc}.
 
@@
 \tcap{\label{tbl:time-point-types}Values for the time point data type.}
-\centerline{\small TIME POINT DATA TYPE}
+\centerline{\small TIME POINT DATA TYPES}
 \vglue-6pt
@@
 \begin{example}
-Consider a historical database with data from diplomatic documents.
+Consider a historical database with data from medieval diplomatic documents.
 The following fields are stored: the digital identifier \emph{ID} which is the primary key  and the estimated time when the document was sent (field \emph{Date}).
@@
 
-\begin{definition}
+%\begin{definition}
 \textbf{Ill-known time interval}\\
 An ill-known time interval denoted by $\left[X, Y\right]$ is a precise time interval whose boundaries are not precisely known.
-The ill-known interval is defined by two ill-known points, namely $X$,$Y$.
-In order to know if all the points in the crisp time interval $I=\left[a, b\right]$ are within the boundaries of the ill-known time interval $\left[X, Y\right]$ we define the following two ill-known constraints $C_{1}, C_{2}$
-\begin{eqnarray}
-\label{eq:constraint-c1}
-C_1\stackrel{\triangle}{=}\left(\geq,X\right)  \\
-\label{eq:constraint-c2}
-C_2\stackrel{\triangle}{=}\left(\leq,Y\right)
-\end{eqnarray}
-Note that the boolean operator applied to the constraints is $\bool=\wedge$. Hence, the construct of both constraints is $C_1 \wedge C_2$
-This means that we want to know if all the points in the interval $I$ are greater or equal to $X$ and, at the same time, smaller or equal to $Y$.
-\end{definition}
+The ill-known time interval is defined by two ill-known  time points, namely $X$,$Y$.
+The evaluation of the ill-known time interval is given by equations \eqref{ill-known-pos},\eqref{ill-known-nec}.
 
-%% equations for the poss and nec!!
-The simplified equations for the possibility and the necessity measures are~\cite{Pons2011}:
-\begin{eqnarray}
-\label{eq:interval-pos}
-\Pos\left(\lambda([a,b])\right)&=&\\
-\nonumber
-\min\bigg(\sup_{a\leq w}\pi_{X}(w),\sup_{b\geq w}\pi_{Y}(w)\bigg)\\
-\label{eq:interval-nec}
-\Nec\left(\lambda([a,b])\right)&=&\\
-\nonumber
-\min\bigg(\inf_{a>w}1-\pi_{X}(w),\inf_{b<w}1-\pi_{Y}(w)\bigg).
-\end{eqnarray}
+% In order to know whether all the points in the crisp time interval $I=\left[a, b\right]$ are within the boundaries of the ill-known time interval $\left[X, Y\right]$ we define the following two ill-known constraints $C_{1}, C_{2}$
+% \begin{eqnarray}
+% \label{eq:constraint-c1}
+% C_1\stackrel{\triangle}{=}\left(\geq,X\right)  \\
+% \label{eq:constraint-c2}
+% C_2\stackrel{\triangle}{=}\left(\leq,Y\right)
+% \end{eqnarray}
+% Note that the boolean operator applied to the constraints is $\bool=\wedge$. Hence, the construct of both constraints is $C_1 \wedge C_2$
+% This means that we want to know if all the points in the interval $I$ are greater than or equal to $X$ and, at the same time, smaller than or equal to $Y$.
+% %\end{definition}
+% 
+% %% equations for the poss and nec!!
+% The simplified equations for the possibility and the necessity measures are~\cite{Pons2011}:
+% \begin{eqnarray}
+% \label{eq:time-interval-pos}
+% \Pos\left(\lambda([a,b])\right)&=&\\
+% \nonumber
+% \min\bigg(\sup_{a\leq w}\pi_{X}(w),\sup_{b\geq w}\pi_{Y}(w)\bigg)\\
+% \label{eq:time-interval-nec}
+% \Nec\left(\lambda([a,b])\right)&=&\\
+% \nonumber
+% \min\bigg(\inf_{a>w}1-\pi_{X}(w),\inf_{b<w}1-\pi_{Y}(w)\bigg).
+% \end{eqnarray}
 
@@
 \subsubsection{\label{subsubsec:representation-semi-open}Representation of semi-open time intervals}
-As mentioned before, the problem resides in the representation of these kind of interval.
-In order to get a short representation for these intervals, two constants are defined: $FB$ \emph{From the Beginning}: it is only valid for the left value of the interval and $UC$ \emph{Until Changed} (this name is usual in the temporal database community, see ~\cite{Jensen1994}) which is only valid for the right value of the interval.
+As mentioned before, the problem resides in the representation of this kind of interval.
+
+% In order to get a short representation for these intervals, two constants are defined: $FB$ -\emph{From the Beginning}- it is only valid for the left value of the interval and $UC$ -\emph{Until Changed}- (this name is usual in the temporal database community, see ~\cite{Jensen1994}) which is only valid for the right value of the interval.
 Note that these two constants are aliases for a function that obtains both possibility and necessity measures.
 
-Because of the ill-known constraints $C_{1}, C_{2}$, a function called \emph{Open} should be defined in order to deal with a proper representation of these intervals:
+Because of the ill-known constraints $C_{1}, C_{2}$, a function called \emph{Open} should be defined in order to deal with a proper representation of these intervals.
 
@@
 $Open(C)$\\
-The function $\Open(C)$ provides both possibility and necessity measures for all the points in the open part of a semi-open ill-known time interval.
-The parameter $C$ is an ill-known constraint $C\stackrel{\triangle}{=}\left(R,T\right)$ as defined in \ref{subsec:interval-evaluation-by-ill-known-constraints}, in the equation \eqref{eq:ill-known-constraint} and in~\cite{Pons2011}.
+Consider an ill-known value $T$, a binary relationship $B_r$ one of $\left(\leq, <, >, \geq\right)$ and the constraint $C\stackrel{\triangle}{=}\left(R,T\right)$.
+The function $\Open(C) = \left(I_p(C), I_n(C) \right)$ provides both possibility and necessity measures for all the points in the open part of a semi-open ill-known time interval.
+
 The possibility and necessity measures are defined by:
@@
 %\begin{table}[htbp]
-\tcap{Relations for the $\Open(C)$ function. Depending on the relation $R = \left(\leq, <, >, \geq\right)$ in the constraint $C$, the values for $\Rp$ and $\Rn$ are shown.}
+\tcap{Relations for the $\Open(C)$ function. Depending on the relation $B_r = \left(\leq, <, >, \geq\right)$ in the constraint $C$, the values for $\Rp$ and $\Rn$ are shown.}
 \centerline{\small Relations}
@@
 \subsubsection{\label{subsubsec:ill-known-time-interval-datatypes}Datatypes}
-In order to represent properly an ill-known time interval in a database, some datatypes are needed.
+In order to properly represent  an ill-known time interval in a database, some datatypes are needed.
 Because of the ill-known constraints, not all the combinations of datatypes for each ill-known time point (see Table \ref{tbl:time-point-types}) are allowed.
-Table \ref{tbl:time-interval-types} shows all the possible combinations for the datatypes for an ill-known time interval denoted by $\left[X, Y\right]$.
+Table \ref{tbl:time-interval-types} shows all the possible datatypes combinations for the  for an ill-known time interval denoted by $\left[X, Y\right]$.
 
@@
 
-\subsection{\label{subsec:fuzzy-allen-relations}Allen's relations}
-The Allen's relations can be represented by a Boolean combination of ill-known constraints.
+% \subsection{\label{subsec:fuzzy-allen-relations}Allen's relations}
+% The Allen's relations can be represented by a Boolean combination of ill-known constraints.
 The constructs for the constraints are shown in Table \ref{tbl:fuzzy-allen-relations}.
 The possibility and necessity measures are obtained for each Boolean function as explained in Section \ref{subsec:interval-evaluation-by-ill-known-constraints} in equations \eqref{eq:conjunctive1}-\eqref{eq:negation}.
-
-\vglue13pt
-%\begin{table}[htbp]
-\tcap{\label{tbl:fuzzy-allen-relations}Constructs of constraints related to their respective Allen relations, as used in the presented work.
+% 
+% 
+% \vglue13pt
+% %\begin{table}[htbp]
+% \tcap{\label{tbl:fuzzy-allen-relations}Constructs of constraints related to their respective Allen relations, as used in the presented work.
 In this table, the ill-known time interval $J = \left[X, Y\right]$ in a record $r$ has a start point described by possibilistic variable $X$ and an end point described by possibilistic variable $Y$.
 The crisp time interval in the user's temporal demand is denoted $I$.}
-\centerline{\small Relations}
-\vglue-6pt
-\centerline{\small\baselineskip=13pt
-\begin{tabular}{p{0.2\columnwidth} p{0.8\columnwidth} }\\
-\hline
-Allen Relation & Construct of constraints\\
-\hline
-I before J & $C_1 \triangleq \left(<,X\right)$ \\
-\hline
-I equal J & $\left(C_1 \triangleq \left(\geq,X\right)\right) \wedge \neg \left(C_2 \triangleq \left(\neq,X\right) \right) \wedge \left(C_3 \triangleq \left(\leq,Y\right) \right) \wedge \neg \left(C_4 \triangleq \left(\neq,Y\right)\right)$ \\
-\hline
-I meets J & $\left(C_1 \triangleq \left(\leq,X\right)\right) \wedge \neg \left(C_2 \triangleq \left(\neq,X\right) \right)$ \\
-\hline
-I overlaps J & $\left(C_1 \triangleq \left(<,Y\right)\right) \wedge \neg \left(C_2 \triangleq \left(\leq,X\right) \right) \wedge \neg \left(C_3 \triangleq \left(\geq,X\right) \right)$ \\
-\hline
-I during J & $\left( \left (C_1 \triangleq \left(>,X\right) \right) \wedge \left(C_2 \triangleq \left(\leq,Y\right) \right) \right) \vee \left( \left(C_3 \triangleq \left(\geq,X\right) \right) \wedge  \left(C_4 \triangleq \left(<,Y\right) \right) \right)$ \\
-\hline
-I starts J & $\left(C_1 \triangleq \left(\geq,X\right) \right) \wedge \neg \left(C_2\triangleq \left(\neq,X\right)\right)$ \\
-\hline
-I finishes J & $\left(C_1 \triangleq \left(\leq,Y\right) \right) \wedge \neg \left(C_2\triangleq \left(\neq,Y\right)\right)$ \\
-\hline\\
-\end{tabular}
-}
+% \centerline{\small Relations}
+% \vglue-6pt
+% \centerline{\small\baselineskip=13pt
+% \begin{tabular}{p{0.2\columnwidth} p{0.8\columnwidth} }\\
+% \hline
+% Allen Relation & Construct of constraints\\
+% \hline
+% I before J & $C_1 \triangleq \left(<,X\right)$ \\
+% \hline
+% I equal J & $\left(C_1 \triangleq \left(\geq,X\right)\right) \wedge \neg \left(C_2 \triangleq \left(\neq,X\right) \right) \wedge \left(C_3 \triangleq \left(\leq,Y\right) \right) \wedge \neg \left(C_4 \triangleq \left(\neq,Y\right)\right)$ \\
+% \hline
+% I meets J & $\left(C_1 \triangleq \left(\leq,X\right)\right) \wedge \neg \left(C_2 \triangleq \left(\neq,X\right) \right)$ \\
+% \hline
+% I overlaps J & $\left(C_1 \triangleq \left(<,Y\right)\right) \wedge \neg \left(C_2 \triangleq \left(\leq,X\right) \right) \wedge \neg \left(C_3 \triangleq \left(\geq,X\right) \right)$ \\
+% \hline
+% I during J & $\left( \left (C_1 \triangleq \left(>,X\right) \right) \wedge \left(C_2 \triangleq \left(\leq,Y\right) \right) \right) \vee \left( \left(C_3 \triangleq \left(\geq,X\right) \right) \wedge  \left(C_4 \triangleq \left(<,Y\right) \right) \right)$ \\
+% \hline
+% I starts J & $\left(C_1 \triangleq \left(\geq,X\right) \right) \wedge \neg \left(C_2\triangleq \left(\neq,X\right)\right)$ \\
+% \hline
+% I finishes J & $\left(C_1 \triangleq \left(\leq,Y\right) \right) \wedge \neg \left(C_2\triangleq \left(\neq,Y\right)\right)$ \\
+% \hline\\
+% \end{tabular}
+% }
 

### 2012-06-13 trunk/ijcis-pos-model/sources/abstract.tex
My contributions so far...
version sent to Olga and Nicolas

--- 2012-05-23 ijcis-pos-model/sources/abstract.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%
-In reality, some objects or concepts have properties with a time-variant or time-related nature.
-Modelling these kinds of objects or concepts in a (relational) database schema is possible, but time-variant and time-related attributes have an impact on the consistency of the entire database.
-Therefore, temporal database models have been proposed to deal with this.
-Time itself can be at the source of imprecision, vagueness and uncertainty, since existing time measuring devices are inherently imperfect.
-Accordingly, human beings manage time using temporal indications and temporal notions, which may contain imprecision, vagueness and uncertainty.
-However, the imperfection in human-used temporal indications is supported by human interpretation, whereas information systems need support for this.
-Several proposals for dealing with such imperfections when modelling temporal aspects exist.
-Some of these proposals transform the temporal expression into a compact representation.
-Other proposals consider the temporal indications in the used temporal expressions to be the source of imperfection.
-In this work we present a novel model to deal with imperfections in valid-time databases. Next to that, the data manipulation language, \emph{DML} is defined and implemented.
+In real world, it is very common that some objects or concepts have properties with a time-variant or time-related nature.
+Modelling this kind of objects or concepts in a (relational) database schema is possible, but time-variant and time-related attributes have an impact on the consistency of the entire database.
+Therefore, temporal database models have been proposed to deal with this problem in the literature.
+Time can be affected by imprecision, vagueness and / or  uncertainty, since existing time measuring devices are inherently imperfect.
+Additionally, human beings manage time using temporal indications and temporal notions, which may also be imprecise.
+However, the imperfection in human-used temporal indications is supported by human interpretation, whereas information systems need appropriate support in order to accomplish this task.
+Several proposals for dealing with such imperfections when modelling temporal data exist.
+Some of these proposals transform the temporal data into a compact representation.
+There is not a model for managing and handling uncertainty regarding temporal information.
+In this work we present a novel model to deal with imprecision in valid-time databases together with the data manipulation language, \emph{DML} is defined and implemented.

### 2012-06-13 trunk/ijcis-pos-model/sources/preliminaries.tex
version sent to Olga and Nicolas

@@
 Indeed, any Boolean proposition $p$ takes just one value in $\mathbb{B}$.
-If the knowledge about which value this proposition $p$ will take is given by a possibility distribution $\pi_p$, the proposition can be seen as a possibilistic variable.
-As the interest lies with the case where the proposition holds, the possibility and necessity that $p$ = $T$ (the proposition holds) demand most attention.
+If the knowledge about which value this proposition $p$ will take is given by a possibility distribution $\pi_p$; the proposition can be seen as a possibilistic variable.
+The possibility and necessity that $p$ = $T$ (the proposition holds) demand more attention.
 This possibility and necessity is noted here as:
 \begin{align}
+\label{propholdsposs}
 \text{Possibility that $p$ = $T$ (p holds):} &\\
 \nonumber
-Pos(p) = \pi_p(T) \label{propholdsposs} \\
+Pos(p) = \pi_p(T)  \\
+\label{propholdsnecc}
 \text{Necessity that $p$ = $T$ (p holds):} & \\
 \nonumber
-Nec(p) = 1-\pi_p(F) \label{propholdsnecc}
+Nec(p) = 1-\pi_p(F) 
 \end{align}
 
-Here, notation \ref{propholdsposs} denotes the possibility that $p$ = $T$ and the proposition holds, notation \ref{propholdsnecc} denotes the necessity that $p$ = $T$ and the proposition holds.
+Here, equation \eqref{propholdsposs} denotes the possibility that $p$ = $T$ and the proposition holds, while equation \eqref{propholdsnecc} denotes the necessity that $p$ = $T$ and the proposition holds.
 
-This work will deal with ill-known intervals.
-These are ill-known sets, defined and represented via a start and end point, which will be ill-known values.
-The elements of the set are the values between the start and end point.
-A closed ill-known interval with start point defined by possibilistic variable $X$ and end point by possibilistic variable $Y$ is noted here $\left[X, Y\right]$.
-The correspondences and transitions between the representations of ill-known sets, between the representations of ill-known intervals and between the representations of an ill-known set and an ill-known interval are part of the authors current research.
+This work deals with ill-known intervals.
+These are ill-known sets, defined and represented via a starting and ending points, which are ill-known values.
+The elements of the set are the values between the starting and ending points.
+A closed ill-known interval with starting point defined by possibilistic variable $X$ and ending point by possibilistic variable $Y$ is noted here $\left[X, Y\right]$.
+%The correspondences and transitions between the representations of ill-known sets, between the representations of ill-known intervals and between the representations of an ill-known set and an ill-known interval are part of the authors current research.
 
@@
 If this modal value $m$ is unique, then $M$ is referred to as a \emph{fuzzy number}. 
-In other words, if the core of a fuzzy interval is a singleton, it is referred to as a fuzzy number (Figure \ref{fig:triangular-dist}).
+%In other words, if the core of a fuzzy interval is a singleton, it is referred to as a fuzzy number (Figure \ref{fig:triangular-dist}).
 
-A simple form of the membership function of a fuzzy interval is a trapezoidal function (Figure \ref{fig:trapezoidal-dist}.
+A simple form of the membership function of a fuzzy interval is a trapezoidal function (Figure \ref{fig:trapezoidal-dist}).
 It can be shown that such a membership function $\mu_T$ for a fuzzy interval $T$ is convex and normalized.
-Four reel values, denoted $\alpha$, $\beta$, $\gamma$ and $\delta$ suffice to represent a trapezoidal membership function of a fuzzy interval.
-In this work, a fuzzy interval defined as such will be noted as $\left[\alpha, \beta, \gamma, \delta\right]$.
-The corresponding membership function definition for this $\mu_T$ is then given by:
+The values $\left[\alpha, \beta, \gamma, \delta\right]$ represent a trapezoidal possibility distribution defined by $\mu_T$:
+
+% Four real values, denoted $\alpha$, $\beta$, $\gamma$ and $\delta$ suffice to represent a trapezoidal membership function of a fuzzy interval.
+In this work, a fuzzy interval defined as such will be noted as .
+The corresponding membership function definition for this  is then given by:
 
@@
 
-\subsection{Interval Evaluation by Ill-known Constraints}
-The problem of interval evaluation is more generally explained in \cite{Pons2011}: the need exists to know if all points in a crisp interval $I$ reside between the boundaries of an ill-known interval $\left[ X , Y \right]$.
-In \cite{Pons2011}, the notion of an \emph{ill-known constraint} is introduced:
+\subsection{\label{subsec:interval-evaluation-by-ill-known-constraints}Interval Evaluation by Ill-known Constraints}
+In our case, we need to know if all points in a crisp interval $I$ reside between the boundaries of an ill-known interval $\left[ X , Y \right]$.
+To compute that, we introduce the concept of \emph{ill-known constraint}\cite{Pons2011}.
 
@@
 \begin{align}
+\label{eq:ill-known-constraint}
 C \triangleq (R,V)
 \end{align}
-Set $A$ now satisfies the constraint if and only if:
+Set $A$ satisfies the constraint if and only if:
 \begin{align}
@@
 \begin{align}
+\label{ill-known-pos}\
 \Pos(A\text{ satisfies }C) & =\\
 \nonumber
-\min_{a \in A}\left(\sup_{(a,w) \in R}\pi_{V}(w)\right) \label{ill-known-pos}\\
+\min_{a \in A}\left(\sup_{(a,w) \in R}\pi_{V}(w)\right) \\
+\label{ill-known-nec}
 \Nec(A\text{ satisfies }C) & =\\
 \nonumber
-\min_{a \in A}\left(\inf_{(a,w) \notin R} 1-\pi_{V}(w)\right) \label{ill-known-nec}
+\min_{a \in A}\left(\inf_{(a,w) \notin R} 1-\pi_{V}(w)\right) 
 \end{align}
 
-So far, we have shown how it can be verified whether or not a set satisfies an ill-known constraint.
-However, from the examples at the beginning of this section, it is observed that Boolean combinations of constraints are required.
-For example, the problem of interval evaluation as explained earlier requires that all elements of an interval $[a,b]$ are larger than a value $X$ and at the same time smaller than a value $Y$, which implies that a conjunctive Boolean combination of both constraints must be satisfied.
+So far, we have shown how it can be verified whether a set satisfies or not an ill-known constraint. The interval evaluation problem is explained in a more general context in \cite{Pons2011}.
+
+\begin{example}
+As example, consider  $X = \left[5, 3, 2 \right]$ and let $C =(\leq, X)$ the ill-known constraint.
+Then, the evaluation of the possibility and the necessity are obtained from \eqref{ill-known-pos} and \eqref{ill-known-nec} respectively.
+(See Figure \ref{fig:example-ill-known}).
+
+\vspace*{13pt}
+\begin{center}
+{
+\includegraphics[scale=1]{./graphs/example-ill-known.pdf}
+}
+\end{center}
+%\centerline{ \psfig{file=./graphs/Y-time-point.eps}}
+\vspace*{10pt}
+\fcaption{\label{fig:example-ill-known}Example of the evaluation of the ill-known constraint $C \triangleq (\leq, X)$. Possibility and necessity measures are shown in grey. }
+%  \label{fig:fuzzy-validity-period}
+\vspace*{13pt}
+
+\begin{align}
+\Pos(A\text{ satisfies }C) & =\\
+\nonumber
+\min_{a \in A}\left(\sup_{a \leq R}\pi_{X}(w)\right)\\
+\Nec(A\text{ satisfies }C) & =\\
+\nonumber
+\min_{a \in A}\left(\inf_{a > w} 1-\pi_{X}(w)\right) 
+\end{align}
+
+\end{example}
+
+It is observed that Boolean combinations of constraints are required.
+For example, the problem of interval evaluation (explained earlier) requires that all elements of an interval $[a,b]$ are larger than a value $X$ and, at the same time, smaller than a value $Y$, which implies that a conjunctive Boolean combination of both constraints must be satisfied.
 To allow Boolean combinations of constraints, the following definitions are introduced.
@@
 
-There is a special boolean combination of constraints that is of particular interest:
+There is a special boolean combination of constraints that is of particular interest; let us see it.
 \begin{definition}
-\emph{Convex combination of ill-known constraints}. Consider the $C_1$ and $C_2$ two ill-known constraints. A convex combination of the constraints is given by:
+\emph{Conjunctive combination of ill-known constraints}.
+Consider a universe $U$.
+Let $R_1, R_2 \subseteq U$ be two binary relations in U.
+Let $X_1, X_2$ be two fixed ill-known values in $U$.
+Let  $C_1 = (R_1, X_1)$ and $C_2 = (R_2, X_2)$ be two ill-known constraints.
+A conjunctive combination of both constraints is given by:
 \begin{align}
@@
 \begin{definition}
-Consider a universe $U$, $X$ an ill-known value and $CC \triangleq \left \lbrace C_{1_{Z_1}}, \ldots, C_{n_{Z_n}}  \right \rbrace$ a convex combination of ill-known constraints over the variables $Z_1, \ldots, Z_n$.
-The value $X$ is defined by means of the convex combination:
+\label{def:convex-combination-ill-known-constraints}
+\emph{Ill-known value defined by conjunctive combination of constraints.}
+Consider a universe $U$, $X$ an ill-known value and $CC \triangleq \left \lbrace C_{1_{Z_1}} \wedge \ldots \wedge C_{n_{Z_n}}  \right \rbrace$ a convex combination of ill-known constraints over the variables $Z_1, \ldots, Z_n$.
+Then, the value $X$ is defined by means of the convex combination:
 \begin{align}
@@
 \end{align} 
-Note that $\pi_{X}$ is convex since $\pi_{CC}$ is convex as demostrated in theorem \ref{th:convex-combination-ill-known-constraints}.
+Note that $\pi_{X}$ is convex since $\pi_{CC}$ is convex as demonstrated in Theorem \ref{th:convex-combination-ill-known-constraints}.
 \end{definition}
@@
 \begin{example}
-Consider the ill-known values $X = \left[12, 2, 2\right]$ and $Y = \left[18, 2, 1 \right]$. The ill-known value $Z$ is defined using the convex combination $CC$ of constraints $C_1$ and $C_2$:
+As an example, consider the hospital database.
+When a patient enters o leaves a medical service in hospital, a record tracking the patient is stored.
+The exact time for this is not precisely known.
+Consider now that $X = \left[12, 2, 2\right]$ is the time in hours when a patient left the emergency service and $Y = \left[18, 2, 1 \right]$ is the time in hours when the patient left the hospital.
+It is possible to obtain $Z$ (the period of time between the emergency service and the left of the patient) by using a conjunctive combination of constraints.
+
 \begin{align}
 \nonumber
-CC \triangleq \left \lbrace C_1\left(>,X\right) \wedge C_2(\leq,Y) \right \rbrace \\
+CC = \left \lbrace C_1\left(>,X\right) \wedge C_2(\leq,Y) \right \rbrace \\
 \nonumber
-Z \triangleq CC
+Z = CC
 \end{align}
-$Z$ is a fuzzy interval defined by a trapezoidal shape given by $\left[12,\ 14,\ 18,\ 19 \right]$
+$Z$ is a fuzzy interval defined by a trapezoidal shape given by $\left[12,\ 14,\ 18,\ 19 \right]$.
 Figure \ref{fig:example-ill-known-by-const} illustrates the relations among the variables $X$, $Y$ and $Z$.

### 2012-06-13 trunk/ijcis-pos-model/sources/temporalModel.tex
version sent to Olga and Nicolas

@@
 
+In this section we will formalize the model for valid-time relational databases. The first subsection is devoted to the formalization of the model. Then, the data manipulation language is defined.
+
 \subsection{\label{subsec:temporal-model}The generalized temporal model}
-The model is based on the GEFRED~\cite{Medina1994} (Generalized Model of Fuzzy Relational DB) model.
-This model is extended by adding valid-time support which will be illustrated with an example.
+The model is based on GEFRED~\cite{Medina1994} (Generalized Model of Fuzzy Relational DB) model.
+This model is extended by adding valid-time support which will be illustrated with examples.
 The information in the system is defined by the following elements:
@@
 \label{def:generalized-fuzzy-domain}
-\textbf{Generalized fuzzy domain.}
-$D$ is the discourse domain, $\tilde \Pow\left(D \right)$ is the set of all possibility distributions defined on $D$, plus the NULL constant. The generalized fuzzy domain $D_G$ is defined as:
+\emph{Generalized fuzzy domain.}
+Let $D$ be the discourse domain, $\tilde \Pow\left(D \right)$ is the set of all possibility distributions defined on $D$, plus the NULL constant. The generalized fuzzy domain $D_G$ is defined as:
 \begin{equation}
@@
 \end{definition}
-The datatypes for $D_G$ are shown in table \ref{tbl:gefred-data-types}.
+The datatypes that can be used to represent $D_G$ are shown in table \ref{tbl:gefred-data-types}.
 
@@
 \label{def:typeof-domain}
-\textbf{Typeof(D).}
-The function typeof$\left(D \right)$ returns the datatype associated with the domain $D$. For datatype, the function returns the associated number as shown in Table \ref{tbl:gefred-data-types}.
+\textbf{Typeof(A).}
+Consider $D_G$ to be a generalized fuzzy domain and let $A \in D_G$ be a value for the domain. 
+The function typeof$\left(A \right)$ returns the datatype associated with the domain $D_G$.
+For a given value $A$, the function returns the associated number as shown in Table \ref{tbl:gefred-data-types}.
 \end{definition}
@@
 \label{def:generalized-fuzzy-temporal-domain}
-\textbf{Generalized fuzzy temporal domain.}
-If $\T$ is the temporal domain, $\tilde \Pow\left( \T\right)$ is the set of all \emph{normalized} possibility distributions (see Section \ref{subsec:possibility-theory}, equation \eqref{NormalizationProperty}) defined on $\T$.
+\emph{Generalized fuzzy temporal domain.}
+Consider $\T$ to be the temporal domain, and let $\tilde \Pow\left( \T\right)$ be the set of all \emph{normalized} possibility distributions (see Section \ref{subsec:possibility-theory}, equation \eqref{NormalizationProperty}) defined on $\T$.
 The Generalized Fuzzy Temporal Domain, $\T_G$ is
@@
 \begin{definition}
-\textbf{Generalized fuzzy temporal relation.}
-A generalized fuzzy temporal relation $R_{FTG}$ is given by:
+\emph{Generalized fuzzy temporal relation.}
+A generalized fuzzy temporal relation $R_{FTG}$ is given by~\cite{Medina1994}:
 \label{def:generalized-fuzzy-temporal-relation}
@@
 \label{ex:sample-ft-database}
- Consider a library database as explained in example \ref{ex:library-database}.
-But now, the time is stored with imprecision.
+ Consider a hospital database as explained in example \ref{ex:library-database}.
+But now, the time is stored with imprecision, because the date of arrival and leave are not precisely known.
 For simplicity, the ill-known time points are represented as triangular fuzzy numbers.
@@
 \nonumber
-&  \left(\text{Title}:D_{\text{title}}\right),\\
-\nonumber
-& \left(\text{Loan}:D_{\text{Loan}}\right) \\
+&  \left(\text{Name}:D_{\text{Name}}\right),\\
+% \nonumber
+% & \left(\text{Loan}:D_{\text{Loan}}\right) \\
 \nonumber
@@
 %\begin{table}[htbp]
-\tcap{\label{tbl:fuzzy-library-sample}Sample library database}
+\tcap{\label{tbl:fuzzy-library-sample}Sample hospital database}
 %\centerline{\small DATA TYPES}
@@
 \centerline{\small\baselineskip=13pt
-\begin{tabular}{c c c |c c}\\
+\begin{tabular}{c c c |p{0.35\columnwidth}}\\
 \hline
-& & & \multicolumn{2}{c}{PVP} \\
-$\Head$ &  \textbf{ID}  & Title  &  Start Loan & End Loan\\
+& & & \centerline{PVP} \\
+$\Head$ &  \textbf{ID}  & Name  &  Arrival , Leave\\
 \hline
-\multirow{3}{*}{$\Body$ \Big\lbrace} &  3 & Dracula & [15/3/2012, 5, 2] & [30/3/2012, 1, 1] \\
- &4 & Frankenstein & [25/3/2012, 3, 2] & [4/4/2012, 1, 7] \\
-&5 & Harry Potter & [18/3/2012, 4, 1] & [2/4/2012, 2, 2]\\
+\multirow{3}{*}{$\Body$ } &  3 & John & [15/3/2012, 5, 2] , [30/3/2012, 1, 1] \\
+ &4 & Antoon & [25/3/2012, 3, 2] , [4/4/2012, 1, 7] \\
+&5 & Daan & [18/3/2012, 4, 1] , [2/4/2012, 2, 2]\\
 % \ 3 & Dracula & [4/4/2012, 3, 3] & UC \\
@@
 \label{def:value-component}
-\textbf{Value component.}
+\emph{Value component.}
 The value component $R^{v}_{FTG}$ of a fuzzy temporal relation $R_{FTG}$ is a set with the value components for both the head and the body of the relation:
@@
 \nonumber
-& \left(\text{Loan}:D_{\text{Loan}}\right), \left(\text{PVP}:D_{\text{PVP}}\right) \rbrace \\
+& \left(\text{PVP}:D_{\text{PVP}}\right) \rbrace \\
 \nonumber
-\Body^ {v} &= \left \lbrace 3:, \text{``Dracula``}: ,  \left[15/3/2012, 30/3/2012\right] \right \rbrace 
+\Body^ {v} &= \left \lbrace 3:, \text{``John``}: ,  \left[15/3/2012, 30/3/2012\right] \right \rbrace 
 \end{align}
@@
 \label{def:compatibility-component}
-\textbf{Compatibility component.}
+\emph{Compatibility component.}
 The compatibility component $R^{c}_{FTG}$ of a fuzzy temporal relation $R_{FTG}$ is a set with the compatibility components for both the head and the body of the relation:
@@
 \label{def:temporal-component}
-\textbf{Temporal component.}
+\emph{Temporal component.}
 The temporal component $R^{t}_{FTG}$ of a fuzzy temporal relation $R_{FTG}$ is a set with the temporal components for both the head and the body of the relation:
@@
 \label{def:generalized-primary-key}
-\textbf{Generalized primary key.}
+\emph{Generalized primary key.}
+Consider $D_G$ to be a fuzzy generalized domain, and let $A_{Gs}:D_{Gs}$ be the attributes and the domain of the attribute for each $s in S \subseteq \left(1, \ldots, n \right)$.
 A generalized primary key, $K_G$ is a subset of the head:
@@
 \label{def:generalized-fuzzy-temporal-key}
-\textbf{Generalized fuzzy temporal key.}
+\emph{Generalized fuzzy temporal key.}
+Consider $D_G$ to be a fuzzy generalized domain, and let $A_{Gs}:D_{Gs}$ be the attributes and the domain of the attribute for each $s in S \subseteq \left(1, \ldots, n \right)$.
 A generalized fuzzy temporal key, $K_{GT}$ is a subset of the head. An attribute, namely \emph{version} $V$ is added when dealing with valid-time.
@@
 \centerline{\small\baselineskip=13pt
-\begin{tabular}{c c c c | c c}\\
+\begin{tabular}{c c c c | p{0.35\columnwidth} }\\
 \hline
-& & & &\multicolumn{2}{c}{PVP} \\
-$\Head$  & \textbf{ID}  & \textbf{V} & Title  & Start Loan & End Loan \\
+% & & & \multicolumn{2}{c}{PVP} \\
+%& & & & \centerline{PVP}\\
+{$\Head$}  &  &  & &(Arrival\\
+& \textbf{ID}  & \textbf{V} & Name& ,Leave )	\\
  \hline
-\multirow{4}{*}{$\Body  \Big\lbrace$}  & 3 & 001 & Dracula & [15/3/2012, 5, 2] & [30/3/2012, 1, 1] \\
-& 4 & 001 & Frankenstein & [25/3/2012, 3, 2] & [4/4/2012, 1, 7]\\
-& 5 & 001 & Harry Potter & [18/3/2012, 4, 1] & [2/4/2012, 2, 2]\\
-& 3 & 002 & Dracula & [4/4/2012, 3, 3] & UC \\
+\multirow{4}{*}{$\Body$}  & 3 & 001 & John & [15/3/2012, 5, 2], [30/3/2012, 1, 1] \\
+& 4 & 001 & Antoon & [25/3/2012, 3, 2] , [4/4/2012, 1, 7]\\
+& 5 & 001 & Daan & [18/3/2012, 4, 1] , [2/4/2012, 2, 2]\\
+ & 3 & 002 & John & [4/4/2012, 3, 3] , UC \\
 \hline\\
@@
 
-The algebra for the data manipulation is defined in \cite{Medina1994}.
-The Generalized Fuzzy Relational Algebra manipulates relations like $R_{FG}$.
+The Generalized Fuzzy Relational Algebra~\cite{Medina1994} manipulates relations like $R_{FG}$.
 The operations defined are: \emph{Union, Intersection, Difference, Cartesian Product, Projection, Join} and \emph{Selection}.
 Thus, in this section we will describe and implement the following operations for temporal databases, as described in \ref{subsubsec:Understanding-valid-time-databases}.
-The operations implemented are: \emph{Insert, Modify, Delete, } and \emph{Revise}.
+The operations implemented are: \emph{Insert, Modify, Delete, }.
 The semantics of the operations will be the same that those defined for a crisp temporal database, whereas the temporal representation is made by the possibilistic valid-time period and the ill-known constraints (see sections \ref{sec:prelim} and \ref{sec:time-rep}.
+%The algebra for the data manipulation is defined in .
 
-It is important to notice that whereas the result of the evaluation of any comparison between crisp time intervals is boolean, the evaluation of any comparison between \emph{PVPs} is in the unit interval.
+It is important to notice that while the result of the evaluation of any comparison between crisp time intervals is boolean, the evaluation of any comparison between \emph{PVPs} is in the unit interval.
 Therefore, for a valid-time object, say $A \in R$, in a crisp temporal database it is not possible an overlapping among any valid-time interval ($\forall I, \not \exists J: \left(A, I \right) \in R \wedge \left(A, J \right) \in R \wedge \lnot \Overlap\left(I, J \right)$).
@@
 This is a natural extension of the crisp temporal databases.
-This approach allows a easy implementation for the DML operations defined in Section \ref{subsubsec:Understanding-valid-time-databases}.
+This approach allows an implementation for the DML operations defined in Section \ref{subsubsec:Understanding-valid-time-databases}.
 
@@
 \nonumber
-\mbox{ if } I = \mbox{ Current } \left(R, A \right) \\ %\wedge J > \mbox{ CloseR } \left(I, J \right)   \\
+\mbox{ if Current } \left(R, A \right) \neq \emptyset \\ %\wedge J > \mbox{ CloseR } \left(I, J \right)   \\
 \nonumber R , \text{ in any other case}
@@
 \begin{verbatim}
-Insert(3,'Dracula',
+Insert(3,'John',
 [[15/3/2012, 5, 2],
         [30/3/2012, 1, 1]]);
-Insert(4,'Frankenstein',
+Insert(4,'Antoon',
 [[25/3/2012, 3, 2], 
         [4/4/2012, 1, 7]]);
-Insert(5,'Harry Potter',
+Insert(5,'Daan',
 [[18/3/2012, 4, 1], 
@@
 Note that here, the result of the Overlaps operator is in the unit interval.
-For example, the librarian loans the book with ID=3 but the book is not returned yet. The insert sentence is the following.
+For example, the patient with ID=3 has not left the hospital yet. The insert sentence is the following.
+
  \begin{verbatim}
-Insert(3,'Dracula',
+Insert(3,'John',
 [[4/4/2012, 3, 3], UC]);
@@
 	\item Modify and close the current version of the entity $A$ and insert $\left(A, I \right)$.
-For example, consider now that the book with ID 3 is loaned  around the 24/4/2012 (this is model by a triangular fuzzy number: $\left[24/4/2012, 1, 1 \right]$).
-Here the problem is that the book with ID 3 was loaned around 4/4/2012, but not returned.
-If the book is already in the library for a new loan, then it is necessary to set the return date and add a new row with the new loan date.
+For example, consider now that the hospital with ID=3 arrived  around the 24/4/2012 (this is model by a triangular fuzzy number: $\left[24/4/2012, 1, 1 \right]$).
+Here the problem is that the patient with ID=3 arrived around 4/4/2012, but it was not stored in the database the leave date.
+If the patient is again in the hospital, then it is necessary to set the leave date and add a new row with the new arrival date.
 %This is illustrated in Table \ref{tbl:library-sample-2}, the insert sentence is:
 	    \begin{verbatim}
-Insert(3,'Dracula',
+Insert(3,'John',
 [[24/4/2012, 1, 1], UC]]);
@@
 	\item Reject the insertion, if the time interval $I$ do overlap with a degree of 1 any existing valid-time interval for the entity $A$ in the relation $R$.
-For example, consider now that the librarian wants to introduce a past loan for the book with ID 3.
-The book was loaned around 6/4/12/2012 and returned around 25/4/2012.
-As this interval does overlaps other time intervals with a degree of 1, it is not possible that the book was loaned in that time interval.Therefore, the insertion is rejected.
+For example, consider now that the hospital manager wants to introduce a past arrival for the patient with ID=3.
+The patient arrived around 6/4/12/2012 and left the hospital around 25/4/2012.
+As this interval does overlaps other time intervals with a degree of 1, it is not possible that the patient was it the hospital at that time interval.
+Therefore, the insertion is rejected.
 The insert sentence is:
@@
 	      \begin{verbatim}
-Insert(3,'Dracula',
+Insert(3,'John',
 [[6/4/12/2012, 1, 1],
@@
 \nonumber
-R \cup \left(A, I \right), & \mbox{ if }  \left(A, I \right) \not \in R \vee \\
-&  \left(A, J \right) \in R \wedge I \text{Overlaps} J < 1\\
+R \cup \left(A, I \right), & \mbox{ if }  \left(A, I \right) \not \in R \mbox{ or }\\
+&  \left(A, J \right) \in R \mbox{ and } I \text{ Overlaps } J < 1\\
 %\big \lbrace \left( \exists J= \left(X', Y' \right): \left(A, J \right) \in R \wedge \right. \\
@@
 \nonumber
-R - \left(A, I \right), &  \forall \left( A, I \right) \in R\\
+R - \left \lbrace \left(A, I_1 \right), \ldots, \left(A, I_n \right) \right \rbrace, &  \forall I_i: \left( A, I_i \right) \in R\\
 R, & \mbox{ otherwise }  
@@
 
-For example, consider that the librarian wants to delete the history for the book with ID = 3. The following sentence deletes all the versions for the book with ID = 3.
+For example, consider that the hospital manager wants to delete the history for the patient with ID = 3. The following sentence deletes all the rows for the patient with ID = 3.
 
 \begin{verbatim}
- Delete(3,'Dracula');
+ Delete(3,'John');
 \end{verbatim}
 
-\subsubsection{\label{subsubsec:revise-fuzzy-temporal}Revise}
-This operation replaces the old values for the entity $A'= \left(a_1', \ldots, a_n' \right) $ with the new values specified by $A = \left(a_1, \ldots, a_n \right)$ in the relation $R$, at a given time interval $I$.
+% \subsubsection{\label{subsubsec:revise-fuzzy-temporal}Revise}
+% This operation replaces the old values for the entity $A'= \left(a_1', \ldots, a_n' \right) $ with the new values specified by $A = \left(a_1, \ldots, a_n \right)$ in the relation $R$, at a given time interval $I$.
 This operation is used when a correction on the values for the entity $A$ should be done.
-\begin{align}
-\label{eq:revise-fuzzy-temporal}
-\REV\left(R, \left(A, I \right)\right)=\\
-\begin{cases}
-\nonumber
- R - \left(A', I \right) \cup \left \lbrace\left(A, I \right)\right \rbrace  & \mbox{ if } \left(A', I \right) \in R \\%: A = \left(a_1, \ldots, a_n \right),\\
-%A'= \left(a_1', \ldots, a_n' \right)   \\   
-R, & \mbox{ otherwise }
-\end{cases}
-\end{align}
-
-For example, consider now that the librarian wants to make a correction in the title for the book with ID = 5. The new title is `Harry Potter IV'. The revise sentence is:
-
-\begin{verbatim}
-Revise(5,'Harry Potter IV', 
-[[18/3/2012, 4, 1], 
-          [2/4/2012, 2, 2]]);
-\end{verbatim}
+% \begin{align}
+% \label{eq:revise-fuzzy-temporal}
+% \REV\left(R, \left(A, I \right)\right)=\\
+% \begin{cases}
+% \nonumber
+%  R - \left(A', I \right) \cup \left \lbrace\left(A, I \right)\right \rbrace  & \mbox{ if } \left(A', I \right) \in R \\%: A = \left(a_1, \ldots, a_n \right),\\
+% %A'= \left(a_1', \ldots, a_n' \right)   \\   
+% R, & \mbox{ otherwise }
+% \end{cases}
+% \end{align} 
+% 
+% 
+% For example, consider now that the librarian wants to make a correction in the title for the book with ID = 5. The new title is `Harry Potter IV'. The revise sentence is:
+% 
+% \begin{verbatim}
+% Revise(5,'Harry Potter IV', 
+% [[18/3/2012, 4, 1], 
+%           [2/4/2012, 2, 2]]);
+% \end{verbatim}

### 2012-06-13 ijcis-pos-model/sources/conclusions.tex
version sent to olga and nicolas.
almost final version

--- 2012-04-15 trunk/ijcis-pos-model/sources/conclusions.tex
@@
-We presented a valid-time model to represent and handle ill-known temporal intervals.
-The semantics and the implementation of the DML operations are described within this work.
-As future work , the definition of the Data Definiton Language as well as the querying will be considered.
+In this work, we presented the first formalization in the literature of a possibilistic valid-time model for a relational databases.
+First we introduced the representation of the ill-known time intervals.
+Then, by using the ill-known constraints we have shown how to manage the relationships among time intervals.
+Then, the valid-time model for relational databases is implemented.
+Finally, the semantics and the implementation of the DML operations are described within this work.
+As future work , the definition of the Data Definition Language as well as the querying will be considered.

### 2012-06-20 ijcis-pos-model/sources/temporalModel.tex
final draft

@@
 
+In this section we will formalize the model for possibilistic valid-time relational databases.
+The first subsection is devoted to the formalization of the model.
+Then, the data manipulation language is defined.
+
 \subsection{\label{subsec:temporal-model}The generalized temporal model}
 The model is based on GEFRED~\cite{Medina1994} (Generalized Model of Fuzzy Relational DB) model.
-This model is extended by adding valid-time support which will be illustrated with an example.
+This model is extended by adding valid-time support which will be illustrated through definitions and examples.
 The information in the system is defined by the following elements:
@@
 \label{def:typeof-domain}
-\textbf{Typeof(D).}
-The function typeof$\left(D \right)$ returns the datatype associated with the domain $D$. For every datatype, the function returns the associated number as shown in Table \ref{tbl:gefred-data-types}.
+\textbf{Typeof(A).}
+Consider $D_G$ to be a generalized fuzzy domain and let $A \in D_G$ be a value for the domain. 
+The function typeof$\left(A \right)$ returns the datatype associated with the domain $D_G$.
+For a given value $A$, the function returns a number in $\left[1, 10 \right]$ as shown in Table \ref{tbl:gefred-data-types}.
 \end{definition}
 
+\begin{samepage}
 \vglue13pt
@@
 } 
+\end{samepage}
 
@@
 
+A generalized fuzzy relation is defined in \cite{Medina1994}. Here, we will extend the definition to a generalized fuzzy temporal relation.
+
 \begin{definition}
-\textbf{Generalized fuzzy temporal relation.}
-A generalized fuzzy temporal relation $R_{FTG}$ is given by:
+\emph{Generalized fuzzy temporal relation.}
+Consider $\A$ to be the set of all the entities, and let $A=\left(A_1, \ldots, A_n \right), A \in \A$ be the set of (fuzzy) attributes that define an entity.
+Let $\I_{PVP}$ be the set of all the ill-known time intervals and let $I = \left(X, Y \right)$ be a PVP, $I \in \I_{PVP}$.
+The pair $\left(A, I\right)$ expresses that the data regarding the entity A are valid during the ill-known time interval I.
+Let $R_{FTG} \subseteq \A \  x\  \I_{PVP}$ be a valid-time relation.
+Then, the following equation indicates that the pair $\left(A, I\right)$ is in the relation R:
+
+\begin{equation}
+\label{eq:fuzzy-rel-def}
+\left( R_{FTG}, \left(A , I\right) \right)
+\end{equation}
+
+A generalized fuzzy temporal relation $R_{FTG}$ can be noted also by:
 \label{def:generalized-fuzzy-temporal-relation}
@@
 \label{ex:sample-ft-database}
- Consider a library database as explained in example \ref{ex:library-database}.
-But now, the time is stored with imprecision.
+ Consider a hospital database as explained in example \ref{ex:library-database}.
+But now, the time is stored with imprecision, because the date of arrival and departure are not precisely known.
 For simplicity, the ill-known time points are represented as triangular fuzzy numbers.
@@
 
+\begin{samepage}
 \vglue13pt
 %\begin{table}[htbp]
-\tcap{\label{tbl:fuzzy-library-sample}Sample library database}
+\tcap{\label{tbl:fuzzy-library-sample}Sample hospital database}
 %\centerline{\small DATA TYPES}
@@
 & & & \centerline{PVP} \\
-$\Head$ &  \textbf{ID}  & Title  &  Start Loan , End Loan\\
+$\Head$ &  \textbf{ID}  & Name  &  Arrival , Leave\\
 \hline
-\multirow{3}{*}{$\Body$ } &  3 & Dracula & [15/3/2012, 5, 2] , [30/3/2012, 1, 1] \\
- &4 & Frankenstein & [25/3/2012, 3, 2] , [4/4/2012, 1, 7] \\
-&5 & Harry Potter & [18/3/2012, 4, 1] , [2/4/2012, 2, 2]\\
+\multirow{3}{*}{$\Body$ } &  3 & John & [15/3/2012, 5, 2] , [30/3/2012, 1, 1] \\
+ &4 & Antoon & [25/3/2012, 3, 2] , [4/4/2012, 1, 7] \\
+&5 & Daan & [18/3/2012, 4, 1] , [2/4/2012, 2, 2]\\
 % \ 3 & Dracula & [4/4/2012, 3, 3] & UC \\
@@
 }
+\end{samepage}
 
@@
 
-Analogously, it is possible to define the the value component for the temporal part and the compatibility component for the temporal part.
+Analogously, it is possible to define both the name value component and the compatibility component for the temporal part.
+
+%the value component for the temporal part and the compatibility component for the temporal part.
 
@@
 \label{def:generalized-primary-key}
-\textbf{Generalized primary key.}
+\emph{Generalized primary key.}
+Consider $D_G$ to be a fuzzy generalized domain, and let $A_{Gs}:D_{Gs}$ be the attributes and the domain of the attribute for each $s \in S \subseteq \left(1, \ldots, n \right)$.
 A generalized primary key, $K_G$ is a subset of the head:
@@
 s\in S \subseteq \left(1, \ldots, n \right) \\
-\nonumber
+\end{align}
+Subject to the following constraints:
+\begin{align}
+%\nonumber
 \forall s \in S, \text{Typeof } \left(D_{Gs} \right) \in \left \lbrace 1, 2 \right \rbrace \\
@@
 The primary key for the table is the attribute \emph{ID}, a unique number.
-Two different books have two different values for the \emph{ID} attribute.
+Two different patients have two different values for the \emph{ID} attribute.
 
 In order to add valid-time support, the primary key should be re-defined.
-E.g., consider the library database.
-If the primary key is the \emph{ID} attribute, a book could not be loaned more than once.
+E.g., consider the hospital database.
+If the primary key is the \emph{ID} attribute, a patient could not arrive to the hospital more than once.
 To resolve this problem, we extend the given primary key with a version identifier.
+
+%a book could not be loaned more than once. 
 %Thus, when a book is loaned, a new entry in the table is added. The \emph{ID} attribute identifies the physical book whereas
@@
 \label{def:generalized-fuzzy-temporal-key}
-\textbf{Generalized fuzzy temporal key.}
-A generalized fuzzy temporal key, $K_{GT}$ is a subset of the head. An attribute, namely \emph{version} $V$ is added when dealing with valid-time.
+\emph{Generalized fuzzy temporal key.}
+Consider $D_G$ to be a fuzzy generalized domain, and let $A_{Gs}:D_{Gs}$ be the attributes and the domain of the attribute for each $s \in S \subseteq \left(1, \ldots, n \right)$.
+Let  $V$ be a new attribute called \emph{version}.
+A generalized fuzzy temporal key, $K_{GT}$ is a subset of the head.
 \begin{align}
@@
 s\in S \subseteq \left(1, \ldots, n \right) \\
-\nonumber
+\end{align}
+Subject to the following constraints:
+\begin{align}
 \forall s \in S, \text{Typeof } \left(D_{Gs} \right) \in \left \lbrace 1, 2 \right \rbrace \\
@@
  \nonumber
-K_{GT} \subseteq \Head, K_{GT} = \left \lbrace  \left(\text{ID}:D_{\text{ID}} \right) \cup  \left(V_{\text{ID}}:D_{\text{ID}} \right) \right \rbrace	
+K_{GT} \subseteq \Head, K_{GT} = \left \lbrace  \left(\text{ID}:D_{\text{ID}} \right) , \left(V_{\text{ID}}:D_{\text{ID}} \right) \right \rbrace	
 %\nonumber
@@
 
+\begin{samepage}
 \vglue13pt
 %\begin{table}[htbp]
-\tcap{\label{tbl:fuzzy-library-sample-valid-time}Sample library database}
+\tcap{\label{tbl:fuzzy-library-sample-valid-time}Sample hospital database}
 %\centerline{\small DATA TYPES}
@@
 %& & & & \centerline{PVP}\\
-{$\Head$}  &  &  & &(Start Loan\\
-& \textbf{ID}  & \textbf{V} & Title& , End Loan)	\\
+{$\Head$}  &  &  & &(Arrival\\
+& \textbf{ID}  & \textbf{V} & Name& ,Leave )	\\
  \hline
-\multirow{4}{*}{$\Body$}  & 3 & 001 & Dracula & [15/3/2012, 5, 2], [30/3/2012, 1, 1] \\
-& 4 & 001 & Frankenstein & [25/3/2012, 3, 2] , [4/4/2012, 1, 7]\\
-& 5 & 001 & Harry Potter & [18/3/2012, 4, 1] , [2/4/2012, 2, 2]\\
- & 3 & 002 & Dracula & [4/4/2012, 3, 3] , UC \\
+\multirow{4}{*}{$\Body$}  & 3 & 001 & John & [15/3/2012, 5, 2], [30/3/2012, 1, 1] \\
+& 4 & 001 & Antoon & [25/3/2012, 3, 2] , [4/4/2012, 1, 7]\\
+& 5 & 001 & Daan & [18/3/2012, 4, 1] , [2/4/2012, 2, 2]\\
+ & 3 & 002 & John & [4/4/2012, 3, 3] , UC \\
 \hline\\
@@
 
-The algebra for the data manipulation is defined in \cite{Medina1994}.
-The Generalized Fuzzy Relational Algebra manipulates relations like $R_{FG}$.
+The Generalized Fuzzy Relational Algebra~\cite{Medina1994} manipulates relations like $R_{FTG}$.
 The operations defined are: \emph{Union, Intersection, Difference, Cartesian Product, Projection, Join} and \emph{Selection}.
 Thus, in this section we will describe and implement the following operations for temporal databases, as described in \ref{subsubsec:Understanding-valid-time-databases}.
-The operations implemented are: \emph{Insert, Modify, Delete, } and \emph{Revise}.
+The operations implemented are: \emph{Insert, Modify, Delete, }.
 The semantics of the operations will be the same that those defined for a crisp temporal database, whereas the temporal representation is made by the possibilistic valid-time period and the ill-known constraints (see sections \ref{sec:prelim} and \ref{sec:time-rep}.
+%The algebra for the data manipulation is defined in .
 
-It is important to notice that whereas the result of the evaluation of any comparison between crisp time intervals is boolean, the evaluation of any comparison between \emph{PVPs} is in the unit interval.
+It is important to notice that while the result of the evaluation of any comparison between crisp time intervals is boolean, the evaluation of any comparison between \emph{PVPs} is a value in the unit interval.
 Therefore, for a valid-time object, say $A \in R$, in a crisp temporal database it is not possible an overlapping among any valid-time interval ($\forall I, \not \exists J: \left(A, I \right) \in R \wedge \left(A, J \right) \in R \wedge \lnot \Overlap\left(I, J \right)$).
-Thus, from the point of view of a fuzzy temporal database, there are the following options (see Figure \ref{fig:tdb-class}):
+Thus, from the point of view of a fuzzy temporal database, there are the following choices (see Figure \ref{fig:tdb-class}):
 
@@
 \begin{definition}
-\textbf{Strictly consistent} It is not possible for two PVPs of the same object to be valid at the same time.
-In other words, if $A$ is an object, $R$ is a relation and $\pi_{I}\left(t \right)$ and $\pi_{J}\left(t \right)$ are the possibilities for a time point $t \in \T$ to be in the PVPs $I$ and $J$ respectively,
+\textbf{Strictly consistent} It is not possible for two PVPs of the same entity to be valid at the same time.
+In other words, if $A$ is an entity, $R_{FTG}$ is a relation and $\pi_{I}\left(t \right)$ and $\pi_{J}\left(t \right)$ are the possibilities for a time point $t \in \T$ to be in the PVPs $I$ and $J$ respectively,
 \begin{align}
@@
 \begin{definition}
- \textbf{Ill-consistent} The overlapping of two PVPs of the same object do overlap. It can be distinguished among the following sub-types:
+ \textbf{Ill-consistent} The overlapping of two PVPs of the same entity do overlap. It can be distinguished among the following sub-types:
 \end{definition}
@@
 	      \begin{definition}
-	      \textbf{Co-existence} Two versions of the same object $A$ may exist at the same time.
+	      \textbf{Co-existence} Two versions of the same entity $A$ may exist at the same time.
 		      \begin{align}
@@
 	      \begin{definition}
-	      \textbf{Weak-consistence} Several versions of the same object $A$ may exist at the same time.
+	      \textbf{Weak-consistence} Several versions of the same entity $A$ may exist at the same time.
 		      \begin{align}
@@
 This is a natural extension of the crisp temporal databases.
-This approach allows a easy implementation for the DML operations defined in Section \ref{subsubsec:Understanding-valid-time-databases}.
+This approach allows an implementation for the DML operations defined in Section \ref{subsubsec:Understanding-valid-time-databases}.
 
-Since the time intervals are now possibilistic valid-time periods, PVPs, the auxiliary functions defined in equations \eqref{eq:close-a-crisp-interval} to \eqref{eq:close-current} should be re-defined.
+Since the time intervals are now possibilistic valid-time periods, PVPs, the auxiliary functions defined in equations \eqref{eq:close-a-crisp-interval} to \eqref{eq:close-current} are the basis for the following auxiliary functions.
 
@@
 \label{def:pvp-current-in-relation}
-\textbf{Current$\left(R, A \right)$.}
-Consider an entity $A$ and a relation $R$ in a fuzzy temporal database.
-The function Current$\left(R, A \right)$ obtains the tuple $\left(A, I \right), I = \left(X, Y \right)$ that is current in the relation $R$, that is it, $I = \left[X , \text{UNKNOWN} \right]$.
+\textbf{Current$\left(R_{FTG}, A \right)$.}
+Consider an entity $A$ and a relation $R_{FTG}$ in a fuzzy temporal database.
+The function Current$\left(R_{FTG}, A \right)$ obtains the tuple $\left(A, I \right), I = \left(X, Y \right)$ that is current in the relation $R_{FTG}$, that is, $I = \left[X , \text{UNKNOWN} \right]$.
 
@@
 \label{eq:pvp-current-in-relation}
-\mbox{Current} \left(R, A \right) &=& \\ 
+\mbox{Current} \left(R_{FTG}, A \right) &=& \\ 
 \begin{cases}
@@
 For example, consider the database in Table \ref{tbl:fuzzy-library-sample-valid-time}.
-The function Current$\left(R, A \right)$ with $A =\left(ID=3\right)$ returns time interval for the current version of $A$, $I = \left[ \left[4/4/2012, 3, 3 \right], UC \right]$.
+The function Current$\left(R_{FTG}, A \right)$ with $A =\left(ID=3\right)$ returns time interval for the current version of $A$, $I = \left[ \left[4/4/2012, 3, 3 \right], UC \right]$.
 Nevertheless, the time interval for the entity with ID = 4 is the empty set.
-Now it is possible to close the current version of an object by using \eqref{eq:close-r-a-pvp} and \eqref{eq:pvp-current-in-relation}.
+Now it is possible to close the current version of an entity by using \eqref{eq:close-r-a-pvp} and \eqref{eq:pvp-current-in-relation}.
 
@@
 \label{def:pvp-close-current-version}
-\textbf{Close-current$\left(R, \left(A, J \right) \right)$.}
-Consider an object $A$, the relation $R$ and a pvp, $J$.
-Then, the function close - current$\left(R, A, J \right)$ closes any current version of the object $A$ if it exists and add the new version $\left(A, J \right)$.
+\textbf{Close-current$\left(R_{FTG}, \left(A, J \right) \right)$.}
+Consider an entity $A$, the relation $R_{FTG}$ and a pvp, $J$.
+Then, the function close - current$\left(R_{FTG}, A, J \right)$ closes any current version of the entity $A$ if it exists and add the new version $\left(A, J \right)$.
 
@@
 \label{eq:pvp-close-current}
-\text{Close-current} \left(R, \left(A, J \right) \right) =\\
+\text{Close-current} \left(R_{FTG}, \left(A, J \right) \right) =\\
 \begin{cases}
 \nonumber
-\big \lbrace R - \left(A, I \right) \cup \left \lbrace \left(A, \mbox{ CloseR } \left(I, J\right) \right) \right \rbrace \cup \left \lbrace \left(A, J\right)\right \rbrace  \big \rbrace \\
+\big \lbrace R_{FTG} - \left(A, I \right) \cup \left \lbrace \left(A, \mbox{ CloseR } \left(I, J\right) \right) \right \rbrace \cup \left \lbrace \left(A, J\right)\right \rbrace  \big \rbrace \\
 \nonumber
-\mbox{ if } I = \mbox{ Current } \left(R, A \right) \\ %\wedge J > \mbox{ CloseR } \left(I, J \right)   \\
+\mbox{ if Current } \left(R_{FTG}, A \right) \neq \emptyset \\ %\wedge J > \mbox{ CloseR } \left(I, J \right)   \\
 \nonumber R , \text{ in any other case}
@@
 For example, consider the database in Table  \ref{tbl:fuzzy-library-sample-valid-time}.
-The function Close-current1$\left(R, \left(ID=3\right) \right.,$ $\left.
-\left[ \left[15/4/2012,2,1\right], UC\right]\right)$ closes the current version of the book with ID=3 and creates a new version with the specified time interval.
+The function Close-current1$\left(R_{FTG}, \left(ID=3\right) \right.,$ $\left.
+\left[ \left[15/4/2012,2,1\right], UC\right]\right)$ closes the current version of for the patient with ID=3 and creates a new version with the specified time interval.
 
 \subsubsection{\label{subsubsec:modify-fuzzy-temporal}Modify}
-This operation adds new information about an existing entity $A$ in the relation $R$. The modify operation does not remove any previous value of the entity $A$.
-Note that the modify operation is only applicable when the object $A$ is current in the relation $R$; $ \left(A, I \right) \in R, I = \left[X, \UC \right]$.
+This operation adds new information about an existing entity $A$ in the relation $R_{FTG}$. The modify operation does not remove any previous value of the entity $A$.
+Note that the modify operation is only applicable when the entity $A$ is current in the relation $R_{FTG}$; $ \left(A, I \right) \in R, I = \left[X, \UC \right]$.
+
+\begin{definition}
+\emph{$\MOD \left( R_{FTG}, \left( A, I \right) \right)$}. 
+Let $R_{FTG} \subseteq \A \  x\  \I_{PVP}$ be a  valid-time relation.
+Let  $A=\left(A_1, \ldots, A_n \right), A \in \A$ be the set of (fuzzy) attributes that define an entity and the PVP given by $I = \left(X, Y\right)$ the set of attributes defining the time interval.
+The algorithm for the modify operation is defined as follows.
+\end{definition}
 
@@
 \label{eq:modify-fuzzy-temporal}
-\MOD \left( R, \left( A, I \right) \right) =\\
+\MOD \left( R_{FTG}, \left( A, I \right) \right) =\\
 %\begin{cases}
 \nonumber
-\mbox{ Close-current }\left(R, \left(A, I\right) \right) 
+\mbox{ Close-current }\left(R_{FTG}, \left(A, I\right) \right) 
 % \mbox{ if Current}\left(R ,A \right) \neq \emptyset \\
@@
 \subsubsection{\label{subsubsec:insert-fuzzy-temporal}Insert}
-The user wants to store an entity $A$ which is valid in the relation $R$ during the time interval specified by the PVP, $I = \left(X, Y \right)$.
+The user wants to store an entity $A$ which is valid in the relation $R_{FTG}$ during the time interval specified by the PVP, $I = \left(X, Y \right)$.
 %
@@
 \begin{enumerate}
-\item The entity $A$ was never in the relation $R$: The entity is added with the valid-time indicated by the PVP, $I$.
-For example, consider the database given by Table \ref{tbl:fuzzy-library-sample-valid-time}. The following sentences correspond with the insertion of the first version of each book.
+\item The entity $A$ was never in the relation $R_{FTG}$: The entity is added with the valid-time indicated by the PVP, $I$.
+For example, consider the database given by Table \ref{tbl:fuzzy-library-sample-valid-time}. The following sentences correspond with the insertion of the first arrival of each patient.
 
 \begin{verbatim}
-Insert(3,'Dracula',
+Insert(3,'John',
 [[15/3/2012, 5, 2],
         [30/3/2012, 1, 1]]);
-Insert(4,'Frankenstein',
+Insert(4,'Antoon',
 [[25/3/2012, 3, 2], 
         [4/4/2012, 1, 7]]);
-Insert(5,'Harry Potter',
+Insert(5,'Daan',
 [[18/3/2012, 4, 1], 
@@
 
-\item The entity $A$ is in the relation $R$. Depending on the value of the time interval $I$, there are three possibilities:
+\item The entity $A$ is in the relation $R_{FTG}$. Depending on the value of the time interval $I$, there are three possibilities:
 	\begin{enumerate}
-	\item Insert $\left(A, I\right)$ in the relation $R$.
-If the time interval $I$ does not overlap any other valid-time interval in the relation $R$.
+	\item Insert $\left(A, I\right)$ in the relation $R_{FTG}$.
+If the time interval $I$ does not overlap any other valid-time interval in the relation $R_{FTG}$.
 Note that here, the result of the Overlaps operator is in the unit interval.
-For example, the librarian loans the book with ID=3 but the book is not returned yet. The insert sentence is the following.
+For example, the patient with ID=3 has not left the hospital yet. The insert sentence is the following.
+
  \begin{verbatim}
-Insert(3,'Dracula',
+Insert(3,'John',
 [[4/4/2012, 3, 3], UC]);
@@
 
-	\item Reject the insertion, if the time interval $I$ do overlap with a degree of 1 any existing valid-time interval for the entity $A$ in the relation $R$.
-For example, consider now that the librarian wants to introduce a past loan for the book with ID 3.
-The book was loaned around 6/4/12/2012 and returned around 25/4/2012.
-As this interval does overlaps other time intervals with a degree of 1, it is not possible that the book was loaned in that time interval.Therefore, the insertion is rejected.
+	\item Reject the insertion, if the time interval $I$ do overlap with a degree of 1 any existing valid-time interval for the entity $A$ in the relation $R_{FTG}$.
+For example, consider now that the hospital manager wants to introduce a past arrival for the patient with ID=3.
+The patient arrived around 6/4/12/2012 and left the hospital around 25/4/2012.
+As this interval does overlaps other time intervals with a degree of 1, it is not possible that the patient was it the hospital at that time interval.
+Therefore, the insertion is rejected.
 The insert sentence is:
@@
 
-The algorithm for the implementation of the insert operation is the following:
+\begin{definition}
+ \emph{$\INS \left(R_{FTG} ,\left(A, I \right) \right)$}.
+Let $R_{FTG} \subseteq \A \  x\  \I_{PVP}$ be a  valid-time relation.
+Let  $A=\left(A_1, \ldots, A_n \right), A \in \A$ be the set of (fuzzy) attributes that define an entity and the PVP given by $I = \left(X, Y\right)$ the set of attributes defining the time interval.
+The algorithm for the implementation of the insert operation is defined as follows.
+\end{definition}
 
@@
 \label{eq:insert-fuzzy-temporal}
-\INS \left(R ,\left(A, I \right) \right) &=\\
+\INS \left(R_{FTG} ,\left(A, I \right) \right) &=\\
 \begin{cases}
 \nonumber
-R \cup \left(A, I \right), & \mbox{ if }  \left(A, I \right) \not \in R \vee \\
-&  \left(A, J \right) \in R \wedge I \text{Overlaps} J < 1\\
+R_{FTG} \cup \left(A, I \right), & \mbox{ if }  \left(A, I \right) \not \in R_{FTG} \mbox{ or }\\
+&  \left(A, J \right) \in R_{FTG} \mbox{ and } I \text{ Overlaps } J < 1\\
 %\big \lbrace \left( \exists J= \left(X', Y' \right): \left(A, J \right) \in R \wedge \right. \\
@@
 %&  J<I<K  \big \rbrace\\
-\MOD \left(R, \left(A, I \right)\right), & \mbox{ otherwise }  
+\MOD \left(R_{FTG}, \left(A, I \right)\right), & \mbox{ otherwise }  
 \end{cases} 	
@@
 \subsubsection{\label{subsubsec:delete-fuzzy-temporal}Delete}
-The delete operation logically removes an entity $A$ which is valid in the relation $R$:
+The delete operation logically removes an entity $A$ which is valid in the relation $R_{FTG}$.
+
+\begin{definition}
+ \emph{$\DEL \left(R_{FTG} ,A \right) $}.
+Let $R_{FTG} \subseteq \A \  x\  \I_{PVP}$ be a  valid-time relation.
+Let  $A=\left(A_1, \ldots, A_n \right), A \in \A$ be the set of (fuzzy) attributes that define an entity and the PVP given by $I = \left(X, Y\right)$ the set of attributes defining the time interval.
+The algorithm for the delete operation is defined as follows.
+\end{definition}
+
 \begin{align}
 \label{eq:delete-fuzzy-temporal}
-\DEL \left(R ,A \right) =\\
+\DEL \left(R_{FTG} ,A \right) =\\
 \begin{cases}
 \nonumber
-R - \left(A, I \right), &  \forall \left( A, I \right) \in R\\
-R, & \mbox{ otherwise }  
+R_{FTG} - \left \lbrace \left(A, I_1 \right), \ldots, \left(A, I_n \right) \right \rbrace, &  \forall I_i: \left( A, I_i \right) \in R_{FTG}\\
+R_{FTG}, & \mbox{ otherwise }  
 \end{cases} 	

### 2012-06-20 ijcis-pos-model/sources/timeRepresentation.tex
final draft

@@
 %The datatypes for an  $X$ are shown in Table \ref{}.
+\begin{samepage}
 \vglue13pt
@@
 \tcap{\label{tbl:time-point-types}Values for the time point data type.}
-\centerline{\small TIME POINT DATA TYPES}
+\centerline{\small DIFFERENT VALUES FOR A TIME POINT}
 \vglue-6pt
@@
 
-%\begin{definition}
-\textbf{Ill-known time interval}\\
 An ill-known time interval denoted by $\left[X, Y\right]$ is a precise time interval whose boundaries are not precisely known.
-The ill-known time interval is defined by two ill-known  time points, namely $X$,$Y$.
+%The ill-known time interval is defined by two ill-known  time points, namely $X$,$Y$.
 The evaluation of the ill-known time interval is given by equations \eqref{ill-known-pos},\eqref{ill-known-nec}.
+
+\begin{definition}
+\emph{Ill-known time interval}
+Let $\T$ be the time domain, and $X$, $Y$ two ill-known values in the time domain.
+An ill-known time interval is given by $\left[X, Y\right]$.
+The evaluation of the ill-known time interval is given by equations \eqref{ill-known-pos},\eqref{ill-known-nec}.
+We will note $\I_{PVP}$ the set of all the ill-known time intervals.
+
+\end{definition}
 
@@
 $Open(C)$\\
-Consider an ill-known value $T$, a binary relationship $B_r$ one of $\left(\leq, <, >, \geq\right)$ and the constraint $C\stackrel{\triangle}{=}\left(R,T\right)$.
+Consider an ill-known value $T$, a binary relationship $B_r \in \left \lbrace \leq, <, >, \geq\right \rbrace$ and the constraint $C\stackrel{\triangle}{=}\left(R,T\right)$.
 The function $\Open(C) = \left(I_p(C), I_n(C) \right)$ provides both possibility and necessity measures for all the points in the open part of a semi-open ill-known time interval.
@@
 \label{eq:open-pos-nec}
-I_p \left(C \right) &=& \Big(\sup_{r\in\T \Rp\  w}\pi_{T}(w)\Big)\\
-I_n \left(C \right) &=& \Big(\inf_{r\in\T \Rn\  w}1-\pi_{T}(w)\Big)
+I_p \left(C \right) &=& \Big(\sup_{r\in\T, r \Rp\  w}\pi_{T}(w)\Big)\\
+I_n \left(C \right) &=& \Big(\inf_{r\in\T, r \Rn\  w}1-\pi_{T}(w)\Big)
 \end{eqnarray}
@@
 
+\begin{samepage}
 \vglue13pt
 %\begin{table}[htbp]
-\tcap{Relations for the $\Open(C)$ function. Depending on the relation $B_r = \left(\leq, <, >, \geq\right)$ in the constraint $C$, the values for $\Rp$ and $\Rn$ are shown.}
+\tcap{Relations for the $\Open(C)$ function. Depending on the relation $B_r \in \left \lbrace \leq, <, >, \geq\right \rbrace$ in the constraint $C$, the values for $\Rp$ and $\Rn$ are shown.}
 \centerline{\small Relations}
@@
 Consider an ill-known time interval given by $\left[\FB, Y \right]$.
-Consider also that in this case, $Y= \left[15/10/2012,3,4 \right] $.
+Consider also that, in this case, $Y= \left[15/10/2012,3,4 \right] $.
 Figure \ref{fig:example-open-time-interval} shows the representation for $Y$.
@@
 
+\begin{samepage}
 \vspace*{13pt}
@@
 \vspace*{13pt}
+\end{samepage}
 
@@
 Because of the ill-known constraints, not all the combinations of datatypes for each ill-known time point (see Table \ref{tbl:time-point-types}) are allowed.
-Table \ref{tbl:time-interval-types} shows all the possible datatypes combinations for the  for an ill-known time interval denoted by $\left[X, Y\right]$.
+Table \ref{tbl:time-interval-types} shows all the possible values that can be used to represent an ill-known time interval denoted by $\left[X, Y\right]$.
 
+\begin{samepage}
 \vglue13pt

### 2012-06-20 trunk/ijcis-pos-model/sources/abstract.tex
final draft

@@
 In real world, it is very common that some objects or concepts have properties with a time-variant or time-related nature.
-Modelling this kind of objects or concepts in a (relational) database schema is possible, but time-variant and time-related attributes have an impact on the consistency of the entire database.
+Modelling this kind of objects or concepts in a (relational) database schema is possible, but time-variant and time-related attributes have an impact on the consistency of the entire database and have to be appropriately managed.
 Therefore, temporal database models have been proposed to deal with this problem in the literature.
@@
 Several proposals for dealing with such imperfections when modelling temporal data exist.
-Some of these proposals transform the temporal data into a compact representation.
-There is not a model for managing and handling uncertainty regarding temporal information.
-In this work we present a novel model to deal with imprecision in valid-time databases together with the data manipulation language, \emph{DML} is defined and implemented.
+Some of these proposals transform the temporal data into a compact representation but there is not a formal model for managing and handling uncertainty regarding temporal information.
+In this work we present a novel model to deal with imprecision in valid-time databases together with the definition and implementation of the data manipulation language, \emph{DML}.

### 2012-06-20 trunk/ijcis-pos-model/sources/conclusions.tex
final draft

@@
-We presented a valid-time model to represent and handle ill-known temporal intervals.
+In this work, we presented a complete valid-time model to represent and handle ill-known temporal intervals.
+The paper includes the formal definition of possibilistic valid-time period in order to manage the time and the formal definition of ill-known constraints to define operators and integrity.
+This is the first formal model in the literature for possibilistic valid-time in relational databases.
 The semantics and the implementation of the DML operations are described within this work.
-As future work , the definition of the Data Definiton Language as well as the querying will be considered.
+As future work , the definition of the Data Definition Language as well as the querying will be considered.

### 2012-06-20 trunk/ijcis-pos-model/sources/intro.tex
version sent to Olga and Nicolas
final draft

@@
 As information systems attempt the modelling of natural objects, concepts or processes, they often require modelling temporal aspects or concepts.
-Thus, several proposals have arose to obtain theoretical models that allow the modelling or representation of time~\cite{Bolour1982,VanderCruyssen1997}.
+Thus, several proposals have arisen to obtain theoretical models that allow the modelling or representation of time~\cite{Bolour1982,VanderCruyssen1997}.
 
@@
 A database contains data representing real objects or concepts.
-In reality, some aspects or properties of objects or concepts are time-variant or time-related.
-e.g., the moment of a bank transaction is traditionally a point in time and thus a time-related notion, the function of an employee in a company can change through the recorded history and is thus time-variant.
-A temporal database schema is a database schema that models objects or concepts with time-related or time-variant properties.
-However, the modelling of temporal aspects has a direct impact on the consistency of the temporal database, because the temporal nature of these aspects imposes extra integrity constraints.
-An example.
+In real world, some aspects or properties of objects are time-variant or time-related.
+e.g., the moment of a bank transaction and the status of an employee in a company, are time-related and time-variant notions, respectively.
+A temporal database schema is a database schema that models objects with time-related or time-variant properties.
+However, the modelling of temporal aspects has a direct impact on the consistency of the temporal database, because the temporal nature of these aspects imposes extra integrity constraints and suitable ways of interactive with the human user.
 
-Consider a hospital database containing data about the state of the patients in the hospital.
+For example, let us consider a hospital database containing data about the state of the patients in a given hospital.
 Two dates are stored: the date when the patient arrives to and leaves from the hospital.
-It is clear that a patient cannot left the hospital if he or she has not arrived.
-Without further precautions, a hospital employee could insert a patient who is already in the hospital.
+It is clear that a patient cannot leave the hospital if he or she has not arrived.
+Without further cautions, a hospital employee could insert a patient who is already in the hospital.
 A temporal database model will typically constrain record insertion and prevent similar modelling inconsistencies.
@@
 Notably, in 1994, ``A Consensus Glossary of Temporal Database Concepts'' was published~\cite{Dyreson1994}.
-For this publication, 44 temporal database researchers, among these some of the main researchers in this field, cooperated to reach a consensus on the nature and definitions of some of the main temporal database concepts and terminology.
+For this publication, 44 temporal database researchers, among them, some of the main researchers in this field, cooperated to reach a consensus on the nature and definitions of some of the main temporal database concepts and terminology.
 This glossary was updated in 1998~\cite{Dyreson1998}.
@@
 In this sense, Allen~\cite{Allen1983} studied temporal relationships between time intervals (and as a special case time points).
-Among others, the querying of temporal databases has greatly profited from these temporal relationships, because they allowed richer and more complex user-specified temporal query demands, by allowing to express more complex relationships between the temporal notions in the temporal expressions in the query.
+Among others, the querying of temporal databases has greatly profited from these temporal relationships, because they allowed more powerful user-specified temporal query demands, by allowing to express more complex relationships between the temporal notions in the temporal expressions in the query.
 For example, a query like `who were the department heads when Thomas worked for the institution' can be evaluated using operators similar to Allen's ones.
@@
 Humans handle temporal information using certain temporal notions like time intervals or time points~\cite{Dyreson1994}, and they often have to deal with imperfections like imprecision, vagueness, uncertainty or inconsistencies possibly contained in the descriptions of these temporal notions.
-Among many others, these possible imperfections in descriptions of temporal notions determine an important issue in temporal modelling.
+These possible imperfections in descriptions of temporal notions determine an important issue in temporal modelling.
 Consider as an example the description of the temporal notion in a sentence like `The Belfry of Bruges was finished between 01/01/1201 A.D.
-and 31/12/1300 A.D.' contains imperfection because of the uncertainty in the used time-related expression.
-It is known that the building was finished on a single day, but it is not precisely known  which day this was.
+and 31/12/1300 A.D.'.
+These sentence contains imperfection because of the uncertainty in the used time-related expression.
+It is known that the building was finished on a single day, but this day is not precisely known.
 
-To allow information systems to cope with these and similar data imperfections, many approaches adopt fuzzy sets~\cite{Zadeh1965} for the representation of temporal information~\cite{Mitra1994,Nagypal2003,Billiet2011,Dubois2003}.
+To allow information systems to cope with these and similar data imperfections, many approaches adopt fuzzy sets~\cite{Zadeh1965} for the representation and management of temporal information~\cite{Mitra1994,Nagypal2003,Billiet2011,Dubois2003}.
 The temporal relationships studied by Allen were fuzzified by several authors~\cite{Ohlbach2004,Nagypal2003,Schockaert2008}.
-Although,  Garrido et al.
-~\cite{Garrido2009} present a compact representation for the time, they defined different relationships among time intervals by using a combination of regular fuzzy comparisons.
+Garrido et al.
+~\cite{Garrido2009} presented a compact representation for the time and defined different relationships among time intervals by using a combination of regular fuzzy comparisons.
 Also ~\cite{Garrido2009,Pons2011} studied uncertainty in temporal expressions concerning time intervals.
@@
 
-In addition to temporal modelling, some attention has been spent on temporal reasoning~\cite{Allen1983}.
+In addition to temporal modelling, some attention has been paid to temporal reasoning~\cite{Allen1983}.
 Although temporal reasoning is not discussed in this paper, it should be noted that, among others, Dubois and Prade et al.~\cite{Dubois2003,DuBois1989} have dealt with fuzziness and uncertainty in temporal reasoning.
@@
 The present work defines and implement a model for properly represent and manage uncertainty in valid-time specification in a relational database.
-None of the previous research offer a database model for acomplish this task.
+Our work is focused on both the proposal of an appropriate formal framework to suitably manage time in databases and the implementation of an DML that allows to the user the transparent use of this proposal.
+None of the previous research offer a database model to accomplish this task.
 
-The aim of this work is to present and explain the main operations of the manipulation language for a temporal database which stores the valid-time periods of the objects affected by imprecision.
+This way together with the theoretical model, we also present and explain the main operations of the manipulation language for a temporal database which stores the valid-time periods of the objects affected by imprecision.
 The rest of the work is organized as follows.
@@
 In section \ref{sec:time-rep} the representation of the valid-time intervals in the database is explained.
-Section \ref{sec:temporal-model} explains the main concepts and behaviour of the temporal model and the Data Manipulation Language (DML) is described and implemented.
-Finally, Section \ref{sec:conclusions} presents the conclusions and some possibilities for future research work.
+Section \ref{sec:temporal-model} explains the main concepts of the temporal Data Manipulation Language (DML) and its implementation.
+Finally, Section \ref{sec:conclusions} presents the conclusions and some guidelines for future research work.
 

### 2012-06-20 trunk/ijcis-pos-model/sources/preliminaries.tex
final draft

@@
 \begin{definition}
-Consider a possibility measure $\Pi$ on $\Omega$. A \emph{possibility distribution} $\pi$ on $\Omega$ underlying the possibility measure $\Pi$ is then a function defined by:
+Consider a possibility measure $\Pi$ on $\Omega$. A \emph{possibility distribution} $\pi$ on $\Omega$ underlying the possibility measure $\Pi$ is a function defined by:
 	\begin{align}
@@
 \begin{definition}
-Consider a confidence measure $N$ on a set of outcomes $\Omega$.
+Consider a confidence measure $N$ on $\Omega$.
 Let $J$ be a countable index set and let $\{ A_{j} | j \in J \wedge A_{j} \subseteq \Omega \}$ be a family of elements of $\wp(\Omega)$.
-$N$ is now a \emph{necessity measure} on $\Omega$ if it satisfies:
+$N$ is a \emph{necessity measure} on $\Omega$ if it satisfies:
 	\begin{align}
@@
 
-It is important to understand the difference between the following two concepts:
-\begin{itemize}
-\item
-A \emph{possibilistic variable} $X$ is bounded to take only one value , but this value is not known due to incomplete knowledge. 
-\item
-An \emph{ill-known set}~\cite{Dubois1988a}: a possibilistic variable defined over the universe $\Pow(U)$.
-\end{itemize}
+% It is important to understand the difference between the following two concepts:
+% \begin{itemize}
+% \item
+% A \emph{possibilistic variable} $X$ is bounded to take only one value , but this value is not known due to incomplete knowledge. 
+% \item
+% An \emph{ill-known set}~\cite{Dubois1988a}: a possibilistic variable defined over the universe $\Pow(U)$.
+% \end{itemize}
 
@@
 This work deals with ill-known intervals.
-These are ill-known sets, defined and represented via a starting and ending points, which are ill-known values.
+These are ill-known sets, defined and represented via a starting and ending points that, in turn are ill-known values.
 The elements of the set are the values between the starting and ending points.
@@
 \vspace*{13pt}
-
+\end{samepage}
 If this modal value $m$ is unique, then $M$ is referred to as a \emph{fuzzy number}. 
@@
 \vspace*{13pt}
+\end{samepage}
+
 %\def\JPicScale{0.5}
@@
 \begin{definition}
-Given a universe $U$, an ill-known constraint $C$ on a set $A \subseteq U$ is specified by means of a binary relation $R \subseteq U^{2}$ and a fixed, ill-known value denoted by its possibilistic variable $V$ over $U$, i.e.:
+Given a universe $U$, an ill-known constraint $C$ on a set $A \subseteq U$ is specified by means of a binary relation $R \subseteq U^{2}$ and a fixed ill-known value denoted by its possibilistic variable $V$ over $U$, i.e.:
 \begin{align}
@@
 
-\begin{example}
-As example, consider  $X = \left[5, 3, 2 \right]$ and let $C =(\leq, X)$ the ill-known constraint.
-Then, the evaluation of the possibility and the necessity are obtained from \eqref{ill-known-pos} and \eqref{ill-known-nec} respectively.
-(See Figure \ref{fig:example-ill-known}).
-
+\begin{samepage}
 \vspace*{13pt}
@@
 \vspace*{10pt}
-\fcaption{\label{fig:example-ill-known}Example of the evaluation of the ill-known constraint $C \triangleq (\leq, X)$. Possibility and necessity measures are shown in grey. }
+\fcaption{\label{fig:example-ill-known}Example of the evaluation of the ill-known constraint $C \triangleq (\leq, X), X = \left[5, 3, 2\right]$. Possibility and necessity measures are shown in grey. }
 %  \label{fig:fuzzy-validity-period}
 \vspace*{13pt}
+\end{samepage}
+
+\begin{example}
+Consider  $X = \left[5, 3, 2 \right]$ and let $C =(\leq, X)$ the ill-known constraint.
+Then, the evaluation of the possibility and the necessity are obtained from \eqref{ill-known-pos} and \eqref{ill-known-nec} respectively.
+(See Figure \ref{fig:example-ill-known}).
 
@@
 \nonumber
-\min_{a \in A}\left(\sup_{a \leq R}\pi_{X}(w)\right)\\
+\min_{a \in A}\left(\sup_{a \leq w}\pi_{X}(w)\right)\\
 \Nec(A\text{ satisfies }C) & =\\
@@
 It is observed that Boolean combinations of constraints are required.
-For example, the problem of interval evaluation (explained earlier) requires that all elements of an interval $[a,b]$ are larger than a value $X$ and, at the same time, smaller than a value $Y$, which implies that a conjunctive Boolean combination of both constraints must be satisfied.
+For example, the problem of interval evaluation (explained earlier) requires that all the elements of an interval $[a,b]$ are larger than a value $X$ and, at the same time, smaller than a value $Y$, which implies that a conjunctive Boolean combination of both constraints must be satisfied.
 To allow Boolean combinations of constraints, the following definitions are introduced.
@@
 \end{definition}
-Definition \ref{def:set-evaluation} presents the definition of an evaluation function that evaluates a Boolean combination of some basic constraints.
+Definition \ref{def:set-evaluation} presents a function that evaluates a Boolean combination of some basic constraints.
 Informally, it states that a set $A$ passes the evaluation made by $\lambda$ if the Boolean combination of some propositions equals $T$.
@@
 \begin{example}
-Consider that we want to check if crisp interval $I = \left[j, k\right]$ is included in $\left[X, Y\right]$. In this situation, two ill-known constraints are constructed.
+Consider that we want to check if the crisp interval $I = \left[j, k\right]$ is included in $\left[X, Y\right]$. In this situation, two ill-known constraints are constructed.
 
@@
 \begin{eqnarray}
+\label{eq:constraint-c1}
 C_1 & \triangleq\left(\geq,X\right)\\
+\label{eq:constraint-c2}
 C_2 & \triangleq\left(\leq,Y\right)
@@
 Consider a universe $U$.
-Let $R_1, R_2 \subseteq U$ be two binary relations in U.
+Let $R_1, R_2$ be two binary relations in U.
 Let $X_1, X_2$ be two fixed ill-known values in $U$.
@@
 \emph{Ill-known value defined by conjunctive combination of constraints.}
-Consider a universe $U$, $X$ an ill-known value and $CC \triangleq \left \lbrace C_{1_{Z_1}} \wedge \ldots \wedge C_{n_{Z_n}}  \right \rbrace$ a convex combination of ill-known constraints over the variables $Z_1, \ldots, Z_n$.
-Then, the value $X$ is defined by means of the convex combination:
+Consider a universe $U$, $CC \triangleq \left \lbrace C_{1_{Z_1}} \wedge \ldots \wedge C_{n_{Z_n}}  \right \rbrace$ a conjunctive combination of ill-known constraints over the variables $Z_1, \ldots, Z_n$.
+The uncertainty about the evaluation of an ill-known value $X$ is given by:
+\begin{align}
+\label{eq:ill-known-value-def-by-const-unc}
+X \in \Pow(U):\pi_{X}=\pi_{CC}
+\end{align}
+
+The definition of the ill-known value $X$ with respect to the conjunctive combination of ill-known constraints, is written as:
 \begin{align}
@@
 
-The uncertainty about the evaluation of an ill-known value $X$ is given by:
-\begin{align}
-\label{eq:ill-known-value-def-by-const-unc}
-X \in \Pow(U):\pi_{X}=\pi_{CC}
-\end{align} 
 Note that $\pi_{X}$ is convex since $\pi_{CC}$ is convex as demonstrated in Theorem \ref{th:convex-combination-ill-known-constraints}.
@@
 Now, we are going to explain the main concepts about the treatment of time and the imperfection related to the time in databases.
-In the following subsections we will show how to represent imperfect time intervals by means of ill-known values.
+This two preliminary analysis will be the pillars of our proposal in section \ref{sec:time-rep}.
+
+%In the following subsections we will show how to represent imperfect time intervals by means of ill-known values.

### 2012-06-20 trunk/ijcis-pos-model/sources/temporaldatabases.tex
version sent to Olga and Nicolas
final draft

@@
 \begin{itemize}
-\item \textbf{Imperfection in the database} the knowledge about the temporal data contains some imperfection. E.g., a database record shows that \emph{`The car was in the garage around April.'}
- \item \textbf{Imprecision in the query specification} denotes the imprecision in the specification of temporal criteria by the user, when querying.
+\item \textbf{Imperfection in the database}; the knowledge about the temporal data contains some imperfection. E.g., a database record shows that \emph{`The car was in the garage around April.'}
+ \item \textbf{Imprecision in the query specification}; it denotes the imprecision in the specification of temporal criteria by the user, when querying.
 E.g., \emph{`The user wants a car which was in the garage around April.'}
@@
 In order to deal with uncertainty in time intervals, several proposals have been made.
-Here, two approaches are introduced: the first one, based on \emph{Fuzzy Validity Periods}~\cite{Garrido2009} and the second one, based on \emph{Possibilistic Valid-time Periods}~\cite{JoseEnriquePons2012}.
+Here, two approaches are described: the first one, based on \emph{Fuzzy Validity Periods}~\cite{Garrido2009} and the second one, based on \emph{Possibilistic Valid-time Periods}~\cite{JoseEnriquePons2012}.
 
 \begin{definition}
-A \textbf{Fuzzy Validity Period} (\emph{FVP}) is defined as a fuzzy time interval specifying when the data regarding an object are valid.
+A \textbf{Fuzzy Validity Period}\cite{Garrido2009} (\emph{FVP}) is defined as a fuzzy time interval specifying when the data regarding an object are valid.
 A fuzzy time interval is then the fuzzification of a crisp time interval.\\
@@
 
-Several options to transform possibility distributions corresponding to the fuzzy starting point and the fuzzy end point into one consistent FVP exist \cite{Garrido2009}, e.g (Fig.
+Several options to transform possibility distributions corresponding to the fuzzy starting point and the fuzzy end point into a consistent FVP exist \cite{Garrido2009}, e.g (Fig.
 \ref{fig:fuzzy-validity-period}):
@@
 %%%%%%
+\begin{samepage}
 \vspace*{13pt}
@@
 \vspace*{10pt}
+\end{samepage}
+
 \fcaption{\label{fig:fuzzy-validity-period}Transformation to obtain the FVP.
@@
 \end{definition}
-A PVP is an ill-known interval in the sense of definition \ref{def;possibilistic-variable}.
-Note that this representation is \emph{disjunctive}: the PVP represents only one crisp time interval, but for some reason, it is (partially) unknown.
+A PVP is an ill-known interval in the sense of definition \ref{def;possibilistic-variable}. Note that the PVP represents only one crisp time interval, but for some reason, it is (partially) unknown.
 
@@
 %%
+\begin{samepage}
 \vglue13pt
@@
 } 
-
+\end{samepage}
 In the rest of the paper we will work only with PVP to represent valid-time intervals.
@@
 For the sake of simplicity, only the three main operations in the Data Manipulation Language (CReate Update, Delete) are shown.
-Usually the DML operations in a temporal database are re-defined (a typical update sentence in SQL could be re-defined by means of several insert and update sentences).
+Usually the DML operations in a temporal database are re-defined (a typical update sentence in SQL could be expressed by means of a couple of insert and update sentences).
 Therefore, for the sake of clarity, these high level primitives in the DML for a valid-time database are usually noted as Insert, Modify and Delete.
@@
 \emph{Valid-time relation.}
-Consider $A=\left(A_1, \ldots, A_n \right)$ be the set of attributes that define an entity and $I = \left(s_i,e_i \right)$ the set of attribute defining the time interval.
-The pair $\left(A, I\right)$ expresses that the data regarding the entities A are valid during the crisp time interval I.
-Let $R \subseteq A\  x\  I$ be a valid-time relation.
+Consider $\A$ be the set of all the entities, and let $A=\left(A_1, \ldots, A_n \right), A \in \A$ be the set of attributes that define an entity.
+Let $I = \left(S_i, E_i \right)$ be the set of attributes defining the time interval.
+The set $\I$ is the set of all the time intervals.
+The pair $\left(A, I\right)$ expresses that the data regarding the entity A are valid during the crisp time interval I.
+Let $R \subseteq \A \  x\  \I$ be a valid-time relation.
 Then, the following equation indicates that the pair $\left(A, I\right)$ is in the relation R:
@@
 \emph{From the beginning (FB).} 
-Consider $A=\left(A_1, \ldots, A_n \right)$ be the set of attributes that define an entity and $I = \left(s_i,e_i \right)$ the set of attribute defining the time interval.
-Let $R \subseteq A\  x\  I$ be a valid-time relation.
-The from the beginning constant is used when the value for $s_i$ is equal to $-\infty$.
+Consider $A=\left(A_1, \ldots, A_n \right)$ be the set of attributes that define an entity and $I = \left(S_i, E_i \right), S_i, E_i \in \T$ the set of attributes defining the time interval.
+Let $R \subseteq \A\  x\  \I$ be a valid-time relation.
+We will say $I = \left(FB, E_i \right)$ when $I = \left(- \infty, E_i \right)$.
+%The from the beginning constant is used when the value for $s_i$ is equal to $-\infty$.
 \end{definition}
@@
 \emph{Until changed (UC).} 
-Consider $A=\left(A_1, \ldots, A_n \right)$ be the set of attributes that define an entity and $I = \left(s_i,e_i \right)$ the set of attribute defining the time interval.
-Let $R \subseteq A\  x\  I$ be a valid-time relation.
-The until changed constant is used when the value for $e_i$ is equal to $+\infty$.
+Consider $A=\left(A_1, \ldots, A_n \right)$ be the set of attributes that define an entity.
+Let $\T$ be the time domain, and $I = \left(S_i, E_i \right), S_i, E_i \in \T$ the set of attributes defining the time interval.
+Let $R \subseteq \A\  x\  \I$ be a valid-time relation.
+We will say  $I = \left(S_i, UC \right)$ when $I = \left(S_i, +\infty \right)$.
+%The until changed constant is used when the value for $e_i$ is equal to $+\infty$.
 \end{definition}
 
-If the pair $\left(A, I \right)$ with $I = \left(c, UC \right), c \in \T$ is in the relation $R$, then, it is said that the pair $\left(A, I \right)$ is \emph{current} in the relation.
+If the pair $\left(A, I \right)$ with $I = \left(S_i, UC \right), S_i \in \T$ is in the relation $R$, then, it is said that the pair $\left(A, I \right)$ is \emph{current} in the relation.
 
-For example, last row in Table \ref{tbl:library-sample}.
+For example, let us consider the last row in Table \ref{tbl:library-sample}.
 The value for the time interval is $I = \left(4/4/2012 , UC \right)$.
-The meaning is that the patient with ID = 3 was in the hospital the  4/4/2012 and it is still there.
+The meaning is that the patient with ID = 3 was in the hospital the  4/4/2012 and he or she is still there.
 The patient with ID = 3 is current in the relation.
@@
 
+\begin{samepage}
 \vglue13pt
@@
 \emph{Current $\left(R, A \right)$.}
-Consider $A=\left(A_1, \ldots, A_n \right)$ be the set of attributes that define an entity and $I = \left(s_i,UC \right)$ the set of attribute defining the time interval.
-Let $R \subseteq A\  x\  I$ be a valid-time relation.
+Consider $A=\left(A_1, \ldots, A_n \right), A \in \A$ be the set of attributes that define an entity and $I = \left(S_i, UC \right)$ the set of attribute defining the time interval.
+Let $R \subseteq \A\  x\  \I$ be a valid-time relation.
 The function Current $\left(R, A \right)$ returns the crisp time interval $I$ if the entity $A$ is currently in the relation $R$ or the empty set, if the entity $A$ is not currently in the relation.
+It is defined as follows:
 
@@
 \nonumber
-I & \mbox{ if } (A,I) \in R \mbox{ and } I = \left(c, UC \right), c \in \T\\
-\emptyset & \mbox{ in any other case }
+I & \mbox{ if } (A,I) \in R \mbox{ and } I = \left(S_i, UC \right), S_i \in \T\\
+\emptyset & \mbox{ otherwise}
 \end{cases}
@@
 
-For example, the function Current(R, (3,`John')) returns the time interval $I = \left(4/4/2012 , UC  \right)$. On the other hand, Current(R, (4,`Antoon')) returns the empty set.
+For example, the function Current(R, (3,`John')) returns the time interval $I = \left(4/4/2012 , UC  \right)$. Conversely, Current(R, (4,`Antoon')) returns the empty set.
 
@@
 \vspace*{13pt}
+\end{samepage}
+
+We will use two of the Allen's relations: Overlaps and During which will be defined as follows.
 
@@
  \label{def:overlaps}
-Given two time intervals $I = \left(c, d \right)$ and $J = \left(e, f\right)$, it is said that I overlaps J if:
+Given two time intervals $I = \left(S_i, E_i \right)$ and $J = \left(S_j, E_j\right)$, it is said that I overlaps J if:
 \begin{equation}
  \label{eqn:overlaps}
-\text{I Overlaps J}  = \left(c > e  \right) \wedge \left(d < f  \right) 
+\text{I Overlaps J}  = \left(S_i > S_j  \right) \wedge \left(E_i < E_j  \right) 
 \end{equation}
@@
  \label{def:during}
-Given two time intervals $I = \left(c, d \right)$ and $J = \left(e, f\right)$, it is said that I during J if:
+Given two time intervals $I = \left(S_i, E_i \right)$ and $J = \left(S_j, E_j\right)$, it is said that I during J if:
 \begin{equation}
  \label{eqn:during}
-\text{I During J}  = \left(c < e  \right) \wedge \left(d > e  \right) \wedge \left(d < f \right)
+\text{I During J}  = \left(S_i < S_j  \right) \wedge \left(E_i > S_j  \right) \wedge \left(E_i < E_j \right)
 \end{equation}
@@
 \label{def:close-a-crisp-interval-r}
-Consider two crisp intervals $I= \left[c ,d \right]$ and $J= \left[e ,f \right]$.
+Consider two crisp intervals $I= \left[S_i ,E_i \right]$ and $J= \left[S_j ,E_j \right]$.
 The CloseR$\left(I, J\right)$ function allows to close the right-open interval $I$ with respect to the first value $e$ in $J$:
@@
 %I=\left[a, c \right[ & \mbox{ if } b = +\infty \wedge J > I
-\left[c, e \right[ & \mbox{ if } d = UC \mbox{ and } J \mbox{ During } I \\
-I & \mbox{ in any other case.}
+\left[S_i, S_j \right[ & \mbox{ if } E_i = UC \mbox{ and } J \mbox{ During } I \\
+I & \mbox{ otherwise}
 \end{cases}
@@
 \emph{Close-current$\left(R, \left(A, J  \right)\right)$}.
-Consider $A=\left(A_1, \ldots, A_n \right)$ be the set of attributes that define an entity and $J = \left(s_j,e_j \right)$ the set of attribute defining the time interval.
-Let $R \subseteq A\  x\  I$ be a valid-time relation.
+Consider $A=\left(A_1, \ldots, A_n \right)$ be the set of attributes that define an entity and $J = \left(S_j,E_j \right)$ the set of attribute defining the time interval.
+Let $R \subseteq \A\  x\  \I$ be a valid-time relation.
 Then, the function close-current$\left(R,\left(A,J\right) \right)$ closes any current version of the object $A$ if it exists and add the new version $\left(A, J\right)$.
@@
 \mbox{ if  Current } \left(R, A \right) \neq \emptyset\\ %\wedge J > \mbox{ CloseR } \left(I, J \right)   \\
-\nonumber R , \text{ in any other case}
+\nonumber R , \text{ otherwise}
 \end{cases}
@@
 
+\begin{definition}
+ \emph{Modify(R, (A,I))}.
+Let $R \subseteq \A\  x\  \I$ be a valid-time relation and  $A=\left(A_1, \ldots, A_n \right), A \in \A$ be the set of attributes that define an entity and $I = \left(S_i,E_i\right)$ the set of attributes defining the time interval.
+The algorithm for the modify operation is defined as follows.
+\end{definition}
+
 \begin{align}
@@
 \subsubsection{\label{subsubsec:insert}Insert}
-The user wants to store an object $A$ which is valid in the relation $R$ during the time interval specified by $I = \left(a, b \right)$.
+The user wants to store an entity $A$ which is valid in the relation $R$ during the time interval specified by $I = \left(a, b \right)$.
 %
@@
 \begin{enumerate}
-\item The object $A$ was never in the relation $R$: The object is added with the valid-time indicated by the crisp interval $I$. 
+\item The entity $A$ was never in the relation $R$: The entity is added with the valid-time indicated by the crisp interval $I$. 
 % In the example, this correspond with the following insertion sentences:
@@
 
-\item The object $A$ is in the relation $R$. Depending on the value of the time interval $I$, there are three possibilities:
+\item The entity $A$ is in the relation $R$. Depending on the value of the time interval $I$, there are three possibilities:
 	\begin{enumerate}
 	\item Insert $\left(A, I\right)$ in the relation $R$.
-If the time interval $I$ does not overlap any other valid-time interval in the relation $R$.
-For example, consider that the book with ID 3 is loaned on 4/4/2012 but the book has not been returned yet.
+If the time interval $I$ does not overlap any other valid-time interval in the relation $R$ for the entity $A$.
+For example, consider that the book with ID = 3 is loaned on 4/4/2012 but the book has not been returned yet.
 This is illustrated in Table \ref{tbl:library-sample}.
@@
 	\item Modify and close the current version of $A$ and insert $\left(A, I\right)$.
-For example, consider now that the book with ID 3 is loaned  the 24/4/2012.
-Here the problem is that the book with ID 3 was loaned on 4/4/2012, but not returned.
-If the book is already in the library for a new loan, then it is necessary to set the return date and add a new row with the new loan date.
+For example, consider now that the patient with ID = 3 arrived to the hospital on 24/4/2012.
+Here the problem is that the patient with ID = 3 arrived on 4/4/2012 but, for some reason, the departure date was not stored.
+If the patient is already at the hospital, then it is necessary to set the departure date and add a new row with the new arrival date.
 This is illustrated in Table \ref{tbl:library-sample-2}.
@@
 % 	      \end{verbatim}
-	\item Reject the insertion, if the time interval $I$ does overlap any existing valid-time interval for the object $A$ in the relation $R$.
-For example, consider now that the librarian wants to introduce a past loan for the book with ID 3.
-The book was loaned on 6/4/12/2012 and returned on 25/4/2012.
-As this interval does overlaps other time intervals, it is not possible that the book was loaned in that time interval.Therefore, the insertion is rejected.
+	\item Reject the insertion, if the time interval $I$ does overlap any existing valid-time interval for the entity $A$ in the relation $R$.
+For example, consider that the hospital manager wants to introduce a past arrival for the patient with ID = 3.
+The patient arrived on 6/4/2012 and left the hospital on 25/4/2012.
+As the dates do overlap, it is not possible that the patient was in the hospital at that time interval.
+Therefore, the insertion is rejected.
 %  The insert sentence is:
@@
 \end{enumerate}
-
+\begin{samepage}
 \vglue13pt
 %\begin{table}[htbp]
-\tcap{\label{tbl:library-sample-2}Sample library database}
+\tcap{\label{tbl:library-sample-2}Sample hospital database}
 %\centerline{\small DATA TYPES}
@@
 }
+\end{samepage}
 
-The algorithm for the implementation of the insert operation is the following:
+\begin{definition}
+ \emph{Insert(R,(A,I)))}.
+Let $R \subseteq \A\  x\  \I$ be a valid-time relation and  $A=\left(A_1, \ldots, A_n \right), A \in \A$ be the set of attributes that define an entity and $I = \left(S_i,E_i\right)$ the set of attributes defining the time interval.
+Then, the algorithm for the implementation of the insert operation is defined as follows.
+\end{definition}
 
@@
 \subsubsection{\label{subsubsec:del}Delete}
-The delete operation logically removes a current entity $A$ which is valid in the relation $R$:
+The delete operation logically removes a current entity $A$ which is valid in the relation $R$.
+\begin{definition}
+ \emph{Delete(R,(A,I))}.
+Let $R \subseteq \A\  x\  \I$ be a valid-time relation and  $A=\left(A_1, \ldots, A_n \right), A \in \A$ be the set of attributes that define an entity and $I = \left(S_i,E_i\right)$ the set of attributes defining the time interval.
+The algorithm for the delete operation is defined as follows.
+\end{definition}
+
 \begin{align}
@@
 \nonumber
-R - \left \lbrace \left(A, I_1 \right),  \ldots, \left(A, I_1 \right)   \right \rbrace, & \forall I_i : \left( A, I_i \right) \in R\\
+R - \left \lbrace \left(A, I_1 \right),  \ldots, \left(A, I_n \right)   \right \rbrace, & \forall I_i : \left( A, I_i \right) \in R\\
 R, & \mbox{ otherwise }  
@@
 
-For example, consider that the hospital wants to delete the history for the patient with ID = 3. The result of this operation is shown in Table \ref{tbl:library-sample-3}.
+For example, consider that the hospital manager wants to delete the history for the patient with ID = 3. The result of this operation is shown in Table \ref{tbl:library-sample-3}.
 

### 2012-06-22 trunk/ijcis-pos-model/sources/preliminaries.tex
last updates

@@
 
-Note that while a possibilistic variable refers to one (partially) unknown value, an ill-known set is a crisp set but, for some reason, (partially) unknown.
+Note that while a single ill-known value refers to one (partially) unknown value, an ill-known set is a crisp set but, for some reason, (partially) unknown.
 
@@
 \begin{example}
-As an example, consider the hospital database.
-When a patient enters o leaves a medical service in hospital, a record tracking the patient is stored.
+As an example, consider an employee's database.
+When an employee changes the status in the company, a record tracking the employee is stored.
 The exact time for this is not precisely known.
-Consider now that $X = \left[12, 2, 2\right]$ is the time in hours when a patient left the emergency service and $Y = \left[18, 2, 1 \right]$ is the time in hours when the patient left the hospital.
-It is possible to obtain $Z$ (the period of time between the emergency service and the left of the patient) by using a conjunctive combination of constraints.
+Consider now that $X = \left[12, 2, 2\right]$ is the day of the month when an employee started a contract as a junior developer and $Y = \left[18, 2, 1 \right]$ is the day of the month when the employee changed the status to senior developer.
+It is possible to obtain $Z$ (the period of time between the junior and the senior contracts) by using a conjunctive combination of constraints.
 
@@
 Now, we are going to explain the main concepts about the treatment of time and the imperfection related to the time in databases.
-This two preliminary analysis will be the pillars of our proposal in section \ref{sec:time-rep}.
+These two preliminary analysis will be the pillars of our proposal in section \ref{sec:time-rep}.
 

### 2012-06-22 ijcis-pos-model/sources/abstract.tex
last updates

@@
 In real world, it is very common that some objects or concepts have properties with a time-variant or time-related nature.
-Modelling this kind of objects or concepts in a (relational) database schema is possible, but time-variant and time-related attributes have an impact on the consistency of the entire database.
+Modelling this kind of objects or concepts in a (relational) database schema is possible, but time-variant and time-related attributes have an impact on the consistency of the entire database and must be appropriately managed.
 Therefore, temporal database models have been proposed to deal with this problem in the literature.
@@
 Additionally, human beings manage time using temporal indications and temporal notions, which may also be imprecise.
-However, the imperfection in human-used temporal indications is supported by human interpretation, whereas information systems need appropiate support in order to accomplish this task.
+However, the imperfection in human-used temporal indications is supported by human interpretation, whereas information systems need appropriate support in order to accomplish this task.
 Several proposals for dealing with such imperfections when modelling temporal data exist.
-Some of these proposals transform the temporal data into a compact representation.
-There is not a model for managing and handling uncertainty regarding temporal information.
-In this work we present a novel model to deal with imprecision in valid-time databases toghether with the data manipulation language, \emph{DML} is defined and implemented.
+Some of these proposals transform the temporal data into a compact representation but there is not a formal model for managing and handling uncertainty regarding temporal information.
+In this work we present a novel model to deal with imprecision in valid-time databases together with the definition and implementation of the data manipulation language, \emph{DML}.

### 2012-06-22 ijcis-pos-model/sources/intro.tex
last updates

--- 2012-01-23 trunk/book/sources/intro.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-The concept of \emph{time} is easy to understand but very complex to define~\cite{klein94}~\cite{Shackle61}. 
-In Information Systems \emph{IS}, several proposals have been concerned with the obtaining of theoretical models that allow the representation of time~\cite{Bolour82}~\cite{Cru97}. 
-More towards temporal databases, a lot of work have been done.
-The first efforts were towards the representation of the historical information related to an object in the database(reference).
-Some works tried to extend the Entity Relationship Model \emph{ERM} ~\cite{Klopprogge:1983} but without an impact on any database standards like SQL~\cite{Sarda:1990:ESH:627277.627409}.
 
-The ``Consensus Glossary of Temporal Database concepts'' \cite{Dyreson1994} is the first publication where the main researchers in temporal databases define the main thesaurus of temporal database concepts.
-This glossary defines the main types of time in a temporal database and is the basis of furthers researches ~\cite{Sarda:1990:ESH:627277.627409} \cite{Jensen94thetsql2}.
+The concept of time is very complex to handle and interpret~\cite{Klein1994,Shackle1961}, though it is very natural and omnipresent in real world data.
+As information systems attempt the modelling of natural objects, concepts or processes, they often require modelling temporal aspects or concepts.
+Thus, several proposals have arisen to obtain theoretical models that allow the modelling or representation of time~\cite{Bolour1982,VanderCruyssen1997}.
 
-Another main issue in temporal databases is the querying.
-The user specifies in the query some temporal constraints, e.g.
-\emph{``before this year''},\emph{``after 1995''}.
-In order to solve these temporal comparisons, between the time specified in the query and the temporal data stored in the database, some basic relationships have been defined.
-Allen~\cite{Allen83} first introduced the temporal relations between time intervals and, to a lesser extent, to time points.
-Some authors~\cite{ohlbach2004},\cite{nagypal2003},\cite{schockaert08} have softened the temporal operators defined by Allen.
-Therefore it is possible to compute the relationships between two ill-known time points or intervals.
-In~\cite{garrido2009}, some different temporal operators are defined by a combination of regular fuzzy comparisons.
-%e.g.
-the \emph{BEFORE} operator is defined by means of the fuzzy less than (FLT) operator.
+A very specific type of information systems are database systems.
+A database contains data representing real objects or concepts.
+In real world, some aspects or properties of objects are time-variant or time-related.
+e.g., the moment of a bank transaction and the status of an employee in a company, are time-related and time-variant notions, respectively.
+A temporal database schema is a database schema that models objects with time-related or time-variant properties.
+However, the modelling of temporal aspects has a direct impact on the consistency of the temporal database, because the temporal nature of these aspects imposes extra integrity constraints and suitable ways of interaction with the human user.
 
-To allow information systems to cope with uncertainty, many approaches adopt fuzzy sets~\cite{Zadeh65} for the representation of temporal information~\cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy},\cite{Dubois:jucs_9_9:fuzziness_and_uncertainty_in},~\cite{devos94}.
-The time points representing the time when the object is valid in the reality might not be precisely known.
-In order to deal with this, several models have been proposed.
-Garrido~\cite{garrido2009} proposes a model to deal with uncertainty in the time interval by means of fuzzy intervals.
-Bronselaer~\cite{Pon11} proposes a consistent framework to deal and represent ill-known time values and their relationships.
-Rough sets~\cite{Pawlak1995} have been also used to represent time intervals~\cite{Qia09} in databases.
+For example, let us consider an employee's database containing data about the positions of employees in a company.
+Two dates are stored: the starting and ending dates respectively of the contract.
+It is clear that an employee cannot have a contract with an ending date without an starting date.
+Without further cautions, a manager in the company could insert a new contract for an employee who is already working for the company.
+A temporal database model will typically constrain record insertion and prevent similar modelling inconsistencies.
 
-%temporal reasoning and fuzzy temporal reasoning.
-Next to the manage of temporal information is the temporal reasoning~\cite{Allen83}.
-Dubois and Prade ~\cite{Dubois:jucs_9_9:fuzziness_and_uncertainty_in},\cite{Dubois89} deal with fuzziness and uncertainty in temporal reasoning, but this topic is outside the scope of this work.
+% let us consider a hospital database containing data about the state of the patients in a given hospital.
+Two dates are stored: the date when the patient arrives to and leaves from the hospital.
+It is clear that a patient cannot leave the hospital if he or she has not arrived.
+Without further cautions, a hospital employee could insert a patient who is already in the hospital.
+A lot of research concerns temporal database models and their approaches to the modelling of time.
+The first efforts were towards the representation of historical information related to objects represented by records in a database~\cite{Clifford1985}.
+Some proposals tried to extend the Entity Relationship Model~\cite{Klopprogge1983}, without impact on any database standards like SQL~\cite{Sarda1990}.
 
-%temporal commercial systems.
+Notably, in 1994, ``A Consensus Glossary of Temporal Database Concepts'' was published~\cite{Dyreson1994}.
+For this publication, 44 temporal database researchers, among them, some of the main researchers in this field, cooperated to reach a consensus on the nature and definitions of some of the main temporal database concepts and terminology.
+This glossary was updated in 1998~\cite{Dyreson1998}.
 
-The work is organized as follows:
+% %From time in general to time in information systems
+% In real world, some aspects or properties of objects, concepts or processes are time-variant or time-related.
+For example, the moment of a bank transaction is traditionally a point in time and thus a time-related notion, the function of an employee in a company can change through recorded history and is thus time-variant.
+As information systems often attempt the modelling of natural objects, concepts or processes, they often require modelling temporal aspects or concepts.
+Thus, several proposals have been concerned with obtaining theoretical models that allow the modelling or representation of time~\cite{Bolour1982},~\cite{VanderCruyssen1997}.
 
-In Section \ref{sec:time-domain} an study about the time and its properties is done.
-Section \ref{sec:temporal-databases} is an overview of the main problems when managing the time in a database, the different proposals for solving them and the temporal database proposals.
-The section concludes with a comparison among the different proposals.
-In Section \ref{sec:fuzzy-temporal-databases} we present a sum up with the proposals for handling the imprecision in temporal databases and the main deficiencies we have detected.
-Finally, there are several commercial temporal database management systems  \emph{TDMBS} like \cite{oracle2009},\cite{posgree2009},\cite{teradata2011},\cite{timedb2005}.
-All of them are analyzed and compared later on this work.
+% %From time in information systems to time in temporal databases
+% Time is a complex concept, which makes modelling time in database systems complex too.
+A \emph{temporal database} is a database that deals with certain time-related aspects in its schema: a \emph{temporal database schema}~\cite{Dyreson1994} is a database schema that models entities (and interactions or relationships between entities) with time-related or time-variant properties.
+However, the modelling of temporal properties or aspects has a direct impact on the consistency of the resulting temporal database, because the temporal nature of these aspects or properties imposes extra integrity constraints.
+An example.
+Consider a relation in a relational library database, representing the physical absence (or presence) of books in the library.
+Every physical book is represented by a unique identifier.
+Every record in the relation contains such an identifier, a date on which the corresponding book was loaned and a date on which it was subsequently returned (if it was returned).
+As such, every record represents a period of time during which the physical book corresponding with the identifier is loand to a library customer.
+Without further precautions, a library employee could add several records with the same book identifier, different `loaned'-dates and no `returned'-dates.
+This group of records would represent the same physical book being loaned several times on different dates and never returned, which is of course impossible.
+A temporal database model will typically constrain record insertion and prevent similar modelling inconsistencies.
+
+An interesting issue in temporal modelling concerns relationships between temporal notions.
+In this sense, Allen~\cite{Allen1983} studied temporal relationships between time intervals (and as a special case time points).
+Among others, the querying of temporal databases has greatly profited from these temporal relationships, because they allowed more powerful user-specified temporal query demands, by allowing to express more complex relationships between the temporal notions in the temporal expressions in the query.
+For example, a query like `who were the department heads when Thomas worked for the institution' can be evaluated using operators similar to Allen's ones.
+
+Humans handle temporal information using certain temporal notions like time intervals or time points~\cite{Dyreson1994}, and they often have to deal with imperfections like imprecision, vagueness, uncertainty or inconsistencies possibly contained in the descriptions of these temporal notions.
+These possible imperfections in descriptions of temporal notions determine an important issue in temporal modelling.
+Consider as an example the description of the temporal notion in a sentence like `The Belfry of Bruges was finished between 01/01/1201 A.D.
+and 31/12/1300 A.D.'.
+This sentence contains imperfection because of the uncertainty in the used time-related expression.
+It is known that the building was finished on a single day, but this day is not precisely known.
+
+To allow information systems to cope with these and similar data imperfections, many approaches adopt fuzzy sets~\cite{Zadeh1965} for the representation and management of temporal information~\cite{Mitra1994,Nagypal2003,Billiet2011,Dubois2003}.
+The temporal relationships studied by Allen were fuzzified by several authors~\cite{Ohlbach2004,Nagypal2003,Schockaert2008}.
+Garrido et al.
+~\cite{Garrido2009} presented a compact representation for the time and defined different relationships among time intervals by using a combination of regular fuzzy comparisons.
+Also ~\cite{Garrido2009,Pons2011} studied uncertainty in temporal expressions concerning time intervals.
+Other approaches, like~\cite{Qiang2009}, use rough sets~\cite{Pawlak1995} to represent time intervals.
+
+%Now introducing imperfections in representations of time...
+% Humans handle temporal information using certain temporal notions like specific time intervals or instants~\cite{Dyreson1994} and they often have to deal with imperfections like imprecisions, vaguenesses, uncertainties or inconsistencies possibly contained in (the descriptions of) these temporal notions.
+Among many others, these possible imperfections in (descriptions of) temporal notions determine an important issue in temporal modelling.
+As an example, consider the temporal notion described in a sentence like `The Belfry of Bruges was finished on a day somewhere between 01/01/1201 A.D.
+and 31/12/1300 A.D.'.
+The description of the temporal notion reads `a day somewhere between 01/01/1201 A.D.
+and 31/12/1300 A.D.'.
+This temporal notion thus contains imperfection in the form of the uncertainty about the exact day the last stone of the building was laid: it is known that the building was finished on a single day, but it is not known which day this was.
+To allow information systems to cope with these and similar temporal data imperfections, many approaches adopt fuzzy sets~\cite{Zadeh1965} and/or fuzzy logic to model temporal information~\cite{Mitra1994},~\cite{Nagypal2003},~\cite{Billiet2011},~\cite{Dubois2003}.
+
+In addition to temporal modelling, some attention has been paid to temporal reasoning~\cite{Allen1983}.
+Although temporal reasoning is not discussed in this paper, it should be noted that, among others, Dubois and Prade et al.~\cite{Dubois2003,DuBois1989} have dealt with fuzziness and uncertainty in temporal reasoning.
+
+The present work defines and implement a model for properly represent and manage uncertainty in valid-time specification in a relational database.
+Our work is focused on both the proposal of an appropriate formal framework to suitably manage time in databases and the implementation of an DML that allows to the user the transparent use of this proposal.
+None of the previous research offer a database model to accomplish this task.
+
+This way together with the theoretical model, we also present and explain the main operations of the manipulation language for a temporal database which stores the valid-time periods of the objects affected by imprecision.
+The rest of the work is organized as follows.
+Section \ref{sec:prelim} presents some background concepts about both possibility theory and temporal databases.
+In section \ref{sec:time-rep} the representation of the valid-time intervals in the database is explained.
+Section \ref{sec:temporal-model} explains the main concepts of the temporal Data Manipulation Language (DML) and its implementation.
+Finally, Section \ref{sec:conclusions} presents the conclusions and some guidelines for future research work.
+
+% 
+% 
+% %From information systems in general to database systems in particular:
+% Generally, \emph{information systems} model the structure and behavior of real objects, concepts or processes.
+A specific type of information systems are \emph{database systems}, which are computer systems designed to manage \emph{databases}.
+A database is basically a collection of (persistent) data.
+These (typically atomic) data represent real objects or concepts.
+In the context of database design, such objects or concepts are typically called \emph{entities} and a collection of similar entities is typically modelled by an \emph{entitytype}, which is basically a combination of a name and a list of \emph{attributes}, which describe properties of the entities.
+Typically, in an early stage of the database development process, the structures of and interactions or relationships between used entities are modelled in a \emph{design schema}, following some design model.
+A database thus contains (typically atomic) data.
+Each (atomic) part of these data is a result value of a measurement of a property of an entity or a description of a property of an entity and will correspond to the attribute of the entity's entitytype, which describes the property.
+% 
+% %From database models in general to the relational model in particular:
+% The structure and behavior of a database, along with some integrity and security restrictions, is dictated by a \emph{database model}.
+A database model is basically a collection of instructions and regulations, used to (logically) describe the structure and behavior of a database, along with some integrity and security restrictions.
+Applying a database model to a concrete design schema results in a \emph{database schema}, which models the logical structure and behavior of a database.
+Several different database models exist, but the most popular is the \emph{relational database model}~\cite{Codd:1970:RMD:362384.362685}.
+Following the relational database model, entitytypes are modelled as \emph{relations}, which comprise a name and a list of attributes, modelling the entitytype's attributes.
+% 
+% %From time in general to time in information systems
+% In reality, some aspects or properties of objects, concepts or processes are time-variant or time-related.
+For example, the moment of a bank transaction is traditionally a moment in time and thus a time-related notion, the function of an employee in a company can change through recorded history and is thus time-variant.
+The concept of time itself is very complex to handle and interpret~\cite{Klein1994},~\cite{Shackle1961}, though it is very natural and omnipresent.
+As information systems often attempt the modelling of natural objects, concepts or processes, they often require modelling temporal aspects or concepts.
+Thus, several proposals have been concerned with obtaining theoretical models that allow the modelling or representation of time~\cite{Bolour1982},~\cite{VanderCruyssen1997}.
+% 
+% %From time in information systems to time in temporal databases
+% Time is a complex concept, which makes modelling time in database systems complex too.
+A \emph{temporal database} is a database that deals with certain time-related aspects in its schema: a \emph{temporal database schema}~\cite{Dyreson1994} is a database schema that models entities (and interactions or relationships between entities) with time-related or time-variant properties.
+However, the modelling of temporal properties or aspects has a direct impact on the consistency of the resulting temporal database, because the temporal nature of these aspects or properties imposes extra integrity constraints.
+An example.
+Consider a relation in a relational library database, representing the physical absence (or presence) of books in the library.
+Every physical book is represented by a unique identifier.
+Every record in the relation contains such an identifier, a date on which the corresponding book was loaned and a date on which it was subsequently returned (if it was returned).
+As such, every record represents a period of time during which the physical book corresponding with the identifier is loand to a library customer.
+Without further precautions, a library employee could add several records with the same book identifier, different `loaned'-dates and no `returned'-dates.
+This group of records would represent the same physical book being loaned several times on different dates and never returned, which is of course impossible.
+A temporal database model will typically constrain record insertion and prevent similar modelling inconsistencies.
+% 
+% %Research about temporal databases
+% A lot of research concerns temporal database models and their approaches to the modelling of time.
+Some of the first proposals concerned the representation of historical information related to entities~\cite{Clifford1985}.
+Some proposals tried to extend the Entity Relationship Model~\cite{Klopprogge1983}, without impact on any database standards like SQL~\cite{Sarda1990}.
+Notably, in 1994, `A Consensus Glossary of Temporal Database Concepts' was published~\cite{Dyreson1994}.
+For this publication, 44 temporal database researchers, among which some of the main researchers in this field, cooperated to reach a consensus on the nature and definitions of some of the main temporal database concepts and their terminology.
+This glossary was subsequently updated in 1998~\cite{Dyreson1998}.
+In the presented work, the concepts and terminology from this glossary are used and followed.
+% 
+% %Research about temporal relationships
+% An interesting issue in temporal modelling concerns the relationships between temporal notions.
+Notably, Allen~\cite{Allen1983} studied temporal relationships between time intervals~\cite{Dyreson1994} (and, as a special case, instants~\cite{Dyreson1994}).
+Among others, the querying of temporal databases has greatly profited from these temporal relationships, because they allow for richer and more complex user-specified temporal query demands, by allowing to express more complex relationships between the temporal notions in the temporal expressions in the query and the temporal indications in the database.
+For example, given a relation representing who was department head of an institution during which time intervals, a query like `Who were the department heads during the time intervals when Thomas worked for the institution?' can be evaluated using similar relationships.
+%TODO: CHECK THIS LAST SENTENCE
+% 
+% %Now introducing imperfections in representations of time...
+% Humans handle temporal information using certain temporal notions like specific time intervals or instants~\cite{Dyreson1994} and they often have to deal with imperfections like imprecisions, vaguenesses, uncertainties or inconsistencies possibly contained in (the descriptions of) these temporal notions.
+Among many others, these possible imperfections in (descriptions of) temporal notions determine an important issue in temporal modelling.
+As an example, consider the temporal notion described in a sentence like `The Belfry of Bruges was finished on a day somewhere between 01/01/1201 A.D.
+and 31/12/1300 A.D.'.
+The description of the temporal notion reads `a day somewhere between 01/01/1201 A.D.
+and 31/12/1300 A.D.'.
+This temporal notion thus contains imperfection in the form of the uncertainty about the exact day the last stone of the building was laid: it is known that the building was finished on a single day, but it is not known which day this was.
+To allow information systems to cope with these and similar temporal data imperfections, many approaches adopt fuzzy sets~\cite{Zadeh1965} and/or fuzzy logic to model temporal information~\cite{Mitra1994},~\cite{Nagypal2003},~\cite{Billiet2011},~\cite{Dubois2003}.
+% 
+% %Now introducing imperfections in temporal relationships
+% The temporal relationships studied by Allen were fuzzified by several authors~\cite{Ohlbach2004},~\cite{Nagypal2003},~\cite{Schockaert2008}.
+Garrido et al.
+~\cite{Garrido2009} present different temporal operators, defined by a combination of fuzzy comparisons.
+Also,~\cite{Garrido2009},~\cite{Pons2011} studied uncertainty in the context of time intervals.
+Other approaches, like~\cite{Qiang2009}, use rough sets~\cite{Pawlak1995} to represent time intervals.
+% 
+% %Imperfections in temporal reasoning
+% Next to temporal modelling, some attention has gone to temporal reasoning~\cite{Allen1983}.
+Although the focus of this paper is temporal modelling, it should be noted that, among others, Dubois and Prade et al.~\cite{Dubois2003},~\cite{DuBois1989} have dealt with fuzziness and uncertainty in temporal reasoning.
+% 
+% %Overview of this work
+% %Include: what is the paper about? (situation and explanation of the problem we will solve and situation of our solution (we use constraints...))
+% %Include: how does the solution work? (short summary of how the solution will solve the problem/attend to the problem)
+% %Include: What exactly will be presented
+% %In all of these: stress every single innovative/novel point of the presented work!
+% %TODO: build the overview when the paper is finished
+% 
+% The aim of this work is to present and explain the main operations of the manipulation language for a temporal database which stores the valid-time periods of the objects with uncertainty.
+The rest of the work is organized as follows.
+Section \ref{sec:prelim} presents some background concepts about both possibility theory and temporal databases.
+In section \ref{sec:time-rep} the representation of the valid-time intervals in the database is explained.
+Section \ref{sec:temporal-model} explains the main concepts and behaviour of the temporal model.
+The Data Manipulation Language (DML) is described and implemented.
+Finally in Section \ref{sec:conclusions} presents the conclusions and some possibilities for future research work.
+%

### 2012-06-22 ijcis-pos-model/sources/temporaldatabases.tex
last updates

--- 2012-05-24 ijcis-pos-model/sources/temporaldatabases.tex
@@
 \subsubsection{Imperfection and time}
-Representing imprecision and its semantics when dealing with time has been studied for a long time.
-Several proposals for representing and computing imprecise time indications can be found in \cite{DeCaluwe1997,DeTre1997}.
+The representation of imprecision and its semantics when dealing with time has been studied for a long time.
+Several proposals for representing and handling imprecise time indications can be found in \cite{DeCaluwe1997,DeTre1997}.
 Also, the changes between several granularities can be seen as a source of imprecision \cite{Devos1998}.
 
-In the proposal section we will consider two kinds of imprecision:
+In this paper we will consider two kinds of imperfection:
 \begin{itemize}
-\item \textbf{Imperfection in the database} denotes the uncertainty that arises when the knowledge about the temporal data in the database is uncertain.
-E.g., a database record shows that \emph{`The car is in the garage around April.'}
- \item \textbf{Imprecision in the query specification} denotes the imprecision in the specification of temporal criteria by the user, when querying.
-E.g., \emph{`The user wants to obtain a car which is red and which is in the garage around April.'}
+\item \textbf{Imperfection in the database}; the knowledge about the temporal data contains some imperfection. E.g., a database record shows that \emph{`The car was in the garage around April.'}
+ \item \textbf{Imprecision in the query specification}; it denotes the imprecision in the specification of temporal criteria by the user, when querying.
+E.g., \emph{`The user wants a car which was in the garage around April.'}
 \end{itemize}
@@
 
-In order to deal with uncertainty in time intervals, several proposals are made.
-Here, two approaches are introduced: the first one, based on \emph{Fuzzy Validity Period}~\cite{Garrido2009} and the second one based on \emph{Possibilistic Valid-time Period}~\cite{JoseEnriquePons2012}.
+In order to deal with uncertainty in time intervals, several proposals have been made.
+Here, two approaches are described: the first one, based on \emph{Fuzzy Validity Periods}~\cite{Garrido2009} and the second one, based on \emph{Possibilistic Valid-time Periods}~\cite{JoseEnriquePons2012}.
 
 \begin{definition}
-A \textbf{Fuzzy Validity Period} (\emph{FVP}) is defined as a fuzzy time interval specifying when the data regarding an object is valid.
-A fuzzy time interval is then the fuzzification of a crisp time interval.
+A \textbf{Fuzzy Validity Period}\cite{Garrido2009} (\emph{FVP}) is defined as a fuzzy time interval specifying when the data regarding an object are valid.
+A fuzzy time interval is then the fuzzification of a crisp time interval.\\
 \end{definition}
-Several options to transform possibility distributions corresponding to the fuzzy starting point and the fuzzy end point into one consistent FVP exist \cite{Garrido2009}, e.g (Fig.
+
+Several options to transform possibility distributions corresponding to the fuzzy starting point and the fuzzy end point into a consistent FVP exist \cite{Garrido2009}, e.g (Fig.
 \ref{fig:fuzzy-validity-period}):
 \begin{itemize}
-\item The \textbf{convex hull} approach is the most intuitive approach. The resulting FVP is the convex hull of the union of both fuzzy sets.
-\item The \textbf{uncertainty preserving} approach is less intuitive but more realistic.
-The amount of uncertainty is maintained at the edges of the possibility distribution representing the FVP \cite{Garrido2009}.
+\item The \textbf{convex hull} approach is the most intuitive approach. The resulting FVP is the convex hull of the union of both possibility distributions.
+\item The \textbf{uncertainty preserving} approach. The amount of uncertainty is maintained at the edges of the possibility distribution representing the FVP \cite{Garrido2009}.
 \end{itemize}
@@
 
+The main feature for the FVP is the optimization for the storage.
+The compact representation is the result of a conjunctive semantic.
+The object is valid within all the time points inside the starting and ending points.
+
 \begin{definition}
-A \textbf{Possibilistic Valid-time Period} (\emph{PVP}) is an ill-known interval in time specifying when the data regarding an object is valid.
+A \textbf{Possibilistic Valid-time Period} (\emph{PVP}) is an ill-known interval of time specifying when the data regarding an object are valid.
 \end{definition}
-A PVP is an ill-known interval in the sense of definition \ref{def;possibilistic-variable}, section \ref{subsec:possibilistic-variables}.
-Note that this representation is \emph{disjunctive}: the PVP represents only one crisp time interval, but that for some reason it is (partially) unknown.
+A PVP is an ill-known interval in the sense of definition \ref{def;possibilistic-variable}. Note that the PVP represents only one crisp time interval, but for some reason, it is (partially) unknown.
 
-The ill-known interval approach has many advantages from the representation by the FVP as demonstrated in \cite{Pons2011},\cite{Pons2012}.
+The main feature for the PVP is that it preserves all the information for both starting and ending points\cite{Pons2011,Pons2012}.
 Table \ref{tbl:comparative-pvp-fvp} is a comparative between PVP and FVP.
@@
 This subsection is devoted to describe the behaviour of a crisp valid-time database.
-For the shake of simplicity, only the three main operations in the Data Manipulation Language \emph{DML} a.k.a \emph{CRUD} (CReate Update, Delete) operations are shown.
-Usually the DML operations in a temporal database are re-defined e.g.
-a typical update sentence in SQL could be re-defined by means of several insert and update sentences.
-Therefore, in order to be clear, these high level primitives in the DML for a valid-time database are usually noted as Insert, Modify and Delete.
+For the sake of simplicity, only the three main operations in the Data Manipulation Language (CReate Update, Delete) are shown.
+Usually the DML operations in a temporal database are re-defined (a typical update sentence in SQL could be expressed by means of a couple of insert and update sentences).
+Therefore, for the sake of clarity, these high level primitives in the DML for a valid-time database are usually noted as Insert, Modify and Delete.
 In the following subsections each primitive is defined and explained.
-Finally a illustrative example is given.
+Finally, an illustrative example is given.
 For a more complete information on the behaviour of a bitemporal database, please refer to ~\cite{Jensen1994}.
@@
 \begin{definition}
-Consider a relation $R$ in a temporal database, an object $A$ given by the attributes $\left(a_1, \ldots, a_n \right)$ and the crisp time interval $I$ given by $\left[a,\ b\right] : a,b \in \T, a \leq b$.
-A valid-time object $A$ in the relation $R$ is specified by:
+\label{def:valid-time-relation}
+\emph{Valid-time relation.}
+Consider the following definitions and notations:
+
+\begin{itemize}
+ \item A set of non-temporal attributes.
 	\begin{equation}
-\label{eq:rel-def}
-\left( R, A , I \right)
+	\label{eq:attribute-set}
+	A = \left \lbrace A_1, A_2, \ldots, A_n \right \rbrace
+	\end{equation}
+      The domain for each attribute $A_1, \ldots, A_n$ is $D_1, \ldots, D_n$ respectively. 
+\item The primary key for the attributes in $A$ is given by $A_K$:
+      \begin{equation}
+       \label{eq:primary-key-a}
+      A_K \subset A
+      \end{equation}
+\item A set of two attributes; $S$  and $E$ the attributes for the starting and ending points respectively. $I$ defines the valid time interval for the data. 
+\begin{equation}
+ \label{eq:attribute-time-interval}
+I = \left \lbrace S, E \right \rbrace
+\end{equation}
+\item Then $R$, the schema for the valid-time relation is:
+\begin{equation}
+ \label{eq:valid-time-relation}
+R = A \cup  I
+\end{equation}
+\item The primary key for the valid-time relation $R$ is:
+\begin{equation}
+ \label{eq:valid-time-temporal-pk}
+PK = A_K \cup I
+\end{equation}
+
+\item We will note by $r$ any valid instance of $R$. 
+      \begin{equation}
+       \label{eq:valid-time-instance}
+      r \subseteq D_1\ x\ \ldots\ x\ D_n
+      \end{equation}
+
+\item Let $t$ be a tuple in the instance $r$; $t \in r$:
+      \begin{itemize}
+      \item We will note the values for the starting and the ending  points,$s$ and $e$  respectively. The value for the time interval is given by $i$.
+      \begin{align}
+       \label{eq:starting-point}
+      s = t\left[S \right]\\
+      e = t\left[E \right]\\
+      i = \left(s, e\right)
+      \end{align}
+
+      \item Let $A_k \subset A$ be the set of the non-temporal attributes that are part of the primary key (equation \eqref{eq:primary-key-a}).
+Then, $a_k$ denotes the values for the non-temporal attributes of the primary key.
+	    \begin{equation}
+	     \label{eq:all-the-versions} 
+	      a_k = t\left[A_K \right]
+	    \end{equation}
+
+    \item Let $PK$ be the primary key for the valid-time relation as given in equation \eqref{eq:valid-time-temporal-pk}. Then, $pk$ denotes the values for the attributes in the primary key.
+	  \begin{equation}
+	   \label{eq:value-pk}
+	  pk = t\left[PK \right]
+	  \end{equation}
+
+      \end{itemize}
+\item $Vt$ is the set with all the versions for a given tuple $t$.
+
+\begin{equation}
+ \label{eq:all-the-versions}
+Vt = r\left(t\left[A_K\right] \right)
+\end{equation}
+If a tuple $t$ contains more than one version, the result of $Vt$ is a set.
+We will use the index $k$ to address the elements of the set.
+E.g.
+$Vt = \left \lbrace t_1, t_2, \ldots, t_n \right \rbrace$.
+Then $t_k , k \in \left \lbrace 1, \ldots, n \right \rbrace$ is an element of $Vt$.
+\end{itemize}
+\end{definition}
+We will illustrate the definitions with an example.
+
+\begin{example}
+Consider the set of attributes $A = \left \lbrace A_1, A_2, A_3 \right \rbrace$.
+The primary key for these attributes is given by $A_K = \left \lbrace A_1, A_2 \right \rbrace$.
+Let $I = \left \lbrace S, E \right \rbrace$ be the set of temporal attributes that define the validity period of the data.
+$R$ is the valid-time relation and $r$ is an instance of the relation.
+The instance $r$ is given by the following elements.
+$r = \left \lbrace \left(a_{11}, a_{12}, a_{13}, s_1 ,e_1 \right) \right.$,  $\left(a_{21}, a_{22}, a_{23}, s_2, e_2 \right)$ , $\left.
+\left(a_{11}, a_{12}, a_{31}, s_3, e_3 \right) \right \rbrace$.
+The instance $r$ is illustrated in Table \ref{tbl:sample-definitions}.
+Consider the tuple $t = \left(a_{11}, a_{12}, a_{13}, s_1 ,e_1 \right)$. Then,
+\begin{align}
+ \nonumber
+s_1 &= t\left[S \right]\\
+ \nonumber
+e_1 &= t[E]\\
+ \nonumber
+i\ \ &= \left(s_1, e_1\right)\\
+ \nonumber
+a_k &= t\left[A_K\right] = \left(a_{11}, a_{12} \right) \\
+ \nonumber
+pk &= t\left[PK\right] =\left(a_{11}, a_{12}, s_1, e_1\right)\\
+ \nonumber
+Vt &= r\left(t\left[A_K\right] \right) = \left \lbrace t_1, t_3 \right \rbrace
+\end{align}
+
+\end{example}
+
+\begin{samepage}
+\vglue13pt
+%\begin{table}[htbp]
+\tcap{\label{tbl:sample-definitions}Sample database containing the instance $r$ of the valid-time relation $R$.}
+%\centerline{\small DATA TYPES}
+\vglue-6pt
+\centerline{\small\baselineskip=13pt
+\begin{tabular}{c c c c c c}\\
+\hline
+& \textbf{A$_1$}  & \textbf{A$_2$}  & $A_3$ & $S$ & $E$ \\
+\hline
+$t_1$&$a_{11}$ & $a_{12}$ & $a_{13}$ & $s_1$ & $e_1$ \\
+$t_1$ & $a_{21}$ & $a_{22}$ & $a_{23}$ &  $s_2$ & $e_2$ \\
+$t_1$ & $a_{11}$ & $a_{12}$ & $a_{31}$ & $s_3$ & $e_3$\\
+\hline\\
+\end{tabular}
+}
+\end{samepage}
+
+% Consider $\A$ be the set of all the entities, and let $A=\left(A_1, \ldots, A_n \right), A \in \A$ be the set of attributes that define an entity.
+Let $I = \left(S_i, E_i \right)$ be the set of attributes defining the time interval.
+The set $\I$ is the set of all the time intervals.
+The pair $\left(A, I\right)$ expresses that the data regarding the entity A are valid during the crisp time interval I.
+Let $R \subseteq \A \  x\  \I$ be a valid-time relation.
+Then, the following equation indicates that the pair $\left(A, I\right)$ is in the relation R:
+% 
+% \begin{equation}
+% \label{eq:rel-def}
+% \left( R, \left(A , I\right) \right)
+% \end{equation}
+% 
+% 
+In order to simplify the algorithms for the manipulation of data, some auxiliary functions and constants are defined:
+
+\begin{definition}
+ \label{def:from-beginning}
+\emph{From the beginning (FB).} 
+Consider the elements in definition \ref{def:valid-time-relation}. We will say $i = \left(FB, e \right)$ when $i = \left(- \infty,e \right)$. 
+% $i = \left(s, e \right)$ be the value for a time interval as given equation \eqref{}.
+% $A=\left(A_1, \ldots, A_n \right)$ be the set of attributes that define an entity and $I = \left(S_i, E_i \right), S_i, E_i \in \T$ the set of attributes defining the time interval.
+Let $R \subseteq \A\  x\  \I$ be a valid-time relation.
+We will say $I = \left(FB, E_i \right)$ when $I = \left(- \infty, E_i \right)$.
+%The from the beginning constant is used when the value for $s_i$ is equal to $-\infty$.
+\end{definition}
+
+\begin{definition}
+ \label{def:until-changed}
+\emph{Until changed (UC).} 
+Consider the elements in definition \ref{def:valid-time-relation}. We will say  $i = \left(s, UC \right)$ when $i = \left(s, +\infty \right)$.
+% Consider $A=\left(A_1, \ldots, A_n \right)$ be the set of attributes that define an entity.
+Let $\T$ be the time domain, and $I = \left(S_i, E_i \right), S_i, E_i \in \T$ the set of attributes defining the time interval.
+Let $R \subseteq \A\  x\  \I$ be a valid-time relation.
+We will say  $I = \left(S_i, UC \right)$ when $I = \left(S_i, +\infty \right)$.
+%The until changed constant is used when the value for $e_i$ is equal to $+\infty$.
+\end{definition}
+
+\begin{definition}
+\emph{Current.} 
+Consider the elements in definition \ref{def:valid-time-relation}. We will say that the tuple $t$ is current in the instance $r$ of the relation $R$ when $i = \left(s, UC \right)$.
+\end{definition}
+
+%If the pair $\left(A, I \right)$ with $I = \left(S_i, UC \right), S_i \in \T$ is in the relation $R$, then, it is said that the pair $\left(A, I \right)$ is \emph{current} in the relation.
+
+For example, let us consider the last row in Table \ref{tbl:library-sample}.
+The value for the time interval is $I = \left(4/4/2012 , UC \right)$.
+The meaning is that the patient with ID = 3 was in the hospital the  4/4/2012 and he or she is still there.
+The patient with ID = 3 is current in the relation.
+
+\begin{example}
+\label{ex:library-database}
+ Consider a hospital database.
+Each patient is stored in the database with an identifier (ID), the name and the dates when the patient entered and left the hospital.
+Table \ref{tbl:library-sample} illustrates the first version of the database, after three insertions.
+In this example, $A = ($ ID, Name $)$ and $I = ($ Arrival, Leave $)$.
+\end{example}
+
+\begin{samepage}
+\vglue13pt
+%\begin{table}[htbp]
+\tcap{\label{tbl:library-sample}Sample hospital database}
+%\centerline{\small DATA TYPES}
+\vglue-6pt
+\centerline{\small\baselineskip=13pt
+\begin{tabular}{c c c c}\\
+\hline
+\textbf{ID}  & Name  & Arrival & Leave \\
+\hline
+3 & John & 15/3/2012 & 30/3/2012 \\
+4 & Antoon & 25/3/2012 & 4/4/2012 \\
+5 & Daan & 18/3/2012 & 2/4/2012\\
+3 & John & 4/4/2012 & UC \\
+\hline\\
+\end{tabular}
+}
+\end{samepage}
+
+\begin{definition}
+\label{def:current-in-relation}
+\emph{Current $\left(a_k \right)$.}
+Consider the elements in definition \ref{def:valid-time-relation}.
+The function Current $\left(a_k \right)$ returns the crisp time interval $i$ if the tuple $t$ with the primary key values given by $a_k$ is currently in the instance $r$ of the relation $R$.
+The function returns the emtyset otherwise.
+% 
+% Consider $A=\left(A_1, \ldots, A_n \right), A \in \A$ be the set of attributes that define an entity and $I = \left(S_i, UC \right)$ the set of attribute defining the time interval.
+Let $R \subseteq \A\  x\  \I$ be a valid-time relation.
+The function Current $\left(R, A \right)$ returns the crisp time interval $I$ if the entity $A$ is currently in the relation $R$ or the empty set, if the entity $A$ is not currently in the relation.
+It is defined as follows:
+
+%Consider an object $A$ and a relation $R$ in a temporal database.
+The function Current$\left(R, A \right)$ obtains the tuple $\left(A, I \right)$ that is current in the relation $R$, that is it, $I = \left[a, +\infty \right]$.
+
+\begin{align}
+\label{eq:current-in-relation}
+\mbox{Current} \left(a_k\right) &=& \\ 
+\begin{cases}
+\nonumber
+t & \mbox{ if } \exists t \in r : i = \left(s, UC \right)\\
+\emptyset & \mbox{ otherwise}
+\end{cases}
+\end{align}
+\end{definition}
+
+For example, the function Current(R, (3,`John')) returns the time interval $I = \left(4/4/2012 , UC  \right)$. Conversely, Current(R, (4,`Antoon')) returns the empty set.
+
+The Allen's relations between two time intervals are shown in Figure \ref{fig:allen-rel}.
+The complete implementation of the relations using only the starting and the ending points is defined in~\cite{Nagypal2003}.
+
+\begin{samepage}
+\vspace*{13pt}
+\begin{center}
+{
+\includegraphics[scale=0.5]{./graphs/allen.pdf}
+
+}
+\end{center}
+%\centerline{ \psfig{file=./graphs/Y-time-point.eps}}
+\vspace*{10pt}
+\fcaption{\label{fig:allen-rel}Allen's relations for two crisp time intervals $I$ and $J$.
+Note that the crisp time interval $I$ is fixed and the different positions of the interval $J$ illustrate the Allen's relations.}
+\vspace*{13pt}
+\end{samepage}
+
+We will use two of the Allen's relations: Overlaps and During which will be defined as follows.
+
+\begin{definition}
+\emph{Overlaps:}
+ \label{def:overlaps}
+Given two time intervals $i_1 = \left(s_1, e_1 \right)$ and $i_2 = \left(s_2 e_2\right)$, it is said that $i_1$ overlaps $i_2$ if:
+\begin{equation}
+ \label{eqn:overlaps}
+i_1 \text{ Overlaps } i_2  = \left(s_1 < s_2  \right) \wedge \left(e_1 < e_2  \right) 
 \end{equation}
@@
 
-In order to simplify the algorithms for the manipulation of data, some auxiliary functions are defined:
+\begin{definition}
+\emph{During:}
+ \label{def:during}
+Given two time intervals $i_1 = \left(s_1, e_1 \right)$ and $i_2 = \left(s_2 e_2\right)$, it is said that $i_1$ during $i_2$ if:
+\begin{align}
+ \label{eqn:during}
+i_1\text{ During } i_2  =& \left(s_1 > s_2  \right) \wedge \left(e_1 > s_2 \right) \wedge \\
+\nonumber & \left(e_1 < e_2 \right)
+\end{align}
+\end{definition}
+
+In the rest of the paper, and without losing generality, we will consider that the time granularity are days. Also, the dates will be given in the format \emph{dd/mm/yyyy}.
 
 \begin{definition}
+\emph{CloseR$\left(i_1, i_2\right)$:}
 \label{def:close-a-crisp-interval-r}
-Consider two crisp intervals $I= \left[a ,b \right]$ and $J= \left[c ,d \right]$, $c > a$.
-The CloseR$\left(I, J\right)$ function allows to close the right-open interval $I$ with respect to the first value $d$ in $J$:
+Consider two crisp intervals $i_1 = \left(s_1, e_1 \right)$ and $i_2 = \left(s_2, e_2\right)$.
+The CloseR$\left(i_1, i_2 \right)$ function allows to close the right-open interval $i_1$ with respect to the first value $s_2$ in $i_2$:
 \begin{align}
 \label{eq:close-a-crisp-interval}
-\mbox{CloseR} \left( I, J \right) &=& \\ 
+\mbox{CloseR} \left( i_1, i_2\right) &=& \\ 
 \begin{cases}
 \nonumber
-I & \mbox{ if } b \neq +\infty \\
-I=\left[a, c \right[ & \mbox{ if } b = +\infty \wedge J > I
+%I & \mbox{ if } b \neq +\infty \\
+%I=\left[a, c \right[ & \mbox{ if } b = +\infty \wedge J > I
+i_1 = \left(s_1,s_2 \right( & \mbox{ if } e_1 = UC \mbox{ and } i_2 \mbox{ During } i_1 \\
+i_1 & \mbox{ otherwise}
 \end{cases}
@@
 \end{definition}
+
+For example, consider two time intervals, $i_1 = \left(4/4/2012, UC \right)$ and $i_2 = \left(24/4/2012, UC \right)$.
+The result of applying the CloseR$\left(i_1, i_2 \right)$ is $i_1 = \left(4/4/2012, 23/4/2012 \right)$.
 
@@
 
-Now it is possible to close the current version of an object by using \eqref{def:close-a-crisp-interval-r} and \eqref{eq:current-in-relation}
+Now it is possible to close the current version of an entity by using \eqref{eq:close-a-crisp-interval} and \eqref{eq:current-in-relation}.
+This functionality is required to add or update new information about an existing entity in the relation.
 
@@
 \label{def:close-current-version}
-Consider an object $A$, the relation $R$ and a time interval $J$.
-Then, the function close-current$\left(R,A,J \right)$ closes any current version of the object $A$ if it exists and add the new version $\left(A, J \right)$.
+\emph{Close-current$\left(r, t\right)$}.
+Consider the elements in definition \ref{def:valid-time-relation}.
+The function Close-current$\left(r, t \right)$ closes any current version $t_k$ of the entity given by $t$ if it exists and add the new version $t$.
+For the implementation the following variables are defined:
+\begin{align}
+ \label{eq:variables-current-version}
+t_{CUR} = \mbox{ Current } \left(t[A_K] \right)\\
+\nonumber
+t_{UP} = t_{CUR}; t_{UP}\left[i \right] = \mbox{Close }\left(t_{UP}\left[i \right], t \right)
+\nonumber
+\end{align}
+Then, $t_{CUR}$ is the current version of the entity given by the tuple $t$, and $t_{UP}$ is the updated version of the tuple $t_{CUR}$.
+In this updated version, the time interval given by $i_{UP}$ is closed with respect to the tuple $t$.
 
@@
 \label{eq:close-current}
-\text{Close-current} \left(R, A, J \right) =\\
+\text{Close-current} \left(r, t \right) =\\
 \begin{cases}
 \nonumber
-\big \lbrace R - \left(A, I \right) \cup \left \lbrace \left(A, \mbox{ CloseR } \left(I, J\right) \right) \cup \left(A, J\right)\right \rbrace  \big \rbrace \\
-\nonumber
-\mbox{ if } I = \mbox{ Current } \left(R, A \right) \wedge J > \mbox{ CloseR } \left(I, J \right)   \\
-\nonumber R , \text{ in any other case}
+r - t_{CUR} \cup \lbrace t_{UP}\rbrace \cup \lbrace t \rbrace
+& \mbox{ if } t_{CUR} \neq \emptyset \\
+% R - \left(A, I \right) \cup \left \lbrace \left(A, \mbox{ CloseR } \left(I, J\right) \right)\right \rbrace \cup \left \lbrace\left(A, J\right) \right \rbrace  \\
+% \nonumber
+% \mbox{ if  Current } \left(R, A \right) \neq \emptyset\\ %\wedge J > \mbox{ CloseR } \left(I, J \right)   \\
+\nonumber r , \text{ otherwise}
 \end{cases}
@@
 \end{definition}
+
+For example, consider the patient with ID = 3 in Table \ref{tbl:library-sample}.
+The current version arrived on 4/4/2012.
+The patient left the hospital but, for some reason, the date  was not registered.
+The patient arrived again at the hospital on 24/4/2012.
+Then, the Close-current function allows to close the current version of the patient and insert a new version.
+First, the function CloseR is applied with $I = \left(4/4/2012, UC \right)$ and $J = \left(24/4/2012, UC \right)$.
+Hence, the value for the time interval $I$ is $ \left(4/4/2012, 23/4/2012 \right)$.
+Then, the modifications on the value for the time interval $I$ are stored.
+Finally, a new row with the current values of the patient and the time interval $J$ is stored.
+The result of this operation is illustrated on Table \ref{tbl:library-sample-2}.
 
@@
 %\end{definition}
+\subsubsection{\label{subsubsec:modify}Modify}
+This operation adds new information about an existing entity $A$ in the relation $R$.
+The modify operation does not remove any previous value of the object $A$.
+It closes the current version for the entity $A$ and adds a new version.
+
+%Note that the modify operation is only applicable when the object $A$ is current in the relation $R$; $ \left(A, I \right) \in R, I = \left[a, +\infty \right]$.
+
+\begin{definition}
+ \emph{Modify(R, (A,I))}.
+Consider the elements in definition \ref{def:valid-time-relation}. The algorithm for the modify operation is defined as follows.
+\end{definition}
+
+\begin{align}
+\label{eq:modify}
+\MOD \left( r, t\right) =\\
+%\begin{cases}
+\nonumber
+\mbox{ Close-current }\left(r, t\right) %\\
+% \mbox{ if Current}\left(R ,A \right) \neq \emptyset \\
+% R,  \mbox{ otherwise }
+%\end{cases}
+\end{align}
 
 \subsubsection{\label{subsubsec:insert}Insert}
-The user wants to store an object $A$ which is valid in the relation $R$ during the time interval specified by $I = \left(a, b \right)$.
+The user wants to store an entity $A$ which is valid in the relation $R$ during the time interval specified by $I = \left(a, b \right)$.
 %
@@
 \begin{enumerate}
-\item The object $A$ was never in the relation $R$: The object is added with the valid-time indicated by the crisp interval $I$.
+\item The entity $A$ was never in the relation $R$: The entity is added with the valid-time indicated by the crisp interval $I$. 
+% In the example, this correspond with the following insertion sentences:
+% \begin{verbatim}
+% Insert(3,'Dracula',
+% (15/3/2012, 30/3/2012));
+% Insert(4,'Frankenstein',
+% (25/3/2012, 4/4/2012));
+% Insert(5,'Harry Potter',
+% (18/3/2012, 2/4/2012));
+% \end{verbatim}
 
-\item The object $A$ is in the relation $R$. Depending on the value of the time interval $I$, there are two possibilities:
+\item The entity $A$ is in the relation $R$. Depending on the value of the time interval $I$, there are three possibilities:
 	\begin{enumerate}
-	\item Insert $\left(A, I\right)$ in the relation $R$. If the time interval $I$ does not overlap any other valid-time interval in the relation $R$.
-	\item Reject the insertion, if the time interval $I$ do overlap any existing valid-time interval for the object $A$ in the relation $R$.
+	\item Insert $\left(A, I\right)$ in the relation $R$.
+If the time interval $I$ does not overlap any other valid-time interval in the relation $R$ for the entity $A$.
+For example, consider that the book with ID = 3 is loaned on 4/4/2012 but the book has not been returned yet.
+This is illustrated in Table \ref{tbl:library-sample}.
+%  The insert sentence is:
+% 	      \begin{verbatim}
+% Insert(3,'Dracula',
+% (4/4/2012, UC));
+% 	      \end{verbatim}
+
+	\item Modify and close the current version of $A$ and insert $\left(A, I\right)$.
+For example, consider now that the patient with ID = 3 arrived to the hospital on 24/4/2012.
+Here the problem is that the patient with ID = 3 arrived on 4/4/2012 but, for some reason, the departure date was not stored.
+If the patient is already at the hospital, then it is necessary to set the departure date and add a new row with the new arrival date.
+This is illustrated in Table \ref{tbl:library-sample-2}.
+%  the insert sentence is:
+% 	    \begin{verbatim}
+% Insert(3,'Dracula',
+% (24/4/2012, UC));
+% 	      \end{verbatim}
+	\item Reject the insertion, if the time interval $I$ does overlap any existing valid-time interval for the entity $A$ in the relation $R$.
+For example, consider that the hospital manager wants to introduce a past arrival for the patient with ID = 3.
+The patient arrived on 6/4/2012 and left the hospital on 25/4/2012.
+As the dates do overlap, it is not possible that the patient was in the hospital at that time interval.
+Therefore, the insertion is rejected.
+%  The insert sentence is:
+% 
+% 	      \begin{verbatim}
+% Insert(3,'Dracula',
+% (6/4/12/2012,25/4/2012));
+% 	    \end{verbatim}
+
 	\end{enumerate}
@@
 \end{enumerate}
+\begin{samepage}
+\vglue13pt
+%\begin{table}[htbp]
+\tcap{\label{tbl:library-sample-2}Sample hospital database}
+%\centerline{\small DATA TYPES}
+\vglue-6pt
+\centerline{\small\baselineskip=13pt
+\begin{tabular}{c c c c}\\
+\hline
+\textbf{ID}  & Name  & Arrival & Leave \\
+\hline
+3 & John & 15/3/2012 & 30/3/2012 \\
+4 & Antoon & 25/3/2012 & 4/4/2012 \\
+5 & Daan Potter & 18/3/2012 & 2/4/2012\\
+3 & John & 4/4/2012 & 23/4/2012 \\
+3 & John & 24/4/2012 & UC \\
+\hline\\
+\end{tabular}
+}
+\end{samepage}
 
-The algorithm for the implementation of the insert operation is the following:
+\begin{definition}
+ \emph{Insert(r, t)}.
+Consider the elements in definition \ref{def:valid-time-relation}. Then, the algorithm for the implementation of the insert operation is defined as follows.
+\end{definition}
 
@@
 \label{eq:insert}
-\INS \left(R ,A, I\right) =\\
+\INS \left(r, t\right) =\\
 \begin{cases}
 \nonumber
-R \cup \left(A, I \right), & \mbox{ if }  A \not \in R\\
-R \cup   \left(A, I \right) & \mbox{ if }  A \in R \wedge \\
-& \big \lbrace \left( \exists J: \left(A, J \right) \in R \wedge J<I  \right) \vee \\
-& \left( \exists K: \left(A, K \right) \in R \wedge K>I \vee \right) \wedge\\
-&  J<I<K  \big \rbrace\\
-R, & \mbox{ otherwise }  
+r \cup \lbrace t \rbrace
+& \mbox{ if }  t \not \in r \mbox{ or }\\
+\nonumber & \forall t_k \in Vt, \lnot \left(i \mbox{ Overlaps } i_k\right)\\
+% 
+% R \cup \left(A, I \right) & \mbox{ if }  (A, I ) \not \in R  \mbox{ or } \\
+%  &   \left(A, J\right) \in R \mbox{ and } \lnot \left(I \mbox{ Overlaps } J\right) \\
+\mbox{modify(r, t)} & \mbox{ otherwise }    %\mbox{ if } \exists J \left(A,J\right) \in R \wedge \mbox{Current(A,J)} \\ 
+%R, &   
 \end{cases} 	
@@
 \subsubsection{\label{subsubsec:del}Delete}
-The delete operation logically removes a current object $A$ which is valid in the relation $R$:
+The delete operation logically removes a current entity $A$ which is valid in the relation $R$.
+\begin{definition}
+ \emph{Delete(r, t)}.
+Consider the elements in definition \ref{def:valid-time-relation}.The algorithm for the delete operation is defined as follows.
+% 
+% Let $R \subseteq \A\  x\  \I$ be a valid-time relation and  $A=\left(A_1, \ldots, A_n \right), A \in \A$ be the set of attributes that define an entity and $I = \left(S_i,E_i\right)$ the set of attributes defining the time interval.
+The algorithm for the delete operation is defined as follows.
+\end{definition}
+
 \begin{align}
 \label{eq:delete}
-\DEL \left(R ,A \right) =\\
+\DEL \left(r, t\right) =
 \begin{cases}
 \nonumber
-R - \left(A, I \right), & \mbox{ if } \left( A, I \right) \in R\\
-R, & \mbox{ otherwise }  
+r - Vt
+% 
+% 
+% R - \left \lbrace \left(A, I_1 \right),  \ldots, \left(A, I_n \right)   \right \rbrace, & \forall I_i : \left( A, I_i \right) \in R\\
+% R, & \mbox{ otherwise }  
 \end{cases} 	
 \end{align}
+The set $Vt$ is computed as explained in definition \ref{def:valid-time-relation} and contains all the versions for the tuple $t$.
 
-\subsubsection{\label{subsubsec:modify}Modify}
-This operation adds new information about an existing object $A$ in the relation $R$. The modify operation does not remove any previous value of the object $A$.
-Note that the modify operation is only applicable when the object $A$ is current in the relation $R$; $ \left(A, I \right) \in R, I = \left[a, +\infty \right]$.
+For example, consider that the hospital manager wants to delete the history for the patient with ID = 3. The result of this operation is shown in Table \ref{tbl:library-sample-3}.
 
-\begin{align}
-\label{eq:modify}
-\MOD \left( R, A, I \right) =\\
-\begin{cases}
-\nonumber
-\mbox{ Close-current }\left(R, A, \mbox{Current}\left(R, A \right) \right) \\
- \mbox{ if Current}\left(R ,A \right) \neq \emptyset \\
- R,  \mbox{ otherwise }
-\end{cases}
-\end{align}
+% The delete sentence is:
+% 
+% \begin{verbatim}
+%  Delete(3,'Dracula');
+% \end{verbatim}
 
-\subsubsection{\label{subsubsec:revise}Revise}
-This operation changes the value of the object $A$ in the relation $R$, at a given time interval $I$. This operation is defined when a correction on the values for the object $A$ should be done.
-\begin{align}
-\label{eq:revise}
-\mbox{ Revise}\left(R, A, I \right)=\\
-\begin{cases}
-\nonumber
- R - \left(A', I \right) \cup \left(A, I \right)  \\
-  \mbox{ if } \left(A', I \right) \in R: A = \left(a_1, \ldots, a_n \right),\\
-A'= \left(a_1', \ldots, a_n' \right)   \\   
-R,  \mbox{ otherwise }
-\end{cases}
-\end{align}
+\begin{samepage}
+\vglue13pt
+%\begin{table}[htbp]
+\tcap{\label{tbl:library-sample-3}Sample hospital database}
+%\centerline{\small DATA TYPES}
+\vglue-6pt
+\centerline{\small\baselineskip=13pt
+\begin{tabular}{c c c c}\\
+\hline
+\textbf{ID}  & Name  & Arrival & Leave\\
+\hline
+4 & Antoon & 25/3/2012 & 4/4/2012 \\
+5 & Daan & 18/3/2012 & 2/4/2012\\
+\hline\\
+\end{tabular}
+}
+\end{samepage}
+
+% \subsubsection{\label{subsubsec:revise}Revise}
+% This operation replaces the old values for the entity $A'= \left(a_1', \ldots, a_n' \right) $ with the new values specified by $A = \left(a_1, \ldots, a_n \right)$ in the relation $R$, at a given time interval $I$.
+This operation is used when a correction on the values for the entity $A$ should be done.
+% \begin{align}
+% \REV\left(R, \left(A, I \right) \right)=\\
+% \begin{cases}
+% \nonumber
+%  R - \left(A', I \right) \cup \left \lbrace \left(A, I \right) \right \rbrace & \mbox{ if } \left(A', I \right) \in R  \\
+%    %: A = \left(a_1, \ldots, a_n \right),\\
+% %A'= \left(a_1', \ldots, a_n' \right)   \\   
+% R & \mbox{ otherwise }
+% \end{cases}
+% \end{align} 
+% 
+% For example, consider now that the librarian wants to make a correction in the title for the book with ID = 5.
+The new title is `Harry Potter IV'.
+The result of this operation is illustrated in Table \ref{tbl:library-sample-3}.
+%  Then, the revise sentence is:
+% 
+% \begin{verbatim}
+% Revise(5,'Harry Potter IV', 
+% (18/3/2012, 2/4/2012));
+% \end{verbatim}

### 2012-06-23 trunk/ijcis-pos-model/sources/intro.tex
more modifications

@@
 A temporal database schema is a database schema that models objects with time-related or time-variant properties.
-However, the modelling of temporal aspects has a direct impact on the consistency of the temporal database, because the temporal nature of these aspects imposes extra integrity constraints and suitable ways of interactive with the human user.
+However, the modelling of temporal aspects has a direct impact on the consistency of the temporal database, because the temporal nature of these aspects imposes extra integrity constraints and suitable ways of interaction with the human user.
 
-For example, let us consider a hospital database containing data about the state of the patients in a given hospital.
+% For example, let us consider an employee's database containing data about the positions of employees in a company.
+Two dates are stored: the starting and ending dates respectively of the contract.
+It is clear that an employee cannot have a contract with an ending date without an starting date.
+Without further cautions, a manager in the company could insert a new contract for an employee who is already working for the company.
+A temporal database model will typically constrain record insertion and prevent similar modelling inconsistencies.
+
+For example, let us consider a library database, modelling the presence of books in the library.
+Two dates are stored: the loan and the return date.
+It is clear that a book cannot be loaned again until it is returned.
+Without further cautions, a library employee could loan the same book several times even if it is not returned.
+A temporal database model will typically constrain record insertion and prevent similar modelling inconsistencies.
+
+% let us consider a hospital database containing data about the state of the patients in a given hospital.
 Two dates are stored: the date when the patient arrives to and leaves from the hospital.
@@
 Without further cautions, a hospital employee could insert a patient who is already in the hospital.
-A temporal database model will typically constrain record insertion and prevent similar modelling inconsistencies.
-
 A lot of research concerns temporal database models and their approaches to the modelling of time.
@@
 and 31/12/1300 A.D.'.
-These sentence contains imperfection because of the uncertainty in the used time-related expression.
+This sentence contains imperfection because of the uncertainty in the used time-related expression.
 It is known that the building was finished on a single day, but this day is not precisely known.

### 2012-06-23 trunk/ijcis-pos-model/sources/preliminaries.tex
more modifications

@@
 \begin{example}
-As an example, consider an employee's database.
-When an employee changes the status in the company, a record tracking the employee is stored.
-The exact time for this is not precisely known.
-Consider now that $X = \left[12, 2, 2\right]$ is the day of the month when an employee started a contract as a junior developer and $Y = \left[18, 2, 1 \right]$ is the day of the month when the employee changed the status to senior developer.
-It is possible to obtain $Z$ (the period of time between the junior and the senior contracts) by using a conjunctive combination of constraints.
+As an example, consider a historical database containing data about diplomatic medieval documents.
+The starting and ending dates when a diplomatic document was valid, are not precisely known.
+Consider now that the time granularity are years.
+Then $X = \left[1112, 2, 2\right]$ is the starting year for the validity of a document.
+A new diplomatic document was valid in the year $Y = \left[1118, 2, 1 \right]$.
+Then, It is possible to obtain $Z$ (the period of time between the starting of both documents) by using a conjunctive combination of constraints.
 
@@
 \nonumber
-CC = \left \lbrace C_1\left(>,X\right) \wedge C_2(\leq,Y) \right \rbrace \\
+CC &= \left \lbrace C_1\left(>,X\right) \wedge C_2(\leq,Y) \right \rbrace \\
 \nonumber
-Z = CC
+Z &= CC
 \end{align}
-$Z$ is a fuzzy interval defined by a trapezoidal shape given by $\left[12,\ 14,\ 18,\ 19 \right]$.
+$Z$ is a fuzzy interval defined by a trapezoidal shape given by $\left[1112,\ 1114,\ 1118,\ 1119 \right]$.
 Figure \ref{fig:example-ill-known-by-const} illustrates the relations among the variables $X$, $Y$ and $Z$.

### 2012-06-23 trunk/ijcis-pos-model/sources/temporalModel.tex
more modifications

@@
 
-In this section we will formalize the model for valid-time relational databases. The first subsection is devoted to the formalization of the model. Then, the data manipulation language is defined.
+In this section we will formalize the model for possibilistic valid-time relational databases.
+The first subsection is devoted to the formalization of the model.
+Then, the data manipulation language is defined.
 
@@
 The model is based on GEFRED~\cite{Medina1994} (Generalized Model of Fuzzy Relational DB) model.
-This model is extended by adding valid-time support which will be illustrated with examples.
+This model is extended by adding valid-time support which will be illustrated through definitions and examples.
 The information in the system is defined by the following elements:
@@
 \label{def:typeof-domain}
-\textbf{Typeof(A).}
-Consider $D_G$ to be a generalized fuzzy domain and let $A \in D_G$ be a value for the domain. 
-The function typeof$\left(A \right)$ returns the datatype associated with the domain $D_G$.
-For a given value $A$, the function returns the associated number as shown in Table \ref{tbl:gefred-data-types}.
+\textbf{Typeof(a).}
+Consider $D_G$ to be a generalized fuzzy domain and the elements in definition \ref{def:valid-time-relation}.
+Let $a$ be the value for the attribute $A$.
+The function typeof$\left(a \right)$ returns the datatype associated with the value $a$ and returns a number in $\left[1, 10 \right]$ as shown in Table \ref{tbl:gefred-data-types}.
+\begin{equation}
+ \label{eq:typeof}
+\mbox{Typeof} \left(a \right) \mapsto \left[1, 10 \right]
+\end{equation}
+
 \end{definition}
 
+\begin{samepage}
 \vglue13pt
@@
 
+A generalized fuzzy relation is defined in \cite{Medina1994}. Here, we will extend the definition to a generalized fuzzy temporal relation.
+
 \begin{definition}
 \emph{Generalized fuzzy temporal relation.}
-A generalized fuzzy temporal relation $R_{FTG}$ is given by~\cite{Medina1994}:
+\label{def:fuzzy-temporal-relation}
+Consider the elements in definition \ref{def:valid-time-relation}. Some of the elements will be extended for the fuzzy case.
+
+\begin{itemize}
+ \item A set of non-temporal fuzzy  or crisp attributes.
+	\begin{equation}
+	\label{eq:fuzzy-attribute-set}
+	A = \left \lbrace A_1, A_2, \ldots, A_n \right \rbrace
+	\end{equation}
+      The domain for each attribute $A_1, \ldots, A_n$ is $D_1, \ldots, D_n$ respectively. 
+\item The primary key $A_K$ is a subset of $A$.
+      \begin{equation}
+       \label{eq:fuzzy-primary-key-a}
+      A_K \subset A
+      \end{equation}
+A formal definition of primary key for fuzzy relational databases will be given later in Definition \ref{def:generalized-primary-key}.
+\item A set of two attributes; $S$  and $E$ the attributes for the starting and ending ill-known points respectively.
+$I$ is a possibilistic validity period \emph{PVP} as explained in Section \ref{subsec:ill-known-interval}.
+\begin{equation}
+ \label{eq:fuzzy-attribute-time-interval}
+I = \left \lbrace S, E \right \rbrace
+\end{equation}
+\item An attribute called version identifier, $V_{ID}$. This attribute is a counter for each different version of the entities. 
+% \begin{equation}
+%  \label{eq:fuzzy-version-identifier}
+% V_{ID} \subset \N
+% \end{equation}
+
+\item Then $R_{FTG}$, the schema for the fuzzy valid-time relation is:
+\begin{equation}
+ \label{eq:fuzzy-valid-time-relation}
+R_{FTG} = A \cup V_{ID} \cup  I
+\end{equation}
+\item The primary key for the fuzzy valid-time relation $R_{FTG}$ is:
+\begin{equation}
+ \label{eq:fuzzy-valid-time-temporal-pk}
+K_{GT} = A_K \cup V_{ID}
+\end{equation}
+A formal definition of the primary key for fuzzy valid-time relations will be given later in Definition \ref{def:generalized-fuzzy-temporal-key}.
+
+\item We will note by $r$ any valid instance of $R_{FTG}$. 
+      \begin{equation}
+       \label{eq:fuzzy-valid-time-instance}
+      r \subseteq D_1\ x\ \ldots\ x\ D_n
+      \end{equation}
+
+\item Let $t$ be a tuple in the instance $r$; $t \in r$:
+      \begin{itemize}
+      \item We will note the values for the starting and the ending  points,$s$ and $e$  respectively. The value for the time interval is given by $i$.
+      \begin{align}
+       \label{eq:fuzzy-starting-point}
+      s = t\left[S \right]\\
+      e = t\left[E \right]\\
+      i = \left(s, e\right)
+      \end{align}
+
+      \item Let $A_k \subset A$ be the set of the non-temporal attributes that are part of the primary key (equation \eqref{eq:fuzzy-primary-key-a}).
+Then, $a_k$ denotes the values for the non-temporal attributes of the primary key.
+	    \begin{equation}
+	     \label{eq:fuzzy-pk-attribute} 
+	      a_k = t\left[A_K \right]
+	    \end{equation}
+
+    \item Let $K_{GT}$ be the primary key for the valid-time relation as given in equation \eqref{eq:fuzzy-valid-time-temporal-pk}. Then, $k$ denotes the values for the attributes in the primary key.
+	  \begin{equation}
+	   \label{eq:fuzzy-value-pk}
+	  k = t\left[K_{GT} \right]
+	  \end{equation}
+
+      \end{itemize}
+\item $Vt$ is the set with all the versions for a given tuple $t$.
+
+\begin{equation}
+ \label{eq:fuzzy-all-the-versions}
+Vt = r\left(t\left[A_K\right] \right)
+\end{equation}
+If a tuple $t$ contains more than one version, the result of $Vt$ is a set.
+We will use the index $k$ to address the elements of the set.
+E.g.
+$Vt = \left \lbrace t_1, t_2, \ldots, t_n \right \rbrace$.
+Then $t_k , k \in \left \lbrace 1, \ldots, n \right \rbrace$ is an element of $Vt$.
+\end{itemize}
+%\end{definition}
+
+We will illustrate the definitions with an example.
+
+\begin{example}
+Consider the set of attributes $A = \left \lbrace A_1, A_2, A_3 \right \rbrace$.
+The primary key for these attributes is given by $A_K = \left \lbrace A_1, A_2 \right \rbrace$.
+Let $I = \left \lbrace S, E \right \rbrace$ be the set of temporal attributes that define the possibilistic validity period of the data.
+$R_{FTG} = A \cup V_{ID} \cup I$ is the valid-time relation and $r$ is an instance of the relation.
+The instance $r$ is given by the following elements.
+$r = \left \lbrace \left(a_{11}, a_{12}, a_{13}, 001, s_1 ,e_1 \right) \right.$,  $\left(a_{21}, a_{22}, a_{23}, 001, s_2, e_2 \right)$ , $\left.
+\left(a_{11}, a_{12}, a_{31}, 002, s_3, e_3 \right) \right \rbrace$.
+The instance $r$ is illustrated in Table \ref{tbl:fuzzy-sample-definitions}.
+Consider the tuple $t = \left(a_{11}, a_{12}, a_{13}, s_1 ,e_1 \right)$. Then,
+\begin{align}
+ \nonumber
+s_1 &= t\left[S \right]\\
+ \nonumber
+e_1 &= t[E]\\
+ \nonumber
+i\ \ &= \left(s_1, e_1\right)\\
+ \nonumber
+a_k &= t\left[A_K\right] = \left(a_{11}, a_{12} \right) \\
+ \nonumber
+k &= t\left[PK\right] =\left(a_{11}, a_{12}, s_1, e_1\right)\\
+ \nonumber
+Vt &= r\left(t\left[A_K\right] \right) = \left \lbrace t_1, t_3 \right \rbrace
+\end{align}
+
+\end{example}
+
+\begin{samepage}
+\vglue13pt
+%\begin{table}[htbp]
+\tcap{\label{tbl:fuzzy-sample-definitions}Sample database containing the instance $r$ of the fuzzy valid-time relation $R_{FTG}$.}
+%\centerline{\small DATA TYPES}
+\vglue-6pt
+\centerline{\small\baselineskip=13pt
+\begin{tabular}{c c c c c c c}\\
+\hline
+& \textbf{A$_1$}  & \textbf{A$_2$}  & $A_3$ & $V_{ID}$ & $S$ & $E$ \\
+\hline
+$t_1$&$a_{11}$ & $a_{12}$ & $a_{13}$ & $001$ & $s_1$ & $e_1$ \\
+$t_2$ & $a_{21}$ & $a_{22}$ & $a_{23}$& $001$ &  $s_2$ & $e_2$ \\
+$t_3$ & $a_{11}$ & $a_{12}$ & $a_{31}$& $002$ & $s_3$ & $e_3$\\
+\hline\\
+\end{tabular}
+}
+\end{samepage}
+
+% 
+% 
+% Consider $\A$ to be the set of all the entities, and let $A=\left(A_1, \ldots, A_n \right), A \in \A$ be the set of (fuzzy) attributes that define an entity.
+Let $\I_{PVP}$ be the set of all the ill-known time intervals and let $I = \left(X, Y \right)$ be a PVP, $I \in \I_{PVP}$.
+The pair $\left(A, I\right)$ expresses that the data regarding the entity A are valid during the ill-known time interval I.
+Let $R_{FTG} \subseteq \A \  x\  \I_{PVP}$ be a valid-time relation.
+Then, the following equation indicates that the pair $\left(A, I\right)$ is in the relation R:
+% 
+% \begin{equation}
+% \label{eq:fuzzy-rel-def}
+% \left( R_{FTG}, \left(A , I\right) \right)
+% \end{equation}
+% 
+ A generalized fuzzy temporal relation $R_{FTG}$ can be noted also by:
 \label{def:generalized-fuzzy-temporal-relation}
@@
 \label{ex:sample-ft-database}
- Consider a hospital database as explained in example \ref{ex:library-database}.
-But now, the time is stored with imprecision, because the date of arrival and leave are not precisely known.
+ Consider a historical database containing diplomatic documents as explained in example \ref{ex:library-database}.
+But now, the documents are from the Medieval Ages.
+Hence, the time is known with imprecision.
 For simplicity, the ill-known time points are represented as triangular fuzzy numbers.
@@
 The values for $a$ and $b$ are integers.
-Thus, the date given by $[15/3/2012,5,2]$ is a triangular fuzzy number with the left bound the 10/3/2012, the core on the 15/3/2012 and the right bound on 17/3/2012.
+Thus, the date given by $[15/3/1012,5,2]$ is a triangular fuzzy number with the left bound the 10/3/1012, the core on the 15/3/1012 and the right bound on 17/3/1012.
 Table \ref{tbl:fuzzy-library-sample} shows the elements Head, $\Head$ and Body, $\Body$. 
@@
 \nonumber
-&  \left(\text{Name}:D_{\text{Name}}\right),\\
+&  \left(\text{Entity}:D_{\text{Entity}}\right),\\
 % \nonumber
@@
 
+\begin{samepage}
 \vglue13pt
 %\begin{table}[htbp]
-\tcap{\label{tbl:fuzzy-library-sample}Sample hospital database}
+\tcap{\label{tbl:fuzzy-library-sample}Sample historical database}
 %\centerline{\small DATA TYPES}
@@
 & & & \centerline{PVP} \\
-$\Head$ &  \textbf{ID}  & Name  &  Arrival , Leave\\
+$\Head$ &  \textbf{ID}  & Entity  &  Start , End \\
 \hline
-\multirow{3}{*}{$\Body$ } &  3 & John & [15/3/2012, 5, 2] , [30/3/2012, 1, 1] \\
- &4 & Antoon & [25/3/2012, 3, 2] , [4/4/2012, 1, 7] \\
-&5 & Daan & [18/3/2012, 4, 1] , [2/4/2012, 2, 2]\\
-% \ 3 & Dracula & [4/4/2012, 3, 3] & UC \\
+\multirow{3}{*}{$\Body$ } &  3 & E.U. & [15/3/1012, 5, 2] , [30/3/1012, 1, 1] \\
+ &4 & N.A.T.O. & [25/3/1012, 3, 2] , [4/4/1012, 1, 7] \\
+&5 & C.E.I. & [18/3/1012, 4, 1] , [2/4/1012, 2, 2]\\
+% \ 3 & Dracula & [4/4/1012, 3, 3] & UC \\
 \hline\\
@@
 
-For example, in the case of the record with ID = 3:
+For example, in the case of the document with ID = 3:
 \begin{align}
@@
 \nonumber
-\Head^{v} &=  \lbrace \left(\text{ID}:D_{\text{ID}}\right), \left(\text{Title}:D_{\text{title}}\right),\\
+\Head^{v} &=  \lbrace \left(\text{ID}:D_{\text{ID}}\right), \left(\text{Entity}:D_{\text{entity}}\right),\\
 \nonumber
@@
 \nonumber
-\Body^ {v} &= \left \lbrace 3:, \text{``John``}: ,  \left[15/3/2012, 30/3/2012\right] \right \rbrace 
+\Body^ {v} &= \left \lbrace 3:, \text{``E.U.''}: ,  \left[15/3/1012, 30/3/1012\right] \right \rbrace 
 \end{align}
@@
 \nonumber
-\Head^{c} = & \lbrace \left(C_\text{ID}\right), \left(C_{\text{title}}\right), \left(C_{\text{Loan}}\right), \left(C_{\text{PVP}}\right) \rbrace \\
+\Head^{c} = & \lbrace \left(C_\text{ID}\right), \left(C_{\text{entity}}\right), \left(C_{\text{PVP}}\right) \rbrace \\
 % \nonumber
@@
 \nonumber
-\Body^ {t} =& \left \lbrace 4/4/2012 , UC \right \rbrace 
+\Body^ {t} =& \left \lbrace 4/4/1012 , UC \right \rbrace 
 \end{align}
 
-Analogously, it is possible to define the the value component for the temporal part and the compatibility component for the temporal part.
+Analogously, it is possible to define both the name value component and the compatibility component for the temporal part.
+
+%the value component for the temporal part and the compatibility component for the temporal part.
 
@@
 \emph{Generalized primary key.}
-Consider $D_G$ to be a fuzzy generalized domain, and let $A_{Gs}:D_{Gs}$ be the attributes and the domain of the attribute for each $s in S \subseteq \left(1, \ldots, n \right)$.
+Consider $D_G$ to be a fuzzy generalized domain, and let $A_{Gs}:D_{Gs}$ be the attributes and the domain of the attribute for each $s \in S \subseteq \left(1, \ldots, n \right)$.
 A generalized primary key, $K_G$ is a subset of the head:
@@
 The primary key for the table is the attribute \emph{ID}, a unique number.
-Two different books have two different values for the \emph{ID} attribute.
+Two different documents have two different values for the \emph{ID} attribute.
 
 In order to add valid-time support, the primary key should be re-defined.
-E.g., consider the library database.
-If the primary key is the \emph{ID} attribute, a book could not be loaned more than once.
+E.g., consider the historical database.
+If the primary key is the \emph{ID} attribute, a document should be valid only during one period of time.
 To resolve this problem, we extend the given primary key with a version identifier.
+
+%a book could not be loaned more than once. 
 %Thus, when a book is loaned, a new entry in the table is added. The \emph{ID} attribute identifies the physical book whereas
@@
 \emph{Generalized fuzzy temporal key.}
-Consider $D_G$ to be a fuzzy generalized domain, and let $A_{Gs}:D_{Gs}$ be the attributes and the domain of the attribute for each $s in S \subseteq \left(1, \ldots, n \right)$.
-A generalized fuzzy temporal key, $K_{GT}$ is a subset of the head. An attribute, namely \emph{version} $V$ is added when dealing with valid-time.
+Consider $D_G$ to be a fuzzy generalized domain, and let $A_{Gs}:D_{Gs}$ be the attributes and the domain of the attribute for each $s \in S \subseteq \left(1, \ldots, n \right)$.
+Let  $V$ be a new attribute called \emph{version}.
+A generalized fuzzy temporal key, $K_{GT}$ is a subset of the head.
 \begin{align}
@@
 
+\begin{samepage}
 \vglue13pt
 %\begin{table}[htbp]
-\tcap{\label{tbl:fuzzy-library-sample-valid-time}Sample library database}
+\tcap{\label{tbl:fuzzy-library-sample-valid-time}Sample historical database}
 %\centerline{\small DATA TYPES}
@@
 %& & & & \centerline{PVP}\\
-{$\Head$}  &  &  & &(Arrival\\
-& \textbf{ID}  & \textbf{V} & Name& ,Leave )	\\
+{$\Head$}  &  &  & &(Start \\
+& \textbf{ID}  & \textbf{V} & Entity& ,End )	\\
  \hline
-\multirow{4}{*}{$\Body$}  & 3 & 001 & John & [15/3/2012, 5, 2], [30/3/2012, 1, 1] \\
-& 4 & 001 & Antoon & [25/3/2012, 3, 2] , [4/4/2012, 1, 7]\\
-& 5 & 001 & Daan & [18/3/2012, 4, 1] , [2/4/2012, 2, 2]\\
- & 3 & 002 & John & [4/4/2012, 3, 3] , UC \\
+\multirow{4}{*}{$\Body$}  & 3 & 001 & E.U. & [15/3/1012, 5, 2], [30/3/1012, 1, 1] \\
+& 4 & 001 & N.A.T.O. & [25/3/1012, 3, 2] , [4/4/1012, 1, 7]\\
+& 5 & 001 & C.E.I. & [18/3/1012, 4, 1] , [2/4/1012, 2, 2]\\
+ & 3 & 002 & E.U. & [4/4/1012, 3, 3] , UC \\
 \hline\\
@@
 
-The Generalized Fuzzy Relational Algebra~\cite{Medina1994} manipulates relations like $R_{FG}$.
+The Generalized Fuzzy Relational Algebra~\cite{Medina1994} manipulates relations like $R_{FTG}$.
 The operations defined are: \emph{Union, Intersection, Difference, Cartesian Product, Projection, Join} and \emph{Selection}.
 Thus, in this section we will describe and implement the following operations for temporal databases, as described in \ref{subsubsec:Understanding-valid-time-databases}.
-The operations implemented are: \emph{Insert, Modify, Delete, }.
+The operations implemented are: \emph{Insert, Modify} and \emph{Delete}.
 The semantics of the operations will be the same that those defined for a crisp temporal database, whereas the temporal representation is made by the possibilistic valid-time period and the ill-known constraints (see sections \ref{sec:prelim} and \ref{sec:time-rep}.
@@
 
-It is important to notice that while the result of the evaluation of any comparison between crisp time intervals is boolean, the evaluation of any comparison between \emph{PVPs} is in the unit interval.
-Therefore, for a valid-time object, say $A \in R$, in a crisp temporal database it is not possible an overlapping among any valid-time interval ($\forall I, \not \exists J: \left(A, I \right) \in R \wedge \left(A, J \right) \in R \wedge \lnot \Overlap\left(I, J \right)$).
-Thus, from the point of view of a fuzzy temporal database, there are the following options (see Figure \ref{fig:tdb-class}):
+It is important to notice that while the result of the evaluation of any comparison between crisp time intervals is boolean, the evaluation of any comparison between \emph{PVPs} is a value in the unit interval.
+Therefore, for a crisp valid-time relation, it is not possible an overlapping among any valid-time interval.
+In other words, a crisp temporal database is always strictly consistent.
+Thus, from the point of view of a fuzzy temporal database, there are the following choices (see Figure \ref{fig:tdb-class}):
 
@@
 \begin{definition}
-\textbf{Strictly consistent} It is not possible for two PVPs of the same object to be valid at the same time.
-In other words, if $A$ is an object, $R$ is a relation and $\pi_{I}\left(t \right)$ and $\pi_{J}\left(t \right)$ are the possibilities for a time point $t \in \T$ to be in the PVPs $I$ and $J$ respectively,
+\textbf{Strictly consistent} It is not possible that two versions of an entity to be valid at the same time.
+Consider a tuple $t$ in the instance $r$ of the fuzzy valid-time relation $R_{FTG}$ and  $Vt = \left \lbrace t_1, \ldots, t_n \right \rbrace$ as defined in \ref{def:fuzzy-temporal-relation}.
+Then, a fuzzy temporal database is strictly consistent if:
 \begin{align}
 \label{eq:stricly-consistent}
-\forall I,J, \forall t \in \T :\\
-%\nonumber
-%\left(A, I \right) \in R \wedge \\
-%\nonumber
-%\left(A, J \right) \in R : \\
-\nonumber
-\pi_{I}(t) + \pi_{J}(t) \leq 1
+\forall k,j \in \left \lbrace 1, \ldots, n \right \rbrace, k < j\\
+\nonumber t_k \mbox{ Overlaps } t_j \leq 1
 \end{align}
@@
 	      \begin{definition}
-	      \textbf{Co-existence} Two versions of the same object $A$ may exist at the same time.
+	      \textbf{Co-existence} Two versions of the same entity may exist at the same time.
 		      \begin{align}
-		      \label{eq:co-existence}
-		      \forall I, \exists ! J, \exists t \in \T : \\
-		      \nonumber
-		      \pi_{I}(t) + \pi_{J}(t) \geq 1
+		      \forall k, \exists !j \in \left \lbrace 1, \ldots, n \right \rbrace, k < j\\
+		      \nonumber t_k \mbox{ Overlaps } t_j \geq 1
 		      \end{align}
@@
 	      \begin{definition}
-	      \textbf{Weak-consistence} Several versions of the same object $A$ may exist at the same time.
+	      \textbf{Weak-consistence} Several versions of the same entity may exist at the same time.
 		      \begin{align}
 		      \label{eq:weak-consistence}
-		      \forall I, \exists J, \exists t \in \T : \\
-		      \nonumber
-		      \pi_{I}(t) + \pi_{J}(t) \geq 1
+		      \forall k, \exists j \in \left \lbrace 1, \ldots, n \right \rbrace, k < j\\
+		      \nonumber t_k \mbox{ Overlaps } t_j \geq 1
 		      \end{align}
@@
 
-Since the time intervals are now possibilistic valid-time periods, PVPs, the auxiliary functions defined in equations \eqref{eq:close-a-crisp-interval} to \eqref{eq:close-current} should be re-defined.
+Since the time intervals are now possibilistic valid-time periods, PVPs, the auxiliary functions defined in equations \eqref{eq:close-a-crisp-interval} to \eqref{eq:close-current} are the basis for the following auxiliary functions.
 
@@
 \label{def:close-r-a-pvp}
-\textbf{CloseR$\left(I, J\right)$.}
-Consider two PVPs given by $I = \left(X, Y \right)$ and $J = \left(X', Y' \right)$.
-The CloseR function closes the PVP given by $I$ with a convex combination of ill-known constraints (see section \ref{subsec:interval-evaluation-by-ill-known-constraints}):
+\textbf{CloseR$\left(i_1, i_2\right)$.}
+Consider the elements in definition \ref{def:generalized-fuzzy-temporal-relation}.
+The CloseR function closes the PVP given by $i_1$ with a conjunctive combination of ill-known constraints (see section \ref{subsec:interval-evaluation-by-ill-known-constraints}):
 \begin{align}
 \label{eq:close-r-a-pvp}
-\text{CloseR}\left(I, J\right) &= \\
+\text{CloseR}\left(i_1, i_2\right) &= \\
 \begin{cases}
 \nonumber
-I   & \mbox{ if } I \neq \left[X, UC \right] \\
-I = \left[X, Z \right] & \mbox{ if } I = \left[X, UC \right] \\
-& Z \triangleq \left \lbrace C_1\left(>, X \right), C_2\left(<, X' \right) \right \rbrace
+i_1 = \left(s_1, e_z \right) & \mbox{ if } i_1 = (s_1, UC), \\
+& e_z \triangleq \left \lbrace C_1\left(>, s_1 \right), C_2\left(<, s_2 \right) \right \rbrace \\
+i_1 & \mbox{ otherwise }
+% \nonumber
+% I   & \mbox{ if } I \neq \left[X, UC \right] \\
+% I = \left[X, Z \right] & \mbox{ if } I = \left[X, UC \right] \\
+% & Z \triangleq \left \lbrace C_1\left(>, X \right), C_2\left(<, X' \right) \right \rbrace
 \end{cases}
@@
 
-For example, consider $I = \left[\left[4/4/2012, 3, 3\right] , UC\right]$ and $J = \left[ \left[15/4/2012,2,1\right], UC\right]$.
-The result for CloseR$\left(I, J\right)$  is $I = \left[ \left[4/4/2012, 3, 3\right], \left[4,7,13,15 \right] \right]$.
-The value that closes $I$ is a trapezoid in the form $\left[ \alpha, \beta, \gamma, \delta \right]$ as defined in Section \ref{subsec:fuzzy-numbers}.
+For example, consider $i_1 = \left[\left[4/4/1012, 3, 3\right] , UC\right]$ and $i_2 = \left[ \left[15/4/1012,2,1\right], UC\right]$.
+The result for CloseR$\left(i_1, i_2\right)$  is $i_1 = \left[ \left[4/4/1012, 3, 3\right], \left[4,7,13,15 \right] \right]$.
+The value that closes $i_1$ is a trapezoid in the form $\left[ \alpha, \beta, \gamma, \delta \right]$ as defined in Section \ref{subsec:fuzzy-numbers}.
 
-\begin{definition}
-\label{def:pvp-current-in-relation}
-\textbf{Current$\left(R, A \right)$.}
-Consider an entity $A$ and a relation $R$ in a fuzzy temporal database.
-The function Current$\left(R, A \right)$ obtains the tuple $\left(A, I \right), I = \left(X, Y \right)$ that is current in the relation $R$, that is it, $I = \left[X , \text{UNKNOWN} \right]$.
-
-\begin{align}
-\label{eq:pvp-current-in-relation}
-\mbox{Current} \left(R, A \right) &=& \\ 
-\begin{cases}
-\nonumber
-I & \mbox{ if } I = \left[X, UC \right] \\
-%&  (A,I) \in R \wedge Y \in \text{UNKNOWN}\\
-\emptyset & \mbox{ in any other case }
-\end{cases}
-\end{align}
-\end{definition}
-
-For example, consider the database in Table \ref{tbl:fuzzy-library-sample-valid-time}.
-The function Current$\left(R, A \right)$ with $A =\left(ID=3\right)$ returns time interval for the current version of $A$, $I = \left[ \left[4/4/2012, 3, 3 \right], UC \right]$.
+% \begin{definition}
+% \label{def:pvp-current-in-relation}
+% \textbf{Current$\left(r, a_k \right)$.}
+% Consider the elements in definition \ref{def:generalized-fuzzy-temporal-relation}.
+The function Current $\left(r,a_k \right)$ returns the PVP $i$, if the tuple $t$ with the primary key values given by $a_k$ is currently in the instance $r$ of the relation $R$.
+The function returns the emtyset otherwise.
+% 
+% \begin{align}
+% \label{eq:pvp-current-in-relation}
+% \mbox{Current} \left(r, a_k\right) &=& \\ 
+% \begin{cases}
+% \nonumber
+% t & \mbox{ if } \exists t \in r : i = \left(s, UC \right) \mbox{ and }\\
+% &  a_k = t[A_K]\\
+% \emptyset & \mbox{ otherwise}
+% \end{cases}
+% \end{align}
+% \end{definition}
+% 
+% For example, consider the database in Table \ref{tbl:fuzzy-library-sample-valid-time}.
+The function Current$\left(R_{FTG}, A \right)$ with $A =\left(ID=3\right)$ returns time interval for the current version of $A$, $I = \left[ \left[4/4/1012, 3, 3 \right], UC \right]$.
 Nevertheless, the time interval for the entity with ID = 4 is the empty set.
-Now it is possible to close the current version of an object by using \eqref{eq:close-r-a-pvp} and \eqref{eq:pvp-current-in-relation}.
+% Now it is possible to close the current version of an entity by using \eqref{eq:close-r-a-pvp} and \eqref{eq:pvp-current-in-relation}.
 
@@
 \label{def:pvp-close-current-version}
-\textbf{Close-current$\left(R, \left(A, J \right) \right)$.}
-Consider an object $A$, the relation $R$ and a pvp, $J$.
-Then, the function close - current$\left(R, A, J \right)$ closes any current version of the object $A$ if it exists and add the new version $\left(A, J \right)$.
+\textbf{Close-current$\left(r, t \right)$.}
+Consider the elements in definition \ref{def:generalized-fuzzy-temporal-relation}.
+The function Close-current$\left(r, t \right)$ closes any current version $t_k$ of the entity given by $t$ if it exists and add the new version $t$.
+In order to implement the functionality, the variables in equation \eqref{eq:variables-current-version} are used and the function Current as given by equation \eqref{eq:current-in-relation}.
 
@@
 \label{eq:pvp-close-current}
-\text{Close-current} \left(R, \left(A, J \right) \right) =\\
+\text{Close-current} \left(r, t \right) =\\
 \begin{cases}
 \nonumber
-\big \lbrace R - \left(A, I \right) \cup \left \lbrace \left(A, \mbox{ CloseR } \left(I, J\right) \right) \right \rbrace \cup \left \lbrace \left(A, J\right)\right \rbrace  \big \rbrace \\
-\nonumber
-\mbox{ if Current } \left(R, A \right) \neq \emptyset \\ %\wedge J > \mbox{ CloseR } \left(I, J \right)   \\
-\nonumber R , \text{ in any other case}
+r - t_{CUR} \cup \lbrace t_{UP}\rbrace \cup \lbrace t \rbrace
+& \mbox{ if } t_{CUR} \neq \emptyset \\
+% R - \left(A, I \right) \cup \left \lbrace \left(A, \mbox{ CloseR } \left(I, J\right) \right)\right \rbrace \cup \left \lbrace\left(A, J\right) \right \rbrace  \\
+% \nonumber
+% \mbox{ if  Current } \left(R, A \right) \neq \emptyset\\ %\wedge J > \mbox{ CloseR } \left(I, J \right)   \\
+\nonumber r , \text{ otherwise}
 \end{cases}
@@
 
+% Then, the function close - current$\left(R_{FTG}, A, J \right)$ closes any current version of the entity $A$ if it exists and add the new version $\left(A, J \right)$.
+
+% \begin{eqnarray}
+% \label{eq:pvp-close-current}
+% \text{Close-current} \left(R_{FTG}, \left(A, J \right) \right) =\\
+% \begin{cases}
+% \nonumber
+% \big \lbrace R_{FTG} - \left(A, I \right) \cup \left \lbrace \left(A, \mbox{ CloseR } \left(I, J\right) \right) \right \rbrace \cup \left \lbrace \left(A, J\right)\right \rbrace  \big \rbrace \\
+% \nonumber
+% \mbox{ if Current } \left(R_{FTG}, A \right) \neq \emptyset \\ %\wedge J > \mbox{ CloseR } \left(I, J \right)   \\
+% \nonumber R , \text{ in any other case}
+% \end{cases}
+% \end{eqnarray}
+% \end{definition}
+
 For example, consider the database in Table  \ref{tbl:fuzzy-library-sample-valid-time}.
-The function Close-current1$\left(R, \left(ID=3\right) \right.,$ $\left.
-\left[ \left[15/4/2012,2,1\right], UC\right]\right)$ closes the current version of the book with ID=3 and creates a new version with the specified time interval.
+The function Close-current1$\left(R_{FTG}, \left(ID=3\right) \right.,$ $\left.
+\left[ \left[15/4/1012,2,1\right], UC\right]\right)$ closes the current version of for the patient with ID=3 and creates a new version with the specified time interval.
 
 \subsubsection{\label{subsubsec:modify-fuzzy-temporal}Modify}
-This operation adds new information about an existing entity $A$ in the relation $R$. The modify operation does not remove any previous value of the entity $A$.
-Note that the modify operation is only applicable when the object $A$ is current in the relation $R$; $ \left(A, I \right) \in R, I = \left[X, \UC \right]$.
+This operation adds new information about an existing entity (given by the tuple $t$) in the instance $r$ of the fuzzy temporal relation $R_{FTG}$.
+The modify operation does not remove any previous value of the entity.
+Note that the modify operation is only applicable when the entity is current in the relation; $ t \in r,  = \left(s, \UC \right)$.
+
+\begin{definition}
+\emph{$\MOD \left( r, t \right)$}. 
+Consider the elements in definition \ref{def:generalized-fuzzy-temporal-relation}. The algorithm for the modify operation is defined as follows.
+\end{definition}
 
@@
 \label{eq:modify-fuzzy-temporal}
-\MOD \left( R, \left( A, I \right) \right) =\\
+\MOD \left( r, t\right) =\\
 %\begin{cases}
 \nonumber
-\mbox{ Close-current }\left(R, \left(A, I\right) \right) 
+\mbox{ Close-current }\left(r, t \right) 
 % \mbox{ if Current}\left(R ,A \right) \neq \emptyset \\
@@
 \subsubsection{\label{subsubsec:insert-fuzzy-temporal}Insert}
-The user wants to store an entity $A$ which is valid in the relation $R$ during the time interval specified by the PVP, $I = \left(X, Y \right)$.
+The user wants to store an entity (given by the tuple $t$) which is valid in the instance $r$ of the fuzzy temporal relation $R_{FTG}$ during the time interval specified by the PVP, $i = \left(s, e \right)$.
 %
@@
 \begin{enumerate}
-\item The entity $A$ was never in the relation $R$: The entity is added with the valid-time indicated by the PVP, $I$.
-For example, consider the database given by Table \ref{tbl:fuzzy-library-sample-valid-time}. The following sentences correspond with the insertion of the first version of each book.
+\item The entity was never in the relation: The entity is added with the valid-time indicated by the PVP, $i$.
+For example, consider the database given by Table \ref{tbl:fuzzy-library-sample-valid-time}. The following sentences correspond with the insertion of the first validity period for each document.
 
 \begin{verbatim}
-Insert(3,'John',
-[[15/3/2012, 5, 2],
-        [30/3/2012, 1, 1]]);
-Insert(4,'Antoon',
-[[25/3/2012, 3, 2], 
-        [4/4/2012, 1, 7]]);
-Insert(5,'Daan',
-[[18/3/2012, 4, 1], 
-        [2/4/2012, 2, 2]]);
+Insert(3,'E.U.',
+[[15/3/1012, 5, 2],
+        [30/3/1012, 1, 1]]);
+Insert(4,'N.A.T.O.',
+[[25/3/1012, 3, 2], 
+        [4/4/1012, 1, 7]]);
+Insert(5,'C.E.I.',
+[[18/3/1012, 4, 1], 
+        [2/4/1012, 2, 2]]);
 \end{verbatim}
 
-\item The entity $A$ is in the relation $R$. Depending on the value of the time interval $I$, there are three possibilities:
+\item The entity is in the relation. Depending on the value of the time interval $i$, there are three possibilities:
 	\begin{enumerate}
-	\item Insert $\left(A, I\right)$ in the relation $R$.
-If the time interval $I$ does not overlap any other valid-time interval in the relation $R$.
+	\item Insert $t$ in the instance $r$ of the relation $R_{FTG}$.
+If the time interval $i$ does not overlap any other valid-time interval in the relation $R_{FTG}$.
 Note that here, the result of the Overlaps operator is in the unit interval.
-For example, the patient with ID=3 has not left the hospital yet. The insert sentence is the following.
+For example, the document with ID=3 is still valid. The insert sentence is the following.
 
  \begin{verbatim}
-Insert(3,'John',
-[[4/4/2012, 3, 3], UC]);
+Insert(3,'E.U.',
+[[4/4/1012, 3, 3], UC]);
 	      \end{verbatim}
 
-	\item Modify and close the current version of the entity $A$ and insert $\left(A, I \right)$.
-For example, consider now that the hospital with ID=3 arrived  around the 24/4/2012 (this is model by a triangular fuzzy number: $\left[24/4/2012, 1, 1 \right]$).
-Here the problem is that the patient with ID=3 arrived around 4/4/2012, but it was not stored in the database the leave date.
-If the patient is again in the hospital, then it is necessary to set the leave date and add a new row with the new arrival date.
+	\item Modify and close the current version of the entity and insert a new version.
+For example, consider now that the document with ID=3 was valid  around the 24/4/1012 (this is model by a triangular fuzzy number: $\left[24/4/1012, 1, 1 \right]$).
+Here the problem is that the document with ID=3 was valid around 4/4/1012, but, for some reason, the ending date was not stored.
+If the document is again valid, then it is necessary to set the ending date and add a new row with the new starting date.
 %This is illustrated in Table \ref{tbl:library-sample-2}, the insert sentence is:
 	    \begin{verbatim}
-Insert(3,'John',
-[[24/4/2012, 1, 1], UC]]);
+Insert(3,'E.U.',
+[[24/4/1012, 1, 1], UC]]);
 	      \end{verbatim}
 
-	\item Reject the insertion, if the time interval $I$ do overlap with a degree of 1 any existing valid-time interval for the entity $A$ in the relation $R$.
-For example, consider now that the hospital manager wants to introduce a past arrival for the patient with ID=3.
-The patient arrived around 6/4/12/2012 and left the hospital around 25/4/2012.
-As this interval does overlaps other time intervals with a degree of 1, it is not possible that the patient was it the hospital at that time interval.
+	\item Reject the insertion, if the time interval $i$ do overlap with a degree of 1 any existing valid-time interval for the entity in the relation.
+For example, consider now that the document manager wants to introduce a past validity period for the document with ID=3.
+The validity starting date for the document was around 6/4/12/1012 and the ending date was around 25/4/1012.
+As this interval does overlaps other time intervals with a degree of 1, it is not possible that the document was valid during two different periods of time.
 Therefore, the insertion is rejected.
@@
 	      \begin{verbatim}
-Insert(3,'John',
-[[6/4/12/2012, 1, 1],
-         [25/4/2012, 1, 1]]);
+Insert(3,'E.U.',
+[[6/4/12/1012, 1, 1],
+         [25/4/1012, 1, 1]]);
 	    \end{verbatim}
@@
 
-The algorithm for the implementation of the insert operation is the following:
+\begin{definition}
+ \emph{$\INS \left(r, t \right)$}.
+Consider the elements in definition \ref{def:generalized-fuzzy-temporal-relation}. The algorithm for the implementation of the insert operation is defined as follows.
+\end{definition}
 
@@
 \label{eq:insert-fuzzy-temporal}
-\INS \left(R ,\left(A, I \right) \right) &=\\
+\INS \left(r, t\right) &=\\
+\INS \left(r, t\right) =\\
 \begin{cases}
 \nonumber
-R \cup \left(A, I \right), & \mbox{ if }  \left(A, I \right) \not \in R \mbox{ or }\\
-&  \left(A, J \right) \in R \mbox{ and } I \text{ Overlaps } J < 1\\
-%\big \lbrace \left( \exists J= \left(X', Y' \right): \left(A, J \right) \in R \wedge \right. \\
-%\left. Y' \in \left \lbrace \left(<, X \right)\right \rbrace  \right) \vee \\
-%\left( \exists K= \left(X'', Y'' \right): \left(A, K \right) \in R \wedge \right. \\
-%\left. X'' \in \left \lbrace \left(>, Y \right)\right \rbrace \vee \right) \big \rbrace\\
-%&  J<I<K  \big \rbrace\\
-\MOD \left(R, \left(A, I \right)\right), & \mbox{ otherwise }  
+r \cup \lbrace t \rbrace
+& \mbox{ if }  t \not \in r \mbox{ or }\\
+\nonumber & \forall t_k \in Vt,  \left(i \mbox{ Overlaps } i_k\right) < 1\\
+% 
+% R \cup \left(A, I \right) & \mbox{ if }  (A, I ) \not \in R  \mbox{ or } \\
+%  &   \left(A, J\right) \in R \mbox{ and } \lnot \left(I \mbox{ Overlaps } J\right) \\
+\mbox{modify(r, t)} & \mbox{ otherwise }    %\mbox{ if } \exists J \left(A,J\right) \in R \wedge \mbox{Current(A,J)} \\ 
+%R, &   
 \end{cases} 	
@@
 \subsubsection{\label{subsubsec:delete-fuzzy-temporal}Delete}
-The delete operation logically removes an entity $A$ which is valid in the relation $R$:
+The delete operation logically removes an entity which is valid in the instance $r$ of the relation $R_{FTG}$.
+
+\begin{definition}
+ \emph{$\DEL \left(r, t \right) $}.
+Consider the elements in definition \ref{def:generalized-fuzzy-temporal-relation}. The algorithm for the delete operation is defined as follows.
+\end{definition}
+
 \begin{align}
 \label{eq:delete-fuzzy-temporal}
-\DEL \left(R ,A \right) =\\
+\\DEL \left(r, t\right) =
 \begin{cases}
 \nonumber
-R - \left \lbrace \left(A, I_1 \right), \ldots, \left(A, I_n \right) \right \rbrace, &  \forall I_i: \left( A, I_i \right) \in R\\
-R, & \mbox{ otherwise }  
+r - Vt
 \end{cases} 	
@@
 
-For example, consider that the hospital manager wants to delete the history for the patient with ID = 3. The following sentence deletes all the rows for the patient with ID = 3.
+For example, consider that the document manager wants to delete the history for the document with ID = 3. The following sentence deletes all the rows for the documents with ID = 3.
 
 \begin{verbatim}
- Delete(3,'John');
+ Delete(3,'E.U.');
 \end{verbatim}
@@
 % Revise(5,'Harry Potter IV', 
-% [[18/3/2012, 4, 1], 
-%           [2/4/2012, 2, 2]]);
+% [[18/3/1012, 4, 1], 
+%           [2/4/1012, 2, 2]]);
 % \end{verbatim}

### 2012-06-23 ijcis-pos-model/sources/temporaldatabases.tex
more modifications

@@
       The domain for each attribute $A_1, \ldots, A_n$ is $D_1, \ldots, D_n$ respectively. 
-\item The primary key for the attributes in $A$ is given by $A_K$:
+\item The primary key $A_K$ is a subset of the attributes in $A$.
       \begin{equation}
@@
 	    \begin{equation}
-	     \label{eq:all-the-versions} 
+	     \label{eq:pk-attribute} 
 	      a_k = t\left[A_K \right]
@@
 $t_1$&$a_{11}$ & $a_{12}$ & $a_{13}$ & $s_1$ & $e_1$ \\
-$t_1$ & $a_{21}$ & $a_{22}$ & $a_{23}$ &  $s_2$ & $e_2$ \\
-$t_1$ & $a_{11}$ & $a_{12}$ & $a_{31}$ & $s_3$ & $e_3$\\
+$t_2$ & $a_{21}$ & $a_{22}$ & $a_{23}$ &  $s_2$ & $e_2$ \\
+$t_3$ & $a_{11}$ & $a_{12}$ & $a_{31}$ & $s_3$ & $e_3$\\
 \hline\\
@@
 The value for the time interval is $I = \left(4/4/2012 , UC \right)$.
-The meaning is that the patient with ID = 3 was in the hospital the  4/4/2012 and he or she is still there.
-The patient with ID = 3 is current in the relation.
+The meaning is that the document with ID = 3 was valid the  4/4/2012 and it is still valid.
+The document with ID = 3 is current in the relation.
 
@@
 \label{ex:library-database}
- Consider a hospital database.
-Each patient is stored in the database with an identifier (ID), the name and the dates when the patient entered and left the hospital.
+ Consider a historical database containing diplomatic documents.
+The starting and the ending dates when the diplomatic document is valid.
+It is possible that a diplomatic document is valid for a period of time and several years later it becomes valid again.
+The following elements are stored: an identifier of the document (ID), the entity that issues the document and the dates when document is valid.
 Table \ref{tbl:library-sample} illustrates the first version of the database, after three insertions.
-In this example, $A = ($ ID, Name $)$ and $I = ($ Arrival, Leave $)$.
+In this example, $A = ($ ID, Entity $)$ and $I = ($ Start, End $)$.
 \end{example}
@@
 %\begin{table}[htbp]
-\tcap{\label{tbl:library-sample}Sample hospital database}
+\tcap{\label{tbl:library-sample}Sample historical database}
 %\centerline{\small DATA TYPES}
@@
 \hline
-\textbf{ID}  & Name  & Arrival & Leave \\
+\textbf{ID}  & Entity  & Start & End \\
 \hline
-3 & John & 15/3/2012 & 30/3/2012 \\
-4 & Antoon & 25/3/2012 & 4/4/2012 \\
-5 & Daan & 18/3/2012 & 2/4/2012\\
-3 & John & 4/4/2012 & UC \\
+3 & E.U. & 15/3/2012 & 30/3/2012 \\
+4 & N.A.T.O.  & 25/3/2012 & 4/4/2012 \\
+5 & C.E.I. & 18/3/2012 & 2/4/2012\\
+3 & E.U. & 4/4/2012 & UC \\
 \hline\\
@@
 \label{def:current-in-relation}
-\emph{Current $\left(a_k \right)$.}
+\emph{Current $\left(r,a_k \right)$.}
 Consider the elements in definition \ref{def:valid-time-relation}.
-The function Current $\left(a_k \right)$ returns the crisp time interval $i$ if the tuple $t$ with the primary key values given by $a_k$ is currently in the instance $r$ of the relation $R$.
-The function returns the emtyset otherwise.
+The function Current $\left(r,a_k \right)$ returns the crisp time interval $i$ if the tuple $t$ with the primary key values given by $a_k$ is currently in the instance $r$ of the relation $R$.
+The function returns the empty set otherwise.
 % 
@@
 \label{eq:current-in-relation}
-\mbox{Current} \left(a_k\right) &=& \\ 
+\mbox{Current} \left(r, a_k\right) &=& \\ 
 \begin{cases}
 \nonumber
-t & \mbox{ if } \exists t \in r : i = \left(s, UC \right)\\
+t & \mbox{ if } \exists t \in r : i = \left(s, UC \right) \mbox{ and }\\
+& a_k = t[A_K] \\
 \emptyset & \mbox{ otherwise}
@@
 
-For example, the function Current(R, (3,`John')) returns the time interval $I = \left(4/4/2012 , UC  \right)$. Conversely, Current(R, (4,`Antoon')) returns the empty set.
+For example, the function Current(r, (3,`E.U.')) returns the time interval $i = \left(4/4/2012 , UC  \right)$. Conversely, Current(r, (4,`N.A.T.O.')) returns the empty set.
 
@@
  \label{eq:variables-current-version}
-t_{CUR} = \mbox{ Current } \left(t[A_K] \right)\\
+t_{CUR} = \mbox{ Current } \left(r, t[A_K] \right)\\
 \nonumber
@@
 
-For example, consider the patient with ID = 3 in Table \ref{tbl:library-sample}.
+For example, consider the document with ID = 3 in Table \ref{tbl:library-sample}.
+The current version of the document started on 4/4/2012.
+The document was not valid anymore but, for some reason, the date  was not registered.
+The document was valid again on 24/4/2012.
+Then, the Close-current function closes the current version of the document and adds a new version.
+First, the function CloseR is applied with $i_1 = \left(4/4/2012, UC \right)$ and $i_2 = \left(24/4/2012, UC \right)$.
+Hence, the value for the time interval $i_1$ is $ \left(4/4/2012, 23/4/2012 \right)$.
+Then, the modifications on the value for the time interval $i_1$ are stored.
+Finally a new row with the current values of the document and the time interval $i_2$ is stored.
+% patient with ID = 3 in Table \ref{tbl:library-sample}.
 The current version arrived on 4/4/2012.
@@
 \subsubsection{\label{subsubsec:modify}Modify}
-This operation adds new information about an existing entity $A$ in the relation $R$.
-The modify operation does not remove any previous value of the object $A$.
-It closes the current version for the entity $A$ and adds a new version.
+This operation adds new information about an existing entity (given by the tuple $t$) in the instance $r$ of the relation $R$.
+The modify operation does not remove any previous value of the entity.
+It closes the current version and adds a new version.
 
@@
 \begin{definition}
- \emph{Modify(R, (A,I))}.
+ \emph{Modify(r, t)}.
 Consider the elements in definition \ref{def:valid-time-relation}. The algorithm for the modify operation is defined as follows.
@@
 \subsubsection{\label{subsubsec:insert}Insert}
-The user wants to store an entity $A$ which is valid in the relation $R$ during the time interval specified by $I = \left(a, b \right)$.
+The user wants to store an entity (given by the tuple $t$) which is valid in the instance $r$ of the relation $R$ during the time interval specified by $i = \left(s, e \right)$.
 %
@@
 \begin{enumerate}
-\item The entity $A$ was never in the relation $R$: The entity is added with the valid-time indicated by the crisp interval $I$. 
+\item The entity was never in the relation: The entity is added with the valid-time indicated by the crisp interval $i$. 
 % In the example, this correspond with the following insertion sentences:
@@
 
-\item The entity $A$ is in the relation $R$. Depending on the value of the time interval $I$, there are three possibilities:
+\item The entity is in the relation. Depending on the value of the time interval, there are three possibilities:
 	\begin{enumerate}
-	\item Insert $\left(A, I\right)$ in the relation $R$.
-If the time interval $I$ does not overlap any other valid-time interval in the relation $R$ for the entity $A$.
-For example, consider that the book with ID = 3 is loaned on 4/4/2012 but the book has not been returned yet.
+	\item Insert $t$ in the instance $r$ of the relation $R$.
+If the time interval $i$ does not overlap any other valid-time interval in the instance $r$ relation $R$ for the entity.
+For example, consider that the document with ID = 3 began again to be valid on 4/4/2012 and it is still currently valid.
 This is illustrated in Table \ref{tbl:library-sample}.
@@
 
-	\item Modify and close the current version of $A$ and insert $\left(A, I\right)$.
-For example, consider now that the patient with ID = 3 arrived to the hospital on 24/4/2012.
-Here the problem is that the patient with ID = 3 arrived on 4/4/2012 but, for some reason, the departure date was not stored.
-If the patient is already at the hospital, then it is necessary to set the departure date and add a new row with the new arrival date.
+	\item Modify and close the current version of $t$ and insert the new version.
+For example, consider now that the document with ID = 3 was valid on 24/4/2012.
+Here the problem is that the document with ID = 3 was valid on 4/4/2012 but, for some reason, the ending date was not stored.
+If the document is again valid, then it is necessary to set the ending date and add a new row with the new starting date.
 This is illustrated in Table \ref{tbl:library-sample-2}.
@@
 % 	      \end{verbatim}
-	\item Reject the insertion, if the time interval $I$ does overlap any existing valid-time interval for the entity $A$ in the relation $R$.
-For example, consider that the hospital manager wants to introduce a past arrival for the patient with ID = 3.
-The patient arrived on 6/4/2012 and left the hospital on 25/4/2012.
-As the dates do overlap, it is not possible that the patient was in the hospital at that time interval.
+	\item Reject the insertion, if the time interval $i$ does overlap any existing valid-time interval for the entity $t$ in the instance $r$ of the relation $R$.
+For example, consider that the document manager wants to introduce a past valid-time for the document with ID = 3.
+The validity period for the document is $[$6/4/2012, 25/4/2012$]$.
+As the dates do overlap, it is not possible that the document was valid at that time interval.
 Therefore, the insertion is rejected.
@@
 %\begin{table}[htbp]
-\tcap{\label{tbl:library-sample-2}Sample hospital database}
+\tcap{\label{tbl:library-sample-2}Sample historical database}
 %\centerline{\small DATA TYPES}
@@
 \hline
-\textbf{ID}  & Name  & Arrival & Leave \\
+\textbf{ID}  & Entity  & Start & End \\
 \hline
-3 & John & 15/3/2012 & 30/3/2012 \\
-4 & Antoon & 25/3/2012 & 4/4/2012 \\
-5 & Daan Potter & 18/3/2012 & 2/4/2012\\
-3 & John & 4/4/2012 & 23/4/2012 \\
-3 & John & 24/4/2012 & UC \\
+3 & E.U. & 15/3/2012 & 30/3/2012 \\
+4 & N.A.T.O. & 25/3/2012 & 4/4/2012 \\
+5 & C.E.I.  & 18/3/2012 & 2/4/2012\\
+3 & E.U. & 4/4/2012 & 23/4/2012 \\
+3 & E.U. & 24/4/2012 & UC \\
 \hline\\
@@
 \subsubsection{\label{subsubsec:del}Delete}
-The delete operation logically removes a current entity $A$ which is valid in the relation $R$.
+The delete operation logically removes a current entity $t$ which is valid in the instance $r$ of the relation $R$.
 \begin{definition}
@@
 
-For example, consider that the hospital manager wants to delete the history for the patient with ID = 3. The result of this operation is shown in Table \ref{tbl:library-sample-3}.
+For example, consider that the document manager wants to delete the history for the document with ID = 3. The result of this operation is shown in Table \ref{tbl:library-sample-3}.
 
@@
 %\begin{table}[htbp]
-\tcap{\label{tbl:library-sample-3}Sample hospital database}
+\tcap{\label{tbl:library-sample-3}Sample historical database}
 %\centerline{\small DATA TYPES}
@@
 \hline
-\textbf{ID}  & Name  & Arrival & Leave\\
+\textbf{ID}  & Entity  & Start & End\\
 \hline
-4 & Antoon & 25/3/2012 & 4/4/2012 \\
-5 & Daan & 18/3/2012 & 2/4/2012\\
+4 & N.A.T.O. & 25/3/2012 & 4/4/2012 \\
+5 & C.E.I. & 18/3/2012 & 2/4/2012\\
 \hline\\

### 2012-06-26 trunk/ijcis-pos-model/sources/temporalModel.tex
final changes

@@
 \label{def:fuzzy-temporal-relation}
-Consider the elements in definition \ref{def:valid-time-relation}. Some of the elements will be extended for the fuzzy case.
+Consider the elements in definition \ref{def:valid-time-relation}. Some of them will be extended for the fuzzy case.
 
 \begin{itemize}
- \item A set of non-temporal fuzzy  or crisp attributes.
-	\begin{equation}
-	\label{eq:fuzzy-attribute-set}
-	A = \left \lbrace A_1, A_2, \ldots, A_n \right \rbrace
-	\end{equation}
-      The domain for each attribute $A_1, \ldots, A_n$ is $D_1, \ldots, D_n$ respectively. 
-\item The primary key $A_K$ is a subset of $A$.
-      \begin{equation}
-       \label{eq:fuzzy-primary-key-a}
-      A_K \subset A
-      \end{equation}
-A formal definition of primary key for fuzzy relational databases will be given later in Definition \ref{def:generalized-primary-key}.
-\item A set of two attributes; $S$  and $E$ the attributes for the starting and ending ill-known points respectively.
+%  \item A set of non-temporal fuzzy  or crisp attributes.
+% 	\begin{equation}
+% 	\label{eq:fuzzy-attribute-set}
+% 	A = \left \lbrace A_1, A_2, \ldots, A_n \right \rbrace
+% 	\end{equation}
+%       The domain for each attribute $A_1, \ldots, A_n$ is $D_1, \ldots, D_n$ respectively. 
+% \item The primary key $A_K$ is a subset of $A$.
+%       \begin{equation}
+%        \label{eq:fuzzy-primary-key-a}
+%       A_K \subset A
+%       \end{equation}
+% A formal definition of primary key for fuzzy relational databases will be given later in Definition \ref{def:generalized-primary-key}.
+% \item A set of two attributes; $S$  and $E$ the attributes for the starting and ending ill-known points respectively.
 $I$ is a possibilistic validity period \emph{PVP} as explained in Section \ref{subsec:ill-known-interval}.
-\begin{equation}
- \label{eq:fuzzy-attribute-time-interval}
-I = \left \lbrace S, E \right \rbrace
-\end{equation}
-\item An attribute called version identifier, $V_{ID}$. This attribute is a counter for each different version of the entities. 
+% \begin{equation}
+%  \label{eq:fuzzy-attribute-time-interval}
+% I = \left \lbrace S, E \right \rbrace
+% \end{equation}
+\item An attribute called version identifier, $V_{ID}$, will be added to the schema. This attribute is a counter for each different version of the entities. 
 % \begin{equation}
@@
 
-\item Let $t$ be a tuple in the instance $r$; $t \in r$:
-      \begin{itemize}
-      \item We will note the values for the starting and the ending  points,$s$ and $e$  respectively. The value for the time interval is given by $i$.
-      \begin{align}
-       \label{eq:fuzzy-starting-point}
-      s = t\left[S \right]\\
-      e = t\left[E \right]\\
-      i = \left(s, e\right)
-      \end{align}
-
-      \item Let $A_k \subset A$ be the set of the non-temporal attributes that are part of the primary key (equation \eqref{eq:fuzzy-primary-key-a}).
+% \item Let $t$ be a tuple in the instance $r$; $t \in r$:
+%       \begin{itemize}
+%       \item We will note the values for the starting and the ending  points,$s$ and $e$  respectively. The value for the time interval is given by $i$.
+%       \begin{align}
+%        \label{eq:fuzzy-starting-point}
+%       s = t\left[S \right]\\
+%       e = t\left[E \right]\\
+%       i = \left(s, e\right)
+%       \end{align}
+% 
+% 
+%       \item Let $A_k \subset A$ be the set of the non-temporal attributes that are part of the primary key (equation \eqref{eq:fuzzy-primary-key-a}).
 Then, $a_k$ denotes the values for the non-temporal attributes of the primary key.
-	    \begin{equation}
-	     \label{eq:fuzzy-pk-attribute} 
-	      a_k = t\left[A_K \right]
-	    \end{equation}
-
+% 	    \begin{equation}
+% 	     \label{eq:fuzzy-pk-attribute} 
+% 	      a_k = t\left[A_K \right]
+% 	    \end{equation}
+% 
     \item Let $K_{GT}$ be the primary key for the valid-time relation as given in equation \eqref{eq:fuzzy-valid-time-temporal-pk}. Then, $k$ denotes the values for the attributes in the primary key.
@@
 	  \end{equation}
-
-      \end{itemize}
-\item $Vt$ is the set with all the versions for a given tuple $t$.
-
-\begin{equation}
- \label{eq:fuzzy-all-the-versions}
-Vt = r\left(t\left[A_K\right] \right)
-\end{equation}
-If a tuple $t$ contains more than one version, the result of $Vt$ is a set.
+% 
+% 
+%       \end{itemize}
+% \item $Vt$ is the set with all the versions for a given tuple $t$.
+% 
+% \begin{equation}
+%  \label{eq:fuzzy-all-the-versions}
+% Vt = r\left(t\left[A_K\right] \right)
+% \end{equation}
+% If a tuple $t$ contains more than one version, the result of $Vt$ is a set.
 We will use the index $k$ to address the elements of the set.
@@
 \INS \left(r, t\right) &=\\
-\INS \left(r, t\right) =\\
 \begin{cases}
@@
 \label{eq:delete-fuzzy-temporal}
-\\DEL \left(r, t\right) =
+\DEL \left(r, t\right) =
 \begin{cases}

### 2012-06-26 trunk/ijcis-pos-model/sources/temporaldatabases.tex
final changes

@@
 \begin{definition}
+\label{def:valid-time-relation}
 \emph{Valid-time relation.}
-Consider $\A$ be the set of all the entities, and let $A=\left(A_1, \ldots, A_n \right), A \in \A$ be the set of attributes that define an entity.
+Consider the following definitions and notations:
+
+\begin{itemize}
+ \item A set of non-temporal attributes.
+	\begin{equation}
+	\label{eq:attribute-set}
+	A = \left \lbrace A_1, A_2, \ldots, A_n \right \rbrace
+	\end{equation}
+      The domain for each attribute $A_1, \ldots, A_n$ is $D_1, \ldots, D_n$ respectively. 
+\item The original primary key $A_K$ is a subset of the attributes in $A$.
+      \begin{equation}
+       \label{eq:primary-key-a}
+      A_K \subset A
+      \end{equation}
+\item A set of two attributes, $S$  and $E$ for the starting and ending points respectively. $I$ defines the valid time interval for the data. 
+\begin{equation}
+ \label{eq:attribute-time-interval}
+I = \left \lbrace S, E \right \rbrace
+\end{equation}
+\item Then $R$, the schema for the valid-time relation is:
+\begin{equation}
+ \label{eq:valid-time-relation}
+R = A \cup  I
+\end{equation}
+\item The primary key for the valid-time relation $R$ is:
+\begin{equation}
+ \label{eq:valid-time-temporal-pk}
+PK = A_K \cup I
+\end{equation}
+
+\item We will note by $r$ any valid instance of $R$. 
+      \begin{equation}
+       \label{eq:valid-time-instance}
+      r \subseteq D_1\ x\ \ldots\ x\ D_n
+      \end{equation}
+
+\item Let $t$ be a tuple in the instance $r$; $t \in r$.
+    % \begin{itemize}
+    %  \item
+Then, we will note the values for the starting and the ending  points, $s$ and $e$  respectively. The value for the time interval is given by $i$.
+      \begin{align}
+       \label{eq:starting-point}
+      s = t\left[S \right]\\
+      e = t\left[E \right]\\
+      i = \left(s, e\right)
+      \end{align}
+
+      \item Let $A_k \subset A$ be as defined in \eqref{eq:primary-key-a}). Then, $a_k$ denotes the values for $A_K$ in tuple $t$.
+	    \begin{equation}
+	     \label{eq:pk-attribute} 
+	      a_k = t\left[A_K \right]
+	    \end{equation}
+
+    \item Let $PK$ be the primary key for the valid-time relation as given in equation \eqref{eq:valid-time-temporal-pk}.
+Then, $pk$ denotes the values for the attributes in the primary key in a tuple $t$.
+	  \begin{equation}
+	   \label{eq:value-pk}
+	  pk = t\left[PK \right]
+	  \end{equation}
+
+     % \end{itemize}
+\item $Vt$ is the set of all the versions for a given tuple $t$. Formally,
+
+\begin{equation}
+ \label{eq:all-the-versions}
+Vt =\left \lbrace t_i \in r, t_i\left[A_K\right] = t\left[A_K\right] \right \rbrace
+\end{equation}
+Obviously, $t$ itself is included in the set.
+% If a tuple $t$ contains more than one version, the result of $Vt$ is a set.
+We will use the index $k$ to address the elements of the set.
+E.g.
+$Vt = \left \lbrace t_1, t_2, \ldots, t_n \right \rbrace$.
+Then $t_k , k \in \left \lbrace 1, \ldots, n \right \rbrace$ is an element of $Vt$.
+\end{itemize}
+\end{definition}
+We will illustrate the definitions with an example.
+
+\begin{example}
+Consider the set of attributes $A = \left \lbrace A_1, A_2, A_3 \right \rbrace$.
+The primary key for these attributes is given by $A_K = \left \lbrace A_1, A_2 \right \rbrace$.
+Let $I = \left \lbrace S, E \right \rbrace$ be the set of temporal attributes that define the validity period of the data.
+$R$ is the valid-time relation and $r$ is an instance of the relation.
+The instance $r$ is given by the following elements.
+$r = \left \lbrace \left(a_{11}, a_{12}, a_{13}, s_1 ,e_1 \right) \right.$,  $\left(a_{21}, a_{22}, a_{23}, s_2, e_2 \right)$ , $\left.
+\left(a_{11}, a_{12}, a_{31}, s_3, e_3 \right) \right \rbrace$.
+The instance $r$ is illustrated in Table \ref{tbl:sample-definitions}.
+Consider the tuple $t = \left(a_{11}, a_{12}, a_{13}, s_1 ,e_1 \right)$. Then,
+\begin{align}
+ \nonumber
+s_1 &= t\left[S \right]\\
+ \nonumber
+e_1 &= t[E]\\
+ \nonumber
+i\ \ &= \left(s_1, e_1\right)\\
+ \nonumber
+a_k &= t\left[A_K\right] = \left(a_{11}, a_{12} \right) \\
+ \nonumber
+pk &= t\left[PK\right] =\left(a_{11}, a_{12}, s_1, e_1\right)\\
+ \nonumber
+Vt &= \left \lbrace t_1, t_3 \right \rbrace
+\end{align}
+
+\end{example}
+
+\begin{samepage}
+\vglue13pt
+%\begin{table}[htbp]
+\tcap{\label{tbl:sample-definitions}Sample database containing the instance $r$ of the valid-time relation $R$.}
+%\centerline{\small DATA TYPES}
+\vglue-6pt
+\centerline{\small\baselineskip=13pt
+\begin{tabular}{c c c c c c}\\
+\hline
+& \textbf{A$_1$}  & \textbf{A$_2$}  & $A_3$ & $S$ & $E$ \\
+\hline
+$t_1$&$a_{11}$ & $a_{12}$ & $a_{13}$ & $s_1$ & $e_1$ \\
+$t_2$ & $a_{21}$ & $a_{22}$ & $a_{23}$ &  $s_2$ & $e_2$ \\
+$t_3$ & $a_{11}$ & $a_{12}$ & $a_{31}$ & $s_3$ & $e_3$\\
+\hline\\
+\end{tabular}
+}
+\end{samepage}
+
+% Consider $\A$ be the set of all the entities, and let $A=\left(A_1, \ldots, A_n \right), A \in \A$ be the set of attributes that define an entity.
 Let $I = \left(S_i, E_i \right)$ be the set of attributes defining the time interval.
@@
 Then, the following equation indicates that the pair $\left(A, I\right)$ is in the relation R:
-
-\begin{equation}
-\label{eq:rel-def}
-\left( R, \left(A , I\right) \right)
-\end{equation}
-\end{definition}
-
+% 
+% \begin{equation}
+% \label{eq:rel-def}
+% \left( R, \left(A , I\right) \right)
+% \end{equation}
+% 
+% 
 In order to simplify the algorithms for the manipulation of data, some auxiliary functions and constants are defined:
@@
 \emph{From the beginning (FB).} 
-Consider $A=\left(A_1, \ldots, A_n \right)$ be the set of attributes that define an entity and $I = \left(S_i, E_i \right), S_i, E_i \in \T$ the set of attributes defining the time interval.
+Consider the elements in definition \ref{def:valid-time-relation}. We will say $i = \left(FB, e \right)$ when $i = \left(- \infty,e \right)$. 
+% $i = \left(s, e \right)$ be the value for a time interval as given equation \eqref{}.
+% $A=\left(A_1, \ldots, A_n \right)$ be the set of attributes that define an entity and $I = \left(S_i, E_i \right), S_i, E_i \in \T$ the set of attributes defining the time interval.
 Let $R \subseteq \A\  x\  \I$ be a valid-time relation.
@@
 \emph{Until changed (UC).} 
-Consider $A=\left(A_1, \ldots, A_n \right)$ be the set of attributes that define an entity.
+Consider the elements in definition \ref{def:valid-time-relation}. We will say  $i = \left(s, UC \right)$ when $i = \left(s, +\infty \right)$.
+% Consider $A=\left(A_1, \ldots, A_n \right)$ be the set of attributes that define an entity.
 Let $\T$ be the time domain, and $I = \left(S_i, E_i \right), S_i, E_i \in \T$ the set of attributes defining the time interval.
@@
 
-If the pair $\left(A, I \right)$ with $I = \left(S_i, UC \right), S_i \in \T$ is in the relation $R$, then, it is said that the pair $\left(A, I \right)$ is \emph{current} in the relation.
+\begin{definition}
+\emph{Current.} 
+Consider the elements in definition \ref{def:valid-time-relation}. We will say that the tuple $t$ is current in the instance $r$ of the relation $R$ when $i = \left(s, UC \right)$.
+\end{definition}
+
+%If the pair $\left(A, I \right)$ with $I = \left(S_i, UC \right), S_i \in \T$ is in the relation $R$, then, it is said that the pair $\left(A, I \right)$ is \emph{current} in the relation.
 
@@
 \label{def:current-in-relation}
-\emph{Current $\left(R, A \right)$.}
-Consider $A=\left(A_1, \ldots, A_n \right), A \in \A$ be the set of attributes that define an entity and $I = \left(S_i, UC \right)$ the set of attribute defining the time interval.
+\emph{Current $\left(r,a_k \right)$.}
+Consider the elements in definition \ref{def:valid-time-relation}.
+The function Current $\left(r,a_k \right)$ returns the crisp time interval $i$ if the tuple $t$ with the primary key values given by $a_k$ is currently in the instance $r$ of the relation $R$.
+The function returns the empty set otherwise.
+% 
+% Consider $A=\left(A_1, \ldots, A_n \right), A \in \A$ be the set of attributes that define an entity and $I = \left(S_i, UC \right)$ the set of attribute defining the time interval.
 Let $R \subseteq \A\  x\  \I$ be a valid-time relation.
@@
 \label{eq:current-in-relation}
-\mbox{Current} \left(R, A \right) &=& \\ 
+\mbox{Current} \left(r, a_k\right) &=& \\ 
 \begin{cases}
 \nonumber
-I & \mbox{ if } (A,I) \in R \mbox{ and } I = \left(S_i, UC \right), S_i \in \T\\
+t & \mbox{ if } \exists t \in r : i = \left(s, UC \right) \mbox{ and }\\
+& a_k = t[A_K] \\
 \emptyset & \mbox{ otherwise}
@@
  \label{def:overlaps}
-Given two time intervals $I = \left(S_i, E_i \right)$ and $J = \left(S_j, E_j\right)$, it is said that I overlaps J if:
+Given two time intervals $i_1 = \left(s_1, e_1 \right)$ and $i_2 = \left(s_2 e_2\right)$, it is said that $i_1$ overlaps $i_2$ if:
 \begin{equation}
  \label{eqn:overlaps}
-\text{I Overlaps J}  = \left(S_i > S_j  \right) \wedge \left(E_i < E_j  \right) 
+i_1 \text{ Overlaps } i_2  = \left(s_1 < s_2  \right) \wedge \left(e_1 < e_2  \right) 
 \end{equation}
@@
  \label{def:during}
-Given two time intervals $I = \left(S_i, E_i \right)$ and $J = \left(S_j, E_j\right)$, it is said that I during J if:
-\begin{equation}
+Given two time intervals $i_1 = \left(s_1, e_1 \right)$ and $i_2 = \left(s_2 e_2\right)$, it is said that $i_1$ during $i_2$ if:
+\begin{align}
  \label{eqn:during}
-\text{I During J}  = \left(S_i < S_j  \right) \wedge \left(E_i > S_j  \right) \wedge \left(E_i < E_j \right)
-\end{equation}
+i_1\text{ During } i_2  =& \left(s_1 > s_2  \right) \wedge \left(e_1 > s_2 \right) \wedge \\
+\nonumber & \left(e_1 < e_2 \right)
+\end{align}
 \end{definition}
@@
 \begin{definition}
-\emph{CloseR$\left( I, J \right)$:}
+\emph{CloseR$\left(i_1, i_2\right)$:}
 \label{def:close-a-crisp-interval-r}
-Consider two crisp intervals $I= \left[S_i ,E_i \right]$ and $J= \left[S_j ,E_j \right]$.
-The CloseR$\left(I, J\right)$ function allows to close the right-open interval $I$ with respect to the first value $e$ in $J$:
+Consider two crisp intervals $i_1 = \left(s_1, e_1 \right)$ and $i_2 = \left(s_2, e_2\right)$.
+The CloseR$\left(i_1, i_2 \right)$ function allows to close the right-open interval $i_1$ with respect to the first value $s_2$ in $i_2$:
 \begin{align}
 \label{eq:close-a-crisp-interval}
-\mbox{CloseR} \left( I, J \right) &=& \\ 
+\mbox{CloseR} \left[ i_1, i_2\right[ &=& \\ 
 \begin{cases}
@@
 %I=\left[a, c \right[ & \mbox{ if } b = +\infty \wedge J > I
-\left[S_i, S_j \right[ & \mbox{ if } E_i = UC \mbox{ and } J \mbox{ During } I \\
-I & \mbox{ otherwise}
+i_1 = \left(s_1,s_2 \right( & \mbox{ if } e_1 = UC \mbox{ and } i_2 \mbox{ During } i_1 \\
+i_1 & \mbox{ otherwise}
 \end{cases}
@@
 
-For example, consider two time intervals, $I = \left(4/4/2012, UC \right)$ and $J = \left(24/4/2012, UC \right)$.
-The result of applying the CloseR$\left(I, J \right)$ is $I = \left(4/4/2012, 23/4/2012 \right)$.
+For example, consider two time intervals, $i_1 = \left(4/4/2012, UC \right)$ and $i_2 = \left(24/4/2012, UC \right)$.
+The result of applying the CloseR$\left(i_1, i_2 \right)$ is $i_1 = \left(4/4/2012, 23/4/2012 \right)$.
 
@@
 \label{def:close-current-version}
-\emph{Close-current$\left(R, \left(A, J  \right)\right)$}.
-Consider $A=\left(A_1, \ldots, A_n \right)$ be the set of attributes that define an entity and $J = \left(S_j,E_j \right)$ the set of attribute defining the time interval.
-Let $R \subseteq \A\  x\  \I$ be a valid-time relation.
-Then, the function close-current$\left(R,\left(A,J\right) \right)$ closes any current version of the object $A$ if it exists and add the new version $\left(A, J\right)$.
+\emph{Close-current$\left(r, t\right)$}.
+Consider the elements in definition \ref{def:valid-time-relation}.
+The function Close-current$\left(r, t \right)$ closes any current version $t_k$ of the entity given by $t$ if it exists and add the new version $t$.
+For the implementation the following variables are defined:
+\begin{align}
+ \label{eq:variables-current-version}
+t_{CUR} = \mbox{ Current } \left(r, t[A_K] \right)\\
+\nonumber
+t_{UP} = t_{CUR}; t_{UP}\left[i \right] = \mbox{Close }\left(t_{UP}\left[i \right], t \right)
+\nonumber
+\end{align}
+Then, $t_{CUR}$ is the current version of the entity given by the tuple $t$, and $t_{UP}$ is the updated version of the tuple $t_{CUR}$.
+In this updated version, the time interval given by $i_{UP}$ is closed with respect to the tuple $t$.
 
@@
 \label{eq:close-current}
-\text{Close-current} \left(R, \left(A, J\right) \right) =\\
+\text{Close-current} \left(r, t \right) =\\
 \begin{cases}
 \nonumber
-R - \left(A, I \right) \cup \left \lbrace \left(A, \mbox{ CloseR } \left(I, J\right) \right)\right \rbrace \cup \left \lbrace\left(A, J\right) \right \rbrace  \\
-\nonumber
-\mbox{ if  Current } \left(R, A \right) \neq \emptyset\\ %\wedge J > \mbox{ CloseR } \left(I, J \right)   \\
-\nonumber R , \text{ otherwise}
+r - t_{CUR} \cup \lbrace t_{UP}\rbrace \cup \lbrace t \rbrace
+& \mbox{ if } t_{CUR} \neq \emptyset \\
+% R - \left(A, I \right) \cup \left \lbrace \left(A, \mbox{ CloseR } \left(I, J\right) \right)\right \rbrace \cup \left \lbrace\left(A, J\right) \right \rbrace  \\
+% \nonumber
+% \mbox{ if  Current } \left(R, A \right) \neq \emptyset\\ %\wedge J > \mbox{ CloseR } \left(I, J \right)   \\
+\nonumber r , \text{ otherwise}
 \end{cases}
@@
 \begin{definition}
- \emph{Modify(R, (A,I))}.
-Let $R \subseteq \A\  x\  \I$ be a valid-time relation and  $A=\left(A_1, \ldots, A_n \right), A \in \A$ be the set of attributes that define an entity and $I = \left(S_i,E_i\right)$ the set of attributes defining the time interval.
-The algorithm for the modify operation is defined as follows.
+ \emph{Modify(r, t)}.
+Consider the elements in definition \ref{def:valid-time-relation}. The algorithm for the modify operation is defined as follows.
 \end{definition}
@@
 \label{eq:modify}
-\MOD \left( R, \left(A, I \right)\right) =\\
+\MOD \left( r, t\right) =\\
 %\begin{cases}
 \nonumber
-\mbox{ Close-current }\left(R, \left(A, I\right) \right) %\\
+\mbox{ Close-current }\left(r, t\right) %\\
 % \mbox{ if Current}\left(R ,A \right) \neq \emptyset \\
@@
 \begin{definition}
- \emph{Insert(R,(A,I)))}.
-Let $R \subseteq \A\  x\  \I$ be a valid-time relation and  $A=\left(A_1, \ldots, A_n \right), A \in \A$ be the set of attributes that define an entity and $I = \left(S_i,E_i\right)$ the set of attributes defining the time interval.
-Then, the algorithm for the implementation of the insert operation is defined as follows.
+ \emph{Insert(r, t)}.
+Consider the elements in definition \ref{def:valid-time-relation}. Then, the algorithm for the implementation of the insert operation is defined as follows.
 \end{definition}
@@
 \label{eq:insert}
-\INS \left(R ,\left(A, I\right)\right) =\\
+\INS \left(r, t\right) =\\
 \begin{cases}
 \nonumber
-R \cup \left(A, I \right) & \mbox{ if }  (A, I ) \not \in R  \mbox{ or } \\
- &   \left(A, J\right) \in R \mbox{ and } \lnot \left(I \mbox{ Overlaps } J\right) \\
-\mbox{modify(R,(A,I))} & \mbox{ otherwise }    %\mbox{ if } \exists J \left(A,J\right) \in R \wedge \mbox{Current(A,J)} \\ 
+r \cup \lbrace t \rbrace
+& \mbox{ if }  t \not \in r \mbox{ or }\\
+\nonumber & \forall t_k \in Vt, \lnot \left(i \mbox{ Overlaps } i_k\right)\\
+% 
+% R \cup \left(A, I \right) & \mbox{ if }  (A, I ) \not \in R  \mbox{ or } \\
+%  &   \left(A, J\right) \in R \mbox{ and } \lnot \left(I \mbox{ Overlaps } J\right) \\
+\mbox{modify(r, t)} & \mbox{ otherwise }    %\mbox{ if } \exists J \left(A,J\right) \in R \wedge \mbox{Current(A,J)} \\ 
 %R, &   
@@
 \subsubsection{\label{subsubsec:del}Delete}
-The delete operation logically removes a current entity $A$ which is valid in the relation $R$.
+The delete operation logically removes a current entity $t$ which is valid in the instance $r$ of the relation $R$.
 \begin{definition}
- \emph{Delete(R,(A,I))}.
-Let $R \subseteq \A\  x\  \I$ be a valid-time relation and  $A=\left(A_1, \ldots, A_n \right), A \in \A$ be the set of attributes that define an entity and $I = \left(S_i,E_i\right)$ the set of attributes defining the time interval.
+ \emph{Delete(r, t)}.
+Consider the elements in definition \ref{def:valid-time-relation}.The algorithm for the delete operation is defined as follows.
+% 
+% Let $R \subseteq \A\  x\  \I$ be a valid-time relation and  $A=\left(A_1, \ldots, A_n \right), A \in \A$ be the set of attributes that define an entity and $I = \left(S_i,E_i\right)$ the set of attributes defining the time interval.
 The algorithm for the delete operation is defined as follows.
@@
 \label{eq:delete}
-\DEL \left(R ,A \right) =\\
+\DEL \left(r, t\right) =
 \begin{cases}
 \nonumber
-R - \left \lbrace \left(A, I_1 \right),  \ldots, \left(A, I_n \right)   \right \rbrace, & \forall I_i : \left( A, I_i \right) \in R\\
-R, & \mbox{ otherwise }  
+r - Vt
+% 
+% 
+% R - \left \lbrace \left(A, I_1 \right),  \ldots, \left(A, I_n \right)   \right \rbrace, & \forall I_i : \left( A, I_i \right) \in R\\
+% R, & \mbox{ otherwise }  
 \end{cases} 	
 \end{align}
+The set $Vt$ is computed as explained in definition \ref{def:valid-time-relation} and contains all the versions for the tuple $t$.
 
-For example, consider that the hospital manager wants to delete the history for the patient with ID = 3. The result of this operation is shown in Table \ref{tbl:library-sample-3}.
+For example, consider that the document manager wants to delete the history for the document with ID = 3. The result of this operation is shown in Table \ref{tbl:library-sample-3}.
 

### 2012-06-26 trunk/ijcis-pos-model/sources/timeRepresentation.tex
final changes

@@
 
-%\begin{definition}
-\textbf{Ill-known time interval}\\
 An ill-known time interval denoted by $\left[X, Y\right]$ is a precise time interval whose boundaries are not precisely known.
-The ill-known time interval is defined by two ill-known  time points, namely $X$,$Y$.
-The evaluation of the ill-known time interval is given iby equations \eqref{ill-known-pos},\eqref{ill-known-nec}.
+%The ill-known time interval is defined by two ill-known  time points, namely $X$,$Y$.
+The evaluation of the ill-known time interval is given by equations \eqref{ill-known-pos},\eqref{ill-known-nec}.
+
+\begin{definition}
+\emph{Ill-known time interval}
+Let $\T$ be the time domain, and $X$, $Y$ two ill-known values in the time domain.
+An ill-known time interval is given by $\left[X, Y\right]$.
+The evaluation of the ill-known time interval is given by equations \eqref{ill-known-pos},\eqref{ill-known-nec}.
+We will note $\I_{PVP}$ the set of all the ill-known time intervals.
+
+\end{definition}
 
@@
 \includegraphics[scale=1]{./graphs/open-time-interval.pdf}
-\label{fig:example-open-time-interval}
+
 }
@@
 \vspace*{10pt}
-\fcaption{Possibility distribution for $Y$, and possibility and necessity measures for the open ill-known point, $X$}
+\fcaption{\label{fig:example-open-time-interval}Possibility distribution for $Y$, and possibility and necessity measures for the open ill-known point, $X$}
 \vspace*{13pt}
+\end{samepage}
 

### 2012-06-26 develop/fuzzyCalculator/src/main/java/es/ugr/decsai/idbis/fuzzycalculator/TrapezoidalFuzzyNumber.java
project to compute things with fuzzy numbers

@@ -1,0 +1,155 @@
+/*
+ * To change this template, choose Tools | Templates
+ * and open the template in the editor.
+ */
+package es.ugr.decsai.idbis.fuzzycalculator;
+
+/**
+ * A Trapezoidal fuzzy number that stores the four values and provides the evaluation function.
+ * As it is natural, alpha <= beta <= delta <= gamma.
+ * @author jose
+ */
+public class TrapezoidalFuzzyNumber {
+    
+    protected Double alpha;
+    protected Double beta;
+    protected Double gamma;
+    protected Double delta;
+    
+
+    public TrapezoidalFuzzyNumber(Double alpha, Double beta, Double gamma,Double delta) {
+        this.alpha = alpha;
+        this.beta = beta;
+        this.delta = delta;
+        this.gamma = gamma;
+    }
+
+    public Double getAlpha() {
+        return alpha;
+    }
+
+    public void setAlpha(Double alpha) {
+        this.alpha = alpha;
+    }
+
+    public Double getBeta() {
+        return beta;
+    }
+
+    public void setBeta(Double beta) {
+        this.beta = beta;
+    }
+
+    public Double getDelta() {
+        return delta;
+    }
+
+    public void setDelta(Double delta) {
+        this.delta = delta;
+    }
+
+    public Double getGamma() {
+        return gamma;
+    }
+
+    public void setGamma(Double gamma) {
+        this.gamma = gamma;
+    }
+    
+    protected void checkConstraints() throws FuzzyConstraintException{
+        if(!(this.alpha <= this.beta ) &&
+                (this.beta <= this.gamma) &&
+                (this.gamma <= this.delta)){
+            throw new FuzzyConstraintException("Values are not consistent");
+        }
+            
+    }
+    /**
+     * Function to obtain the membership of a value d with respect to the trapezoid.
+     * @param d
+     * @return
+     * @throws FuzzyConstraintException 
+     */
+    public Double evaluate(Double d) throws FuzzyConstraintException{
+        if(d==null ||
+                this.alpha == null ||
+                this.beta == null ||
+                this.gamma == null ||
+                this.delta == null
+                ){
+            throw new FuzzyConstraintException("values not initialized");
+        }
+        // cases d < alpha or d > delta
+        Double value = 0D;
+        //d belongs to the core
+        if(d >= this.beta && d <= this.gamma){
+            value = 1D;
+        }else if(d >= this.alpha && d < this.beta){
[…]

### 2012-06-26 trunk/develop/fuzzyCalculator/src/main/java/es/ugr/decsai/idbis/fuzzycalculator/App.java
project to compute things with fuzzy numbers

@@ -1,0 +1,33 @@
+package es.ugr.decsai.idbis.fuzzycalculator;
+
+import java.util.logging.Level;
+import java.util.logging.Logger;
+
+/**
+ * Hello world!
+ *
+ */
+public class App 
+{
+    public static void main( String[] args )
+    {
+        System.out.println( "Hello World!" );
+        TrapezoidalFuzzyNumber t = new TrapezoidalFuzzyNumber(0D, 10D, 15D, 20D);
+        TrapezoidalFuzzyNumber t1 = new TrapezoidalFuzzyNumber(15D, 20D, 25D, 30D);
+        
+        
+        for(double i=0;i< 30;i++){
+            try {
+                System.out.println("i: "+i + " eval "+t.evaluate(i)); 
+            } catch (FuzzyConstraintException ex) {
+                Logger.getLogger(App.class.getName()).log(Level.SEVERE, null, ex);
+            }
+        }
+        try {
+            System.out.println("FEQ = "+TrapezoidalFuzzyNumber.Feq(t, t1)); 
+        } catch (FuzzyConstraintException ex) {
+            Logger.getLogger(App.class.getName()).log(Level.SEVERE, null, ex);
+        }
+        
+    }
+}

### 2012-06-26 trunk/develop/fuzzyCalculator/src/main/java/es/ugr/decsai/idbis/fuzzycalculator/FuzzyConstraintException.java

@@ -1,0 +1,20 @@
+/*
+ * To change this template, choose Tools | Templates
+ * and open the template in the editor.
+ */
+package es.ugr.decsai.idbis.fuzzycalculator;
+
+/**
+ *
+ * @author jose
+ */
+class FuzzyConstraintException extends Exception {
+
+    public FuzzyConstraintException(String string) {
+        super(string);
+    }
+
+    public FuzzyConstraintException() {
+    }
+    
+}

### 2012-06-26 trunk/develop/fuzzyCalculator/src/test/java/es/ugr/decsai/idbis/fuzzycalculator/AppTest.java

@@ -1,0 +1,38 @@
+package es.ugr.decsai.idbis.fuzzycalculator;
+
+import junit.framework.Test;
+import junit.framework.TestCase;
+import junit.framework.TestSuite;
+
+/**
+ * Unit test for simple App.
+ */
+public class AppTest 
+    extends TestCase
+{
+    /**
+     * Create the test case
+     *
+     * @param testName name of the test case
+     */
+    public AppTest( String testName )
+    {
+        super( testName );
+    }
+
+    /**
+     * @return the suite of tests being tested
+     */
+    public static Test suite()
+    {
+        return new TestSuite( AppTest.class );
+    }
+
+    /**
+     * Rigourous Test :-)
+     */
+    public void testApp()
+    {
+        assertTrue( true );
+    }
+}

### 2012-08-01 ijcis-pos-model/sources/intro.tex
preprint version

@@
 In real world, some aspects or properties of objects are time-variant or time-related.
-e.g., the moment of a bank transaction and the status of an employee in a company, are time-related and time-variant notions, respectively.
+E.g., the moment of a bank transaction and the status of an employee in a company, are time-related and time-variant notions, respectively.
 A temporal database schema is a database schema that models objects with time-related or time-variant properties.
@@
 
-For example, let us consider an employee's database containing data about the positions of employees in a company.
+% For example, let us consider an employee's database containing data about the positions of employees in a company.
 Two dates are stored: the starting and ending dates respectively of the contract.
@@
 Without further cautions, a manager in the company could insert a new contract for an employee who is already working for the company.
+A temporal database model will typically constrain record insertion and prevent similar modelling inconsistencies.
+
+For example, let us consider a library database and, concretely, the modelling of the presence of books in the library.
+Two dates are stored: the loan and the return date.
+It is clear that a book cannot be loaned again until it is returned.
+Without further cautions, a library employee could loan the same book several times even if it is not returned.
 A temporal database model will typically constrain record insertion and prevent similar modelling inconsistencies.
@@
 
-Notably, in 1994, ``A Consensus Glossary of Temporal Database Concepts'' was published~\cite{Dyreson1994}.
+%Notably, in 1994, ``A Consensus Glossary of Temporal Database Concepts'' was published~\cite{Dyreson1994}.
 For this publication, 44 temporal database researchers, among them, some of the main researchers in this field, cooperated to reach a consensus on the nature and definitions of some of the main temporal database concepts and terminology.
@@
 In this sense, Allen~\cite{Allen1983} studied temporal relationships between time intervals (and as a special case time points).
-Among others, the querying of temporal databases has greatly profited from these temporal relationships, because they allowed more powerful user-specified temporal query demands, by allowing to express more complex relationships between the temporal notions in the temporal expressions in the query.
+Among others, the querying of temporal databases has greatly profited from these temporal relationships, because they allow more powerful user-specified temporal query demands, by allowing to express more complex relationships between the temporal notions in the temporal expressions in the query.
 For example, a query like `who were the department heads when Thomas worked for the institution' can be evaluated using operators similar to Allen's ones.
@@
 Humans handle temporal information using certain temporal notions like time intervals or time points~\cite{Dyreson1994}, and they often have to deal with imperfections like imprecision, vagueness, uncertainty or inconsistencies possibly contained in the descriptions of these temporal notions.
-These possible imperfections in descriptions of temporal notions determine an important issue in temporal modelling.
+These imperfections in descriptions of temporal notions determine an important issue in temporal modelling.
 Consider as an example the description of the temporal notion in a sentence like `The Belfry of Bruges was finished between 01/01/1201 A.D.
@@
 
-To allow information systems to cope with these and similar data imperfections, many approaches adopt fuzzy sets~\cite{Zadeh1965} for the representation and management of temporal information~\cite{Mitra1994,Nagypal2003,Billiet2011,Dubois2003}.
+To allow information systems to cope with these and similar data imperfections, many approaches adopt fuzzy sets~\cite{Zadeh1965} for the representation and management of temporal information~\cite{Mitra1994,Nagypal2003,Billiet2011,Dubois1983,Dubois2003}.
 The temporal relationships studied by Allen were fuzzified by several authors~\cite{Ohlbach2004,Nagypal2003,Schockaert2008}.
@@
 Although temporal reasoning is not discussed in this paper, it should be noted that, among others, Dubois and Prade et al.~\cite{Dubois2003,DuBois1989} have dealt with fuzziness and uncertainty in temporal reasoning.
+Finally, in \cite{rita1,rita2}, an approach to the linguistic summarization of data with time dimension is presented.
 
-The present work defines and implement a model for properly represent and manage uncertainty in valid-time specification in a relational database.
-Our work is focused on both the proposal of an appropriate formal framework to suitably manage time in databases and the implementation of an DML that allows to the user the transparent use of this proposal.
-None of the previous research offer a database model to accomplish this task.
+The present work defines and implements a model for properly representing and managing uncertainty in valid-time specification in a relational database.
+Our work is focused on both the proposal of an appropriate formal framework to suitably manage time in databases and the implementation of a DML that allows to the user the transparent use of this proposal.
+None of the previous research offers a database model to accomplish this task.
 
-This way together with the theoretical model, we also present and explain the main operations of the manipulation language for a temporal database which stores the valid-time periods of the objects affected by imprecision.
+This way, together with the theoretical model, we also present and explain the main operations of the manipulation language for a temporal database which stores the valid-time periods of the objects affected by imprecision.
 The rest of the work is organized as follows.

### 2012-08-01 ijcis-pos-model/sources/temporaldatabases.tex
preprint version

@@
 The time when the employee's contract is stored in the database is the TT.
-The time when the decision for hiring this employee was made is the DT.
+The time when the decision for hiring this employee was made is a DT.
 
@@
 Some proposals work with rough sets \cite{Qiang2009}, and some others rely on possibility distributions for representing uncertainty in time \cite{Dyreson1998,Garrido2009,Galindo2001}.
-In order to compare temporal possibility distributions, the extensions of the classical Allen's operators \cite{Allen1983} are defined in \cite{Ohlbach2004,Nagypal2003,Dubois2003a,Schockaert2008}.
+To compare temporal possibility distributions, the extensions of the classical Allen's operators \cite{Allen1983} are defined in \cite{Ohlbach2004,Nagypal2003,Dubois2003a,Schockaert2008}.
 %In the proposal section, we will follow the representation by means of possibility distributions, in order to work with both satisfaction and dissatisfaction degrees.
@@
 \end{definition}
-A PVP is an ill-known interval in the sense of definition \ref{def;possibilistic-variable}. Note that the PVP represents only one crisp time interval, but for some reason, it is (partially) unknown.
 
-The main feature for the PVP is that it preserves all the information for both starting and ending points\cite{Pons2011,Pons2012}.
+Note that the PVP represents only one crisp time interval, but for some reason, it is (partially) unknown.
+
+The main advantage for the PVP is that it preserves all the information for both starting and ending points\cite{Pons2011,Pons2012}.
 Table \ref{tbl:comparative-pvp-fvp} is a comparative between PVP and FVP.
@@
 \item Storage: The way the data are stored in the database.
-\item Possibility measures: Does the framework provide always a possibility measure for any relation between the temporal elements?
-\item Necessity measures:  Does the framework provide always a necessity measure for any relation between the temporal elements?
+\item Possibility measures: Does the framework provide a possibility measure for any relation between the temporal elements?
+\item Necessity measures:  Does the framework provide a necessity measure for any relation between the temporal elements?
 \end{enumerate}
@@
       The domain for each attribute $A_1, \ldots, A_n$ is $D_1, \ldots, D_n$ respectively. 
-\item The primary key $A_K$ is a subset of the attributes in $A$.
+\item The original primary key $A_K$ is a subset of the attributes in $A$.
       \begin{equation}
        \label{eq:primary-key-a}
-      A_K \subset A
+      A_K \subseteq A
       \end{equation}
-\item A set of two attributes; $S$  and $E$ the attributes for the starting and ending points respectively. $I$ defines the valid time interval for the data. 
+\item Two attributes, $S$  and $E$ for the starting and ending points respectively. $I$ defines the valid time interval for the data. 
 \begin{equation}
@@
 \end{equation}
+
+$\T$ is the time domain.
+
 \item Then $R$, the schema for the valid-time relation is:
@@
        \label{eq:valid-time-instance}
-      r \subseteq D_1\ x\ \ldots\ x\ D_n
+      r \subseteq D_1\ x\ \ldots\ x\ D_n x\ T x\ T
       \end{equation}
 
-\item Let $t$ be a tuple in the instance $r$; $t \in r$:
-      \begin{itemize}
-      \item We will note the values for the starting and the ending  points,$s$ and $e$  respectively. The value for the time interval is given by $i$.
-      \begin{align}
-       \label{eq:starting-point}
-      s = t\left[S \right]\\
-      e = t\left[E \right]\\
-      i = \left(s, e\right)
-      \end{align}
-
-      \item Let $A_k \subset A$ be the set of the non-temporal attributes that are part of the primary key (equation \eqref{eq:primary-key-a}).
-Then, $a_k$ denotes the values for the non-temporal attributes of the primary key.
-	    \begin{equation}
-	     \label{eq:pk-attribute} 
-	      a_k = t\left[A_K \right]
-	    \end{equation}
-
-    \item Let $PK$ be the primary key for the valid-time relation as given in equation \eqref{eq:valid-time-temporal-pk}. Then, $pk$ denotes the values for the attributes in the primary key.
-	  \begin{equation}
-	   \label{eq:value-pk}
-	  pk = t\left[PK \right]
-	  \end{equation}
-
-      \end{itemize}
-\item $Vt$ is the set with all the versions for a given tuple $t$.
+     % \end{itemize}
+\item $V(t)$ is the set of all the versions for a given tuple $t$. Formally,
 
@@
  \label{eq:all-the-versions}
-Vt = r\left(t\left[A_K\right] \right)
+V(t) =\left \lbrace t_i \in r, t_i\left[A_K\right] = t\left[A_K\right] \right \rbrace
 \end{equation}
-If a tuple $t$ contains more than one version, the result of $Vt$ is a set.
+Obviously, $t$ itself is included in the set.
+% If a tuple $t$ contains more than one version, the result of $Vt$ is a set.
 We will use the index $k$ to address the elements of the set.
@@
 The instance $r$ is illustrated in Table \ref{tbl:sample-definitions}.
-Consider the tuple $t = \left(a_{11}, a_{12}, a_{13}, s_1 ,e_1 \right)$. Then,
+Consider the tuple $t_1 = \left(a_{11}, a_{12}, a_{13}, s_1 ,e_1 \right)$. Then,
 \begin{align}
  \nonumber
-s_1 &= t\left[S \right]\\
+ t_1\left[S \right]&=s_1\\
  \nonumber
-e_1 &= t[E]\\
+t_1[E]&=e_1\\
  \nonumber
-i\ \ &= \left(s_1, e_1\right)\\
+t_1[S,E]&= \left(s_1, e_1\right)\\
  \nonumber
-a_k &= t\left[A_K\right] = \left(a_{11}, a_{12} \right) \\
+t_1\left[A_K\right] &= \left(a_{11}, a_{12} \right) \\
  \nonumber
-pk &= t\left[PK\right] =\left(a_{11}, a_{12}, s_1, e_1\right)\\
+t_1\left[PK\right] &=\left(a_{11}, a_{12}, s_1, e_1\right)\\
  \nonumber
-Vt &= r\left(t\left[A_K\right] \right) = \left \lbrace t_1, t_3 \right \rbrace
+V(t_1) &= \left \lbrace t_1, t_3 \right \rbrace
 \end{align}
@@
 \emph{From the beginning (FB).} 
-Consider the elements in definition \ref{def:valid-time-relation}. We will say $i = \left(FB, e \right)$ when $i = \left(- \infty,e \right)$. 
+Consider the elements in definition \ref{def:valid-time-relation} and a tuple t. We will say $t[S]=FB$ when $t[S]=- \infty$. 
 % $i = \left(s, e \right)$ be the value for a time interval as given equation \eqref{}.
@@
 \emph{Until changed (UC).} 
-Consider the elements in definition \ref{def:valid-time-relation}. We will say  $i = \left(s, UC \right)$ when $i = \left(s, +\infty \right)$.
+Consider the elements in definition \ref{def:valid-time-relation} and a tuple t. We will say $t[E]=UC$ when $t[E]=+ \infty$. 
 % Consider $A=\left(A_1, \ldots, A_n \right)$ be the set of attributes that define an entity.
@@
 \emph{Current.} 
-Consider the elements in definition \ref{def:valid-time-relation}. We will say that the tuple $t$ is current in the instance $r$ of the relation $R$ when $i = \left(s, UC \right)$.
+Consider the elements in definition \ref{def:valid-time-relation}. We will say that the tuple $t$ is current in the instance $r$ of the relation $R$ when $t[E]=UC$.
 \end{definition}
@@
  Consider a historical database containing diplomatic documents.
-The starting and the ending dates when the diplomatic document is valid.
+The starting and the ending dates say when the diplomatic document is valid.
 It is possible that a diplomatic document is valid for a period of time and several years later it becomes valid again.
-The following elements are stored: an identifier of the document (ID), the entity that issues the document and the dates when document is valid.
+The following elements are stored: an identifier of the document (ID), the entity that issues the document and the dates when the document is valid.
 Table \ref{tbl:library-sample} illustrates the first version of the database, after three insertions.
@@
 \emph{Current $\left(r,a_k \right)$.}
-Consider the elements in definition \ref{def:valid-time-relation}.
-The function Current $\left(r,a_k \right)$ returns the crisp time interval $i$ if the tuple $t$ with the primary key values given by $a_k$ is currently in the instance $r$ of the relation $R$.
-The function returns the empty set otherwise.
+Consider the elements in definition \ref{def:valid-time-relation}. The function Current $\left(r,a_k \right)$ returns the crisp time interval of a tuple $t$ with primary key $a_k$ as follows:
 % 
@@
 \nonumber
-t & \mbox{ if } \exists t \in r : i = \left(s, UC \right) \mbox{ and }\\
-& a_k = t[A_K] \\
+t[S,E] & \mbox{ if } \exists t \in r : t[E]=UC \mbox{ and }\\
+& t[A_K]=a_k \\
 \emptyset & \mbox{ otherwise}
@@
 
-For example, the function Current(r, (3,`E.U.')) returns the time interval $i = \left(4/4/2012 , UC  \right)$. Conversely, Current(r, (4,`N.A.T.O.')) returns the empty set.
+For example, the function Current(r, 3) returns the time interval $\left(4/4/2012 , UC  \right)$. Conversely, Current(r, (4,`N.A.T.O.')) returns the empty set.
 
@@
  \label{def:overlaps}
-Given two time intervals $i_1 = \left(s_1, e_1 \right)$ and $i_2 = \left(s_2 e_2\right)$, it is said that $i_1$ overlaps $i_2$ if:
-\begin{equation}
+Given two time intervals defined by the couples of values $i_1 = \left(s_1, e_1 \right)$ and $i_2 = \left(s_2, e_2\right)$, it is said that $i_1$ overlaps $i_2$ if:
+\begin{align}
  \label{eqn:overlaps}
-i_1 \text{ Overlaps } i_2  = \left(s_1 < s_2  \right) \wedge \left(e_1 < e_2  \right) 
-\end{equation}
+i_1 \text{ overlaps } i_2  = \left((\left(s_1 < s_2  \right) \wedge \left(e_1 < e_2  \right)\right)\\
+\nonumber
+\vee \left((\left(s_1 > s_2  \right) \wedge \left(e_1 > e_2  \right)\right)
+\end{align}
+
 \end{definition}
@@
  \label{def:during}
-Given two time intervals $i_1 = \left(s_1, e_1 \right)$ and $i_2 = \left(s_2 e_2\right)$, it is said that $i_1$ during $i_2$ if:
+Given two time intervals defined by the couples of values $i_1 = \left(s_1, e_1 \right)$ and $i_2 = \left(s_2, e_2\right)$, it is said that $i_1$ during $i_2$ if:
 \begin{align}
  \label{eqn:during}
-i_1\text{ During } i_2  =& \left(s_1 > s_2  \right) \wedge \left(e_1 > s_2 \right) \wedge \\
-\nonumber & \left(e_1 < e_2 \right)
+i_1\text{ during } i_2  =& \left(s_1 > s_2  \right) \wedge \left(e_1 < e_2 \right)
 \end{align}
@@
 \label{def:close-a-crisp-interval-r}
-Consider two crisp intervals $i_1 = \left(s_1, e_1 \right)$ and $i_2 = \left(s_2, e_2\right)$.
+Consider two crisp intervals defined by the couples of values $i_1 = \left(s_1, e_1 \right)$ and $i_2 = \left(s_2, e_2\right)$.
 The CloseR$\left(i_1, i_2 \right)$ function allows to close the right-open interval $i_1$ with respect to the first value $s_2$ in $i_2$:
@@
 %I=\left[a, c \right[ & \mbox{ if } b = +\infty \wedge J > I
-i_1 = \left(s_1,s_2 \right( & \mbox{ if } e_1 = UC \mbox{ and } i_2 \mbox{ During } i_1 \\
+\left(s_1,s_2-1\right) & \mbox{ if } e_1 = UC \mbox{ and } i_2 \mbox{ During } i_1 \\
 i_1 & \mbox{ otherwise}
@@
 Consider the elements in definition \ref{def:valid-time-relation}.
-The function Close-current$\left(r, t \right)$ closes any current version $t_k$ of the entity given by $t$ if it exists and add the new version $t$.
-For the implementation the following variables are defined:
+The function Close-current$\left(r, t \right)$ closes any current version $t_k$ of the entity given by $t$ and adds the new version $t$.
+For the implementation $t_{CUR}$ and $t_{UP}$ variables are defined:
 \begin{align}
  \label{eq:variables-current-version}
-t_{CUR} = \mbox{ Current } \left(r, t[A_K] \right)\\
-\nonumber
-t_{UP} = t_{CUR}; t_{UP}\left[i \right] = \mbox{Close }\left(t_{UP}\left[i \right], t \right)
-\nonumber
+t_{CUR}\left[A_k\right] &= t[A_K]\\\nonumber
+t_{CUR}\left[S,E \right] &= \mbox{Current } \left(r, t[A_K] \right)\\\nonumber
+t_{UP}[A_k] &= t[A_k]\\
+ t_{UP}\left[S,E \right] &= \mbox{CloseR }\left(t_{CUR}\left[S,E \right], t[S,E] \right)\nonumber
 \end{align}
@@
 r - t_{CUR} \cup \lbrace t_{UP}\rbrace \cup \lbrace t \rbrace
-& \mbox{ if } t_{CUR} \neq \emptyset \\
+& \mbox{ if } \mbox{Current } \left(r, t[A_K] \right) \neq \emptyset \\
 % R - \left(A, I \right) \cup \left \lbrace \left(A, \mbox{ CloseR } \left(I, J\right) \right)\right \rbrace \cup \left \lbrace\left(A, J\right) \right \rbrace  \\
@@
 % \mbox{ if  Current } \left(R, A \right) \neq \emptyset\\ %\wedge J > \mbox{ CloseR } \left(I, J \right)   \\
-\nonumber r , \text{ otherwise}
+r , \text{ otherwise}
 \end{cases}
@@
 \nonumber
+r & \mbox{ if }  \exists t_k \in V(t), \left(t[S,E] \mbox{ overlaps } t_k[S,E]\right)\\
+\nonumber
 r \cup \lbrace t \rbrace
-& \mbox{ if }  t \not \in r \mbox{ or }\\
-\nonumber & \forall t_k \in Vt, \lnot \left(i \mbox{ Overlaps } i_k\right)\\
+& \mbox{ if }  V(t)=\emptyset \mbox{ or }\\
+\nonumber & \forall t_k \in V(t), \lnot \left(t[S,E] \mbox{ overlaps } t_k[S,E]\right)\\
 % 
@@
 \DEL \left(r, t\right) =
-\begin{cases}
+%\begin{cases}
 \nonumber
-r - Vt
+r - V(t)
 % 
@@
 % R, & \mbox{ otherwise }  
-\end{cases} 	
+%\end{cases} 	
 \end{align}
-The set $Vt$ is computed as explained in definition \ref{def:valid-time-relation} and contains all the versions for the tuple $t$.
+The set $V(t)$ is computed as explained in definition \ref{def:valid-time-relation} and contains all the versions for the tuple $t$.
 

### 2012-08-01 ijcis-pos-model/sources/timeRepresentation.tex
preprint version

@@
 This section is devoted to specify the representation of time within the framework of the possibility theory.
-First of all, the specification for a single ill-known temporal point will be explained.
-Then, the formal specification and the related constraints are given for an ill-known valid-time interval.
+
+First of all, the specification for a single ill-known temporal point will be explained. Then, the formal specification and the related constraints are given for an ill-known valid-time interval.
 
 \subsection{\label{subsec:ill-known-point-rep}Ill-known time point}
-An ill-known time point $X$ is a precise time point that, for some reason, is not fully specified. Note that $X$ has only one possible value but that value is unspecified.
+An ill-known time point $X$ is an atomic time point that, for some reason, is not fully specified.
+
+Note that $X$ has only one possible value but that value is unspecified.
 
@@
 \textbf{Domain for an ill-known time point.}\\
-Consider $\Pow(\T)$ the set of all the possibility distributions over $\T$, and the three fuzzy constants \emph{UNKNOWN} $=\left\lbrace 1/t, \forall t \in \T \right\rbrace$, \emph{UNDEFINED} $=\left\lbrace 0/t,\forall t \in \T \right\rbrace$ and \emph{NULL} $=\lbrace 1/$ \emph{UNKNOWN}, $1/$ \emph{UNDEFINED} $\rbrace$.
+Consider $\tilde\Pow(\T)$ the set of all the possibility distributions over $\T$, and the three fuzzy constants:
+
+\begin{itemize}
+
+\item \emph{UNKNOWN} $=\left\lbrace 1/t, \forall t \in \T \right\rbrace$, ç
+\item \emph{UNDEFINED} $=\left\lbrace 0/t,\forall t \in \T \right\rbrace$ and \item \emph{NULL} $=\lbrace 1/$ \emph{UNKNOWN}, $1/$ \emph{UNDEFINED} $\rbrace$. 
+\end{itemize}
+
 The domain for an ill-known time point $X$ is given by: 
@@
 \label{eq:ill-known-domain}
-\Dom (X) =  \lbrace \Pow(\T) & \cup \mbox{\emph{UNKNOWN}} \\
+\Dom (X) =  \lbrace \tilde\Pow(\T) & \cup \mbox{\emph{UNKNOWN}} \\
 \nonumber
@@
 \subsubsection{\label{subsubsec:ill-known-time-datatypes}Datatypes}
-The data type for the representation of an ill-known time point allows the representation of the values shown in Table \ref{tbl:time-point-types}
+The data type for the representation of an ill-known time point allows the representation of the values shown in Table \ref{tbl:time-point-types}.
 %The datatypes for an  $X$ are shown in Table \ref{}.
@@
 Consider a historical database with data from medieval diplomatic documents.
+
 The following fields are stored: the digital identifier \emph{ID} which is the primary key  and the estimated time when the document was sent (field \emph{Date}).
+
 Table \ref{tbl:sample-time-point} contains some example data from this database. 
 \end{example}
-
-\vglue13pt
+\newpage
+%\vglue13pt
 %\begin{table}[htbp]
@@
 
-An ill-known time interval denoted by $\left[X, Y\right]$ is a precise time interval whose boundaries are not precisely known.
+An ill-known time interval denoted by $\left[X, Y\right]$ is a time interval whose boundaries are not precisely known.
 %The ill-known time interval is defined by two ill-known  time points, namely $X$,$Y$.
@@
 \begin{definition}
-\textbf{Semi-open time interval}: Only one of the two ill-known endpoints for the time interval $\left[X, Y\right]$ is unknown.
+\textbf{Semi-open time interval}: Only one of the two ill-known boundaries for the time interval $\left[X, Y\right]$ is unknown.
 Example \ref{ex:open-time-interval} and Figure \ref{fig:example-open-time-interval} illustrates a left open time interval.
@@
 $Open(C)$\\
-Consider an ill-known value $T$, a binary relationship $B_r \in \left \lbrace \leq, <, >, \geq\right \rbrace$ and the constraint $C\stackrel{\triangle}{=}\left(R,T\right)$.
+Consider an ill-known value $T$, a binary relationship $B_r \in \left \lbrace \leq, <, >, \geq\right \rbrace$ and the constraint $C\stackrel{\triangle}{=}\left(B_r,T\right)$.
 The function $\Open(C) = \left(I_p(C), I_n(C) \right)$ provides both possibility and necessity measures for all the points in the open part of a semi-open ill-known time interval.
@@
 \vspace*{10pt}
-\fcaption{Possibility distribution for $Y$, and possibility and necessity measures for the open ill-known point, $X$}
+\fcaption{\label{fig:example-open-time-interval}Possibility distribution for $Y$, and possibility and necessity measures for the open ill-known point, $X$}
 \vspace*{13pt}

### 2012-08-01 trunk/ijcis-pos-model/sources/preliminaries.tex
preprint version

@@
 %Introductory text:
-In this section, we introduce some basic concepts concerning possibilistic variables and fuzzy numbers and intervals.
-Then, the framework of set evaluation by ill-known constraints \cite{Pons2011} is explained.
-The section concludes with a brief introduction to temporal databases.
+In this section, the framework of set evaluation by ill-known constraints \cite{Pons2011} is explained. The section also includes a brief introduction to temporal databases.
 
@@
 
-\subsection{\label{subsec:possibility-theory}Possibility Theory}
-Possibility theory is devoted to the handling of incomplete information. In order to capture partial ignorance, two measures are provided: possibility and necessity.
-
-\begin{definition}
-Consider a set of outcomes $\Omega$.
-Let $\wp(\Omega)$ denote the powerset of $\Omega$ and let $A$ and $B$ be elements of $\wp(\Omega)$.
+%\subsection{\label{subsec:possibility-theory}Possibility Theory}
+%Possibility theory is devoted to the handling of incomplete information. In order to capture partial ignorance, two measures are provided: possibility and necessity.
+%
+%\begin{definition}
+%Consider a set of outcomes $\Omega$.
+Let $\Pow(\Omega)$ denote the powerset of $\Omega$ and let $A$ and $B$ be elements of $\Pow(\Omega)$.
 A \emph{confidence measure on $\Omega$} is defined by a function
-	\begin{align}
-	g : \wp(\Omega) & \rightarrow \left[0,1\right]
-	\end{align}
-that satisfies
-	\begin{align}
-	g(\emptyset) &= 0 \\
-	g(\Omega) &= 1 	\label{NormalizationProperty} \\
-	A \subseteq B &\Rightarrow g(A) \leq g(B) \label{MonotonicityProperty}
-	\end{align}
-\end{definition}
-
-Both possibility and necessity measures are special cases of confidence measures.
-
-\begin{definition}
-Consider a confidence measure $\Pi$ on a set of outcomes $\Omega$.
-Let $J$ be a countable index set and let $\{ A_{j} | j \in J \wedge A_{j} \subseteq \Omega \}$ be a family of elements of $\wp(\Omega)$.
+%	\begin{align}
+%	g : \Pow(\Omega) & \rightarrow \left[0,1\right]
+%	\end{align}
+%that satisfies
+%	\begin{align}
+%	g(\emptyset) &= 0 \\
+%	g(\Omega) &= 1 	\label{NormalizationProperty} \\
+%	A \subseteq B &\Rightarrow g(A) \leq g(B) \label{MonotonicityProperty}
+%	\end{align}
+%\end{definition}
+%
+%Both possibility and necessity measures are special cases of confidence measures.
+%
+%
+%
+%\begin{definition}
+%Consider a confidence measure $\Pi$ on a set of outcomes $\Omega$.
+Let $J$ be a countable index set and let $\{ A_{j} | j \in J \wedge A_{j} \subseteq \Omega \}$ be a family of elements of $\Pow(\Omega)$.
 $\Pi$ is now a \emph{possibility measure on $\Omega$} if it satisfies:
-	\begin{align}
-	\Pi\left(\bigcup_{j \in J} A_{j} \right) = \sup_{j \in J} \Pi(A_{j})
-	\end{align}
-\end{definition}
-
-In this work, the interpretation is as follows.
+%	\begin{align}
+%	\Pi\left(\bigcup_{j \in J} A_{j} \right) = \sup_{j \in J} \Pi(A_{j})
+%	\end{align}
+%\end{definition}
+%
+%In this work, the interpretation is as follows.
 The possibility of an event expresses how plausible the occurrence of the event seems to an observer of the experiment, given the (partial) knowledge of the observer about the experiment.
-
-Information on the possibility of distinct elements of the universe of discourse $\Omega$ can now be given by a \emph{possibility distribution} $\pi$ on $\Omega$, defined by:
-
-\begin{definition}
-Consider a possibility measure $\Pi$ on $\Omega$. A \emph{possibility distribution} $\pi$ on $\Omega$ underlying the possibility measure $\Pi$ is a function defined by:
-	\begin{align}
-	\pi : \Omega \rightarrow \left[0, 1\right] : \pi(u) = \Pi(\{u\})
-	\end{align}
-\end{definition}
-
-\begin{definition}
-Consider a confidence measure $N$ on $\Omega$.
-Let $J$ be a countable index set and let $\{ A_{j} | j \in J \wedge A_{j} \subseteq \Omega \}$ be a family of elements of $\wp(\Omega)$.
+%
+%Information on the possibility of distinct elements of the universe of discourse $\Omega$ can now be given by a \emph{possibility distribution} $\pi$ on $\Omega$, defined by:
+%
+%\begin{definition}
+%Consider a possibility measure $\Pi$ on $\Omega$. A \emph{possibility distribution} $\pi$ on $\Omega$ underlying the possibility measure $\Pi$ is a function defined by:
+%	\begin{align}
+%	\pi : \Omega \rightarrow \left[0, 1\right] : \pi(u) = \Pi(\{u\})
+%	\end{align}
+%\end{definition}
+%
+%\begin{definition}
+%Consider a confidence measure $N$ on $\Omega$.
+Let $J$ be a countable index set and let $\{ A_{j} | j \in J \wedge A_{j} \subseteq \Omega \}$ be a family of elements of $\Pow(\Omega)$.
 $N$ is a \emph{necessity measure} on $\Omega$ if it satisfies:
-	\begin{align}
-	N\left(\bigcap_{j \in J} A_{j} \right) = \inf_{j \in J} N(A_{j})
-	\end{align}
-\end{definition}
-
-In this work, the interpretation is as follows.
+%	\begin{align}
+%	N\left(\bigcap_{j \in J} A_{j} \right) = \inf_{j \in J} N(A_{j})
+%	\end{align}
+%\end{definition}
+%
+%In this work, the interpretation is as follows.
 The necessity of an event expresses how necessary the occurrence of the event seems to an observer of the experiment, given the (partial) knowledge of the observer about the experiment.
-
-Possibility and necessity measures are dual in the sense that:
-
-\begin{align}
-\forall A \subseteq \Omega : N(A) = 1 - \Pi(\bar{A})
-\end{align}
-
-That is, the degree to which an event is necessary is the extent to every other possible event is not plausible.
-
-\subsection{\label{subsec:possibilistic-variables}Possibilistic Variables}
-Possibilistic variables rely on possibility theory \cite{Dubois1988a} and are defined as follows \cite{Pons2011}.
-
-\begin{definition}
-\label{def;possibilistic-variable}
-A possibilistic variable $X$ over a universe $U$ is defined as a variable taking exactly one value in $U$, but for which this value is (partially) unknown.
+%
+%Possibility and necessity measures are dual in the sense that:
+%
+%\begin{align}
+%\forall A \subseteq \Omega : N(A) = 1 - \Pi(\bar{A})
+%\end{align}
+%
+%That is, the degree to which an event is necessary is the extent to which every other possible event is not plausible.
+%
+%\subsection{\label{subsec:possibilistic-variables}Possibilistic Variables}
+%Possibilistic variables rely on possibility theory \cite{Dubois1988a} and are defined as follows \cite{Pons2011}.
+%
+%\begin{definition}
+%\label{def;possibilistic-variable}
+%A possibilistic variable $X$ over a universe $U$ is defined as a variable taking exactly one value in $U$, but for which this value is (partially) unknown.
 Possibility distribution $\pi_X$ on $U$ models the available knowledge about the value $X$ takes; for each $u\in U$, $\pi_X(u)$ represents the possibility that $X$ takes the value $u$.
 In this work, this possibility is interpreted as a measure of how plausible it is that $X$ takes the value $u$, given (partial) knowledge about the value $X$ takes.
-\end{definition}
-
-The exact value a possibilistic variable takes, which is (partially) unknown, is called an \emph{ill-known value} in this work \cite{Dubois1988a}.
-
-When a possibilistic variable is defined on the powerset $\Pow(U)$ of some universe $U$, the unique value the variable takes will be a crisp set and its possibility distribution on the powerset $\Pow(U)$ will describe the possibility of each crisp subset of $U$ to be the value the variable takes.
+%\end{definition}
+%
+%The exact value a possibilistic variable takes, which is (partially) unknown, is called an \emph{ill-known value} in this work \cite{Dubois1988a}.
+%
+%When a possibilistic variable is defined on the powerset $\Pow(U)$ of some universe $U$, the unique value the variable takes will be a crisp set and its possibility distribution on the powerset $\Pow(U)$ will describe the possibility of each crisp subset of $U$ to be the value the variable takes.
 This exact value (a crisp set) the variable takes, is now called an \emph{ill-known set} \cite{Dubois1988a}.
-
-% It is important to understand the difference between the following two concepts:
-% \begin{itemize}
-% \item
-% A \emph{possibilistic variable} $X$ is bounded to take only one value , but this value is not known due to incomplete knowledge. 
-% \item
-% An \emph{ill-known set}~\cite{Dubois1988a}: a possibilistic variable defined over the universe $\Pow(U)$.
-% \end{itemize}
-
-Note that while a single ill-known value refers to one (partially) unknown value, an ill-known set is a crisp set but, for some reason, (partially) unknown.
-
-A specific application of possibilistic variables is obtained when the universe under consideration is the set of Boolean values $\mathbb{B}$ = $\{T,F\}$.
+%
+%% It is important to understand the difference between the following two concepts:
+%% \begin{itemize}
+%% \item
+%% A \emph{possibilistic variable} $X$ is bounded to take only one value , but this value is not known due to incomplete knowledge. 
+%% \item
+%% An \emph{ill-known set}~\cite{Dubois1988a}: a possibilistic variable defined over the universe $\Pow(U)$.
+%% \end{itemize}
+%
+%Note that while a single ill-known value refers to one (partially) unknown value, an ill-known set is a crisp set but, for some reason, (partially) unknown.
+%
+%A specific application of possibilistic variables is obtained when the universe under consideration is the set of Boolean values $\mathbb{B}$ = $\{T,F\}$.
 Indeed, any Boolean proposition $p$ takes just one value in $\mathbb{B}$.
@@
 The possibility and necessity that $p$ = $T$ (the proposition holds) demand more attention.
-This possibility and necessity is noted here as:
-\begin{align}
-\label{propholdsposs}
-\text{Possibility that $p$ = $T$ (p holds):} &\\
-\nonumber
-Pos(p) = \pi_p(T)  \\
-\label{propholdsnecc}
-\text{Necessity that $p$ = $T$ (p holds):} & \\
-\nonumber
-Nec(p) = 1-\pi_p(F) 
-\end{align}
-
-Here, equation \eqref{propholdsposs} denotes the possibility that $p$ = $T$ and the proposition holds, while equation \eqref{propholdsnecc} denotes the necessity that $p$ = $T$ and the proposition holds.
-
-This work deals with ill-known intervals.
-These are ill-known sets, defined and represented via a starting and ending points that, in turn are ill-known values.
+This possibility and necessity are noted here as:
+%\begin{align}
+%\label{propholdsposs}
+%\text{Possibility that $p$ = $T$ (p holds):} &\\
+%\nonumber
+%Pos(p) = \pi_p(T)  \\
+%\label{propholdsnecc}
+%\text{Necessity that $p$ = $T$ (p holds):} & \\
+%\nonumber
+%Nec(p) = 1-\pi_p(F) 
+%\end{align}
+%
+%Here, equation \eqref{propholdsposs} denotes the possibility that $p$ = $T$ and the proposition holds, while equation \eqref{propholdsnecc} denotes the necessity that $p$ = $T$ and the proposition holds.
+%
+%This work deals with ill-known intervals.
+These are ill-known sets, defined and represented via a starting and ending points that, in turn, are ill-known values.
 The elements of the set are the values between the starting and ending points.
-A closed ill-known interval with starting point defined by possibilistic variable $X$ and ending point by possibilistic variable $Y$ is noted here $\left[X, Y\right]$.
+A closed ill-known interval with starting point defined by possibilistic variable $X$ and ending point by possibilistic variable $Y$ is noted  $\left[X, Y\right]$ here.
 %The correspondences and transitions between the representations of ill-known sets, between the representations of ill-known intervals and between the representations of an ill-known set and an ill-known interval are part of the authors current research.
-
-\subsection{\label{subsec:fuzzy-numbers}Fuzzy Numbers and Fuzzy Intervals}
-Among others, Dubois and Prade~\cite{Dubois1983} use fuzzy sets \cite{Zadeh1965} to define a \emph{fuzzy interval}:
-\begin{definition}
-A fuzzy interval is a fuzzy set $M$, defined by a membership function $\mu_{M}$ on the set of real numbers $\mathbb{R}$ such that:
-\begin{eqnarray}
-\mu_{M} :  \mathbb{R} \rightarrow \left[0,1\right]  \\ 
-\nonumber
-\forall (u,v)\in\mathbb{R}^2: \forall w \in [u,v]:\\
-\nonumber
-\mu_M(w) \geq\min(\mu_M(u),\mu_M(v))  \\
-\nonumber
-\exists m \in \mathbb{R} :  \mu_M(m)=1 
-\end{eqnarray}
-\end{definition}
-
-\begin{samepage}
-\vspace*{13pt}
-\begin{center}
-{
-\includegraphics[scale=0.25]{./graphs/triangular.pdf}
-
-}
-\end{center}
-%\centerline{ \psfig{file=./graphs/Y-time-point.eps}}
-\vspace*{10pt}
-\fcaption{\label{fig:triangular-dist}Example of fuzzy number.}
-%  \label{fig:fuzzy-validity-period}
-\vspace*{13pt}
-\end{samepage}
-If this modal value $m$ is unique, then $M$ is referred to as a \emph{fuzzy number}. 
-%In other words, if the core of a fuzzy interval is a singleton, it is referred to as a fuzzy number (Figure \ref{fig:triangular-dist}).
-
-A simple form of the membership function of a fuzzy interval is a trapezoidal function (Figure \ref{fig:trapezoidal-dist}).
+%
+%\subsection{\label{subsec:fuzzy-numbers}Fuzzy Numbers and Fuzzy Intervals}
+%Among others, Dubois and Prade~\cite{Dubois1983} use fuzzy sets \cite{Zadeh1965} to define a \emph{fuzzy interval}:
+%\begin{definition}
+%A fuzzy interval is a fuzzy set $M$, defined by a membership function $\mu_{M}$ on the set of real numbers $\mathbb{R}$ such that:
+%\begin{eqnarray}
+%\mu_{M} :  \mathbb{R} \rightarrow \left[0,1\right]  \\ 
+%\nonumber
+%\forall (u,v)\in\mathbb{R}^2: \forall w \in [u,v]:\\
+%\nonumber
+%\mu_M(w) \geq\min(\mu_M(u),\mu_M(v))  \\
+%\nonumber
+%\exists m \in \mathbb{R} :  \mu_M(m)=1 
+%\end{eqnarray}
+%\end{definition}
+%
+%\begin{samepage}
+%\vspace*{13pt}
+%\begin{center}
+%{
+%\includegraphics[scale=0.25]{./graphs/triangular.pdf}
+%
+%}
+%\end{center}
+%%\centerline{ \psfig{file=./graphs/Y-time-point.eps}}
+%\vspace*{10pt}
+%\fcaption{\label{fig:triangular-dist}Example of fuzzy number.}
+%%  \label{fig:fuzzy-validity-period}
+%\vspace*{13pt}
+%\end{samepage}
+%If this modal value $m$ is unique, then $M$ is referred to as a \emph{fuzzy number}. 
+%%In other words, if the core of a fuzzy interval is a singleton, it is referred to as a fuzzy number (Figure \ref{fig:triangular-dist}).
+%
+%A simple form of the membership function of a fuzzy interval is a trapezoidal function (Figure \ref{fig:trapezoidal-dist}).
 It can be shown that such a membership function $\mu_T$ for a fuzzy interval $T$ is convex and normalized.
 The values $\left[\alpha, \beta, \gamma, \delta\right]$ represent a trapezoidal possibility distribution defined by $\mu_T$:
-
-% Four real values, denoted $\alpha$, $\beta$, $\gamma$ and $\delta$ suffice to represent a trapezoidal membership function of a fuzzy interval.
+%
+%% Four real values, denoted $\alpha$, $\beta$, $\gamma$ and $\delta$ suffice to represent a trapezoidal membership function of a fuzzy interval.
 In this work, a fuzzy interval defined as such will be noted as .
 The corresponding membership function definition for this  is then given by:
-
-\begin{align}
-\mu_T : & \quad \mathbb{R} \rightarrow \left[0,1\right] \\
-\nonumber
- : & \quad x \rightarrow
-\begin{cases}
-1 & \mbox{ if } x \in [\beta,\gamma] \\
-0 & \mbox{ if } x > \delta \vee x < \alpha \\
-\frac{x-\alpha}{\beta - \alpha} & \mbox{ if } x \in [\alpha,\beta[ \\
-\frac{\delta -x}{\delta - \gamma} & \mbox{ if } x \in ]\gamma,\delta] \\
-\end{cases}
-\end{align}
-
-\begin{samepage}
-\vspace*{13pt}
-\begin{center}
-{
-\includegraphics[scale=0.25]{./graphs/trapezoidalDistribution.pdf}
-
-}
-\end{center}
-%\centerline{ \psfig{file=./graphs/Y-time-point.eps}}
-\vspace*{10pt}
-\fcaption{\label{fig:trapezoidal-dist}Example of fuzzy interval.}
-%  \label{fig:fuzzy-validity-period}
-\vspace*{13pt}
-\end{samepage}
-
-%\def\JPicScale{0.5}
-%\begin{figure}[h!]
-%  \centering
-%  \input{./graphs/trapezoidal.tex}
-%  \caption{Trapezoidal membership function}
-%  \label{fig:trapezoidal}
-%\end{figure}
-
-The representation for a triangular possibility distribution is given by three values $\left[D,a,b \right]$.
+%
+%\begin{align}
+%\mu_T : & \quad \mathbb{R} \rightarrow \left[0,1\right] \\
+%\nonumber
+% : & \quad x \rightarrow
+%\begin{cases}
+%1 & \mbox{ if } x \in [\beta,\gamma] \\
+%0 & \mbox{ if } x > \delta \vee x < \alpha \\
+%\frac{x-\alpha}{\beta - \alpha} & \mbox{ if } x \in [\alpha,\beta[ \\
+%\frac{\delta -x}{\delta - \gamma} & \mbox{ if } x \in ]\gamma,\delta] \\
+%\end{cases}
+%\end{align}
+%
+%\begin{samepage}
+%\vspace*{13pt}
+%\begin{center}
+%{
+%\includegraphics[scale=0.25]{./graphs/trapezoidalDistribution.pdf}
+%
+%}
+%\end{center}
+%%\centerline{ \psfig{file=./graphs/Y-time-point.eps}}
+%\vspace*{10pt}
+%\fcaption{\label{fig:trapezoidal-dist}Example of fuzzy interval.}
+%%  \label{fig:fuzzy-validity-period}
+%\vspace*{13pt}
+%\end{samepage}
+%
+%%\def\JPicScale{0.5}
+%%\begin{figure}[h!]
+%%  \centering
+%%  \input{./graphs/trapezoidal.tex}
+%%  \caption{Trapezoidal membership function}
+%%  \label{fig:trapezoidal}
+%%\end{figure}
+%
+%
+%The representation for a triangular possibility distribution is given by three values $\left[D,a,b \right]$.
 The distribution is obtained from a trapezoidal possibility distribution by replacing the values $\left[D-a, D, D, D+b \right]$.
-
-% The most convenient form of the membership function of a fuzzy number is a triangular form.
+%
+%
+%% The most convenient form of the membership function of a fuzzy number is a triangular form.
 It can be shown that such a membership function $\mu_M$ for a fuzzy number $M$ is convex and normalized.
@@
 Here:
-% \begin{itemize}
-% \item
-% $D$ denotes the single value in the core of $M$
-% \item
-% $D-a$ is then $\inf \{u \in \mathbb{R} : \mu_{M}(u) > 0\}$
-% \item
-% $D+b$ is then $\sup \{u \in \mathbb{R} : \mu_{M}(u) > 0\}$
-% \end{itemize}
+%% \begin{itemize}
+%% \item
+%% $D$ denotes the single value in the core of $M$
+%% \item
+%% $D-a$ is then $\inf \{u \in \mathbb{R} : \mu_{M}(u) > 0\}$
+%% \item
+%% $D+b$ is then $\sup \{u \in \mathbb{R} : \mu_{M}(u) > 0\}$
+%% \end{itemize}
+%% 
+%% The membership function of such a fuzzy number is then given by:
+%% 
+%% \vspace{-10pt}
+%% 
+%% \begin{align}
+%% \mu_M :   \mathbb{R} \rightarrow \left[0,1\right] :  x \rightarrow \nonumber\\
+%% \begin{cases}
+%% \nonumber
+%% 0 & \mbox{ if } \left(x < D-a\right) \vee \left(x > D+b\right) \\
+%% \frac{\left[x-\left(D-a\right)\right]}{a} & \mbox{ if } \left(x \geq D-a\right) \wedge (x \leq D)  \\
+%% -\frac{\left[x-\left(D+b\right)\right]}{b} & \mbox{ if } \left(x \geq D\right) \wedge (x \leq D+b)
+%% \end{cases}
+%% \end{align}
 %
-% The membership function of such a fuzzy number is then given by:
+%%\begin{figure}[h!]
+%%  \centering
+%%  \input{./graphs/triangular.tex}
+%%  \caption{Triangular membership function.}
+%%  \label{fig:triangular}
+%%\end{figure}
 %
-% \vspace{-10pt}
-% 
-% \begin{align}
-% \mu_M :   \mathbb{R} \rightarrow \left[0,1\right] :  x \rightarrow \nonumber\\
-% \begin{cases}
-% \nonumber
-% 0 & \mbox{ if } \left(x < D-a\right) \vee \left(x > D+b\right) \\
-% \frac{\left[x-\left(D-a\right)\right]}{a} & \mbox{ if } \left(x \geq D-a\right) \wedge (x \leq D)  \\
-% -\frac{\left[x-\left(D+b\right)\right]}{b} & \mbox{ if } \left(x \geq D\right) \wedge (x \leq D+b)
-% \end{cases}
-% \end{align}
-
-%\begin{figure}[h!]
-%  \centering
-%  \input{./graphs/triangular.tex}
-%  \caption{Triangular membership function.}
-%  \label{fig:triangular}
-%\end{figure}
 
 \subsection{\label{subsec:interval-evaluation-by-ill-known-constraints}Interval Evaluation by Ill-known Constraints}
-In our case, we need to know if all points in a crisp interval $I$ reside between the boundaries of an ill-known interval $\left[ X , Y \right]$.
+In our case, we need to know if all points in a crisp interval $I$ reside between the boundaries of an ill-known interval $\left[ X , Y \right]$, where X e Y are possibilistic variables represented by triangular possibility distributions \[D,a,b\], where D is the modal value and a and b are the left and right spreads respectively.
 To compute that, we introduce the concept of \emph{ill-known constraint}\cite{Pons2011}.
@@
 \begin{definition}
-Given a universe $U$, an ill-known constraint $C$ on a set $A \subseteq U$ is specified by means of a binary relation $R \subseteq U^{2}$ and a fixed ill-known value denoted by its possibilistic variable $V$ over $U$, i.e.:
+Given a universe $U$, an ill-known constraint $C$ on a set $A \subseteq U$ is specified by means of a binary relation $R \subseteq  \tilde\Pow(U)^{2}$ and a fixed ill-known value denoted by its possibilistic variable $V$ over $U$, i.e.:
 \begin{align}
@@
 \vspace*{10pt}
-\fcaption{\label{fig:example-ill-known}Example of the evaluation of the ill-known constraint $C \triangleq (\leq, X), X = \left[5, 3, 2\right]$. Possibility and necessity measures are shown in grey. }
+\fcaption{\label{fig:example-ill-known}Example of the evaluation of the ill-known constraint $C \triangleq (\leq, X), X = \left[5, 3, 2\right]$.  }
 %  \label{fig:fuzzy-validity-period}
@@
 Consider a universe $U$.
-Let $R_1, R_2$ be two binary relations in U.
+Let $R_1, R_2$ be two binary relations in  $\tilde\Pow(U)$.
 Let $X_1, X_2$ be two fixed ill-known values in $U$.
@@
 Sometimes, an ill-known value might be specified by a convex combination of ill-known constraints.
-This allow to define ill-known values by means of relationships with respect to other ill-known points.
+This allows to define ill-known values by means of relationships with respect to other ill-known points.
 
@@
 \label{eq:ill-known-value-def-by-const-unc}
-X \in \Pow(U):\pi_{X}=\pi_{CC}
+X \in \tilde\Pow(U):\pi_{X}=\pi_{CC}
 \end{align}
@@
 A new diplomatic document was valid in the year $Y = \left[1118, 2, 1 \right]$.
-Then, It is possible to obtain $Z$ (the period of time between the starting of both documents) by using a conjunctive combination of constraints.
+Then, it is possible to obtain $Z$ (the period of time between the starting of both documents) by using a conjunctive combination of constraints.
 

### 2012-08-01 trunk/ijcis-pos-model/sources/temporalModel.tex
preprint version

@@
 The first subsection is devoted to the formalization of the model.
-Then, the data manipulation language is defined.
+Then, a data manipulation language is defined.
 
@@
 The model is based on GEFRED~\cite{Medina1994} (Generalized Model of Fuzzy Relational DB) model.
-This model is extended by adding valid-time support which will be illustrated through definitions and examples.
+This model is extended by adding valid-time support which will be illustrated through the following definitions and examples.
 The information in the system is defined by the following elements:
@@
 \emph{Generalized fuzzy temporal domain.}
-Consider $\T$ to be the temporal domain, and let $\tilde \Pow\left( \T\right)$ be the set of all \emph{normalized} possibility distributions (see Section \ref{subsec:possibility-theory}, equation \eqref{NormalizationProperty}) defined on $\T$.
+Consider $\T$ to be the temporal domain, and let $\tilde \Pow\left( \T\right)$ be the set of all \emph{normalized} possibility distributions defined on $\T$.
 The Generalized Fuzzy Temporal Domain, $\T_G$ is
@@
        \label{eq:fuzzy-valid-time-instance}
-      r \subseteq D_1\ x\ \ldots\ x\ D_n
+      r \subseteq D_1\ x\ \ldots\ x\ D_n x\ \mathbb{N}  x\  \T_G\  x\ \T_G 
       \end{equation}
@@
 
-We will illustrate the definitions with an example.
-
-\begin{example}
-Consider the set of attributes $A = \left \lbrace A_1, A_2, A_3 \right \rbrace$.
-The primary key for these attributes is given by $A_K = \left \lbrace A_1, A_2 \right \rbrace$.
-Let $I = \left \lbrace S, E \right \rbrace$ be the set of temporal attributes that define the possibilistic validity period of the data.
-$R_{FTG} = A \cup V_{ID} \cup I$ is the valid-time relation and $r$ is an instance of the relation.
-The instance $r$ is given by the following elements.
-$r = \left \lbrace \left(a_{11}, a_{12}, a_{13}, 001, s_1 ,e_1 \right) \right.$,  $\left(a_{21}, a_{22}, a_{23}, 001, s_2, e_2 \right)$ , $\left.
-\left(a_{11}, a_{12}, a_{31}, 002, s_3, e_3 \right) \right \rbrace$.
-The instance $r$ is illustrated in Table \ref{tbl:fuzzy-sample-definitions}.
-Consider the tuple $t = \left(a_{11}, a_{12}, a_{13}, s_1 ,e_1 \right)$. Then,
-\begin{align}
- \nonumber
-s_1 &= t\left[S \right]\\
- \nonumber
-e_1 &= t[E]\\
- \nonumber
-i\ \ &= \left(s_1, e_1\right)\\
- \nonumber
-a_k &= t\left[A_K\right] = \left(a_{11}, a_{12} \right) \\
- \nonumber
-k &= t\left[PK\right] =\left(a_{11}, a_{12}, s_1, e_1\right)\\
- \nonumber
-Vt &= r\left(t\left[A_K\right] \right) = \left \lbrace t_1, t_3 \right \rbrace
-\end{align}
-
-\end{example}
+Table \ref{tbl:fuzzy-sample-definitions} contains an example instance.
 
@@
 \end{equation}
-Where $\Head$ is the Head of the relation and consist on a fixed set of triplets attribute- domain - compatibility with an optional the valid-time attribute:
+Where $\Head$ is the Head of the relation and consists on a fixed set of triplets attribute- domain - compatibility with an optional the valid-time attribute:
 
@@
 
-$\Body$ is the body of the relation and it consists on a set of tuples. Each tuple is a triplet attribute- value- degree with an optional valid-time attribute:
+$\Body$ is the body of the relation and it consists on a set of tuples. Each tuple is a set of triplets attribute- value- degree with an optional valid-time attribute:
 
@@
 \label{eq:body-valid-time}
-\Body = \big \lbrace \left(A_{G1}:\tilde{d}_{i1}\left[,c_{i1} \right] \right),\\
+\Body = \big \lbrace \ldots \{ \left(A_{G1}:\tilde{d}_{i1}\left[,c_{i1} \right] \right),\\
 \nonumber
@@
   \nonumber
-   \Big[  \left( \text{PVP}, \tilde{d}_{\text{PVP}} \left[,C_{A_{\text{PVP}}} \right] \right)  \Big] \big \rbrace
+   \Big[  \left( \text{PVP}, \tilde{d}_{\text{PVP}} \left[,C_{A_{\text{PVP}}} \right] \right)  \Big] \} \ldots \big \rbrace
 \end{align}
@@
 \nonumber
-\Head^{v} = \left \lbrace A_{G1}:D_{G1}, \ldots,  A_{Gn}:D_{Gn} \right \rbrace \\
+\Head^{v} = \left \lbrace (A_{G1}:D_{G1}), \ldots,  (A_{Gn}:D_{Gn}) \right \rbrace \\
 \nonumber
-\Body^ {v} = \left \lbrace A_{G1}:\tilde{d}_{i1}, \ldots,  A_{Gn}:\tilde{d}_{in}\ \right \rbrace 
+\Body^ {v} = \left \lbrace (A_{G1}:\tilde{d}_{i1}), \ldots,  (A_{Gn}:\tilde{d}_{in})\ \right \rbrace 
 \end{align}
@@
 \nonumber
-\Body^ {v} &= \left \lbrace 3:, \text{``E.U.''}: ,  \left[15/3/1012, 30/3/1012\right] \right \rbrace 
+\Body^ {v} &= \{\ldots \{ (\text{ID}:3), (\text{Entity}:\text{``E.U.''}) ,\\
+\nonumber
+& t(\text{PVP}:\left[15/3/1012,5,2\right], \left[30/3/1012,1,1\right]) \} \ldots\}
 \end{align}
@@
 \nonumber
-\Head^{t} = & \lbrace \left( \text{PVP}, D_{\text{PVP}}\right),\\
+\Head^{t} = & \lbrace \left( \text{PVP}, D_{\text{PVP}}\right)\}\\
 \nonumber
-\Body^ {t} =& \left \lbrace 4/4/1012 , UC \right \rbrace 
+\Body^ {t} =& \{ \ldots (\text{PVP}:[\left[15/3/1012,5,2\right],\\
+\nonumber
+& \left[30/3/1012,1,1\right]]) \ldots\} 
 \end{align}
@@
 A generalized fuzzy temporal key, $K_{GT}$ is a subset of the head.
+
 \begin{align}
 \label{eq:generalized-fuzzy-temporal-key}
-K_{GT} \subseteq \Head, K_{GT} = \left \lbrace  \left(A_{Gs}:D_{Gs} \right) \right.  \\
+K_{GT} \subseteq \Head, K_{GT} = \left \lbrace \left(A_{Gs}:D_{Gs} \right)  \right \rbrace \\
 \nonumber
- \left. \cup  \left(V_{ID}:D_{ID} \right)	\right \rbrace \\
+\cup  \{\left(V_{ID}:\mathbb{N}  \right)\}\\
 \nonumber
-\text{Typeof }\left(D_{ID} \right) = \\
-\nonumber
-s\in S \subseteq \left(1, \ldots, n \right) \\
+s\in S \subseteq \left(1, \ldots, n \right)
 \end{align}
@@
 \nonumber
-\forall i, i' \in \left \lbrace 1, \ldots, m\right \rbrace , \exists s \in S: \\
+\forall i, i' \in \left \lbrace 1, \ldots, m \right \rbrace , \exists s \in S: \\
 \nonumber
@@
 The operations implemented are: \emph{Insert, Modify} and \emph{Delete}.
-The semantics of the operations will be the same that those defined for a crisp temporal database, whereas the temporal representation is made by the possibilistic valid-time period and the ill-known constraints (see sections \ref{sec:prelim} and \ref{sec:time-rep}.
+The semantics of the operations will be the same that those defined for a crisp temporal database, whereas the temporal representation is made by the possibilistic valid-time period and the ill-known constraints (see sections \ref{sec:prelim} and \ref{sec:time-rep}).
 %The algebra for the data manipulation is defined in .
@@
 It is important to notice that while the result of the evaluation of any comparison between crisp time intervals is boolean, the evaluation of any comparison between \emph{PVPs} is a value in the unit interval.
-Therefore, for a crisp valid-time relation, it is not possible an overlapping among any valid-time interval.
-In other words, a crisp temporal database is always strictly consistent.
-Thus, from the point of view of a fuzzy temporal database, there are the following choices (see Figure \ref{fig:tdb-class}):
-
-% \begin{itemize}
-% \item 
-\begin{definition}
-\textbf{Strictly consistent} It is not possible that two versions of an entity to be valid at the same time.
-Consider a tuple $t$ in the instance $r$ of the fuzzy valid-time relation $R_{FTG}$ and  $Vt = \left \lbrace t_1, \ldots, t_n \right \rbrace$ as defined in \ref{def:fuzzy-temporal-relation}.
-Then, a fuzzy temporal database is strictly consistent if:
-\begin{align}
-\label{eq:stricly-consistent}
-\forall k,j \in \left \lbrace 1, \ldots, n \right \rbrace, k < j\\
-\nonumber t_k \mbox{ Overlaps } t_j \leq 1
-\end{align}
-\end{definition}
-
-%\item 
-\begin{definition}
- \textbf{Ill-consistent} The overlapping of two PVPs of the same entity do overlap. It can be distinguished among the following sub-types:
-\end{definition}
-% 	\begin{itemize}
-% 	\item 
-	      \begin{definition}
-	      \textbf{Co-existence} Two versions of the same entity may exist at the same time.
-		      \begin{align}
-		      \forall k, \exists !j \in \left \lbrace 1, \ldots, n \right \rbrace, k < j\\
-		      \nonumber t_k \mbox{ Overlaps } t_j \geq 1
-		      \end{align}
-	      \end{definition}
-
-	%\item 
-	      \begin{definition}
-	      \textbf{Weak-consistence} Several versions of the same entity may exist at the same time.
-		      \begin{align}
-		      \label{eq:weak-consistence}
-		      \forall k, \exists j \in \left \lbrace 1, \ldots, n \right \rbrace, k < j\\
-		      \nonumber t_k \mbox{ Overlaps } t_j \geq 1
-		      \end{align}
-	      \end{definition}
-
-	%\end{itemize}
-
-\begin{samepage}
-\vspace*{13pt}
-\begin{center}
-{
-\includegraphics[scale=1.25]{./graphs/tdb-clas.pdf}
-
-}
-\end{center}
-%\centerline{ \psfig{file=./graphs/Y-time-point.eps}}
-\vspace*{10pt}
-\fcaption{\label{fig:tdb-class}Classification for a fuzzy temporal database.
-The validity for an entity is represented by a trapezoid.
-Graphic (a) shows an example of strictly consistent database.
-Between two versions of the entity it is possible some amount of overlapping, but always holds equation \eqref{eq:stricly-consistent}.
-Graphic (b) shows an example of a co-existence.
-Two versions of the same entity may overlap in any degree, but not more than two versions.
-Graphic (c) shows an example of weak-consistence.
-The different versions of the same entity are shown in an unfolded way, to clarify the multiple overlapping.
-}
-\vspace*{13pt}
-\end{samepage}
-
-%\end{itemize}
-
-In the implementation of the DML operations we will consider a strictly consistent temporal database.
-This is a natural extension of the crisp temporal databases.
-This approach allows an implementation for the DML operations defined in Section \ref{subsubsec:Understanding-valid-time-databases}.
 
@@
 \textbf{CloseR$\left(i_1, i_2\right)$.}
-Consider the elements in definition \ref{def:generalized-fuzzy-temporal-relation}.
+Consider the elements in definition \ref{def:generalized-fuzzy-temporal-relation} and ill-known intervals given by $i_1=(s_1,e_1)$ and $i_2=(s_2,e2)$.
 The CloseR function closes the PVP given by $i_1$ with a conjunctive combination of ill-known constraints (see section \ref{subsec:interval-evaluation-by-ill-known-constraints}):
@@
 \nonumber
-i_1 = \left(s_1, e_z \right) & \mbox{ if } i_1 = (s_1, UC), \\
+\left(s_1, e_z \right) & \mbox{ if } i_1 = (s_1, UC), \\
 & e_z \triangleq \left \lbrace C_1\left(>, s_1 \right), C_2\left(<, s_2 \right) \right \rbrace \\
@@
 The result for CloseR$\left(i_1, i_2\right)$  is $i_1 = \left[ \left[4/4/1012, 3, 3\right], \left[4,7,13,15 \right] \right]$.
-The value that closes $i_1$ is a trapezoid in the form $\left[ \alpha, \beta, \gamma, \delta \right]$ as defined in Section \ref{subsec:fuzzy-numbers}.
+The value that closes $i_1$ is a trapezoid in the form $\left[ \alpha, \beta, \gamma, \delta \right]$.
 
@@
 If the time interval $i$ does not overlap any other valid-time interval in the relation $R_{FTG}$.
-Note that here, the result of the Overlaps operator is in the unit interval.
+Note that here, the result of the overlaps operator is in the unit interval.
 For example, the document with ID=3 is still valid. The insert sentence is the following.
@@
 
-	\item Reject the insertion, if the time interval $i$ do overlap with a degree of 1 any existing valid-time interval for the entity in the relation.
+	\item Reject the insertion, if the time interval $i$ does overlap with a degree of 1 any existing valid-time interval for the entity in the relation.
 For example, consider now that the document manager wants to introduce a past validity period for the document with ID=3.
@@
 \label{eq:insert-fuzzy-temporal}
-\INS \left(r, t\right) &=\\
+\INS \left(r, t\right) =\\
 \begin{cases}
 \nonumber
+r & \mbox{ if }  \exists t_k \in V(t), \\
+\nonumber &\left(t[S,E] \mbox{ overlaps } t_k[S,E]=1\right)\\
+\nonumber
 r \cup \lbrace t \rbrace
-& \mbox{ if }  t \not \in r \mbox{ or }\\
-\nonumber & \forall t_k \in Vt,  \left(i \mbox{ Overlaps } i_k\right) < 1\\
+& \mbox{ if }  V(t)=\emptyset \mbox{ or }\\
+\nonumber & \forall t_k \in V(t), \\
+\nonumber & \left(t[S,E] \mbox{ overlaps } t_k[S,E]\right) < 1\\
 % 
@@
 \end{align}
+
 \subsubsection{\label{subsubsec:delete-fuzzy-temporal}Delete}
@@
 \DEL \left(r, t\right) =
-\begin{cases}
 \nonumber
-r - Vt
-\end{cases} 	
+r - V(t)
 \end{align}

### 2012-08-27 book/sources/timeDomain.tex
My last (implemented) changes on the chapter.

@@
 
-Data models used by information systems (and in specific, temporal database systems) may represent these underlying time domains using \emph{chronons}~\cite{Dyreson1994}, which can informally be described as the smallest distinguishable time units that can be used in the system.
+Data models used by information systems (and in specific, temporal database systems) may represent an underlying time axis using \emph{chronons}~\cite{Dyreson1994}, which can informally be described as the smallest distinguishable time units that can be used in the system.
 However, to explain what chronons are, an explanation of some other temporal concepts is necessary.
@@
 
-Thus, an instant is basically an instantaneous time point in a time model.
+Thus, an instant is basically an instantaneous time point on the time axis underlying a time model. The term is used in the context of the time model too.
 
-Orthogonal to the classification of underlying time models as linear or cyclic, they can be classified as \emph{discrete}, \emph{dense} or \emph{continuous} models\cite{Dyreson1994}.
-In a discrete model~\cite{Clifford:1985:AHR:971699.318922}, the notion exists that every instant has a unique successor and the set of instants is seen as a discrete one.
+Orthogonal to the classification of time models as linear or cyclic, they can be classified as \emph{discrete}, \emph{dense} or \emph{continuous} models\cite{Dyreson1994},~\cite{Jensen:Dyreson:1998:TheConsensusGlossary}.
+In a discrete model~\cite{Clifford:1985:AHR:971699.318922}, the notion exists that every instant has a unique successor and the set of (modelled) instants is seen as a discrete one.
 Here, intuitively, the set of instants can be seen as isomorphic to the set of natural numbers $\Nat$.
 In a dense model, the notion exists that between any two instants always lies another.
-Here, intuitively, the set of instants can be seen as isomorphic to the set of rational numbers $\Q$ (when the set of instants is a discrete one) or the set of real numbers $\R$ (when the set of instants is a continuous one).
-In a continuous model, the notion also exists that between any two instants always lies another one, but the set of instants is always seen as continuous and there are no ``gaps'' between successive instants.
+Here, intuitively, the set of instants can be seen as isomorphic to the set of rational numbers $\Q$ (when the set of (modelled) instants is a discrete one) or the set of real numbers $\R$ (when the set of (modelled) instants is a continuous one).
+In a continuous model, the notion also exists that between any two instants always lies another one, but the set of (modelled) instants is always seen as continuous and there are no ``gaps'' between successive instants.
 
@@
 
-Data models for time modelling might now represent an underlying time domain using chronons:
+Data models used for time modelling might now represent an underlying time axis using chronons:
 
@@
 
-A data model may now represent a time model or time domain by a sequence of consecutive chronons.
+A time model contained in a data model may now represent an underlying time axis by a sequence of consecutive chronons.
 These chronons have identical durations.
@@
 The fact that chronons are actually time intervals has a particular effect on the representation of instants and time intervals.
-In a data model representing a time domain using chronons, an instant is of course represented by a chronon.
+In a time model using chronons, an instant is of course represented by a chronon.
 A time interval may be represented by a set of contiguous chronons, depending on the amount of time the time interval comprises.
@@
 \includegraphics[scale=0.5]{graphs/granularityGraph.eps}
-\caption{The granularity graph corresponding to example~\ref{examplegrangraph}}
+\caption{The granularity graph corresponding to example~\ref{examplegrangraph}.}
 \label{fig:granularity-graph-example}
@@
 
-Several (groups of) operators have been proposed in order to compare temporal elements and model the temporal relationships between them.
+Several (collections of) operators have been proposed in order to compare temporal elements and model the temporal relationships between them.
 Allen~\cite{Allen83} most notably described such relationships between time intervals and as a special case, between instants.

### 2012-09-03 trunk/book/sources/fuzzyTimeDomain.tex
update
bibliography update and new allen's relation graph.

@@
 %INTRO
-As explained in the introduction, humans handle temporal information using temporal notions like time intervals or time points \cite{Dyreson1994}.
+As explained in the introduction, humans handle temporal information using temporal notions like time intervals or time points~\cite{Dyreson1994}.
 While the used temporal notions may contain imperfections~\cite{Dev98},~\cite{Dubois:jucs_9_9:fuzziness_and_uncertainty_in},~\cite{nagypal2003},~\cite{Dubois89}, humans often gracefully deal with these, as their inherent interpretation capability accounts for a lot of them.
@@
 In this section, some main concepts and issues concerning this combination are presented.
-
-%\subsection{\label{subsec:classification}Classification of imperfect temporal information}
 
@@
 \item \emph{Possibility theory}:
-Using possibility theory~\cite{Dubois:Prade:1988:PossibilityTheory}, a possibility degree is associated to the temporal fact or event.
+Using possibility theory~\cite{DidierDubois1988a}, a possibility degree is associated to the temporal fact or event.
 Possibility theory is widely used to model uncertainty and vaguenes in time~\cite{Dubois:jucs_9_9:fuzziness_and_uncertainty_in},\cite{Dubois89},\cite{devos94},\cite{nagypal2003}.
@@
 The aim of these works is generally to obtain a flexible way to compare uncertain, ill-known temporal intervals by means of temporal relationships.
+The study of imperfect temporal metadata is done in \cite{Bordogna2009}, \cite{Bordogna2011}.
+In ~\cite{Soysangwarn2009} a proposal to use in fuzzy databases temporal fuzzy linguistic terms is studied.
+Burney ~\cite{burney2010a}, \cite{burney2010b} has studied recently the combination of fuzzy databases with temporal data.
 
@@
 The two dimensional representation of time intervals and the temporal relationships between them has been studied in~\cite{Qia09}.
+In ~\cite{Burney2009} a rough set-based model for temporal databases is presented.
+The study of temporal relationships between rough time interval is studied in ~\cite{Bassiri2009}.
 \end{itemize}
@@
 These approaches are often based on Allen's operators~\cite{Allen83}.
-Example \ref{ex:fuzzy-allen-relation} presents a short example concerning one of Allen's relationships.
+Example~\ref{ex:fuzzy-allen-relation} presents a short example concerning one of Allen's relationships.
 
@@
 Consider an event which takes place between time points $A$ and $B$.
-Thus, the event comprises time interval $[A,B]$ (this is visualized in part (1) of figure \ref{fig:fuzzy-allen-relationship}).
-The classical Allen relationship `after' returns an interval $]B,\infty[$ as shown in part (2) of figure \ref{fig:fuzzy-allen-relationship}.
-A fuzzified version of Allen's `after' operator is illustrated in part (3) of figure \ref{fig:fuzzy-allen-relationship}.
+Thus, the event comprises time interval $[A,B]$ (this is visualized in part (1) of figure~\ref{fig:fuzzy-allen-relationship}).
+The classical Allen relationship `after' returns an interval $]B,\infty[$ as shown in part (2) of figure~\ref{fig:fuzzy-allen-relationship}.
+A fuzzified version of Allen's `after' operator is illustrated in part (3) of figure~\ref{fig:fuzzy-allen-relationship}.
 The comparison between two time intervals results in a possibility degree in the unit interval.
-The shape of the possibility distribution is shown in part (3) of figure \ref{fig:fuzzy-allen-relationship}.
+The shape of the possibility distribution is shown in part (3) of figure~\ref{fig:fuzzy-allen-relationship}.
 Note that all the points strictly after the point $B$ results in a possibility degree of 1 whereas there is an area near the point $B$ in which the possibility degree runs smoothly between 0 and 1.
 
-Consider now the interval given by $[C,D]$, illustrated in part (4) of figure \ref{fig:fuzzy-allen-relationship}.
+Consider now the interval given by $[C,D]$, illustrated in part (4) of figure~\ref{fig:fuzzy-allen-relationship}.
 The user wants to know if $[C,D]$ is after $[A,B]$.

### 2012-09-07 temporaldml/main.tex
first draft of the temporal dml

\documentclass{llncs}
%
%
%espero que no se enfaden por incluir este paquete:
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[dvips]{graphicx}
\usepackage{makeidx}  % allows for indexgeneration
\usepackage{color}

\hyphenation{pro-per-ties}
\hyphenation{ge-ne-ral-ly}
\hyphenation{pre-fe-ren-ces}
\hyphenation{u-sing}
\hyphenation{pu-nish-ment}
\newcommand{\Pow}{\mathcal{P}}
\newcommand{\N}{\operatorname{N}}
\newcommand{\bool}{\operatorname*{\mathcal{B}}}
\newcommand{\Pos}{\operatorname{Pos}}
\newcommand{\Nec}{\operatorname{Nec}}
\newcommand{\Open}{\operatorname{Open}}
\newcommand{\Rp}{\operatorname{Rp}}
\newcommand{\Rn}{\operatorname{Rn}}
\newcommand{\FB}{\operatorname{FB}}
\newcommand{\UC}{\operatorname{UC}}
\newcommand{\T}{\mathcal{T}}
\newcommand{\Rel}{\mathcal{R}}
\newcommand{\C}{\mathcal{C}}
\newcommand{\Nat}{\mathbb{N}}
\newcommand{\Boolean}{\mathbb{B}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\Head}{\mathcal{H}}
\newcommand{\Body}{\mathcal{B}}
\newcommand{\Dom}{\mathcal{D}}
\newcommand{\INS}{\mbox{\textbf{insert}}}
\newcommand{\DEL}{\mbox{\textbf{delete}}}
\newcommand{\MOD}{\mbox{\textbf{modify}}}
\newcommand{\REV}{\mbox{\textbf{revise}}}
\newcommand{\Overlap}{\mbox{\textbf{Overlap}}}
\newcommand{\A}{\mathcal{A}}
\newcommand{\I}{\mathcal{I}}
\begin{document}

\mainmatter              % start of the contributions
%
\title{Data Manipulation Language for fuzzy temporal databases.}
%
%\titlerunning{A Fuzzy Valid-Time Model}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{Jos\'e Enrique Pons \and Olga Pons Capote}
%Jeffrey Dean \and David Grove \and Craig Chambers \and Kim~B.~Bruce \and
%Elsa Bertino}
%
\authorrunning{Jos\'e Enrique Pons et al.} % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
%\tocauthor{Ivar Ekeland, Roger Temam, Jeffrey Dean, David Grove,
%Craig Chambers, Kim B. Bruce, and Elisa Bertino}
%
\institute{
Department of Computer Science and Artificial Intelligence \\
Universidad de Granada \\
Escuela T\'ecnica Superior de Ingenier\'ia Inform\'atica \\
C/Periodista Daniel Saucedo Aranda s/n \\
E-18071 (Granada-Spain)\\
\email{jpons,opc@decsai.ugr.es}\\ 
%WWW home page:
%\texttt{http://users/\homedir iekeland/web/welcome.html}
%\and
%Universit\'{e} de Paris-Sud,
%Laboratoire d'Analyse Num\'{e}rique, B\^{a}timent 425,\\
%F-91405 Orsay Cedex, France
}

\maketitle              % typeset the title of the contribution

\begin{abstract}

\end{abstract}

\section{\label{sec:selection}Selection}
The selection operator $\sigma$ obtain a subset of tuples that fulfill a set of constraints from a given relation $R$. The set of constraints is usually a boolean combination of atomic constraints. The selection operator is noted as follows:
\begin{equation}
 \label{eq:selection}
\sigma_{S} \left( R \right)
\end{equation}

Where $R$ is the relation, and $S$ is the selection formula. The selection formula is a set of two elements:
% \begin{itemize}
% \item relation.attribute boolean relation relation.attribute
% \item relation.temporal-attribute allen-relation relation.temporal-attribute 
% \end{itemize}
\begin{equation}
 \label{eq:selection_formula}
S = \left \lbrace Q, Q^{t}\right \rbrace
\end{equation}

Where $q$ is a boolean combination of (possibly) fuzzy (and bipolar) non-temporal constraints and $q^{t}$ is also a boolean combination of fuzzy temporal constraints.

\begin{equation}
 \label{eq:non-temporal-constraints}
Q = \left \lbrace q^{\left \lbrace \mbox{pos,neg} \right \rbrace}_{\mbox{attribute}_1}  \theta \mbox{ val }, \ldots, q^{\left \lbrace \mbox{pos,neg} \right \rbrace}_{\mbox{attribute}_n}  \theta \mbox{ val } \right \rbrace
\end{equation}
Where $\theta$ is an artihmetic operator; one of $\left \lbrace =, \neq, <, \leq, >, \geq \right \rbrace$ and val is a value in the domain of the attribute.

Then, consider a boolean function $\bool:\Boolean^{n}  \rightarrow \Boolean$. The evaluation function for the constraints in $Q$  is given by:

\begin{equation}
 \label{eq:evaluation-function}
\lambda : Q \rightarrow \Boolean : \lambda (Q) \rightarrow \bool \left(q^{\left \lbrace \mbox{pos,neg} \right \rbrace}_{\mbox{attribute}_1}  \theta \mbox{ val }, \ldots, q^{\left \lbrace \mbox{pos,neg} \right \rbrace}_{\mbox{attribute}_n}  \theta \mbox{ val } \right)
\end{equation}

Now, the uncertainty about the evaluation of $Q$ and a boolean function $\bool:\Boolean^{n}  \rightarrow \Boolean$ is given by:

\begin{equation}
 \label{eq:evaluation-lambda-function}
\pi_{\lambda \left(Q \right)} = \tilde{\bool} \left(\pi_{q^{\left \lbrace \mbox{pos,neg} \right \rbrace}_{\mbox{attribute}_1}  \theta \mbox{ val }}, \ldots, \pi_{q^{\left \lbrace \mbox{pos,neg} \right \rbrace}_{\mbox{attribute}_n}  \theta \mbox{ val }} \right)
\end{equation}

The temporal constraints, $Q^t$ are provided in a similar way. The main difference is that, instead of comparations like $\left \lbrace =, \neq, <, \leq, >, \geq \right \rbrace$, we use the allen's relations. Hence, the representation of the temporal constraints is expressed as follows:

\begin{equation}
\label{eq:temporal-constratins}
Q^{t} = \left \lbrace q_{\mbox{attribute}_1}  ar \mbox{ val }, \ldots, q_{\mbox{attribute}_n}  ar \mbox{ val } \right \rbrace
\end{equation}
Here, $ar$ is one of the thirdteen Allen's relations (equal, before, overlaps, starts, finishes, meets, during) and their respective reverses.

Now the combination of temporal and non-temporal results should be explicited here.

\section{\label{sec:join}Join}
The join operator $\Join$ builds a new relation from two given relations, namely $R$ and $S$. This new relation is a set with all the possible combinations of tuples in both $R$ and $S$. It is usually noted as $R \Join_{r \theta s} S$, where $r,s$ are attributes from $R$ and $S$ respectively and $\theta$ an arithmetic operator.

The following example illustrates the behaviour of the join operator in a crisp temporal database.

\begin{example}
Consider the employees database shown in Table \ref{}.  
\end{example}

\section*{Acknowledgements}
%
The researchers are supported by the grant BES-2009-013805 within the research project TIN2008-02066: \emph{Fuzzy Temporal Information treatment in relational DBMS}.

\bibliographystyle{splncs03}
\bibliography{biblio}

\end{document}

### 2012-09-14 temporaldml/main.tex
new version

@@
 \usepackage{amssymb}
+%\usepackage{amsthm}
+
 \usepackage[dvips]{graphicx}
@@
 \usepackage{color}
+\usepackage[table]{xcolor}
 
@@
 \newcommand{\I}{\mathcal{I}}
+\renewcommand{\Join}{\bowtie}
+%\theoremstyle{definition}
+%\newtheorem{definition}{Definition}
 \begin{document}
@@
 
+\section{\label{sec:preliminaries}Preliminares}
+
+% \begin{definition}
+%  \label{def:temporal-schema}
+% A crisp temporal schema is given by the following:
+% 
+% \begin{equation}
+% \label{eq:temporal-relation}
+% R = \left(A_1, \ldots, A_n, T_s, T_e \right) 
+% \end{equation}
+% Where $A_1, \ldots, A_n$ are the attributes of the relation $R$ and $T = \left \lbrace T_s, T_e\right \rbrace$ is the staring and ending points of the valid-time interval.
+% We define $r$ as an instance of the relation $R$.
+% \end{definition}
+
+\begin{definition}
+\label{def:valid-time-relation}
+\emph{Valid-time relation.}
+Consider the following definitions and notations:
+
+\begin{itemize}
+ \item A set of non-temporal attributes.
+	\begin{equation}
+	\label{eq:attribute-set}
+	A = \left \lbrace A_1, A_2, \ldots, A_n \right \rbrace
+	\end{equation}
+      The domain for each attribute $A_1, \ldots, A_n$ is $D_1, \ldots, D_n$ respectively. 
+\item The original primary key $A_K$ is a subset of the attributes in $A$.
+      \begin{equation}
+       \label{eq:primary-key-a}
+      A_K \subseteq A
+      \end{equation}
+\item Two attributes, $S$  and $E$ for the starting and ending points respectively. $I$ defines the valid time interval for the data. 
+\begin{equation}
+ \label{eq:attribute-time-interval}
+I = \left \lbrace S, E \right \rbrace
+\end{equation}
+
+$\T$ is the time domain.
+
+\item Then $R$, the schema for the valid-time relation is:
+\begin{equation}
+ \label{eq:valid-time-relation}
+R = A \cup  I
+\end{equation}
+\item The primary key for the valid-time relation $R$ is:
+\begin{equation}
+ \label{eq:valid-time-temporal-pk}
+PK = A_K \cup I
+\end{equation}
+
+\item We will note by $r$ any valid instance of $R$. 
+      \begin{equation}
+       \label{eq:valid-time-instance}
+      r \subseteq D_1\ x\ \ldots\ x\ D_n x\ T x\ T
+      \end{equation}
+
+     % \end{itemize}
+\item $V(t)$ is the set of all the versions for a given tuple $t$. Formally,
+
+\begin{equation}
+ \label{eq:all-the-versions}
+V(t) =\left \lbrace t_i \in r, t_i\left[A_K\right] = t\left[A_K\right] \right \rbrace
+\end{equation}
+Obviously, $t$ itself is included in the set.
+% If a tuple $t$ contains more than one version, the result of $Vt$ is a set.
+We will use the index $k$ to address the elements of the set.
+E.g.
+$Vt = \left \lbrace t_1, t_2, \ldots, t_n \right \rbrace$.
+Then $t_k , k \in \left \lbrace 1, \ldots, n \right \rbrace$ is an element of $Vt$.
+\end{itemize}
+\end{definition}
+We will illustrate the definitions with an example.
+
+\begin{example}
+Consider the set of attributes $A = \left \lbrace A_1, A_2, A_3 \right \rbrace$.
+The primary key for these attributes is given by $A_K = \left \lbrace A_1, A_2 \right \rbrace$.
+Let $I = \left \lbrace S, E \right \rbrace$ be the set of temporal attributes that define the validity period of the data.
+$R$ is the valid-time relation and $r$ is an instance of the relation.
+The instance $r$ is given by the following elements.
+$r = \left \lbrace \left(a_{11}, a_{12}, a_{13}, s_1 ,e_1 \right) \right.$,  $\left(a_{21}, a_{22}, a_{23}, s_2, e_2 \right)$ , $\left.
+\left(a_{11}, a_{12}, a_{31}, s_3, e_3 \right) \right \rbrace$.
+The instance $r$ is illustrated in Table \ref{tbl:sample-definitions}.
+Consider the tuple $t_1 = \left(a_{11}, a_{12}, a_{13}, s_1 ,e_1 \right)$. Then,
+\begin{align}
+ \nonumber
+ t_1\left[S \right]&=s_1\\
+ \nonumber
+t_1[E]&=e_1\\
+ \nonumber
+t_1[S,E]&= \left(s_1, e_1\right)\\
+ \nonumber
+t_1\left[A_K\right] &= \left(a_{11}, a_{12} \right) \\
+ \nonumber
+t_1\left[PK\right] &=\left(a_{11}, a_{12}, s_1, e_1\right)\\
+ \nonumber
+V(t_1) &= \left \lbrace t_1, t_3 \right \rbrace
+\end{align}
+
+\begin{table}
+\centering
+\label{tbl:sample-definitions}
+\caption{Sample database containing the instance $r$ of the valid-time relation $R$.}
+\begin{tabular}{c c c c c c}\\
+\hline
+& \textbf{A$_1$}  & \textbf{A$_2$}  & $A_3$ & $S$ & $E$ \\
+\hline
+$t_1$&$a_{11}$ & $a_{12}$ & $a_{13}$ & $s_1$ & $e_1$ \\
+$t_2$ & $a_{21}$ & $a_{22}$ & $a_{23}$ &  $s_2$ & $e_2$ \\
+$t_3$ & $a_{11}$ & $a_{12}$ & $a_{31}$ & $s_3$ & $e_3$\\
+\hline\\
+\end{tabular}
+\end{table}
+
+\end{example}
+
+% \begin{definition}
+%  \label{def:fuzzy-temporal-schema}
+% A fuzzy temporal schema is given by the following:
+% 
+% % \begin{equation}
+% %  \label{eq:fuzzy-temporal-relation}
+% % R = \left(A_1, \ldots, A_n, T_s, T_e \right) 
+% % \end{equation}
+% % Where 
+%  \end{definition}
+
+\begin{definition}
+\label{def:generalized-fuzzy-temporal-domain}
+\emph{Generalized fuzzy temporal domain.}
+Consider $\T$ to be the temporal domain, and let $\tilde \Pow\left( \T\right)$ be the set of all \emph{normalized} possibility distributions defined on $\T$.
+The Generalized Fuzzy Temporal Domain, $\T_G$ is
+\begin{equation}
+\T_G \subseteq \left \lbrace \tilde \Pow\left( \T\right) \cup \text{NULL} \right \rbrace
+\end{equation}
+\end{definition}
+
+Note that $\T_G \subseteq D_G$.
+The datatypes for this domain have been studied previously in section \ref{sec:time-rep} and are shown in tables \ref{tbl:time-point-types} and \ref{tbl:time-interval-types}.
+
+A generalized fuzzy relation is defined in \cite{Medina1994}. Here, we will extend the definition to a generalized fuzzy temporal relation.
+
+\begin{definition}
+\emph{Generalized fuzzy temporal relation.}
+\label{def:fuzzy-temporal-relation}
+Consider the elements in definition \ref{def:valid-time-relation}. Some of them will be extended for the fuzzy case.
+
+\begin{itemize}
+%  \item A set of non-temporal fuzzy  or crisp attributes.
+% 	\begin{equation}
+% 	\label{eq:fuzzy-attribute-set}
+% 	A = \left \lbrace A_1, A_2, \ldots, A_n \right \rbrace
+% 	\end{equation}
+%       The domain for each attribute $A_1, \ldots, A_n$ is $D_1, \ldots, D_n$ respectively. 
+% \item The primary key $A_K$ is a subset of $A$.
+%       \begin{equation}
+%        \label{eq:fuzzy-primary-key-a}
+%       A_K \subset A
+%       \end{equation}
+% A formal definition of primary key for fuzzy relational databases will be given later in Definition \ref{def:generalized-primary-key}.
+% \item A set of two attributes; $S$  and $E$ the attributes for the starting and ending ill-known points respectively.
+$I$ is a possibilistic validity period \emph{PVP} as explained in Section \ref{subsec:ill-known-interval}.
+% \begin{equation}
+%  \label{eq:fuzzy-attribute-time-interval}
+% I = \left \lbrace S, E \right \rbrace
+% \end{equation}
+\item An attribute called version identifier, $V_{ID}$, will be added to the schema. This attribute is a counter for each different version of the entities. 
+% \begin{equation}
+%  \label{eq:fuzzy-version-identifier}
+% V_{ID} \subset \N
+% \end{equation}
+
+\item Then $R_{FTG}$, the schema for the fuzzy valid-time relation is:
+\begin{equation}
+ \label{eq:fuzzy-valid-time-relation}
+R_{FTG} = A \cup V_{ID} \cup  I
+\end{equation}
+\item The primary key for the fuzzy valid-time relation $R_{FTG}$ is:
+\begin{equation}
+ \label{eq:fuzzy-valid-time-temporal-pk}
+K_{GT} = A_K \cup V_{ID}
+\end{equation}
+A formal definition of the primary key for fuzzy valid-time relations will be given later in Definition \ref{def:generalized-fuzzy-temporal-key}.
+
+\item We will note by $r$ any valid instance of $R_{FTG}$. 
+      \begin{equation}
+       \label{eq:fuzzy-valid-time-instance}
+      r \subseteq D_1\ x\ \ldots\ x\ D_n x\ \mathbb{N}  x\  \T_G\  x\ \T_G 
+      \end{equation}
+
+% \item Let $t$ be a tuple in the instance $r$; $t \in r$:
+%       \begin{itemize}
+%       \item We will note the values for the starting and the ending  points,$s$ and $e$  respectively. The value for the time interval is given by $i$.
+%       \begin{align}
+%        \label{eq:fuzzy-starting-point}
+%       s = t\left[S \right]\\
+%       e = t\left[E \right]\\
+%       i = \left(s, e\right)
+%       \end{align}
+% 
+% 
+%       \item Let $A_k \subset A$ be the set of the non-temporal attributes that are part of the primary key (equation \eqref{eq:fuzzy-primary-key-a}).
+Then, $a_k$ denotes the values for the non-temporal attributes of the primary key.
+% 	    \begin{equation}
+% 	     \label{eq:fuzzy-pk-attribute} 
+% 	      a_k = t\left[A_K \right]
+% 	    \end{equation}
+% 
+    \item Let $K_{GT}$ be the primary key for the valid-time relation as given in equation \eqref{eq:fuzzy-valid-time-temporal-pk}. Then, $k$ denotes the values for the attributes in the primary key.
+	  \begin{equation}
+	   \label{eq:fuzzy-value-pk}
+	  k = t\left[K_{GT} \right]
+	  \end{equation}
+% 
+% 
+%       \end{itemize}
+% \item $Vt$ is the set with all the versions for a given tuple $t$.
+% 
+% \begin{equation}
+%  \label{eq:fuzzy-all-the-versions}
+% Vt = r\left(t\left[A_K\right] \right)
+% \end{equation}
+% If a tuple $t$ contains more than one version, the result of $Vt$ is a set.
+We will use the index $k$ to address the elements of the set.
+E.g.
+$Vt = \left \lbrace t_1, t_2, \ldots, t_n \right \rbrace$.
+Then $t_k , k \in \left \lbrace 1, \ldots, n \right \rbrace$ is an element of $Vt$.
+\end{itemize}
+%\end{definition}
+
+Table \ref{tbl:fuzzy-sample-definitions} contains an example instance.
+
+\begin{table}
+\caption{\label{tbl:fuzzy-sample-definitions}Sample database containing the instance $r$ of the fuzzy valid-time relation $R_{FTG}$.}
+\centering
+%\centerline{\small DATA TYPES}
+\begin{tabular}{c c c c c c c}\\
+\hline
+& \textbf{A$_1$}  & \textbf{A$_2$}  & $A_3$ & $V_{ID}$ & $S$ & $E$ \\
+\hline
+$t_1$&$a_{11}$ & $a_{12}$ & $a_{13}$ & $001$ & $s_1$ & $e_1$ \\
+$t_2$ & $a_{21}$ & $a_{22}$ & $a_{23}$& $001$ &  $s_2$ & $e_2$ \\
+$t_3$ & $a_{11}$ & $a_{12}$ & $a_{31}$& $002$ & $s_3$ & $e_3$\\
+\hline\\
+\end{tabular}
+\end{table}
+
+% 
+% 
+% Consider $\A$ to be the set of all the entities, and let $A=\left(A_1, \ldots, A_n \right), A \in \A$ be the set of (fuzzy) attributes that define an entity.
+Let $\I_{PVP}$ be the set of all the ill-known time intervals and let $I = \left(X, Y \right)$ be a PVP, $I \in \I_{PVP}$.
+The pair $\left(A, I\right)$ expresses that the data regarding the entity A are valid during the ill-known time interval I.
+Let $R_{FTG} \subseteq \A \  x\  \I_{PVP}$ be a valid-time relation.
+Then, the following equation indicates that the pair $\left(A, I\right)$ is in the relation R:
+% 
+% \begin{equation}
+% \label{eq:fuzzy-rel-def}
+% \left( R_{FTG}, \left(A , I\right) \right)
+% \end{equation}
+% 
+ A generalized fuzzy temporal relation $R_{FTG}$ can be noted also by:
+\label{def:generalized-fuzzy-temporal-relation}
+\begin{equation}
+\label{eq:generalized-fuzzy-temporal-relation}
+R_{FTG} = \left(\Head, \Body \right)
+\end{equation}
+Where $\Head$ is the Head of the relation and consists on a fixed set of triplets attribute- domain - compatibility with an optional the valid-time attribute:
+
+\begin{align}
+\label{eq:head-valid-time}
+\Head = \big \lbrace \left(A_{G1}:D_{G1}\left[,C_{A_{G1}} \right] \right),\\
+\nonumber
+ \ldots,\\
+ \nonumber
+  \left(A_{Gn}:D_{Gn}\left[,C_{A_{Gn}} \right] \right),\\
+  \nonumber
+  \Big[  \left( \text{PVP}, D_{\text{PVP}}\left[,C_{A_{\text{PVP}}} \right] \right) \Big] \big \rbrace
+\end{align}
+Note that $D_{Gj}$ ($j = 1, \ldots, n$) is the domain for the attribute $A_{Gj}$. $C_{A_{Gj}}$ is the compatibility attribute in the unit interval $\left[0, 1 \right]$.
+
+$\Body$ is the body of the relation and it consists on a set of tuples. Each tuple is a set of triplets attribute- value- degree with an optional valid-time attribute:
+
+\begin{align}
+\label{eq:body-valid-time}
+\Body = \big \lbrace \ldots \{ \left(A_{G1}:\tilde{d}_{i1}\left[,c_{i1} \right] \right),\\
+\nonumber
+ \ldots,\\
+ \nonumber
+  \left(A_{Gn}:\tilde{d}_{in}\left[,c_{in} \right] \right),\\
+  \nonumber
+   \Big[  \left( \text{PVP}, \tilde{d}_{\text{PVP}} \left[,C_{A_{\text{PVP}}} \right] \right)  \Big] \} \ldots \big \rbrace
+\end{align}
+
+\end{definition}
+
 \section{\label{sec:selection}Selection}
@@
  \label{eq:selection}
-\sigma_{S} \left( R \right)
+\sigma_{P} \left( R \right)
 \end{equation}
 
-Where $R$ is the relation, and $S$ is the selection formula. The selection formula is a set of two elements:
+Where $R$ is the relation, and $P$ is the selection formula. The selection formula is a set of two elements:
 % \begin{itemize}
@@
  \label{eq:selection_formula}
-S = \left \lbrace Q, Q^{t}\right \rbrace
+P = \left \lbrace Q, Q^{t}\right \rbrace
 \end{equation}
@@
 Now the combination of temporal and non-temporal results should be explicited here.
+\begin{example}
+Consider an employee's database.
+The data are stored in two tables; Table \ref{table:employees} (relation $R$) contains temporal data about the employees and Table \ref{table:address} (relation $S$) contains temporal  data about the addresses for each emploree.
+Consider now that the user wants to obtain all the employees who are a profesor and that worked during the time interval $\left[7, 10 \right]$.
+%\vspace{-10pt} \ref{table:employees} \ref{table:address}
+
+\begin{table}
+\centering
+\caption{Employee's database. Relation $R$.}
+
+\begin{tabular}{c c c c c c }
+\hline
+ID & Job & Works for & Start & Finish \\ \hline
+1 & Prof. & 4 & 5 & 10 \\
+2 & Tech. & 4 & 3 & 7 \\
+3 & Account. & 4  & 4 & 10 \\
+4 & Adm. & - & 1 & UC \\
+1 & Prof. & 4 & 11 & UC \\
+\hline 
+\end{tabular}
+\label{table:employees}
+
+%\vspace{10pt}
+
+\end{table}
+
+%\vspace{-25pt}
+
+%\vspace{-10pt}
+
+\begin{table}
+\centering
+\caption{Address database. Relation $S$.}
+
+\begin{tabular}{c c c c }
+\hline
+ID & Adr. & Start & Finish \\ \hline
+1 & C/ Camino de ronda & FB & 12 \\
+2 & C/ Recogidas & FB & UC \\ 
+3 & C/ Pintor Maldonado & FB & UC \\
+4 & C/ Mesones & FB & UC \\
+1 & C/ Manuel de Falla & 12 & UC \\
+\hline 
+\end{tabular}
+\label{table:address}
+
+%\vspace{10pt}
+
+\end{table}
+
+%\vspace{-25pt}
+
+The selection formula is the following:
+
+\begin{equation}
+ \label{eq:selection-example}
+\sigma_{r.Job = 'Profesor', r.T \mbox{ During } \left[7, 10\right]} \left(R \right)
+\end{equation}
+The resultset for the selection is shown in Table \ref{table:example-selection}.
+\end{example}
+
+%\vspace{-10pt}
+
+\begin{table}
+\centering
+\caption{Resultset table for the selection in equation \eqref{eq:selection-example}}
+
+\begin{tabular}{c c c c c c }
+\hline
+ID & Job & Works for & Start & Finish \\ \hline
+1 & Prof. & 4 & 5 & 10 \\
+\hline 
+\end{tabular}
+\label{table:example-selection}
+
+%\vspace{10pt}
+
+\end{table}
+
+%\vspace{-25pt}
+
+\section{\label{sec:cartesian-product}Cartesian Product}
+The relational cartesian product is defined as all the possible combinations between two given relations $R$ and $S$. It is usually noted as $R \times S$.
+The temporal cartesian product operator is defined as the relational cartesian product with a predicate on the temporal valid-time intervals \cite{DengfengGao2002}.
+The temporal cartesian product is defined then as $R \times^T S$.
+
+We will re-define the predicate given in \cite{DengfengGao2002} in terms of the Allen's relations. Two auxiliary functions should be defined.
+
+\begin{definition}
+ \label{def:crisp-intersect}
+Intersect$\left(I, J \right)$.
+Given two intervals $I$ and $J$, the Intersect function returns a boolean value showing whether the two intervals intersect or not.
+In terms of the Allen's relations, two intervals $I$ and $J$ intersect if between them exist any of the Allen's relations except Before or After.
+In other words, only if the Allen's relations Before or After do not hold for the two intervals, they will intersect.
+
+\begin{equation}
+ \label{eq:crisp-intersect}
+\mbox{Intersect}\left( I, J \right) = \left \lbrace \neg \left(I_e < J_s \right) \vee \neg \left(J_e < I_s \right) \right \rbrace
+\end{equation}
+
+\end{definition}
+
+\begin{example}
+Consider the following intervals $I = \left[5, 10 \right]$ and  $J = \left[0, 12 \right]$. The instersect operation is calculated as follows:
+\begin{align}
+\mbox{Intersect}\left( I, J \right) &=& \left \lbrace \neg \left(10 < 0 \right)  \vee \neg \left(12 < 5 \right) \right \rbrace \\
+\nonumber
+&=& \mbox{True} \vee \mbox{True} \\
+\nonumber &=& \mbox{True}
+\end{align}
+
+\end{example}
+
+The other auxiliary function that allows to define the predicate for the cartesian product, needs some previous definitios as well.
+
+\begin{definition}
+ \label{def:crisp-last}
+Last$\left(I_n, J_m \right)$. Given two time points $I_n$ and $J_m$, this function obtains the greatest time point.
+\begin{equation}
+ \label{eq:last-crisp}
+\mbox{Last}\left(I_n, J_m \right) = 
+\begin{cases}
+I_n & \mbox{ if } I_n > J_m \\
+J_m & \mbox{ otherwise } 
+\end{cases}
+\end{equation}
+\end{definition}
+
+\begin{definition}
+ \label{def:crisp-first}
+First$\left(I_n, J_m \right)$. Given two time points $I_n$ and $J_m$, this function obtains the least time point.
+\begin{equation}
+ \label{eq:first-crisp}
+\mbox{First}\left(I_n, J_m \right) =
+\begin{cases}
+I_n & \mbox{ if } I_n < J_m \\
+J_m & \mbox{ otherwise } 
+\end{cases}
+\end{equation}
+\end{definition}
+
+\begin{definition}
+ \label{def:crisp-overlapping-interval}
+Overlapping-interval$\left(I, J \right)$. Given two intervals $I$ and $J$, this function returns an overlapping interval from the original two intervals.
+\begin{equation}
+ \label{eq:crisp-overlapping-interval}
+\mbox{OvInt}\left(I, J \right) = 
+\begin{cases}
+\left[\mbox{Last}\left( I_s, J_s \right), \mbox{First}\left(I_e, J_e \right) \right] & \mbox{ if } \mbox{Last}\left( I_s, J_s \right) \leq \mbox{First}\left(I_e, J_e \right) \\
+\emptyset & \mbox{ otherwise } 
+\end{cases}
+\end{equation}
+\end{definition}
+
+Now it is possible to define the temporal cartesian product.
+
+\begin{definition}
+ \label{def:crisp-temporal-cartesian-product}
+Temporal Cartesian Product.\\
+The temporal cartesian product of two temporal relations $r$ and $s$ is noted as $r \times^T s$ and is defined by the following formula:
+
+\begin{align}
+ \label{eq:crisp-temporal-cartesian-product}
+r \times^T s =  \lbrace z^{\left(n+m+2 \right)} | \exists x \in r, \exists y \in s \\
+\nonumber
+\mbox{intersect}\left(x[T], y[T] \right) \wedge \\
+\nonumber
+z\left[A\right] = x\left[A\right] \wedge z\left[B\right] = y \left[B \right] \wedge \\
+\nonumber
+ z\left[T \right] = \mbox{OvInt}\left(x\left[T\right], y\left[T\right] \right) \wedge z\left[T\right] \neq \emptyset  \rbrace
+\end{align}
+
+\end{definition}
+
+The following example illustrates the temporal cartesian product.
+\begin{example}
+ \label{ex:crisp-temporal-cartesian-product}
+Consider the employee's database mentioned before (Tables \ref{table:employees} and \ref{table:address}).
+Table \ref{table:example-crisp-cartesian-product} illustrates an intermediate step to compute the temporal cartesian  product.
+
+\end{example}
+
+\begin{table}
+\centering
+\caption{Intermediate calculations for the temporal cartesian product.}
+
+\begin{tabular}{c c c c c c c c}
+\hline
+r.id & s.id & $x\left[T\right]$ & $y\left[T\right]$ & intersect$\left(x\left[T\right], y\left[T\right]\right)$ & First$\left(I_e, J_e \right)$& Last$\left( I_s, J_s \right)$ &$z\left[T \right] = $OvInt$\left(x\left[T\right], y\left[T\right] \right) $  \\ \hline
+1 & 1 & $\left[5, 10 \right]$  & $\left[FB, 11 \right]$  & True  & 11 &  5 & $\left[5, 11 \right]$ \\
+1 & 2 & $\left[5, 10 \right]$  & $\left[FB, UC \right]$  & True  & UC &  5 & $\left[5, UC \right]$ \\
+1 & 3 & $\left[5, 10 \right]$  & $\left[FB, UC \right]$  & True  & UC &  5 & $\left[5, UC \right]$ \\
+1 & 4 & $\left[5, 10 \right]$  & $\left[FB, UC \right]$  & True  & UC &  5 & $\left[5, UC \right]$ \\
+1 & 1 & $\left[5, 10 \right]$  & $\left[12, UC \right]$  & False & UC & 12 & -                     \\
+2 & 1 & $\left[3, 7 \right]$   & $\left[FB, 11 \right]$  & True  & 11 &  3 & $\left[3, 11 \right]$ \\
+2 & 2 & $\left[3, 7 \right]$   & $\left[FB, UC \right]$  & True  & UC &  3 & $\left[3, UC \right]$ \\
+2 & 3 & $\left[3, 7 \right]$   & $\left[FB, UC \right]$  & True  & UC &  3 & $\left[3, UC \right]$ \\
+2 & 4 & $\left[3, 7 \right]$   & $\left[FB, UC \right]$  & True  & UC &  3 & $\left[3, UC \right]$ \\
+2 & 1 & $\left[3, 7 \right]$   & $\left[12, UC \right]$  & False & UC & 12 &  -                    \\
+3 & 1 & $\left[4, 10 \right]$  & $\left[FB, 11 \right]$  & True  & 11 &  4 & $\left[4, 11 \right]$ \\
+3 & 2 & $\left[4, 10 \right]$  & $\left[FB, UC \right]$  & True  & UC &  4 & $\left[4, 11 \right]$ \\
+3 & 3 &  $\left[4, 10 \right]$ & $\left[FB, UC \right]$  & True  & UC &  4 & $\left[4, 11 \right]$ \\
+3 & 4 & $\left[4, 10 \right]$  & $\left[FB, UC \right]$  & True  & UC &  4 & $\left[4, 11 \right]$ \\
+3 & 1 & $\left[4, 10 \right]$  & $\left[12, UC \right]$  & False & 12 & 12 & -                     \\
+4 & 1 & $\left[1, UC \right]$   & $\left[FB, 11 \right]$ & True  & UC &  1 & $\left[1, UC \right]$ \\
+4 & 2 & $\left[1, UC \right]$   & $\left[FB, UC \right]$ & True  & UC &  1 & $\left[1, UC \right]$ \\
+4 & 3 & $\left[1, UC \right]$   & $\left[FB, UC \right]$ & True  & UC &  1 & $\left[1, UC \right]$ \\
+4 & 4 & $\left[1, UC \right]$   & $\left[FB, UC \right]$ & True  & UC &  1 & $\left[1, UC \right]$ \\
+4 & 1 & $\left[1, UC \right]$   & $\left[12, UC \right]$ & True  & UC & 12 & $\left[12, UC \right]$ \\
+1 & 1 & $\left[11, UC \right]$  & $\left[FB, 11 \right]$ & False & UC & 11 & -                      \\
+1 & 2 & $\left[11, UC \right]$  & $\left[FB, UC \right]$ & True  & UC & 11 &  $\left[11, UC \right]$ \\
+1 & 3 & $\left[11, UC \right]$  & $\left[FB, UC \right]$ & True  & UC & 11 & $\left[11, UC \right]$ \\
+1 & 4 & $\left[11, UC \right]$  & $\left[FB, UC \right]$ & True  & UC & 11 & $\left[11, UC \right]$ \\
+1 & 1 & $\left[11, UC \right]$  & $\left[12, UC \right]$ & True  & UC & 12 &  $\left[12, UC \right]$ \\
+\hline 
+\end{tabular}
+\label{table:example-crisp-cartesian-product}
+\end{table}
+
+\subsection{\label{subsec:cartesian-product-ikc}Temporal cartesian product by ill-known constraints}
+In this subsection we are going to extend the crisp version of the temporal cartesian product to deal with ill-known time intervals.
+First of all, we need to define the ill-known counterpart functions of the previously defined \emph{intersects, first, last, OvIn}.
+
+\begin{definition}
+ \label{def:ik-intersects}
+Two ill-known time intervals intersect if one or both of the before and after relationships do not hold.
+Consider $I = \left[I_s, I_e \right]$ and $J = \left[J_s, J_e \right]$ be two ill-known time intervals with $I_s, J_s$ the starting ill-known points and $I_e, J_e$ the ending ill-known points of the intervals.
+Let the following ill-known constraints:
+\begin{align}
+ \label{eq:ik-intersects}
+C_1 &=& \left(<, I_s \right)  \\
+C_2 &=& \left(=, J_s \right) \\
+C_3 &=& \left(<, J_s \right) \\
+C_4 &=& \left(=, I_s \right) \\
+\mbox{Intersects}\left(I, J\right) &=& \left \lbrace \neg \left(C_1 \wedge C_2 \right) \vee  \neg \left( C_3 \wedge C_4 \right)  \right \rbrace \\
+\nonumber
+&=& \left \lbrace \neg \left( \left(C_1 \wedge C_2 \right) \wedge \left(C_3 \wedge C_4 \right)  \right)\right \rbrace
+\end{align}
+\end{definition}
+
+\begin{definition}
+ \label{def:ik-last}
+Last$\left(I_n, J_m \right)$. Given two ill-known points $I_n$ and $I_m$, the last operator is a combination of the following ill-known constraints:
+\begin{align}
+ \label{eq:il-last}
+C_1 &=& \left(>, I_n \right) \\
+C_2 &=& \left(=, J_m \right) \\
+C_3 &=& \left(>, J_m \right) \\
+C_4 &=& \left(=, I_n \right) \\
+\mbox{Last}\left(I_n, J_m \right) &=& \left(C_1 \wedge C_2  \right) \vee \left(C_3 \wedge C_4 \right)
+\end{align}
+\end{definition}
+
+Analogously, the first operator is defined as follows.
+\begin{definition}
+ \label{def:ik-first}
+First$\left(I_n, J_m \right)$. Given two ill-known points $I_n$ and $I_m$, this operator is a combination of the following ill-known constraints:
+\begin{align}
+ \label{eq:ik-first}
+C_1 &=& \left(<, I_n \right)\\
+C_2 &=& \left(=, J_m \right)\\
+C_3 &=& \left(<, J_m \right) \\
+C_4 &=& \left(=, I_n \right) \\
+\mbox{First}\left(I_n, J_m \right) &=& \left(C_1 \wedge C_2  \right) \vee \left(C_3 \wedge C_4 \right)
+\end{align}
+\end{definition}
+
+Now it is possible to define the function overlapping interval for the ill-known case.
+
+\begin{definition}
+ \label{def:ik-overlapping-interval}
+Given two ill-known time intervals $I$ and $J$, this function returns an overlapping interval from the original two intervals. It is necessary to define two ill-known constraints:
+
+\begin{align}
+ \label{eq:ilc-smaller-than}
+C_1 &=& \left(\leq,\mbox{First}\left(I_e, J_e \right)\right) \\
+C_2 &=& \left(=, \mbox{Last}\left( I_s, J_s \right)\right)
+\end{align}
+
+\begin{equation}
+ \mbox{OvInt}\left(I, J \right) = 
+\begin{cases}
+\left[\mbox{Last}\left( I_s, J_s \right), \mbox{First}\left(I_e, J_e \right) \right] & \mbox{ if } C_1 \wedge C_2 \neq 0.  \\
+\emptyset & \mbox{ otherwise } 
+\end{cases}
+\end{equation}
+\end{definition}
+
+Finally, the temporal cartesian product for ill-known time intervals is defined as follows.
+
+\begin{definition}
+ \label{def:temporal-cartesian-product-ik}
+The temporal cartesian product is notated $r \times^T s$ and is given by the following equation:
+\begin{align}
+ \label{eq:ik-temporal-cartesian-product}
+r \times^T s = \left \lbrace z^{\left(n+m+2\right)}  | \exists x \in r, \exists y \in s \right. \\
+\nonumber
+\mbox{intersect}\left(x[T], y[T] \right) > 0 \wedge \\
+\nonumber
+z\left[A\right] = x\left[A\right] \wedge z\left[B\right] = y \left[B \right] \wedge \\
+\nonumber
+\left.  z\left[T \right] = \mbox{OvInt}\left(x\left[T\right], y\left[T\right] \right) \wedge z\left[T\right] \neq \emptyset  \right \rbrace
+\end{align}
+\end{definition}
+
+The temporal cartesian product by ill-known constraints is illustrated with an example.
+\begin{example}
+ \label{ex:temporal-ill-known-constraint-ik}
+Consider a ill-known version of the employee's database, as illustrated by tables \ref{table:ill-known-employees} and \ref{table:ill-known-addresses}
+\end{example}
+
+\begin{table}
+\centering
+\caption{Relation for the employees with an ill-known valid-time interval.}
+\begin{tabular}{c c c c c c }
+\hline
+ID & Job & Works for & Start & Finish \\ \hline
+1 & Prof. & 4 & $\left[5, 1, 1 \right]$ & $\left[10, 1, 1 \right]$ \\
+2 & Tech. & 4 & $\left[3, 1, 1 \right]$ & $\left[7, 1, 1 \right]$ \\
+\hline 
+\end{tabular}
+\label{table:ill-known-employees}
+\end{table}
+
+\begin{table}
+\centering
+\caption{Relation for the addresses with an ill-known valid-time interval.}
+\begin{tabular}{c c c c }
+\hline
+ID & Address & Start & Finish \\ \hline
+1 & C/ Camino de Ronda & $\left[1, 1, 1 \right]$ & $\left[12, 1, 1 \right]$ \\
+2 & C/ Recogidas &  $\left[1, 1, 1 \right]$ & $\left[5, 1, 1 \right]$ \\
+\hline 
+\end{tabular}
+\label{table:ill-known-addresses}
+\end{table}
+
+\begin{table}
+\centering
+\caption{Intermediate calculations for the temporal cartesian product.}
+
+\begin{tabular}{c c c c c }
+\hline
+r.id & s.id &  intersect$\left(x\left[T\right], y\left[T\right]\right)$ &  Last$\left( I_s, J_s \right)$ & First$\left(I_e, J_e \right)$  \\ \hline
+1 & 1 & 1 & $\left[5, 1, 1 \right]$ & $\left[10, 1, 1 \right]$ \\
+1 & 2 & 1 & $\left[5, 1, 1 \right]$ & $\left[5, 1, 1 \right]$ \\
+2 & 1 & 1 & $\left[3, 1, 1 \right]$ & $\left[7, 1, 1 \right]$ \\
+2 & 2 & 1 & $\left[3, 1, 1 \right]$ & $\left[5, 1, 1 \right]$ \\
+
+\hline 
+\end{tabular}
+\label{table:example-ik-cartesian-product}
+\end{table}
 
@@
 This new relation is a set with all the possible combinations of tuples in both $R$ and $S$.
-It is usually noted as $R \Join_{r \theta s} S$, where $r,s$ are attributes from $R$ and $S$ respectively and $\theta$ an arithmetic operator.
+It is usually noted as $R \Join_{r \theta s} S$ and called (theta) join, where $r,s$ are attributes from $R$ and $S$ respectively and $\theta$ an arithmetic operator.
+The temporal join is defined on the temporal cartesian product.
 
-The following example illustrates the behaviour of the join operator in a crisp temporal database.
+\begin{definition}
+ \label{def:temporal-theta-join}
+Temporal theta join. Let $R$ and $S$ be two relations and $r, s$ be the instances of each relation respectively. Then the temporal theta join is defined as follows:
+\begin{equation}
+ \label{eq:temporal-theta-join}
+R \Join_{r \theta s}^{T} S= \sigma_{r \theta s} \left(R \times^T S \right)
+\end{equation}
 
-\begin{example}
-Consider the employees database shown in Table \ref{}.  
-\end{example}
+\end{definition}
+
+\begin{figure}
+ \includegraphics[scale=0.60]{graphs/innerjoin.eps}
+\caption{\label{fi:innerjoin} Comparison between the relational (non-temporal) inner join (left graph) and the temporal inner join (right graph).}
+\end{figure}
+
+Figure \ref{fi:innerjoin} compares the relational inner join with respect to the temporal inner join.
+The right graph illustrates the temporal case.
+It is clear that the temporal inner join is a subset or is equal to the relational inner join.
+
+\subsection{\label{subsec:Equijoin}Equijoin}
+The equijoin operator for relational databases enforces equality between specified subsets of the attributes of the relations.
+Then, the temporal equijoin operator is defined as the temporal join operator.
+
+\begin{definition}
+ \label{def:equijoin}
+Equijoin.
+Let $R$ and $S$ be two relations and $r, s$ be the instances of each relation respectively.
+$A , B$ are the sets of the attributes for the relations $R$ and $D$ respectively.
+And let $A' \subseteq A$ and $B' \subseteq B$.
+Then, the temporal equijoin is defined as follows:
+\begin{equation}
+ \label{eq:equijoin}
+R \Join_{r\left[A' \right] = s\left[B' \right]}^{T} S =  \sigma_{r\left[A' \right] = s\left[B' \right]} \left(R \times^T S \right)
+\end{equation}
+\end{definition}
+
+\subsubsection{\label{subsubsec:temporal-equijoin}Temporal Equijoin}
+The temporal equijoin is an equijoin operator in which the subset of attributes in the equijoin condition, are part of the primary key. Hence, the operator is defined as follows:
+
+\begin{definition}
+\label{def:temporal-equijoin}
+Temporal equijoin ($R$ TE-join $S$).
+Let $R, S$ be two relations and $r, s$ the instances of each relation respectively.
+The sets $A, B$ are the attributes of each relation.
+And $ID \subseteq A; ID \subseteq B$ is the primary key of both relations.
+The temporal equijoin operator is defined as follows:
+\begin{equation}
+ \label{eq:temporal-equijoin} 
+R \mbox{TE-join} S \equiv R \Join_{r\left[ID\right] = s\left[ID\right]}^{T} S
+\end{equation}
+\end{definition}
+
+\subsection{\label{subsec:natural-join}Natural Join}
+The temporal natural join is a temporal equijoin on identically named attributes.
+
+\begin{definition}
+ \label{def:temporal-natural-join}
+Temporal natural join.
+Consider the following relations $R = \left(A_1, \ldots, A_n, C_1, \ldots, C_k, T_s, T_e \right)$ and $S = \left(B_1, \ldots, B_m, C_1, \ldots, C_k, T_s, T_e \right)$.
+The temporal natural join is defined as follows:
+\begin{equation}
+ \label{eq:temporal-natural-join}
+R \Join^T S = R \Join_{r\left[C_1\right] = s\left[C_1\right] \wedge \ldots \wedge r\left[C_k\right] = s\left[C_k\right]} S
+\end{equation}
+\end{definition}
+
+\subsection{\label{subsec:outer-joins}Outer Joins}
+
+\begin{figure}
+ \includegraphics[scale=0.60]{graphs/outerjoin.eps}
+\caption{\label{fi:outerjoin} Comparison between a left outer join (left graph) and a right inner join (right graph).}
+\end{figure}
+
+%The following example illustrates the behaviour of the join operator in a crisp temporal database.
+% \begin{table}
+% \centering
+% \caption{Join Employees - Address.}
+% 
+% \begin{tabular}{c c c c c c c c c }
+% \hline
+% ID & Job & Works for & Start & Finish & Adr. & Start & Finish \\ \hline
+% 1 & Prof. & 4 & 5 & 10 & C/ Camino de ronda & - & 12 \\
+% \cellcolor[gray]{0.9}
+% 1 & \cellcolor[gray]{0.9} Prof.
+&\cellcolor[gray]{0.9} 4 &\cellcolor[gray]{0.9} 5 &\cellcolor[gray]{0.9} 10 &\cellcolor[gray]{0.9} C/ Manuel de Falla &\cellcolor[gray]{0.9} 12 &\cellcolor[gray]{0.9} - \\
+% 2 & Tech. & 4 & 3 & 7  & C/ Recogidas & - & - \\ 
+% 3 & Account. & 4  & 4 & 10 & C/ Pintor Maldonado & - & - \\
+% 4 & Adm. & - & 1 & -& C/ Mesones & - & - \\
+% 1 & Prof. & 4 & 11 & - & C/ Manuel de Falla & 12 & - \\
+% 1 & Prof. & 4 & 11 & - & C/ Camino de ronda & - & 12 \\
+% \hline 
+% \end{tabular}
+% \label{table:join-employees-address}
+% 
+% \vspace{10pt}
+% 
+% 
+% \end{table}
+% 
+% \vspace{-25pt}
+
+\newpage
 

### 2012-10-22 trunk/develop/testQuerying/src/main/java/es/jpons/persistence/query/CriteriaAllenRelations.java
querying: during, contains, equals, starts and finishes

@@ -16,9 +16,10 @@
  * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston,
  * MA 02110-1301  USA
  */
-
 package es.jpons.persistence.query;
 
+import com.sun.org.apache.bcel.internal.generic.GETSTATIC;
+import es.jpons.temporal.types.PossibilisticTP;
 import es.jpons.temporal.types.PossibilisticVTP;
 import java.sql.Types;
 import javassist.bytecode.analysis.Type;
@@ -28,353 +29,513 @@ import org.hibernate.Hibernate;
 import org.hibernate.criterion.Restrictions;
 import org.hibernate.type.LongType;
 
-
 /**
  * Class for implementing the corresponding criteria for allen relations.
- * @author Jose Enrique Pons Frías <jpons@decsai.ugr.es>
- * First version 15/10/2012
- * 
+ *
+ * @author Jose Enrique Pons Frías <jpons@decsai.ugr.es> First version
+ * 15/10/2012
+ *
  */
 public class CriteriaAllenRelations {
+
     protected static Logger log = Logger.getLogger(CriteriaAllenRelations.class);
+
     /**
-     * Returns a criteria for the before allen relation.
-     * This models the relation:
-     * record.pvp before query.pvp 
+     * Returns a criteria for the before allen relation. This models the
+     * relation: record.pvp before query.pvp
+     *
      * @param initializedCriteria an initialized instance of a criteria
      * @param pvp a possibilistic valid time period
      * @return A criteria with the before operation implemented.
      */
-    public static Criteria before(Criteria initializedCriteria, PossibilisticVTP pvp){
+    public static Criteria before(Criteria initializedCriteria, PossibilisticVTP pvp) {
         initializedCriteria = initializedCriteria.add(Restrictions.lt("pvp.endMP", pvp.getStartMP()));
-        
+
         return initializedCriteria;
     }
-    
-    
+
     public static Double computeBeforeSatisfactionDegree(PossibilisticVTP i, PossibilisticVTP j) {
         Double satisfaction = 0.0;
 
-        Double a = 0.0+ i.getEndLeft();
-        Double b = 0.0+i.getEndRight();
-        Double d = 0.0+i.getEndMP();
-        
+        Double a = 0.0 + i.getEndLeft();
+        Double b = 0.0 + i.getEndRight();
+        Double d = 0.0 + i.getEndMP();
 
[…]

### 2012-10-22 trunk/develop/testQuerying/src/main/java/es/jpons/persistence/query/TemporalQuery.java

@@ -88,8 +88,27 @@ public class TemporalQuery {
                 break;
             case meets: this.temporalCriteria = CriteriaAllenRelations.meets(this.temporalCriteria, pvp);
                 break;
-            case meet_by: 
+            case meet_by: this.temporalCriteria = CriteriaAllenRelations.meet_by(temporalCriteria, pvp);
                 break;
+            case overlaps: this.temporalCriteria = CriteriaAllenRelations.overlaps(temporalCriteria, pvp);
+                break;
+            case overlapped_by: this.temporalCriteria = CriteriaAllenRelations.overlapped_by(temporalCriteria, pvp);
+                break;
+            case during: this.temporalCriteria = CriteriaAllenRelations.during(temporalCriteria, pvp);
+                break;
+            case contains: this.temporalCriteria  = CriteriaAllenRelations.contains(temporalCriteria, pvp);
+                break;
+            case equal: this.temporalCriteria = CriteriaAllenRelations.equals(temporalCriteria, pvp);
+                break;
+            case starts: 
+            case started_by:
+                this.temporalCriteria = CriteriaAllenRelations.starts(temporalCriteria, pvp);
+                break;
+            case finishes:
+            case finished_by:
+                this.temporalCriteria = CriteriaAllenRelations.ends(temporalCriteria, pvp);
+                break;
+            
         }
     }
     
@@ -118,6 +137,16 @@ public class TemporalQuery {
                 case meets: q.temporalSatisfaction = CriteriaAllenRelations.computeMeetsSatisfactionDegree(q.vtp, pvp);
                     break;
                 case meet_by: q.temporalSatisfaction = CriteriaAllenRelations.computeMeetsSatisfactionDegree(pvp, q.vtp);
+                    break;
+                case overlaps: q.temporalSatisfaction = CriteriaAllenRelations.computeOverlapsSatisfactionDegree(q.vtp, pvp);
+                    break;
+                case overlapped_by: q.temporalSatisfaction = CriteriaAllenRelations.computeOverlapsSatisfactionDegree(pvp,q.vtp);
+                    break;
+                case during: q.temporalSatisfaction = CriteriaAllenRelations.computeDuringSatisfactionDegree(q.vtp, pvp);
+                    break;
+                case contains: q.temporalSatisfaction = CriteriaAllenRelations.computeContainsSatisfactionDegree(q.vtp, pvp);
+                    break;
+                
             }
            
             q.aggregation = weight*(q.crispSatisfaction) + (1-weight)*q.temporalSatisfaction;

### 2012-10-24 develop/testQuerying/src/main/java/es/jpons/persistence/query/CriteriaAllenRelations.java
querying: during, contains, equals, starts and finishes

@@ -56,36 +56,49 @@ public class CriteriaAllenRelations {
 
     public static Double computeBeforeSatisfactionDegree(PossibilisticVTP i, PossibilisticVTP j) {
         Double satisfaction = 0.0;
-
-        Double a = 0.0 + i.getEndLeft();
-        Double b = 0.0 + i.getEndRight();
-        Double d = 0.0 + i.getEndMP();
-
-
-        Double a1 = 0.0 + j.getStartLeft();
-        Double b1 = 0.0 + j.getStartRight();
-        Double d1 = 0.0 + j.getStartMP();
-        if (log.isDebugEnabled()) {
-            log.debug("a1: " + a1 + " b1: " + b1 + " d1: " + d1);
+        boolean inside = false;
+        
+        if(i.getEndMP() < j.getStartMP()){
+            inside = true;
         }
-        Double y1 = 0.0, y2 = 0.0, x1 = 0.0, x2 = 0.0;
-        x1 += (a * (d1 - a1) + a1 * d) / (a1 + a);
-        y1 += -(-d1 + d - a) / (a1 + a);
-
-        if (Math.abs(b - a1) > 0) {
-            x2 += (b * (a1 - d1) + a1 * d) / (b - a1);
-            y2 += (-d1 + d + b) / (b - a1);
-        }
-
-        Double max = 0.0;
-        if (y1 > 0 && y1 <= 1 && y2 >= 0 && y2 <= 1) {
-            satisfaction += (y1 >= y2) ? y1 : y2;
-        } else {
+        
+        
+//
+//        Double a = 0.0 + i.getEndLeft();
+//        Double b = 0.0 + i.getEndRight();
+//        Double d = 0.0 + i.getEndMP();
+//
+//
+//        Double a1 = 0.0 + j.getStartLeft();
+//        Double b1 = 0.0 + j.getStartRight();
+//        Double d1 = 0.0 + j.getStartMP();
+//        if (log.isDebugEnabled()) {
+//            log.debug("a1: " + a1 + " b1: " + b1 + " d1: " + d1);
+//        }
+//        Double y1 = 0.0, y2 = 0.0, x1 = 0.0, x2 = 0.0;
+//        x1 += (a * (d1 - a1) + a1 * d) / (a1 + a);
+//        y1 += -(-d1 + d - a) / (a1 + a);
+//
+//        if (Math.abs(b - a1) > 0) {
[…]

### 2012-10-24 develop/testQuerying/src/main/java/es/jpons/temporal/types/PossibilisticVTP.java

@@ -20,6 +20,7 @@ package es.jpons.temporal.types;
 
 import es.jpons.persistence.constants.OpenInterval;
 import java.io.Serializable;
+import java.rmi.server.Skeleton;
 import java.util.Date;
 import javax.persistence.Embeddable;
 import javax.persistence.Transient;
@@ -132,16 +133,28 @@ public class PossibilisticVTP implements Serializable {
         }
     }
     
-    
-    public PossibilisticVTP(Integer sday,Integer smonth,Integer syear,Integer eday,Integer emonth,Integer eyear){
+    /**
+     * Constructor
+     * @param sday Starting day of month
+     * @param smonth Starting month of year
+     * @param syear Starting year
+     * @param sLeftDays Number of days of the left starting margin.
+     * @param sRightDays Number of days of the right starting margin.
+     * @param eday Ending day of month.
+     * @param emonth Ending month of year.
+     * @param eyear Ending year.
+     * @param eLeftDays Number of days of the ending margin.
+     * @param eRightDays Number of days of the right ending margin.
+     */
+    public PossibilisticVTP(Integer sday,Integer smonth,Integer syear,Integer sLeftDays, Integer sRightDays,Integer eday,Integer emonth,Integer eyear,Integer eLeftDays, Integer eRightDays){
         DateTime dateStartingMainPoint,dateStartingLeft,dateStartingRight;
             DateTime dateEndingMainPoint,dateEndingLeft,dateEndingRight;
             
             dateStartingMainPoint = new DateTime(syear,smonth, sday, 0, 0);
             
           
-            dateStartingLeft = dateStartingMainPoint.minusDays(3);
-            dateStartingRight = dateStartingMainPoint.plusDays(3);
+            dateStartingLeft = dateStartingMainPoint.minusDays(sLeftDays);
+            dateStartingRight = dateStartingMainPoint.plusDays(sRightDays);
             
             Duration d = new Duration(dateStartingLeft, dateStartingMainPoint);
             Duration d1 = new Duration(dateStartingMainPoint, dateStartingRight);
@@ -150,16 +163,17 @@ public class PossibilisticVTP implements Serializable {
             
             
             dateEndingMainPoint = new DateTime(eyear,emonth, eday, 0, 0);
-            dateEndingLeft = dateEndingMainPoint.minusDays(3);
-            dateEndingRight = dateEndingMainPoint.plusDays(3);
[…]

### 2012-10-24 develop/testQuerying/src/test/java/es/jpons/persistence/query/CriteriaAllenRelationsTest.java

@@ -1,0 +1,586 @@
+/*
+ * To change this template, choose Tools | Templates
+ * and open the template in the editor.
+ */
+package es.jpons.persistence.query;
+
+import es.jpons.temporal.types.PossibilisticVTP;
+import java.util.Date;
+import java.util.GregorianCalendar;
+import junit.framework.TestCase;
+import org.hibernate.Criteria;
+
+/**
+ * Tests for Criteria Allen Relations class
+ * @author Jose Enrique Pons Frías <jpons@decsai.ugr.es>
+ */
+public class CriteriaAllenRelationsTest extends TestCase {
+    
+    public CriteriaAllenRelationsTest(String testName) {
+        super(testName);
+    }
+    
+    @Override
+    protected void setUp() throws Exception {
+        super.setUp();
+    }
+    
+    @Override
+    protected void tearDown() throws Exception {
+        super.tearDown();
+    }
+
+    /**
+     * Test of before method, of class CriteriaAllenRelations.
+     */
+    public void testBefore() {
+        System.out.println("before");
+        Criteria initializedCriteria = null;
+        PossibilisticVTP pvp = null;
+        Criteria expResult = null;
+        Criteria result = CriteriaAllenRelations.before(initializedCriteria, pvp);
+        assertEquals(expResult, result);
+        // TODO review the generated test code and remove the default call to fail.
+        fail("The test case is a prototype.");
+    }
+
+    /**
+     * Test of computeBeforeSatisfactionDegree method, of class CriteriaAllenRelations.
+     */
+    public void testComputeBeforeSatisfactionDegree() {
+        System.out.println("computeBeforeSatisfactionDegree");
+        
+        System.out.println("Computing degree in [0,1]");
+        PossibilisticVTP i = new PossibilisticVTP(10, 1,2012, 1, 1, 6, 1, 2012, 2,2);
+        PossibilisticVTP j = new PossibilisticVTP(6,1,2012,1,1,9,1,2012,1,1);
+        
+        Double expResult = 2.0/3.0;
+        Double result = CriteriaAllenRelations.computeBeforeSatisfactionDegree(i, j);
+        System.out.println("result: "+result);
+        assertEquals(expResult, result);
+        
+        System.out.println("Computing degree = 1");
+        
+        i = new PossibilisticVTP(10, 1,2012, 1, 1, 16, 1, 2012, 2,2);
[…]

### 2013-02-05 joins/crisp-algebra.tex
second draft

\subsection{\label{sec:selection}Selection}
The selection operator is very useful when the user wants to obtain a subset of tuples that fulfill a certain constraint. More formally,

\begin{definition}
\label{def:crisp-selection}
 Consider the elements in definition \ref{def:valid-time-relation}.
The selection operator $\sigma$ obtains a subset of tuples that fulfill a set of constraints $P$ from an instance $r$ of the relation $R$. The set of constraints is usually a boolean combination of atomic constraints. The selection operator is noted as follows:
\end{definition}

\begin{equation}
 \label{eq:selection}
\sigma_{P} \left( r \right)
\end{equation}

Where $r \in R$ is the relation, and $P$ is the selection formula. The selection formula is a set of two elements:
% \begin{itemize}
% \item relation.attribute boolean relation relation.attribute
% \item relation.temporal-attribute allen-relation relation.temporal-attribute 
% \end{itemize}
\begin{equation}
 \label{eq:selection_formula}
P = \left \lbrace Q, Q^{T}\right \rbrace
\end{equation}

Where $Q$ is the set of non-temporal constraints and $Q^{T}$ the set of temporal constraints.
The component $Q$ is a set composed by atomic constraints:

\begin{equation}
 \label{eq:non-temporal-constraints}
Q = \left \lbrace q_{\mbox{a}_1}  \theta \mbox{ val}_1, \ldots, q_{\mbox{a}_n}  \theta \mbox{ val}_n \right \rbrace
\end{equation}

Where:
\begin{itemize}
 \item $q_{a \in A }$ is an atomic constraint. The constraint refers to an attribute $a$ that belongs to the set of attributes $A \subseteq A_1 \ldots A_n $ of the relation $R$.
\item $\theta$ is a relational operator; usually one of $\left \lbrace =, \neq, <, \leq, >, \geq \right \rbrace$.
\item \emph{val}$_1 \ldots$\emph{val}$_n$ are values in the domain of the queried attribute. 
 \end{itemize}

The temporal constraints, $Q^t$ are provided in a similar way. The main difference is that, instead of comparisons like $\left \lbrace =, \neq, <, \leq, >, \geq \right \rbrace$, we use the Allen's relations~\cite{Allen1983}. Hence, the representation of the temporal constraints is expressed as follows:

\begin{equation}
\label{eq:temporal-constratins}
Q^{t} = \left \lbrace q_{v_1}  \AR i_1, \ldots, q_{v_n}  \AR i_n \right \rbrace
\end{equation}
Where
\begin{itemize}
\item $q_{v \in I}$ is an attribute representing a time interval. ($q_v =  \left[s_v, e_v \right]$).
\item $\AR$ is one of the thirteen Allen's relations (equal, before, overlaps, starts, finishes, meets, during) and their respective reverses (See Figure \ref{fig:allen} ).
\item $i_{x \in \left \lbrace 1 \ldots n\right \rbrace}$ is a crisp time interval with starting and ending points ($i_x = \left[s_x, e_x \right]$).

\end{itemize}

\subsubsection{\label{crisp-query-eval}Query Evaluation}
In a crisp relational database, the modelling of the query satisfaction is a boolean result. The evaluation of the query requirements results in one of two possibilities; accept the record if it satisfies all the constraints or reject the record otherwise. The evaluation of the selection formula $P$ given in equation \eqref{eq:selection_formula} is handled as follows. For each record $r$ in the database, two things happen independently:

\begin{itemize}
 \item The crisp constraints expressed in $Q$ are evaluated and aggregated. The result of this is a boolean value. We will note as $e_Q(r)$ value for the evaluation of the constraints in  $Q$ for the record $r$.
\item The temporal constraints expressed in $Q$ are evaluated and aggregated. The result of this is again a boolean value. We will note as $e_Q^{T}(r)$ the value for the evaluation of the constraints in $Q^T$ for the record $r$.
\end{itemize}

The results from $e_Q(r)$ and $e_Q^{T}(r)$ are aggregated using a boolean combination. For valid-time intervals, the preferred combination is the \emph{'and'} ($\wedge$) operator . The function $e_{\mbox{final}}\left(r\right)$ is given by:

\begin{equation}
 \label{eq:e-final-crisp}
e_{\mbox{final}}\left(r\right) = e_Q(r) \wedge e_Q^{T}(r)
\end{equation}

% Quizás mejor dejar la agregación para el caso difuso
% In order to present the results to the user, it is necessary to provide methods for both aggregation and ranking. This is explained in the following subsection.
% 
% 
% \subsubsection{\label{crisp-ranking}Aggregation and Ranking}
% In the crisp case we need to provide a method to combine the output from $Q$ and $Q^T$. That method selects or rejects a tuple

\begin{example}
Consider an employee's database. The data are stored in two tables; Table \ref{table:employees} ($r \in R$) contains temporal data about the employees and Table \ref{table:address} ($s \in S$) contains temporal  data about the addresses for each employee. Consider now that the user wants to obtain all the employees who are a professor and that worked in the time interval $\left[7, 10 \right]$.
%\vspace{-10pt} \ref{table:employees} \ref{table:address}

\begin{table}
\centering
\caption{Employee's table. Instance $r$ of relation $R$.}
\vspace{2mm}

\begin{tabular}{c c c c c c }
\hline
ID & Job & Works for & Start & Finish \\ \hline
1 & Professor & 4 & 5 & 10 \\
2 & Technician & 4 & 3 & 7 \\
3 & Accountant & 4  & 4 & 10 \\
4 & Administrator & - & 1 & UC \\
1 & Professor & 4 & 11 & UC \\
\hline 
\end{tabular}
\label{table:employees}

%\vspace{10pt}

\end{table}

%\vspace{-25pt}

%\vspace{-10pt}

\begin{table}
\centering
r\caption{Address table. Instance $s$ of relation $S$.}
\vspace{2mm}
\begin{tabular}{c c c c }
\hline
ID & Adr. & Start & Finish \\ \hline
1 & C/ Camino de ronda & FB & 12 \\
2 & C/ Recogidas & FB & UC \\ 
3 & C/ Pintor Maldonado & FB & UC \\
4 & C/ Mesones & FB & UC \\
1 & C/ Manuel de Falla & 12 & UC \\
\hline 
\end{tabular}
\label{table:address}

%\vspace{10pt}

\end{table}

%\vspace{-25pt}

The selection formula is the following:

\begin{equation}
 \label{eq:selection-example}
\sigma_{r.Job = 'Professor', r.\left[S, E \right] \mbox{ Contains } \left[7, 10\right]} \left(r \right)
\end{equation}
The values for $e_Q(r)$ and $e_Q^T(r)$ for each record, as well as the final aggregation (see Equation \eqref{eq:e-final-crisp}) are illustrated in Table \ref{table:crisp-intermediate-calculations-employees}.
The resultset for the selection is shown in Table \ref{table:example-selection}.
\end{example}

\begin{table}
\centering
\caption[Intermediate calculations.]{Intermediate calculations for selection formula given in equation \eqref{eq:selection-example}}
\vspace{2mm}

\begin{tabular}{c c c c c c c c }
\hline
ID & Job & Works for & Start & Finish & $e_Q(r)$& $e_Q(r)$ & $e_{final}(r)$\\ \hline
1 & Professor & 4 & 5 & 10 & True & True & True\\
2 & Technician & 4 & 3 & 7 & False & False & False\\
3 & Accountant & 4  & 4 & 10 & False & True & False\\
4 & Administrator & - & 1 & UC & False & True & False\\
1 & Professor & 4 & 11 & UC & True & False & False\\
\hline 
\end{tabular}
\label{table:crisp-intermediate-calculations-employees}

%\vspace{10pt}

\end{table}

%\vspace{-10pt}

\begin{table}
\centering
\caption{Resultset table for the selection in equation \eqref{eq:selection-example}}
\vspace{2mm}
\begin{tabular}{c c c c c c }
\hline
ID & Job & Works for & Start & Finish \\ \hline
1 & Prof. & 4 & 5 & 10 \\
\hline 
\end{tabular}
\label{table:example-selection}

%\vspace{10pt}

\end{table}

%\todo[caption={crisp cartesian product}]{Maybe the cartesian product should be inside the selection??}
\subsection{\label{sec:cartesian-product}Cartesian Product}
The relational cartesian product is defined as all the possible combinations between tuples of  $r, s$ from two given relations $R$ and $S$ respectively. It is usually noted as $r \times s$.
The temporal cartesian product operator is defined as the relational cartesian product with a predicate on the temporal valid-time intervals \cite{DengfengGao2002}. The temporal cartesian product is defined then as $r \times^T s$.

We will re-define the predicate given in \cite{DengfengGao2002} in terms of the Allen's relations. To do this, two auxiliary functions should be defined.

%\todo[caption={Crisp temporal operators}]{Instead of define inline the crisp operators for intervals, Could it be more interesting to have an appendix with the implementation?}

\begin{definition}
 \label{def:crisp-intersect}
Intersect$\left(i_1, i_2 \right)$. Given two intervals $i_1$ and $i_2$, the Intersect function returns a boolean value showing whether the two intervals intersect or not. In terms of the Allen's relations, two intervals $i_1$ and $i_2$ intersect if between them exist any of the Allen's relations except Before or After. In other words, only if the Allen's relations Before or After do not hold for the two intervals, they will intersect.

\begin{equation}
 \label{eq:crisp-intersect}
\mbox{Intersect}\left( i_1, i_2 \right) = \left \lbrace \neg \left(e_1 < s_2 \right) \vee \neg \left(e_2 < s_1 \right) \right \rbrace
\end{equation}

\end{definition}

\begin{example}
Consider the following intervals $i_1 = \left[5, 10 \right]$ and  $i_2 = \left[0, 12 \right]$. The intersect operation is calculated as follows:
\begin{align}
\mbox{Intersect}\left( i_1, i_2 \right) &=& \left \lbrace \neg \left(10 < 0 \right)  \vee \neg \left(12 < 5 \right) \right \rbrace \\
\nonumber
&=& \mbox{True} \vee \mbox{True} \\
\nonumber &=& \mbox{True}
\end{align}

\end{example}

The other auxiliary function that allows to define the predicate for the cartesian product, needs some previous definitions as well.

\begin{definition}
 \label{def:crisp-last}
Last$\left(i_p, i_q \right)$. Given two time points $i_p$ and $i_q$, this function obtains the greatest time point.
\begin{equation}
 \label{eq:last-crisp}
\mbox{Last}\left(i_p, i_q \right) = 
\begin{cases}
i_p & \mbox{ if } i_p \mbox{ after } i_q \\
i_q & \mbox{ otherwise } 
\end{cases}
\end{equation}
Note that because both $i_p, i_q$ are time points, the after relation is implemented as $i_p > i_q$.
\end{definition}

\begin{definition}
 \label{def:crisp-first}
First$\left(i_p, i_q \right)$. Given two time points $i_p$ and $i_q$, this function obtains the least time point.
\begin{equation}
 \label{eq:first-crisp}
\mbox{First}\left(i_p, i_q \right) =
\begin{cases}
i_p & \mbox{ if } i_p \mbox{ before } i_q \\
i_q & \mbox{ otherwise } 
\end{cases}
\end{equation}
Note that because both $i_p, i_q$ are time points, the before relation is implemented as  $i_p < i_q$.
\end{definition}

\begin{definition}
 \label{def:crisp-overlapping-interval}
Overlapping-interval$\left(i_1, i_2 \right)$. Given two intervals $i_1 = \left[s_1, e_1 \right]$ and $i_2 = \left[s_2, e_2\right]$, this function returns an overlapping interval from the original two intervals.
\begin{equation}
 \label{eq:crisp-overlapping-interval}
\mbox{OvInt}\left(i_1, i_2 \right) = 
\begin{cases}
\left[\mbox{Last}\left( s_1, s_2 \right), \mbox{First}\left(e_1, e_2 \right) \right] & \mbox{ if } \mbox{Last}\left( s_1, s_2 \right) \leq \mbox{First}\left(e_1, e_2 \right) \\
\emptyset & \mbox{ otherwise } 
\end{cases}
\end{equation}
\end{definition}

Now it is possible to define the temporal cartesian product.

\begin{definition}
 \label{def:crisp-temporal-cartesian-product}
Temporal Cartesian Product. Consider the elements in definition \ref{def:valid-time-relation}.
The temporal cartesian product of two temporal relations $r \in R = \left(A_1, \ldots A_n, S, E \right)$ and $s \in S = \left(B_1 \ldots, B_m, S, E \right)$ is noted as $r \times^T s$ and is defined by the following formula:

\begin{align}
 \label{eq:crisp-temporal-cartesian-product}
r \times^T s =  \lbrace z^{\left(n+m+2 \right)} | \exists x \in r, \exists y \in s \\
\nonumber
\mbox{intersect}\left(x[S,E], y[S,E] \right) \wedge \\
\nonumber
z\left[A\right] = x\left[A\right] \wedge z\left[B\right] = y \left[B \right] \wedge \\
\nonumber
 z\left[S, E \right] = \mbox{OvInt}\left(x\left[S, E\right], y\left[S, E\right] \right) \wedge z\left[S, E\right] \neq \emptyset  \rbrace
\end{align}
Where $A$ and $B$ are a shorthand for $\left \lbrace A_1, \ldots A_n\right \rbrace$ and $\left \lbrace B_1 \ldots, B_m\right \rbrace$ respectively.
\end{definition}

The following example illustrates the temporal cartesian product.
\begin{example}
 \label{ex:crisp-temporal-cartesian-product}
Consider the employee's database mentioned before (Tables \ref{table:employees} and \ref{table:address}). Table \ref{table:example-crisp-cartesian-product} illustrates an intermediate step to compute the temporal cartesian  product.

\end{example}

\begin{table}
\begin{center}
\caption{Intermediate calculations for the temporal cartesian product.}
\vspace{2mm}
\begin{tabular}{c c c c c c c c}
\hline
r.id & s.id & $x\left[S, E\right]$ & $y\left[S, E\right]$ & $^1$ & $^2$ & $^3$  &$z\left[S, E \right]$  \\ \hline
1 & 1 & $\left[5, 10 \right]$  & $\left[FB, 11 \right]$  & True  & 11 &  5 & $\left[5, 11 \right]$ \\
1 & 2 & $\left[5, 10 \right]$  & $\left[FB, UC \right]$  & True  & UC &  5 & $\left[5, UC \right]$ \\
1 & 3 & $\left[5, 10 \right]$  & $\left[FB, UC \right]$  & True  & UC &  5 & $\left[5, UC \right]$ \\
1 & 4 & $\left[5, 10 \right]$  & $\left[FB, UC \right]$  & True  & UC &  5 & $\left[5, UC \right]$ \\
1 & 1 & $\left[5, 10 \right]$  & $\left[12, UC \right]$  & False & UC & 12 & -                     \\
2 & 1 & $\left[3, 7 \right]$   & $\left[FB, 11 \right]$  & True  & 11 &  3 & $\left[3, 11 \right]$ \\
2 & 2 & $\left[3, 7 \right]$   & $\left[FB, UC \right]$  & True  & UC &  3 & $\left[3, UC \right]$ \\
2 & 3 & $\left[3, 7 \right]$   & $\left[FB, UC \right]$  & True  & UC &  3 & $\left[3, UC \right]$ \\
2 & 4 & $\left[3, 7 \right]$   & $\left[FB, UC \right]$  & True  & UC &  3 & $\left[3, UC \right]$ \\
2 & 1 & $\left[3, 7 \right]$   & $\left[12, UC \right]$  & False & UC & 12 &  -                    \\
3 & 1 & $\left[4, 10 \right]$  & $\left[FB, 11 \right]$  & True  & 11 &  4 & $\left[4, 11 \right]$ \\
3 & 2 & $\left[4, 10 \right]$  & $\left[FB, UC \right]$  & True  & UC &  4 & $\left[4, 11 \right]$ \\
3 & 3 &  $\left[4, 10 \right]$ & $\left[FB, UC \right]$  & True  & UC &  4 & $\left[4, 11 \right]$ \\
3 & 4 & $\left[4, 10 \right]$  & $\left[FB, UC \right]$  & True  & UC &  4 & $\left[4, 11 \right]$ \\
3 & 1 & $\left[4, 10 \right]$  & $\left[12, UC \right]$  & False & 12 & 12 & -                     \\
4 & 1 & $\left[1, UC \right]$   & $\left[FB, 11 \right]$ & True  & UC &  1 & $\left[1, UC \right]$ \\
4 & 2 & $\left[1, UC \right]$   & $\left[FB, UC \right]$ & True  & UC &  1 & $\left[1, UC \right]$ \\
4 & 3 & $\left[1, UC \right]$   & $\left[FB, UC \right]$ & True  & UC &  1 & $\left[1, UC \right]$ \\
4 & 4 & $\left[1, UC \right]$   & $\left[FB, UC \right]$ & True  & UC &  1 & $\left[1, UC \right]$ \\
4 & 1 & $\left[1, UC \right]$   & $\left[12, UC \right]$ & True  & UC & 12 & $\left[12, UC \right]$ \\
1 & 1 & $\left[11, UC \right]$  & $\left[FB, 11 \right]$ & False & UC & 11 & -                      \\
1 & 2 & $\left[11, UC \right]$  & $\left[FB, UC \right]$ & True  & UC & 11 &  $\left[11, UC \right]$ \\
1 & 3 & $\left[11, UC \right]$  & $\left[FB, UC \right]$ & True  & UC & 11 & $\left[11, UC \right]$ \\
1 & 4 & $\left[11, UC \right]$  & $\left[FB, UC \right]$ & True  & UC & 11 & $\left[11, UC \right]$ \\
1 & 1 & $\left[11, UC \right]$  & $\left[12, UC \right]$ & True  & UC & 12 &  $\left[12, UC \right]$ \\
\hline 
\end{tabular}
\label{table:example-crisp-cartesian-product}
\end{center}
$^1$, intersect$\left(x\left[S, E\right], y\left[S, E\right]\right)$\\
$^2$, First$\left(x[E], y[E] \right)$ \\
$^3$, Last$\left(x[S], y[S] \right)$
\end{table}

\subsubsection{\label{sec:join}Join}
The join operator $\Join$ builds a new relation from two given relations, namely $r \in R$ and $s \in S$. This new relation is a set with all the possible combinations of tuples in both $r$ and $s$ that fulfill a predicate. It is usually noted as $r \Join_{a \theta b} s$ and called (theta) join, where $a,b$ are attributes from $r$ and $s$ respectively and $\theta$ a relational operator. The temporal join definition is based on the temporal cartesian product.

\begin{definition}
 \label{def:temporal-theta-join}
Temporal theta join. Let $r$ and $s$ be two  instances of relations $R, S$ respectively. Then the temporal theta join is defined as follows:
\begin{equation}
 \label{eq:temporal-theta-join}
r \Join_{a \theta b}^{T} s= \sigma_{a \theta b} \left(r \times^T s \right)
\end{equation}

\end{definition}

% _{\begin{figure}
%  \includegraphics[scale=0.60]{graphs/innerjoin.eps}
% \caption[Comparison inner join vs temporal inner join.]{\label{fi:innerjoin} Comparison between the relational (non-temporal) inner join (left graph) and the temporal inner join (right graph).}
% \end{figure}
% 
% 
% 
% Figure \ref{fi:innerjoin} compares the relational inner join with respect to the temporal inner join. The right graph illustrates the temporal case. It is clear that the temporal inner join is a subset or is equal to the relational inner join.}

% \subsubsection{\label{subsec:Equijoin}Equijoin}
% The equijoin operator for relational databases enforces equality between specified subsets of the attributes of the relations. Then, the temporal equijoin operator is defined as the temporal join operator.
% 
% \begin{definition}
%  \label{def:equijoin}
% Equijoin. Let $R$ and $S$ be two relations and $r, s$ be the instances of each relation respectively. $A , B$ are the sets of the attributes for the relations $R$ and $S$ respectively. And let $A' \subseteq A$ and $B' \subseteq B$. Then, the temporal equijoin is defined as follows:
% \begin{equation}
%  \label{eq:equijoin}
% r \Join_{r\left[A' \right] = s\left[B' \right]}^{T} s =  \sigma_{r\left[A' \right] = s\left[B' \right]} \left(r \times^T s \right)
% \end{equation}
% \end{definition}
% 
% 
% \subsubsection{\label{subsubsec:temporal-equijoin}Temporal Equijoin}
% The temporal equijoin is an equijoin operator in which the subset of attributes in the equijoin condition, are part of the primary key. Hence, the operator is defined as follows:
% 
% \begin{definition}
% \label{def:temporal-equijoin}
% Temporal equijoin ($r \mbox{ TE-join } s$). Let $R, S$ be two relations and $r, s$ the instances of each relation respectively. Consider that $PK$ is the primary key of both relations. The temporal equijoin operator is defined as follows:
% \begin{equation}
%  \label{eq:temporal-equijoin} 
% r \mbox{ TE-join } s \equiv r \Join_{r\left[PK\right] = s\left[PK\right]}^{T} s
% \end{equation}
% \end{definition}
% 
% \subsubsection{\label{subsec:natural-join}Natural Join}
% The temporal natural join is a temporal equijoin on identically named attributes.
% 
% \begin{definition}
%  \label{def:temporal-natural-join}
% Temporal natural join. Consider the following relations $R = \left(A_1, \ldots, A_n, C_1, \ldots, C_k, S, E \right)$ and $S = \left(B_1, \ldots, B_m, C_1, \ldots, C_k, S, E \right)$. Let $r,s$ be instances of $R, S$ respectively. The temporal natural join is defined as follows:
% \begin{equation}
%  \label{eq:temporal-natural-join}
% r \Join^T s = r \Join_{r\left[C_1\right] = s\left[C_1\right] \wedge \ldots \wedge r\left[C_k\right] = s\left[C_k\right]} s
% \end{equation}
% \end{definition}

% \todo[caption={outer-joins}]{Still have to define the outer joins implementation...}
% \subsubsection{\label{subsec:outer-joins}Outer Joins}
% 
% 
% \begin{figure}
%  \includegraphics[scale=0.60]{graphs/outerjoin.eps}
% \caption{\label{fi:outerjoin} Comparison between a left outer join (left graph) and a right inner join (right graph).}
% \end{figure}

\subsection{Projection}
\subsection{Cartesian Product}
\subsection{Union}
\subsection{Difference}
\subsection{Intersection}
\subsection{Join}
\subsection{Division}

### 2013-02-05 joins/fuzzy-select.tex
second draft

\subsection{\label{subsection:fuzzy-selection}Selection}
The selection operator is defined analogously as the crisp selection operator given in Definition \ref{def:crisp-selection} and the formal specification given by equation \eqref{eq:selection}. 
%In order to distinguish the possibilistic selection from the crisp selection, the possibilistic operators will be written with the superscript FT.

\begin{definition}
\label{def:poss-selection}
 Consider the elements in definition \ref{def:fuzzy-temporal-relation}.
The selection operator $\tilde\sigma^{T}$ returns a subset of tuples that fulfill a set of constraints $P$ from an instance $r$ of the relation $R$. The set of constraints is usually a boolean combination of atomic constraints. The selection operator is noted as follows:
\end{definition}

\begin{equation}
 \label{eq:poss-selection}
\tilde\sigma^{\mbox{T}}_{P} \left( r \right)
\end{equation}

Where $r \in R$ is the relation, and $P$ is the selection formula.

In the possibilistic model, the main modification is that the selection predicate $P$ is not a boolean function. The selection predicate $P$ returns a satisfaction degree (in the unit interval). 
The selection formula has the same appearance than the crisp selection formula in equation \eqref{eq:selection_formula}:
\begin{equation}
 \label{eq:fuzzy-selection_formula}
P = \left \lbrace Q, Q^{T}\right \rbrace
\end{equation}

In this case, in order to evaluate both, where  $Q$ and $Q^{T}$ could be fuzzy or crisp predicates, it is necessary to define the evaluation functions for both $Q$ and $Q^{T}$. We will illustrate the possibilistic evaluation of $Q$. The possibilistic evaluation of $Q^T$ is analogous. First of all, consider a boolean function $\bool:\Boolean^{n}  \rightarrow \Boolean$. The evaluation function for the constraints in $Q$  is given by:

\begin{equation}
 \label{eq:evaluation-function}
\lambda : Q \rightarrow \Boolean : \lambda (Q) \rightarrow \bool \left(q_{a_1}  \theta \mbox{ val }, \ldots, q_{a_n}  \theta \mbox{ val } \right)
\end{equation}

Now, the uncertainty about the evaluation of $Q$ and a boolean function $\bool:\Boolean^{n}  \rightarrow \Boolean$ is given by:

\begin{equation}
 \label{eq:evaluation-lambda-function}
\pi_{\lambda \left(Q \right)} = \tilde{\bool} \left(\pi_{q_{a_1}  \theta \mbox{ val }}, \ldots, \pi_{q_{a_n}  \theta \mbox{ val }} \right)
\end{equation}

% The temporal constraints, $Q^t$ are provided in a similar way. The main difference is that, instead of comparations like $\left \lbrace =, \neq, <, \leq, >, \geq \right \rbrace$, we use the allen's relations. Hence, the representation of the temporal constraints is expressed as follows:
% 
% \begin{equation}
% \label{eq:temporal-constratins}
% Q^{t} = \left \lbrace q_{v_1}  ar \mbox{ val }, \ldots, q_{v_n}  ar \mbox{ val } \right \rbrace
% \end{equation}
% Here, $ar$ is one of the thirdteen Allen's relations (equal, before, overlaps, starts, finishes, meets, during) and their respective reverses.

\subsubsection{Query Evaluation}
In fuzzy querying of regular (relational) databases, the modelling of query satisfaction is a matter of degree. Usually, the evaluation of the query requirements for a record results in a satisfaction degree $s$, where $s$ lies in $\left[0,1\right]$, where 0 denotes total dissatisfaction and 1 denotes complete satisfaction. In crisp querying, the evaluation of query requirements for a record results in the accepting or rejecting of the record as a part of the result set. This can be modelled using satisfaction degrees, by assigning rejection a degree of $0$ and acceptance a degree of $1$ and not using any other value in $\left[0,1\right]$.

The evaluation of the predicate $P = \left( Q, Q^T \right)$, is now handled as follows. For each record $r$ in the database, with the valid-time notion of $r$ being specified by a PVP $J$, two events happen independently:

\begin{itemize}
\item
The preferences expressed in $Q$ are evaluated, resulting in a satisfaction degree denoted here $e_{Q}(r)$. The presented model accepts any sound way of calculating this evaluation, as long as $e_{Q}(r) \in \left[0,1\right]$. 
\item
Depending on the Allen Relation selected ($\AR$), a specific set of ill-known constraints is considered. The possibility and necessity that $r$ fulfills all these constraints are calculated using formulas based on equations \eqref{ill-known-pos} respectively \eqref{ill-known-nec} and aggregated using the $\min$ operator. 
\end{itemize}

\subsubsection{Aggregation and Ranking}
In order to present the results to the user, a crude ranking method is used: for every record $r$, the sum of $\Pos_{Q^{time}}(r)$ and $\Nec_{Q^{time}}(r)$ gives an evaluation score $e'_{Q^{time}}(r)$ in interval $\left[0,2\right]$. Because necessity cannot exceed $0$ unless possibility is $1$, this gives a natural ranking score. Some authors ~\cite{Bosc2010a} mentioned before that the possibility and necessity measures result in a total order in the set of events.  This $e'_{Q^{time}}(r)$ is then rescaled to the unit interval, resulting in $e_{Q^{time}}(r)$. The final ranking $e_{final}(r)$ is now given by a convex combination:

\begin{equation}
\label{eq:convex-comb}
e_{final}(r)\ =\ \omega*e_{Q}(r)\ +\ (1-\omega)*e_{Q^{time}}(r), \omega \in \left[0, 1 \right]
\end{equation}

The use of this convex combination allows a record to make up for a low score for the temporal constraint by a good score for the non-temporal constraint (or vice versa). Changing $\omega$ also allows granting the temporal constraint more weight with respect to the non-temporal constraint (or vice versa).

\begin{example} 
Consider the example relation $c \in C$ given in Table \ref{tb:car-models} describing car models, containing general attributes (model name, manufacturer, car segment) and one PVP  describing the approximate period of time during which the car model was sold. In this example, the value for $D$ is stored in \emph{yyyy} format and $a$ and $b$ are represented by an integer. The ID field identifies a car model while the field Instance ID (IID) identifies the instance for a car model, thus a car model in a certain state.
\end{example}

\begin{table}[h]
\centering
\caption{Example database, instance $c \in C$}
\vspace{2mm}
\begin{tabular}{c c c c c c c}
\hline
ID & IID & Segment & Manufacturer & Name & Start & End  \\ [0.5ex]
\hline
001 & 1 & B & Peugeot & 205 & [1985,2,3] & [1997,2,1] \\
002 & 1 & C & Peugeot & 305 & [1977,2,2] & [1989,2,3] \\
003 & 1 & B & Citroen & C2 & [2001,1,1] & [2005,1,1] \\
001 & 2 & B & Peugeot & 206 & [2000,1,2] & [2011,2,1] \\
001 & 3 & B & Peugeot & 207 & [2006,1,1] & [2011,1,1]\\
\hline
\end{tabular}
%\vspace{10pt}
\label{tb:car-models}
%\vspace{-30pt}
\end{table}

Consider the following query:
\begin{center}
\emph{The user wants to obtain a list of models from segment B, sold by manufacturer Peugeot before the Citroen C2.}\\
\end{center}

Using the introduced notations in \eqref{eq:poss-selection},\eqref{eq:fuzzy-selection_formula}, the query is translated to: 
\begin{equation}
\tilde \sigma^{\mbox{T}}_{\left \lbrace Q, Q^T\right \rbrace} \left( c \right)\
\end{equation}

The query constraints are the following:

\begin{align}
Q & = \left(c.Segment =  B\right) \wedge \left(c.Manufacturer = Peugeot\right)\\
Q^{T} & = c.\left[S, E \right] \mbox{Before} \left[2001,1,1\right], \left[2005,1 ,1 \right]
\end{align}

\begin{table}[ht]
\caption{Result table and ranking}
\centering
\begin{tabular}{c c c c c c c}
\hline
ID & IID &  $\Pos_{Q^{T}}$ & $\Nec_{Q^{T}}$ & $e_{Q^{T}} (rescaled)$ & $Q$ & $e_{final}$ ($\omega=0.5$) \\ [0.5ex]
\hline
001 & 1 & 1 &  1 & 1 & 1 & 1 \\
002 & 1 & 1 & 1 & 1 & 0.5 & 0.75 \\
003 & 1 & 1 & 0.5 & 0.75 &0 & 0.375\\
001 & 2 & 1 & 0 & 0.5 &1 & 0.75 \\
001 & 3 & 0 & 0 & 0 &1 & 0.5\\
\hline
\end{tabular}
\label{tb:results}
\end{table}

Table \ref{tb:results} shows a natural and gradual ranking for the results. The last record, (ID 001, IID 3) shows also that with $\omega$ = 0.5 both temporal and regular criteria have the same importance.

\subsection{\label{subsection:fuzzy-cartesian-product}Cartesian Product}

In this subsection we are going to extend the crisp version of the temporal cartesian product to deal with ill-known time intervals. First of all, we need to define the ill-known counterpart functions of the previously defined \emph{intersects, first, last, OvIn}.

\begin{definition}
 \label{def:ik-intersects}
Two ill-known time intervals intersect if one or both of the before and after relationships do not hold. Consider $I = \left[I_s, I_e \right]$ and $J = \left[J_s, J_e \right]$ be two ill-known time intervals with $I_s, J_s$ the starting ill-known points and $I_e, J_e$ the ending ill-known points of the intervals. Let the following ill-known constraints:
\begin{align}
 \label{eq:ik-intersects}
C_1 &=& \left(<, I_s \right)  \\
C_2 &=& \left(=, J_s \right) \\
C_3 &=& \left(<, J_s \right) \\
C_4 &=& \left(=, I_s \right) \\
\mbox{Intersects}\left(I, J\right) &=& \left \lbrace \neg \left(C_1 \wedge C_2 \right) \vee  \neg \left( C_3 \wedge C_4 \right)  \right \rbrace \\
\nonumber
&=& \left \lbrace \neg \left( \left(C_1 \wedge C_2 \right) \wedge \left(C_3 \wedge C_4 \right)  \right)\right \rbrace
\end{align}
\end{definition}

\begin{definition}
 \label{def:ik-last}
Last$\left(I_n, J_m \right)$. Given two ill-known points $I_n$ and $I_m$, the last operator is a combination of the following ill-known constraints:
\begin{align}
 \label{eq:il-last}
C_1 &=& \left(>, I_n \right) \\
C_2 &=& \left(=, J_m \right) \\
C_3 &=& \left(>, J_m \right) \\
C_4 &=& \left(=, I_n \right) \\
\mbox{Last}\left(I_n, J_m \right) &=& \left(C_1 \wedge C_2  \right) \vee \left(C_3 \wedge C_4 \right)
\end{align}
\end{definition}

Analogously, the first operator is defined as follows.
\begin{definition}
 \label{def:ik-first}
First$\left(I_n, J_m \right)$. Given two ill-known points $I_n$ and $I_m$, this operator is a combination of the following ill-known constraints:
\begin{align}
 \label{eq:ik-first}
C_1 &=& \left(<, I_n \right)\\
C_2 &=& \left(=, J_m \right)\\
C_3 &=& \left(<, J_m \right) \\
C_4 &=& \left(=, I_n \right) \\
\mbox{First}\left(I_n, J_m \right) &=& \left(C_1 \wedge C_2  \right) \vee \left(C_3 \wedge C_4 \right)
\end{align}
\end{definition}

Now it is possible to define the function overlapping interval for the ill-known case.

\begin{definition}
 \label{def:ik-overlapping-interval}
Given two ill-known time intervals $I$ and $J$, this function returns an overlapping interval from the original two intervals. It is necessary to define two ill-known constraints:

\begin{align}
 \label{eq:ilc-smaller-than}
C_1 &=& \left(\leq,\mbox{First}\left(I_e, J_e \right)\right) \\
C_2 &=& \left(=, \mbox{Last}\left( I_s, J_s \right)\right)
\end{align}

\begin{equation}
 \mbox{OvInt}\left(I, J \right) = 
\begin{cases}
\left[\mbox{Last}\left( I_s, J_s \right), \mbox{First}\left(I_e, J_e \right) \right] & \mbox{ if } C_1 \wedge C_2 \neq 0.  \\
\emptyset & \mbox{ otherwise } 
\end{cases}
\end{equation}
\end{definition}

Finally, the temporal cartesian product for ill-known time intervals is defined as follows.

\begin{definition}
 \label{def:temporal-cartesian-product-ik}Temporal cartesian product.Consider the elements in definition \ref{def:generalized-fuzzy-temporal-relation}.
The temporal cartesian product is notated $r \times^{FT} s$ and is given by the following equation:
\begin{align}
 \label{eq:ik-temporal-cartesian-product}
r \tilde \times^{T} s = \left \lbrace z^{\left(n+m+2\right)}  | \exists x \in r, \exists y \in s \right. \\
\nonumber
\mbox{intersect}\left(x[S, E], y[S, E] \right) > 0 \wedge \\
\nonumber
z\left[A\right] = x\left[A\right] \wedge z\left[B\right] = y \left[B \right] \wedge \\
\nonumber
\left.  z\left[S, E \right] = \mbox{OvInt}\left(x\left[S, E\right], y\left[S, E\right] \right) \wedge z\left[S, E\right] \neq \emptyset  \right \rbrace
\end{align}
\end{definition}

The temporal cartesian product by ill-known constraints is illustrated in the following example.
\begin{example}
 \label{ex:temporal-ill-known-constraint-ik}
Consider a ill-known version of the employee's database, as illustrated by tables \ref{table:ill-known-employees} and \ref{table:ill-known-addresses}
\end{example}

\begin{table}[h]
\centering
\caption{Relation for the employees with an ill-known valid-time interval.}
\vspace{2mm}
\begin{tabular}{c c c c c c }
\hline
ID & Job & Works for & Start & Finish \\ \hline
1 & Prof. & 4 & $\left[5, 1, 1 \right]$ & $\left[10, 1, 1 \right]$ \\
2 & Tech. & 4 & $\left[3, 1, 1 \right]$ & $\left[7, 1, 1 \right]$ \\
\hline 
\end{tabular}
\label{table:ill-known-employees}
\end{table}

\begin{table}[h]
\centering
\caption{Relation for the addresses with an ill-known valid-time interval.}
\vspace{2mm}
\begin{tabular}{c c c c }
\hline
ID & Address & Start & Finish \\ \hline
1 & C/ Camino de Ronda & $\left[1, 1, 1 \right]$ & $\left[12, 1, 1 \right]$ \\
2 & C/ Recogidas &  $\left[1, 1, 1 \right]$ & $\left[5, 1, 1 \right]$ \\
\hline 
\end{tabular}
\label{table:ill-known-addresses}
\end{table}

\begin{table}[h]
\centering
\caption{Intermediate calculations for the temporal cartesian product.}
\vspace{2mm}
\begin{tabular}{c c c c c }
\hline
r.id & s.id &  intersect$\left(x\left[T\right], y\left[T\right]\right)$ &  Last$\left( I_s, J_s \right)$ & First$\left(I_e, J_e \right)$  \\ \hline
1 & 1 & 1 & $\left[5, 1, 1 \right]$ & $\left[10, 1, 1 \right]$ \\
1 & 2 & 1 & $\left[5, 1, 1 \right]$ & $\left[5, 1, 1 \right]$ \\
2 & 1 & 1 & $\left[3, 1, 1 \right]$ & $\left[7, 1, 1 \right]$ \\
2 & 2 & 1 & $\left[3, 1, 1 \right]$ & $\left[5, 1, 1 \right]$ \\
\hline 
\end{tabular}
\label{table:example-ik-cartesian-product}
\end{table}

\subsubsection{\label{sec:poss-join}Join}
The possibilistic join operator $\tilde \Join$ builds a new fuzzy temporal relation from two given relations, namely $r \in R$ and $s \in S$. This new relation is a set with all the possible combinations of tuples in both $r$ and $s$. It is usually noted as $r \tilde \Join_{a \theta b} s$ and called (theta) join, where $a,b$ are attributes from $r$ and $s$ respectively and $\theta$ a relational operator. The possibilistic temporal join definition is based on the temporal cartesian product.

\begin{definition}
 \label{def:poss-temporal-theta-join}
Temporal theta join. Let $r$ and $s$ be two  instances of relations $R, S$ respectively. Then the temporal theta join is defined as follows:
\begin{equation}
 \label{eq:poss-temporal-theta-join}
r \tilde \Join_{a \theta b}^{T} s= \sigma_{a \theta b} \left(r \tilde \times^{T} s \right)
\end{equation}

\end{definition}

% \begin{figure}
%  \includegraphics[scale=0.60]{graphs/innerjoin.eps}
% \caption{\label{fi:innerjoin} Comparison between the relational (non-temporal) inner join (left graph) and the temporal inner join (right graph).}
% \end{figure}

% 
% Figure \ref{fi:innerjoin} compares the relational inner join with respect to the temporal inner join. The right graph illustrates the temporal case. It is clear that the temporal inner join is a subset or is equal to the relational inner join.

% \subsubsection{\label{subsec:poss-Equijoin}Equijoin}
% The equijoin operator for relational databases enforces equality between specified subsets of the attributes of the relations. Then, the temporal equijoin operator is defined as the possbilistic temporal join operator.
% 
% \begin{definition}
%  \label{def:poss-equijoin}
% Equijoin. Let $R$ and $S$ be two relations and $r, s$ be the instances of each relation respectively. $A , B$ are the sets of the attributes for the relations $R$ and $S$ respectively. And let $A' \subseteq A$ and $B' \subseteq B$. Then, the temporal equijoin is defined as follows:
% \begin{equation}
%  \label{eq:poss-equijoin}
% r \Join_{r\left[A' \right] = s\left[B' \right]}^{FT} s =  \sigma^{FT}_{r\left[A' \right] = s\left[B' \right]} \left(r \times^{FT} s \right)
% \end{equation}
% \end{definition}
% 
% 
% \subsubsection{\label{subsubsec:poss-temporal-equijoin}Temporal Equijoin}
% The temporal equijoin is an equijoin operator in which the subset of attributes in the equijoin condition, are part of the primary key. Hence, the operator is defined as follows:
% 
% \begin{definition}
% \label{def:poss-temporal-equijoin}
% Temporal equijoin ($r \mbox{ TE-join }^{FT} s$). Let $R, S$ be two relations and $r, s$ the instances of each relation respectively. Consider that $PK$ is the primary key of both relations. The temporal equijoin operator is defined as follows:
% \begin{equation}
%  \label{eq:poss-temporal-equijoin} 
% r \mbox{ TE-join }^{FT} s \equiv r \Join_{r\left[PK\right] = s\left[PK\right]}^{FT} s
% \end{equation}
% \end{definition}
% 
% \subsubsection{\label{subsec:poss-natural-join}Natural Join}
% The temporal natural join is a temporal equijoin on identically named attributes.
% 
% \begin{definition}
%  \label{def:poss-temporal-natural-join}
% Temporal natural join. Consider the following relations $R = \left(A_1, \ldots, A_n, C_1, \ldots, C_k, S, E \right)$ and $S = \left(B_1, \ldots, B_m, C_1, \ldots, C_k, S, E \right)$. Let $r,s$ be instances of $R, S$ respectively. The temporal natural join is defined as follows:
% \begin{equation}
%  \label{eq:poss-temporal-natural-join}
% r \Join^{FT} s = r \Join^{FT}_{r\left[C_1\right] = s\left[C_1\right] \wedge \ldots \wedge r\left[C_k\right] = s\left[C_k\right]} s
% \end{equation}
% \end{definition}

% \todo[caption={outer-joins}]{Still have to define the outer joins implementation...}
% \subsubsection{\label{subsec:poss-outer-joins}Outer Joins}
% 
% 
% % \begin{figure}
% %  \includegraphics[scale=0.60]{graphs/outerjoin.eps}
% % \caption{\label{fi:outerjoin} Comparison between a left outer join (left graph) and a right inner join (right graph).}
% % \end{figure}

\subsection{Selection}
\subsection{Projection}
\subsection{Cartesian Product}
\subsection{Union}
\subsection{Difference}
\subsection{Intersection}
\subsection{Join}
\subsection{Division}

### 2013-02-05 trunk/joins/fuzzy-model.tex
second draft

--- 2012-09-07 temporaldml/main.tex
@@
-\documentclass{llncs}
+\begin{definition}
+\label{def:generalized-fuzzy-temporal-domain}
+\emph{Generalized fuzzy temporal domain.}
+Consider $\T$ to be the temporal domain, and let $\tilde \Pow\left( \T\right)$ be the set of all \emph{normalized} possibility distributions (see Section \ref{subsec:possibility-theory}, equation \eqref{NormalizationProperty}) defined on $\T$.
+The Generalized Fuzzy Temporal Domain, $\T_G$ is
+\begin{equation}
+\T_G \subseteq \left \lbrace \tilde \Pow\left( \T\right) \cup \text{NULL} \right \rbrace
+\end{equation}
+\end{definition}
+
+Note that $\T_G \subseteq D_G$.
+The datatypes for this domain have been studied previously in section \ref{sec:time-rep} and are shown in tables \ref{tbl:time-point-types} and \ref{tbl:time-interval-types}.
+
+A generalized fuzzy relation is defined in \cite{Medina1994}. Here, we will extend the definition to a generalized fuzzy temporal relation.
+
+\begin{definition}
+\emph{Generalized fuzzy temporal relation.}
+\label{def:fuzzy-temporal-relation}
+Consider the elements in definition \ref{def:valid-time-relation}. Some of them will be extended for the fuzzy case.
+
+\begin{itemize}
+%  \item A set of non-temporal fuzzy  or crisp attributes.
+% 	\begin{equation}
+% 	\label{eq:fuzzy-attribute-set}
+% 	A = \left \lbrace A_1, A_2, \ldots, A_n \right \rbrace
+% 	\end{equation}
+%       The domain for each attribute $A_1, \ldots, A_n$ is $D_1, \ldots, D_n$ respectively. 
+% \item The primary key $A_K$ is a subset of $A$.
+%       \begin{equation}
+%        \label{eq:fuzzy-primary-key-a}
+%       A_K \subset A
+%       \end{equation}
+% A formal definition of primary key for fuzzy relational databases will be given later in Definition \ref{def:generalized-primary-key}.
+% \item A set of two attributes; $S$  and $E$ the attributes for the starting and ending ill-known points respectively.
+$I$ is a possibilistic validity period \emph{PVP} as explained in Section \ref{subsec:ill-known-interval}.
+% \begin{equation}
+%  \label{eq:fuzzy-attribute-time-interval}
+% I = \left \lbrace S, E \right \rbrace
+% \end{equation}
+\item An attribute called version identifier, $V_{ID}$, will be added to the schema. This attribute is a counter for each different version of the entities. 
+% \begin{equation}
+%  \label{eq:fuzzy-version-identifier}
+% V_{ID} \subset \N
+% \end{equation}
+
+\item Then $R_{FTG}$, the schema for the fuzzy valid-time relation is:
+\begin{equation}
+ \label{eq:fuzzy-valid-time-relation}
+R_{FTG} = A \cup V_{ID} \cup  I
+\end{equation}
+\item The primary key for the fuzzy valid-time relation $R_{FTG}$ is:
+\begin{equation}
+ \label{eq:fuzzy-valid-time-temporal-pk}
+K_{GT} = A_K \cup V_{ID}
+\end{equation}
+A formal definition of the primary key for fuzzy valid-time relations will be given later in Definition \ref{def:generalized-fuzzy-temporal-key}.
+
+\item We will note by $r$ any valid instance of $R_{FTG}$. 
+      \begin{equation}
+       \label{eq:fuzzy-valid-time-instance}
+      r \subseteq D_1\ x\ \ldots\ x\ D_n
+      \end{equation}
+
+% \item Let $t$ be a tuple in the instance $r$; $t \in r$:
+%       \begin{itemize}
+%       \item We will note the values for the starting and the ending  points,$s$ and $e$  respectively. The value for the time interval is given by $i$.
+%       \begin{align}
+%        \label{eq:fuzzy-starting-point}
+%       s = t\left[S \right]\\
+%       e = t\left[E \right]\\
+%       i = \left(s, e\right)
+%       \end{align}
 % 
 % 
-%espero que no se enfaden por incluir este paquete:
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage[dvips]{graphicx}
-\usepackage{makeidx}  % allows for indexgeneration
-\usepackage{color}
+%       \item Let $A_k \subset A$ be the set of the non-temporal attributes that are part of the primary key (equation \eqref{eq:fuzzy-primary-key-a}).
+Then, $a_k$ denotes the values for the non-temporal attributes of the primary key.
+% 	    \begin{equation}
+% 	     \label{eq:fuzzy-pk-attribute} 
+% 	      a_k = t\left[A_K \right]
+% 	    \end{equation}
+% 
+    \item Let $K_{GT}$ be the primary key for the valid-time relation as given in equation \eqref{eq:fuzzy-valid-time-temporal-pk}. Then, $k$ denotes the values for the attributes in the primary key.
+	  \begin{equation}
+	   \label{eq:fuzzy-value-pk}
+	  k = t\left[K_{GT} \right]
+	  \end{equation}
+% 
+% 
+%       \end{itemize}
+% \item $Vt$ is the set with all the versions for a given tuple $t$.
+% 
+% \begin{equation}
+%  \label{eq:fuzzy-all-the-versions}
+% Vt = r\left(t\left[A_K\right] \right)
+% \end{equation}
+% If a tuple $t$ contains more than one version, the result of $Vt$ is a set.
+We will use the index $k$ to address the elements of the set.
+E.g.
+$Vt = \left \lbrace t_1, t_2, \ldots, t_n \right \rbrace$.
+Then $t_k , k \in \left \lbrace 1, \ldots, n \right \rbrace$ is an element of $Vt$.
+\end{itemize}
+%\end{definition}
 
-\hyphenation{pro-per-ties}
-\hyphenation{ge-ne-ral-ly}
-\hyphenation{pre-fe-ren-ces}
-\hyphenation{u-sing}
-\hyphenation{pu-nish-ment}
-\newcommand{\Pow}{\mathcal{P}}
-\newcommand{\N}{\operatorname{N}}
-\newcommand{\bool}{\operatorname*{\mathcal{B}}}
-\newcommand{\Pos}{\operatorname{Pos}}
-\newcommand{\Nec}{\operatorname{Nec}}
-\newcommand{\Open}{\operatorname{Open}}
-\newcommand{\Rp}{\operatorname{Rp}}
-\newcommand{\Rn}{\operatorname{Rn}}
-\newcommand{\FB}{\operatorname{FB}}
-\newcommand{\UC}{\operatorname{UC}}
-\newcommand{\T}{\mathcal{T}}
-\newcommand{\Rel}{\mathcal{R}}
-\newcommand{\C}{\mathcal{C}}
-\newcommand{\Nat}{\mathbb{N}}
-\newcommand{\Boolean}{\mathbb{B}}
-\newcommand{\Q}{\mathbb{Q}}
-\newcommand{\R}{\mathbb{R}}
-\newcommand{\Z}{\mathbb{Z}}
-\newcommand{\Head}{\mathcal{H}}
-\newcommand{\Body}{\mathcal{B}}
-\newcommand{\Dom}{\mathcal{D}}
-\newcommand{\INS}{\mbox{\textbf{insert}}}
-\newcommand{\DEL}{\mbox{\textbf{delete}}}
-\newcommand{\MOD}{\mbox{\textbf{modify}}}
-\newcommand{\REV}{\mbox{\textbf{revise}}}
-\newcommand{\Overlap}{\mbox{\textbf{Overlap}}}
-\newcommand{\A}{\mathcal{A}}
-\newcommand{\I}{\mathcal{I}}
-\begin{document}
-
-\mainmatter              % start of the contributions
-%
-\title{Data Manipulation Language for fuzzy temporal databases.}
-%
-%\titlerunning{A Fuzzy Valid-Time Model}  % abbreviated title (for running head)
-%                                     also used for the TOC unless
-%                                     \toctitle is used
-%
-\author{Jos\'e Enrique Pons \and Olga Pons Capote}
-%Jeffrey Dean \and David Grove \and Craig Chambers \and Kim~B.~Bruce \and
-%Elsa Bertino}
-%
-\authorrunning{Jos\'e Enrique Pons et al.} % abbreviated author list (for running head)
-%
-%%%% list of authors for the TOC (use if author list has to be modified)
-%\tocauthor{Ivar Ekeland, Roger Temam, Jeffrey Dean, David Grove,
-%Craig Chambers, Kim B. Bruce, and Elisa Bertino}
-%
-\institute{
-Department of Computer Science and Artificial Intelligence \\
-Universidad de Granada \\
-Escuela T\'ecnica Superior de Ingenier\'ia Inform\'atica \\
-C/Periodista Daniel Saucedo Aranda s/n \\
-E-18071 (Granada-Spain)\\
-\email{jpons,opc@decsai.ugr.es}\\ 
-%WWW home page:
-%\texttt{http://users/\homedir iekeland/web/welcome.html}
-%\and
-%Universit\'{e} de Paris-Sud,
-%Laboratoire d'Analyse Num\'{e}rique, B\^{a}timent 425,\\
-%F-91405 Orsay Cedex, France
-}
-
-\maketitle              % typeset the title of the contribution
-
-\begin{abstract}
-
-\end{abstract}
-
-\section{\label{sec:selection}Selection}
-The selection operator $\sigma$ obtain a subset of tuples that fulfill a set of constraints from a given relation $R$.
-The set of constraints is usually a boolean combination of atomic constraints.
-The selection operator is noted as follows:
-\begin{equation}
- \label{eq:selection}
-\sigma_{S} \left( R \right)
-\end{equation}
-
-Where $R$ is the relation, and $S$ is the selection formula. The selection formula is a set of two elements:
-% \begin{itemize}
-% \item relation.attribute boolean relation relation.attribute
-% \item relation.temporal-attribute allen-relation relation.temporal-attribute 
-% \end{itemize}
-\begin{equation}
- \label{eq:selection_formula}
-S = \left \lbrace Q, Q^{t}\right \rbrace
-\end{equation}
-
-Where $q$ is a boolean combination of (possibly) fuzzy (and bipolar) non-temporal constraints and $q^{t}$ is also a boolean combination of fuzzy temporal constraints.
-
-\begin{equation}
- \label{eq:non-temporal-constraints}
-Q = \left \lbrace q^{\left \lbrace \mbox{pos,neg} \right \rbrace}_{\mbox{attribute}_1}  \theta \mbox{ val }, \ldots, q^{\left \lbrace \mbox{pos,neg} \right \rbrace}_{\mbox{attribute}_n}  \theta \mbox{ val } \right \rbrace
-\end{equation}
-Where $\theta$ is an artihmetic operator; one of $\left \lbrace =, \neq, <, \leq, >, \geq \right \rbrace$ and val is a value in the domain of the attribute.
-
-Then, consider a boolean function $\bool:\Boolean^{n}  \rightarrow \Boolean$. The evaluation function for the constraints in $Q$  is given by:
-
-\begin{equation}
- \label{eq:evaluation-function}
-\lambda : Q \rightarrow \Boolean : \lambda (Q) \rightarrow \bool \left(q^{\left \lbrace \mbox{pos,neg} \right \rbrace}_{\mbox{attribute}_1}  \theta \mbox{ val }, \ldots, q^{\left \lbrace \mbox{pos,neg} \right \rbrace}_{\mbox{attribute}_n}  \theta \mbox{ val } \right)
-\end{equation}
-
-Now, the uncertainty about the evaluation of $Q$ and a boolean function $\bool:\Boolean^{n}  \rightarrow \Boolean$ is given by:
-
-\begin{equation}
- \label{eq:evaluation-lambda-function}
-\pi_{\lambda \left(Q \right)} = \tilde{\bool} \left(\pi_{q^{\left \lbrace \mbox{pos,neg} \right \rbrace}_{\mbox{attribute}_1}  \theta \mbox{ val }}, \ldots, \pi_{q^{\left \lbrace \mbox{pos,neg} \right \rbrace}_{\mbox{attribute}_n}  \theta \mbox{ val }} \right)
-\end{equation}
-
-The temporal constraints, $Q^t$ are provided in a similar way.
-The main difference is that, instead of comparations like $\left \lbrace =, \neq, <, \leq, >, \geq \right \rbrace$, we use the allen's relations.
-Hence, the representation of the temporal constraints is expressed as follows:
-
-\begin{equation}
-\label{eq:temporal-constratins}
-Q^{t} = \left \lbrace q_{\mbox{attribute}_1}  ar \mbox{ val }, \ldots, q_{\mbox{attribute}_n}  ar \mbox{ val } \right \rbrace
-\end{equation}
-Here, $ar$ is one of the thirdteen Allen's relations (equal, before, overlaps, starts, finishes, meets, during) and their respective reverses.
-
-Now the combination of temporal and non-temporal results should be explicited here.
-
-\section{\label{sec:join}Join}
-The join operator $\Join$ builds a new relation from two given relations, namely $R$ and $S$.
-This new relation is a set with all the possible combinations of tuples in both $R$ and $S$.
-It is usually noted as $R \Join_{r \theta s} S$, where $r,s$ are attributes from $R$ and $S$ respectively and $\theta$ an arithmetic operator.
-
-The following example illustrates the behaviour of the join operator in a crisp temporal database.
+We will illustrate the definitions with an example.
 
 \begin{example}
-Consider the employees database shown in Table \ref{}.  
+Consider the set of attributes $A = \left \lbrace A_1, A_2, A_3 \right \rbrace$.
+The primary key for these attributes is given by $A_K = \left \lbrace A_1, A_2 \right \rbrace$.
+Let $I = \left \lbrace S, E \right \rbrace$ be the set of temporal attributes that define the possibilistic validity period of the data.
+$R_{FTG} = A \cup V_{ID} \cup I$ is the valid-time relation and $r$ is an instance of the relation.
+The instance $r$ is given by the following elements.
+$r = \left \lbrace \left(a_{11}, a_{12}, a_{13}, 001, s_1 ,e_1 \right) \right.$,  $\left(a_{21}, a_{22}, a_{23}, 001, s_2, e_2 \right)$ , $\left.
+\left(a_{11}, a_{12}, a_{31}, 002, s_3, e_3 \right) \right \rbrace$.
+The instance $r$ is illustrated in Table \ref{tbl:fuzzy-sample-definitions}.
+Consider the tuple $t = \left(a_{11}, a_{12}, a_{13}, s_1 ,e_1 \right)$. Then,
+\begin{align}
+ \nonumber
+s_1 &= t\left[S \right]\\
+ \nonumber
+e_1 &= t[E]\\
+ \nonumber
+i\ \ &= \left(s_1, e_1\right)\\
+ \nonumber
+a_k &= t\left[A_K\right] = \left(a_{11}, a_{12} \right) \\
+ \nonumber
+k &= t\left[PK\right] =\left(a_{11}, a_{12}, s_1, e_1\right)\\
+ \nonumber
+Vt &= r\left(t\left[A_K\right] \right) = \left \lbrace t_1, t_3 \right \rbrace
+\end{align}
+
 \end{example}
 
-\section*{Acknowledgements}
+\begin{table}[h]
+\caption[Example of fuzzy valid-time relation.]{\label{tbl:fuzzy-sample-definitions}Sample database containing the instance $r$ of the fuzzy valid-time relation $R_{FTG}$.}
+\centering
+\begin{tabular}{c c c c c c c}\\ \hline
+& \textbf{A$_1$}  & \textbf{A$_2$}  & $A_3$ & $V_{ID}$ & $S$ & $E$ \\
+\hline
+$t_1$&$a_{11}$ & $a_{12}$ & $a_{13}$ & $001$ & $s_1$ & $e_1$ \\
+$t_2$ & $a_{21}$ & $a_{22}$ & $a_{23}$& $001$ &  $s_2$ & $e_2$ \\
+$t_3$ & $a_{11}$ & $a_{12}$ & $a_{31}$& $002$ & $s_3$ & $e_3$\\
+\hline\\
+\end{tabular}
+\end{table}
+
 % 
-The researchers are supported by the grant BES-2009-013805 within the research project TIN2008-02066: \emph{Fuzzy Temporal Information treatment in relational DBMS}.
+% 
+% Consider $\A$ to be the set of all the entities, and let $A=\left(A_1, \ldots, A_n \right), A \in \A$ be the set of (fuzzy) attributes that define an entity.
+Let $\I_{PVP}$ be the set of all the ill-known time intervals and let $I = \left(X, Y \right)$ be a PVP, $I \in \I_{PVP}$.
+The pair $\left(A, I\right)$ expresses that the data regarding the entity A are valid during the ill-known time interval I.
+Let $R_{FTG} \subseteq \A \  x\  \I_{PVP}$ be a valid-time relation.
+Then, the following equation indicates that the pair $\left(A, I\right)$ is in the relation R:
+% 
+% \begin{equation}
+% \label{eq:fuzzy-rel-def}
+% \left( R_{FTG}, \left(A , I\right) \right)
+% \end{equation}
+% 
+ A generalized fuzzy temporal relation $R_{FTG}$ can be noted also by:
+\label{def:generalized-fuzzy-temporal-relation}
+\begin{equation}
+\label{eq:generalized-fuzzy-temporal-relation}
+R_{FTG} = \left(\Head, \Body \right)
+\end{equation}
+Where $\Head$ is the Head of the relation and consist on a fixed set of triplets attribute- domain - compatibility with an optional the valid-time attribute:
 
-\bibliographystyle{splncs03}
-\bibliography{biblio}
+\begin{align}
+\label{eq:head-valid-time}
+\Head = \big \lbrace \left(A_{G1}:D_{G1}\left[,C_{A_{G1}} \right] \right),\\
+\nonumber
+ \ldots,\\
+ \nonumber
+  \left(A_{Gn}:D_{Gn}\left[,C_{A_{Gn}} \right] \right),\\
+  \nonumber
+  \Big[  \left( \text{PVP}, D_{\text{PVP}}\left[,C_{A_{\text{PVP}}} \right] \right) \Big] \big \rbrace
+\end{align}
+Note that $D_{Gj}$ ($j = 1, \ldots, n$) is the domain for the attribute $A_{Gj}$. $C_{A_{Gj}}$ is the compatibility attribute in the unit interval $\left[0, 1 \right]$.
 
-\end{document}
+$\Body$ is the body of the relation and it consists on a set of tuples. Each tuple is a triplet attribute- value- degree with an optional valid-time attribute:
+
+\begin{align}
+\label{eq:body-valid-time}
+\Body = \big \lbrace \left(A_{G1}:\tilde{d}_{i1}\left[,c_{i1} \right] \right),\\
+\nonumber
+ \ldots,\\
+ \nonumber
+  \left(A_{Gn}:\tilde{d}_{in}\left[,c_{in} \right] \right),\\
+  \nonumber
+   \Big[  \left( \text{PVP}, \tilde{d}_{\text{PVP}} \left[,C_{A_{\text{PVP}}} \right] \right)  \Big] \big \rbrace
+\end{align}
+
+\end{definition}
+
+The definition in \cite{Medina1994} for $R_{FTG}$ shows that classical relations are a particular case of this model.

### 2013-02-05 trunk/joins/temporal-model.tex
second draft

--- 2012-09-07 temporaldml/main.tex
@@
-\documentclass{llncs}
-%
-%
-%espero que no se enfaden por incluir este paquete:
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage[dvips]{graphicx}
-\usepackage{makeidx}  % allows for indexgeneration
-\usepackage{color}
+We will define the main concepts and elements in order to implement a model for temporal relational databases. First of all, we will introduce some definitions and notations.
 
-\hyphenation{pro-per-ties}
-\hyphenation{ge-ne-ral-ly}
-\hyphenation{pre-fe-ren-ces}
-\hyphenation{u-sing}
-\hyphenation{pu-nish-ment}
-\newcommand{\Pow}{\mathcal{P}}
-\newcommand{\N}{\operatorname{N}}
-\newcommand{\bool}{\operatorname*{\mathcal{B}}}
-\newcommand{\Pos}{\operatorname{Pos}}
-\newcommand{\Nec}{\operatorname{Nec}}
-\newcommand{\Open}{\operatorname{Open}}
-\newcommand{\Rp}{\operatorname{Rp}}
-\newcommand{\Rn}{\operatorname{Rn}}
-\newcommand{\FB}{\operatorname{FB}}
-\newcommand{\UC}{\operatorname{UC}}
-\newcommand{\T}{\mathcal{T}}
-\newcommand{\Rel}{\mathcal{R}}
-\newcommand{\C}{\mathcal{C}}
-\newcommand{\Nat}{\mathbb{N}}
-\newcommand{\Boolean}{\mathbb{B}}
-\newcommand{\Q}{\mathbb{Q}}
-\newcommand{\R}{\mathbb{R}}
-\newcommand{\Z}{\mathbb{Z}}
-\newcommand{\Head}{\mathcal{H}}
-\newcommand{\Body}{\mathcal{B}}
-\newcommand{\Dom}{\mathcal{D}}
-\newcommand{\INS}{\mbox{\textbf{insert}}}
-\newcommand{\DEL}{\mbox{\textbf{delete}}}
-\newcommand{\MOD}{\mbox{\textbf{modify}}}
-\newcommand{\REV}{\mbox{\textbf{revise}}}
-\newcommand{\Overlap}{\mbox{\textbf{Overlap}}}
-\newcommand{\A}{\mathcal{A}}
-\newcommand{\I}{\mathcal{I}}
-\begin{document}
+\begin{definition}
+\label{def:valid-time-relation}
+\emph{Valid-time relation.}
+Consider the following definitions and notations:
 
-\mainmatter              % start of the contributions
-%
-\title{Data Manipulation Language for fuzzy temporal databases.}
-%
-%\titlerunning{A Fuzzy Valid-Time Model}  % abbreviated title (for running head)
-%                                     also used for the TOC unless
-%                                     \toctitle is used
-%
-\author{Jos\'e Enrique Pons \and Olga Pons Capote}
-%Jeffrey Dean \and David Grove \and Craig Chambers \and Kim~B.~Bruce \and
-%Elsa Bertino}
-%
-\authorrunning{Jos\'e Enrique Pons et al.} % abbreviated author list (for running head)
-%
-%%%% list of authors for the TOC (use if author list has to be modified)
-%\tocauthor{Ivar Ekeland, Roger Temam, Jeffrey Dean, David Grove,
-%Craig Chambers, Kim B. Bruce, and Elisa Bertino}
-%
-\institute{
-Department of Computer Science and Artificial Intelligence \\
-Universidad de Granada \\
-Escuela T\'ecnica Superior de Ingenier\'ia Inform\'atica \\
-C/Periodista Daniel Saucedo Aranda s/n \\
-E-18071 (Granada-Spain)\\
-\email{jpons,opc@decsai.ugr.es}\\ 
-%WWW home page:
-%\texttt{http://users/\homedir iekeland/web/welcome.html}
-%\and
-%Universit\'{e} de Paris-Sud,
-%Laboratoire d'Analyse Num\'{e}rique, B\^{a}timent 425,\\
-%F-91405 Orsay Cedex, France
-}
-
-\maketitle              % typeset the title of the contribution
-
-\begin{abstract}
-
-\end{abstract}
-
-\section{\label{sec:selection}Selection}
-The selection operator $\sigma$ obtain a subset of tuples that fulfill a set of constraints from a given relation $R$.
-The set of constraints is usually a boolean combination of atomic constraints.
-The selection operator is noted as follows:
+\begin{itemize}
+ \item A set of non-temporal attributes.
 	\begin{equation}
- \label{eq:selection}
-\sigma_{S} \left( R \right)
+	\label{eq:attribute-set}
+	A = \left \lbrace A_1, A_2, \ldots, A_n \right \rbrace
+	\end{equation}
+      The domain for each attribute $A_1, \ldots, A_n$ is $D_1, \ldots, D_n$ respectively. 
+\item The original primary key $A_K$ is a subset of the attributes in $A$.
+      \begin{equation}
+       \label{eq:primary-key-a}
+      A_K \subseteq A
+      \end{equation}
+\item Two attributes, $S$  and $E$ for the starting and ending points respectively. $I$ defines the valid time interval for the data. 
+\begin{equation}
+ \label{eq:attribute-time-interval}
+I = \left \lbrace S, E \right \rbrace
 \end{equation}
 
-Where $R$ is the relation, and $S$ is the selection formula. The selection formula is a set of two elements:
-% \begin{itemize}
-% \item relation.attribute boolean relation relation.attribute
-% \item relation.temporal-attribute allen-relation relation.temporal-attribute 
-% \end{itemize}
+$\T$ is the time domain.
+
+\item Then $R$, the schema for the valid-time relation is:
 \begin{equation}
- \label{eq:selection_formula}
-S = \left \lbrace Q, Q^{t}\right \rbrace
+ \label{eq:valid-time-relation}
+R = A \cup  I
+\end{equation}
+\item The primary key for the valid-time relation $R$ is:
+\begin{equation}
+ \label{eq:valid-time-temporal-pk}
+PK = A_K \cup I
 \end{equation}
 
-Where $q$ is a boolean combination of (possibly) fuzzy (and bipolar) non-temporal constraints and $q^{t}$ is also a boolean combination of fuzzy temporal constraints.
+\item We will note by $r$ any valid instance of $R$. 
+      \begin{equation}
+       \label{eq:valid-time-instance}
+      r \subseteq D_1\ x\ \ldots\ x\ D_n x\ T x\ T
+      \end{equation}
+
+     % \end{itemize}
+\item $V(t)$ is the set of all the versions for a given tuple $t$. Formally,
 
 \begin{equation}
- \label{eq:non-temporal-constraints}
-Q = \left \lbrace q^{\left \lbrace \mbox{pos,neg} \right \rbrace}_{\mbox{attribute}_1}  \theta \mbox{ val }, \ldots, q^{\left \lbrace \mbox{pos,neg} \right \rbrace}_{\mbox{attribute}_n}  \theta \mbox{ val } \right \rbrace
+ \label{eq:all-the-versions}
+V(t) =\left \lbrace t_i \in r, t_i\left[A_K\right] = t\left[A_K\right] \right \rbrace
 \end{equation}
-Where $\theta$ is an artihmetic operator; one of $\left \lbrace =, \neq, <, \leq, >, \geq \right \rbrace$ and val is a value in the domain of the attribute.
-
-Then, consider a boolean function $\bool:\Boolean^{n}  \rightarrow \Boolean$. The evaluation function for the constraints in $Q$  is given by:
-
-\begin{equation}
- \label{eq:evaluation-function}
-\lambda : Q \rightarrow \Boolean : \lambda (Q) \rightarrow \bool \left(q^{\left \lbrace \mbox{pos,neg} \right \rbrace}_{\mbox{attribute}_1}  \theta \mbox{ val }, \ldots, q^{\left \lbrace \mbox{pos,neg} \right \rbrace}_{\mbox{attribute}_n}  \theta \mbox{ val } \right)
-\end{equation}
-
-Now, the uncertainty about the evaluation of $Q$ and a boolean function $\bool:\Boolean^{n}  \rightarrow \Boolean$ is given by:
-
-\begin{equation}
- \label{eq:evaluation-lambda-function}
-\pi_{\lambda \left(Q \right)} = \tilde{\bool} \left(\pi_{q^{\left \lbrace \mbox{pos,neg} \right \rbrace}_{\mbox{attribute}_1}  \theta \mbox{ val }}, \ldots, \pi_{q^{\left \lbrace \mbox{pos,neg} \right \rbrace}_{\mbox{attribute}_n}  \theta \mbox{ val }} \right)
-\end{equation}
-
-The temporal constraints, $Q^t$ are provided in a similar way.
-The main difference is that, instead of comparations like $\left \lbrace =, \neq, <, \leq, >, \geq \right \rbrace$, we use the allen's relations.
-Hence, the representation of the temporal constraints is expressed as follows:
-
-\begin{equation}
-\label{eq:temporal-constratins}
-Q^{t} = \left \lbrace q_{\mbox{attribute}_1}  ar \mbox{ val }, \ldots, q_{\mbox{attribute}_n}  ar \mbox{ val } \right \rbrace
-\end{equation}
-Here, $ar$ is one of the thirdteen Allen's relations (equal, before, overlaps, starts, finishes, meets, during) and their respective reverses.
-
-Now the combination of temporal and non-temporal results should be explicited here.
-
-\section{\label{sec:join}Join}
-The join operator $\Join$ builds a new relation from two given relations, namely $R$ and $S$.
-This new relation is a set with all the possible combinations of tuples in both $R$ and $S$.
-It is usually noted as $R \Join_{r \theta s} S$, where $r,s$ are attributes from $R$ and $S$ respectively and $\theta$ an arithmetic operator.
-
-The following example illustrates the behaviour of the join operator in a crisp temporal database.
+Obviously, $t$ itself is included in the set.
+% If a tuple $t$ contains more than one version, the result of $Vt$ is a set.
+We will use the index $k$ to address the elements of the set.
+E.g.
+$Vt = \left \lbrace t_1, t_2, \ldots, t_n \right \rbrace$.
+Then $t_k , k \in \left \lbrace 1, \ldots, n \right \rbrace$ is an element of $Vt$.
+\end{itemize}
+\end{definition}
+We will illustrate the definitions with an example.
 
 \begin{example}
-Consider the employees database shown in Table \ref{}.  
+Consider the set of attributes $A = \left \lbrace A_1, A_2, A_3 \right \rbrace$.
+The primary key for these attributes is given by $A_K = \left \lbrace A_1, A_2 \right \rbrace$.
+Let $I = \left \lbrace S, E \right \rbrace$ be the set of temporal attributes that define the validity period of the data.
+$R$ is the valid-time relation and $r$ is an instance of the relation.
+The instance $r$ is given by the following elements.
+$r = \left \lbrace \left(a_{11}, a_{12}, a_{13}, s_1 ,e_1 \right) \right.$,  $\left(a_{21}, a_{22}, a_{23}, s_2, e_2 \right)$ , $\left.
+\left(a_{11}, a_{12}, a_{31}, s_3, e_3 \right) \right \rbrace$.
+The instance $r$ is illustrated in Table \ref{tbl:sample-definitions}.
+Consider the tuple $t_1 = \left(a_{11}, a_{12}, a_{13}, s_1 ,e_1 \right)$. Then,
+\begin{align}
+ \nonumber
+ t_1\left[S \right]&=s_1\\
+ \nonumber
+t_1[E]&=e_1\\
+ \nonumber
+t_1[S,E]&= \left(s_1, e_1\right)\\
+ \nonumber
+t_1\left[A_K\right] &= \left(a_{11}, a_{12} \right) \\
+ \nonumber
+t_1\left[PK\right] &=\left(a_{11}, a_{12}, s_1, e_1\right)\\
+ \nonumber
+V(t_1) &= \left \lbrace t_1, t_3 \right \rbrace
+\end{align}
+
 \end{example}
 
-\section*{Acknowledgements}
-%
-The researchers are supported by the grant BES-2009-013805 within the research project TIN2008-02066: \emph{Fuzzy Temporal Information treatment in relational DBMS}.
-
-\bibliographystyle{splncs03}
-\bibliography{biblio}
-
-\end{document}
+\begin{table}[h]
+\centering
+%\begin{table}[htbp]
+\caption[Example database for $r \in R$]{\label{tbl:sample-definitions}Example database containing the instance $r$ of the valid-time relation $R$.}
+\vspace{2mm}
+%\centerline{\small DATA TYPES}
+\begin{tabular}{c c c c c c}\\
+\hline
+& \textbf{A$_1$}  & \textbf{A$_2$}  & $A_3$ & $S$ & $E$ \\
+\hline
+$t_1$&$a_{11}$ & $a_{12}$ & $a_{13}$ & $s_1$ & $e_1$ \\
+$t_2$ & $a_{21}$ & $a_{22}$ & $a_{23}$ &  $s_2$ & $e_2$ \\
+$t_3$ & $a_{11}$ & $a_{12}$ & $a_{31}$ & $s_3$ & $e_3$\\
+\hline\\
+\end{tabular}
+\end{table}

### 2013-02-05 trunk/joins/time.tex
second draft

--- 2012-06-12 trunk/ijcis-pos-model/sources/temporaldatabases.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{\label{subsec:time-in-databases}Time in Databases}
-The concept of time has been studied in the database framework for a long time.
+The concept of time has been studied in databases for a long time.
 A true standard for adding temporal aspects to relational databases does not exist, but there is a consensus in the literature \cite{Dyreson1994} on what is called a \emph{temporal database}: a temporal database is a database dealing with some aspects of time in its schema.
@@
 	\item
-	\textbf{Transaction time} (\emph{TT}) \cite{Rowe1987,Jensen1991} denotes the time when the fact (object) is stored in the database.
-It is usually append-only: as the past cannot be changed, TT cannot be changed either.
+	\textbf{Transaction time} (\emph{TT}) \cite{L.Rowe1987},\cite{Jensen1991a} denotes the time when the fact (object) is stored in the database.
+It is usually append-only: as the past can not be changed, TT can not be changed neither.
 Furthermore, at the moment of insertion, a TT can be neither in the past nor in the future.
 	\item
-	\textbf{Valid time} (\emph{VT}) \cite{Jensen1994,Sarda1990,McKenzie1981} denotes the time when the fact (object) is true in the modelled reality.
-A fuzzy extension has been proposed by \cite{Garrido2009}.
+	\textbf{Valid time} (\emph{VT}) \cite{Jensen1994},\cite{Sarda1990} denotes the time when the fact (object) is true in the modelled reality. A fuzzy extension has been proposed by \cite{Garrido2009}. 
 %	\item
@@
 	\item
-	\textbf{Decision time} (\emph{DT}) \cite{Nascimento1995,Chakravarthy1994,Etzion1992,Ozsoyoglu1995} denotes the time when an event was decided to happen. 
+	\textbf{Decision time} (\emph{DT}, proposed in \cite{Nascimento1995}) denotes the time when an event was decided to happen. 
 	\end{itemize}
 
-	For example, consider a database containing employee contract descriptions.
-The time when the employee's contract is valid, represented by an interval, is the VT.
+	E.g., consider a database containing employee contract descriptions.
+The time when the employee's contract is valid, represented by an interval, is VT.
 The time when the employee's contract is stored in the database is the TT.
@@
 This means that the attribute uses the date/time domain, but the database model does not treat the attribute differently from non-temporal attributes.
-Depending on the time managed, a database is classified as either a \textbf{Valid Time Database} (\emph{VTDB}), a \textbf{Transaction Time Database} (\emph{TTDB}), a \textbf{bi-temporal database} (both valid and transaction time are managed) or a \textbf{tri-temporal} or \emph{multitemporal} database (valid time, transaction time and decision time are managed).
+Depending on the time managed, a database is classified as either a \textbf{Valid Time Database} (\emph{VTDB}), a \textbf{Transaction Time Database} (\emph{TTDB}), a \textbf{bi-temporal database} (both valid and transaction time are managed) or a \textbf{tri-temporal database} (valid time, transaction time and decision time are managed).
 
@@
 %%	
-%	In order to define a temporal database model, some basic elements should be defined \cite{Dyreson:1994:CGT:181550.181560}:
+%	In order to define a temporal database model, some basic elements should be defined \cite{Dyreson1994}:
 %	\begin{itemize}
@@
 \subsubsection{Imperfection and time}
-The representation of imprecision and its semantics when dealing with time has been studied for a long time.
-Several proposals for representing and handling imprecise time indications can be found in \cite{DeCaluwe1997,DeTre1997}.
+Representing imprecision and its semantics when dealing with time has been studied for a long time.
+Several proposals for representing and computing imprecise time indications can be found in \cite{DeCaluwe1997} and \cite{DeTre1997a}.
 Also, the changes between several granularities can be seen as a source of imprecision \cite{Devos1998}.
 
-In this paper we will consider two kinds of imperfection:
+In the proposal section we will consider two kinds of imprecision:
 \begin{itemize}
-\item \textbf{Imperfection in the database} the knowledge about the temporal data contains some imperfection. E.g., a database record shows that \emph{`The car was in the garage around April.'}
+\item \textbf{Uncertainty in the database} denotes the uncertainty that arises when the knowledge about the temporal data in the database is uncertain.
+E.g., a database record shows that \emph{`The car is in the garage around April.'}
  \item \textbf{Imprecision in the query specification} denotes the imprecision in the specification of temporal criteria by the user, when querying.
-E.g., \emph{`The user wants a car which was in the garage around April.'}
+E.g., \emph{`The user wants to obtain a car which is red and which is in the garage around April.'}
 \end{itemize}
@@
 Several proposals for managing uncertain time in a database exist.
-Some proposals work with rough sets \cite{Qiang2009}, and some others rely on possibility distributions for representing uncertainty in time \cite{Dyreson1998,Garrido2009,Galindo2001}.
-In order to compare temporal possibility distributions, the extensions of the classical Allen's operators \cite{Allen1983} are defined in \cite{Ohlbach2004,Nagypal2003,Dubois2003a,Schockaert2008}.
-%In the proposal section, we will follow the representation by means of possibility distributions, in order to work with both satisfaction and dissatisfaction degrees.
+Some proposals work with rough sets \cite{Qiang2009}, other proposals rely on possibility distributions for representing uncertainty in time \cite{Garrido2009}, \cite{Galindo2001}.
+In order to compare temporal possibility distributions, extensions of the classical Allen's operators \cite{Allen1983} are defined in \cite{Ohlbach2004} and \cite{Nagypal2003}.
+In the proposal section, we will follow the representation by means of possibility distributions, in order to work with both satisfaction and dissatisfaction degrees.
 Also, in order to work properly with fuzzy operators, the underlying domain should be numeric.
-% Take a look into this:
-%In this paper, the representation for the dates will follow the Julian Day Number (JDN) representation \cite{Dir96}.
+In this paper, the representation for the dates will follow the Julian Day Number (JDN) representation \cite{Husfeld1996}.
 
-%If the starting points and/or the end points of the interval representing the time are not known precisely, it is easy to fuzzify them, using, e.g., two triangular possibility distributions.
-
-In order to deal with uncertainty in time intervals, several proposals have been made.
-Here, two approaches are introduced: the first one, based on \emph{Fuzzy Validity Periods}~\cite{Garrido2009} and the second one, based on \emph{Possibilistic Valid-time Periods}~\cite{JoseEnriquePons2012}.
-
-\begin{definition}
-A \textbf{Fuzzy Validity Period} (\emph{FVP}) is defined as a fuzzy time interval specifying when the data regarding an object are valid.
-A fuzzy time interval is then the fuzzification of a crisp time interval.\\
-\end{definition}
-
+If the starting points and/or the end points of the interval representing the time are not known precisely, it is easy to fuzzify them, using, e.g., two triangular possibility distributions.
+A \textbf{Fuzzy Validity Period} (\emph{FVP}) is defined as a fuzzy time interval specifying when an object is valid.
+A fuzzy time interval is then the fuzzification of a crisp time interval.
 Several options to transform possibility distributions corresponding to the fuzzy starting point and the fuzzy end point into one consistent FVP exist \cite{Garrido2009}, e.g (Fig.
@@
 \begin{itemize}
-\item The \textbf{convex hull} approach is the most intuitive approach. The resulting FVP is the convex hull of the union of both possibility distributions.
-\item The \textbf{uncertainty preserving} approach. The amount of uncertainty is maintained at the edges of the possibility distribution representing the FVP \cite{Garrido2009}.
+\item The \textbf{convex hull} approach is the most intuitive approach. The resulting FVP is the convex hull of the union of both fuzzy sets.
+\item The \textbf{uncertainty preserving} approach is less intuitive but more realistic.
+The amount of uncertainty is maintained at the edges of the possibility distribution representing the FVP \cite{Garrido2009}.
 \end{itemize}
@@
 %%%%%%
-\vspace*{13pt}
-\begin{center}
-{
-\includegraphics[scale=0.25]{./graphs/comparisoncv.pdf}
-
-}
-\end{center}
-%\centerline{ \psfig{file=./graphs/Y-time-point.eps}}
-\vspace*{10pt}
-\fcaption{\label{fig:fuzzy-validity-period}Transformation to obtain the FVP.
+\begin{figure}[h!]
+  \centering
+  \includegraphics[scale=0.3]{graphs/trapezoidalDistribution.eps}
+  \caption{Transformation to obtain the FVP.
 The top graph shows the two triangular possibility distributions.
 The middle graph shows the convex hull validity period, the bottom one shows the result of the second transformation, which maintains the imprecision.}
-%  \label{fig:fuzzy-validity-period}
-\vspace*{13pt}
-
-The main feature for the FVP is the optimization for the storage.
-The compact representation is the result of a conjunctive semantic.
-The object is valid within all the time points inside the starting and ending points.
-
-\begin{definition}
-A \textbf{Possibilistic Valid-time Period} (\emph{PVP}) is an ill-known interval of time specifying when the data regarding an object are valid.
-\end{definition}
-A PVP is an ill-known interval in the sense of definition \ref{def;possibilistic-variable}.
-Note that this representation is \emph{disjunctive}: the PVP represents only one crisp time interval, but for some reason, it is (partially) unknown.
-
-The main feature for the PVP is that it preserves all the information for both starting and ending points\cite{Pons2011,Pons2012}.
-Table \ref{tbl:comparative-pvp-fvp} is a comparative between PVP and FVP.
-The following list defines the items in the comparative:
-
-\begin{enumerate}
-\item Domain: The domain of the possibility distribution modelled by the approach.
-\item Implementation of relationships: How to implement a relationship.
-\item Allen's relations: Are the Allen's relations defined?
-\item Storage: The way the data are stored in the database.
-\item Possibility measures: Does the framework provide always a possibility measure for any relation between the temporal elements?
-\item Necessity measures:  Does the framework provide always a necessity measure for any relation between the temporal elements?
-\end{enumerate}
-
-%%
-%% comparison table between pvp and fvp.
-%%
-\vglue13pt
-%\begin{table}[htbp]
-\tcap{Comparative PVP vs FVP}
-%\centerline{\small DATA TYPES}
-\vglue-6pt
-\centerline{\small\baselineskip=13pt
-\begin{tabular}{c p{2cm} p{2cm}}\\
-Item & PVP & FVP\\
-\hline
-(1) & $\Pow(\R)$ & $\R$ \\
-(2) & Ill-known constraints. & Ad-hoc operators. \\
-(3) & $\checkmark$ & - \\
-(4) & Two distributions one for each endpoint. & Only one distribution. \\ 
-(5) & $\checkmark$ & $\checkmark$ \\
-(6) & $\checkmark$ & - \\
-\hline\\
-\end{tabular}
-\label{tbl:comparative-pvp-fvp}
-}
-
-In the rest of the paper we will work only with PVP to represent valid-time intervals.
-
-\subsection{\label{subsubsec:Understanding-valid-time-databases}Understanding Valid-time Databases}
-This subsection is devoted to describe the behaviour of a crisp valid-time database.
-For the sake of simplicity, only the three main operations in the Data Manipulation Language (CReate Update, Delete) are shown.
-Usually the DML operations in a temporal database are re-defined (a typical update sentence in SQL could be re-defined by means of several insert and update sentences).
-Therefore, for the sake of clarity, these high level primitives in the DML for a valid-time database are usually noted as Insert, Modify and Delete.
-In the following subsections each primitive is defined and explained.
-Finally, an illustrative example is given.
-For a more complete information on the behaviour of a bitemporal database, please refer to ~\cite{Jensen1994}.
-
-\begin{definition}
-\emph{Valid-time relation.}
-Consider $A=\left(A_1, \ldots, A_n \right)$ be the set of attributes that define an entity and $I = \left(s_i,e_i \right)$ the set of attribute defining the time interval.
-The pair $\left(A, I\right)$ expresses that the data regarding the entities A are valid during the crisp time interval I.
-Let $R \subseteq A\  x\  I$ be a valid-time relation.
-Then, the following equation indicates that the pair $\left(A, I\right)$ is in the relation R:
-
-\begin{equation}
-\label{eq:rel-def}
-\left( R, \left(A , I\right) \right)
-\end{equation}
-\end{definition}
-
-In order to simplify the algorithms for the manipulation of data, some auxiliary functions and constants are defined:
-
-\begin{definition}
- \label{def:from-beginning}
-\emph{From the beginning (FB).} 
-Consider $A=\left(A_1, \ldots, A_n \right)$ be the set of attributes that define an entity and $I = \left(s_i,e_i \right)$ the set of attribute defining the time interval.
-Let $R \subseteq A\  x\  I$ be a valid-time relation.
-The from the beginning constant is used when the value for $s_i$ is equal to $-\infty$.
-\end{definition}
-
-\begin{definition}
- \label{def:until-changed}
-\emph{Until changed (UC).} 
-Consider $A=\left(A_1, \ldots, A_n \right)$ be the set of attributes that define an entity and $I = \left(s_i,e_i \right)$ the set of attribute defining the time interval.
-Let $R \subseteq A\  x\  I$ be a valid-time relation.
-The until changed constant is used when the value for $e_i$ is equal to $+\infty$.
-\end{definition}
-
-If the pair $\left(A, I \right)$ with $I = \left(c, UC \right), c \in \T$ is in the relation $R$, then, it is said that the pair $\left(A, I \right)$ is \emph{current} in the relation.
-
-For example, last row in Table \ref{tbl:library-sample}.
-The value for the time interval is $I = \left(4/4/2012 , UC \right)$.
-The meaning is that the patient with ID = 3 was in the hospital the  4/4/2012 and it is still there.
-The patient with ID = 3 is current in the relation.
-
-\begin{example}
-\label{ex:library-database}
- Consider a hospital database.
-Each patient is stored in the database with an identifier (ID), the name and the dates when the patient entered and left the hospital.
-Table \ref{tbl:library-sample} illustrates the first version of the database, after three insertions.
-In this example, $A = ($ ID, Name $)$ and $I = ($ Arrival, Leave $)$.
-\end{example}
-
-\vglue13pt
-%\begin{table}[htbp]
-\tcap{\label{tbl:library-sample}Sample hospital database}
-%\centerline{\small DATA TYPES}
-\vglue-6pt
-\centerline{\small\baselineskip=13pt
-\begin{tabular}{c c c c}\\
-\hline
-\textbf{ID}  & Name  & Arrival & Leave \\
-\hline
-3 & John & 15/3/2012 & 30/3/2012 \\
-4 & Antoon & 25/3/2012 & 4/4/2012 \\
-5 & Daan & 18/3/2012 & 2/4/2012\\
-3 & John & 4/4/2012 & UC \\
-\hline\\
-\end{tabular}
-}
-
-\begin{definition}
-\label{def:current-in-relation}
-\emph{Current $\left(R, A \right)$.}
-Consider $A=\left(A_1, \ldots, A_n \right)$ be the set of attributes that define an entity and $I = \left(s_i,UC \right)$ the set of attribute defining the time interval.
-Let $R \subseteq A\  x\  I$ be a valid-time relation.
-The function Current $\left(R, A \right)$ returns the crisp time interval $I$ if the entity $A$ is currently in the relation $R$ or the empty set, if the entity $A$ is not currently in the relation.
-
-%Consider an object $A$ and a relation $R$ in a temporal database.
-The function Current$\left(R, A \right)$ obtains the tuple $\left(A, I \right)$ that is current in the relation $R$, that is it, $I = \left[a, +\infty \right]$.
-
-\begin{align}
-\label{eq:current-in-relation}
-\mbox{Current} \left(R, A \right) &=& \\ 
-\begin{cases}
-\nonumber
-I & \mbox{ if } (A,I) \in R \mbox{ and } I = \left(c, UC \right), c \in \T\\
-\emptyset & \mbox{ in any other case }
-\end{cases}
-\end{align}
-\end{definition}
-
-For example, the function Current(R, (3,`John')) returns the time interval $I = \left(4/4/2012 , UC  \right)$. On the other hand, Current(R, (4,`Antoon')) returns the empty set.
-
-The Allen's relations between two time intervals are shown in Figure \ref{fig:allen-rel}.
-The complete implementation of the relations using only the starting and the ending points is defined in~\cite{Nagypal2003}.
-
-\vspace*{13pt}
-\begin{center}
-{
-\includegraphics[scale=0.5]{./graphs/allen.pdf}
-
-}
-\end{center}
-%\centerline{ \psfig{file=./graphs/Y-time-point.eps}}
-\vspace*{10pt}
-\fcaption{\label{fig:allen-rel}Allen's relations for two crisp time intervals $I$ and $J$.
-Note that the crisp time interval $I$ is fixed and the different positions of the interval $J$ illustrate the Allen's relations.}
-\vspace*{13pt}
-
-\begin{definition}
-\emph{Overlaps:}
- \label{def:overlaps}
-Given two time intervals $I = \left(c, d \right)$ and $J = \left(e, f\right)$, it is said that I overlaps J if:
-\begin{equation}
- \label{eqn:overlaps}
-\text{I Overlaps J}  = \left(c > e  \right) \wedge \left(d < f  \right) 
-\end{equation}
-\end{definition}
-
-\begin{definition}
-\emph{During:}
- \label{def:during}
-Given two time intervals $I = \left(c, d \right)$ and $J = \left(e, f\right)$, it is said that I during J if:
-\begin{equation}
- \label{eqn:during}
-\text{I During J}  = \left(c < e  \right) \wedge \left(d > e  \right) \wedge \left(d < f \right)
-\end{equation}
-\end{definition}
-
-In the rest of the paper, and without losing generality, we will consider that the time granularity are days. Also, the dates will be given in the format \emph{dd/mm/yyyy}.
-
-\begin{definition}
-\emph{CloseR$\left( I, J \right)$:}
-\label{def:close-a-crisp-interval-r}
-Consider two crisp intervals $I= \left[c ,d \right]$ and $J= \left[e ,f \right]$.
-The CloseR$\left(I, J\right)$ function allows to close the right-open interval $I$ with respect to the first value $e$ in $J$:
-\begin{align}
-\label{eq:close-a-crisp-interval}
-\mbox{CloseR} \left( I, J \right) &=& \\ 
-\begin{cases}
-\nonumber
-%I & \mbox{ if } b \neq +\infty \\
-%I=\left[a, c \right[ & \mbox{ if } b = +\infty \wedge J > I
-\left[c, e \right[ & \mbox{ if } d = UC \mbox{ and } J \mbox{ During } I \\
-I & \mbox{ in any other case.}
-\end{cases}
-\end{align}
-\end{definition}
-
-For example, consider two time intervals, $I = \left(4/4/2012, UC \right)$ and $J = \left(24/4/2012, UC \right)$.
-The result of applying the CloseR$\left(I, J \right)$ is $I = \left(4/4/2012, 23/4/2012 \right)$.
-
-%Analogously, it is defined the function to close a left-open interval: 
-%\begin{definition}
-%\label{def:close-a-crisp-interval-l}
-%CloseL$\left(I, J\right)$ 
-%\begin{align}
-%\mbox{CloseL} \left( I, J \right) &=& \\ 
-%\begin{cases}
-%\nonumber
-%I & \mbox{ if } a \neq -\infty \\
-%I=\left]d, b \right] & \mbox{ if } a = -\infty \wedge J < I
-%\end{cases}
-%\end{align}
-%\end{definition}
-
-%\begin{definition}
-%\label{def:first-in-relation}
-%Analogously, the function First$\left(R, A \right)$ obtains the tuple $\left(A, I \right)$ where $I$ is a right-open interval and which is the first version in the relation $R$, that is it, $I = \left[-\infty, b \right]$.
-%\begin{align}
-%\label{eq:first-in-relation}
-%\mbox{First} \left(R, A \right) &=& \\ 
-%\begin{cases}
-%\nonumber
-%I & \mbox{ if } \exists b \in I : (A,I) \in R \wedge a = -\infty \\
-%\emptyset & \mbox{ in any other case }
-%\end{cases}
-%\end{align}
-%
-%\end{definition}
-
-Now it is possible to close the current version of an entity by using \eqref{eq:close-a-crisp-interval} and \eqref{eq:current-in-relation}.
-This functionality is required to add or update new information about an existing entity in the relation.
-
-\begin{definition}
-\label{def:close-current-version}
-\emph{Close-current$\left(R, \left(A, J  \right)\right)$}.
-Consider $A=\left(A_1, \ldots, A_n \right)$ be the set of attributes that define an entity and $J = \left(s_j,e_j \right)$ the set of attribute defining the time interval.
-Let $R \subseteq A\  x\  I$ be a valid-time relation.
-Then, the function close-current$\left(R,\left(A,J\right) \right)$ closes any current version of the object $A$ if it exists and add the new version $\left(A, J\right)$.
-
-\begin{eqnarray}
-\label{eq:close-current}
-\text{Close-current} \left(R, \left(A, J\right) \right) =\\
-\begin{cases}
-\nonumber
-R - \left(A, I \right) \cup \left \lbrace \left(A, \mbox{ CloseR } \left(I, J\right) \right)\right \rbrace \cup \left \lbrace\left(A, J\right) \right \rbrace  \\
-\nonumber
-\mbox{ if  Current } \left(R, A \right) \neq \emptyset\\ %\wedge J > \mbox{ CloseR } \left(I, J \right)   \\
-\nonumber R , \text{ in any other case}
-\end{cases}
-\end{eqnarray}
-\end{definition}
-
-For example, consider the patient with ID = 3 in Table \ref{tbl:library-sample}.
-The current version arrived on 4/4/2012.
-The patient left the hospital but, for some reason, the date  was not registered.
-The patient arrived again at the hospital on 24/4/2012.
-Then, the Close-current function allows to close the current version of the patient and insert a new version.
-First, the function CloseR is applied with $I = \left(4/4/2012, UC \right)$ and $J = \left(24/4/2012, UC \right)$.
-Hence, the value for the time interval $I$ is $ \left(4/4/2012, 23/4/2012 \right)$.
-Then, the modifications on the value for the time interval $I$ are stored.
-Finally, a new row with the current values of the patient and the time interval $J$ is stored.
-The result of this operation is illustrated on Table \ref{tbl:library-sample-2}.
-
-%It is also possible to append at the beggining by using \eqref{def:close-a-crisp-interval-l} and \eqref{def:first-in-relation}.
-%\begin{definition}
-%\label{def:append-first-version}
-%Consider an object $A$, the relation $R$ and a time interval $J$.
-Then, the function append-first$\left(R,A,J \right)$ closes any right open version of the object $A$ if it exists and add the new version $\left(A, J \right)$.
-%
-%\begin{eqnarray}
-%%\label{eq:close-current}
-%\text{Append-first} \left(R, A, J \right) =\\
-%\begin{cases}
-%\nonumber
-%\big \lbrace R - \left(A, I \right) \cup \left \lbrace \left(A, \mbox{ CloseL } \left(I, J\right) \right) \cup \left(A, J\right)\right \rbrace  \big \rbrace \\
-%\nonumber
-%\mbox{ if } I = \mbox{ First } \left(R, A \right) \wedge J < \mbox{ CloseL } \left(I, J \right)   \\
-%\nonumber R , \text{ in any other case}
-%\end{cases}
-%\end{eqnarray}
-%\end{definition}
-\subsubsection{\label{subsubsec:modify}Modify}
-This operation adds new information about an existing entity $A$ in the relation $R$.
-The modify operation does not remove any previous value of the object $A$.
-It closes the current version for the entity $A$ and adds a new version.
-
-%Note that the modify operation is only applicable when the object $A$ is current in the relation $R$; $ \left(A, I \right) \in R, I = \left[a, +\infty \right]$.
-
-\begin{align}
-\label{eq:modify}
-\MOD \left( R, \left(A, I \right)\right) =\\
-%\begin{cases}
-\nonumber
-\mbox{ Close-current }\left(R, \left(A, I\right) \right) %\\
-% \mbox{ if Current}\left(R ,A \right) \neq \emptyset \\
-% R,  \mbox{ otherwise }
-%\end{cases}
-\end{align}
-
-\subsubsection{\label{subsubsec:insert}Insert}
-The user wants to store an object $A$ which is valid in the relation $R$ during the time interval specified by $I = \left(a, b \right)$.
-%
-%
-%The interpretation of the insert operation is the following: The user wants to store in the database that an object $A$  is true for some period(s) of time $I$ in the given relation $R$.
-To indicate that the object $A$ is current in the relation, the special value Until Current, $\UC$  is used.
-There are two main cases when performing a create operation:
-\begin{enumerate}
-\item The object $A$ was never in the relation $R$: The object is added with the valid-time indicated by the crisp interval $I$. 
-% In the example, this correspond with the following insertion sentences:
-% \begin{verbatim}
-% Insert(3,'Dracula',
-% (15/3/2012, 30/3/2012));
-% Insert(4,'Frankenstein',
-% (25/3/2012, 4/4/2012));
-% Insert(5,'Harry Potter',
-% (18/3/2012, 2/4/2012));
-% \end{verbatim}
-
-\item The object $A$ is in the relation $R$. Depending on the value of the time interval $I$, there are three possibilities:
-	\begin{enumerate}
-	\item Insert $\left(A, I\right)$ in the relation $R$.
-If the time interval $I$ does not overlap any other valid-time interval in the relation $R$.
-For example, consider that the book with ID 3 is loaned on 4/4/2012 but the book has not been returned yet.
-This is illustrated in Table \ref{tbl:library-sample}.
-%  The insert sentence is:
-% 	      \begin{verbatim}
-% Insert(3,'Dracula',
-% (4/4/2012, UC));
-% 	      \end{verbatim}
-
-	\item Modify and close the current version of $A$ and insert $\left(A, I\right)$.
-For example, consider now that the book with ID 3 is loaned  the 24/4/2012.
-Here the problem is that the book with ID 3 was loaned on 4/4/2012, but not returned.
-If the book is already in the library for a new loan, then it is necessary to set the return date and add a new row with the new loan date.
-This is illustrated in Table \ref{tbl:library-sample-2}.
-%  the insert sentence is:
-% 	    \begin{verbatim}
-% Insert(3,'Dracula',
-% (24/4/2012, UC));
-% 	      \end{verbatim}
-	\item Reject the insertion, if the time interval $I$ does overlap any existing valid-time interval for the object $A$ in the relation $R$.
-For example, consider now that the librarian wants to introduce a past loan for the book with ID 3.
-The book was loaned on 6/4/12/2012 and returned on 25/4/2012.
-As this interval does overlaps other time intervals, it is not possible that the book was loaned in that time interval.Therefore, the insertion is rejected.
-%  The insert sentence is:
-% 
-% 	      \begin{verbatim}
-% Insert(3,'Dracula',
-% (6/4/12/2012,25/4/2012));
-% 	    \end{verbatim}
-
-	\end{enumerate}
-
-\end{enumerate}
-
-\vglue13pt
-%\begin{table}[htbp]
-\tcap{\label{tbl:library-sample-2}Sample library database}
-%\centerline{\small DATA TYPES}
-\vglue-6pt
-\centerline{\small\baselineskip=13pt
-\begin{tabular}{c c c c}\\
-\hline
-\textbf{ID}  & Name  & Arrival & Leave \\
-\hline
-3 & John & 15/3/2012 & 30/3/2012 \\
-4 & Antoon & 25/3/2012 & 4/4/2012 \\
-5 & Daan Potter & 18/3/2012 & 2/4/2012\\
-3 & John & 4/4/2012 & 23/4/2012 \\
-3 & John & 24/4/2012 & UC \\
-\hline\\
-\end{tabular}
-}
-
-The algorithm for the implementation of the insert operation is the following:
-
-\begin{align}
-\label{eq:insert}
-\INS \left(R ,\left(A, I\right)\right) =\\
-\begin{cases}
-\nonumber
-R \cup \left(A, I \right) & \mbox{ if }  (A, I ) \not \in R  \mbox{ or } \\
- &   \left(A, J\right) \in R \mbox{ and } \lnot \left(I \mbox{ Overlaps } J\right) \\
-\mbox{modify(R,(A,I))} & \mbox{ otherwise }    %\mbox{ if } \exists J \left(A,J\right) \in R \wedge \mbox{Current(A,J)} \\ 
-%R, &   
-\end{cases} 	
-\end{align}
-
-\subsubsection{\label{subsubsec:del}Delete}
-The delete operation logically removes a current entity $A$ which is valid in the relation $R$:
-\begin{align}
-\label{eq:delete}
-\DEL \left(R ,A \right) =\\
-\begin{cases}
-\nonumber
-R - \left \lbrace \left(A, I_1 \right),  \ldots, \left(A, I_1 \right)   \right \rbrace, & \forall I_i : \left( A, I_i \right) \in R\\
-R, & \mbox{ otherwise }  
-\end{cases} 	
-\end{align}
-
-For example, consider that the hospital wants to delete the history for the patient with ID = 3. The result of this operation is shown in Table \ref{tbl:library-sample-3}.
-
-% The delete sentence is:
-% 
-% \begin{verbatim}
-%  Delete(3,'Dracula');
-% \end{verbatim}
-
-\vglue13pt
-%\begin{table}[htbp]
-\tcap{\label{tbl:library-sample-3}Sample hospital database}
-%\centerline{\small DATA TYPES}
-\vglue-6pt
-\centerline{\small\baselineskip=13pt
-\begin{tabular}{c c c c}\\
-\hline
-\textbf{ID}  & Name  & Arrival & Leave\\
-\hline
-4 & Antoon & 25/3/2012 & 4/4/2012 \\
-5 & Daan & 18/3/2012 & 2/4/2012\\
-\hline\\
-\end{tabular}
-}
-
-% \subsubsection{\label{subsubsec:revise}Revise}
-% This operation replaces the old values for the entity $A'= \left(a_1', \ldots, a_n' \right) $ with the new values specified by $A = \left(a_1, \ldots, a_n \right)$ in the relation $R$, at a given time interval $I$.
-This operation is used when a correction on the values for the entity $A$ should be done.
-% \begin{align}
-% \REV\left(R, \left(A, I \right) \right)=\\
-% \begin{cases}
-% \nonumber
-%  R - \left(A', I \right) \cup \left \lbrace \left(A, I \right) \right \rbrace & \mbox{ if } \left(A', I \right) \in R  \\
-%    %: A = \left(a_1, \ldots, a_n \right),\\
-% %A'= \left(a_1', \ldots, a_n' \right)   \\   
-% R & \mbox{ otherwise }
-% \end{cases}
-% \end{align} 
-% 
-% For example, consider now that the librarian wants to make a correction in the title for the book with ID = 5.
-The new title is `Harry Potter IV'.
-The result of this operation is illustrated in Table \ref{tbl:library-sample-3}.
-%  Then, the revise sentence is:
-% 
-% \begin{verbatim}
-% Revise(5,'Harry Potter IV', 
-% (18/3/2012, 2/4/2012));
-% \end{verbatim}
+  \label{fig:fuzzy-validity-period}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% FVP
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2013-02-06 trunk/joins/intro.tex
second version

The task of modelling the reality in an information system is usually complex. The modelled reality could be composed by natural objects, processes or concepts. Some of them may present time-related or time-variant properties. Modelling these time-dependent objects  have an impact on the consistency. Depending on the temporal nature of these objects, some extra integrity constraints are imposed.

For example, consider a system which stores the temperature in a room every hour. The temperature per hour is a time-related attribute. Hence, several integrity constraints could be imposed. There can not be stored two different temperatures at the same time.

An specific type of information system is a database management system. More specifically, a temporal database deals with time-dependent objects. Several mechansims have been defined to achieve the extra integrity constraints imposed by the temporal nature of the modelled objects.

Abundant reseach have been done to deal with time in databases  \cite{Jensen1991,Snodgrass1984,Nascimento1995,VanderCruyssen1997a,Jensen1995} which consider the time to be crisp. But time, as well as other attributes is not always perfectly known.

Then it arises the necessity of modelling imperfections in both temporal and non-temporal attributes. There are several types of imperfections such as imprecision, uncertainty or vagueness. To deal with these imperfections in information systems, several theoretical frameworks have been proposed. Among them, probability theory(\cite{Kolmogorov1933,Finetti1937}), possibility theory (\cite{Zadeh1965,LotfiZadeh1978,HenriPrade1982,DidierDubois1988a,Dubois2005}) and rough sets theory (\cite{Pawlak1995}).

In database modelling, several approaches have been proposed to extend the relational model~\cite{Codd1970,Ullman1982} to support imperfections. The main proposals work with possibility theory \cite{Buckles1982,Prade1984,Umano1980,Zemankova-Leech1984,Tre2004,Medina1994}. But none of them deal with time in the sense of time dependent objects.

There are several proposals to represent imperfect time intervals in a relational database like \cite{Garrido2009,Pons2011,Pons2012,JoseEnriquePons2012}. A model that implements the integrity constraints imposed by the nature of time-dependent objects is presented in \cite{Pons2012a}. The model implements the data manipulation language (DML) sentences and keeps the consistency of the time-dependent data.

In this work we will define a relational algebra to deal with time-dependent entities. It is an algeabra for valid-time entities for which the time is not known precisely. The algebra defines the main operators like selection, projection, cartesian product and join.

The remainder of the paper is organized as follows. In section \ref{sec:preliminaries}, some preliminaries about possibility theory and temporal databases is introduced. Next to that, section \ref{sec:proposal} introduces the algebra for crisp temporal databases. Then the algebra is extended to deal with imperfections in time. Finally, section \ref{sec:conc} presents the main conclusions and some lines for future work in this research.

% In order to compare different time intervals we will use  a possibilistic version \cite{Pons2011} of the classical Allen's relations \cite{Allen1983}.

% 
% To allow information systems to cope with these imperfections, there are some approaches that work with probability~\cite{Dekhtyar2001},~\cite{Parisi2010} whereas some other approaches adopt fuzzy sets for the representation of temporal information~\cite{Billiet2011},\cite{Dubois2003}. The temporal relations studied by Allen were recently fuzzified by several authors~\cite{Ohlbach2004},\cite{Schockaert2008}.

% 
% Dealing with time in information systems is usually a difficult task~\cite{Bolour1982}.
% First of all, the concept of \emph{time} itself is very complex. More towards information systems, several proposals have been concerned with obtaining theoretical models that allow the representation of time in information systems~\cite{VanderCruyssen1997a}. Notably, temporal relations between time intervals and to a lesser extent time points, were studied by Allen~\cite{Allen1983}.
% 
% One issue in these attempts at temporal modelling is of course the inherent complexities of the concept of time. Another is the possible inherent imperfections in the description of notions of time. The description of the temporal notion in a sentence like `The Alhambra, which I can see from my room, is a few hundred years old.' contains imperfection caused by the vagueness of the used expression. The description of the temporal notion in a sentence like `The Alhambra was finished somewhere between 1301 A.D. and 1400 A.D.' contains imperfection caused by the imprecision in the used expression about the exact finishing day of the building. Though precise boundaries are given, the exact day on which the building was finished is unknown. An example of uncertainty is given by the following sentence: `I believe that the Alhambra was finished in 1300 A.D.'
% 
% The transition between temporal granularities~\cite{Lin1997a} is also considered as a source of imprecision in temporal modelling and some proposals consider granularity as the base of the temporal model~\cite{VanderCruyssen1997a}.
% 
% To allow information systems to cope with these imperfections, there are some approaches that work with probability~\cite{Dekhtyar2001},~\cite{Parisi2010} whereas some other approaches adopt fuzzy sets for the representation of temporal information~\cite{Billiet2011},\cite{Dubois2003}. The temporal relations studied by Allen were recently fuzzified by several authors~\cite{Ohlbach2004},\cite{Schockaert2008}.
% 
% 
% Next to the issues in temporal modelling, the inclusion of time in a relational database system leads to several more practical problems, as the modelling of the temporal aspects of a time-variant object has an impact on the database consistency. Different models have been proposed \cite{Jensen1991},\cite{Snodgrass1984},\cite{Nascimento1995} that consider time as a crisp notion. Some of these models uses TSQL~\cite{Jensen1995} (a temporal extension to SQL) as query language. In some specific works, authors consider the necessity of allowing imprecision in temporal models~\cite{VanderCruyssen1997a},\cite{Garrido2009}.
% 
% Most of these models use the concepts of fuzzy numbers and fuzzy intervals to represent imprecise temporal notions and considerable attention has thus been given to the problem of transforming two fuzzy numbers which represent the boundaries of an imprecise time interval into one fuzzy interval representing this interval. There are several proposals for this transformation, but some, like those presented in~\cite{Garrido2009} are misleading. 
% 
% In this work we will propose a possibilistic valid-time model for relational databases, based on the framework presented in~\cite{Pons2011}, which deals with the inference of uncertainty about the evaluation of sets from the uncertainty about the exact values used in evaluation. 
% 
% The remainder of the paper is organized as follows. In section \ref{sec:preliminaries}, some preliminaries and concepts are presented. 
% Section \ref{sec:proposal} presents the possibilistic model for valid-time databases. In this section, both representation and querying are introduced.
% Finally, section \ref{sec:futher-research} presents the main conclusions and some lines for future work in this research.

### 2013-02-06 trunk/joins/possibilistic.tex
second version

--- 2012-04-09 ipmu2012/preliminaries.tex
@@
-In this section, some basic concepts are introduced concerning possibilistic variables and fuzzy numbers and intervals, after which the framework of set evaluation by ill-known constraints \cite{Pons2011} is explained.
-The section concludes with a brief introduction to temporal databases.
-
 \subsection{\label{subsec:possibilistic-variables}Possibilistic Variables}
@@
 \end{align}
-
-\subsection{Temporal Databases}
-A \emph{temporal database} is a database that manages some aspects of time in its schema \cite{Dyreson1994}.
-The reality a temporal database tries to model, contains some temporal notions which have to be handled specifically in order to maintain a consistent modelling behavior.
-A \emph{chronon} is the shortest duration of time supported by the database.
-Time can be represented either as points~\cite{DuBois1989} or intervals ~\cite{Garrido2009},\cite{Bohlen1998} that may be subject to imperfection.
-
-The temporal notions in temporal databases can be classified into four types based on their interpretation and modelling purpose.
-\emph{User-defined time} has no specific impact on the database consistency, but the other types do:
-
-\begin{itemize}
-	\item
-	\emph{transaction time}~\cite{Jensen1991}: time when the fact is stored in the database.
-	\item
-	\emph{valid time}~\cite{Snodgrass1984}: time when the fact is true in the modelled reality.
-	\item
-	\emph{decision time}~\cite{Nascimento1995}: time when an event was decided to happen. 
-\end{itemize}
-
-Database models can also be classified into \emph{bi-temporal} (both valid and transaction-time) or \emph{tri-temporal}  (bi-temporal and decision time) models.

### 2013-02-06 trunk/joins/time.tex
second version

@@
 %Guy's suggestion: instead of imprecision, use imperfection which is more general.
-\subsubsection{Imperfection and time}
-Representing imprecision and its semantics when dealing with time has been studied for a long time.
+% \subsubsection{Imperfection and time}
+% Representing imprecision and its semantics when dealing with time has been studied for a long time.
 Several proposals for representing and computing imprecise time indications can be found in \cite{DeCaluwe1997} and \cite{DeTre1997a}.
 Also, the changes between several granularities can be seen as a source of imprecision \cite{Devos1998}.
-
-In the proposal section we will consider two kinds of imprecision:
-\begin{itemize}
-\item \textbf{Uncertainty in the database} denotes the uncertainty that arises when the knowledge about the temporal data in the database is uncertain.
+% 
+% In the proposal section we will consider two kinds of imprecision:
+% \begin{itemize}
+% \item \textbf{Uncertainty in the database} denotes the uncertainty that arises when the knowledge about the temporal data in the database is uncertain.
 E.g., a database record shows that \emph{`The car is in the garage around April.'}
- \item \textbf{Imprecision in the query specification} denotes the imprecision in the specification of temporal criteria by the user, when querying.
+%  \item \textbf{Imprecision in the query specification} denotes the imprecision in the specification of temporal criteria by the user, when querying.
 E.g., \emph{`The user wants to obtain a car which is red and which is in the garage around April.'}
-\end{itemize}
+% \end{itemize}
 

### 2013-02-15 fqas2013/paper.tex
The start of the paper.

%%FQAS 2013 Contribution

\documentclass[runningheads,a4paper]{llncs}

\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx}

\begin{document}

\mainmatter  %start of the contribution

\title{Bipolar Valid-time Querying}

%Short form of title for running heads
\titlerunning{Bipolar Valid-time Querying}

\author{Christophe Billiet\inst{1} \and Jos\'{e} Enrique Pons\inst{2} \and Olga Pons\inst{2} \and Guy De Tr\'{e}\inst{1}}

\authorrunning{Christophe Billiet et al.}

\institute{
Department of Telecommunications and Information Processing, Ghent University,\\
Sint-Pietersnieuwstraat 41, B-9000, Ghent, Belgium\\
Christophe.Billiet@UGent.be, Guy.DeTr\'{e}@UGent.be\\
\and
Department of Computer Science and Artificial Intelligence, University of Granada,\\
C/Periodista Daniel Saucedo Aranda, S/N, E-18071, Granada, Spain\\
jpons@decsai.ugr.es, opc@decsai.ugr.es
}

\maketitle

\begin{abstract}
Here goes the abstract.
\keywords{keyword1, keyword2}
\end{abstract}

\section{Introduction}

\bibliographystyle{splncs}
\bibliography{paper}

\end{document}

### 2013-02-15 joins/time.tex
update!

--- 2013-02-06 trunk/joins/time.tex
@@
 \subsubsection{Representation}
-Several proposals for managing uncertain time in a database exist.
+Several proposals for representing uncertain time in a database exist.
 Some proposals work with rough sets \cite{Qiang2009}, other proposals rely on possibility distributions for representing uncertainty in time \cite{Garrido2009}, \cite{Galindo2001}.
@@
 If the starting points and/or the end points of the interval representing the time are not known precisely, it is easy to fuzzify them, using, e.g., two triangular possibility distributions.
+\begin{definition}
 A \textbf{Fuzzy Validity Period} (\emph{FVP}) is defined as a fuzzy time interval specifying when an object is valid.
@@
 \end{itemize}
+\end{definition}
+
+\begin{definition}
+A \emph{Possibilistic Valid-time Period} is an ill-known interval in time, which models a time period during which an object in a certain state is valid.
+\end{definition}
+
+Because a PVP is an ill-known interval, it allows modelling the uncertainty about the start and/or end point of a time interval (and thus about the time interval itself) if such uncertainty exists.
+The interpretation is \emph{disjunctive}: the PVP represents exactly one valid-time interval, but precisely \emph{which} interval is represented, is (partially) unknown.
+In the presented model, only PVPs are considered of which the possibility distributions of the possibilistic variables defining the start and end point of the ill-known interval have exactly the same characteristics as the membership functions of fuzzy numbers.
+A perfectly known start or end point can then be modelled by such an ill-known value defined by a possibilistic variable $P$ for which $\exists !
+x : \mu_{P}(x) > 0$.
+
+As mentioned in \cite{Pons2011}, this approach differs from the one where a valid-time period is represented by one fuzzy set.
+Such a fuzzy set is seen as a possibility distribution on $\mathbb{R}$ and thus defines just one ill-known value.
+However, in the presented approach, a time period is modelled using an ill-known set, which is defined by a possibility distribution on $\Pow(\mathbb{R})$.
 

### 2013-02-15 temporaldml/main.tex
update!

@@
        \label{eq:valid-time-instance}
-      r \subseteq D_1\ x\ \ldots\ x\ D_n x\ T x\ T
+      r \subseteq D_1\times \ \ldots\ \times \ D_n \times \ T \times \ T
       \end{equation}
@@
 \section{\label{sec:selection}Selection}
-The selection operator $\sigma$ obtain a subset of tuples that fulfill a set of constraints from a given relation $R$.
-The set of constraints is usually a boolean combination of atomic constraints.
-The selection operator is noted as follows:
+The selection operator $\sigma$ obtain a subset of tuples that fulfill a set of constraints $P$ from an instance $r$ of the relation $R \subseteq
+ A \times I$. The set of constraints is usually a boolean combination of atomic constraints. The selection operator is noted as follows:
 \begin{equation}
  \label{eq:selection}
-\sigma_{P} \left( R \right)
+\sigma_{P} \left( r \right)
 \end{equation}
 
-Where $R$ is the relation, and $P$ is the selection formula. The selection formula is a set of two elements:
+Where $r \in R$ is the relation, and $P$ is the selection formula. The selection formula is a set of two elements:
 % \begin{itemize}
@@
  \label{eq:selection_formula}
-P = \left \lbrace Q, Q^{t}\right \rbrace
+P = \left \lbrace Q, Q^{T}\right \rbrace
 \end{equation}
 
-Where $q$ is a boolean combination of (possibly) fuzzy (and bipolar) non-temporal constraints and $q^{t}$ is also a boolean combination of fuzzy temporal constraints.
+Where $Q$ is the set of (possibly) fuzzy (and bipolar) non-temporal constraints and $Q^{T}$ is also a boolean combination of fuzzy temporal constraints.
 
@@
  \label{eq:non-temporal-constraints}
-Q = \left \lbrace q^{\left \lbrace \mbox{pos,neg} \right \rbrace}_{\mbox{attribute}_1}  \theta \mbox{ val }, \ldots, q^{\left \lbrace \mbox{pos,neg} \right \rbrace}_{\mbox{attribute}_n}  \theta \mbox{ val } \right \rbrace
+Q = \left \lbrace q^{\left \lbrace \mbox{pos,neg} \right \rbrace}_{\mbox{a}_1}  \theta \mbox{ val }, \ldots, q^{\left \lbrace \mbox{pos,neg} \right \rbrace}_{\mbox{a}_n}  \theta \mbox{ val } \right \rbrace
 \end{equation}
-Where $\theta$ is an artihmetic operator; one of $\left \lbrace =, \neq, <, \leq, >, \geq \right \rbrace$ and val is a value in the domain of the attribute.
+
+Where:
+\begin{itemize}
+ \item $q^{\left \lbrace \mbox{pos,neg} \right \rbrace}_{a \in A }$ is an atomic constraint.
+A possitive preference is notated by $q^{\left \lbrace \mbox{pos} \right \rbrace}$ whether a negative preference is notated by $q^{\left \lbrace \mbox{neg} \right \rbrace}$.
+The constraint refers to an attribute $a$ that belongs to the set of attributes $A \subseteq A_1 \ldots A_n $ of the relation $R$.
+\item $\theta$ is an artihmetic operator; one of $\left \lbrace =, \neq, <, \leq, >, \geq \right \rbrace$.
+\item \emph{val} is a value in the domain of the queried attribute. 
+ \end{itemize}
 

### 2013-02-15 develop/fuzzyCalculator/src/main/java/es/ugr/decsai/idbis/fuzzycalculator/TrapezoidalFuzzyNumber.java

@@ -18,7 +18,7 @@ public class TrapezoidalFuzzyNumber extends FuzzyNumber{
     protected Double beta;
     protected Double gamma;
     protected Double delta;
-    
+
 
     public TrapezoidalFuzzyNumber(Double alpha, Double beta, Double gamma,Double delta) {
         this.alpha = alpha;
@@ -100,6 +100,27 @@ public class TrapezoidalFuzzyNumber extends FuzzyNumber{
         
         
     }
+    
+     protected static Double getMin(TrapezoidalFuzzyNumber f1,TrapezoidalFuzzyNumber f2){
+         Double min;
+         if(f1.getAlpha() < f2.getAlpha()){
+            min = f1.getAlpha();
+        }else{
+            min = f2.getAlpha();
+        }
+         return min;
+     }
+     
+     
+     protected static Double getMax(TrapezoidalFuzzyNumber f1,TrapezoidalFuzzyNumber f2){
+         Double max;
+         if(f1.getDelta() > f2.getDelta()){
+            max = f1.getDelta();
+        }else{
+            max = f2.getDelta();
+        }
+         return max;
+     }
     /**
      * Fuzzy equality implemented as
      * sup min eval(f1),eval(f2))
@@ -250,6 +271,26 @@ public class TrapezoidalFuzzyNumber extends FuzzyNumber{
     public static Double Neq(TrapezoidalFuzzyNumber f1,TrapezoidalFuzzyNumber f2) throws FuzzyConstraintException{
         return Neq(f1, f2, 0.01D);
     }
+    
+    
+    public static Double Fgt(TrapezoidalFuzzyNumber f1,TrapezoidalFuzzyNumber f2,Double precission) throws FuzzyConstraintException{
+        Double result = 0D;
+        
+        Double min = getMin(f1, f2);
+        Double max = getMax(f1, f2);
+        Double val1, val2, maxval, inf;
+        inf = Double.POSITIVE_INFINITY;
+        for(double i = min; i< max;i+= precission){
+            val1 = f1.evaluate(i);
+            val2 = f2.evaluate(i);
+            
+            
+        }
+        
+        return result;
+    }
+    
+    
     /**
      * Discrete implementation for the Fuzzy Greater Than operator.
      * @param f1 the first trapezoidal number.

### 2013-02-15 develop/testApplication/src/main/java/es/ugr/decsai/jpons/testapplication/Main.java

@@ -345,6 +345,18 @@ TemporalPersistenceManager manager = TemporalPersistenceManager.getInstance(prop
             
             
             
+            //test for the cartesian product:
+            tq1.createTemporalQuery(e1);
+            System.out.println("---- CARTESIAN PRODUCT ----");
+            List<QueryResult> list2 = tq1.cartesianProduct();
+            
+            for(QueryResult qr:list2){
+                System.out.println("aggregation: "+qr.getAggregation()
+                        + " temporal "+qr.getTemporalSatisfaction()+
+                        " endmp: " + qr.getVtp().getEndMP() + " vid: " + qr.getKeys());
+            }
+            
+            
 //            String tableName = e3.getClass().getSimpleName();
 //            TemporalPersistenceManager.getSession().createQuery("Delete from "+ tableName + " where tid.id = :idval" ).setParameter("idval", e3.getTid().getId()).executeUpdate();
 ////            //query:

### 2013-02-15 develop/testQuerying/src/main/java/es/jpons/persistence/query/CriteriaAllenRelations.java

@@ -388,6 +388,73 @@ public class CriteriaAllenRelations {
         res = computeIntersections(i.getEndLeft(), i.getEndRight(), i.getEndMP(), j.getEndLeft(), j.getEndRight(), j.getEndMP());
         return res;
     }
+    
+    /**
+     * Intersection of two possibilistic valid time points.
+     * 
+     * @param i
+     * @param j
+     * @return 
+     */
+    public static double computeIntersectsSatisfactionDegree(PossibilisticVTP i, PossibilisticVTP j){
+        Double res,a,b;
+       a = computeIntersectionLt(i.getStartLeft(),i.getStartRight(),i.getStartMP(),j.getStartLeft(),j.getStartMP());
+       b = computeIntersectionLt(j.getStartLeft(),j.getStartRight(),j.getStartMP(),i.getStartLeft(),i.getStartMP());
+       
+       res = (a<b)?a:b;
+       
+       return res;
+    }
+    
+    /**
+     * Computes de overlapping interval for two possibilistic valid-time points
+     * @param i
+     * @param j
+     * @return 
+     */
+   public static PossibilisticVTP overlappingInterval(PossibilisticVTP i, PossibilisticVTP j){
+       PossibilisticVTP result = new PossibilisticVTP();
+       // 1st step: compute the lower bound:
+       Double beforeij = computeIntersectionLt(i.getStartLeft(),i.getStartRight(),i.getStartMP(),j.getStartLeft(),j.getStartMP());
+       Double beforeji = computeIntersectionLt(j.getStartLeft(), j.getStartRight(), j.getStartMP(), i.getStartLeft(), i.getStartMP());
+       
+       if(beforeij > beforeji ){
+           // i before j holds, then j should be taken as the lower bound:
+           result.setStartDate(j.getStartDate());
+           result.setStartLeft(j.getStartLeft());
+           result.setStartRight(j.getStartRight());
+       }else{
+           // j before i holds then i should be taken as the lower bound:
+           result.setStartDate(i.getStartDate());
+           result.setStartLeft(i.getStartLeft());
+           result.setStartRight(i.getStartRight());
+       }
+       
+       //2nd step: compute the higher bound:
+       beforeij = computeIntersectionLt(i.getEndLeft(), i.getEndRight(), i.getEndMP(), j.getEndLeft(), j.getEndMP());
+       beforeji = computeIntersectionLt(j.getEndLeft(), j.getEndRight(), j.getEndMP(), i.getEndLeft(),i.getEndMP());
+       
[…]

### 2013-02-15 trunk/book/sources/fuzzyTimeDomain.tex
update!

@@
 	\item \emph{Uncertainty}: Temporal information or data may contain uncertainty.
-This means that the exact temporal value is (partially) unknown, however, generally some knowledge is present anyhow, possibly describing the value~\cite{Dubois:jucs_9_9:fuzziness_and_uncertainty_in},~\cite{nagypal2003},~\cite{Dubois89},~\cite{343607}.
+This means that the exact temporal value is (partially) unknown, however, generally some knowledge is present anyhow, possibly describing the value~\cite{Dubois2003},~\cite{nagypal2003},~\cite{Dubois1989},~\cite{343607}.
 E.g., the temporal notion described in a sentence like `The Belfry of Bruges was finished on a day somewhere between 01/01/1201 A.D.
@@
 Indefinite temporal information contains information describing a situation in which the time indication associated with some fact has not been fully defined.
-E.g., consider an event that in fact occured but it is not known exactly when.
+E.g., consider an event that in fact occurred but it is not known exactly when.
 
@@
 E.g., consider meetings that take place every Wednesday at noon.
-Some systems (usually with different information providers) may dispute the occurence and/or the duration of a fact.
+Some systems (usually with different information providers) may dispute the occurrence and/or the duration of a fact.
 \end{itemize}
@@
 \begin{example}
-Consider a speaker and a hearer. The speaker wants to make an appointment with the hearer. Now, consider the speaker saysing: \\
+Consider a speaker and a hearer. The speaker wants to make an appointment with the hearer. Now, consider the speaker saying: \\
 \begin{center}
@@
 The use of probability theory is very usual in logistics information systems.
-E.g., \emph{`The package will arrive at its destination on monday morning with a probability of 0.8'}.
+E.g., \emph{`The package will arrive at its destination on Monday morning with a probability of 0.8'}.
 
@@
 Using possibility theory~\cite{DidierDubois1988a}, a possibility degree is associated to the temporal fact or event.
-Possibility theory is widely used to model uncertainty and vaguenes in time~\cite{Dubois:jucs_9_9:fuzziness_and_uncertainty_in},\cite{Dubois89},\cite{devos94},\cite{nagypal2003}.
+Possibility theory is widely used to model uncertainty and vagueness in time~\cite{Dubois:jucs_9_9:fuzziness_and_uncertainty_in},\cite{Dubois89},\cite{devos94},\cite{nagypal2003}.
 Several works~\cite{schockaert08},~\cite{ohlbach2004} present fuzzy versions of the temporal relations proposed by Allen~\cite{Allen83}.

### 2013-02-15 trunk/book/sources/intro.tex
update!

@@
 
-Humans handle temporal information using certain temporal notions like time intervals or time points~\cite{Dyreson1994}, and they often have to deal with imperfections like imprecisions, vaguenesses, uncertainties or inconsistencies possibly contained in the descriptions of these temporal notions.
+Humans handle temporal information using certain temporal notions like time intervals or time points~\cite{Dyreson1994}, and they often have to deal with imperfections like imprecision's, vagueness, uncertainties or inconsistencies possibly contained in the descriptions of these temporal notions.
 Among many others, these possible imperfections in descriptions of temporal notions determine an important issue in temporal modelling.

### 2013-02-15 trunk/develop/fuzzyCalculator/src/main/java/es/ugr/decsai/idbis/fuzzycalculator/App.java

@@ -1,6 +1,7 @@
 package es.ugr.decsai.idbis.fuzzycalculator;
 
 import es.ugr.decsai.idbis.fuzzy.FuzzyConstraintException;
+import es.ugr.decsai.idbis.fuzzy.relationship.FuzzyRelationship;
 import java.util.ArrayList;
 import java.util.logging.Level;
 import java.util.logging.Logger;
@@ -22,6 +23,30 @@ public class App
         TrapezoidalFuzzyNumber t3 = new TrapezoidalFuzzyNumber(20D, 30D, 30D, 35D);
         
         
+        TriangularFN start = new TriangularFN(5D, 2D, 2D);
+        TriangularFN end = new TriangularFN(10D, 2D, 2D);
+        
+        IllKnownConstraint ikc = new IllKnownConstraint(start, FuzzyRelationship.FGT);
+        IllKnownConstraint ik2 = new IllKnownConstraint(end, FuzzyRelationship.FLEQ);
+        
+        TriangularFN a = new TriangularFN(6D, 0D, 0D);
+        TriangularFN b = new TriangularFN(7D, 0D, 0D);
+        try {
+            Double vala = ikc.evaluate(a);
+            Double valb = ikc.evaluate(b);
+            
+            System.out.println("Minimo:" + ((vala<valb)?vala:valb));
+            
+        } catch (FuzzyConstraintException ex) {
+            Logger.getLogger(App.class.getName()).log(Level.SEVERE, null, ex);
+        }
+        
+        
+        
+        
+        
+        
+        
 //        for(double i=0;i< 30;i++){
 //            try {
 //                System.out.println("i: "+i + " eval "+t.evaluate(i)); 
@@ -31,22 +56,11 @@ public class App
 //        }
         try {
             
-            ArrayList<Double> feq,nfeq,fgt,nfgt,fgeq,nfgeq,flt,nflt,fleq,nfleq,mgt,nmgt,mlt,nmlt;
+            ArrayList<Double> ikcList;
             ArrayList<TrapezoidalFuzzyNumber> numbers;
             
-            feq = new ArrayList<Double>();
-            nfeq = new ArrayList<Double>();
-            fgt = new ArrayList<Double>();
-            nfgt = new ArrayList<Double>();
-            fgeq = new ArrayList<Double>();
-            nfgeq = new ArrayList<Double>();
-            flt = new ArrayList<Double>();
-            nflt = new ArrayList<Double>();
[…]

### 2013-02-15 trunk/develop/testQuerying/src/main/java/es/jpons/persistence/query/QueryResult.java

@@ -16,41 +16,120 @@
  * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston,
  * MA 02110-1301  USA
  */
-
 package es.jpons.persistence.query;
 
 import es.jpons.persistence.TimeUtils;
 import es.jpons.persistence.exception.TemporalException;
 import es.jpons.temporal.types.PossibilisticVTP;
 import es.jpons.temporal.types.TemporalPK;
+import java.util.ArrayList;
+import java.util.Arrays;
+import java.util.List;
 
 /**
- * Represents the result of a query.
- * It is stored both crisp and temporal satisfaction.
- * @author Jose Enrique Pons Frías <jpons@decsai.ugr.es>
- * First version 15/10/2012
- * 
+ * Represents the result of a query. It is stored both crisp and temporal
+ * satisfaction.
+ *
+ * @author Jose Enrique Pons Frías <jpons@decsai.ugr.es> First version
+ * 15/10/2012 First version.
+ * 30/01/2013 Modification for the cartesian product.
+ *
  */
 public class QueryResult {
-    protected Object entity;
-    protected TemporalPK key;
+
+    protected List entity;
+    protected List<TemporalPK> key;
     protected PossibilisticVTP vtp;
     protected Double crispSatisfaction;
     protected Double temporalSatisfaction;
     protected Double aggregation;
-    
 
+    /**
+     * Entity constructor
+     *
+     * @param entity
+     * @throws TemporalException
+     */
     public QueryResult(Object entity) throws TemporalException {
-        this.entity = entity;
-        if(entity!=null){
-            try{
-            this.key = TimeUtils.getPK(entity);
-            this.vtp = TimeUtils.getValidTime(entity);
-            }catch(Exception e){
+        this.entity = new ArrayList();
+        this.key = new ArrayList<TemporalPK>();
+        this.crispSatisfaction = 0D;
+        this.temporalSatisfaction = 0D;
+        this.aggregation = 0D;
+        
+        if (entity != null && isList(entity)) {
+            List entities = (List) entity;
+            addEntities(entities);
+        } else {
+            addEntity(entity);
+        }
+
+
+
+
+//        if (entity != null) {
+//            try {
+//                this.key.add(TimeUtils.getPK(entity));
+//                this.vtp = TimeUtils.getValidTime(entity);
+//            } catch (Exception e) {
+//                throw new TemporalException(e);
+//            }
+
+
+
+
+//        }
+    }
+
+    
[…]

### 2013-02-15 trunk/develop/testQuerying/src/main/java/es/jpons/persistence/query/TemporalQuery.java

@@ -45,10 +45,13 @@ public class TemporalQuery {
     protected AllenRelation  ar;
     protected PossibilisticVTP pvp;
     
+    protected List<TemporalQuery> temporalQueryList;
+    
     public TemporalQuery(Session session, Object entity){
         this.session = session;
         crispCriteria = session.createCriteria(entity.getClass());
         temporalCriteria = session.createCriteria(entity.getClass());
+        this.temporalQueryList = new ArrayList<TemporalQuery>();
         
     }
 
@@ -126,8 +129,9 @@ public class TemporalQuery {
         
         
         for(QueryResult q:merge){
+            q.temporalSatisfaction = 0.0D;
             
-            
+            if(ar!=null){
             switch(ar){
                 case before: q.temporalSatisfaction = CriteriaAllenRelations.computeBeforeSatisfactionDegree(q.vtp, pvp);
                     break;
@@ -147,6 +151,7 @@ public class TemporalQuery {
                 case contains: q.temporalSatisfaction = CriteriaAllenRelations.computeContainsSatisfactionDegree(q.vtp, pvp);
                     break;
                 
+            }
             }
            
             q.aggregation = weight*(q.crispSatisfaction) + (1-weight)*q.temporalSatisfaction;
@@ -215,7 +220,68 @@ public class TemporalQuery {
         
         
         return this;
+        
+        
     }
+    
+    
+    
+    /**
+     * Function to create a new temporal query.
+     * This is used to make cartesian products and temporal joins.
+     * @param entity An entity
+     * @return An instance of the temporal query object.
+     */
+    public TemporalQuery createTemporalQuery(Object entity){
+        TemporalQuery tq = new TemporalQuery(session, entity);
+        temporalQueryList.add(tq);
+        return tq;
+    }
+    
+    
+    public List<QueryResult> cartesianProduct() throws TemporalException{
+        List<QueryResult> queryResult = this.getList();
+        if(this.temporalQueryList.size()>0){
+//            List<QueryResult> queryResult = this.getList();
+            
+            for(TemporalQuery tq:temporalQueryList){
+                if(!tq.equals(this)){
+                    List<QueryResult> queryResult2 = tq.getList();
+                    queryResult = mergeCartesianProduct(queryResult, queryResult2);
+                    
+                }
+            }
+            
+            
+        }
+        return queryResult;
+        
+    }
+    
[…]

### 2013-02-22 trunk/develop/javagnuplot/src/main/java/es/ugr/decsai/jpons/javagnuplot/App.java
interface for gnuplot in java, maven project

@@ -1,0 +1,13 @@
+package es.ugr.decsai.jpons.javagnuplot;
+
+/**
+ * Hello world!
+ *
+ */
+public class App 
+{
+    public static void main( String[] args )
+    {
+        System.out.println( "Hello World!" );
+    }
+}

### 2013-02-22 trunk/develop/javagnuplot/src/test/java/es/ugr/decsai/jpons/javagnuplot/AppTest.java

@@ -1,0 +1,38 @@
+package es.ugr.decsai.jpons.javagnuplot;
+
+import junit.framework.Test;
+import junit.framework.TestCase;
+import junit.framework.TestSuite;
+
+/**
+ * Unit test for simple App.
+ */
+public class AppTest 
+    extends TestCase
+{
+    /**
+     * Create the test case
+     *
+     * @param testName name of the test case
+     */
+    public AppTest( String testName )
+    {
+        super( testName );
+    }
+
+    /**
+     * @return the suite of tests being tested
+     */
+    public static Test suite()
+    {
+        return new TestSuite( AppTest.class );
+    }
+
+    /**
+     * Rigourous Test :-)
+     */
+    public void testApp()
+    {
+        assertTrue( true );
+    }
+}

### 2013-02-25 trunk/eusflat2013/preliminaries.tex
first version paper eusflat

This section introduces the framework of the ill-known contraints. First of all, some preliminary concepts about possibility theory is explained. Next to that, the representation of time intervals by using the presented framework is done. Finally, the evaluation of the Allen's relationships between a crisp time interval and a possibilistic time interval is provided.

\subsection{\label{subsec:possibilistic-variables}Possibilistic Variables}
Informally speaking, a possibilistic variable is a variable that takes only one value but that, for some reason, this value is unknown. This is rather usual for historical events. For example, the conquest of Granada by the Christians' kings took place on February of 1492, but the exact day is not precissely known.
A more formal definition of possibilistic variables is given as follows.
%Possibilistic variables rely on possibility theory \cite{Dubois1988a}. A \emph{possibilistic variable} is defined as follows \cite{Pons2011}.

\begin{definition}
\label{def:ill-known-value}
A possibilistic variable \cite{Dubois1988a}, \cite{Pons2011} $X$ over a universe $U$ is defined as a variable taking exactly one value in $U$. A possibility distribution $\pi_X$ on $U$ is defined and models the available knowledge about the value that $X$ takes: for each $u\in U$, $\pi_X(u)$ represents the possibility that $X$ takes the value $u$. In this work, this possibility is interpreted as a measure of how plausible it is that $X$ takes the value $u$, given (partial) knowledge about the value $X$ takes.
\end{definition}

The exact value a possibilistic variable takes, which is (partially) unknown, is called an \emph{ill-known value} in this work \cite{Dubois1988a}.

An \emph{ill-known-set} \cite{Dubois1988a} is a possibilistic variable defined on the powerset $\Pow(U)$ of the universe $U$.

%When a possibilistic variable is defined on the powerset $\Pow(R)$ of some universe $R$, the unique value the variable takes will be a crisp set and its possibility distribution on the powerset $\Pow(R)$ will describe the possibility of each crisp subset of $R$ to be the value the variable takes. This exact value (a crisp set) the variable takes, is now called an \emph{ill-known set} \cite{Dubois1988a}.

% A specific application of possibilistic variables is obtained when the universe under consideration is the set of Boolean values $\mathbb{B}$ = $\{T,F\}$. Indeed, any Boolean proposition $p$ takes just one value in $\mathbb{B}$. If the knowledge about which value this proposition $p$ will take is given by a possibility distribution $\pi_p$, the proposition can be seen as a possibilistic variable. As the interest lies with the case where the proposition holds, the possibility and necessity that $p$ = $T$ (the proposition holds) demand most attention. This possibility and necessity is noted here as:
% \begin{align}
% \text{Possibility that $p$ = $T$ (p holds):}  & Pos(p) = \pi_p(T) \label{propholdsposs} \\
% \text{Necessity that $p$ = $T$ (p holds):}  & Nec(p) = 1-\pi_p(F) \label{propholdsnecc}
% \end{align}

% This work will deal with ill-known intervals. These are ill-known sets, defined and represented via a start and end point, which will be ill-known values. The elements of the set are the values between the start and end point. A closed ill-known interval with start point defined by possibilistic variable $X$ and end point by possibilistic variable $Y$ is noted here $\left[X, Y\right]$. The correspondences and transitions between the representations of ill-known sets, between the representations of ill-known intervals and between the representations of an ill-known set and an ill-known interval are part of the authors current research.

\subsection{\label{subsec:fuzzy-numbers}Fuzzy Numbers and Fuzzy Intervals}
A fuzzy set is a set for which its elements have a degree of membership. Among others, Dubois and Prade~\cite{Dubois1983} use fuzzy sets \cite{Zadeh1965} to define a \emph{fuzzy interval}:
\begin{definition}
A fuzzy interval is a fuzzy set $M$, defined by a membership function $\mu_{M}$, on the set of real numbers $\mathbb{R}$ such that:
\begin{eqnarray}
\mu_{M} : & \mathbb{R} \rightarrow \left[0,1\right] \nonumber \\ 
\label{eq:convexity}
\forall (u,v)\in\mathbb{R}^2: \forall w \in [u,v]:& \\
\nonumber
\mu_M(w) \geq\min(\mu_M(u),\mu_M(v))  \\
\label{eq:unicity}
\exists m \in \mathbb{R} : &  \mu_M(m)=1 
\end{eqnarray}
\end{definition}
If this modal value $m$ is unique, then $M$ is referred to as a \emph{fuzzy number}. In other words, if the core of a fuzzy interval is a singleton, it is referred to as a fuzzy number.
Equation \eqref{eq:convexity} and \eqref{eq:unicity} are known to be the convexity and unicity properties respectively.

The most convenient form of the membership function of a fuzzy number is a triangular form. It can be shown that such a membership function $\mu_M$ for a fuzzy number $M$ is convex and normalized. Three real values, denoted by $a$, $b$ and $D$, suffice to represent a triangular membership function of a fuzzy number and in this work, a fuzzy number defined as such will be noted as $\left[D, a, b \right]$. Here $D$ is the core of $M$. $D-a$ and $D+b$ are respectively the lower and higher bounds of the possibility distribution.
% \begin{itemize}
% \item
% $D$ denotes the single value in the core of $M$
% \item
% $D-a$ is then $\inf \{u \in \mathbb{R} : \mu_{M}(u) > 0\}$
% \item
% $D+b$ is then $\sup \{u \in \mathbb{R} : \mu_{M}(u) > 0\}$
% \end{itemize}

%%
%% Estructure: 
%% 2.3 The IKC framework.
%%.2.4 Representation of time intervals: Table with the IKC.

\subsection{\label{subsec:ikc-framework}The ill-known constraint framework}
A constraint is a restriction in the values that a given variable can take. An ill-known constraint is a constraint applied on a ill-known value~\cite{Pons2011}. More formally, an ill-known constraint is defined as follows.
\begin{definition}
Given a universe $U$, an ill-known constraint $C$ on a set $A \subseteq U$ is specified by means of a binary relation $R \subseteq U^{2}$ and a fixed, ill-known value denoted by its possibilistic variable $V$ over $U$, i.e.:
\begin{align}
C \triangleq (R,V)
\end{align}
Set $A$ now satisfies the constraint if and only if:
\begin{align}
\forall a \in A : (a,V) \in R
\end{align}
\end{definition}
A possibility distribution $\pi_{V}$ is defined on the variable $V$. Then, the  possibility and necessity that $A$ satisfies $C$ is computed by:

\begin{align}
\Pos(A\text{ sat. }C) & = \min_{a \in A}\left(\sup_{(a,w) \in R}\pi_{V}(w)\right) \label{ill-known-pos}\\
\Nec(A\text{ sat. }C) & = \min_{a \in A}\left(\inf_{(a,w) \notin R} 1-\pi_{V}(w)\right) \label{ill-known-nec}
\end{align}

For example, consider an ill-known value $X$ given by a triangular possibility distribution $\left[15, 1, 1\right]$. The ill-known contraint for all the points greater than $X$ is given by $C \triangleq (>, X)$. See Figure \ref{fig:ikc-greater}.

\begin{figure}[h]
   \centering
   \includegraphics[scale=1.5]{graphs/gt.eps}
   \caption{An ill-known value $X = \left[15, 1, 1\right]$ and an ill-known contraint $C \triangleq (>, X)$. }
   \label{fig:ikc-greater}
 \end{figure}

Several ill-known constraints can be aggregated by using boolean operators: and ($\wedge$), or ($\vee$), not ($\neg$). The possibility and neccesity for these operators are computed as follows.

\begin{align}
\Pos(C_1 \wedge \ldots \wedge C_n) & = \min_{i= 1}^{n} \left(\Pos(C_i)\right) \label{and-ill-known-pos}\\
\Nec(C_1 \wedge \ldots \wedge C_n) & = \min_{i= 1}^{n} \left( \Nec(C_i )\right) \label{and-ill-known-nec}\\
%\end{align}
%\begin{align}
\Pos(C_1 \vee \ldots \vee C_n) & = \max_{i= 1}^{n} \left(\Pos(C_i)\right) \label{or-ill-known-pos}\\
\Nec(C_1 \vee \ldots \vee C_n) & = \max_{i= 1}^{n} \left( \Nec(C_i )\right) \label{or-ill-known-nec}\\
%\end{align}
%\begin{align}
\Pos( \neg C) & =  \left(1-\Nec(C)\right) \label{not-ill-known-pos}\\
\Nec(\neg C) & = \left(1- \Pos(C)\right) \label{not-ill-known-nec}
\end{align}

 \begin{figure}[h]
   \centering
   \includegraphics[scale=1.5]{graphs/ill-known-ti.eps}
   \caption{An ill-known time interval $J = \left[X, Y  \right]$ given by two ill-known values. }
   \label{fig:ill-known-ti}
 \end{figure}

An ill-known time interval is defined by two ill-known values (see Figure \ref{fig:ill-known-ti}). Therefore, both starting and ending points are not precissely known. The following is the notation for an ill-known time interval.

\begin{align}
\label{eq:ill-known-time-interval}
J = \left[X, Y \right] 
\end{align}

Where both $X$ and $Y$ are ill-known time values.

The Allen's Relations between a crisp time interval $I = [a,b]$ ($a, b$ are crisp time points) and an ill-known time interval given by $J = [X,Y]$ ($X, Y$ are ill-known values, see Definition \ref{def:ill-known-value}) can be defined by using aggregations of ill-known constraints. Table \ref{tab:allen-relations} shows the corresponding ill-known constraints for each Allen's Relation. These ill-known constraints are aggregated by using the equations \eqref{and-ill-known-pos} to \eqref{not-ill-known-nec}.

\begin{table}[h]
\centering
\begin{tabular}{|c|c|c|}
\hline
Allen Relation & Constraints & Aggregation \\
\hline
I before J & $C_1\stackrel{\triangle}{=} \left(<,X\right)$ & $C_1(I)$ \\
\hline
\multirow{4}{*}
{I equal J} & $C_1\stackrel{\triangle}{=} \left(\geq,X\right)$ & $C_1(I)\wedge$\\
 & $C_2\stackrel{\triangle}{=} \left(\neq,X\right)$ & $\neg C_2(I)\wedge$\\
 & $C_3\stackrel{\triangle}{=} \left(\leq,Y\right)$ &  $C_3(I)\wedge$ \\
 & $C_4\stackrel{\triangle}{=} \left(\neq,Y\right)$ & $\neg C_4(I)$\\
\hline
\multirow{2}{*}
{I meets J} & $C_1\stackrel{\triangle}{=} \left(\leq,X\right)$ & $C_1(I)\wedge$\\
 & $C_2\stackrel{\triangle}{=} \left(\neq,X\right)$ & $\neg C_2(I)$\\
\hline
\multirow{3}{*}
{I overlaps J} & $C_1\stackrel{\triangle}{=} \left(<,Y\right)$ & $C_1(I)\wedge$\\
 & $C_2\stackrel{\triangle}{=} \left(\leq,X\right)$ & $\neg C_2(I)\wedge$\\
 & $C_3\stackrel{\triangle}{=} \left(\geq,X\right)$ & $\neg C_3(I)$\\
\hline
\multirow{4}{*}
{I during J} & $C_1\stackrel{\triangle}{=} \left(>,X\right)$ & $\big(C_1(I)\wedge$\\
 & $C_2\stackrel{\triangle}{=} \left(\leq,Y\right)$ & $ C_2(I)\big)\vee$ \\
 & $C_3\stackrel{\triangle}{=} \left(\geq,X\right)$ & $\big(C_3(I)\wedge$ \\
 & $C_4\stackrel{\triangle}{=} \left(<,Y\right)$ & $C_4(I)\big)$\\
\hline
\multirow{2}{*}
{I starts J} & $C_1\stackrel{\triangle}{=} \left(\geq,X\right)$ & $C_1(I)\wedge$\\
 &  $C_2\stackrel{\triangle}{=} \left(\neq,X\right)$ & $\neg C_2(I)$\\
\hline
\multirow{2}{*}
{I finishes J} & $C_1\stackrel{\triangle}{=} \left(\leq,Y\right)$ & $C_1(I)\wedge$\\
 & $C_2\stackrel{\triangle}{=} \left(\neq,Y\right)$ & $\neg C_2(I)$\\
\hline
\end{tabular}
\caption{Allen's relations represented in the framework.}
\label{tab:allen-relations}
\end{table}

% \subsection{\label{subsec:ill-known-time-interval}Ill-known time intervals}
% The representation of an ill-known time interval is made by using two ill-known values. Therefore, both starting and ending points are not precissely known. Usually, a crisp interval $[a, b]$ is defined by using a constraint ($a \leq b$). In the case of ill-known intervals, an ill-known constraint is defined.

%%%
%%% Define ill-known time interval in an informal way.
%%%
%%%

% \subsection{Interval Evaluation by Ill-known Constraints}
% The problem of interval evaluation is more generally explained in \cite{Pons2011}: the need exists to know if all points in a crisp interval $I$ reside between the boundaries of an ill-known interval $\left[ X , Y \right]$. In \cite{Pons2011}, the notion of an \emph{ill-known constraint} is introduced:
% 
% \begin{definition}
% Given a universe $U$, an ill-known constraint $C$ on a set $A \subseteq U$ is specified by means of a binary relation $R \subseteq U^{2}$ and a fixed, ill-known value denoted by its possibilistic variable $V$ over $U$, i.e.:
% \begin{align}
% C \triangleq (R,V)
% \end{align}
% Set $A$ now satisfies the constraint if and only if:
% \begin{align}
% \forall a \in A : (a,V) \in R
% \end{align}
% \end{definition}
% 
% An example of an ill-known constraint is $C_{ex} \triangleq (<, X)$. Some set $A$ then satisfies $C_{ex}$ if $\forall a \in A : a < X$, given possibilistic variable $X$.
% 
% The satisfaction of a constraint $C \triangleq (R,V)$ by a set $A$ is basically still a Boolean matter, but due to the uncertainty about the ill-known value $V$, it can be uncertain whether $C$ is satisfied by $A$ or not \cite{Pons2011}. In fact, this satisfaction now behaves as a proposition. Based on the possibility distribution $\pi_{V}$ of $V$, the possibility and necessity that $A$ satisfies $C$ can be found. This proposition can thus be seen as a possibilistic variable on $\mathbb{B}$. The required possibility and necessity are:
% 
% \vspace{-10pt}
% 
% \begin{align}
% \Pos(A\text{ satisfies }C) & = \min_{a \in A}\left(\sup_{(a,w) \in R}\pi_{V}(w)\right) \label{ill-known-pos}\\
% \Nec(A\text{ satisfies }C) & = \min_{a \in A}\left(\inf_{(a,w) \notin R} 1-\pi_{V}(w)\right) \label{ill-known-nec}
% \end{align}
% 
% Now, e.g., to check if crisp interval $I = \left[j, k\right]$ is included in $\left[X, Y\right]$, 2 ill-known constraints are constructed:
% 
% \vspace{-10pt}
% 
% \begin{eqnarray}
% C_1 & \triangleq\left(\geq,X\right)\\
% C_2 & \triangleq\left(\leq,Y\right)
% \end{eqnarray}
% 
% To calculate the possibility and necessity concerning a conjunction of constraints, the $\min$ operator can be used. The possibility and necessity of $I$ being included in $\left[X, Y\right]$ are now: 
% 
% \vspace{-10pt}
% 
% \begin{align}
% \label{eq:interval-pos}
% \Pos(I\text{ satisfies }C_1\ and\ C_2) & = \min_{a \in I}\left(\sup_{a \geq w}\pi_{X}(w),\sup_{a \leq v}\pi_{Y}(v)\right)\\
% \label{eq:interval-nec}
% \Nec(I\text{ satisfies }C_1\ and\ C_2) & = \min_{a \in I}\left(\inf_{a < w} 1-\pi_{X}(w),\inf_{a > v} 1-\pi_{Y}(v)\right).
% \end{align}

\subsection{\label{subsec:triangular-model}The triangular model}

\subsection{\label{subsec:uncertainty-about-time-intervals}Uncertainty about Time Intervals}
Time is a complex natural phenomenon that, among others, allows humans to organise life and plan activities. Time also plays an important role in information management as it allows to register facts with different kinds of timestamps and make use of these timestamps during querying and information retrieval. Dealing with time in an information system requires a time model. In this paper, it is implicitly assumed that time is modelled as being discrete, linear and finite. This means that it is considered that time can only be/ observed using a limited precision, say $\delta$. Observing time using a maximum precision $\delta$ and with respect to a given origin $t_0$ involves a crisp discrete (countable) set of time points, given by
\begin{equation}
 \label{eq:t0-delta}
T_{0,\delta} = \lbrace t_K \in T | t_k = t_0 + k\delta, k \in \Z \rbrace
\end{equation}

where $Z = \left \lbrace 0, 1, -1, 2, -2, \ldots \rbrace \right. T_{0, \delta}$ is a proper subset of the continuum $T$ of the physical time points. The discretisation can be described as a surjective mapping from $T$ on $T_{0,\delta}$, which maps each $\tau \in T$ on the element $t_k \in T_{0,\delta}$ that lies in the interval
\begin{equation}
\gamma_{0,k} = \left[ t_0 + k\delta, t_0 + (k+1)\delta \right[, k \in \Z 
\end{equation}

Discretisation implies that all time points within an interval $\gamma_{0,k}$ are considered to be indistinguishable and mapped to the same time point $t_k$. This is not really a limitation, on condition that the precision $\delta$ is chosen sufficiently accurate. The discretisation is necessary to circumvent the density problem (i.e., the fact that between any two distinct points, there always exists at least one other point). Linearity implies a total order over the time points. Finally, our model is chosen to be finite in view of a computer representation. This implies that all values that exceed the determined upper and lower bound will be handled by introducing two special values $(- \infty $ and  $+\infty)$.

The intervals $\gamma_{0,k}$ with $\delta$ as length are usually called `chronons' in literature (cf. \cite{Dyreson1994}). The introduction of chronons eliminates the necessity of making a distinction between points and intervals, because even a chronon, the smallest unit in the model, is essentially an interval.

A time interval $\left[t_s, t_e \right]$ is defined by a start time point $t_s$  and a length $l$, which defines the distance (number of chronons) from ts to the end time point $t_e$, i.e., $l = t_e - t_s$.

A time point is equivalent to a time interval with $l = 0$ (or $t_e = t_s$ ). Note that the length of an interval is not the same as the duration of the interval: the length of a time point is 0, its duration is 1 chronon ($\delta$).

In practice, it often occurs that time intervals cannot be exactly specified. Especially, in historical data it is often the case that either the starting date or the ending date (or both) of a time interval are partially or completely unknown. In such cases, the best thing to do is modelling the available knowledge about (the uncertainty of) the interval as accurate as possible, hereby avoiding information loss. Soft computing techniques, more specifically possibility theory \cite{DidierDubois1988a}, can be used for that purpose. In \cite{Billiet2012}, \cite{JoseEnriquePons2012} a possibilistic valid time model able to cope with time intervals that have an uncertain start and/or end point has been presented. The basics of this model are used in this paper and briefly described as follows. Each (Un)certain Time Interval (UTI) is defined by a pair
\begin{equation}
 \label{eq:uncertain-time-interval}
\left(\pi_s, \pi_e \right)
\end{equation}

where $\pi_s$ and $\pi_e$ are two convex possibility distributions, respectively reflecting the knowledge about the start and end point of the UTI (as illustrated in Fig. \ref{}(a)). In case of certainty, the possibility distribution is characterised by a fuzzy set with singleton support and core containing the crisp date. In case of uncertainty, the possibility distribution is characterised by a fuzzy set containing all possible candidate values for the date and their associated degree of possibility. 
% figura 2

Together, both possibility distributions reflect the available knowledge about the start and end of the UTI they model. This implies that $\pi$s and $\pi$e together represent another possibility distribution $\pi$I consisting of all possible time intervals that can be constructed from $\pi$s and $\pi$e . Some of the intervals in $\pi_I$ are depicted in Fig. \ref{}(b). For each time interval $\left[t_s , t_e \right]$, its associated possibility grade in $\pi_I$ is computed by
\begin{equation}
 \label{eq:possibility-uti}
\pi_{I} \left(\left[t_s, t_e\right] \right) = 
\begin{cases}
 \min \left(\pi_s (t_s), \pi_e (t_e) \right), & \mbox{ if } t_s \leq t_e\\
0, & \mbox{ otherwise. }
\end{cases}
\end{equation}

Eq. \eqref{eq:possibility-uti} reflects that given the uncertainty about the start point, modelled by $\pi$s and the uncertainty about the end point, modelled by $\pi_e$ $\pi$e , the possibility that $\left[t_s, t_e \right]$ is the actual value of the UTI equals the possibility that ts is the actual start point (i.e., $\pi_s (t_s)$) and te is the actual end point (i.e., $\pi_e (t_e)$) of $I$, where the conjunction of both conditions is modelled by the minimum t-norm. In the inconsistent case where $t_s > t_e$ the interval is considered to be completely impossible. In the remainder of this paper we will denote an UTI as $\pi_I$, i.e., $\pi_I = \left( \pi_s, \pi_e \right)$.

\subsection{\label{subsec:the-triangular-model}The triangular model}
Mainly due to its linear characteristic, time is traditionally visualised using a linear representation where several time intervals are depicted as linear segments on top of a horizontal time axis. This approach has also been used in Fig. \ref{}(b). One can read the start and end point of each interval the horizontal scale. The vertical dimension is solely used to differentiate multiple overlapping intervals, if used at all. The linear representation works well as long as the number of represented time intervals remains low. As soon as a large number of overlapping intervals have to be represented the representation becomes overloaded and accurate visual data analysis becomes almost impossible.

For that reason, an alternative Triangular time Model (TM) has been presented in \cite{Weghe2007}. This model is an extension of Kulpa's triangular model for the modelling of crisp time intervals \cite{Kulpa1997}. Basically, intervals are modelled as interval points in a two-dimensional space as follows. First, a horizontal time axis is considered. Second, for each time interval $\left[t_s, t_e \right]$ two straight lines $L_s$ and $L_e$ are constructed as depicted in Fig. \ref{}. Line $L_s$ is passing through $t_s$ and makes a fixed angle $\alpha$ with the horizontal time axis, whereas Le is passing through te making a fixed angle $-\alpha$ with the horizontal time axis. The interval $\left[t_s, t_e \right]$ is then represented by the intersection of $L_s$ and $L_e$ , which is called the interval point of $\left[t_s, t_e \right]$. For the ease of use, $\alpha$ is here chosen to be $45^{\circ}$ . In TM all time intervals are represented by their corresponding interval point. So, because $\alpha$ is fixed for all interval points, the position of the interval point completely determines the start and end of the interval. The two-dimensional space of interval points is called the interval space and is denoted by IR \cite{Kulpa2006}.

%Figure 3

From the observation that the start points and end points of two crisp time intervals $I^{1} = \left[t_s^1, t_e^1 \right]$ and $I^2 = \left[t_s^2, t_e^2 \right]$ can be smaller than ($<$), equal to ($=$) or larger than ($>$) each other, Allen proposed the thirteen possible relations between two crisp time intervals given in Table \ref{} \cite{Allen1983}.

In TM, these thirteen Allen relations each correspond to a specific zone in the interval space IR \cite{Kulpa1997}. These thirteen zones are called Crisp Relational Zones (CRZs). The CRZs with respect to a reference interval point $I$ are depicted in Fig. \ref{}. Hereby, the following shorthand notations are used: equals (E), starts (S), started by (SB), finishes (F ), finished by (F B), meets (M ), met by (M B), overlaps (O), overlapped by (OB), during (D), contains (C), before (B) and after (A).

% TABLE I
% T HE THIRTEEN POSSIBLE A LLEN RELATIONS BETWEEN CRISP TIME
% INTERVALS
% 
% I 1 equals I 2
% I 1 starts I 2
% I 1 started by I 2
% I 1 f inishes I 2
% I 1 f inished by I 2
% I 1 meets I 2
% I 1 met by I 2
% I 1 overlaps I 2
% I 1 overlapped by I 2
% I 1 during I 2
% I 1 contains I 2
% I 1 bef ore I 2
% I 1 af ter I 2

Each CRZ represents the set of interval points of the crisp time intervals that are in the corresponding Allen relation with respect to the reference interval $I$. For example, all interval points of intervals that come before $I$ are located in the CRZ represented by the leftmost lower triangle in the interval space. Likewise, if an interval point is located in the top quadrangle above $I$, then its corresponding interval contains $I$. CRZs allow to visually analyse relations between (large) sets of interval points.

### 2013-02-25 trunk/eusflat2013/proposal.tex
first version paper eusflat

In this section we will analyze first the calculation of all the possible Allen's relations between a crisp time point and an ill-known time point and between a crisp time point and an uncertain time indication (UTI). Then, we will compare two ill-known time points and two uncertain time indications.

\subsection{\label{subsec:prop-crisp-vs-ill}Calculation of crisp time intervals}
In this section we first study how uncertainty about wheter crisp interval is enclosed by the ill-known interval given by $[X,Y]$ or by the UTI.

\subsubsection{Crisp time interval enclosed by an ill-known time interval}
Consider a crisp time interval given by $I=[a, b]$. We want to know whether all the points in this interval are within the boundaries of the ill-known interval given by $[X, Y]$. The calculation is done by means of two ill-known constraints as explained in \cite{}:
\begin{eqnarray}
C_1\stackrel{\triangle}{=}\left(\geq,X\right)\\
C_2\stackrel{\triangle}{=}\left(\leq,Y\right).
\end{eqnarray}

The possibility and the necessity are given by:
\begin{eqnarray}
\Pos\left(\lambda([a,b])\right)&=&\min\bigg(\Pos(C_1([a,b])),\Pos(C_2([a,b]))\bigg)\\
\Nec\left(\lambda([a,b])\right)&=&\min\bigg(\Nec(C_1([a,b])),\Nec(C_2([a,b]))\bigg).
\end{eqnarray}
Which can be expanded to:
\begin{eqnarray}
\label{eq:interval-pos}
\Pos\left(\lambda([a,b])\right)&=&\min\bigg(\sup_{a\leq w}\pi_{X}(w),\sup_{b\geq w}\pi_{Y}(w)\bigg)\\
\label{eq:interval-nec}
\Nec\left(\lambda([a,b])\right)&=&\min\bigg(\inf_{a>w}1-\pi_{X}(w),\inf_{b<w}1-\pi_{Y}(w)\bigg).
\end{eqnarray}

\subsubsection{\label{subsubsec:crisp-uti}Crisp time interval enclosed by an UTI}
Consider now a crisp time interval given by $I=[a, b]$ and an Uncertain Time Interval given by $\pi_I = \left(\pi_s, \pi_e \right)$. Again, we want to know if all the points in the interval are within the boundaries of the UTI. The calculation is done as explained in \cite{}:
\begin{equation}
 \pi_I \left( \left[a, b  \right] \right) =
\begin{cases}
 \min \left(\pi_s(a),\pi_e(b)\right) & \mbox{ if } a \leq b\\
0 				     & \mbox{ otherwise. }
\end{cases}
\end{equation}

### 2013-02-26 eusflat2013/preliminaries.tex
updated version

--- 2013-02-25 trunk/eusflat2013/preliminaries.tex
@@
 
-\subsection{\label{subsec:ikc-framework}The ill-known constraint framework}
+\subsection{\label{subsec:ikc-framework}Ill-known constraints}
 A constraint is a restriction in the values that a given variable can take.
@@
 % \end{align}
-
-\subsection{\label{subsec:triangular-model}The triangular model}
-
-\subsection{\label{subsec:uncertainty-about-time-intervals}Uncertainty about Time Intervals}
-Time is a complex natural phenomenon that, among others, allows humans to organise life and plan activities.
-Time also plays an important role in information management as it allows to register facts with different kinds of timestamps and make use of these timestamps during querying and information retrieval.
-Dealing with time in an information system requires a time model.
-In this paper, it is implicitly assumed that time is modelled as being discrete, linear and finite.
-This means that it is considered that time can only be/ observed using a limited precision, say $\delta$.
-Observing time using a maximum precision $\delta$ and with respect to a given origin $t_0$ involves a crisp discrete (countable) set of time points, given by
-\begin{equation}
- \label{eq:t0-delta}
-T_{0,\delta} = \lbrace t_K \in T | t_k = t_0 + k\delta, k \in \Z \rbrace
-\end{equation}
-
-where $Z = \left \lbrace 0, 1, -1, 2, -2, \ldots \rbrace \right.
-T_{0, \delta}$ is a proper subset of the continuum $T$ of the physical time points.
-The discretisation can be described as a surjective mapping from $T$ on $T_{0,\delta}$, which maps each $\tau \in T$ on the element $t_k \in T_{0,\delta}$ that lies in the interval
-\begin{equation}
-\gamma_{0,k} = \left[ t_0 + k\delta, t_0 + (k+1)\delta \right[, k \in \Z 
-\end{equation}
-
-Discretisation implies that all time points within an interval $\gamma_{0,k}$ are considered to be indistinguishable and mapped to the same time point $t_k$.
-This is not really a limitation, on condition that the precision $\delta$ is chosen sufficiently accurate.
-The discretisation is necessary to circumvent the density problem (i.e., the fact that between any two distinct points, there always exists at least one other point).
-Linearity implies a total order over the time points.
-Finally, our model is chosen to be finite in view of a computer representation.
-This implies that all values that exceed the determined upper and lower bound will be handled by introducing two special values $(- \infty $ and  $+\infty)$.
-
-The intervals $\gamma_{0,k}$ with $\delta$ as length are usually called `chronons' in literature (cf.
-\cite{Dyreson1994}).
-The introduction of chronons eliminates the necessity of making a distinction between points and intervals, because even a chronon, the smallest unit in the model, is essentially an interval.
-
-A time interval $\left[t_s, t_e \right]$ is defined by a start time point $t_s$  and a length $l$, which defines the distance (number of chronons) from ts to the end time point $t_e$, i.e., $l = t_e - t_s$.
-
-A time point is equivalent to a time interval with $l = 0$ (or $t_e = t_s$ ).
-Note that the length of an interval is not the same as the duration of the interval: the length of a time point is 0, its duration is 1 chronon ($\delta$).
-
-In practice, it often occurs that time intervals cannot be exactly specified.
-Especially, in historical data it is often the case that either the starting date or the ending date (or both) of a time interval are partially or completely unknown.
-In such cases, the best thing to do is modelling the available knowledge about (the uncertainty of) the interval as accurate as possible, hereby avoiding information loss.
-Soft computing techniques, more specifically possibility theory \cite{DidierDubois1988a}, can be used for that purpose.
-In \cite{Billiet2012}, \cite{JoseEnriquePons2012} a possibilistic valid time model able to cope with time intervals that have an uncertain start and/or end point has been presented.
-The basics of this model are used in this paper and briefly described as follows.
-Each (Un)certain Time Interval (UTI) is defined by a pair
-\begin{equation}
- \label{eq:uncertain-time-interval}
-\left(\pi_s, \pi_e \right)
-\end{equation}
-
-where $\pi_s$ and $\pi_e$ are two convex possibility distributions, respectively reflecting the knowledge about the start and end point of the UTI (as illustrated in Fig.
-\ref{}(a)).
-In case of certainty, the possibility distribution is characterised by a fuzzy set with singleton support and core containing the crisp date.
-In case of uncertainty, the possibility distribution is characterised by a fuzzy set containing all possible candidate values for the date and their associated degree of possibility.
-% figura 2
-
-Together, both possibility distributions reflect the available knowledge about the start and end of the UTI they model.
-This implies that $\pi$s and $\pi$e together represent another possibility distribution $\pi$I consisting of all possible time intervals that can be constructed from $\pi$s and $\pi$e .
-Some of the intervals in $\pi_I$ are depicted in Fig.
-\ref{}(b).
-For each time interval $\left[t_s , t_e \right]$, its associated possibility grade in $\pi_I$ is computed by
-\begin{equation}
- \label{eq:possibility-uti}
-\pi_{I} \left(\left[t_s, t_e\right] \right) = 
-\begin{cases}
- \min \left(\pi_s (t_s), \pi_e (t_e) \right), & \mbox{ if } t_s \leq t_e\\
-0, & \mbox{ otherwise. }
-\end{cases}
-\end{equation}
-
-Eq.
-\eqref{eq:possibility-uti} reflects that given the uncertainty about the start point, modelled by $\pi$s and the uncertainty about the end point, modelled by $\pi_e$ $\pi$e , the possibility that $\left[t_s, t_e \right]$ is the actual value of the UTI equals the possibility that ts is the actual start point (i.e., $\pi_s (t_s)$) and te is the actual end point (i.e., $\pi_e (t_e)$) of $I$, where the conjunction of both conditions is modelled by the minimum t-norm.
-In the inconsistent case where $t_s > t_e$ the interval is considered to be completely impossible.
-In the remainder of this paper we will denote an UTI as $\pi_I$, i.e., $\pi_I = \left( \pi_s, \pi_e \right)$.
-
-\subsection{\label{subsec:the-triangular-model}The triangular model}
-Mainly due to its linear characteristic, time is traditionally visualised using a linear representation where several time intervals are depicted as linear segments on top of a horizontal time axis.
-This approach has also been used in Fig.
-\ref{}(b).
-One can read the start and end point of each interval the horizontal scale.
-The vertical dimension is solely used to differentiate multiple overlapping intervals, if used at all.
-The linear representation works well as long as the number of represented time intervals remains low.
-As soon as a large number of overlapping intervals have to be represented the representation becomes overloaded and accurate visual data analysis becomes almost impossible.
-
-For that reason, an alternative Triangular time Model (TM) has been presented in \cite{Weghe2007}.
-This model is an extension of Kulpa's triangular model for the modelling of crisp time intervals \cite{Kulpa1997}.
-Basically, intervals are modelled as interval points in a two-dimensional space as follows.
-First, a horizontal time axis is considered.
-Second, for each time interval $\left[t_s, t_e \right]$ two straight lines $L_s$ and $L_e$ are constructed as depicted in Fig.
-\ref{}.
-Line $L_s$ is passing through $t_s$ and makes a fixed angle $\alpha$ with the horizontal time axis, whereas Le is passing through te making a fixed angle $-\alpha$ with the horizontal time axis.
-The interval $\left[t_s, t_e \right]$ is then represented by the intersection of $L_s$ and $L_e$ , which is called the interval point of $\left[t_s, t_e \right]$.
-For the ease of use, $\alpha$ is here chosen to be $45^{\circ}$ .
-In TM all time intervals are represented by their corresponding interval point.
-So, because $\alpha$ is fixed for all interval points, the position of the interval point completely determines the start and end of the interval.
-The two-dimensional space of interval points is called the interval space and is denoted by IR \cite{Kulpa2006}.
-
-%Figure 3
-
-From the observation that the start points and end points of two crisp time intervals $I^{1} = \left[t_s^1, t_e^1 \right]$ and $I^2 = \left[t_s^2, t_e^2 \right]$ can be smaller than ($<$), equal to ($=$) or larger than ($>$) each other, Allen proposed the thirteen possible relations between two crisp time intervals given in Table \ref{} \cite{Allen1983}.
-
-In TM, these thirteen Allen relations each correspond to a specific zone in the interval space IR \cite{Kulpa1997}.
-These thirteen zones are called Crisp Relational Zones (CRZs).
-The CRZs with respect to a reference interval point $I$ are depicted in Fig.
-\ref{}.
-Hereby, the following shorthand notations are used: equals (E), starts (S), started by (SB), finishes (F ), finished by (F B), meets (M ), met by (M B), overlaps (O), overlapped by (OB), during (D), contains (C), before (B) and after (A).
-
-% TABLE I
-% T HE THIRTEEN POSSIBLE A LLEN RELATIONS BETWEEN CRISP TIME
-% INTERVALS
-% 
-% I 1 equals I 2
-% I 1 starts I 2
-% I 1 started by I 2
-% I 1 f inishes I 2
-% I 1 f inished by I 2
-% I 1 meets I 2
-% I 1 met by I 2
-% I 1 overlaps I 2
-% I 1 overlapped by I 2
-% I 1 during I 2
-% I 1 contains I 2
-% I 1 bef ore I 2
-% I 1 af ter I 2
-
-Each CRZ represents the set of interval points of the crisp time intervals that are in the corresponding Allen relation with respect to the reference interval $I$.
-For example, all interval points of intervals that come before $I$ are located in the CRZ represented by the leftmost lower triangle in the interval space.
-Likewise, if an interval point is located in the top quadrangle above $I$, then its corresponding interval contains $I$.
-CRZs allow to visually analyse relations between (large) sets of interval points.

### 2013-02-26 trunk/eusflat2013/triangularmodel.tex
updated version

In this section we will explain a graphical visualization of temporal data called the triangular model (TM). This has been widely used mainly on rough sets for the visualization of historical data. More recently, the model has been extended on possibility theory to model Uncertain Time Intervals (UTI).
First of all we will explain the main concepts and properties of the triangular model to model Uncertain Time Intervals (UTI). Then, we will detail how to compute the Allen's Relations between a crisp time point and an UTI.

\subsection{\label{subsec:utis}Uncertain Time Intervals}
As mentioned in the previous section, the boundaries of time intervals are not always well-known. The available knowledge about the interval can be modelled by using possibility distributions ~\cite{Billiet2012}, ~\cite{JoseEnriquePons2012} on both starting and ending points as mentioned in the previous section. Informally, an Uncertain Time Interval UTI is a time interval for which one or both boundaries are not precisely known.

\begin{definition}
 \label{def:uncertain-time-interval}
An uncertain time interval UTI $I$ is composed by an starting point $S$, and an ending point $E$. For both points, two different convex possibility distributions are defined, namely $\pi_s$, $\pi_e$. These two possibility distributions represent all the possible time intervals that can be constructed from $I$. The following formula obtains the possibility $\pi_{I}$ that a crisp interval  given by  $[a, b]$ is constructed from $I$. 
\begin{equation}
 \label{eq:possibility-uti}
\pi_{I} \left(\left[t_s, t_e\right] \right) = 
\begin{cases}
 \min \left(\pi_s (t_s), \pi_e (t_e) \right), & \mbox{ if } t_s \leq t_e\\
0, & \mbox{ otherwise. }
\end{cases}
\end{equation}

In the following we will note an UTI as $\pi_{I} = \left(\pi_s, \pi_e \right)$.
% It is important to notice that, the main diference with respect to the ill-known time interval (see Definition \ref{}) is that an UTI only consider as possible starting and ending points

\end{definition}

\subsection{\label{subsec:visualization-tm}Visualization within the triangular model}
Time intervals are usually represented in a one dimensional scale. This is due to its linear characteristic. This approach has also been used in Figure \ref{fig:ill-known-ti} and \ref{}. The dificulty of accurate visual analisys on the temporal data arises when the number of time intervals increase.

The triangular model (TM) ~\cite{Kulpa1997}, \cite{Weghe2007} represents time intervals in a two dimensional space. The procedure to represent a time interval $[t_s, t_e]$ as an interval point in a two dimensional space is the following:

\begin{enumerate}
 \item Draw each time point of the interval in the horizontal line.
\item Construct one straight line from $t_s$ with an angle $\alpha$.
\item Construct one straight line from $t_e$ with an angle $-\alpha$.
\item The intersection of these two lines is the interval point.
\end{enumerate}

For the shake of simplicity, in this work, the angle for $\alpha$ is $45^{\circ}$ .

% Mainly due to its linear characteristic, time is traditionally visualised using a linear representation where several time intervals are depicted as linear segments on top of a horizontal time axis. This approach has also been used in Fig. \ref{fig:ill-known-ti}. One can read the start and end point of each interval the horizontal scale. The vertical dimension is solely used to differentiate multiple overlapping intervals, if used at all. The linear representation works well as long as the number of represented time intervals remains low. As soon as a large number of overlapping intervals have to be represented the representation becomes overloaded and accurate visual data analysis becomes almost impossible.

For that reason, an alternative Triangular time Model (TM) has been presented in \cite{Weghe2007}. This model is an extension of Kulpa's triangular model for the modelling of crisp time intervals \cite{Kulpa1997}. Basically, intervals are modelled as interval points in a two-dimensional space as follows. First, a horizontal time axis is considered. Second, for each time interval $\left[t_s, t_e \right]$ two straight lines $L_s$ and $L_e$ are constructed as depicted in Fig. \ref{}. Line $L_s$ is passing through $t_s$ and makes a fixed angle $\alpha$ with the horizontal time axis, whereas Le is passing through te making a fixed angle $-\alpha$ with the horizontal time axis. The interval $\left[t_s, t_e \right]$ is then represented by the intersection of $L_s$ and $L_e$ , which is called the interval point of $\left[t_s, t_e \right]$. For the ease of use, $\alpha$ is here chosen to be $45^{\circ}$ . In TM all time intervals are represented by their corresponding interval point. So, because $\alpha$ is fixed for all interval points, the position of the interval point completely determines the start and end of the interval. The two-dimensional space of interval points is called the interval space and is denoted by IR \cite{Kulpa2006}.

%Figure 3

From the observation that the start points and end points of two crisp time intervals $I^{1} = \left[t_s^1, t_e^1 \right]$ and $I^2 = \left[t_s^2, t_e^2 \right]$ can be smaller than ($<$), equal to ($=$) or larger than ($>$) each other, Allen proposed the thirteen possible relations between two crisp time intervals given in Table \ref{} \cite{Allen1983}.

% \subsection{\label{subsec:uncertainty-about-time-intervals}Uncertainty about Time Intervals}
% Time is a complex natural phenomenon that, among others, allows humans to organise life and plan activities. Time also plays an important role in information management as it allows to register facts with different kinds of timestamps and make use of these timestamps during querying and information retrieval. Dealing with time in an information system requires a time model. In this paper, it is implicitly assumed that time is modelled as being discrete, linear and finite. This means that it is considered that time can only be/ observed using a limited precision, say $\delta$. Observing time using a maximum precision $\delta$ and with respect to a given origin $t_0$ involves a crisp discrete (countable) set of time points, given by
% \begin{equation}
%  \label{eq:t0-delta}
% T_{0,\delta} = \lbrace t_K \in T | t_k = t_0 + k\delta, k \in \Z \rbrace
% \end{equation}
% 
% where $Z = \left \lbrace 0, 1, -1, 2, -2, \ldots \rbrace \right. T_{0, \delta}$ is a proper subset of the continuum $T$ of the physical time points. The discretisation can be described as a surjective mapping from $T$ on $T_{0,\delta}$, which maps each $\tau \in T$ on the element $t_k \in T_{0,\delta}$ that lies in the interval
% \begin{equation}
% \gamma_{0,k} = \left[ t_0 + k\delta, t_0 + (k+1)\delta \right[, k \in \Z 
% \end{equation}
% 
% Discretisation implies that all time points within an interval $\gamma_{0,k}$ are considered to be indistinguishable and mapped to the same time point $t_k$. This is not really a limitation, on condition that the precision $\delta$ is chosen sufficiently accurate. The discretisation is necessary to circumvent the density problem (i.e., the fact that between any two distinct points, there always exists at least one other point). Linearity implies a total order over the time points. Finally, our model is chosen to be finite in view of a computer representation. This implies that all values that exceed the determined upper and lower bound will be handled by introducing two special values $(- \infty $ and  $+\infty)$.
% 
% The intervals $\gamma_{0,k}$ with $\delta$ as length are usually called `chronons' in literature (cf. \cite{Dyreson1994}). The introduction of chronons eliminates the necessity of making a distinction between points and intervals, because even a chronon, the smallest unit in the model, is essentially an interval. 
% 
% A time interval $\left[t_s, t_e \right]$ is defined by a start time point $t_s$  and a length $l$, which defines the distance (number of chronons) from ts to the end time point $t_e$, i.e., $l = t_e - t_s$.
% 
% 
% A time point is equivalent to a time interval with $l = 0$ (or $t_e = t_s$ ). Note that the length of an interval is not the same as the duration of the interval: the length of a time point is 0, its duration is 1 chronon ($\delta$). 
% 
% In practice, it often occurs that time intervals cannot be exactly specified. Especially, in historical data it is often the case that either the starting date or the ending date (or both) of a time interval are partially or completely unknown. In such cases, the best thing to do is modelling the available knowledge about (the uncertainty of) the interval as accurate as possible, hereby avoiding information loss. Soft computing techniques, more specifically possibility theory \cite{DidierDubois1988a}, can be used for that purpose. In \cite{Billiet2012}, \cite{JoseEnriquePons2012} a possibilistic valid time model able to cope with time intervals that have an uncertain start and/or end point has been presented. The basics of this model are used in this paper and briefly described as follows. Each (Un)certain Time Interval (UTI) is defined by a pair
% \begin{equation}
%  \label{eq:uncertain-time-interval}
% \left(\pi_s, \pi_e \right)
% \end{equation}
% 
% where $\pi_s$ and $\pi_e$ are two convex possibility distributions, respectively reflecting the knowledge about the start and end point of the UTI (as illustrated in Fig. \ref{}(a)). In case of certainty, the possibility distribution is characterised by a fuzzy set with singleton support and core containing the crisp date. In case of uncertainty, the possibility distribution is characterised by a fuzzy set containing all possible candidate values for the date and their associated degree of possibility. 
% % figura 2
% 
% Together, both possibility distributions reflect the available knowledge about the start and end of the UTI they model. This implies that $\pi$s and $\pi$e together represent another possibility distribution $\pi$I consisting of all possible time intervals that can be constructed from $\pi$s and $\pi$e . Some of the intervals in $\pi_I$ are depicted in Fig. \ref{}(b). For each time interval $\left[t_s , t_e \right]$, its associated possibility grade in $\pi_I$ is computed by
% \begin{equation}
%  \label{eq:possibility-uti}
% \pi_{I} \left(\left[t_s, t_e\right] \right) = 
% \begin{cases}
%  \min \left(\pi_s (t_s), \pi_e (t_e) \right), & \mbox{ if } t_s \leq t_e\\
% 0, & \mbox{ otherwise. }
% \end{cases}
% \end{equation}
% 
% Eq. \eqref{eq:possibility-uti} reflects that given the uncertainty about the start point, modelled by $\pi$s and the uncertainty about the end point, modelled by $\pi_e$ $\pi$e , the possibility that $\left[t_s, t_e \right]$ is the actual value of the UTI equals the possibility that ts is the actual start point (i.e., $\pi_s (t_s)$) and te is the actual end point (i.e., $\pi_e (t_e)$) of $I$, where the conjunction of both conditions is modelled by the minimum t-norm. In the inconsistent case where $t_s > t_e$ the interval is considered to be completely impossible. In the remainder of this paper we will denote an UTI as $\pi_I$, i.e., $\pi_I = \left( \pi_s, \pi_e \right)$.

\subsection{\label{subsec:the-triangular-model}The triangular model}
Mainly due to its linear characteristic, time is traditionally visualised using a linear representation where several time intervals are depicted as linear segments on top of a horizontal time axis. This approach has also been used in Fig. \ref{}(b). One can read the start and end point of each interval the horizontal scale. The vertical dimension is solely used to differentiate multiple overlapping intervals, if used at all. The linear representation works well as long as the number of represented time intervals remains low. As soon as a large number of overlapping intervals have to be represented the representation becomes overloaded and accurate visual data analysis becomes almost impossible.

For that reason, an alternative Triangular time Model (TM) has been presented in \cite{Weghe2007}. This model is an extension of Kulpa's triangular model for the modelling of crisp time intervals \cite{Kulpa1997}. Basically, intervals are modelled as interval points in a two-dimensional space as follows. First, a horizontal time axis is considered. Second, for each time interval $\left[t_s, t_e \right]$ two straight lines $L_s$ and $L_e$ are constructed as depicted in Fig. \ref{}. Line $L_s$ is passing through $t_s$ and makes a fixed angle $\alpha$ with the horizontal time axis, whereas Le is passing through te making a fixed angle $-\alpha$ with the horizontal time axis. The interval $\left[t_s, t_e \right]$ is then represented by the intersection of $L_s$ and $L_e$ , which is called the interval point of $\left[t_s, t_e \right]$. For the ease of use, $\alpha$ is here chosen to be $45^{\circ}$ . In TM all time intervals are represented by their corresponding interval point. So, because $\alpha$ is fixed for all interval points, the position of the interval point completely determines the start and end of the interval. The two-dimensional space of interval points is called the interval space and is denoted by IR \cite{Kulpa2006}.

%Figure 3

From the observation that the start points and end points of two crisp time intervals $I^{1} = \left[t_s^1, t_e^1 \right]$ and $I^2 = \left[t_s^2, t_e^2 \right]$ can be smaller than ($<$), equal to ($=$) or larger than ($>$) each other, Allen proposed the thirteen possible relations between two crisp time intervals given in Table \ref{} \cite{Allen1983}.

In TM, these thirteen Allen relations each correspond to a specific zone in the interval space IR \cite{Kulpa1997}. These thirteen zones are called Crisp Relational Zones (CRZs). The CRZs with respect to a reference interval point $I$ are depicted in Fig. \ref{}. Hereby, the following shorthand notations are used: equals (E), starts (S), started by (SB), finishes (F ), finished by (F B), meets (M ), met by (M B), overlaps (O), overlapped by (OB), during (D), contains (C), before (B) and after (A).

% TABLE I
% T HE THIRTEEN POSSIBLE A LLEN RELATIONS BETWEEN CRISP TIME
% INTERVALS
% 
% I 1 equals I 2
% I 1 starts I 2
% I 1 started by I 2
% I 1 f inishes I 2
% I 1 f inished by I 2
% I 1 meets I 2
% I 1 met by I 2
% I 1 overlaps I 2
% I 1 overlapped by I 2
% I 1 during I 2
% I 1 contains I 2
% I 1 bef ore I 2
% I 1 af ter I 2

Each CRZ represents the set of interval points of the crisp time intervals that are in the corresponding Allen relation with respect to the reference interval $I$. For example, all interval points of intervals that come before $I$ are located in the CRZ represented by the leftmost lower triangle in the interval space. Likewise, if an interval point is located in the top quadrangle above $I$, then its corresponding interval contains $I$. CRZs allow to visually analyse relations between (large) sets of interval points.

### 2013-02-27 trunk/eusflat2013/introduction.tex
new version

\begin{table}[h]
\centering
\begin{tabular}{|c|l|}
\hline
Name & Implementation \\ \hline 
$I$ equals $J$ & if $s_i = s_j \wedge e_i = e_j $ \\
$I$ starts $J$ & if $s_i = s_j \wedge e_i < e_j $ \\
$I$ started by $J$ & if $s_i = s_j \wedge e_i > e_j $ \\
$I$ finishes $J$ & if $s_i > s_j \wedge e_i = e_j $ \\
$I$ finished by $J$ & if $s_i < s_j \wedge e_i = e_j $ \\
$I$ meets $J$ & if $e_i = s_j $ \\
$I$ met by $J$ & if $s_i = e_j $ \\
$I$ overlaps $J$ & if $s_i < s_j \wedge e_i < e_j \wedge e_i > s_j $ \\
$I$ overlapped by $J$ & if $s_i > s_j \wedge e_j < e_i \wedge s_i < e_j  $ \\
$I$ during $J$ & if $s_i > s_j \wedge e_i < e_j $ \\
$I$ contains $J$ & if $  s_i < s_j \wedge e_i > e_j$ \\
$I$ before $J$ & if $e_i < s_j $ \\
$I$ after $J$ & if $s_i > e_j $ \\
\hline
\end{tabular}
\caption{Allen's relations represented in the framework. $I = \left[s_i, e_i\right]$, $J=  \left[s_j, e_j\right]$}
\label{tab:allen-relations}
\end{table}

\begin{figure}[h]
   \centering
   \includegraphics[width=0.8\columnwidth]{graphs/allen.eps}
   \caption{The Allen relationships between two crisp intervals $I$ and $J$.  }
   \label{fig:allen-relationships}
 \end{figure}

### 2013-02-27 trunk/eusflat2013/preliminaries.tex
new version

@@
 
-Several ill-known constraints can be aggregated by using boolean operators: and ($\wedge$), or ($\vee$), not ($\neg$). The possibility and neccesity for these operators are computed as follows.
+Several ill-known constraints can be aggregated by using the possibilistic extension of boolean operators: and ($\wedge$), or ($\vee$), not ($\neg$).
+The possibility and neccesity for these operators are computed as follows.
 

### 2013-02-27 trunk/eusflat2013/triangularmodel.tex
new version

@@
 This is due to its linear characteristic.
-This approach has also been used in Figure \ref{fig:ill-known-ti} and \ref{}.
+This approach has also been used in Figure\ref{fig:allen-relationships} and \ref{fig:ill-known-ti}.
 The dificulty of accurate visual analisys on the temporal data arises when the number of time intervals increase.
 
-The triangular model (TM) ~\cite{Kulpa1997}, \cite{Weghe2007} represents time intervals in a two dimensional space.
+The triangular model (TM) ~\cite{Kulpa1997}, \cite{Weghe2007} represents time intervals in a two dimensional space called IRterval point completely determines the start and end of the interval.
+The two-dimensional space of interval points is called the interval space and is denoted by IR ~\cite{Kulpa2006}.
 The procedure to represent a time interval $[t_s, t_e]$ as an interval point in a two dimensional space is the following:
@@
 \begin{enumerate}
- \item Draw each time point of the interval in the horizontal line.
-\item Construct one straight line from $t_s$ with an angle $\alpha$.
-\item Construct one straight line from $t_e$ with an angle $-\alpha$.
+\item Draw each time point of the interval $[t_s, t_e]$ in the horizontal line.
+\item Construct one straight line $L_s$ from $t_s$ with an angle $\alpha$.
+\item Construct one straight line $L_e$ from $t_e$ with an angle $-\alpha$.
 \item The intersection of these two lines is the interval point.
@@
 
-For the shake of simplicity, in this work, the angle for $\alpha$ is $45^{\circ}$ .
+For the shake of simplicity, in this work, the angle for $\alpha$ is chosen to be $45^{\circ}$.
+
+In TM, these thirteen Allen relations each correspond to a specific zone in the interval space IR \cite{Kulpa1997}.
+These thirteen zones are called Crisp Relational Zones (CRZs).
+The CRZs with respect to a reference interval point $I$ are depicted in Figure \ref{fig:allencrz}.
+
+\begin{figure}[h]
+   \centering
+   \includegraphics[width=0.8\columnwidth]{graphs/allencrc.eps}
+   \caption{CRZs corresponding to the thirdteen Allen's relations.
+The following shorthand notations are used: equals (E), starts (S), started by (SB), finishes (F), finished by (FB), meets (M), met by (MB), overlaps (O), overlapped by (OB), during (D), contains (C), before (B) and after (A).
+}
+   \label{fig:allencrz}
+ \end{figure}
+
+Each CRZ represents the set of interval points of the crisp time intervals that are in the corresponding Allen relation with respect to the reference interval $I$.
+For example, all interval points of intervals that come before $I$ are located in the CRZ represented by the leftmost lower triangle in the interval space.
+Likewise, if an interval point is located in the top quadrangle above $I$, then its corresponding interval contains $I$.
+CRZs allow to visually analyse relations between (large) sets of interval points.
+
+\subsection{\label{subsec:utis-in-tm}Representation of Uncertain Time Intervals}
+In \cite{DeTre2012} a technique to represent UTIs in the triangular model has been presented.
+Basically, the procedure is an extension of the method to represent crisp time intervals in the triangular space.
+An UTI $\pi_I = \left(\pi_s, \pi_e \right)$ can be represented in the triangular model by the following procedure, considering that both $\pi_s, \pi_e$ are given by triangular membership functions (see Section \ref{subsec:fuzzy-numbers}).
+
+\begin{enumerate}
+ \item Let $\pi_s = \left[D_s, a_s, b_s\right]$ a triangular membership function.
+\item Let $\pi_e = \left[D_e, a_e, b_e \right]$ a triangular membership function.
+\item Let $t_s^{-} = D_s - a_s$, and $t_s^{+} = D_s + b_s$.
+\item Let $t_e^{-} = D_e -a_e$, and $t_e^{+} = D_e + b_e$.
+\item Construct two straight lines $L_s^{-}, L_s^{-}$ from $t_s^{-}, t_s^{+}$ with an angle $\alpha$.
+\item Construct two straight lines $L_e^{-}, L_e^{-}$ from $t_e^{-}, t_e^{+}$ with an angle $-\alpha$.
+\end{enumerate}
+
+The intersection of these lines is a quadrangle zone where the UTI is contained.
+Note that not all the interval points in this zone are to the same extent representing candidate intervals for the UTI.
+A grey scaled color gradient is obtained from Eq.
+\eqref{eq:possibility-uti}.
+As shown in Fig.
+\ref{fig:utitm} the color black is associated with a possibility degree of 1, whereas the color white is used when the possibility is 0.
+%Equation \eqref{eq:possibility-uti} computes the possibility obtained for the UTI with respect to a candidate interval.
+% This formula is used to calculate a grey scaled gradient in which the color black denotes the possibility degree 1 and the color white denotes the possibility degree 0.
+
+\begin{figure}[h]
+   \centering
+   \includegraphics[width=0.8\columnwidth]{graphs/utitm.eps}
+   \caption{Representation in the triangular model for an UTI $\pi_I$.  }
+   \label{fig:utitm}
+ \end{figure}
+
+\subsection{\label{subsec:relational-info-uti}Obtaining Relational Information}
+In order to make temporal reasoning with UTIs, in the triangular model, the thirteen Allen's Relations, and its respectives Crisp Relational Zones (CRZ) have been generalized.
+In this case, we obtain fifteen Uncertain Relational Zones (URZ) , shown in Fig.
+\ref{fig:urz} and in Table \ref{tab:urz}.
+
+\begin{figure}[h]
+   \centering
+   \includegraphics[width=0.8\columnwidth]{graphs/urz.eps}
+   \caption{Uncertain Relational Zones (URZ) for a given UTI.  }
+   \label{fig:urz}
+ \end{figure}
+
+\begin{table}[h]
+\centering
+\begin{tabular}{|l|l|l|}
+\hline
+Symbol & Name & Possible Relations \\
+\hline
+B    & Before & B \\
+O    & Overlaps & O \\
+C    & Contains & C \\
+D    & During & D \\
+OB   & Overlapped by & OB \\
+A    & After & A \\
+PM   & Possibly meets & M, B, O \\
+PS   & Possibly starts & O, S, D \\
+PFB  & Possibly finished by & O, FB, C \\
+\multirow{2}{*}
+{PE}   & Possibly equal  & E, FB, SB, F,\\
+       &                 & S, C, D,O, OB \\
+%       &                 &  \\
+PSB  & Possibly started by & C, OB, SB \\
+PF   & Possibly finishes & OB, MB, A \\
+PMB  & Possibly met by & B, M, O, S, D \\
+UB   & Uncertain beginning & B, M, O, S, D \\
+UE   & Uncertain ending & D, F, OB, MB, A \\
+\hline
+\end{tabular}
+\caption{The fifteen possible relations between an interval point $J$ and an UTI $\pi_I$.}
+\label{tab:urz}
+\end{table}
+
+The way to obtain the relational information between an interval point $J = \left[s_j, e_j \right]$, and an UTI $\pi_I$ depends on the location of $J$ with respect to the UTI $\pi_I$:
+
+\begin{itemize}
+ \item If $J$ is located in a zone with only one possible Allen relation, then it is true that $J$ is in the relation. (And therefore, the possibility is always 1).
+\begin{equation}
+ \label{eq:possibility-iz-uti-certain}
+\mu_{\left(J R \pi_I \right)} \left(R \right) = 1.
+\end{equation}
+
+\item If $J$ is located in a zone with multiple candidate Allen's relations, then there is uncertainty about the Allen relation that applies between $J$ and the UTI $\pi_I$.
+The procedure to obtain the possibility degree is the following:
+
+\begin{enumerate}
+ \item Draw the lines $L_s, L_e$ that represent the interval point $J$, with the procedure shown un Section \ref{subsec:visualization-tm}. 
+\item Draw the orthogonal lines from $L_s, L_e$. 
+\item One or several of these lines could intersect with the UIZ of $\pi_I$.
+Then, the UIZ is divided into as many different subzones $\pi_I^{R'}$ as possible candidate Allen relations $R'$ between $J$ and $\pi_I$ exist.
+\item $\pi_I^{R'}$ is the zone that correspond to a candidate Allen Relation $R'$.
+\end{enumerate}
+The possibility that a candidate relation $R'$ is the actual relation between $J$ and $\pi_I$ is given by:
+\begin{equation}
+ \label{eq:possibility-iz-uti}
+\mu_{\left(J R \pi_I \right)} \left(R' \right) = \sup_{I \in \pi_I^{R'}} \pi_I(I).
+\end{equation}
+\end{itemize}
+
+\begin{figure}[h]
+   \centering
+   \includegraphics[width=0.8\columnwidth]{graphs/calculation.eps}
+   \caption{Uncertain Relational Zones (URZ) for a given UTI.  }
+   \label{fig:calculation}
+ \end{figure}
 
@@
 
-For that reason, an alternative Triangular time Model (TM) has been presented in \cite{Weghe2007}.
+% For that reason, an alternative Triangular time Model (TM) has been presented in \cite{Weghe2007}.
 This model is an extension of Kulpa's triangular model for the modelling of crisp time intervals \cite{Kulpa1997}.
@@
 
-From the observation that the start points and end points of two crisp time intervals $I^{1} = \left[t_s^1, t_e^1 \right]$ and $I^2 = \left[t_s^2, t_e^2 \right]$ can be smaller than ($<$), equal to ($=$) or larger than ($>$) each other, Allen proposed the thirteen possible relations between two crisp time intervals given in Table \ref{} \cite{Allen1983}.
+% From the observation that the start points and end points of two crisp time intervals $I^{1} = \left[t_s^1, t_e^1 \right]$ and $I^2 = \left[t_s^2, t_e^2 \right]$ can be smaller than ($<$), equal to ($=$) or larger than ($>$) each other, Allen proposed the thirteen possible relations between two crisp time intervals given in Table \ref{} \cite{Allen1983}.
+%
 
@@
 
-\subsection{\label{subsec:the-triangular-model}The triangular model}
-Mainly due to its linear characteristic, time is traditionally visualised using a linear representation where several time intervals are depicted as linear segments on top of a horizontal time axis.
+% 
+% \subsection{\label{subsec:the-triangular-model}The triangular model}
+% Mainly due to its linear characteristic, time is traditionally visualised using a linear representation where several time intervals are depicted as linear segments on top of a horizontal time axis.
 This approach has also been used in Fig.
@@
 As soon as a large number of overlapping intervals have to be represented the representation becomes overloaded and accurate visual data analysis becomes almost impossible.
-
-For that reason, an alternative Triangular time Model (TM) has been presented in \cite{Weghe2007}.
+% 
+% For that reason, an alternative Triangular time Model (TM) has been presented in \cite{Weghe2007}.
 This model is an extension of Kulpa's triangular model for the modelling of crisp time intervals \cite{Kulpa1997}.
@@
 The two-dimensional space of interval points is called the interval space and is denoted by IR \cite{Kulpa2006}.
-
-%Figure 3
-
-From the observation that the start points and end points of two crisp time intervals $I^{1} = \left[t_s^1, t_e^1 \right]$ and $I^2 = \left[t_s^2, t_e^2 \right]$ can be smaller than ($<$), equal to ($=$) or larger than ($>$) each other, Allen proposed the thirteen possible relations between two crisp time intervals given in Table \ref{} \cite{Allen1983}.
-
-In TM, these thirteen Allen relations each correspond to a specific zone in the interval space IR \cite{Kulpa1997}.
+% 
+% %Figure 3
+% 
+% From the observation that the start points and end points of two crisp time intervals $I^{1} = \left[t_s^1, t_e^1 \right]$ and $I^2 = \left[t_s^2, t_e^2 \right]$ can be smaller than ($<$), equal to ($=$) or larger than ($>$) each other, Allen proposed the thirteen possible relations between two crisp time intervals given in Table \ref{} \cite{Allen1983}.
+% 
+% In TM, these thirteen Allen relations each correspond to a specific zone in the interval space IR \cite{Kulpa1997}.
 These thirteen zones are called Crisp Relational Zones (CRZs).
@@
 Hereby, the following shorthand notations are used: equals (E), starts (S), started by (SB), finishes (F ), finished by (F B), meets (M ), met by (M B), overlaps (O), overlapped by (OB), during (D), contains (C), before (B) and after (A).
-
-% TABLE I
-% T HE THIRTEEN POSSIBLE A LLEN RELATIONS BETWEEN CRISP TIME
-% INTERVALS
 % 
-% I 1 equals I 2
-% I 1 starts I 2
-% I 1 started by I 2
-% I 1 f inishes I 2
-% I 1 f inished by I 2
-% I 1 meets I 2
-% I 1 met by I 2
-% I 1 overlaps I 2
-% I 1 overlapped by I 2
-% I 1 during I 2
-% I 1 contains I 2
-% I 1 bef ore I 2
-% I 1 af ter I 2
-
-Each CRZ represents the set of interval points of the crisp time intervals that are in the corresponding Allen relation with respect to the reference interval $I$.
+% % TABLE I
+% % T HE THIRTEEN POSSIBLE A LLEN RELATIONS BETWEEN CRISP TIME
+% % INTERVALS
+% % 
+% % I 1 equals I 2
+% % I 1 starts I 2
+% % I 1 started by I 2
+% % I 1 f inishes I 2
+% % I 1 f inished by I 2
+% % I 1 meets I 2
+% % I 1 met by I 2
+% % I 1 overlaps I 2
+% % I 1 overlapped by I 2
+% % I 1 during I 2
+% % I 1 contains I 2
+% % I 1 bef ore I 2
+% % I 1 af ter I 2
+% 
+% Each CRZ represents the set of interval points of the crisp time intervals that are in the corresponding Allen relation with respect to the reference interval $I$.
 For example, all interval points of intervals that come before $I$ are located in the CRZ represented by the leftmost lower triangle in the interval space.

### 2013-02-27 eusflat2013/proposal.tex
new version

\subsection{\label{subsec:sim-dif}Similarities and differences}

In this section we will analyze first the calculation of all the possible Allen's relations between a crisp time point and an ill-known time point and between a crisp time point and an uncertain time indication (UTI). Then, we will compare two ill-known time points and two uncertain time indications.

% \subsection{\label{subsec:prop-crisp-vs-ill}Calculation of crisp time intervals}
% In this section we first study how uncertainty about wheter crisp interval is enclosed by the ill-known interval given by $[X,Y]$ or by the UTI.
% 
% \subsubsection{Crisp time interval enclosed by an ill-known time interval}
% Consider a crisp time interval given by $I=[a, b]$. We want to know whether all the points in this interval are within the boundaries of the ill-known interval given by $[X, Y]$. The calculation is done by means of two ill-known constraints as explained in \cite{}:
% \begin{eqnarray}
% C_1\stackrel{\triangle}{=}\left(\geq,X\right)\\
% C_2\stackrel{\triangle}{=}\left(\leq,Y\right).
% \end{eqnarray}
% 
% The possibility and the necessity are given by:
% \begin{eqnarray}
% \Pos\left(\lambda([a,b])\right)&=&\min\bigg(\Pos(C_1([a,b])),\Pos(C_2([a,b]))\bigg)\\
% \Nec\left(\lambda([a,b])\right)&=&\min\bigg(\Nec(C_1([a,b])),\Nec(C_2([a,b]))\bigg).
% \end{eqnarray}
% Which can be expanded to:
% \begin{eqnarray}
% \label{eq:interval-pos}
% \Pos\left(\lambda([a,b])\right)&=&\min\bigg(\sup_{a\leq w}\pi_{X}(w),\sup_{b\geq w}\pi_{Y}(w)\bigg)\\
% \label{eq:interval-nec}
% \Nec\left(\lambda([a,b])\right)&=&\min\bigg(\inf_{a>w}1-\pi_{X}(w),\inf_{b<w}1-\pi_{Y}(w)\bigg).
% \end{eqnarray}
% 
% 
% \subsubsection{\label{subsubsec:crisp-uti}Crisp time interval enclosed by an UTI}
% Consider now a crisp time interval given by $I=[a, b]$ and an Uncertain Time Interval given by $\pi_I = \left(\pi_s, \pi_e \right)$. Again, we want to know if all the points in the interval are within the boundaries of the UTI. The calculation is done as explained in \cite{}:
% \begin{equation}
%  \pi_I \left( \left[a, b  \right] \right) =
% \begin{cases}
%  \min \left(\pi_s(a),\pi_e(b)\right) & \mbox{ if } a \leq b\\
% 0 				     & \mbox{ otherwise. }
% \end{cases}
% \end{equation}

### 2013-03-01 eusflat2013/preliminaries.tex
eusflat2013 new version

@@
    \includegraphics[scale=1.5]{graphs/ill-known-ti.eps}
-   \caption{An ill-known time interval $J = \left[X, Y  \right]$ given by two ill-known values. }
+   \caption{An ill-known time interval IKTI $J = \left[X, Y  \right]$ given by two ill-known values. }
    \label{fig:ill-known-ti}
@@
 
-An ill-known time interval is defined by two ill-known values (see Figure \ref{fig:ill-known-ti}).
+\begin{definition}
+An ill-known time interval \emph{(IKTI)} is defined by two ill-known values (see Figure \ref{fig:ill-known-ti}).
 Therefore, both starting and ending points are not precissely known.
-The following is the notation for an ill-known time interval.
+The following is the notation for an IKTI.
 
@@
 Where both $X$ and $Y$ are ill-known time values.
+\end{definition}
 
-The Allen's Relations between a crisp time interval $I = [a,b]$ ($a, b$ are crisp time points) and an ill-known time interval given by $J = [X,Y]$ ($X, Y$ are ill-known values, see Definition \ref{def:ill-known-value}) can be defined by using aggregations of ill-known constraints.
+The Allen's Relations between a crisp time interval $I = [a,b]$ ($a, b$ are crisp time points) and an IKTI given by $J = [X,Y]$ ($X, Y$ are ill-known values, see Definition \ref{def:ill-known-value}) can be defined by using aggregations of ill-known constraints.
 Table \ref{tab:allen-relations} shows the corresponding ill-known constraints for each Allen's Relation.

### 2013-03-01 eusflat2013/proposal.tex
eusflat2013 new version

@@
+Two frameworks for dealing with uncertainty in time intervals have been presented.
+In this section, we will study the equivalences between an uncertain time interval (UTI) and an ill-known time interval (IKTI).
+Nevertheless, there are some diferences in the semantics of UTIs and IKTIs which are also explained.
+Next to that, the computation of the Allen's Relations are compared.
+
 \subsection{\label{subsec:sim-dif}Similarities and differences}
+A time interval is defined by a starting and an ending point. When uncertainty arises, in one or both starting and / or ending points, then either an UTI $\pi_I$ or an IKTI can be used.
+
+\begin{itemize}
+ \item An IKTI $J = \left[X, Y \right]$ is defined by two ill-known values $X, Y$ which are respectively, the starting and ending points of the time interval.
+Two possibility distributions $\pi_X, \pi_Y$ are defined for $X$ and $Y$ respectively.
+\item An UTI $\pi_I$ is defined by two possibility distributions, $\pi_S$, $\pi_E$ for the starting  and ending point respectively.
+\end{itemize}
+
+In the following, we will assume that:
+
+\begin{align}
+\pi_X &= \pi_S = \mu_S \\
+\pi_Y &= \pi_E = \mu_E
+\end{align}
+
+Where $\mu_S, \mu_E$ are triangular membership functions.
+Notice that the main diference between an UTI and an IKTI is a matter of interpretation.
+IKTIs  keep both possibility distributions for both starting and ending poins, while UTIs take a t-norm (the minimum) of both possibility distributions (see Eq.
+\eqref{eq:possibility-uti}).
+That is, not all the crisp intervals between the boundaries of the starting and ending point are candidate intervals for the UTI.
+
+\begin{figure}[h]
+   \centering
+   \includegraphics[scale=1.5]{graphs/equivalences.eps}
+   \caption{Possibility degrees for two candidate intervals with respect to the UTI $\pi_I$.
+The upper crisp time interval has a possibility degree of 1, whereas the lower crisp interval has a possibility degree of 0.
+}
+   \label{fig:ikc-greater}
+ \end{figure}
+
+It is possible to obtain the same interpretation for an IKTI, but some ill-known constraints must be defined over the starting and ending ill-known values $X$, $Y$.
+In \cite{Pons2011}, two constraints are defined to determinte whether a crisp time interval $[a, b]$ resides between the boundaries of $J$.
+In this case, all the crisp intervals whose boundaries are within the boundaries of $X$ and $Y$ are considered to be a candidate interval.
+Nevertheless, if we want to obtain the same interpretation as in the UTI, the following two ill-known constraints must be defined:
+
+\begin{align}
+ \label{eq:ikc-ikti-uti}
+C_1 &\triangleq \left(=, X \right) \\
+C_2 &\triangleq \left(=, Y \right)\\
+\Pos \left(\left[a, b\right] \right) &= C_1 \left(a \right) \wedge C_2\left(b \right) =\\
+\label{eq:equivalence-ikti-uti}
+\Pos \left(\left[a, b\right] \right) &= \min \left( \pi_X(a), \pi_Y(b) \right)
+\end{align}
+
+In this case, Eq. \eqref{eq:equivalence-ikti-uti} is equivalent to Eq. \eqref{eq:possibility-uti}.
+In the following subsection, we will study the connection between the Allen's relations in each framework.
+
+\subsection{\label{subsec:eval-allen-relations}Evaluation of the Allen's Relations}
 

### 2013-03-01 trunk/eusflat2013/proposal.tex
updated

@@
-In this section we will analyze first the calculation of all the possible Allen's relations between a crisp time point and an ill-known time point and between a crisp time point and an uncertain time indication (UTI).
+Two frameworks for dealing with uncertainty in time intervals have been presented.
+In this section, we will study the equivalences between an uncertain time interval (UTI) and an ill-known time interval (IKTI).
+Nevertheless, there are some diferences in the semantics of UTIs and IKTIs which are also explained.
+Next to that, the computation of the Allen's Relations are compared.
+
+\subsection{\label{subsec:sim-dif}Similarities and differences}
+A time interval is defined by a starting and an ending point. When uncertainty arises, in one or both starting and / or ending points, then either an UTI $\pi_I$ or an IKTI can be used.
+
+\begin{itemize}
+ \item An IKTI $J = \left[X, Y \right]$ is defined by two ill-known values $X, Y$ which are respectively, the starting and ending points of the time interval.
+Two possibility distributions $\pi_X, \pi_Y$ are defined for $X$ and $Y$ respectively.
+\item An UTI $\pi_I$ is defined by two possibility distributions, $\pi_S$, $\pi_E$ for the starting  and ending point respectively.
+\end{itemize}
+
+In the following, we will assume that:
+
+\begin{align}
+\pi_X &= \pi_S = \mu_S \\
+\pi_Y &= \pi_E = \mu_E
+\end{align}
+
+Where $\mu_S, \mu_E$ are triangular membership functions.
+Notice that the main diference between an UTI and an IKTI is a matter of interpretation.
+IKTIs  keep both possibility distributions for both starting and ending poins, while UTIs take a t-norm (the minimum) of both possibility distributions (see Eq.
+\eqref{eq:possibility-uti}).
+That is, not all the crisp intervals between the boundaries of the starting and ending point are candidate intervals for the UTI.
+
+\begin{figure}[h]
+   \centering
+   \includegraphics[scale=1.5]{graphs/equivalences.eps}
+   \caption{Possibility degrees for two candidate intervals with respect to the UTI $\pi_I$.
+The upper crisp time interval has a possibility degree of 1, whereas the lower crisp interval has a possibility degree of 0.
+}
+   \label{fig:ikc-greater}
+ \end{figure}
+
+It is possible to obtain the same interpretation for an IKTI, but some ill-known constraints must be defined over the starting and ending ill-known values $X$, $Y$.
+In \cite{Pons2011}, two constraints are defined to determinte whether a crisp time interval $[a, b]$ resides between the boundaries of $J$.
+In this case, all the crisp intervals whose boundaries are within the boundaries of $X$ and $Y$ are considered to be a candidate interval.
+Nevertheless, if we want to obtain the same interpretation as in the UTI, the following two ill-known constraints must be defined:
+
+\begin{align}
+ \label{eq:ikc-ikti-uti}
+C_1 &\triangleq \left(=, X \right) \\
+C_2 &\triangleq \left(=, Y \right)\\
+\Pos \left(\left[a, b\right] \right) &= C_1 \left(a \right) \wedge C_2\left(b \right) =\\
+\label{eq:equivalence-ikti-uti}
+\Pos \left(\left[a, b\right] \right) &= \min \left( \pi_X(a), \pi_Y(b) \right)
+\end{align}
+
+In this case, Eq. \eqref{eq:equivalence-ikti-uti} is equivalent to Eq. \eqref{eq:possibility-uti}.
+In the following subsection, we will study the connection between the Allen's relations in each framework.
+
+\subsection{\label{subsec:eval-allen-relations}Evaluation of the Allen's Relations}
+As it has been shown in Table \ref{tab:urz}, when using UTIs in the triangular model, it is possible to obtain a more detailed relational information (there are fifteen possible relations) than with the ill-known constraint approach.
+The following observations should be taken into account when extracting relational information from both frameworks:
+
+\begin{itemize}
+ \item In TM, the Allen's Relations $R \in \left \lbrace B, O, C, D, OB, A \right \rbrace$ between an interval point $J$  and an UTI $\pi_I$ are computed by Eq.
+\eqref{eq:possibility-iz-uti-certain}.
+Therefore, for the Allen's Relations in $R$,
+ possibility that an Allen Relation between an interval point $J$  and an UTI $\pi_I$ is $1$ if the interval point $J$ is in the Allen Relation, or $0$ in other case.
+\item In the TM, the other nine Allen's Relations between an interval point $J$ and an UTI $\pi_I$ are graded from $1$ to $0$ and computed by using the Eq. \eqref{eq:possibility-iz-uti}. 
+\item In the ill-known constraint framework, all the thirdteen Allen's Relations are computed by using an aggregation of ill-known constraints (see Table \ref{tab:allen-relations}).
+Therefore, the possibility that a time point $J$ is in a given Allen Relation is always graded from $1$ to $0$.
+\end{itemize}
+
+For example, consider that we want to know whether the relation Before (B) holds between a time point $J$ and an IKTI $I$.
+We will note $\Pos \left(C_B (J)  \right)$ the possibility that the relation before holds.
+There are three possibilities:
+\begin{enumerate}
+ \item $\Pos \left(C_B (J)  \right)$ = 1: The time point $J$ is before the IKTI $I$.
+\item $\Pos \left(C_B (J)  \right)$ = 0: The time point $J$ is not before the IKTI $I$.
+\item $0 < \Pos \left(C_B (J)  \right) < 1$: The time point $J$ is to some extent before the IKTI $I$.
+\end{enumerate}
+
+The last case corresponds in the TM to case when the before relation is one of the possible relations (See Table \ref{tab:urz}). That is, one of $\left \lbrace PM, PMB, UB \right \rbrace$.
+
+The following table stablishes the correspondences between the Allen's relations in the TM and in the ill-known constraints framework.
+
+\begin{table}[h]
+\centering
+\begin{tabular}{|l|l|l|}
+\hline
+Symbol & Name & Possible Relations \\
+\hline
+B & Before & B, PM, PMB, UB \\
+\multirow{2}{*}
+{O} & Overlaps & O, PM, PS, PFB,\\
+     &          &  PE, PMB, UB \\
+C & Contains & C, PFB, PE, PSB \\
+\multirow{2}{*}
+D & During & D, PS, PE, \\
+  &         &  PMB, UB  \\
+OB & Overlapped by & OB, PE, PSB, PF \\
+A & After & A, PF, UE \\
+M & Meets & M, PM, PMB, UB \\
+MB & Met By & MB, PF, UE \\
+E & Equals & E, PE \\
+\multirow{2}{*}
+S & Starts & S, PS, PE, \\
+  &         &   PMB, UB \\
+SB & Started by & SB, PE, PSB \\
+F & Finishes & F, PE \\
+FB & Finished By & FB, PE, PFB\\
+\hline
+\end{tabular}
+\caption{Correspondences between the Allen's Relations in the ill-known constraint framework and the triangular model,}
+\label{tab:correspondences-tm-ikcf}
+\end{table}
+
+% 
+% In this section we will analyze first the calculation of all the possible Allen's relations between a crisp time point and an ill-known time point and between a crisp time point and an uncertain time indication (UTI).
 Then, we will compare two ill-known time points and two uncertain time indications.
 
-\subsection{\label{subsec:prop-crisp-vs-ill}Calculation of crisp time intervals}
-In this section we first study how uncertainty about wheter crisp interval is enclosed by the ill-known interval given by $[X,Y]$ or by the UTI.
-
-\subsubsection{Crisp time interval enclosed by an ill-known time interval}
-Consider a crisp time interval given by $I=[a, b]$.
+% \subsection{\label{subsec:prop-crisp-vs-ill}Calculation of crisp time intervals}
+% In this section we first study how uncertainty about wheter crisp interval is enclosed by the ill-known interval given by $[X,Y]$ or by the UTI.
+% 
+% \subsubsection{Crisp time interval enclosed by an ill-known time interval}
+% Consider a crisp time interval given by $I=[a, b]$.
 We want to know whether all the points in this interval are within the boundaries of the ill-known interval given by $[X, Y]$.
 The calculation is done by means of two ill-known constraints as explained in \cite{}:
-\begin{eqnarray}
-C_1\stackrel{\triangle}{=}\left(\geq,X\right)\\
-C_2\stackrel{\triangle}{=}\left(\leq,Y\right).
-\end{eqnarray}
-
-The possibility and the necessity are given by:
-\begin{eqnarray}
-\Pos\left(\lambda([a,b])\right)&=&\min\bigg(\Pos(C_1([a,b])),\Pos(C_2([a,b]))\bigg)\\
-\Nec\left(\lambda([a,b])\right)&=&\min\bigg(\Nec(C_1([a,b])),\Nec(C_2([a,b]))\bigg).
-\end{eqnarray}
-Which can be expanded to:
-\begin{eqnarray}
-\label{eq:interval-pos}
-\Pos\left(\lambda([a,b])\right)&=&\min\bigg(\sup_{a\leq w}\pi_{X}(w),\sup_{b\geq w}\pi_{Y}(w)\bigg)\\
-\label{eq:interval-nec}
-\Nec\left(\lambda([a,b])\right)&=&\min\bigg(\inf_{a>w}1-\pi_{X}(w),\inf_{b<w}1-\pi_{Y}(w)\bigg).
-\end{eqnarray}
-
-\subsubsection{\label{subsubsec:crisp-uti}Crisp time interval enclosed by an UTI}
-Consider now a crisp time interval given by $I=[a, b]$ and an Uncertain Time Interval given by $\pi_I = \left(\pi_s, \pi_e \right)$.
+% \begin{eqnarray}
+% C_1\stackrel{\triangle}{=}\left(\geq,X\right)\\
+% C_2\stackrel{\triangle}{=}\left(\leq,Y\right).
+% \end{eqnarray}
+% 
+% The possibility and the necessity are given by:
+% \begin{eqnarray}
+% \Pos\left(\lambda([a,b])\right)&=&\min\bigg(\Pos(C_1([a,b])),\Pos(C_2([a,b]))\bigg)\\
+% \Nec\left(\lambda([a,b])\right)&=&\min\bigg(\Nec(C_1([a,b])),\Nec(C_2([a,b]))\bigg).
+% \end{eqnarray}
+% Which can be expanded to:
+% \begin{eqnarray}
+% \label{eq:interval-pos}
+% \Pos\left(\lambda([a,b])\right)&=&\min\bigg(\sup_{a\leq w}\pi_{X}(w),\sup_{b\geq w}\pi_{Y}(w)\bigg)\\
+% \label{eq:interval-nec}
+% \Nec\left(\lambda([a,b])\right)&=&\min\bigg(\inf_{a>w}1-\pi_{X}(w),\inf_{b<w}1-\pi_{Y}(w)\bigg).
+% \end{eqnarray}
+% 
+% 
+% \subsubsection{\label{subsubsec:crisp-uti}Crisp time interval enclosed by an UTI}
+% Consider now a crisp time interval given by $I=[a, b]$ and an Uncertain Time Interval given by $\pi_I = \left(\pi_s, \pi_e \right)$.
 Again, we want to know if all the points in the interval are within the boundaries of the UTI.
 The calculation is done as explained in \cite{}:
-\begin{equation}
- \pi_I \left( \left[a, b  \right] \right) =
-\begin{cases}
- \min \left(\pi_s(a),\pi_e(b)\right) & \mbox{ if } a \leq b\\
-0 				     & \mbox{ otherwise. }
-\end{cases}
-\end{equation}
+% \begin{equation}
+%  \pi_I \left( \left[a, b  \right] \right) =
+% \begin{cases}
+%  \min \left(\pi_s(a),\pi_e(b)\right) & \mbox{ if } a \leq b\\
+% 0 				     & \mbox{ otherwise. }
+% \end{cases}
+% \end{equation}

### 2013-03-02 trunk/eusflat2013/preliminaries.tex
updated eusflat

@@
-This section introduces the framework of the ill-known contraints.
+This section introduces the framework of the ill-known constraints.
 First of all, some preliminary concepts about possibility theory is explained.
@@
 This is rather usual for historical events.
-For example, the conquest of Granada by the Christians' kings took place on February of 1492, but the exact day is not precissely known.
+For example, the conquest of Granada by the Christians' kings took place on February of 1492, but the exact day is not precisely known.
 A more formal definition of possibilistic variables is given as follows.
@@
 For example, consider an ill-known value $X$ given by a triangular possibility distribution $\left[15, 1, 1\right]$.
-The ill-known contraint for all the points greater than $X$ is given by $C \triangleq (>, X)$.
+The ill-known constraint for all the points greater than $X$ is given by $C \triangleq (>, X)$.
 See Figure \ref{fig:ikc-greater}.
@@
    \includegraphics[scale=1.5]{graphs/gt.eps}
-   \caption{An ill-known value $X = \left[15, 1, 1\right]$ and an ill-known contraint $C \triangleq (>, X)$. }
+   \caption{An ill-known value $X = \left[15, 1, 1\right]$ and an ill-known constraint $C \triangleq (>, X)$. }
    \label{fig:ikc-greater}
@@
 Several ill-known constraints can be aggregated by using the possibilistic extension of boolean operators: and ($\wedge$), or ($\vee$), not ($\neg$).
-The possibility and neccesity for these operators are computed as follows.
+The possibility and necessity for these operators are computed as follows.
 
@@
 
-An ill-known time interval is defined by two ill-known values (see Figure \ref{fig:ill-known-ti}).
-Therefore, both starting and ending points are not precissely known.
-The following is the notation for an ill-known time interval.
+\begin{definition}
+An ill-known time interval \emph{(IKTI)} is defined by two ill-known values (see Figure \ref{fig:ill-known-ti}).
+Therefore, both starting and ending points are not precisely known.
+The following is the notation for an IKTI.
 

### 2013-03-02 trunk/eusflat2013/proposal.tex
updated eusflat

@@
 In this section, we will study the equivalences between an uncertain time interval (UTI) and an ill-known time interval (IKTI).
-Nevertheless, there are some diferences in the semantics of UTIs and IKTIs which are also explained.
+Nevertheless, there are some differences in the semantics of UTIs and IKTIs which are also explained.
 Next to that, the computation of the Allen's Relations are compared.
@@
 Where $\mu_S, \mu_E$ are triangular membership functions.
-Notice that the main diference between an UTI and an IKTI is a matter of interpretation.
+Notice that the main difference between an UTI and an IKTI is a matter of interpretation.
 IKTIs  keep both possibility distributions for both starting and ending poins, while UTIs take a t-norm (the minimum) of both possibility distributions (see Eq.
@@
 It is possible to obtain the same interpretation for an IKTI, but some ill-known constraints must be defined over the starting and ending ill-known values $X$, $Y$.
-In \cite{Pons2011}, two constraints are defined to determinte whether a crisp time interval $[a, b]$ resides between the boundaries of $J$.
+In \cite{Pons2011}, two constraints are defined to determinate whether a crisp time interval $[a, b]$ resides between the boundaries of $J$.
 In this case, all the crisp intervals whose boundaries are within the boundaries of $X$ and $Y$ are considered to be a candidate interval.
@@
 \item In the TM, the other nine Allen's Relations between an interval point $J$ and an UTI $\pi_I$ are graded from $1$ to $0$ and computed by using the Eq. \eqref{eq:possibility-iz-uti}. 
-\item In the ill-known constraint framework, all the thirdteen Allen's Relations are computed by using an aggregation of ill-known constraints (see Table \ref{tab:allen-relations}).
+\item In the ill-known constraint framework, all the thirteen Allen's Relations are computed by using an aggregation of ill-known constraints (see Table \ref{tab:allen-relations}).
 Therefore, the possibility that a time point $J$ is in a given Allen Relation is always graded from $1$ to $0$.
@@
 
-The following table stablishes the correspondences between the Allen's relations in the TM and in the ill-known constraints framework.
+The following table establishes the correspondences between the Allen's relations in the TM and in the ill-known constraints framework.
 
@@
 \end{table}
+
+We will illustrate the calculation of one Allen Relation with an example.
+\begin{example}
+Consider a crisp time interval $J = [1, 3]$ the two ill-known time points $S, E$ that define a time interval. 
+We want to obtain the possibility that $J$ is before the time interval.
+The points are defined by two triangular membership functions $\mu_S$, $\mu_E$.
+
+\begin{align}
+\label{eq:ex-point}
+\pi_S &= \mu_S = \left[2,5,7 \right] \\
+\pi_E &= \mu_E = \left[9,12,15 \right]
+\end{align}
+
+Then, it is possible to define both an IKTI $I$ and an UTI $\pi_I$.
+
+\begin{align}
+\label{eq:ex-uti-ikti}
+I = \left[S, E \right]\\
+\mu_I = \left(\mu_S, \mu_E \right)
+\end{align}
+
+First we will obtain the possibility for the UTI $\pi_I$. The relation possibly meets has three candidates: Before, Meets and Overlaps.
+
+\begin{align}
+  \mu_{\left(J R \pi_I \right)}(B) &= \sup_{a < e} \pi_S(a)\\
+\mu_{\left(J R \pi_I \right)}(M) &= \pi_S(e)\\
+\mu_{\left(J R \pi_I \right)}(O) &= \sup_{a > e} \pi_S(a)
+\end{align}
+
+\begin{align}
+\mu_{\left(J R \pi_I \right)}(B) &= 0.67\\
+\mu_{\left(J R \pi_I \right)}(M) &= 0.33\\
+\mu_{\left(J R \pi_I \right)}(O) &= 0
+\end{align}
+
+In order to compute the possibility  for the before relation between the crisp interval $J$ and the IKTI $I$, a ill-known constraint is needed.
+
+\begin{equation}
+ C_1\stackrel{\triangle}{=} \left(<,X\right)
+\end{equation}
+
+Then the possibility is computed as follows:
+
+\begin{align}
+\Pos \left(C_1 (J) \right) &= \sup_{a < e} \pi_S(a)\\
+\Pos \left(C_1 (J) \right) &= 0.67 
+\end{align}
+
+\end{example}
 

### 2013-03-02 trunk/eusflat2013/triangularmodel.tex
updated eusflat

@@
 This approach has also been used in Figure\ref{fig:allen-relationships} and \ref{fig:ill-known-ti}.
-The dificulty of accurate visual analisys on the temporal data arises when the number of time intervals increase.
+The difficulty of accurate visual analysis on the temporal data arises when the number of time intervals increase.
 
@@
    \includegraphics[width=0.8\columnwidth]{graphs/allencrc.eps}
-   \caption{CRZs corresponding to the thirdteen Allen's relations.
+   \caption{CRZs corresponding to the thirteen Allen's relations.
 The following shorthand notations are used: equals (E), starts (S), started by (SB), finishes (F), finished by (FB), meets (M), met by (MB), overlaps (O), overlapped by (OB), during (D), contains (C), before (B) and after (A).
@@
 \subsection{\label{subsec:relational-info-uti}Obtaining Relational Information}
-In order to make temporal reasoning with UTIs, in the triangular model, the thirteen Allen's Relations, and its respectives Crisp Relational Zones (CRZ) have been generalized.
+In order to make temporal reasoning with UTIs, in the triangular model, the thirteen Allen's Relations, and its respective Crisp Relational Zones (CRZ) have been generalized.
 In this case, we obtain fifteen Uncertain Relational Zones (URZ) , shown in Fig.
@@
 \begin{enumerate}
- \item Draw the lines $L_s, L_e$ that represent the interval point $J$, with the procedure shown un Section \ref{subsec:visualization-tm}. 
+ \item Draw the lines $L_s, L_e$ that represent the interval point $J$, with the procedure shown in Section \ref{subsec:visualization-tm}. 
 \item Draw the orthogonal lines from $L_s, L_e$. 

### 2013-03-04 eusflat2013/proposal.tex
new version for the eusflat2013

@@
 \subsection{\label{subsec:eval-allen-relations}Evaluation of the Allen's Relations}
+As it has been shown in Table \ref{tab:urz}, when using UTIs in the triangular model, it is possible to obtain a more detailed relational information (there are fifteen possible relations) than with the ill-known constraint approach.
+The following observations should be taken into account when extracting relational information from both frameworks:
 
-In this section we will analyze first the calculation of all the possible Allen's relations between a crisp time point and an ill-known time point and between a crisp time point and an uncertain time indication (UTI).
+\begin{itemize}
+ \item In TM, the Allen's Relations $R \in \left \lbrace B, O, C, D, OB, A \right \rbrace$ between an interval point $J$  and an UTI $\pi_I$ are computed by Eq.
+\eqref{eq:possibility-iz-uti-certain}.
+Therefore, for the Allen's Relations in $R$,
+ possibility that an Allen Relation between an interval point $J$  and an UTI $\pi_I$ is $1$ if the interval point $J$ is in the Allen Relation, or $0$ in other case.
+\item In the TM, the other nine Allen's Relations between an interval point $J$ and an UTI $\pi_I$ are graded from $1$ to $0$ and computed by using the Eq. \eqref{eq:possibility-iz-uti}. 
+\item In the ill-known constraint framework, all the thirteen Allen's Relations are computed by using an aggregation of ill-known constraints (see Table \ref{tab:allen-relations}).
+Therefore, the possibility that a time point $J$ is in a given Allen Relation is always graded from $1$ to $0$.
+\end{itemize}
+
+For example, consider that we want to know whether the relation Before (B) holds between a time point $J$ and an IKTI $I$.
+We will note $\Pos \left(C_B (J)  \right)$ the possibility that the relation before holds.
+There are three possibilities:
+\begin{enumerate}
+ \item $\Pos \left(C_B (J)  \right)$ = 1: The time point $J$ is before the IKTI $I$.
+\item $\Pos \left(C_B (J)  \right)$ = 0: The time point $J$ is not before the IKTI $I$.
+\item $0 < \Pos \left(C_B (J)  \right) < 1$: The time point $J$ is to some extent before the IKTI $I$.
+\end{enumerate}
+
+The last case corresponds in the TM to case when the before relation is one of the possible relations (See Table \ref{tab:urz}). That is, one of $\left \lbrace PM, PMB, UB \right \rbrace$.
+
+The following table establishes the correspondences between the Allen's relations in the TM and in the ill-known constraints framework.
+
+\begin{table}[h]
+\centering
+\begin{tabular}{|l|l|l|}
+\hline
+Symbol & Name & Possible Relations \\
+\hline
+B & Before & B, PM, PMB, UB \\
+\multirow{2}{*}
+{O} & Overlaps & O, PM, PS, PFB,\\
+     &          &  PE, PMB, UB \\
+C & Contains & C, PFB, PE, PSB \\
+\multirow{2}{*}
+D & During & D, PS, PE, \\
+  &         &  PMB, UB  \\
+OB & Overlapped by & OB, PE, PSB, PF \\
+A & After & A, PF, UE \\
+M & Meets & M, PM, PMB, UB \\
+MB & Met By & MB, PF, UE \\
+E & Equals & E, PE \\
+\multirow{2}{*}
+S & Starts & S, PS, PE, \\
+  &         &   PMB, UB \\
+SB & Started by & SB, PE, PSB \\
+F & Finishes & F, PE \\
+FB & Finished By & FB, PE, PFB\\
+\hline
+\end{tabular}
+\caption{Correspondences between the Allen's Relations in the ill-known constraint framework and the triangular model,}
+\label{tab:correspondences-tm-ikcf}
+\end{table}
+
+We will illustrate the calculation of one Allen Relation with an example.
+\begin{example}
+Consider a crisp time interval $J = [1, 3]$ the two ill-known time points $S, E$ that define a time interval. 
+We want to obtain the possibility that $J$ is before the time interval.
+The points are defined by two triangular membership functions $\mu_S$, $\mu_E$.
+
+\begin{align}
+\label{eq:ex-point}
+\pi_S &= \mu_S = \left[2,5,7 \right] \\
+\pi_E &= \mu_E = \left[9,12,15 \right]
+\end{align}
+
+Then, it is possible to define both an IKTI $I$ Eq. \eqref{eq:ex-uti-ikti} and an UTI $\pi_I$ Eq. \eqref{eq:ex-uti-uti}. The intervals are illustrated in Figure \ref{fig:example}.
+
+\begin{figure}[h]
+   \centering
+   \includegraphics[scale=1.5]{graphs/example.eps}
+   \caption{In the top, the crisp time interval $J$ and the IKTI given by $I = [S, E]$.
+The ill-known constraint for the before relation is illustrated in grey color.
+In the lower graphic, the interval point $J$ and the UTI $\pi_I$.
+}
+   \label{fig:example}
+ \end{figure}
+
+\begin{align}
+\label{eq:ex-uti-ikti}
+I = \left[S, E \right]\\
+\label{eq:ex-uti-uti}
+\mu_I = \left(\mu_S, \mu_E \right)
+\end{align}
+
+First we will obtain the possibility between $J$ and the UTI $\pi_I$. The relation possibly meets has three candidates: Before, Meets and Overlaps.
+
+\begin{align}
+  \mu_{\left(J R \pi_I \right)}(B) &= \sup_{a < e} \pi_S(a)\\
+\mu_{\left(J R \pi_I \right)}(M) &= \pi_S(e)\\
+\mu_{\left(J R \pi_I \right)}(O) &= \sup_{a > e} \pi_S(a)
+\end{align}
+
+\begin{align}
+\mu_{\left(J R \pi_I \right)}(B) &= 0.67\\
+\mu_{\left(J R \pi_I \right)}(M) &= 0.33\\
+\mu_{\left(J R \pi_I \right)}(O) &= 0
+\end{align}
+
+In order to compute the possibility  for the before relation between the crisp interval $J$ and the IKTI $I$, a ill-known constraint is needed.
+
+\begin{equation}
+ C_1\stackrel{\triangle}{=} \left(<,X\right)
+\end{equation}
+
+Then the possibility is computed as follows:
+
+\begin{align}
+\Pos \left(C_1 (J) \right) &= \sup_{a < e} \pi_S(a)\\
+\Pos \left(C_1 (J) \right) &= 0.67 
+\end{align}
+
+\end{example}
+
+% 
+% In this section we will analyze first the calculation of all the possible Allen's relations between a crisp time point and an ill-known time point and between a crisp time point and an uncertain time indication (UTI).
 Then, we will compare two ill-known time points and two uncertain time indications.

### 2013-03-06 trunk/eusflat2013/introduction.tex
introduction, abstract and conclusions

@@
+The concept of the time has been widely studied \cite{Benthem1982}, \cite{Shackle1961}, \cite{Klein1994}.
+Moreover, humans beings manage temporal indications in an imprecise way \cite{Devos1998}.
+But when dealing with time in an Information System, some simplifications must to be done.
+The first thing to do is the discretization of the time line into time points or time intervals.
+Both approaches have been proved to be equivalent, although due to the nature of the smallest unit of time in a computer, a \emph{chronon}, the discretization of a time point returns a time interval.
+
+All the possible positions between two time intervals were studied by Allen \cite{Allen1983}, \cite{Allen1985}.
+As result, the thirteen Allen's Relations were obtained (See Table \ref{tab:allen-relations}) which are illustrated in Figure \ref{fig:allen-relationships}.
+In order to achieve a more intelligent processing of time, some theoretical frameworks are used to reason with time.
+First the possibility theory was employed in the reasoning with temporal information \cite{Dubois2003a}.
+Then, several proposals \cite{Schockaert2008}, \cite{nagypal2003}, \cite{Ohlbach2004}, \cite{Pons2011} studied how to extend the thirteen Allen's relations to the possibilistic case.
+Also rough set theory \cite{Pawlak1995} has been used to represent and reason about imperfect time intervals \cite{Qiang2009}, \cite{Qiang2010}.
+
+Although there exists several proposals to represent and visualize imperfect time intervals, in this paper we will focus on two: the ill-known constraint framework \cite{Pons2011} and the triangular model \cite{DeTre2012}.
+The first one is a theoretical framework that deals with temporal reasoning and the second one is a visual framework that represent imperfect time intervals in a two dimensional space.
+Both frameworks can be used to represent imperfect time intervals as well as temporal reasoning.
+
+The structure of this paper is the following.
+Section  \ref{sec:preliminaries} introduces the Ill-known constraint framework.
+Section \ref{sec:triangular-model} introduces the triangular model.
+Section \ref{sec:proposal} analyze both frameworks and shows the correspondences and differences between both frameworks.
+The section concludes with an example showing that the calculations made in both frameworks are equivalent.
+Finally, Section \ref{sec:conclusions} presents the main conclusions and the future work.
+
 \begin{table}[h]

### 2013-03-06 trunk/eusflat2013/proposal.tex
introduction, abstract and conclusions

@@
 Notice that the main difference between an UTI and an IKTI is a matter of interpretation.
-IKTIs  keep both possibility distributions for both starting and ending poins, while UTIs take a t-norm (the minimum) of both possibility distributions (see Eq.
+IKTIs  keep both possibility distributions for both starting and ending points, while UTIs take a t-norm (the minimum) of both possibility distributions (see Eq.
 \eqref{eq:possibility-uti}).
@@
 
-Then, it is possible to define both an IKTI $I$ and an UTI $\pi_I$.
+Then, it is possible to define both an IKTI $I$ Eq. \eqref{eq:ex-uti-ikti} and an UTI $\pi_I$ Eq. \eqref{eq:ex-uti-uti}. The intervals are illustrated in Figure \ref{fig:example}.
+
+\begin{figure}[h]
+   \centering
+   \includegraphics[scale=1.5]{graphs/example.eps}
+   \caption{In the top, the crisp time interval $J$ and the IKTI given by $I = [S, E]$.
+The ill-known constraint for the before relation is illustrated in grey color.
+In the lower graphic, the interval point $J$ and the UTI $\pi_I$.
+}
+   \label{fig:example}
+ \end{figure}
 
@@
 I = \left[S, E \right]\\
+\label{eq:ex-uti-uti}
 \mu_I = \left(\mu_S, \mu_E \right)
@@
 
-First we will obtain the possibility for the UTI $\pi_I$. The relation possibly meets has three candidates: Before, Meets and Overlaps.
+First we will obtain the possibility between $J$ and the UTI $\pi_I$. The relation possibly meets has three candidates: Before, Meets and Overlaps.
 

### 2013-03-06 eusflat2013/conclusions.tex
introduction, abstract and conclusions

In this work we have presented an analysis between two different frameworks that deal and represent imperfect temporal information. Besides the differences in the visualization of imperfect time intervals, we obtained that the temporal reasoning in both frameworks is equivalent. It is also important to notice that whereas the ill-known constraint framework deal with the thirteen Allen's relations, the triangular model distinguishes fifteen possible relational zones and therefore, it is necessary to set a correspondence between the Allen's relations in both frameworks.

Future research will deal with the analysis between two imperfect time points in each framework and with the representation of open time intervals.

### 2013-03-09 eusflat2013/introduction.tex
Some slight changes. I wrote an email about them...

%1. Introduction
Information systems (IS) in general and database systems (DS) in particular have always tried to model parts of reality. To achieve this modelling, such systems contain data representing properties of real-world objects or concepts. As time is an essential aspect of many of these, information systems often contain data representing temporal notions which describe their temporal properties. Such temporal notions usually take the form of either instants (time points)~\cite{Dyreson1994},~\cite{Jensen1998} or time intervals~\cite{Dyreson1994},~\cite{Jensen1998}.
%TODO: more references in the above part?

Data are often produced by humans, but human-made data are prone to imperfections: some data may be vague, imprecise, incomplete, contradictory or uncertain. Data representing temporal notions may contain such uncertainties too.

Generally, one of the most important purposes of an IS is to allow the retrieval of information or knowledge deduced from its data. Information or knowledge is usually retrieved from an IS by querying its contents and examining the query results or by visualizing its contents, querying this visualization and examining these query results.
%TODO: include some form of proof or previous work with these statements!

Of course, when temporal information is contained in an IS, it can be used in the querying process. Usually, querying an IS containing data representing temporal notions is done by specifying one or more time indications and requesting information that is related to these notions, where the nature and semantics of these relationships are specifically temporal. Thus, specific temporal relationships are necessary. Notably, Allen~\cite{Allen1983} presented a reasoning framework containing all semantically usefull basic temporal relationships between time intervals (and as a special case instants). The resulting basic Allen relationships are shown in figure \ref{fig:allen-relationships}.

To be able to query IS and DS containing uncertain data representing uncertain time indications, a framework is necessary, able to represent uncertainties in time indications without (much) information loss and able to temporally reason with such uncertain time indications. Although many proposals for such frameworks exist, the work presented in this paper focusses on two: the \emph{ill-known constraint \emph{(IKC)} framework}~\cite{Pons2011} and the \emph{triangular model \emph{(TM)} framework}~\cite{DeTre2012}.

The work presented in this paper consists of a comparison of the approaches, used by both frameworks, to representing time intervals subject to uncertainty and to reasoning about the temporal relationships between such intervals and perfectly defined time intervals.

The structure of this paper is as follows: Section \ref{sec:general-preliminaries} presents some general preliminaries and naming and notation conventions used in this paper. Section \ref{sec:ikc} introduces the IKC framework and Section \ref{sec:tm} introduces the TM framework. In Section \ref{sec:proposal}, both frameworks are compared. Finally, Section \ref{sec:conclusions} presents the principal conclusions of the presented work and some possible future work.

%The concept of the time has been widely studied \cite{Benthem1982}, \cite{Shackle1961}, \cite{Klein1994}. Moreover, humans beings manage temporal indications in an imprecise way \cite{Devos1998}. But when dealing with time in an Information System, some simplifications must to be done. The first thing to do is the discretization of the time line into time points or time intervals. Both approaches have been proved to be equivalent, although due to the nature of the smallest unit of time in a computer, a \emph{chronon}, the discretization of a time point returns a time interval.

%All the possible positions between two time intervals were studied by Allen \cite{Allen1983}, \cite{Allen1985}. As result, the thirteen Allen's Relations were obtained (See Table \ref{tab:allen-relations}) which are illustrated in Figure \ref{fig:allen-relationships}. In order to achieve a more intelligent processing of time, some theoretical frameworks are used to reason with time. First the possibility theory was employed in the reasoning with temporal information \cite{Dubois2003a}. Then, several proposals \cite{Schockaert2008}, \cite{nagypal2003}, \cite{Ohlbach2004}, \cite{Pons2011} studied how to extend the thirteen Allen's relations to the possibilistic case. Also rough set theory \cite{Pawlak1995} has been used to represent and reason about imperfect time intervals \cite{Qiang2009}, \cite{Qiang2010}.

%Although there exists several proposals to represent and visualize imperfect time intervals, in this paper we will focus on two: the ill-known constraint framework \cite{Pons2011} and the triangular model \cite{DeTre2012}. The first one is a theoretical framework that deals with temporal reasoning and the second one is a visual framework that represent imperfect time intervals in a two dimensional space. Both frameworks can be used to represent imperfect time intervals as well as temporal reasoning.

%The structure of this paper is the following. Section  \ref{sec:preliminaries} introduces the Ill-known constraint framework. Section \ref{sec:triangular-model} introduces the triangular model. Section \ref{sec:proposal} analyze both frameworks and shows the correspondences and differences between both frameworks. The section concludes with an example showing that the calculations made in both frameworks are equivalent. Finally, Section \ref{sec:conclusions} presents the main conclusions and the future work.

%\begin{table}[h]
%\centering
%\begin{tabular}{|c|l|}
%\hline
%Name & Implementation \\ \hline 
%$I$ equals $J$ & if $s_i = s_j \wedge e_i = e_j $ \\
%$I$ starts $J$ & if $s_i = s_j \wedge e_i < e_j $ \\
%$I$ started by $J$ & if $s_i = s_j \wedge e_i > e_j $ \\
%$I$ finishes $J$ & if $s_i > s_j \wedge e_i = e_j $ \\
%$I$ finished by $J$ & if $s_i < s_j \wedge e_i = e_j $ \\
%$I$ meets $J$ & if $e_i = s_j $ \\
%$I$ met by $J$ & if $s_i = e_j $ \\
%$I$ overlaps $J$ & if $s_i < s_j \wedge e_i < e_j \wedge e_i > s_j $ \\
%$I$ overlapped by $J$ & if $s_i > s_j \wedge e_j < e_i \wedge s_i < e_j  $ \\
%$I$ during $J$ & if $s_i > s_j \wedge e_i < e_j $ \\
%$I$ contains $J$ & if $  s_i < s_j \wedge e_i > e_j$ \\
%$I$ before $J$ & if $e_i < s_j $ \\
%$I$ after $J$ & if $s_i > e_j $ \\
%\hline
%\end{tabular}
%\caption{Allen's relations represented in the framework. $I = \left[s_i, e_i\right]$, $J=  \left[s_j, e_j\right]$}
%\label{tab:allen-relations}
%\end{table}

\begin{figure}[h]
   \centering
   \includegraphics[width=0.8\columnwidth]{graphs/allen.eps}
   \caption{The Allen relationships between two crisp intervals $I$ and $J$.  }
   \label{fig:allen-relationships}
 \end{figure}

### 2013-03-09 trunk/eusflat2013/general_preliminaries.tex
Some slight changes. I wrote an email about them...

%2. General Preliminaries
As time is often thought of as following an axis (a time axis), it is often (as in the presented work) seen as an ordered set of time points, called \emph{instants}~\cite{Dyreson1994},~\cite{Jensen1998}. In this paper, instants will be noted using lowercase letters, whereas time intervals will be noted using uppercase letters. A \emph{time interval} $I$ is then an interval subset of this ordered set of instants. In this paper, such a time interval starting at (and containing) instant $s$ and ending at (and containing) instant $e$ is noted $I = \left[s, e\right]$.

The presented work allows time intervals to be subject to uncertainties. These uncertainties are always seen as caused by a (partial) lack of knowledge: the exact time interval is not known, eventhough there is only one correct time interval and as such no variability. Confidence about which interval is the correct one in the context of such uncertainties is modelled using possibility theory~\cite{DidierDubois1988a}. Possibility is always interpreted as plausibility given a (partial) lack of knowledge in the presented work.

To be able to distinguish easily, time intervals not subject to uncertainties will be called \emph{crisp time intervals} in the presented work.

### 2013-03-11 eusflat2013/introduction.tex
I updated the part about the IKC. I'll send an email explaining my choices.

@@
 Notably, Allen~\cite{Allen1983} presented a reasoning framework containing all semantically usefull basic temporal relationships between time intervals (and as a special case instants).
-The resulting basic Allen relationships are shown in figure \ref{fig:allen-relationships}.
+The resulting Allen relationships are shown in figure \ref{fig:allen-relationships}.
 
@@
 
-The structure of this paper is as follows: Section \ref{sec:general-preliminaries} presents some general preliminaries and naming and notation conventions used in this paper.
-Section \ref{sec:ikc} introduces the IKC framework and Section \ref{sec:tm} introduces the TM framework.
-In Section \ref{sec:proposal}, both frameworks are compared.
-Finally, Section \ref{sec:conclusions} presents the principal conclusions of the presented work and some possible future work.
+The structure of this paper is as follows: section \ref{sec:general-preliminaries} presents some general preliminaries and naming and notation conventions used in this paper.
+Sections \ref{sec:ikc} and \ref{sec:tm} introduce the IKC and TM frameworks respectively: both sections first introduce some preliminary concepts and techniques specific to the framework under consideration, then explain how the representation of time intervals is done and finally show how the evaluation of Allen relationships between a crisp interval and one subject to uncertainty is done.
+In section \ref{sec:proposal}, both frameworks are compared: first their approaches to representing time intervals are compared, next their approaches to evaluating Allen relationships are compared.
+Finally, section \ref{sec:conclusions} presents the principal conclusions of the presented work and some possible future work.
 

### 2013-03-12 trunk/eusflat2013/tm.tex
An update, mainly about the TM part.

--- 2013-03-02 trunk/eusflat2013/triangularmodel.tex
@@
-In this section we will explain a graphical visualization of temporal data called the triangular model (TM).
-This has been widely used mainly on rough sets for the visualization of historical data.
-More recently, the model has been extended on possibility theory to model Uncertain Time Intervals (UTI).
-First of all we will explain the main concepts and properties of the triangular model to model Uncertain Time Intervals (UTI).
-Then, we will detail how to compute the Allen's Relations between a crisp time point and an UTI.
+In~\cite{Weghe2007}, the triangular model framework is introduced, based on~\cite{Kulpa1997}.
+In~\cite{DeTre2012}, the framework is generalized to allow the representation of and reasoning with time intervals subject to uncertainty.
+In this section, this generalization's main aspects are briefly presented.
 
-\subsection{\label{subsec:utis}Uncertain Time Intervals}
-As mentioned in the previous section, the boundaries of time intervals are not always well-known.
-The available knowledge about the interval can be modelled by using possibility distributions ~\cite{Billiet2012}, ~\cite{JoseEnriquePons2012} on both starting and ending points as mentioned in the previous section.
-Informally, an Uncertain Time Interval UTI is a time interval for which one or both boundaries are not precisely known.
+\subsection{\label{subsec:tm-preliminaries}Preliminary Concepts}
+Usually, different time intervals are visualized as different parallel line segments in the image plane.
+However, as this linear approach introduces a few difficulties, the TM framework was conceived to visualize time intervals as points in the image plane.
 
-\begin{definition}
- \label{def:uncertain-time-interval}
-An uncertain time interval UTI $I$ is composed by an starting point $S$, and an ending point $E$.
-For both points, two different convex possibility distributions are defined, namely $\pi_s$, $\pi_e$.
-These two possibility distributions represent all the possible time intervals that can be constructed from $I$.
-The following formula obtains the possibility $\pi_{I}$ that a crisp interval  given by  $[a, b]$ is constructed from $I$.
-\begin{equation}
- \label{eq:possibility-uti}
-\pi_{I} \left(\left[t_s, t_e\right] \right) = 
-\begin{cases}
- \min \left(\pi_s (t_s), \pi_e (t_e) \right), & \mbox{ if } t_s \leq t_e\\
-0, & \mbox{ otherwise. }
-\end{cases}
-\end{equation}
+To achieve this, on the image plane, a horizontal line segment visualizing the used (part of the) time axis is drawn.
+Then, a triangle is drawn, using this line segment as a side of which the angles with the other two sides have sizes $\alpha$ and $-\alpha$ respectively.
+The area contained in this triangle is called the \emph{interval space} and will contain all visualizations of intervals.
+Now, the procedure used to visualize a time interval $\left[s, e\right]$ in the interval space is the following:
 
-In the following we will note an UTI as $\pi_{I} = \left(\pi_s, \pi_e \right)$.
-% It is important to notice that, the main diference with respect to the ill-known time interval (see Definition \ref{}) is that an UTI only consider as possible starting and ending points
+\begin{enumerate}
+	\item Visualize $s$ as a point on the visualization of the time axis.
+	\item Construct a straight half-line $L_s$ from this point, with an angle of size $\alpha$ with the visualization of the time axis.
+	\item Visualize $e$ as a point on the visualization of the time axis.
+	\item Construct a straight half-line $L_e$ from this point, with an angle of size $-\alpha$ with the visualization of the time axis.
+	\item The intersection point of $L_s$ and $L_e$ is the visualization of interval $\left[s, e\right]$.
+\end{enumerate}
 
-\end{definition}
+The resulting point visualizing an interval is generally called an \emph{interval point}.
+In this work, as it is usually done, the size of $\alpha$ is chosen to be $45^{\circ}$.
+An example of such construction is shown in figure \ref{fig:tm-const-ex}.
 
-\subsection{\label{subsec:visualization-tm}Visualization within the triangular model}
-Time intervals are usually represented in a one dimensional scale.
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.8\columnwidth]{graphs/TM_model_several.eps}
+	\caption{The visualization of several time intervals using the TM framework.}
+	\label{fig:tm-const-ex}
+\end{figure}
+
+%Time intervals are usually represented in a one dimensional scale.
 This is due to its linear characteristic.
@@
 
-The triangular model (TM) ~\cite{Kulpa1997}, \cite{Weghe2007} represents time intervals in a two dimensional space called IRterval point completely determines the start and end of the interval.
+%The triangular model (TM) ~\cite{Kulpa1997}, \cite{Weghe2007} represents time intervals in a two dimensional space called IRterval point completely determines the start and end of the interval.
 The two-dimensional space of interval points is called the interval space and is denoted by IR ~\cite{Kulpa2006}.
@@
 
-\begin{enumerate}
-\item Draw each time point of the interval $[t_s, t_e]$ in the horizontal line.
-\item Construct one straight line $L_s$ from $t_s$ with an angle $\alpha$.
-\item Construct one straight line $L_e$ from $t_e$ with an angle $-\alpha$.
-\item The intersection of these two lines is the interval point.
-\end{enumerate}
+%\begin{enumerate}
+%\item Draw each time point of the interval $[t_s, t_e]$ in the horizontal line.
+%\item Construct one straight line $L_s$ from $t_s$ with an angle $\alpha$.
+%\item Construct one straight line $L_e$ from $t_e$ with an angle $-\alpha$.
+%\item The intersection of these two lines is the interval point.
+%\end{enumerate}
 
-For the shake of simplicity, in this work, the angle for $\alpha$ is chosen to be $45^{\circ}$.
+%For the shake of simplicity, in this work, the angle for $\alpha$ is chosen to be $45^{\circ}$.
 
-In TM, these thirteen Allen relations each correspond to a specific zone in the interval space IR \cite{Kulpa1997}.
+%In TM, these thirteen Allen relations each correspond to a specific zone in the interval space IR \cite{Kulpa1997}.
 These thirteen zones are called Crisp Relational Zones (CRZs).
@@
 
-\begin{figure}[h]
-   \centering
-   \includegraphics[width=0.8\columnwidth]{graphs/allencrc.eps}
-   \caption{CRZs corresponding to the thirteen Allen's relations.
+%\begin{figure}[h]
+%   \centering
+%   \includegraphics[width=0.8\columnwidth]{graphs/allencrc.eps}
+%   \caption{CRZs corresponding to the thirteen Allen's relations.
 The following shorthand notations are used: equals (E), starts (S), started by (SB), finishes (F), finished by (FB), meets (M), met by (MB), overlaps (O), overlapped by (OB), during (D), contains (C), before (B) and after (A).
 }
-   \label{fig:allencrz}
- \end{figure}
+%   \label{fig:allencrz}
+% \end{figure}
 
-Each CRZ represents the set of interval points of the crisp time intervals that are in the corresponding Allen relation with respect to the reference interval $I$.
+%Each CRZ represents the set of interval points of the crisp time intervals that are in the corresponding Allen relation with respect to the reference interval $I$.
 For example, all interval points of intervals that come before $I$ are located in the CRZ represented by the leftmost lower triangle in the interval space.
@@
 
-\subsection{\label{subsec:utis-in-tm}Representation of Uncertain Time Intervals}
-In \cite{DeTre2012} a technique to represent UTIs in the triangular model has been presented.
+\subsection{\label{subsec:tm-interval}Representation of Uncertainty in Intervals}
+The TM framework allows uncertainty in time intervals by supporting uncertain time intervals:
+
+\begin{definition}
+Given an ordered set of instants $T$, an \emph{uncertain time interval} (UTI) $J$ in $T$ is here a time interval defined by a pair
+\begin{align}
+J \triangleq (\pi_s, \pi_e)
+\end{align}
+where $\pi_s$ and $\pi_e$ are two convex possibility distributions on $T$.
+\end{definition}
+
+For a UTI $J$ defined by $(\pi_s, \pi_e)$ with $\pi_s$ and $\pi_e$ two possibility distributions on an ordered set of instants $T$, $\pi_s$ and $\pi_e$ define the confidence that an instant in $T$ is the starting instant, respectively the ending instant of $J$.
+Because for such UTI $J$, $\pi_s$ and $\pi_e$ are required to be convex, the possibility $\pi_J(I)$ that a given crisp time interval $I = \left[s_i, e_i\right]$ is the exact time interval represented by $J$ can be uniquely determined by:
+
+\begin{align}
+\pi_J(I) = \begin{cases}
+ \min \left(\pi_s (s_i), \pi_e (e_i) \right), & \mbox{ if } s_i \leq e_i\\
+0, & \mbox{ otherwise }
+\end{cases}
+\end{align}
+
+An UTI $J$ defined by a pair $(\pi_s, \pi_e)$ is now visualized as an area in the image plane, using a single-color greyscale.
+Every crisp time interval $I$ for which the possibility $\pi_J(I)$ of $I$ being the time interval represented by $J$ is higher than zero, is considered.
+For every such $I$, the interval point is drawn and is given the greyscale color which reflects the possibility that $I$ is the time interval represented by $J$.
+For this, a single-color greyscale is used where the saturation of the color is linearly related to the possibility.
+An example of such construction is shown in figure \ref{fig:tm-ill-const-ex}.
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.8\columnwidth]{graphs/TM_model_ill_known.eps}
+	\caption{The visualization of a UTI defined by a pair $(\pi_s, \pi_e)$ using the TM framework.}
+	\label{fig:tm-ill-const-ex}
+\end{figure}
+
+\subsection{\label{subsec:tm-evaluation}Evaluation of Allen Relationships}
+
+%In this section we will explain a graphical visualization of temporal data called the triangular model (TM).
+This has been widely used mainly on rough sets for the visualization of historical data.
+More recently, the model has been extended on possibility theory to model Uncertain Time Intervals (UTI).
+%First of all we will explain the main concepts and properties of the triangular model to model Uncertain Time Intervals (UTI).
+Then, we will detail how to compute the Allen's Relations between a crisp time point and an UTI.
+
+%\subsection{\label{subsec:utis}Uncertain Time Intervals}
+%As mentioned in the previous section, the boundaries of time intervals are not always well-known.
+The available knowledge about the interval can be modelled by using possibility distributions ~\cite{Billiet2012}, ~\cite{JoseEnriquePons2012} on both starting and ending points as mentioned in the previous section.
+Informally, an Uncertain Time Interval UTI is a time interval for which one or both boundaries are not precisely known.
+
+%\begin{definition}
+% \label{def:uncertain-time-interval}
+%An uncertain time interval UTI $I$ is composed by an starting point $S$, and an ending point $E$.
+For both points, two different convex possibility distributions are defined, namely $\pi_s$, $\pi_e$.
+These two possibility distributions represent all the possible time intervals that can be constructed from $I$.
+The following formula obtains the possibility $\pi_{I}$ that a crisp interval  given by  $[a, b]$ is constructed from $I$.
+%\begin{equation}
+% \label{eq:possibility-uti}
+%\pi_{I} \left(\left[t_s, t_e\right] \right) = 
+%\begin{cases}
+% \min \left(\pi_s (t_s), \pi_e (t_e) \right), & \mbox{ if } t_s \leq t_e\\
+%0, & \mbox{ otherwise. }
+%\end{cases}
+%\end{equation}
+
+%In the following we will note an UTI as $\pi_{I} = \left(\pi_s, \pi_e \right)$.
+% It is important to notice that, the main diference with respect to the ill-known time interval (see Definition \ref{}) is that an UTI only consider as possible starting and ending points
+
+%\end{definition}
+
+%\subsection{\label{subsec:visualization-tm}Visualization within the triangular model}
+
+%\subsection{\label{subsec:utis-in-tm}Representation of Uncertain Time Intervals}
+%In \cite{DeTre2012} a technique to represent UTIs in the triangular model has been presented.
 Basically, the procedure is an extension of the method to represent crisp time intervals in the triangular space.
@@
 
-\begin{enumerate}
- \item Let $\pi_s = \left[D_s, a_s, b_s\right]$ a triangular membership function.
-\item Let $\pi_e = \left[D_e, a_e, b_e \right]$ a triangular membership function.
-\item Let $t_s^{-} = D_s - a_s$, and $t_s^{+} = D_s + b_s$.
-\item Let $t_e^{-} = D_e -a_e$, and $t_e^{+} = D_e + b_e$.
-\item Construct two straight lines $L_s^{-}, L_s^{-}$ from $t_s^{-}, t_s^{+}$ with an angle $\alpha$.
-\item Construct two straight lines $L_e^{-}, L_e^{-}$ from $t_e^{-}, t_e^{+}$ with an angle $-\alpha$.
-\end{enumerate}
+%\begin{enumerate}
+% \item Let $\pi_s = \left[D_s, a_s, b_s\right]$ a triangular membership function.
+%\item Let $\pi_e = \left[D_e, a_e, b_e \right]$ a triangular membership function.
+%\item Let $t_s^{-} = D_s - a_s$, and $t_s^{+} = D_s + b_s$.
+%\item Let $t_e^{-} = D_e -a_e$, and $t_e^{+} = D_e + b_e$.
+%\item Construct two straight lines $L_s^{-}, L_s^{-}$ from $t_s^{-}, t_s^{+}$ with an angle $\alpha$.
+%\item Construct two straight lines $L_e^{-}, L_e^{-}$ from $t_e^{-}, t_e^{+}$ with an angle $-\alpha$.
+%\end{enumerate}
 
-The intersection of these lines is a quadrangle zone where the UTI is contained.
+%The intersection of these lines is a quadrangle zone where the UTI is contained.
 Note that not all the interval points in this zone are to the same extent representing candidate intervals for the UTI.
@@
 
-\begin{figure}[h]
-   \centering
-   \includegraphics[width=0.8\columnwidth]{graphs/utitm.eps}
-   \caption{Representation in the triangular model for an UTI $\pi_I$.  }
-   \label{fig:utitm}
- \end{figure}
+%\begin{figure}[h]
+%   \centering
+%   \includegraphics[width=0.8\columnwidth]{graphs/utitm.eps}
+%   \caption{Representation in the triangular model for an UTI $\pi_I$.  }
+%   \label{fig:utitm}
+% \end{figure}
 

### 2013-03-14 develop/excel2sql/src/main/java/es/ugr/decsai/excel2sql/ExcelReader.java
First version of the excel 2 sql program to translate excel database into sql database.

@@ -1,0 +1,151 @@
+/*
+ * To change this template, choose Tools | Templates
+ * and open the template in the editor.
+ */
+package es.ugr.decsai.excel2sql;
+
+import java.io.FileInputStream;
+import java.io.FileNotFoundException;
+import java.io.IOException;
+import java.io.InputStream;
+import java.util.ArrayList;
+import java.util.Iterator;
+import java.util.List;
+import org.apache.poi.openxml4j.exceptions.InvalidFormatException;
+import org.apache.poi.ss.usermodel.Cell;
+import org.apache.poi.ss.usermodel.Row;
+import org.apache.poi.ss.usermodel.Sheet;
+import org.apache.poi.ss.usermodel.Workbook;
+import org.apache.poi.ss.usermodel.WorkbookFactory;
+
+/**
+ * class to read an excel file
+ *
+ * @author Jose
+ */
+public class ExcelReader {
+
+    Workbook wb;
+    private int MAX_ITER = 10;
+
+    /**
+     * Open an excel file
+     *
+     * @param fileName
+     * @throws FileNotFoundException
+     * @throws IOException
+     * @throws InvalidFormatException
+     */
+    public ExcelReader(String fileName) throws FileNotFoundException, IOException, InvalidFormatException {
+
+        InputStream is = new FileInputStream(fileName);
+        wb = WorkbookFactory.create(is);
+
+    }
+
+    public void ShowFields() {
+
+        Sheet sheet = wb.getSheetAt(0);
+        // no iterations at row level:
+        Row row = sheet.getRow(0);
+//        Cell cell = row.getCell(0);
+        Iterator<Cell> cellIterator = row.cellIterator();
+
+        while (cellIterator.hasNext()) {
+            Cell cell = cellIterator.next();
+            switch (cell.getCellType()) {
+                case Cell.CELL_TYPE_STRING:
+                    System.out.println(cell.getRichStringCellValue().getString());
+                    break;
+                case Cell.CELL_TYPE_NUMERIC:
+//                    if (DateUtil.isCellDateFormatted(cell)) {
+//                        System.out.println(cell.getDateCellValue());
+//                    } else {
+                    System.out.println(cell.getNumericCellValue());
+//                    }
+                    break;
+                case Cell.CELL_TYPE_BOOLEAN:
+                    System.out.println(cell.getBooleanCellValue());
+                    break;
+                case Cell.CELL_TYPE_FORMULA:
+                    System.out.println(cell.getCellFormula());
[…]

### 2013-03-14 develop/excel2sql/src/test/java/es/ugr/decsai/excel2sql/AppTest.java

@@ -1,0 +1,38 @@
+package es.ugr.decsai.excel2sql;
+
+import junit.framework.Test;
+import junit.framework.TestCase;
+import junit.framework.TestSuite;
+
+/**
+ * Unit test for simple App.
+ */
+public class AppTest 
+    extends TestCase
+{
+    /**
+     * Create the test case
+     *
+     * @param testName name of the test case
+     */
+    public AppTest( String testName )
+    {
+        super( testName );
+    }
+
+    /**
+     * @return the suite of tests being tested
+     */
+    public static Test suite()
+    {
+        return new TestSuite( AppTest.class );
+    }
+
+    /**
+     * Rigourous Test :-)
+     */
+    public void testApp()
+    {
+        assertTrue( true );
+    }
+}

### 2013-03-14 trunk/develop/excel2sql/src/main/java/es/ugr/decsai/excel2sql/App.java
First version of the excel 2 sql program to translate excel database into sql database.

@@ -1,0 +1,45 @@
+package es.ugr.decsai.excel2sql;
+
+import java.util.List;
+import java.util.logging.Level;
+import java.util.logging.Logger;
+import org.hibernate.type.Type;
+
+/**
+ * Hello world!
+ *
+ */
+public class App {
+
+    public static void main(String[] args) {
+        //  if(args.length >0){
+        //String filename = args[0];
+        String filename = "E:\\Users\\Jose\\Documents\\sources\\mr-t\\develop\\excel2sql\\documents\\Output_Test.xlsx";
+        System.out.println("Starting with argument: " + filename);
+
+        try {
+            ExcelReader er = new ExcelReader(filename);
+            System.out.println("File read");
+            List<String> fieldList = er.getFieldList();
+            System.out.println("Computing field list done");
+            List<String> heuristicHibernateTypes = er.heuristicHibernateTypes();
+            System.out.println("Computing types done");
+
+            SqlFormat sqlf = new SqlFormat();
+            String createSentece = sqlf.createSentece("Contents", fieldList, heuristicHibernateTypes);
+            System.out.println(createSentece);
+//            for(int i = 0; i<fieldList.size();i++){
+//                System.out.println(fieldList.get(i)+ " type :"+heuristicHibernateTypes.get(i));
+//                
+//            }
+
+
+        } catch (Exception ex) {
+            Logger.getLogger(App.class.getName()).log(Level.SEVERE, null, ex);
+        }
+
+//        }else{
+//            System.err.println("Wrong call");
+//        }
+    }
+}

### 2013-03-14 trunk/develop/excel2sql/src/main/java/es/ugr/decsai/excel2sql/SqlFormat.java

@@ -1,0 +1,38 @@
+/*
+ * To change this template, choose Tools | Templates
+ * and open the template in the editor.
+ */
+package es.ugr.decsai.excel2sql;
+
+import java.util.List;
+
+/**
+ * Class to write sql code
+ * @author Jose
+ */
+public class SqlFormat {
+    private static final String BLANK = " ";
+    private static final String LEFT_PARENTHESIS = " ( ";
+    private static final String RIGHT_PARENTHESIS = " ) ";
+    private static final String TOKEN_CREATE = "CREATE";
+    private static final String TOKEN_TABLE = "TABLE";
+    private static final String COMMA = " , ";
+    private static final String SIMPLE_COMMA = "'";
+    private static final String END = ";";
+    
+    public String createSentece(String tableName, List<String> fieldName, List<String> fieldType){
+        String commandText = TOKEN_CREATE + BLANK + TOKEN_TABLE + BLANK + tableName+ BLANK + LEFT_PARENTHESIS;
+        
+        for(int i=0;i< fieldName.size();i++){
+            commandText += SIMPLE_COMMA + fieldName.get(i)+SIMPLE_COMMA+BLANK+fieldType.get(i);
+            if(i<fieldName.size()-1){
+                commandText += COMMA;
+            }
+        }
+        
+        
+        commandText += BLANK + RIGHT_PARENTHESIS + END;
+        return commandText;
+    }
+    
+}

### 2013-03-14 develop/excel2sql/src/main/java/es/ugr/decsai/excel2sql/App.java
update with sqlite dialect

@@ -1,5 +1,8 @@
 package es.ugr.decsai.excel2sql;
 
+import es.ugr.decsai.excel2sql.sql.SqlDatatypes;
+import es.ugr.decsai.excel2sql.sql.SqlFormat;
+import es.ugr.decsai.excel2sql.sql.sqlite.SqLiteTypeRender;
 import java.util.List;
 import java.util.logging.Level;
 import java.util.logging.Logger;
@@ -22,10 +25,10 @@ public class App {
             System.out.println("File read");
             List<String> fieldList = er.getFieldList();
             System.out.println("Computing field list done");
-            List<String> heuristicHibernateTypes = er.heuristicHibernateTypes();
+            List<SqlDatatypes> heuristicHibernateTypes = er.heuristicHibernateTypes();
             System.out.println("Computing types done");
 
-            SqlFormat sqlf = new SqlFormat();
+            SqlFormat sqlf = new SqlFormat(new SqLiteTypeRender());
             String createSentece = sqlf.createSentece("Contents", fieldList, heuristicHibernateTypes);
             System.out.println(createSentece);
 //            for(int i = 0; i<fieldList.size();i++){

### 2013-03-14 develop/excel2sql/src/main/java/es/ugr/decsai/excel2sql/ExcelReader.java

@@ -4,6 +4,7 @@
  */
 package es.ugr.decsai.excel2sql;
 
+import es.ugr.decsai.excel2sql.sql.SqlDatatypes;
 import java.io.FileInputStream;
 import java.io.FileNotFoundException;
 import java.io.IOException;
@@ -92,8 +93,8 @@ public class ExcelReader {
         return fieldList;
     }
 
-    public List<String> heuristicHibernateTypes() {
-        List<String> typeList = new ArrayList<String>();
+    public List<SqlDatatypes> heuristicHibernateTypes() {
+        List<SqlDatatypes> typeList = new ArrayList<SqlDatatypes>();
         Sheet s = wb.getSheetAt(0);
         Row row = s.getRow(0);
 
@@ -135,11 +136,11 @@ public class ExcelReader {
            
 
             if (candidateInt >= candidateDouble && candidateInt >= candidateString) {
-                typeList.add("INTEGER");
+                typeList.add(SqlDatatypes.INTEGER);
             } else if (candidateDouble >= candidateString) {
-                typeList.add("DOUBLE");
+                typeList.add(SqlDatatypes.DOUBLE);
             } else {
-                typeList.add("STRING");
+                typeList.add(SqlDatatypes.STRING);
             }
 
 

### 2013-03-14 develop/excel2sql/src/main/java/es/ugr/decsai/excel2sql/sql/SqlDatatypes.java

@@ -1,0 +1,21 @@
+/*
+ * To change this template, choose Tools | Templates
+ * and open the template in the editor.
+ */
+package es.ugr.decsai.excel2sql.sql;
+
+/**
+ * Small subset of datatypes.
+ *
+ * @author Jose Enrique Pons Frías <jpons@decsai.ugr.es>
+ */
+public enum SqlDatatypes {
+
+    INTEGER,
+    DOUBLE,
+    STRING;
+    
+    
+    
+  
+}

### 2013-03-14 develop/excel2sql/src/main/java/es/ugr/decsai/excel2sql/sql/sqlite/SqLiteDatatypes.java

@@ -1,0 +1,17 @@
+/*
+ * To change this template, choose Tools | Templates
+ * and open the template in the editor.
+ */
+package es.ugr.decsai.excel2sql.sql.sqlite;
+
+/**
+ * Datatypes in SqLite
+ * @author Jose Enrique Pons Frías <jpons@decsai.ugr.es>
+ */
+public enum SqLiteDatatypes {
+    NULL,
+    INTEGER,
+    REAL,
+    TEXT,
+    BLOB;
+}

### 2013-03-14 develop/excel2sql/src/main/java/es/ugr/decsai/excel2sql/sql/sqlite/SqLiteTypeRender.java

@@ -1,0 +1,38 @@
+/*
+ * To change this template, choose Tools | Templates
+ * and open the template in the editor.
+ */
+package es.ugr.decsai.excel2sql.sql.sqlite;
+
+import es.ugr.decsai.excel2sql.sql.AbstractSQLRenderer;
+import es.ugr.decsai.excel2sql.sql.SqlDatatypes;
+
+/**
+ *
+ * @author Jose Enrique Pons Frías <jpons@decsai.ugr.es>
+ */
+public class SqLiteTypeRender extends AbstractSQLRenderer{
+
+    /**
+     * Implementation of the interface TypeRender.
+     * Returns the datatypes in SqLite.
+     * @param type
+     * @return 
+     */
+    public String render(SqlDatatypes type){
+        String render="";
+        switch(type){
+            case DOUBLE: render = SqLiteDatatypes.REAL.name();
+                break;
+            case INTEGER: render = SqLiteDatatypes.INTEGER.name();
+                break;
+            case STRING: render = SqLiteDatatypes.TEXT.name();
+                break;
+            default:
+                render = SqLiteDatatypes.TEXT.name();
+        }
+        
+        
+        return render;
+    }
+}

### 2013-03-14 trunk/develop/excel2sql/src/main/java/es/ugr/decsai/excel2sql/sql/AbstractSQLRenderer.java
update with sqlite dialect

@@ -1,0 +1,36 @@
+/*
+ * Copyright (C) 2013 Jose Enrique Pons Frías <jpons@decsai.ugr.es>.
+ *
+ * This library is free software; you can redistribute it and/or
+ * modify it under the terms of the GNU Lesser General Public
+ * License as published by the Free Software Foundation; either
+ * version 2.1 of the License, or (at your option) any later version.
+ *
+ * This library is distributed in the hope that it will be useful,
+ * but WITHOUT ANY WARRANTY; without even the implied warranty of
+ * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the GNU
+ * Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with this library; if not, write to the Free Software
+ * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston,
+ * MA 02110-1301  USA
+ */
+package es.ugr.decsai.excel2sql.sql;
+
+/**
+ * Abstract class for rendering sql code
+ * @author Jose Enrique Pons Frías <jpons@decsai.ugr.es> First version
+ * 14/03/2013
+ *
+ */
+public abstract class AbstractSQLRenderer {
+
+    /**
+     * Function that returns the rendered type for that SQL type.
+     *
+     * @param type
+     * @return an string with the given datatype.
+     */
+    public abstract String render(SqlDatatypes type);
+}

### 2013-03-14 trunk/develop/excel2sql/src/main/java/es/ugr/decsai/excel2sql/sql/SqlFormat.java

@@ -2,7 +2,7 @@
  * To change this template, choose Tools | Templates
  * and open the template in the editor.
  */
-package es.ugr.decsai.excel2sql;
+package es.ugr.decsai.excel2sql.sql;
 
 import java.util.List;
 
@@ -20,11 +20,20 @@ public class SqlFormat {
     private static final String SIMPLE_COMMA = "'";
     private static final String END = ";";
     
-    public String createSentece(String tableName, List<String> fieldName, List<String> fieldType){
+    protected AbstractSQLRenderer sqlTypeRender;
+
+    public SqlFormat(AbstractSQLRenderer sqlTypeRender) {
+        this.sqlTypeRender = sqlTypeRender;
+    }
+    
+    
+    
+    
+    public String createSentece(String tableName, List<String> fieldName, List<SqlDatatypes> fieldType){
         String commandText = TOKEN_CREATE + BLANK + TOKEN_TABLE + BLANK + tableName+ BLANK + LEFT_PARENTHESIS;
         
         for(int i=0;i< fieldName.size();i++){
-            commandText += SIMPLE_COMMA + fieldName.get(i)+SIMPLE_COMMA+BLANK+fieldType.get(i);
+            commandText += SIMPLE_COMMA + fieldName.get(i)+SIMPLE_COMMA+BLANK+ sqlTypeRender.render(fieldType.get(i));
             if(i<fieldName.size()-1){
                 commandText += COMMA;
             }

### 2013-03-14 trunk/develop/excel2sql/src/main/java/es/ugr/decsai/excel2sql/App.java
First version of the excel 2 sql program to translate excel database into sql database.

@@ -31,10 +31,12 @@ public class App {
             SqlFormat sqlf = new SqlFormat(new SqLiteTypeRender());
             String createSentece = sqlf.createSentece("Contents", fieldList, heuristicHibernateTypes);
             System.out.println(createSentece);
-//            for(int i = 0; i<fieldList.size();i++){
-//                System.out.println(fieldList.get(i)+ " type :"+heuristicHibernateTypes.get(i));
-//                
-//            }
+            String insertSentence = sqlf.insertSentence("Contents",er.getRowContents(1) , heuristicHibernateTypes);
+            System.out.println(insertSentence);
+            //            for(int i = 0; i<fieldList.size();i++){
+            //                System.out.println(fieldList.get(i)+ " type :"+heuristicHibernateTypes.get(i));
+            //
+            //            }
 
 
         } catch (Exception ex) {

### 2013-03-14 trunk/develop/excel2sql/src/main/java/es/ugr/decsai/excel2sql/sql/SqlFormat.java

@@ -19,6 +19,8 @@ public class SqlFormat {
     private static final String COMMA = " , ";
     private static final String SIMPLE_COMMA = "'";
     private static final String END = ";";
+    private static final String INSERT = "INSERT INTO";
+    private static final String VALUES = "VALUES";
     
     protected AbstractSQLRenderer sqlTypeRender;
 
@@ -40,6 +42,18 @@ public class SqlFormat {
         }
         
         
+        commandText += BLANK + RIGHT_PARENTHESIS + END;
+        return commandText;
+    }
+    
+    public String insertSentence(String tableName,List<String> values,List<SqlDatatypes> fieldType){
+        String commandText = INSERT + BLANK + tableName + BLANK + VALUES + LEFT_PARENTHESIS;
+        for(int i =0;i< values.size();i++){
+            commandText += sqlTypeRender.renderValue(fieldType.get(i), values.get(i));
+            if(i<values.size()-1){
+                commandText += COMMA;
+            }
+        }
         commandText += BLANK + RIGHT_PARENTHESIS + END;
         return commandText;
     }

### 2013-03-14 develop/excel2sql/src/main/java/es/ugr/decsai/excel2sql/ExcelReader.java
First version of the excel 2 sql program to translate excel database into sql database.

@@ -28,7 +28,7 @@ public class ExcelReader {
 
     Workbook wb;
     private int MAX_ITER = 10;
-
+    
     /**
      * Open an excel file
      *
@@ -148,5 +148,21 @@ public class ExcelReader {
             j++;
         }
         return typeList;
+    }
+    
+    public List<String> getRowContents(int index){
+        List<String> valueList = new ArrayList<String>();
+          Sheet s = wb.getSheetAt(0);
+        Row row = s.getRow(index);
+
+        Iterator<Cell> cellIterator = row.cellIterator();
+       
+        while (cellIterator.hasNext()) {
+
+            Cell cell = cellIterator.next();
+            String content = cell.getRichStringCellValue().getString();
+            valueList.add(content);
+        }
+        return valueList;
     }
 }

### 2013-03-14 develop/excel2sql/src/main/java/es/ugr/decsai/excel2sql/sql/AbstractSQLRenderer.java

@@ -25,7 +25,8 @@ package es.ugr.decsai.excel2sql.sql;
  *
  */
 public abstract class AbstractSQLRenderer {
-
+    private static final String COMMA = "'";
+    private static final String NULL = "NULL";
     /**
      * Function that returns the rendered type for that SQL type.
      *
@@ -33,4 +34,25 @@ public abstract class AbstractSQLRenderer {
      * @return an string with the given datatype.
      */
     public abstract String render(SqlDatatypes type);
+    
+    public String renderValue(SqlDatatypes type, String value){
+        String renderedText = value;
+        boolean isNull = (value==null || value.compareTo("")==0)?true:false;
+        if(!isNull){
+        switch(type){
+            case DOUBLE:
+                break;
+            case INTEGER:
+                break;
+            case STRING: renderedText = COMMA + value + COMMA;
+                break;
+                default:
+                    renderedText = COMMA + value + COMMA;
+                   
+        }
+        }else{
+            renderedText = NULL;
+        }
+        return renderedText;
+    }
 }

### 2013-03-15 trunk/eusflat2013/notations_and_nomenclature.tex
Another update that adds a few things and dots the i's and crosses the t's in the abstract and sections 1, 2 and 3.

Often, time is thought of as following an axis (a time axis) and thus it is often seen as an ordered set of infinitely short `moments' or `points' in time, called \emph{instants}~\cite{Dyreson1994},~\cite{Jensen1998}. This is also the case in the presented work. A \emph{time interval} is then an interval subset of this ordered set of instants. In this paper, instants are noted using lowercase letters and time intervals using uppercase letters. Also, a time interval $I$ starting at (and containing) instant $s$ and ending at (and containing) instant $e$ is noted using square brackets: $I = \left[s, e\right]$.

The presented work allows time intervals to be subject to uncertainties. These uncertainties are always seen as caused by a (partial) lack of knowledge: the exact, intended time interval is not known, eventhough there is only one correct intended time interval and as such no variability. Confidence about exactly which interval is the intended one in the context of such uncertainties is modelled using possibility theory~\cite{DidierDubois1988a}. In the presented work, possibility is always interpreted as plausibility given a (partial) lack of knowledge.

To be able to distinguish easily, time intervals not subject to uncertainties will be called \emph{crisp time intervals} in the presented work.

### 2013-03-15 trunk/eusflat2013/tm.tex
Another update that adds a few things and dots the i's and crosses the t's in the abstract and sections 1, 2 and 3.

@@
 The area contained in this triangle is called the \emph{interval space} and will contain all visualizations of intervals.
-Now, the procedure used to visualize a time interval $\left[s, e\right]$ in the interval space is the following:
+Now, the procedure used to visualize a crisp time interval $\left[s, e\right]$ in the interval space is the following:
 
@@
 For this, a single-color greyscale is used where the saturation of the color is linearly related to the possibility.
+Thus, every interval point in this visualization corresponds to an interval with a higher-than-zero possibility of being the time interval represented by the UTI.
 An example of such construction is shown in figure \ref{fig:tm-ill-const-ex}.
+In this figure, a darker color indicates a higher possibility and a lighter color indicates a lower possibility.
 
@@
 \subsection{\label{subsec:tm-evaluation}Evaluation of Allen Relationships}
+At the base of the evaluation of Allen relationships in the TM framework is the concept of uncertain relational zones.
+
+\begin{definition}
+For a given UTI $J$, the \emph{uncertain relational zones} (URZ) are fixed areas in the interval space.
+\end{definition}
+
+Now, given an UTI $J$, the procedure to visualize its URZ is the following:
+
+\begin{enumerate}
+	\item Determine the visualizations (the points) of both the smallest and greatest instant with a non-zero possibility of being the starting instant of $J$.
+	\item Construct two straight half-lines from each of these points: one with an angle of size $\alpha$ with the visualization of the time axis and one with an angle of size $-\alpha$ with the visualization of the time axis.
+The parts of these half-lines that lie outside of the interval space are not of interest.
+	\item Determine the visualizations (the points) of both the smallest and greatest instant with a non-zero possibility of being the ending instant of $J$.
+	\item Construct two straight half-lines from each of these points: one with an angle of size $\alpha$ with the visualization of the time axis and one with an angle of size $-\alpha$ with the visualization of the time axis.
+The parts of these half-lines that lie outside of the interval space are not of interest.
+	\item The thusly constructed lines divide the interval space in at most fifteen different areas, including the area corresponding to the visualization of $J$ itself. These areas are $J$'s URZ.
+\end{enumerate}
+
+In this procedure, $\alpha$ is the same angle as in the procedure for the visualization of crisp time intervals.
+The results of this procedure for an example UTI $J$ (as wel as the $J$ itself) are shown in figure \ref{fig:tm-urz}.
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.8\columnwidth]{graphs/TM_model_URZ.eps}
+	\caption{The visualization of the URZ for a given UTI $J$.}
+	\label{fig:tm-urz}
+\end{figure}
+
+Given a UTI $J$ and its URZ, every URZ is assigned a symbol and a name and corresponds to a set of Allen relationships.
+The symbol and name are used to uniquely identify each of $J$'s URZ.
+The set of Allen relationships corresponding to an URZ contains every Allen relationship in which the intervals visualized by the URZ's interval points could be with the interval represented by $J$, depending on what time interval $J$ actually represents.
+In figure \ref{fig:tm-urz}, the URZ symbols are also shown.
+In table \ref{tab:urz}, every different row corresponds to a different URZ for a given UTI.
+In the first two columns, the symbol and name of the URZ are given, the last column contains the set of corresponding Allen relationships.
+For these, the following notations are used:
+
+\begin{itemize}
+	\item `E' denotes `equals'.
+	\item `S' denotes `starts'.
+	\item `SB' denotes `started by'.
+	\item `F' denotes `finishes'.
+	\item `FB' denotes `finished by'.
+	\item `M' denotes `meets'.
+	\item `MB' denotes `met by'.
+	\item `O' denotes `overlaps'.
+	\item `OB' denotes `overlapped by'.
+	\item `D' denotes `during'.
+	\item `C' denotes `contains'.
+	\item `B' denotes `before'.
+	\item `A' denotes `after'.
+\end{itemize}
+
+\begin{table}[h]
+\centering
+\begin{tabular}{|l|l|l|}
+\hline
+Symbol & Name & Relationships \\
+\hline
+B    & Before & B \\
+O    & Overlaps & O \\
+C    & Contains & C \\
+D    & During & D \\
+OB   & Overlapped by & OB \\
+A    & After & A \\
+PM   & Possibly meets & M, B, O \\
+PS   & Possibly starts & O, S, D \\
+PFB  & Possibly finished by & O, FB, C \\
+\multirow{2}{*}
+{PE}   & Possibly equal  & E, FB, SB, F,\\
+       &                 & S, C, D,O, OB \\
+PSB  & Possibly started by & C, OB, SB \\
+PF   & Possibly finishes & OB, F, D \\
+PMB  & Possibly met by & OB, MB A \\
+UB   & Uncertain beginning & B, M, O, S, D \\
+UE   & Uncertain ending & D, F, OB, MB, A \\
+\hline
+\end{tabular}
+\caption{The fifteen possible URZ for a given UTI.}
+\label{tab:urz}
+\end{table}
+
+The TM framework now allows evaluating all possible Allen relationships between a given crisp time interval $I$ and a given UTI $J$.
+This is done as follows.
+First, $J$ and its URZ are visualized.
+Next, $I$'s interval point and its construction half-lines $L_s$ and $L_e$ are drawn in the same interval space.
+Then, evaluation is a matter of positioning (in the following, the possibility that $I$ is in Allen relationship $AR$ with $J$ is denoted  $\Pos(I\ AR\ J)$):
+
+\begin{itemize}
+	\item if $I$'s interval point is located in an URZ corresponding with a singleton of Allen relationships $\{R\}$, it is in relationship $R$ with the interval represented by $J$, whichever this interval is.
+Thus: $\Pos(I\ AR\ J) = 1$.
+	\item if $I$'s interval point is located in an URZ corresponding with a non-singleton set of $n$ Allen relationships $\{R_1, R_2, \ldots, R_n\}$, two more half-lines are drawn: one (denoted $L_s'$) from the intersection point of $L_s$ and the visualization of the time axis, its angle with this visualization having size $-\alpha$ and one (denoted $L_e'$) from the intersection point of $L_e$ and the visualization of the time axis, having an angle $\alpha$ with this visualization.
+Now, some of these half-lines $L_s$, $L_s'$, $L_e$ and $L_e'$ may intersect with the area visualizing $J$.
+These will divide the area visualizing $J$ in $n$ parts $p_i, 1 \leq i \leq n$ (intersection line segments also count as parts) and each of these parts will correspond to an Allen relationship $R_i$.
+Now, for each $R_i$:
+	\begin{equation}
+		\Pos(I\ R_i\ J) = \sup_{I \in p_i} \pi_J(I)
+	\end{equation}
+\end{itemize}
+
+%TODO: example!
 
@@
 
-\subsection{\label{subsec:relational-info-uti}Obtaining Relational Information}
-In order to make temporal reasoning with UTIs, in the triangular model, the thirteen Allen's Relations, and its respective Crisp Relational Zones (CRZ) have been generalized.
+%\subsection{\label{subsec:relational-info-uti}Obtaining Relational Information}
+%In order to make temporal reasoning with UTIs, in the triangular model, the thirteen Allen's Relations, and its respective Crisp Relational Zones (CRZ) have been generalized.
 In this case, we obtain fifteen Uncertain Relational Zones (URZ) , shown in Fig.
@@
 
-\begin{figure}[h]
-   \centering
-   \includegraphics[width=0.8\columnwidth]{graphs/urz.eps}
-   \caption{Uncertain Relational Zones (URZ) for a given UTI.  }
-   \label{fig:urz}
- \end{figure}
+%\begin{figure}[h]
+%   \centering
+%   \includegraphics[width=0.8\columnwidth]{graphs/urz.eps}
+%   \caption{Uncertain Relational Zones (URZ) for a given UTI.  }
+%   \label{fig:urz}
+% \end{figure}
 
-\begin{table}[h]
-\centering
-\begin{tabular}{|l|l|l|}
-\hline
-Symbol & Name & Possible Relations \\
-\hline
-B    & Before & B \\
-O    & Overlaps & O \\
-C    & Contains & C \\
-D    & During & D \\
-OB   & Overlapped by & OB \\
-A    & After & A \\
-PM   & Possibly meets & M, B, O \\
-PS   & Possibly starts & O, S, D \\
-PFB  & Possibly finished by & O, FB, C \\
-\multirow{2}{*}
-{PE}   & Possibly equal  & E, FB, SB, F,\\
-       &                 & S, C, D,O, OB \\
-%       &                 &  \\
-PSB  & Possibly started by & C, OB, SB \\
-PF   & Possibly finishes & OB, MB, A \\
-PMB  & Possibly met by & B, M, O, S, D \\
-UB   & Uncertain beginning & B, M, O, S, D \\
-UE   & Uncertain ending & D, F, OB, MB, A \\
-\hline
-\end{tabular}
-\caption{The fifteen possible relations between an interval point $J$ and an UTI $\pi_I$.}
-\label{tab:urz}
-\end{table}
+%The way to obtain the relational information between an interval point $J = \left[s_j, e_j \right]$, and an UTI $\pi_I$ depends on the location of $J$ with respect to the UTI $\pi_I$:
 
-The way to obtain the relational information between an interval point $J = \left[s_j, e_j \right]$, and an UTI $\pi_I$ depends on the location of $J$ with respect to the UTI $\pi_I$:
+%\begin{itemize}
+% \item If $J$ is located in a zone with only one possible Allen relation, then it is true that $J$ is in the relation. (And therefore, the possibility is always 1).
+%\begin{equation}
+% \label{eq:possibility-iz-uti-certain}
+%\mu_{\left(J R \pi_I \right)} \left(R \right) = 1.
+%\end{equation}
 
-\begin{itemize}
- \item If $J$ is located in a zone with only one possible Allen relation, then it is true that $J$ is in the relation. (And therefore, the possibility is always 1).
-\begin{equation}
- \label{eq:possibility-iz-uti-certain}
-\mu_{\left(J R \pi_I \right)} \left(R \right) = 1.
-\end{equation}
-
-\item If $J$ is located in a zone with multiple candidate Allen's relations, then there is uncertainty about the Allen relation that applies between $J$ and the UTI $\pi_I$.
+%\item If $J$ is located in a zone with multiple candidate Allen's relations, then there is uncertainty about the Allen relation that applies between $J$ and the UTI $\pi_I$.
 The procedure to obtain the possibility degree is the following:
 
-\begin{enumerate}
- \item Draw the lines $L_s, L_e$ that represent the interval point $J$, with the procedure shown in Section \ref{subsec:visualization-tm}. 
-\item Draw the orthogonal lines from $L_s, L_e$. 
-\item One or several of these lines could intersect with the UIZ of $\pi_I$.
+%\begin{enumerate}
+% \item Draw the lines $L_s, L_e$ that represent the interval point $J$, with the procedure shown in Section \ref{subsec:visualization-tm}. 
+%\item Draw the orthogonal lines from $L_s, L_e$. 
+%\item One or several of these lines could intersect with the UIZ of $\pi_I$.
 Then, the UIZ is divided into as many different subzones $\pi_I^{R'}$ as possible candidate Allen relations $R'$ between $J$ and $\pi_I$ exist.
-\item $\pi_I^{R'}$ is the zone that correspond to a candidate Allen Relation $R'$.
-\end{enumerate}
-The possibility that a candidate relation $R'$ is the actual relation between $J$ and $\pi_I$ is given by:
-\begin{equation}
- \label{eq:possibility-iz-uti}
-\mu_{\left(J R \pi_I \right)} \left(R' \right) = \sup_{I \in \pi_I^{R'}} \pi_I(I).
-\end{equation}
-\end{itemize}
+%\item $\pi_I^{R'}$ is the zone that correspond to a candidate Allen Relation $R'$.
+%\end{enumerate}
+%The possibility that a candidate relation $R'$ is the actual relation between $J$ and $\pi_I$ is given by:
+%\begin{equation}
+% \label{eq:possibility-iz-uti}
+%\mu_{\left(J R \pi_I \right)} \left(R' \right) = \sup_{I \in \pi_I^{R'}} \pi_I(I).
+%\end{equation}
+%\end{itemize}
 
-\begin{figure}[h]
-   \centering
-   \includegraphics[width=0.8\columnwidth]{graphs/calculation.eps}
-   \caption{Uncertain Relational Zones (URZ) for a given UTI.  }
-   \label{fig:calculation}
- \end{figure}
+%\begin{figure}[h]
+%   \centering
+%   \includegraphics[width=0.8\columnwidth]{graphs/calculation.eps}
+%   \caption{Uncertain Relational Zones (URZ) for a given UTI.  }
+%   \label{fig:calculation}
+% \end{figure}
 

### 2013-03-15 eusflat2013/introduction.tex
Another update that adds a few things and dots the i's and crosses the t's in the abstract and sections 1, 2 and 3.

@@
-%1. Introduction
-Information systems (IS) in general and database systems (DS) in particular have always tried to model parts of reality.
+Information systems (IS) have always tried to model parts of reality.
 To achieve this modelling, such systems contain data representing properties of real-world objects or concepts.
-As time is an essential aspect of many of these, information systems often contain data representing temporal notions which describe their temporal properties.
-Such temporal notions usually take the form of either instants (time points)~\cite{Dyreson1994},~\cite{Jensen1998} or time intervals~\cite{Dyreson1994},~\cite{Jensen1998}.
-%TODO: more references in the above part?
+As time is an essential aspect of many real-world objects or concepts, information systems often contain data representing temporal notions which describe such temporal properties.
+Such temporal notions usually take the form of either instants~\cite{Dyreson1994},~\cite{Jensen1998}, which can informally be seen as infinitely short `moments' or `points' in time, or time intervals~\cite{Dyreson1994},~\cite{Jensen1998}.
 
 Data are often produced by humans, but human-made data are prone to imperfections: some data may be vague, imprecise, incomplete, contradictory or uncertain.
-Data representing temporal notions may contain such uncertainties too.
+Data representing temporal notions may contain such imprecisions too.
+The work presented in this paper is specifically concerned with time intervals (and as a special case: instants) subject to uncertainty.
 
 Generally, one of the most important purposes of an IS is to allow the retrieval of information or knowledge deduced from its data.
-Information or knowledge is usually retrieved from an IS by querying its contents and examining the query results or by visualizing its contents, querying this visualization and examining these query results.
-%TODO: include some form of proof or previous work with these statements!
+Information or knowledge is usually retrieved from an IS by querying it and examining or analyzing the query results or by visualizing the contents of the IS, querying this visualization and examining or analyzing the resulting visualization(s).
 
-Of course, when temporal information is contained in an IS, it can be used in the querying process.
-Usually, querying an IS containing data representing temporal notions is done by specifying one or more time indications and requesting information that is related to these notions, where the nature and semantics of these relationships are specifically temporal.
-Thus, specific temporal relationships are necessary.
+Of course, when temporal information is represented in an IS, querying this IS may have a temporal aspect too.
+Usually, querying an IS containing data representing temporal notions is conceptually done by specifying one or more time indications and requesting information that is in a specific relationship with these indications, where the semantics of these relationships are specifically temporal.
+Thus, some existing proposals have considered groups of basic relationships between time indications used to construct and express specific temporal relationships.
 Notably, Allen~\cite{Allen1983} presented a reasoning framework containing all semantically usefull basic temporal relationships between time intervals (and as a special case instants).
-The resulting Allen relationships are shown in figure \ref{fig:allen-relationships}.
+The resulting relationships are shown in figure \ref{fig:allen-relationships}.
+These Allen relationships are used in the presented work.
 
-To be able to query IS and DS containing uncertain data representing uncertain time indications, a framework is necessary, able to represent uncertainties in time indications without (much) information loss and able to temporally reason with such uncertain time indications.
-Although many proposals for such frameworks exist, the work presented in this paper focusses on two: the \emph{ill-known constraint \emph{(IKC)} framework}~\cite{Pons2011} and the \emph{triangular model \emph{(TM)} framework}~\cite{DeTre2012}.
+\begin{figure}[h]
+   \centering
+   \includegraphics[width=0.8\columnwidth]{graphs/allen.eps}
+   \caption{The Allen relationships between two crisp intervals $I$ and $J$.  }
+   \label{fig:allen-relationships}
+ \end{figure}
 
-The work presented in this paper consists of a comparison of the approaches, used by both frameworks, to representing time intervals subject to uncertainty and to reasoning about the temporal relationships between such intervals and perfectly defined time intervals.
+To be able to query IS containing data representing time indications subject to uncertainty, a framework is necessary, able to represent uncertainties in time indications in a semantically sound way, without (much) information loss and able to temporally reason with such time indications in a semantically sound and usefull way.
+Although more proposals for such frameworks exist, the work presented in this paper focusses on just two: the \emph{ill-known constraint \emph{(IKC)} framework}~\cite{Pons2011} and the \emph{triangular model \emph{(TM)} framework}~\cite{DeTre2012}.
+
+The work presented in this paper consists of a comparison of both frameworks, about the approaches they use to represent time intervals subject to uncertainty and to reason about the temporal relationships between such intervals and time intervals without uncertainty.
 
 The structure of this paper is as follows: section \ref{sec:general-preliminaries} presents some general preliminaries and naming and notation conventions used in this paper.
-Sections \ref{sec:ikc} and \ref{sec:tm} introduce the IKC and TM frameworks respectively: both sections first introduce some preliminary concepts and techniques specific to the framework under consideration, then explain how the representation of time intervals is done and finally show how the evaluation of Allen relationships between a crisp interval and one subject to uncertainty is done.
+Sections \ref{sec:ikc} and \ref{sec:tm} introduce the IKC and TM frameworks respectively: both sections first introduce some preliminary concepts and techniques specific to the framework under consideration, then explain how the representation of time intervals by the framework is done and finally show how the evaluation of Allen relationships between an interval subject to uncertainty and one without uncertainty is done.
 In section \ref{sec:proposal}, both frameworks are compared: first their approaches to representing time intervals are compared, next their approaches to evaluating Allen relationships are compared.
 Finally, section \ref{sec:conclusions} presents the principal conclusions of the presented work and some possible future work.
+
+%TODO: more references in the above part?
+%TODO: include some form of proof or previous work with these statements?!
 
@@
 %\end{table}
-
-\begin{figure}[h]
-   \centering
-   \includegraphics[width=0.8\columnwidth]{graphs/allen.eps}
-   \caption{The Allen relationships between two crisp intervals $I$ and $J$.  }
-   \label{fig:allen-relationships}
- \end{figure}

### 2013-03-15 develop/excel2sql/src/main/java/es/ugr/decsai/excel2sql/App.java
Executable version for the excel2sql program

@@ -3,6 +3,9 @@ package es.ugr.decsai.excel2sql;
 import es.ugr.decsai.excel2sql.sql.SqlDatatypes;
 import es.ugr.decsai.excel2sql.sql.SqlFormat;
 import es.ugr.decsai.excel2sql.sql.sqlite.SqLiteTypeRender;
+import java.sql.Connection;
+import java.sql.DriverManager;
+import java.sql.Statement;
 import java.util.List;
 import java.util.logging.Level;
 import java.util.logging.Logger;
@@ -17,11 +20,12 @@ public class App {
     public static void main(String[] args) {
         //  if(args.length >0){
         //String filename = args[0];
-        String filename = "E:\\Users\\Jose\\Documents\\sources\\mr-t\\develop\\excel2sql\\documents\\Output_Test.xlsx";
+        String path = "E:\\Users\\Jose\\Documents\\sources\\mr-t\\develop\\excel2sql\\documents\\";
+        String filename = "Output_Test.xlsx";
         System.out.println("Starting with argument: " + filename);
 
         try {
-            ExcelReader er = new ExcelReader(filename);
+            ExcelReader er = new ExcelReader(path+filename);
             System.out.println("File read");
             List<String> fieldList = er.getFieldList();
             System.out.println("Computing field list done");
@@ -32,7 +36,43 @@ public class App {
             String createSentece = sqlf.createSentece("Contents", fieldList, heuristicHibernateTypes);
             System.out.println(createSentece);
             String insertSentence = sqlf.insertSentence("Contents",er.getRowContents(1) , heuristicHibernateTypes);
+            String deleteSentence = "DROP TABLE Contents";
             System.out.println(insertSentence);
+            
+            //try to connect to sqlitedb:
+            String driverName = "org.sqlite.JDBC";
+            Class.forName(driverName);
+            
+            String databaseName = "history.db";
+            String connectionString = "jdbc:sqlite:" + path+ databaseName;
+            Connection conn = DriverManager.getConnection(connectionString);
+           // Statement stm;
+            try{
+               Statement stm = conn.createStatement();
+               try{
+                   stm.executeUpdate(deleteSentence);
+                   stm.executeUpdate(createSentece);
+                  // stm.executeUpdate(insertSentence);
+                   for(int i=1;i<er.getRows();i++){
+                       sqlf.insertDatabase("Contents", er.getRowContents(i), heuristicHibernateTypes, fieldList, conn);
+                       System.out.println("Inserting row"+i+"...");
[…]

### 2013-03-15 develop/excel2sql/src/main/java/es/ugr/decsai/excel2sql/ExcelReader.java

@@ -165,4 +165,9 @@ public class ExcelReader {
         }
         return valueList;
     }
+    
+    public int getRows(){
+        Sheet s = wb.getSheetAt(0);
+       return s.getLastRowNum();
+    }
 }

### 2013-03-15 develop/excel2sql/src/main/java/es/ugr/decsai/excel2sql/sql/SqlFormat.java

@@ -4,13 +4,21 @@
  */
 package es.ugr.decsai.excel2sql.sql;
 
+import java.sql.Connection;
+import java.sql.PreparedStatement;
+import java.sql.SQLException;
+import java.sql.Types;
 import java.util.List;
+import java.util.logging.Level;
+import java.util.logging.Logger;
 
 /**
  * Class to write sql code
+ *
  * @author Jose
  */
 public class SqlFormat {
+
     private static final String BLANK = " ";
     private static final String LEFT_PARENTHESIS = " ( ";
     private static final String RIGHT_PARENTHESIS = " ) ";
@@ -19,43 +27,135 @@ public class SqlFormat {
     private static final String COMMA = " , ";
     private static final String SIMPLE_COMMA = "'";
     private static final String END = ";";
-    private static final String INSERT = "INSERT INTO";
+    private static final String INSERT = "INSERT";
+    private static final String INTO = "INTO";
     private static final String VALUES = "VALUES";
-    
+    private static final String PARAM = "?";
+    private static final String EMPTY_STRING = "";
     protected AbstractSQLRenderer sqlTypeRender;
 
     public SqlFormat(AbstractSQLRenderer sqlTypeRender) {
         this.sqlTypeRender = sqlTypeRender;
     }
-    
-    
-    
-    
-    public String createSentece(String tableName, List<String> fieldName, List<SqlDatatypes> fieldType){
-        String commandText = TOKEN_CREATE + BLANK + TOKEN_TABLE + BLANK + tableName+ BLANK + LEFT_PARENTHESIS;
-        
-        for(int i=0;i< fieldName.size();i++){
-            commandText += SIMPLE_COMMA + fieldName.get(i)+SIMPLE_COMMA+BLANK+ sqlTypeRender.render(fieldType.get(i));
-            if(i<fieldName.size()-1){
+
+    public String createSentece(String tableName, List<String> fieldName, List<SqlDatatypes> fieldType) {
+        String commandText = TOKEN_CREATE + BLANK + TOKEN_TABLE + BLANK + tableName + BLANK + LEFT_PARENTHESIS;
+
+        for (int i = 0; i < fieldName.size(); i++) {
+            commandText += SIMPLE_COMMA + fieldName.get(i) + SIMPLE_COMMA + BLANK + sqlTypeRender.render(fieldType.get(i));
+            if (i < fieldName.size() - 1) {
                 commandText += COMMA;
             }
         }
-        
-        
+
+
[…]

### 2013-03-17 trunk/develop/utiCalculator/src/main/java/es/ugr/decsai/jpons/uticalculator/App.java
library for utis

@@ -1,0 +1,45 @@
+package es.ugr.decsai.jpons.uticalculator;
+
+import java.util.List;
+
+/**
+ * Hello world!
+ *
+ */
+public class App 
+{
+    public static void main( String[] args )
+    {
+        if(args.length < 6){
+            System.err.println("Sintax error: Incorrect parameter number");
+            System.err.println("parameters: start1 start start2 end1 end end2");
+            System.exit(-1);
+        }
+        int precision = 25;
+        if(args.length == 7){
+            precision = Integer.valueOf(args[6]);
+        }
+        
+        
+        double start1,start,start2,end1,end,end2;
+        
+        start1 = Double.valueOf(args[0]);
+        start = Double.valueOf(args[1]);
+        start2 = Double.valueOf(args[2]);
+        
+        end1 = Double.valueOf(args[3]);
+        end = Double.valueOf(args[4]);
+        end2 = Double.valueOf(args[5]);
+        
+     //   System.out.println("Values: "+ start1 + ", " + start + " start2");
+        
+        Uti u = new Uti(start1, start, start2, end1, end, end2);
+        u.setIterations(precision);
+        List<IntervalPoint> intervalPoints = u.getIntervalPoints();
+        
+        for(IntervalPoint p:intervalPoints){
+            System.out.println(p.tx+ "\t"+ p.ty + "\t"+ p.membership);
+        }
+        
+    }
+}

### 2013-03-17 trunk/develop/utiCalculator/src/test/java/es/ugr/decsai/jpons/uticalculator/AppTest.java

@@ -1,0 +1,38 @@
+package es.ugr.decsai.jpons.uticalculator;
+
+import junit.framework.Test;
+import junit.framework.TestCase;
+import junit.framework.TestSuite;
+
+/**
+ * Unit test for simple App.
+ */
+public class AppTest 
+    extends TestCase
+{
+    /**
+     * Create the test case
+     *
+     * @param testName name of the test case
+     */
+    public AppTest( String testName )
+    {
+        super( testName );
+    }
+
+    /**
+     * @return the suite of tests being tested
+     */
+    public static Test suite()
+    {
+        return new TestSuite( AppTest.class );
+    }
+
+    /**
+     * Rigourous Test :-)
+     */
+    public void testApp()
+    {
+        assertTrue( true );
+    }
+}

### 2013-03-17 develop/utiCalculator/src/main/java/es/ugr/decsai/jpons/uticalculator/IntervalPoint.java
library for utis

@@ -1,0 +1,46 @@
+/*
+ * Copyright (C) 2013 Jose Enrique Pons Frías <jpons@decsai.ugr.es>.
+ *
+ * This library is free software; you can redistribute it and/or
+ * modify it under the terms of the GNU Lesser General Public
+ * License as published by the Free Software Foundation; either
+ * version 2.1 of the License, or (at your option) any later version.
+ *
+ * This library is distributed in the hope that it will be useful,
+ * but WITHOUT ANY WARRANTY; without even the implied warranty of
+ * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the GNU
+ * Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with this library; if not, write to the Free Software
+ * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston,
+ * MA 02110-1301  USA
+ */
+package es.ugr.decsai.jpons.uticalculator;
+
+/**
+ *
+ * @author Jose Enrique Pons Frías <jpons@decsai.ugr.es> First version
+ * 06/03/2013
+ *
+ */
+public class IntervalPoint {
+
+    double start;
+    double end;
+    double tx;
+    double ty;
+    double membership;
+
+    public IntervalPoint(double start, double end, double membership) {
+        this.start = start;
+        this.end = end;
+        this.membership = membership;
+        toTM();
+    }
+
+    private void toTM() {
+        tx = (start + end) / 2;
+        ty = (end - start) / 2;
+    }
+}

### 2013-03-17 develop/utiCalculator/src/main/java/es/ugr/decsai/jpons/uticalculator/Uti.java

@@ -1,0 +1,144 @@
+/*
+ * Copyright (C) 2013 Jose Enrique Pons Frías <jpons@decsai.ugr.es>.
+ *
+ * This library is free software; you can redistribute it and/or
+ * modify it under the terms of the GNU Lesser General Public
+ * License as published by the Free Software Foundation; either
+ * version 2.1 of the License, or (at your option) any later version.
+ *
+ * This library is distributed in the hope that it will be useful,
+ * but WITHOUT ANY WARRANTY; without even the implied warranty of
+ * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the GNU
+ * Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with this library; if not, write to the Free Software
+ * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston,
+ * MA 02110-1301  USA
+ */
+package es.ugr.decsai.jpons.uticalculator;
+
+import java.util.ArrayList;
+import java.util.List;
+
+/**
+ * Uncertain time interval
+ *
+ * @author Jose Enrique Pons Frías <jpons@decsai.ugr.es> First version
+ * 06/03/2013
+ *
+ */
+public class Uti {
+
+    /**
+     * Coordinates in the cartesian model
+     */
+    protected double start;
+    protected double end;
+    protected double start1;
+    protected double start2;
+    protected double end1;
+    protected double end2;
+    protected  int iterations = 100;
+    
+    public void setIterations(int iter){
+        iterations = iter;
+    }
+
+    /**
+     * computes the coordinates in the tm model
+     */
+    protected double getTx(double a, double b) {
+        return (a + b) / 2;
+    }
+
+    protected double getTy(double a, double b) {
+        return (b - a) / 2;
+    }
+
+    /**
+     * Function that computes a triangular membership function
+     *
+     * @param start usually D-a
+     * @param core usually D
+     * @param end usually D+b
+     * @param value The value to compute the membership
+     * @return a value in the unit interval [0,1] that represents the membership
+     */
+    protected double triangularMF(double start, double core, double end, double value) {
+        double membership = 0D;
+        if (value == core) {
+            membership = 1D;
+        } else if (value >= start && value < core) {
+            membership = (value - start) / (core - start);
+        } else if (value >= core && value < end) {
+            membership = (end - value) / (end - core);
+        }
[…]

### 2013-03-20 trunk/eusflat2013/tm.tex
Everything ok up untill section 5 (maybe some details ...).

@@
-In~\cite{Weghe2007}, the triangular model framework is introduced, based on~\cite{Kulpa1997}.
+In~\cite{Weghe2007}, the TM framework for crisp intervals is introduced, based on~\cite{Kulpa1997}.
 In~\cite{DeTre2012}, the framework is generalized to allow the representation of and reasoning with time intervals subject to uncertainty.
-In this section, this generalization's main aspects are briefly presented.
+In this section, this generalization's main aspects towards interval representation and Allen relationship evaluation are briefly presented.
 
 \subsection{\label{subsec:tm-preliminaries}Preliminary Concepts}
-Usually, different time intervals are visualized as different parallel line segments in the image plane.
-However, as this linear approach introduces a few difficulties, the TM framework was conceived to visualize time intervals as points in the image plane.
+In the triangular model framework, the approaches to interval representation and Allen relationship evaluation are closely linked with the approach to interval visualization.
 
-To achieve this, on the image plane, a horizontal line segment visualizing the used (part of the) time axis is drawn.
+Usually, different time intervals are visualized as different parallel line segments in the same image plane.
+However, as this linear approach introduces a few issues, time intervals are visualized as points in the image plane in the TM framework.
+
+To achieve this, a horizontal line segment visualizing the used (part of the) time axis is drawn in the image plane.
 Then, a triangle is drawn, using this line segment as a side of which the angles with the other two sides have sizes $\alpha$ and $-\alpha$ respectively.
@@
 \begin{enumerate}
-	\item Visualize $s$ as a point on the visualization of the time axis.
-	\item Construct a straight half-line $L_s$ from this point, with an angle of size $\alpha$ with the visualization of the time axis.
-	\item Visualize $e$ as a point on the visualization of the time axis.
-	\item Construct a straight half-line $L_e$ from this point, with an angle of size $-\alpha$ with the visualization of the time axis.
-	\item The intersection point of $L_s$ and $L_e$ is the visualization of interval $\left[s, e\right]$.
+	\item $s$ is visualized as a point on the visualization of the time axis.
+	\item A straight half-line $L_s$ is constructed from this point, having, in this point, an angle of size $\alpha$ with the visualization of the time axis.
+	\item $e$ is visualized as a point on the visualization of the time axis.
+	\item A straight half-line $L_e$ is constructed from this point, having, in this point, an angle of size $-\alpha$ with the visualization of the time axis.
+	\item The intersection point of $L_s$ and $L_e$ is the intended visualization of interval $\left[s, e\right]$.
 \end{enumerate}
 
-The resulting point visualizing an interval is generally called an \emph{interval point}.
+A similar point visualizing an interval is generally called an \emph{interval point}.
 In this work, as it is usually done, the size of $\alpha$ is chosen to be $45^{\circ}$.
@@
 	\includegraphics[width=0.8\columnwidth]{graphs/TM_model_several.eps}
-	\caption{The visualization of several time intervals using the TM framework.}
+	\caption{The visualization of several crisp time intervals using the TM framework.}
 	\label{fig:tm-const-ex}
@@
 
-%Time intervals are usually represented in a one dimensional scale.
-This is due to its linear characteristic.
-This approach has also been used in Figure\ref{fig:allen-relationships} and \ref{fig:ill-known-ti}.
-The difficulty of accurate visual analysis on the temporal data arises when the number of time intervals increase.
-
-%The triangular model (TM) ~\cite{Kulpa1997}, \cite{Weghe2007} represents time intervals in a two dimensional space called IRterval point completely determines the start and end of the interval.
-The two-dimensional space of interval points is called the interval space and is denoted by IR ~\cite{Kulpa2006}.
-The procedure to represent a time interval $[t_s, t_e]$ as an interval point in a two dimensional space is the following:
-
-%\begin{enumerate}
-%\item Draw each time point of the interval $[t_s, t_e]$ in the horizontal line.
-%\item Construct one straight line $L_s$ from $t_s$ with an angle $\alpha$.
-%\item Construct one straight line $L_e$ from $t_e$ with an angle $-\alpha$.
-%\item The intersection of these two lines is the interval point.
-%\end{enumerate}
-
-%For the shake of simplicity, in this work, the angle for $\alpha$ is chosen to be $45^{\circ}$.
-
-%In TM, these thirteen Allen relations each correspond to a specific zone in the interval space IR \cite{Kulpa1997}.
-These thirteen zones are called Crisp Relational Zones (CRZs).
-The CRZs with respect to a reference interval point $I$ are depicted in Figure \ref{fig:allencrz}.
-
-%\begin{figure}[h]
-%   \centering
-%   \includegraphics[width=0.8\columnwidth]{graphs/allencrc.eps}
-%   \caption{CRZs corresponding to the thirteen Allen's relations.
-The following shorthand notations are used: equals (E), starts (S), started by (SB), finishes (F), finished by (FB), meets (M), met by (MB), overlaps (O), overlapped by (OB), during (D), contains (C), before (B) and after (A).
-}
-%   \label{fig:allencrz}
-% \end{figure}
-
-%Each CRZ represents the set of interval points of the crisp time intervals that are in the corresponding Allen relation with respect to the reference interval $I$.
-For example, all interval points of intervals that come before $I$ are located in the CRZ represented by the leftmost lower triangle in the interval space.
-Likewise, if an interval point is located in the top quadrangle above $I$, then its corresponding interval contains $I$.
-CRZs allow to visually analyse relations between (large) sets of interval points.
-
-\subsection{\label{subsec:tm-interval}Representation of Uncertainty in Intervals}
+\subsection{\label{subsec:tm-interval}Representation of Time Intervals Subject to Uncertainty}
 The TM framework allows uncertainty in time intervals by supporting uncertain time intervals:
@@
 
-For a UTI $J$ defined by $(\pi_s, \pi_e)$ with $\pi_s$ and $\pi_e$ two possibility distributions on an ordered set of instants $T$, $\pi_s$ and $\pi_e$ define the confidence that an instant in $T$ is the starting instant, respectively the ending instant of $J$.
-Because for such UTI $J$, $\pi_s$ and $\pi_e$ are required to be convex, the possibility $\pi_J(I)$ that a given crisp time interval $I = \left[s_i, e_i\right]$ is the exact time interval represented by $J$ can be uniquely determined by:
+Consider an ordered set of instants $T$ and an UTI $J$ in $T$, defined by the pair $(\pi_s, \pi_e)$.
+Given an instant $x \in T$, $\pi_s(x)$ now expresses the possibility that $x$ is the starting instant of $J$ and $\pi_e(x)$ expresses the possibility that $x$ is the ending instant of $J$.
+Thus, $J$ intends to indicate just one crisp time interval, but due to a (partial) lack of knowledge, it is (partially) unknown exactly which time interval this is.
+Because $\pi_s$ and $\pi_e$ are required to be convex, the possibility $\pi_J(I)$ that a given crisp time interval $I = \left[s_i, e_i\right]$ is the exact time interval intended by $J$ can now be uniquely determined by:
 
@@
 
-An UTI $J$ defined by a pair $(\pi_s, \pi_e)$ is now visualized as an area in the image plane, using a single-color greyscale.
-Every crisp time interval $I$ for which the possibility $\pi_J(I)$ of $I$ being the time interval represented by $J$ is higher than zero, is considered.
-For every such $I$, the interval point is drawn and is given the greyscale color which reflects the possibility that $I$ is the time interval represented by $J$.
-For this, a single-color greyscale is used where the saturation of the color is linearly related to the possibility.
-Thus, every interval point in this visualization corresponds to an interval with a higher-than-zero possibility of being the time interval represented by the UTI.
+Now consider an ordered set of instants $T$ and a visualization of this in an image plane, using the TM framework.
+An UTI $J$ in $T$ defined by a pair $(\pi_s, \pi_e)$ is now visualized as an area in the image plane.
+This area is constructed as follows: for every crisp time interval $K$ for which the possibility $\pi_J(K)$ of $K$ being the time interval represented by $J$ is higher than zero, the interval point is drawn and is given a color.
+This color is part of a linear single-color greyscale and its saturation is linearly related to the degree of possibility that $K$ is the time interval intended by $J$.
+Thus, every drawn interval point in this visualization corresponds to an interval with a higher-than-zero possibility of being the time interval intended by $J$.
+Together, these drawn interval points form the visualization of $J$.
 An example of such construction is shown in figure \ref{fig:tm-ill-const-ex}.
-In this figure, a darker color indicates a higher possibility and a lighter color indicates a lower possibility.
+In this figure, a darker color of interval point indicates a higher and a lighter color a lower possibility that the corresponding interval is the intended interval.
 
@@
 	\includegraphics[width=0.8\columnwidth]{graphs/TM_model_ill_known.eps}
-	\caption{The visualization of a UTI defined by a pair $(\pi_s, \pi_e)$ using the TM framework.}
+	\caption{The visualization of one UTI defined by a pair $(\pi_s, \pi_e)$ using the TM framework.}
 	\label{fig:tm-ill-const-ex}
@@
 \subsection{\label{subsec:tm-evaluation}Evaluation of Allen Relationships}
-At the base of the evaluation of Allen relationships in the TM framework is the concept of uncertain relational zones.
+At the base of the framework's approach to evaluating Allen relationships is the concept of uncertain relational zones:
 
@@
 
-Now, given an UTI $J$, the procedure to visualize its URZ is the following:
+Now, given an UTI $J$ and its visualization in an interval space, the procedure to visualize its URZ is the following:
 
 \begin{enumerate}
-	\item Determine the visualizations (the points) of both the smallest and greatest instant with a non-zero possibility of being the starting instant of $J$.
-	\item Construct two straight half-lines from each of these points: one with an angle of size $\alpha$ with the visualization of the time axis and one with an angle of size $-\alpha$ with the visualization of the time axis.
-The parts of these half-lines that lie outside of the interval space are not of interest.
-	\item Determine the visualizations (the points) of both the smallest and greatest instant with a non-zero possibility of being the ending instant of $J$.
-	\item Construct two straight half-lines from each of these points: one with an angle of size $\alpha$ with the visualization of the time axis and one with an angle of size $-\alpha$ with the visualization of the time axis.
-The parts of these half-lines that lie outside of the interval space are not of interest.
-	\item The thusly constructed lines divide the interval space in at most fifteen different areas, including the area corresponding to the visualization of $J$ itself. These areas are $J$'s URZ.
+	\item The visualizations (the points) on the visualization of the time axis of both the smallest and greatest instant with a non-zero possibility of being the starting instant of $J$ are determined.
+	\item Two straight half-lines are constructed from each of these points: one having, in the point, an angle of size $\alpha$ with the visualization of the time axis and one having, in the point, an angle of size $-\alpha$ with the visualization of the time axis.
+The parts of these half-lines that lie outside of the interval space are not drawn.
+	\item The visualizations (the points) on the visualization of the time axis of both the smallest and greatest instant with a non-zero possibility of being the ending instant of $J$ are determined.
+	\item Two straight half-lines are constructed from each of these points: one having, in the point, an angle of size $\alpha$ with the visualization of the time axis and one having, in the point, an angle of size $-\alpha$ with the visualization of the time axis.
+The parts of these half-lines that lie outside of the interval space are not drawn.
+	\item The lines constructed this way divide the interval space in at most fifteen different areas, including the area corresponding to the visualization of $J$ itself.
+These areas are the intended visualizations of $J$'s URZ.
 \end{enumerate}
 
-In this procedure, $\alpha$ is the same angle as in the procedure for the visualization of crisp time intervals.
-The results of this procedure for an example UTI $J$ (as wel as the $J$ itself) are shown in figure \ref{fig:tm-urz}.
+In this procedure, $\alpha$ is the same angle size as in the procedure for the visualization of crisp time intervals.
+The results of this procedure for an example UTI $J$ (as wel as the visualization of $J$ itself) are shown in figure \ref{fig:tm-urz}.
 
@@
 	\includegraphics[width=0.8\columnwidth]{graphs/TM_model_URZ.eps}
-	\caption{The visualization of the URZ for a given UTI $J$.}
+	\caption{The visualization of the URZ for a given UTI $J$, using the TM framework.}
 	\label{fig:tm-urz}
@@
 
-Given a UTI $J$ and its URZ, every URZ is assigned a symbol and a name and corresponds to a set of Allen relationships.
-The symbol and name are used to uniquely identify each of $J$'s URZ.
-The set of Allen relationships corresponding to an URZ contains every Allen relationship in which the intervals visualized by the URZ's interval points could be with the interval represented by $J$, depending on what time interval $J$ actually represents.
-In figure \ref{fig:tm-urz}, the URZ symbols are also shown.
-In table \ref{tab:urz}, every different row corresponds to a different URZ for a given UTI.
+Given an UTI $J$ and its URZ, every URZ is assigned a symbol and a name and corresponds to a set of Allen relationships.
+These symbols and names are used to uniquely identify each of $J$'s URZ.
+The set of Allen relationships corresponding to an URZ contains every Allen relationship in which any interval visualized by one of the URZ's interval points could be with the interval intended by $J$, depending on exactly which time interval $J$ actually intends to represent.
+In figure \ref{fig:tm-urz}, the URZ's symbols are noted inside the areas visualizing their respective URZ's.
+In table \ref{tab:urz}, every different row corresponds to a different URZ.
 In the first two columns, the symbol and name of the URZ are given, the last column contains the set of corresponding Allen relationships.
@@
 
-The TM framework now allows evaluating all possible Allen relationships between a given crisp time interval $I$ and a given UTI $J$.
+The TM framework now allows evaluating the Allen relationships a given crisp time interval $I = \left[s, e\right]$ could be in with a given UTI $J$, depending on the exact interval intended by $J$.
 This is done as follows.
-First, $J$ and its URZ are visualized.
-Next, $I$'s interval point and its construction half-lines $L_s$ and $L_e$ are drawn in the same interval space.
-Then, evaluation is a matter of positioning (in the following, the possibility that $I$ is in Allen relationship $AR$ with $J$ is denoted  $\Pos(I\ AR\ J)$):
+First, $J$ and its URZ are visualized in an interval space.
+Next, $I$'s interval point, its construction half-lines $L_s$ and $L_e$ and the visualization of its starting and ending instants on the visualization of the time are drawn in the same interval space.
+Then, evaluation is a matter of position (in the following, the possibility that $I$ is the interval intended by $J$ is denoted $\pi_J(I)$ and the possibility that $I$ is in Allen relationship $AR$ with $J$ is denoted  $\Pos(I\ AR\ J)$):
 
 \begin{itemize}
-	\item if $I$'s interval point is located in an URZ corresponding with a singleton of Allen relationships $\{R\}$, it is in relationship $R$ with the interval represented by $J$, whichever this interval is.
+	\item if $I$'s interval point is located in an URZ corresponding with a singleton of Allen relationships $\{R\}$, it is in relationship $R$ with the interval intended by $J$, whichever this interval is.
 Thus: $\Pos(I\ AR\ J) = 1$.
-	\item if $I$'s interval point is located in an URZ corresponding with a non-singleton set of $n$ Allen relationships $\{R_1, R_2, \ldots, R_n\}$, two more half-lines are drawn: one (denoted $L_s'$) from the intersection point of $L_s$ and the visualization of the time axis, its angle with this visualization having size $-\alpha$ and one (denoted $L_e'$) from the intersection point of $L_e$ and the visualization of the time axis, having an angle $\alpha$ with this visualization.
+	\item if $I$'s interval point is located in an URZ corresponding to a non-singleton set of $n$ Allen relationships $\{R_1, R_2, \ldots, R_n\}$, two more half-lines are drawn: one (denoted $L_s'$) from the visualization of $s$ on the visualization of the time axis, having, in this point, an angle of size $-\alpha$ with this visualization and one (denoted $L_e'$) from the visualization of $e$ on the visualization of the time axis, having, in this point, an angle of size $\alpha$ with this visualization.
 Now, some of these half-lines $L_s$, $L_s'$, $L_e$ and $L_e'$ may intersect with the area visualizing $J$.
-These will divide the area visualizing $J$ in $n$ parts $p_i, 1 \leq i \leq n$ (intersection line segments also count as parts) and each of these parts will correspond to an Allen relationship $R_i$.
+These intersection line segments will divide the area visualizing $J$ in $n$ area parts $p_i, 1 \leq i \leq n$ (intersection line segments also count as area parts), each containing only interval points visualizing intervals with a non-zero possibility of being the interval intended by $J$.
+Each of these parts $p_i$ will now correspond to an Allen relationship $R_i$.
 Now, for each $R_i$:
 	\begin{equation}
-		\Pos(I\ R_i\ J) = \sup_{I \in p_i} \pi_J(I)
+		\Pos(I\ R_i\ J) = \sup_{K \in p_i} \pi_J(K)
 	\end{equation}
 \end{itemize}
+
+Here, the intervals $K$ are crisp time intervals and the expression `$K \in p_i$' is a notation to express that $K$ is visualized by an interval point contained in area part $p_i$.
+
+%Time intervals are usually represented in a one dimensional scale.
+This is due to its linear characteristic.
+This approach has also been used in Figure\ref{fig:allen-relationships} and \ref{fig:ill-known-ti}.
+The difficulty of accurate visual analysis on the temporal data arises when the number of time intervals increase.
+
+%The triangular model (TM) ~\cite{Kulpa1997}, \cite{Weghe2007} represents time intervals in a two dimensional space called IRterval point completely determines the start and end of the interval.
+The two-dimensional space of interval points is called the interval space and is denoted by IR ~\cite{Kulpa2006}.
+The procedure to represent a time interval $[t_s, t_e]$ as an interval point in a two dimensional space is the following:
+
+%\begin{enumerate}
+%\item Draw each time point of the interval $[t_s, t_e]$ in the horizontal line.
+%\item Construct one straight line $L_s$ from $t_s$ with an angle $\alpha$.
+%\item Construct one straight line $L_e$ from $t_e$ with an angle $-\alpha$.
+%\item The intersection of these two lines is the interval point.
+%\end{enumerate}
+
+%For the shake of simplicity, in this work, the angle for $\alpha$ is chosen to be $45^{\circ}$.
+
+%In TM, these thirteen Allen relations each correspond to a specific zone in the interval space IR \cite{Kulpa1997}.
+These thirteen zones are called Crisp Relational Zones (CRZs).
+The CRZs with respect to a reference interval point $I$ are depicted in Figure \ref{fig:allencrz}.
+
+%\begin{figure}[h]
+%   \centering
+%   \includegraphics[width=0.8\columnwidth]{graphs/allencrc.eps}
+%   \caption{CRZs corresponding to the thirteen Allen's relations.
+The following shorthand notations are used: equals (E), starts (S), started by (SB), finishes (F), finished by (FB), meets (M), met by (MB), overlaps (O), overlapped by (OB), during (D), contains (C), before (B) and after (A).
+}
+%   \label{fig:allencrz}
+% \end{figure}
+
+%Each CRZ represents the set of interval points of the crisp time intervals that are in the corresponding Allen relation with respect to the reference interval $I$.
+For example, all interval points of intervals that come before $I$ are located in the CRZ represented by the leftmost lower triangle in the interval space.
+Likewise, if an interval point is located in the top quadrangle above $I$, then its corresponding interval contains $I$.
+CRZs allow to visually analyse relations between (large) sets of interval points.
 

### 2013-03-20 develop/excel2sql/src/main/java/es/ugr/decsai/excel2sql/sql/hsqldb/HSqlDBDatatypes.java
Executable version for the excel2sql program

@@ -1,0 +1,16 @@
+/*
+ * To change this template, choose Tools | Templates
+ * and open the template in the editor.
+ */
+package es.ugr.decsai.excel2sql.sql.hsqldb;
+
+/**
+ *
+ * @author Jose Enrique Pons <jpons@decsai.ugr.es>
+ */
+public enum HSqlDBDatatypes {
+    VARCHAR,
+    LONGVARCHAR,
+    INTEGER,
+    DOUBLE
+}

### 2013-03-20 develop/excel2sql/src/main/java/es/ugr/decsai/excel2sql/sql/hsqldb/HSqlDBTypeRender.java

@@ -1,0 +1,37 @@
+/*
+ * To change this template, choose Tools | Templates
+ * and open the template in the editor.
+ */
+package es.ugr.decsai.excel2sql.sql.hsqldb;
+
+import es.ugr.decsai.excel2sql.sql.AbstractSQLRenderer;
+import es.ugr.decsai.excel2sql.sql.SqlDatatypes;
+import static es.ugr.decsai.excel2sql.sql.SqlDatatypes.DOUBLE;
+import static es.ugr.decsai.excel2sql.sql.SqlDatatypes.INTEGER;
+import static es.ugr.decsai.excel2sql.sql.SqlDatatypes.STRING;
+
+/**
+ *
+ * @author Jose Enrique Pons <jpons@decsai.ugr.es>
+ */
+public class HSqlDBTypeRender extends AbstractSQLRenderer{
+
+    @Override
+    public String render(SqlDatatypes type) {
+         String render;
+        switch(type){
+            case DOUBLE: render = HSqlDBDatatypes.DOUBLE.name();
+                break;
+            case INTEGER: render = HSqlDBDatatypes.INTEGER.name();
+                break;
+            case STRING: render = HSqlDBDatatypes.LONGVARCHAR.name();
+                break;
+            default:
+                render = HSqlDBDatatypes.LONGVARCHAR.name();
+        }
+        
+        
+        return render;
+    }
+    
+}

### 2013-03-20 trunk/develop/excel2sql/src/main/java/es/ugr/decsai/excel2sql/App.java
Executable version for the excel2sql program

@@ -2,89 +2,201 @@ package es.ugr.decsai.excel2sql;
 
 import es.ugr.decsai.excel2sql.sql.SqlDatatypes;
 import es.ugr.decsai.excel2sql.sql.SqlFormat;
+import es.ugr.decsai.excel2sql.sql.hsqldb.HSqlDBTypeRender;
 import es.ugr.decsai.excel2sql.sql.sqlite.SqLiteTypeRender;
+import java.io.FileInputStream;
+import java.io.IOException;
 import java.sql.Connection;
 import java.sql.DriverManager;
 import java.sql.Statement;
 import java.util.List;
+import java.util.Properties;
 import java.util.logging.Level;
 import java.util.logging.Logger;
-import org.hibernate.type.Type;
 
 /**
- * Hello world!
+ * Base class application
  *
  */
 public class App {
 
-    public static void main(String[] args) {
-        //  if(args.length >0){
-        //String filename = args[0];
-        String path = "E:\\Users\\Jose\\Documents\\sources\\mr-t\\develop\\excel2sql\\documents\\";
-        String filename = "Output_Test.xlsx";
-        System.out.println("Starting with argument: " + filename);
+    private static final String PROPERTIES_FILENAME = "configuration.properties";
+    private static final String PATH = "path";
+    private static final String INFILE = "input";
+    private static final String OUTFILE = "output";
+    private static final String CREATE = "create";
+    private static final String EXPORT = "export_data";
+    private static final String DRIVER = "driver_name";
+    private static final String CONNECTION = "connection_string";
+    private static final String MAX = "max_items";
 
+    /* 
+     * 
+     * create=true
+     # ¿export the database?
+     export_data=true
+     */
+    public static void main(String[] args) {
+
+        Properties p = new Properties();
         try {
-            ExcelReader er = new ExcelReader(path+filename);
+            p.load(ClassLoader.getSystemResourceAsStream(PROPERTIES_FILENAME));
+            String path = p.getProperty(PATH);
+            String filename = p.getProperty(INFILE);
+            String outputFile = p.getProperty(OUTFILE);
+
+            String driver = p.getProperty(DRIVER);
+            String url = p.getProperty(CONNECTION);
+
+            boolean create = Boolean.valueOf(p.getProperty(CREATE));
+            boolean export = Boolean.valueOf(p.getProperty(EXPORT));
+            
+            int maxItems = Integer.valueOf(p.getProperty(MAX));
+
+
+            System.out.println("Starting with argument: " + filename);
[…]

### 2013-03-20 trunk/develop/excel2sql/src/main/java/es/ugr/decsai/excel2sql/ExcelReader.java

@@ -28,7 +28,8 @@ public class ExcelReader {
 
     Workbook wb;
     private int MAX_ITER = 10;
-    
+    private static final String EMPTY_STRING = "";
+
     /**
      * Open an excel file
      *
@@ -111,29 +112,27 @@ public class ExcelReader {
                 Row row1 = s.getRow(i);
                 Cell cell2 = row1.getCell(j);
 
-        
+
 
                 try {
-                    if(cell2.getCellType()!=Cell.CELL_TYPE_BLANK){
-                    String content = cell2.getRichStringCellValue().getString();
-                    candidateString++;
-                    new Integer(content);
-                    candidateInt++;
-                    new Double(content);
-                    candidateDouble++;
+                    if (cell2.getCellType() != Cell.CELL_TYPE_BLANK) {
+                        String content = cell2.getRichStringCellValue().getString();
+                        if (content.trim().compareTo(EMPTY_STRING) != 0) {
+                            candidateString++;
+                            new Integer(content);
+                            candidateInt++;
+                            new Double(content);
+                            candidateDouble++;
+                        }
                     }
 
                 } catch (Exception ex) {
-            
-                    
                 } finally {
-                    
-
                 }
 
             }
             //check the max:
-           
+
 
             if (candidateInt >= candidateDouble && candidateInt >= candidateString) {
                 typeList.add(SqlDatatypes.INTEGER);
@@ -149,14 +148,14 @@ public class ExcelReader {
         }
         return typeList;
     }
-    
-    public List<String> getRowContents(int index){
+
+    public List<String> getRowContents(int index) {
         List<String> valueList = new ArrayList<String>();
-          Sheet s = wb.getSheetAt(0);
+        Sheet s = wb.getSheetAt(0);
         Row row = s.getRow(index);
 
         Iterator<Cell> cellIterator = row.cellIterator();
-       
+
         while (cellIterator.hasNext()) {
 
             Cell cell = cellIterator.next();
@@ -165,9 +164,9 @@ public class ExcelReader {
         }
         return valueList;
     }
-    
-    public int getRows(){
+
+    public int getRows() {
         Sheet s = wb.getSheetAt(0);
-       return s.getLastRowNum();
+        return s.getLastRowNum();
     }
 }

### 2013-03-20 trunk/develop/excel2sql/src/main/java/es/ugr/decsai/excel2sql/sql/SqlFormat.java

@@ -32,6 +32,14 @@ public class SqlFormat {
     private static final String VALUES = "VALUES";
     private static final String PARAM = "?";
     private static final String EMPTY_STRING = "";
+    private static final String PRIMARY = "PRIMARY";
+    private static final String KEY = "KEY";
+    private static final String ASC = "ASC";
+    private static final String DEC = "DESC";
+    private static final String CACHED = "CACHED";
+    private static final String original = "áàäéèëíìïóòöúùuñÁÀÄÉÈËÍÌÏÓÒÖÚÙÜÑçÇ-";
+    // Cadena de caracteres ASCII que reemplazarán los originales.
+    private static final String ascii = "aaaeeeiiiooouuunAAAEEEIIIOOOUUUNcC ";
     protected AbstractSQLRenderer sqlTypeRender;
 
     public SqlFormat(AbstractSQLRenderer sqlTypeRender) {
@@ -39,10 +47,18 @@ public class SqlFormat {
     }
 
     public String createSentece(String tableName, List<String> fieldName, List<SqlDatatypes> fieldType) {
-        String commandText = TOKEN_CREATE + BLANK + TOKEN_TABLE + BLANK + tableName + BLANK + LEFT_PARENTHESIS;
+        String commandText = TOKEN_CREATE + BLANK + "TEXT" + BLANK + TOKEN_TABLE + BLANK + tableName + BLANK + LEFT_PARENTHESIS;
 
         for (int i = 0; i < fieldName.size(); i++) {
-            commandText += SIMPLE_COMMA + fieldName.get(i) + SIMPLE_COMMA + BLANK + sqlTypeRender.render(fieldType.get(i));
+            //commandText += SIMPLE_COMMA + cleanString(fieldName.get(i)) + SIMPLE_COMMA + BLANK + sqlTypeRender.render(fieldType.get(i));
+            commandText += cleanString(fieldName.get(i)) + BLANK + sqlTypeRender.render(fieldType.get(i));
+
+            //by default first field is primary key:
+            if (i == 0) {
+                commandText += BLANK + PRIMARY + BLANK + KEY + BLANK; //+ ASC+BLANK;
+
+            }
+
             if (i < fieldName.size() - 1) {
                 commandText += COMMA;
             }
@@ -71,7 +87,8 @@ public class SqlFormat {
 
 
         for (int i = 0; i < values.size(); i++) {
-            commandText += sqlTypeRender.renderValue(SqlDatatypes.STRING, attributes.get(i));
[…]

### 2013-03-20 trunk/develop/historyDB/src/main/java/es/ugr/decsai/historydb/App.java
database utils

@@ -1,0 +1,58 @@
+package es.ugr.decsai.historydb;
+
+import es.ugr.decsai.historydb.dao.ContentDao;
+import es.ugr.decsai.historydb.model.Contents;
+import es.ugr.decsai.jpons.uticalculator.Uti;
+import java.util.List;
+import org.hibernate.Query;
+import org.hibernate.Session;
+import org.hibernate.Transaction;
+import org.joda.time.Chronology;
+import org.joda.time.DateTime;
+import org.joda.time.chrono.GJChronology;
+
+/**
+ * Hello world!
+ *
+ */
+public class App {
+
+    public static void main(String[] args) {
+        Session session = HibernateUtil.getSession();
+        Transaction t = session.beginTransaction();
+
+        //Query query = session.createQuery("FROM Contents c WHERE c.dateMillesimeValeur1 >600 AND c.dateMillesimeValeur1 <700");
+//        Query query = session.createQuery("FROM Contents c ");
+//        List list = query.list();
+//        for(Object o:list){
+//            Contents c = (Contents) o;
+//            System.out.println("Contents n: "+c.getNumerowauters()+" year: "+c.getDatemillesimevaleur1());
+//        }
+//        t.commit();
+//        session.close();
+
+        ContentDao dao = new ContentDao(session);
+        Chronology chrono = GJChronology.getInstance();
+        DateTime start = new DateTime(620, 1, 1, 0, 0, chrono);
+        DateTime end = new DateTime(710, 1, 1, 0, 0, chrono);
+
+        DateTime start1 = new DateTime(610, 1, 1, 0, 0, chrono);
+        DateTime start2 = new DateTime(630, 1, 1, 0, 0, chrono);
+
+        DateTime end1 = new DateTime(700, 1, 1, 0, 0, chrono);
+        DateTime end2 = new DateTime(720, 1, 1, 0, 0, chrono);
+
+
+
+
+        Uti queriedTime = new Uti(start1, start, start2, end1, end, end2);
+        List<Contents> contentBetweenDates = dao.getContentBetweenDates(queriedTime);
+        if(contentBetweenDates!=null){
+            for(Contents c:contentBetweenDates){
+                System.out.println("Date: "+c.getDatemillesimevaleur1() + " month: "+c.getDatemoisvaleur1()+ " day "+c.getDatejourvaleur1());
+            }
+        }
+
+
+    }
+}

### 2013-03-20 trunk/develop/historyDB/src/main/java/es/ugr/decsai/historydb/HibernateUtil.java

@@ -1,0 +1,100 @@
+/*
+ * Copyright (C) 2012 University of Granada, Department of Computer Science and Artificial Intelligence.
+ *
+ * This library is free software; you can redistribute it and/or
+ * modify it under the terms of the GNU Lesser General Public
+ * License as published by the Free Software Foundation; either
+ * version 2.1 of the License, or (at your option) any later version.
+ *
+ * This library is distributed in the hope that it will be useful,
+ * but WITHOUT ANY WARRANTY; without even the implied warranty of
+ * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the GNU
+ * Lesser General Public License for more details.
+ *
+ * You should have received a copy of the GNU Lesser General Public
+ * License along with this library; if not, write to the Free Software
+ * Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston,
+ * MA 02110-1301  USA
+ */
+
+package es.ugr.decsai.historydb;
+
+
+import es.ugr.decsai.historydb.model.Contents;
+import java.util.Properties;
+import org.hibernate.Session;
+import org.hibernate.SessionFactory;
+import org.hibernate.cfg.AnnotationConfiguration;
+
+/**
+ *
+ * @author Jose Enrique Pons Frías <jpons@decsai.ugr.es>
+ * First version 27/09/2012
+ * 
+ */
+public class HibernateUtil {
+private static  SessionFactory factory;
+private static final String PROPERTIES_FILENAME = "configuration.properties";
+//private static final String DIALECT="hibernate.dialect";
+//private static final String DRIVER="hibernate.connection.driver_class";
+//private static final String URL="hibernate.connection.url";
+//private static final String USER="hibernate.connection.username";
+//private static final String PWD="hibernate.connection.password";
+//private static final String POOL="hibernate.connection.pool_size";
+//private static final String AUTOCOMMIT="hibernate.connection.autocommit";
+//private static final String HBMDDL="hibernate.hbm2ddl.auto";
+//private static final String SHOWSQL="hibernate.show_sql";
+
+    private static  AnnotationConfiguration config; 
+//= new AnnotationConfiguration().setProperty("hibernate.dialect", "org.hibernate.dialect.HSQLDialect").
+//                setProperty("hibernate.connection.driver_class", "org.hsqldb.jdbcDriver").
[…]

### 2013-03-20 trunk/develop/historyDB/src/main/java/es/ugr/decsai/historydb/util/ContentsUtil.java

@@ -1,0 +1,145 @@
+/*
+ * To change this template, choose Tools | Templates
+ * and open the template in the editor.
+ */
+package es.ugr.decsai.historydb.util;
+
+import es.ugr.decsai.historydb.model.Contents;
+import es.ugr.decsai.jpons.uticalculator.Uti;
+import java.lang.reflect.Field;
+import java.lang.reflect.Method;
+import java.util.ArrayList;
+import java.util.List;
+import org.joda.time.Chronology;
+import org.joda.time.DateTime;
+import org.joda.time.chrono.GJChronology;
+
+/**
+ *
+ * @author Jose Enrique Pons <jpons@decsai.ugr.es>
+ */
+public class ContentsUtil {
+
+    public static Uti buildUtiFromContent(Contents c) {
+        if (c == null) {
+            return null;
+        }
+        List<DateTime> dateTimeFromContents = getDateTimeFromContents(c);
+        DateTime min, max;
+        min = max = dateTimeFromContents.get(0);
+
+
+
+
+        for (DateTime t : dateTimeFromContents) {
+            if (min.isAfter(t)) {
+                min = t;
+            }
+            if (max.isBefore(t)) {
+                max = t;
+            }
+        }
+
+
+
+        Uti u = new Uti(min.minusSeconds(1000), min, min.plusSeconds(1000), max.minusSeconds(1000), max, max.plusSeconds(1000));
+
+        return u;
+    }
+
+    public static List<DateTime> getDateTimeFromContents(Contents c) {
+        if (c == null) {
+            return null;
+        }
+
+
+        Class<?> clazz = c.getClass();
+        Chronology chrono = GJChronology.getInstance();
+        List<DateTime> dtList = new ArrayList<DateTime>(15);
+
+        int d1, m1, y1;
+
+        try {
+            y1 = c.getDatemillesimevaleur1();
+        } catch (NullPointerException ex) {
+            y1 = 0;
+        }
+        try {
+            m1 = c.getDatemoisvaleur1();
+            if (m1 < 1 || m1 > 12) {
+                m1 = 1;
+            }
+        } catch (NullPointerException ex) {
+            m1 = 1;
+        }
+
+        try {
+            d1 = c.getDatejourvaleur1();
+            if (d1 < 1 || d1 > 31) {
+                d1 = 1;
+            }
+        } catch (NullPointerException ex) {
+            d1 = 1;
+        }
+
[…]

### 2013-03-20 develop/historyDB/src/main/java/es/ugr/decsai/historydb/dao/ContentDao.java
database utils

@@ -1,0 +1,49 @@
+/*
+ * To change this template, choose Tools | Templates
+ * and open the template in the editor.
+ */
+package es.ugr.decsai.historydb.dao;
+
+import es.ugr.decsai.historydb.model.Contents;
+import es.ugr.decsai.jpons.uticalculator.Uti;
+import java.util.List;
+import org.hibernate.Query;
+import org.hibernate.Session;
+import org.joda.time.Chronology;
+import org.joda.time.DateTime;
+import org.joda.time.chrono.GJChronology;
+
+/**
+ * Class to deal with Contents
+ * @author Jose Enrique Pons <jpons@decsai.ugr.es>
+ */
+public class ContentDao {
+    private  Session session;
+    
+   public ContentDao(Session s){
+        this.session =s;
+    }
+    
+    
+    public List<Contents> getContentBetweenDates(Uti queriedTime){
+        if(queriedTime==null){
+            return null;
+        }
+//         Chronology chrono = GJChronology.getInstance();
+//        DateTime start = new DateTime();
+//        start.withChronology(chrono);
+//        start.withMillis(((Double) queriedTime.getStart1()).longValue());
+//        DateTime end = new DateTime();
+//        end.withChronology(chrono);
+//        end.withMillis(((Double) queriedTime.getEnd2()).longValue());
+        Query query = session.createQuery("SELECT c FROM Contents c WHERE c.datemillesimevaleur1 > :start AND c.datemillesimevaleur1 < :end ");
+        query.setParameter("start", queriedTime.getDtStart1().getYear());
+        query.setParameter("end", queriedTime.getDtEnd().getYear());
+        List<Contents> contentList = query.list();
+        
+        return contentList;
+        
+        
+    }
+    
+}

### 2013-03-20 develop/historyDB/src/main/java/es/ugr/decsai/historydb/model/Contents.java

@@ -1,0 +1,4803 @@
+/*
+ * To change this template, choose Tools | Templates
+ * and open the template in the editor.
+ */
+package es.ugr.decsai.historydb.model;
+
+import java.io.Serializable;
+import javax.persistence.Basic;
+import javax.persistence.Column;
+import javax.persistence.Entity;
+import javax.persistence.Id;
+import javax.persistence.NamedQueries;
+import javax.persistence.NamedQuery;
+import javax.persistence.Table;
+import javax.xml.bind.annotation.XmlRootElement;
+
+/**
+ *
+ * @author Jose Enrique Pons <jpons@decsai.ugr.es>
+ */
+@Entity
+@Table(name = "CONTENTS")
+@XmlRootElement
+@NamedQueries({
+    @NamedQuery(name = "Contents.findAll", query = "SELECT c FROM Contents c")})
+public class Contents implements Serializable {
+    private static final long serialVersionUID = 1L;
+    @Id
+    @Basic(optional = false)
+    @Column(name = "NUMEROWAUTERS")
+    private Integer numerowauters;
+    @Column(name = "GENRE")
+    private String genre;
+    @Column(name = "LANGUE")
+    private String langue;
+    @Column(name = "VERACITE")
+    private String veracite;
+    @Column(name = "DATEMILLESIMEVALEUR1")
+    private Integer datemillesimevaleur1;
+    @Column(name = "DATEMOISVALEUR1")
+    private Integer datemoisvaleur1;
+    @Column(name = "DATEJOURVALEUR1")
+    private Integer datejourvaleur1;
+    @Column(name = "DATECODESVALEUR1")
+    private Integer datecodesvaleur1;
+    @Column(name = "DATEJUSTIFICATIONSVALEUR1")
+    private String datejustificationsvaleur1;
+    @Column(name = "MIXDAT1VALEUR1")
+    private Integer mixdat1valeur1;
+    @Column(name = "MIXDAT2VALEUR1")
+    private Integer mixdat2valeur1;
+    @Column(name = "MIXDAT3VALEUR1")
+    private Integer mixdat3valeur1;
+    @Column(name = "MIXDATCCODEVALEUR1")
+    private Integer mixdatccodevaleur1;
[…]

### 2013-03-20 develop/historyDB/src/main/java/es/ugr/decsai/historydb/model/Test.java

@@ -1,0 +1,83 @@
+/*
+ * To change this template, choose Tools | Templates
+ * and open the template in the editor.
+ */
+package es.ugr.decsai.historydb.model;
+
+import java.io.Serializable;
+import javax.persistence.Column;
+import javax.persistence.Entity;
+import javax.persistence.Id;
+import javax.persistence.NamedQueries;
+import javax.persistence.NamedQuery;
+import javax.persistence.Table;
+import javax.xml.bind.annotation.XmlRootElement;
+
+/**
+ *
+ * @author Jose Enrique Pons <jpons@decsai.ugr.es>
+ */
+@Entity
+@Table(name = "test")
+@XmlRootElement
+@NamedQueries({
+    @NamedQuery(name = "Test.findAll", query = "SELECT t FROM Test t"),
+    @NamedQuery(name = "Test.findByR1", query = "SELECT t FROM Test t WHERE t.r1 = :r1"),
+    @NamedQuery(name = "Test.findByR2", query = "SELECT t FROM Test t WHERE t.r2 = :r2")})
+public class Test implements Serializable {
+    private static final long serialVersionUID = 1L;
+    @Id
+    @Column(name = "r1")
+    private Integer r1;
+    @Column(name = "r2")
+    private String r2;
+
+    public Test() {
+    }
+
+    public Test(Integer r1) {
+        this.r1 = r1;
+    }
+
+    public Integer getR1() {
+        return r1;
+    }
+
+    public void setR1(Integer r1) {
+        this.r1 = r1;
+    }
+
+    public String getR2() {
+        return r2;
+    }
+
+    public void setR2(String r2) {
+        this.r2 = r2;
+    }
+
+    @Override
+    public int hashCode() {
+        int hash = 0;
+        hash += (r1 != null ? r1.hashCode() : 0);
+        return hash;
+    }
+
+    @Override
+    public boolean equals(Object object) {
+        // TODO: Warning - this method won't work in the case the id fields are not set
+        if (!(object instanceof Test)) {
+            return false;
+        }
+        Test other = (Test) object;
+        if ((this.r1 == null && other.r1 != null) || (this.r1 != null && !this.r1.equals(other.r1))) {
+            return false;
+        }
+        return true;
+    }
+
+    @Override
+    public String toString() {
+        return "es.ugr.decsai.historydb.model.Test[ r1=" + r1 + " ]";
+    }
+    
+}

### 2013-03-20 develop/historyDB/src/test/java/es/ugr/decsai/historydb/AppTest.java

@@ -1,0 +1,38 @@
+package es.ugr.decsai.historydb;
+
+import junit.framework.Test;
+import junit.framework.TestCase;
+import junit.framework.TestSuite;
+
+/**
+ * Unit test for simple App.
+ */
+public class AppTest 
+    extends TestCase
+{
+    /**
+     * Create the test case
+     *
+     * @param testName name of the test case
+     */
+    public AppTest( String testName )
+    {
+        super( testName );
+    }
+
+    /**
+     * @return the suite of tests being tested
+     */
+    public static Test suite()
+    {
+        return new TestSuite( AppTest.class );
+    }
+
+    /**
+     * Rigourous Test :-)
+     */
+    public void testApp()
+    {
+        assertTrue( true );
+    }
+}

### 2013-03-20 trunk/develop/utiCalculator/src/main/java/es/ugr/decsai/jpons/uticalculator/Uti.java
update uti calculator

@@ -20,6 +20,7 @@ package es.ugr.decsai.jpons.uticalculator;
 
 import java.util.ArrayList;
 import java.util.List;
+import org.joda.time.DateTime;
 
 /**
  * Uncertain time interval
@@ -31,7 +32,7 @@ import java.util.List;
 public class Uti {
 
     /**
-     * Coordinates in the cartesian model
+     * Coordinates in the time line
      */
     protected double start;
     protected double end;
@@ -39,6 +40,46 @@ public class Uti {
     protected double start2;
     protected double end1;
     protected double end2;
+    
+    DateTime dtStart;
+    DateTime dtStart1;
+    DateTime dtStart2;
+    DateTime dtEnd;
+    DateTime dtEnd1;
+    DateTime dtEnd2;
+    // screen coordinates:
+    double scs;
+    double scs1;
+    double scs2;
+    double sce;
+    double sce1;
+    double sce2;
+    
+    protected void toScreenCoords(){
+        if(dtStart == null 
+                || dtStart1 == null
+                || dtStart2 == null
+                || dtEnd == null
+                || dtEnd1 == null
+                || dtEnd2 == null){
+            return;
+        }
+        scs = convertUnits(dtStart);
+        scs1 = convertUnits(dtStart1);
+        scs2 = convertUnits(dtStart2);
+        
+        sce = convertUnits(dtEnd);
+        sce1 = convertUnits(dtEnd1);
+        sce2 = convertUnits(dtEnd2);
+        
+    }
+    
+    protected double convertUnits(DateTime t){
+        double cents = t.getDayOfYear()*100/365;
+        return t.getYear() + 0.01 * cents;
+        
+    }
+    
     protected  int iterations = 100;
     
     public void setIterations(int iter){
@@ -56,6 +97,24 @@ public class Uti {
         return (b - a) / 2;
     }
 
+    
+    public double getTx(){
+        return getTx(start, end);
+    }
+    
+    public double getTy(){
+        return getTy(start,end);
+    }
+    
+    public double getScreenX(){
+        return getTx(scs,sce);
+    }
+    
+    public double getScreenY(){
+        return getTy(scs,sce);
+    }
+    
+    
     /**
      * Function that computes a triangular membership function
      *
@@ -96,13 +155,48 @@ public class Uti {
      * @param end1
      * @param end2
      */
[…]

### 2013-03-20 develop/utiCalculator/src/main/java/es/ugr/decsai/jpons/uticalculator/TriangularDistribution.java
update uti calculator

@@ -1,0 +1,47 @@
+/*
+ * To change this template, choose Tools | Templates
+ * and open the template in the editor.
+ */
+package es.ugr.decsai.jpons.uticalculator;
+
+/**
+ *
+ * @author Jose Enrique Pons <jpons@decsai.ugr.es>
+ */
+public class TriangularDistribution<T extends Number> {
+    T leftSide;
+    T core;
+    T rightSide;
+
+    public TriangularDistribution(T leftSide, T core, T rightSide) {
+        this.leftSide = leftSide;
+        this.core = core;
+        this.rightSide = rightSide;
+    }
+
+    public T getLeftSide() {
+        return leftSide;
+    }
+
+    public void setLeftSide(T leftSide) {
+        this.leftSide = leftSide;
+    }
+
+    public T getCore() {
+        return core;
+    }
+
+    public void setCore(T core) {
+        this.core = core;
+    }
+
+    public T getRightSide() {
+        return rightSide;
+    }
+
+    public void setRightSide(T rightSide) {
+        this.rightSide = rightSide;
+    }
+
+  
+}

### 2013-03-20 develop/utiCalculator/src/main/java/es/ugr/decsai/jpons/uticalculator/Triplet.java

@@ -1,0 +1,36 @@
+/*
+ * To change this template, choose Tools | Templates
+ * and open the template in the editor.
+ */
+package es.ugr.decsai.jpons.uticalculator;
+
+import java.util.Comparator;
+
+/**
+ * Class for dealing with triplets of numbers
+ * @author Jose Enrique Pons <jpons@decsai.ugr.es>
+ */
+public class Triplet<T extends Number> implements Comparable<Triplet>{
+    T a,b,c; 
+    public int compareTo(Triplet o) {
+//       if(o==null){
+//           return -1;
+//       }
+//       int ca,cb,cc;
+//       
+//       ca = cb = cc = 0;
+//       if(a.doubleValue()<o.a.doubleValue()){
+//           ca++;
+//       }else{
+//           ca++;
+//       }
+//       
+//       if(b!=null){
+//           if(o.b==null)
+//           
+//           
+//       }
+//        
+    return 0;}
+    
+}

### 2013-03-20 develop/utiCalculator/src/main/java/es/ugr/decsai/jpons/uticalculator/UtiUtil.java

@@ -1,0 +1,75 @@
+/*
+ * To change this template, choose Tools | Templates
+ * and open the template in the editor.
+ */
+package es.ugr.decsai.jpons.uticalculator;
+
+import java.util.List;
+import org.joda.time.Chronology;
+import org.joda.time.DateTime;
+import org.joda.time.chrono.GJChronology;
+
+/**
+ * Utility functions to deal with utis
+ * @author Jose Enrique Pons <jpons@decsai.ugr.es>
+ */
+public  class UtiUtil {
+    
+   public static TriangularDistribution getDistributionFromTime(Integer day,Integer month,Integer year){
+      // double leftSide, core,rightSide;
+       DateTime leftSide = null, core = null,rightSide = null;
+       Chronology chrono = GJChronology.getInstance();
+       if(year!=null){
+           if(month==null && day==null){
+               leftSide = new DateTime(year, 1, 1, 0, 0,0,chrono);
+               core = new DateTime(year,6,1,0,0,0,chrono);
+               rightSide = new DateTime(year,12,31,0,0,0,chrono);
+                       
+           }else if(month!=null && day ==null){
+               leftSide = new DateTime(year, month, 1, 0, 0,0,chrono);
+               int maximumValue = leftSide.dayOfMonth().getMaximumValue();
+               core = new DateTime(year,month,15,0,0,0,chrono);
+               rightSide = new DateTime(year,month,maximumValue,0,0,0,chrono);
+               
+           }else if(month!=null && day!=null){
+               leftSide = new DateTime(year, month, day, 0, 0,0,chrono);
+               core = new DateTime(year,month, day,12,0,0,chrono);
+               rightSide = new DateTime(year,month, day,23,59,59,chrono);
+               
+           }
+       }
+       
+       TriangularDistribution<Double> t = new TriangularDistribution(leftSide.getMillis(), core.getMillis(), rightSide.getMillis());
+       return t;
+   }
+   
+   public static Uti utiBuilder(Integer sday,Integer smonth,Integer syear,Integer eday,Integer emonth,Integer eyear ){
+       TriangularDistribution<Double> start, end;
+       
+       start = getDistributionFromTime(sday,smonth, syear);
+       end = getDistributionFromTime(eday,emonth, eyear);
+       
+       Uti result = new Uti(start.leftSide, start.core, start.rightSide, end.leftSide, end.core, end.rightSide);
+       return result;
+   }
+   
[…]

### 2013-03-20 develop/UtiDisplayFX/src/es/ugr/decsai/tm/UtiChart.java
display uti in java fx

@@ -1,0 +1,216 @@
+/*
+ * Copyright (C) 2013 Jose Enrique Pons <jpons@decsai.ugr.es>
+ *
+ * This program is free software: you can redistribute it and/or modify
+ * it under the terms of the GNU General Public License as published by
+ * the Free Software Foundation, either version 3 of the License, or
+ * (at your option) any later version.
+ *
+ * This program is distributed in the hope that it will be useful,
+ * but WITHOUT ANY WARRANTY; without even the implied warranty of
+ * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ * GNU General Public License for more details.
+ *
+ * You should have received a copy of the GNU General Public License
+ * along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+package es.ugr.decsai.tm;
+
+import es.ugr.decsai.jpons.uticalculator.Uti;
+import java.util.ArrayList;
+import java.util.List;
+import javafx.geometry.Point2D;
+import javafx.scene.Group;
+import javafx.scene.Scene;
+import javafx.scene.chart.NumberAxis;
+import javafx.scene.chart.ScatterChart;
+import javafx.scene.chart.XYChart;
+import javafx.scene.paint.Color;
+import javafx.scene.paint.CycleMethod;
+import javafx.scene.paint.RadialGradient;
+import javafx.scene.paint.RadialGradientBuilder;
+import javafx.scene.paint.Stop;
+import javafx.scene.shape.Ellipse;
+import javafx.stage.Stage;
+
+/**
+ * A class for visualization of utis in javafx
+ *
+ * @author Jose Enrique Pons <jpons@decsai.ugr.es>
+ */
+public class UtiChart {
+
+    ScatterChart<Number, Number> chart;
+    ScatterChart.Series<Number, Number> series;
+    NumberAxis xAxis;
+    NumberAxis yAxis;
+    List<Uti> utiList;
+    List<Ellipse> ellipseList;
+    Uti queriedTime;
+
+    public Uti getQueriedTime() {
+        return queriedTime;
+    }
+
+    public void setQueriedTime(Uti queriedTime) {
+        this.queriedTime = queriedTime;
+        series.getData().add(new ScatterChart.Data<Number, Number>(queriedTime.getScreenX(), queriedTime.getScreenY()));
+    }
+
+    private void init() {
+        xAxis = new NumberAxis();
+        yAxis = new NumberAxis();
[…]

### 2013-03-20 trunk/develop/UtiDisplayFX/src/es/ugr/decsai/tm/AbstractUti.java
display uti in java fx

@@ -1,0 +1,27 @@
+/*
+ * Copyright (C) 2013 Jose Enrique Pons <jpons@decsai.ugr.es>
+ *
+ * This program is free software: you can redistribute it and/or modify
+ * it under the terms of the GNU General Public License as published by
+ * the Free Software Foundation, either version 3 of the License, or
+ * (at your option) any later version.
+ *
+ * This program is distributed in the hope that it will be useful,
+ * but WITHOUT ANY WARRANTY; without even the implied warranty of
+ * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ * GNU General Public License for more details.
+ *
+ * You should have received a copy of the GNU General Public License
+ * along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+package es.ugr.decsai.tm;
+
+import javafx.scene.shape.Shape;
+
+/**
+ *
+ * @author Jose Enrique Pons <jpons@decsai.ugr.es>
+ */
+public abstract class AbstractUti extends Shape{
+    
+}

### 2013-03-20 trunk/develop/UtiDisplayFX/src/es/ugr/decsai/tm/EllipseUti.java

@@ -1,0 +1,32 @@
+/*
+ * Copyright (C) 2013 Jose Enrique Pons <jpons@decsai.ugr.es>
+ *
+ * This program is free software: you can redistribute it and/or modify
+ * it under the terms of the GNU General Public License as published by
+ * the Free Software Foundation, either version 3 of the License, or
+ * (at your option) any later version.
+ *
+ * This program is distributed in the hope that it will be useful,
+ * but WITHOUT ANY WARRANTY; without even the implied warranty of
+ * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ * GNU General Public License for more details.
+ *
+ * You should have received a copy of the GNU General Public License
+ * along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+package es.ugr.decsai.tm;
+
+import com.sun.javafx.geom.Shape;
+
+/**
+ *
+ * @author Jose Enrique Pons <jpons@decsai.ugr.es>
+ */
+public class EllipseUti extends AbstractUti{
+
+    @Override
+    public Shape impl_configShape() {
+        throw new UnsupportedOperationException("Not supported yet."); //To change body of generated methods, choose Tools | Templates.
+    }
+    
+}

### 2013-03-20 trunk/develop/UtiDisplayFX/src/utidisplayfx/UtiDisplayFX.java

@@ -1,0 +1,223 @@
+/*
+ * To change this template, choose Tools | Templates
+ * and open the template in the editor.
+ */
+package utidisplayfx;
+
+import es.ugr.decsai.historydb.HibernateUtil;
+import es.ugr.decsai.historydb.dao.ContentDao;
+import es.ugr.decsai.historydb.model.Contents;
+import es.ugr.decsai.historydb.util.ContentsUtil;
+import es.ugr.decsai.jpons.uticalculator.Uti;
+import es.ugr.decsai.tm.UtiChart;
+import java.util.ArrayList;
+import java.util.List;
+import javafx.animation.SequentialTransition;
+import javafx.application.Application;
+import javafx.scene.chart.NumberAxis;
+import javafx.scene.chart.ScatterChart;
+import javafx.stage.Stage;
+import org.hibernate.Session;
+import org.joda.time.Chronology;
+import org.joda.time.DateTime;
+import org.joda.time.chrono.GJChronology;
+
+/**
+ *
+ * @author Jose
+ */
+public class UtiDisplayFX extends Application {
+
+    private ScatterChart.Series<Number, Number> series;
+    private double nextX = 0;
+    private SequentialTransition animation;
+    final NumberAxis xAxis = new NumberAxis(0, 10, 1);
+    final NumberAxis yAxis = new NumberAxis(0, 10, 1);
+    final ScatterChart<Number, Number> sc = new ScatterChart<Number, Number>(xAxis, yAxis);
+
+    protected ScatterChart<Number, Number> createChart() {
+
+        xAxis.setForceZeroInRange(false);
+
+
+        // setup chart
+        sc.setId("liveScatterChart");
+        sc.setTitle("UTI visualizer");
+        xAxis.setLabel("X Axis");
+        xAxis.setAnimated(false);
+        yAxis.setLabel("Y Axis");
+        yAxis.setAutoRanging(false);
+
+
+        // add starting data
+        series = new ScatterChart.Series<Number, Number>();
+        series.setName("Sine Wave");
+        series.getData().add(new ScatterChart.Data<Number, Number>(5d, 5d));
+        series.getData().add(new ScatterChart.Data<Number, Number>(0d, 0d));
+        sc.getData().add(series);
+
+
+
+
+
+
+        return sc;
+    }
+
+    @Override
+    public void start(Stage primaryStage) {
+        Session s = HibernateUtil.getSession();
+        s.beginTransaction();
+
[…]

### 2013-03-21 develop/JavaFXApplication4/src/javafxapplication4/JavaFXApplication4.java
display uti in java fx

@@ -1,0 +1,40 @@
+/*
+ * To change this template, choose Tools | Templates
+ * and open the template in the editor.
+ */
+package javafxapplication4;
+
+import javafx.application.Application;
+import javafx.fxml.FXMLLoader;
+import javafx.scene.Parent;
+import javafx.scene.Scene;
+import javafx.stage.Stage;
+
+/**
+ *
+ * @author Jose Enrique Pons <jpons@decsai.ugr.es>
+ */
+public class JavaFXApplication4 extends Application {
+    
+    @Override
+    public void start(Stage stage) throws Exception {
+        Parent root = FXMLLoader.load(getClass().getResource("Sample.fxml"));
+        
+        Scene scene = new Scene(root);
+        
+        stage.setScene(scene);
+        stage.show();
+    }
+
+    /**
+     * The main() method is ignored in correctly deployed JavaFX application.
+     * main() serves only as fallback in case the application can not be
+     * launched through deployment artifacts, e.g., in IDEs with limited FX
+     * support. NetBeans ignores main().
+     *
+     * @param args the command line arguments
+     */
+    public static void main(String[] args) {
+        launch(args);
+    }
+}

### 2013-03-21 develop/JavaFXApplication4/src/javafxapplication4/SampleController.java

@@ -1,0 +1,91 @@
+/*
+ * To change this template, choose Tools | Templates
+ * and open the template in the editor.
+ */
+package javafxapplication4;
+
+import es.ugr.decsai.jpons.uticalculator.Uti;
+import java.net.URL;
+import java.util.ResourceBundle;
+import javafx.fxml.FXML;
+import javafx.fxml.Initializable;
+import javafx.scene.chart.NumberAxis;
+import javafx.scene.chart.ScatterChart;
+import javafx.scene.control.Label;
+import javafx.scene.control.Slider;
+import javafx.scene.input.MouseEvent;
+import org.joda.time.Chronology;
+import org.joda.time.DateTime;
+import org.joda.time.chrono.GJChronology;
+
+/**
+ *
+ * @author Jose Enrique Pons <jpons@decsai.ugr.es>
+ */
+public class SampleController implements Initializable {
+    
+    @FXML
+    private Label label;
+    @FXML
+    private Slider startYear;
+    @FXML
+    private Slider endYear;
+    @FXML
+    private ScatterChart<Number, Number> sc;
+    
+    ScatterChart.Series<Number, Number> series;
+    @FXML
+    NumberAxis xAxis;
+    @FXML
+    NumberAxis yAxis;
+    Uti queriedTime;
+    
+     
+    
+    @Override
+    public void initialize(URL url, ResourceBundle rb) {
+        // TODO
+//        xAxis = new NumberAxis();
+//        yAxis = new NumberAxis();
+////        sc = new ScatterChart<Number,Number>(xAxis,yAxis);
+        series = new ScatterChart.Series<>();
+         Chronology chrono = GJChronology.getInstance();
+        DateTime start = new DateTime(600,1,1,0,0,chrono);
+        DateTime start1 = start.minusDays(100);
+        DateTime start2 = start.plusDays(100);
+        DateTime end = start.plusYears(10);
+        DateTime end1 = end.minusDays(100);
+        DateTime end2 = end.plusDays(100);
+        queriedTime = new Uti(start1,start, start2, end1, end, end2);
+        
+         series.getData().add(new ScatterChart.Data<Number, Number>(queriedTime.getScreenX(), queriedTime.getScreenY()));
+         sc.getData().add(series);
+    }    
+
+    @FXML
+    private void startDragDone3(MouseEvent event) {
+        createUti(startYear.getValue(), endYear.getValue());
+    }
+
+    @FXML
+    private void onMouseEnd(MouseEvent event) {
[…]

### 2013-03-23 eusflat2013/introduction.tex
Almost the last changes! Text is too long though :s

@@
 Information systems (IS) have always tried to model parts of reality.
-To achieve this modelling, such systems contain data representing properties of real-world objects or concepts.
-As time is an essential aspect of many real-world objects or concepts, information systems often contain data representing temporal notions which describe such temporal properties.
-Such temporal notions usually take the form of either instants~\cite{Dyreson1994},~\cite{Jensen1998}, which can informally be seen as infinitely short `moments' or `points' in time, or time intervals~\cite{Dyreson1994},~\cite{Jensen1998}.
+To achieve this modelling, such systems contain data representing properties of real-world objects or concepts~\cite{JoseEnriquePons2012},~\cite{Billiet2012}.
+As time is an essential aspect of many real-world objects or concepts, information systems often contain data representing temporal notions which describe such temporal properties~\cite{Bolour1982},~\cite{VanderCruyssen1997}.
+Such temporal notions usually take the form of either instants~\cite{Jensen1998}, which can informally be seen as infinitely short `moments' or `points' in time, or time intervals~\cite{Jensen1998}.
 
 Data are often produced by humans, but human-made data are prone to imperfections: some data may be vague, imprecise, incomplete, contradictory or uncertain.
-Data representing temporal notions may contain such imprecisions too.
+Data representing temporal notions may contain such imprecisions too~\cite{Devos1994},~\cite{Dubois2003},~\cite{JoseEnriquePons2012},~\cite{Billiet2012}.
 The work presented in this paper is specifically concerned with time intervals (and as a special case: instants) subject to uncertainty.
@@
 Of course, when temporal information is represented in an IS, querying this IS may have a temporal aspect too.
-Usually, querying an IS containing data representing temporal notions is conceptually done by specifying one or more time indications and requesting information that is in a specific relationship with these indications, where the semantics of these relationships are specifically temporal.
-Thus, some existing proposals have considered groups of basic relationships between time indications used to construct and express specific temporal relationships.
+Usually, querying an IS containing data representing temporal notions is conceptually done by specifying one or more time indications and requesting information that is in a specific relationship with these indications, where the semantics of these relationships are specifically temporal~\cite{Billiet2012},~\cite{JoseEnriquePons2012},~\cite{Pons2012a}.
+Thus, some existing proposals have considered groups of basic relationships between time indications used to construct and express specific temporal relationships~\cite{Medina1994},~\cite{Schockaert2008}.
 Notably, Allen~\cite{Allen1983} presented a reasoning framework containing all semantically usefull basic temporal relationships between time intervals (and as a special case instants).
@@
    \centering
-   \includegraphics[width=0.8\columnwidth]{graphs/allen.eps}
+   \includegraphics[width=0.9\columnwidth]{graphs/allen.eps}
    \caption{The Allen relationships between two crisp intervals $I$ and $J$.  }
@@
 
-To be able to query IS containing data representing time indications subject to uncertainty, a framework is necessary, able to represent uncertainties in time indications in a semantically sound way, without (much) information loss and able to temporally reason with such time indications in a semantically sound and usefull way.
+To be able to query IS containing data representing time indications subject to uncertainty, a framework is necessary, able to represent uncertainties in time indications in a semantically sound way, without (much) information loss and able to temporally reason with such time indications in a semantically sound and usefull way~\cite{Dubois1983},~\cite{Dubois2003}.
 Although more proposals for such frameworks exist, the work presented in this paper focusses on just two: the \emph{ill-known constraint \emph{(IKC)} framework}~\cite{Pons2011} and the \emph{triangular model \emph{(TM)} framework}~\cite{DeTre2012}.

### 2013-03-23 eusflat2013/tm.tex
Almost the last changes! Text is too long though :s

--- 2013-03-20 trunk/eusflat2013/tm.tex
@@
 \subsection{\label{subsec:tm-preliminaries}Preliminary Concepts}
-In the triangular model framework, the approaches to interval representation and Allen relationship evaluation are closely linked with the approach to interval visualization.
-
+In the triangular model framework, the approaches to interval representation and Allen relationship evaluation are closely related to the approach to interval visualization.
 Usually, different time intervals are visualized as different parallel line segments in the same image plane.
-However, as this linear approach introduces a few issues, time intervals are visualized as points in the image plane in the TM framework.
-
+However, as this linear approach introduces a few issues, time intervals are visualized as points in the image plane in the TM framework~\cite{Weghe2007},~\cite{DeTre2012}.
 To achieve this, a horizontal line segment visualizing the used (part of the) time axis is drawn in the image plane.
@@
 The area contained in this triangle is called the \emph{interval space} and will contain all visualizations of intervals.
-Now, the procedure used to visualize a crisp time interval $\left[s, e\right]$ in the interval space is the following:
+Now, the procedure used to visualize a CTI $\left[s, e\right]$ in the interval space is the following~\cite{Weghe2007},~\cite{DeTre2012}:
 
@@
 
-A similar point visualizing an interval is generally called an \emph{interval point}.
+A similar point visualizing an interval is generally called an \emph{interval point}~\cite{Weghe2007},~\cite{DeTre2012}.
 In this work, as it is usually done, the size of $\alpha$ is chosen to be $45^{\circ}$.
@@
 	\centering
-	\includegraphics[width=0.8\columnwidth]{graphs/TM_model_several.eps}
-	\caption{The visualization of several crisp time intervals using the TM framework.}
+	\includegraphics[width=0.9\columnwidth]{graphs/TM_model_several.eps}
+	\caption{The visualization of several CTI using the TM framework.}
 	\label{fig:tm-const-ex}
@@
 \subsection{\label{subsec:tm-interval}Representation of Time Intervals Subject to Uncertainty}
-The TM framework allows uncertainty in time intervals by supporting uncertain time intervals:
+The TM framework allows uncertainty in time intervals by supporting uncertain time intervals~\cite{DeTre2012}:
 
@@
 Given an instant $x \in T$, $\pi_s(x)$ now expresses the possibility that $x$ is the starting instant of $J$ and $\pi_e(x)$ expresses the possibility that $x$ is the ending instant of $J$.
-Thus, $J$ intends to indicate just one crisp time interval, but due to a (partial) lack of knowledge, it is (partially) unknown exactly which time interval this is.
-Because $\pi_s$ and $\pi_e$ are required to be convex, the possibility $\pi_J(I)$ that a given crisp time interval $I = \left[s_i, e_i\right]$ is the exact time interval intended by $J$ can now be uniquely determined by:
+Thus, $J$ intends to indicate just one CTI, but due to a (partial) lack of knowledge, it is (partially) unknown exactly which time interval this is.
+Because $\pi_s$ and $\pi_e$ are required to be convex, the possibility $\pi_J(I)$ that a given CTI $I = \left[s_i, e_i\right]$ is the exact time interval intended by $J$ can now be uniquely determined by~\cite{DeTre2012}:
 
@@
 An UTI $J$ in $T$ defined by a pair $(\pi_s, \pi_e)$ is now visualized as an area in the image plane.
-This area is constructed as follows: for every crisp time interval $K$ for which the possibility $\pi_J(K)$ of $K$ being the time interval represented by $J$ is higher than zero, the interval point is drawn and is given a color.
+This area is constructed as follows: for every CTI $K$ for which the possibility $\pi_J(K)$ of $K$ being the time interval represented by $J$ is higher than zero, the interval point is drawn and is given a color.
 This color is part of a linear single-color greyscale and its saturation is linearly related to the degree of possibility that $K$ is the time interval intended by $J$.
 Thus, every drawn interval point in this visualization corresponds to an interval with a higher-than-zero possibility of being the time interval intended by $J$.
-Together, these drawn interval points form the visualization of $J$.
+Together, these drawn interval points form the visualization of $J$~\cite{DeTre2012}.
 An example of such construction is shown in figure \ref{fig:tm-ill-const-ex}.
@@
 	\centering
-	\includegraphics[width=0.8\columnwidth]{graphs/TM_model_ill_known.eps}
+	\includegraphics[width=0.9\columnwidth]{graphs/TM_model_ill_known.eps}
 	\caption{The visualization of one UTI defined by a pair $(\pi_s, \pi_e)$ using the TM framework.}
@@
 \subsection{\label{subsec:tm-evaluation}Evaluation of Allen Relationships}
-At the base of the framework's approach to evaluating Allen relationships is the concept of uncertain relational zones:
+At the base of the framework's approach to evaluating Allen relationships is the concept of uncertain relational zones~\cite{DeTre2012}:
 
@@
 
-Now, given an UTI $J$ and its visualization in an interval space, the procedure to visualize its URZ is the following:
+Now, given an UTI $J$ and its visualization in an interval space, the procedure to visualize its URZ is the following~\cite{DeTre2012}:
 
@@
 
-In this procedure, $\alpha$ is the same angle size as in the procedure for the visualization of crisp time intervals.
+In this procedure, $\alpha$ is the same angle size as in the procedure for the visualization of CTI.
 The results of this procedure for an example UTI $J$ (as wel as the visualization of $J$ itself) are shown in figure \ref{fig:tm-urz}.
@@
 	\centering
-	\includegraphics[width=0.8\columnwidth]{graphs/TM_model_URZ.eps}
+	\includegraphics[width=0.9\columnwidth]{graphs/TM_model_URZ.eps}
 	\caption{The visualization of the URZ for a given UTI $J$, using the TM framework.}
@@
 
-Given an UTI $J$ and its URZ, every URZ is assigned a symbol and a name and corresponds to a set of Allen relationships.
+Given an UTI $J$ and its URZ, every URZ is assigned a symbol and a name and corresponds to a set of Allen relationships~\cite{DeTre2012}.
 These symbols and names are used to uniquely identify each of $J$'s URZ.
@@
 D    & During & D \\
-OB   & Overlapped by & OB \\
+OB   & Overlapped By & OB \\
 A    & After & A \\
-PM   & Possibly meets & M, B, O \\
-PS   & Possibly starts & O, S, D \\
-PFB  & Possibly finished by & O, FB, C \\
+PM   & Possibly Meets & M, B, O \\
+PS   & Possibly Starts & O, S, D \\
+PFB  & Possibly Finished By & O, FB, C \\
 \multirow{2}{*}
-{PE}   & Possibly equal  & E, FB, SB, F,\\
+{PE}   & Possibly Equal  & E, FB, SB, F,\\
        &                 & S, C, D,O, OB \\
-PSB  & Possibly started by & C, OB, SB \\
-PF   & Possibly finishes & OB, F, D \\
-PMB  & Possibly met by & OB, MB A \\
-UB   & Uncertain beginning & B, M, O, S, D \\
-UE   & Uncertain ending & D, F, OB, MB, A \\
+PSB  & Possibly Started By & C, OB, SB \\
+PF   & Possibly Finishes & OB, F, D \\
+PMB  & Possibly Met By & OB, MB A \\
+UB   & Uncertain Beginning & B, M, O, S, D \\
+UE   & Uncertain Ending & D, F, OB, MB, A \\
 \hline
@@
 
-The TM framework now allows evaluating the Allen relationships a given crisp time interval $I = \left[s, e\right]$ could be in with a given UTI $J$, depending on the exact interval intended by $J$.
+The TM framework now allows evaluating the Allen relationships a given CTI $I = \left[s, e\right]$ could be in with a given UTI $J$, depending on the exact interval intended by $J$.
 This is done as follows.
@@
 Next, $I$'s interval point, its construction half-lines $L_s$ and $L_e$ and the visualization of its starting and ending instants on the visualization of the time are drawn in the same interval space.
-Then, evaluation is a matter of position (in the following, the possibility that $I$ is the interval intended by $J$ is denoted $\pi_J(I)$ and the possibility that $I$ is in Allen relationship $AR$ with $J$ is denoted  $\Pos(I\ AR\ J)$):
+Then, evaluation is a matter of position (in the following, the possibility that $I$ is the interval intended by $J$ is denoted $\pi_J(I)$ and the possibility that $I$ is in Allen relationship $AR$ with $J$ is denoted  $\Pos(I\ AR\ J)$)~\cite{DeTre2012}:
 
@@
 
-Here, the intervals $K$ are crisp time intervals and the expression `$K \in p_i$' is a notation to express that $K$ is visualized by an interval point contained in area part $p_i$.
+Here, the intervals $K$ are CTI and the expression `$K \in p_i$' is a notation to express that $K$ is visualized by an interval point contained in area part $p_i$.
 

### 2013-03-23 trunk/eusflat2013/conclusions.tex
Almost the last changes! Text is too long though :s

In this paper, a comparison is presented between two different frameworks designed to represent time intervals subject to uncertainty and evaluate temporal relationships between such intervals and crisp intervals: the triangular model (TM) framework and the ill-known constraint (IKC) framework. It is concluded that:

\begin{itemize}
	\item With respect to representation, both frameworks differ only slightly, with the TM framework allowing easier human assessments, due to its approach including visualization.
	\item With respect to temporal relationship evaluation, the TM framework allows easy human assessments in several situations, but the IKC framework seems more fitted for complex reasoning, due to its modular build.
\end{itemize}

Future research might deal with the evaluation of temporal relationships between several time intervals subject to uncertainty.

TODO: MORE FUTURE WORK

%In this work we have presented an analysis between two different frameworks that deal and represent imperfect temporal information. Besides the differences in the visualization of imperfect time intervals, we obtained that the temporal reasoning in both frameworks is equivalent. It is also important to notice that whereas the ill-known constraint framework deal with the thirteen Allen's relations, the triangular model distinguishes fifteen possible relational zones and therefore, it is necessary to set a correspondence between the Allen's relations in both frameworks.

%Future research will deal with the analysis between two imperfect time points in each framework and with the representation of open time intervals.

### 2013-03-23 trunk/eusflat2013/notations_and_nomenclature.tex
Almost the last changes! Text is too long though :s

@@
-Often, time is thought of as following an axis (a time axis) and thus it is often seen as an ordered set of infinitely short `moments' or `points' in time, called \emph{instants}~\cite{Dyreson1994},~\cite{Jensen1998}.
+Often, time is thought of as following an axis (a time axis) and thus it is often seen as an ordered set of infinitely short `moments' or `points' in time, called \emph{instants}~\cite{Jensen1998}.
 This is also the case in the presented work.
@@
 These uncertainties are always seen as caused by a (partial) lack of knowledge: the exact, intended time interval is not known, eventhough there is only one correct intended time interval and as such no variability.
-Confidence about exactly which interval is the intended one in the context of such uncertainties is modelled using possibility theory~\cite{DidierDubois1988a}.
+Confidence about exactly which interval is the intended one in the context of such uncertainties is modelled using possibility theory~\cite{Dubois1983}.
 In the presented work, possibility is always interpreted as plausibility given a (partial) lack of knowledge.
 
-To be able to distinguish easily, time intervals not subject to uncertainties will be called \emph{crisp time intervals} in the presented work.
+To be able to distinguish easily, time intervals not subject to uncertainties will be called \emph{crisp time intervals} (CTI) in the presented work.

### 2013-03-23 trunk/eusflat2013/proposal.tex
Almost the last changes! Text is too long though :s

--- 2013-03-01 trunk/eusflat2013/proposal.tex
@@
-Two frameworks for dealing with uncertainty in time intervals have been presented.
-In this section, we will study the equivalences between an uncertain time interval (UTI) and an ill-known time interval (IKTI).
-Nevertheless, there are some diferences in the semantics of UTIs and IKTIs which are also explained.
-Next to that, the computation of the Allen's Relations are compared.
+In this section, first the specific approaches of the IKC and TM frameworks towards interval representation are compared.
+Next, their specific approaches towards the evaluation of Allen relationships between a crisp interval and an interval subject to uncertainty are compared.
 
-\subsection{\label{subsec:sim-dif}Similarities and differences}
-A time interval is defined by a starting and an ending point. When uncertainty arises, in one or both starting and / or ending points, then either an UTI $\pi_I$ or an IKTI can be used.
+\subsection{\label{subsec:comp-interval}Comparison of Approaches to Interval Representation}
+The IKC framework uses IKTI to represent time intervals subject to uncertainty, whereas the TM framework uses UTI for this.
+In both approaches, to consider uncertainty about the exact CTI which is intended, uncertainty about the exact starting and ending instants of the intended interval is considered.
+In both approaches, confidence in the context of this uncertainty is expressed using two possibility distributions.
+In an IKTI, these possibility distributions each define one IKV.
+One of these then defines the IKTI's starting instant and the other defines its ending instant.
+In an UTI, one of these possibility distributions is directly meant to define the possibility that a given instant is the UTI's starting instant and the other is directly meant to define the possibility that a given instant is the UTI's ending instant.
+It is obvious that the concepts of IKTI and UTI are exactly the same, except for the usage of the explicit concept of IKV in IKTI to define and describe uncertainty about starting or ending instants.
+This equality means that both approaches have the same basic restriction: they cannot represent every form of interval subject to uncertainty imaginable.
+For example, imagine an ordered set of instants coinciding with $\mathbb{N}_0$ and imagine an interval subject to uncertainty in this set, where the intended interval is either $\left[2, 4\right]$ (possibility $1$) or $\left[3, 7\right]$ (possibility $1$).
+This interval can not be modelled by a possibility distribution defining the starting instant combined with one defining the ending instant without making $\left[2, 7\right]$ and $\left[3, 4\right]$ also possible as intended intervals to some extent.
+This issue was first suggested about IKTI in~\cite{Billiet2012}.
 
-\begin{itemize}
- \item An IKTI $J = \left[X, Y \right]$ is defined by two ill-known values $X, Y$ which are respectively, the starting and ending points of the time interval.
-Two possibility distributions $\pi_X, \pi_Y$ are defined for $X$ and $Y$ respectively.
-\item An UTI $\pi_I$ is defined by two possibility distributions, $\pi_S$, $\pi_E$ for the starting  and ending point respectively.
-\end{itemize}
+Now, a major difference between the two frameworks, concerning their approaches towards the representation of time intervals in general, is that the TM framework includes visualization in its approach.
 
-In the following, we will assume that:
+A first consequence of this is that the TM framework allows the visualization of multiple CTI in the same image plane.
+In theory, it should also allow the visualization of multiple UTI in the same image plane.
+However, if the interval space area's visualizing these UTI overlap, it is not yet researched which greyscale (or other) color and intensity each interval point in this overlapping area should have, as the appearance of such interval point should both reflect the possibility of it being the interval intended by the one UTI and the possibility of it being the interval intended by the other UTI.
+The advantage of the ability to visualize multiple time intervals in the same image plane is that a human observer could easily assess certain characteristics of a distribution of time intervals from an image containing their visualization.
+The IKC framework, on the other hand, doesn't incorporate any form of visualization and doesn't allow such human assessments.
+
+A second consequence of this is that the TM framework requires that an UTI can be visualized, using a visualization method which actually visualizes the possibility of a CTI of being the interval intended by this UTI for every CTI that has a non-zero possibility of this.
+Thus, a method is required, to calculate the possibility of a given CTI of being the interval intended by an UTI based on the possibility distributions defining this UTI, which express the possibility of an instant being the starting or ending instant of the UTI.
+This method is found by demanding that the given CTI's starting instant is the intended interval's starting instant \emph{and} that the given CTI's ending instant is the intended interval's ending instant and by determining the possibility of the conjunction of both demands using the standard possibility theory conjunction operator `minimum'.
+Although not necessary for the correct and consistent functioning of the TM framework, it appears to be the intention of the TM framework that possibility about the exact starting or ending instant of the interval intended by the UTI could be derived from the possibility distribution defining the possibility that a given CTI is the interval intended by the UTI.
+For this derivation to be consistent, the possibility distributions defining the UTI must be convex.
+Given an ordered set $T$ and an UTI $J = (\pi_{J_s}, \pi_{J_e})$ in $T$ with possibility $\pi_J(I)$ that a given CTI $I = \left[s_i, e_i\right]$ is the exact time interval intended by $J$, the derivations can be calculated as follows:
 
 \begin{align}
-\pi_X &= \pi_S = \mu_S \\
-\pi_Y &= \pi_E = \mu_E
+\pi_{J_s}(s_i) = \max_{K = \left[s_i, k\right], k \in T, k > s_i}(\pi_J(K)) \nonumber \\
+\pi_{J_e}(e_i) = \max_{K = \left[k, e_i\right], k \in T, k < e_i}(\pi_J(K)) \nonumber
 \end{align}
 
-Where $\mu_S, \mu_E$ are triangular membership functions.
-Notice that the main diference between an UTI and an IKTI is a matter of interpretation.
-IKTIs  keep both possibility distributions for both starting and ending poins, while UTIs take a t-norm (the minimum) of both possibility distributions (see Eq.
-\eqref{eq:possibility-uti}).
-That is, not all the crisp intervals between the boundaries of the starting and ending point are candidate intervals for the UTI.
+With respect to this convexity demand, the IKC framework is similar: the possibility distributions defining the starting and ending IKV of an IKTI are also demanded to be convex by the IKC framework, although this appears not to be necessary for the framework to function correctly and consistently.
+
+\subsection{\label{subsec:comp-eval}Comparison of Approaches to Allen Relationship Evaluation}
+As mentioned before, a major difference between the two frameworks, concerning their approaches towards the representation of time intervals in general, is that the TM framework includes visualization in its approach.
+As a result, it also includes visualization in its approach towards the evaluation of Allen relationships between a CTI and an UTI.
+
+A first consequence of this is that, given a CTI, an UTI, its URZ and their visualizations in the same image, the TM framework allows a visual, human assessment of the degree of possibility of this CTI being in an Allen relationship with this UTI, for every Allen relationship with a non-zero such degree, based on this image.
+Moreover, those with possibility degree equal to zero can be easily found before any calculation is done: they are the relationships not contained in the set corresponding to the URZ containing the CTI's interval point.
+On the other hand, to examine the respective possibilities of a given CTI of being in several different Allen relationships with a given IKTI using the IKC framework, a new collection of specific IKC and a specific aggregation of these should be constructed for every Allen relationship under consideration, allowing to calculate its exact possibility and necessity.
+In contrast to the TM framework, using the IKC framework a human assessment before any calculation is not possible and it is not known before any calculation which Allen relationships will result in a possibility degree equal to zero.
+
+A second consequence is that, given an UTI and its URZ, multiple CTI can be visualized in the same image as the UTI and its URZ.
+Thus, the same image could provide a visual, human assessment of the possibilities with which multiple CTI are in Allen relationships with the UTI, before any calculation is done.
+Again, the IKC framework would need a different collection of specific IKC and a specific aggregation of these for every Allen relationship under consideration, but calculating the possibilities for several CTI to be in a given Allen relationship with a given IKTI would not provide much extra work.
+
+A third consequence is that, in the TM framework, given an UTI, its URZ and a visualization of these in the same image, a given Allen relationship could correspond with several URZ.
+Thus, given a distribution of CTI, it is not easy to visually assess which CTI have a non-zero possibility of being in the given Allen relationship with the given UTI.
+On the other hand, assessing this using the IKC framework is impossible without any calculation, but the necessary calculations are pretty straightforward.
+
+Although this has not been rigorously researched yet, it is the conviction of the authors that a great strength of the IKC framework lies in its modular approach towards the evaluation of temporal relationships, resulting in a flexibility and an easy handling of complex temporal relationships, while the visualization of these using the TM framework could become complex and heavy.
+This would give the IKC framework a major advantage over the TM framework, when used in reasoning systems like e.g.
+decision support systems.
+Moreover, in some cases the visualization step used in the TM framework could be a redundant step.
+
+%Both above statements say something about `visual assessment'.
+However, to actually calculate the plausibilities in question, the IKC framework requires a rather complex numerical (?) calculation based on the set of IKC and its combination corresponding to the Allen relationship and the CTI in question.
+In this, the TM framework is similar, as it also requires a rather complex geometrical construction and a calculation based on this, for every Allen relationship and CTI in question, IF the crisp interval point is in an URZ corresponding with a non-singleton set of Allen relationships.
+Else, the plausibility for the Allen relationship in the singleton is 1, all others are 0.
+
+%Although it has not been rigorously researched yet, it is the conviction of the authors that both frameworks can also be used to evaluate combinations of Allen relationships between crisp intervals and an UTI/IKTI.
+For example: `what is the possibility that $J$ is before $K$ but after $I$?'.
+
+\subsection{\label{subsec:example}An Example}
+In this section, some of the comparison results presented above will be illustrated using a simple example.
+
+\begin{example}
+Consider an ordered set of instants $T$ coinciding with $\mathbb{R}$, a CTI $I = \left[1, 3\right]$ and a time interval $J$ of which the starting instant is defined by a triangular possibility distribution on $T$ with core $\{5\}$ and support $\left[2, 7\right]$ and of which the ending instant is defined by a triangular possibility distribution on $T$ with core $\{12\}$ and support $\left[9, 15\right]$.
+The visualization of this example situation using the TM framework is shown in figure \ref{fig:ex}.
+The interval point for $I$ lies in the URZ `PM'.
+Thus, possibility is higher than zero for $I$ to be in a `meets', `before', or `overlaps' relationship with $J$.
+As the darkest point in the `overlaps' area part is very light, the darkest point in the `meets' line segment is reasonably dark and the darkest point in the `before' area part is perfectly black, it can be estimated that $I$ overlaps $J$ with a low possibility, $I$ meets $J$ with a high but non-zero possibility and $I$ is before $J$ with possibility $1$.
+
+TODO: IKC
+
+\end{example}
 
@@
 	\centering
-   \includegraphics[scale=1.5]{graphs/equivalences.eps}
-   \caption{Possibility degrees for two candidate intervals with respect to the UTI $\pi_I$.
-The upper crisp time interval has a possibility degree of 1, whereas the lower crisp interval has a possibility degree of 0.
-}
-   \label{fig:ikc-greater}
+	\includegraphics[width=0.9\columnwidth]{graphs/example_image.eps}
+	\caption{The visualization of the example using the TM framework.}
+	\label{fig:ex}
 \end{figure}
 
-It is possible to obtain the same interpretation for an IKTI, but some ill-known constraints must be defined over the starting and ending ill-known values $X$, $Y$.
-In \cite{Pons2011}, two constraints are defined to determinte whether a crisp time interval $[a, b]$ resides between the boundaries of $J$.
-In this case, all the crisp intervals whose boundaries are within the boundaries of $X$ and $Y$ are considered to be a candidate interval.
-Nevertheless, if we want to obtain the same interpretation as in the UTI, the following two ill-known constraints must be defined:
+%We will illustrate the calculation of one Allen Relation with an example.
+%\begin{example}
+%Consider a CTI $J = [1, 3]$ the two ill-known time points $S, E$ that define a time interval. 
+%We want to obtain the possibility that $J$ is before the time interval.
+%The points are defined by two triangular membership functions $\mu_S$, $\mu_E$.
 
-\begin{align}
- \label{eq:ikc-ikti-uti}
-C_1 &\triangleq \left(=, X \right) \\
-C_2 &\triangleq \left(=, Y \right)\\
-\Pos \left(\left[a, b\right] \right) &= C_1 \left(a \right) \wedge C_2\left(b \right) =\\
-\label{eq:equivalence-ikti-uti}
-\Pos \left(\left[a, b\right] \right) &= \min \left( \pi_X(a), \pi_Y(b) \right)
-\end{align}
+%\begin{align}
+%\label{eq:ex-point}
+%\pi_S &= \mu_S = \left[2,5,7 \right] \\
+%\pi_E &= \mu_E = \left[9,12,15 \right]
+%\end{align}
 
-In this case, Eq. \eqref{eq:equivalence-ikti-uti} is equivalent to Eq. \eqref{eq:possibility-uti}.
-In the following subsection, we will study the connection between the Allen's relations in each framework.
+%Then, it is possible to define both an IKTI $I$ Eq. \eqref{eq:ex-uti-ikti} and an UTI $\pi_I$ Eq. \eqref{eq:ex-uti-uti}. The intervals are illustrated in Figure \ref{fig:example}.
 
-\subsection{\label{subsec:eval-allen-relations}Evaluation of the Allen's Relations}
-As it has been shown in Table \ref{tab:urz}, when using UTIs in the triangular model, it is possible to obtain a more detailed relational information (there are fifteen possible relations) than with the ill-known constraint approach.
-The following observations should be taken into account when extracting relational information from both frameworks:
+%\begin{figure}[h]
+%   \centering
+%   \includegraphics[scale=1.5]{graphs/example.eps}
+%   \caption{In the top, the crisp time interval $J$ and the IKTI given by $I = [S, E]$.
+The ill-known constraint for the before relation is illustrated in grey color.
+In the lower graphic, the interval point $J$ and the UTI $\pi_I$.
+}
+%   \label{fig:example}
+% \end{figure}
 
-\begin{itemize}
- \item In TM, the Allen's Relations $R \in \left \lbrace B, O, C, D, OB, A \right \rbrace$ between an interval point $J$  and an UTI $\pi_I$ are computed by Eq.
-\eqref{eq:possibility-iz-uti-certain}.
-Therefore, for the Allen's Relations in $R$,
- possibility that an Allen Relation between an interval point $J$  and an UTI $\pi_I$ is $1$ if the interval point $J$ is in the Allen Relation, or $0$ in other case.
-\item In the TM, the other nine Allen's Relations between an interval point $J$ and an UTI $\pi_I$ are graded from $1$ to $0$ and computed by using the Eq. \eqref{eq:possibility-iz-uti}. 
-\item In the ill-known constraint framework, all the thirdteen Allen's Relations are computed by using an aggregation of ill-known constraints (see Table \ref{tab:allen-relations}).
-Therefore, the possibility that a time point $J$ is in a given Allen Relation is always graded from $1$ to $0$.
-\end{itemize}
+%\begin{align}
+%\label{eq:ex-uti-ikti}
+%I = \left[S, E \right]\\
+%\label{eq:ex-uti-uti}
+%\mu_I = \left(\mu_S, \mu_E \right)
+%\end{align}
 
-For example, consider that we want to know whether the relation Before (B) holds between a time point $J$ and an IKTI $I$.
-We will note $\Pos \left(C_B (J)  \right)$ the possibility that the relation before holds.
-There are three possibilities:
-\begin{enumerate}
- \item $\Pos \left(C_B (J)  \right)$ = 1: The time point $J$ is before the IKTI $I$.
-\item $\Pos \left(C_B (J)  \right)$ = 0: The time point $J$ is not before the IKTI $I$.
-\item $0 < \Pos \left(C_B (J)  \right) < 1$: The time point $J$ is to some extent before the IKTI $I$.
-\end{enumerate}
+%First we will obtain the possibility between $J$ and the UTI $\pi_I$. The relation possibly meets has three candidates: Before, Meets and Overlaps.
 
-The last case corresponds in the TM to case when the before relation is one of the possible relations (See Table \ref{tab:urz}). That is, one of $\left \lbrace PM, PMB, UB \right \rbrace$.
+%\begin{align}
+%  \mu_{\left(J R \pi_I \right)}(B) &= \sup_{a < e} \pi_S(a)\\
+%\mu_{\left(J R \pi_I \right)}(M) &= \pi_S(e)\\
+%\mu_{\left(J R \pi_I \right)}(O) &= \sup_{a > e} \pi_S(a)
+%\end{align}
 
-The following table stablishes the correspondences between the Allen's relations in the TM and in the ill-known constraints framework.
+%\begin{align}
+%\mu_{\left(J R \pi_I \right)}(B) &= 0.67\\
+%\mu_{\left(J R \pi_I \right)}(M) &= 0.33\\
+%\mu_{\left(J R \pi_I \right)}(O) &= 0
+%\end{align}
 
-\begin{table}[h]
-\centering
-\begin{tabular}{|l|l|l|}
-\hline
-Symbol & Name & Possible Relations \\
-\hline
-B & Before & B, PM, PMB, UB \\
-\multirow{2}{*}
-{O} & Overlaps & O, PM, PS, PFB,\\
-     &          &  PE, PMB, UB \\
-C & Contains & C, PFB, PE, PSB \\
-\multirow{2}{*}
-D & During & D, PS, PE, \\
-  &         &  PMB, UB  \\
-OB & Overlapped by & OB, PE, PSB, PF \\
-A & After & A, PF, UE \\
-M & Meets & M, PM, PMB, UB \\
-MB & Met By & MB, PF, UE \\
-E & Equals & E, PE \\
-\multirow{2}{*}
-S & Starts & S, PS, PE, \\
-  &         &   PMB, UB \\
-SB & Started by & SB, PE, PSB \\
-F & Finishes & F, PE \\
-FB & Finished By & FB, PE, PFB\\
-\hline
-\end{tabular}
-\caption{Correspondences between the Allen's Relations in the ill-known constraint framework and the triangular model,}
-\label{tab:correspondences-tm-ikcf}
-\end{table}
+%In order to compute the possibility  for the before relation between the crisp interval $J$ and the IKTI $I$, a ill-known constraint is needed.
+
+%\begin{equation}
+% C_1\stackrel{\triangle}{=} \left(<,X\right)
+%\end{equation}
+
+%Then the possibility is computed as follows:
+
+%\begin{align}
+%\Pos \left(C_1 (J) \right) &= \sup_{a < e} \pi_S(a)\\
+%\Pos \left(C_1 (J) \right) &= 0.67 
+%\end{align}
+
+%\end{example}
 
@@
 % \end{equation}
+
+%Two frameworks for dealing with uncertainty in time intervals have been presented.
+In this section, we will study the equivalences between an uncertain time interval (UTI) and an ill-known time interval (IKTI).
+Nevertheless, there are some differences in the semantics of UTIs and IKTIs which are also explained.
+Next to that, the computation of the Allen's Relations are compared.
+
+%\subsection{\label{subsec:sim-dif}Similarities and differences}
+%A time interval is defined by a starting and an ending point. When uncertainty arises, in one or both starting and / or ending points, then either an UTI $\pi_I$ or an IKTI can be used.
+
+%\begin{itemize}
+% \item An IKTI $J = \left[X, Y \right]$ is defined by two ill-known values $X, Y$ which are respectively, the starting and ending points of the time interval.
+Two possibility distributions $\pi_X, \pi_Y$ are defined for $X$ and $Y$ respectively.
+%\item An UTI $\pi_I$ is defined by two possibility distributions, $\pi_S$, $\pi_E$ for the starting  and ending point respectively.
+%\end{itemize}
+
+%In the following, we will assume that:
+
+%\begin{align}
+%\pi_X &= \pi_S = \mu_S \\
+%\pi_Y &= \pi_E = \mu_E
+%\end{align}
+
+%Where $\mu_S, \mu_E$ are triangular membership functions.
+Notice that the main difference between an UTI and an IKTI is a matter of interpretation.
+IKTIs  keep both possibility distributions for both starting and ending points, while UTIs take a t-norm (the minimum) of both possibility distributions (see Eq.
+\eqref{eq:possibility-uti}).
+That is, not all the crisp intervals between the boundaries of the starting and ending point are candidate intervals for the UTI.
+
+%\begin{figure}[h]
+%   \centering
+%   \includegraphics[scale=1.5]{graphs/equivalences.eps}
+%   \caption{Possibility degrees for two candidate intervals with respect to the UTI $\pi_I$.
+The upper crisp time interval has a possibility degree of 1, whereas the lower crisp interval has a possibility degree of 0.
+}
+%   \label{fig:ikc-greater}
+% \end{figure}
+
+%It is possible to obtain the same interpretation for an IKTI, but some ill-known constraints must be defined over the starting and ending ill-known values $X$, $Y$.
+In \cite{Pons2011}, two constraints are defined to determinate whether a crisp time interval $[a, b]$ resides between the boundaries of $J$.
+In this case, all the crisp intervals whose boundaries are within the boundaries of $X$ and $Y$ are considered to be a candidate interval.
+Nevertheless, if we want to obtain the same interpretation as in the UTI, the following two ill-known constraints must be defined:
+
+%\begin{align}
+% \label{eq:ikc-ikti-uti}
+%C_1 &\triangleq \left(=, X \right) \\
+%C_2 &\triangleq \left(=, Y \right)\\
+%\Pos \left(\left[a, b\right] \right) &= C_1 \left(a \right) \wedge C_2\left(b \right) =\\
+%\label{eq:equivalence-ikti-uti}
+%\Pos \left(\left[a, b\right] \right) &= \min \left( \pi_X(a), \pi_Y(b) \right)
+%\end{align}
+
+%In this case, Eq. \eqref{eq:equivalence-ikti-uti} is equivalent to Eq. \eqref{eq:possibility-uti}.
+%In the following subsection, we will study the connection between the Allen's relations in each framework.
+
+%\subsection{\label{subsec:eval-allen-relations}Evaluation of the Allen's Relations}
+%As it has been shown in Table \ref{tab:urz}, when using UTIs in the triangular model, it is possible to obtain a more detailed relational information (there are fifteen possible relations) than with the ill-known constraint approach.
+The following observations should be taken into account when extracting relational information from both frameworks:
+
+%\begin{itemize}
+% \item In TM, the Allen's Relations $R \in \left \lbrace B, O, C, D, OB, A \right \rbrace$ between an interval point $J$  and an UTI $\pi_I$ are computed by Eq.
+\eqref{eq:possibility-iz-uti-certain}.
+Therefore, for the Allen's Relations in $R$,
+% possibility that an Allen Relation between an interval point $J$  and an UTI $\pi_I$ is $1$ if the interval point $J$ is in the Allen Relation, or $0$ in other case.
+%\item In the TM, the other nine Allen's Relations between an interval point $J$ and an UTI $\pi_I$ are graded from $1$ to $0$ and computed by using the Eq. \eqref{eq:possibility-iz-uti}. 
+%\item In the ill-known constraint framework, all the thirteen Allen's Relations are computed by using an aggregation of ill-known constraints (see Table \ref{tab:allen-relations}).
+Therefore, the possibility that a time point $J$ is in a given Allen Relation is always graded from $1$ to $0$.
+%\end{itemize}
+
+%For example, consider that we want to know whether the relation Before (B) holds between a time point $J$ and an IKTI $I$.
+We will note $\Pos \left(C_B (J)  \right)$ the possibility that the relation before holds.
+There are three possibilities:
+%\begin{enumerate}
+% \item $\Pos \left(C_B (J)  \right)$ = 1: The time point $J$ is before the IKTI $I$.
+%\item $\Pos \left(C_B (J)  \right)$ = 0: The time point $J$ is not before the IKTI $I$.
+%\item $0 < \Pos \left(C_B (J)  \right) < 1$: The time point $J$ is to some extent before the IKTI $I$.
+%\end{enumerate}
+
+%The last case corresponds in the TM to case when the before relation is one of the possible relations (See Table \ref{tab:urz}). That is, one of $\left \lbrace PM, PMB, UB \right \rbrace$.
+
+%The following table establishes the correspondences between the Allen's relations in the TM and in the ill-known constraints framework.
+
+%\begin{table}[h]
+%\centering
+%\begin{tabular}{|l|l|l|}
+%\hline
+%Symbol & Name & Possible Relations \\
+%\hline
+%B & Before & B, PM, PMB, UB \\
+%\multirow{2}{*}
+%{O} & Overlaps & O, PM, PS, PFB,\\
+%     &          &  PE, PMB, UB \\
+%C & Contains & C, PFB, PE, PSB \\
+%\multirow{2}{*}
+%D & During & D, PS, PE, \\
+%  &         &  PMB, UB  \\
+%OB & Overlapped by & OB, PE, PSB, PF \\
+%A & After & A, PF, UE \\
+%M & Meets & M, PM, PMB, UB \\
+%MB & Met By & MB, PF, UE \\
+%E & Equals & E, PE \\
+%\multirow{2}{*}
+%S & Starts & S, PS, PE, \\
+%  &         &   PMB, UB \\
+%SB & Started by & SB, PE, PSB \\
+%F & Finishes & F, PE \\
+%FB & Finished By & FB, PE, PFB\\
+%\hline
+%\end{tabular}
+%\caption{Correspondences between the Allen's Relations in the ill-known constraint framework and the triangular model,}
+%\label{tab:correspondences-tm-ikcf}
+%\end{table}

### 2013-03-23 trunk/eusflat2013/tm.tex
Almost final changes.

@@
 \subsection{\label{subsec:tm-preliminaries}Preliminary Concepts}
-In the triangular model framework, the approaches to interval representation and Allen relationship evaluation are closely linked with the approach to interval visualization.
-
+In the triangular model framework, the approaches to interval representation and Allen relationship evaluation are closely related to the approach to interval visualization.
 Usually, different time intervals are visualized as different parallel line segments in the same image plane.
-However, as this linear approach introduces a few issues, time intervals are visualized as points in the image plane in the TM framework.
-
-To achieve this, a horizontal line segment visualizing the used (part of the) time axis is drawn in the image plane.
+However, as this linear approach introduces a few issues, time intervals are visualized as points in the image plane in the TM framework~\cite{Weghe2007},~\cite{DeTre2012}.
+To achieve this, a horizontal line segment visualizing the used (part of the) time axis is drawn in the image plane (this is called the \emph{time line}).
 Then, a triangle is drawn, using this line segment as a side of which the angles with the other two sides have sizes $\alpha$ and $-\alpha$ respectively.
 The area contained in this triangle is called the \emph{interval space} and will contain all visualizations of intervals.
-Now, the procedure used to visualize a crisp time interval $\left[s, e\right]$ in the interval space is the following:
+Now, the procedure used to visualize a CTI $\left[s, e\right]$ in the interval space is the following~\cite{Weghe2007},~\cite{DeTre2012}:
 
 \begin{enumerate}
-	\item $s$ is visualized as a point on the visualization of the time axis.
-	\item A straight half-line $L_s$ is constructed from this point, having, in this point, an angle of size $\alpha$ with the visualization of the time axis.
-	\item $e$ is visualized as a point on the visualization of the time axis.
-	\item A straight half-line $L_e$ is constructed from this point, having, in this point, an angle of size $-\alpha$ with the visualization of the time axis.
+	\item $s$ is visualized as a point on the time line.
+	\item A straight half-line $L_s$ is constructed from this point, having, in this point, an angle of size $\alpha$ with the time line.
+	\item $e$ is visualized as a point on the time line.
+	\item A straight half-line $L_e$ is constructed from this point, having, in this point, an angle of size $-\alpha$ with the time line.
 	\item The intersection point of $L_s$ and $L_e$ is the intended visualization of interval $\left[s, e\right]$.
@@
 Given an instant $x \in T$, $\pi_s(x)$ now expresses the possibility that $x$ is the starting instant of $J$ and $\pi_e(x)$ expresses the possibility that $x$ is the ending instant of $J$.
-Thus, $J$ intends to indicate just one crisp time interval, but due to a (partial) lack of knowledge, it is (partially) unknown exactly which time interval this is.
-Because $\pi_s$ and $\pi_e$ are required to be convex, the possibility $\pi_J(I)$ that a given crisp time interval $I = \left[s_i, e_i\right]$ is the exact time interval intended by $J$ can now be uniquely determined by:
+Thus, $J$ intends to indicate just one CTI, but it is (partially) unknown exactly which time interval this is.
+The possibility $\pi_J(I)$ that a given CTI $I = \left[s_i, e_i\right]$ is the exact time interval intended by $J$ can now be uniquely determined by~\cite{DeTre2012}:
 
@@
 An UTI $J$ in $T$ defined by a pair $(\pi_s, \pi_e)$ is now visualized as an area in the image plane.
-This area is constructed as follows: for every crisp time interval $K$ for which the possibility $\pi_J(K)$ of $K$ being the time interval represented by $J$ is higher than zero, the interval point is drawn and is given a color.
-This color is part of a linear single-color greyscale and its saturation is linearly related to the degree of possibility that $K$ is the time interval intended by $J$.
-Thus, every drawn interval point in this visualization corresponds to an interval with a higher-than-zero possibility of being the time interval intended by $J$.
-Together, these drawn interval points form the visualization of $J$.
+This area is constructed as follows: for every CTI $K$ for which the possibility $\pi_J(K)$ of $K$ being the CTI intended by $J$ is higher than zero, the interval point is drawn and is given a color.
+This color is part of a linear single-color greyscale and its saturation is linearly related to the degree of possibility that $K$ is the CTI intended by $J$.
+Together, these drawn interval points form the visualization of $J$~\cite{DeTre2012}.
 An example of such construction is shown in figure \ref{fig:tm-ill-const-ex}.
-In this figure, a darker color of interval point indicates a higher and a lighter color a lower possibility that the corresponding interval is the intended interval.
+In this figure, a darker color of interval point indicates a higher and a lighter color a lower possibility that the corresponding CTI is the intended interval.
 
@@
 
-Now, given an UTI $J$ and its visualization in an interval space, the procedure to visualize its URZ is the following:
+Now, given an UTI $J$ and its visualization in an interval space, the procedure to visualize its URZ is the following~\cite{DeTre2012}:
 
 \begin{enumerate}
-	\item The visualizations (the points) on the visualization of the time axis of both the smallest and greatest instant with a non-zero possibility of being the starting instant of $J$ are determined.
-	\item Two straight half-lines are constructed from each of these points: one having, in the point, an angle of size $\alpha$ with the visualization of the time axis and one having, in the point, an angle of size $-\alpha$ with the visualization of the time axis.
-The parts of these half-lines that lie outside of the interval space are not drawn.
-	\item The visualizations (the points) on the visualization of the time axis of both the smallest and greatest instant with a non-zero possibility of being the ending instant of $J$ are determined.
-	\item Two straight half-lines are constructed from each of these points: one having, in the point, an angle of size $\alpha$ with the visualization of the time axis and one having, in the point, an angle of size $-\alpha$ with the visualization of the time axis.
-The parts of these half-lines that lie outside of the interval space are not drawn.
+	\item The visualizations (the points) on the time line of both the smallest and greatest instant with a non-zero possibility of being the starting instant of $J$ are determined.
+	\item Two straight half-lines are constructed from each of these points: one having, in the point, an angle of size $\alpha$ with the time line and one having, in the point, an angle of size $-\alpha$ with the time line.
+	\item The visualizations (the points) on the time line of both the smallest and greatest instant with a non-zero possibility of being the ending instant of $J$ are determined.
+	\item Two straight half-lines are constructed from each of these points: one having, in the point, an angle of size $\alpha$ with the time line and one having, in the point, an angle of size $-\alpha$ with the time line.
 	\item The lines constructed this way divide the interval space in at most fifteen different areas, including the area corresponding to the visualization of $J$ itself.
@@
 
-Given an UTI $J$ and its URZ, every URZ is assigned a symbol and a name and corresponds to a set of Allen relationships.
+Given an UTI $J$ and its URZ, every URZ is assigned a symbol and a name and corresponds to a set of Allen relationships~\cite{DeTre2012}.
 These symbols and names are used to uniquely identify each of $J$'s URZ.
-The set of Allen relationships corresponding to an URZ contains every Allen relationship in which any interval visualized by one of the URZ's interval points could be with the interval intended by $J$, depending on exactly which time interval $J$ actually intends to represent.
+The set of Allen relationships corresponding to an URZ contains every Allen relationship in which any interval visualized by one of the URZ's interval points could be with the interval intended by $J$, depending on exactly which CTI $J$ actually intends to represent.
 In figure \ref{fig:tm-urz}, the URZ's symbols are noted inside the areas visualizing their respective URZ's.
@@
 
-The TM framework now allows evaluating the Allen relationships a given crisp time interval $I = \left[s, e\right]$ could be in with a given UTI $J$, depending on the exact interval intended by $J$.
+The TM framework now allows evaluating the Allen relationships a given CTI $I = \left[s, e\right]$ could be in with a given UTI $J$, depending on the exact CTI intended by $J$.
 This is done as follows.
 First, $J$ and its URZ are visualized in an interval space.
-Next, $I$'s interval point, its construction half-lines $L_s$ and $L_e$ and the visualization of its starting and ending instants on the visualization of the time are drawn in the same interval space.
-Then, evaluation is a matter of position (in the following, the possibility that $I$ is the interval intended by $J$ is denoted $\pi_J(I)$ and the possibility that $I$ is in Allen relationship $AR$ with $J$ is denoted  $\Pos(I\ AR\ J)$):
+Next, $I$'s interval point, its construction half-lines $L_s$ and $L_e$ and the visualization of its starting and ending instants on the time line are drawn in the same interval space.
+Then, evaluation is a matter of position (in the following, the possibility that $I$ is the interval intended by $J$ is denoted $\pi_J(I)$ and the possibility that $I$ is in Allen relationship $AR$ with $J$ is denoted  $\Pos(I\ AR\ J)$)~\cite{DeTre2012}:
 
@@
 Thus: $\Pos(I\ AR\ J) = 1$.
-	\item if $I$'s interval point is located in an URZ corresponding to a non-singleton set of $n$ Allen relationships $\{R_1, R_2, \ldots, R_n\}$, two more half-lines are drawn: one (denoted $L_s'$) from the visualization of $s$ on the visualization of the time axis, having, in this point, an angle of size $-\alpha$ with this visualization and one (denoted $L_e'$) from the visualization of $e$ on the visualization of the time axis, having, in this point, an angle of size $\alpha$ with this visualization.
+	\item if $I$'s interval point is located in an URZ corresponding to a non-singleton set of $n$ Allen relationships $\{R_1, R_2, \ldots, R_n\}$, two more half-lines are drawn: one (denoted $L_s'$) from the visualization of $s$ on the time line, having, in this point, an angle of size $-\alpha$ with the time line and one (denoted $L_e'$) from the visualization of $e$ on the time line, having, in this point, an angle of size $\alpha$ with the time line.
 Now, some of these half-lines $L_s$, $L_s'$, $L_e$ and $L_e'$ may intersect with the area visualizing $J$.

### 2013-03-23 eusflat2013/introduction.tex
Almost final changes.

@@
 Information systems (IS) have always tried to model parts of reality.
-To achieve this modelling, such systems contain data representing properties of real-world objects or concepts~\cite{JoseEnriquePons2012},~\cite{Billiet2012}.
-As time is an essential aspect of many real-world objects or concepts, information systems often contain data representing temporal notions which describe such temporal properties~\cite{Bolour1982},~\cite{VanderCruyssen1997}.
+To achieve this modelling, IS contain data representing properties of real-world objects or concepts~\cite{JoseEnriquePons2012},~\cite{Billiet2012}.
+As time is an essential aspect of many real-world objects or concepts, IS often contain data representing temporal notions which describe such temporal properties~\cite{Bolour1982},~\cite{VanderCruyssen1997}.
 Such temporal notions usually take the form of either instants~\cite{Jensen1998}, which can informally be seen as infinitely short `moments' or `points' in time, or time intervals~\cite{Jensen1998}.
@@
 Generally, one of the most important purposes of an IS is to allow the retrieval of information or knowledge deduced from its data.
-Information or knowledge is usually retrieved from an IS by querying it and examining or analyzing the query results or by visualizing the contents of the IS, querying this visualization and examining or analyzing the resulting visualization(s).
+Such retrieval is usually done by querying the IS and examining or analyzing the query results or by visualizing the contents of the IS, querying this visualization and examining or analyzing the resulting visualization(s).
 
 Of course, when temporal information is represented in an IS, querying this IS may have a temporal aspect too.
-Usually, querying an IS containing data representing temporal notions is conceptually done by specifying one or more time indications and requesting information that is in a specific relationship with these indications, where the semantics of these relationships are specifically temporal~\cite{Billiet2012},~\cite{JoseEnriquePons2012},~\cite{Pons2012a}.
+Usually, querying such an IS is conceptually done by specifying one or more time indications and requesting information that is in a specific relationship with these indications, where the semantics of these relationships are specifically temporal~\cite{Billiet2012},~\cite{JoseEnriquePons2012},~\cite{Pons2012a}.
 Thus, some existing proposals have considered groups of basic relationships between time indications used to construct and express specific temporal relationships~\cite{Medina1994},~\cite{Schockaert2008}.
@@
 The structure of this paper is as follows: section \ref{sec:general-preliminaries} presents some general preliminaries and naming and notation conventions used in this paper.
-Sections \ref{sec:ikc} and \ref{sec:tm} introduce the IKC and TM frameworks respectively: both sections first introduce some preliminary concepts and techniques specific to the framework under consideration, then explain how the representation of time intervals by the framework is done and finally show how the evaluation of Allen relationships between an interval subject to uncertainty and one without uncertainty is done.
-In section \ref{sec:proposal}, both frameworks are compared: first their approaches to representing time intervals are compared, next their approaches to evaluating Allen relationships are compared.
+Sections \ref{sec:ikc} and \ref{sec:tm} introduce the IKC and TM frameworks respectively: both sections first introduce some specific preliminary concepts and techniques, then explain how the representation of time intervals by the framework is done and finally show how the evaluation of Allen relationships between an interval subject to uncertainty and one without uncertainty is done.
+In section \ref{sec:proposal}, both frameworks are compared: first their approaches to representing time intervals are compared, next their approaches to evaluating Allen relationships.
 Finally, section \ref{sec:conclusions} presents the principal conclusions of the presented work and some possible future work.

### 2013-03-23 eusflat2013/proposal.tex
Almost final changes.

--- 2013-02-27 eusflat2013/proposal.tex
@@
-\subsection{\label{subsec:sim-dif}Similarities and differences}
+In this section, first the approaches of the IKC and TM frameworks towards interval representation are compared.
+Next, their approaches towards the evaluation of Allen relationships between a crisp interval and an interval subject to uncertainty are compared.
 
-In this section we will analyze first the calculation of all the possible Allen's relations between a crisp time point and an ill-known time point and between a crisp time point and an uncertain time indication (UTI).
+\subsection{\label{subsec:comp-interval}Comparison of Approaches to Interval Representation}
+The IKC framework uses IKTI to represent time intervals subject to uncertainty, whereas the TM framework uses UTI for this.
+In both approaches, to consider uncertainty about the exact CTI which is intended, uncertainty about the exact starting and ending instants of the intended interval is considered and confidence in the context of this uncertainty is expressed using two possibility distributions.
+In an IKTI, these possibility distributions each define one IKV.
+One of these then defines the IKTI's starting instant and the other defines its ending instant.
+In an UTI, one of these possibility distributions is directly meant to define the UTI's starting instant and the other is directly meant to define the UTI's ending instant.
+It is obvious that the concepts of IKTI and UTI are exactly the same, except for the explicit usage of the concept of IKV in IKTI to define and describe uncertainty about starting and ending instants.
+This equality implies that both approaches have the same basic restriction: they cannot represent every kind of interval subject to uncertainty imaginable.
+For example, imagine an ordered set of instants coinciding with $\mathbb{N}_0$ and imagine an interval subject to uncertainty in this set, where the intended interval is either $\left[2, 4\right]$ (possibility $1$) or $\left[3, 7\right]$ (possibility $1$).
+This interval can not be modelled by a combination of possibility distributions defining the starting and ending instants without making $\left[2, 7\right]$ and $\left[3, 4\right]$ also possible as intended intervals to some extent.
+This issue was first suggested about IKTI in~\cite{Billiet2012}.
+
+Now, a major difference between the two frameworks, concerning their approaches towards the representation of time intervals in general, is that the TM framework includes visualization in its approach.
+
+A first consequence of this is that the TM framework allows the visualization of multiple CTI in the same image plane.
+In theory, it should also allow the visualization of multiple UTI in the same image plane.
+However, if the interval space area's visualizing these UTI overlap, it is not yet researched which greyscale (or other) color and intensity each interval point in this overlapping area should have, as the appearance of such interval point should both reflect the possibility of it being the interval intended by the one UTI and the possibility of it being the interval intended by the other UTI.
+The advantage of the ability to visualize multiple time intervals in the same image plane is that a human observer could easily assess certain characteristics of a distribution of time intervals from an image containing their visualization.
+The IKC framework, on the other hand, doesn't incorporate any form of visualization and doesn't allow such human assessments.
+
+A second consequence of this is that the TM framework requires that an UTI can be visualized, using a visualization method which actually visualizes the possibility of a CTI of being the interval intended by this UTI for every CTI that has a non-zero possibility of this.
+Thus, a method is required, to calculate the possibility of a given CTI of being the interval intended by an UTI based on the possibility distributions defining this UTI.
+This method is found by demanding that the given CTI's starting instant is the intended interval's starting instant \emph{and} that the given CTI's ending instant is the intended interval's ending instant and by determining the possibility of the conjunction of both demands using the standard possibility theory conjunction operator `minimum'.
+Although not necessary for the correct and consistent functioning of the TM framework, it appears to be the intention of the TM framework that possibility about the exact starting or ending instant of the interval intended by the UTI could be derived from the possibility distribution defining the possibility that a given CTI is the interval intended by the UTI.
+For this derivation to be consistent, the possibility distributions defining the UTI must be convex.
+Given an ordered set $T$ and an UTI $J = (\pi_{J_s}, \pi_{J_e})$ in $T$ with possibility $\pi_J(I)$ that a given CTI $I = \left[s_i, e_i\right]$ is the exact time interval intended by $J$, the derivations can be calculated as follows:
+
+\begin{align}
+\pi_{J_s}(s_i) = \max_{K = \left[s_i, k\right], k \in T, k > s_i}(\pi_J(K)) \nonumber \\
+\pi_{J_e}(e_i) = \max_{K = \left[k, e_i\right], k \in T, k < e_i}(\pi_J(K)) \nonumber
+\end{align}
+
+With respect to this convexity demand, the IKC framework is similar: the possibility distributions defining the starting and ending IKV of an IKTI are also demanded to be convex by the IKC framework, although this appears not to be necessary for the framework to function correctly and consistently.
+
+\subsection{\label{subsec:comp-eval}Comparison of Approaches to Allen Relationship Evaluation}
+As mentioned before, a major difference between the two frameworks, concerning their approaches towards the representation of time intervals in general, is that the TM framework includes visualization in its approach.
+As a result, it also includes visualization in its approach towards the evaluation of Allen relationships between a CTI and an UTI.
+
+A first consequence of this is that, given a CTI, an UTI, its URZ and their visualizations in the same image, the TM framework allows a visual, human assessment of the degree of possibility of this CTI being in an Allen relationship with this UTI, for every Allen relationship with a non-zero such degree, based on this image.
+Moreover, those with possibility degree equal to zero can be easily found before any calculation is done: they are the relationships not contained in the set corresponding to the URZ containing the CTI's interval point.
+On the other hand, to examine the respective possibilities of a given CTI of being in several different Allen relationships with a given IKTI using the IKC framework, a new collection of specific IKC and a specific aggregation of these should be constructed for every Allen relationship under consideration, allowing to calculate its exact possibility and necessity.
+In contrast to the TM framework, using the IKC framework a human assessment before any calculation is not possible and it is not known before any calculation which Allen relationships will result in a possibility degree equal to zero.
+
+A second consequence is that, given an UTI and its URZ, multiple CTI can be visualized in the same image as the UTI and its URZ.
+Thus, the same image could provide a visual, human assessment of the possibilities with which multiple CTI are in Allen relationships with the UTI, before any calculation is done.
+Again, the IKC framework would need a different collection of specific IKC and a specific aggregation of these for every Allen relationship under consideration, but calculating the possibilities for several CTI to be in a given Allen relationship with a given IKTI would not provide much extra work.
+
+A third consequence is that, in the TM framework, given an UTI, its URZ and a visualization of these in the same image, a given Allen relationship could correspond with several URZ.
+Thus, given a distribution of CTI, it is not easy to visually assess which CTI have a non-zero possibility of being in the given Allen relationship with the given UTI.
+On the other hand, assessing this using the IKC framework is impossible without any calculation, but the necessary calculations are pretty straightforward.
+
+Although this has not been rigorously researched yet, it is the conviction of the authors that a great strength of the IKC framework lies in its modular approach towards the evaluation of temporal relationships, resulting in a flexibility and an easy handling of complex temporal relationships, while the visualization of these using the TM framework could become complex and heavy.
+This would give the IKC framework a major advantage over the TM framework, when used in reasoning systems like e.g.
+decision support systems.
+Moreover, in some cases the visualization step used in the TM framework could be a redundant step.
+
+%Both above statements say something about `visual assessment'.
+However, to actually calculate the plausibilities in question, the IKC framework requires a rather complex numerical (?) calculation based on the set of IKC and its combination corresponding to the Allen relationship and the CTI in question.
+In this, the TM framework is similar, as it also requires a rather complex geometrical construction and a calculation based on this, for every Allen relationship and CTI in question, IF the crisp interval point is in an URZ corresponding with a non-singleton set of Allen relationships.
+Else, the plausibility for the Allen relationship in the singleton is 1, all others are 0.
+
+%Although it has not been rigorously researched yet, it is the conviction of the authors that both frameworks can also be used to evaluate combinations of Allen relationships between crisp intervals and an UTI/IKTI.
+For example: `what is the possibility that $J$ is before $K$ but after $I$?'.
+
+\subsection{\label{subsec:example}An Example}
+In this section, some of the comparison results presented above will be illustrated using a simple example.
+
+\begin{example}
+Consider an ordered set of instants $T$ coinciding with $\mathbb{R}$, a CTI $I = \left[1, 3\right]$ and a time interval $J$ of which the starting instant is defined by a triangular possibility distribution on $T$ with core $\{5\}$ and support $\left[2, 7\right]$ and of which the ending instant is defined by a triangular possibility distribution on $T$ with core $\{12\}$ and support $\left[9, 15\right]$.
+The visualization of this example situation using the TM framework is shown in figure \ref{fig:ex}.
+The interval point for $I$ lies in the URZ `PM'.
+Thus, possibility is higher than zero for $I$ to be in a `meets', `before', or `overlaps' relationship with $J$.
+As the darkest point in the `overlaps' area part is very light, the darkest point in the `meets' line segment is reasonably dark and the darkest point in the `before' area part is perfectly black, it can be estimated that $I$ overlaps $J$ with a low possibility, $I$ meets $J$ with a high but non-zero possibility and $I$ is before $J$ with possibility $1$.
+
+TODO: IKC
+
+\end{example}
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.9\columnwidth]{graphs/example_image.eps}
+	\caption{The visualization of the example using the TM framework.}
+	\label{fig:ex}
+\end{figure}
+
+%We will illustrate the calculation of one Allen Relation with an example.
+%\begin{example}
+%Consider a CTI $J = [1, 3]$ the two ill-known time points $S, E$ that define a time interval. 
+%We want to obtain the possibility that $J$ is before the time interval.
+%The points are defined by two triangular membership functions $\mu_S$, $\mu_E$.
+
+%\begin{align}
+%\label{eq:ex-point}
+%\pi_S &= \mu_S = \left[2,5,7 \right] \\
+%\pi_E &= \mu_E = \left[9,12,15 \right]
+%\end{align}
+
+%Then, it is possible to define both an IKTI $I$ Eq. \eqref{eq:ex-uti-ikti} and an UTI $\pi_I$ Eq. \eqref{eq:ex-uti-uti}. The intervals are illustrated in Figure \ref{fig:example}.
+
+%\begin{figure}[h]
+%   \centering
+%   \includegraphics[scale=1.5]{graphs/example.eps}
+%   \caption{In the top, the crisp time interval $J$ and the IKTI given by $I = [S, E]$.
+The ill-known constraint for the before relation is illustrated in grey color.
+In the lower graphic, the interval point $J$ and the UTI $\pi_I$.
+}
+%   \label{fig:example}
+% \end{figure}
+
+%\begin{align}
+%\label{eq:ex-uti-ikti}
+%I = \left[S, E \right]\\
+%\label{eq:ex-uti-uti}
+%\mu_I = \left(\mu_S, \mu_E \right)
+%\end{align}
+
+%First we will obtain the possibility between $J$ and the UTI $\pi_I$. The relation possibly meets has three candidates: Before, Meets and Overlaps.
+
+%\begin{align}
+%  \mu_{\left(J R \pi_I \right)}(B) &= \sup_{a < e} \pi_S(a)\\
+%\mu_{\left(J R \pi_I \right)}(M) &= \pi_S(e)\\
+%\mu_{\left(J R \pi_I \right)}(O) &= \sup_{a > e} \pi_S(a)
+%\end{align}
+
+%\begin{align}
+%\mu_{\left(J R \pi_I \right)}(B) &= 0.67\\
+%\mu_{\left(J R \pi_I \right)}(M) &= 0.33\\
+%\mu_{\left(J R \pi_I \right)}(O) &= 0
+%\end{align}
+
+%In order to compute the possibility  for the before relation between the crisp interval $J$ and the IKTI $I$, a ill-known constraint is needed.
+
+%\begin{equation}
+% C_1\stackrel{\triangle}{=} \left(<,X\right)
+%\end{equation}
+
+%Then the possibility is computed as follows:
+
+%\begin{align}
+%\Pos \left(C_1 (J) \right) &= \sup_{a < e} \pi_S(a)\\
+%\Pos \left(C_1 (J) \right) &= 0.67 
+%\end{align}
+
+%\end{example}
+
+% 
+% In this section we will analyze first the calculation of all the possible Allen's relations between a crisp time point and an ill-known time point and between a crisp time point and an uncertain time indication (UTI).
 Then, we will compare two ill-known time points and two uncertain time indications.

### 2013-03-23 eusflat2013/conclusions.tex
Some finalization of the conclusions.

@@
-In this work we have presented an analysis between two different frameworks that deal and represent imperfect temporal information.
+In this paper, a comparison is presented between two different frameworks designed to represent time intervals subject to uncertainty and evaluate temporal relationships between such intervals and crisp intervals: the triangular model (TM) framework and the ill-known constraint (IKC) framework.
+It is concluded that:
+
+\begin{itemize}
+	\item With respect to representation, both frameworks differ only slightly, with the TM framework allowing easier human assessments, due to its approach including visualization.
+	\item With respect to temporal relationship evaluation, the TM framework allows easy human assessments in several situations, but the IKC framework seems more fitted for complex reasoning, due to its modular build.
+\end{itemize}
+
+Future research will deal with the evaluation of temporal relationships between several time intervals subject to uncertainty and with querying aspects in both frameworks.
+
+%In this work we have presented an analysis between two different frameworks that deal and represent imperfect temporal information.
 Besides the differences in the visualization of imperfect time intervals, we obtained that the temporal reasoning in both frameworks is equivalent.
@@
 
-Future research will deal with the analysis between two imperfect time points in each framework and with the representation of open time intervals.
+%Future research will deal with the analysis between two imperfect time points in each framework and with the representation of open time intervals.

### 2013-03-23 trunk/eusflat2013/introduction.tex
Final draft.

--- 2013-02-27 trunk/eusflat2013/introduction.tex
@@
-\begin{table}[h]
-\centering
-\begin{tabular}{|c|l|}
-\hline
-Name & Implementation \\ \hline 
-$I$ equals $J$ & if $s_i = s_j \wedge e_i = e_j $ \\
-$I$ starts $J$ & if $s_i = s_j \wedge e_i < e_j $ \\
-$I$ started by $J$ & if $s_i = s_j \wedge e_i > e_j $ \\
-$I$ finishes $J$ & if $s_i > s_j \wedge e_i = e_j $ \\
-$I$ finished by $J$ & if $s_i < s_j \wedge e_i = e_j $ \\
-$I$ meets $J$ & if $e_i = s_j $ \\
-$I$ met by $J$ & if $s_i = e_j $ \\
-$I$ overlaps $J$ & if $s_i < s_j \wedge e_i < e_j \wedge e_i > s_j $ \\
-$I$ overlapped by $J$ & if $s_i > s_j \wedge e_j < e_i \wedge s_i < e_j  $ \\
-$I$ during $J$ & if $s_i > s_j \wedge e_i < e_j $ \\
-$I$ contains $J$ & if $  s_i < s_j \wedge e_i > e_j$ \\
-$I$ before $J$ & if $e_i < s_j $ \\
-$I$ after $J$ & if $s_i > e_j $ \\
-\hline
-\end{tabular}
-\caption{Allen's relations represented in the framework. $I = \left[s_i, e_i\right]$, $J=  \left[s_j, e_j\right]$}
-\label{tab:allen-relations}
-\end{table}
+Information systems (IS) have always tried to model parts of reality.
+To achieve this modelling, IS contain data representing properties of real-world objects or concepts~\cite{JoseEnriquePons2012},~\cite{Billiet2012}.
+As time is an essential aspect of many real-world objects or concepts, IS often contain data representing temporal notions which describe such temporal properties~\cite{Bolour1982},~\cite{VanderCruyssen1997}.
+Such temporal notions usually take the form of either instants~\cite{Jensen1998}, which can informally be seen as infinitely short `moments' or `points' in time, or time intervals~\cite{Jensen1998}.
+
+Data are often produced by humans, but human-made data are prone to imperfections: some data may be vague, imprecise, incomplete, contradictory or uncertain.
+Data representing temporal notions may contain such imprecisions too~\cite{Devos1994},~\cite{Dubois2003},~\cite{JoseEnriquePons2012},~\cite{Billiet2012}.
+The work presented in this paper is specifically concerned with time intervals (and as a special case: instants) subject to uncertainty.
+
+Generally, one of the most important purposes of an IS is to allow the retrieval of information or knowledge deduced from its data.
+Such retrieval is usually done by querying the IS and examining or analyzing the query results or by visualizing the contents of the IS, querying this visualization and examining or analyzing the resulting visualization(s).
+
+Of course, when temporal information is represented in an IS, querying this IS may have a temporal aspect too.
+Usually, querying such an IS is conceptually done by specifying one or more time indications and requesting information that is in a specific relationship with these indications, where the semantics of these relationships are specifically temporal~\cite{Billiet2012},~\cite{JoseEnriquePons2012},~\cite{Pons2012a}.
+Thus, some existing proposals have considered groups of basic relationships between time indications used to construct and express specific temporal relationships~\cite{Medina1994},~\cite{Schockaert2008}.
+Notably, Allen~\cite{Allen1983} presented a reasoning framework containing all semantically usefull basic temporal relationships between time intervals (and as a special case instants).
+The resulting relationships are shown in figure \ref{fig:allen-relationships}.
+These Allen relationships are used in the presented work.
 
@@
  \end{figure}
+
+To be able to query IS containing data representing time indications subject to uncertainty, a framework is necessary, able to represent uncertainties in time indications in a semantically sound way, without (much) information loss and able to temporally reason with such time indications in a semantically sound and usefull way~\cite{Dubois1983},~\cite{Dubois2003}.
+Although more proposals for such frameworks exist, the work presented in this paper focusses on just two: the \emph{ill-known constraint \emph{(IKC)} framework}~\cite{Pons2011} and the \emph{triangular model \emph{(TM)} framework}~\cite{DeTre2012}.
+
+The work presented in this paper consists of a comparison of both frameworks, about the approaches they use to represent time intervals subject to uncertainty and to reason about the temporal relationships between such intervals and time intervals without uncertainty.
+
+The structure of this paper is as follows: section \ref{sec:general-preliminaries} presents some general preliminaries and naming and notation conventions used in this paper.
+Sections \ref{sec:ikc} and \ref{sec:tm} introduce the IKC and TM frameworks respectively: both sections first introduce some specific preliminary concepts and techniques, then explain how the representation of time intervals by the framework is done and finally show how the evaluation of Allen relationships between an interval subject to uncertainty and one without uncertainty is done.
+In section \ref{sec:proposal}, both frameworks are compared: first their approaches to representing time intervals are compared, next their approaches to evaluating Allen relationships.
+Also, an example is given.
+Finally, section \ref{sec:conclusions} presents the principal conclusions of the presented work and some possible future work.
+
+%TODO: more references in the above part?
+%TODO: include some form of proof or previous work with these statements?!
+
+%The concept of the time has been widely studied \cite{Benthem1982}, \cite{Shackle1961}, \cite{Klein1994}.
+Moreover, humans beings manage temporal indications in an imprecise way \cite{Devos1998}.
+But when dealing with time in an Information System, some simplifications must to be done.
+The first thing to do is the discretization of the time line into time points or time intervals.
+Both approaches have been proved to be equivalent, although due to the nature of the smallest unit of time in a computer, a \emph{chronon}, the discretization of a time point returns a time interval.
+
+%All the possible positions between two time intervals were studied by Allen \cite{Allen1983}, \cite{Allen1985}.
+As result, the thirteen Allen's Relations were obtained (See Table \ref{tab:allen-relations}) which are illustrated in Figure \ref{fig:allen-relationships}.
+In order to achieve a more intelligent processing of time, some theoretical frameworks are used to reason with time.
+First the possibility theory was employed in the reasoning with temporal information \cite{Dubois2003a}.
+Then, several proposals \cite{Schockaert2008}, \cite{nagypal2003}, \cite{Ohlbach2004}, \cite{Pons2011} studied how to extend the thirteen Allen's relations to the possibilistic case.
+Also rough set theory \cite{Pawlak1995} has been used to represent and reason about imperfect time intervals \cite{Qiang2009}, \cite{Qiang2010}.
+
+%Although there exists several proposals to represent and visualize imperfect time intervals, in this paper we will focus on two: the ill-known constraint framework \cite{Pons2011} and the triangular model \cite{DeTre2012}.
+The first one is a theoretical framework that deals with temporal reasoning and the second one is a visual framework that represent imperfect time intervals in a two dimensional space.
+Both frameworks can be used to represent imperfect time intervals as well as temporal reasoning.
+
+%The structure of this paper is the following.
+Section  \ref{sec:preliminaries} introduces the Ill-known constraint framework.
+Section \ref{sec:triangular-model} introduces the triangular model.
+Section \ref{sec:proposal} analyze both frameworks and shows the correspondences and differences between both frameworks.
+The section concludes with an example showing that the calculations made in both frameworks are equivalent.
+Finally, Section \ref{sec:conclusions} presents the main conclusions and the future work.
+
+%\begin{table}[h]
+%\centering
+%\begin{tabular}{|c|l|}
+%\hline
+%Name & Implementation \\ \hline 
+%$I$ equals $J$ & if $s_i = s_j \wedge e_i = e_j $ \\
+%$I$ starts $J$ & if $s_i = s_j \wedge e_i < e_j $ \\
+%$I$ started by $J$ & if $s_i = s_j \wedge e_i > e_j $ \\
+%$I$ finishes $J$ & if $s_i > s_j \wedge e_i = e_j $ \\
+%$I$ finished by $J$ & if $s_i < s_j \wedge e_i = e_j $ \\
+%$I$ meets $J$ & if $e_i = s_j $ \\
+%$I$ met by $J$ & if $s_i = e_j $ \\
+%$I$ overlaps $J$ & if $s_i < s_j \wedge e_i < e_j \wedge e_i > s_j $ \\
+%$I$ overlapped by $J$ & if $s_i > s_j \wedge e_j < e_i \wedge s_i < e_j  $ \\
+%$I$ during $J$ & if $s_i > s_j \wedge e_i < e_j $ \\
+%$I$ contains $J$ & if $  s_i < s_j \wedge e_i > e_j$ \\
+%$I$ before $J$ & if $e_i < s_j $ \\
+%$I$ after $J$ & if $s_i > e_j $ \\
+%\hline
+%\end{tabular}
+%\caption{Allen's relations represented in the framework. $I = \left[s_i, e_i\right]$, $J=  \left[s_j, e_j\right]$}
+%\label{tab:allen-relations}
+%\end{table}

### 2013-03-23 eusflat2013/proposal.tex
Final draft.

@@
 Although not necessary for the correct and consistent functioning of the TM framework, it appears to be the intention of the TM framework that possibility about the exact starting or ending instant of the interval intended by the UTI could be derived from the possibility distribution defining the possibility that a given CTI is the interval intended by the UTI.
-For this derivation to be consistent, the possibility distributions defining the UTI must be convex.
+For this derivation to be consistent, the possibility distributions defining the UTI must be convex~\cite{Dubois1983}.
 Given an ordered set $T$ and an UTI $J = (\pi_{J_s}, \pi_{J_e})$ in $T$ with possibility $\pi_J(I)$ that a given CTI $I = \left[s_i, e_i\right]$ is the exact time interval intended by $J$, the derivations can be calculated as follows:
@@
 Thus, possibility is higher than zero for $I$ to be in a `meets', `before', or `overlaps' relationship with $J$.
-As the darkest point in the `overlaps' area part is very light, the darkest point in the `meets' line segment is reasonably dark and the darkest point in the `before' area part is perfectly black, it can be estimated that $I$ overlaps $J$ with a low possibility, $I$ meets $J$ with a high but non-zero possibility and $I$ is before $J$ with possibility $1$.
+As the darkest point in the `overlaps' area part is very light, the darkest point in the `meets' line segment is of almost exactly the same lightness and the darkest point in the `before' area part is perfectly black, it can be estimated that $I$ overlaps $J$ with a low possibility, $I$ meets $J$ with the same low possibility and $I$ is before $J$ with possibility $1$.
+Calculation using the IKC framework now learns that:
 
-TODO: IKC
+\begin{align}
+\Pos(I overlaps J) & = \min(1, 1/3, 1) = 1/3 \nonumber \\
+\Pos(I meets J) & = \min(1/3, 1) = 1/3\nonumber \\
+\Pos(I before J) & = \min(1) = 1\nonumber
+\end{align}
 

### 2013-03-26 trunk/fqas2013/paper.tex
The base file.

%%FQAS 2013 Contribution

\documentclass[runningheads,a4paper]{llncs}

\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx}

\begin{document}

\mainmatter  %start of the contribution

\title{Bipolar Querying of Valid-time Intervals Subject to Uncertainty}

%Short form of title for running heads
\titlerunning{Bipolar Valid-time Querying}

\author{Christophe Billiet\inst{1} \and Jos\'{e} Enrique Pons\inst{2} \and Olga Pons\inst{2} \and Guy De Tr\'{e}\inst{1}}

\authorrunning{Christophe Billiet et al.}

\institute{
Department of Telecommunications and Information Processing, Ghent University,\\
Sint-Pietersnieuwstraat 41, B-9000, Ghent, Belgium\\
Christophe.Billiet@UGent.be, Guy.DeTr\'{e}@UGent.be\\
\and
Department of Computer Science and Artificial Intelligence, University of Granada,\\
C/Periodista Daniel Saucedo Aranda, S/N, E-18071, Granada, Spain\\
jpons@decsai.ugr.es, opc@decsai.ugr.es
}

\maketitle

\begin{abstract}
Here goes the abstract.
\keywords{keyword1, keyword2}
\end{abstract}

\section{Introduction}
%1.1. On databases and their purpose
Databases are collections of data. These data are usually the result of measurements or descriptions of aspects or properties of real-world objects or concepts. As these data represent properties of objects or concepts in reality, the database itself represents a part of reality~\cite{Billiet2012ipmu},~\cite{Pons2012ipmu}.

%1.2. Time in databases
To many real-world concepts or objects, time is an essential aspect. E.g., certain historical events are meaningless without time frame. Therefore, many databases contain data representing temporal notions which describe the temporal properties of real-world objects or concepts. In the context of database systems, such temporal values~\cite{Dyreson1994sigmod},~\cite{Bohlen1998lncs} can be classified into different categories, based on their interpretation and purpose:

\begin{itemize}
	\item \textbf{valid-time indications} describe when a database fact or data is a true or valid representation of the reality modelled by the database~\cite{Dyreson1994sigmod},~\cite{Bohlen1998lncs}.
	\item \textbf{transaction-time indications} describe when a fact or data is current in the database, which means it is not logically deleted and can thus be retrieved~\cite{Dyreson1994sigmod},~\cite{Bohlen1998lncs}.
	\item \textbf{decision-time indications} describe when certain events were decided to happen~\cite{Nascimento1995siwtrr}.
\end{itemize}

%1.3. Imperfect data
%1.3.1. General
%1.3.2. In time
A lot of data is made by humans. Human-made data is prone to imperfections: it can be vague or imprecise, may contain contradictions, be incomplete or contain uncertainties~\cite{Pons2012ipmu},~\cite{Pons2012ijcis}. Uncertainties in data may be caused by variability in the outcomes of an experiment and confidence in the context of such uncertainty may be modelled using probability theory~\cite{Pons2013ijufkbs}. Uncertainties in data may also be caused by a (partial) lack of knowledge: the exact value which is the answer to a certain question might not be known, even if there is just one correct answer and as such no variability. Confidence in the context of this kind of uncertainty may be modelled using possibility theory~\cite{Zadeh1978fss},~\cite{Dubois1988},~\cite{Pons2013ijufkbs}, ~\cite{Pons2012ipmu},~\cite{Pons2012ijcis}. Uncertainties caused by a (partial) lack of knowledge may exist in temporal notions~\cite{Pons2012ijcis},~\cite{Pons2012ipmu},~\cite{Pons2013ijufkbs}. E.g., consider a medieval document defining a legal act. The document contains the date defining the day on which the legal act took effect. Consider the document damaged, making this date unreadable. In this case, there is uncertainty about the exact day on which the legal act took effect and this uncertainty is caused by the lack of some knowledge: it is known that there is exactly one day on which the act took effect, but it is not known which one.

%1.4. Querying databases
%1.4.1. Crisp querying of databases
%1.4.2. Fuzzy querying of databases
%1.4.3. Bipolar querying of databases
One of the most important purposes of a database is of course to allow the retrieval of information and knowledge deduced from its data. Often, this is done by querying the database. Databases can be queried in a `regular' way: the user describes the data which are desired or satisfactory to him or her and which he or she thus wants to retrieve, by perfectly describing the allowed values of this data for certain attributes in clearly stated user preferences. However, databases can also be queried in a `fuzzy' way: the user describes the data which are desired or satisfactory to him or her by imperfectly describing the allowed values of this data for certain attributes~\cite{DeCaluwe2007ijis}. These imperfect descriptions may contain vagueness or imprecision, often through the use of linguistic terms~\cite{Kacprzyk2001is}. Databases can also be queried in what is called a `bipolar' way. Generally, there are two main approaches to this. One is for the user to describe the data which are required (acceptable) to him or her and to describe the data among the required data, which are really desired (wished-for) to him or her, both by describing the allowed values of this data for certain attributes~\cite{Dubois2002lnai}. In the other, the user describes the data which are desired or satisfactory to him or her and the data which are undesired or unsatisfactory to him or her, both by describing the allowed values of this data for certain attributes~\cite{DeTre2010ieeetfs},~\cite{Matthe2011ijis}. These descriptions might or might not contain imperfections. The presented work will only consider the latter approach to bipolar querying.

%1.5. Querying temporal data
Generally, in querying, temporal data are handled specifically. Because of the temporal capacity and interpretation of temporal data, users usually like expressing their temporal constraints or preferences using a specific set of temporal operators~\cite{Galindo2001},~\cite{Pons2012ijcis},~\cite{Schockaert2008ieeetfs}. Often, such temporal operators are based on the possible temporal relationships between two temporal values~\cite{Galindo2001},~\cite{Pons2012ijcis},~\cite{Schockaert2008ieeetfs}. Such temporal relationships express semantically meaningful relationships between two temporal values. In~\cite{Allen1983cacm} a groundbreaking collection of temporal relationships between two time intervals (and as a special case time instants or time points)~\cite{Dyreson1994sigmod},~\cite{Bohlen1998lncs} is presented. To query temporal data in an imperfect way, fuzzy variants of temporal operators are of course necessary. Thus, several proposals have considered fuzzy variants of temporal relationships~\cite{Schockaert2008ieeetfs},~\cite{Pons2013ijufkbs} and of temporal operators used for the fuzzy querying of temporal databases~\cite{Galindo2001},~\cite{Pons2012ipmu}.

%1.5. About the presented work
%1.5.1. Situating among others
%1.5.2. content
Several existing proposals have covered the regular or fuzzy querying of valid-time relational databases containing uncertain valid-time da\-ta~\cite{Pons2012ijcis},~\cite{Pons2012ipmu},~\cite{Pons2013ijufkbs}. However, to the knowledge of the authors, few proposals have considered the bipolar querying of valid-time databases and even fewer the bipolar querying of valid-time databases containing temporal data subject to uncertainty. The presented work tries to fill part of this gap by presenting a novel technique to query a valid-time relation containing uncertain valid-time data in a bipolar way. This novel technique allows the specification of a valid-time constraint related to each bipolar elementary query constraint. This novel technique is presented in Section~\ref{sec:proposal}. In Section~\ref{sec:preliminaries}, some necessary preliminary concepts and techniques are clarified and in Section~\ref{sec:conclusions}, the conclusions of this paper and some directions for future research are given.

\section{Preliminaries}

\subsection{Relational Databases and Time}

\subsection{Valid-time Relations Subject to Uncertainty}

\subsection{Ill-known Time Intervals}

\subsection{Convexity}

\subsection{Evaluation of Temporal Relationships}

\subsection{Bipolar Querying}

\section{A Novel Querying Approach}

\section{Conclusions}

\bibliographystyle{splncs}
\bibliography{sources}

\end{document}

### 2013-03-27 fqas2013/paper.tex
Small suggested change.
The abstract and introduction are fine now, the bibliography is extensive.

@@
 
-\title{Bipolar Valid-time Querying}
+\title{Bipolar Querying of Valid-time Intervals Subject to Uncertainty}
 
@@
 \begin{abstract}
-Here goes the abstract.
-\keywords{keyword1, keyword2}
+Databases model parts of reality by containing data representing properties of real-world objects or concepts.
+Often, some of these properties are time-related.
+Thus, databases often contain data representing temporal values.
+However, as they may be produced by humans, such data may contain imperfections like uncertainties.
+Now, an important purpose of databases is to allow their data to be queried to allow access to the information these data represent.
+In queries, users describe their preferences concerning the data they are interested in.
+Because users may have both positive and negative preferences, they may want to query databases in a bipolar way.
+Such preferences may also have a temporal nature.
+In this paper, a novel technique is presented to query a valid-time relation containing uncertain valid-time data in a bipolar way, which allows the query to have a single bipolar temporal query constraint.
+
+\keywords{Bipolar Querying, Valid-time Relation, Uncertainty, Temporal Databases, Ill-known Time Intervals}
 \end{abstract}
@@
 \section{Introduction}
+The general purpose of database systems (DS) is to model (a part of) reality.
+Therefore, databases contain data representing properties of real-world objects or concepts~\cite{Pons2012ijcis},~\cite{Billiet2012ipmu},~\cite{Pons2012ipmu}.
+Some essential properties of real-world objects or concepts are time-related.
+Thus, databases often contain data representing temporal values which describe such properties~\cite{Pons2012ijcis},~\cite{Bolour1982acm},~\cite{Jensen1999ieeetkde},~\cite{Galindo2001}.
+These temporal values are usually either time intervals~\cite{Bohlen1998lncs} or instants~\cite{Bohlen1998lncs}, which may informally be seen as infinitesimally short `points' in time.
+Based on their interpretation and purpose, temporal values can be classified into several categories, but the presented work will only consider valid-time notions, which indicate when database data is a true representation of the reality modelled by the database~\cite{Bohlen1998lncs},~\cite{Pons2012ijcis},~\cite{Billiet2012ipmu},~\cite{Pons2012ipmu}.
+
+A lot of data in DS are produced by humans, but human-made data are prone to imperfections: some data may be vague, imprecise, contradictory, incomplete or uncertain~\cite{Billiet2012ipmu},~\cite{Pons2012ipmu},~\cite{Medina1994is},~\cite{Bosc2010ijufkbs},~\cite{Bosc2009sum}.
+Of course, such imperfections may arise in data representing temporal notions too~\cite{Billiet2012ipmu},~\cite{Pons2012ipmu},~\cite{Pons2012ijcis},~\cite{Dyreson1998acm}.
+The presented work will consider DS containing data representing time intervals subject to uncertainty and will assume all non-temporal data in these DS to contain no imperfections.
+
+One of the most important purposes of a database is allowing its data to be queried, which allows the information or knowledge represented by the data to be retrieved.
+A database may be queried in a `regular' way: the user describes the data which he or she finds desired or satisfactory and thus wants to retrieve, by perfectly describing the preferred values of this data.
+A database may also be queried in a `fuzzy' way: the user describes the data which he or she finds desired or satisfactory by imperfectly describing the preferred values of this data ~\cite{Zadrozny2008hrfipd},~\cite{DeCaluwe2007ijis}.
+These imperfect descriptions may contain vagueness or imprecision, often through the use of linguistic terms ~\cite{Devos1998jql},~\cite{Kacprzyk2001is}.
+A database may also be queried in a `bipolar' way.
+Generally, two main approaches to this exist.
+One is for the user to describe the data which he or she finds acceptable and to describe the data among this acceptable data, which he or she finds really desired (wished-for), both by describing the preferred values of this data ~\cite{Dubois2008hrfipd},~\cite{Dubois2002lnai}.
+The other is for the user to describe both the data which he or she finds desired or satisfactory and the data which he or she finds undesired or unsatisfactory, both by describing the preferred values of this data~\cite{DeTre2010ieeetfs},~\cite{Matthe2011ijis}.
+The descriptions used in bipolar querying might or might not contain imperfections.
+The presented work will only consider the latter approach to bipolar querying and will allow a simple form of imprecision in non-temporal elementary query conditions.
+
+In querying, temporal preferences are usually expressed specifically.
+Because of their temporal nature and interpretation, users usually prefer using specific temporal operators to express temporal preferences.
+Hence, several proposals have considered specific sets of temporal operators~\cite{Galindo2001},~\cite{Pons2012ijcis},~\cite{Schockaert2008ieeetfs}, often based on the possible temporal relationships between two time indications~\cite{Galindo2001},~\cite{Pons2012ijcis},~\cite{Schockaert2008ieeetfs}.
+Such temporal relationships define semantically meaningful relationships with a temporal nature, between two time indications.
+In~\cite{Allen1983cacm}, a groundbreaking collection of temporal relationships between two time intervals (and as a special case instants) is introduced.
+Of course, to query temporal data in a fuzzy way, fuzzy variants of temporal operators are necessary and several proposals have thus introduced such variants~\cite{Galindo2001},~\cite{Pons2012ipmu}, often based on fuzzy variants of temporal relationships~\cite{Schockaert2008ieeetfs},~\cite{Pons2013ijufkbs}.
+
+Techniques for the regular or fuzzy querying of valid-time databases containing uncertain valid-time data are considered by several existing proposals~\cite{Pons2012ijcis},~\cite{Pons2012ipmu},~\cite{Pons2013ijufkbs}.
+However, to the knowledge of the authors, few proposals have considered the bipolar querying of valid-time databases~\cite{Billiet2011fqas} and even fewer the bipolar querying of valid-time databases containing temporal data subject to uncertainty.
+This paper presents a novel technique to query a valid-time relation containing uncertain valid-time data in a bipolar way, which allows the query to have a single bipolar temporal query constraint.
+This paper is structured as follows: in section \ref{sec:preliminaries}, some necessary preliminary concepts and techniques are described, in section \ref{sec:proposal} the novel technique which is the contribution of the work presented in this paper is presented and in section \ref{sec:conclusions}, the conclusions of this paper and some directions for future research are given.
+
+\section{Preliminaries\label{sec:preliminaries}}
+
+\subsection{Time in Relational Databases}
+TODO: explain everything about time in relational databases (categories of time indications, mostly, and such basic things)
+
+\subsection{Valid-time Relations}
+TODO: 'formally' explain what we will call a relation
+TODO: 'formally' explain what we will cal a valid-time relation
+
+\subsection{Uncertainty in Valid-time Intervals}
+TODO: explain all theory for IKTI
+TODO: explain that we will require the poss. dist. to be convex and what this means
+
+\subsection{Evaluation of Temporal Relationships}
+TODO: explain how temporal relationships boil down to combinations of Allen relationships and what the crisp Allen relationships are
+TODO: explain all theory for IKC
+TODO: explain
+
+\subsection{Bipolar Querying}
+
+\section{A Novel Querying Approach\label{sec:proposal}}
+
+\subsection{Valid-time Relations Subject to Uncertainty}
+TODO: mention that we will consider valid-time relations subject to uncertainty and how they are composed (refer to earlier work)
+TODO: present the example table
+
+\subsection{Querying Using Bipolar Valid-time Constraints}
+TODO: present the used query structure
+
+\subsection{Evaluation}
+TODO: explain how temporal query evaluation requires the use of Allen Relationships, so in this case a poss. equivalent has to be found
+TODO: explain how this equivalent is provided by the IKC framework.
+TODO: present the used evaluation technique
+
+\subsection{Aggregation and Ranking}
+TODO: figure correct interpretation and subsequent aggregation out
+
+\section{Conclusions and Future Work\label{sec:conclusions}}
 
 \bibliographystyle{splncs}
-\bibliography{paper}
+\bibliography{sources}
 

### 2013-04-04 trunk/fqas2013/paper.tex
A reworking of the abstract, introduction and the first 3 paragraphs of the preliminaries.

@@
 \usepackage{amssymb}
+\usepackage{amsmath}
+\usepackage{graphicx}
+\usepackage{multirow}
 \setcounter{tocdepth}{3}
-\usepackage{graphicx}
+
+\newcommand{\Pos}{\operatorname{Pos}}
+\newcommand{\Nec}{\operatorname{Nec}}
+
+\newcommand{\mytitlestring}{Bipolar Querying of Valid-time Intervals Subject to Uncertainty}
 
@@
 
-\title{Bipolar Querying of Valid-time Intervals Subject to Uncertainty}
+\title{\mytitlestring}
 
 %Short form of title for running heads
-\titlerunning{Bipolar Valid-time Querying}
+\titlerunning{\mytitlestring}
 
@@
 Sint-Pietersnieuwstraat 41, B-9000, Ghent, Belgium\\
-Christophe.Billiet@UGent.be, Guy.DeTr\'{e}@UGent.be\\
+Christophe.Billiet@UGent.be, Guy.DeTre@UGent.be\\
 \and
@@
 \begin{abstract}
-Here goes the abstract.
-\keywords{keyword1, keyword2}
+Databases model parts of reality by containing data representing properties of real-world objects or concepts.
+Often, some of these properties are time-related.
+Thus, databases often contain data representing time-related information.
+However, as they may be produced by humans, such data or information may contain imperfections like uncertainties.
+An important purpose of databases is to allow their data to be queried, to allow access to the information these data represent.
+Users may do this using queries, in which they describe their preferences concerning the data they are (not) interested in.
+Because users may have both positive and negative such preferences, they may want to query databases in a bipolar way.
+Such preferences may also have a temporal nature, but, traditionally, temporal query constraints are handled specifically.
+In this paper, a novel technique is presented to query a valid-time relation containing uncertain valid-time data in a bipolar way, which allows the query to have a single bipolar temporal query constraint.
+
+\keywords{Bipolar Querying, Valid-time Relation, Valid Time, Temporal Databases, Uncertainty, Possibility Theory, Ill-known Intervals}
 \end{abstract}
@@
 \section{Introduction}
-%1.1. On databases and their purpose
-Databases are collections of data.
-These data are usually the result of measurements or descriptions of aspects or properties of real-world objects or concepts.
-As these data represent properties of objects or concepts in reality, the database itself represents a part of reality~\cite{Billiet2012ipmu},~\cite{Pons2012ipmu}.
+Generally, database systems model (parts of) reality.
+For this, their databases contain data representing properties of real-world objects or concepts~\cite{Pons2012ijcis},~\cite{Billiet2012ipmu},~\cite{Pons2012ipmu}.
+Some essential properties of real-world objects or concepts are time-related.
+Thus, databases often contain data representing temporal values~\cite{Bohlen1998lncs}, which are basically indications of time and describe such properties~\cite{Pons2012ijcis},~\cite{Jensen1999ieeetkde},~\cite{Galindo2001}.
+These temporal values are usually either time intervals~\cite{Bohlen1998lncs} or instants~\cite{Bohlen1998lncs}, which may informally be seen as infinitesimally short `periods' or `points' in time.
+Based on their interpretation and purpose, temporal values can be classified into several categories, but the presented work will only consider valid-time indications, which indicate when corresponding data is a valid or true representation of the reality modelled by its database~\cite{Bohlen1998lncs},~\cite{Pons2012ijcis},~\cite{Billiet2012ipmu},~\cite{Pons2012ipmu}.
 
-%1.2. Time in databases
-To many real-world concepts or objects, time is an essential aspect.
-E.g., certain historical events are meaningless without time frame.
-Therefore, many databases contain data representing temporal notions which describe the temporal properties of real-world objects or concepts.
-In the context of database systems, such temporal values~\cite{Dyreson1994sigmod},~\cite{Bohlen1998lncs} can be classified into different categories, based on their interpretation and purpose:
+A lot of database data are produced by humans, but human-made data are prone to imperfections, as some of these data may be vague, imprecise, contradictory, incomplete or uncertain~\cite{Billiet2012ipmu},~\cite{Pons2012ipmu},~\cite{Medina1994is},~\cite{Bosc2010ijufkbs}.
+Of course, data representing temporal values may contain such imperfections too~\cite{Billiet2012ipmu},~\cite{Pons2012ipmu},~\cite{Pons2012ijcis},~\cite{Dyreson1998acm}.
+The work presented in this paper will consider databases containing data representing valid-time intervals subject to uncertainty and will assume all non-temporal data in these databases to contain no imperfections.
 
-\begin{itemize}
-	\item \textbf{valid-time indications} describe when a database fact or data is a true or valid representation of the reality modelled by the database~\cite{Dyreson1994sigmod},~\cite{Bohlen1998lncs}.
-	\item \textbf{transaction-time indications} describe when a fact or data is current in the database, which means it is not logically deleted and can thus be retrieved~\cite{Dyreson1994sigmod},~\cite{Bohlen1998lncs}.
-	\item \textbf{decision-time indications} describe when certain events were decided to happen~\cite{Nascimento1995siwtrr}.
-\end{itemize}
+One of the most important purposes of a database is to allow its data to be queried, to allow the information or knowledge represented by this data to be retrieved.
+A user may query a database in a `regular' way: the user describes the data which he or she finds desired or satisfactory and thus wants to retrieve, by perfectly describing the allowed values of these data.
+A user may also query a database in a `fuzzy' way: the user describes the data which he or she finds desired or satisfactory by imperfectly describing the allowed values of these data~\cite{Zadrozny2008hrfipd}.
+These imperfect descriptions may contain vagueness or imprecision, often through the use of linguistic terms ~\cite{Devos1998jql},~\cite{Kacprzyk2001is}.
+A user may also query a database in a `bipolar' way.
+Generally, two main approaches to this exist.
+One is for the user to describe the data which he or she finds acceptable and to describe the data among this acceptable data, which he or she finds really desired, both by describing the allowed values of these data~\cite{Dubois2002lnai}.
+The other is for the user to independently describe both the data which he or she finds desired or satisfactory and the data which he or she finds undesired or unsatisfactory, both by describing the allowed values of these data~\cite{DeTre2010ieeetfs},~\cite{Matthe2011ijis}.
+The descriptions used in bipolar querying may contain imperfections.
+The presented work will only consider the latter approach to bipolar querying and will allow a simple form of imprecision in non-temporal elementary query conditions.
 
-%1.3. Imperfect data
-%1.3.1. General
-%1.3.2. In time
-A lot of data is made by humans.
-Human-made data is prone to imperfections: it can be vague or imprecise, may contain contradictions, be incomplete or contain uncertainties~\cite{Pons2012ipmu},~\cite{Pons2012ijcis}.
-Uncertainties in data may be caused by variability in the outcomes of an experiment and confidence in the context of such uncertainty may be modelled using probability theory~\cite{Pons2013ijufkbs}.
-Uncertainties in data may also be caused by a (partial) lack of knowledge: the exact value which is the answer to a certain question might not be known, even if there is just one correct answer and as such no variability.
-Confidence in the context of this kind of uncertainty may be modelled using possibility theory~\cite{Zadeh1978fss},~\cite{Dubois1988},~\cite{Pons2013ijufkbs}, ~\cite{Pons2012ipmu},~\cite{Pons2012ijcis}.
-Uncertainties caused by a (partial) lack of knowledge may exist in temporal notions~\cite{Pons2012ijcis},~\cite{Pons2012ipmu},~\cite{Pons2013ijufkbs}.
-E.g., consider a medieval document defining a legal act.
-The document contains the date defining the day on which the legal act took effect.
-Consider the document damaged, making this date unreadable.
-In this case, there is uncertainty about the exact day on which the legal act took effect and this uncertainty is caused by the lack of some knowledge: it is known that there is exactly one day on which the act took effect, but it is not known which one.
+Compared to non-temporal user preferences, temporal user preferences usually have an uncommon nature and interpretation and thus expressing them relies on uncommon mechanics in querying: users usually prefer using specific temporal operators to express temporal preferences.
+Hence, several proposals have considered specific sets of temporal operators~\cite{Galindo2001},~\cite{Pons2012ijcis},~\cite{Schockaert2008ieeetfs}, often based on the possible temporal relationships between two time indications~\cite{Galindo2001},~\cite{Pons2012ijcis},~\cite{Schockaert2008ieeetfs}.
+Such temporal relationships define semantically meaningful relationships with a temporal nature, between two time indications.
+In~\cite{Allen1983cacm}, a collection of temporal relationships between two time intervals (and as a special case instants) is introduced and this collection is considered groundbreaking.
+Of course, to query temporal data in a fuzzy way, fuzzy variants of such temporal operators are necessary and several proposals have thus introduced such variants~\cite{Galindo2001},~\cite{Pons2012ipmu}, often based on fuzzy variants of such temporal relationships~\cite{Schockaert2008ieeetfs},~\cite{Pons2013ijufkbs}.
 
-%1.4. Querying databases
-%1.4.1. Crisp querying of databases
-%1.4.2. Fuzzy querying of databases
-%1.4.3. Bipolar querying of databases
-One of the most important purposes of a database is of course to allow the retrieval of information and knowledge deduced from its data.
-Often, this is done by querying the database.
-Databases can be queried in a `regular' way: the user describes the data which are desired or satisfactory to him or her and which he or she thus wants to retrieve, by perfectly describing the allowed values of this data for certain attributes in clearly stated user preferences.
-However, databases can also be queried in a `fuzzy' way: the user describes the data which are desired or satisfactory to him or her by imperfectly describing the allowed values of this data for certain attributes~\cite{DeCaluwe2007ijis}.
-These imperfect descriptions may contain vagueness or imprecision, often through the use of linguistic terms~\cite{Kacprzyk2001is}.
-Databases can also be queried in what is called a `bipolar' way.
-Generally, there are two main approaches to this.
-One is for the user to describe the data which are required (acceptable) to him or her and to describe the data among the required data, which are really desired (wished-for) to him or her, both by describing the allowed values of this data for certain attributes~\cite{Dubois2002lnai}.
-In the other, the user describes the data which are desired or satisfactory to him or her and the data which are undesired or unsatisfactory to him or her, both by describing the allowed values of this data for certain attributes~\cite{DeTre2010ieeetfs},~\cite{Matthe2011ijis}.
-These descriptions might or might not contain imperfections.
-The presented work will only consider the latter approach to bipolar querying.
+Techniques for the regular or fuzzy querying of valid-time databases containing valid-time data subject to uncertainty are considered by several existing proposals~\cite{Pons2012ijcis},~\cite{Pons2012ipmu},~\cite{Pons2013ijufkbs}.
+However, to the knowledge of the authors, few proposals have considered the bipolar querying of valid-time databases~\cite{Billiet2011fqas} and even fewer the bipolar querying of valid-time databases containing temporal data subject to uncertainty.
+Thus, this paper presents a novel technique to query a valid-time relation containing valid-time data subject to uncertainty in a bipolar way, which allows the user to specify a single bipolar temporal query constraint.
+This paper is structured as follows: in section \ref{sec:preliminaries}, some preliminary concepts and techniques are described, in section \ref{sec:proposal}, the novel technique which is the main contribution of the work presented in this paper, is presented and in section \ref{sec:conclusions}, the conclusions of this paper and some directions for future research are given.
 
-%1.5. Querying temporal data
-Generally, in querying, temporal data are handled specifically.
-Because of the temporal capacity and interpretation of temporal data, users usually like expressing their temporal constraints or preferences using a specific set of temporal operators~\cite{Galindo2001},~\cite{Pons2012ijcis},~\cite{Schockaert2008ieeetfs}.
-Often, such temporal operators are based on the possible temporal relationships between two temporal values~\cite{Galindo2001},~\cite{Pons2012ijcis},~\cite{Schockaert2008ieeetfs}.
-Such temporal relationships express semantically meaningful relationships between two temporal values.
-In~\cite{Allen1983cacm} a groundbreaking collection of temporal relationships between two time intervals (and as a special case time instants or time points)~\cite{Dyreson1994sigmod},~\cite{Bohlen1998lncs} is presented.
-To query temporal data in an imperfect way, fuzzy variants of temporal operators are of course necessary.
-Thus, several proposals have considered fuzzy variants of temporal relationships~\cite{Schockaert2008ieeetfs},~\cite{Pons2013ijufkbs} and of temporal operators used for the fuzzy querying of temporal databases~\cite{Galindo2001},~\cite{Pons2012ipmu}.
+\section{Preliminaries\label{sec:preliminaries}}
 
-%1.5. About the presented work
-%1.5.1. Situating among others
-%1.5.2. content
-Several existing proposals have covered the regular or fuzzy querying of valid-time relational databases containing uncertain valid-time da\-ta~\cite{Pons2012ijcis},~\cite{Pons2012ipmu},~\cite{Pons2013ijufkbs}.
-However, to the knowledge of the authors, few proposals have considered the bipolar querying of valid-time databases and even fewer the bipolar querying of valid-time databases containing temporal data subject to uncertainty.
-The presented work tries to fill part of this gap by presenting a novel technique to query a valid-time relation containing uncertain valid-time data in a bipolar way.
-This novel technique allows the specification of a valid-time constraint related to each bipolar elementary query constraint.
-This novel technique is presented in Section~\ref{sec:proposal}.
-In Section~\ref{sec:preliminaries}, some necessary preliminary concepts and techniques are clarified and in Section~\ref{sec:conclusions}, the conclusions of this paper and some directions for future research are given.
+\subsection{General Preliminaries, Notations and Nomenclature}
+Databases may contain data representing temporal values.
+Based on their purpose and interpretation, such time indications can be classified into different categories~\cite{Bohlen1998lncs},~\cite{Billiet2012ipmu}.
+The work presented in this paper only considers temporal values of the category \emph{valid time}.
+Their purpose or interpretation is for every valid-time indication to correspond to a collection of data and to indicate a period of time during which this data is a valid or true representation of reality.
 
-\section{Preliminaries}
+The work presented in this paper concerns time indications subject to uncertainty.
+This uncertainty is always assumed to be caused by a (partial) lack of knowledge: the exact, intended time indication is not known, eventhough there is only one time indication intended and as such no variability.
+Confidence about exactly which time indication is the intended one in the context of such uncertainty is modelled using possibility theory~\cite{Dubois1988},~\cite{Pons2013ijufkbs}.
+In the presented work, `possibility' and `necessity' are always interpreted as measures of plausibility, respectively necessity, given all available knowledge.
+Time intervals not subject to any imperfection are called \emph{crisp time intervals} (CTI) in this paper.
 
-\subsection{Relational Databases and Time}
+\subsection{Valid-time Relations}
+The presented work uses the relational database model~\cite{Pons2012ijcis},~\cite{Galindo2001}.
+Here, a \emph{relation} models a collection of similar real-world concepts or objects.
+To achieve this, such a relation has a set of $n$ \emph{attributes}, of which each describes a different property shared by all of these concepts or objects, and a set of $n$-tuples, of which each represents one of these real-world objects or concepts and contains $n$ values, each representing the capacity of the object or concept corresponding to the tuple, of the property described by the attribute corresponding to the value~\cite{Billiet2012ipmu},~\cite{Pons2012ipmu}.
+
+A \emph{valid-time relation} (VTR) always has \emph{valid-time attributes}.
+These are attributes describing a single valid time~\cite{Bohlen1998lncs} for the objects or concepts represented by the VTR's tuple.
+A VTR may contain different tuples corresponding to the same real-world concept or object.
+The non-valid-time attribute values of such a tuple represent the capacities of their corresponding properties which were or are true or valid for the object or concept corresponding to the tuple during the period in time indicated by the valid-time indication represented by the tuple's valid-time attribute values.
+In the presented work, such valid-time indications will always be time intervals~\cite{Bohlen1998lncs} and will always be referred to as \emph{valid-time intervals} (VTI).
+
+\subsection{Uncertainty in Valid-time Intervals}
+The presented work allows VTI to be subject to uncertainty, by allowing them to be \emph{ill-known valid-time intervals} (IKVTI).
+The concept of IKVTI is based on the concepts of \emph{possibilistic variables} (PV) and \emph{ill-known intervals} (IKI)~\cite{Dubois1988cma},~\cite{Pons2013ijufkbs},~\cite{Billiet2012ipmu}:
+
+\begin{definition}
+\label{def:poss-variable}
+A \emph{possibilistic variable} (PV) $X$ on a universe $U$ is a variable taking exactly one value in $U$, but for which this value is (partially) unknown.
+The possibility distribution $\pi_X$ on $U$ models the available knowledge about the value that $X$ takes: for each $u \in U$, $\pi_X(u)$ represents the possibility that $X$ takes the value $u$.
+\end{definition}
+
+Now consider a set $U$ containing single values (and not collections of values).
+When a PV $X_{v}$ is defined on such a set $U$, the unique value $X_{v}$ takes, which is (partially) unknown, will be a single value in $U$ and is called an \emph{ill-known value} (IKV) in $U$ in the presented work~\cite{Dubois1988cma},~\cite{Pons2013ijufkbs},~\cite{Billiet2012ipmu}.
+In this paper, IKV will be denoted using lower-case letters.
+The work presented in this paper uses a specific kind of IKI, defined as follows, although other definitions exist~\cite{Pons2012ipmu},~\cite{Billiet2012ipmu}:
+
+\begin{definition}
+Consider an ordered set $U$. An \emph{ill-known interval} (IKI) in $U$ is an interval in $U$ of which both boundary values are IKV in $U$.
+\end{definition}
+
+Specifically concerning valid time, an IKI in a time domain modelled by the domains of a VTR's valid-time attributes is called an \emph{ill-known valid-time interval} (IKVTI) in the presented work.
+The work presented in this paper requires the possibility distributions defining an IKVTI's IKV to be convex~\cite{Pons2013ijufkbs}.
+In this paper, an IKVTI with boundary IKV $s$ and $e$ will be noted $\left[s, e\right]$.
+
+\subsection{Evaluation of Temporal Relationships}
+In~\cite{Allen1983cacm}, Allen presents a collection of basic temporal relationships.
+Using these, every meaningful temporal relationship between time intervals can be expressed.
+Thus, it suffices to consider the evaluation of these Allen relationships.
+When uncertainty is involved, evaluation can be done using the \emph{ill-known constraints} (IKC) framework presented in~\cite{Pons2013ijufkbs}.
+
+\begin{definition}
+Given an ordered set $U$, an \emph{ill-known constraint} (IKC) $C$ on $U$ is specified by means of a binary relation $R \subseteq U^{2}$ and a fixed IKV $v$ in $U$ defined by its possibilistic variable $V$ on $U$, i.e.:
+\begin{align}
+C \triangleq (R,v) \nonumber
+\end{align}
+Any set $A \subseteq U$ now satisfies IKC $C = (R,v)$ if and only if:
+\begin{align}
+\forall a \in A : (a,v) \in R \nonumber
+\end{align}
+\end{definition}
+
+The satisfaction of an IKC $C$ by a set $A$ will be noted $C(A)$ in this paper.
+Consider an ordered set $U$, an IKC $C \triangleq (R,v)$ on $U$ and a set $A \subseteq U$.
+Due to the uncertainty about the exact value of $v$, it is uncertain whether $A$ satisfies $C$ or not.
+Confidence about the plausibility that $A$ satisfies $C$ can be expressed using a possibility degree $\Pos(C(A))$ and a necessity degree $\Nec(C(A))$, which are shown to be calculated as follows:
+
+\begin{align}
+\Pos(C(A)) & = \min_{a \in A}\left(\sup_{(a,w) \in R}\pi_{V}(w)\right) \label{ill-known-pos}\\
+\Nec(C(A)) & = \min_{a \in A}\left(\inf_{(a,w) \notin R} 1-\pi_{V}(w)\right) \label{ill-known-nec}
+\end{align}
+
+Given an ordered set $U$, possibility and necessity degrees expressing confidence about the plausibility that a set $A \subseteq U$ satisfies a boolean aggregation of IKC on $U$ can be found by using the possibilistic extensions of boolean operators `and' ($\wedge$), `or' ($\vee$) and `not' ($\neg$)~\cite{Pons2013ijufkbs},~\cite{Pons2012ipmu},~\cite{Billiet2012ipmu}.
+
+The IKC framework now allows evaluating a given Allen relationship $AR$ between a given CTI $I$ and a given IKVTI $J = \left[s, e\right]$ by allowing to calculate the possibility and necessity that $I$ $AR$ $J$ holds.
+For this, the combination of $AR$ and $J$ is translated to a specific boolean aggregation of specific IKC.
+These translations are shown in table \ref{tab:allen-relations}.
+Every row of this table corresponds to an Allen relationship.
+The Allen relationships the rows correspond to, are shown in the `Allen Relationship' column, the collections of specific IKC for given Allen relationships are shown in the `Constraints' column (every $C_i, i \in \{1, 2, 3, 4\}$ denotes an IKC) and the specific aggregation of these IKC used for evaluation of the Allen relationships are shown in column `Aggregation'.
+Finally, the possibility and necessity degrees expressing confidence about the plausibility that $I$ $AR$ $J$ holds are then the possibility and necessity expressing confidence about the plausibility that $I$ satisfies the specific aggregation of specific IKC found as translation of the combination of $AR$ and $J$.
+Using the formulas shown above, the requested possibility and necessity degrees can be calculated from these.
+
+%TODO: show this table and explain these things?
+
+\begin{table}[ht]
+\caption{The translations of Allen relationships to the IKC framework.}
+\centering
+\begin{tabular}{|c|c|c|}
+\hline
+Allen Relationship & Constraints & Aggregation \\
+\hline
+I before J & $C_1\stackrel{\triangle}{=} \left(<,s\right)$ & $C_1(I)$ \\
+\hline
+\multirow{2}{*}
+{I equal J} & $C_1\stackrel{\triangle}{=} \left(\geq,s\right)$, $C_2\stackrel{\triangle}{=} \left(\neq,s\right)$ & $C_1(I)\wedge$ $\neg C_2(I)\wedge$\\
+& $C_3\stackrel{\triangle}{=} \left(\leq,e\right)$, $C_4\stackrel{\triangle}{=} \left(\neq,e\right)$ & $C_3(I)\wedge$ $\neg C_4(I)$\\
+\hline
+I meets J & $C_1\stackrel{\triangle}{=} \left(\leq,s\right)$ $C_2\stackrel{\triangle}{=} \left(\neq,s\right)$ & $C_1(I)\wedge$ $\neg C_2(I)$\\
+\hline
+I overlaps J & $C_1\stackrel{\triangle}{=} \left(<,e\right)$, $C_2\stackrel{\triangle}{=} \left(\leq,s\right)$, $C_3\stackrel{\triangle}{=} \left(\geq,s\right)$ & $C_1(I)\wedge$ $\neg C_2(I)\wedge$ $\neg C_3(I)$\\
+\hline
+\multirow{2}{*}
+{I during J} & $C_1\stackrel{\triangle}{=} \left(>,s\right)$, $C_2\stackrel{\triangle}{=} \left(\leq,e\right)$ & $\big(C_1(I)\wedge$ $ C_2(I)\big)\vee$\\
+ & $C_3\stackrel{\triangle}{=} \left(\geq,s\right)$, $C_4\stackrel{\triangle}{=} \left(<,e\right)$ & $\big(C_3(I)\wedge$ $C_4(I)\big)$\\
+\hline
+I starts J & $C_1\stackrel{\triangle}{=} \left(\geq,s\right)$, $C_2\stackrel{\triangle}{=} \left(\neq,s\right)$ & $C_1(I)\wedge$, $\neg C_2(I)$\\
+\hline
+I finishes J & $C_1\stackrel{\triangle}{=} \left(\leq,e\right)$, $C_2\stackrel{\triangle}{=} \left(\neq,e\right)$ & $C_1(I)\wedge$ $\neg C_2(I)$\\
+\hline
+\end{tabular}
+\label{tab:allen-relations}
+\end{table}
+
+\subsection{Bipolar Querying}
+As mentioned before, people may express their query preferences using both positive and negative query conditions.
+If the semantics of these conditions are non-symmetric, meaning that the positive preferences can not be derived from the negative or vice versa, the bipolarity in this query is called \emph{heterogenous}.
+The presented work will consider such heterogeneous bipolar query specification.
+
+Bipolarity in a query can either be specified between or inside the elementary query conditions.
+In~\cite{Matthe2011ijis}, it is shown that combining both approaches makes no sense and that the approach where bipolarity is specified inside elementary query conditions, using intuitionistic fuzzy sets~\cite{Atanassov1986fss}, is a more intuitive one.
+In the presented work, we will only use the latter approach.
+In this approach, elementary query conditions express both what is accepted and what is not accepted by the query, at once.
+Such conditions are called \emph{bipolar query conditions}~\cite{Matthe2011ijis}.
+
+Consider an attribute $A$.
+Let $dom_{A}$ be the domain of $A$'s data type, let $\mu_{c_{A}}$ and $\nu_{c_{A}}$ be membership functions on $dom_{A}$, where $\mu_{c_{A}}(x)$ represents to what extent $x \in dom_{A}$ is satisfactory and $\nu_{c_{A}}(x)$ to what extent $x$ is unsatisfactory to the user, then a bipolar query condition $c_{A}$ expressing a user's preferences about the values of an attribute $A$ can be modelled by an Intuitionistic Fuzzy Set (IFS)~\cite{Atanassov1986fss} as~\cite{Matthe2011ijis},~\cite{DeTre2010ieeetfs}:
+\begin{equation}
+c_{A} = \{(x, \mu_{c_{A}}(x), \nu_{c_{A}}(x)) : x \in dom_{A}\} \nonumber
+\end{equation}
+
+Note that to allow overspecification of the user's preferences, the IFS's consistency condition can be relaxed, which means that 
+\begin{equation}
+\forall x \in dom_{A}: \mu_{c_{A}}(x) + \nu_{c_{A}}(x) \leq 1 \nonumber
+\end{equation}
+does not necessarily have to hold.
+
+If the user explicitely defines $\mu_{c_{A}}(x)$, but doesn't define $\nu_{c_{A}}(x)$, then the non-membership function will be assumed to be the inverse of the membership function, i.e., 
+\begin{equation}
+\nu_{c_{A}}(x) = 1 - \mu_{c_{A}}(x), \forall x \in dom_{A} \nonumber
+\end{equation}
+
+If the user explicitely defines $\nu_{c_{A}}(x)$, but doesn't define $\mu_{c_{A}}(x)$, then the membership function will be assumed to be the inverse of the non-membership function, i.e.,
+\begin{equation}
+\mu_{c_{A}}(x) = 1 - \nu_{c_{A}}(x), \forall x \in dom_{A} \nonumber
+\end{equation}
+
+Thus, in the absence of clear heterogeneousness of the bipolarity, the bipolarity will be assumed homogeneous.
+
+The evaluation of a bipolar query condition $c_A$ results in a so-called {\em bipolar satisfaction degree} (BSD)~\cite{Matthe2011ijis}, which is a pair
+\begin{equation}
+(s,d),\; s,d \in [0,1] \nonumber
+\end{equation}
+where $s$ is called the \emph{satisfaction degree} and $d$ is called the \emph{dissatisfaction degree}~\cite{Matthe2011ijis}.
+Here, $s$ and $d$ are independent from each other and denote to which extent the BSD respectively represents `satisfied' and `dissatisfied'~\cite{Matthe2011ijis}.
+For this, both $s$ and $d$ take values in the unit interval $[0,1]$: extreme values for $s$ and $d$ are 0 (`not at all') and 1 (`fully').
+For example: the BSD $(1,0)$ represents `fully satisfied, not dissatisfied at all', whereas $(0,1)$ represents `not satisfied at all, fully dissatisfied'.
+
+As with bipolar query conditions, there is no consistency condition for BSD's, i.e., a condition like $0\leq s+d\leq 1$ is missing~\cite{Matthe2011ijis}.
+Indeed, because $s$ and $d$ are considered to be completely independent from each other, allowed $s+d>1$ is allowed.
+The motivation is that BSD's try to reflect heterogeneous bipolarity in human reasoning, which can sometimes be inconsistent.
+
+In general, the evaluation of a bipolar query condition $c_{A}$ on attribute $A$ for a relation tuple $R$ will result in a BSD, which is calculated as follows.
+Let $R[A]$ denote the value of tuple $R$ for attribute $A$, then~\cite{Matthe2011ijis}:
+\begin{equation}\label{eq:bsdeval}
+(s_{c_A}^R, d_{c_A}^R)=(\mu_{c_A}(R[A]), \nu_{c_A}(R[A]))
+\end{equation}
+with $s_{c_A}^R$ and $d_{c_A}^R$ the satisfaction degree, respectively dissatisfaction degree, of tuple $R$ for condition $c_A$.
+
+Remark that the traditional approach to fuzzy querying using regular fuzzy sets can be obtained from this as a special case, where the bipolarity is symmetric.
+In that case, a user only specifies positive query preferences~\cite{Matthe2011ijis}.
+%If only positive information is given by the user ($\mu_{c_A}$), this is reduced to \[(s_{c_A}^R, d_{c_A}^R) = (\mu_{c_A}(R[A]),1-\mu_{c_A}(R[A]))\]
+%Analogously, when only negative information is given ($\nu_{c_A}$), this is reduced to \[(s_{c_A}^R, d_{c_A}^R) = (1-\nu_{c_A}(R[A]),\nu_{c_A}(R[A]))\]
+
+\section{A Novel Querying Approach\label{sec:proposal}}
 
 \subsection{Valid-time Relations Subject to Uncertainty}
+TODO: mention that we will consider valid-time relations subject to uncertainty and how they are composed (refer to earlier work)
+TODO: present the example table
 
-\subsection{Ill-known Time Intervals}
+\subsection{Querying Using Bipolar Valid-time Constraints}
+TODO: present the used query structure
 
-\subsection{Convexity}
+\subsection{Evaluation}
+TODO: explain how temporal query evaluation requires the use of Allen Relationships, so in this case a poss. equivalent has to be found
+TODO: explain how this equivalent is provided by the IKC framework.
+TODO: present the used evaluation technique
 
-\subsection{Evaluation of Temporal Relationships}
+\subsection{Aggregation and Ranking}
+TODO: figure correct interpretation and subsequent aggregation out
 
-\subsection{Bipolar Querying}
-
-\section{A Novel Querying Approach}
-
-\section{Conclusions}
+\section{Conclusions and Future Work\label{sec:conclusions}}
 

### 2013-04-12 trunk/fqas2013/paper.tex
Everything finished, only a shortening required (this is the next and final step).

@@
 Because users may have both positive and negative such preferences, they may want to query databases in a bipolar way.
-Such preferences may also have a temporal nature, but, traditionally, temporal query constraints are handled specifically.
-In this paper, a novel technique is presented to query a valid-time relation containing uncertain valid-time data in a bipolar way, which allows the query to have a single bipolar temporal query constraint.
+Such preferences may also have a temporal nature, but, traditionally, temporal query conditions are handled specifically.
+In this paper, a novel technique is presented to query a valid-time relation containing uncertain valid-time data in a bipolar way, which allows the query to have a single bipolar temporal query condition.
 
@@
 However, to the knowledge of the authors, few proposals have considered the bipolar querying of valid-time databases~\cite{Billiet2011fqas} and even fewer the bipolar querying of valid-time databases containing temporal data subject to uncertainty.
-Thus, this paper presents a novel technique to query a valid-time relation containing valid-time data subject to uncertainty in a bipolar way, which allows the user to specify a single bipolar temporal query constraint.
+Thus, this paper presents a novel technique to query a valid-time relation containing valid-time data subject to uncertainty in a bipolar way, which allows the user to specify a single bipolar temporal query condition.
 This paper is structured as follows: in section \ref{sec:preliminaries}, some preliminary concepts and techniques are described, in section \ref{sec:proposal}, the novel technique which is the main contribution of the work presented in this paper, is presented and in section \ref{sec:conclusions}, the conclusions of this paper and some directions for future research are given.
@@
 A \emph{valid-time relation} (VTR) always has \emph{valid-time attributes}.
-These are attributes describing a single valid time~\cite{Bohlen1998lncs} for the objects or concepts represented by the VTR's tuple.
+These are attributes describing a single valid time~\cite{Bohlen1998lncs} for the objects or concepts represented by the VTR's tuples.
 A VTR may contain different tuples corresponding to the same real-world concept or object.
-The non-valid-time attribute values of such a tuple represent the capacities of their corresponding properties which were or are true or valid for the object or concept corresponding to the tuple during the period in time indicated by the valid-time indication represented by the tuple's valid-time attribute values.
+The non-valid-time attribute values of such a tuple represent the capacities of the properties described by their corresponding attributes which were, are or will be true or valid for the object or concept corresponding to the tuple during the period in time indicated by the valid-time indication represented by the tuple's valid-time attribute values.
+Thus, such a tuple represents the `version' of the object or concept corresponding to this tuple which was, is or will be real or valid or the `state' this object or concept was, is or will be in, during the period in time indicated by the valid-time indication represented by the tuple's valid-time attribute values.
 In the presented work, such valid-time indications will always be time intervals~\cite{Bohlen1998lncs} and will always be referred to as \emph{valid-time intervals} (VTI).
@@
 
-Specifically concerning valid time, an IKI in a time domain modelled by the domains of a VTR's valid-time attributes is called an \emph{ill-known valid-time interval} (IKVTI) in the presented work.
+Specifically concerning valid time, an IKI in a time domain represented by the domains of a VTR's valid-time attributes is called an \emph{ill-known valid-time interval} (IKVTI) in the presented work.
 The work presented in this paper requires the possibility distributions defining an IKVTI's IKV to be convex~\cite{Pons2013ijufkbs}.
@@
 \subsection{Evaluation of Temporal Relationships}
-In~\cite{Allen1983cacm}, Allen presents a collection of basic temporal relationships.
-Using these, every meaningful temporal relationship between time intervals can be expressed.
-Thus, it suffices to consider the evaluation of these Allen relationships.
-When uncertainty is involved, evaluation can be done using the \emph{ill-known constraints} (IKC) framework presented in~\cite{Pons2013ijufkbs}.
+To express temporal elementary query conditions, operators based on temporal relationships are necessary.
+In the presented work, only Allen relationships~\cite{Allen1983cacm} between a CTI and a IKVTI are considered.
+To evaluate such relationships, the \emph{ill-known constraints} (IKC) framework presented in~\cite{Pons2013ijufkbs} is used.
+It relies on the concept of IKC.
 
 \begin{definition}
-Given an ordered set $U$, an \emph{ill-known constraint} (IKC) $C$ on $U$ is specified by means of a binary relation $R \subseteq U^{2}$ and a fixed IKV $v$ in $U$ defined by its possibilistic variable $V$ on $U$, i.e.:
+Given an ordered set $U$, an \emph{ill-known constraint} (IKC) $C$ on $U$ is specified by means of a binary relation $R \subseteq U^{2}$ and a fixed IKV $v$ in $U$, i.e.:
 \begin{align}
@@
 Consider an ordered set $U$, an IKC $C \triangleq (R,v)$ on $U$ and a set $A \subseteq U$.
-Due to the uncertainty about the exact value of $v$, it is uncertain whether $A$ satisfies $C$ or not.
-Confidence about the plausibility that $A$ satisfies $C$ can be expressed using a possibility degree $\Pos(C(A))$ and a necessity degree $\Nec(C(A))$, which are shown to be calculated as follows:
+Due to the uncertainty inherent to $v$, it is uncertain whether $A$ satisfies $C$ or not.
+The degree of possibility $\Pos(C(A))$ that $A$ satisfies $C$ and the degree of necessity $\Nec(C(A))$ that $A$ satisfies $C$, can be calculated as follows~\cite{Pons2013ijufkbs}:
 
+\vspace{-10pt}
 \begin{align}
@@
 
-Given an ordered set $U$, possibility and necessity degrees expressing confidence about the plausibility that a set $A \subseteq U$ satisfies a boolean aggregation of IKC on $U$ can be found by using the possibilistic extensions of boolean operators `and' ($\wedge$), `or' ($\vee$) and `not' ($\neg$)~\cite{Pons2013ijufkbs},~\cite{Pons2012ipmu},~\cite{Billiet2012ipmu}.
+Given an ordered set $U$, degrees of possibility and necessity that a set $A \subseteq U$ satisfies a boolean combination of IKC on $U$ can be found by using the possibilistic extensions of boolean operators `and' ($\wedge$), `or' ($\vee$) and `not' ($\neg$), as described in~\cite{Pons2013ijufkbs},~\cite{Pons2012ipmu},~\cite{Billiet2012ipmu}.
 
-The IKC framework now allows evaluating a given Allen relationship $AR$ between a given CTI $I$ and a given IKVTI $J = \left[s, e\right]$ by allowing to calculate the possibility and necessity that $I$ $AR$ $J$ holds.
-For this, the combination of $AR$ and $J$ is translated to a specific boolean aggregation of specific IKC.
+The IKC framework now allows evaluating a given Allen relationship $AR$ between a given CTI $I$ and a given IKVTI $J = \left[s, e\right]$ by allowing the calculation of the degrees of possibility and necessity that $I$ $AR$ $J$ holds.
+For this, the combination of $AR$ and $J$ is translated to a specific boolean combination of specific IKC.
 These translations are shown in table \ref{tab:allen-relations}.
-Every row of this table corresponds to an Allen relationship.
-The Allen relationships the rows correspond to, are shown in the `Allen Relationship' column, the collections of specific IKC for given Allen relationships are shown in the `Constraints' column (every $C_i, i \in \{1, 2, 3, 4\}$ denotes an IKC) and the specific aggregation of these IKC used for evaluation of the Allen relationships are shown in column `Aggregation'.
-Finally, the possibility and necessity degrees expressing confidence about the plausibility that $I$ $AR$ $J$ holds are then the possibility and necessity expressing confidence about the plausibility that $I$ satisfies the specific aggregation of specific IKC found as translation of the combination of $AR$ and $J$.
+Every row of this table corresponds to a given Allen relationship between $I$ and $J$, indicated by the row's value in the `Allen Relationship' column.
+The collections of specific IKC for given Allen relationships are shown in the `Constraints' column (every $C_i, i \in \{1, 2, 3, 4\}$ denotes an IKC) and the specific combination of these IKC used for evaluation of the Allen relationships are shown in the `Combination' column.
+Finally, the degrees of possibility and necessity that $I$ $AR$ $J$ holds are then the degrees of possibility, respectively necessity that $I$ satisfies the specific aggregation of specific IKC found as translation of the combination of $AR$ and $J$.
 Using the formulas shown above, the requested possibility and necessity degrees can be calculated from these.
-
-%TODO: show this table and explain these things?
 
@@
 \hline
-Allen Relationship & Constraints & Aggregation \\
+\textbf{Allen Relationship} & \textbf{Constraints} & \textbf{Combination} \\
+\hline
 \hline
@@
 \subsection{Bipolar Querying}
-As mentioned before, people may express their query preferences using both positive and negative query conditions.
-If the semantics of these conditions are non-symmetric, meaning that the positive preferences can not be derived from the negative or vice versa, the bipolarity in this query is called \emph{heterogenous}.
-The presented work will consider such heterogeneous bipolar query specification.
+As mentioned before, humans may express their query preferences using both positive and negative query conditions~\cite{Matthe2011ijis},~\cite{DeTre2010ieeetfs}.
+If the semantics of these conditions are non-symmetric, meaning that the positive preferences can not be derived from the negative or vice versa, the bipolarity in this query is called \emph{heterogenous}~\cite{Matthe2011ijis}.
+The presented work will concern only such heterogenous query bipolarity.
 
@@
 In~\cite{Matthe2011ijis}, it is shown that combining both approaches makes no sense and that the approach where bipolarity is specified inside elementary query conditions, using intuitionistic fuzzy sets~\cite{Atanassov1986fss}, is a more intuitive one.
-In the presented work, we will only use the latter approach.
-In this approach, elementary query conditions express both what is accepted and what is not accepted by the query, at once.
-Such conditions are called \emph{bipolar query conditions}~\cite{Matthe2011ijis}.
+In the presented work, only the latter approach is used.
+In this approach, elementary query conditions express both what is accepted and what is not accepted by the query, at once, and are called \emph{bipolar query conditions}~\cite{Matthe2011ijis}.
 
-Consider an attribute $A$.
-Let $dom_{A}$ be the domain of $A$'s data type, let $\mu_{c_{A}}$ and $\nu_{c_{A}}$ be membership functions on $dom_{A}$, where $\mu_{c_{A}}(x)$ represents to what extent $x \in dom_{A}$ is satisfactory and $\nu_{c_{A}}(x)$ to what extent $x$ is unsatisfactory to the user, then a bipolar query condition $c_{A}$ expressing a user's preferences about the values of an attribute $A$ can be modelled by an Intuitionistic Fuzzy Set (IFS)~\cite{Atanassov1986fss} as~\cite{Matthe2011ijis},~\cite{DeTre2010ieeetfs}:
+Consider a relation attribute $A$.
+Let $dom_{A}$ be the domain of $A$'s data type, let $\mu_{c_{A}}$ and $\nu_{c_{A}}$ be membership functions from $dom_{A}$ to the unit interval $\left[0, 1\right]$, where $\mu_{c_{A}}(x)$ represents to what extent $x \in dom_{A}$ is satisfactory and $\nu_{c_{A}}(x)$ to what extent $x$ is unsatisfactory to a user, then a bipolar query condition $c_{A}$ expressing this user's preferences concerning $A$ can be modelled by an Intuitionistic Fuzzy Set (IFS)~\cite{Atanassov1986fss} as~\cite{Matthe2011ijis},~\cite{DeTre2010ieeetfs}:
 \begin{equation}
-c_{A} = \{(x, \mu_{c_{A}}(x), \nu_{c_{A}}(x)) : x \in dom_{A}\} \nonumber
+c_{A} = \{(v, \mu_{c_{A}}(v), \nu_{c_{A}}(v)) : v \in dom_{A}\}
 \end{equation}
 
-Note that to allow overspecification of the user's preferences, the IFS's consistency condition can be relaxed, which means that 
-\begin{equation}
-\forall x \in dom_{A}: \mu_{c_{A}}(x) + \nu_{c_{A}}(x) \leq 1 \nonumber
-\end{equation}
-does not necessarily have to hold.
+Note that to allow overspecification of the user's preferences, the IFS's consistency condition can be relaxed, which means that there may exist values $v \in dom_{A}$ for which $\mu_{c_{A}}(v) + \nu_{c_{A}}(v) > 1$~\cite{Matthe2011ijis},~\cite{DeTre2010ieeetfs}.
 
-If the user explicitely defines $\mu_{c_{A}}(x)$, but doesn't define $\nu_{c_{A}}(x)$, then the non-membership function will be assumed to be the inverse of the membership function, i.e., 
-\begin{equation}
-\nu_{c_{A}}(x) = 1 - \mu_{c_{A}}(x), \forall x \in dom_{A} \nonumber
-\end{equation}
+If the user explicitely defines $\mu_{c_{A}}$, but doesn't define $\nu_{c_{A}}$, then $\nu_{c_{A}}$ will be assumed to be the inverse of $\mu_{c_{A}}$~\cite{Matthe2011ijis}.
+If the user explicitely defines $\nu_{c_{A}}$, but doesn't define $\mu_{c_{A}}$, then $\mu_{c_{A}}$ will be assumed to be the inverse of $\nu_{c_{A}}$~\cite{Matthe2011ijis}.
+Thus, in the absence of clear heterogenousness of the bipolarity in a query condition, the bipolarity will be assumed homogenous~\cite{Matthe2011ijis}.
 
-If the user explicitely defines $\nu_{c_{A}}(x)$, but doesn't define $\mu_{c_{A}}(x)$, then the membership function will be assumed to be the inverse of the non-membership function, i.e.,
-\begin{equation}
-\mu_{c_{A}}(x) = 1 - \nu_{c_{A}}(x), \forall x \in dom_{A} \nonumber
-\end{equation}
-
-Thus, in the absence of clear heterogeneousness of the bipolarity, the bipolarity will be assumed homogeneous.
-
-The evaluation of a bipolar query condition $c_A$ results in a so-called {\em bipolar satisfaction degree} (BSD)~\cite{Matthe2011ijis}, which is a pair
+The evaluation of a bipolar query condition results in a so-called {\em bipolar satisfaction degree} (BSD)~\cite{Matthe2011ijis}, which is a pair
 \begin{equation}
@@
 where $s$ is called the \emph{satisfaction degree} and $d$ is called the \emph{dissatisfaction degree}~\cite{Matthe2011ijis}.
-Here, $s$ and $d$ are independent from each other and denote to which extent the BSD respectively represents `satisfied' and `dissatisfied'~\cite{Matthe2011ijis}.
-For this, both $s$ and $d$ take values in the unit interval $[0,1]$: extreme values for $s$ and $d$ are 0 (`not at all') and 1 (`fully').
-For example: the BSD $(1,0)$ represents `fully satisfied, not dissatisfied at all', whereas $(0,1)$ represents `not satisfied at all, fully dissatisfied'.
+Here, $s$ and $d$ are independent from each other and express to which extent the BSD respectively represents `satisfied' and `dissatisfied'~\cite{Matthe2011ijis}.
+Extreme values for $s$ and $d$ are 0 (`not at all') and 1 (`fully').
+For example: the BSD $(1,0)$ represents `fully satisfied, not dissatisfied at all'~\cite{Matthe2011ijis}.
 
-As with bipolar query conditions, there is no consistency condition for BSD's, i.e., a condition like $0\leq s+d\leq 1$ is missing~\cite{Matthe2011ijis}.
-Indeed, because $s$ and $d$ are considered to be completely independent from each other, allowed $s+d>1$ is allowed.
-The motivation is that BSD's try to reflect heterogeneous bipolarity in human reasoning, which can sometimes be inconsistent.
+As explained in~\cite{Matthe2011ijis}, there is no consistency condition for BSD's and for a BSD $(s, d)$, $s+d>1$ is allowed.
+The motivation is that BSD's try to reflect heterogenous bipolarity in human reasoning, which can sometimes be inconsistent.
 
-In general, the evaluation of a bipolar query condition $c_{A}$ on attribute $A$ for a relation tuple $R$ will result in a BSD, which is calculated as follows.
+In general, the evaluation of a bipolar query condition $c_{A}$ on relation attribute $A$ for a relation tuple $R$ will result in a BSD $(s_{c_A}^R, d_{c_A}^R)$, which is calculated as follows.
 Let $R[A]$ denote the value of tuple $R$ for attribute $A$, then~\cite{Matthe2011ijis}:
@@
 \end{equation}
-with $s_{c_A}^R$ and $d_{c_A}^R$ the satisfaction degree, respectively dissatisfaction degree, of tuple $R$ for condition $c_A$.
 
-Remark that the traditional approach to fuzzy querying using regular fuzzy sets can be obtained from this as a special case, where the bipolarity is symmetric.
+Remark that the traditional approach to fuzzy querying using regular fuzzy sets can be obtained from this as a special case, where the bipolarity involved is homogenous.
 In that case, a user only specifies positive query preferences~\cite{Matthe2011ijis}.
-%If only positive information is given by the user ($\mu_{c_A}$), this is reduced to \[(s_{c_A}^R, d_{c_A}^R) = (\mu_{c_A}(R[A]),1-\mu_{c_A}(R[A]))\]
-%Analogously, when only negative information is given ($\nu_{c_A}$), this is reduced to \[(s_{c_A}^R, d_{c_A}^R) = (1-\nu_{c_A}(R[A]),\nu_{c_A}(R[A]))\]
 
@@
 \subsection{Valid-time Relations Subject to Uncertainty}
-TODO: mention that we will consider valid-time relations subject to uncertainty and how they are composed (refer to earlier work)
-TODO: present the example table
+The presented proposal will concern VTR of which the VTI are IKVTI.
+Generally, such a VTR $R$ can be seen as constructed in the following way.
+Let $R$ have $n$ non-temporal attributes $A_{i}, 1 \leq i \leq n, i \in \mathbb{N}$ and two valid-time attributes $VST$ and $VET$.
+Every tuple $T$ of $R$ represents an object or concept version or state which is valid during the time period indicated by the tuple's IKVTI $I_T$.
+This IKVTI $I_T$ is now defined by two IKV, which respectively describe the starting and ending instants of $I_T$.
+The IKV's describing the tuple's IKVTI's starting, respectively ending instants are represented by the tuple's VST, respectively VET values.
+The interpretation here is that the object or concept version or state corresponding to a tuple was, is or will be valid during a period of time, but exactly which period this is intended to be, is unknown.
+Confidence about exactly which period is intended, is modelled by the tuple's IKVTI~\cite{Pons2012ijcis},~\cite{Pons2012ipmu}.
 
-\subsection{Querying Using Bipolar Valid-time Constraints}
-TODO: present the used query structure
+Throughout the rest of this paper, the relation shown in table~\ref{tab:ex-relation} will be used as example relation.
+This relation models the being in effect of legal acts written down in medieval charters.
+However, since the properties of a legal act cannot change once it has taken effect, every legal act has only one version or state.
+This absence of multiple versions or states was deliberately chosen for simplicity in the example.
+Thus, every tuple of the relation corresponds with a legal act.
+A tuple's value for attribute `ID' is a number uniquely identifying the legal act corresponding to the tuple.
+A tuple's value for attribute `Author' is a character string containing only the name of the author of the charter containing the legal act corresponding to the tuple.
+A tuple's IKVTI represents the time period during which the act corresponding to the tuple was in effect.
+For this, every value for $VST$ and $VET$ now defines an IKV describing the day on which the legal act respectively took effect and stopped taking effect.
+For this, every value for $VST$ or $VET$ is a triple $\left[d_1, d_2, d_3\right]$, where $d_1, d_2, d_3$ are elements of the ordered set $\mathbb{T}$ of days in history.
+This triple now defines a triangular (and thus convex) possibility distribution $\pi$ which defines the mentioned IKV and which is defined by (differences in dates in this function prescription are expressed in amounts of days):
 
-\subsection{Evaluation}
-TODO: explain how temporal query evaluation requires the use of Allen Relationships, so in this case a poss. equivalent has to be found
-TODO: explain how this equivalent is provided by the IKC framework.
-TODO: present the used evaluation technique
+\vspace{-10pt}
+\begin{align}
+\pi(x) =
+\begin{cases}
+\frac{x - d_1}{d_2 - d_1}, & \text{if } d_1 \leq x < d_2 \\
+\frac{d_3 - x}{d_3 - d_2}, & \text{if } d_2 \leq x \leq d_3 \\
+0, & \text{else}
+\end{cases}
+\end{align}
+
+\vspace{-10pt}
+
+\begin{table}[ht]
+\caption{The example relation used in this paper.}
+\centering
+\begin{tabular}{|c|c|c|c|}
+\hline
+\textbf{ID} & \textbf{Author} & \textbf{VST} & \textbf{VET} \\
+\hline
+\hline
+1 & Alo\"{i}sius & $\left[4/4/1208, 6/4/1208, 16/4/1208\right]$ & $\left[10/12/1208, 1/1/1209, 26/1/1209\right]$ \\
+\hline
+2 & Theofilus & $\left[2/4/1209, 12/4/1209, 22/4/1209\right]$ & $\left[21/12/1209, 1/1/1210, 21/1/1210\right]$ \\
+\hline
+3 & Gerardus & $\left[14/1/1209, 15/1/1209, 16/1/1209\right]$ & $\left[21/12/1209, 15/1/1210, 25/1/1210\right]$ \\
+\hline
+4 & Euforius & $\left[21/12/1210, 1/1/1211, 11/1/1211\right]$ & $\left[21/12/1211, 1/1/1212, 11/1/1212\right]$ \\
+\hline
+5 & Ambrosius & $\left[11/12/1213, 21/12/1213, 15/1/1214\right]$ & $\left[9/10/1216, 10/10/1216, 15/10/1216\right]$ \\
+\hline
+6 & Alo\"{i}sius & $\left[21/12/1213, 1/1/1214, 11/1/1214\right]$ & $\left[9/6/1217, 9/6/1217, 12/6/1217\right]$ \\
+\hline
+7 & Gerardus & $\left[29/12/1214, 1/1/1215, 8/1/1215\right]$ & $\left[9/6/1217, 10/6/1217, 12/6/1217\right]$ \\
+\hline
+\end{tabular}
+\label{tab:ex-relation}
+\end{table}
+
+\subsection{Querying Using Bipolar Valid-time Conditions}
+The presented work introduces a novel querying technique. Using this technique, a user query $Q$ consists of two separate parts $Q_N$ and $Q_T$:
+
+\begin{equation}
+Q = (Q_N, Q_T) \nonumber
+\end{equation}
+
+Here, $Q_N$ is a boolean combination of bipolar query conditions on non-valid-time attributes, expressing the user's non-temporal demands.
+$Q_T$ expresses the user's valid-time demands and is in fact a single crisp temporal bipolar query condition $((AR_{pos}, I_{pos}), (AR_{neg}, I_{neg}))$, where both $AR_{pos}$ and $AR_{neg}$ are Allen relationships and both $I_{pos}$ and $I_{neg}$ are CTI.:
+
+\begin{equation}
+Q_T = ((AR_{pos}, I_{pos}), (AR_{neg}, I_{neg})) \nonumber
+\end{equation}
+
+The interpretation here is that the user requires an object or concept that has a version or state that complies with his or her non-temporal demands (both positive and negative), that was, is or will be valid during a time interval which is in Allen relationship $AR_{pos}$ with CTI $I_{pos}$ and that wasn't, isn't or won't be valid during a time interval which is in Allen relationship $AR_{neg}$ with CTI $I_{neg}$.
+
+Consider the example relation shown in table~\ref{tab:ex-relation}.
+Now assume a user is interested in the writing style of authors of charters containing legal acts which actually took effect in a certain period of time.
+The user could then query the relation to find all legal acts of which the charter's author is preferably named Alo\"isius, less preferably Euforius and perhaps Eugenius, and of which the author is preferably not named Ambrosius, rather not Theofilus and perhaps not Antonius and which took effect preferably before $2/1/1210$ and preferably not after $1/1/1214$.
+These demands can now be translated to a query $Q_{ex}$ in the following way:
+
+\vspace{-15pt}
+
+\begin{align}
+Q_{ex} = & (Q_{N,ex}, ((AR_{pos,ex}, I_{pos,ex}), (AR_{neg,ex}, I_{neg,ex}))), \\
+Q_{N,ex} = & \{(x, \mu_{ex}(x), \nu_{ex}(x)\}, \forall x \in \mathbb{S} \nonumber \\
+AR_{pos,ex}, AR_{neg,ex} = & DURING \nonumber \\
+I_{pos,ex} = & \left]- infinity, 1/1/1210\right] \nonumber \\
+I_{neg,ex} = & \left[1/1/1214, + infinity\right[ \nonumber
+\end{align}
+\\
+and $\mu_{ex}$ and $\nu_{ex}$ are the membership functions of the fuzzy sets:
+\begin{align}
+\{(Aloisius, 1), (Euforius, 0.7), (Eugenius, 0.1)\} &\text{, respectively} \nonumber \\
+\{(Ambrosius, 1), (Theofilus, 0.7), (Antonius, 0.1)\} & \nonumber
+\end{align}
+
+\subsection{Elementary Query Condition Evaluation}
+Generally, a first step in determining which objects or concepts corresponding to relation tuples to present to a user as answer to his or her query towards this relation, is evaluating the elementary query conditions in the user query for every relation tuple.
+In the presented work, for every tuple $T$ of VTR $R$, the following is done separately:
+
+\begin{itemize}
+\item every non-temporal bipolar query condition in $Q_N$ is evaluated as described in~\cite{Matthe2011ijis}, resulting in a BSD for each.
+The interpretation is as described in~\cite{Matthe2011ijis}: the BSD's satisfaction, respectively dissatisfaction degrees express to which extent $T$ satisfies, respectively dissatisfies the user preferences expressed by the query condition.
+\item the query's crisp temporal bipolar query condition $Q_T$ is evaluated as follows: let $I_T$ be the tuple's IKVTI, then separately, both $I_T$ $AR_{pos}$ $I_{pos}$ and $I_T$ $AR_{neg}$ $I_{neg}$ are evaluated using the IKC framework, exactly as explained in~\cite{Pons2013ijufkbs},~\cite{Pons2012ijcis},~\cite{Pons2012ipmu}, resulting in a possibility degree $pos_{I_{pos}}(T)$ and a necessity degree $nec_{I_{pos}}(T)$, respectively a possibility degree $pos_{I_{neg}}(T)$ and a necessity degree $nec_{I_{neg}}(T)$.
+The interpretation here is that $pos_{I_{pos}}(T)$ and $nec_{I_{pos}}(T)$ express the plausibility, respectively necessity, given available knowledge, that the time interval during which the object or concept version or state represented by $T$ is actually valid and which is intended by $I_T$, is in relationship $AR_{pos}$ with $I_{pos}$ and thus complies with the user's positive temporal demand and is preferred, while $pos_{I_{neg}}(T)$ and $nec_{I_{neg}}(T)$ express the plausibility, respectively necessity, given a (partial) lack of knowledge, that the time interval during which the object or concept version or state represented by $T$ is actually valid and which is intended by $I_T$, is in relationship $AR_{neg}$ with $I_{neg}$ and thus complies with the user's negative temporal demand and is not preferred.
+\end{itemize}
+
+Table \ref{tab:ex-evaluation} shows the results of the evaluation of the example query's elementary query conditions for the tuples of the example relation.
+
+\begin{table}[ht]
+\caption{Elementary query condition evaluation results for the example relation.}
+\centering
+\begin{tabular}{|c|c|c|c|}
+\hline
+\textbf{ID} & \textbf{$Q_{N,ex}$} & \textbf{$(pos_{I_{pos}}, nec_{I_{pos}})$} & \textbf{$(pos_{I_{neg}}, nec_{I_{neg}})$} \\
+\hline
+\hline
+1 & $(1, 0)$ & $(1,1)$ & $(0,0)$ \\
+2 & $(0, 0.7)$ & $(1,0)$ & $(0,0)$ \\
+3 & $(0, 0)$ & $(11/25,0)$ & $(0,0)$ \\
+4 & $(0.7, 0)$ & $(0,0)$ & $(0,0)$ \\
+5 & $(0, 1)$ & $(0,0)$ & $(14/25,0)$ \\
+6 & $(1, 0)$ & $(0,0)$ & $(1,0)$ \\
+7 & $(0, 0)$ & $(0,0)$ & $(1,1)$ \\
+\hline
+\end{tabular}
+\label{tab:ex-evaluation}
+\end{table}
+\vspace{-10pt}
 
 \subsection{Aggregation and Ranking}
-TODO: figure correct interpretation and subsequent aggregation out
+Generally, a second step in determining which objects or concepts corresponding to relation tuples to present to a user as answer to his or her query towards this relation, is aggregating, for every tuple in the relation, the tuple's evaluation results for the elementary query conditions, to determine a single `score' expressing how well the tuple complies with the general user request expressed by the combination of the elementary query conditions in the user query.
+Usually, these evaluation results are quantifications of (dis)satisfaction.
+However, in the presented proposal, two different types of evaluation results can be discerned:
+
+\begin{itemize}
+	\item the BSD (dis)satisfaction degrees constitute quantifications of (dis)satisfaction: they quantify to which extent a tuple's attribute values (dis)satisfy a user's non-temporal preferences and thus assess an answer to the question: `To what extent does an object or concept represented in the relation (dis)satisfy the user's request and could thus be a (un)wanted result?'
+	\item the possibility and necessity degrees $pos_{I_{pos}}$, $nec_{I_{pos}}$, respectively $pos_{I_{neg}}$, $nec_{I_{neg}}$ constitute quantifications of plausibility and necessity, given a (partial) lack of knowledge: they quantify the plausibility that a tuple's intended crisp VTI (dis)satisfies the user's temporal preferences by quantifying confidence about exactly which crisp VTI is the tuple's intended VTI and thus assess an answer to the question: `How plausible is it that an object or concept represented in the relation that may or may not (dis)satisfy the user's request actually existed, exists or will exist during the time period requested by the user?'
+\end{itemize}
+
+A fundamental question arises now: should one consider combining quantifications of these different categories?
+On one hand, quantifications of these different categories have clearly different and even orthogonal semantics and it would not be clear what exactly the meaning would be of the result of such a combination or what the semantically most coherent ways would be to further process such combination results.
+Thus, it is important, for every query result tuple presented to the user, to certainly keep both different types of evaluations as separate (meta)data.
+On the other hand, without an unambiguous and straightforward ranking of the query results, the user cannot clearly discern the result tuples which comply well with his or her demands from those which don't, which defeats the purpose of querying.
+Reasonably, such ranking should be based on the elementary query condition evaluation results.
+As there cannot already exist a ranking between quantifications of categories with different semantics, a combination seems to be required.
+In most existing proposals requiring a combination of quantifications of satisfaction and plausibility, both quantifications are combined as to restrict one another.
+The result is usually seen as a quantification of plausibility.
+Below, this approach is translated to the specific situation encountered in the presented work.
+
+For every tuple $T$, the BSD's which are the evaluations of the non-temporal bipolar query conditions are combined to a single BSD $(s_N(T), d_N(T))$ as described in~\cite{Matthe2011ijis}.
+In this combination method, satisfaction degrees expressing how well $T$ fits the user's positive demands are combined with each other and separately, dissatisfaction degrees expressing how well $T$ fits the user's negative demands are combined with each other.
+This reasoning is extended to include $pos_{I_{pos}}(T)$, $nec_{I_{pos}}(T)$, $pos_{I_{neg}}(T)$ and $nec_{I_{neg}}(T)$: a couple $(pos_{pos}(T), nec_{pos}(T))$, respectively $(pos_{neg}(T), nec_{neg}(T))$ is calculated, expressing the plausibility and necessity, given a (partial) lack of knowledge, that the object or concept version or state corresponding to $T$ complies with the user's positive, respectively negative demands.
+This calculation is done as follows:
+
+\vspace{-10pt}
+\begin{align}
+pos_{pos}(T) = & \min(s_N(T), pos_{I_{pos}}(T)) \nonumber \\
+nec_{pos}(T) = &
+	\begin{cases}
+	0, & \text{if } pos_{pos}(T) < 1 \\
+	\min(s_N(T), nec_{I_{pos}}(T)), & \text{else}
+	\end{cases}
+	\nonumber \\
+pos_{neg}(T) = & \min(d_N(T), pos_{I_{neg}}(T)) \nonumber \\
+nec_{neg}(T) = &
+	\begin{cases}
+	0, & \text{if } pos_{neg}(T) < 1 \\
+	\min(d_N(T), nec_{I_{neg}}(T)), & \text{else}
+	\end{cases}
+	\nonumber
+\end{align}
+
+Next, for every object or concept $o$ which is being represented in $R$, a couple $(pos_{pos}(o), nec_{pos}(o))$, respectively $(pos_{neg}(o), nec_{neg}(o))$ is calculated, expressing the plausibility and necessity, given a (partial) lack of knowledge, that $o$ complies with the user's positive, respectively negative demands.
+This calculation is done as follows:
+
+\vspace{-10pt}
+\begin{align}
+(pos_{pos}(o), nec_{pos}(o)) = (\max_{T_{o,i}, 1 \leq i \leq k}(pos_{pos}(T_{o,i})), \max_{T_{o,i}, 1 \leq i \leq k}(nec_{pos}(T_{o,i}))) \nonumber \\
+(pos_{neg}(o), nec_{neg}(o)) = (\max_{T_{o,i}, 1 \leq i \leq k}(pos_{neg}(T)_{o,i}), \max_{T_{o,i}, 1 \leq i \leq k}(nec_{neg}(T)_{o,i})) \nonumber
+\end{align}
+
+Here, $\{T_{o,i}, 1 \leq i \leq k, i \in \mathbb{N}\}$ is the set of tuples $T_{o,i}, 1 \leq i \leq k, i \in \mathbb{N}$ which represent the $k$ versions or states of $o$.
+Next, a tie-break approach is used to rank the objects or concepts $o$ represented in $R$: ranking is done based on the value of $pos_{pos}(o)$, with $nec_{pos}(o)$ as tie-breaker for $pos_{pos}(o)$, with $pos_{neg}(o)$ as tie-breaker for $( pos_{pos}(o), pos_{neg}(o))$ and finally $nec_{neg}(o)$ as tiebreaker for $(pos_{pos}(o), pos_{neg}(o), pos_{neg}(o))$.
+The results of this aggregation and ranking approach for the example relation and query are shown in table \ref{tab:ex-ranking}.
+
+\begin{table}[ht]
+\caption{Aggregation and ranking results for the example relation and query.}
+\centering
+\begin{tabular}{|c|c|c|c|c||c|c|c|}
+\hline
+\textbf{ID} & \textbf{$pos_{pos}(o)$} & \textbf{$nec_{pos}(o)$} & \textbf{$pos_{neg}(o)$} & \textbf{$nec_{neg}(o)$} & \textbf{$Q_{N,ex}$} & \textbf{$(pos_{I_{pos}}, nec_{I_{pos}})$} & \textbf{$(pos_{I_{neg}}, nec_{I_{neg}})$} \\
+\hline
+\hline
+1 & 1 & 1 & 0 & 0 & $(1, 0)$ & $(1,1)$ & $(0,0)$ \\
+2 & 0 & 0 & 0 & 0 & $(0, 0.7)$ & $(1,0)$ & $(0,0)$ \\
+3 & 0 & 0 & 0 & 0 & $(0, 0)$ & $(11/25,0)$ & $(0,0)$ \\
+4 & 0 & 0 & 0 & 0 & $(0.7, 0)$ & $(0,0)$ & $(0,0)$ \\
+5 & 0 & 0 & 14/25 & 0 & $(0, 1)$ & $(0,0)$ & $(14/25,0)$ \\
+6 & 0 & 0 & 0 & 0 & $(1, 0)$ & $(0,0)$ & $(1,0)$ \\
+7 & 0 & 0 & 0 & 0 & $(0, 0)$ & $(0,0)$ & $(1,1)$ \\
+\hline
+\end{tabular}
+\label{tab:ex-ranking}
+\end{table}
 
 \section{Conclusions and Future Work\label{sec:conclusions}}
+In this paper, a novel technique is presented, to query a valid-time relation containing valid-time data subject to uncertainty in a bipolar way, which allows the user to specify a single valid-time bipolar query condition.
+A major issue concerning the necessity to combine quantifications of (dis)satisfaction with quantifications of plausibility resulting from this technique and a solution for this is presented and shortly discussed.
+In the near future, the interactions between these types of quantifications and between uncertainty in valid-time data and bipolar querying will be further studied.
+Also, an approach to allow the valid-time bipolar query condition to be fuzzy will be considered.
 

### 2013-04-12 trunk/fqas2013/paper.tex
Final draft (should be).

@@
 In the presented work, only the latter approach is used.
-In this approach, elementary query conditions express both what is accepted and what is not accepted by the query, at once, and are called \emph{bipolar query conditions}~\cite{Matthe2011ijis}.
+In this approach, elementary query conditions express both what is accepted and what is not accepted by the query, at once, and are called \emph{bipolar query conditions} (BQC)~\cite{Matthe2011ijis}.
 
 Consider a relation attribute $A$.
-Let $dom_{A}$ be the domain of $A$'s data type, let $\mu_{c_{A}}$ and $\nu_{c_{A}}$ be membership functions from $dom_{A}$ to the unit interval $\left[0, 1\right]$, where $\mu_{c_{A}}(x)$ represents to what extent $x \in dom_{A}$ is satisfactory and $\nu_{c_{A}}(x)$ to what extent $x$ is unsatisfactory to a user, then a bipolar query condition $c_{A}$ expressing this user's preferences concerning $A$ can be modelled by an Intuitionistic Fuzzy Set (IFS)~\cite{Atanassov1986fss} as~\cite{Matthe2011ijis},~\cite{DeTre2010ieeetfs}:
+Let $dom_{A}$ be the domain of $A$'s data type, let $\mu_{c_{A}}$ and $\nu_{c_{A}}$ be membership functions from $dom_{A}$ to the unit interval $\left[0, 1\right]$, where $\mu_{c_{A}}(x)$ represents to what extent $x \in dom_{A}$ is satisfactory and $\nu_{c_{A}}(x)$ to what extent $x$ is unsatisfactory to a user, then a BQC $c_{A}$ expressing this user's preferences concerning $A$ can be modelled by an Intuitionistic Fuzzy Set (IFS)~\cite{Atanassov1986fss} as~\cite{Matthe2011ijis},~\cite{DeTre2010ieeetfs}:
+\vspace{-5pt}
 \begin{equation}
@@
 
-The evaluation of a bipolar query condition results in a so-called {\em bipolar satisfaction degree} (BSD)~\cite{Matthe2011ijis}, which is a pair
+The evaluation of a BQC results in a so-called {\em bipolar satisfaction degree} (BSD)~\cite{Matthe2011ijis}, which is a pair
+\vspace{-10pt}
 \begin{equation}
@@
 
-In general, the evaluation of a bipolar query condition $c_{A}$ on relation attribute $A$ for a relation tuple $R$ will result in a BSD $(s_{c_A}^R, d_{c_A}^R)$, which is calculated as follows.
+In general, the evaluation of a BQC $c_{A}$ on relation attribute $A$ for a relation tuple $R$ will result in a BSD $(s_{c_A}^R, d_{c_A}^R)$, which is calculated as follows.
 Let $R[A]$ denote the value of tuple $R$ for attribute $A$, then~\cite{Matthe2011ijis}:
@@
 \subsection{Valid-time Relations Subject to Uncertainty}
-The presented proposal will concern VTR of which the VTI are IKVTI.
+The presented proposal will concern VTR where the VTI are IKVTI.
 Generally, such a VTR $R$ can be seen as constructed in the following way.
@@
 Every tuple $T$ of $R$ represents an object or concept version or state which is valid during the time period indicated by the tuple's IKVTI $I_T$.
-This IKVTI $I_T$ is now defined by two IKV, which respectively describe the starting and ending instants of $I_T$.
-The IKV's describing the tuple's IKVTI's starting, respectively ending instants are represented by the tuple's VST, respectively VET values.
-The interpretation here is that the object or concept version or state corresponding to a tuple was, is or will be valid during a period of time, but exactly which period this is intended to be, is unknown.
+This IKVTI $I_T$ is now defined by two IKV, which respectively describe the starting and ending instants of $I_T$ and are represented by the tuple's VST, respectively VET values.
+The interpretation is that the version or state corresponding to a tuple was, is or will be valid during a period of time, but exactly which period this is intended to be, is unknown.
 Confidence about exactly which period is intended, is modelled by the tuple's IKVTI~\cite{Pons2012ijcis},~\cite{Pons2012ipmu}.
 
-Throughout the rest of this paper, the relation shown in table~\ref{tab:ex-relation} will be used as example relation.
-This relation models the being in effect of legal acts written down in medieval charters.
-However, since the properties of a legal act cannot change once it has taken effect, every legal act has only one version or state.
-This absence of multiple versions or states was deliberately chosen for simplicity in the example.
-Thus, every tuple of the relation corresponds with a legal act.
-A tuple's value for attribute `ID' is a number uniquely identifying the legal act corresponding to the tuple.
-A tuple's value for attribute `Author' is a character string containing only the name of the author of the charter containing the legal act corresponding to the tuple.
-A tuple's IKVTI represents the time period during which the act corresponding to the tuple was in effect.
-For this, every value for $VST$ and $VET$ now defines an IKV describing the day on which the legal act respectively took effect and stopped taking effect.
-For this, every value for $VST$ or $VET$ is a triple $\left[d_1, d_2, d_3\right]$, where $d_1, d_2, d_3$ are elements of the ordered set $\mathbb{T}$ of days in history.
-This triple now defines a triangular (and thus convex) possibility distribution $\pi$ which defines the mentioned IKV and which is defined by (differences in dates in this function prescription are expressed in amounts of days):
-
 \vspace{-10pt}
-\begin{align}
-\pi(x) =
-\begin{cases}
-\frac{x - d_1}{d_2 - d_1}, & \text{if } d_1 \leq x < d_2 \\
-\frac{d_3 - x}{d_3 - d_2}, & \text{if } d_2 \leq x \leq d_3 \\
-0, & \text{else}
-\end{cases}
-\end{align}
-
-\vspace{-10pt}
-
 \begin{table}[ht]
@@
 \end{table}
+\vspace{-10pt}
+
+In this paper, the relation shown in table~\ref{tab:ex-relation} will be used as example relation.
+This relation models the being in effect of medieval legal acts.
+Since the properties of a legal act cannot change once it has taken effect, every legal act has only one version or state.
+This was deliberately done to simplify the example.
+Thus, every tuple of the relation corresponds to a legal act.
+A tuple's value for attribute `ID' is a number uniquely identifying the legal act corresponding to the tuple.
+A tuple's value for attribute `Author' is a character string representation of the name of the author of the legal act corresponding to the tuple.
+A tuple's IKVTI represents the time period during which the act corresponding to the tuple was in effect.
+For this, every value for $VST$, respectively $VET$ represents an IKV describing the day on which the legal act respectively took effect and stopped taking effect.
+For this, every value for $VST$ or $VET$ is a triple $\left[d_1, d_2, d_3\right]$, where $d_1, d_2, d_3$ are elements of the ordered set of days in history.
+Such a triple $\left[d_1, d_2, d_3\right]$ now defines a triangular (and thus convex) possibility distribution $\pi$ which defines the mentioned IKV and which is defined by (differences in dates in this function prescription are expressed in amounts of days):
+
+\vspace{-10pt}
+\begin{align}
+\pi(x) =
+\begin{cases}
+\frac{x - d_1}{d_2 - d_1}, & \text{if } d_1 \leq x < d_2 \\
+\frac{d_3 - x}{d_3 - d_2}, & \text{if } d_2 \leq x \leq d_3 \\
+0, & \text{else}
+\end{cases}
+\end{align}
 
 \subsection{Querying Using Bipolar Valid-time Conditions}
-The presented work introduces a novel querying technique. Using this technique, a user query $Q$ consists of two separate parts $Q_N$ and $Q_T$:
-
-\begin{equation}
-Q = (Q_N, Q_T) \nonumber
-\end{equation}
-
-Here, $Q_N$ is a boolean combination of bipolar query conditions on non-valid-time attributes, expressing the user's non-temporal demands.
-$Q_T$ expresses the user's valid-time demands and is in fact a single crisp temporal bipolar query condition $((AR_{pos}, I_{pos}), (AR_{neg}, I_{neg}))$, where both $AR_{pos}$ and $AR_{neg}$ are Allen relationships and both $I_{pos}$ and $I_{neg}$ are CTI.:
-
-\begin{equation}
-Q_T = ((AR_{pos}, I_{pos}), (AR_{neg}, I_{neg})) \nonumber
-\end{equation}
-
-The interpretation here is that the user requires an object or concept that has a version or state that complies with his or her non-temporal demands (both positive and negative), that was, is or will be valid during a time interval which is in Allen relationship $AR_{pos}$ with CTI $I_{pos}$ and that wasn't, isn't or won't be valid during a time interval which is in Allen relationship $AR_{neg}$ with CTI $I_{neg}$.
+The presented work introduces a novel querying technique.
+According to this technique, a user query $Q$ consists of two separate parts $Q_N$ and $Q_T$: $Q = (Q_N, Q_T)$.
+Here, $Q_N$ is a boolean combination of BQC on non-valid-time attributes, expressing the user's non-temporal demands.
+$Q_T$ expresses the user's valid-time demands and is a single crisp temporal BQC $((AR_{pos}, I_{pos}), (AR_{neg}, I_{neg}))$, where both $AR_{pos}$ and $AR_{neg}$ are Allen relationships and both $I_{pos}$ and $I_{neg}$ are CTI.: $Q_T = ((AR_{pos}, I_{pos}), (AR_{neg}, I_{neg}))$.
+The interpretation is that the user requires an object or concept that has a version or state that complies with his or her non-temporal demands and was, is or will be valid during a time interval which is in Allen relationship $AR_{pos}$ with $I_{pos}$ and wasn't, isn't or won't be valid during a time interval which is in Allen relationship $AR_{neg}$ with $I_{neg}$.
 
 Consider the example relation shown in table~\ref{tab:ex-relation}.
-Now assume a user is interested in the writing style of authors of charters containing legal acts which actually took effect in a certain period of time.
-The user could then query the relation to find all legal acts of which the charter's author is preferably named Alo\"isius, less preferably Euforius and perhaps Eugenius, and of which the author is preferably not named Ambrosius, rather not Theofilus and perhaps not Antonius and which took effect preferably before $2/1/1210$ and preferably not after $1/1/1214$.
+Now assume a user queries the relation to find all legal acts of which the author is preferably named Alo\"isius, less preferably Euforius and perhaps Eugenius, and of which the author is preferably not named Ambrosius, rather not Theofilus and perhaps not Antonius and which took effect preferably before $2/1/1210$ and preferably not after $1/1/1214$.
 These demands can now be translated to a query $Q_{ex}$ in the following way:
@@
 \vspace{-15pt}
-
 \begin{align}
-Q_{ex} = & (Q_{N,ex}, ((AR_{pos,ex}, I_{pos,ex}), (AR_{neg,ex}, I_{neg,ex}))), \\
-Q_{N,ex} = & \{(x, \mu_{ex}(x), \nu_{ex}(x)\}, \forall x \in \mathbb{S} \nonumber \\
-AR_{pos,ex}, AR_{neg,ex} = & DURING \nonumber \\
-I_{pos,ex} = & \left]- infinity, 1/1/1210\right] \nonumber \\
-I_{neg,ex} = & \left[1/1/1214, + infinity\right[ \nonumber
+Q_{ex} = &\ (Q_{N,ex}, ((AR_{pos,ex}, I_{pos,ex}), (AR_{neg,ex}, I_{neg,ex}))), \\
+Q_{N,ex} = &\ \{(x, \mu_{ex}(x), \nu_{ex}(x)\}, \forall x \in \mathbb{S} \nonumber \\
+AR_{pos,ex} = &\ \ AR_{neg,ex}\ =\ DURING \nonumber \\
+I_{pos,ex} = &\ \left]- infinity, 1/1/1210\right],\ \ I_{neg,ex}\ =\ \left[1/1/1214, + infinity\right[ \nonumber
 \end{align}
-\\
+\vspace{-10pt}\\
 and $\mu_{ex}$ and $\nu_{ex}$ are the membership functions of the fuzzy sets:
@@
 \end{align}
+\vspace{-25pt}
 
 \subsection{Elementary Query Condition Evaluation}
-Generally, a first step in determining which objects or concepts corresponding to relation tuples to present to a user as answer to his or her query towards this relation, is evaluating the elementary query conditions in the user query for every relation tuple.
+Generally, a first step in determining which objects or concepts corresponding to relation tuples to present to a user as answer to his or her query, is evaluating the query's elementary conditions for every relation tuple.
 In the presented work, for every tuple $T$ of VTR $R$, the following is done separately:
@@
 \begin{itemize}
-\item every non-temporal bipolar query condition in $Q_N$ is evaluated as described in~\cite{Matthe2011ijis}, resulting in a BSD for each.
-The interpretation is as described in~\cite{Matthe2011ijis}: the BSD's satisfaction, respectively dissatisfaction degrees express to which extent $T$ satisfies, respectively dissatisfies the user preferences expressed by the query condition.
-\item the query's crisp temporal bipolar query condition $Q_T$ is evaluated as follows: let $I_T$ be the tuple's IKVTI, then separately, both $I_T$ $AR_{pos}$ $I_{pos}$ and $I_T$ $AR_{neg}$ $I_{neg}$ are evaluated using the IKC framework, exactly as explained in~\cite{Pons2013ijufkbs},~\cite{Pons2012ijcis},~\cite{Pons2012ipmu}, resulting in a possibility degree $pos_{I_{pos}}(T)$ and a necessity degree $nec_{I_{pos}}(T)$, respectively a possibility degree $pos_{I_{neg}}(T)$ and a necessity degree $nec_{I_{neg}}(T)$.
-The interpretation here is that $pos_{I_{pos}}(T)$ and $nec_{I_{pos}}(T)$ express the plausibility, respectively necessity, given available knowledge, that the time interval during which the object or concept version or state represented by $T$ is actually valid and which is intended by $I_T$, is in relationship $AR_{pos}$ with $I_{pos}$ and thus complies with the user's positive temporal demand and is preferred, while $pos_{I_{neg}}(T)$ and $nec_{I_{neg}}(T)$ express the plausibility, respectively necessity, given a (partial) lack of knowledge, that the time interval during which the object or concept version or state represented by $T$ is actually valid and which is intended by $I_T$, is in relationship $AR_{neg}$ with $I_{neg}$ and thus complies with the user's negative temporal demand and is not preferred.
+\item every non-temporal BQC in $Q_N$ is evaluated as described in~\cite{Matthe2011ijis}, resulting in a BSD for each.
+The interpretation is as described in~\cite{Matthe2011ijis}: the BSD's satisfaction, respectively dissatisfaction degrees express to which extent $T$ satisfies, respectively dissatisfies the user preferences expressed by the BQC.
+\item the query's temporal BQC $Q_T$ is evaluated as follows: let $I_T$ be the tuple's IKVTI, then separately, $I_T$ $AR_{pos}$ $I_{pos}$, respectively $I_T$ $AR_{neg}$ $I_{neg}$ are evaluated using the IKC framework as described in~\cite{Pons2013ijufkbs},~\cite{Pons2012ijcis},~\cite{Pons2012ipmu}, resulting in a possibility degree $pos_{I_{pos}}(T)$ and a necessity degree $nec_{I_{pos}}(T)$, respectively a possibility degree $pos_{I_{neg}}(T)$ and a necessity degree $nec_{I_{neg}}(T)$.
+The interpretation is that $pos_{I_{pos}}(T)$ and $nec_{I_{pos}}(T)$ express the possibility, respectively necessity, that the time interval during which the version or state represented by $T$ is actually valid and which is intended by $I_T$, is in relationship $AR_{pos}$ with $I_{pos}$ and thus complies with the user's positive temporal demand and is preferred, while $pos_{I_{neg}}(T)$ and $nec_{I_{neg}}(T)$ express the possibility, respectively necessity, that the time interval during which the version or state represented by $T$ is actually valid and which is intended by $I_T$, is in relationship $AR_{neg}$ with $I_{neg}$ and thus complies with the user's negative temporal demand and is not preferred.
 \end{itemize}
@@
 Table \ref{tab:ex-evaluation} shows the results of the evaluation of the example query's elementary query conditions for the tuples of the example relation.
-
+\vspace{-10pt}
 \begin{table}[ht]
@@
 1 & $(1, 0)$ & $(1,1)$ & $(0,0)$ \\
+\hline
 2 & $(0, 0.7)$ & $(1,0)$ & $(0,0)$ \\
+\hline
 3 & $(0, 0)$ & $(11/25,0)$ & $(0,0)$ \\
+\hline
 4 & $(0.7, 0)$ & $(0,0)$ & $(0,0)$ \\
+\hline
 5 & $(0, 1)$ & $(0,0)$ & $(14/25,0)$ \\
+\hline
 6 & $(1, 0)$ & $(0,0)$ & $(1,0)$ \\
+\hline
 7 & $(0, 0)$ & $(0,0)$ & $(1,1)$ \\
@@
 \end{table}
-\vspace{-10pt}
+\vspace{-30pt}
 
 \subsection{Aggregation and Ranking}
-Generally, a second step in determining which objects or concepts corresponding to relation tuples to present to a user as answer to his or her query towards this relation, is aggregating, for every tuple in the relation, the tuple's evaluation results for the elementary query conditions, to determine a single `score' expressing how well the tuple complies with the general user request expressed by the combination of the elementary query conditions in the user query.
+Generally, a second step in determining which objects or concepts corresponding to relation tuples to present to a user as answer to his or her query, is aggregating, for every relation tuple, the tuple's evaluation results for the query's elementary conditions, in order to to determine how well the tuple complies with the entire user request expressed by the combination of the elementary query conditions.
 Usually, these evaluation results are quantifications of (dis)satisfaction.
@@
 \begin{itemize}
-	\item the BSD (dis)satisfaction degrees constitute quantifications of (dis)satisfaction: they quantify to which extent a tuple's attribute values (dis)satisfy a user's non-temporal preferences and thus assess an answer to the question: `To what extent does an object or concept represented in the relation (dis)satisfy the user's request and could thus be a (un)wanted result?'
-	\item the possibility and necessity degrees $pos_{I_{pos}}$, $nec_{I_{pos}}$, respectively $pos_{I_{neg}}$, $nec_{I_{neg}}$ constitute quantifications of plausibility and necessity, given a (partial) lack of knowledge: they quantify the plausibility that a tuple's intended crisp VTI (dis)satisfies the user's temporal preferences by quantifying confidence about exactly which crisp VTI is the tuple's intended VTI and thus assess an answer to the question: `How plausible is it that an object or concept represented in the relation that may or may not (dis)satisfy the user's request actually existed, exists or will exist during the time period requested by the user?'
+	\item the BSD's (dis)satisfaction degrees constitute quantifications of (dis)satisfaction: they quantify to which extent a tuple's attribute values (dis)satisfy a user's non-temporal preferences and thus assess an answer to the question: `To what extent does a version or state represented in the relation (dis)satisfy the user's request and could thus be a (un)wanted result?'
+	\item the possibility and necessity degrees $pos_{I_{pos}}$, $nec_{I_{pos}}$, respectively $pos_{I_{neg}}$, $nec_{I_{neg}}$ constitute quantifications of possibility and necessity: they quantify the possibility that a tuple's intended crisp VTI (dis)satisfies the user's temporal preferences by quantifying confidence about exactly which crisp VTI is the tuple's intended VTI and thus assess an answer to the question: `How plausible is it that a version or state represented in the relation (that may or may not (dis)satisfy the user's request) actually existed, exists or will exist during the time period indicated by the user?'
 \end{itemize}
@@
 A fundamental question arises now: should one consider combining quantifications of these different categories?
-On one hand, quantifications of these different categories have clearly different and even orthogonal semantics and it would not be clear what exactly the meaning would be of the result of such a combination or what the semantically most coherent ways would be to further process such combination results.
-Thus, it is important, for every query result tuple presented to the user, to certainly keep both different types of evaluations as separate (meta)data.
-On the other hand, without an unambiguous and straightforward ranking of the query results, the user cannot clearly discern the result tuples which comply well with his or her demands from those which don't, which defeats the purpose of querying.
-Reasonably, such ranking should be based on the elementary query condition evaluation results.
-As there cannot already exist a ranking between quantifications of categories with different semantics, a combination seems to be required.
-In most existing proposals requiring a combination of quantifications of satisfaction and plausibility, both quantifications are combined as to restrict one another.
-The result is usually seen as a quantification of plausibility.
+On one hand, such quantifications have clearly different semantics and it would not be clear what exactly the meaning would be of the result of such a combination or what the semantically most coherent ways would be to further process such combination results.
+Thus, it is important, for every query result tuple presented to the user, to certainly keep both different types of evaluation results as separate (meta)data.
+On the other hand, without an unambiguous and straightforward ranking of the query result tuples, the user cannot clearly discern the result tuples which comply well with his or her demands from those which don't, which defeats the purpose of querying.
+Reasonably, such a ranking should be based on the elementary query condition evaluation results.
+As there cannot exist a ranking between quantifications of categories with different semantics, a combination seems to be required.
+In most existing proposals requiring a combination of quantifications of satisfaction and possibility, both quantifications are combined as to restrict one another.
+The result is usually seen as a quantification of possibility.
 Below, this approach is translated to the specific situation encountered in the presented work.
 
-For every tuple $T$, the BSD's which are the evaluations of the non-temporal bipolar query conditions are combined to a single BSD $(s_N(T), d_N(T))$ as described in~\cite{Matthe2011ijis}.
-In this combination method, satisfaction degrees expressing how well $T$ fits the user's positive demands are combined with each other and separately, dissatisfaction degrees expressing how well $T$ fits the user's negative demands are combined with each other.
-This reasoning is extended to include $pos_{I_{pos}}(T)$, $nec_{I_{pos}}(T)$, $pos_{I_{neg}}(T)$ and $nec_{I_{neg}}(T)$: a couple $(pos_{pos}(T), nec_{pos}(T))$, respectively $(pos_{neg}(T), nec_{neg}(T))$ is calculated, expressing the plausibility and necessity, given a (partial) lack of knowledge, that the object or concept version or state corresponding to $T$ complies with the user's positive, respectively negative demands.
+For every tuple $T$, the BSD's which are the evaluation results of the non-temporal BQC are combined to a single BSD $(s_N(T), d_N(T))$ as described in~\cite{Matthe2011ijis}.
+In this combination method, satisfaction degrees are combined with each other and separately, dissatisfaction degrees are combined with each other.
+This reasoning is extended to include $pos_{I_{pos}}(T)$, $nec_{I_{pos}}(T)$, $pos_{I_{neg}}(T)$ and $nec_{I_{neg}}(T)$: a couple $(pos_{pos}(T), nec_{pos}(T))$, respectively $(pos_{neg}(T), nec_{neg}(T))$ is calculated, expressing the possibility and necessity that the version or state corresponding to $T$ complies with the user's positive, respectively negative demands.
 This calculation is done as follows:
@@
 
-Next, for every object or concept $o$ which is being represented in $R$, a couple $(pos_{pos}(o), nec_{pos}(o))$, respectively $(pos_{neg}(o), nec_{neg}(o))$ is calculated, expressing the plausibility and necessity, given a (partial) lack of knowledge, that $o$ complies with the user's positive, respectively negative demands.
-This calculation is done as follows:
+Next, a tie-break approach is used to rank the versions or states represented in $R$: ranking is done based on the value of $pos_{pos}(T)$, with $nec_{pos}(T)$ as tie-breaker for $pos_{pos}(T)$, with $pos_{neg}(T)$ as tie-breaker for $(pos_{pos}(T), nec_{pos}(T))$ and finally $nec_{neg}(T)$ as tiebreaker for $(pos_{pos}(T), nec_{pos}(5), pos_{neg}(5))$.
+The results of this aggregation and ranking approach for the example relation and query are shown in table \ref{tab:ex-ranking}.
+
+The introduced technique for aggregating elementary query condition evaluation results and determining a ranking does not require the combination of quantifications with different interpretations and still presents the query result tuples in a ordered manner, where this ordering is consistent with the expected extend to which each tuple is usefull to the user.
+However, the technique still requires the user to decide which objects or concepts constitute the best query answer, although this decision is heavily supported.
 
 \vspace{-10pt}
-\begin{align}
-(pos_{pos}(o), nec_{pos}(o)) = (\max_{T_{o,i}, 1 \leq i \leq k}(pos_{pos}(T_{o,i})), \max_{T_{o,i}, 1 \leq i \leq k}(nec_{pos}(T_{o,i}))) \nonumber \\
-(pos_{neg}(o), nec_{neg}(o)) = (\max_{T_{o,i}, 1 \leq i \leq k}(pos_{neg}(T)_{o,i}), \max_{T_{o,i}, 1 \leq i \leq k}(nec_{neg}(T)_{o,i})) \nonumber
-\end{align}
-
-Here, $\{T_{o,i}, 1 \leq i \leq k, i \in \mathbb{N}\}$ is the set of tuples $T_{o,i}, 1 \leq i \leq k, i \in \mathbb{N}$ which represent the $k$ versions or states of $o$.
-Next, a tie-break approach is used to rank the objects or concepts $o$ represented in $R$: ranking is done based on the value of $pos_{pos}(o)$, with $nec_{pos}(o)$ as tie-breaker for $pos_{pos}(o)$, with $pos_{neg}(o)$ as tie-breaker for $( pos_{pos}(o), pos_{neg}(o))$ and finally $nec_{neg}(o)$ as tiebreaker for $(pos_{pos}(o), pos_{neg}(o), pos_{neg}(o))$.
-The results of this aggregation and ranking approach for the example relation and query are shown in table \ref{tab:ex-ranking}.
-
 \begin{table}[ht]
@@
 1 & 1 & 1 & 0 & 0 & $(1, 0)$ & $(1,1)$ & $(0,0)$ \\
+\hline
 2 & 0 & 0 & 0 & 0 & $(0, 0.7)$ & $(1,0)$ & $(0,0)$ \\
+\hline
 3 & 0 & 0 & 0 & 0 & $(0, 0)$ & $(11/25,0)$ & $(0,0)$ \\
+\hline
 4 & 0 & 0 & 0 & 0 & $(0.7, 0)$ & $(0,0)$ & $(0,0)$ \\
+\hline
+6 & 0 & 0 & 0 & 0 & $(1, 0)$ & $(0,0)$ & $(1,0)$ \\
+\hline
+7 & 0 & 0 & 0 & 0 & $(0, 0)$ & $(0,0)$ & $(1,1)$ \\
+\hline
 5 & 0 & 0 & 14/25 & 0 & $(0, 1)$ & $(0,0)$ & $(14/25,0)$ \\
-6 & 0 & 0 & 0 & 0 & $(1, 0)$ & $(0,0)$ & $(1,0)$ \\
-7 & 0 & 0 & 0 & 0 & $(0, 0)$ & $(0,0)$ & $(1,1)$ \\
 \hline
@@
 \end{table}
+\vspace{-15pt}
 
@@
 In this paper, a novel technique is presented, to query a valid-time relation containing valid-time data subject to uncertainty in a bipolar way, which allows the user to specify a single valid-time bipolar query condition.
-A major issue concerning the necessity to combine quantifications of (dis)satisfaction with quantifications of plausibility resulting from this technique and a solution for this is presented and shortly discussed.
+A major issue concerning the necessity to combine quantifications of (dis)satisfaction with quantifications of possibility resulting from this technique and a solution for this is presented and shortly discussed.
 In the near future, the interactions between these types of quantifications and between uncertainty in valid-time data and bipolar querying will be further studied.

### 2013-05-30 fqas2013/paper.tex
Some changes based on reviews.

--- 2013-02-15 fqas2013/paper.tex
@@
 
-\title{Bipolar Valid-time Querying}
+\title{\mytitlestring}
 
 %Short form of title for running heads
-\titlerunning{Bipolar Valid-time Querying}
+\titlerunning{\mytitlestring}
 
@@
 \begin{abstract}
-Here goes the abstract.
-\keywords{keyword1, keyword2}
+Databases model parts of reality by containing data representing properties of real-world objects or concepts.
+Often, some of these properties are time-related.
+Thus, databases often contain data representing time-related information.
+However, as they may be produced by humans, such data or information may contain imperfections like uncertainties.
+An important purpose of databases is to allow their data to be queried, to allow access to the information these data represent.
+Users may do this using queries, in which they describe their preferences concerning the data they are (not) interested in.
+Because users may have both positive and negative such preferences, they may want to query databases in a bipolar way.
+Such preferences may also have a temporal nature, but, traditionally, temporal query conditions are handled specifically.
+In this paper, a novel technique is presented to query a valid-time relation containing uncertain valid-time data in a bipolar way, which allows the query to have a single bipolar temporal query condition.
+
+\keywords{Bipolar Querying, Valid-time Relation, Valid Time, Temporal Databases, Uncertainty, Possibility Theory, Ill-known Intervals}
 \end{abstract}
@@
 \section{Introduction}
+Generally, database systems model (parts of) reality.
+For this, their databases contain data representing properties of real-world objects or concepts.
+Some essential properties of real-world objects or concepts are time-related.
+Thus, databases often contain data representing temporal values~\cite{Bohlen1998lncs}, which are basically indications of time and describe such properties.
+These temporal values are usually either time intervals~\cite{Bohlen1998lncs} or instants~\cite{Bohlen1998lncs}, which may informally be seen as infinitesimally short `periods' or `points' in time.
+Based on their interpretation and purpose, temporal values can be classified into several categories, but the presented work will only consider valid-time indications, which indicate when corresponding data is a valid or true representation of the reality modelled by its database~\cite{Bohlen1998lncs},~\cite{Pons2012ijcis},~\cite{Billiet2012ipmu},~\cite{Pons2012ipmu}.
+
+A lot of database data are produced by humans, but human-made data are prone to imperfections, as some of these data may be vague or imprecise~\cite{Medina1994is}, contradictory, incomplete or uncertain~\cite{Billiet2012ipmu},~\cite{Pons2012ipmu},~\cite{Bosc2010ijufkbs}.
+Of course, data representing temporal values may contain such imperfections too~\cite{Pons2012ijcis},~\cite{Billiet2012ipmu},~\cite{Pons2012ipmu},~\cite{Dyreson1998acm}.
+The work presented in this paper will consider databases containing data representing valid-time intervals subject to uncertainty and will assume all non-temporal data in these databases to contain no imperfections.
+
+One of the most important purposes of a database is to allow its data to be queried, to allow the information or knowledge represented by this data to be retrieved.
+A user may query a database in a `regular' way: the user describes the data which he or she finds desired or satisfactory and thus wants to retrieve, by perfectly describing the allowed values of these data.
+A user may also query a database in a `fuzzy' way: the user describes the data which he or she finds desired or satisfactory by imperfectly describing the allowed values of these data~\cite{Zadrozny2008hrfipd}.
+These imperfect descriptions may contain vagueness or imprecision, often through the use of linguistic terms~\cite{Devos1998jql},~\cite{Kacprzyk2001is}.
+A user may also query a database in a `bipolar' way.
+Generally, two main approaches to this exist.
+One is for the user to describe the data which he or she finds acceptable and to describe the data among this acceptable data, which he or she finds really desired, both by describing the allowed values of these data~\cite{Dubois2002lnai}.
+The other is for the user to independently describe both the data which he or she finds desired or satisfactory and the data which he or she finds undesired or unsatisfactory, both by describing the allowed values of these data~\cite{DeTre2010ieeetfs},~\cite{Matthe2011ijis}.
+The descriptions used in bipolar querying may contain imperfections.
+The presented work will only consider the latter approach to bipolar querying and will allow a simple form of imprecision in non-temporal elementary query conditions.
+
+Compared to non-temporal user preferences, temporal user preferences usually have an uncommon nature and interpretation and thus expressing them relies on uncommon mechanics in querying: users usually prefer using specific temporal operators to express temporal preferences.
+Hence, several proposals have considered specific sets of temporal operators, often based on the possible temporal relationships between two time indications~\cite{Pons2012ijcis},~\cite{Galindo2001},~\cite{Schockaert2008ieeetfs}.
+Such temporal relationships define semantically meaningful relationships with a temporal nature, between two time indications.
+In~\cite{Allen1983cacm}, a collection of temporal relationships between two time intervals (and as a special case instants) is introduced and this collection is considered groundbreaking.
+Of course, to query temporal data in a fuzzy way, fuzzy variants of such temporal operators are necessary and several proposals have thus introduced such variants~\cite{Pons2012ipmu},~\cite{Galindo2001}, often based on fuzzy variants of such temporal relationships~\cite{Schockaert2008ieeetfs},~\cite{Pons2013ijufkbs}.
+
+Techniques for the regular or fuzzy querying of valid-time databases containing valid-time data subject to uncertainty are considered by several existing proposals~\cite{Pons2012ijcis},~\cite{Pons2012ipmu},~\cite{Pons2013ijufkbs}.
+However, to the knowledge of the authors, only one proposal has considered the bipolar querying of valid-time databases (in~\cite{Billiet2011fqas}, a technique is proposed to query a valid-time database containing temporal data subject to imprecision in a bipolar way) and none have considered the bipolar querying of valid-time databases containing temporal data subject to uncertainty.
+Thus, this paper presents a novel technique to query a valid-time relation containing valid-time data subject to uncertainty in a bipolar way, which allows the user to specify a single bipolar temporal query condition.
+This paper is structured as follows: in section \ref{sec:preliminaries}, some preliminary concepts and techniques are described, in section \ref{sec:proposal}, the novel technique which is the main contribution of the work presented in this paper, is explained and in section \ref{sec:conclusions}, the conclusions of this paper and some directions for future research are given.
+
+\section{Preliminaries\label{sec:preliminaries}}
+
+\subsection{General Preliminaries, Notations and Nomenclature}
+Databases may contain data representing temporal values.
+Based on their purpose and interpretation, such time indications can be classified into different categories~\cite{Bohlen1998lncs},~\cite{Billiet2012ipmu}.
+The work presented in this paper only considers temporal values of the category \emph{valid time}.
+Their purpose or interpretation is for every valid-time indication to correspond to a collection of data and to indicate a period of time during which this data is a valid or true representation of reality.
+
+The work presented in this paper concerns time indications subject to uncertainty.
+This uncertainty is always assumed to be caused by a (partial) lack of knowledge: the exact, intended time indication is not known, eventhough there is only one time indication intended and as such no variability.
+Confidence about exactly which time indication is the intended one in the context of such uncertainty is modelled using possibility theory~\cite{Pons2013ijufkbs},~\cite{Dubois1988}.
+In the presented work, `possibility' and `necessity' are always interpreted as measures of plausibility, respectively necessity, given all available knowledge.
+Time intervals not subject to any imperfection are called \emph{crisp time intervals} (CTI) in this paper.
+
+\subsection{Valid-time Relations}
+%The presented work uses the relational database model~\cite{Pons2012ijcis},~\cite{Galindo2001}.
+Here, a \emph{relation} models a collection of similar real-world concepts or objects.
+To achieve this, such a relation has a set of $n$ \emph{attributes}, of which each describes a different property shared by all of these concepts or objects, and a set of $n$-tuples, of which each represents one of these real-world objects or concepts and contains $n$ values, each representing the capacity of the object or concept corresponding to the tuple, of the property described by the attribute corresponding to the value~\cite{Billiet2012ipmu},~\cite{Pons2012ipmu}.
+
+A \emph{valid-time relation} (VTR) always has \emph{valid-time attributes}.
+These are attributes describing a single valid time~\cite{Bohlen1998lncs} for the objects or concepts represented by the VTR's tuples.
+A VTR may contain different tuples corresponding to the same real-world concept or object.
+The non-valid-time attribute values of such a tuple represent the capacities of the properties described by their corresponding attributes which were, are or will be true or valid for the object or concept corresponding to the tuple during the period in time indicated by the valid-time indication represented by the tuple's valid-time attribute values.
+Thus, such a tuple represents the `version' of the object or concept corresponding to this tuple which was, is or will be real or valid or the `state' this object or concept was, is or will be in, during the period in time indicated by the valid-time indication represented by the tuple's valid-time attribute values.
+In the presented work, such valid-time indications will always be time intervals~\cite{Bohlen1998lncs} and will always be referred to as \emph{valid-time intervals} (VTI).
+
+\subsection{Uncertainty in Valid-time Intervals}
+The presented work allows VTI to be subject to uncertainty, by allowing them to be \emph{ill-known valid-time intervals} (IKVTI).
+The concept of IKVTI is based on the concepts of \emph{possibilistic variables} (PV) and \emph{ill-known intervals} (IKI)~\cite{Billiet2012ipmu},~\cite{Pons2013ijufkbs},~\cite{Dubois1988cma}.
+
+\begin{definition}
+\label{def:poss-variable}
+A \emph{possibilistic variable} (PV) $X$ on a universe $U$ is a variable taking exactly one value in $U$, but for which this value is (partially) unknown.
+The possibility distribution $\pi_X$ on $U$ models the available knowledge about the value that $X$ takes: for each $u \in U$, $\pi_X(u)$ represents the possibility that $X$ takes the value $u$.
+\end{definition}
+
+Now consider a set $U$ containing single values (and not collections of values).
+When a PV $X_{v}$ is defined on such a set $U$, the unique value $X_{v}$ takes, which is (partially) unknown, will be a single value in $U$ and is called an \emph{ill-known value} (IKV) in $U$~\cite{Billiet2012ipmu},~\cite{Pons2013ijufkbs},~\cite{Dubois1988cma}.
+In this paper, IKV will be denoted using lower-case letters.
+The work presented in this paper uses a specific kind of IKI, defined as follows, although other definitions exist~\cite{Billiet2012ipmu},~\cite{Pons2012ipmu}.
+
+\begin{definition}
+Consider an ordered set $U$. An \emph{ill-known interval} (IKI) in $U$ is an interval in $U$ of which both boundary values are IKV in $U$.
+\end{definition}
+
+Specifically concerning valid time, an IKI in a time domain represented by the domains of a VTR's valid-time attributes is called an \emph{ill-known valid-time interval} (IKVTI).
+The work presented in this paper requires the possibility distributions defining an IKVTI's IKV to be convex~\cite{Pons2013ijufkbs}.
+In this paper, an IKVTI with boundary IKV $s$ and $e$ will be noted $\left[s, e\right]$.
+
+\subsection{Evaluation of Temporal Relationships}
+To express temporal elementary query conditions, operators based on temporal relationships are necessary.
+In the presented work, only Allen relationships~\cite{Allen1983cacm} between a CTI and a IKVTI are considered.
+To evaluate such relationships, the \emph{ill-known constraints} (IKC) framework presented in~\cite{Pons2013ijufkbs} is used.
+It relies on the concept of IKC.
+
+\begin{definition}
+Given an ordered set $U$, an \emph{ill-known constraint} (IKC) $C = (R,v)$ on $U$ is specified by means of a binary relation $R \subseteq U^{2}$ and a fixed IKV $v$ in $U$.
+Any set $A \subseteq U$ now satisfies IKC $C = (R,v)$ if and only if:
+\begin{align}
+\forall a \in A : (a,v) \in R \nonumber
+\end{align}
+\end{definition}
+
+The satisfaction of an IKC $C$ by a set $A$ will be noted $C(A)$ in this paper.
+Consider an ordered set $U$, an IKC $C = (R,v)$ on $U$ and a set $A \subseteq U$.
+Due to the uncertainty inherent to $v$, it is uncertain whether $A$ satisfies $C$ or not.
+The degree of possibility $\Pos(C(A))$ that $A$ satisfies $C$ and the degree of necessity $\Nec(C(A))$ that $A$ satisfies $C$, can be calculated as follows~\cite{Pons2013ijufkbs}:
+
+\vspace{-10pt}
+\begin{align}
+\Pos(C(A)) & = \min_{a \in A}\left(\sup_{(a,w) \in R}\pi_{X_{v}}(w)\right) \label{ill-known-pos}\\
+\Nec(C(A)) & = \min_{a \in A}\left(\inf_{(a,w) \notin R} 1-\pi_{X_{v}}(w)\right) \label{ill-known-nec}
+\end{align}
+
+Given an ordered set $U$, degrees of possibility and necessity that a set $A \subseteq U$ satisfies a boolean combination of IKC on $U$ can be found by using the possibilistic extensions of boolean operators `and' ($\wedge$), `or' ($\vee$) and `not' ($\neg$), as described in~\cite{Billiet2012ipmu},~\cite{Pons2012ipmu},~\cite{Pons2013ijufkbs}.
+
+The IKC framework now allows evaluating a given Allen relationship $AR$ between a given CTI $I$ and a given IKVTI $J = \left[s, e\right]$ by allowing the calculation of the degrees of possibility and necessity that $I$ $AR$ $J$ holds.
+For this, the combination of $AR$ and $J$ is translated to a specific boolean combination of specific IKC.
+These translations are shown in table \ref{tab:allen-relations}.
+Every row of this table corresponds to a given Allen relationship between $I$ and $J$, indicated by the row's value in the `Allen Relationship' column.
+The collections of specific IKC for given Allen relationships are shown in the `Constraints' column (every $C_i, i \in \{1, 2, 3, 4\}$ denotes an IKC) and the specific combination of these IKC used for evaluation of the Allen relationships are shown in the `Combination' column.
+Finally, the degrees of possibility and necessity that $I$ $AR$ $J$ holds are then the degrees of possibility, respectively necessity that $I$ satisfies the specific aggregation of specific IKC found as translation of the combination of $AR$ and $J$.
+Using the formulas shown above, the requested possibility and necessity degrees can be calculated from these.
+
+\vspace{-10pt}
+\begin{table}[ht]
+\caption{The translations of Allen relationships to the IKC framework.}
+\centering
+\begin{tabular}{|c|c|c|}
+\hline
+\textbf{Allen Relationship} & \textbf{Constraints} & \textbf{Combination} \\
+\hline
+\hline
+I before J & $C_1\stackrel{\triangle}{=} \left(<,s\right)$ & $C_1(I)$ \\
+\hline
+\multirow{2}{*}
+{I equal J} & $C_1\stackrel{\triangle}{=} \left(\geq,s\right)$, $C_2\stackrel{\triangle}{=} \left(\neq,s\right)$ & $C_1(I)\wedge$ $\neg C_2(I)\wedge$\\
+& $C_3\stackrel{\triangle}{=} \left(\leq,e\right)$, $C_4\stackrel{\triangle}{=} \left(\neq,e\right)$ & $C_3(I)\wedge$ $\neg C_4(I)$\\
+\hline
+I meets J & $C_1\stackrel{\triangle}{=} \left(\leq,s\right)$ $C_2\stackrel{\triangle}{=} \left(\neq,s\right)$ & $C_1(I)\wedge$ $\neg C_2(I)$\\
+\hline
+I overlaps J & $C_1\stackrel{\triangle}{=} \left(<,e\right)$, $C_2\stackrel{\triangle}{=} \left(\leq,s\right)$, $C_3\stackrel{\triangle}{=} \left(\geq,s\right)$ & $C_1(I)\wedge$ $\neg C_2(I)\wedge$ $\neg C_3(I)$\\
+\hline
+\multirow{2}{*}
+{I during J} & $C_1\stackrel{\triangle}{=} \left(>,s\right)$, $C_2\stackrel{\triangle}{=} \left(\leq,e\right)$ & $\big(C_1(I)\wedge$ $ C_2(I)\big)\vee$\\
+ & $C_3\stackrel{\triangle}{=} \left(\geq,s\right)$, $C_4\stackrel{\triangle}{=} \left(<,e\right)$ & $\big(C_3(I)\wedge$ $C_4(I)\big)$\\
+\hline
+I starts J & $C_1\stackrel{\triangle}{=} \left(\geq,s\right)$, $C_2\stackrel{\triangle}{=} \left(\neq,s\right)$ & $C_1(I)\wedge$, $\neg C_2(I)$\\
+\hline
+I finishes J & $C_1\stackrel{\triangle}{=} \left(\leq,e\right)$, $C_2\stackrel{\triangle}{=} \left(\neq,e\right)$ & $C_1(I)\wedge$ $\neg C_2(I)$\\
+\hline
+\end{tabular}
+\label{tab:allen-relations}
+\end{table}
+\vspace{-10pt}
+
+\subsection{Bipolar Querying}
+As mentioned before, humans may express their query preferences using both positive and negative query conditions~\cite{DeTre2010ieeetfs},~\cite{Matthe2011ijis}.
+If the semantics of these conditions are non-symmetric, meaning that the positive preferences can not be derived from the negative or vice versa, the bipolarity in this query is called \emph{heterogenous}~\cite{Matthe2011ijis}.
+The presented work will concern only such heterogenous query bipolarity.
+
+A query usually takes the form of a boolean combination of elementary query conditions.
+Every elementary query condition then expresses the user's demands concerning a single attribute.
+Bipolarity in a query can either be specified between or inside the elementary query conditions.
+In~\cite{Matthe2011ijis}, it is shown that combining both approaches makes no sense and that the approach where bipolarity is specified inside elementary query conditions, using intuitionistic fuzzy sets~\cite{Atanassov1986fss}, is a more intuitive one.
+In the presented work, only the latter approach is used.
+In this approach, elementary query conditions express both what is accepted and what is not accepted by the query, at once, and are called \emph{bipolar query conditions} (BQC)~\cite{Matthe2011ijis}.
+
+Consider a relation attribute $A$.
+Let $dom_{A}$ be the domain of $A$'s data type, let $\mu_{c_{A}}$ and $\nu_{c_{A}}$ be membership functions from $dom_{A}$ to the unit interval $\left[0, 1\right]$, where $\mu_{c_{A}}(x)$ represents to what extent $x \in dom_{A}$ is satisfactory and $\nu_{c_{A}}(x)$ to what extent $x$ is unsatisfactory to a user, then a BQC $c_{A}$ expressing this user's preferences concerning $A$ can be modelled by an Intuitionistic Fuzzy Set (IFS)~\cite{Atanassov1986fss} as~\cite{DeTre2010ieeetfs},~\cite{Matthe2011ijis}:
+\vspace{-5pt}
+\begin{equation}
+c_{A} = \{(v, \mu_{c_{A}}(v), \nu_{c_{A}}(v)) : v \in dom_{A}\}
+\end{equation}
+
+Note that to allow overspecification of the user's preferences, the IFS's consistency condition can be relaxed, which means that there may exist values $v \in dom_{A}$ for which $\mu_{c_{A}}(v) + \nu_{c_{A}}(v) > 1$~\cite{DeTre2010ieeetfs},~\cite{Matthe2011ijis}.
+
+If the user explicitely defines $\mu_{c_{A}}$, but doesn't define $\nu_{c_{A}}$, then $\nu_{c_{A}}$ will be assumed to be the inverse of $\mu_{c_{A}}$~\cite{Matthe2011ijis}.
+If the user explicitely defines $\nu_{c_{A}}$, but doesn't define $\mu_{c_{A}}$, then $\mu_{c_{A}}$ will be assumed to be the inverse of $\nu_{c_{A}}$~\cite{Matthe2011ijis}.
+Thus, in the absence of clear heterogenousness of the bipolarity in a query condition, the bipolarity will be assumed homogenous~\cite{Matthe2011ijis}.
+
+The evaluation of a BQC results in a so-called {\em bipolar satisfaction degree} (BSD)~\cite{Matthe2011ijis}, which is a pair
+\vspace{-10pt}
+\begin{equation}
+(s,d),\; s,d \in [0,1] \nonumber
+\end{equation}
+where $s$ is called the \emph{satisfaction degree} and $d$ is called the \emph{dissatisfaction degree}~\cite{Matthe2011ijis}.
+Here, $s$ and $d$ are independent from each other and express to which extent the BSD respectively represents `satisfied' and `dissatisfied'~\cite{Matthe2011ijis}.
+Extreme values for $s$ and $d$ are 0 (`not at all') and 1 (`fully').
+For example: the BSD $(1,0)$ represents `fully satisfied, not dissatisfied at all'~\cite{Matthe2011ijis}.
+
+As explained in~\cite{Matthe2011ijis}, there is no consistency condition for BSD's and for a BSD $(s, d)$, $s+d>1$ is allowed.
+The motivation is that BSD's try to reflect heterogenous bipolarity in human reasoning, which can sometimes be inconsistent.
+
+In general, the evaluation of a BQC $c_{A}$ on relation attribute $A$ for a tuple $r$ will result in a BSD $(s_{c_A}^r, d_{c_A}^r)$, which is calculated as follows.
+Let $r[A]$ denote the value of tuple $r$ for attribute $A$, then~\cite{Matthe2011ijis}:
+\begin{equation}\label{eq:bsdeval}
+(s_{c_A}^r, d_{c_A}^r)=(\mu_{c_A}(r[A]), \nu_{c_A}(r[A]))
+\end{equation}
+
+Remark that the traditional approach to fuzzy querying using regular fuzzy sets can be obtained from this as a special case, where the bipolarity involved is homogenous.
+In that case, a user only specifies positive query preferences~\cite{Matthe2011ijis}.
+
+\section{A Novel Querying Approach\label{sec:proposal}}
+
+\subsection{Valid-time Relations Subject to Uncertainty}
+The presented proposal will concern VTR where the VTI are IKVTI.
+Generally, such a VTR $R$ can be seen as constructed in the following way.
+Let $R$ have $n$ non-temporal attributes $A_{i}, 1 \leq i \leq n, i \in \mathbb{N}$ and two valid-time attributes $VST$ and $VET$.
+Every tuple $T$ of $R$ represents an object or concept version or state which is valid during the time period indicated by the tuple's IKVTI $I_T$.
+This IKVTI $I_T$ is now defined by two IKV, which respectively describe the starting and ending instants of $I_T$ and are represented by the tuple's VST, respectively VET values.
+The interpretation is that the version or state corresponding to a tuple was, is or will be valid during a period of time, but exactly which period this is intended to be, is unknown.
+Confidence about exactly which period is intended, is modelled by the tuple's IKVTI~\cite{Pons2012ijcis},~\cite{Pons2012ipmu}.
+
+\vspace{-10pt}
+\begin{table}[ht]
+\caption{The example relation used in this paper.}
+\centering
+\begin{tabular}{|c|c|c|c|}
+\hline
+\textbf{ID} & \textbf{Author} & \textbf{VST} & \textbf{VET} \\
+\hline
+\hline
+1 & Alo\"{i}sius & $\left[4/4/1208, 6/4/1208, 16/4/1208\right]$ & $\left[10/12/1208, 1/1/1209, 26/1/1209\right]$ \\
+\hline
+2 & Theofilus & $\left[2/4/1209, 12/4/1209, 22/4/1209\right]$ & $\left[21/12/1209, 1/1/1210, 21/1/1210\right]$ \\
+\hline
+3 & Gerardus & $\left[14/1/1209, 15/1/1209, 16/1/1209\right]$ & $\left[21/12/1209, 15/1/1210, 25/1/1210\right]$ \\
+\hline
+4 & Euforius & $\left[21/12/1210, 1/1/1211, 11/1/1211\right]$ & $\left[21/12/1211, 1/1/1212, 11/1/1212\right]$ \\
+\hline
+5 & Ambrosius & $\left[11/12/1213, 21/12/1213, 15/1/1214\right]$ & $\left[9/10/1216, 10/10/1216, 15/10/1216\right]$ \\
+\hline
+6 & Alo\"{i}sius & $\left[21/12/1213, 1/1/1214, 11/1/1214\right]$ & $\left[9/6/1217, 9/6/1217, 12/6/1217\right]$ \\
+\hline
+7 & Gerardus & $\left[29/12/1214, 1/1/1215, 8/1/1215\right]$ & $\left[9/6/1217, 10/6/1217, 12/6/1217\right]$ \\
+\hline
+\end{tabular}
+\label{tab:ex-relation}
+\end{table}
+\vspace{-10pt}
+
+In this paper, the relation shown in table~\ref{tab:ex-relation} will be used as example relation.
+This relation models the being in effect of medieval legal acts.
+Since the properties of a legal act cannot change once it has taken effect, every legal act has only one version or state.
+This was deliberately done to simplify the example.
+Thus, every tuple of the relation corresponds to a legal act.
+A tuple's value for attribute `ID' is a number uniquely identifying the legal act corresponding to the tuple.
+A tuple's value for attribute `Author' is a character string representation of the name of the author of the legal act corresponding to the tuple.
+A tuple's IKVTI represents the time period during which the act corresponding to the tuple was in effect.
+For this, every value for $VST$, respectively $VET$ represents an IKV describing the day on which the legal act respectively took effect and stopped taking effect.
+For this, every value for $VST$ or $VET$ is a triple $\left[d_1, d_2, d_3\right]$, where $d_1, d_2, d_3$ are elements of the ordered set of days in history.
+Such a triple $\left[d_1, d_2, d_3\right]$ now defines a triangular (and thus convex) possibility distribution $\pi$ which defines the mentioned IKV and which is defined by (differences in dates in this function prescription are expressed in amounts of days):
+
+\vspace{-10pt}
+\begin{align}
+\pi(x) =
+\begin{cases}
+\frac{x - d_1}{d_2 - d_1}, & \text{if } d_1 \leq x < d_2 \\
+\frac{d_3 - x}{d_3 - d_2}, & \text{if } d_2 \leq x \leq d_3 \\
+0, & \text{else}
+\end{cases}
+\end{align}
+\vspace{-15pt}
+
+\subsection{Querying Using Bipolar Valid-time Conditions \label{sec:query}}
+The presented work introduces a novel querying technique.
+The most interesting aspect of this technique is that it allows the user to specify a bipolar valid-time demand.
+According to this technique, a user query $Q$ consists of two separate parts $Q_n$ and $Q_t$: $Q = (Q_n, Q_t)$.
+Here, $Q_n$ is a boolean combination of BQC on non-valid-time attributes, expressing the user's non-temporal demands.
+$Q_t$ expresses the user's valid-time demands and is a single crisp temporal BQC $((AR_{+}, I_{+}), (AR_{-}, I_{-}))$, where both $AR_{+}$ and $AR_{-}$ are Allen relationships and both $I_{+}$ and $I_{-}$ are CTI.
+The interpretation is that the user requires an object or concept that has a version or state that complies with his or her non-temporal demands and was, is or will be valid during a time interval which is in Allen relationship $AR_{+}$ with $I_{+}$ and wasn't, isn't or won't be valid during a time interval which is in Allen relationship $AR_{-}$ with $I_{-}$.
+
+Consider the example relation shown in table~\ref{tab:ex-relation}.
+Now assume a user queries the relation to find all legal acts of which the author is preferably named Alo\"isius, less preferably Euforius and perhaps Eugenius, and of which the author is preferably not named Ambrosius, rather not Theofilus and perhaps not Antonius and which took effect preferably before $2/1/1210$ and preferably not after $1/1/1214$.
+These demands can now be translated to a query $Q_{ex}$ in the following way:
+
+\vspace{-15pt}
+\begin{align}
+Q_{ex} = &\ (Q_{n,ex}, Q_{t,ex})\ =\ (Q_{n,ex}, ((AR_{+,ex}, I_{+,ex}), (AR_{-,ex}, I_{-,ex}))), \nonumber\\
+Q_{n,ex} = &\ \{(x, \mu_{ex}(x), \nu_{ex}(x)), \forall x \in \mathbb{S} \}\nonumber \\
+AR_{+,ex} = &\ \ AR_{-,ex}\ =\ DURING \nonumber \\
+I_{+,ex} = &\ \left]- infinity, 1/1/1210\right],\ \ I_{-,ex}\ =\ \left[1/1/1214, + infinity\right[ \nonumber
+\end{align}
+\vspace{-10pt}\\
+and $\mu_{ex}$ and $\nu_{ex}$ are the membership functions of the fuzzy sets:
+\begin{align}
+\{(Aloisius, 1), (Euforius, 0.7), (Eugenius, 0.1)\} &\text{, respectively} \nonumber \\
+\{(Ambrosius, 1), (Theofilus, 0.7), (Antonius, 0.1)\} & \nonumber
+\end{align}
+and $\mathbb{S}$ is the set of all author names in the example relation's `Author' attribute domain.
+%\vspace{-25pt}
+
+\subsection{Elementary Query Condition Evaluation}
+Generally, a first step in determining which objects or concepts corresponding to tuples to present to a user as answer to his or her query, is evaluating the query's elementary conditions for every tuple.
+In the presented work, given a user query $Q$ constructed as proposed in section \ref{sec:query} and using the same notations, the following is done separately for every tuple $T$ of VTR $R$:
+
+\begin{itemize}
+\item every non-temporal BQC in $Q_n$ is evaluated as described in~\cite{Matthe2011ijis}, resulting in a BSD for each.
+The interpretation is as described in~\cite{Matthe2011ijis}: the BSD's satisfaction, respectively dissatisfaction degrees express to which extent $T$ satisfies, respectively dissatisfies the user preferences expressed by the BQC.
+\item the query's temporal BQC $Q_t$ is evaluated as follows.
+Let $I_T$ be the tuple's IKVTI.
+Then, independently, the statements `$I_T$ $AR_{+}$ $I_{+}$', respectively `$I_T$ $AR_{-}$ $I_{-}$' are evaluated using the IKC framework as described in~\cite{Pons2012ijcis},~\cite{Pons2012ipmu},~\cite{Pons2013ijufkbs}.
+These evaluations result in a possibility degree $Pos_{+}(I_{T})$ and a necessity degree $Nec_{+}(I_{T})$, respectively a possibility degree $Pos_{-}(I_{T})$ and a necessity degree $Nec_{-}(I_{T})$.
+The interpretation is that $Pos_{+}(I_{T})$ and $Nec_{+}(I_{T})$ express the possibility, respectively necessity, that the time interval during which the version or state represented by $T$ is valid and which is intended by $I_T$, is in relationship $AR_{+}$ with $I_{+}$ and thus complies with the user's positive temporal demand.
+Furthermore, the interpretation is that $Pos_{-}(I_{T})$ and $Nec_{-}(I_{T})$ express the possibility, respectively necessity, that the time interval during which the version or state represented by $T$ is valid and which is intended by $I_T$, is in relationship $AR_{-}$ with $I_{-}$ and thus complies with the user's negative temporal demand.
+\end{itemize}
+
+Table \ref{tab:ex-evaluation} shows the results of the evaluation of the example query's elementary query conditions for the tuples of the example relation.
+\vspace{-10pt}
+\begin{table}[ht]
+\caption{Elementary query condition evaluation results for the example relation.}
+\centering
+\begin{tabular}{|c|c|c|c|}
+\hline
+\textbf{ID} & \textbf{$BSD_{Q_{N,ex}}(T)$} & \textbf{$(Pos_{+}(I_{T}), Nec_{+}(I_{T})$} & \textbf{$(Pos_{-}(I_T), Nec_{-}(I_T)$} \\
+\hline
+\hline
+1 & $(1, 0)$ & $(1,1)$ & $(0,0)$ \\
+\hline
+2 & $(0, 0.7)$ & $(1,0)$ & $(0,0)$ \\
+\hline
+3 & $(0, 0)$ & $(11/25,0)$ & $(0,0)$ \\
+\hline
+4 & $(0.7, 0)$ & $(0,0)$ & $(0,0)$ \\
+\hline
+5 & $(0, 1)$ & $(0,0)$ & $(14/25,0)$ \\
+\hline
+6 & $(1, 0)$ & $(0,0)$ & $(1,0)$ \\
+\hline
+7 & $(0, 0)$ & $(0,0)$ & $(1,1)$ \\
+\hline
+\end{tabular}
+\label{tab:ex-evaluation}
+\end{table}
+\vspace{-30pt}
+
+\subsection{Aggregation and Ranking}
+Generally, a second step in determining which objects or concepts corresponding to tuples to present to a user as answer to his or her query, is aggregating, for every tuple, the tuple's evaluation results for the query's elementary conditions, in order to determine how well the tuple complies with the entire user request expressed by the combination of the elementary query conditions.
+Usually, these evaluation results are quantifications of (dis)satisfaction.
+However, in the presented proposal, two different types of evaluation results can be discerned:
+
+\begin{itemize}
+	\item the BSD's (dis)satisfaction degrees constitute quantifications of (dis)satisfaction: they quantify to which extent a tuple's attribute values (dis)satisfy a user's non-temporal preferences and thus assess an answer to the question: `To what extent does a version or state represented in the relation (dis)satisfy the user's request and could thus be a (un)wanted result?'
+	\item the possibility and necessity degrees $Pos_{+}$, $Nec_{+}$, respectively $Pos_{-}$, $Nec_{-}$ constitute quantifications of possibility and necessity: they quantify the possibility that a tuple's intended crisp VTI does (not) comply with the user's temporal demands by quantifying confidence about exactly which crisp VTI is the tuple's intended VTI.
+Thus, they assess an answer to the question: `Given all available knowledge, how plausible is it that a version or state represented in the relation (that may or may not (dis)satisfy the user's non-temporal preferences) actually existed, exists or will exist during the time period indicated by the user?'
+\end{itemize}
+
+A fundamental question arises now: should one consider combining quantifications of these different categories?
+On one hand, such quantifications have clearly different semantics and it would not be clear what exactly the meaning would be of the result of such a combination or what the semantically most coherent ways would be to further process such combination results.
+Thus, it is important, for every query result tuple presented to the user, to certainly keep both different types of evaluation results as separate (meta)data.
+On the other hand, without an unambiguous and straightforward ranking of the query result tuples, the user cannot clearly discern the result tuples which comply well with his or her demands from those which don't.
+This would defeat the purpose of querying.
+Reasonably, such a ranking should be based on the elementary query condition evaluation results.
+As there cannot exist a ranking between quantifications of categories with different semantics, a combination of quantifications of satisfaction and possibility seems to be required.
+In most existing proposals requiring a combination of quantifications of satisfaction and possibility, both quantifications are combined as to restrict one another.
+The result is usually seen as a quantification of possibility.
+Below, this approach is translated to the specific situation encountered in the presented work.
+
+For every tuple $T$, the BSD's which are the evaluation results of the non-temporal BQC are combined to a single BSD $(s_n(T), d_n(T))$ as described in~\cite{Matthe2011ijis}.
+In this combination method, satisfaction degrees are combined with each other and separately, dissatisfaction degrees are combined with each other.
+This reasoning is now extended to include $Pos_{+}(I_{T})$, $Nec_{+}(I_{T})$, $Pos_{-}(I_{T})$ and $Nec_{-}(I_{T})$: a couple $(Pos_{+}(T), Nec_{+}(T))$, respectively $(Pos_{-}(T), Nec_{-}(T))$ is calculated, expressing the possibility and necessity that the version or state corresponding to $T$ complies with all of the user's positive, respectively negative demands.
+This calculation is done as follows:
+
+\vspace{-10pt}
+\begin{align}
+Pos_{+}(T) = & \min(s_n(T), Pos_{+}(I_{T})) \nonumber \\
+Nec_{+}(T) = &
+	\begin{cases}
+	0, & \text{if } Pos_{+}(T) < 1 \\
+	\min(s_n(T), Nec_{+}(I_{T})), & \text{else}
+	\end{cases}
+	\nonumber \\
+Pos_{-}(T) = & \min(d_n(T), Pos_{-}(I_{T})) \nonumber \\
+Nec_{-}(T) = &
+	\begin{cases}
+	0, & \text{if } Pos_{-}(T) < 1 \\
+	\min(d_n(T), Nec_{-}(I_{T})), & \text{else}
+	\end{cases}
+	\nonumber
+\end{align}
+
+Next, a tie-break approach is used to rank the versions or states represented in $R$: ranking is done based on the value of $Pos_{+}(T)$, with $Nec_{+}(T)$ as tie-breaker for $Pos_{+}(T)$, with $Pos_{-}(T)$ as tie-breaker for $(Pos_{+}(T), Nec_{+}(T))$ and finally $Nec_{-}(T)$ as tiebreaker for $(Pos_{+}(T), Nec_{+}(T), Pos_{-}(T))$.
+The results of this aggregation and ranking approach for the example relation and query are shown in table \ref{tab:ex-ranking}.
+
+The introduced technique for aggregating elementary query condition evaluation results and determining a ranking does not require the combination of quantifications with different interpretations and still presents the query result tuples in a ordered manner, where this ordering is consistent with the expected extend to which each tuple is usefull to the user.
+However, the technique still requires the user to decide which objects or concepts constitute the best query answer, although this decision is heavily supported.
+
+\vspace{-10pt}
+\begin{table}[ht]
+\caption{Aggregation and ranking results for the example relation and query.}
+\centering
+\begin{tabular}{|c|c|c|c|c||c|c|c|}
+\hline
+\textbf{ID} & \textbf{$Pos_{+}(T)$} & \textbf{$Nec_{+}(T)$} & \textbf{$Pos_{-}(T)$} & \textbf{$Nec_{-}(T)$} & \textbf{$BSD_{Q_{n,ex}}$} & \textbf{$(Pos_{+}(I_T), Nec_{+}(I_T))$} & \textbf{$(Pos_{-}(I_T), Nec_{-}(I_T))$} \\
+\hline
+\hline
+1 & 1 & 1 & 0 & 0 & $(1, 0)$ & $(1,1)$ & $(0,0)$ \\
+\hline
+2 & 0 & 0 & 0 & 0 & $(0, 0.7)$ & $(1,0)$ & $(0,0)$ \\
+\hline
+3 & 0 & 0 & 0 & 0 & $(0, 0)$ & $(11/25,0)$ & $(0,0)$ \\
+\hline
+4 & 0 & 0 & 0 & 0 & $(0.7, 0)$ & $(0,0)$ & $(0,0)$ \\
+\hline
+6 & 0 & 0 & 0 & 0 & $(1, 0)$ & $(0,0)$ & $(1,0)$ \\
+\hline
+7 & 0 & 0 & 0 & 0 & $(0, 0)$ & $(0,0)$ & $(1,1)$ \\
+\hline
+5 & 0 & 0 & 14/25 & 0 & $(0, 1)$ & $(0,0)$ & $(14/25,0)$ \\
+\hline
+\end{tabular}
+\label{tab:ex-ranking}
+\end{table}
+\vspace{-15pt}
+
+\section{Conclusions and Future Work\label{sec:conclusions}}
+In this paper, a novel technique to query a valid-time relation containing valid-time data subject to uncertainty in a bipolar way, is presented.
+This technique allows the user to specify a single valid-time bipolar query condition.
+A major issue concerning the need to combine quantifications of (dis)satisfaction with quantifications of possibility resulting from this technique is presented and shortly discussed, along with a solution for this issue.
+In the near future, the interactions between these types of quantifications and between uncertainty in valid-time data and bipolar querying will be further studied.
+Also, an approach to allow the valid-time bipolar query condition to be fuzzy will be considered.
 
 \bibliographystyle{splncs}
-\bibliography{paper}
+\bibliography{sources}
 

### 2013-05-31 trunk/fqas2013/paper.tex
Laatste kleine aanpassingen voor indienen.

@@
 
-\newcommand{\mytitlestring}{Bipolar Querying of Valid-time Intervals Subject to Uncertainty}
+\newcommand{\mytitlestring}{Bipolar Querying of Valid-Time Intervals Subject to Uncertainty}
 
@@
 Generally, database systems model (parts of) reality.
-For this, their databases contain data representing properties of real-world objects or concepts~\cite{Pons2012ijcis},~\cite{Billiet2012ipmu},~\cite{Pons2012ipmu}.
+For this, their databases contain data representing properties of real-world objects or concepts.
 Some essential properties of real-world objects or concepts are time-related.
-Thus, databases often contain data representing temporal values~\cite{Bohlen1998lncs}, which are basically indications of time and describe such properties~\cite{Pons2012ijcis},~\cite{Jensen1999ieeetkde},~\cite{Galindo2001}.
+Thus, databases often contain data representing temporal values~\cite{Bohlen1998lncs}, which are basically indications of time and describe such properties.
 These temporal values are usually either time intervals~\cite{Bohlen1998lncs} or instants~\cite{Bohlen1998lncs}, which may informally be seen as infinitesimally short `periods' or `points' in time.
-Based on their interpretation and purpose, temporal values can be classified into several categories, but the presented work will only consider valid-time indications, which indicate when corresponding data is a valid or true representation of the reality modelled by its database~\cite{Bohlen1998lncs},~\cite{Pons2012ijcis},~\cite{Billiet2012ipmu},~\cite{Pons2012ipmu}.
+Based on their interpretation and purpose, temporal values can be classified into several categories, but the presented work will only consider valid-time indications, which indicate when corresponding data is a valid or true representation of the reality modelled by its database~\cite{Bohlen1998lncs,Pons2012ijcis,Billiet2012ipmu,Pons2012ipmu}.
 
-A lot of database data are produced by humans, but human-made data are prone to imperfections, as some of these data may be vague, imprecise, contradictory, incomplete or uncertain~\cite{Billiet2012ipmu},~\cite{Pons2012ipmu},~\cite{Medina1994is},~\cite{Bosc2010ijufkbs}.
-Of course, data representing temporal values may contain such imperfections too~\cite{Billiet2012ipmu},~\cite{Pons2012ipmu},~\cite{Pons2012ijcis},~\cite{Dyreson1998acm}.
+A lot of database data are produced by humans, but human-made data are prone to imperfections, as some of these data may be vague or imprecise~\cite{Medina1994is}, contradictory, incomplete or uncertain~\cite{Billiet2012ipmu},~\cite{Pons2012ipmu},~\cite{Bosc2010ijufkbs}.
+Of course, data representing temporal values may contain such imperfections too~\cite{Pons2012ijcis,Billiet2012ipmu,Pons2012ipmu},~\cite{Dyreson1998acm}.
 The work presented in this paper will consider databases containing data representing valid-time intervals subject to uncertainty and will assume all non-temporal data in these databases to contain no imperfections.
@@
 A user may also query a database in a `fuzzy' way: the user describes the data which he or she finds desired or satisfactory by imperfectly describing the allowed values of these data~\cite{Zadrozny2008hrfipd}.
-These imperfect descriptions may contain vagueness or imprecision, often through the use of linguistic terms ~\cite{Devos1998jql},~\cite{Kacprzyk2001is}.
+These imperfect descriptions may contain vagueness or imprecision, often through the use of linguistic terms~\cite{Devos1998jql},~\cite{Kacprzyk2001is}.
 A user may also query a database in a `bipolar' way.
@@
 Compared to non-temporal user preferences, temporal user preferences usually have an uncommon nature and interpretation and thus expressing them relies on uncommon mechanics in querying: users usually prefer using specific temporal operators to express temporal preferences.
-Hence, several proposals have considered specific sets of temporal operators~\cite{Galindo2001},~\cite{Pons2012ijcis},~\cite{Schockaert2008ieeetfs}, often based on the possible temporal relationships between two time indications~\cite{Galindo2001},~\cite{Pons2012ijcis},~\cite{Schockaert2008ieeetfs}.
+Hence, several proposals have considered specific sets of temporal operators, often based on the possible temporal relationships between two time indications~\cite{Pons2012ijcis},~\cite{Galindo2001},~\cite{Schockaert2008ieeetfs}.
 Such temporal relationships define semantically meaningful relationships with a temporal nature, between two time indications.
 In~\cite{Allen1983cacm}, a collection of temporal relationships between two time intervals (and as a special case instants) is introduced and this collection is considered groundbreaking.
-Of course, to query temporal data in a fuzzy way, fuzzy variants of such temporal operators are necessary and several proposals have thus introduced such variants~\cite{Galindo2001},~\cite{Pons2012ipmu}, often based on fuzzy variants of such temporal relationships~\cite{Schockaert2008ieeetfs},~\cite{Pons2013ijufkbs}.
+Of course, to query temporal data in a fuzzy way, fuzzy variants of such temporal operators are necessary and several proposals have thus introduced such variants~\cite{Pons2012ipmu},~\cite{Galindo2001}, often based on fuzzy variants of such temporal relationships~\cite{Schockaert2008ieeetfs},~\cite{Pons2013ijufkbs}.
 
 Techniques for the regular or fuzzy querying of valid-time databases containing valid-time data subject to uncertainty are considered by several existing proposals~\cite{Pons2012ijcis},~\cite{Pons2012ipmu},~\cite{Pons2013ijufkbs}.
-However, to the knowledge of the authors, few proposals have considered the bipolar querying of valid-time databases~\cite{Billiet2011fqas} and even fewer the bipolar querying of valid-time databases containing temporal data subject to uncertainty.
+However, to the knowledge of the authors, only one proposal has considered the bipolar querying of valid-time databases (in~\cite{Billiet2011fqas}, a technique is proposed to query a valid-time database containing temporal data subject to imprecision in a bipolar way) and none have considered the bipolar querying of valid-time databases containing temporal data subject to uncertainty.
 Thus, this paper presents a novel technique to query a valid-time relation containing valid-time data subject to uncertainty in a bipolar way, which allows the user to specify a single bipolar temporal query condition.
-This paper is structured as follows: in section \ref{sec:preliminaries}, some preliminary concepts and techniques are described, in section \ref{sec:proposal}, the novel technique which is the main contribution of the work presented in this paper, is presented and in section \ref{sec:conclusions}, the conclusions of this paper and some directions for future research are given.
+This paper is structured as follows: in section \ref{sec:preliminaries}, some preliminary concepts and techniques are described, in section \ref{sec:proposal}, the novel technique which is the main contribution of the work presented in this paper, is explained and in section \ref{sec:conclusions}, the conclusions of this paper and some directions for future research are given.
 
@@
 This uncertainty is always assumed to be caused by a (partial) lack of knowledge: the exact, intended time indication is not known, eventhough there is only one time indication intended and as such no variability.
-Confidence about exactly which time indication is the intended one in the context of such uncertainty is modelled using possibility theory~\cite{Dubois1988},~\cite{Pons2013ijufkbs}.
+Confidence about exactly which time indication is the intended one in the context of such uncertainty is modelled using possibility theory~\cite{Pons2013ijufkbs},~\cite{Dubois1988}.
 In the presented work, `possibility' and `necessity' are always interpreted as measures of plausibility, respectively necessity, given all available knowledge.
@@
 
-\subsection{Valid-time Relations}
-The presented work uses the relational database model~\cite{Pons2012ijcis},~\cite{Galindo2001}.
-Here, a \emph{relation} models a collection of similar real-world concepts or objects.
-To achieve this, such a relation has a set of $n$ \emph{attributes}, of which each describes a different property shared by all of these concepts or objects, and a set of $n$-tuples, of which each represents one of these real-world objects or concepts and contains $n$ values, each representing the capacity of the object or concept corresponding to the tuple, of the property described by the attribute corresponding to the value~\cite{Billiet2012ipmu},~\cite{Pons2012ipmu}.
-
+\subsection{Valid-Time Relations}
 A \emph{valid-time relation} (VTR) always has \emph{valid-time attributes}.
@@
 
-\subsection{Uncertainty in Valid-time Intervals}
+\subsection{Uncertainty in Valid-Time Intervals}
 The presented work allows VTI to be subject to uncertainty, by allowing them to be \emph{ill-known valid-time intervals} (IKVTI).
-The concept of IKVTI is based on the concepts of \emph{possibilistic variables} (PV) and \emph{ill-known intervals} (IKI)~\cite{Dubois1988cma},~\cite{Pons2013ijufkbs},~\cite{Billiet2012ipmu}:
+The concept of IKVTI is based on the concepts of \emph{possibilistic variables} (PV) and \emph{ill-known intervals} (IKI)~\cite{Billiet2012ipmu},~\cite{Pons2013ijufkbs},~\cite{Dubois1988cma}.
 
@@
 Now consider a set $U$ containing single values (and not collections of values).
-When a PV $X_{v}$ is defined on such a set $U$, the unique value $X_{v}$ takes, which is (partially) unknown, will be a single value in $U$ and is called an \emph{ill-known value} (IKV) in $U$ in the presented work~\cite{Dubois1988cma},~\cite{Pons2013ijufkbs},~\cite{Billiet2012ipmu}.
+When a PV $X_{v}$ is defined on such a set $U$, the unique value $X_{v}$ takes, which is (partially) unknown, will be a single value in $U$ and is called an \emph{ill-known value} (IKV) in $U$~\cite{Billiet2012ipmu},~\cite{Pons2013ijufkbs},~\cite{Dubois1988cma}.
 In this paper, IKV will be denoted using lower-case letters.
-The work presented in this paper uses a specific kind of IKI, defined as follows, although other definitions exist~\cite{Pons2012ipmu},~\cite{Billiet2012ipmu}:
+The work presented in this paper uses a specific kind of IKI, defined as follows, although other definitions exist~\cite{Billiet2012ipmu},~\cite{Pons2012ipmu}.
 
@@
 
-Specifically concerning valid time, an IKI in a time domain represented by the domains of a VTR's valid-time attributes is called an \emph{ill-known valid-time interval} (IKVTI) in the presented work.
+Specifically concerning valid time, an IKI in a time domain represented by the domains of a VTR's valid-time attributes is called an \emph{ill-known valid-time interval} (IKVTI).
 The work presented in this paper requires the possibility distributions defining an IKVTI's IKV to be convex~\cite{Pons2013ijufkbs}.
@@
 \begin{definition}
-Given an ordered set $U$, an \emph{ill-known constraint} (IKC) $C$ on $U$ is specified by means of a binary relation $R \subseteq U^{2}$ and a fixed IKV $v$ in $U$, i.e.:
-\begin{align}
-C \triangleq (R,v) \nonumber
-\end{align}
+Given an ordered set $U$, an \emph{ill-known constraint} (IKC) $C = (R,v)$ on $U$ is specified by means of a binary relation $R \subseteq U^{2}$ and a fixed IKV $v$ in $U$.
 Any set $A \subseteq U$ now satisfies IKC $C = (R,v)$ if and only if:
@@
 The satisfaction of an IKC $C$ by a set $A$ will be noted $C(A)$ in this paper.
-Consider an ordered set $U$, an IKC $C \triangleq (R,v)$ on $U$ and a set $A \subseteq U$.
+Consider an ordered set $U$, an IKC $C = (R,v)$ on $U$ and a set $A \subseteq U$.
 Due to the uncertainty inherent to $v$, it is uncertain whether $A$ satisfies $C$ or not.
@@
 
-\vspace{-10pt}
 \begin{align}
-\Pos(C(A)) & = \min_{a \in A}\left(\sup_{(a,w) \in R}\pi_{V}(w)\right) \label{ill-known-pos}\\
-\Nec(C(A)) & = \min_{a \in A}\left(\inf_{(a,w) \notin R} 1-\pi_{V}(w)\right) \label{ill-known-nec}
+\Pos(C(A)) & = \min_{a \in A}\left(\sup_{(a,w) \in R}\pi_{X_{v}}(w)\right) \label{ill-known-pos}\\
+\Nec(C(A)) & = \min_{a \in A}\left(\inf_{(a,w) \notin R} 1-\pi_{X_{v}}(w)\right) \label{ill-known-nec}
 \end{align}
 
-Given an ordered set $U$, degrees of possibility and necessity that a set $A \subseteq U$ satisfies a boolean combination of IKC on $U$ can be found by using the possibilistic extensions of boolean operators `and' ($\wedge$), `or' ($\vee$) and `not' ($\neg$), as described in~\cite{Pons2013ijufkbs},~\cite{Pons2012ipmu},~\cite{Billiet2012ipmu}.
-
-The IKC framework now allows evaluating a given Allen relationship $AR$ between a given CTI $I$ and a given IKVTI $J = \left[s, e\right]$ by allowing the calculation of the degrees of possibility and necessity that $I$ $AR$ $J$ holds.
-For this, the combination of $AR$ and $J$ is translated to a specific boolean combination of specific IKC.
-These translations are shown in table \ref{tab:allen-relations}.
-Every row of this table corresponds to a given Allen relationship between $I$ and $J$, indicated by the row's value in the `Allen Relationship' column.
-The collections of specific IKC for given Allen relationships are shown in the `Constraints' column (every $C_i, i \in \{1, 2, 3, 4\}$ denotes an IKC) and the specific combination of these IKC used for evaluation of the Allen relationships are shown in the `Combination' column.
-Finally, the degrees of possibility and necessity that $I$ $AR$ $J$ holds are then the degrees of possibility, respectively necessity that $I$ satisfies the specific aggregation of specific IKC found as translation of the combination of $AR$ and $J$.
-Using the formulas shown above, the requested possibility and necessity degrees can be calculated from these.
-
-\begin{table}[ht]
-\caption{The translations of Allen relationships to the IKC framework.}
+\begin{table}[b]
+\caption{The translations of Allen relationships to the IKC framework}
 \centering
@@
 
+Given an ordered set $U$, degrees of possibility and necessity that a set $A \subseteq U$ satisfies a boolean combination of IKC on $U$ can be found by using the possibilistic extensions of boolean operators `and' ($\wedge$), `or' ($\vee$) and `not' ($\neg$), as described in~\cite{Billiet2012ipmu},~\cite{Pons2012ipmu},~\cite{Pons2013ijufkbs}.
+
+The IKC framework now allows evaluating a given Allen relationship $AR$ between a given CTI $I$ and a given IKVTI $J = \left[s, e\right]$ by allowing the calculation of the degrees of possibility and necessity that $I$ $AR$ $J$ holds.
+For this, the combination of $AR$ and $J$ is translated to a specific boolean combination of specific IKC.
+These translations are shown in table \ref{tab:allen-relations}.
+Every row of this table corresponds to a given Allen relationship between $I$ and $J$, indicated by the row's value in the `Allen Relationship' column.
+The collections of specific IKC for given Allen relationships are shown in the `Constraints' column (every $C_i, i \in \{1, 2, 3, 4\}$ denotes an IKC) and the specific combination of these IKC used for evaluation of the Allen relationships are shown in the `Combination' column.
+Finally, the degrees of possibility and necessity that $I$ $AR$ $J$ holds are then the degrees of possibility, respectively necessity that $I$ satisfies the specific aggregation of specific IKC found as translation of the combination of $AR$ and $J$.
+Using the formulas shown above, the requested possibility and necessity degrees can be calculated from these.
+
 \subsection{Bipolar Querying}
-As mentioned before, humans may express their query preferences using both positive and negative query conditions~\cite{Matthe2011ijis},~\cite{DeTre2010ieeetfs}.
+As mentioned before, humans may express their query preferences using both positive and negative query conditions~\cite{DeTre2010ieeetfs},~\cite{Matthe2011ijis}.
 If the semantics of these conditions are non-symmetric, meaning that the positive preferences can not be derived from the negative or vice versa, the bipolarity in this query is called \emph{heterogenous}~\cite{Matthe2011ijis}.
@@
 
+A query usually takes the form of a boolean combination of elementary query conditions.
+Every elementary query condition then expresses the user's demands concerning a single attribute.
 Bipolarity in a query can either be specified between or inside the elementary query conditions.
@@
 Consider a relation attribute $A$.
-Let $dom_{A}$ be the domain of $A$'s data type, let $\mu_{c_{A}}$ and $\nu_{c_{A}}$ be membership functions from $dom_{A}$ to the unit interval $\left[0, 1\right]$, where $\mu_{c_{A}}(x)$ represents to what extent $x \in dom_{A}$ is satisfactory and $\nu_{c_{A}}(x)$ to what extent $x$ is unsatisfactory to a user, then a BQC $c_{A}$ expressing this user's preferences concerning $A$ can be modelled by an Intuitionistic Fuzzy Set (IFS)~\cite{Atanassov1986fss} as~\cite{Matthe2011ijis},~\cite{DeTre2010ieeetfs}:
+Let $dom_{A}$ be the domain of $A$'s data type, let $\mu_{c_{A}}$ and $\nu_{c_{A}}$ be membership functions from $dom_{A}$ to the unit interval $\left[0, 1\right]$, where $\mu_{c_{A}}(x)$ represents to what extent $x \in dom_{A}$ is satisfactory and $\nu_{c_{A}}(x)$ to what extent $x$ is unsatisfactory to a user, then a BQC $c_{A}$ expressing this user's preferences concerning $A$ can be modelled by an Intuitionistic Fuzzy Set (IFS)~\cite{Atanassov1986fss} as~\cite{DeTre2010ieeetfs},~\cite{Matthe2011ijis}:
 \vspace{-5pt}
@@
 
-Note that to allow overspecification of the user's preferences, the IFS's consistency condition can be relaxed, which means that there may exist values $v \in dom_{A}$ for which $\mu_{c_{A}}(v) + \nu_{c_{A}}(v) > 1$~\cite{Matthe2011ijis},~\cite{DeTre2010ieeetfs}.
+Note that to allow overspecification of the user's preferences, the IFS's consistency condition can be relaxed, which means that there may exist values $v \in dom_{A}$ for which $\mu_{c_{A}}(v) + \nu_{c_{A}}(v) > 1$~\cite{DeTre2010ieeetfs},~\cite{Matthe2011ijis}.
 
@@
 
-In general, the evaluation of a BQC $c_{A}$ on relation attribute $A$ for a relation tuple $R$ will result in a BSD $(s_{c_A}^R, d_{c_A}^R)$, which is calculated as follows.
-Let $R[A]$ denote the value of tuple $R$ for attribute $A$, then~\cite{Matthe2011ijis}:
+In general, the evaluation of a BQC $c_{A}$ on relation attribute $A$ for a tuple $r$ will result in a BSD $(s_{c_A}^r, d_{c_A}^r)$, which is calculated as follows.
+Let $r[A]$ denote the value of tuple $r$ for attribute $A$, then~\cite{Matthe2011ijis}:
 \begin{equation}\label{eq:bsdeval}
-(s_{c_A}^R, d_{c_A}^R)=(\mu_{c_A}(R[A]), \nu_{c_A}(R[A]))
+(s_{c_A}^r, d_{c_A}^r)=(\mu_{c_A}(r[A]), \nu_{c_A}(r[A]))
 \end{equation}
@@
 
-\subsection{Valid-time Relations Subject to Uncertainty}
+\subsection{Valid-Time Relations Subject to Uncertainty}
 The presented proposal will concern VTR where the VTI are IKVTI.
@@
 \begin{table}[ht]
-\caption{The example relation used in this paper.}
+\caption{The example relation used in this paper}
 \centering
@@
 \end{table}
-\vspace{-10pt}
+\vspace{-15pt}
 
@@
 \end{align}
+\vspace{-15pt}
 
-\subsection{Querying Using Bipolar Valid-time Conditions}
+\subsection{Querying Using Bipolar Valid-Time Conditions \label{sec:query}}
 The presented work introduces a novel querying technique.
-According to this technique, a user query $Q$ consists of two separate parts $Q_N$ and $Q_T$: $Q = (Q_N, Q_T)$.
-Here, $Q_N$ is a boolean combination of BQC on non-valid-time attributes, expressing the user's non-temporal demands.
-$Q_T$ expresses the user's valid-time demands and is a single crisp temporal BQC $((AR_{pos}, I_{pos}), (AR_{neg}, I_{neg}))$, where both $AR_{pos}$ and $AR_{neg}$ are Allen relationships and both $I_{pos}$ and $I_{neg}$ are CTI.: $Q_T = ((AR_{pos}, I_{pos}), (AR_{neg}, I_{neg}))$.
-The interpretation is that the user requires an object or concept that has a version or state that complies with his or her non-temporal demands and was, is or will be valid during a time interval which is in Allen relationship $AR_{pos}$ with $I_{pos}$ and wasn't, isn't or won't be valid during a time interval which is in Allen relationship $AR_{neg}$ with $I_{neg}$.
+The most interesting aspect of this technique is that it allows the user to specify a bipolar valid-time demand.
+According to this technique, a user query $Q$ consists of two separate parts $Q_n$ and $Q_t$: $Q = (Q_n, Q_t)$.
+Here, $Q_n$ is a boolean combination of BQC on non-valid-time attributes, expressing the user's non-temporal demands.
+$Q_t$ expresses the user's valid-time demands and is a single crisp temporal BQC $((AR_{+}, I_{+}), (AR_{-}, I_{-}))$, where both $AR_{+}$ and $AR_{-}$ are Allen relationships and both $I_{+}$ and $I_{-}$ are CTI.
+The interpretation is that the user requires an object or concept that has a version or state that complies with his or her non-temporal demands and was, is or will be valid during a time interval which is in Allen relationship $AR_{+}$ with $I_{+}$ and wasn't, isn't or won't be valid during a time interval which is in Allen relationship $AR_{-}$ with $I_{-}$.
 
@@
 \begin{align}
-Q_{ex} = &\ (Q_{N,ex}, ((AR_{pos,ex}, I_{pos,ex}), (AR_{neg,ex}, I_{neg,ex}))), \\
-Q_{N,ex} = &\ \{(x, \mu_{ex}(x), \nu_{ex}(x)\}, \forall x \in \mathbb{S} \nonumber \\
-AR_{pos,ex} = &\ \ AR_{neg,ex}\ =\ DURING \nonumber \\
-I_{pos,ex} = &\ \left]- infinity, 1/1/1210\right],\ \ I_{neg,ex}\ =\ \left[1/1/1214, + infinity\right[ \nonumber
+Q_{ex} = &\ (Q_{n,ex}, Q_{t,ex})\ =\ (Q_{n,ex}, ((AR_{+,ex}, I_{+,ex}), (AR_{-,ex}, I_{-,ex}))), \nonumber\\
+Q_{n,ex} = &\ \{(x, \mu_{ex}(x), \nu_{ex}(x)), \forall x \in \mathbb{S} \}\nonumber \\
+AR_{+,ex} = &\ \ AR_{-,ex}\ =\ DURING \nonumber \\
+I_{+,ex} = &\ \left]- infinity, 1/1/1210\right],\ \ I_{-,ex}\ =\ \left[1/1/1214, + infinity\right[ \nonumber
 \end{align}
@@
 \end{align}
-\vspace{-25pt}
+and $\mathbb{S}$ is the set of all author names in the example relation's `Author' attribute domain.
+%\vspace{-25pt}
 
 \subsection{Elementary Query Condition Evaluation}
-Generally, a first step in determining which objects or concepts corresponding to relation tuples to present to a user as answer to his or her query, is evaluating the query's elementary conditions for every relation tuple.
-In the presented work, for every tuple $T$ of VTR $R$, the following is done separately:
+Generally, a first step in determining which objects or concepts corresponding to tuples to present to a user as answer to his or her query, is evaluating the query's elementary conditions for every tuple.
+In the presented work, given a user query $Q$ constructed as proposed in section \ref{sec:query} and using the same notations, the following is done separately for every tuple $T$ of VTR $R$:
 
 \begin{itemize}
-\item every non-temporal BQC in $Q_N$ is evaluated as described in~\cite{Matthe2011ijis}, resulting in a BSD for each.
+\item every non-temporal BQC in $Q_n$ is evaluated as described in~\cite{Matthe2011ijis}, resulting in a BSD for each.
 The interpretation is as described in~\cite{Matthe2011ijis}: the BSD's satisfaction, respectively dissatisfaction degrees express to which extent $T$ satisfies, respectively dissatisfies the user preferences expressed by the BQC.
-\item the query's temporal BQC $Q_T$ is evaluated as follows: let $I_T$ be the tuple's IKVTI, then separately, $I_T$ $AR_{pos}$ $I_{pos}$, respectively $I_T$ $AR_{neg}$ $I_{neg}$ are evaluated using the IKC framework as described in~\cite{Pons2013ijufkbs},~\cite{Pons2012ijcis},~\cite{Pons2012ipmu}, resulting in a possibility degree $pos_{I_{pos}}(T)$ and a necessity degree $nec_{I_{pos}}(T)$, respectively a possibility degree $pos_{I_{neg}}(T)$ and a necessity degree $nec_{I_{neg}}(T)$.
-The interpretation is that $pos_{I_{pos}}(T)$ and $nec_{I_{pos}}(T)$ express the possibility, respectively necessity, that the time interval during which the version or state represented by $T$ is actually valid and which is intended by $I_T$, is in relationship $AR_{pos}$ with $I_{pos}$ and thus complies with the user's positive temporal demand and is preferred, while $pos_{I_{neg}}(T)$ and $nec_{I_{neg}}(T)$ express the possibility, respectively necessity, that the time interval during which the version or state represented by $T$ is actually valid and which is intended by $I_T$, is in relationship $AR_{neg}$ with $I_{neg}$ and thus complies with the user's negative temporal demand and is not preferred.
+\item the query's temporal BQC $Q_t$ is evaluated as follows.
+Let $I_T$ be the tuple's IKVTI.
+Then, independently, the statements `$I_T$ $AR_{+}$ $I_{+}$', respectively `$I_T$ $AR_{-}$ $I_{-}$' are evaluated using the IKC framework as described in~\cite{Pons2012ijcis},~\cite{Pons2012ipmu},~\cite{Pons2013ijufkbs}.
+These evaluations result in a possibility degree $Pos_{+}(I_{T})$ and a necessity degree $Nec_{+}(I_{T})$, respectively a possibility degree $Pos_{-}(I_{T})$ and a necessity degree $Nec_{-}(I_{T})$.
+The interpretation is that $Pos_{+}(I_{T})$ and $Nec_{+}(I_{T})$ express the possibility, respectively necessity, that the time interval during which the version or state represented by $T$ is valid and which is intended by $I_T$, is in relationship $AR_{+}$ with $I_{+}$ and thus complies with the user's positive temporal demand.
+Furthermore, the interpretation is that $Pos_{-}(I_{T})$ and $Nec_{-}(I_{T})$ express the possibility, respectively necessity, that the time interval during which the version or state represented by $T$ is valid and which is intended by $I_T$, is in relationship $AR_{-}$ with $I_{-}$ and thus complies with the user's negative temporal demand.
 \end{itemize}
@@
 \begin{table}[ht]
-\caption{Elementary query condition evaluation results for the example relation.}
+\caption{Elementary query condition evaluation results for the example relation}
 \centering
@@
 \hline
-\textbf{ID} & \textbf{$Q_{N,ex}$} & \textbf{$(pos_{I_{pos}}, nec_{I_{pos}})$} & \textbf{$(pos_{I_{neg}}, nec_{I_{neg}})$} \\
+\textbf{ID} & \textbf{$BSD_{Q_{N,ex}}(T)$} & \textbf{$(Pos_{+}(I_{T}), Nec_{+}(I_{T})$} & \textbf{$(Pos_{-}(I_T), Nec_{-}(I_T)$} \\
 \hline
@@
 \subsection{Aggregation and Ranking}
-Generally, a second step in determining which objects or concepts corresponding to relation tuples to present to a user as answer to his or her query, is aggregating, for every relation tuple, the tuple's evaluation results for the query's elementary conditions, in order to to determine how well the tuple complies with the entire user request expressed by the combination of the elementary query conditions.
+Generally, a second step in determining which objects or concepts corresponding to tuples to present to a user as answer to his or her query, is aggregating, for every tuple, the tuple's evaluation results for the query's elementary conditions, in order to determine how well the tuple complies with the entire user request expressed by the combination of the elementary query conditions.
 Usually, these evaluation results are quantifications of (dis)satisfaction.
@@
 	\item the BSD's (dis)satisfaction degrees constitute quantifications of (dis)satisfaction: they quantify to which extent a tuple's attribute values (dis)satisfy a user's non-temporal preferences and thus assess an answer to the question: `To what extent does a version or state represented in the relation (dis)satisfy the user's request and could thus be a (un)wanted result?'
-	\item the possibility and necessity degrees $pos_{I_{pos}}$, $nec_{I_{pos}}$, respectively $pos_{I_{neg}}$, $nec_{I_{neg}}$ constitute quantifications of possibility and necessity: they quantify the possibility that a tuple's intended crisp VTI (dis)satisfies the user's temporal preferences by quantifying confidence about exactly which crisp VTI is the tuple's intended VTI and thus assess an answer to the question: `How plausible is it that a version or state represented in the relation (that may or may not (dis)satisfy the user's request) actually existed, exists or will exist during the time period indicated by the user?'
+	\item the possibility and necessity degrees $Pos_{+}$, $Nec_{+}$, respectively $Pos_{-}$, $Nec_{-}$ constitute quantifications of possibility and necessity: they quantify the possibility that a tuple's intended crisp VTI does (not) comply with the user's temporal demands by quantifying confidence about exactly which crisp VTI is the tuple's intended VTI.
+Thus, they assess an answer to the question: `Given all available knowledge, how plausible is it that a version or state represented in the relation (that may or may not (dis)satisfy the user's non-temporal preferences) actually existed, exists or will exist during the time period indicated by the user?'
 \end{itemize}
@@
 Thus, it is important, for every query result tuple presented to the user, to certainly keep both different types of evaluation results as separate (meta)data.
-On the other hand, without an unambiguous and straightforward ranking of the query result tuples, the user cannot clearly discern the result tuples which comply well with his or her demands from those which don't, which defeats the purpose of querying.
+On the other hand, without an unambiguous and straightforward ranking of the query result tuples, the user cannot clearly discern the result tuples which comply well with his or her demands from those which don't.
+This would defeat the purpose of querying.
 Reasonably, such a ranking should be based on the elementary query condition evaluation results.
-As there cannot exist a ranking between quantifications of categories with different semantics, a combination seems to be required.
+As there cannot exist a ranking between quantifications of categories with different semantics, a combination of quantifications of satisfaction and possibility seems to be required.
 In most existing proposals requiring a combination of quantifications of satisfaction and possibility, both quantifications are combined as to restrict one another.
@@
 
-For every tuple $T$, the BSD's which are the evaluation results of the non-temporal BQC are combined to a single BSD $(s_N(T), d_N(T))$ as described in~\cite{Matthe2011ijis}.
+For every tuple $T$, the BSD's which are the evaluation results of the non-temporal BQC are combined to a single BSD $(s_n(T), d_n(T))$ as described in~\cite{Matthe2011ijis}.
 In this combination method, satisfaction degrees are combined with each other and separately, dissatisfaction degrees are combined with each other.
-This reasoning is extended to include $pos_{I_{pos}}(T)$, $nec_{I_{pos}}(T)$, $pos_{I_{neg}}(T)$ and $nec_{I_{neg}}(T)$: a couple $(pos_{pos}(T), nec_{pos}(T))$, respectively $(pos_{neg}(T), nec_{neg}(T))$ is calculated, expressing the possibility and necessity that the version or state corresponding to $T$ complies with the user's positive, respectively negative demands.
+This reasoning is now extended to include $Pos_{+}(I_{T})$, $Nec_{+}(I_{T})$, $Pos_{-}(I_{T})$ and $Nec_{-}(I_{T})$: a couple $(Pos_{+}(T), Nec_{+}(T))$, respectively $(Pos_{-}(T), Nec_{-}(T))$ is calculated, expressing the possibility and necessity that the version or state corresponding to $T$ complies with all of the user's positive, respectively negative demands.
 This calculation is done as follows:
@@
 \begin{align}
-pos_{pos}(T) = & \min(s_N(T), pos_{I_{pos}}(T)) \nonumber \\
-nec_{pos}(T) = &
+Pos_{+}(T) = & \min(s_n(T), Pos_{+}(I_{T})) \nonumber \\
+Nec_{+}(T) = &
 	\begin{cases}
-	0, & \text{if } pos_{pos}(T) < 1 \\
-	\min(s_N(T), nec_{I_{pos}}(T)), & \text{else}
+	0, & \text{if } Pos_{+}(T) < 1 \\
+	\min(s_n(T), Nec_{+}(I_{T})), & \text{else}
 	\end{cases}
 	\nonumber \\
-pos_{neg}(T) = & \min(d_N(T), pos_{I_{neg}}(T)) \nonumber \\
-nec_{neg}(T) = &
+Pos_{-}(T) = & \min(d_n(T), Pos_{-}(I_{T})) \nonumber \\
+Nec_{-}(T) = &
 	\begin{cases}
-	0, & \text{if } pos_{neg}(T) < 1 \\
-	\min(d_N(T), nec_{I_{neg}}(T)), & \text{else}
+	0, & \text{if } Pos_{-}(T) < 1 \\
+	\min(d_n(T), Nec_{-}(I_{T})), & \text{else}
 	\end{cases}
@@
 
-Next, a tie-break approach is used to rank the versions or states represented in $R$: ranking is done based on the value of $pos_{pos}(T)$, with $nec_{pos}(T)$ as tie-breaker for $pos_{pos}(T)$, with $pos_{neg}(T)$ as tie-breaker for $(pos_{pos}(T), nec_{pos}(T))$ and finally $nec_{neg}(T)$ as tiebreaker for $(pos_{pos}(T), nec_{pos}(5), pos_{neg}(5))$.
+Next, a tie-break approach is used to rank the versions or states represented in $R$: ranking is done based on the value of $Pos_{+}(T)$, with $Nec_{+}(T)$ as tie-breaker for $Pos_{+}(T)$, with $Pos_{-}(T)$ as tie-breaker for $(Pos_{+}(T), Nec_{+}(T))$ and finally $Nec_{-}(T)$ as tiebreaker for $(Pos_{+}(T), Nec_{+}(T), Pos_{-}(T))$.
 The results of this aggregation and ranking approach for the example relation and query are shown in table \ref{tab:ex-ranking}.
@@
 
-\vspace{-10pt}
+\vspace{-15pt}
 \begin{table}[ht]
-\caption{Aggregation and ranking results for the example relation and query.}
+\caption{Aggregation and ranking results for the example relation and query}
 \centering
@@
 \hline
-\textbf{ID} & \textbf{$pos_{pos}(o)$} & \textbf{$nec_{pos}(o)$} & \textbf{$pos_{neg}(o)$} & \textbf{$nec_{neg}(o)$} & \textbf{$Q_{N,ex}$} & \textbf{$(pos_{I_{pos}}, nec_{I_{pos}})$} & \textbf{$(pos_{I_{neg}}, nec_{I_{neg}})$} \\
+\textbf{ID} & \textbf{$Pos_{+}(T)$} & \textbf{$Nec_{+}(T)$} & \textbf{$Pos_{-}(T)$} & \textbf{$Nec_{-}(T)$} & \textbf{$BSD_{Q_{n,ex}}$} & \textbf{$(Pos_{+}(I_T), Nec_{+}(I_T))$} & \textbf{$(Pos_{-}(I_T), Nec_{-}(I_T))$} \\
 \hline
@@
 \end{table}
-\vspace{-15pt}
+\vspace{-20pt}
 
 \section{Conclusions and Future Work\label{sec:conclusions}}
-In this paper, a novel technique is presented, to query a valid-time relation containing valid-time data subject to uncertainty in a bipolar way, which allows the user to specify a single valid-time bipolar query condition.
-A major issue concerning the necessity to combine quantifications of (dis)satisfaction with quantifications of possibility resulting from this technique and a solution for this is presented and shortly discussed.
+In this paper, a novel technique to query a valid-time relation containing valid-time data subject to uncertainty in a bipolar way, is presented.
+This technique allows the user to specify a single valid-time bipolar query condition.
+A major issue concerning the need to combine quantifications of (dis)satisfaction with quantifications of possibility resulting from this technique is presented and shortly discussed, along with a solution for this issue.
 In the near future, the interactions between these types of quantifications and between uncertainty in valid-time data and bipolar querying will be further studied.
 Also, an approach to allow the valid-time bipolar query condition to be fuzzy will be considered.
+
+\vspace{-10pt}
 

### 2013-06-19 eusflat2013/tm.tex
Reviews included. Now only shortening is required.

@@
 Usually, different time intervals are visualized as different parallel line segments in the same image plane.
-However, as this linear approach introduces a few issues, time intervals are visualized as points in the image plane in the TM framework~\cite{Weghe2007},~\cite{DeTre2012}.
-To achieve this, a horizontal line segment visualizing the used (part of the) time axis is drawn in the image plane.
+However, as this linear approach introduces a few issues, time intervals are visualized as points in the image plane in the TM framework~\cite{DeTre2012},~\cite{Weghe2007}.
+To achieve this, a horizontal line segment visualizing the used (part of the) time axis is drawn in the image plane (this is called the \emph{time line}).
 Then, a triangle is drawn, using this line segment as a side of which the angles with the other two sides have sizes $\alpha$ and $-\alpha$ respectively.
 The area contained in this triangle is called the \emph{interval space} and will contain all visualizations of intervals.
-Now, the procedure used to visualize a CTI $\left[s, e\right]$ in the interval space is the following~\cite{Weghe2007},~\cite{DeTre2012}:
+Now, the procedure used to visualize a CTI $\left[s, e\right]$ in the interval space is the following~\cite{DeTre2012},~\cite{Weghe2007}:
 
 \begin{enumerate}
-	\item $s$ is visualized as a point on the visualization of the time axis.
-	\item A straight half-line $L_s$ is constructed from this point, having, in this point, an angle of size $\alpha$ with the visualization of the time axis.
-	\item $e$ is visualized as a point on the visualization of the time axis.
-	\item A straight half-line $L_e$ is constructed from this point, having, in this point, an angle of size $-\alpha$ with the visualization of the time axis.
+	\item $s$ is visualized as a point on the time line.
+	\item A straight half-line $L_s$ is constructed from this point, having, in this point, an angle of size $\alpha$ with the time line.
+	\item $e$ is visualized as a point on the time line.
+	\item A straight half-line $L_e$ is constructed from this point, having, in this point, an angle of size $-\alpha$ with the time line.
 	\item The intersection point of $L_s$ and $L_e$ is the intended visualization of interval $\left[s, e\right]$.
@@
 
-A similar point visualizing an interval is generally called an \emph{interval point}~\cite{Weghe2007},~\cite{DeTre2012}.
+A similar point visualizing an interval is generally called an \emph{interval point}~\cite{DeTre2012},~\cite{Weghe2007}.
 In this work, as it is usually done, the size of $\alpha$ is chosen to be $45^{\circ}$.
@@
 \begin{align}
-J \triangleq (\pi_s, \pi_e)
+J \triangleq (\pi_s, \pi_e) \nonumber
 \end{align}
-where $\pi_s$ and $\pi_e$ are two convex possibility distributions on $T$.
+where $\pi_s$ and $\pi_e$ are two convex~\cite{Dubois1983} possibility distributions on $T$.
 \end{definition}
@@
 Given an instant $x \in T$, $\pi_s(x)$ now expresses the possibility that $x$ is the starting instant of $J$ and $\pi_e(x)$ expresses the possibility that $x$ is the ending instant of $J$.
-Thus, $J$ intends to indicate just one CTI, but due to a (partial) lack of knowledge, it is (partially) unknown exactly which time interval this is.
-Because $\pi_s$ and $\pi_e$ are required to be convex, the possibility $\pi_J(I)$ that a given CTI $I = \left[s_i, e_i\right]$ is the exact time interval intended by $J$ can now be uniquely determined by~\cite{DeTre2012}:
+Thus, $J$ intends to indicate just one CTI, but it is (partially) unknown exactly which time interval this is.
+The possibility $\pi_J(I)$ that a given CTI $I = \left[s_i, e_i\right]$ is the exact time interval intended by $J$ can now be uniquely determined by~\cite{DeTre2012}:
 
@@
 An UTI $J$ in $T$ defined by a pair $(\pi_s, \pi_e)$ is now visualized as an area in the image plane.
-This area is constructed as follows: for every CTI $K$ for which the possibility $\pi_J(K)$ of $K$ being the time interval represented by $J$ is higher than zero, the interval point is drawn and is given a color.
-This color is part of a linear single-color greyscale and its saturation is linearly related to the degree of possibility that $K$ is the time interval intended by $J$.
-Thus, every drawn interval point in this visualization corresponds to an interval with a higher-than-zero possibility of being the time interval intended by $J$.
+This area is constructed as follows: for every CTI $K$ for which the possibility $\pi_J(K)$ of $K$ being the CTI intended by $J$ is higher than zero, the interval point is drawn and is given a color.
+This color is part of a linear single-color greyscale and its saturation is linearly related to the degree of possibility that $K$ is the CTI intended by $J$.
 Together, these drawn interval points form the visualization of $J$~\cite{DeTre2012}.
 An example of such construction is shown in figure \ref{fig:tm-ill-const-ex}.
-In this figure, a darker color of interval point indicates a higher and a lighter color a lower possibility that the corresponding interval is the intended interval.
+In this figure, a darker color of interval point indicates a higher and a lighter color a lower possibility that the corresponding CTI is the intended interval.
 
@@
 \begin{enumerate}
-	\item The visualizations (the points) on the visualization of the time axis of both the smallest and greatest instant with a non-zero possibility of being the starting instant of $J$ are determined.
-	\item Two straight half-lines are constructed from each of these points: one having, in the point, an angle of size $\alpha$ with the visualization of the time axis and one having, in the point, an angle of size $-\alpha$ with the visualization of the time axis.
-The parts of these half-lines that lie outside of the interval space are not drawn.
-	\item The visualizations (the points) on the visualization of the time axis of both the smallest and greatest instant with a non-zero possibility of being the ending instant of $J$ are determined.
-	\item Two straight half-lines are constructed from each of these points: one having, in the point, an angle of size $\alpha$ with the visualization of the time axis and one having, in the point, an angle of size $-\alpha$ with the visualization of the time axis.
-The parts of these half-lines that lie outside of the interval space are not drawn.
+	\item The visualizations (the points) on the time line of both the smallest and greatest instant with a non-zero possibility of being the starting instant of $J$ are determined.
+	\item Two straight half-lines are constructed from each of these points: one having, in the point, an angle of size $\alpha$ with the time line and one having, in the point, an angle of size $-\alpha$ with the time line.
+	\item The visualizations (the points) on the time line of both the smallest and greatest instant with a non-zero possibility of being the ending instant of $J$ are determined.
+	\item Two straight half-lines are constructed from each of these points: one having, in the point, an angle of size $\alpha$ with the time line and one having, in the point, an angle of size $-\alpha$ with the time line.
 	\item The lines constructed this way divide the interval space in at most fifteen different areas, including the area corresponding to the visualization of $J$ itself.
@@
 These symbols and names are used to uniquely identify each of $J$'s URZ.
-The set of Allen relationships corresponding to an URZ contains every Allen relationship in which any interval visualized by one of the URZ's interval points could be with the interval intended by $J$, depending on exactly which time interval $J$ actually intends to represent.
+The set of Allen relationships corresponding to an URZ contains every Allen relationship in which any interval visualized by one of the URZ's interval points could be with the interval intended by $J$, depending on exactly which CTI $J$ actually intends to represent.
 In figure \ref{fig:tm-urz}, the URZ's symbols are noted inside the areas visualizing their respective URZ's.
@@
 
-The TM framework now allows evaluating the Allen relationships a given CTI $I = \left[s, e\right]$ could be in with a given UTI $J$, depending on the exact interval intended by $J$.
+The TM framework now allows evaluating the Allen relationships a given CTI $I = \left[s, e\right]$ could be in with a given UTI $J$, depending on the exact CTI intended by $J$.
 This is done as follows.
 First, $J$ and its URZ are visualized in an interval space.
-Next, $I$'s interval point, its construction half-lines $L_s$ and $L_e$ and the visualization of its starting and ending instants on the visualization of the time are drawn in the same interval space.
+Next, $I$'s interval point, its construction half-lines $L_s$ and $L_e$ and the visualization of its starting and ending instants on the time line are drawn in the same interval space.
 Then, evaluation is a matter of position (in the following, the possibility that $I$ is the interval intended by $J$ is denoted $\pi_J(I)$ and the possibility that $I$ is in Allen relationship $AR$ with $J$ is denoted  $\Pos(I\ AR\ J)$)~\cite{DeTre2012}:
@@
 Here, the intervals $K$ are CTI and the expression `$K \in p_i$' is a notation to express that $K$ is visualized by an interval point contained in area part $p_i$.
-
-%Time intervals are usually represented in a one dimensional scale.
-This is due to its linear characteristic.
-This approach has also been used in Figure\ref{fig:allen-relationships} and \ref{fig:ill-known-ti}.
-The difficulty of accurate visual analysis on the temporal data arises when the number of time intervals increase.
-
-%The triangular model (TM) ~\cite{Kulpa1997}, \cite{Weghe2007} represents time intervals in a two dimensional space called IRterval point completely determines the start and end of the interval.
-The two-dimensional space of interval points is called the interval space and is denoted by IR ~\cite{Kulpa2006}.
-The procedure to represent a time interval $[t_s, t_e]$ as an interval point in a two dimensional space is the following:
-
-%\begin{enumerate}
-%\item Draw each time point of the interval $[t_s, t_e]$ in the horizontal line.
-%\item Construct one straight line $L_s$ from $t_s$ with an angle $\alpha$.
-%\item Construct one straight line $L_e$ from $t_e$ with an angle $-\alpha$.
-%\item The intersection of these two lines is the interval point.
-%\end{enumerate}
-
-%For the shake of simplicity, in this work, the angle for $\alpha$ is chosen to be $45^{\circ}$.
-
-%In TM, these thirteen Allen relations each correspond to a specific zone in the interval space IR \cite{Kulpa1997}.
-These thirteen zones are called Crisp Relational Zones (CRZs).
-The CRZs with respect to a reference interval point $I$ are depicted in Figure \ref{fig:allencrz}.
-
-%\begin{figure}[h]
-%   \centering
-%   \includegraphics[width=0.8\columnwidth]{graphs/allencrc.eps}
-%   \caption{CRZs corresponding to the thirteen Allen's relations.
-The following shorthand notations are used: equals (E), starts (S), started by (SB), finishes (F), finished by (FB), meets (M), met by (MB), overlaps (O), overlapped by (OB), during (D), contains (C), before (B) and after (A).
-}
-%   \label{fig:allencrz}
-% \end{figure}
-
-%Each CRZ represents the set of interval points of the crisp time intervals that are in the corresponding Allen relation with respect to the reference interval $I$.
-For example, all interval points of intervals that come before $I$ are located in the CRZ represented by the leftmost lower triangle in the interval space.
-Likewise, if an interval point is located in the top quadrangle above $I$, then its corresponding interval contains $I$.
-CRZs allow to visually analyse relations between (large) sets of interval points.
-
-%TODO: example!
-
-%In this section we will explain a graphical visualization of temporal data called the triangular model (TM).
-This has been widely used mainly on rough sets for the visualization of historical data.
-More recently, the model has been extended on possibility theory to model Uncertain Time Intervals (UTI).
-%First of all we will explain the main concepts and properties of the triangular model to model Uncertain Time Intervals (UTI).
-Then, we will detail how to compute the Allen's Relations between a crisp time point and an UTI.
-
-%\subsection{\label{subsec:utis}Uncertain Time Intervals}
-%As mentioned in the previous section, the boundaries of time intervals are not always well-known.
-The available knowledge about the interval can be modelled by using possibility distributions ~\cite{Billiet2012}, ~\cite{JoseEnriquePons2012} on both starting and ending points as mentioned in the previous section.
-Informally, an Uncertain Time Interval UTI is a time interval for which one or both boundaries are not precisely known.
-
-%\begin{definition}
-% \label{def:uncertain-time-interval}
-%An uncertain time interval UTI $I$ is composed by an starting point $S$, and an ending point $E$.
-For both points, two different convex possibility distributions are defined, namely $\pi_s$, $\pi_e$.
-These two possibility distributions represent all the possible time intervals that can be constructed from $I$.
-The following formula obtains the possibility $\pi_{I}$ that a crisp interval  given by  $[a, b]$ is constructed from $I$.
-%\begin{equation}
-% \label{eq:possibility-uti}
-%\pi_{I} \left(\left[t_s, t_e\right] \right) = 
-%\begin{cases}
-% \min \left(\pi_s (t_s), \pi_e (t_e) \right), & \mbox{ if } t_s \leq t_e\\
-%0, & \mbox{ otherwise. }
-%\end{cases}
-%\end{equation}
-
-%In the following we will note an UTI as $\pi_{I} = \left(\pi_s, \pi_e \right)$.
-% It is important to notice that, the main diference with respect to the ill-known time interval (see Definition \ref{}) is that an UTI only consider as possible starting and ending points
-
-%\end{definition}
-
-%\subsection{\label{subsec:visualization-tm}Visualization within the triangular model}
-
-%\subsection{\label{subsec:utis-in-tm}Representation of Uncertain Time Intervals}
-%In \cite{DeTre2012} a technique to represent UTIs in the triangular model has been presented.
-Basically, the procedure is an extension of the method to represent crisp time intervals in the triangular space.
-An UTI $\pi_I = \left(\pi_s, \pi_e \right)$ can be represented in the triangular model by the following procedure, considering that both $\pi_s, \pi_e$ are given by triangular membership functions (see Section \ref{subsec:fuzzy-numbers}).
-
-%\begin{enumerate}
-% \item Let $\pi_s = \left[D_s, a_s, b_s\right]$ a triangular membership function.
-%\item Let $\pi_e = \left[D_e, a_e, b_e \right]$ a triangular membership function.
-%\item Let $t_s^{-} = D_s - a_s$, and $t_s^{+} = D_s + b_s$.
-%\item Let $t_e^{-} = D_e -a_e$, and $t_e^{+} = D_e + b_e$.
-%\item Construct two straight lines $L_s^{-}, L_s^{-}$ from $t_s^{-}, t_s^{+}$ with an angle $\alpha$.
-%\item Construct two straight lines $L_e^{-}, L_e^{-}$ from $t_e^{-}, t_e^{+}$ with an angle $-\alpha$.
-%\end{enumerate}
-
-%The intersection of these lines is a quadrangle zone where the UTI is contained.
-Note that not all the interval points in this zone are to the same extent representing candidate intervals for the UTI.
-A grey scaled color gradient is obtained from Eq.
-\eqref{eq:possibility-uti}.
-As shown in Fig.
-\ref{fig:utitm} the color black is associated with a possibility degree of 1, whereas the color white is used when the possibility is 0.
-%Equation \eqref{eq:possibility-uti} computes the possibility obtained for the UTI with respect to a candidate interval.
-% This formula is used to calculate a grey scaled gradient in which the color black denotes the possibility degree 1 and the color white denotes the possibility degree 0.
-
-%\begin{figure}[h]
-%   \centering
-%   \includegraphics[width=0.8\columnwidth]{graphs/utitm.eps}
-%   \caption{Representation in the triangular model for an UTI $\pi_I$.  }
-%   \label{fig:utitm}
-% \end{figure}
-
-%\subsection{\label{subsec:relational-info-uti}Obtaining Relational Information}
-%In order to make temporal reasoning with UTIs, in the triangular model, the thirteen Allen's Relations, and its respective Crisp Relational Zones (CRZ) have been generalized.
-In this case, we obtain fifteen Uncertain Relational Zones (URZ) , shown in Fig.
-\ref{fig:urz} and in Table \ref{tab:urz}.
-
-%\begin{figure}[h]
-%   \centering
-%   \includegraphics[width=0.8\columnwidth]{graphs/urz.eps}
-%   \caption{Uncertain Relational Zones (URZ) for a given UTI.  }
-%   \label{fig:urz}
-% \end{figure}
-
-%The way to obtain the relational information between an interval point $J = \left[s_j, e_j \right]$, and an UTI $\pi_I$ depends on the location of $J$ with respect to the UTI $\pi_I$:
-
-%\begin{itemize}
-% \item If $J$ is located in a zone with only one possible Allen relation, then it is true that $J$ is in the relation. (And therefore, the possibility is always 1).
-%\begin{equation}
-% \label{eq:possibility-iz-uti-certain}
-%\mu_{\left(J R \pi_I \right)} \left(R \right) = 1.
-%\end{equation}
-
-%\item If $J$ is located in a zone with multiple candidate Allen's relations, then there is uncertainty about the Allen relation that applies between $J$ and the UTI $\pi_I$.
-The procedure to obtain the possibility degree is the following:
-
-%\begin{enumerate}
-% \item Draw the lines $L_s, L_e$ that represent the interval point $J$, with the procedure shown in Section \ref{subsec:visualization-tm}. 
-%\item Draw the orthogonal lines from $L_s, L_e$. 
-%\item One or several of these lines could intersect with the UIZ of $\pi_I$.
-Then, the UIZ is divided into as many different subzones $\pi_I^{R'}$ as possible candidate Allen relations $R'$ between $J$ and $\pi_I$ exist.
-%\item $\pi_I^{R'}$ is the zone that correspond to a candidate Allen Relation $R'$.
-%\end{enumerate}
-%The possibility that a candidate relation $R'$ is the actual relation between $J$ and $\pi_I$ is given by:
-%\begin{equation}
-% \label{eq:possibility-iz-uti}
-%\mu_{\left(J R \pi_I \right)} \left(R' \right) = \sup_{I \in \pi_I^{R'}} \pi_I(I).
-%\end{equation}
-%\end{itemize}
-
-%\begin{figure}[h]
-%   \centering
-%   \includegraphics[width=0.8\columnwidth]{graphs/calculation.eps}
-%   \caption{Uncertain Relational Zones (URZ) for a given UTI.  }
-%   \label{fig:calculation}
-% \end{figure}
-
-% Mainly due to its linear characteristic, time is traditionally visualised using a linear representation where several time intervals are depicted as linear segments on top of a horizontal time axis.
-This approach has also been used in Fig.
-\ref{fig:ill-known-ti}.
-One can read the start and end point of each interval the horizontal scale.
-The vertical dimension is solely used to differentiate multiple overlapping intervals, if used at all.
-The linear representation works well as long as the number of represented time intervals remains low.
-As soon as a large number of overlapping intervals have to be represented the representation becomes overloaded and accurate visual data analysis becomes almost impossible.
-
-% For that reason, an alternative Triangular time Model (TM) has been presented in \cite{Weghe2007}.
-This model is an extension of Kulpa's triangular model for the modelling of crisp time intervals \cite{Kulpa1997}.
-Basically, intervals are modelled as interval points in a two-dimensional space as follows.
-First, a horizontal time axis is considered.
-Second, for each time interval $\left[t_s, t_e \right]$ two straight lines $L_s$ and $L_e$ are constructed as depicted in Fig.
-\ref{}.
-Line $L_s$ is passing through $t_s$ and makes a fixed angle $\alpha$ with the horizontal time axis, whereas Le is passing through te making a fixed angle $-\alpha$ with the horizontal time axis.
-The interval $\left[t_s, t_e \right]$ is then represented by the intersection of $L_s$ and $L_e$ , which is called the interval point of $\left[t_s, t_e \right]$.
-For the ease of use, $\alpha$ is here chosen to be $45^{\circ}$ .
-In TM all time intervals are represented by their corresponding interval point.
-So, because $\alpha$ is fixed for all interval points, the position of the interval point completely determines the start and end of the interval.
-The two-dimensional space of interval points is called the interval space and is denoted by IR \cite{Kulpa2006}.
-
-%Figure 3
-
-% From the observation that the start points and end points of two crisp time intervals $I^{1} = \left[t_s^1, t_e^1 \right]$ and $I^2 = \left[t_s^2, t_e^2 \right]$ can be smaller than ($<$), equal to ($=$) or larger than ($>$) each other, Allen proposed the thirteen possible relations between two crisp time intervals given in Table \ref{} \cite{Allen1983}.
-%
-
-% \subsection{\label{subsec:uncertainty-about-time-intervals}Uncertainty about Time Intervals}
-% Time is a complex natural phenomenon that, among others, allows humans to organise life and plan activities.
-Time also plays an important role in information management as it allows to register facts with different kinds of timestamps and make use of these timestamps during querying and information retrieval.
-Dealing with time in an information system requires a time model.
-In this paper, it is implicitly assumed that time is modelled as being discrete, linear and finite.
-This means that it is considered that time can only be/ observed using a limited precision, say $\delta$.
-Observing time using a maximum precision $\delta$ and with respect to a given origin $t_0$ involves a crisp discrete (countable) set of time points, given by
-% \begin{equation}
-%  \label{eq:t0-delta}
-% T_{0,\delta} = \lbrace t_K \in T | t_k = t_0 + k\delta, k \in \Z \rbrace
-% \end{equation}
-% 
-% where $Z = \left \lbrace 0, 1, -1, 2, -2, \ldots \rbrace \right.
-T_{0, \delta}$ is a proper subset of the continuum $T$ of the physical time points.
-The discretisation can be described as a surjective mapping from $T$ on $T_{0,\delta}$, which maps each $\tau \in T$ on the element $t_k \in T_{0,\delta}$ that lies in the interval
-% \begin{equation}
-% \gamma_{0,k} = \left[ t_0 + k\delta, t_0 + (k+1)\delta \right[, k \in \Z 
-% \end{equation}
-% 
-% Discretisation implies that all time points within an interval $\gamma_{0,k}$ are considered to be indistinguishable and mapped to the same time point $t_k$.
-This is not really a limitation, on condition that the precision $\delta$ is chosen sufficiently accurate.
-The discretisation is necessary to circumvent the density problem (i.e., the fact that between any two distinct points, there always exists at least one other point).
-Linearity implies a total order over the time points.
-Finally, our model is chosen to be finite in view of a computer representation.
-This implies that all values that exceed the determined upper and lower bound will be handled by introducing two special values $(- \infty $ and  $+\infty)$.
-% 
-% The intervals $\gamma_{0,k}$ with $\delta$ as length are usually called `chronons' in literature (cf.
-\cite{Dyreson1994}).
-The introduction of chronons eliminates the necessity of making a distinction between points and intervals, because even a chronon, the smallest unit in the model, is essentially an interval.
-% 
-% A time interval $\left[t_s, t_e \right]$ is defined by a start time point $t_s$  and a length $l$, which defines the distance (number of chronons) from ts to the end time point $t_e$, i.e., $l = t_e - t_s$.
-% 
-% 
-% A time point is equivalent to a time interval with $l = 0$ (or $t_e = t_s$ ).
-Note that the length of an interval is not the same as the duration of the interval: the length of a time point is 0, its duration is 1 chronon ($\delta$).
-% 
-% In practice, it often occurs that time intervals cannot be exactly specified.
-Especially, in historical data it is often the case that either the starting date or the ending date (or both) of a time interval are partially or completely unknown.
-In such cases, the best thing to do is modelling the available knowledge about (the uncertainty of) the interval as accurate as possible, hereby avoiding information loss.
-Soft computing techniques, more specifically possibility theory \cite{DidierDubois1988a}, can be used for that purpose.
-In \cite{Billiet2012}, \cite{JoseEnriquePons2012} a possibilistic valid time model able to cope with time intervals that have an uncertain start and/or end point has been presented.
-The basics of this model are used in this paper and briefly described as follows.
-Each (Un)certain Time Interval (UTI) is defined by a pair
-% \begin{equation}
-%  \label{eq:uncertain-time-interval}
-% \left(\pi_s, \pi_e \right)
-% \end{equation}
-% 
-% where $\pi_s$ and $\pi_e$ are two convex possibility distributions, respectively reflecting the knowledge about the start and end point of the UTI (as illustrated in Fig.
-\ref{}(a)).
-In case of certainty, the possibility distribution is characterised by a fuzzy set with singleton support and core containing the crisp date.
-In case of uncertainty, the possibility distribution is characterised by a fuzzy set containing all possible candidate values for the date and their associated degree of possibility.
-% % figura 2
-% 
-% Together, both possibility distributions reflect the available knowledge about the start and end of the UTI they model.
-This implies that $\pi$s and $\pi$e together represent another possibility distribution $\pi$I consisting of all possible time intervals that can be constructed from $\pi$s and $\pi$e .
-Some of the intervals in $\pi_I$ are depicted in Fig.
-\ref{}(b).
-For each time interval $\left[t_s , t_e \right]$, its associated possibility grade in $\pi_I$ is computed by
-% \begin{equation}
-%  \label{eq:possibility-uti}
-% \pi_{I} \left(\left[t_s, t_e\right] \right) = 
-% \begin{cases}
-%  \min \left(\pi_s (t_s), \pi_e (t_e) \right), & \mbox{ if } t_s \leq t_e\\
-% 0, & \mbox{ otherwise. }
-% \end{cases}
-% \end{equation}
-% 
-% Eq.
-\eqref{eq:possibility-uti} reflects that given the uncertainty about the start point, modelled by $\pi$s and the uncertainty about the end point, modelled by $\pi_e$ $\pi$e , the possibility that $\left[t_s, t_e \right]$ is the actual value of the UTI equals the possibility that ts is the actual start point (i.e., $\pi_s (t_s)$) and te is the actual end point (i.e., $\pi_e (t_e)$) of $I$, where the conjunction of both conditions is modelled by the minimum t-norm.
-In the inconsistent case where $t_s > t_e$ the interval is considered to be completely impossible.
-In the remainder of this paper we will denote an UTI as $\pi_I$, i.e., $\pi_I = \left( \pi_s, \pi_e \right)$.
-
-% 
-% \subsection{\label{subsec:the-triangular-model}The triangular model}
-% Mainly due to its linear characteristic, time is traditionally visualised using a linear representation where several time intervals are depicted as linear segments on top of a horizontal time axis.
-This approach has also been used in Fig.
-\ref{}(b).
-One can read the start and end point of each interval the horizontal scale.
-The vertical dimension is solely used to differentiate multiple overlapping intervals, if used at all.
-The linear representation works well as long as the number of represented time intervals remains low.
-As soon as a large number of overlapping intervals have to be represented the representation becomes overloaded and accurate visual data analysis becomes almost impossible.
-% 
-% For that reason, an alternative Triangular time Model (TM) has been presented in \cite{Weghe2007}.
-This model is an extension of Kulpa's triangular model for the modelling of crisp time intervals \cite{Kulpa1997}.
-Basically, intervals are modelled as interval points in a two-dimensional space as follows.
-First, a horizontal time axis is considered.
-Second, for each time interval $\left[t_s, t_e \right]$ two straight lines $L_s$ and $L_e$ are constructed as depicted in Fig.
-\ref{}.
-Line $L_s$ is passing through $t_s$ and makes a fixed angle $\alpha$ with the horizontal time axis, whereas Le is passing through te making a fixed angle $-\alpha$ with the horizontal time axis.
-The interval $\left[t_s, t_e \right]$ is then represented by the intersection of $L_s$ and $L_e$ , which is called the interval point of $\left[t_s, t_e \right]$.
-For the ease of use, $\alpha$ is here chosen to be $45^{\circ}$ .
-In TM all time intervals are represented by their corresponding interval point.
-So, because $\alpha$ is fixed for all interval points, the position of the interval point completely determines the start and end of the interval.
-The two-dimensional space of interval points is called the interval space and is denoted by IR \cite{Kulpa2006}.
-% 
-% %Figure 3
-% 
-% From the observation that the start points and end points of two crisp time intervals $I^{1} = \left[t_s^1, t_e^1 \right]$ and $I^2 = \left[t_s^2, t_e^2 \right]$ can be smaller than ($<$), equal to ($=$) or larger than ($>$) each other, Allen proposed the thirteen possible relations between two crisp time intervals given in Table \ref{} \cite{Allen1983}.
-% 
-% In TM, these thirteen Allen relations each correspond to a specific zone in the interval space IR \cite{Kulpa1997}.
-These thirteen zones are called Crisp Relational Zones (CRZs).
-The CRZs with respect to a reference interval point $I$ are depicted in Fig.
-\ref{}.
-Hereby, the following shorthand notations are used: equals (E), starts (S), started by (SB), finishes (F ), finished by (F B), meets (M ), met by (M B), overlaps (O), overlapped by (OB), during (D), contains (C), before (B) and after (A).
-% 
-% % TABLE I
-% % T HE THIRTEEN POSSIBLE A LLEN RELATIONS BETWEEN CRISP TIME
-% % INTERVALS
-% % 
-% % I 1 equals I 2
-% % I 1 starts I 2
-% % I 1 started by I 2
-% % I 1 f inishes I 2
-% % I 1 f inished by I 2
-% % I 1 meets I 2
-% % I 1 met by I 2
-% % I 1 overlaps I 2
-% % I 1 overlapped by I 2
-% % I 1 during I 2
-% % I 1 contains I 2
-% % I 1 bef ore I 2
-% % I 1 af ter I 2
-% 
-% Each CRZ represents the set of interval points of the crisp time intervals that are in the corresponding Allen relation with respect to the reference interval $I$.
-For example, all interval points of intervals that come before $I$ are located in the CRZ represented by the leftmost lower triangle in the interval space.
-Likewise, if an interval point is located in the top quadrangle above $I$, then its corresponding interval contains $I$.
-CRZs allow to visually analyse relations between (large) sets of interval points.

### 2013-06-19 trunk/eusflat2013/introduction.tex
Reviews included. Now only shortening is required.

--- 2013-02-27 trunk/eusflat2013/introduction.tex
@@
-\begin{table}[h]
-\centering
-\begin{tabular}{|c|l|}
-\hline
-Name & Implementation \\ \hline 
-$I$ equals $J$ & if $s_i = s_j \wedge e_i = e_j $ \\
-$I$ starts $J$ & if $s_i = s_j \wedge e_i < e_j $ \\
-$I$ started by $J$ & if $s_i = s_j \wedge e_i > e_j $ \\
-$I$ finishes $J$ & if $s_i > s_j \wedge e_i = e_j $ \\
-$I$ finished by $J$ & if $s_i < s_j \wedge e_i = e_j $ \\
-$I$ meets $J$ & if $e_i = s_j $ \\
-$I$ met by $J$ & if $s_i = e_j $ \\
-$I$ overlaps $J$ & if $s_i < s_j \wedge e_i < e_j \wedge e_i > s_j $ \\
-$I$ overlapped by $J$ & if $s_i > s_j \wedge e_j < e_i \wedge s_i < e_j  $ \\
-$I$ during $J$ & if $s_i > s_j \wedge e_i < e_j $ \\
-$I$ contains $J$ & if $  s_i < s_j \wedge e_i > e_j$ \\
-$I$ before $J$ & if $e_i < s_j $ \\
-$I$ after $J$ & if $s_i > e_j $ \\
-\hline
-\end{tabular}
-\caption{Allen's relations represented in the framework. $I = \left[s_i, e_i\right]$, $J=  \left[s_j, e_j\right]$}
-\label{tab:allen-relations}
-\end{table}
+Information systems (IS) have always tried to model parts of reality.
+To achieve this modelling, IS contain data representing properties of real-world objects or concepts~\cite{JoseEnriquePons2012},~\cite{Billiet2012}.
+As time is an essential aspect of many real-world objects or concepts, IS often contain data representing temporal notions which describe such temporal properties~\cite{Bolour1982},~\cite{VanderCruyssen1997}.
+Such temporal notions usually take the form of either instants~\cite{Jensen1998}, which can informally be seen as infinitely short `moments' or `points' in time, or time intervals~\cite{Jensen1998}.
+
+Data are often produced by humans, but human-made data are prone to imperfections: some data may be vague, imprecise, incomplete, contradictory or uncertain.
+Data representing temporal notions may contain such imprecisions too~\cite{JoseEnriquePons2012},~\cite{Billiet2012},~\cite{VanderCruyssen1997},~\cite{Devos1994},~\cite{Dubois2003}.
+Although The work presented in this paper is specifically concerned with time intervals (and as a special case: instants) subject to uncertainty.
+
+Generally, one of the most important purposes of an IS is to allow the retrieval of information or knowledge deduced from its data.
+Such retrieval is usually done by querying the IS and examining or analyzing the query results or by visualizing the contents of the IS, querying this visualization and examining or analyzing the resulting visualization(s).
+
+Of course, when temporal information is represented in an IS, querying this IS may have a temporal aspect too.
+Usually, querying such an IS is conceptually done by specifying one or more time indications and requesting information that is in a specific relationship with these indications, where the semantics of these relationships are specifically temporal~\cite{JoseEnriquePons2012},~\cite{Billiet2012},~\cite{Pons2012a}.
+Thus, some existing proposals have considered groups of basic relationships between time indications used to construct and express specific temporal relationships~\cite{Medina1994},~\cite{Schockaert2008},~\cite{Dubois2007}.
+Notably, Allen~\cite{Allen1983} presented a reasoning framework containing all semantically useful basic temporal relationships between time intervals (and as a special case instants) and subsequently used these in a framework to represent and reason with time-related actions~\cite{Allen1984} and compared them with other approaches~\cite{Allen1991}.
+The resulting relationships are shown in figure \ref{fig:allen-relationships}.
+These Allen relationships are used in the presented work.
 
@@
  \end{figure}
+
+To be able to query IS containing data representing time indications subject to uncertainty, a framework is necessary, able to represent uncertainties in time indications in a semantically sound way, without (much) information loss and able to temporally reason with such time indications in a semantically sound and useful way~\cite{Dubois2003},~\cite{Dubois2007},~\cite{Dubois1983}.
+Although more proposals for such frameworks exist, the work presented in this paper focusses on just two: the \emph{ill-known constraint \emph{(IKC)} framework}~\cite{Pons2011} and the \emph{triangular model \emph{(TM)} framework}~\cite{DeTre2012}.
+
+The work presented in this paper consists of a comparison of both frameworks, about the approaches they use to represent time intervals subject to uncertainty and to reason about the temporal relationships between such intervals and time intervals without uncertainty.
+
+The structure of this paper is as follows: section \ref{sec:general-preliminaries} presents some general preliminaries and naming and notation conventions used in this paper.
+Sections \ref{sec:ikc} and \ref{sec:tm} introduce the IKC and TM frameworks respectively: both sections first introduce some specific preliminary concepts and techniques, then explain how the representation of time intervals by the framework is done and finally show how the evaluation of Allen relationships between an interval subject to uncertainty and one without uncertainty is done.
+In section \ref{sec:proposal}, both frameworks are compared: first their approaches to representing time intervals are compared, next their approaches to evaluating Allen relationships.
+Also, an example is given.
+Finally, section \ref{sec:conclusions} presents the principal conclusions of the presented work and some possible future work.

### 2013-06-19 trunk/eusflat2013/tm.tex
Final draft.
Final version.

@@
 Usually, different time intervals are visualized as different parallel line segments in the same image plane.
-However, as this linear approach introduces a few issues, time intervals are visualized as points in the image plane in the TM framework~\cite{Weghe2007},~\cite{DeTre2012}.
+However, as this linear approach introduces a few issues, time intervals are visualized as points in the image plane in the TM framework~\cite{DeTre2012},~\cite{Weghe2007}.
 To achieve this, a horizontal line segment visualizing the used (part of the) time axis is drawn in the image plane (this is called the \emph{time line}).
@@
 The area contained in this triangle is called the \emph{interval space} and will contain all visualizations of intervals.
-Now, the procedure used to visualize a CTI $\left[s, e\right]$ in the interval space is the following~\cite{Weghe2007},~\cite{DeTre2012}:
-
+Now, the procedure used to visualize a CTI $\left[s, e\right]$ in the interval space is the following~\cite{DeTre2012},~\cite{Weghe2007}:
+\vspace{-5pt}
 \begin{enumerate}
@@
 \end{enumerate}
-
-A similar point visualizing an interval is generally called an \emph{interval point}~\cite{Weghe2007},~\cite{DeTre2012}.
+\vspace{-5pt}
+A similar point visualizing an interval is generally called an \emph{interval point}~\cite{DeTre2012},~\cite{Weghe2007}.
 In this work, as it is usually done, the size of $\alpha$ is chosen to be $45^{\circ}$.
@@
 \end{figure}
+\vspace{-10pt}
 
@@
 The possibility $\pi_J(I)$ that a given CTI $I = \left[s_i, e_i\right]$ is the exact time interval intended by $J$ can now be uniquely determined by~\cite{DeTre2012}:
-
+\vspace{-5pt}
 \begin{align}
@@
 \end{itemize}
-
+\vspace{-10pt}
 \begin{table}[h]
@@
 \end{table}
-
+\vspace{-5pt}
 The TM framework now allows evaluating the Allen relationships a given CTI $I = \left[s, e\right]$ could be in with a given UTI $J$, depending on the exact CTI intended by $J$.
@@
 Then, evaluation is a matter of position (in the following, the possibility that $I$ is the interval intended by $J$ is denoted $\pi_J(I)$ and the possibility that $I$ is in Allen relationship $AR$ with $J$ is denoted  $\Pos(I\ AR\ J)$)~\cite{DeTre2012}:
-
+\vspace{-5pt}
 \begin{itemize}

### 2013-06-19 eusflat2013/conclusions.tex
Reviews included. Now only shortening is required.
Final version.

@@
 It is concluded that:
-
+\vspace{-5pt}
 \begin{itemize}
@@
 \end{itemize}
-
+\vspace{-5pt}
 Future research will deal with the evaluation of temporal relationships between several time intervals subject to uncertainty and with querying aspects in both frameworks.
-
-%In this work we have presented an analysis between two different frameworks that deal and represent imperfect temporal information.
-Besides the differences in the visualization of imperfect time intervals, we obtained that the temporal reasoning in both frameworks is equivalent.
-It is also important to notice that whereas the ill-known constraint framework deal with the thirteen Allen's relations, the triangular model distinguishes fifteen possible relational zones and therefore, it is necessary to set a correspondence between the Allen's relations in both frameworks.
-
-%Future research will deal with the analysis between two imperfect time points in each framework and with the representation of open time intervals.
